### 2017-10-28 proposal/proposal.tex
add initial ramblings

\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage[margin=0.5in]{geometry}

\begin{document}

\title{Competing Bandits}
\author{Guy Aridor, Kevin Liu}
\maketitle

In this project we hope to explore the use of bandit models in explaining product selection and innovation in firms. To begin our project, we will focus on a simulation project based on  \hyperref{https://arxiv.org/abs/1702.08533}''Mansour, Slivkins, and Wu (2017) and then proceed to investigating alternative specifications of the model. Thus, our project is will start as a simulation project and, depending on the time and success of the simulation, will consider exploring original research in this topic. Our intention is to proceed as follows:
\begin{enumerate}
\item We hope to reproduce the results of the aforementioned paper via simulation. The method by which we intend to proceed for this is to take the theoretical results regarding the optimal algorithm chosen conditional on the decision-rule of the agents and the assumptions on the prior as given and then to write code for the optimal bandit algorithms for the principals and the decision-rules of the agents. We hope to reproduce the inverted-U relationship between competition and innovation.
\item We will continue utilizing the same model, but investigate the effects of different parameterizations of the model. Namely, we will use the simulation code in the first part of the project to examine the robustness of the results for different distributions for the arms and priors. In doing so we hope to understand what the effects of distributional assumptions on the tradeoff between competition and innovation in this model.
\item After this we will attempt to both modify and extend the model to be focused on a particular setting where we interpret the principals as firms and the agents as consumers. We do this to investigate the motivating question of utilizing a bandit model to explain product selection and research and development between competing firms. This part of the project will be a mix of simulations and theoretical exercises:
\begin{enumerate}
\item The sequence of decisions is taken to be the same as the original model. However, we consider an alternative modelling assumption that breaks up the consumer and firm payoffs. Here we interpret the different arms as being different products (bundles of characteristics) offered by the firm and the firms' payoffs (reward) are interpreted as being the profitability of the arm. Where we deviate from the original model is to model consumers as having stochastic heterogenous utility for different products and not (exactly) linked to the stochastic reward from the arms. This modelling assumption comes from the fact that different arms will have different costs associated with them for the firm (the original model effectively assumes no cost) and thus the optimal arm for the consumer is not necessarily the optimal arm for the firm.
\item The sequence of events in the original model are a good model for online web services where a service determines in realtime which arm to pull. However, we will consider a model where a firm decides which product to produce for a fixed time period and then observes its reward over that time period (the number of agents it attracts) and then decides which product to produce (arm to pull) in the next time period. This kind of modelling decision is more in line with firms that are not online web services.
\item We will likely not have the time to undertake this potential extension of the model as it is will likely be quite difficult to tackle in a month, but we note it here since, if time permits, at least thinking about the intuition and rudimentary simulations for it would be useful. We consider an extension of the model where instead of committing to a MAB algorithm in the first period, the firms update their beliefs not only from the rewards they acrue from consumers who buy from them but also from their observations of the actions of the other firms. The intuition as to why this might be a reasonable extension is that the actions of firms reveal private information about their beliefs about the reward distributions and thus if the actions of the firm are observable then this is also information the other firm can use to update its belief. Thus, this kind of model is more in line with strategic experimentation models.
\end{enumerate}

\end{enumerate}

\end{document}

### 2017-10-29 proposal/proposal.tex
make things make sense

@@
 \usepackage{float}
+\usepackage{amsfonts}
 \usepackage[margin=0.5in]{geometry}
@@
 Namely, we will use the simulation code in the first part of the project to examine the robustness of the results for different distributions for the arms and priors.
-In doing so we hope to understand what the effects of distributional assumptions on the tradeoff between competition and innovation in this model.
-\item After this we will attempt to both modify and extend the model to be focused on a particular setting where we interpret the principals as firms and the agents as consumers.
-We do this to investigate the motivating question of utilizing a bandit model to explain product selection and research and development between competing firms.
-This part of the project will be a mix of simulations and theoretical exercises:
+In doing so we hope to understand what the effects of distributional assumptions on the degree of the tradeoff between competition and innovation in this model.
+\item After this we will attempt to modify the model and ask if we can derive the same tradeoff between competition and innovation but maintain the assumption of rationality.
+We do this to investigate the motivating question of utilizing a bandit model to explain product selection and research and development between competing firms and thus the proposed modelling approach will attempt to follow in this direction.
+This part of the project will be a mix of simulations and theoretical exercises.
+Here are some possible directions we are considering taking:
 \begin{enumerate}
@@
 This modelling assumption comes from the fact that different arms will have different costs associated with them for the firm (the original model effectively assumes no cost) and thus the optimal arm for the consumer is not necessarily the optimal arm for the firm.
-\item The sequence of events in the original model are a good model for online web services where a service determines in realtime which arm to pull.
+For instance, one possible idea is to have heterogeneity in consumer preferences with respect to the characteristics of the action.
+For instance, we could view the payoffs of the arms as being elements in $\mathbb{R}^{k}$ (viewed as, for instance, characteristics of a good) and having different consumers prefer different characteristics.
+Thus, the intuition is that perhaps instead of general deviations in rationality causing exploration, heterogeneity in consumer tastes may generate similar results.
+\item The sequence of events in the original model are a good model for online web services where a principal determines in realtime which arm to pull and agents need to decide principals first.
 However, we will consider a model where a firm decides which product to produce for a fixed time period and then observes its reward over that time period (the number of agents it attracts) and then decides which product to produce (arm to pull) in the next time period.
 This kind of modelling decision is more in line with firms that are not online web services.
+Keeping everything else constant, it will be interesting to see what happens to the simulation results (and to theoretical results) when we change the sequence of decisions to be as such.
 \item We will likely not have the time to undertake this potential extension of the model as it is will likely be quite difficult to tackle in a month, but we note it here since, if time permits, at least thinking about the intuition and rudimentary simulations for it would be useful.
-We consider an extension of the model where instead of committing to a MAB algorithm in the first period, the firms update their beliefs not only from the rewards they acrue from consumers who buy from them but also from their observations of the actions of the other firms.
+We consider an extension of the model where instead of committing to a MAB algorithm in the first period, the firms update their beliefs not only from the rewards they acrue from consumers who buy from them but also from their observations of the actions of the other firms and thus at each period make a strategic decision.
 The intuition as to why this might be a reasonable extension is that the actions of firms reveal private information about their beliefs about the reward distributions and thus if the actions of the firm are observable then this is also information the other firm can use to update its belief.
-Thus, this kind of model is more in line with strategic experimentation models.
+Thus, this kind of model is more in line with standard strategic experimentation models but extended to a dynamic setting.
 \end{enumerate} 
-
 \end{enumerate}

### 2017-12-02 Writeup/writeup.tex
begin writeup

\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{amsfonts}
\usepackage[margin=0.5in]{geometry}

\begin{document}

\title{Competing Bandits}
\author{Guy Aridor, Kevin Liu}
\maketitle

\section*{Introduction / Setup}

In this project we mainly look at the following sequence of actions:
\begin{enumerate}
\item Two principals compete against each other for agents and each can only get reward if an agent picks them over the other.
\item At each $t = 1,...,T$, a single agent enters (and lives only for one period) and selects a principal based on some belief they have about the expected payoff from picking each principal.
\end{enumerate}

This is based off the setup of the model in Mansour, Slivkins, and Wu (2017). We wrote a variety of simulations to look to see if we could reproduce the original results in the paper as well as a few experiments testing certain pertubations of the model.

The first step was to take the model in the paper and make it amenable to simulation as the methodology in the theoretical analysis of the paper made it hard to simulate the beliefs of the agent. To tackle this we assumed that the agent had access to a ``reputation" score that came from a sliding window average of the past $n$ agents that had picked this principal.

Then, we implemented the following four behavioral algorithms that describe the decision rule of the agent given the reputation score:
\begin{enumerate}
\item HardMax where the agent chooses the firm with a larger score, breaking ties uniformly
\item HardMaxRandom: Each agent uses HardMax with probability $1 - \epsilon$ and an action chosen uniformly at random with probability $\epsilon$
\item SoftMax: Each firm is chosen with probability $e^{\alpha score}$ for some given $\alpha$
\item SoftMaxRandom: Each agent uses SoftMax with probability $1 - \epsilon$ and an action chosen uniformly at random with probability $\epsilon$
\end{enumerate}

In the simulations we maintained the assumption in the paper that the principals commit to a learning algorithm in the first period and thus we implemented standard bandit learning algorithms for the principals. In general, we were interested in thinking about the differences that would arise when the principals were running greedy algorithms, non-adaptive exploration algorithms, or adaptive exploration algorithms.

\section*{Experiment 1 - Finite Memory}
Recall that in our simulations we encoded agents' beliefs about the principals as coming from a sliding window average. One set of simulations that we ran considered the consequences of changing the size of this window. This can be interpreted as agents having finite memory so that ``bad" actions by the principals in the past will stop mattering at some point. Intuitively one would expect this to increase exploration even if we fixed agents as playing $HardMax$ and thus being expected utility maximizers. In the paper the main result around $HardMax$ was that ANY deviation from dynamic greedy would cause a principal to lose all remaining agents. However, if there is finite memory we don't necessarily expect this to be the case.

There are two possible ways to define finite memory and we report results from both. The results we report are both the market share that the principals get as well as the regret that they incur since we want to determine both if there is more exploration and if there are differences in the resulting share of agents they get (i.e. if they can recover from ``bad"' choices).

The two definitions of finite memory we explore are:
\begin{enumerate}
\item Agents remember only the last $n$ periods, no matter who was picked. If a principal wasn't picked in the last $n$ periods, then we default back to the original prior.
\item Agents form their score for a principal based on the last $n$ observations they have from that principal.
\end{enumerate}

[Insert some plots]

[Insert some analysis]

\section*{Experiment 2 - Which Algorithm Wins?}
This experiment mainly looked at what happens if we fix the priors of the principals and the agents to be identical at the beginning of the simulation.

We looked at each behavioral assumption and asked, fixing that behavioral assumption, is there any case where playing a better learning algorithm significantly helps the principal get market share?

[insert plots]

\section*{Experiment 3 - Tuning SoftMax}
This experiment mainly looked at what happens if we do a grid search over alpha parameters in tuning SoftMax.

[insert plots]

\section*{Experiment 4 - Prior or Algorithm}
A natural question to ask in this context is whether the correctness of the initial information that the principals have matters more than the learning algorithm they employ. Specifically, we want to test what happens if we fix one principal as having a better prior than the other but that principal plays $StaticGreedy$ and thus only uses their original priors to decide on their actions.   Suppose the other principal has a more ``incorrect" set of initial beliefs but employs a more sophisticated, adaptive exploration algorithm such as Thompson Sampling. Will the principal playing a smarter learning algorithm catch up eventually (for some behavioral assumptions) or is the original prior more important?

\section*{Experiment 5 - Warm Start and HardMax}
What is the effect of a ``warm start" on the performance of HardMax? Our simulations show that with HardMax the initial rounds matter a lot in a particular simulation and whoever wins in the first few rounds takes the entire market. We explore if having a warm start makes it so that the market is more evenly split.

[results]

\section*{Conclusion}
conclude

\end{document}

### 2017-12-03 Writeup/writeup.tex
tweaks to regret + epsilon probs

@@
 \item Two principals compete against each other for agents and each can only get reward if an agent picks them over the other.
+Before the beginning of the game, each principal selects a bandit algorithm to use.
 \item At each $t = 1,...,T$, a single agent enters (and lives only for one period) and selects a principal based on some belief they have about the expected payoff from picking each principal.
+\item The chosen principal runs its bandit algorithm to select an arm. The reward generated by the arm is given to the agent. 
 \end{enumerate}
@@
 \item HardMax where the agent chooses the firm with a larger score, breaking ties uniformly
-\item HardMaxRandom: Each agent uses HardMax with probability $1 - \epsilon$ and an action chosen uniformly at random with probability $\epsilon$
+\item HardMaxRandom: Each agent uses HardMax with probability $1 - \epsilon$, and chooses between the principals uniformly at random with probability $\epsilon$
 \item SoftMax: Each firm is chosen with probability $e^{\alpha score}$ for some given $\alpha$
-\item SoftMaxRandom: Each agent uses SoftMax with probability $1 - \epsilon$ and an action chosen uniformly at random with probability $\epsilon$
+\item SoftMaxRandom: Each agent uses SoftMax with probability $1 - \epsilon$ and chooses between the principals uniformly at random with probability $\epsilon$ %can we remove this?
+SoftMaxWithRandom doesn't exist in the paper, and I think it's pointless because SoftMax already gives a baseline probability to each principal to be chosen
+\item Uniform: Each agent chooses between the two principals uniformly at random
 \end{enumerate}
@@
 In general, we were interested in thinking about the differences that would arise when the principals were running greedy algorithms, non-adaptive exploration algorithms, or adaptive exploration algorithms.
+Bandit algorithms that we implemented include:
+\begin{enumerate}
+\item ExploreThenExploit
+\item StaticGreedy
+\item DynamicGreedy
+\item DynamicEpsilonGreedy
+\item ThompsonSampling
+\item UCB
+\end{enumerate}
 
@@
 
-[Insert some plots]
+\begin{center}
+Agent: HardMax \\Principals: Thompson Sampling vs Thompson Sampling
+ \begin{tabular}{||c c c c c||} 
+ \hline
+ Memory Size & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ \hline\hline
+ 1 & 0.498 & 0.502 & 0.021 & 0.021 \\ 
+ \hline
+ 5 & 0.5219 & 0.478 &  0.119 &  0.132 \\
+ \hline
+ 10 & 0.407 & 0.593 & 0.165  & 0.113 \\
+ \hline
+ 50 & 0.420 & 0.580 & 0.168 & 0.131 \\
+ \hline
+ 100 & 0.460 & 0.540 & 0.161 & 0.160 \\ [1ex] 
+ \hline
+\end{tabular}
 
-[Insert some analysis]
+\vspace{0.75cm}
+Agent: HardMax \\Principals: Thompson Sampling vs Dynamic $\epsilon$-Greedy
+ \begin{tabular}{||c c c c c||} 
+ \hline
+ Memory Size & Thompson Sampling MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ \hline\hline
+ 1 & 0.508 & 0.491 & 0.020 & 0.029 \\ 
+ \hline
+ 5 & 0.464 & 0.536 &  0.128 &  0.109 \\
+ \hline
+ 10 & 0.248 & 0.751 & 0.218  & 0.062 \\
+ \hline
+ 50 & 0.160 & 0.839 & 0.259 & 0.0338 \\
+ \hline
+ 100 & 0.161192 & 0.834 & n/a & n/a \\ [1ex] 
+ \hline
+\end{tabular}
+\end{center}
+For the case of 100, there were many cases when the principal was never chosen!
 
 \section*{Experiment 2 - Which Algorithm Wins?}
-This experiment mainly looked at what happens if we fix the priors of the principals and the agents to be identical at the beginning of the simulation.
+This experiment mainly looked at what happens if we fix the priors of the principals to be identical over the arms and the agents to be identical over the principals at the beginning of the simulation.
 
@@
 
-\section*{Experiment 3 - Tuning SoftMax}
-This experiment mainly looked at what happens if we do a grid search over alpha parameters in tuning SoftMax.
+\section*{Experiment 3 - Tuning SoftMax} % might want to put this one at the top
+This experiment mainly looked at what happens if we do a grid search over alpha parameters in tuning SoftMax. % you have to tune both alpha and epsilon for SoftMax
 
@@
 
+We simulate this in the following scenario. We set $K = 10$
+
 \section*{Experiment 5 - Warm Start and HardMax}
@@
 Our simulations show that with HardMax the initial rounds matter a lot in a particular simulation and whoever wins in the first few rounds takes the entire market.
-We explore if having a warm start makes it so that the market is more evenly split.
+We explore if having a warm start makes it so that the market is more evenly split in a given simulation.\\
 
-[results]
+We experimented with giving a warm start of 0, 5, 25, 50, and 100 observations to each principal.
+We ran the following competing algorithms: UCB vs UCB, UCB vs DynamicGreedy, StaticGreedy vs UCB and observed little effect.
+Regardless of the algorithm that was played or the number of free observations we gave to the principals, the sequence of simulations lead to market shares in each simulation looking roughly the same.
+Namely, in a given simulation one of the principals would take the entire market but since the principals had the same priors over the arms and the agents had the same priors over the principals, it was random who would take the entire market.
+Thus, it seems that free observations for the principal do not impact the resulting market share.
 

### 2017-12-03 BanditAlgorithm.py

@@ -31,7 +31,13 @@ class BanditAlgorithm:
     self.rewardHistory = []
     self.armCounts = [0.0 for i in range(self.banditProblemInstance.K)]
     self.rewardTotal = [0.0 for i in range(self.banditProblemInstance.K)]
-    self.regret = 0
+    self.regret = 0.0
+
+  def getAverageRegret(self):
+    if self.n == 0:
+      print('no luck')
+      return 100
+    return self.regret / float(self.n)
   
   def executeStep(self):
     arm = self.pickAnArm()

### 2017-12-03 DynamicEpsilonGreedy.py

@@ -17,4 +17,4 @@ class DynamicEpsilonGreedy(BanditAlgorithm):
     if rand.random() > self.epsilon:
       return chosenArm
     else:
-      return rand.choice(range(0, self.banditProblemInstance.K))
+      return rand.choice(range(self.banditProblemInstance.K))

### 2017-12-03 HardMaxWithRandom.py

@@ -4,7 +4,7 @@ from HardMax import HardMax
 class HardMaxWithRandom(HardMax):
   # epsilon: the baseline probablility that each principal will be picked with. must be in (0, 0.5).
   # if epsilon=0.1, for example, then 20% of the time, the agent picks randomly between the 2 principals
-  def __init__(self, principals, K, priors=None, epsilon=0.1, memory=50):
+  def __init__(self, principals, K, priors=None, epsilon=0.025, memory=50):
     super(HardMaxWithRandom, self).__init__(principals, K, priors, memory=memory)
     self.epsilon = epsilon
 

### 2017-12-03 Writeup/writeup.tex
update

@@
 
-We simulate this in the following scenario. We set $K = 10$
+We simulate this in the following scenario.
+We set $K = 10$ where 8 of the arms have approximately identical means ($Bernoulli(0.45)$), one of the arms (denote it as arm $B$) has a higher mean ($Bernoulli(0.55)$), and we vary the mean of the ``best" arm (denote it as arm $A$) in increments of 0.1.
+Thus, we consider a simulation where $Bernoulli(0.55), Bernoulli(0.65), Bernoulli(0.75)$.
+
+We suppose that the ``dumb" principal 2 has relatively correct beliefs in that she has a prior such that arm $B$ is the best arm.
+We suppose that the ``smart" principal 1 has perverse beliefs such that she has a prior that the 8 worst arms have the best mean reward ($Beta(0.45, 0.55)$), arm $B$ has slightly worse mean reward ($Beta(0.4, 0.6$), and arm $A$ has substantially worse mean reward ($Beta(0.1, 0.9)$).
+
+We run simulations under this setup to see if the ``smart" principal's learning algorithm can let her overcome the bad beliefs.
+
+\begin{center}
+Agent: HardMax, T = 5000
+ \begin{tabular}{||c c c c c c||} 
+ \hline
+ Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ \hline\hline
+ 0.55 & ThompsonSampling & 0.440 & 0.560 & 0.049 & 0.00 \\ 
+ \hline
+ 0.65 & ThompsonSampling & 0.647 & 0.353 &  0.057 &  0.100 \\
+ \hline
+ 0.75 & ThompsonSampling & 0.710 & 0.291 & 0.059  & 0.200 \\
+ \hline
+ 0.55 & Dynamic $\epsilon$-greedy & 0.298 & 0.656 & 0.057 & 0.000 \\
+ \hline
+  0.65 & Dynamic $\epsilon$-greedy & 0.335 & 0.664 & 0.138 & 0.100 \\
+ \hline
+  0.75 & Dynamic $\epsilon$-greedy & 0.442 & 0.558 & 0.154 & 0.200 \\[1ex]
+  \hline
+\end{tabular}
+\end{center}
+\vspace{0.75cm}
+\begin{center}
+Agent: HardMaxWithRandom, T = 5000
+ \begin{tabular}{||c c c c c c||} 
+ \hline
+ Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ \hline\hline
+ 0.55 & ThompsonSampling & 0.317 & 0.683 & 0.031 & 0.00 \\ 
+ \hline
+ 0.65 & ThompsonSampling & 0.630 & 0.369 &  0.028 &  0.100 \\
+ \hline
+ 0.75 & ThompsonSampling & 0.787 & 0.291 & 0.015  & 0.200 \\
+ \hline
+ 0.55 & Dynamic $\epsilon$-greedy & 0.343 & 0.702 & 0.038 & 0.000 \\
+ \hline
+  0.65 & Dynamic $\epsilon$-greedy & 0.400 & 0.600 & 0.076 & 0.100 \\
+ \hline
+  0.75 & Dynamic $\epsilon$-greedy &0.615 & 0.385 & 0.074 & 0.200 \\[1ex]
+  \hline
+\end{tabular}
+\end{center}
+
+\vspace{0.75cm}
+\begin{center}
+Agent: SoftMax, T = 5000
+ \begin{tabular}{||c c c c c c||} 
+ \hline
+ Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ \hline\hline
+ 0.55 & ThompsonSampling & 0.442 & 0.558 & 0.027 & 0.00 \\ 
+ \hline
+ 0.65 & ThompsonSampling & 0.600 & 0.400 &  0.025 &  0.100 \\
+ \hline
+ 0.75 & ThompsonSampling & 0.721 & 0.280 & 0.026  & 0.200 \\
+ \hline
+ 0.55 & Dynamic $\epsilon$-greedy & 0.442 & 0.556 & 0.038 & 0.000 \\
+ \hline
+  0.65 & Dynamic $\epsilon$-greedy & 0.400 & 0.600 & 0.021 & 0.100 \\
+ \hline
+  0.75 & Dynamic $\epsilon$-greedy & 0.666 & 0.334 & 0.050 & 0.200 \\[1ex]
+  \hline
+\end{tabular}
+\end{center}
+\vspace{0.25cm}
+Regardless of the behavioral model of the agent, we see that the ``dumb" principal 2 wins a higher share of the market than principal 1 when she happens to have correct prior beliefs about the best arm.
+This happens whether principal 1 plays an adaptive or a non-adaptive exploration algorithm.
+Intuitively this makes sense as the ``dumb" principal will end up getting regret 0 simply because of her correct beliefs and it is incredibly hard for a smarter learning algorithm to make up for this.
+\\
+\\
+However, as we increase the reward of arm $A$ such that principal 2 no longer has completely correct beliefs (she believes the second-best arm is the best arm), we see that when principal 1 plays an adaptive exploration algorithm she catches up very quickly and takes most of the market.
+However, when principal 1 plays a non-adaptive exploration algorithm we see that the ``dumb" algorithm with better initial information still wins out, especially when the agent is HardMax.
+As we make the agent ``more behavioral", the non-adaptive exploration algorithm becomes progressively better.\\
+\\
+The previous results show that, especially when playing an adaptive exploration algorithm, it is possible that a better learning algorithm helps overcome initial bad information even if the opponent has almost correct information.
+However, one expects this takes time and that perhaps if we set $T$ to be lower, the ``dumb" principal would still win.
+To test this, we re-run the same simulations as above but set $T=500$
+
+\begin{center}
+Agent: HardMax, T = 500
+ \begin{tabular}{||c c c c c c||} 
+ \hline
+ Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ \hline\hline
+ 0.55 & ThompsonSampling & 0.181 & 0.812 & 0.088 & 0.0 \\
+ \hline
+ 0.65 & ThompsonSampling & 0.287 & 0.712 & 0.178 &  0.100 \\
+ \hline
+ 0.75 & ThompsonSampling & 0.274 & 0.723 & 0.244  & 0.200 \\
+ \hline
+ 0.55 & Dynamic $\epsilon$-greedy & 0.183 & 0.817 & 0.100 & 0.000 \\
+ \hline
+  0.65 & Dynamic $\epsilon$-greedy & 0.169 & 0.831 & 0.190 & 0.100 \\
+ \hline
+  0.75 & Dynamic $\epsilon$-greedy & 0.297 & 0.703 & 0.251 & 0.200 \\[1ex]
+  \hline
+\end{tabular}
+\end{center}
+\vspace{0.25cm}
+This demonstrates that part of the results from above were driven by the large $T$ and that with a low $T$ of 500, the ``smart" learning algorithm does not have enough time to overcome the bad initial information.
+Thus we conclude that the ``smart" learning algorithm will matter if the number of rounds is sufficiently high but for a low number of rounds, better initial information may be more important than a better learning algorithm.
 
@@
 What is the effect of a ``warm start" on the performance of HardMax?
+In this case we define a ``warm start" as giving the principal free observations (but do not change the information set of the agent).
 Our simulations show that with HardMax the initial rounds matter a lot in a particular simulation and whoever wins in the first few rounds takes the entire market.

### 2017-12-03 UCB.py

@@ -17,5 +17,6 @@ class UCB(BanditAlgorithm):
     totalCounts = sum(self.armCounts)
     for a in range(numArms):
       bonus = math.sqrt((2 * math.log(totalCounts)) / float(self.armCounts[a]))
-      ucbValues[a] = (self.rewardTotal[a] / self.armCounts[a]) + bonus
+      meanReward = self.rewardTotal[a] / float(self.armCounts[a])
+      ucbValues[a] = meanReward + bonus
     return np.argmax(ucbValues)

### 2017-12-04 Writeup/writeup.tex
commit pres

@@
-\documentclass[a4paper]{article}
+\documentclass[11pt,letterpaper]{article}
+\usepackage{fullpage}
 
@@
 
-In this project we mainly look at the following sequence of actions:
+In this project we look at a game played between two competing bandit algorithms:
 \begin{enumerate}
-\item Two principals compete against each other for agents and each can only get reward if an agent picks them over the other.
-Before the beginning of the game, each principal selects a bandit algorithm to use.
-\item At each $t = 1,...,T$, a single agent enters (and lives only for one period) and selects a principal based on some belief they have about the expected payoff from picking each principal.
-\item The chosen principal runs its bandit algorithm to select an arm. The reward generated by the arm is given to the agent. 
+\item Agents (who can be thought of as customers or users) choose between two competing principals (who can be thought of as firms).
+Each agent does so based on running an algorithm and some beliefs about the rewards they will receive from each principal.
+Agents seek to maximize their reward.
+In our implementation, an agent's beliefs are based on the realized rewards that prior agents received, which differs from the paper (in which each agent has no information on what happened before them).
+There are $T$ agents in each simulation, so the simulation lasts for $T$ rounds.
+\item Before the beginning of the game, each principal selects a bandit algorithm to use, and starts with some prior beliefs on the distributions of the $K$ arms.
+Each time the principal is chosen by an agent, the principal runs its bandit algorithm and selects an arm.
+The reward (drawn from a Bernoulli distribution in our simulations) is given to the agent, and the $(arm, reward)$ data point is used to Bayesian update the principal's posterior (which is a Beta distribution) on that arm.
+Each principal seeks to maximize its market share (the fraction of agents that chooses them), and can do so by choosing arms with high rewards so that agents will choose them.
+\item At each $t = 1,...,T$, a single agent enters (and lives only for one period) and selects a principal according to the Agent algorithm (throughout the simulation, every agent uses the same algorithm) and the agent's beliefs (which, in our implementation, varies between agents according to past information).
+\item The chosen principal runs its bandit algorithm to select an arm. The reward generated by the arm is given to the agent. The principal not chosen is not aware of the $(arm, reward)$ pair.
 \end{enumerate}
@@
 This is based off the setup of the model in Mansour, Slivkins, and Wu (2017).
+For more specific details on the model, please refer to their paper.
+We note any meaningful departures we took from their paper in this paper.
 We wrote a variety of simulations to look to see if we could reproduce the original results in the paper as well as a few experiments testing certain pertubations of the model.
@@
 \begin{enumerate}
-\item HardMax where the agent chooses the firm with a larger score, breaking ties uniformly
+\item HardMax: The agent chooses the firm with a larger score, breaking ties uniformly
 \item HardMaxRandom: Each agent uses HardMax with probability $1 - \epsilon$, and chooses between the principals uniformly at random with probability $\epsilon$
-\item SoftMax: Each firm is chosen with probability $e^{\alpha score}$ for some given $\alpha$
-\item SoftMaxRandom: Each agent uses SoftMax with probability $1 - \epsilon$ and chooses between the principals uniformly at random with probability $\epsilon$ %can we remove this?
-SoftMaxWithRandom doesn't exist in the paper, and I think it's pointless because SoftMax already gives a baseline probability to each principal to be chosen
-\item Uniform: Each agent chooses between the two principals uniformly at random
+\item SoftMax: The principals are chosen randomly according to a logistic function, whose steepness is controlled by $\alpha$, and each principal has a minimum probability $\epsilon$ of being chosen.
+Further detail is given in a section below.
+\item Uniform: Each agent chooses between the two principals uniformly at random. Though we implemented this algorithm, we did not end up using it. 
 \end{enumerate}
@@
 \begin{enumerate}
-\item ExploreThenExploit
-\item StaticGreedy
-\item DynamicGreedy
-\item DynamicEpsilonGreedy
-\item ThompsonSampling
-\item UCB
+\item ExploreThenExploit: The algorithm takes $X$ exploration steps divided evenly among the $K$ arms, according to a modulus function.
+After the $X$ exploration steps, it always selects the arm it believes has the highest mean.
+\item StaticGreedy: Always select the arm whose mean is highest according to the prior distributions.
+\item DynamicGreedy: Each round it's run, it selects the arm whose mean is highest according to the posterior distributions at that moment in time.
+\item DynamicEpsilonGreedy: With probability $\epsilon$, picks an arm at random. With probability $1-\epsilon$, runs DyanmicGreedy according to the above. 
+\item ThompsonSampling: For each of the $K$ posterior distributions, sample a point from each distribution.
+Take the highest point, and choose the arm corresponding to the distribution that the point came from.
+\item UCB: Select the arm whose confidence interval has the highest upper bound. 
 \end{enumerate}
 
+\section*{Summary of Results}
+In this project, we looked at a competing bandits framework and explored several experiments using it.
+In our first experiment we looked at the effects on exploration from imposing finite memory on the agents so that they quickly forget the past.
+In this experiment we showed that with sufficiently low memory we can generate exploration even when agents are perfect expected utility maximizers.
+In our second experiment we ran an exhaustive experiment comparing the results of various learning algorithms competing against each other.
+[Talk about results].
+In our third experiment we explored the consequences of different parameterizations of the SoftMax behavioral model.
+[Talk about results].
+In our fourth experiment we tried to answer the question of if the learning algorithm or correctness of the initial information is more important in determining which principal gets higher market share.
+Our simulations indicated that with enough time, adaptive exploration with poor initial information will catch up with a static greedy algorithm with better initial information.
+Non-adaptive exploration however will catch up at a slower pace.
+Our fifth experiment attempted to explore what the consequences of giving the principals free information was if the agents were expected utility maximizers and we found that this has virtually no effect.
+
 \section*{Experiment 1 - Finite Memory}
-Recall that in our simulations we encoded agents' beliefs about the principals as coming from a sliding window average.
+In our simulations we encoded agents' beliefs about the principals as coming from a sliding window average of the rewards previous agents had experienced.
 One set of simulations that we ran considered the consequences of changing the size of this window.
-This can be interpreted as agents having finite memory so that ``bad" actions by the principals in the past will stop mattering at some point.
+This can be interpreted as agents having finite memory so that ``bad" actions by the principals in the past will be forgotten at some point.
 Intuitively one would expect this to increase exploration even if we fixed agents as playing $HardMax$ and thus being expected utility maximizers.
@@
 
-There are two possible ways to define finite memory and we report results from both.
+There are two reasonable ways to define finite memory in the context of this model, but we report results from only one.
 The results we report are both the market share that the principals get as well as the regret that they incur since we want to determine both if there is more exploration and if there are differences in the resulting share of agents they get (i.e.
@@
 
-The two definitions of finite memory we explore are:
+The two definitions of finite memory are:
 \begin{enumerate}
@@
 \end{enumerate}
+
+We only simulate definition 2 and the results are as follows:
 
@@
  \hline
- 5 & 0.5219 & 0.478 &  0.119 &  0.132 \\
+ 5 & 0.566 & 0.433 & 0.041 & 0.068 \\
  \hline
- 10 & 0.407 & 0.593 & 0.165  & 0.113 \\
+ 10 & 0.540 & 0.460 & 0.062 & 0.080 \\
  \hline
- 50 & 0.420 & 0.580 & 0.168 & 0.131 \\
+ 50 & 0.480 & 0.520 & 0.109 & 0.103 \\
  \hline
- 100 & 0.460 & 0.540 & 0.161 & 0.160 \\ [1ex] 
+ 100 & 0.320 & 0.680 & 0.130 & 0.067 \\ [1ex] 
  \hline
@@
  \hline
- Memory Size & Thompson Sampling MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ Memory Size & Dynamic $\epsilon$-Greedy MS  & Thompson Sampling MS & D $\epsilon$ Avg Regret & TS Avg Regret  \\ [0.5ex] 
  \hline\hline
- 1 & 0.508 & 0.491 & 0.020 & 0.029 \\ 
+ 1 & 0.495 & 0.505 & 0.035 & 0.032 \\ 
  \hline
- 5 & 0.464 & 0.536 &  0.128 &  0.109 \\
+ 5 & 0.607 & 0.3926 & 0.052 & 0.079 \\
  \hline
- 10 & 0.248 & 0.751 & 0.218  & 0.062 \\
+ 10 & 0.531 & 0.469 & 0.061 & 0.065 \\
  \hline
- 50 & 0.160 & 0.839 & 0.259 & 0.0338 \\
+ 50 & 0.402 & 0.598 & 0.131 & 0.0801 \\
  \hline
- 100 & 0.161192 & 0.834 & n/a & n/a \\ [1ex] 
+ 100 & 0.520 & 0.480 & 0.111 & 0.109 \\ [1ex] 
  \hline
@@
 \end{center}
-For the case of 100, there were many cases when the principal was never chosen!
+
+These results show that regret roughly increases as the size of memory increases.
+One of the results from the paper our setting is based off is that in HardMax the principals cannot do any exploration otherwise they'll lose for the remaining periods.
+In this simulation, when memory is large, we see that in a particular simulation (as opposed to on aggregate) one of the principals always ends up getting almost every agent that comes into the market and there is little exploration, which is in line with the results from the paper.
+However, when memory is small, there is a more even split of the market in a given simulation and we have that average regret is lower (and thus there is more exploration on average in a given simulation).
+This makes sense given the fact that agents have a short-term memory and thus principals may not necessarily be punished for too long for sub-optimal decisions.
+Thus we see that we can get ``innovation" even without any behavioral deviations by assuming that agents forget the past eventually.
 
 \section*{Experiment 2 - Which Algorithm Wins?}
-This experiment mainly looked at what happens if we fix the priors of the principals to be identical over the arms and the agents to be identical over the principals at the beginning of the simulation.
+This experiment mainly looks at which learning algorithm ``wins" when the two firms do not play the same algorithm.
+In this experiment we fix the priors of the principals to be identical over the arms and the agents to be identical over the principals at the beginning of the simulation.
 
-We looked at each behavioral assumption and asked, fixing that behavioral assumption, is there any case where playing a better learning algorithm significantly helps the principal get market share?
+We looked at each behavioral assumption and asked, fixing that behavioral assumption, is there any case where playing a better learning algorithm helps the principal get a higher market share?
 
-[insert plots]
+\begin{center}
+Agent: HardMax, T = 5000
+ \begin{tabular}{||c c c c c c||} 
+ \hline
+ Principal 1 Alg & Principal 2 Alg & Principal 1 MS & Principal 2 MS & P1 Avg Regret & P2 Avg Regret  \\ [0.5ex] 
+ \hline\hline
+ StaticGreedy & ExploreThenExploit & 0.352 & 0.648 & 0.200 & 0.084 \\ 
+ \hline
+ StaticGreedy & Dynamic $\epsilon$-greedy & 0.510 & 0.490 & 0.2 & 0.098\\
+ \hline
+ DynamicGreedy & ExploreThenExploit & 0.444 & 0.556 & 0.129 & 0.093 \\ 
+ \hline
+ DynamicGreedy & Dynamic $\epsilon$-greedy & 0.486 & 0.514 & 0.124 & 0.111 \\
+ \hline
+  UCB1 & ExploreThenExploit & 0.360 & 0.640 & 0.156 & 0.080 \\
+  \hline
+  UCB1 & Dynamic $\epsilon$-greedy & 0.335 & 0.664 & 0.138 & 0.100 \\
+  \hline
+  ThompsonSampling & ExploreThenExploit & 0.480 & 0.512 & 0.094 & 0.102 \\
+  \hline
+  ThompsonSampling & Dynamic $\epsilon$-greedy & 0.559 & 0.441 & 0.086 & 0.118 \\
+  \hline
+    ThompsonSampling & StaticGreedy & 0.335 & 0.664 & 0.138 & 0.100 \\
+  \hline
+    ThompsonSampling & DynamicGreedy & 0.508 & 0.492 & 0.085 & 0.137 \\
+  \hline
+  UCB1 & StaticGreedy & 0.640& 0.360 & 0.119 & 0.2 \\
+  \hline
+      UCB1 & DynamicGreedy & 0.713 & 0.287 & 0.111 & 0.180\\
+  \hline
+   UCB1 & ThompsonSampling & 0.40 & 0.600 & 0.148 & 0.086 \\
+  \hline
+\end{tabular}
+\end{center}
 
@@
 [insert plots]
-
 \section*{Experiment 4 - Prior or Algorithm}
@@
 Suppose the other principal has a more ``incorrect" set of initial beliefs but employs a more sophisticated, adaptive exploration algorithm such as Thompson Sampling.
-Will the principal playing a smarter learning algorithm catch up eventually (for some behavioral assumptions) or is the original prior more important?
+Will the principal playing a smarter learning algorithm catch up eventually (for any of the behavioral assumptions) or is the original prior more important?
 
@@
 We set $K = 10$ where 8 of the arms have approximately identical means ($Bernoulli(0.45)$), one of the arms (denote it as arm $B$) has a higher mean ($Bernoulli(0.55)$), and we vary the mean of the ``best" arm (denote it as arm $A$) in increments of 0.1.
-Thus, we consider a simulation where $Bernoulli(0.55), Bernoulli(0.65), Bernoulli(0.75)$.
+Thus, we consider a simulation where arm $A$ has true distributions of $Bernoulli(0.55), Bernoulli(0.65), Bernoulli(0.75)$.
 
@@
  \hline
- Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & P1 Avg Regret & P2 Avg Regret  \\ [0.5ex] 
  \hline\hline
- 0.55 & ThompsonSampling & 0.440 & 0.560 & 0.049 & 0.00 \\ 
+ 0.55 & ThompsonSampling & 0.440 & 0.560 & 0.049 & 0.000 \\ 
  \hline
@@
  \hline
- Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & P1 Avg Regret & P2 Avg Regret  \\ [0.5ex] 
  \hline\hline
- 0.55 & ThompsonSampling & 0.317 & 0.683 & 0.031 & 0.00 \\ 
+ 0.55 & ThompsonSampling & 0.317 & 0.683 & 0.031 & 0.000 \\ 
  \hline
@@
  \hline
- Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & P1 Avg Regret & P2 Avg Regret  \\ [0.5ex] 
  \hline\hline
- 0.55 & ThompsonSampling & 0.442 & 0.558 & 0.027 & 0.00 \\ 
+ 0.55 & ThompsonSampling & 0.442 & 0.558 & 0.027 & 0.000 \\ 
  \hline
@@
 \vspace{0.25cm}
-Regardless of the behavioral model of the agent, we see that the ``dumb" principal 2 wins a higher share of the market than principal 1 when she happens to have correct prior beliefs about the best arm.
+Regardless of the behavioral model of the agent, we see that the ``dumb" principal 2 wins a higher share of the market than principal 1 when she happens to have correct prior beliefs about the best arm (when the distribution of arm $A$ is $Bernoulli(0.55)$).
 This happens whether principal 1 plays an adaptive or a non-adaptive exploration algorithm.
@@
  \hline
- Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
+ Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & P1 Avg Regret & P2 Avg Regret  \\ [0.5ex] 
  \hline\hline
@@
 What is the effect of a ``warm start" on the performance of HardMax?
-In this case we define a ``warm start" as giving the principal free observations (but do not change the information set of the agent).
 Our simulations show that with HardMax the initial rounds matter a lot in a particular simulation and whoever wins in the first few rounds takes the entire market.
@@
 
-\section*{Conclusion}
-conclude
+\section*{References}
+\begin{enumerate}
+\item Mansour Y., Slivkins A., Wu S. (2017) Competing bandits: Learning under Competition, The 9th Innovations in Theoretical Computer Science (ITCS'18)
+\end{enumerate}
 

### 2017-12-04 pres/pres.tex
commit pres

%% LyX 2.2.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{beamer}
\usepackage[utf8x]{inputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{amstext}
\usepackage{esint}
\usepackage{stackengine}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
 % this default might be overridden by plain title style
 \newcommand\makebeamertitle{\frame{\maketitle}}%
 % (ERT) argument for the TOC
 \AtBeginDocument{%
   \let\origtableofcontents=\tableofcontents
   \def\tableofcontents{\@ifnextchar[{\origtableofcontents}{\gobbletableofcontents}}
   \def\gobbletableofcontents#1{\origtableofcontents}
 }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\AtBeginDocument{%
   \let\origtableofcontents=\tableofcontents
   \def\tableofcontents{\@ifnextchar[{\origtableofcontents}{\gobbletableofcontents}}
   \def\gobbletableofcontents#1{\origtableofcontents}
 }% Choose how your presentation looks.
%
% For more themes, color themes and font themes, see:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
%
\mode<presentation> {
  \usetheme{default}      % or try Darmstadt, Madrid, Warsaw, ...
  \usecolortheme{default} % or try albatross, beaver, crane, ...
  \usefonttheme{default}  % or try serif, structurebold, ...
  \setbeamertemplate{navigation symbols}{}
  \setbeamertemplate{caption}[numbered]
}

\usepackage[english]{babel}
\usepackage{dsfont}

\newtheorem{prop}{Proposition}

\title[]{Competing Bandits}
\author{Guy Aridor, Kevin Liu}
\institute{}
\date{\today}

\usepackage{babel}

\makeatother

\usepackage{babel}
\begin{document}
\newcommand\barbelow[1]{\stackunder[1.2pt]{$#1$}{\rule{.8ex}{.075ex}}}
\begin{frame}
\titlepage 
\end{frame}

\section{Introduction}
\begin{frame}
{Introduction} 
\begin{itemize}
\item Focused on a ``Competing Bandits" model similar to Mansour, Slivkins, Wu (2017) where we have two principals that compete for agents
\item Principals can only learn if the agents select them and aim to maximize the number of agents which select them
\item We look at this model from a simulation perspective and run several experiments asking how certain variants of the setup change the results
\end{itemize}
\end{frame}
%
\begin{frame}
{Model Setup}
\begin{enumerate}
\item Agents choose between two competing principals (who can be thought of as firms). Agents seek to maximize their reward. In our implementation, an agent's beliefs are based on the realized rewards that prior agents received.
\item Before the beginning of the game, each principal selects a bandit algorithm to use, and starts with some prior beliefs on the distributions of the $K$ arms. Each time the principal is chosen by an agent, the principal runs its bandit algorithm and selects an arm. Each principal seeks to maximize its market share, and can do so by choosing arms with high rewards so that agents will choose them. 
\item At each $t = 1,...,T$, a single agent enters (and lives only for one period) and selects a principal according to the Agent algorithm and the agent's beliefs.
\item The chosen principal runs its bandit algorithm to select an arm. The reward generated by the arm is given to the agent. The principal not chosen is not aware of the $(arm, reward)$ pair.
\end{enumerate}
\end{frame}
%
\begin{frame}
{Algorithm Preliminaries}
\begin{itemize}
\item Principals:
\begin{enumerate}
\item Adaptive Exploration 
\begin{itemize}
\item UCB1
\item Thompson Sampling
\end{itemize}
\item Non-Adaptive Exploration
\begin{itemize}
\item Dynamic $\epsilon$-greedy
\item Explore Then Exploit
\end{itemize}
\item Greedy
\begin{itemize}
\item Bayesian Greedy
\item Static Greedy
\end{itemize}
\end{enumerate}
\item Agents:
\begin{enumerate}
\item HardMax - Perfect expected utility maximizers based on beliefs over principals
\item HardMaxWithRandom - With probability $\epsilon$ play random actions, otherwise play HardMax chosen actions
\item SoftMax - The principals are chosen randomly according to a logistic function (higher scores $\implies$ higher probability of being chosen)
\end{enumerate}
\end{itemize}
\end{frame}
%
\begin{frame}
{Simulation Details}
\begin{itemize}
\item Considered only Bernoulli rewards with Beta Priors
\item Simulations have $K = 10$
\item Reputation score generated by a sliding window average (fix memory sufficiently high except for limited memory experiment)
\end{itemize}
\end{frame}
%
\begin{frame}
{Experiments - Overview}
\begin{itemize}
\item In our simulations we considered the following five experiments:
\begin{enumerate}
\item Competing algorithms
\item Limited memory
\item Tuning SoftMax
\item Prior or Algorithm?
\item Effects of ``Warm start"
\end{enumerate}
\end{itemize}
\end{frame}
%
\frame{\frametitle{Experiment 1 - Competing Algorithms}
\begin{itemize}
\item In this experiment, we ran several competing algorithms against each other.
\item results
\end{itemize}
}
%
\begin{frame}
{Experiment 2 - Limited Memory}
\begin{itemize}
\item In this experiment we consider what happens when agents are perfect expected utility maximizers, but have limited memory.
\item The driving question is can you get sufficient exploration in any given simulation simply by limiting how much agents remember?
\includegraphics[scale=0.5]{memory_exp}
\end{itemize}
\end{frame}
%
\begin{frame}
{Experiment 3 - Tuning SoftMax}
\begin{itemize}
\item SoftMax is great!
\end{itemize}
\end{frame}
%
\begin{frame}
{Experiment 4 - Prior or Algorithm}
\begin{itemize}
\item The purpose of this experiment was to attempt to figure out if it's better to use a better learning algorithm matters more or have better initial information.
\item Ran simulations competing StaticGreedy (with almost correct beliefs) vs Thompson Sampling (with perverse beliefs) with $K = 10$
\item StaticGreedy had highest expected mean on the 2nd best arm, but Thompson Sampling had priors that thought the worst arms were the best and that the best arms were very bad. Who wins?
\includegraphics[scale=0.5]{prior_vs_alg_1}
\end{itemize}
\end{frame}
%
\begin{frame}
{Experiment 4 -  Prior or Algorithm Cont'd}
\begin{itemize}
\item What if we drastically reduce the number of periods in the simulation? \\
\includegraphics[scale=0.5]{prior_vs_alg_2}
\item Given enough time, the algorithms that explore will catch up, but if we have sufficiently few periods then better initial information seems to win (according to our parameterization)
\end{itemize}
\end{frame}
%
\begin{frame}
{Additional Considerations}
\begin{itemize}
\item What if we gave the principals free information at the start by giving them $N$ agents? Some simulations showed that there was no difference.
\item What if we made the rewards of the agents separate from the rewards of the principals?
\end{itemize}
\end{frame}
\end{document}

### 2017-12-05 Writeup/writeup.tex
commit pres + writeup

@@
 Agents seek to maximize their reward.
-In our implementation, an agent's beliefs are based on the realized rewards that prior agents received, which differs from the paper (in which each agent has no information on what happened before them).
+In our implementation, an agent's beliefs are based on the realized rewards that prior agents received, which differs from the paper (in which each agent has no information on what happened before them but has beliefs about what has happened).
 There are $T$ agents in each simulation, so the simulation lasts for $T$ rounds.
@@
 In our second experiment we ran an exhaustive experiment comparing the results of various learning algorithms competing against each other.
-[Talk about results].
+Our results show that Thompson Sampling wins against the other algorithms.
+As well, the results of which algorithm wins seem to be consistent across all the behavioral models we considered.
 In our third experiment we explored the consequences of different parameterizations of the SoftMax behavioral model.
-[Talk about results].
+We found that the results are sensitive to the $\alpha$ that we pick, with $\alpha = 10$ generating results we expected.
+We found that once we set $\alpha = 10$, the value of $\epsilon$ does not make a big difference.
 In our fourth experiment we tried to answer the question of if the learning algorithm or correctness of the initial information is more important in determining which principal gets higher market share.
@@
 Our fifth experiment attempted to explore what the consequences of giving the principals free information was if the agents were expected utility maximizers and we found that this has virtually no effect.
+
+A few notes on the simulations / reported results:
+\begin{enumerate}
+\item MS in the figures means market share (\% of agents that picked the principal)
+\item Unless otherwise noted, the reported results are an average over 25 simulations
+\item The distribution of the arms in all of the simulations is a Bernoulli distribution and the principals have Beta priors.
+\end{enumerate}
 
@@
  \hline
- Memory Size & Dynamic $\epsilon$-Greedy MS  & Thompson Sampling MS & D $\epsilon$ Avg Regret & TS Avg Regret  \\ [0.5ex] 
+ Memory Size & Dynamic $\epsilon$-Greedy MS  & Thompson Sampling MS & D-$\epsilon$ Avg Regret & TS Avg Regret  \\ [0.5ex] 
  \hline\hline
@@
 This experiment mainly looks at which learning algorithm ``wins" when the two firms do not play the same algorithm.
-In this experiment we fix the priors of the principals to be identical over the arms and the agents to be identical over the principals at the beginning of the simulation.
+In this experiment we fix the priors of the principals to be identical over the arms ($Beta(0.5, 0.5)$ prior distributions over all arms) and the agent priors to be identical ($Beta(0.5, 0.5)$) over the principals at the beginning of the simulation.
 
 We looked at each behavioral assumption and asked, fixing that behavioral assumption, is there any case where playing a better learning algorithm helps the principal get a higher market share?
+In all of the simulations below, $K=10$ and $T=5000$.
+8 arms have distributions $Beta(0.45, 0.55)$, one arm has a distribution $Beta(0.55, 0.45)$, and the last arm has a distribution of $Beta(0.65, 0.35)$.
 
 \begin{center}
-Agent: HardMax, T = 5000
+Agent: HardMax
  \begin{tabular}{||c c c c c c||} 
@@
  \hline
- StaticGreedy & Dynamic $\epsilon$-greedy & 0.510 & 0.490 & 0.2 & 0.098\\
+ StaticGreedy & Dynamic $\epsilon$-greedy & 0.510 & 0.490 & 0.200 & 0.098\\
  \hline
@@
   \hline
-    ThompsonSampling & StaticGreedy & 0.335 & 0.664 & 0.138 & 0.100 \\
+    ThompsonSampling & StaticGreedy &  0.517 & 0.482 & 0.101 & 0.200 \\
   \hline
@@
   \hline
-  UCB1 & StaticGreedy & 0.640& 0.360 & 0.119 & 0.2 \\
+  UCB1 & StaticGreedy & 0.640& 0.360 & 0.119 & 0.20 \\
   \hline
@@
 
-\section*{Experiment 3 - Tuning SoftMax} % might want to put this one at the top
-This experiment mainly looked at what happens if we do a grid search over alpha parameters in tuning SoftMax. % you have to tune both alpha and epsilon for SoftMax
+\begin{center}
+Agent: HardMaxWithRandom, $\epsilon=0.1$
+ \begin{tabular}{||c c c c c c||} 
+ \hline
+ Principal 1 Alg & Principal 2 Alg & Principal 1 MS & Principal 2 MS & P1 Avg Regret & P2 Avg Regret  \\ [0.5ex] 
+ \hline\hline
+ StaticGreedy & ExploreThenExploit & 0.168 & 0.832 & 0.200 & 0.040 \\ 
+ \hline
+ StaticGreedy & Dynamic $\epsilon$-greedy & 0.154 & 0.846 & 0.200 & 0.045\\
+ \hline
+ DynamicGreedy & ExploreThenExploit & 0.408 & 0.592 & 0.093 & 0.053 \\ 
+ \hline
+ DynamicGreedy & Dynamic $\epsilon$-greedy & 0.485 & 0.515 & 0.100 & 0.083 \\
+ \hline
+  UCB1 & ExploreThenExploit & 0.248 & 0.752 & 0.130 & 0.042 \\
+  \hline
+  UCB1 & Dynamic $\epsilon$-greedy & 0.265 & 0.735 & 0.132 & 0.045 \\
+  \hline
+  ThompsonSampling & ExploreThenExploit & 0.538 & 0.462 & 0.038 & 0.036 \\
+  \hline
+  ThompsonSampling & Dynamic $\epsilon$-greedy & 0.557 & 0.443 & 0.042 & 0.064 \\
+  \hline
+  ThompsonSampling & StaticGreedy & 0.903 & 0.097 & 0.021 & 0.200 \\
+  \hline
+  ThompsonSampling & DynamicGreedy & 0.603 & 0.397 & 0.035 & 0.089 \\
+  \hline
+  UCB1 & StaticGreedy & 0.828 & 0.172 & 0.088 & 0.200 \\
+  \hline
+  UCB1 & DynamicGreedy & 0.388 & 0.612 & 0.125 & 0.082 \\
+  \hline
+   UCB1 & ThompsonSampling & 0.242 & 0.758 & 0.132 & 0.039 \\
+  \hline
+\end{tabular}
+\end{center}
 
-[insert plots]
+\begin{center}
+Agent: SoftMax, $\alpha=10$, $\epsilon=0.1$
+ \begin{tabular}{||c c c c c c||} 
+ \hline
+ Principal 1 Alg & Principal 2 Alg & Principal 1 MS & Principal 2 MS & P1 Avg Regret & P2 Avg Regret  \\ [0.5ex] 
+ \hline\hline
+ StaticGreedy & ExploreThenExploit & 0.247 & 0.753 & 0.200 & 0.027 \\ 
+ \hline
+ StaticGreedy & Dynamic $\epsilon$-greedy & 0.250 & 0.750 & 0.200 & 0.032 \\
+ \hline
+ DynamicGreedy & ExploreThenExploit & 0.485 & 0.515 & 0.055 & 0.043 \\ 
+ \hline
+ DynamicGreedy & Dynamic $\epsilon$-greedy & 0.422 & 0.578 & 0.110 & 0.061 \\
+ \hline
+  UCB1 & ExploreThenExploit & 0.344 & 0.656 & 0.122 & 0.029 \\
+  \hline
+  UCB1 & Dynamic $\epsilon$-greedy & 0.392 & 0.608 & 0.116 & 0.053 \\
+  \hline
+  ThompsonSampling & ExploreThenExploit & 0.521 & 0.479 & 0.034 & 0.049 \\
+  \hline
+  ThompsonSampling & Dynamic $\epsilon$-greedy & 0.496 & 0.504 & 0.041 & 0.047 \\
+  \hline
+  ThompsonSampling & StaticGreedy & 0.752 & 0.248 & 0.026 & 0.200 \\
+  \hline
+  ThompsonSampling & DynamicGreedy & 0.585 & 0.415 & 0.030 & 0.093 \\
+  \hline
+  UCB1 & StaticGreedy & 0.658 & 0.342 & 0.010 & 0.200 \\
+  \hline
+  UCB1 & DynamicGreedy & 0.442 & 0.558 & 0.114 & 0.079 \\
+  \hline
+  UCB1 & ThompsonSampling & 0.345 & 0.655 & 0.120 & 0.027 \\
+  \hline
+\end{tabular}
+\end{center}
+
+We note several learnings from this experiment:
+\begin{enumerate}
+\item Regardless of the behavioral assumption, Thompson Sampling seems to get a larger market share regardless of what the competing algorithm is.
+\item In general, non-adaptive exploration algorithms seem to ``beat'' greedy algorithms.
+\item UCB does surprisingly poorly against every algorithm except for StaticGreedy
+\item There does not appear to be a large advantage to using an adaptive exploration algorithm compared to a non-adaptive exploration algorithm (as previously noted, Thompson Sampling does slightly better but UCB does worse).
+\item The results do not seem to qualitatively vary across different behavioral assumptions.
+\end{enumerate}
+
+\section*{Experiment 3 - Tuning SoftMax}
+We implemented SoftMax using a logistic function.
+There are two parameters that we need to tune: $alpha$, which controls the steepness of the curve, and $\epsilon$, which controls the minimum and maximum values that the logistic function can take.
+A typical logistic function takes the form
+
+\[f(x)=\frac{1}{1+e^{-\alpha(x)}}\]
+
+In our case, $x$ corresponds in the difference between principal 1's score and principal 2's score (we calculated score as the mean of the past 50 rewards each principal attained).
+The range of this function lies in $(0,1)$, but we want to compress the range to $(\epsilon, 1-\epsilon)$.
+That way, when we draw uniformly a $z$ from the range $[0,1]$, we can compare it to $f(x)$ and pick principal 1 if $z<f(x)$, and pick principal 2 otherwise.
+So, we instead use the following $f(x)$, which achieves the desired logistic function:
+
+\[f(x)=\epsilon + \frac{1-2\epsilon}{1+e^{-\alpha(x)}}\]
+
+To tune $\alpha$, we construct the following situation: $K=10$, $T=5000$.
+8 arms have distributions $Beta(0.45, 0.55)$, 1 arm has distribution $Beta(0.55, 0.45)$, and 1 arm has distribution $Beta(0.8, 0.2)$.
+Both principals start with prior distributions over all 10 arms of $Beta(0.5, 0.5)$, and agent priors over the two principals start as $Beta(0.5, 0.5)$, so the agents do not initially favor either principal over the other.
+Then, we run $[StaticGreedy, DynamicGreedy]$ pairwise against $[UCB, ThompsonSampling]$, for $\alpha \in (1, 5, 10, 15)$, and $\epsilon=0.1$.
+A reasonable value for $\alpha$ would be one in which the adaptive algorithms take a reasonable chunk of the market.
+The results are below:
+
+\begin{center}
+
+ \begin{tabular}{||c c c c c||} 
+ \hline
+ Alpha & Principal 1 Alg & Principal 2 Alg & Principal 1 MS & Principal 2 MS \\ [0.5ex] 
+ \hline\hline
+ 1 & StaticGreedy & ThompsonSampling & 0.435 & 0.565 \\ 
+ \hline
+ 1 & StaticGreedy & UCB1 & 0.451 & 0.549 \\ 
+ \hline
+ 1 & DynamicGreedy & ThompsonSampling & 0.472 & 0.528 \\ 
+ \hline
+ 1 & DynamicGreedy & UCB1 & 0.495 & 0.505 \\ 
+ \hline
+ 5 & StaticGreedy & ThompsonSampling & 0.234 & 0.766 \\ 
+ \hline
+ 5 & StaticGreedy & UCB1 & 0.279 & 0.721 \\ 
+ \hline
+ 5 & DynamicGreedy & ThompsonSampling & 0.443 & 0.557 \\ 
+ \hline
+ 5 & DynamicGreedy & UCB1 & 0.548 & 0.452 \\ 
+ \hline
+ 10 & StaticGreedy & ThompsonSampling & 0.145 & 0.855 \\ 
+ \hline
+ 10 & StaticGreedy & UCB1 & 0.180 & 0.820 \\ 
+ \hline
+ 10 & DynamicGreedy & ThompsonSampling & 0.392 & 0.608 \\ 
+ \hline
+ 10 & DynamicGreedy & UCB1 & 0.472 & 0.528 \\ 
+ \hline
+ 15 & StaticGreedy & ThompsonSampling & 0.115 & 0.885 \\ 
+ \hline
+ 15 & StaticGreedy & UCB1 & 0.145 & 0.855 \\ 
+ \hline
+ 15 & DynamicGreedy & ThompsonSampling & 0.310 & 0.690 \\ 
+ \hline
+ 15 & DynamicGreedy & UCB1 & 0.603 & 0.397 \\ 
+ \hline
+\end{tabular}
+\end{center}
+
+In the $\alpha=1$ case, the market is split roughly evenly.
+This makes sense because if we graph the logistic function, it is very flat, especially in the $[-1,1]$ domain that $x$ lies in.
+For $\alpha$ values 10 and 15, the adaptive algorithms reliably beat the greedy algorithms, but there wasn't a big difference between 10 and 15, so we decided to go with $\alpha=10$.
+
+To tune $\epsilon$, we want to construct a situation where the epsilon baseline allows a ``smarter" algorithm to ``catch up” and beat a ``dumber" algorithm that starts with an advantage.
+We run simulations pitting StaticGreedy against DynamicGreedy, and construct the following circumstance: $K=10$, $T=5000$.
+8 arms are distributed according to $Beta(0.45, 0.55)$, 1 arm is distributed according to $Beta(0.55, 0.45)$, and 1 arm is distributed according to $Beta(0.8, 0.2)$.
+StaticGreedy starts with Beta priors of mean 0.5 over all arms except the arm that has mean 0.55.
+StaticGreedy’s prior on that is $Beta(0.55, 045)$, so StaticGreedy will always choose the arm whose true mean is 0.55 (in other words, it always chooses the second-best arm).
+DynamicGreedy, on the other hand, starts at a disadvantage: it initially chooses one of the worst arms, because it starts with Beta priors of mean 0.5 over all arms except for one arm (whose real mean is 0.45), for which its prior is $Beta(0.7, 0.3)$.
+
+When we run HardMax, StaticGreedy beats DynamicGreedy, taking, on average over our simulations, 81.4\% of the market share.
+But when we switch to SoftMax (with $\alpha=10$), even an epsilon of 0.01 allows DynamicGreedy to win.
+
+\begin{center}
+
+ \begin{tabular}{||c c c c c||} 
+ \hline
+ Epsilon & Principal 1 Alg & Principal 2 Alg & Principal 1 MS & Principal 2 MS \\ [0.5ex] 
+ \hline\hline
+ 0.01 & StaticGreedy & DynamicGreedy & 0.331 & 0.669 \\ 
+ \hline
+ 0.05 & StaticGreedy & DynamicGreedy & 0.349 & 0.651 \\ 
+ \hline
+ 0.10 & StaticGreedy & DynamicGreedy & 0.344 & 0.656 \\ 
+ \hline
+\end{tabular}
+\end{center}
+
+Based on this, we concluded that the particular value of epsilon didn't matter all that much, as long as it is positive so that a ``smart" algorithm has a chance to get back in the game and prove itself to the agents.
+Thus, we considered 0.05 or 0.10 to both be ok.
+
 \section*{Experiment 4 - Prior or Algorithm}
@@
  \hline
- Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & P1 Avg Regret & P2 Avg Regret  \\ [0.5ex] 
+ Mean of Arm A & Principal 1 Alg & Principal 1 MS & Principal 2 MS & Principal 1 Avg Regret & Principal 2 Avg Regret  \\ [0.5ex] 
  \hline\hline
@@
 What is the effect of a ``warm start" on the performance of HardMax?
-Our simulations show that with HardMax the initial rounds matter a lot in a particular simulation and whoever wins in the first few rounds takes the entire market.
+A ``warm start" is defined as giving principals free observations without including it in the information set of the agents.
+Our previous simulations show that with HardMax the initial rounds matter a lot in a particular simulation and whoever wins in the first few rounds takes the entire market.
 We explore if having a warm start makes it so that the market is more evenly split in a given simulation.\\
@@
 
+\section*{Appendix}
+
+Code for the simulations can be found at https://github.com/rawls238/bandits-rl-project
+
 \end{document}

### 2017-12-05 pres/pres.tex
commit pres + writeup

@@
 \item In this experiment, we ran several competing algorithms against each other.
-\item results
+\item Several things we learned from this experiment:
+\begin{enumerate}
+\item Regardless of the behavioral assumption, Thompson Sampling seems to get a larger market share regardless of what the competing algorithm is.
+\item In general, non-adaptive exploration algorithms seem to ``beat'' greedy algorithms.
+\item UCB does surprisingly poorly against every algorithm except for StaticGreedy
+\item There does not appear to be a large advantage to using an adaptive exploration algorithm compared to a non-adaptive exploration algorithm (as previously noted, Thompson Sampling does slightly better but UCB does worse).
+\item The results do not seem to qualitatively vary across different behavioral assumptions.
+\end{enumerate}
 \end{itemize}

### 2018-01-03 pres/pres.tex
final pres + writeup

@@
 \item Reputation score generated by a sliding window average (fix memory sufficiently high except for limited memory experiment)
+\item Unless specified otherwise, reported results found by averaging results from 25 simulations
 \end{itemize}
@@
 \begin{itemize}
-\item SoftMax is great!
+\item The logistic function takes the form \[f(x)=\epsilon + \frac{1-2\epsilon}{1+e^{-\alpha(x)}}\] where  $\epsilon$ controls the baseline probability each principal is chosen, and $\alpha$ controls the steepness of the curve.
+$x$ is the difference in score between the two principals.
+\item We set up a situation where a "smarter" algorithm starts at a disadvantage to a "dumber" algorithm due to different prior beliefs, and tune values of $\epsilon$ and $\alpha$ that result in the "smarter" algorithm winning
+\item We settled on $\epsilon=0.05$ and $\alpha=10$ .
 \end{itemize}
@@
 \item What if we gave the principals free information at the start by giving them $N$ agents? Some simulations showed that there was no difference.
-\item What if we made the rewards of the agents separate from the rewards of the principals?
 \end{itemize}

### 2018-01-03 main.py

@@ -120,7 +120,7 @@ N = 25
 numCores = multiprocessing.cpu_count()
 
 # AGENT_ALGS = [HardMax, SoftMax, HardMaxWithRandom, SoftMaxWithRandom]
-AGENT_ALGS = [SoftMax] #i do SoftMax and HardMaxWithRandom
+AGENT_ALGS = [HardMax] #i do SoftMax and HardMaxWithRandom
 
 #greedy algorithms: [StaticGreedy, DynamicGreedy]
 #non-adaptive exploration: [DynamicEpsilonGreedy, ExploreThenExploit]
@@ -132,8 +132,8 @@ AGENT_ALGS = [SoftMax] #i do SoftMax and HardMaxWithRandom
 # UCB vs ThompsonSampling = 1
 
 # valid principal algs are: [StaticGreedy, UCB, DynamicEpsilonGreedy, DynamicGreedy, ExploreThenExploit, ThompsonSampling]
-PRINCIPAL1_ALGS = [StaticGreedy, UCB, DynamicEpsilonGreedy, DynamicGreedy, ExploreThenExploit, ThompsonSampling]
-PRINCIPAL2_ALGS = [StaticGreedy, UCB, DynamicEpsilonGreedy, DynamicGreedy, ExploreThenExploit, ThompsonSampling]
+PRINCIPAL1_ALGS = [ThompsonSampling]
+PRINCIPAL2_ALGS = [StaticGreedy]
 
 
 results = {}

### 2018-01-17 preliminary/preliminary_report.tex
commit a bunch of plots

\documentclass[11pt,letterpaper]{article}
\usepackage{fullpage}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{amsfonts}

\graphicspath{{../results}}

\begin{document}

\title{Preliminary Experiment}
\maketitle

Considered Bandit Prior/Instances that were considered (only for 10 arms). :
\begin{enumerate}
\item Uniform: Draw the mean rewards for the arms from [0.25, 0.75]
\item ``Heavy-tailed": For ease we simply took the 10 arms to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$)
\item Needle-in-haystack
\begin{enumerate}
\item Low Mean
\begin{enumerate}
\item Low - 9 arms with mean 0.01, 1 arm with mean 0.02 (+ 0.01)
\item Medium - 9 arms with mean 0.01, 1 arm with mean 0.06 (+ 0.05)
\item High - 9 arms with mean 0.01, 1 arm with mean 0.21 (+ 0.20)
\end{enumerate}
\item Medium Mean
\begin{enumerate}
\item Low - 9 arms with mean 0.50, 1 arm with mean 0.01 (+ 0.01)
\item Medium - 9 arms with mean 0.50, 1 arm with mean 0.55 (+ 0.05)
\item High - 9 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
\end{enumerate}
\item High Mean
\begin{enumerate}
\item Low - 9 arms with mean 0.80, 1 arm with mean 0.81 (+ 0.01)
\item Medium - 9 arms with mean 0.80, 1 arm with mean 0.85 (+ 0.05)
\item High - 9 arms with mean 0.80, 1 arm with mean 1.0 (+ 0.20)
\end{enumerate}
\end{enumerate}
\end{enumerate}

Algorithms considered:
\begin{enumerate}
\item ThompsonSampling with priors of $Beta(5, 5)$ for every arm.
\item ExploreThenExploit with 100 exploration steps
\item DynamicGreedy with priors of $Beta(5, 5)$ for every arm
\item Dynamic $\epsilon$-greedy with $\epsilon$=0.05
\item UCB1
\end{enumerate}

The first run considered $N=50$ runs which didn't seem to be able to differentiate between the learning rates of the particular algorithms. However, it was helpful for reducing the space of needle in haystack algorithms. Originally we considered three regimes of ``needle-in-haystack" instances since we weren't sure if learning would be different between the different regimes. We considered so many needle-in-haystack instances to see if there would be meaningful qualitative differences between learning when the means were at extremes vs when most means were in the middle of the range of possible means. \\

The plots for all the algorithms are reported below. The ``low" $\Delta$ in the needle-in-haystack appears uninteresting across all considered means, likely becuase the $\Delta$ of $0.01$ is too small to be meaningful. For the ``medium" instances it appears as though there is a qualitative difference between the ``low" mean and the others. Though this may be partially due to low $N$, it appears as though there is some divergence particularly between UCB and Thompson Sampling. For the high $\Delta$ it the ``low" mean and ``medium" mean seem to have qualitatively similar results. The ``high" mean has starker differences in terms of the learning rates of the algorithms.

First, looking at the medium range: \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle01 - Medium 1"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - Medium 1"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle80 - Medium 1"}
\\
The low range: \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle01 - Low 1"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - Low 1"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle80 - Low 1"}
\\

The high range: \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle01 - High 1"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High 1"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle80 - High 1"}
\\

Surprisingly with $N=50$ there appeared to be considerable variability in the trajectories and so the experiment was reran with $N=250$ and $T=5000$. To reduce the amount of time this took, only Uniform, HeavyTail, Needle50 - Medium, and Needle50 - High were considered. Though there were some qualitative differences between the different needle in haystack instances discussed above, they were omitted here but we can run this again with those if we decide we want some more fine-grain information.

The plots with all the algorithms are reported below:
The high range: \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - Medium 3"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High 1"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail 3"}
\\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Uniform 3"}
\\

A few particular striking comparative plots: \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High UCB"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail UCB"} \\

How does UCB compare against Dynamic $\epsilon$-greedy in these cases? \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High UCB, Dyn eps"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail UCB, Dyn eps"} \\

The rate of learning for UCB seems slower than for the other algorithms. Is this expected (or do we need to tune the UCB1 constant)? It appears that UCB does not catch up to Dynamic $\epsilon$-greedy until $t = 1500$. This potentially explains the results we saw that UCB1 loses in the competing bandits game because it's rate of learning growth for these bandit instances is slower than the other algorithms.

Comparing ThompsonSampling to Dynamic $\epsilon$-greedy: \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High TS, Dyn eps"} \\
\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail TS, Dyn eps"} \\

\end{document}

### 2018-01-17 plots.R

@@ -2,16 +2,26 @@ library("ggplot2")
 library("dplyr")
 library("reshape")
 
-dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_plots_2.csv")
+
+# Three datasets that currently exist are:
+# preliminary_plots.csv includes more distributions but the resulting reward means are only averaged over fifty rounds.
+# preliminary_plots_2.csv was run with N=250, T=1000
+# preliminary_plots_3.csv was run with N=250, T=5000
+
+WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
+
+dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_plots_3.csv", sep=""))
 algs <- as.list(unique(dat['Algorithm']))$Algorithm
+filter_algs <- c("ThompsonSampling", "DynamicEpsilonGreedy")
+dat <- filter(dat, Algorithm %in% filter_algs)
 dists <- as.list(unique(dat['Distribution']))$Distribution
 
 filter_by_dist_and_plot <- function(dist) {
   d <- filter(dat, Distribution == dist)
-  title <- paste("Reward Trajectory for", dist)
+  title <- paste("Reward Trajectory for", dist, "TS, Dyn eps")
   q <- ggplot(data=d) + ggtitle(title) + ylab("Mean Reward") + xlab("time") +
     geom_path(aes(x=t, y=Reward.Mean, colour=Algorithm))
-  ggsave(width=8, height=8, dpi=300, filename=paste("/Users/garidor/Desktop/bandits-rl-project/results/", title, ".pdf", sep=""), plot=q)
+  ggsave(width=8, height=8, dpi=300, filename=paste(WORKING_PATH, "/results/", title, ".pdf", sep=""), plot=q)
   
   return(d)
 }

### 2018-01-17 preliminary.py

@@ -23,7 +23,7 @@ from scipy.stats import bernoulli, beta
 numCores = multiprocessing.cpu_count()
 
 
-T = 1000
+T = 5000
 N = 250
 K = 10
 
@@ -57,7 +57,6 @@ BANDIT_DISTR = {
   'Uniform': uniform_real_distr, 
   'Needle50 - High': needle_in_haystack_50_high, 
   'Needle50 - Medium': needle_in_haystack_50_medium, 
-  'Needle50 - Low': needle_in_haystack_50_low,
   'Heavy Tail' :heavy_tailed
 }
 
@@ -66,7 +65,7 @@ BANDIT_DISTR = {
 
 FIELDNAMES = ['Algorithm', 'K', 'Distribution', 't', 'Reward Mean', 'Reward Std', 'Best Arm Mean']
 
-with open('results/preliminary_plots_2.csv', 'w') as csvfile:
+with open('results/preliminary_plots_3.csv', 'w') as csvfile:
   writer = csv.DictWriter(csvfile, fieldnames=FIELDNAMES)
   writer.writeheader()
   for alg in ALGS:

### 2018-01-18 preliminary/preliminary_report.tex

@@
 
-Considered Bandit Prior/Instances that were considered (only for 10 arms). :
-\begin{enumerate}
+Bandit Prior/Instances that were considered (only for 10 arms):
+\begin{itemize}
 \item Uniform: Draw the mean rewards for the arms from [0.25, 0.75]
-\item ``Heavy-tailed": For ease we simply took the 10 arms to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$)
+\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$).
+With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
 \item Needle-in-haystack
@@
 \end{enumerate}
-\end{enumerate}
+\end{itemize}
 
@@
 
-The first run considered $N=50$ runs which didn't seem to be able to differentiate between the learning rates of the particular algorithms.
-However, it was helpful for reducing the space of needle in haystack algorithms.
-Originally we considered three regimes of ``needle-in-haystack" instances since we weren't sure if learning would be different between the different regimes.
+The first run considered $N=50$ runs.
+The main use of this run was to help us detect any qualitative differences between the different needle-in-haystack regimes we  Originally we considered three regimes of ``needle-in-haystack" instances since we weren't sure if learning would be different between the different regimes.
 We considered so many needle-in-haystack instances to see if there would be meaningful qualitative differences between learning when the means were at extremes vs when most means were in the middle of the range of possible means.
@@
 The plots for all the algorithms are reported below.
-The ``low" $\Delta$ in the needle-in-haystack appears uninteresting across all considered means, likely becuase the $\Delta$ of $0.01$ is too small to be meaningful.
+The ``low" $\Delta$ in the needle-in-haystack appears uninteresting across all considered means, likely because the $\Delta$ of $0.01$ is too small to be meaningful.
 For the ``medium" instances it appears as though there is a qualitative difference between the ``low" mean and the others.
@@
 The ``high" mean has starker differences in terms of the learning rates of the algorithms.
+\\
 
@@
 
-The plots with all the algorithms are reported below:
-The high range: \\
+The plots with all the algorithms are reported below: \\
 \includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - Medium 3"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High 1"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High 3"} \\
 \includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail 3"}
@@
 The rate of learning for UCB seems slower than for the other algorithms.
-Is this expected (or do we need to tune the UCB1 constant)?
-It appears that UCB does not catch up to Dynamic $\epsilon$-greedy until $t = 1500$.
+Is this expected (or do we need to tune the UCB1 constant - I don't think so)?
 This potentially explains the results we saw that UCB1 loses in the competing bandits game because it's rate of learning growth for these bandit instances is slower than the other algorithms.
@@
 
+As well, a version of the plots that smooths the trajectory: \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - Medium 3_smooth"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High 3_smooth"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail 3_smooth"}
+\\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Uniform 3_smooth"}
+\\
+
 \end{document}

### 2018-01-29 preliminary/preliminary_report.tex
add prelim report

@@
 \usepackage{amsfonts}
+\usepackage{algorithm,algpseudocode}
 
@@
 
-Bandit Prior/Instances that were considered (only for 10 arms):
+\section*{Simulation Details}
+
+Considered $K = 3$, $T = 6000$. \\
+\textbf{The Bandit priors that were considered}:
 \begin{itemize}
@@
 \begin{enumerate}
-\item Low Mean
-\begin{enumerate}
-\item Low - 9 arms with mean 0.01, 1 arm with mean 0.02 (+ 0.01)
-\item Medium - 9 arms with mean 0.01, 1 arm with mean 0.06 (+ 0.05)
-\item High - 9 arms with mean 0.01, 1 arm with mean 0.21 (+ 0.20)
-\end{enumerate}
-\item Medium Mean
-\begin{enumerate}
 \item Low - 9 arms with mean 0.50, 1 arm with mean 0.01 (+ 0.01)
@@
 \end{enumerate}
-\item High Mean
+\end{itemize}
+\textbf{Algorithms considered}:
 \begin{enumerate}
-\item Low - 9 arms with mean 0.80, 1 arm with mean 0.81 (+ 0.01)
-\item Medium - 9 arms with mean 0.80, 1 arm with mean 0.85 (+ 0.05)
-\item High - 9 arms with mean 0.80, 1 arm with mean 1.0 (+ 0.20)
+\item ThompsonSampling with priors of $Beta(1, 1)$ for every arm.
+\item DynamicGreedy with priors of $Beta(1, 1)$ for every arm
+\item Bayesian Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm
+\begin{enumerate}
+\item $\epsilon=t^{-1/3}$
+\item $\epsilon=T^{-1/3}$
+\item $\epsilon=0.05$
+\end{enumerate}
+\item Non-Bayesian $\epsilon$-greedy - the greedy decision was made based on empirical mean. When there were zero observations, assumed that the empirical mean was 0 (this seems questionable).
+\begin{enumerate}
+\item $\epsilon=t^{-1/3}$
+\item $\epsilon=T^{-1/3}$
+\item $\epsilon=0.05$
+\end{enumerate}
+\item UCB1
+\begin{enumerate}
+\item UCB1 with constant $\sqrt{2log(t)}$
+\item UCB1 with constant $1$
 \end{enumerate}
 \end{enumerate}
-\end{itemize}
+\pagebreak
+\textbf{Simulation Procedure}
+\begin{algorithm}
+\begin{algorithmic}[1]
+\For{Each prior $p$}
+	\For{Each experiment $i$}
+		\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
+		\State Generate realizations for each arm and round $t$ and instantiate bandit instance
+		\For{Each algorithm $alg$}
+			\State Run simulation for $T$ periods
+		\EndFor
+	\EndFor
+\EndFor
+\end{algorithmic}
+\end{algorithm}
 
-Algorithms considered:
-\begin{enumerate}
-\item ThompsonSampling with priors of $Beta(5, 5)$ for every arm.
-\item ExploreThenExploit with 100 exploration steps
-\item DynamicGreedy with priors of $Beta(5, 5)$ for every arm
-\item Dynamic $\epsilon$-greedy with $\epsilon$=0.05
-\item UCB1
-\end{enumerate}
+One caveat to the reported results is that the Bayesian greedy results were run separately and thus have different true distributions and realizations than the rest of the algorithms.
+If we decide to include the Bayesian version instead of the non-Bayesian version we'll have to re-run these simulations.
 
-The first run considered $N=50$ runs.
-The main use of this run was to help us detect any qualitative differences between the different needle-in-haystack regimes we  Originally we considered three regimes of ``needle-in-haystack" instances since we weren't sure if learning would be different between the different regimes.
-We considered so many needle-in-haystack instances to see if there would be meaningful qualitative differences between learning when the means were at extremes vs when most means were in the middle of the range of possible means.
+\section*{Results}
+
+The confidence bands come from the standard error of the loess regression used to generate the curves.
+The actual datapoints, when plotted, are incredibly volatile but this volatility decreases as we increase N.
+Part of this comes from the fact that what is plotted is the actual realized reward and not the mean reward \textbf{*}.
+We briefly discussed this last time and perhaps we should be reporting this instead (or in addition).
 \\
+\subsection*{Instantaneous Reward Plots}
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack Low inst"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack Medium inst"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack High inst"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Uniform inst"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail inst"} \\
 
-The plots for all the algorithms are reported below.
-The ``low" $\Delta$ in the needle-in-haystack appears uninteresting across all considered means, likely because the $\Delta$ of $0.01$ is too small to be meaningful.
-For the ``medium" instances it appears as though there is a qualitative difference between the ``low" mean and the others.
-Though this may be partially due to low $N$, it appears as though there is some divergence particularly between UCB and Thompson Sampling.
-For the high $\Delta$ it the ``low" mean and ``medium" mean seem to have qualitatively similar results.
-The ``high" mean has starker differences in terms of the learning rates of the algorithms.
-\\
+\subsection*{Cumulative Reward Plots}
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack Low cumulative"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack Medium cumulative"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack High cumulative"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Uniform cumulative"} \\
+\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail cumulative"} \\
 
-First, looking at the medium range: \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle01 - Medium 1"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - Medium 1"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle80 - Medium 1"}
-\\
-The low range: \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle01 - Low 1"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - Low 1"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle80 - Low 1"}
-\\
+\subsection*{Empirical Standard Deviation}
 
-The high range: \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle01 - High 1"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High 1"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle80 - High 1"}
-\\
-
-Surprisingly with $N=50$ there appeared to be considerable variability in the trajectories and so the experiment was reran with $N=250$ and $T=5000$.
-To reduce the amount of time this took, only Uniform, HeavyTail, Needle50 - Medium, and Needle50 - High were considered.
-Though there were some qualitative differences between the different needle in haystack instances discussed above, they were omitted here but we can run this again with those if we decide we want some more fine-grain information.
-
-The plots with all the algorithms are reported below: \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - Medium 3"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High 3"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail 3"}
-\\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Uniform 3"}
-\\
-
-A few particular striking comparative plots: \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High UCB"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail UCB"} \\
-
-How does UCB compare against Dynamic $\epsilon$-greedy in these cases? \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High UCB, Dyn eps"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail UCB, Dyn eps"} \\
-
-The rate of learning for UCB seems slower than for the other algorithms.
-Is this expected (or do we need to tune the UCB1 constant - I don't think so)?
-This potentially explains the results we saw that UCB1 loses in the competing bandits game because it's rate of learning growth for these bandit instances is slower than the other algorithms.
-
-Comparing ThompsonSampling to Dynamic $\epsilon$-greedy: \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High TS, Dyn eps"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail TS, Dyn eps"} \\
-
-As well, a version of the plots that smooths the trajectory: \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - Medium 3_smooth"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle50 - High 3_smooth"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail 3_smooth"}
-\\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Uniform 3_smooth"}
-\\
+Another thing we discussed previously was reporting and plotting the empirical standard deviation of the realized rewards.
+These tend to be quite large.
+Across almost every distribution and algorithm, the empirical standard deviation of the vector of realized rewards was on average ~0.4-0.5.
+This was calculated, for each algorithm and prior, as follows: we have a vector of $N$ realized rewards for each $t$ and we can find the empirical standard deviation of this vector.
+Averaging across each $t$ for each algorithm and prior, all of them have an average empirical standard deviation between 0.4-0.5.
+An alternative statistic of potential interest would be to look at the empirical standard of the mean rewards of the selected arms which would likely be smaller.
+Not sure which makes more sense.
 

### 2018-02-15 experiment_1/experiment_1.tex
yay

\documentclass[11pt,letterpaper]{article}
\usepackage{fullpage}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{algorithm,algpseudocode}

\graphicspath{{../results}}

\begin{document}

\title{Experiment 1}
\maketitle

\section*{Simulation Details}

Considered $K = 3$, $T = 5005$. Report statistics at $t = 1000, 3000, 5000$ \\
\textbf{The Bandit priors that were considered}:
\begin{itemize}
\item Uniform: Draw the mean rewards for the arms from [0.25, 0.75]
\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$). With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
\item Needle-in-haystack
\begin{enumerate}
\item Medium - 9 arms with mean 0.50, 1 arm with mean 0.55 (+ 0.05)
\item High - 9 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
\end{enumerate}
\end{itemize}
\textbf{Algorithms considered}:
\begin{enumerate}
\item ThompsonSampling with priors of $Beta(1, 1)$ for every arm.
\item DynamicGreedy with priors of $Beta(1, 1)$ for every arm
\item Bayesian Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm and $\epsilon=0.05$
\end{enumerate}
\textbf{Agent Algorithms considered}:
\begin{enumerate}
\item HardMax
\item HardMaxWithRandom
\item SoftMax
\end{enumerate}
\textbf{Memory Sizes}
\begin{enumerate}
\item 10
\item 25
\item 100
\end{enumerate}
\pagebreak
\textbf{Simulation Procedure}
\begin{algorithm}
\begin{algorithmic}[1]
\For{Each prior $p$}
	\For{Each agent algorithm $agent_alg$}
		\For{Each principal algorithm pair $principal_alg1$, $principal_alg2$}
			\For{Each simulation $i$}
				\State Generate true distribution from $p$ (except for needle-in-		haystack, just use $p$ itself)
				\State Run simulation for T periods
			\EndFor
		\EndFor
	\EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

\pagebreak
\section*{Results}

One thing which is ambiguous to define is the regret value to use when a principal never gets chosen in a given simulation. When calculating any of the aggregate regret statistics we drop these simulations, but we do record how many rounds have an undefined regret. \\
\vspace{1cm}

First, we'll restrict focus to $t = 5000$ and look at the performance of ThompsonSampling. Note that the y axis here represents the market share that the ThompsonSampling principal gets.
\vspace{0.5cm}
Performance of ThompsonSampling vs DynamicGreedy and DynamicEpsilonGreedy across all agent models, memory sizes, and priors: \\
\includegraphics[scale=0.75]{ts_perf_dg} \\
\includegraphics[scale=0.75]{ts_perf_deg} \\

Now, looking across different memory sizes for each agent model (still using each prior): \\
\textbf{HardMax} \\
\includegraphics[scale=0.75]{hm_ts_deg} \\
\includegraphics[scale=0.75]{hm_ts_dg} \\

\pagebreak
\textbf{HardMaxWithRandom} \\
\includegraphics[scale=0.75]{hmr_ts_deg} \\
\includegraphics[scale=0.75]{hmr_ts_dg} \\

\pagebreak
\textbf{SoftMax} \\
\includegraphics[scale=0.75]{sm_ts_deg} \\
\includegraphics[scale=0.75]{sm_ts_dg} \\

\pagebreak
We'll now look fix memory size to be 100 and look at the performance across priors.\\
\textbf{HardMax} \\
\includegraphics[scale=0.75]{hm_ts_deg_prior} \\
\includegraphics[scale=0.75]{hm_ts_dg_prior} \\

\pagebreak
\textbf{HardMaxWithRandom} \\
\includegraphics[scale=0.75]{hmr_ts_deg_prior} \\
\includegraphics[scale=0.75]{hmr_ts_dg_prior} \\

\pagebreak
\textbf{SoftMax} \\
\includegraphics[scale=0.75]{sm_ts_deg_prior} \\
\includegraphics[scale=0.75]{sm_ts_dg_prior} \\

\end{document}

### 2018-02-16 experiment_1/experiment_1.tex
prelim experiment
latest

@@
 
-Considered $K = 3$, $T = 5005$. Report statistics at $t = 1000, 3000, 5000$ \\
+Considered $K = 3$, $T = 5005$, $N = 125$. Report statistics at $t = 1000, 3000, 5000$ \\
 \textbf{The Bandit priors that were considered}:
@@
 \For{Each prior $p$}
-	\For{Each agent algorithm $agent_alg$}
-		\For{Each principal algorithm pair $principal_alg1$, $principal_alg2$}
-			\For{Each simulation $i$}
-				\State Generate true distribution from $p$ (except for needle-in-		haystack, just use $p$ itself)
+	\For{Each agent algorithm $agent alg$}
+		\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
+			\For{$N$ simulations}
+				\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
+				\State Give the agents 5 observations from each principal
 				\State Run simulation for T periods
@@
 \\
-\vspace{1cm}
+\vspace{0.5cm}
 
-First, we'll restrict focus to $t = 5000$ and look at the performance of ThompsonSampling. Note that the y axis here represents the market share that the ThompsonSampling principal gets.
+First, we'll restrict focus to $t = 5000$ and look at the performance of ThompsonSampling. Note that the y axis here represents the market share that the ThompsonSampling principal gets. \\
 \vspace{0.5cm}
+
 Performance of ThompsonSampling vs DynamicGreedy and DynamicEpsilonGreedy across all agent models, memory sizes, and priors: \\
@@
 \textbf{HardMax} \\
-\includegraphics[scale=0.75]{hm_ts_deg} \\
-\includegraphics[scale=0.75]{hm_ts_dg} \\
+\includegraphics[scale=0.9]{hm_ts_deg} \\
+\includegraphics[scale=0.9]{hm_ts_dg} \\
+Under HardMax it's clear that there is high variability in market share that is increasing in memory size and indicative of less competition but the median across memory sizes still has ThompsonSampling winning.
 
@@
 \textbf{HardMaxWithRandom} \\
-\includegraphics[scale=0.75]{hmr_ts_deg} \\
-\includegraphics[scale=0.75]{hmr_ts_dg} \\
+\includegraphics[scale=0.9]{hmr_ts_deg} \\
+\includegraphics[scale=0.9]{hmr_ts_dg} \\
+Under HardMaxWithRandom, the amount of variability decreases (more competition) but is still increasing in memory size and ThompsonSampling still wins in most simulations.
 
@@
 \textbf{SoftMax} \\
-\includegraphics[scale=0.75]{sm_ts_deg} \\
-\includegraphics[scale=0.75]{sm_ts_dg} \\
+\includegraphics[scale=0.9]{sm_ts_deg} \\
+\includegraphics[scale=0.9]{sm_ts_dg} \\
+Under SoftMax there is less variability (more competition) that does not change much with memory size and here ThompsonSampling wins more in the median simulation but only a bit.
 
@@
 \textbf{HardMax} \\
-\includegraphics[scale=0.75]{hm_ts_deg_prior} \\
-\includegraphics[scale=0.75]{hm_ts_dg_prior} \\
+\includegraphics[scale=0.9]{hm_ts_deg_prior} \\
+\includegraphics[scale=0.9]{hm_ts_dg_prior} \\
+With HardMax there is significant variability (as noted before), but we have that in the needle in haystack medium case the median is approximately 0.5 while in Uniform + Needle in Haystack High the median is 1 and heavy tail it is almost 0.
+I don't know if this is something interesting or just noise due to the variability in HardMax.
+Will have to look into this more.
 
@@
 \textbf{HardMaxWithRandom} \\
-\includegraphics[scale=0.75]{hmr_ts_deg_prior} \\
-\includegraphics[scale=0.75]{hmr_ts_dg_prior} \\
+\includegraphics[scale=0.9]{hmr_ts_deg_prior} \\
+\includegraphics[scale=0.9]{hmr_ts_dg_prior} \\
+With TS vs DEG we see roughly the same performance across priors, but for TS vs DG we see as with HardMax that in the needle in haystack high instance we have that TS wins by a lot but not so much in the heavy tail case.
+This makes sense as TS (and DEG) should generally find the ``good" arm while DG may sometimes wrongly identify the best arm, leading to this disparity.
 
@@
 \textbf{SoftMax} \\
-\includegraphics[scale=0.75]{sm_ts_deg_prior} \\
-\includegraphics[scale=0.75]{sm_ts_dg_prior} \\
+\includegraphics[scale=0.9]{sm_ts_deg_prior} \\
+\includegraphics[scale=0.9]{sm_ts_dg_prior} \\
+Nothing horribly interesting here besides noting that the same TS vs DG difference appears in the Needle in haystack high prior.
+
+\pagebreak
+Now, a couple of plots giving a sense of regret values for ThompsonSampling across different agent algorithms and memory values.\\
+\includegraphics[scale=1]{ts_regret_agent_algs} \\
+\includegraphics[scale=1]{ts_regret_memory}
 

### 2018-02-20 experiment_2/experiment_2.tex
add experiment 2

--- 2018-02-15 experiment_1/experiment_1.tex
@@
 
-\title{Experiment 1}
+\title{Experiment 2}
 \maketitle
 
+\section*{Differences from Experiment 1}
+
+The only difference between this experiment and experiment 1 is that we give principal 2 a ``headstart" in the sense of letting them be the only firm in the market for the first 1000 periods.
+Thus, the principals are effectively given 1000 free observations for learning and the agents are given 1000 free observations in order to form a reputation score for principal 2.
+The question this simulation then attempts to answer is if \textit{information advantage} can be viewed as a barrier to entry.
+Namely, it attempts to begin to answer if the first mover advantage manifests itself in this model by giving the incumbent firm a better prior than any entrants (who we assume come in with only ``fake" priors).
+Under what assumptions about the agent rationality (response functions or memory sizes) and what algorithms do we see the entering firm able to gain any market share?
 \section*{Simulation Details}
 
-Considered $K = 3$, $T = 5005$. Report statistics at $t = 1000, 3000, 5000$ \\
+Considered $K = 3$, $T = 5005$, $N = 60$. Report statistics at $t = 1000, 3000, 5000$ \\
 \textbf{The Bandit priors that were considered}:
@@
 \end{enumerate}
-\pagebreak
+
+The simulation procedure is the only thing that has changed compared to Experiment 1:
 \textbf{Simulation Procedure}
@@
 \For{Each prior $p$}
-	\For{Each agent algorithm $agent_alg$}
-		\For{Each principal algorithm pair $principal_alg1$, $principal_alg2$}
-			\For{Each simulation $i$}
-				\State Generate true distribution from $p$ (except for needle-in-		haystack, just use $p$ itself)
+	\For{Each agent algorithm $agent alg$}
+		\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
+			\For{$N$ simulations}
+				\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
+				\State Give the agents 5 observations from each principal
+				\State Give principal 2 1000 free observations (and give the agents this information too)
 				\State Run simulation for T periods
@@
 \section*{Results}
-
-One thing which is ambiguous to define is the regret value to use when a principal never gets chosen in a given simulation.
-When calculating any of the aggregate regret statistics we drop these simulations, but we do record how many rounds have an undefined regret.
-\\
-\vspace{1cm}
-
-First, we'll restrict focus to $t = 5000$ and look at the performance of ThompsonSampling. Note that the y axis here represents the market share that the ThompsonSampling principal gets.
-\vspace{0.5cm}
-Performance of ThompsonSampling vs DynamicGreedy and DynamicEpsilonGreedy across all agent models, memory sizes, and priors: \\
-\includegraphics[scale=0.75]{ts_perf_dg} \\
-\includegraphics[scale=0.75]{ts_perf_deg} \\
-
-Now, looking across different memory sizes for each agent model (still using each prior): \\
 \textbf{HardMax} \\
-\includegraphics[scale=0.75]{hm_ts_deg} \\
-\includegraphics[scale=0.75]{hm_ts_dg} \\
+\includegraphics[scale=0.9]{hm_ts_dg_memory_5000} \\ \includegraphics[scale=0.9]{hm_ts_deg_memory_5000} \\
+The main interesting result here is that for ThompsonSampling vs DEG, with low memory sizes we have that ThompsonSampling almost catches up to DEG, but as we increase memory size TS gets almost no market share!
+Should re-run this experiment with higher N to validate this.
 
@@
 \textbf{HardMaxWithRandom}  \\
-\includegraphics[scale=0.75]{hmr_ts_deg} \\
-\includegraphics[scale=0.75]{hmr_ts_dg} \\
+\includegraphics[scale=0.9]{hmr_ts_dg_memory_5000} \\ \includegraphics[scale=0.9]{hmr_ts_deg_memory_5000}
 
@@
 \textbf{SoftMax} \\
-\includegraphics[scale=0.75]{sm_ts_deg} \\
-\includegraphics[scale=0.75]{sm_ts_dg} \\
+\includegraphics[scale=0.9]{sm_ts_dg_memory_5000} \\ \includegraphics[scale=0.9]{sm_ts_deg_memory_5000}
 
 \pagebreak
-We'll now look fix memory size to be 100 and look at the performance across priors.\\
-\textbf{HardMax} \\
-\includegraphics[scale=0.75]{hm_ts_deg_prior} \\
-\includegraphics[scale=0.75]{hm_ts_dg_prior} \\
+Zooming in on the HardMax results and looking at them over time we get (for memory = 25): \\
+\includegraphics[scale=0.9]{hm_ts_deg_25_timehorizon} \\ \includegraphics[scale=0.9]{hm_ts_dg_25_timehorizon}
 
 \pagebreak
-\textbf{HardMaxWithRandom} \\
-\includegraphics[scale=0.75]{hmr_ts_deg_prior} \\
-\includegraphics[scale=0.75]{hmr_ts_dg_prior} \\
-
-\pagebreak
-\textbf{SoftMax} \\
-\includegraphics[scale=0.75]{sm_ts_deg_prior} \\
-\includegraphics[scale=0.75]{sm_ts_dg_prior} \\
+\textbf{Memory = 100} \\
+Take note of the scale on the y-axis on this graph. \\
+\includegraphics[scale=0.9]{hm_ts_deg_100_timehorizon} \\ \includegraphics[scale=0.9]{hm_ts_dg_100_timehorizon}
 

### 2018-02-28 analysis_code/agents_table.Rmd
reorganization of results + some hacky table generation

```{r echo=FALSE,message=FALSE}

library(dplyr)
library(knitr)
WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"

dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_raw_results_4.csv", sep=""))
p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
priors <- as.list(unique(dat['Prior']))$Prior
EXPERIMENT_1_CV_HIGH <- 156.71410383 #chisquared critical values for 124 degrees of freedom
EXPERIMENT_1_CV_LOW <- 95.07008897

EXPERIMENT_2_CV_HIGH <- 128.42198864
EXPERIMENT_2_CV_LOW <- 73.36108019

ZERO_ONE_CUTOFF <- 0.1
```

```{r echo=FALSE, xtable, results="asis"}

concise_rep <- function(alg) {
  if (alg == "ThompsonSampling") {
    return("TS")
  } else if (alg == "DynamicEpsilonGreedy") {
    return("DEG")
  } else if (alg == "DynamicGreedy") {
    return("DG")
  }
}

algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
options(xtable.sanitize.text.function=identity)
options(xtable.caption.placement="top")

alg_pairs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "ThompsonSampling", "DynamicGreedy", "DynamicGreedy", "DynamicEpsilonGreedy")
print("Reported means are for principal 1 which is the first algorithm listed in the pair for each column<br>")
print("In each cell, first line is mean, second line is variance, third line is % of the time that market share > 0.1 or market share < 0.9")
library(xtable)
for (prior in priors) {
  for (j in 1:length(time_horizons)) {
    results <- matrix(nrow=length(agent_algs), ncol=length(algs))
    colnames(results) <- c(paste(concise_rep(alg_pairs[1]), concise_rep(alg_pairs[2]), sep=" vs "), paste(concise_rep(alg_pairs[3]), concise_rep(alg_pairs[4]), sep=" vs "), paste(concise_rep(alg_pairs[5]), concise_rep(alg_pairs[6]), sep=" vs "))
    rownames(results) <- agent_algs
    for (i in 1:length(agent_algs)) {
      agent_alg <- agent_algs[i]
      time <- time_horizons[j]
      for (k in 1:length(alg_pairs)) {
        if (k %% 2 == 0) { next }
          p1alg = alg_pairs[k]
          p2alg = alg_pairs[k+1]
          filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
          if (nrow(filtered_dat) == 0) {
            next
          }
          share <- filtered_dat$Market.Share.for.P1
          test <- t.test(share)
          cin <- signif(test$conf.int[2] - mean(share), digits=2)
          shares <- sum(share >= (1 - ZERO_ONE_CUTOFF) | share <= ZERO_ONE_CUTOFF, na.rm=TRUE) / nrow(filtered_dat)
          shares_percent <- signif(shares * 100, digits=2)
          lower_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_HIGH, digits=2)
          high_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_LOW, digits=2)
          results[i, (k+1)/2] <- paste("<b>",signif(mean(share), digits=2), "</b> +/-", cin, "<br>", signif(var(share), digits=2), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", shares_percent, "%")
        }
      }
      tab <-xtable(results, caption=paste("Results for", time, prior))
      print(tab, type="html")
  }
  print('<br><br><br><br><br><br>')
  }

```

### 2018-02-28 analysis_code/tournament_table.Rmd
reorganization of results + some hacky table generation

```{r echo=FALSE,message=FALSE}

library(dplyr)
library(knitr)
WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"

dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/memory_experiment_raw_results_2.csv", sep=""))
p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
priors <- as.list(unique(dat['Prior']))$Prior
EXPERIMENT_1_CV_HIGH <- 156.71410383 #chisquared critical values for 124 degrees of freedom
EXPERIMENT_1_CV_LOW <- 95.07008897

EXPERIMENT_2_CV_HIGH <- 128.42198864
EXPERIMENT_2_CV_LOW <- 73.36108019

ZERO_ONE_CUTOFF <- 0.1
```

```{r echo=FALSE, xtable, results="asis"}
algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
options(xtable.sanitize.text.function=identity)
options(xtable.caption.placement="top")
library(xtable)
for (prior in priors) {
  for (i in 1:length(agent_algs)) {
    agent_alg <- agent_algs[i]
    for (j in 1:length(time_horizons)) {
      time <- time_horizons[j]
      results <- matrix(nrow=length(algs), ncol=length(algs))
      colnames(results) <- algs
      rownames(results) <- algs
      for (k in 1:length(algs)) {
        for (l in 1:length(algs)) {
          p1alg = algs[k]
          p2alg = algs[l]
          if (p1alg == p2alg) { next }
          filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
          if (nrow(filtered_dat) == 0) {
            next
          }
          share <- filtered_dat$Market.Share.for.P1
          test <- t.test(share)
          cin <- signif(test$conf.int[2] - mean(share), digits=4)
          shares <- sum(share >= (1 - ZERO_ONE_CUTOFF) | share <= ZERO_ONE_CUTOFF, na.rm=TRUE) / nrow(filtered_dat)
          lower_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_HIGH, digits=4)
          high_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_LOW, digits=4)
          results[k, l] <- paste("<b>",signif(mean(share), digits=4), "</b> +/-", cin, "<br>", signif(var(share), digits=4), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", shares)
        }
      }
      tab <-xtable(results, caption=paste("Results for", agent_alg, time, prior))
      print(tab, type="html")
    }
    print('<br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br>')
  }
}

```

### 2018-02-28 analysis_code/bands.R

@@ -1,51 +1,0 @@
-
-#mean and its conf range,
-#sample variance and its conf range
-#fraction of “degenerate outcomes” (0-1 split).
-
-library(dplyr)
-WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
-
-
-dat <- read.csv(file=paste(WORKING_PATH, "/results/memory_experiment_raw_results_2.csv", sep=""))
-p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
-p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
-agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
-time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-priors <- as.list(unique(dat['Prior']))$Prior
-EXPERIMENT_1_CV_HIGH <- 156.71410383 #chisquared critical values for 124 degrees of freedom
-EXPERIMENT_1_CV_LOW <- 95.07008897
-
-EXPERIMENT_2_CV_HIGH <- 128.42198864
-EXPERIMENT_2_CV_LOW <- 73.36108019
-
-
-
-ZERO_ONE_CUTOFF <- 0.1
-for (prior in priors) {
-  for (p1alg in p1_algs) {
-    for (p2alg in p2_algs) {
-      if (p1alg == p2alg) { next }
-      results <- matrix(nrow=length(agent_algs), ncol=length(time_horizons))
-      colnames(results) <- c("Time 1000", "Time 3000", "Time 5000")
-      rownames(results) <- agent_algs
-      for (i in 1:length(agent_algs)) {
-        agent_alg <- agent_algs[i]
-        for (j in 1:length(time_horizons)) {
-          time <- time_horizons[j]
-          filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
-          share <- filtered_dat$Market.Share.for.P1
-          test <- t.test(share)
-          cin <- signif(test$conf.int[2] - mean(share), digits=4)
-          shares <- sum(share >= (1 - ZERO_ONE_CUTOFF) | share <= ZERO_ONE_CUTOFF, na.rm=TRUE) / nrow(filtered_dat)
-          lower_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_HIGH, digits=6)
[…]

### 2018-02-28 analysis_code/kevin_plots.R

@@ -1,7 +1,7 @@
 library(data.table)
 library(ggplot2)
 
-#dt.data <- data.table(read.csv("~/Desktop/bandits-rl-project/results/preliminary_plots.csv", header=T, stringsAsFactors=T))
+#dt.data <- data.table(read.csv("~/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots.csv", header=T, stringsAsFactors=T))
 
 
 # dt.data[(Algorithm == 'ThompsonSampling' & K==10)]
@@ -22,8 +22,8 @@ print(q)
 
 WORKING_PATH <- '/Users/garidor/Desktop/bandits-rl-project'
 
-dat <- data.table(read.csv(file=paste(WORKING_PATH, "/results/preliminary_plots_3_arms.csv", sep="")))
-dat2 <- data.table(read.csv(file=paste(WORKING_PATH, "/results/preliminary_plots_3_arms_2.csv", sep="")))
+dat <- data.table(read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms.csv", sep="")))
+dat2 <- data.table(read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_2.csv", sep="")))
 dt.data <- rbind(dat, dat2)
 rm(dat)
 rm(dat2)
@@ -61,7 +61,7 @@ ggplot(dt.data[(Distribution=="Needle In Haystack Medium")], aes(x=t, y=Instanta
 
 # free_obs_experiment_raw_results.csv
 # memory_experiment_raw_results_2.csv
-dt.data <- data.table(read.csv("~/Desktop/bandits-rl-project/results/free_obs_experiment_raw_results_2.csv", header=T, stringsAsFactors=T))
+dt.data <- data.table(read.csv("~/Desktop/bandits-rl-project/free_obs_raw_results/results/free_obs_experiment_raw_results_4.csv", header=T, stringsAsFactors=T))
 
 unique(dt.data$Memory.Size)
 

### 2018-02-28 analysis_code/non_md_tables.R

@@ -1,0 +1,55 @@
+
+library(dplyr)
+library(knitr)
+WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
+
+
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_raw_results_2.csv", sep=""))
+p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
+p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
+agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
+time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
+priors <- as.list(unique(dat['Prior']))$Prior
+EXPERIMENT_1_CV_HIGH <- 156.71410383 #chisquared critical values for 124 degrees of freedom
+EXPERIMENT_1_CV_LOW <- 95.07008897
+
+EXPERIMENT_2_CV_HIGH <- 128.42198864
+EXPERIMENT_2_CV_LOW <- 73.36108019
+
+
+ZERO_ONE_CUTOFF <- 0.1
+
+alg_pairs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "ThompsonSampling", "DynamicGreedy", "DynamicGreedy", "DynamicEpsilonGreedy")
+library(xtable)
+for (prior in priors) {
+  for (j in 1:length(time_horizons)) {
+    results <- matrix(nrow=3, ncol=3)
+    #colnames(results) <- c(paste(alg_pairs[1], alg_pairs[2], sep=", "), paste(alg_pairs[3], alg_pairs[4], sep=", "), paste(alg_pairs[5], alg_pairs[6], sep=", "))    
+    #rownames(results) <- agent_algs
+    print(agent_algs)
+    for (i in 1:length(agent_algs)) {
+      agent_alg <- agent_algs[i]
+      time <- time_horizons[j]
+      for (k in 1:length(alg_pairs)) {
+        if (k %% 2 == 0) { next }
+        p1alg = alg_pairs[k]
+        p2alg = alg_pairs[k+1]
+        filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+        if (nrow(filtered_dat) == 0) {
+          next
+        }
+        print((k+1)/2)
[…]

### 2018-02-28 analysis_code/preliminary_plots.R

@@ -8,8 +8,8 @@ library("reshape")
 
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_plots_3_arms.csv", sep=""))
-dat2 <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_plots_3_arms_2.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms.csv", sep=""))
+dat2 <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_2.csv", sep=""))
 dat <- rbind(dat, dat2)
 algs <- as.list(unique(dat['Algorithm']))$Algorithm
 print(algs)
@@ -24,7 +24,7 @@ filter_by_dist_and_plot <- function(dist) {
     #geom_path() # just plot the raw trajectory
     geom_smooth(method="loess") #+ smooths the trajectory 
     #geom_errorbar(aes(ymin=Instantaneous.Reward.Mean-1.96*Instantaneous.Reward.Std, ymax=Instantaneous.Reward.Mean+1.96*Instantaneous.Reward.Std))
-  ggsave(width=8, height=8, dpi=300, filename=paste(WORKING_PATH, "/results/", title, ".pdf", sep=""), plot=q)
+  ggsave(width=8, height=8, dpi=300, filename=paste(WORKING_PATH, "/results/preliminary_figures/", title, ".pdf", sep=""), plot=q)
   
   return(d)
 }

### 2018-02-28 analysis_code/agents_table.Rmd
tweaks

@@
+ <style>
+   th,td{
+     padding:10px;
+   }
+ </style>
 ```{r echo=FALSE,message=FALSE}
@@
 
-concise_rep <- function(alg) {
+concise_alg_rep <- function(alg) {
   if (alg == "ThompsonSampling") {
@@
   }
+}
+
+concise_agent_rep <- function(agent_alg) {
+  if (agent_alg == "HardMax") {
+    return("HM")
+  } else if (agent_alg == "HardMaxWithRandom") {
+    return("HMR")
+  } else if (agent_alg == "SoftMax") {
+    return("SM")
+  }
+}
+
+get_agent_rep <- function(repre) {
+  res <- c()
+  for(a in repre) {
+    res <- c(res, concise_agent_rep(a))
+  }
+  return(res)
 }
@@
     results <- matrix(nrow=length(agent_algs), ncol=length(algs))
-    colnames(results) <- c(paste(concise_rep(alg_pairs[1]), concise_rep(alg_pairs[2]), sep=" vs "), paste(concise_rep(alg_pairs[3]), concise_rep(alg_pairs[4]), sep=" vs "), paste(concise_rep(alg_pairs[5]), concise_rep(alg_pairs[6]), sep=" vs "))
-    rownames(results) <- agent_algs
+    colnames(results) <- c(paste(concise_alg_rep(alg_pairs[1]), concise_alg_rep(alg_pairs[2]), sep=" vs "), paste(concise_alg_rep(alg_pairs[3]), concise_alg_rep(alg_pairs[4]), sep=" vs "), paste(concise_alg_rep(alg_pairs[5]), concise_alg_rep(alg_pairs[6]), sep=" vs "))
+    rownames(results) <- get_agent_rep(agent_algs)
     for (i in 1:length(agent_algs)) {

### 2018-02-28 main.py

@@ -59,7 +59,6 @@ BANDIT_DISTR = {
 AGGREGATE_FIELD_NAMES = ['P1 Number of NaNs', 'P2 Number of NaNs', 'Prior', 'P1 Alg', 'P2 Alg', 'Time Horizon', 'Agent Alg', 'Market Share for P1', 'P1 Regret Mean', 'P1 Regret Std', 'P2 Regret Mean', 'P2 Regret Std', 'Abs Average Delta Regret']
 INDIVIDUAL_FIELD_NAMES =['Prior', 'P1 Alg', 'P2 Alg', 'Time Horizon', 'Agent Alg', 'Market Share for P1', 'P1 Regret', 'P2 Regret', 'Abs Delta Regret']
 
-
 def run_finite_memory_experiment(memory_sizes):
   results = {}
   with open('results/free_obs_experiment_aggregate_results_3.csv', 'w') as aggregate_csv:

### 2018-03-03 analysis_code/tournament_table.Rmd
updates

@@
-```{r echo=FALSE,message=FALSE}
+```{r echo=FALSE,message=FALSE,xtable, results="asis"}
 
@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/memory_experiment_raw_results_2.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_raw_results.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 ZERO_ONE_CUTOFF <- 0.1
-```
-
-```{r echo=FALSE, xtable, results="asis"}
 algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
@@
           p2alg = algs[l]
-          if (p1alg == p2alg) { next }
           filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)

### 2018-03-03 analysis_code/kevin_plots.R

@@ -61,13 +61,13 @@ ggplot(dt.data[(Distribution=="Needle In Haystack Medium")], aes(x=t, y=Instanta
 
 # free_obs_experiment_raw_results.csv
 # memory_experiment_raw_results_2.csv
-dt.data <- data.table(read.csv("~/Desktop/bandits-rl-project/free_obs_raw_results/results/free_obs_experiment_raw_results_4.csv", header=T, stringsAsFactors=T))
+dt.data <- data.table(read.csv("/Users/garidor/Desktop/bandits-rl-project/results/free_obs_raw_results/free_obs_experiment_raw_results_6.csv", header=T, stringsAsFactors=T))
 
 unique(dt.data$Memory.Size)
 
 dt.data_subset = dt.data[(Memory.Size==100)]
 
-ggplot(dt.data, aes(x=as.factor(Memory.Size), y=Market.Share.for.P1, color=as.factor(Time.Horizon))) + 
+ggplot(dt.data_subset, aes(x=as.factor(Memory.Size), y=Market.Share.for.P1, color=as.factor(Time.Horizon))) + 
   facet_grid(Prior + P1.Alg ~ Agent.Alg + P2.Alg) + 
   geom_boxplot(alpha=0.5) +
   geom_hline(yintercept=0.5, alpha=0.6) +

### 2018-03-03 fix_data.py

@@ -1,26 +1,0 @@
-import csv
-import numpy as np
-
-AGGREGATE_FIELD_NAMES = ['P1 Number of NaNs', 'P2 Number of NaNs', 'Prior', 'P1 Alg', 'P2 Alg', 'Time Horizon', 'Agent Alg', 'Market Share for P1', 'P1 Regret Mean', 'P1 Regret Std', 'P2 Regret Mean', 'P2 Regret Std', 'Abs Average Delta Regret', 'Memory Size']
-INDIVIDUAL_FIELD_NAMES =['Prior', 'P1 Alg', 'P2 Alg', 'Time Horizon', 'Agent Alg', 'Market Share for P1', 'P1 Regret', 'P2 Regret', 'Abs Delta Regret']
-
-with open('/Users/garidor/Desktop/bandits-rl-project/results/cur_results/memory_experiment_aggregate_results_2.csv', 'w') as write_csv:
-  csvwriter = csv.writer(write_csv)
-  csvwriter.writerow(AGGREGATE_FIELD_NAMES)
-  results = {}
-  with open('/Users/garidor/Desktop/bandits-rl-project/results/cur_results/memory_experiment_raw_results_2.csv', 'rb') as read_csv:
-    csvreader = csv.reader(read_csv)
-    next(csvreader)
-    for row in csvreader:
-      key = (row[1], row[2], row[3], row[4], row[9])
-      if key not in results:
-        results[key] = []
-      results[key].append(float(row[5]))
-  with open('/Users/garidor/Desktop/bandits-rl-project/results/cur_results/memory_experiment_aggregate_results.csv', 'rb') as read_csv_2:
-    csvreader= csv.reader(read_csv_2)
-    next(csvreader)
-    for row in csvreader:
-      key = (row[3], row[4], row[5], row[6], row[13])
-      row[7] = np.mean(results[key])
-      csvwriter.writerow(row)
-

### 2018-03-06 analysis_code/agents_table.Rmd
commit a bunch of results + some tweaks to analysis scripts

@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_raw_results_4.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_raw_results.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
+memory_sizes <- sort(as.list(unique(dat['Memory.Size']))$Memory.Size, decreasing=TRUE)
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
+time_horizons <- c(1000, 5000, 10000, 15000)
 priors <- as.list(unique(dat['Prior']))$Prior
-EXPERIMENT_1_CV_HIGH <- 156.71410383 #chisquared critical values for 124 degrees of freedom
-EXPERIMENT_1_CV_LOW <- 95.07008897
-
-EXPERIMENT_2_CV_HIGH <- 128.42198864
-EXPERIMENT_2_CV_LOW <- 73.36108019
 
@@
 
-alg_pairs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "ThompsonSampling", "DynamicGreedy", "DynamicGreedy", "DynamicEpsilonGreedy")
-print("Reported means are for principal 1 which is the first algorithm listed in the pair for each column<br>")
-print("In each cell, first line is mean, second line is variance, third line is % of the time that market share > 0.1 or market share < 0.9")
+alg_pairs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "ThompsonSampling", "DynamicGreedy", "DynamicGreedy", "DynamicEpsilonGreedy") #"ThompsonSampling", "ThompsonSampling", "DynamicEpsilonGreedy", "DynamicEpsilonGreedy", "DynamicGreedy", "DynamicGreedy")
+#print("Reported means are for principal 1 which is the first algorithm listed in the pair for each column<br>")
+#print("In each cell, first line is mean, second line is variance, third line is % of the time that market share > 0.1 or market share < 0.9")
 library(xtable)
+for (memory in memory_sizes) {
 for (prior in priors) {
     for (j in 1:length(time_horizons)) {
-    results <- matrix(nrow=length(agent_algs), ncol=length(algs))
-    colnames(results) <- c(paste(concise_alg_rep(alg_pairs[1]), concise_alg_rep(alg_pairs[2]), sep=" vs "), paste(concise_alg_rep(alg_pairs[3]), concise_alg_rep(alg_pairs[4]), sep=" vs "), paste(concise_alg_rep(alg_pairs[5]), concise_alg_rep(alg_pairs[6]), sep=" vs "))
+      results <- matrix(nrow=length(agent_algs), ncol=(length(alg_pairs) / 2))
+    colnames(results) <- c(paste(concise_alg_rep(alg_pairs[1]), concise_alg_rep(alg_pairs[2]), sep=" vs "), paste(concise_alg_rep(alg_pairs[3]), concise_alg_rep(alg_pairs[4]), sep=" vs "), paste(concise_alg_rep(alg_pairs[5]), concise_alg_rep(alg_pairs[6]), sep=" vs ")) #paste(concise_alg_rep(alg_pairs[7]), concise_alg_rep(alg_pairs[8]), sep=" vs "), paste(concise_alg_rep(alg_pairs[9]), concise_alg_rep(alg_pairs[10]), sep=" vs "), paste(concise_alg_rep(alg_pairs[11]), concise_alg_rep(alg_pairs[12]), sep=" vs "))
     rownames(results) <- get_agent_rep(agent_algs)
@@
           p2alg = alg_pairs[k+1]
-          filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+          filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == memory & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
           if (nrow(filtered_dat) == 0) {
@@
           test <- t.test(share)
-          cin <- signif(test$conf.int[2] - mean(share), digits=2)
+          cin <- signif(test$conf.int[2] - mean(share), digits=1)
           shares <- sum(share >= (1 - ZERO_ONE_CUTOFF) | share <= ZERO_ONE_CUTOFF, na.rm=TRUE) / nrow(filtered_dat)
           shares_percent <- signif(shares * 100, digits=2)
-          lower_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_HIGH, digits=2)
-          high_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_LOW, digits=2)
-          results[i, (k+1)/2] <- paste("<b>",signif(mean(share), digits=2), "</b> +/-", cin, "<br>", signif(var(share), digits=2), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", shares_percent, "%")
+          cv_high <- qchisq(.975, df=length(share) - 1)
+          cv_low <- qchisq(.025, df=length(share) - 1)
+          lower_var <- signif((length(share) - 1) * var(share) / cv_high, digits=1)
+          high_var <- signif((length(share) - 1) * var(share) / cv_low, digits=1)
+          results[i, (k+1)/2] <- paste("<b>",signif(mean(share), digits=2), "</b> +/-", cin, "<br>Var:", signif(var(share), digits=2), "<br>", shares_percent, "%")
       }
     }
-      tab <-xtable(results, caption=paste("Results for", time, prior))
+      tab <-xtable(results, caption=paste("Results for", "t=",time, prior, "Memory=", memory))
       print(tab, type="html")
   }
-  print('<br><br><br><br><br><br>')
+  }
 }

### 2018-03-06 analysis_code/tournament_table.Rmd
commit a bunch of results + some tweaks to analysis scripts

@@
+ <style>
+   th,td{
+     padding:10px;
+   }
+ </style>
+
 ```{r echo=FALSE,message=FALSE,xtable, results="asis"}
@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_raw_results.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_raw_results.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
 p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
+memory_sizes <- sort(as.list(unique(dat['Memory.Size']))$Memory.Size, decreasing=TRUE)
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
+time_horizons <- c(5000, 10000, 15000)
 priors <- as.list(unique(dat['Prior']))$Prior
-EXPERIMENT_1_CV_HIGH <- 156.71410383 #chisquared critical values for 124 degrees of freedom
-EXPERIMENT_1_CV_LOW <- 95.07008897
-
-EXPERIMENT_2_CV_HIGH <- 128.42198864
-EXPERIMENT_2_CV_LOW <- 73.36108019
 
@@
 library(xtable)
+for (j in 1:length(time_horizons)) {
   for (prior in priors) {
@@
       agent_alg <- agent_algs[i]
-    for (j in 1:length(time_horizons)) {
+      print_breaks <- FALSE
         time <- time_horizons[j]
@@
             if (nrow(filtered_dat) == 0) {
+              if (p1alg == p2alg) { print_breaks <- TRUE }
               next
@@
             test <- t.test(share)
-          cin <- signif(test$conf.int[2] - mean(share), digits=4)
+            cin <- signif(test$conf.int[2] - mean(share), digits=2)
             shares <- sum(share >= (1 - ZERO_ONE_CUTOFF) | share <= ZERO_ONE_CUTOFF, na.rm=TRUE) / nrow(filtered_dat)
-          lower_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_HIGH, digits=4)
-          high_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_LOW, digits=4)
-          results[k, l] <- paste("<b>",signif(mean(share), digits=4), "</b> +/-", cin, "<br>", signif(var(share), digits=4), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", shares)
+            cv_high <- qchisq(.975, df=length(share) - 1)
+            cv_low <- qchisq(.025, df=length(share) - 1)
+            lower_var <- signif((length(share) - 1) * var(share) / cv_high, digits=2)
+            high_var <- signif((length(share) - 1) * var(share) / cv_low, digits=2)
+            results[k, l] <- paste("<b>", signif(mean(share), digits=2), "</b> +/-", cin, "<br>", signif(var(share), digits=2), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", shares)
           }
         }
-      tab <-xtable(results, caption=paste("Results for", agent_alg, time, prior))
+        tab <-xtable(results, caption=paste("Results for", agent_alg, "t =", time, prior))
         print(tab, type="html")
+        if (print_breaks) { print("<br>") }
       }
-    print('<br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br>')
     }

### 2018-03-06 analysis_code/kevin_plots.R

@@ -61,11 +61,11 @@ ggplot(dt.data[(Distribution=="Needle In Haystack Medium")], aes(x=t, y=Instanta
 
 # free_obs_experiment_raw_results.csv
 # memory_experiment_raw_results_2.csv
-dt.data <- data.table(read.csv("/Users/garidor/Desktop/bandits-rl-project/results/free_obs_raw_results/free_obs_experiment_raw_results_6.csv", header=T, stringsAsFactors=T))
+dt.data <- data.table(read.csv("/Users/garidor/Desktop/bandits-rl-project/results/free_obs_raw_results/free_obs_experiment_raw_results_1.csv", header=T, stringsAsFactors=T))
 
 unique(dt.data$Memory.Size)
 
-dt.data_subset = dt.data[(Memory.Size==100)]
+dt.data_subset = dt.data[(Memory.Size==1000)]
 
 ggplot(dt.data_subset, aes(x=as.factor(Memory.Size), y=Market.Share.for.P1, color=as.factor(Time.Horizon))) + 
   facet_grid(Prior + P1.Alg ~ Agent.Alg + P2.Alg) + 

### 2018-03-06 lib/agent/MultinomialLogit.py

@@ -1,0 +1,16 @@
+import random as rand
+import numpy as np
+from Agent import Agent
+
+class MultinomialLogit(Agent):
+  def selectPrincipal(self):
+    scores = self.informationSet.getScores()
+    total = sum([np.exp(score) for score in scores.values()])
+    probs = { principal: np.exp(score) / total for (principal, score) in scores.iteritems() }
+    
+    threshold = rand.random()
+    cumProb = 0.0
+    for (principal, prob) in probs.iteritems():
+      cumProb += prob
+      if threshold <= cumProb
+        return (principal, self.principals[principal])

### 2018-03-13 analysis_code/agents_table.Rmd
new results

--- 2018-02-28 analysis_code/agents_table.Rmd
@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_raw_results_4.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/full_tournament_raw_results.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
+memory_sizes <- sort(as.list(unique(dat['Memory.Size']))$Memory.Size, decreasing=TRUE)
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
+time_horizons <- c(1000)
 priors <- as.list(unique(dat['Prior']))$Prior
-EXPERIMENT_1_CV_HIGH <- 156.71410383 #chisquared critical values for 124 degrees of freedom
-EXPERIMENT_1_CV_LOW <- 95.07008897
-
-EXPERIMENT_2_CV_HIGH <- 128.42198864
-EXPERIMENT_2_CV_LOW <- 73.36108019
 
@@
 
-alg_pairs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "ThompsonSampling", "DynamicGreedy", "DynamicGreedy", "DynamicEpsilonGreedy")
-print("Reported means are for principal 1 which is the first algorithm listed in the pair for each column<br>")
-print("In each cell, first line is mean, second line is variance, third line is % of the time that market share > 0.1 or market share < 0.9")
+alg_pairs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "ThompsonSampling", "DynamicGreedy", "DynamicGreedy", "DynamicEpsilonGreedy", "ThompsonSampling", "ThompsonSampling", "DynamicEpsilonGreedy", "DynamicEpsilonGreedy", "DynamicGreedy", "DynamicGreedy")
+#print("Reported means are for principal 1 which is the first algorithm listed in the pair for each column<br>")
+#print("In each cell, first line is mean, second line is variance, third line is % of the time that market share > 0.1 or market share < 0.9")
 library(xtable)
+for (memory in memory_sizes) {
 for (prior in priors) {
     for (j in 1:length(time_horizons)) {
-    results <- matrix(nrow=length(agent_algs), ncol=length(algs))
-    colnames(results) <- c(paste(concise_alg_rep(alg_pairs[1]), concise_alg_rep(alg_pairs[2]), sep=" vs "), paste(concise_alg_rep(alg_pairs[3]), concise_alg_rep(alg_pairs[4]), sep=" vs "), paste(concise_alg_rep(alg_pairs[5]), concise_alg_rep(alg_pairs[6]), sep=" vs "))
+      results <- matrix(nrow=length(agent_algs), ncol=(length(alg_pairs) / 2))
+    colnames(results) <- c(paste(concise_alg_rep(alg_pairs[1]), concise_alg_rep(alg_pairs[2]), sep=" vs "), paste(concise_alg_rep(alg_pairs[3]), concise_alg_rep(alg_pairs[4]), sep=" vs "), paste(concise_alg_rep(alg_pairs[5]), concise_alg_rep(alg_pairs[6]), sep=" vs "), paste(concise_alg_rep(alg_pairs[7]), concise_alg_rep(alg_pairs[8]), sep=" vs "), paste(concise_alg_rep(alg_pairs[9]), concise_alg_rep(alg_pairs[10]), sep=" vs "), paste(concise_alg_rep(alg_pairs[11]), concise_alg_rep(alg_pairs[12]), sep=" vs "))
     rownames(results) <- get_agent_rep(agent_algs)
@@
           test <- t.test(share)
-          cin <- signif(test$conf.int[2] - mean(share), digits=2)
+          cin <- signif(test$conf.int[2] - mean(share), digits=1)
           shares <- sum(share >= (1 - ZERO_ONE_CUTOFF) | share <= ZERO_ONE_CUTOFF, na.rm=TRUE) / nrow(filtered_dat)
           shares_percent <- signif(shares * 100, digits=2)
-          lower_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_HIGH, digits=2)
-          high_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_LOW, digits=2)
-          results[i, (k+1)/2] <- paste("<b>",signif(mean(share), digits=2), "</b> +/-", cin, "<br>", signif(var(share), digits=2), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", shares_percent, "%")
+          cv_high <- qchisq(.975, df=length(share) - 1)
+          cv_low <- qchisq(.025, df=length(share) - 1)
+          lower_var <- signif((length(share) - 1) * var(share) / cv_high, digits=1)
+          high_var <- signif((length(share) - 1) * var(share) / cv_low, digits=1)
+          results[i, (k+1)/2] <- paste("<b>",signif(mean(share), digits=2), "</b> +/-", cin, "<br>Var:", signif(var(share), digits=2), "<br>Share:", shares_percent, "%")
       }
     }
-      tab <-xtable(results, caption=paste("Results for", time, prior))
+      tab <-xtable(results, caption=paste("Results for", "t=",time, prior, "Memory=", memory))
       print(tab, type="html")
   }
-  print('<br><br><br><br><br><br>')
+  }
 }

### 2018-03-13 analysis_code/tournament_table.Rmd
new results

@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_raw_results.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/full_tournament_raw_results.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(5000, 10000, 15000)
+time_horizons <- c(1000)
 priors <- as.list(unique(dat['Prior']))$Prior
@@
             high_var <- signif((length(share) - 1) * var(share) / cv_low, digits=2)
-            results[k, l] <- paste("<b>", signif(mean(share), digits=2), "</b> +/-", cin, "<br>", signif(var(share), digits=2), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", shares)
+            results[k, l] <- paste("<b>", signif(mean(share), digits=2), "</b> +/-", cin, "<br>", signif(var(share), digits=2), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", 100*signif(shares, digits=2), "%")
           }

### 2018-03-13 experiment_1/experiment_1.tex
new results

--- 2018-02-15 experiment_1/experiment_1.tex
@@
 \usepackage{algorithm,algpseudocode}
+\usepackage{pdfpages}
 
@@
 
-Considered $K = 3$, $T = 5005$. Report statistics at $t = 1000, 3000, 5000$ \\
+Considered $K = 3$, $T = 1001$, $N = 1200$. Report statistics at $t = 1000$ \\
 \textbf{The Bandit priors that were considered}:
@@
 \begin{enumerate}
-\item Medium - 9 arms with mean 0.50, 1 arm with mean 0.55 (+ 0.05)
-\item High - 9 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
+\item High - 2 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
 \end{enumerate}
@@
 \item HardMaxWithRandom
-\item SoftMax
+\item SoftMax ($\alpha = 30$)
 \end{enumerate}
@@
 \begin{enumerate}
-\item 10
-\item 25
 \item 100
@@
 \For{Each prior $p$}
-	\For{Each agent algorithm $agent_alg$}
-		\For{Each principal algorithm pair $principal_alg1$, $principal_alg2$}
-			\For{Each simulation $i$}
-				\State Generate true distribution from $p$ (except for needle-in-		haystack, just use $p$ itself)
+\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
+\State Generate $T \times K$ realizations for the arms 
+	\For{Each agent algorithm $agent alg$}
+		\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
+			\For{$N$ simulations}
+
+				\State Give the agents 5 observations from each principal
 				\State Run simulation for T periods
@@
 
-\pagebreak
 \section*{Results}
+All results are reported for memory size = 100. \\
+\vspace{0.25cm}
 
-One thing which is ambiguous to define is the regret value to use when a principal never gets chosen in a given simulation.
-When calculating any of the aggregate regret statistics we drop these simulations, but we do record how many rounds have an undefined regret.
+The rows represent the agent algorithms (defined as above).
+The columns represent principal 1 vs principal 2 (e.g.
+TS vs DG means Thompson Sampling is principal 1 and Dynamic Greedy is principal 2).
 \\
-\vspace{1cm}
+\vspace{0.1cm}
 
-First, we'll restrict focus to $t = 5000$ and look at the performance of ThompsonSampling. Note that the y axis here represents the market share that the ThompsonSampling principal gets.
-\vspace{0.5cm}
-Performance of ThompsonSampling vs DynamicGreedy and DynamicEpsilonGreedy across all agent models, memory sizes, and priors: \\
-\includegraphics[scale=0.75]{ts_perf_dg} \\
-\includegraphics[scale=0.75]{ts_perf_deg} \\
+Within each cell, the following data are presented: \\
 
-Now, looking across different memory sizes for each agent model (still using each prior): \\
-\textbf{HardMax} \\
-\includegraphics[scale=0.75]{hm_ts_deg} \\
-\includegraphics[scale=0.75]{hm_ts_dg} \\
+\begin{enumerate}
+\item The first row displays the sample mean market share for principal 1 as well as a 95\% confidence band of the mean.
 
-\pagebreak
-\textbf{HardMaxWithRandom} \\
-\includegraphics[scale=0.75]{hmr_ts_deg} \\
-\includegraphics[scale=0.75]{hmr_ts_dg} \\
+\item The second row displays the sample variance of the market share and in parentheses are 95 \% confidence bands for the variance.
 
-\pagebreak
-\textbf{SoftMax} \\
-\includegraphics[scale=0.75]{sm_ts_deg} \\
-\includegraphics[scale=0.75]{sm_ts_dg} \\
-
-\pagebreak
-We'll now look fix memory size to be 100 and look at the performance across priors.\\
-\textbf{HardMax} \\
-\includegraphics[scale=0.75]{hm_ts_deg_prior} \\
-\includegraphics[scale=0.75]{hm_ts_dg_prior} \\
-
-\pagebreak
-\textbf{HardMaxWithRandom} \\
-\includegraphics[scale=0.75]{hmr_ts_deg_prior} \\
-\includegraphics[scale=0.75]{hmr_ts_dg_prior} \\
-
-\pagebreak
-\textbf{SoftMax} \\
-\includegraphics[scale=0.75]{sm_ts_deg_prior} \\
-\includegraphics[scale=0.75]{sm_ts_dg_prior} \\
+\item The third row displays the \% of simulations that resulted in ``extreme" market shares.
+Extreme market shares are defined as being simulations were one of the principals (either principal 1 or 2) ended up with 90\% or more of the market.
+\end{enumerate}
+\includepdf[pages={-}]{agents_table}
+Comments: Almost everything is just a 50-50 split!
+\begin{enumerate}
+\item Almost all of the same algorithms pitted against each other ended with getting a 50/50 market share. This seems to be a good sanity check.
+\item Strange things: focus on TS vs DEG (Heavy Tail). DEG wins for HM and HMR, but TS wins for SM. This seems weird? TS also seems to lose on HMR with Uniform prior.
+\item Why does DynamicGreedy do so well on the HeavyTail prior? Does this make sense?
+\end{enumerate}
 

### 2018-03-13 experiment_2/experiment_2.tex
updates
new results

--- 2018-02-15 experiment_1/experiment_1.tex
@@
 
-\title{Experiment 1}
+\title{Experiment 2}
 \maketitle
@@
 
-Considered $K = 3$, $T = 5005$. Report statistics at $t = 1000, 3000, 5000$ \\
+Considered $K = 3$, $T = 1001$, $N = 500$. Report statistics at $t = 1000$ \\
 \textbf{The Bandit priors that were considered}:
@@
 \item HardMaxWithRandom
-\item SoftMax
+\item SoftMax ($\alpha = 10$)
 \end{enumerate}
@@
 \For{Each prior $p$}
-	\For{Each agent algorithm $agent_alg$}
-		\For{Each principal algorithm pair $principal_alg1$, $principal_alg2$}
-			\For{Each simulation $i$}
-				\State Generate true distribution from $p$ (except for needle-in-		haystack, just use $p$ itself)
+\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
+\State Generate $T \times K$ realizations for the arms 
+	\For{Each agent algorithm $agent alg$}
+		\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
+			\For{$N$ simulations}
+
+				\State Give the agents 5 observations from each principal
+				\State Give principal 2 200 free observations (the agents also get these observations)
 				\State Run simulation for T periods
@@
 
-\pagebreak
 \section*{Results}
+All results are reported for memory size = 100. \\
+\vspace{0.25cm}
 
-One thing which is ambiguous to define is the regret value to use when a principal never gets chosen in a given simulation.
-When calculating any of the aggregate regret statistics we drop these simulations, but we do record how many rounds have an undefined regret.
+The rows represent Principal 1 and the columns represent Principal 2.
+Thus, the cell (1, 2) represents principal 1 playing Thompson Sampling and principal 2 playing DynamicEpsilonGreedy.
+In this experiment remember that principal 2 gets 200 free observations.
 \\
-\vspace{1cm}
+\vspace{0.1cm}
 
-First, we'll restrict focus to $t = 5000$ and look at the performance of ThompsonSampling. Note that the y axis here represents the market share that the ThompsonSampling principal gets.
-\vspace{0.5cm}
-Performance of ThompsonSampling vs DynamicGreedy and DynamicEpsilonGreedy across all agent models, memory sizes, and priors: \\
-\includegraphics[scale=0.75]{ts_perf_dg} \\
-\includegraphics[scale=0.75]{ts_perf_deg} \\
+Within each cell, the following data are presented: \\
 
-Now, looking across different memory sizes for each agent model (still using each prior): \\
-\textbf{HardMax} \\
-\includegraphics[scale=0.75]{hm_ts_deg} \\
-\includegraphics[scale=0.75]{hm_ts_dg} \\
+\begin{enumerate}
+\item The first row displays the sample mean market share for principal 1 as well as a 95\% confidence band of the mean.
 
-\pagebreak
-\textbf{HardMaxWithRandom} \\
-\includegraphics[scale=0.75]{hmr_ts_deg} \\
-\includegraphics[scale=0.75]{hmr_ts_dg} \\
+\item The second row displays the sample variance of the market share and in parentheses are 95 \% confidence bands for the variance.
 
-\pagebreak
-\textbf{SoftMax} \\
-\includegraphics[scale=0.75]{sm_ts_deg} \\
-\includegraphics[scale=0.75]{sm_ts_dg} \\
+\item The third row displays the \% of simulations that resulted in ``extreme" market shares.
+Extreme market shares are defined as being simulations were one of the principals (either principal 1 or 2) ended up with 90\% or more of the market.
+\end{enumerate}
+\includepdf[pages={-}]{tournament_table}
 
-\pagebreak
-We'll now look fix memory size to be 100 and look at the performance across priors.\\
-\textbf{HardMax} \\
-\includegraphics[scale=0.75]{hm_ts_deg_prior} \\
-\includegraphics[scale=0.75]{hm_ts_dg_prior} \\
-
-\pagebreak
-\textbf{HardMaxWithRandom} \\
-\includegraphics[scale=0.75]{hmr_ts_deg_prior} \\
-\includegraphics[scale=0.75]{hmr_ts_dg_prior} \\
-
-\pagebreak
-\textbf{SoftMax} \\
-\includegraphics[scale=0.75]{sm_ts_deg_prior} \\
-\includegraphics[scale=0.75]{sm_ts_dg_prior} \\
+Comments: The story I seem to get from this is (roughly) that the worse the algorithm that the incumbent plays is, the better the entrant will do.
+However, in order for the entrant to do as well as possible, the entrant should play a worse algorithm!
+This seems counterintuitive, it means that the worst case for the entrant is when we're both playing ThompsonSampling and the best case for the entrant is when we're both playing DynamicGreedy.
 

### 2018-03-13 preliminary/preliminary_report.tex
new results

@@
 
-Considered $K = 3$, $T = 6000$. \\
+$N = 200$ simulations were run. \\
 \textbf{The Bandit priors that were considered}:
 \begin{itemize}
-\item Uniform: Draw the mean rewards for the arms from [0.25, 0.75]
-\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$).
-With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
 \item Needle-in-haystack
 \begin{enumerate}
-\item Low - 9 arms with mean 0.50, 1 arm with mean 0.01 (+ 0.01)
-\item Medium - 9 arms with mean 0.50, 1 arm with mean 0.55 (+ 0.05)
-\item High - 9 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
+\item High (with three arms) - 2 arms with mean 0.50, 1 arm with mean 0.7 (+ 0.2)
+\item High (with ten arms) - 9 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
+\item Medium-High (with ten arms) - 8 arms with mean 0.50, 1 arm with 0.60, 1 arm 0.80
 \end{enumerate}
@@
 \item DynamicGreedy with priors of $Beta(1, 1)$ for every arm
-\item Bayesian Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm
-\begin{enumerate}
-\item $\epsilon=t^{-1/3}$
-\item $\epsilon=T^{-1/3}$
-\item $\epsilon=0.05$
-\end{enumerate}
-\item Non-Bayesian $\epsilon$-greedy - the greedy decision was made based on empirical mean. When there were zero observations, assumed that the empirical mean was 0 (this seems questionable).
-\begin{enumerate}
-\item $\epsilon=t^{-1/3}$
-\item $\epsilon=T^{-1/3}$
-\item $\epsilon=0.05$
-\end{enumerate}
-\item UCB1
-\begin{enumerate}
-\item UCB1 with constant $\sqrt{2log(t)}$
-\item UCB1 with constant $1$
-\end{enumerate}
+\item Bayesian Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm and $\epsilon = 0.05$
+\item Non-Bayesian $\epsilon$-greedy - the greedy decision was made based on empirical mean.
+When there were zero observations, assumed that the empirical mean was 0 (this seems questionable), $\epsilon = 0.05$
+\item UCB1 (with constant $1$)
 \end{enumerate}
@@
 	\For{Each experiment $i$}
-		\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
+		\State Use $p$ as the true distribution of the arms
 		\State Generate realizations for each arm and round $t$ and instantiate bandit instance
@@
 
-One caveat to the reported results is that the Bayesian greedy results were run separately and thus have different true distributions and realizations than the rest of the algorithms.
-If we decide to include the Bayesian version instead of the non-Bayesian version we'll have to re-run these simulations.
-
 \section*{Results}
@@
 The actual datapoints, when plotted, are incredibly volatile but this volatility decreases as we increase N.
-Part of this comes from the fact that what is plotted is the actual realized reward and not the mean reward \textbf{*}.
-We briefly discussed this last time and perhaps we should be reporting this instead (or in addition).
+Part of this comes from the fact that what is plotted is the actual realized reward and not the mean reward.
 \\
-\subsection*{Instantaneous Reward Plots}
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack Low inst"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack Medium inst"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack High inst"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Uniform inst"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail inst"} \\
+\subsection*{Instantaneous Realized Reward Plots}
+These plots contain the fitted lines for the instantaneous realized reward. \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Needle In Haystack 1 High 10 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Needle In Haystack 1 Medium 1 High 10 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Needle In Haystack 1 High 3 arms"} \\
 
-\subsection*{Cumulative Reward Plots}
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack Low cumulative"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack Medium cumulative"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Needle In Haystack High cumulative"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Uniform cumulative"} \\
-\includegraphics[scale=0.5]{"../results/Reward Trajectory for Heavy Tail cumulative"} \\
+\subsection*{Instantaneous True Mean Reward Plots}
+These plots contain the fitted lines for the TRUE mean reward for the selected arms. \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous True Mean Reward Trajectory for Needle In Haystack 1 High 10 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous True Mean Reward Trajectory for Needle In Haystack 1 Medium 1 High 10 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous True Mean Reward Trajectory for Needle In Haystack 1 High 3 arms"} \\
 
-\subsection*{Empirical Standard Deviation}
-
-Another thing we discussed previously was reporting and plotting the empirical standard deviation of the realized rewards.
-These tend to be quite large.
-Across almost every distribution and algorithm, the empirical standard deviation of the vector of realized rewards was on average ~0.4-0.5.
-This was calculated, for each algorithm and prior, as follows: we have a vector of $N$ realized rewards for each $t$ and we can find the empirical standard deviation of this vector.
-Averaging across each $t$ for each algorithm and prior, all of them have an average empirical standard deviation between 0.4-0.5.
-An alternative statistic of potential interest would be to look at the empirical standard of the mean rewards of the selected arms which would likely be smaller.
-Not sure which makes more sense.
-
+Comments: It does appear that using $10$ arms instead of $3$ in this case does not appear too different.
+It appears that Thompson Sampling and UCB learn quickly (roughly seem to flatline at $t = 2000$) and DynamicGreedy slowly improves but does not seem to get to convergence even by $t = 10000$.
+However, the difference between the TS and DG trajectories is not vastly different between the 3 arm and 10 arm case.
+One difference to note is that it seems like one thing that does change between the two is that Dynamic $\epsilon$-greedy seems to take longer to learn (doesn't even appear to converge by $t = 10000$) in the more complex cases but seems to converge as quickly as TS and UCB in the 3 arm case.
 \end{document}

### 2018-03-13 analysis_code/kevin_plots.R

@@ -61,11 +61,11 @@ ggplot(dt.data[(Distribution=="Needle In Haystack Medium")], aes(x=t, y=Instanta
 
 # free_obs_experiment_raw_results.csv
 # memory_experiment_raw_results_2.csv
-dt.data <- data.table(read.csv("/Users/garidor/Desktop/bandits-rl-project/results/free_obs_raw_results/free_obs_experiment_raw_results_1.csv", header=T, stringsAsFactors=T))
+dt.data <- data.table(read.csv("/Users/garidor/Desktop/bandits-rl-project/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", header=T, stringsAsFactors=T))
 
 unique(dt.data$Memory.Size)
 
-dt.data_subset = dt.data[(Memory.Size==1000)]
+dt.data_subset = dt.data[(Memory.Size==100)]
 
 ggplot(dt.data_subset, aes(x=as.factor(Memory.Size), y=Market.Share.for.P1, color=as.factor(Time.Horizon))) + 
   facet_grid(Prior + P1.Alg ~ Agent.Alg + P2.Alg) + 

### 2018-03-13 analysis_code/preliminary_plots.R

@@ -8,19 +8,23 @@ library("reshape")
 
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms.csv", sep=""))
-dat2 <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_2.csv", sep=""))
-dat <- rbind(dat, dat2)
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_20_arms.csv", sep=""))
+
+dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_3.csv", sep=""))
+#dat2 <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_2.csv", sep=""))
+#dat <- rbind(dat, dat2)
+dat$Algorithm <- replace(as.character(dat$Algorithm), dat$Algorithm == "DynamicEpsilonGreedy", "DynamicEpsilonGreedy, 0.05")
+dat$Algorithm <- replace(as.character(dat$Algorithm), dat$Algorithm == "NonBayesianEpsilonGreedy", "NonBayesianEpsilonGreedy, 0.05")
 algs <- as.list(unique(dat['Algorithm']))$Algorithm
-print(algs)
 filter_algs <- c("ThompsonSampling", "UCB1WithConstantOne", "NonBayesianEpsilonGreedy, 0.05", "DynamicEpsilonGreedy, 0.05", "DynamicGreedy")
 dat <- filter(dat, Algorithm %in% filter_algs)
 dists <- as.list(unique(dat['Distribution']))$Distribution
+dists <- c("Needle In Haystack 1 High")
 
 filter_by_dist_and_plot <- function(dist) {
   d <- filter(dat, Distribution == dist)
-  title <- paste("Reward Trajectory for", dist, "inst")
-  q <- ggplot(data=d, aes(x=t, y=Instantaneous.Reward.Mean, colour=Algorithm)) + ggtitle(title) + ylab("Instantaneous Mean Reward") + xlab("time") +
+  title <- paste("Instantaneous Realized Reward Trajectory for", dist, "3 arms")
[…]

### 2018-03-13 preliminary.py

@@ -21,7 +21,7 @@ from scipy.stats import bernoulli, beta
 
 T = 10000
 N = 200
-K = 20
+K = 3
 
 
 DEFAULT_COMMON_PRIOR = [beta(1, 1) for k in xrange(K)]
@@ -38,20 +38,21 @@ needle_in_haystack = [bernoulli(default_mean) for i in xrange(K)]
 needle_in_haystack_50_high = deepcopy(needle_in_haystack)
 needle_in_haystack_50_high[int(K/2)] = bernoulli(default_mean + 0.2)
 
-needle_in_haystack_50_one_medium_one_high = deepcopy(needle_in_haystack)
-needle_in_haystack_50_one_medium_one_high[int(K/2)] = bernoulli(default_mean + 0.3)
-needle_in_haystack_50_one_medium_one_high[int(K/2) + 1] = bernoulli(default_mean + 0.1)
+#needle_in_haystack_50_one_medium_one_high = deepcopy(needle_in_haystack)
+#needle_in_haystack_50_one_medium_one_high[int(K/2)] = bernoulli(default_mean + 0.3)
+#needle_in_haystack_50_one_medium_one_high[int(K/2) + 1] = bernoulli(default_mean + 0.1)
 
 
 
-#heavy_tail_prior = beta(0.6, 0.6)
-#heavy_tailed = [bernoulli(heavy_tail_prior.rvs()) for i in xrange(K)]
+heavy_tail_prior = beta(0.6, 0.6)
+heavy_tailed = [bernoulli(heavy_tail_prior.rvs()) for i in xrange(K)]
 
 
 ALGS = [NonBayesianEpsilonGreedy, ThompsonSampling, UCB1WithConstantOne, DynamicEpsilonGreedy, DynamicGreedy]
 BANDIT_DISTR = {
   'Needle In Haystack 1 High': needle_in_haystack_50_high,
-  'Needle In Haystack 1 Medium 1 High': needle_in_haystack_50_one_medium_one_high
+  'Uniform': None,
+  'Heavy Tail': None
 }
 
 
@@ -65,7 +66,7 @@ simResults = {}
 This runs N iterations of the experiment. In each iteration a set of true distributions is chosen and then for each algorithm we record the reward history. After this is done we aggregate the results and report them.
 '''
 
-with open('results/preliminary_raw_results/preliminary_plots_20_arms.csv', 'w') as csvfile:
[…]

### 2018-03-15 experiment_2/experiment_2.tex
updated experiment 2

@@
 Comments: The story I seem to get from this is (roughly) that the worse the algorithm that the incumbent plays is, the better the entrant will do.
-However, in order for the entrant to do as well as possible, the entrant should play a worse algorithm!
-This seems counterintuitive, it means that the worst case for the entrant is when we're both playing ThompsonSampling and the best case for the entrant is when we're both playing DynamicGreedy.
+Thus, the incumbent should want to play a better algorithm.
+However, in order for the entrant to do as well as possible, the entrant should in fact play a worse algorithm!
+This is interesting but I'm not sure it is intuitive.
 

### 2018-03-16 experiment_2/experiment_2.tex
updated comments

@@
 However, in order for the entrant to do as well as possible, the entrant should in fact play a worse algorithm!
-This is interesting but I'm not sure it is intuitive.
+If you think of it from the incumbent point of view, the incumbent is in the market first and so wants to learn as much as possible before the entrant comes in (thus does better with a better algorithm since it should have learned more by the time the entrant enters).
+From the entrant point of view, since the incumbent has already learned something in order to have any chance to compete you do better by purely exploiting than by engaging in potentially suboptimal exploration.
+Two follow-up questions: Will this result hold if we run the simulations for more periods?
+What will happen to this result if we increase memory size?
 

### 2018-03-16 experiment_1/experiment_1.tex
fix typo

@@
 \item Almost all of the same algorithms pitted against each other ended with getting a 50/50 market share. This seems to be a good sanity check.
-\item Strange things: focus on TS vs DEG (Heavy Tail). DEG wins for HM and HMR, but TS wins for SM. This seems weird? TS also seems to lose on HMR with Uniform prior.
-\item Why does DynamicGreedy do so well on the HeavyTail prior? Does this make sense?
+\item Things that stand out: DynamicGreedy does surprisingly well on the Heavy Tail prior.
+DG seems to beat both DEG and TS with the Heavy Tail prior.
+DEG also beats TS in the heavy tail prior, though TS seems to beat DG when the agent response function is SoftMax.
+Does this make sense?
+\item Almost everything in both Uniform and Needle In Haystack - high ends up more or less being a tie.
 \end{enumerate}

### 2018-03-16 preliminary/preliminary_report.tex
fix typo

@@
 \item High (with three arms) - 2 arms with mean 0.50, 1 arm with mean 0.7 (+ 0.2)
-\item High (with ten arms) - 9 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
-\item Medium-High (with ten arms) - 8 arms with mean 0.50, 1 arm with 0.60, 1 arm 0.80
+\item High (with ten arms) - 19 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
+\item Medium-High (with ten arms) - 18 arms with mean 0.50, 1 arm with 0.60, 1 arm 0.80
 \end{enumerate}

### 2018-03-16 preliminary/preliminary_report.tex
another typo

--- 2018-03-13 preliminary/preliminary_report.tex
@@
 \item High (with three arms) - 2 arms with mean 0.50, 1 arm with mean 0.7 (+ 0.2)
-\item High (with ten arms) - 9 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
-\item Medium-High (with ten arms) - 8 arms with mean 0.50, 1 arm with 0.60, 1 arm 0.80
+\item High (with twenty arms) - 19 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
+\item Medium-High (with twenty arms) - 18 arms with mean 0.50, 1 arm with 0.60, 1 arm 0.80
 \end{enumerate}

### 2018-03-16 preliminary/preliminary_report.tex
another typo

@@
 
-Comments: It does appear that using $10$ arms instead of $3$ in this case does not appear too different.
+Comments: It does appear that using $20$ arms instead of $3$ in this case does not appear too different.
 It appears that Thompson Sampling and UCB learn quickly (roughly seem to flatline at $t = 2000$) and DynamicGreedy slowly improves but does not seem to get to convergence even by $t = 10000$.
-However, the difference between the TS and DG trajectories is not vastly different between the 3 arm and 10 arm case.
+However, the difference between the TS and DG trajectories is not vastly different between the 3 arm and 20 arm case.
 One difference to note is that it seems like one thing that does change between the two is that Dynamic $\epsilon$-greedy seems to take longer to learn (doesn't even appear to converge by $t = 10000$) in the more complex cases but seems to converge as quickly as TS and UCB in the 3 arm case.

### 2018-03-20 experiment_2/experiment_2.tex
update

@@
 What will happen to this result if we increase memory size?
+Another question is whether or not the effect is driven by the better information the incumbent has or the effect of reputation.
+\vspace{0.5cm}
+
+This motivates the following experiment which has an identical setup as before (principal 2 gets to be in the game for 200 periods before principal 1), however after the 200 rounds we are going to erase the memory of the agents.
+Thus, when principal 1 enters the market the incumbent (principal 2) does not have an advantage in terms of reputation, but only in terms of the information s/he possesses.
+The results don't change much qualitatively from the case we first considered.
+\includepdf[pages={-}]{no_reputation_tournament_table}
+
+Another check of the robustness of the results: The memory size is increased to 250 so that when principal 1 enters the market the entire history of the incumbent is contained in the information set of the incumbent.
+\includepdf[pages={-}]{more_memory_tournament_table}
 

### 2018-03-26 analysis_code/agents_table.Rmd
updates

--- 2018-02-28 analysis_code/agents_table.Rmd
@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_raw_results_4.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_exp_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
           test <- t.test(share)
-          cin <- signif(test$conf.int[2] - mean(share), digits=2)
+          cin <- signif(test$conf.int[2] - mean(share), digits=1)
           shares <- sum(share >= (1 - ZERO_ONE_CUTOFF) | share <= ZERO_ONE_CUTOFF, na.rm=TRUE) / nrow(filtered_dat)
           shares_percent <- signif(shares * 100, digits=2)
-          lower_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_HIGH, digits=2)
-          high_var <- signif((length(share) - 1) * var(share) / EXPERIMENT_1_CV_LOW, digits=2)
-          results[i, (k+1)/2] <- paste("<b>",signif(mean(share), digits=2), "</b> +/-", cin, "<br>", signif(var(share), digits=2), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", shares_percent, "%")
+          cv_high <- qchisq(.975, df=length(share) - 1)
+          cv_low <- qchisq(.025, df=length(share) - 1)
+          lower_var <- signif((length(share) - 1) * var(share) / cv_high, digits=1)
+          high_var <- signif((length(share) - 1) * var(share) / cv_low, digits=1)
+          results[i, (k+1)/2] <- paste("<b>",signif(mean(share), digits=2), "</b> +/-", cin, "<br>Var:", signif(var(share), digits=2), "<br>Share:", shares_percent, "%")
       }
     }
-      tab <-xtable(results, caption=paste("Results for", time, prior))
+      tab <-xtable(results, caption=paste("Results for", "t=",time, prior, "Memory=", memory), row.header='Entrant', col.header='Incumbent')
       print(tab, type="html")
   }
-  print('<br><br><br><br><br><br>')
+  }
 }

### 2018-03-26 analysis_code/tournament_table.Rmd
updates

@@
 ```{r echo=FALSE,message=FALSE,xtable, results="asis"}
-
 library(dplyr)
@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/full_tournament_raw_results.csv", sep=""))
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/full_tournament_raw_results.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_test_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg

### 2018-03-26 lib/agent/Agent.py

@@ -33,5 +33,8 @@ class Agent:
     self.numRounds += 1
     self.informationSet.updateInformationSet(reward, arm, principalName)
 
+  def resetInformationSet(self):
+    self.informationSet = InformationSet(self.principals, self.K, self.priors, memory=self.memory, discount_factor=self.discount_factor, score=self.score)
+
   def printMeanBeliefs(self):
     print(self.informationSet.getScores())

### 2018-03-26 lib/agent/HardMaxWithRandom.py

@@ -4,7 +4,7 @@ from HardMax import HardMax
 class HardMaxWithRandom(HardMax):
   # epsilon: the baseline probablility that each principal will be picked with. must be in (0, 0.5).
   # if epsilon=0.1, for example, then 20% of the time, the agent picks randomly between the 2 principals
-  def __init__(self, principals, K, priors=None, epsilon=0.025, memory=50):
+  def __init__(self, principals, K, priors=None, epsilon=0.15, memory=50):
     super(HardMaxWithRandom, self).__init__(principals, K, priors, memory=memory)
     self.epsilon = epsilon
 

### 2018-03-26 main.py

@@ -27,17 +27,19 @@ import random
 numCores = multiprocessing.cpu_count()
 
 K = 3
-T = 1002
-NUM_SIMULATIONS = 500
+T = 8000
+NUM_SIMULATIONS = 100
 
-FREE_OBS = True
+FREE_OBS = False
+FREE_OBS_NUM = 100
+exp_name = 'large_epsilon'
 
-AGENT_ALGS = [HardMax, HardMaxWithRandom, SoftMax]
+AGENT_ALGS = [HardMaxWithRandom]
 
 # valid principal algs are: [StaticGreedy, UCB, DynamicEpsilonGreedy, DynamicGreedy, ExploreThenExploit, ThompsonSampling]
 ALG_PAIRS = [(ThompsonSampling, DynamicEpsilonGreedy),(ThompsonSampling, DynamicGreedy), (DynamicGreedy, DynamicEpsilonGreedy),
-            (ThompsonSampling, ThompsonSampling), (DynamicGreedy, DynamicGreedy), (DynamicEpsilonGreedy, DynamicEpsilonGreedy),
-            (DynamicGreedy, ThompsonSampling), (DynamicEpsilonGreedy, ThompsonSampling), (DynamicEpsilonGreedy, DynamicGreedy)]
+            (ThompsonSampling, ThompsonSampling), (DynamicGreedy, DynamicGreedy), (DynamicEpsilonGreedy, DynamicEpsilonGreedy)]
+            #(DynamicGreedy, ThompsonSampling), (DynamicEpsilonGreedy, ThompsonSampling), (DynamicEpsilonGreedy, DynamicGreedy)]
 default_mean = 0.5
 needle_in_haystack = [bernoulli(default_mean) for i in xrange(K)]
 
@@ -55,17 +57,30 @@ BANDIT_DISTR = {
   'Needle In Haystack High': needle_in_haystack_50_high
 }
 
-WORKING_DIRECTORY = '/rigel/home/ga2449/bandits-rl-project/'
-#WORKING_DIRECTORY = ''
+#WORKING_DIRECTORY = '/rigel/home/ga2449/bandits-rl-project/'
+WORKING_DIRECTORY = ''
+
+if FREE_OBS:
+  dir_name = WORKING_DIRECTORY + 'results/free_obs_raw_results/'
+  base_name = dir_name + 'free_obs_experiment' + exp_name
+else:
+  dir_name = WORKING_DIRECTORY + 'results/tournament_raw_results/'
[…]

### 2018-03-26 simulate.py

@@ -32,7 +32,8 @@ def simulate(principalAlg1, principalAlg2, agentAlg, K, T,
   principal1Priors=None,
   principal2Priors=None,
   realizations=None,
-  freeObsForP2=False):
+  freeObsForP2=False,
+  freeObsNum=100):
 
   if principal1Priors is None:
     principal1Priors = getDefaultPrior(K)
@@ -61,8 +62,8 @@ def simulate(principalAlg1, principalAlg2, agentAlg, K, T,
     principal.resetStats()
 
   if freeObsForP2:
-    for j in range(200):
-      (reward, arm) = principals['principal2'].executeStep(i)
+    for j in range(freeObsNum):
+      (reward, arm) = principals['principal2'].executeStep(j)
       agents.updateInformationSet(reward, arm, 'principal2')
     principals['principal2'].resetStats()
 

### 2018-04-08 analysis_code/agents_table.Rmd
updates

@@
 #dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_exp_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_large_horizon_3_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(1000)
+time_horizons <- c(2000, 15000)
 priors <- as.list(unique(dat['Prior']))$Prior
@@
   } else if (alg == "DynamicEpsilonGreedy") {
-    return("DEG")
+    return("BEG")
   } else if (alg == "DynamicGreedy") {
-    return("DG")
+    return("BG")
   }
@@
   if (agent_alg == "HardMax") {
-    return("HM")
+    return("EU")
   } else if (agent_alg == "HardMaxWithRandom") {
-    return("HMR")
+    return("Noisy EU")
   } else if (agent_alg == "SoftMax") {
-    return("SM")
+    return("Logit")
   }

### 2018-04-08 analysis_code/tournament_table.Rmd
updates

@@
 #dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/full_tournament_raw_results.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_test_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
+alg_names <- c("Thompson Sampling", "Bayesian Epsilon Greedy", "Bayesian Greedy")
 options(xtable.sanitize.text.function=identity)
@@
         results <- matrix(nrow=length(algs), ncol=length(algs))
-        colnames(results) <- algs
-        rownames(results) <- algs
+        colnames(results) <- alg_names
+        rownames(results) <- alg_names
         for (k in 1:length(algs)) {

### 2018-04-08 overview/overview.tex
updates

\documentclass[11pt,letterpaper]{article}
\usepackage{fullpage}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{algorithm,algpseudocode}
\usepackage{pdfpages}

\graphicspath{{../results}}

\begin{document}

\title{Overview}
\maketitle

\section*{Simulation Details}

Considered $K = 3$, $Memory = 100$ (unless otherwise noted) \\
\textbf{The Bandit priors that were considered}:
\begin{itemize}
\item Uniform: Draw the mean rewards for the arms from [0.25, 0.75]
\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$). With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
\item Needle-in-haystack
\begin{enumerate}
\item High - 2 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
\end{enumerate}
\end{itemize}
\textbf{Algorithms considered}:
\begin{enumerate}
\item ThompsonSampling with priors of $Beta(1, 1)$ for every arm.
\item DynamicGreedy with priors of $Beta(1, 1)$ for every arm
\item Bayesian Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm and $\epsilon=0.05$
\end{enumerate}
\textbf{Agent Algorithms considered}:
\begin{enumerate}
\item HardMax
\item HardMaxWithRandom ($\epsilon = 0.05$)
\item SoftMax ($\alpha = 30$)
\end{enumerate}
\textbf{Memory Sizes}
\begin{enumerate}
\item 100
\end{enumerate}
\pagebreak
\textbf{Simulation Procedure}
\begin{algorithm}
\begin{algorithmic}[1]
\For{Each prior $p$}
\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
\State Generate $T \times K$ realizations for the arms 
	\For{Each agent algorithm $agent alg$}
		\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
			\For{$N$ simulations}

				\State Give the agents $k$ observations from each principal
				\State Give principal 2 $X$ free observations (the agents also get these observations)
				\State Run simulation for T periods
			\EndFor
		\EndFor
	\EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

\section*{Results}

First, let's look at the preliminary simulation results on the instances we considered:

\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Needle In Haystack 1 High 3 arms"} \\
\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Uniform 3 arms"} \\
\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Heavy Tail 3 arms"}

The first set of results we will consider are $X = 0$ and $k = 5$ so that there is simultaneous entry and only a small cold start. Recall to read the tables: \\
\begin{enumerate}
\item In bold is the average market share for principal 1 and with a 95\% confidence interval
\item The var term is the variance of the market shares
\item The ``share" line means the \% of simulations that ended up with one principal getting more than 90\% of the market.
\end{enumerate}

\section*{Simultaneous Entry}

\textbf{Question of Interest}: What algorithms should win under different agent response models?
\textbf{Conjecture}: HardMaxWithRandom and SoftMax should eventually lead to the better algorithm winning. \\
\textbf{Why?} Looking at the preliminary calibration plots, since there is some randomness in the choice rule of the agents, each principal should always be getting some free observations (our parameters such as the $\alpha$ in SoftMax and $\epsilon$ in HardMaxWithRandom tune how many expected observations a principal should get for a fixed $T$).  Let's fix attention on the case of ThompsonSampling vs Bayesian/Dynamic Greedy. Looking at each of the preliminary plots we see that ThompsonSampling beats Dynamic / Bayesian Greedy for sufficiently large $T$. Thus, if we calibrate the time horizon long enough so that the principal playing Thompson Sampling should, in expectation, get sufficiently many samples to be on the part of the curve where learning is almost done then it should be able to accumulate a higher reputation score (by having better arm selection than DynamicGreedy) and eventually get more of the market. \\

This seems to be confirmed in the simulations. Using the calibration from above ($\epsilon = 0.05$ for HMR and $\alpha = 30$ for SoftMax), the following are the results looking at $t = 2000$ and $t = 15000$ for each of the priors.

\includepdf[pages={-}]{tournament_results}

On the next page are the results for large number of simulations ($N=1200$), but low $T$ \\

\includepdf[pages={-}]{low_time_high_num_sim}

The conjecture for HMR and SM seems to be consistent with the data generated from the simulation. However, what is going on with HardMax appears confusing. A conjecture is that since there is very little initial information, the algorithm choice matters very little and random noise in the early rounds seems to lead to one firm taking the market. To get some evidence for this, we re-run simulations but track market share over time. In most cases, we get evidence of this and the vast majority of market shares over time look as follows:

\includegraphics[scale=0.25]{decreasing_ms_over_time} \\
\includegraphics[scale=0.25]{standard_ms_over_time} \\
\vspace{0.25cm}

However, there are some simulations where we do not observe this. For instance, \\
\includegraphics[scale=0.25]{odd_ms_over_time} \\

I am not sure what happens in these cases, but one conjecture is that these arise from random noise from a low memory size (memory here is 100, as we have been using in our experiment). I've re-run this experiment for a memory size of 1000 and there are fewer cases which have odd switches as shown above. However, would need to look into this more if we think this is worth digging into more.\\
As a comparison, what do the trajectories for the HardMaxWithRandom agent response functions look like? Haven't dug too deeply into the data for these simulations, but here is a sample one (ThompsonSampling vs DEG) that is similar to many in those that I have looked at: \\
\includegraphics[scale=0.25]{hmr_over_time} \\

\section*{Incumbent Experiment}
Now, we change $X = 200$. In words, we give principal 2 $X$ free observations (the ``incumbent") and this updates her information set and reputation amongst the agents. In this setting what should the incumbent do? What should the entrant do? Does this depend on the agent model? I would conjecture that when we have sufficient randomness in the agent model the same conclusions as before should hold - namely that the entrant, even if getting beat upon entry, should eventually get enough ``free" observations that she can catch up and take more of the market. However, the time horizon, in contrast to the simultaneous entry case, seems like it may be way longer! So far, most of the simulations have simply looked at $t = 1000$, but it may make sense to look at longer time horizons due to the conjecture from above. \\
\vspace{0.25cm}

What about with HardMax? The only intuition I have is for the incumbent. For the entrant, it is not clear to me what the entrant should do. For the incumbent, in our simulations we fix a $X$ when the entrant will come in and the incumbent should want to maximize both the information that she accrues before then and the reputation that she amongst the agents. However, this seems exactly like the standard exploration exploitation dilemma where reputation is simply the cumulative reward and ``smart" algorithms like Thompson Sampling ought to be better than the ``dumb" algorithms at maximizing information gain while at the same time optimizing cumulative reward. Thus, I expect that Thompson Sampling should be in a better starting position at the start of the competing bandits game (we can view the incumbent experiment as being the simultaneous entry game where the incumbent starts with a head start in terms of information and reputation) and thus control more of the market. This seems to be the case in the reported simulations, both with large $X = 200$ and $X = 100$. Below are the results, the first for $X = 200$ and the second for $X = 100$.

To read the tables: the columns represent the algorithm played by the incumbent and the row represents the algorithm played by the incumbent. For instance, looking at the second cell in the first row we have that the Incumbent played Dynamic (Bayesian) Epsilon Greedy and the entrant played Thompson Sampling and the reported mean is the average market share for the entrant.

\includepdf[pages={-}]{free_obs_tournament}

Now, look at $X = 100$

\includepdf[pages={-}]{less_free_obs_tournament_table}

\end{document}

### 2018-04-08 analysis_code/preliminary_plots.R

@@ -10,8 +10,8 @@ WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
 #dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_20_arms.csv", sep=""))
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_3.csv", sep=""))
-#dat2 <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_2.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_long.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_3.csv", sep=""))
 #dat <- rbind(dat, dat2)
 dat$Algorithm <- replace(as.character(dat$Algorithm), dat$Algorithm == "DynamicEpsilonGreedy", "DynamicEpsilonGreedy, 0.05")
 dat$Algorithm <- replace(as.character(dat$Algorithm), dat$Algorithm == "NonBayesianEpsilonGreedy", "NonBayesianEpsilonGreedy, 0.05")
@@ -19,12 +19,11 @@ algs <- as.list(unique(dat['Algorithm']))$Algorithm
 filter_algs <- c("ThompsonSampling", "UCB1WithConstantOne", "NonBayesianEpsilonGreedy, 0.05", "DynamicEpsilonGreedy, 0.05", "DynamicGreedy")
 dat <- filter(dat, Algorithm %in% filter_algs)
 dists <- as.list(unique(dat['Distribution']))$Distribution
-dists <- c("Needle In Haystack 1 High")
 
 filter_by_dist_and_plot <- function(dist) {
   d <- filter(dat, Distribution == dist)
-  title <- paste("Instantaneous Realized Reward Trajectory for", dist, "3 arms")
-  q <- ggplot(data=d, aes(x=t, y=Instantaneous.Realized.Reward.Mean, colour=Algorithm)) + ggtitle(title) + ylab("Instantaneous Mean Reward") + xlab("time") +
+  title <- paste("Reputation Trajectory for", dist, "3 arms")
[…]

### 2018-04-08 analysis_code/trajectory.R

@@ -1,0 +1,14 @@
+library(dplyr)
+
+WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
+f <- "tournament_experiment_everything_hmr_raw.csv"
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/", f, sep=""))
+filtered_dat <- filter(dat, Prior == "Heavy Tail")
+
+WINDOW <- 5000
+for (i in 220:240) {
+  begin <- (i*WINDOW) + 1
+  end_window <- (i + 1) * WINDOW
+  one_sim <- filtered_dat[begin:end_window,]
+  plot(one_sim[,"Time.Horizon"], one_sim[, "Market.Share.for.P1"], main="HMR Market Share over Time", ylab="Market Share", xlab="Time")
+}

### 2018-04-08 lib/InformationSet.py

@@ -79,7 +79,7 @@ class InformationSet:
 
   # We initialize an information class for each principal, initializing each principal as though it has been picked once 
   # and the mean of its prior distribution is considered to be the only observed reward
-  def __init__(self, principals, K, priors, memory, score, discount_factor):
+  def __init__(self, principals, K, memory, score, discount_factor):
     # note: here, principal is actually a string variable ("principal1") and v is a variable of class BanditAlgorithm
     self.infoSet  = { principal: Info(principal, K, memory=memory, discount_factor=discount_factor) for (principal, v) in principals.iteritems() }
     self.score = score

### 2018-04-08 lib/agent/Agent.py

@@ -17,24 +17,22 @@ class Agent:
     self.principals = principals
     self.priors = priors
     self.score = score
-    self.numRounds = 0
     self.K = K
     self.memory = memory
     self.discount_factor = discount_factor
 
     # for each principal store the number of times selected and total reward
-    self.informationSet = InformationSet(principals, K, priors, memory=memory, discount_factor=discount_factor, score=score)
+    self.informationSet = InformationSet(principals, K, memory=memory, discount_factor=discount_factor, score=score)
 
   @abc.abstractmethod
   def selectPrincipal(self):
     return
 
   def updateInformationSet(self, reward, arm, principalName):
-    self.numRounds += 1
     self.informationSet.updateInformationSet(reward, arm, principalName)
 
   def resetInformationSet(self):
-    self.informationSet = InformationSet(self.principals, self.K, self.priors, memory=self.memory, discount_factor=self.discount_factor, score=self.score)
+    self.informationSet = InformationSet(self.principals, self.K, discount_factor=self.discount_factor, memory=self.memory, score=self.score)
 
   def printMeanBeliefs(self):
     print(self.informationSet.getScores())

### 2018-04-08 preliminary.py

@@ -50,13 +50,15 @@ heavy_tailed = [bernoulli(heavy_tail_prior.rvs()) for i in xrange(K)]
 
 ALGS = [NonBayesianEpsilonGreedy, ThompsonSampling, UCB1WithConstantOne, DynamicEpsilonGreedy, DynamicGreedy]
 BANDIT_DISTR = {
-  'Heavy Tail': None
+  'Heavy Tail': heavy_tail_prior,
+  'Uniform': None,
+  'Needle In Haystack High': needle_in_haystack_50_high
 }
 
 
 # Algorithm, Arms, Prior, t, n, reward
 
-FIELDNAMES = ['Algorithm', 'K', 'Distribution', 't', 'Instantaneous Realized Reward Mean', 'Instantaneous Realized Reward Std', 'Cumulative Realized Reward Mean', 'Cumulative Realized Reward Std', 'Instantaneous Mean Reward Mean', 'Instantaneous Mean Reward Std', 'Cumulative Mean Reward Mean', 'Cumulative Mean Reward Std', 'Best Arm Mean']
+FIELDNAMES = ['Algorithm', 'K', 'Distribution', 't', 'Instantaneous Realized Reward Mean', 'Instantaneous Realized Reward Std', 'Cumulative Realized Reward Mean', 'Reputation', 'Cumulative Realized Reward Std', 'Instantaneous Mean Reward Mean', 'Instantaneous Mean Reward Std', 'Cumulative Mean Reward Mean', 'Cumulative Mean Reward Std', 'Best Arm Mean']
 simResults = {}
 
 
@@ -64,7 +66,7 @@ simResults = {}
 This runs N iterations of the experiment. In each iteration a set of true distributions is chosen and then for each algorithm we record the reward history. After this is done we aggregate the results and report them.
 '''
 
-with open('results/preliminary_raw_results/preliminary_plots_3_arms_heavy_tail.csv', 'w') as csvfile:
+with open('results/preliminary_raw_results/preliminary_plots_3_arms_reputation.csv', 'w') as csvfile:
   writer = csv.DictWriter(csvfile, fieldnames=FIELDNAMES)
   writer.writeheader()
   for (banditDistrName, banditDistr) in BANDIT_DISTR.iteritems():
@@ -99,7 +101,9 @@ with open('results/preliminary_raw_results/preliminary_plots_3_arms_heavy_tail.c
         instantaneous_realized = []
         cumulative_mean = []
         instantaneous_mean = []
[…]

### 2018-04-08 simulate.py

@@ -33,7 +33,8 @@ def simulate(principalAlg1, principalAlg2, agentAlg, K, T,
   alpha=DEFAULT_ALPHA,
   warmStartNumObservations=DEFAULT_WARM_START_NUM_OBSERVATIONS,
   principal1Priors=None,
-  principal2Priors=None
+  principal2Priors=None,
+  recordStatsAt=RECORD_STATS_AT
 ):
 
   if principal1Priors is None:
@@ -78,7 +79,7 @@ def simulate(principalAlg1, principalAlg2, agentAlg, K, T,
   principalHistory = []
   results = []
   for t in xrange(int(T)):
-    if t in RECORD_STATS_AT:
+    if t in recordStatsAt:
       results.append({
         'marketShare1' : principal1.n / float(t),
         'marketShare2' : principal2.n / float(t),
@@ -93,7 +94,7 @@ def simulate(principalAlg1, principalAlg2, agentAlg, K, T,
     (reward, arm) = principal.executeStep(t)
     trueMeanOfArm = banditProblemInstance.getMeanOfArm(arm)
     principal.regret += (bestArmMean - trueMeanOfArm)
-    agents.updateInformationSet(reward, arm, principalName)
+    agents.updateInformationSet(trueMeanOfArm, arm, principalName)
   
   return results
 

### 2018-04-08 tests/AgentsTest.py

@@ -1,0 +1,49 @@
+import unittest
+
+import sys 
+sys.path.append('..')
+from lib.agent.HardMax import HardMax
+from lib.agent.HardMaxWithRandom import HardMaxWithRandom
+from lib.InformationSet import InformationSet
+
+class TestAgents(unittest.TestCase):
+
+  def test_moving_average(self):
+    principals = { 'principal1': None, 'principal2': None }
+    info = InformationSet(principals, 3, memory=2, score='moving_average', discount_factor=0.99)
+    self.assertEqual(info.getScores().values(), [0, 0])
+    info.updateInformationSet(5, 1, 'principal1')
+    self.assertEqual(info.getMaxPrincipalsAndScores(), (['principal1'], 5.0))
+    info.updateInformationSet(3, 1, 'principal1')
+    self.assertEqual(info.getMaxPrincipalsAndScores(), (['principal1'], 4.0))
+    info.updateInformationSet(1, 1, 'principal1')
+    self.assertEqual(info.getMaxPrincipalsAndScores(), (['principal1'], 2.0))
+    info.updateInformationSet(-100, 1, 'principal1')
+    self.assertEqual(info.getMaxPrincipalsAndScores(), (['principal2'], 0.0))
+
+  def test_hard_max(self):
+    principals = { 'principal1': None, 'principal2': None }
+    agent = HardMax(principals, 3, memory=2)
+    agent.updateInformationSet(5, 1, 'principal1')
+    self.assertEqual(agent.selectPrincipal()[0], 'principal1')
+    agent.updateInformationSet(3, 1, 'principal1')
+    self.assertEqual(agent.selectPrincipal()[0], 'principal1')
+    agent.updateInformationSet(-8, 1, 'principal1')
+    self.assertEqual(agent.selectPrincipal()[0], 'principal2')
+    agent.resetInformationSet()
+    agent.updateInformationSet(1, 1, 'principal1')
+    self.assertEqual(agent.selectPrincipal()[0], 'principal1')
+
+  def test_hard_max_random(self):
+    principals = { 'principal1': None, 'principal2': None }
+    agent = HardMaxWithRandom(principals, 3, memory=2)
+    agent.updateInformationSet(5, 1, 'principal1')
+    N = 1000
+    c = 0
+    for i in xrange(N):
+      if agent.selectPrincipal()[0] == 'principal2':
+        c += 1
+    self.assertAlmostEqual(c, 25, delta=10)
+
+if __name__ == '__main__':
+    unittest.main()

### 2018-04-08 tests/BanditTest.py

@@ -1,0 +1,12 @@
+import unittest
+
+import sys 
+sys.path.append('..')
+from lib.BanditProblemInstance import BanditProblemInstance
+from lib.bandit.StaticGreedy import StaticGreedy
+from lib.bandit.DynamicGreedy import DynamicGreedy
+
+class TestBandits(unittest.TestCase):
+
+if __name__ == '__main__':
+    unittest.main()

### 2018-04-08 tests/SimulateTest.py

@@ -1,0 +1,48 @@
+import unittest
+
+import sys 
+sys.path.append('..')
+from lib.bandit.DynamicGreedy import DynamicGreedy
+from lib.bandit.StaticGreedy import StaticGreedy
+from lib.agent.HardMax import HardMax
+from lib.BanditProblemInstance import BanditProblemInstance
+from simulate import simulate
+
+from scipy.stats import bernoulli, beta
+
+class TestSimulate(unittest.TestCase):
+  def test_simple_simulate(self):
+    T = 2
+    K = 2
+    realizations = [[1, 1], [1, 1]]
+    heavy_tail_prior = beta(0.6, 0.6)
+    distributions = [bernoulli(heavy_tail_prior.rvs()) for i in xrange(K)]
+    N = 1000
+    total_static = 0
+    total_dynamic = 0
+    for i in xrange(N):
+      res = simulate(DynamicGreedy, DynamicGreedy, HardMax, K, T, realDistributions=distributions, realizations=realizations, warmStartNumObservations=0, recordStatsAt=[1])
+      res = simulate(StaticGreedy, StaticGreedy, HardMax, K, T, realDistributions=distributions, realizations=realizations, warmStartNumObservations=0, recordStatsAt=[1])
+      total_static += res[0]['marketShare1']
+      total_dynamic += res[0]['marketShare1']
+    self.assertAlmostEqual(total_static, N/2, delta=N/20)
+    self.assertAlmostEqual(total_dynamic, N/2, delta=N/20)
+
+  def test_free_obs(self):
+    T = 2
+    K = 2
+    realizations = [[1, 1], [1, 1]]
+    heavy_tail_prior = beta(0.6, 0.6)
+    distributions = [bernoulli(heavy_tail_prior.rvs()) for i in xrange(K)]
+    N = 500
+    total_dynamic = 0
+    for i in xrange(N):
+      res = simulate(DynamicGreedy, DynamicGreedy, HardMax, K, T, realDistributions=distributions, realizations=realizations, warmStartNumObservations=0, freeObsForP2=True,recordStatsAt=[1])
+      total_dynamic += res[0]['marketShare2']
+    self.assertEqual(total_dynamic, N)
+
+
+
+
+if __name__ == '__main__':
+    unittest.main()

### 2018-04-10 overview/overview.tex
commit overview

@@
 
-First, let's look at the preliminary simulation results on the instances we considered:
+First, let's look at the preliminary simulation results on the instances we considered.
+What is plotted here is the reputation score for a memory size of $100$ for each of the priors and each of the algorithms:
 
-\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Needle In Haystack 1 High 3 arms"} \\
-\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Uniform 3 arms"} \\
-\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Heavy Tail 3 arms"}
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Needle In Haystack High 3 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Uniform 3 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Heavy Tail 3 arms"}
 
@@
 
-\textbf{Question of Interest}: What algorithms should win under different agent response models?
-\textbf{Conjecture}: HardMaxWithRandom and SoftMax should eventually lead to the better algorithm winning. \\
+\textbf{Question of Interest}: What algorithms should win under different agent response models? \\
+\textbf{Conjecture}: HardMaxWithRandom and SoftMax should eventually lead to the better algorithm winning as long as, on the instance we are considering there is a gap in the long-term reward of the better algorithms.
+\\
 \textbf{Why?} Looking at the preliminary calibration plots, since there is some randomness in the choice rule of the agents, each principal should always be getting some free observations (our parameters such as the $\alpha$ in SoftMax and $\epsilon$ in HardMaxWithRandom tune how many expected observations a principal should get for a fixed $T$).
@@
 
-This seems to be confirmed in the simulations.
-Using the calibration from above ($\epsilon = 0.05$ for HMR and $\alpha = 30$ for SoftMax), the following are the results looking at $t = 2000$ and $t = 15000$ for each of the priors.
+\textbf{Conjecture}: Under HardMax, the early rounds should dictate almost everything since we suppose the algorithms start with little initial information and thus the difference between the algorithms should not make much of a difference.
+Since the rewards are randomly drawn, lucky early round draws may dictate the course of the game.
+Thus a conjecture is that the mean market share will be 50/50.
+
+The first conjecture seems to be confirmed in the simulations.
+Using the calibration from above ($\epsilon = 0.05$ for HMR and $\alpha = 30$ for SoftMax).
+However, the HardMax results require some more validation.
+First, we look at the results for low $T$ (1000) but high $N$ (1200).
+Then we extend the time horizon, but look at lower $N$.
+
+\includepdf[pages={-}]{low_time_high_num_sim}
 
@@
 
-On the next page are the results for large number of simulations ($N=1200$), but low $T$ \\
-
-\includepdf[pages={-}]{low_time_high_num_sim}
-
 The conjecture for HMR and SM seems to be consistent with the data generated from the simulation.
 However, what is going on with HardMax appears confusing.
-A conjecture is that since there is very little initial information, the algorithm choice matters very little and random noise in the early rounds seems to lead to one firm taking the market.
-To get some evidence for this, we re-run simulations but track market share over time.
+To get some more evidence for the original conjecture, we re-run simulations but track market share over time.
 In most cases, we get evidence of this and the vast majority of market shares over time look as follows:
 
-\includegraphics[scale=0.25]{decreasing_ms_over_time} \\
-\includegraphics[scale=0.25]{standard_ms_over_time} \\
+\includegraphics[scale=0.30]{decreasing_ms_over_time} \\
+\includegraphics[scale=0.30]{standard_ms_over_time} \\
 \vspace{0.25cm}
@@
 However, there are some simulations where we do not observe this. For instance, \\
-\includegraphics[scale=0.25]{odd_ms_over_time} \\
+\includegraphics[scale=0.30]{odd_ms_over_time} \\
 
@@
 As a comparison, what do the trajectories for the HardMaxWithRandom agent response functions look like?
-Haven't dug too deeply into the data for these simulations, but here is a sample one (ThompsonSampling vs DEG) that is similar to many in those that I have looked at: \\
+Haven't dug too deeply into the data for these simulations, but here is a sample one (ThompsonSampling vs DEG - plotted market share is for ThompsonSampling) that is similar to many in those that I have looked at: \\
 \includegraphics[scale=0.25]{hmr_over_time} \\
+\textbf{Puzzles}: Inconsistent with this interpretation of the HardMax results is the result for the Heavy Tail prior.
+Why does Thompson Sampling do so poorly!
+Looking at the reputation plots we see that in the early rounds there is very little difference between the algorithms but Thompson Sampling still does better, even if only slightly.
+Thus, it's kind of confusing as to what is actually going on.
+Perhaps worth looking more closely at some of the trajectory plots for this prior.\\
+
+In general, it seems that differences in algorithm performance in early rounds is important for determining results in HardMax and differences in algorithm performance for large $T$ is important for determining results in HardMaxWithRandom and SoftMax.
+
+Look at the performance on using the TRUE means:
+
+[insert plot]
 
@@
 
-Now, look at $X = 100$
-
-\includepdf[pages={-}]{less_free_obs_tournament_table}
-
+What next? Have re-run this for lower values of $X$ and higher $T$ (3000) and seen qualitatively similar results. Some potential follow-up questions of interest are:
+\begin{enumerate}
+\item Does reputation or information play a bigger role in the dominance of the incumbent?
+Can test this by re-initializing either reputation or information of the incumbent when the entrant wins the market (and re-running simulations of that).
+\item Even for HardMax it seems that the entrant gets a non-zero market share (the average seems to be between .05 and .25 usually).
+What do the market share trajectories look like?
+Are there a lot of simulations where the entrant simply gets zero and some where the entrant seems to get lucky upon entry and has an increasing market share over time?
+\item Run for a larger time horizon for the models with agent randomness to see if better algorithms win in the long-run.
+\end{enumerate}
 \end{document}

### 2018-04-10 overview/overview.tex
overview draft

@@
 
-\title{Overview}
+\title{Competing Bandits Overview}
 \maketitle
@@
 \textbf{Question of Interest}: What algorithms should win under different agent response models? \\
+\textbf{Overall Conjecture}: For the agent response functions with randomness (HMR and SoftMax), the algorithm performance in the later rounds of the single agent case should dictate which algorithm does better in the long-run in the competing bandits game.
+For the agent response functions without randomness (HM), the algorithm that does better in the very short-run of the single agent case should do better in the long-run in the competing bandits game.
+\\
 \textbf{Conjecture}: HardMaxWithRandom and SoftMax should eventually lead to the better algorithm winning as long as, on the instance we are considering there is a gap in the long-term reward of the better algorithms.
@@
 Then we extend the time horizon, but look at lower $N$.
+One annoying note, it is hard to compare SoftMax across these next two simulations since in the first $\alpha = 10$ and the in the second $\alpha = 30$
 
@@
 I've re-run this experiment for a memory size of 1000 and there are fewer cases which have odd switches as shown above.
-However, would need to look into this more if we think this is worth digging into more.\\
+However, would need to look into this more if we think this is worth digging into more.
+Another thing that might be interesting here is to ask what happens with HardMax if we get rid of any random noise in the reputation score and make it purely dependent on arm selection.
+In other words, what happens if the reputation score is constructed by using the true means of the selected arms.
+This of course is not a realistic model but may be useful for thinking about what is going on in this case.
+The results for this are reported in the appendix.\\
+
 As a comparison, what do the trajectories for the HardMaxWithRandom agent response functions look like?
@@
 
-In general, it seems that differences in algorithm performance in early rounds is important for determining results in HardMax and differences in algorithm performance for large $T$ is important for determining results in HardMaxWithRandom and SoftMax.
-
-Look at the performance on using the TRUE means:
-
-[insert plot]
+In general, the conjecture that differences in algorithm performance in early rounds is important for determining results in HardMax and differences in algorithm performance for large $T$ is important for determining results in HardMaxWithRandom and SoftMax seems to be roughly validated.
 
@@
 This seems to be the case in the reported simulations, both with large $X = 200$ and $X = 100$.
-Below are the results, the first for $X = 200$ and the second for $X = 100$.
+Below are the result for $X = 200$ ($X = 100$ is in the appendix figures)
 
@@
 \end{enumerate}
+
+\section*{Appendix}
+
+This contains some figures which are useful, but seem second-order to me.
+
+\subsection*{Preliminary Simulations}
+
+The plots for the realized rewards for each of the priors:
+
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Needle In Haystack 1 High 3 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Uniform 3 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Instantaneous Realized Reward Trajectory for Heavy Tail 3 arms"}
+
+\subsection*{Simultaneous Entry}
+
+The following is the result of running simulations where instead of the realized reward being utilized to form the reputation score, the TRUE mean of the selected arm was included in the reputation score.\\
+\includegraphics[scale=0.5]{"reputation_true_mean"}
+
+\subsection*{Asymmetric Entry}
+
+The following is the result of running the incumbent experiment with $X = 100$
+
+\includepdf[pages={-}]{less_free_obs_tournament_table}
+
 \end{document}

### 2018-05-08 analysis_code/agents_table.Rmd
add additional plots to the overview

@@
 #dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_large_horizon_3_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_multiple_haystack_2_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(2000, 15000)
+time_horizons <- c(1000, 15000)
 priors <- as.list(unique(dat['Prior']))$Prior

### 2018-05-08 analysis_code/tournament_table.Rmd
add additional plots to the overview

--- 2018-03-26 analysis_code/tournament_table.Rmd
@@
 
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/full_tournament_raw_results.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_test_raw.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_reputation_window_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_greedy_thompson_2_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(1000)
+time_horizons <- c(3000)
 priors <- as.list(unique(dat['Prior']))$Prior
@@
 ZERO_ONE_CUTOFF <- 0.1
-algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
+algs <- c("ThompsonSampling", "ThompsonStaticGreedy", "ThompsonDynamicGreedy")
+alg_names <- c("Thompson Sampling", "Thompson Static Greedy", "Thompson Dynamci Greedy")
 options(xtable.sanitize.text.function=identity)

### 2018-05-08 overview/overview.tex
add additional plots to the overview

@@
 \pagebreak
-\textbf{Simulation Procedure}
+\textbf{Competing Bandits Simulation Procedure}
 \begin{algorithm}
@@
 
-\section*{Results}
+\section*{Preliminary Calibration}
 
 First, let's look at the preliminary simulation results on the instances we considered.
-What is plotted here is the reputation score for a memory size of $100$ for each of the priors and each of the algorithms:
+In these simulations we simply run the selected bandit algorithm on the selected prior (no competition).
+What is plotted here is the reputation score for a memory size of $100$ for each of the priors and each of the algorithms (recall that reputation score is the average of the last 100 observations):
 
@@
 
-The first set of results we will consider are $X = 0$ and $k = 5$ so that there is simultaneous entry and only a small cold start. Recall to read the tables: \\
+One thing which may be problematic with these plots is that for Heavy Tail and Uniform the means for the instances are randomly drawn for each simulation but these plots just take averages over all of the simulations.
+
+Some additional preliminary calibration plots.
+These plots show the \% of simulations over $N=200$ that identified the best arm over time (based on looking at the posterior distributions at a given time): \\
+\includegraphics[scale=0.5]{"needle_in_haystack_best_arm_ident"} \\
+\includegraphics[scale=0.5]{"heavy_tail_best_arm_ident"} \\
+\includegraphics[scale=0.5]{"uniform_best_arm_identification"}
+
+\section*{Simultaneous Entry}
+
+The first set of results we will consider are $X = 0$ and $k = 5$ so that there is simultaneous entry. Recall to read the tables: \\
 \begin{enumerate}
@@
 \end{enumerate}
-
-\section*{Simultaneous Entry}
 
@@
 
-The conjecture for HMR and SM seems to be consistent with the data generated from the simulation.
+The conjecture for HMR and SM seems to be consistent with the data generated from the simulation (the one exception seems to be HeavyTail for BG vs BEG - more on this later).
 However, what is going on with HardMax appears confusing.
@@
 Thus, it's kind of confusing as to what is actually going on.
-Perhaps worth looking more closely at some of the trajectory plots for this prior.\\
+Perhaps worth looking more closely at some of the trajectory plots for this prior.
+As well, Bayesian Greedy seems to beat Bayesian Epsilon Greedy on this prior even when there is randomness in the agent response model which seems puzzling given the preliminary plots.
+\\
 
-In general, the conjecture that differences in algorithm performance in early rounds is important for determining results in HardMax and differences in algorithm performance for large $T$ is important for determining results in HardMaxWithRandom and SoftMax seems to be roughly validated.
+In general, the conjecture that differences in algorithm performance in early rounds is important for determining results in HardMax and differences in algorithm performance for large $T$ is important for determining results in HardMaxWithRandom and SoftMax seems to be roughly validated (but there are some cases we need to think more about what is happening).
 

### 2018-05-08 analysis_code/preliminary_2.R

@@ -1,0 +1,32 @@
+library(dplyr)
+library(ggplot2)
+
+best_arm <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_best_arm.csv")
+
+algorithms <- unique(best_arm$Algorithm)
+dists <- unique(best_arm$Distribution)
+times <- unique(best_arm$t)
+dat <- data.frame(matrix(vector(), 0, 4,
+                                   dimnames=list(c(), c("Algorithm", "Distribution", "t", "mean_correct"))),
+                            stringsAsFactors=F)
+count <- 1
+for (dist in dists) {
+  filtered_dist <- filter(best_arm, Distribution == UQ(dist))
+  for (algorithm in algorithms) {
+    for (time in times) {
+      cur <- filter(filtered_dist, Algorithm == UQ(algorithm) & t == UQ(time))
+      avg_correct <- mean(as.numeric(cur$Best.Arm.Identification) - 1)
+      item <- c(algorithm, dist, time, avg_correct)
+      dat[count, ] <- item
+      count <- count + 1
+    }
+  }
+}
+dat$t <- as.numeric(dat$t)
+dat$mean_correct <- as.numeric(dat$mean_correct)
+
+for (dist in dists ) {
+  d <- filter(dat, Distribution == UQ(dist))
+  p <- ggplot(d, aes(x=t, y=mean_correct)) + geom_line(aes(col=Algorithm)) + labs(title=dist) + ylab("% that Identified Best Arm")
+  print(p)
+}

### 2018-05-08 analysis_code/preliminary_best_arm.R

@@ -1,0 +1,1 @@
+res <- read.csv("/Users/garidor/Desktop/bandits-rl-projects/results/preliminary_raw_results/preliminary_plots_best_arm.csv")

### 2018-05-08 analysis_code/preliminary_plots.R

@@ -10,20 +10,20 @@ WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
 #dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_20_arms.csv", sep=""))
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_long.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_reputation_2.csv", sep=""))
 #dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_3.csv", sep=""))
 #dat <- rbind(dat, dat2)
 dat$Algorithm <- replace(as.character(dat$Algorithm), dat$Algorithm == "DynamicEpsilonGreedy", "DynamicEpsilonGreedy, 0.05")
 dat$Algorithm <- replace(as.character(dat$Algorithm), dat$Algorithm == "NonBayesianEpsilonGreedy", "NonBayesianEpsilonGreedy, 0.05")
 algs <- as.list(unique(dat['Algorithm']))$Algorithm
-filter_algs <- c("ThompsonSampling", "UCB1WithConstantOne", "NonBayesianEpsilonGreedy, 0.05", "DynamicEpsilonGreedy, 0.05", "DynamicGreedy")
+filter_algs <- c("ThompsonSampling", "UCB1WithConstantOne", "DynamicEpsilonGreedy, 0.05", "DynamicGreedy")
 dat <- filter(dat, Algorithm %in% filter_algs)
 dists <- as.list(unique(dat['Distribution']))$Distribution
 
 filter_by_dist_and_plot <- function(dist) {
   d <- filter(dat, Distribution == dist)
   title <- paste("Reputation Trajectory for", dist, "3 arms")
-  q <- ggplot(data=d, aes(x=t, y=Reputation, colour=Algorithm)) + ggtitle(title) + ylab("Reputation") + xlab("time") +
+  q <- ggplot(data=d, aes(x=t, y=Realized.Reputation, colour=Algorithm)) + ggtitle(title) + ylab("Reputation") + xlab("time") +
     #geom_path() # just plot the raw trajectory
     geom_smooth(method="loess") #+ smooths the trajectory 
[…]

### 2018-05-27 analysis_code/agents_table.Rmd
commit a bunch of stuff

--- 2018-03-26 analysis_code/agents_table.Rmd
@@
 #dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_exp_raw.csv", sep=""))
-p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
-p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
-agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
-memory_sizes <- sort(as.list(unique(dat['Memory.Size']))$Memory.Size, decreasing=TRUE)
-time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(1000)
-priors <- as.list(unique(dat['Prior']))$Prior
+#dat_3 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_large_horizon_3_raw.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_many_arms_haystack_raw.csv", sep=""))
+#dat_2 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_many_arms_2_raw.csv", sep=""))
+#dat_10 <- rbind(dat, dat_2)
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_heavy_tail_many_sim_raw.csv", sep=""))
+dat_3 <- dat
+
+p1_algs <- as.list(unique(dat_3['P1.Alg']))$P1.Alg
+p2_algs <- as.list(unique(dat_3['P2.Alg']))$P2.Alg
+agent_algs <- as.list(unique(dat_3['Agent.Alg']))$Agent.Alg
+memory_sizes <- sort(as.list(unique(dat_3['Memory.Size']))$Memory.Size, decreasing=TRUE)
+time_horizons <-  as.list(unique(dat_3['Time.Horizon']))$Time.Horizon
+priors <- as.list(unique(dat_3['Prior']))$Prior
 
@@
     rownames(results) <- get_agent_rep(agent_algs)
+  for (q in 1:1) {
     for (i in 1:length(agent_algs)) {
+
       agent_alg <- agent_algs[i]
@@
           p2alg = alg_pairs[k+1]
+          if (q == 1) {
+            K <- 3
             filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == memory & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+          } else {
+            K <- 10
+            filtered_dat <- filter(dat_10, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == memory & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+          }
           if (nrow(filtered_dat) == 0) {
@@
     }
-      tab <-xtable(results, caption=paste("Results for", "t=",time, prior, "Memory=", memory), row.header='Entrant', col.header='Incumbent')
+      tab <-xtable(results, caption=paste("Results for", "t=",time, prior, "Memory=", memory, "K=", K), row.header='Entrant', col.header='Incumbent')
       print(tab, type="html")
+    }
       }

### 2018-05-27 analysis_code/reputation_vs_information.Rmd
commit a bunch of stuff

 <style>
   th,td{
     padding:10px;
   }
 </style>

```{r echo=FALSE,message=FALSE,xtable, results="asis"}
library(dplyr)
library(knitr)
WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"

#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_reputation_window_raw.csv", sep=""))
dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
erase_rep_dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_erase_reputation_raw.csv", sep=""))
erase_info_dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_erase_information_raw.csv", sep=""))
p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
memory_sizes <- sort(as.list(unique(dat['Memory.Size']))$Memory.Size, decreasing=TRUE)
agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
time_horizons <- c(1000)
priors <- as.list(unique(dat['Prior']))$Prior

ZERO_ONE_CUTOFF <- 0.1
algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
alg_names <- c("Thompson Sampling", "Dynamic Epsilon Greedy", " DynamicGreedy")
options(xtable.sanitize.text.function=identity)
options(xtable.caption.placement="top")
library(xtable)
for (j in 1:length(time_horizons)) {
  for (prior in priors) {
    for (i in 1:length(agent_algs)) {
      agent_alg <- agent_algs[i]
      print_breaks <- FALSE
      for (q in 1:3) {
        time <- time_horizons[j]
        results <- matrix(nrow=length(algs), ncol=length(algs))
        colnames(results) <- alg_names
        rownames(results) <- alg_names
        for (k in 1:length(algs)) {
          for (l in 1:length(algs)) {
            p1alg = algs[k]
            p2alg = algs[l]
            filtered_dat <- NULL
            title <- NULL
            if (q == 1) {
              title <- "Info + Rep"
              filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
            } else if (q == 2) {
              title <- "Info"
              filtered_dat <- filter(erase_rep_dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
            } else if (q == 3) {
              title <- "Rep"
              filtered_dat <- filter(erase_info_dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
            }
            if (nrow(filtered_dat) == 0) {
              if (p1alg == p2alg) { print_breaks <- TRUE }
              next
            }
            share <- filtered_dat$Market.Share.for.P1
            test <- t.test(share)
            cin <- signif(test$conf.int[2] - mean(share), digits=2)
            shares <- sum(share >= (1 - ZERO_ONE_CUTOFF) | share <= ZERO_ONE_CUTOFF, na.rm=TRUE) / nrow(filtered_dat)
            cv_high <- qchisq(.975, df=length(share) - 1)
            cv_low <- qchisq(.025, df=length(share) - 1)
            lower_var <- signif((length(share) - 1) * var(share) / cv_high, digits=2)
            high_var <- signif((length(share) - 1) * var(share) / cv_low, digits=2)
            results[k, l] <- paste("<b>", signif(mean(share), digits=2), "</b> +/-", cin, "<br>", signif(var(share), digits=2), paste("(", lower_var, ", ", high_var, ")", sep=""), "<br>", "Extreme Shares:", 100*signif(shares, digits=2), "%")
          }
        }
        tab <-xtable(results, caption=paste("Results for", agent_alg, "t =", time, prior, title))
        print(tab, type="html")
        if (print_breaks) { print("<br>") }
      }
    }
  }
}

```

### 2018-05-27 new_overview/new_experiments.tex
commit a bunch of stuff

\documentclass[11pt,letterpaper]{article}
\usepackage{fullpage}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{algorithm,algpseudocode}
\usepackage{pdfpages}

\graphicspath{{../results}}

\begin{document}

\title{New Experiments}
\maketitle

\section*{More Complex Instances}
We consider the same instances as before, but we consider $K = 10$ instead of $K = 3$. Previously we had conjectured that increasing $K$ only changes the time it takes to learn, but now we want to test if difficulty to learn the best arm can impact what happens in competition. Why would we think this should happen? There are two potentially counteracting effects one can think of:
\begin{enumerate}
\item Making the learning problem harder can benefit better algorithms in the long-run since they'll be more likely to identify the best arm than the worse algorithms and thus we expect there to be a bigger difference in the reputation scores between a good algorithm and a bad algorithm
\item Making the learning problem harder also forces the better algorithms to engage in more sub-optimal exploration which could have perverse short-run effects on reputation.
\end{enumerate}

We summarize the results for each instance and comment on the difference from the $K=3$ case:
\begin{enumerate}
\item Needle in Haystack:
\begin{enumerate}
\item HMR / SM: The results are roughly the same. $TS > DEG > DG$
\item HM: In $K=3$, we had that the results were effectively 50/50 between TS and DEG/DG. However, with $K=10$, we see that TS wins by a lot.
\end{enumerate}
\item Uniform:
\begin{enumerate}
\item HMR/SM: The results are for the most part qualitatively the same. The one difference is that for $K = 10$ we see that $DEG > DG$ slightly more clearly.
\item HM: The results are also qualitatively similar, with no algorithm clearly dominating.
\end{enumerate}
\item Heavy Tail:
\begin{enumerate}
\item HMR/SM: The results are starker than with $K = 3$ arms with TS dominating DEG and DG by a lot. However, we no longer see that $DG > DEG$ as we did in $K = 3$
\item HM: The results here are starker than with $K = 3$ arms, with DEG and DG dominating TS more than in the $K = 3$ case. 
\end{enumerate}
\end{enumerate}
\includepdf[pages={-}]{many_arms_vs_few_arms}

\section*{Information vs Reputation - Incumbent}

The following tables contain three separate runs of the incumbent experiment (the realizations are different each time, but we should change this in future runs). The first is the one we have looked at before where one of the principals gets 200 free observations. In the second we artificially eliminate the reputation gain of the incumbent upon the entrant entering. In the third we artificially eliminate the information gain of the incumbent upon the entrant entering. Thus, the overall goal here is to ask what the effect reputation and information have individually as barriers to entry for learning.

Overall, the results are as follows:
\begin{enumerate}
\item The effect of reputation in HardMax seems to be relatively large for the incumbent. Looking at the Info table with the HardMax response function shows that the entrant gets a significant amount of the market back when the reputation gain is erased. When the information gain is erased, with HardMax the incumbent doesn't lose much market share. It appears somewhat similar in HardMaxWithRandom, but the difference is not so stark.
\item I am not sure what to make of the SoftMax case. It seems that losing reputation and information can actually \textit{benefit} the incumbent. I can't think of why this would make any sense, so I'm going to rerun these again using the same realizations. I did not record the parametrization for SoftMax during each of the runs, so it is possible that this is why we see these results. As a result, take the SoftMax results with a grain of salt until I re-run this.
\end{enumerate}

\includepdf[pages={-}]{reputation_vs_information}

\end{document}

### 2018-05-27 lib/bandit/ThompsonDynamicGreedy.py

@@ -7,7 +7,7 @@ class ThompsonDynamicGreedy(BanditAlgorithm):
     return 'TSDG'
 
   def __init__(self, banditProblemInstance, priors, distr='beta'):
-    super(ExploreThenExploit, self).__init__(banditProblemInstance, priors, distr)
+    super(ThompsonDynamicGreedy, self).__init__(banditProblemInstance, priors, distr)
     self.greedy = False
 
   def switchAlgorithm(self):

### 2018-05-27 lib/bandit/ThompsonStaticGreedy.py

@@ -1,14 +1,14 @@
 import numpy as np
 from BanditAlgorithm import BanditAlgorithm
-from deepcopy import deepcopy
+from copy import deepcopy
 
-class ThompsonDynamicGreedy(BanditAlgorithm):
+class ThompsonStaticGreedy(BanditAlgorithm):
   @staticmethod
   def shorthand():
-    return 'TSDG'
+    return 'TSSG'
 
   def __init__(self, banditProblemInstance, priors, distr='beta'):
-    super(ExploreThenExploit, self).__init__(banditProblemInstance, priors, distr)
+    super(ThompsonStaticGreedy, self).__init__(banditProblemInstance, priors, distr)
     self.greedy = False
 
   def switchAlgorithm(self):

### 2018-05-27 main.py

@@ -15,6 +15,7 @@ from lib.agent.SoftMax import SoftMax
 from lib.agent.HardMaxWithRandom import HardMaxWithRandom
 
 # library imports
+from joblib import Parallel, delayed
 from scipy.stats import bernoulli, beta
 from copy import copy, deepcopy
 from collections import Counter
@@ -24,12 +25,14 @@ import pickle
 import random
 
 K = 3
-T = 15002
-NUM_SIMULATIONS = 150
+T = 5002
+NUM_SIMULATIONS = 100
 
 FREE_OBS = False
 FREE_OBS_NUM = 100
-exp_name = 'large_horizon_2'
+exp_name = 'test_2'
+REALIZATIONS_NAME = '' #if you want to pull in past realizations, fill this in with the realizations base name
+numCores = 1
 
 AGENT_ALGS = [HardMax, HardMaxWithRandom, SoftMax]
 
@@ -48,26 +51,52 @@ needle_in_haystack_50_medium[int(K/2)] = bernoulli(default_mean + 0.05)
 heavy_tail_prior = beta(0.6, 0.6)
 
 BANDIT_DISTR = {
-  'Needle In Haystack High': needle_in_haystack_50_high
+  'Uniform': None,
+  'Needle In Haystack High': needle_in_haystack_50_high,
 }
 
-WORKING_DIRECTORY = '/rigel/home/ga2449/bandits-rl-project/'
-#WORKING_DIRECTORY = ''
+#WORKING_DIRECTORY = '/rigel/home/ga2449/bandits-rl-project/'
+WORKING_DIRECTORY = ''
 
 if FREE_OBS:
   dir_name = WORKING_DIRECTORY + 'results/free_obs_raw_results/'
-  base_name = dir_name + 'free_obs_experiment_' + exp_name
+  base_name = dir_name + 'free_obs_experiment_'
 else:
   dir_name = WORKING_DIRECTORY + 'results/tournament_raw_results/'
-  base_name = dir_name + 'tournament_experiment_' + exp_name
+  base_name = dir_name + 'tournament_experiment_'
 
-aggregate_name = base_name + '_aggregate.csv'
-raw_name = base_name + '_raw.csv'
-realizations_name = base_name + '_realizations.csv'
-dist_name = base_name + '_dist.csv'
+exp_base_name = base_name + exp_name
+aggregate_name = exp_base_name + '_aggregate.csv'
+raw_name = exp_base_name + '_raw.csv'
[…]

### 2018-05-27 simulate.py

@@ -36,7 +36,7 @@ def simulate(principalAlg1, principalAlg2, agentAlg, K, T,
   principal2Priors=None,
   recordStatsAt=RECORD_STATS_AT
 ):
-
+  np.random.seed(None)
   if principal1Priors is None:
     principal1Priors = getDefaultPrior(K)
   if principal2Priors is None:

### 2018-05-29 overview/overview.tex
commit a bunch of stuff
updated overview

@@
 
-Considered $K = 3$, $Memory = 100$ (unless otherwise noted) \\
+Considered $K = 3, 10$, $Memory = 100$ (unless otherwise noted) \\
 \textbf{The Bandit priors that were considered}:
@@
 \begin{enumerate}
-\item High - 2 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
+\item High - $K-1$ arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
 \end{enumerate}
@@
 \item DynamicGreedy with priors of $Beta(1, 1)$ for every arm
-\item Bayesian Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm and $\epsilon=0.05$
+\item Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm and $\epsilon=0.05$
 \end{enumerate}
@@
 			\For{$N$ simulations}
-
+				\State Give principal 2 $X$ free observations (the agents also get these observations)				
 				\State Give the agents $k$ observations from each principal
-				\State Give principal 2 $X$ free observations (the agents also get these observations)
 				\State Run simulation for T periods
@@
 
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Needle In Haystack High 10 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Uniform 10 arms"} \\
+\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Heavy Tail 10 arms"}
+
 One thing which may be problematic with these plots is that for Heavy Tail and Uniform the means for the instances are randomly drawn for each simulation but these plots just take averages over all of the simulations.
@@
 These plots show the \% of simulations over $N=200$ that identified the best arm over time (based on looking at the posterior distributions at a given time): \\
-\includegraphics[scale=0.5]{"needle_in_haystack_best_arm_ident"} \\
-\includegraphics[scale=0.5]{"heavy_tail_best_arm_ident"} \\
-\includegraphics[scale=0.5]{"uniform_best_arm_identification"}
+\includegraphics[scale=0.5]{"preliminary_nih_3_arms"} \\
+\includegraphics[scale=0.5]{"preliminary_best_arm_heavy_tail_3"}
+\includegraphics[scale=0.5]{"uniform_best_arm_identification_3_arms"} \\
+\includegraphics[scale=0.5]{preliminary_nih_10_arms} \\
+\includegraphics[scale=0.5]{preliminary_best_arm_uniform_10} 
+\includegraphics[scale=0.5]{preliminary_best_arm_heavy_tail_10}
 
@@
 
+However, this is not always the case.
+For Heavy Tail with $K = 3$ we have that $DG > DEG$.
+This does not hold however for $K = 10$.
+This is possibly because learning is ``easy" for DG in $K = 3$ (it identifies the best arm very quickly and in many cases - see preliminary calibration plots).
+As well, since DEG engages in non-adaptive exploration and Heavy Tail possibly has some very bad arms (i.e.
+means close to 0), this non-adaptive exploration could have large effects on reputation.
+Learning appears to be difficult when $K = 10$ and the reputation gap appears to be slightly higher between DEG and DG but it is unclear what leads to the results we see.
+
 \textbf{Conjecture}: Under HardMax, the early rounds should dictate almost everything since we suppose the algorithms start with little initial information and thus the difference between the algorithms should not make much of a difference.
@@
 
-The first conjecture seems to be confirmed in the simulations.
-Using the calibration from above ($\epsilon = 0.05$ for HMR and $\alpha = 30$ for SoftMax).
-However, the HardMax results require some more validation.
-First, we look at the results for low $T$ (1000) but high $N$ (1200).
-Then we extend the time horizon, but look at lower $N$.
-One annoying note, it is hard to compare SoftMax across these next two simulations since in the first $\alpha = 10$ and the in the second $\alpha = 30$
+This also does not hold. See the puzzles section after the results.
 
-\includepdf[pages={-}]{low_time_high_num_sim}
+\includepdf[pages={-}]{many_arms_vs_few_arms}
 
-\includepdf[pages={-}]{tournament_results}
+\textbf{Puzzles}: There are two puzzles with respect to HardMax.
+Why does Needle in a Haystack ($K = 10$) lead to $TS > DG$ and why does Heavy Tail (both $K = 3$ and $K = 10$) lead to $DG, DEG > TS$?
+Clearly the original conjecture is flawed in some way.
+What are things that stand out?
+Under Heavy Tail we see that the difference in algorithm performance without competition is not that large (looking at reputation plots) and that $DG$ identifies the best arm in a large fraction of runs ($\approx 0.8$).
+Under Needle in a Haystack we see that the difference between algorithm performance is large and that TS identifies the best arm with only a few samples, but what in particular leads it to win under competition with the HardMax response function?
 
-The conjecture for HMR and SM seems to be consistent with the data generated from the simulation (the one exception seems to be HeavyTail for BG vs BEG - more on this later).
-However, what is going on with HardMax appears confusing.
-To get some more evidence for the original conjecture, we re-run simulations but track market share over time.
-In most cases, we get evidence of this and the vast majority of market shares over time look as follows:
-
-\includegraphics[scale=0.30]{decreasing_ms_over_time} \\
-\includegraphics[scale=0.30]{standard_ms_over_time} \\
-\vspace{0.25cm}
-
-However, there are some simulations where we do not observe this. For instance, \\
-\includegraphics[scale=0.30]{odd_ms_over_time} \\
-
-I am not sure what happens in these cases, but one conjecture is that these arise from random noise from a low memory size (memory here is 100, as we have been using in our experiment).
-I've re-run this experiment for a memory size of 1000 and there are fewer cases which have odd switches as shown above.
-However, would need to look into this more if we think this is worth digging into more.
-Another thing that might be interesting here is to ask what happens with HardMax if we get rid of any random noise in the reputation score and make it purely dependent on arm selection.
-In other words, what happens if the reputation score is constructed by using the true means of the selected arms.
-This of course is not a realistic model but may be useful for thinking about what is going on in this case.
-The results for this are reported in the appendix.\\
-
-As a comparison, what do the trajectories for the HardMaxWithRandom agent response functions look like?
-Haven't dug too deeply into the data for these simulations, but here is a sample one (ThompsonSampling vs DEG - plotted market share is for ThompsonSampling) that is similar to many in those that I have looked at: \\
-\includegraphics[scale=0.25]{hmr_over_time} \\
-\textbf{Puzzles}: Inconsistent with this interpretation of the HardMax results is the result for the Heavy Tail prior.
-Why does Thompson Sampling do so poorly!
-Looking at the reputation plots we see that in the early rounds there is very little difference between the algorithms but Thompson Sampling still does better, even if only slightly.
-Thus, it's kind of confusing as to what is actually going on.
-Perhaps worth looking more closely at some of the trajectory plots for this prior.
-As well, Bayesian Greedy seems to beat Bayesian Epsilon Greedy on this prior even when there is randomness in the agent response model which seems puzzling given the preliminary plots.
-\\
-
-In general, the conjecture that differences in algorithm performance in early rounds is important for determining results in HardMax and differences in algorithm performance for large $T$ is important for determining results in HardMaxWithRandom and SoftMax seems to be roughly validated (but there are some cases we need to think more about what is happening).
+In general, some properties of the instances and the algorithms seem to drive who wins when.
+We do not have consistency as in the single instance case where $TS > DEG > DG$, but it's unclear under what conditions we have who wins.
+Should think more about the properties of the instances.
 
@@
 However, the time horizon, in contrast to the simultaneous entry case, seems like it may be way longer!
-So far, most of the simulations have simply looked at $t = 1000$, but it may make sense to look at longer time horizons due to the conjecture from above.
+So far, most of the simulations have simply looked at $t = 2000$, but it may make sense to look at longer time horizons due to the conjecture from above.
 \\
@@
 
-\includepdf[pages={-}]{free_obs_tournament}
+In this version of the simulations we additionally have simulations where the information, reputation, or both are artificially erased from the incumbent. The timing of events is as follows:
+\begin{enumerate}
+\item The incumbent gets $X$ free observations
+\item Either information gains are erased (the posterior beliefs of the incumbent are reset to the fake priors) or the reputation gains are erased (the agents' reputation score for the incumbent is reset) or both
+\item Both the incumbent and the entrant get $k=5$ free observations
+\end{enumerate}
 
-What next? Have re-run this for lower values of $X$ and higher $T$ (3000) and seen qualitatively similar results. Some potential follow-up questions of interest are:
+\includepdf[pages={-}]{reputation_vs_information}
+
+What next? Have re-run this for lower values of $X$ and higher $T$ (3000) and seen qualitatively similar results (see Appendix). Some potential follow-up questions of interest are:
 \begin{enumerate}
-\item Does reputation or information play a bigger role in the dominance of the incumbent?
-Can test this by re-initializing either reputation or information of the incumbent when the entrant wins the market (and re-running simulations of that).
 \item Even for HardMax it seems that the entrant gets a non-zero market share (the average seems to be between .05 and .25 usually).
@@
 \item Run for a larger time horizon for the models with agent randomness to see if better algorithms win in the long-run.
+\item What results are surprising?
 \end{enumerate}

### 2018-06-07 new_overview/new_experiments.tex
more experiments

@@
 
-\section*{More Complex Instances}
-We consider the same instances as before, but we consider $K = 10$ instead of $K = 3$.
-Previously we had conjectured that increasing $K$ only changes the time it takes to learn, but now we want to test if difficulty to learn the best arm can impact what happens in competition.
-Why would we think this should happen?
-There are two potentially counteracting effects one can think of:
-\begin{enumerate}
-\item Making the learning problem harder can benefit better algorithms in the long-run since they'll be more likely to identify the best arm than the worse algorithms and thus we expect there to be a bigger difference in the reputation scores between a good algorithm and a bad algorithm
-\item Making the learning problem harder also forces the better algorithms to engage in more sub-optimal exploration which could have perverse short-run effects on reputation.
-\end{enumerate}
+\section*{.5/.7 Instance}
 
-We summarize the results for each instance and comment on the difference from the $K=3$ case:
-\begin{enumerate}
-\item Needle in Haystack:
-\begin{enumerate}
-\item HMR / SM: The results are roughly the same. $TS > DEG > DG$
-\item HM: In $K=3$, we had that the results were effectively 50/50 between TS and DEG/DG. However, with $K=10$, we see that TS wins by a lot.
-\end{enumerate}
-\item Uniform:
-\begin{enumerate}
-\item HMR/SM: The results are for the most part qualitatively the same. The one difference is that for $K = 10$ we see that $DEG > DG$ slightly more clearly.
-\item HM: The results are also qualitatively similar, with no algorithm clearly dominating.
-\end{enumerate}
-\item Heavy Tail:
-\begin{enumerate}
-\item HMR/SM: The results are starker than with $K = 3$ arms with TS dominating DEG and DG by a lot. However, we no longer see that $DG > DEG$ as we did in $K = 3$
-\item HM: The results here are starker than with $K = 3$ arms, with DEG and DG dominating TS more than in the $K = 3$ case. 
-\end{enumerate}
-\end{enumerate}
-\includepdf[pages={-}]{many_arms_vs_few_arms}
+plots plots
 
-\section*{Information vs Reputation - Incumbent}
+\section*{Vary Warm Start + Effective end of game}
 
-The following tables contain three separate runs of the incumbent experiment (the realizations are different each time, but we should change this in future runs).
-The first is the one we have looked at before where one of the principals gets 200 free observations.
-In the second we artificially eliminate the reputation gain of the incumbent upon the entrant entering.
-In the third we artificially eliminate the information gain of the incumbent upon the entrant entering.
-Thus, the overall goal here is to ask what the effect reputation and information have individually as barriers to entry for learning.
+Here we vary
+\includepdf[pages={-}]{eeog_old_priors}
 
-Overall, the results are as follows:
-\begin{enumerate}
-\item The effect of reputation in HardMax seems to be relatively large for the incumbent.
-Looking at the Info table with the HardMax response function shows that the entrant gets a significant amount of the market back when the reputation gain is erased.
-When the information gain is erased, with HardMax the incumbent doesn't lose much market share.
-It appears somewhat similar in HardMaxWithRandom, but the difference is not so stark.
-\item I am not sure what to make of the SoftMax case.
-It seems that losing reputation and information can actually \textit{benefit} the incumbent.
-I can't think of why this would make any sense, so I'm going to rerun these again using the same realizations.
-I did not record the parametrization for SoftMax during each of the runs, so it is possible that this is why we see these results.
-As a result, take the SoftMax results with a grain of salt until I re-run this.
-\end{enumerate}
+\section*{Various Needles in a Haystack}
 
-\includepdf[pages={-}]{reputation_vs_information}
+\includepdf[pages={-}]{haystack_eeog}
 

### 2018-06-11 new_overview/new_experiments.tex
commit new results

@@
 
+\section*{Vary Warm Start + Effective end of game}
+
+Here we vary the size of the warm start (the number of observations given to the principals before the competing bandits game begins).
+We also track a new quantity - the ``effective end of game" which we define as the last round where there is a ``switch" in the decision of the agent.
+So, for instance, if in $t-1$ the agent picks principal 1 and in $t$ the agent picks principal 2 then this is a ``switch." The last time this happens in a simulation is defined as effective end of game (EEOG).
+
+Below we report the results of running on the same priors we have been considering, but varying the warm start and recording market share + EEOG.
+\includepdf[pages={-}]{eeog_old_priors}
+
 \section*{.5/.7 Instance}
 
-plots plots
+This is an instance where the arms are uniformly either 0.5 or 0.7 (as opposed to Needle in a Haystack where 9 arms have mean 0.5 and 1 arm has mean 0.7).
 
-\section*{Vary Warm Start + Effective end of game}
+This is what the preliminary calibration for this instance looks like:
 
-Here we vary
-\includepdf[pages={-}]{eeog_old_priors}
+\includegraphics[scale=0.5]{"Reputation Trajectory for 5_7"}
+
+The results of the simulations:
+
+\includepdf[pages={-}]{5_7_results}
 
 \section*{Various Needles in a Haystack}
+In this set of experiments we simply shift the needle in a haystack distribution.
+
+We consider the following instances:
+\begin{enumerate}
+\item Needle In Haystack 0.1 - 9 arms with mean 0.1, 1 arm with mean 0.3
+\item Needle In Haystack 0.3 - 9 arms with mean 0.3, 1 arm with mean 0.5
+\item Needle In Haystack 0.5 - 9 arms with mean 0.5, 1 arm with mean 0.7
+\item Needle In Haystack 0.7 - 9 arms with mean 0.5, 1 arm with mean 0.7
+\end{enumerate}
+
+Preliminary plots for 0.1 and 0.3: \\
+\includegraphics[scale=0.4]{"./Reputation Trajectory for Needle In Haystack 1 10 arms"} \\
+\includegraphics[scale=0.4]{"./Reputation Trajectory for Needle In Haystack 3 10 arms"}
 

### 2018-06-27 analysis_code/agents_table.Rmd
commit new results

@@
 #dat_10 <- rbind(dat, dat_2)
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_heavy_tail_many_sim_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ws_small_erase_rep_raw.csv", sep=""))
 dat_3 <- dat
@@
 agent_algs <- as.list(unique(dat_3['Agent.Alg']))$Agent.Alg
-memory_sizes <- sort(as.list(unique(dat_3['Memory.Size']))$Memory.Size, decreasing=TRUE)
 time_horizons <-  as.list(unique(dat_3['Time.Horizon']))$Time.Horizon
+time_horizons <- c(5000)
 priors <- as.list(unique(dat_3['Prior']))$Prior
+warm_starts <-  sort(as.list(unique(dat_3['Warm.Start']))$Warm.Start, decreasing=FALSE)
 
@@
 library(xtable)
-for (memory in memory_sizes) {
 for (prior in priors) {
@@
   for (q in 1:1) {
+    for (start in warm_starts) {
     for (i in 1:length(agent_algs)) {
@@
           p2alg = alg_pairs[k+1]
+          K <- 10
           if (q == 1) {
-            K <- 3
-            filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == memory & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+            filtered_dat <- filter(dat, P1.Alg == p1alg & Warm.Start == start & P2.Alg == p2alg & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
           } else {
-            K <- 10
-            filtered_dat <- filter(dat_10, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == memory & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+            filtered_dat <- filter(dat_10, P1.Alg == p1alg & Warm.Start == start & P2.Alg == p2alg & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
           }
@@
       }
-      tab <-xtable(results, caption=paste("Results for", "t=",time, prior, "Memory=", memory, "K=", K), row.header='Entrant', col.header='Incumbent')
+      tab <-xtable(results, caption=paste("Results for", "t=",time, prior, "Warm Start=", start, "K=", K), row.header='Entrant', col.header='Incumbent')
       print(tab, type="html")

### 2018-06-27 analysis_code/eeog.Rmd
commit new results

 <style>
   th,td{
     padding:10px;
   }
 </style>

```{r echo=FALSE, message=FALSE, xtable, results="asis"}

library(dplyr)
library(knitr)
library(xtable)

options(xtable.sanitize.text.function=identity)
options(xtable.caption.placement="top")

concise_alg_rep <- function(alg) {
  if (alg == "ThompsonSampling") {
    return("TS")
  } else if (alg == "DynamicEpsilonGreedy") {
    return("DEG")
  } else if (alg == "DynamicGreedy") {
    return("DG")
  }
}

WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/"
f <- "tournament_experiment_ws_small_erase_rep_raw.csv"
dat <- read.csv(file=paste(WORKING_PATH, f, sep=""))
#dat_2 <- read.csv(file=paste(WORKING_PATH, "tournament_experiment_tourn_.5_.7_raw.csv", sep=""))
#dat <- rbind(dat, dat_2)
dat <- filter(dat, Time.Horizon == 5000)
priors <- as.list(unique(dat['Prior']))$Prior
warm_starts <- as.list(unique(dat['Warm.Start']))$Warm.Start
algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")

for (prior in priors) {
  results <- matrix(nrow=3, ncol=length(warm_starts))
  rownames(results) <- c("TS vs DG", "TS vs DEG", "DG vs DEG")
  colnames(results) <- lapply(warm_starts, function (start) { return(paste("WS =", start))})
  for (i in 1:length(warm_starts)) {
    warm_start = warm_starts[i]
    for (k in 1:length(algs)) {
      for (l in 1:length(algs)) {
        p1alg = algs[k]
        p2alg = algs[l]
        filtered_dat <- filter(dat, Prior == prior & Agent.Alg == "HardMax" & Warm.Start == warm_start & P1.Alg == p1alg & P2.Alg == p2alg)
        if (nrow(filtered_dat) == 0) { next }
        share <- filtered_dat$Market.Share.for.P1
        test <- t.test(share)
        cin <- signif(test$conf.int[2] - mean(share), digits=1)
        eeog_mean <- signif(mean(filtered_dat$EEOG), digits=2)
        eeog_summary <- summary(filtered_dat$EEOG)
        res_text <- paste("<b>",signif(mean(share), digits=2), "</b> (", cin, ")<br><u>eeog</u><br> avg: ", eeog_mean, "<br>med: ", eeog_summary["Median"] , sep="")
        if (p1alg == "ThompsonSampling" && p2alg == "DynamicGreedy") {
          results[1, i] <- res_text
        } else if (p1alg == "ThompsonSampling" && p2alg == "DynamicEpsilonGreedy") {
          results[2, i] <- res_text
        } else if (p1alg == "DynamicGreedy" && p2alg == "DynamicEpsilonGreedy") {
          results[3, i] <- res_text
        }
      }
    }
  }
  tab <-xtable(results, caption=paste("Results for", prior, "HardMax", "K=10, erased reputation"))
  print(tab, type="html")
  if (prior != ".5/.7 Random Draw") {   print("<br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br>") }

}
```

### 2018-06-27 analysis_code/tournament_table.Rmd
commit new results

--- 2018-03-26 analysis_code/tournament_table.Rmd
@@
 
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/full_tournament_raw_results.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_test_raw.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_parallel_test_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
 p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
-memory_sizes <- sort(as.list(unique(dat['Memory.Size']))$Memory.Size, decreasing=TRUE)
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(1000)
+time_horizons <- c(5000)
 priors <- as.list(unique(dat['Prior']))$Prior
@@
 algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
+alg_names <- c("Thompson Sampling", "Dynamic Epsilon Greedy", " DynamicGreedy")
 options(xtable.sanitize.text.function=identity)
@@
             p2alg = algs[l]
-            filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Memory.Size == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+            filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Warm.Start == 5 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
             if (nrow(filtered_dat) == 0) {
@@
         }
-        tab <-xtable(results, caption=paste("Results for", agent_alg, "t =", time, prior))
+        tab <-xtable(results, caption=paste("Results for", agent_alg, "t =", time, prior, "200 free obs", "K =", 10))
         print(tab, type="html")

### 2018-06-27 new_conjectures/new_results.tex
commit new results

\documentclass[11pt,letterpaper]{article}
\usepackage{fullpage}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{algorithm,algpseudocode}
\usepackage{pdfpages}

\graphicspath{{../results}}

\begin{document}

\title{Competing Bandits}
\maketitle

\section*{Overview of Old Simulation Results}

First, summarizing the simulation results that we have seen:

\textbf{Priors}:
\begin{enumerate}
\item Needle In a Haystack (x, K) = (K-1) arms with a mean of x and 1 arm with mean x + 0.2
\item Heavy Tail - means of arms randomly drawn from beta(0.6, 0.6)
\item Uniform - means of arms randomly drawn from [0.25, 0.75]
\item .5/.7 Prior - means of arms randomly drawn from {0.5, 0.7}
\end{enumerate}

\textbf{Question}: What algorithms win when both firms start at the same time (simultaneous entry)? How does this vary across agent algorithms?

\begin{enumerate}
\item For HM
\begin{enumerate}
\item $TS > DEG$, DG for Needle In Haystack (0.5, 10), (0.7, 10)
\item 50/50 for Needle In Haystack (0.5, 3), (0.1, 10), (0.3, 10),  Uniform (K = 3, 10 - low warm start), .5/.7 prior
\item $DEG, DG > TS$ for Heavy Tail (K = 3, 10), Uniform (K = 10 - moderate warm start of 15, 20)
\item $DG > DEG$ for Heavy Tail (this holds for really large time horizons!)
\item Results sensitive to the warm start (see latest experiments)
\item Effective End of Game is skewed to the right with many instances having a median of  0, which means that many of the games end simply from the choices made in the warm start, but a relatively larger mean.
\item Many extreme shares - one principal takes most of the market
\end{enumerate}
\item For HMR:
\begin{enumerate}
\item $TS > DEG, DG$ wins with sufficiently large time horizon 
\item One exception is the .5/.7 prior (where means of the arms are randomly drawn from {0.5, 0.7}) where even this is 50/50
\item We see that $DG > DEG$ for HeavyTail prior, with K = 3
\item Moderate variance in the market shares
\end{enumerate}
\item For SM

\begin{enumerate}
\item Qualitatively similar to HMR where $TS > DEG, DG$ except in the .5/.7 prior.
\item However, $DG > DEG$ does not hold for HeavyTail prior.
\item Very low variance in the shares
\end{enumerate}

\end{enumerate}

\textbf{Question}: What happens if there is asymmetric entry so that one principal has an incumbency advantage? When the other principal enters the incumbent has both an informational and reputation advantage. Which serves as bigger barriers to entry?

\begin{enumerate}
\item For the incumbent TS is a dominant strategy across priors and agent models that we have tested out.
\item For SM, TS is not better by much
\item For HMR, TS is better by more than SM
\item For HM, TS is better by more than HMR
\item For the entrant, it is ambiguous still what is the best strategy
\end{enumerate}

\textbf{Information or reputation?}

\begin{enumerate}
\item For HM erasing reputation hurts incumbent more (reputational advantage is more important than data advantage)
\item For HMR/SM - erasing either hurts a bit, but erasing both hurts a lot implying that repetitional and data advantage substitute for one another
\end{enumerate}

\pagebreak

\section*{Conjectures}

Our goal in this section is to be able to have a coherent story to explain the results discussed in the previous section as well as guide what we ought to validate in order to confirm the story. In our model we introduce a notion of reputation in the competing bandits game by allowing agents to learn about the past performance of the firms. The consequence of this for the firms is that now the principals face a dilemma between exploration and reputation since exploration today may give me better information for tomorrow but may also hurt my reputation for tomorrow so that I lose users in competition tomorrow while exploration may lead me to not gain much information for tomorrow but may help my reputation for tomorrow. This tradeoff is especially pertinent when just starting out and neither principal has strong reputation nor sufficient information so that there is a value to exploration.

This leads to following question: Under what conditions are principals incentivized to adopt smarter learning algorithms?

To begin, focus on a standard multi-armed bandit problem. If we compare the performance of a greedy algorithm and an adaptive exploration algorithm in isolation, then on sufficiently hard learning problems where the adaptive exploration algorithm has to explore for a while, we expect the performance of the greedy algorithm to be better than the adaptive exploration algorithm in the early rounds and the performance of the adaptive exploration algorithm to be better than the greedy algorithm in the later rounds. Thus, for a given learning problem, there is some threshold $\bar{t}$ where, if the instance runs for $T > \bar{t}$ then the adaptive exploration algorithm would, on average, perform better. Additionally, we expect that the ``performance gap" between a ``good" algorithm such as TS and a ``bad" algorithm such as DG will be greater for ``harder" learning instances.

When we move to the competition game in our model we introduce a notion of reputation so that suboptimal exploration does not only impact my reward today but my reputation tomorrow. In comparing the performance of the algorithms in isolation we only had what was effectively a ``learning" phase and an ``exploitation" phase. When adding reputation we conjecture that we can view this as adding a third phase in between learning and exploitation which is a reputation-recovery phase. The main effect that this has is that the threshold at which a ``better" algorithm is expected to ``win" in competition with reputation is greater than $\bar{t}$ from before since it is not only necessary to have better instantaneous regret than the other algorithm but the algorithm also needs additional rounds to recover the reputation loss suffered from exploration (let's call this time $t^{w}(B)$).

\textbf{Main Conjecture}: Fix a bandit instance $B$, then for that bandit instance there is some $t^{w}(B)$ where a ``better" algorithm will on average win so long as the time horizon is sufficiently long and somehow it eventually gets $t^{w}(B)$ observations.

In this context, we for now say an algorithm $a$ is ``better" than an algorithm $b$ on $B$ if, in isolation, past some $t^{w}$, algorithm $a$ gets a weakly higher reputation $b$ more than 50\% of the time when run on $I$\footnote{Note that previously we were looking at the mean in the preliminary plots, see later discussion. It would be better to have a definition that is not dependent on relative performance but just on the instance itself }.

Further, we conjecture that $t^{w}(B)$ is increasing in the difficulty of the learning problem. We conjecture that the hardness, $H$, of an instance can be defined by \\ $H = \mathbb{E}[\sum\limits_{\mu_{i} < \mu^{*}} \frac{1}{\mu^{*} - \mu_{i}}]$, where $\mu^{*}$ is the mean of the best arm.

There are several ways in our model in which a principal playing a good algorithm can get $t^{w}(B)$ observations:
\begin{enumerate}
\item There is any randomness in the agent response function, such as the HardMaxWithRandom or SoftMax agent response function. Depending on the parameterization, at some point the good algorithm should get enough random agents to pass the threshold and ``win" the game on average.
\item A sufficiently long incumbency advantage (not sure what sufficiently long is here)
\item A sufficiently long ``warm start" where principals get free users (``sufficiently long" here should be $t^{w}(B)$ especially since the median EEOG is $0$)
\end{enumerate}

Thus, whether better algorithms do better in competition or not depends on whether there is any randomness in the agent response rule or how many free observations we give the firms at the start of the game. Counterintuitively, under the HardMax decision rule, when firms enter at the same time they are \textit{not} incentivized to play the better algorithm when there are reputational costs to exploration. However, if we give the firms a sufficient number of rounds as a monopolist or give them a sufficient number of free customers then they \textit{are} incentivized to play the better algorithm. It's interesting to relate this to the R \& D literature where generally the development of R \& D is costly but the ability to patent after invention allows the firms to extract monopoly profits ex-post and thus recoup the R \& D costs and this incentives R \& D ex-ante. In our model we have no explicit development costs but rather have \textit{reputational} costs due to the exploration involved in better algorithms. Instead of allowing the firm to recoup the R \& D costs \textit{after} invention as a monopolist via patents we have that it is better for the firm to be a monopolist \textit{at the start} sufficiently long to recoup the costs of exploration while having the threat that there will be an entrant at some fixed point in the future. In some sense we can view the ``monopolist" or ``free consumer" case here as saying that we relax the incentive constraint on the consumer in the early rounds in order to incentivize the firm to employ a better algorithm when the incentive constraints for the consumer kick back in when the entrant comes into the market.

Crucially, this result comes from the fact that firms cannot observe the behavior of other firms and learn from the behavior of other firms since there are two advantages that the initial ``free" rounds gives the firms: reputation and information. At the start of the game there is a tradeoff between exploration and reputation. Initial exploration leads to bad initial reputation but improved information. Since consumers are myopic and don't care about the information gain of the firm but only the reputation of the firm, the firm employing algorithms that involve purposeful exploration suffer but with sufficiently many initial free rounds, the firm will be able to recoup these reputation losses and thus will have both better information and reputation than an entrant. If the entrant could observe the actions of the other firm, then the entrant could potentially use this as information about the bandit instance and learn not only from consumers but from the other firms. However, we require that firms can only directly learn from consumers that pick them. Thus, upon entry, the information gained by the incumbent is the incumbent's alone and thus the incumbency gives the incumbent a data advantage.

A natural follow-up question is then, conditional on having a firm early in the market, what serves as stronger barriers to entry in our model: reputation or data advantage? Our results show that, for HardMax, the reputation advantage is more useful but that the data advantage still is significant relative to the results from the original game. I don't really have a good intuition for why this should be the case, but wonder if this has anything to do with good information can't help you if you get unlucky with your rewards.

\section*{DEG vs DG}

The first thing that is at odds with the above conjectures is that we've observed that $DEG > DG$ for Heavy Tail. However, the preliminary plots that we had been looking at were looked at the reputation curves for the \textit{mean} reputation across simulations. For instance, the mean plot for Heavy Tail looked as follows: \\
\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Heavy Tail 3 arms"}

but if, instead of looking at mean reputation, we look at the fraction of rounds where the reputation of DEG is higher than DG we see that $DG > DEG$ on this instance.

\includegraphics[scale=0.5]{deg_dg_ht_3_prelim}

\section*{Large Warm Start}

We want to see if, for sufficiently large warm start, we indeed see that TS wins. First, let's take a look at the relative reputation plots for Heavy Tail and Uniform. Notice that, for both, $DEG, DG > TS$ in the earlier time horizons and $TS > DEG, DG$ in the later time horizons, with Uniform taking longer for $TS > DEG, DG$ than Heavy Tail. \\

\includegraphics[scale=0.5]{ts_dg_ht_prelim_10} \\
\includegraphics[scale=0.5]{ts_deg_ht_10_prelim} \\
\includegraphics[scale=0.4]{deg_dg_ht_10_prelim} \\
\includegraphics[scale=0.5]{ts_dg_uniform_10_prelim} \\
\includegraphics[scale=0.5]{ts_deg_unif_10_prelim} \\
\includegraphics[scale=0.4]{deg_dg_unif_10_prelim} \\

Given this we conjectured that in the competition game, $DEG, DG > TS$ in the early rounds but $TS > DEG, DG$. This largely seems to be the case. See the results below. Additionally, here are two more preliminary plots with smaller time gaps.

\includegraphics[scale=0.5]{ts_dg_unif_fine} \\
\includegraphics[scale=0.5]{ts_dg_heavy_fine} \\

\includepdf[pages={-}]{large_ws}

\section*{Learning Complexity}

Using the same instances that we used for the plots generated in the preliminary plots, we calculate the empirical ``learning complexity" for the $K = 10$ instances according to the hardness metric defined above.

\vspace{0.5cm}
\noindent \underline{Needle In Haystack - 0.5} \\ \textbf{Mean}:  45 \textbf{Median} : 45 \\
\underline{Heavy Tail} \\ \textbf{Mean}:  196.897  \textbf{Median} : 49.08325 \\
\underline{.5/.7 Random Draw} \\ \textbf{Mean}:  23.82 \textbf{Median}: 25  \\
\underline{Uniform} \\ \textbf{Mean}:  171.7698 \textbf{Median} : 79.59571  \\

\vspace{0.5cm}
This is promising given that Heavy Tail and Uniform are the two instances where we see that $DEG, DG > TS$ for HardMax since before we conjectured that on harder learning instances, for small enough warm start observations, TS will do purposeful exploration and get bad reputation and thus lose the competition game. There is probably more we should validate about this metric but need to think about it more.

One puzzling thing about it is, for instance, on the easy instance of 0.5 / 0.7 Random Draw, we see that things are roughly 50 / 50 but we see that $TS > DEG, DG$ on Needle In Haystack - 0.5? One idea is that on 0.5 / 0.7 learning is so easy that there is no value to smarter exploration and that on Needle In Haystack learning is not trivial but is very quick  so that only a few samples are needed and thus the reputational cost is negligible but since $DG$ can't trivially find the best arm there is value to smarter exploration. Not sure though.

\section*{Incumbent Experiment - Varying the number of free observations}

Another implication of the proposed conjectures is that TS ought to be the dominant strategy for the incumbent as long as they are an incumbent for sufficiently long before the entrant enters the market. We fix the agent response model as HardMax. Then, for sufficiently hard learning problems it ought to be the case that if the incumbent only has relatively few periods as the monopolist then TS will not necessarily win. However, with sufficiently long time as the monopolist then TS ought to win.

We verify this in two experiments by first looking where the incumbent gets 50 free observations and then where the incumbent gets 200 free observations. We observe that in the case of 50 free observations, TS does not beat DEG and DG from the ``harder" instances defined by the complexity metric defined above (i.e. Uniform and Heavy Tail), wins in the relatively easier but not trivial learning problem NIH, and algorithm choice makes little difference in the 0.5 / 0.7 Prior.

When we increase the number of free observations to 200 we see that TS is a better algorithm choice for the incumbent (one exception - Uniform against DEG but this is likely due to the fact that we need more free observations, see the preliminary plots).

\includepdf[pages={-}]{free_obs_50_10_arms}

\section*{Appendix}

The rest of the relative reputation plots: \\
\includegraphics[scale=0.5]{ts_dg_ht_3_prelim} \\
\includegraphics[scale=0.5]{ts_deg_ht_3_prelim} \\
\includegraphics[scale=0.5]{ts_deg_unif_3_prelim} \\
\includegraphics[scale=0.5]{ts_dg_unif_3_prelim} \\
\includegraphics[scale=0.5]{deg_dg_unif_3_prelim} \\

\includegraphics[scale=0.5]{ts_deg_nih_10_prelim} \\
\includegraphics[scale=0.5]{ts_dg_nih_10_prelim} \\
\includegraphics[scale=0.5]{deg_dg_nih_10_prelim} \\
\includegraphics[scale=0.5]{ts_deg_nih_3_prelim} \\
\includegraphics[scale=0.5]{ts_dg_nih_3_prelim} \\
\includegraphics[scale=0.5]{deg_dg_nih_3_prelim} \\

\includegraphics[scale=0.5]{ts_deg_5_7_10_prelim} \\
\includegraphics[scale=0.5]{ts_dg_5_7_10_prelim} \\
\includegraphics[scale=0.5]{deg_dg_5_7_10_prelim} \\

Finally, here is the warm start experiment redone with reputation erased after the warm start:

\includepdf[pages={-}]{large_ws_erase_rep}

\end{document}

### 2018-06-27 analysis_code/preliminary_3.R

@@ -1,0 +1,95 @@
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+
+dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_10_arms_10_int.csv")
+#dat <- read.csv("/Volumes/Mac/bandits/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_10_arms.csv")
+
+
+concise_alg_rep <- function(alg) {
+  if (alg == "ThompsonSampling") {
+    return("TS")
+  } else if (alg == "DynamicEpsilonGreedy") {
+    return("DEG")
+  } else if (alg == "DynamicGreedy") {
+    return("DG")
+  }
+}
+print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
+K <- unique(dat$K)[1]
+dist_dat <- filter(dat, Distribution == dist & Realized.Complexity >= minComplexity)
+deg <- filter(dist_dat, Algorithm == alg1)
+dg <- filter(dist_dat, Algorithm == alg2)
+n_vals <- unique(dist_dat$n)
+t_vals <- seq(10, 1500, 10)
+
+counts <- c()
+for (t in t_vals) {
+  count <- 0
+  for (n in n_vals) {
+    cur_delayed_t <- t - 200
+    cur_dg <- filter(dg, n == UQ(n) & t == UQ(t))
+    cur_deg <- filter(deg, n == UQ(n) & t == UQ(t))
+    comp <- cur_deg$Realized.Reputation >= cur_dg$Realized.Reputation
+    if (comp) {
+      count <- count + 1
+    }
+  }
+  counts <- c(counts, (count / length(n_vals)))
+}
+plot_title <- paste("Re Rep -", dist, "K =", K, "Comp >=", minComplexity)
+q <- qplot(t_vals, counts) + 
+  ggtitle(plot_title) + xlab("t") + 
+  ylab(paste(concise_alg_rep(alg1), " >= ", concise_alg_rep(alg2), "(reputation)"))
+
+print(q)
+}
+
+print_mean_graphs <- function (dist, alg1, alg2, minComplexity) {
+  K <- unique(dat$K)[1]
+  dist_dat <- filter(dat, Distribution == dist & Realized.Complexity >= minComplexity)
+  deg <- filter(dist_dat, Algorithm == alg1)
[…]

### 2018-06-27 analysis_code/preliminary_plots.R

@@ -2,17 +2,10 @@ library("ggplot2")
 library("dplyr")
 library("reshape")
 
-# Three datasets that currently exist are:
-# preliminary_plots.csv was run with N=250, T=6000
-# preliminary_plots_2.csv includes the Bayesian Dynamic Epsilon Greedy runs
-
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_20_arms.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_all_haystack.csv", sep=""))
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_reputation_2.csv", sep=""))
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_3.csv", sep=""))
-#dat <- rbind(dat, dat2)
 dat$Algorithm <- replace(as.character(dat$Algorithm), dat$Algorithm == "DynamicEpsilonGreedy", "DynamicEpsilonGreedy, 0.05")
 dat$Algorithm <- replace(as.character(dat$Algorithm), dat$Algorithm == "NonBayesianEpsilonGreedy", "NonBayesianEpsilonGreedy, 0.05")
 algs <- as.list(unique(dat['Algorithm']))$Algorithm
@@ -22,7 +15,7 @@ dists <- as.list(unique(dat['Distribution']))$Distribution
 
 filter_by_dist_and_plot <- function(dist) {
   d <- filter(dat, Distribution == dist)
-  title <- paste("Reputation Trajectory for", dist, "3 arms")
+  title <- paste("Reputation Trajectory for", dist, "10 arms")
   q <- ggplot(data=d, aes(x=t, y=Realized.Reputation, colour=Algorithm)) + ggtitle(title) + ylab("Reputation") + xlab("time") +
     #geom_path() # just plot the raw trajectory
     geom_smooth(method="loess") #+ smooths the trajectory 

### 2018-06-27 analysis_code/trajectory.R

@@ -1,14 +1,44 @@
 library(dplyr)
+library(xtable)
 
-WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
-f <- "tournament_experiment_everything_hmr_raw.csv"
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/", f, sep=""))
-filtered_dat <- filter(dat, Prior == "Heavy Tail")
+WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/"
+f <- "tournament_experiment_effective_end_of_game_warm_start_raw.csv"
+dat <- read.csv(file=paste(WORKING_PATH, f, sep=""))
+dat <- filter(dat, Time.Horizon == 2000)
+priors <- as.list(unique(dat['Prior']))$Prior
+warm_starts <- as.list(unique(dat['Warm.Start']))$Warm.Start
+algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
 
-WINDOW <- 5000
-for (i in 220:240) {
-  begin <- (i*WINDOW) + 1
-  end_window <- (i + 1) * WINDOW
-  one_sim <- filtered_dat[begin:end_window,]
-  plot(one_sim[,"Time.Horizon"], one_sim[, "Market.Share.for.P1"], main="HMR Market Share over Time", ylab="Market Share", xlab="Time")
+for (prior in priors) {
+  results <- matrix(nrow=3, ncol=length(warm_starts))
+  for (i in 1:length(warm_starts)) {
+    warm_start = warm_starts[i]
+    for (k in 1:length(algs)) {
+      for (l in 1:length(algs)) {
+        p1alg = algs[k]
+        p2alg = algs[l]
+        filtered_dat <- filter(dat, Prior == prior & Warm.Start == warm_start & P1.Alg == p1alg & P2.Alg == p2alg)
+        if (nrow(filtered_dat) == 0) { next }
+        share <- filtered_dat$Market.Share.for.P1
+        test <- t.test(share)
+        cin <- signif(test$conf.int[2] - mean(share), digits=1)
+        if (p1alg == "ThompsonSampling" && p2alg == "DynamicGreedy") {
+          results[1, i] <- mean(filtered_dat$EEOG)
+        } else if (p1alg == "ThompsonSampling" && p2alg == "DynamicEpsilonGreedy") {
[…]

### 2018-07-31 new_conjectures/new_results.tex
check in bunch of updates...

@@
 
-\section*{Overview of Old Simulation Results}
+\section*{Overview of Results}
 
@@
 \item Uniform - means of arms randomly drawn from [0.25, 0.75]
-\item .5/.7 Prior - means of arms randomly drawn from {0.5, 0.7}
+\item .5/.7 Prior - means of arms randomly drawn from $\{0.5, 0.7\}$
 \end{enumerate}
@@
 \begin{enumerate}
-\item For HM
+\item Can we use any metrics of the bandit instances run in isolation that help us predict who wins in the competion game?
+A metric that seems to work well is the relative reputation plot that simply computes the proportion of instances when an algorithm $a$ has at least as high of a reputation than algorithm $b$ at fixed time step $t$ and on the same bandit instance (with the same realizations).
 \begin{enumerate}
-\item $TS > DEG$, DG for Needle In Haystack (0.5, 10), (0.7, 10)
-\item 50/50 for Needle In Haystack (0.5, 3), (0.1, 10), (0.3, 10),  Uniform (K = 3, 10 - low warm start), .5/.7 prior
-\item $DEG, DG > TS$ for Heavy Tail (K = 3, 10), Uniform (K = 10 - moderate warm start of 15, 20)
-\item $DG > DEG$ for Heavy Tail (this holds for really large time horizons!)
-\item Results sensitive to the warm start (see latest experiments)
-\item Effective End of Game is skewed to the right with many instances having a median of  0, which means that many of the games end simply from the choices made in the warm start, but a relatively larger mean.
+\item For HM - fixing a warm start value $t$ (number of free observations at the start), whichever algorithm has $>0.5$ proportion of ``victories" at time $t$ according to the relative reputation plot should, on average, win more in competition
+\item For HMR and SM, whichever algorithm has $>0.5$ proportion of ``victories" for sufficiently large $t$ (i.e. when learning is more or less done) should win in competition
+\item \textbf{Question}: Does this always line up with looking at the mean reputation plots?
+In one case, Heavy Tail with $K = 3$, we see that the mean plots show $DEG > DG$, but our simulation results show that $DG > DEG$ in competition and $DG > DEG$ in the relative reputation plots.
+\end{enumerate}
+\item What algorithms win when fixing the HM response model?
+\begin{enumerate}
+\item We see that there is a low effective end of game\footnote{Effective End of Game (EEOG) is defined as the last round when the agent ``switched" the firm she chose}, which is skewed with many instances having a median of 0, meaning that many of the games end simply from the choices made in the warm start
 \item Many extreme shares - one principal takes most of the market
+\item What algorithm wins depends on the warm start parameter (number of initial free observations). 
+\item For low warm start of 5 we have the following results
+\begin{enumerate}
+\item $TS > DEG, DG$ for Needle In Haystack (0.5, 10), (0.7, 10)
+\item 50/50 for Needle In Haystack (0.5, 3), (0.1, 10), (0.3, 10),  Uniform (K = 3, 10), .5/.7 prior
+\item $DEG, DG > TS$ for Heavy Tail (K = 3, 10)
+\item $DG > DEG$ for Heavy Tail (K = 3)
 \end{enumerate}
+\item For moderate warm start (20, 50)
+\begin{enumerate}
+\item $DEG, DG > TS$ for Uniform (K = 10), Heavy Tail (K = 10)
+\item $TS > DEG, DG$ for Needle In Haystack (0.5, 10), (0.7, 10)
+\end{enumerate}
+\item For large warm start
+\begin{enumerate}
+\item $TS > DEG, DG$ for everything except .5/.7 prior
+\end{enumerate}
+\end{enumerate}
+
 \item For HMR:
 \begin{enumerate}
-\item $TS > DEG, DG$ wins with sufficiently large time horizon 
+\item ``Better algorithm" wins with sufficiently large time horizon, where ``better algorithm" is defined as the algorithm that has a larger proportion of wins in the relative reputation plot
 \item One exception is the .5/.7 prior (where means of the arms are randomly drawn from {0.5, 0.7}) where even this is 50/50
-\item We see that $DG > DEG$ for HeavyTail prior, with K = 3
 \item Moderate variance in the market shares
@@
 \begin{enumerate}
-\item Qualitatively similar to HMR where $TS > DEG, DG$ except in the .5/.7 prior.
-\item However, $DG > DEG$ does not hold for HeavyTail prior.
+\item Qualitatively similar to HMR where better algorithm wins for sufficiently large time horizon
+\item Results closer to 50/50 than HMR
 \item Very low variance in the shares
@@
 \textbf{Question}: What happens if there is asymmetric entry so that one principal has an incumbency advantage?
-When the other principal enters the incumbent has both an informational and reputation advantage.
-Which serves as bigger barriers to entry?
 
 \begin{enumerate}
-\item For the incumbent TS is a dominant strategy across priors and agent models that we have tested out.
+\item If the incumbency advantage is sufficiently long, then TS is a dominant strategy across priors and agent models that we have tested out.
+\item If the incumbency advantage is not sufficiently long, TS is no longer a dominant strategy
 \item For SM, TS is not better by much
@@
 \item For HM, TS is better by more than HMR
-\item For the entrant, it is ambiguous still what is the best strategy
+\item For the entrant, it is ambiguous what is the best strategy
 \end{enumerate}
 
-\textbf{Information or reputation?}
+\textbf{Question}: Incumbency gives both an informational and a reputational advantage.
+If we artificially erase one upon entry of the entrant, does information or reputation play a bigger role as barriers to entry?
 
@@
 \item For HM erasing reputation hurts incumbent more (reputational advantage is more important than data advantage)
-\item For HMR/SM - erasing either hurts a bit, but erasing both hurts a lot implying that repetitional and data advantage substitute for one another
+\item For HMR/SM - erasing either hurts a bit, but erasing both hurts a lot implying that reputational and data advantage substitute for one another
 \end{enumerate}
+
+\section*{Story}
+
+\begin{itemize}
+\item Competition in learning environments with no prices - compete for users based on quality alone but quality of different actions is unknown and need to learn the quality of actions.
+Learn by committing to a learning algorithm at the start of the world.
+Want the adoption of better algorithms since it is socially beneficial but when does this happen (identifies best arm, maximizes social welfare - i.e.
+minimizes regret)?
+\item \textbf{Problem}: Competition imposes incentive compatibility constraints on myopic consumers (i.e. firm needs to choose actions so that they actually get selected by users in competition)
+\item We introduce a simple metric of reputation - agents want to maximize their reward and use a reputation score as a proxy for that.
+Reputation score is a sliding window average of the reward experienced by past agents that had selected this firm.
+Agents are myopic and non-strategic so they do not consider other factors such as the algorithm employed by the firms.
+\item Firms face a tradeoff between reputation and exploration.
+Exploration allows them to gather information and implement the first-best but comes at the cost of reputation.
+Consumers don't care about information gain but do care about reputation.
+Thus, algorithms that engage in pure exploration suffer reputational costs and lose in the competition game to algorithms that engage in no exploration.
+\item How to incentivize better algorithms? Sufficient number of consumers whose incentive compatibility constraints are effectively removed. A few ways this can happen
+\begin{enumerate}
+\item ``Random agents" or probabilistic decision rules (HMR or SM response)
+\item ``Warm start" or free agents at the start of the game
+\item Some period of monopoly power
+\end{enumerate}
+\item Takeaway: competition can stifle ``innovation" if learning has reputational effects. Monopoly has
+
+\end{itemize}
 

### 2018-07-31 new_conjectures/results_overview.tex
check in bunch of updates...

--- 2018-07-31 new_conjectures/new_results.tex
@@
 
-\title{Competing Bandits}
+\title{Competing Bandits - Summary}
 \maketitle
@@
 \section*{Overview of Results}
-
-First, summarizing the simulation results that we have seen:
 
@@
 
-\textbf{Question}: What algorithms win when both firms start at the same time (simultaneous entry)? How does this vary across agent algorithms?
+\textbf{Question}: What algorithms win when both firms start at the same time (simultaneous entry)? How does this vary across agent response functions?
 
@@
 \item For HM - fixing a warm start value $t$ (number of free observations at the start), whichever algorithm has $>0.5$ proportion of ``victories" at time $t$ according to the relative reputation plot should, on average, win more in competition
-\item For HMR and SM, whichever algorithm has $>0.5$ proportion of ``victories" for sufficiently large $t$ (i.e. when learning is more or less done) should win in competition
+\item For HMR and SM, whichever algorithm has $>0.5$ proportion of ``victories" for sufficiently large $t$ (i.e.
+when learning is more or less done) should win in competition.
+This is the case since the reputation curve of a given algorithm eventually converges and we simply need to ask which reputation curve is relatively higher at the convergence value.
 \item \textbf{Question}: Does this always line up with looking at the mean reputation plots?
 In one case, Heavy Tail with $K = 3$, we see that the mean plots show $DEG > DG$, but our simulation results show that $DG > DEG$ in competition and $DG > DEG$ in the relative reputation plots.
+Relation to Dueling Algorithms paper (?).
+One intuition to this which plays out in the data is that DG either:
+\begin{enumerate}
+\item identifies the best arm and thus beats or equals reputation of DEG since DEG explores randomly with $\epsilon$ probability.
+\item does not identify the best arm and does substantially worse than DEG.
+\end{enumerate}
+As a result, it will have a worse average reputation due to the cases where it identifies the wrong best arm but there may be enough cases where it does so that when only comparing relative performance in a given simulation it does better.
 \end{enumerate}
@@
 \begin{enumerate}
-\item $TS > DEG, DG$ for everything except .5/.7 prior
+\item $TS > DEG, DG$ for everything except .5/.7 prior (why not for .5/.7 prior? Will run more simulations to see if it narrowly wins)
+\end{enumerate}
+\item \textbf{Question}: Why do results vary by warm start value?
+\begin{enumerate}
+\item On ``harder" learning instances, learning takes longer and pure exploration has early reputation costs.
+Warm start needs to be sufficiently high to have recovered the reputation costs of exploration.
+\item On ``harder" learning instances, there is a benefit to exploration in the sense that it allows the principal to differentiate itself in reputation against an algorithm like dynamic greedy in the long run.
+\item On ``easier" learning instances, little reputation cost but also little benefit to deploying a smarter learning algorithm?
+\item How to define ``harder?" Use  $H = \mathbb{E}[\sum\limits_{\mu_{i} < \mu^{*}} \frac{1}{\mu^{*} - \mu_{i}}]$, where $\mu^{*}$ is the mean of the best arm to define instance difficulty.
+On the ``harder" instances we seem to see larger warm start values required in order for the better algorithm to win but not sure this explains everything.
 \end{enumerate}
@@
 \item If the incumbency advantage is not sufficiently long, TS is no longer a dominant strategy
-\item For SM, TS is not better by much
-\item For HMR, TS is better by more than SM
-\item For HM, TS is better by more than HMR
+\item For HM, TS leads to one firm (the incumbent) dominating the market.
+\item For HMR, TS leads to one firm dominating the market but not as dramatically compared to HM.
+\item For SM, TS still wins but closer to 50/50. than HMR/HM.
 \item For the entrant, it is ambiguous what is the best strategy
@@
 \begin{enumerate}
-\item For HM erasing reputation hurts incumbent more (reputational advantage is more important than data advantage)
+\item For HM erasing reputation hurts incumbent more (reputational advantage is more important than data advantage) (Why?)
 \item For HMR/SM - erasing either hurts a bit, but erasing both hurts a lot implying that reputational and data advantage substitute for one another
@@
 \begin{itemize}
-\item Competition in learning environments with no prices - compete for users based on quality alone but quality of different actions is unknown and need to learn the quality of actions.
+\item Competition in learning environments - compete for users based on quality alone but quality of different actions is unknown and need to learn the quality of actions (modeled as an iid MAB).
 Learn by committing to a learning algorithm at the start of the world.
-Want the adoption of better algorithms since it is socially beneficial but when does this happen (identifies best arm, maximizes social welfare - i.e.
-minimizes regret)?
-\item \textbf{Problem}: Competition imposes incentive compatibility constraints on myopic consumers (i.e. firm needs to choose actions so that they actually get selected by users in competition)
+Want the adoption of traditionally ``better" algorithms since it is socially beneficial but when does this happen?
 \item We introduce a simple metric of reputation - agents want to maximize their reward and use a reputation score as a proxy for that.
@@
 Agents are myopic and non-strategic so they do not consider other factors such as the algorithm employed by the firms.
-\item Firms face a tradeoff between reputation and exploration.
-Exploration allows them to gather information and implement the first-best but comes at the cost of reputation.
-Consumers don't care about information gain but do care about reputation.
-Thus, algorithms that engage in pure exploration suffer reputational costs and lose in the competition game to algorithms that engage in no exploration.
-\item How to incentivize better algorithms? Sufficient number of consumers whose incentive compatibility constraints are effectively removed. A few ways this can happen
+\item \textbf{Problem}: Competition imposes incentive compatibility constraints on myopic consumers (i.e.
+firm needs to choose actions so that they actually get selected by users in competition).
+Start with consumers using the HardMax decision rule.
+\item Firms face a tradeoff between picking algorithms that maximize short-term reputation and those engage in pure exploration (which assists in maximizing long-term reputation).
+Consumers don't care about information gain from exploration but do care about reputation.
+Thus, algorithms that engage in pure exploration suffer short-term reputational costs and lose in the competition game to algorithms that maximize short-term reputation.
+Another aspect of this is that it depends on the hardness of the learning instance - the ``harder" learning instances have larger (short-term) reputational costs but potentially larger (long-run) reputational advantage.
+\item How to incentivize firms better algorithms? Sufficient number of consumers whose incentive compatibility constraints are effectively removed. A few ways this can happen
 \begin{enumerate}
+\item Sufficient number of periods as the monopolist in the market (incumbent experiment results)
 \item ``Random agents" or probabilistic decision rules (HMR or SM response)
 \item ``Warm start" or free agents at the start of the game
-\item Some period of monopoly power
 \end{enumerate}
-\item Takeaway: competition can stifle ``innovation" if learning has reputational effects. Monopoly has
+\item Focusing on the monopoly case and HM decision rule, the firm accumulates both a reputational advantage and a data advantage.
+Which is more important in the competition game for serving as barriers to entry.
+We see that it is reputation (I have no good intuition for this yet).
+\item \textbf{Takeaway}: competition can stifle ``innovation" if learning has reputational effects due to incentive constraints on consumers.
+Counterintuitively, monopoly can have benefit for innovation since it allows for exploration by not having to worry about incentive constraint of consumers.
+However, sometimes even a traditionally ``better" algorithm according to mean regret cannot win in competition since relative reputation matters more.
 
 \end{itemize}
-
-\pagebreak
-
-\section*{Conjectures}
-
-Our goal in this section is to be able to have a coherent story to explain the results discussed in the previous section as well as guide what we ought to validate in order to confirm the story.
-In our model we introduce a notion of reputation in the competing bandits game by allowing agents to learn about the past performance of the firms.
-The consequence of this for the firms is that now the principals face a dilemma between exploration and reputation since exploration today may give me better information for tomorrow but may also hurt my reputation for tomorrow so that I lose users in competition tomorrow while exploration may lead me to not gain much information for tomorrow but may help my reputation for tomorrow.
-This tradeoff is especially pertinent when just starting out and neither principal has strong reputation nor sufficient information so that there is a value to exploration.
-
-This leads to following question: Under what conditions are principals incentivized to adopt smarter learning algorithms?
-
-To begin, focus on a standard multi-armed bandit problem.
-If we compare the performance of a greedy algorithm and an adaptive exploration algorithm in isolation, then on sufficiently hard learning problems where the adaptive exploration algorithm has to explore for a while, we expect the performance of the greedy algorithm to be better than the adaptive exploration algorithm in the early rounds and the performance of the adaptive exploration algorithm to be better than the greedy algorithm in the later rounds.
-Thus, for a given learning problem, there is some threshold $\bar{t}$ where, if the instance runs for $T > \bar{t}$ then the adaptive exploration algorithm would, on average, perform better.
-Additionally, we expect that the ``performance gap" between a ``good" algorithm such as TS and a ``bad" algorithm such as DG will be greater for ``harder" learning instances.
-
-When we move to the competition game in our model we introduce a notion of reputation so that suboptimal exploration does not only impact my reward today but my reputation tomorrow.
-In comparing the performance of the algorithms in isolation we only had what was effectively a ``learning" phase and an ``exploitation" phase.
-When adding reputation we conjecture that we can view this as adding a third phase in between learning and exploitation which is a reputation-recovery phase.
-The main effect that this has is that the threshold at which a ``better" algorithm is expected to ``win" in competition with reputation is greater than $\bar{t}$ from before since it is not only necessary to have better instantaneous regret than the other algorithm but the algorithm also needs additional rounds to recover the reputation loss suffered from exploration (let's call this time $t^{w}(B)$).
-
-\textbf{Main Conjecture}: Fix a bandit instance $B$, then for that bandit instance there is some $t^{w}(B)$ where a ``better" algorithm will on average win so long as the time horizon is sufficiently long and somehow it eventually gets $t^{w}(B)$ observations.
-
-In this context, we for now say an algorithm $a$ is ``better" than an algorithm $b$ on $B$ if, in isolation, past some $t^{w}$, algorithm $a$ gets a weakly higher reputation $b$ more than 50\% of the time when run on $I$\footnote{Note that previously we were looking at the mean in the preliminary plots, see later discussion.
-It would be better to have a definition that is not dependent on relative performance but just on the instance itself }.
-
-Further, we conjecture that $t^{w}(B)$ is increasing in the difficulty of the learning problem.
-We conjecture that the hardness, $H$, of an instance can be defined by \\ $H = \mathbb{E}[\sum\limits_{\mu_{i} < \mu^{*}} \frac{1}{\mu^{*} - \mu_{i}}]$, where $\mu^{*}$ is the mean of the best arm.
-
-There are several ways in our model in which a principal playing a good algorithm can get $t^{w}(B)$ observations:
-\begin{enumerate}
-\item There is any randomness in the agent response function, such as the HardMaxWithRandom or SoftMax agent response function.
-Depending on the parameterization, at some point the good algorithm should get enough random agents to pass the threshold and ``win" the game on average.
-\item A sufficiently long incumbency advantage (not sure what sufficiently long is here)
-\item A sufficiently long ``warm start" where principals get free users (``sufficiently long" here should be $t^{w}(B)$ especially since the median EEOG is $0$)
-\end{enumerate}
-
-Thus, whether better algorithms do better in competition or not depends on whether there is any randomness in the agent response rule or how many free observations we give the firms at the start of the game.
-Counterintuitively, under the HardMax decision rule, when firms enter at the same time they are \textit{not} incentivized to play the better algorithm when there are reputational costs to exploration.
-However, if we give the firms a sufficient number of rounds as a monopolist or give them a sufficient number of free customers then they \textit{are} incentivized to play the better algorithm.
-It's interesting to relate this to the R \& D literature where generally the development of R \& D is costly but the ability to patent after invention allows the firms to extract monopoly profits ex-post and thus recoup the R \& D costs and this incentives R \& D ex-ante.
-In our model we have no explicit development costs but rather have \textit{reputational} costs due to the exploration involved in better algorithms.
-Instead of allowing the firm to recoup the R \& D costs \textit{after} invention as a monopolist via patents we have that it is better for the firm to be a monopolist \textit{at the start} sufficiently long to recoup the costs of exploration while having the threat that there will be an entrant at some fixed point in the future.
-In some sense we can view the ``monopolist" or ``free consumer" case here as saying that we relax the incentive constraint on the consumer in the early rounds in order to incentivize the firm to employ a better algorithm when the incentive constraints for the consumer kick back in when the entrant comes into the market.
-
-Crucially, this result comes from the fact that firms cannot observe the behavior of other firms and learn from the behavior of other firms since there are two advantages that the initial ``free" rounds gives the firms: reputation and information.
-At the start of the game there is a tradeoff between exploration and reputation.
-Initial exploration leads to bad initial reputation but improved information.
-Since consumers are myopic and don't care about the information gain of the firm but only the reputation of the firm, the firm employing algorithms that involve purposeful exploration suffer but with sufficiently many initial free rounds, the firm will be able to recoup these reputation losses and thus will have both better information and reputation than an entrant.
-If the entrant could observe the actions of the other firm, then the entrant could potentially use this as information about the bandit instance and learn not only from consumers but from the other firms.
-However, we require that firms can only directly learn from consumers that pick them.
-Thus, upon entry, the information gained by the incumbent is the incumbent's alone and thus the incumbency gives the incumbent a data advantage.
-
-A natural follow-up question is then, conditional on having a firm early in the market, what serves as stronger barriers to entry in our model: reputation or data advantage?
-Our results show that, for HardMax, the reputation advantage is more useful but that the data advantage still is significant relative to the results from the original game.
-I don't really have a good intuition for why this should be the case, but wonder if this has anything to do with good information can't help you if you get unlucky with your rewards.
-
-\section*{DEG vs DG}
-
-The first thing that is at odds with the above conjectures is that we've observed that $DEG > DG$ for Heavy Tail.
-However, the preliminary plots that we had been looking at were looked at the reputation curves for the \textit{mean} reputation across simulations.
-For instance, the mean plot for Heavy Tail looked as follows: \\
-\includegraphics[scale=0.5]{"../results/preliminary_figures/Reputation Trajectory for Heavy Tail 3 arms"}
-
-but if, instead of looking at mean reputation, we look at the fraction of rounds where the reputation of DEG is higher than DG we see that $DG > DEG$ on this instance.
-
-\includegraphics[scale=0.5]{deg_dg_ht_3_prelim}
-
-\section*{Large Warm Start}
-
-We want to see if, for sufficiently large warm start, we indeed see that TS wins.
-First, let's take a look at the relative reputation plots for Heavy Tail and Uniform.
-Notice that, for both, $DEG, DG > TS$ in the earlier time horizons and $TS > DEG, DG$ in the later time horizons, with Uniform taking longer for $TS > DEG, DG$ than Heavy Tail.
-\\
-
-\includegraphics[scale=0.5]{ts_dg_ht_prelim_10} \\
-\includegraphics[scale=0.5]{ts_deg_ht_10_prelim} \\
-\includegraphics[scale=0.4]{deg_dg_ht_10_prelim} \\
-\includegraphics[scale=0.5]{ts_dg_uniform_10_prelim} \\
-\includegraphics[scale=0.5]{ts_deg_unif_10_prelim} \\
-\includegraphics[scale=0.4]{deg_dg_unif_10_prelim} \\
-
-Given this we conjectured that in the competition game, $DEG, DG > TS$ in the early rounds but $TS > DEG, DG$.
-This largely seems to be the case.
-See the results below.
-Additionally, here are two more preliminary plots with smaller time gaps.
-
-\includegraphics[scale=0.5]{ts_dg_unif_fine} \\
-\includegraphics[scale=0.5]{ts_dg_heavy_fine} \\
-
-\includepdf[pages={-}]{large_ws}
-
-\section*{Learning Complexity}
-
-Using the same instances that we used for the plots generated in the preliminary plots, we calculate the empirical ``learning complexity" for the $K = 10$ instances according to the hardness metric defined above.
-
-\vspace{0.5cm}
-\noindent \underline{Needle In Haystack - 0.5} \\ \textbf{Mean}:  45 \textbf{Median} : 45 \\
-\underline{Heavy Tail} \\ \textbf{Mean}:  196.897  \textbf{Median} : 49.08325 \\
-\underline{.5/.7 Random Draw} \\ \textbf{Mean}:  23.82 \textbf{Median}: 25  \\
-\underline{Uniform} \\ \textbf{Mean}:  171.7698 \textbf{Median} : 79.59571  \\
-
-\vspace{0.5cm}
-This is promising given that Heavy Tail and Uniform are the two instances where we see that $DEG, DG > TS$ for HardMax since before we conjectured that on harder learning instances, for small enough warm start observations, TS will do purposeful exploration and get bad reputation and thus lose the competition game.
-There is probably more we should validate about this metric but need to think about it more.
-
-One puzzling thing about it is, for instance, on the easy instance of 0.5 / 0.7 Random Draw, we see that things are roughly 50 / 50 but we see that $TS > DEG, DG$ on Needle In Haystack - 0.5?
-One idea is that on 0.5 / 0.7 learning is so easy that there is no value to smarter exploration and that on Needle In Haystack learning is not trivial but is very quick  so that only a few samples are needed and thus the reputational cost is negligible but since $DG$ can't trivially find the best arm there is value to smarter exploration.
-Not sure though.
-
-\section*{Incumbent Experiment - Varying the number of free observations}
-
-Another implication of the proposed conjectures is that TS ought to be the dominant strategy for the incumbent as long as they are an incumbent for sufficiently long before the entrant enters the market.
-We fix the agent response model as HardMax.
-Then, for sufficiently hard learning problems it ought to be the case that if the incumbent only has relatively few periods as the monopolist then TS will not necessarily win.
-However, with sufficiently long time as the monopolist then TS ought to win.
-
-We verify this in two experiments by first looking where the incumbent gets 50 free observations and then where the incumbent gets 200 free observations.
-We observe that in the case of 50 free observations, TS does not beat DEG and DG from the ``harder" instances defined by the complexity metric defined above (i.e.
-Uniform and Heavy Tail), wins in the relatively easier but not trivial learning problem NIH, and algorithm choice makes little difference in the 0.5 / 0.7 Prior.
-
-When we increase the number of free observations to 200 we see that TS is a better algorithm choice for the incumbent (one exception - Uniform against DEG but this is likely due to the fact that we need more free observations, see the preliminary plots).
-
-\includepdf[pages={-}]{free_obs_50_10_arms}
-
-\section*{Appendix}
-
-The rest of the relative reputation plots: \\
-\includegraphics[scale=0.5]{ts_dg_ht_3_prelim} \\
-\includegraphics[scale=0.5]{ts_deg_ht_3_prelim} \\
-\includegraphics[scale=0.5]{ts_deg_unif_3_prelim} \\
-\includegraphics[scale=0.5]{ts_dg_unif_3_prelim} \\
-\includegraphics[scale=0.5]{deg_dg_unif_3_prelim} \\
-
-\includegraphics[scale=0.5]{ts_deg_nih_10_prelim} \\
-\includegraphics[scale=0.5]{ts_dg_nih_10_prelim} \\
-\includegraphics[scale=0.5]{deg_dg_nih_10_prelim} \\
-\includegraphics[scale=0.5]{ts_deg_nih_3_prelim} \\
-\includegraphics[scale=0.5]{ts_dg_nih_3_prelim} \\
-\includegraphics[scale=0.5]{deg_dg_nih_3_prelim} \\
-
-\includegraphics[scale=0.5]{ts_deg_5_7_10_prelim} \\
-\includegraphics[scale=0.5]{ts_dg_5_7_10_prelim} \\
-\includegraphics[scale=0.5]{deg_dg_5_7_10_prelim} \\
-
-Finally, here is the warm start experiment redone with reputation erased after the warm start:
-
-\includepdf[pages={-}]{large_ws_erase_rep}
-
 \end{document}

### 2018-07-31 analysis_code/agents_table.R

@@ -1,0 +1,106 @@
+library(dplyr)
+library(knitr)
+WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
+
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
+#dat_3 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_large_horizon_3_raw.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_many_arms_haystack_raw.csv", sep=""))
+#dat_2 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_many_arms_2_raw.csv", sep=""))
+#dat_10 <- rbind(dat, dat_2)
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ws_small_erase_rep_raw.csv", sep=""))
+#dat_3 <- dat
+
+
+p1_algs <- as.list(unique(dat_3['P1.Alg']))$P1.Alg
+p2_algs <- as.list(unique(dat_3['P2.Alg']))$P2.Alg
+agent_algs <- as.list(unique(dat_3['Agent.Alg']))$Agent.Alg
+agent_algs <- c("HardMax")
+time_horizons <-  as.list(unique(dat_3['Time.Horizon']))$Time.Horizon
+time_horizons <- c(5000)
+priors <- as.list(unique(dat_3['Prior']))$Prior
+priors <- c("Heavy Tail", "Needle In Haystack - 0.5")
+warm_starts <- c(5)
+
+ZERO_ONE_CUTOFF <- 0.1
+
+
+concise_alg_rep <- function(alg) {
+  if (alg == "ThompsonSampling") {
+    return("TS")
+  } else if (alg == "DynamicEpsilonGreedy") {
+    return("DEG")
+  } else if (alg == "DynamicGreedy") {
+    return("DG")
+  }
+}
+
+concise_agent_rep <- function(agent_alg) {
+  if (agent_alg == "HardMax") {
+    return("HM")
+  } else if (agent_alg == "HardMaxWithRandom") {
+    return("HMR")
+  } else if (agent_alg == "SoftMax") {
+    return("SM")
+  }
+}
+
[…]

### 2018-07-31 analysis_code/eeog.R

@@ -1,10 +1,4 @@
- <style>
-   th,td{
-     padding:10px;
-   }
- </style>
-
-```{r echo=FALSE, message=FALSE, xtable, results="asis"}
+#```{r echo=FALSE, message=FALSE, xtable, results="asis"}
 
 library(dplyr)
 library(knitr)
@@ -24,13 +18,13 @@ concise_alg_rep <- function(alg) {
 }
 
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/"
-f <- "tournament_experiment_hm_large_ws_raw.csv"
+f <- "tournament_experiment_ws_small_raw.csv"
 dat <- read.csv(file=paste(WORKING_PATH, f, sep=""))
 #dat_2 <- read.csv(file=paste(WORKING_PATH, "tournament_experiment_tourn_.5_.7_raw.csv", sep=""))
 #dat <- rbind(dat, dat_2)
 dat <- filter(dat, Time.Horizon == 5000)
-#priors <- as.list(unique(dat['Prior']))$Prior
-priors <- c("Heavy Tail")
+priors <- as.list(unique(dat['Prior']))$Prior
+priors <- c("Needle In Haystack - 0.5")
 #warm_starts <- as.list(unique(dat['Warm.Start']))$Warm.Start
 warm_starts <- c(20, 100, 200)
 algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
@@ -52,7 +46,7 @@ for (prior in priors) {
         cin <- signif(test$conf.int[2] - mean(share), digits=1)
         eeog_mean <- signif(mean(filtered_dat$EEOG), digits=2)
         eeog_summary <- summary(filtered_dat$EEOG)
-        res_text <- paste("<b>",signif(mean(share), digits=2), "</b> (", cin, ")<br><u>eeog</u><br> avg: ", eeog_mean, "<br>med: ", eeog_summary["Median"] , sep="")
+        res_text <- paste("\\makecell{\\textbf{",signif(mean(share), digits=2), "} $\\pm$", cin, "\\\ eeog \\\ avg: ", eeog_mean, "\\\ med: ", eeog_summary["Median"], "}" , sep="")
         if (p1alg == "ThompsonSampling" && p2alg == "DynamicGreedy") {
           results[1, i] <- res_text
         } else if (p1alg == "ThompsonSampling" && p2alg == "DynamicEpsilonGreedy") {
[…]

### 2018-07-31 analysis_code/reputation_vs_information.R

@@ -1,0 +1,72 @@
+library(dplyr)
+library(knitr)
+WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
+
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_reputation_window_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
+erase_rep_dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_erase_reputation_raw.csv", sep=""))
+erase_info_dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_erase_information_raw.csv", sep=""))
+p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
+p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
+memory_sizes <- sort(as.list(unique(dat['Memory.Size']))$Memory.Size, decreasing=TRUE)
+agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
+agent_algs <- c("HardMax")
+time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
+time_horizons <- c(1000)
+priors <- as.list(unique(dat['Prior']))$Prior
+priors <- c("Heavy Tail")
+
+ZERO_ONE_CUTOFF <- 0.1
+algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
+alg_names <- c("TS", "DEG", " DG")
+options(xtable.sanitize.text.function=identity)
+options(xtable.caption.placement="top")
+library(xtable)
+for (j in 1:length(time_horizons)) {
+  for (prior in priors) {
+    for (i in 1:length(agent_algs)) {
+      agent_alg <- agent_algs[i]
+      print_breaks <- FALSE
+      for (q in 1:3) {
+        time <- time_horizons[j]
+        results <- matrix(nrow=length(algs), ncol=length(algs))
+        colnames(results) <- alg_names
+        rownames(results) <- alg_names
+        for (k in 1:length(algs)) {
+          for (l in 1:length(algs)) {
+            p1alg = algs[k]
+            p2alg = algs[l]
+            filtered_dat <- NULL
[…]

### 2018-07-31 analysis_code/tournament_table.R

@@ -1,0 +1,58 @@
+
+library(dplyr)
+library(knitr)
+WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
+
+
+
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_parallel_test_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
+p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
+p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
+agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
+agent_algs - c("HardMax")
+time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
+time_horizons <- c(5000)
+priors <- as.list(unique(dat['Prior']))$Prior
+priors <- c("Needle In Haystack - 0.5", "Heavy Tail")
+
+ZERO_ONE_CUTOFF <- 0.1
+algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
+alg_names <- c("TS", "DEG", " DG")
+options(xtable.sanitize.text.function=identity)
+options(xtable.caption.placement="top")
+library(xtable)
+for (j in 1:length(time_horizons)) {
+  for (prior in priors) {
+    for (i in 1:length(agent_algs)) {
+      agent_alg <- agent_algs[i]
+      print_breaks <- FALSE
+      time <- time_horizons[j]
+      results <- matrix(nrow=length(algs), ncol=length(algs))
+      colnames(results) <- alg_names
+      rownames(results) <- alg_names
+      for (k in 1:length(algs)) {
+        for (l in 1:length(algs)) {
+          p1alg = algs[k]
+          p2alg = algs[l]
+          filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Warm.Start == 5 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+          if (nrow(filtered_dat) == 0) {
+            if (p1alg == p2alg) { print_breaks <- TRUE }
+            next
+          }
+          share <- filtered_dat$Market.Share.for.P1
+          test <- t.test(share)
[…]

### 2018-07-31 lib/agent/HardMaxWithRandom.py

@@ -1,11 +1,12 @@
 import numpy as np
 from HardMax import HardMax
+from constants import DEFAULT_MEMORY
 
 class HardMaxWithRandom(HardMax):
   # epsilon: the baseline probablility that each principal will be picked with. must be in (0, 0.5).
   # if epsilon=0.1, for example, then 20% of the time, the agent picks randomly between the 2 principals
-  def __init__(self, principals, K, priors=None, epsilon=0.05):
-    super(HardMaxWithRandom, self).__init__(principals, K, priors)
+  def __init__(self, principals, K, priors=None, epsilon=0.05, memory=DEFAULT_MEMORY):
+    super(HardMaxWithRandom, self).__init__(principals, K, priors, memory=memory)
     self.epsilon = epsilon
 
   def selectPrincipal(self):

### 2018-07-31 lib/constants.py

@@ -6,4 +6,4 @@ DEFAULT_ALPHA = 30
 DEFAULT_WARM_START_NUM_OBSERVATIONS = 5
 
 
-RECORD_STATS_AT = [1, 500, 1000, 2000, 4000, 5000, 6000, 8000, 12000, 15000]
+RECORD_STATS_AT = [1, 50, 100, 500, 1000, 2000, 4000, 5000, 6000, 8000, 12000, 15000]

### 2018-07-31 main.py

@@ -6,6 +6,7 @@ from lib.bandit.DynamicGreedy import DynamicGreedy
 from lib.bandit.UCB import UCB1WithConstantT
 from lib.bandit.ThompsonSampling import ThompsonSampling
 from lib.bandit.ExploreThenExploit import ExploreThenExploit
+from lib.bandit.MixedGreedy import MixedGreedy
 from lib.constants import RECORD_STATS_AT, DEFAULT_WARM_START_NUM_OBSERVATIONS
 from simulate import simulate, getRealDistributionsFromPrior, initialResultDict
 
@@ -25,12 +26,12 @@ import sys
 import pickle
 
 K = 10
-T = 5002
-NUM_SIMULATIONS = 300
+T = 1001
+NUM_SIMULATIONS = 50
 
 FREE_OBS = False
 FREE_OBS_NUM = 200
-exp_name = 'true_mean_hardmax'
+exp_name = 'complexity_exp'
 print('Exp name', exp_name)
 REALIZATIONS_NAME = '' #if you want to pull in past realizations, fill this in with the realizations base name
 ERASE_REPUTATION = False
@@ -50,13 +51,13 @@ def get_needle_in_haystack(starting_mean):
   needle_in_haystack[int(K/2)] = bernoulli(starting_mean + 0.2)
   return needle_in_haystack
 
+def gen_rand_instance():
+  return [bernoulli(np.random.rand()) for i in xrange(np.random.randint(10, 20))]
+
 heavy_tail_prior = beta(0.6, 0.6)
 
 BANDIT_DISTR = {
-  'Needle In Haystack - 0.5': get_needle_in_haystack(0.5),
-  'Heavy Tail': heavy_tail_prior,
-  'Uniform': None,
-  '.5/.7 Random Draw': None
+  'Uniform': None
 }
 
 WORKING_DIRECTORY = ''
@@ -76,6 +77,7 @@ dist_name = exp_base_name + '_dist.csv'
 
 INDIVIDUAL_FIELD_NAMES =['Prior', 'P1 Alg', 'EEOG', 'Instance Complexity', 'Reputation Erased', 'P2 Alg', 'Time Horizon', 'Agent Alg', 'Market Share for P1', 'P1 Regret', 'P2 Regret', 'P1 Reputation', 'P2 Reputation', 'Abs Delta Regret']
 
+# fetch distributions from previous run
 def fetch_distributions(filename, priorname):
   realDistributions = {}
   count = 0
[…]

### 2018-07-31 preliminary_reputation.py

@@ -19,7 +19,7 @@ from scipy.stats import bernoulli, beta
 
 
 T = 5001
-N = 150
+N = 500
 K = 10
 
 
@@ -54,27 +54,33 @@ needle_in_haystack_50_high[int(K/2)] = bernoulli(default_mean + 0.2)
 #needle_in_haystack_50_one_medium_one_high[int(K/2)] = bernoulli(default_mean + 0.3)
 #needle_in_haystack_50_one_medium_one_high[int(K/2) + 1] = bernoulli(default_mean + 0.1)
 
-
-
 heavy_tail_prior = beta(0.6, 0.6)
 heavy_tailed = [bernoulli(heavy_tail_prior.rvs()) for i in xrange(K)]
 
+def get_needle_in_haystack(starting_mean):
+  needle_in_haystack = [bernoulli(starting_mean) for i in xrange(K)]
+  needle_in_haystack[int(K/2)] = bernoulli(starting_mean + 0.2)
+  return needle_in_haystack
 
 ALGS = [ThompsonSampling, DynamicEpsilonGreedy, DynamicGreedy]
 BANDIT_DISTR = {
-  'Needle In Haystack High': needle_in_haystack_50_high,
-  'Uniform': None,
-  'Heavy Tail': heavy_tail_prior
+  '.5/.7 Random Draw': None,
+  'Needle In Haystack - 0.1': get_needle_in_haystack(0.1),
+  'Needle In Haystack - 0.3': get_needle_in_haystack(0.3)
 }
 
-
+WORKING_DIRECTORY = ''
+#WORKING_DIRECTORY = '/rigel/home/ga2449/bandits-rl-project/'
 # Algorithm, Arms, Prior, t, n, reward
+
+RESULTS_DIR = WORKING_DIRECTORY + 'results/preliminary_raw_results/'
+FILENAME = 'preliminary_plots_all_haystack.csv'
 
 
 FIELDNAMES = ['True Mean Reputation', 'Realized Reputation', 'Algorithm', 'K', 'Distribution', 't', 'Instantaneous Realized Reward Mean', 'Instantaneous Realized Reward Std', 'Cumulative Realized Reward Mean', 'Cumulative Realized Reward Std', 'Instantaneous Mean Reward Mean', 'Instantaneous Mean Reward Std', 'Cumulative Mean Reward Mean', 'Cumulative Mean Reward Std', 'Best Arm Mean']
 simResults = {}
 
[…]

### 2018-07-31 preliminary_reputation_2.py

@@ -20,10 +20,10 @@ from lib.bandit.ExploreThenExploit import ExploreThenExploit
 from scipy.stats import bernoulli, beta
 
 
-T = 5001
-N = 750
-K = 3
-numCores = 12
+T = 1000
+N = 250
+K = 10
+numCores = 2
 
 
 DEFAULT_COMMON_PRIOR = [beta(1, 1) for k in xrange(K)]
@@ -59,18 +59,16 @@ def get_needle_in_haystack(starting_mean):
 ALGS = [DynamicGreedy, ThompsonSampling, DynamicEpsilonGreedy]
 BANDIT_DISTR = {
   'Heavy Tail': heavy_tail_prior,
-  'Uniform': None,
-  '.5/.7 Random Draw': None,
-  'Needle In Haystack - 0.5': get_needle_in_haystack(0.5)
+  'Uniform': None
 }
 
 
 WORKING_DIRECTORY = ''
-WORKING_DIRECTORY = '/rigel/home/ga2449/bandits-rl-project/'
+#WORKING_DIRECTORY = '/rigel/home/ga2449/bandits-rl-project/'
 # Algorithm, Arms, Prior, t, n, reward
 
 RESULTS_DIR = WORKING_DIRECTORY + 'results/preliminary_raw_results/'
-FILENAME = 'preliminary_plots_full_path.csv'
+FILENAME = 'preliminary_plots_10_arms_small_window.csv'
 
 
 FIELDNAMES = ['Realized Complexity', 'n', 'True Mean Reputation', 'Realized Reputation', 'Algorithm', 'K', 'Distribution', 't', 'Instantaneous Realized Reward Mean', 'Instantaneous Mean Reward Mean', 'Arm Means']
@@ -94,7 +92,7 @@ with open(RESULTS_DIR + FILENAME, 'w') as csvfile:
         banditDistrs[i] = [bernoulli(np.random.choice([0.5, 0.7])) for j in xrange(K)]
       else:
         banditDistrs[i] = banditDistr
-      
+
       realizations[i] = {}
       for t in xrange(T):
         realizations[i][t] = {}
@@ -104,7 +102,7 @@ with open(RESULTS_DIR + FILENAME, 'w') as csvfile:
     for a in xrange(len(ALGS)):
       alg = ALGS[a]
       simResults[alg] = Parallel(n_jobs=numCores)(delayed(sim)(alg, banditDistrs[j], realizations[j], j) for j in xrange(N))
[…]

### 2018-07-31 simulate.py

@@ -21,11 +21,44 @@ def getRealDistributionsFromPrior(priorName, prior, K):
     return [bernoulli(prior.rvs()) for i in xrange(K)]
   elif priorName == '.5/.7 Random Draw':
     return [bernoulli(random.choice([0.5, 0.7])) for i in xrange(K)]
+  elif priorName == 'Complexity':
+    return [bernoulli(np.random.rand()) for i in xrange(K)]
   return prior
 
-# realDistributions - the true distribution of the arms
-# principalPriors - the priors the principals have over the arms
-# agentPriors - the priors the agents have for the distribution of rewards from each principal
+"""
+Parameters
+  ----------
+  principalAlg1 / principalAlg2 : BanditAlgorithm object
+    The algorithm played by principal 1 and 2, respectively
+  agentAlg : Agent object
+    The decision rule utilized by the agents
+  K : Int
+    The number of arms in the bandit instance
+  T : Int
+    The time horizon for the competition game
+  memory: Int
+    Sliding window length for the reputation score
+  discountFactor: Float
+    Discount factor used in reputation score calculation
+  realDistributions: Distributions vector
+    The distributions of the K arms
+  warmStartRealizations:
+    The realizations for the warm start - realizations for each arm in each warm start round
+  realizations:
+    Realizations for each arm for T rounds
+  freeObsForP2: Boolean
+    Used for the incumbent experiment - does principal 2 (the incumbent) get free agents before the competition game?
+  freeObsNum: Int
+    For how many rounds is principal 2 the incumbent?
+  principal1Priors / principal2Priors: Vector of Distributions
+    Initial beliefs of the principals
+  recordStatsAt: Vector of times
+    We record the state of the game at these times
+  seed: Float
+    Used to reinitialize the seed for the random draws - important when using parallelization
+  eraseReputation: Boolean
+    A boolen flag on whether to erase reputation or not
+"""
 def simulate(principalAlg1, principalAlg2, agentAlg, K, T,
   memory=DEFAULT_MEMORY,
   discountFactor=DEFAULT_DISCOUNT_FACTOR,
@@ -34,7 +67,6 @@ def simulate(principalAlg1, principalAlg2, agentAlg, K, T,
   realizations=None,
   freeObsForP2=False,
   freeObsNum=100,
-  alpha=DEFAULT_ALPHA,
   warmStartNumObservations=DEFAULT_WARM_START_NUM_OBSERVATIONS,
   principal1Priors=None,
   principal2Priors=None,

### 2018-07-31 tests/AgentsTest.py

@@ -1,10 +1,14 @@
 import unittest
 
-import sys 
+import sys
+import numpy as np
+from scipy.stats import bernoulli, beta
 sys.path.append('..')
 from lib.agent.HardMax import HardMax
 from lib.agent.HardMaxWithRandom import HardMaxWithRandom
 from lib.InformationSet import InformationSet
+from lib.bandit.DynamicGreedy import DynamicGreedy
+from lib.BanditProblemInstance import BanditProblemInstance
 
 class TestAgents(unittest.TestCase):
 
@@ -20,6 +24,17 @@ class TestAgents(unittest.TestCase):
     self.assertEqual(info.getMaxPrincipalsAndScores(), (['principal1'], 2.0))
     info.updateInformationSet(-100, 1, 'principal1')
     self.assertEqual(info.getMaxPrincipalsAndScores(), (['principal2'], 0.0))
+
+  def test_reputation(self):
+    T = 1000
+    prob_instance = BanditProblemInstance(1, T, [bernoulli(0.5)])
+    principal = { 'principal1': DynamicGreedy(prob_instance, [beta(1, 1)]) }
+    agent = HardMax(principal, 1)
+    for t in xrange(T):
+        (reward, arm) = principal['principal1'].executeStep(t)
+        agent.updateInformationSet(reward, arm, 'principal1')
+    self.assertAlmostEqual(np.sum(principal['principal1'].realizedRewardHistory), T/2, delta=50)
+    self.assertAlmostEqual(agent.informationSet.getScores()['principal1'], 0.5, delta=0.05)
 
   def test_hard_max(self):
     principals = { 'principal1': None, 'principal2': None }

### 2018-08-03 paper/competing_bandits.tex
commit skeleton of paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}

% ready for submission
\usepackage{aaai}

% to compile a preprint version, e.g., for submission to arXiv, add
% add the [preprint] option:
% \usepackage[preprint]{nips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2018}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amssymb}
\usepackage{natbib}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{float}
\usepackage{sgame, tikz} % Game theory packages
\usepackage{caption} 
\usepackage{algorithm,algpseudocode}
\usepackage{makecell}
 \usepackage{multirow}
 \usepackage{graphicx}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\captionsetup{font=footnotesize}
\bibliographystyle{aaai}

\title{Competing Bandits}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

 \author{-}
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\

\begin{document}

\maketitle

\begin{abstract}
...
\end{abstract}

\section{Introduction}
\label{S:1}

...

\section{Related Work}
\label{S:2}

\citealt*{mansour2017competing}, \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}, etc.

\section{Model}
\label{S:3}

...

\section{Simulation Details}
\label{S:4}

\subsection{Bandit Priors}
We study the implications of our model via simulation. For computational tractability we focus our attention on Bernoulli bandits with beta priors. We consider the following bandit ``priors" from which the bandit instances faced in simulation are drawn:
\begin{enumerate}
\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$). With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
\item Needle-in-haystack High - $K-1$ arms with mean 0.50, 1 arm with mean 0.70. 
\end{enumerate}

[We can't feasibly include all the priors. I think Heavy Tail is interesting because of the relative reputation plots, but I'm not sure what other one to include that would be meaningful. For now I've just used NIH]

In the appendix we consider the performance of additional priors, but qualitatively the results presented are unchanged.

\subsection{Learning Algorithms}

For our simulations we evaluate the performance of three algorithms, each endowed with a ``fake prior" of $Beta(1, 1)$. In general, there are three different classes of learning algorithms and we select one representative algorithm from each class: 
\begin{enumerate}
\item Adaptive exploration algorithms that engage in purposeful exploration in a ``smart" way by adapting to the previous history of results. We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is ``optimal" in the sense of having the highest mean reward.
\item Non-adaptive exploration algorithms that engage in purposeful exploration without considering the past results. We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability. For our experiments we keep $\epsilon = 0.05$ fixed.
\item Greedy / Myopic algorithms that engage in no purposeful exploration and take the best short-sighted action. We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
\end{enumerate}

Run in isolation we expect that the performance of the algorithms (according to cumulative regret) should be \\
 Adaptive Exploration $>=$ Non-adaptive exploration $>=$ Greedy. We verify that, for the chosen algorithms and chosen priors, this is indeed the case (with some caveats, discussed in the next section).

The primary question that we are interested in is under what conditions in competition are the firms in our model incentivized to commit to adaptive exploration algorithms.

\subsection{Simulation of Competition Game}
Algorithm \ref{alg_1} (FIX TO BE FIGURE) describes the simulation procedure utilized to evaluate the results of the competition game. $T$ is the finite time horizon of the competition game and $K$ is the number of arms in the bandit instance that is considered.

\begin{algorithm}
\caption{Simulation Pseudo-Code}
\begin{algorithmic}[1]
\For{Each prior $p$}
\State Generate true distribution from $p$
\State Generate $T \times K$ realizations for the arms 
	\For{Each agent algorithm $agent alg$}
		\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
			\For{$N$ simulations}
				\State Give principal 2 $X$ free observations (the agents also get these observations)				
				\State Give the agents $k$ observations from each principal (warm start)
				\State Run simulation for $T$ periods
			\EndFor
		\EndFor
	\EndFor
\EndFor
\end{algorithmic}
\label{alg_1}
\end{algorithm}

\section{Performance in Isolation}
\label{S:5}

We look at the performance of each of our learning algorithms in isolation, meaning that we simply run the learning algorithm for $T$ steps on a realization of the prior instances that we consider and record the realized rewards and the arm that is pulled. In our model the reputation score is an input to the decision rule of the agents and thus this statistic is an important property of the learning algorithm selected by the arm. Figure \ref{prelim_means} displays the mean average reputation score of each of the algorithms across the different simulations.

\begin{figure}
\caption{Reputation Trajectories in Isolation}
\includegraphics[scale=0.2]{"figures/Reputation Trajectory for Heavy Tail 10 arms"}
\includegraphics[scale=0.2]{"figures/Reputation Trajectory for Needle In Haystack High 10 arms"}
\label{prelim_means}
\caption*{\tiny{The plots contain the average reputation over $250$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs. TODO: get rid of smoothing on plot}}
\end{figure}

\begin{figure}
\caption{Relative Reputation Plots}
\includegraphics[scale=0.25]{"figures/deg_dg_ht_10_prelim"}
\includegraphics[scale=0.25]{"figures/ts_dg_ht_10_prelim"} \\
\includegraphics[scale=0.25]{"figures/deg_dg_nih_10_prelim"}
\includegraphics[scale=0.25]{"figures/ts_dg_nih_10_prelim"} \\
\caption*{\tiny{The plots contain the proportion of simulations over $250$ runs for a memory size of $100$ where, for a given $t$, the reputation of one algorithm is at least as high as the reputation of another on the same instance.}}
\label{relative_rep}
\end{figure}

As expected we see that according to the mean reputation plots $TS > DEG > DG$ for sufficiently large values of $t$. In the competition game we have defined, however, another relevant statistic is the relative comparison of reputation values across algorithms. Figure \ref{prelim_comparison} displays the proportion of rounds where one algorithm has at least as high of a reputation as another. Interestingly we observe that, on the Heavy Tail prior, $DG > DEG$ according to this metric even though $DEG > DG$ according to the mean reputation. The intuition for this is simple. $DG$ will either find the best arm or not. If it does then since $DEG$ engages in non-adaptive exploration $DG$ will, with $\epsilon$ probability, pick an arm uniformly at random which potentially hurts its reputation if the selected arm is not the best arm. $DEG$ eventually converges on the best arm with a sufficiently large time horizon, but $DG$ does not necessarily do so and so when $DG$ does not find the best arm it may potentially have substantially worse reputation which drags down its mean reputation. However, if it is able to find the best arm in sufficiently many simulations then it may have slightly higher reputation than $DEG$ in those simulations and thus do better according to the relative reputation proportion but, due to the simulations where it does not find the best arm it may then have worse mean reputation.

[Would be nice to get at what is the driving factor behind this. Right now, we simply observe it.]

\section{Simultaneous Entry Experiment}
\label{S:6}

In this section we present simulation results of the model discussed in Section \ref{S:2} where $X = 0$ so that both firms enter at the same time. We want to understand what algorithms the firms are incentivized to commit to in the competition game and so we take the algorithms as given and study the consequences of playing them in the competition game. For this experiment we fix the $HardMax$ agent decision rule. Under this decision rule we expect that one firm will end up taking the market in most of the simulations and so when reasoning about the decision rule we track the effective end of game.
[Reword this]

\begin{definition}
\textit{Effective End of Game (EEOG)} - the last round, $t$ in a simulation where there is a difference between the decision of the agent alive at $t-1$ and the agent alive at $t$ (todo: state this formally using notation of model)
\end{definition}

\begin{table}[ht]
\centering
\caption{Needle In Haystack Prior} 
\begin{tabular}{rlll}
  \hline
Alg 1 vs 2 & k = 20 & k = 100 & k = 200 \\ 
  \hline
TS vs DG & \makecell{\textbf{0.66} $\pm$0.05\\ eeog \\ avg: 330\\ med: 31} & \makecell{\textbf{0.61} $\pm$0.05\\ eeog \\ avg: 590\\ med: 61.5} &  \\
\hline 
  TS vs DEG & \makecell{\textbf{0.58} $\pm$0.06\\ eeog \\ avg: 180\\ med: 1} & \makecell{\textbf{0.54} $\pm$0.05\\ eeog \\ avg: 670\\ med: 60} &  \\ 
  \hline 
  DG vs DEG & \makecell{\textbf{0.41} $\pm$0.05\\ eeog \\ avg: 650\\ med: 182} & \makecell{\textbf{0.41} $\pm$0.05\\ eeog \\ avg: 1500\\ med: 672} &  \\ 
   \hline
   \label{hm_nih}
   \end{tabular}
\caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95 \% confidence band. For example, the cell in the top left indicates that TS gets on average 32\% of the market when played against DG. The next line contain the average and median effective end of game for this set of simulations.}}
\end{table}

\begin{table}[ht]
\centering
\caption{Heavy Tail Prior} 
\begin{tabular}{rlll}
  \hline
Alg 1 vs 2 & k = 20 & k = 100 & k = 200 \\ 
  \hline
TS vs DG & \makecell{\textbf{0.32} $\pm$0.05\\ eeog \\ avg: 150\\ med: 0} & \makecell{\textbf{0.55} $\pm$0.05\\ eeog \\ avg: 300\\ med: 0} & \makecell{\textbf{0.73} $\pm$0.04\\ eeog \\ avg: 1400\\ med: 226.5} \\ 
\hline
  TS vs DEG & \makecell{\textbf{0.34} $\pm$0.05\\ eeog \\ avg: 36\\ med: 0} & \makecell{\textbf{0.65} $\pm$0.05\\ eeog \\ avg: 330\\ med: 0} & \makecell{\textbf{0.91} $\pm$0.02\\ eeog \\ avg: 1100\\ med: 107.5} \\ 
  \hline
  DG vs DEG & \makecell{\textbf{0.62} $\pm$0.05\\ eeog \\ avg: 800\\ med: 10} & \makecell{\textbf{0.6} $\pm$0.05\\ eeog \\ avg: 1500\\ med: 866.5} & \makecell{\textbf{0.62} $\pm$0.05\\ eeog \\ avg: 1900\\ med: 1627.5} \\ 
   \hline
   \label{hm_ht}
\end{tabular}
\end{table}

Tables \ref{hm_nih} and \ref{hm_ht} display the results. Under Heavy Tail we see that for small warm start of $k = 20$, $DG$ gives the firm more of the market than $TS$ and $DEG$. As we increase the warm start to $k= 100, 200$ we observe that $TS$ does better than $DG$ and $DEG$. Note that, in particular for the simulations involving $TS$, the effective end of game had a median of 0. This means that the result of the simulation was \textit{completely determined by the warm start period}. Thus, when reasoning about what algorithm would win in the competition game under the HardMax decision rule we need to think about the performance of the algorithm on $k$ samples, which may be small. For low warm start, it is thus not necessarily the case that $TS$ is the best algorithm. The intuition for this is that $TS$ needs to engage in pure exploration early on in order to identify the best arm. A consequence of exploration in our model is that it may lead to a reputational cost relative to the greedy algorithm since it involves experimenting with potentially sub-optimal arms relative to the current belief of the best arm. However, a sufficiently large warm start allows $TS$ to incur this reputational cost without the informational gain.

Another interesting result is that $DG$ still does better than $DEG$ even with a large warm start under the Heavy Tail prior. In general, we observe that even under other agent response functions besides HardMax that $DG$ does better than $DEG$. Thus, even though the mean reputation plot in Figure \ref{prelim_means} shows that $DEG$ does better than $DG$ in terms of average mean reputation, this does not lead to $DEG$ doing better under competition. Rather, the relative reputation plot shown in Figure \ref{relative_rep} is more predictive in this case of what algorithm wins under competition.

Looking at the results for Needle In a Haystack we observe that $TS$ always does better than $DEG, DG$. T

[TODO: is there something more interesting / general that we can discuss here besides this? Why does TS win? Learning is easy? Something completely unrelated to learning hardness?]

[TODO: Can we say anything general about this using the ``learning hardness" metric that we have started looking at?]

\section{Incumbent Experiment}
\label{S:7}

We now re-run the same simulations but let $X = 200$ so that firm 2 has an ``incumbency" advantage. Recall that $X$ controls the number of rounds the firm is in the market by itself before the other firm enters and so in these rounds agents arrive and must choose firm 2. After $X$ rounds, firm 1 enters the market and each agent that enters in subsequent periods must select between the two firms.

\begin{table}[ht]
\label{ht_incum}
\centering
\caption{Incumbent Experiment for Heavy Tail}
\begin{tabular}{rlll}
\multicolumn{3}{c}{Incumbent Algorithm} \\
\multirow{3}{*}{\rotatebox{90}{Entrant Algorithm}} \\
\hline
 & TS & DEG &  DG \\ 
 \hline
TS & \makecell{\textbf{ 0.0067 } $\pm$ 0.0092 \\Variance: 0.007 \\ ES: 100 \%} & \makecell{\textbf{ 0.023 } $\pm$ 0.017 \\Variance: 0.02 \\ ES: 99 \%} & \makecell{\textbf{ 0.064 } $\pm$ 0.027 \\Variance: 0.05 \\ ES: 97 \%} \\ 
\hline
  DEG & \makecell{\textbf{ 0.024 } $\pm$ 0.015 \\Variance: 0.02 \\ ES: 98 \%} & \makecell{\textbf{ 0.13 } $\pm$ 0.034 \\Variance: 0.09 \\ ES: 86 \%} & \makecell{\textbf{ 0.14 } $\pm$ 0.036 \\Variance: 0.1 \\ ES: 89 \%} \\ 
\hline   
   DG & \makecell{\textbf{ 0.063 } $\pm$ 0.024 \\Variance: 0.04 \\ ES: 93 \%} & \makecell{\textbf{ 0.19 } $\pm$ 0.041 \\Variance: 0.1 \\ ES: 91 \%} & \makecell{\textbf{ 0.15 } $\pm$ 0.032 \\Variance: 0.08 \\ ES: 77 \%} \\ 
\end{tabular}
\caption*{\tiny{TODO: fix table names and general table alignment. The first line in each cell contains the average market share for the entrant over $N=250$ simulations as well as a 95\% confidence interval. The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one principal getting $> 90\%$ of the market. Note that smaller values in the table are better for the incumbent. Market shares are calculated as the fraction of users selecting a particular firm \textit{after} the entrant has already entered (i.e. the free rounds to firm 2 do not count towards the share)}}
\end{table}

Table \ref{ht_incum} shows the results of the simulations for the Heavy Tail prior. $TS$ is a dominant strategy for the incumbent when $X$ is sufficiently high. In the appendix we include additional simulations where $X = 50$ and observe that $TS$ is not always a dominant strategy in this case. These results are robust across the different priors we considered as well as across agent response models. Namely, for sufficiently large $X$, $TS$ is a dominant strategy for the incumbent.

Whereas with simultaneous entry $TS$ was a worse strategy than $DG$ (for small warm start), by allowing one firm to have a first mover advantage and be a monopolist in the market in the early rounds we incentivize the monopolist to play the ``best" algorithm. The intuition for this is that competition forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration. In some sense one can view allowing one firm to temporarily be a monopolist as temporarily relaxing the ``incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation. The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition. As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.

[This section merits much more explanation]

\section{Data and Reputation \\ as Barriers to Entry}
\label{S:8}

As the simulation results of the incumbent experiment in Table \ref{ht_incum} and in the appendix show, the incumbent usually ends up with a substantial share of the market over the entrant. The rounds where the incumbent was a monopolist provided the incumbent with free data for learning (data advantage) as well as a chance to establish a reputational advantage over a potential entrant (reputational advantage). Which plays a bigger role in preventing the entrant from being able to establish market share? We run two additional experiments, modifying the previous incumbent experiment so that in one set of simulations the reputation of the incumbent is artificially erased and another in which the information gained by the incumbent is artificially erased so that the posterior is reset to the prior.

\begin{table}[ht]
\centering
\caption{Reputation Erased Experiment} 
\begin{tabular}{rlll}
  \hline
 & TS & DEG &  DG \\ 
  \hline
TS & \makecell{ \textbf{ 0.18 } $\pm$ 0.042 \\Var:  0.1 \\ ES: 97 \% } & \makecell{ \textbf{ 0.17 } $\pm$ 0.042 \\Var:  0.1 \\ ES: 98 \% } & \makecell{ \textbf{ 0.23 } $\pm$ 0.047 \\Var:  0.2 \\ ES: 98 \% } \\ 
  DEG & \makecell{ \textbf{ 0.23 } $\pm$ 0.045 \\Var:  0.2 \\ ES: 92 \% } & \makecell{ \textbf{ 0.32 } $\pm$ 0.049 \\Var:  0.2 \\ ES: 86 \% } & \makecell{ \textbf{ 0.25 } $\pm$ 0.046 \\Var:  0.2 \\ ES: 90 \% } \\ 
   DG & \makecell{ \textbf{ 0.28 } $\pm$ 0.047 \\Var:  0.2 \\ ES: 87 \% } & \makecell{ \textbf{ 0.33 } $\pm$ 0.051 \\Var:  0.2 \\ ES: 91 \% } & \makecell{ \textbf{ 0.3 } $\pm$ 0.047 \\Var:  0.2 \\ ES: 83 \% } \\ 
   \hline
   \label{rep_erase}
\end{tabular}
\end{table}
% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Sun Jul 22 14:23:06 2018
\begin{table}[ht]
\centering
\caption{Information Erased Experiment} 
\begin{tabular}{rlll}
  \hline
 & TS & DEG &  DG \\ 
  \hline
TS & \makecell{ \textbf{ 0.059 } $\pm$ 0.025 \\Var:  0.05 \\ ES: 97 \% } & \makecell{ \textbf{ 0.11 } $\pm$ 0.033 \\Var:  0.09 \\ ES: 94 \% } & \makecell{ \textbf{ 0.15 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 92 \% } \\ 
  DEG & \makecell{ \textbf{ 0.11 } $\pm$ 0.031 \\Var:  0.08 \\ ES: 89 \% } & \makecell{ \textbf{ 0.19 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 82 \% } & \makecell{ \textbf{ 0.18 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 82 \% } \\ 
   DG & \makecell{ \textbf{ 0.17 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 87 \% } & \makecell{ \textbf{ 0.22 } $\pm$ 0.042 \\Var:  0.1 \\ ES: 82 \% } & \makecell{ \textbf{ 0.26 } $\pm$ 0.043 \\Var:  0.1 \\ ES: 77 \% } \\ 
   \hline
   \label{info_erase}
\end{tabular}
\end{table}

Tables \ref{rep_erase} and \ref{info_erase} display the results of these experiments. Even with the data advantage or reputation advantage alone the incumbent ends up getting a substantial portion of the market, regardless of the algorithm played. However, fixing the same algorithm played by the incumbent and the entrant, the market share for the incumbent is always higher when the incumbent retains her reputation instead of her information. The intuition for this is that the reputational advantage protects the incumbent from ``bad" decisions or bad luck on rewards that adversely affect reputation and thus, especially in the earlier rounds, agents in the earlier rounds will still choose the incumbent which will allow the incumbent to regain the information they lost. However, with reputation re-initialized the better information does not protect the incumbent from getting unlucky with her rewards. Looking at the variability in the shares in the two experiments indicates that not only does the incumbent get a smaller market share on average in the reputation erased experiment, but the variability in the shares is also higher.

Though the setup is purely experimental, it is nonetheless interesting to look at if $TS$ still remains as a dominant strategy for the incumbent. In the information erased experiment, $TS$ is still a dominant strategy while in the reputation erased experiment it is not a dominant strategy. The intuition for this is that in the reputation erased experiment, the incumbent still has the data that she accumulated during her time as a monopolist and as a result, there is less value to an algorithm that engages in pure exploration as she wants to rebuild her reputation and thus might be better off exploiting the information she has in order to rebuild reputation instead of engaging in exploration and potentially damage her reputation. However, in the information erased experiment, she has a reputational cushion that allows for exploration so that she can regain the information that she lost and thus $TS$ remains a dominant strategy.

In the appendix we include the results of the same experiment but with different agent response models. In these experiments we get that information and reputation serve as substitutes for each other

[TODO: leave this, expound on this, or throw it out?]

Thus, we see that both information and reputation advantages alone can serve as substantial barriers to entry in our model. However, under the HardMax agent response model, we see that reputation serves as more of a barrier to entry compared to information.

\section{Varying the agent response model}
\label{S:9}

Thus far we have largely presented simulation results utilizing the deterministic HardMax rule. This is a natural decision rule where agents simply use the reputation score as a strict decision-making rule. In this section we consider the implications of adding randomness to the agent rule. What are the implications on what algorithm the firms are incentivized to play? How does the distribution of market share change?

[Discuss that, under random agent models, firms that play better learning algorithms do better in the long-run since they get enough free agents to move to the part of the relative reputation curve where better algorithms win]

[Discuss that, under HMR, market share variance is lower and see less extreme market shares.]

[Discuss that, under SM, market share variance is very low and market shares are around 50/50].

\section{Conclusion}
\label{S:10}

...

\pagebreak
\bibliography{refs}

\end{document}

### 2018-08-03 paper/competing_bandits.tex
updates to paper

@@
 With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
+[We should rename this prior instance]
 \item Needle-in-haystack High - $K-1$ arms with mean 0.50, 1 arm with mean 0.70. 
 \end{enumerate}
+
+[Need some discussion of what makes these priors at all ``representative" - why can we just focus on them?]
 
@@
 We look at the performance of each of our learning algorithms in isolation, meaning that we simply run the learning algorithm for $T$ steps on a realization of the prior instances that we consider and record the realized rewards and the arm that is pulled.
-In our model the reputation score is an input to the decision rule of the agents and thus this statistic is an important property of the learning algorithm selected by the arm.
+In our simulations the realizations of each of the arms were pre-drawn and the same across all algorithms and so the differences in algorithm performance come from the difference in the arms selected by the algorithms and not from differences in realizations.
+When looking at the results of each of the learning algorithms in isolation we want to confirm that, for the prior instances that we consider, ``better" algorithms win according to ``absolute" measures of performance (i.e.
+looking at mean reputation).
+Additionally, we want to understand what statistics we can look at in isolation to understand what wins under competition.
+
+In our model the reputation score is an input to the decision rule of the agents and thus it is important to understand how this score differs across algorithms when run outside of competition.
 Figure \ref{prelim_means} displays the mean average reputation score of each of the algorithms across the different simulations.
@@
 As expected we see that according to the mean reputation plots $TS > DEG > DG$ for sufficiently large values of $t$.
-In the competition game we have defined, however, another relevant statistic is the relative comparison of reputation values across algorithms.
+While looking at the mean reputation is a natural statistic to consider when attempting to understand what algorithm wins under competition, we consider an additional statistic when attempting to understand what wins under competition.
+In the competition game the decision rule of the agents compares the relative reputation of the two firms so that if the reputation of firm $A$ is higher than that of firm $B$, the agents will (be more likely to) select firm $A$.
+Thus, another natural statistic to look at is to look at the two algorithms in isolation and consider the fraction of simulations where, for a fixed time $t$, the reputation of firm $A$ is at least as high as that of firm $B$.
 Figure \ref{prelim_comparison} displays the proportion of rounds where one algorithm has at least as high of a reputation as another.
+
 Interestingly we observe that, on the Heavy Tail prior, $DG > DEG$ according to this metric even though $DEG > DG$ according to the mean reputation.
@@
 However, if it is able to find the best arm in sufficiently many simulations then it may have slightly higher reputation than $DEG$ in those simulations and thus do better according to the relative reputation proportion but, due to the simulations where it does not find the best arm it may then have worse mean reputation.
-
-[Would be nice to get at what is the driving factor behind this. Right now, we simply observe it.]
+In the next section we move to considering the competition game and revisit if this statistic is more useful for predicting who wins in competition.
 
@@
 In this section we present simulation results of the model discussed in Section \ref{S:2} where $X = 0$ so that both firms enter at the same time.
-We want to understand what algorithms the firms are incentivized to commit to in the competition game and so we take the algorithms as given and study the consequences of playing them in the competition game.
-For this experiment we fix the $HardMax$ agent decision rule.
-Under this decision rule we expect that one firm will end up taking the market in most of the simulations and so when reasoning about the decision rule we track the effective end of game.
-[Reword this]
+We want to understand what algorithms the firms are incentivized to commit to in the competition game but the payoffs associated with each algorithm are too difficult to compete analytically and thus, we begin by simply imposing that the algorithm played by the firm is taken as given and study the consequences of playing them in the competition game.
+Additionally, for this experiment we present results for the $HardMax$ agent decision rule and postpone discussion of alternative decision rules.
+Besides keeping track of market share, we also will keep track of the effective end of game in order to better understand what happens in the game.
 
@@
 
+\begin{figure}
+\label{rel_rep_ht_fine}
+\caption{Relative Reputation - Heavy Tail}
+\includegraphics[scale=0.3]{figures/ts_dg_heavy_fine}
+\caption*{TODO: Reduce the range of the x-axis of this so that you can understand when TS starts to do better than DG}
+\end{figure}
+
 Tables \ref{hm_nih} and \ref{hm_ht} display the results.
 Under Heavy Tail we see that for small warm start of $k = 20$, $DG$ gives the firm more of the market than $TS$ and $DEG$.
-As we increase the warm start to $k= 100, 200$ we observe that $TS$ does better than $DG$ and $DEG$.
+As we increase the warm start to $k = 100, 200$ we observe that $TS$ does better than $DG$ and $DEG$.
 Note that, in particular for the simulations involving $TS$, the effective end of game had a median of 0.
 This means that the result of the simulation was \textit{completely determined by the warm start period}.
-Thus, when reasoning about what algorithm would win in the competition game under the HardMax decision rule we need to think about the performance of the algorithm on $k$ samples, which may be small.
+Thus, when reasoning about what algorithm would win in the competition game under the HardMax decision rule we need to think about the performance of the algorithm on $k$ samples instead of $T$ samples, which may be small.
 For low warm start, it is thus not necessarily the case that $TS$ is the best algorithm.
-The intuition for this is that $TS$ needs to engage in pure exploration early on in order to identify the best arm.
-A consequence of exploration in our model is that it may lead to a reputational cost relative to the greedy algorithm since it involves experimenting with potentially sub-optimal arms relative to the current belief of the best arm.
-However, a sufficiently large warm start allows $TS$ to incur this reputational cost without the informational gain.
+However, we see that under the Needle In Haystack prior $TS$ is always the best algorithm across all warm start values.
 
-Another interesting result is that $DG$ still does better than $DEG$ even with a large warm start under the Heavy Tail prior.
-In general, we observe that even under other agent response functions besides HardMax that $DG$ does better than $DEG$.
-Thus, even though the mean reputation plot in Figure \ref{prelim_means} shows that $DEG$ does better than $DG$ in terms of average mean reputation, this does not lead to $DEG$ doing better under competition.
-Rather, the relative reputation plot shown in Figure \ref{relative_rep} is more predictive in this case of what algorithm wins under competition.
+[Missing here is a discussion of the fact that with HardMax the market gets taken by one firm completely (in most simulations) as well as why this is the case.]
 
-Looking at the results for Needle In a Haystack we observe that $TS$ always does better than $DEG, DG$. T
+First, we note that the relative reputation plots were better at predicting the result of the competition game than the ``absolute" mean reputation plots.
+For the Needle In Haystack prior, both plots showed that $TS > DEG > DG$ in isolation and this is also the result of the competition game.
+However, for the Heavy Tail prior, we see that $DG, DEG > TS$ for low warm start and that, in general, $DEG > DG$ \footnote{The result that $DEG > DG$ for this prior is robust across decision rules as well.}.
+While the ``absolute" reputation plots imply that $TS > DEG > DG$, we see that the relative reputation plot in Figure \ref{relative_rep} implies that $DEG > DG$ and in Figure \ref{rel_rep_ht_fine} we can see that in the early rounds, $DG$ does relatively better than $TS$ and that, around $t = 100$, $TS$ does relatively better than $DG$.
+Note that since the median effective end of game is $0$ in \ref{hm_ht} and the agents are using the $HardMax$ decision rule, this means that the relative reputation between the two firms after $k$ samples ends up determining who takes the market in the simulations and this is precisely what the relative reputation plots tell us.
+Thus, looking at the relative reputation plot at time $k$ gives us a good predictor of which algorithm will take more of the market in the competition game.
 
-[TODO: is there something more interesting / general that we can discuss here besides this? Why does TS win? Learning is easy? Something completely unrelated to learning hardness?]
+Why is it that we see that $TS$ always wins under Needle in Haystack but, for lower warm start, $DG > TS$?
+We know that for a sufficiently long time horizon, $TS$ leads to a higher reputation than $DG$.
+However, in our model, firms need to incentivize the agents to select them over their competition.
+One possible intuition for the results here are that, in the early rounds, $TS$ engages in purposeful exploration whereas $DG$ is myopic.
+A consequence of exploration in our model is that it may lead to a reputational cost relative to the greedy algorithm since the sub-optimal arms pulled during exploration can harm reputation relative to the greedy algorithm.
+$TS$ thus incurs a relative reputational cost in the early rounds, but once it has explored enough and gained enough information it is able to recover the relative reputational loss and acquire a larger reputation than the greedy algorithm.
+A sufficiently large warm start allows $TS$ to explore in the early rounds and recover the reputational loss and thus win the competition game whereas a low warm start is not sufficient for this and thus $DG$ beats $TS$.
+Why do we only see this on Heavy Tail and not Needle in Haystack?
+One possible explanation for this is that learning on Heavy Tail is harder (use learning hardness metric here) and learning on Needle In Haystack is easier and thus TS needs more samples in the Heavy Tail instances compared to the Needle In Haystack instances in order to be able to recover the relative reputation loss [TODO: Explain this a bit more and with more formality.
+Can we have some experimental results that ``prove" this or generalize it?]
 
@@
 As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
-
-[This section merits much more explanation]
 

### 2018-08-04 paper/competing_bandits.tex
updates to paper

@@
 
-We now re-run the same simulations but let $X = 200$ so that firm 2 has an ``incumbency" advantage.
+We now fix the same parameters as before, fixing a warm start of $k = 20$ [we should re-run this] but let $X = 200$ so that firm 2 has an ``incumbency" advantage.
 Recall that $X$ controls the number of rounds the firm is in the market by itself before the other firm enters and so in these rounds agents arrive and must choose firm 2.
-After $X$ rounds, firm 1 enters the market and each agent that enters in subsequent periods must select between the two firms.
+Importantly, we treat $X$ as an exogenous parameter (and not something optimally chosen by the entrant) and study the consequences of a fixed value of $X$.
+Thus, firm 2 is a monopolist for $X$ periods and, after $X$ rounds, firm 1 enters the market and each agent that enters in subsequent periods must select between the two firms.
+We want to understand what algorithms the incumbent (and the entrant) are incentivized to play.
+In particular, we are interested if introducing this asymmetry in entry induces firms who were previously incentivized to play $DG$ to play $TS$.
+As a result, we present the results of this experiment for $k = 20$ and on the Heavy Tail prior, where we saw that $DG$ was preferred to $TS$.
 
@@
 
-Table \ref{ht_incum} shows the results of the simulations for the Heavy Tail prior.
-$TS$ is a dominant strategy for the incumbent when $X$ is sufficiently high.
+Table \ref{ht_incum} shows the results of the simulations.
+We observe that, for the incumbent, $TS$ is a dominant strategy.
 In the appendix we include additional simulations where $X = 50$ and observe that $TS$ is not always a dominant strategy in this case.
-These results are robust across the different priors we considered as well as across agent response models.
+Qualitatively, these results are robust across the different priors we considered as well as across agent response models.
 Namely, for sufficiently large $X$, $TS$ is a dominant strategy for the incumbent.
 
-Whereas with simultaneous entry $TS$ was a worse strategy than $DG$ (for small warm start), by allowing one firm to have a first mover advantage and be a monopolist in the market in the early rounds we incentivize the monopolist to play the ``best" algorithm.
-The intuition for this is that competition forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration.
+Why do we observe this reversal?
+The intuition for this is that competition forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
+This intuition is very similar to that observed in the previous section by increasing warm start.
 In some sense one can view allowing one firm to temporarily be a monopolist as temporarily relaxing the ``incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
 The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.
-As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
+As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
+Thus, counterintuitively, by allowing one firm to be a monopoly and dominate the market, we can incentivize them to play $TS$ and thus get higher social welfare than under competition.
+[add some more economic intuition here]
+
+[Include some discussion about the entrant]
 

### 2018-08-07 paper/competing_bandits.tex
update some figures + some more writing

@@
 \centering
-\caption{Needle In Haystack Prior} 
+\caption{Simultaneous Entry Needle In Haystack} 
 \begin{tabular}{rlll}
   \hline
-Alg 1 vs 2 & k = 20 & k = 100 & k = 200 \\ 
+ & k = 20 & k = 100 & k = 200 \\ 
   \hline
-TS vs DG & \makecell{\textbf{0.66} $\pm$0.05\\ eeog \\ avg: 330\\ med: 31} & \makecell{\textbf{0.61} $\pm$0.05\\ eeog \\ avg: 590\\ med: 61.5} &  \\
-\hline 
-  TS vs DEG & \makecell{\textbf{0.58} $\pm$0.06\\ eeog \\ avg: 180\\ med: 1} & \makecell{\textbf{0.54} $\pm$0.05\\ eeog \\ avg: 670\\ med: 60} &  \\ 
-  \hline 
-  DG vs DEG & \makecell{\textbf{0.41} $\pm$0.05\\ eeog \\ avg: 650\\ med: 182} & \makecell{\textbf{0.41} $\pm$0.05\\ eeog \\ avg: 1500\\ med: 672} &  \\ 
+TS vs DG & \makecell{\textbf{0.86} $\pm$0.02\\ eeog \\ avg: 14\\ med: 0} & \makecell{\textbf{0.65} $\pm$0.02\\ eeog \\ avg: 11\\ med: 0} & \makecell{\textbf{0.59} $\pm$0.02\\ eeog \\ avg: 6\\ med: 0} \\
+  TS vs DEG & \makecell{\textbf{0.84} $\pm$0.02\\ eeog \\ avg: 20\\ med: 0} & \makecell{\textbf{0.62} $\pm$0.02\\ eeog \\ avg: 8\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 5.4\\ med: 0} \\
+  DG vs DEG & \makecell{\textbf{0.41} $\pm$0.02\\ eeog \\ avg: 400\\ med: 115} & \makecell{\textbf{0.47} $\pm$0.02\\ eeog \\ avg: 320\\ med: 0} & \makecell{\textbf{0.5} $\pm$0.02\\ eeog \\ avg: 290\\ med: 0} \\
    \hline
@@
 \centering
-\caption{Heavy Tail Prior} 
+\caption{Simultaneous Entry Heavy Tail} 
 \begin{tabular}{rlll}
   \hline
-Alg 1 vs 2 & k = 20 & k = 100 & k = 200 \\ 
+ & k = 20 & k = 100 & k = 200 \\ 
   \hline
-TS vs DG & \makecell{\textbf{0.32} $\pm$0.05\\ eeog \\ avg: 150\\ med: 0} & \makecell{\textbf{0.55} $\pm$0.05\\ eeog \\ avg: 300\\ med: 0} & \makecell{\textbf{0.73} $\pm$0.04\\ eeog \\ avg: 1400\\ med: 226.5} \\
-\hline
-  TS vs DEG & \makecell{\textbf{0.34} $\pm$0.05\\ eeog \\ avg: 36\\ med: 0} & \makecell{\textbf{0.65} $\pm$0.05\\ eeog \\ avg: 330\\ med: 0} & \makecell{\textbf{0.91} $\pm$0.02\\ eeog \\ avg: 1100\\ med: 107.5} \\
-  \hline
-  DG vs DEG & \makecell{\textbf{0.62} $\pm$0.05\\ eeog \\ avg: 800\\ med: 10} & \makecell{\textbf{0.6} $\pm$0.05\\ eeog \\ avg: 1500\\ med: 866.5} & \makecell{\textbf{0.62} $\pm$0.05\\ eeog \\ avg: 1900\\ med: 1627.5} \\
+TS vs DG & \makecell{\textbf{0.22} $\pm$0.02\\ eeog \\ avg: 0.27\\ med: 0} & \makecell{\textbf{0.41} $\pm$0.02\\ eeog \\ avg: 0.039\\ med: 0} & \makecell{\textbf{0.58} $\pm$0.02\\ eeog \\ avg: 12\\ med: 0} \\
+  TS vs DEG & \makecell{\textbf{0.24} $\pm$0.02\\ eeog \\ avg: 3.2\\ med: 0} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 39\\ med: 0} & \makecell{\textbf{0.91} $\pm$0.01\\ eeog \\ avg: 84\\ med: 0} \\
+  DG vs DEG & \makecell{\textbf{0.8} $\pm$0.02\\ eeog \\ avg: 16\\ med: 0} & \makecell{\textbf{0.86} $\pm$0.02\\ eeog \\ avg: 4.7\\ med: 0} & \makecell{\textbf{0.83} $\pm$0.02\\ eeog \\ avg: 2.7\\ med: 0} \\
    \hline
@@
 \end{figure}
+
+[IDEA: take a look at the distribution of reputation induced by each algorithm under each prior]
 
@@
 
-[Discuss that, under random agent models, firms that play better learning algorithms do better in the long-run since they get enough free agents to move to the part of the relative reputation curve where better algorithms win]
+Under the HardMax decision rule, firms need to incentivize consumers to select them over their competition by having a higher reputation.
+First, we consider the $HardMaxWithRandom$ decision rule discussed in the Section \ref{S:3}.
+Similar to the result in \citet{mansour2017competing} we find that this induces firms to play the ``better" algorithm for sufficiently long time horizon (see appendix).
+The intuition for this is the same as before, where, even though the firm initially suffers reputation loss from exploration, eventually the firm gets sufficiently many agents that don't need to be incentivized to select the firm in order to get a higher reputation than a ``worse" algorithm.
+We have a similar observation for the $SoftMax$ decision rule discussed in Section \ref{S:3}.
 
-[Discuss that, under HMR, market share variance is lower and see less extreme market shares.]
+\begin{figure}
+\label{agent_variances}
+\end{figure}
 
-[Discuss that, under SM, market share variance is very low and market shares are around 50/50].
+We see that better algorithms are incentivized when we consider random agent response rules, but there are differences in the overall market composition.
+As Figure \ref{agent_variances} shows, the variance of market shares across simulations varies greatly.
+Under $HardMax$, we see that, in a single simulation, the market is generally ``winner take all" where one firm generally gets a large share of the market and there is a high variance.
+Under $HardMaxWithRandom$ we see that, in a single simulation, the market is mostly taken by one firm (but to a lower extent than under $HardMax$) and there is a lower, though still high, variance.
+Under $SoftMax$ we see that the market variance is very low and market shares are generally close to 50/50.
 

### 2018-08-09 analysis_code/eeog.Rmd
update analysis code + pdf results

@@
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/"
-f <- "tournament_experiment_ws_small_erase_rep_raw.csv"
+f <- "tournament_experiment_test_raw.csv"
 dat <- read.csv(file=paste(WORKING_PATH, f, sep=""))
@@
 #dat <- rbind(dat, dat_2)
-dat <- filter(dat, Time.Horizon == 5000)
-priors <- as.list(unique(dat['Prior']))$Prior
-warm_starts <- as.list(unique(dat['Warm.Start']))$Warm.Start
+dat <- filter(dat, Time.Horizon == 2000)
+#priors <- as.list(unique(dat['Prior']))$Prior
+priors <- c("Heavy Tail")
+#warm_starts <- as.list(unique(dat['Warm.Start']))$Warm.Start
+warm_starts <- c(20, 50, 100)
 algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
@@
   rownames(results) <- c("TS vs DG", "TS vs DEG", "DG vs DEG")
-  colnames(results) <- lapply(warm_starts, function (start) { return(paste("WS =", start))})
+  colnames(results) <- lapply(warm_starts, function (start) { return(paste("k =", start))})
   for (i in 1:length(warm_starts)) {
@@
   }
-  tab <-xtable(results, caption=paste("Results for", prior, "HardMax", "K=10, erased reputation"))
+  tab <-xtable(results, caption=paste("Results for", prior, "HardMax", "K=10"))
   print(tab, type="html")
-  if (prior != ".5/.7 Random Draw") {   print("<br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br>") }
 

### 2018-08-09 analysis_code/tournament_table.Rmd
update analysis code + pdf results

@@
 
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_parallel_test_raw.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_testing_prelim_raw.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(5000)
+time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior
@@
             p2alg = algs[l]
-            filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Warm.Start == 5 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+            filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Warm.Start == 100 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
             if (nrow(filtered_dat) == 0) {

### 2018-08-09 paper/competing_bandits.tex
update analysis code + pdf results

@@
 
- \author{-}
+\author{Guy Aridor \\
+g.aridor@columbia.edu \\
+Columbia University \\ Department of Economics\\
+	\And Kevin Liu\\
+	kevin.liu@columbia.edu \\
+	Columbia University \\Department of Computer Science\\
+	\AND Alex Slivkins \\
+	slivkins@microsoft.edu\\
+	Microsoft Research NYC\\
+	 \\ 
+	\And Steven Wu\\
+	zsw@umn.edu\\	
+	University of Minnesota - Twin Cities \\Department of Computer Science\\	
+}
   %% Affiliation \\
@@
   \hline
-TS vs DG & \makecell{\textbf{0.86} $\pm$0.02\\ eeog \\ avg: 14\\ med: 0} & \makecell{\textbf{0.65} $\pm$0.02\\ eeog \\ avg: 11\\ med: 0} & \makecell{\textbf{0.59} $\pm$0.02\\ eeog \\ avg: 6\\ med: 0} \\
-  TS vs DEG & \makecell{\textbf{0.84} $\pm$0.02\\ eeog \\ avg: 20\\ med: 0} & \makecell{\textbf{0.62} $\pm$0.02\\ eeog \\ avg: 8\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 5.4\\ med: 0} \\
-  DG vs DEG & \makecell{\textbf{0.41} $\pm$0.02\\ eeog \\ avg: 400\\ med: 115} & \makecell{\textbf{0.47} $\pm$0.02\\ eeog \\ avg: 320\\ med: 0} & \makecell{\textbf{0.5} $\pm$0.02\\ eeog \\ avg: 290\\ med: 0} \\
+TS vs DG & \makecell{\textbf{0.66} $\pm$0.02\\ eeog \\ avg: 190\\ med: 26} & \makecell{\textbf{0.62} $\pm$0.02\\ eeog \\ avg: 240\\ med: 32} &  \\ 
+  TS vs DEG & \makecell{\textbf{0.61} $\pm$0.02\\ eeog \\ avg: 150\\ med: 12} & \makecell{\textbf{0.55} $\pm$0.02\\ eeog \\ avg: 240\\ med: 26} &  \\ 
+  DG vs DEG & \makecell{\textbf{0.44} $\pm$0.02\\ eeog \\ avg: 360\\ med: 106} & \makecell{\textbf{0.44} $\pm$0.02\\ eeog \\ avg: 590\\ med: 312} &  \\ 
    \hline
@@
   \hline
-TS vs DG & \makecell{\textbf{0.22} $\pm$0.02\\ eeog \\ avg: 0.27\\ med: 0} & \makecell{\textbf{0.41} $\pm$0.02\\ eeog \\ avg: 0.039\\ med: 0} & \makecell{\textbf{0.58} $\pm$0.02\\ eeog \\ avg: 12\\ med: 0} \\
-  TS vs DEG & \makecell{\textbf{0.24} $\pm$0.02\\ eeog \\ avg: 3.2\\ med: 0} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 39\\ med: 0} & \makecell{\textbf{0.91} $\pm$0.01\\ eeog \\ avg: 84\\ med: 0} \\
-  DG vs DEG & \makecell{\textbf{0.8} $\pm$0.02\\ eeog \\ avg: 16\\ med: 0} & \makecell{\textbf{0.86} $\pm$0.02\\ eeog \\ avg: 4.7\\ med: 0} & \makecell{\textbf{0.83} $\pm$0.02\\ eeog \\ avg: 2.7\\ med: 0} \\
+TS vs DG & \makecell{\textbf{0.33} $\pm$0.02\\ eeog \\ avg: 49\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.02\\ eeog \\ avg: 120\\ med: 0} &  \\ 
+  TS vs DEG & \makecell{\textbf{0.3} $\pm$0.02\\ eeog \\ avg: 26\\ med: 0} & \makecell{\textbf{0.58} $\pm$0.02\\ eeog \\ avg: 160\\ med: 0} &  \\ 
+  DG vs DEG & \makecell{\textbf{0.62} $\pm$0.02\\ eeog \\ avg: 350\\ med: 4} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 650\\ med: 398.5} &  \\ 
    \hline

### 2018-08-09 analysis_code/agents_table.R

@@ -2,9 +2,9 @@ library(dplyr)
 library(knitr)
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
 #dat_3 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_large_horizon_3_raw.csv", sep=""))
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_many_arms_haystack_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_full_ws_many_sim_raw.csv", sep=""))
 #dat_2 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_many_arms_2_raw.csv", sep=""))
 #dat_10 <- rbind(dat, dat_2)
 #dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ws_small_erase_rep_raw.csv", sep=""))

### 2018-08-09 analysis_code/bin_scatter.R

@@ -1,0 +1,76 @@
+
+binscatter <- function(formula, key_var, data, bins=20, partial=FALSE){
+  
+  require(lfe)
+  require(ggplot2)
+  
+  if(partial==TRUE){
+    y <- unlist(strsplit(formula, "~"))[1]
+    x <- unlist(strsplit(formula, "~"))[2]
+    controls <- gsub(paste("[[:punct:]]*",key_var,"[[:space:]]*[[:punct:]]*",sep=""),
+                     "",x)
+    
+    reg_all <- felm(formula(formula),data=data)
+    reg_y <- felm(formula(paste(y, "~", controls, sep="")), data=data)
+    reg_x <- felm(formula(paste(key_var, "~", controls, sep="")), data=data)
+    resid_all <- resid(reg_all)
+    resid_y <- resid(reg_y)
+    resid_x <- resid(reg_x)
+    df <- data.frame(resid_y, resid_x)
+    cluster_grp <- trimws(unlist(strsplit(formula, "\\|"))[4])
+    if(is.na(cluster_grp)){
+      reg <- felm(resid_y ~ resid_x)  
+    } else{
+      data$resid_y <- resid_y
+      data$resid_x <- resid_x
+      reg <- felm(formula(paste("resid_y ~ resid_x | 0 | 0 |",
+                                cluster_grp, sep="")), data=data)
+    }
+    newdata <- df
+    colnames(df) <- c(paste("residual",names(df)[1]), paste("residual",names(df)[2]))
+    
+  } else if(partial==FALSE){
+    reg <- felm(formula(formula),data=data)
+    y <- trimws(unlist(strsplit(formula, "~"))[1])
+    df <- data[,c(y,key_var)]
+    newdata <- df  # To calculate CI of predicted mean
+    #  colnames(df) <- c("resid_y", "resid_x")
+  }
+  intercept <- coef(reg)[1]
+  slope <- coef(reg)[2]
+  
+  ### Use cluster vcov from the correct model, if not available, use robust
+  if(is.null(reg$clustervcv)){
+    vcov <- reg$robustvcv
+    se_type <- "robust"
+  } else {
+    vcov <- reg$clustervcv
+    se_type <- "cluster"
+  }
+  
+  Terms <- terms(reg)
+  m.mat <- model.matrix(Terms,data=newdata)
[…]

### 2018-08-09 analysis_code/complexity_exp_analysis.R

@@ -1,0 +1,28 @@
+source("bin_scatter.R")
+
+dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/tournament_experiment_fixed_complexity_raw.csv")
+dat <- filter(dat, Time.Horizon == 1000)
+alg1 <- "ThompsonSampling"
+alg2 <- "DynamicEpsilonGreedy"
+ws <- 100
+#filtered_dat <- filter(dat, P1.Alg == alg1 & P2.Alg == alg2 & Warm.Start == ws & Instance.Complexity < 10000)
+#grouped_dat <- filtered_dat %>% group_by(Instance.Complexity, N) %>% summarize(mshare=mean(Market.Share.for.P1))
+#binscatter(formula="mshare ~ Instance.Complexity", key_var = "Instance.Complexity", data=grouped_dat, bins=5, partial=FALSE)
+
+n_vals <- unique(dat$N)
+complexities <- c(50, 150, 200, 250, 300, 350, 400)
+final_dat <- as.data.frame(matrix(nrow=length(complexities)*length(n_vals), ncol=2))
+for (i in 1:length(complexities)) {
+  complexity <- complexities[i]
+  lower_complex <- complexity - 10
+  upper_complex <- complexity + 10
+  for (j in 1:length(n_vals)) {
+    n = n_vals[j]
+    filtered_dat <- filter(dat, P1.Alg == alg1 & P2.Alg == alg2 & Warm.Start == ws & Instance.Complexity <= upper_complex & Instance.Complexity >= lower_complex & N == n)
+    final_dat[(i-1)*length(n_vals)+j, 1] <- complexity
+    final_dat[(i-1)*length(n_vals)+j, 2] <- mean(filtered_dat$Market.Share.for.P1)
+  }
+}
+
+colnames(final_dat) <- c("Complexity", "Market.Share")
+binscatter(formula="Market.Share ~ Complexity", key_var = "Complexity", data=final_dat, bins=7, partial=FALSE)

### 2018-08-09 analysis_code/eeog.R

@@ -18,14 +18,14 @@ concise_alg_rep <- function(alg) {
 }
 
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/"
-f <- "tournament_experiment_ws_small_raw.csv"
+f <- "tournament_experiment_full_ws_many_sim_raw.csv"
 dat <- read.csv(file=paste(WORKING_PATH, f, sep=""))
 #dat_2 <- read.csv(file=paste(WORKING_PATH, "tournament_experiment_tourn_.5_.7_raw.csv", sep=""))
 #dat <- rbind(dat, dat_2)
-dat <- filter(dat, Time.Horizon == 5000)
+dat <- filter(dat, Time.Horizon == 2000)
 priors <- as.list(unique(dat['Prior']))$Prior
-priors <- c("Needle In Haystack - 0.5")
-#warm_starts <- as.list(unique(dat['Warm.Start']))$Warm.Start
+priors <- c("Needle In Haystack", "Heavy Tail")
+warm_starts <- as.list(unique(dat['Warm.Start']))$Warm.Start
 warm_starts <- c(20, 100, 200)
 algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
 
@@ -46,7 +46,7 @@ for (prior in priors) {
         cin <- signif(test$conf.int[2] - mean(share), digits=1)
         eeog_mean <- signif(mean(filtered_dat$EEOG), digits=2)
         eeog_summary <- summary(filtered_dat$EEOG)
-        res_text <- paste("\\makecell{\\textbf{",signif(mean(share), digits=2), "} $\\pm$", cin, "\\\ eeog \\\ avg: ", eeog_mean, "\\\ med: ", eeog_summary["Median"], "}" , sep="")
+        res_text <- paste("\\makecell{\\textbf{",signif(mean(share), digits=2), "} $\\pm$", cin, "\\\\ eeog \\\\ avg: ", eeog_mean, "\\\\ med: ", eeog_summary["Median"], "}" , sep="")
         if (p1alg == "ThompsonSampling" && p2alg == "DynamicGreedy") {
           results[1, i] <- res_text
         } else if (p1alg == "ThompsonSampling" && p2alg == "DynamicEpsilonGreedy") {
@@ -57,7 +57,7 @@ for (prior in priors) {
       }
     }
   }
[…]

### 2018-08-09 analysis_code/kevin_plots.R

@@ -1,84 +1,0 @@
-library(data.table)
-library(ggplot2)
-
-#dt.data <- data.table(read.csv("~/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots.csv", header=T, stringsAsFactors=T))
-
-
-# dt.data[(Algorithm == 'ThompsonSampling' & K==10)]
-
-distribution_names = unique(dt.data[, Distribution])
-
-q <- ggplot(dt.data[(Distribution=="Needle50 - High")], aes(x=t, y=Reward.Mean, color=Algorithm)) +
-  # geom_point(alpha=0.1) +
-  geom_smooth(method="loess", se=TRUE) +
-  scale_x_continuous("Time") +
-  scale_y_continuous('Average instantaneous reward') +
-  theme_bw() +
-  ggtitle("Average reward over time")
-print(q)
-
-
-##########
-
-WORKING_PATH <- '/Users/garidor/Desktop/bandits-rl-project'
-
-dat <- data.table(read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms.csv", sep="")))
-dat2 <- data.table(read.csv(file=paste(WORKING_PATH, "/results/preliminary_raw_results/preliminary_plots_3_arms_2.csv", sep="")))
-dt.data <- rbind(dat, dat2)
-rm(dat)
-rm(dat2)
-
-unique(dt.data$SubAlgorithm)
-
-# renaming so that we can get the desired color vs linetype aesthetics for easy visual identification
-dt.data[, SubAlgorithm := Algorithm]
-dt.data[SubAlgorithm %in% c("NonBayesianEpsilonGreedy, 0.05", 
-                            "NonBayesianEpsilonGreedy, T^(-1/3)", 
-                            "NonBayesianEpsilonGreedy, (t+1)^(-1/3)"), 
-        Algorithm := "NonBayesianEpsilonGreedy"]
-dt.data[SubAlgorithm %in% c("DynamicGreedy", 
-                            "DynamicEpsilonGreedy, 0.05", 
-                            "DynamicEpsilonGreedy, T^(-1/3)", 
-                            "DynamicEpsilonGreedy, (t+1)^(-1/3)"), 
-        Algorithm := "DynamicEpsilonGreedy"]
-dt.data[SubAlgorithm %in% c("UCB1WithConstantOne", 
[…]

### 2018-08-09 analysis_code/preliminary_3.R

@@ -2,7 +2,8 @@ library(dplyr)
 library(ggplot2)
 library(reshape2)
 
-dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_10_arms_10_int.csv")
+#dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
+
 #dat <- read.csv("/Volumes/Mac/bandits/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_10_arms.csv")
 
 
@@ -21,7 +22,7 @@ dist_dat <- filter(dat, Distribution == dist & Realized.Complexity >= minComplex
 deg <- filter(dist_dat, Algorithm == alg1)
 dg <- filter(dist_dat, Algorithm == alg2)
 n_vals <- unique(dist_dat$n)
-t_vals <- seq(10, 1500, 10)
+t_vals <- seq(10, 2500, 10)
 
 counts <- c()
 for (t in t_vals) {
@@ -92,4 +93,12 @@ alg_1_rep <- alg_1_dat$Realized.Reputation
 alg_2_rep <- filter(filtered_dat, Algorithm == alg2)$Realized.Reputation
 rep_diff <- alg_2_rep > alg_1_rep
 print(qplot(complexity_vals, rep_diff))
+
+tmp_ts <- function(n) {
+  return(filter(ts, n == UQ(n)))
+}
+
+tmp_dg <- function(n) {
+  return(filter(dg, n == UQ(n)))
+}
 

### 2018-08-09 preliminary_best_arm.py

@@ -1,118 +1,0 @@
-import numpy as np
-import csv
-import random
-
-
-from copy import deepcopy
-
-
-from lib.BanditProblemInstance import BanditProblemInstance
-
-from lib.bandit.StaticGreedy import StaticGreedy
-from lib.bandit.DynamicEpsilonGreedy import DynamicEpsilonGreedy
-from lib.bandit.NonBayesianEpsilonGreedy import NonBayesianEpsilonGreedy
-from lib.bandit.DynamicGreedy import DynamicGreedy
-from lib.bandit.UCB import UCB1WithConstantOne, UCB1WithConstantT
-from lib.bandit.ThompsonSampling import ThompsonSampling
-from lib.bandit.ExploreThenExploit import ExploreThenExploit
-
-from scipy.stats import bernoulli, beta
-
-
-T = 1501
-N = 100
-K = 10
-
-
-DEFAULT_COMMON_PRIOR = [beta(1, 1) for k in xrange(K)]
-
-
-def sim(alg):
-  for t in xrange(T):
-    banditAlg.executeStep(t)
-  return (banditAlg.realizedRewardHistory, banditAlg.realizedCumulativeRewardHistory, banditAlg.meanRewardHistory, banditAlg.meanCumulativeRewardHistory)
-
-def sim_best_arm(alg):
-  bestArm = banditAlg.banditProblemInstance.getBestArm()
-  res = []
-  look_at = [100, 200, 250, 500, 1000, 1500]
-  for t in xrange(T):
-    if t in look_at:
-      identifiedBestArm = np.argmax([banditAlg.posteriors[i].mean() for i in xrange(K)])
-      res.append((t,identifiedBestArm == bestArm))
-    banditAlg.executeStep(t)
-  identifiedBestArm = np.argmax([banditAlg.posteriors[i].mean() for i in xrange(K)])
-  res.append((T, identifiedBestArm == bestArm))
-  return res
-
-default_mean = 0.5
-needle_in_haystack = [bernoulli(default_mean) for i in xrange(K)]
-
-needle_in_haystack_50_high = deepcopy(needle_in_haystack)
-needle_in_haystack_50_high[int(K/2)] = bernoulli(default_mean + 0.2)
-
-#needle_in_haystack_50_one_medium_one_high = deepcopy(needle_in_haystack)
-#needle_in_haystack_50_one_medium_one_high[int(K/2)] = bernoulli(default_mean + 0.3)
[…]

### 2018-08-09 preliminary_reputation.py

@@ -1,146 +1,0 @@
-import numpy as np
-import csv
-import random
-
-
-from copy import deepcopy
-
-from lib.BanditProblemInstance import BanditProblemInstance
-
-from lib.bandit.StaticGreedy import StaticGreedy
-from lib.bandit.DynamicEpsilonGreedy import DynamicEpsilonGreedy
-from lib.bandit.NonBayesianEpsilonGreedy import NonBayesianEpsilonGreedy
-from lib.bandit.DynamicGreedy import DynamicGreedy
-from lib.bandit.UCB import UCB1WithConstantOne, UCB1WithConstantT
-from lib.bandit.ThompsonSampling import ThompsonSampling
-from lib.bandit.ExploreThenExploit import ExploreThenExploit
-
-from scipy.stats import bernoulli, beta
-
-
-T = 5001
-N = 500
-K = 10
-
-
-DEFAULT_COMMON_PRIOR = [beta(1, 1) for k in xrange(K)]
-
-
-def sim(alg):
-  for t in xrange(T):
-    banditAlg.executeStep(t)
-  return (banditAlg.realizedRewardHistory, banditAlg.realizedCumulativeRewardHistory, banditAlg.meanRewardHistory, banditAlg.meanCumulativeRewardHistory)
-
-def sim_best_arm(alg):
-  bestArm = banditAlg.banditProblemInstance.getBestArm()
-  res = []
-  look_at = [100, 200, 250, 500, 1000, 1500]
-  for t in xrange(T):
-    if t in look_at:
-      identifiedBestArm = np.argmax([banditAlg.posteriors[i].mean() for i in xrange(K)])
-      res.append((t,identifiedBestArm == bestArm))
-    banditAlg.executeStep(t)
-  identifiedBestArm = np.argmax([banditAlg.posteriors[i].mean() for i in xrange(K)])
-  res.append((T, identifiedBestArm == bestArm))
-  return res
-
-default_mean = 0.5
-needle_in_haystack = [bernoulli(default_mean) for i in xrange(K)]
-
-needle_in_haystack_50_high = deepcopy(needle_in_haystack)
-needle_in_haystack_50_high[int(K/2)] = bernoulli(default_mean + 0.2)
-
-#needle_in_haystack_50_one_medium_one_high = deepcopy(needle_in_haystack)
-#needle_in_haystack_50_one_medium_one_high[int(K/2)] = bernoulli(default_mean + 0.3)
[…]

### 2018-08-09 preliminary_reputation_2.py

@@ -1,141 +1,0 @@
-import numpy as np
-import csv
-import random
-import time
-from joblib import Parallel, delayed
-
-
-from copy import deepcopy
-
-from lib.BanditProblemInstance import BanditProblemInstance
-
-from lib.bandit.StaticGreedy import StaticGreedy
-from lib.bandit.DynamicEpsilonGreedy import DynamicEpsilonGreedy
-from lib.bandit.NonBayesianEpsilonGreedy import NonBayesianEpsilonGreedy
-from lib.bandit.DynamicGreedy import DynamicGreedy
-from lib.bandit.UCB import UCB1WithConstantOne, UCB1WithConstantT
-from lib.bandit.ThompsonSampling import ThompsonSampling
-from lib.bandit.ExploreThenExploit import ExploreThenExploit
-from simulate import getRealDistributionsFromPrior
-
-from scipy.stats import bernoulli, beta
-
-
-T = 2001
-N = 150
-K = 10
-numCores = 10
-WARM_START_SIZE = 100
-
-
-DEFAULT_COMMON_PRIOR = [beta(1, 1) for k in xrange(K)]
-
-
-def sim(alg, banditDistr, realizations, warmStartRealizations, seed):
-  seed = int(seed)
-  np.random.seed(seed)
-
-  banditProblemInstance = BanditProblemInstance(K, T, banditDistr, warmStartRealizations)
-  banditAlg = alg(banditProblemInstance, DEFAULT_COMMON_PRIOR)
-  for t in xrange(WARM_START_SIZE):
-    banditAlg.executeStep(t)
-
-  banditProblemInstance = BanditProblemInstance(K, T, banditDistr, realizations)
-  banditAlg.setBanditInstance(banditProblemInstance)
-  for t in xrange(T):
-    banditAlg.executeStep(t)
-  return (banditAlg.realizedRewardHistory, banditAlg.realizedCumulativeRewardHistory, banditAlg.meanRewardHistory, banditAlg.meanCumulativeRewardHistory, banditProblemInstance.getArmMeans(), banditProblemInstance.getComplexityMetric())
-
-default_mean = 0.5
-needle_in_haystack = [bernoulli(default_mean) for i in xrange(K)]
-
-needle_in_haystack_50_high = deepcopy(needle_in_haystack)
-needle_in_haystack_50_high[int(K/2)] = bernoulli(default_mean + 0.2)
-
[…]

### 2018-08-10 analysis_code/agents_table.Rmd
update analysis

@@
 #dat_10 <- rbind(dat, dat_2)
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ws_small_erase_rep_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv", sep=""))
 dat_3 <- dat
@@
 time_horizons <-  as.list(unique(dat_3['Time.Horizon']))$Time.Horizon
-time_horizons <- c(5000)
+time_horizons <- c(2000)
 priors <- as.list(unique(dat_3['Prior']))$Prior

### 2018-08-10 analysis_code/tournament_table.Rmd
update analysis

--- 2018-06-27 analysis_code/tournament_table.Rmd
@@
 
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_parallel_test_raw.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_testing_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(5000)
+time_horizons <- c(1000)
 priors <- as.list(unique(dat['Prior']))$Prior
@@
             p2alg = algs[l]
-            filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Warm.Start == 5 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+            filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Warm.Start == 20 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
             if (nrow(filtered_dat) == 0) {

### 2018-08-10 paper/competing_bandits.tex
update analysis

--- 2018-08-07 paper/competing_bandits.tex
@@
 \caption{Relative Reputation Plots}
-\includegraphics[scale=0.25]{"figures/deg_dg_ht_10_prelim"}
-\includegraphics[scale=0.25]{"figures/ts_dg_ht_10_prelim"} \\
-\includegraphics[scale=0.25]{"figures/deg_dg_nih_10_prelim"}
-\includegraphics[scale=0.25]{"figures/ts_dg_nih_10_prelim"} \\
+\includegraphics[scale=0.2]{"figures/deg_dg_ht_10_prelim"}
+\includegraphics[scale=0.2]{"figures/ts_dg_ht_10_prelim"} \\
+\includegraphics[scale=0.2]{"figures/deg_dg_nih_10_prelim"}
+\includegraphics[scale=0.2]{"figures/ts_dg_nih_10_prelim"} \\
 \caption*{\tiny{The plots contain the proportion of simulations over $250$ runs for a memory size of $100$ where, for a given $t$, the reputation of one algorithm is at least as high as the reputation of another on the same instance.}}
@@
   \hline
-TS vs DG & \makecell{\textbf{0.86} $\pm$0.02\\ eeog \\ avg: 14\\ med: 0} & \makecell{\textbf{0.65} $\pm$0.02\\ eeog \\ avg: 11\\ med: 0} & \makecell{\textbf{0.59} $\pm$0.02\\ eeog \\ avg: 6\\ med: 0} \\
-  TS vs DEG & \makecell{\textbf{0.84} $\pm$0.02\\ eeog \\ avg: 20\\ med: 0} & \makecell{\textbf{0.62} $\pm$0.02\\ eeog \\ avg: 8\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 5.4\\ med: 0} \\
-  DG vs DEG & \makecell{\textbf{0.41} $\pm$0.02\\ eeog \\ avg: 400\\ med: 115} & \makecell{\textbf{0.47} $\pm$0.02\\ eeog \\ avg: 320\\ med: 0} & \makecell{\textbf{0.5} $\pm$0.02\\ eeog \\ avg: 290\\ med: 0} \\
+TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 210\\ med: 28} & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 260\\ med: 68} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 320\\ med: 0} \\
+  TS vs DEG & \makecell{\textbf{0.56} $\pm$0.03\\ eeog \\ avg: 150\\ med: 11} & \makecell{\textbf{0.54} $\pm$0.03\\ eeog \\ avg: 240\\ med: 24.5} & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 420\\ med: 36.5} \\
+  DG vs DEG & \makecell{\textbf{0.43} $\pm$0.03\\ eeog \\ avg: 330\\ med: 97.5} & \makecell{\textbf{0.42} $\pm$0.03\\ eeog \\ avg: 550\\ med: 229.5} & \makecell{\textbf{0.41} $\pm$0.02\\ eeog \\ avg: 650\\ med: 420} \\
    \hline
@@
   \hline
-TS vs DG & \makecell{\textbf{0.22} $\pm$0.02\\ eeog \\ avg: 0.27\\ med: 0} & \makecell{\textbf{0.41} $\pm$0.02\\ eeog \\ avg: 0.039\\ med: 0} & \makecell{\textbf{0.58} $\pm$0.02\\ eeog \\ avg: 12\\ med: 0} \\
-  TS vs DEG & \makecell{\textbf{0.24} $\pm$0.02\\ eeog \\ avg: 3.2\\ med: 0} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 39\\ med: 0} & \makecell{\textbf{0.91} $\pm$0.01\\ eeog \\ avg: 84\\ med: 0} \\
-  DG vs DEG & \makecell{\textbf{0.8} $\pm$0.02\\ eeog \\ avg: 16\\ med: 0} & \makecell{\textbf{0.86} $\pm$0.02\\ eeog \\ avg: 4.7\\ med: 0} & \makecell{\textbf{0.83} $\pm$0.02\\ eeog \\ avg: 2.7\\ med: 0} \\
+TS vs DG & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 65\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 130\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 500\\ med: 0} \\
+  TS vs DEG & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 26\\ med: 0} & \makecell{\textbf{0.59} $\pm$0.03\\ eeog \\ avg: 160\\ med: 0} & \makecell{\textbf{0.86} $\pm$0.02\\ eeog \\ avg: 420\\ med: 0} \\
+  DG vs DEG & \makecell{\textbf{0.63} $\pm$0.03\\ eeog \\ avg: 390\\ med: 10} & \makecell{\textbf{0.62} $\pm$0.02\\ eeog \\ avg: 670\\ med: 500.5} & \makecell{\textbf{0.61} $\pm$0.02\\ eeog \\ avg: 770\\ med: 670.5} \\
    \hline

### 2018-08-10 analysis_code/agents_table.R

@@ -2,14 +2,8 @@ library(dplyr)
 library(knitr)
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
-#dat_3 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_large_horizon_3_raw.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_full_ws_many_sim_raw.csv", sep=""))
-#dat_2 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_many_arms_2_raw.csv", sep=""))
-#dat_10 <- rbind(dat, dat_2)
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ws_small_erase_rep_raw.csv", sep=""))
-#dat_3 <- dat
-
+dat_3 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_"))
 
 p1_algs <- as.list(unique(dat_3['P1.Alg']))$P1.Alg
 p2_algs <- as.list(unique(dat_3['P2.Alg']))$P2.Alg

### 2018-08-10 analysis_code/compare_across.R

@@ -1,0 +1,32 @@
+library(dplyr)
+
+dat_1 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
+dat_2 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv")
+
+dist <- "Heavy Tail"
+ws_time <- 20
+iso_dat <- filter(dat_1, Distribution == dist)
+compet_dat <- filter(dat_2, Prior == dist & Warm.Start == 20)
+
+alg1 <- "ThompsonSampling"
+alg2 <- "DynamicGreedy"
+alg_1 <- filter(iso_dat, Algorithm == alg1)
+alg_2 <- filter(iso_dat, Algorithm == alg2)
+n_vals <- unique(iso_dat$n)
+winning_n <- c()
+losing_n <- c()
+for (n in n_vals) {
+  cur_alg1 <- filter(alg_1,  t == ws_time & n == UQ(n))
+  cur_alg2 <- filter(alg_2, t == ws_time & n == UQ(n))
+  if (cur_alg1$Realized.Reputation > cur_alg2$Realized.Reputation) {
+    winning_n <- c(winning_n, n)
+  } else {
+    losing_n <- c(losing_n, n)
+  }
+}
+
+compet_dat_filtered <- filter(compet_dat, Time.Horizon == 2000 & Agent.Alg == "HardMax" & P1.Alg == alg1 & P2.Alg == alg2)
+winning_dat <- filter(compet_dat_filtered, row_number() %in% winning_n)
+losing_dat <- filter(compet_dat_filtered, row_number() %in% losing_n)
+cat("Winning dat", nrow(winning_dat), "Market share", mean(winning_dat$Market.Share.for.P1))
+cat("Losing dat", nrow(losing_dat), "Market share", mean(losing_dat$Market.Share.for.P1))

### 2018-08-10 analysis_code/eeog.R

@@ -1,7 +1,4 @@
-#```{r echo=FALSE, message=FALSE, xtable, results="asis"}
-
 library(dplyr)
-library(knitr)
 library(xtable)
 
 options(xtable.sanitize.text.function=identity)
@@ -18,10 +15,8 @@ concise_alg_rep <- function(alg) {
 }
 
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/"
-f <- "tournament_experiment_full_ws_many_sim_raw.csv"
+f <- "tournament_experiment_full_sim_with_realizations_raw.csv"
 dat <- read.csv(file=paste(WORKING_PATH, f, sep=""))
-#dat_2 <- read.csv(file=paste(WORKING_PATH, "tournament_experiment_tourn_.5_.7_raw.csv", sep=""))
-#dat <- rbind(dat, dat_2)
 dat <- filter(dat, Time.Horizon == 2000)
 priors <- as.list(unique(dat['Prior']))$Prior
 priors <- c("Needle In Haystack", "Heavy Tail")

### 2018-08-10 analysis_code/relative_reputation_plots.R

@@ -2,10 +2,7 @@ library(dplyr)
 library(ggplot2)
 library(reshape2)
 
-#dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
-
-#dat <- read.csv("/Volumes/Mac/bandits/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_10_arms.csv")
-
+dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
 
 concise_alg_rep <- function(alg) {
   if (alg == "ThompsonSampling") {

### 2018-08-10 analysis_code/tournament_table.R

@@ -6,15 +6,15 @@ WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
 
 #dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_parallel_test_raw.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_with_realizations_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
 p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
 agent_algs - c("HardMax")
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(5000)
+time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior
-priors <- c("Needle In Haystack - 0.5", "Heavy Tail")
+priors <- c("Needle In Haystack", "Heavy Tail")
 
 ZERO_ONE_CUTOFF <- 0.1
 algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")

### 2018-08-10 analysis_code/compare_across.R
update figures in the paper to include confidence intervals and use the same data

@@ -3,7 +3,7 @@ library(dplyr)
 dat_1 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
 dat_2 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv")
 
-dist <- "Heavy Tail"
+dist <- "Needle In Haystack"
 ws_time <- 20
 iso_dat <- filter(dat_1, Distribution == dist)
 compet_dat <- filter(dat_2, Prior == dist & Warm.Start == 20)

### 2018-08-10 analysis_code/relative_reputation_plots.R

@@ -2,7 +2,7 @@ library(dplyr)
 library(ggplot2)
 library(reshape2)
 
-dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
+#dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
 
 concise_alg_rep <- function(alg) {
   if (alg == "ThompsonSampling") {
@@ -40,34 +40,39 @@ print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
   q <- ggplot(df, aes(x=t, y=relative_rep)) + geom_line() + geom_point() +
     geom_errorbar(aes(ymin=relative_rep-1.96*se, ymax=relative_rep+1.96*se), width=.2) +
     ggtitle(plot_title) + xlab("t") + 
-    ylab(paste(concise_alg_rep(alg1), " >= ", concise_alg_rep(alg2), "(reputation)"))
+    ylab(paste(concise_alg_rep(alg1), " >= ", concise_alg_rep(alg2), "(reputation)")) +
+    theme(plot.title = element_text(hjust = 0.5))
   
   print(q)
 }
 
-print_mean_graphs <- function (dist, alg1, alg2, minComplexity) {
-  K <- unique(dat$K)[1]
+print_mean_graphs <- function (dist, alg1, alg2, alg3, minComplexity) {
   dist_dat <- filter(dat, Distribution == dist & Realized.Complexity >= minComplexity)
-  deg <- filter(dist_dat, Algorithm == alg1)
-  dg <- filter(dist_dat, Algorithm == alg2)
+  alg_1 <- filter(dist_dat, Algorithm == alg1)
+  alg_2 <- filter(dist_dat, Algorithm == alg2)
+  alg_3 <- filter(dist_dat, Algorithm == alg3)
   n_vals <- unique(dat$n)
   t_vals <- seq(10, 2000, 10)
   
-  means_deg <- c()
-  means_dg <- c()
+  df <- as.data.frame(matrix(nrow=0, ncol=4))
+  colnames(df) <- c("t", "mean_rep", "ci", "alg")
   for (t in t_vals) {
-    cur_dg <- filter(dg, t == UQ(t))
[…]

### 2018-08-10 paper/competing_bandits.tex
update figures in the paper to include confidence intervals and use the same data
further updates to figures

@@
 \begin{figure}
-\caption{Reputation Trajectories in Isolation}
-\includegraphics[scale=0.2]{"figures/Reputation Trajectory for Heavy Tail 10 arms"}
-\includegraphics[scale=0.2]{"figures/Reputation Trajectory for Needle In Haystack High 10 arms"}
+\caption{Mean Reputation Trajectories in Isolation}
+\includegraphics[scale=0.2]{"figures/heavy_tail_iso_mean"}
+\includegraphics[scale=0.2]{"figures/nih_iso_mean"}
 \label{prelim_means}
@@
 \end{table}
-
-\begin{figure}
-\label{rel_rep_ht_fine}
-\caption{Relative Reputation - Heavy Tail}
-\includegraphics[scale=0.3]{figures/ts_dg_heavy_fine}
-\caption*{TODO: Reduce the range of the x-axis of this so that you can understand when TS starts to do better than DG}
-\end{figure}
 

### 2018-08-10 analysis_code/relative_reputation_plots.R

@@ -18,7 +18,7 @@ print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
   alg_1 <- filter(dist_dat, Algorithm == alg1)
   alg_2 <- filter(dist_dat, Algorithm == alg2)
   n_vals <- unique(dist_dat$n)
-  t_vals <- seq(100, 2000, 100)
+  t_vals <- seq(10, 2000, 10)
   df <- as.data.frame(matrix(nrow=length(t_vals), ncol=3))
   colnames(df) <- c("t", "relative_rep")
   df$t <- t_vals
@@ -78,10 +78,8 @@ print_mean_graphs <- function (dist, alg1, alg2, alg3, minComplexity) {
 }
 
 dists <- unique(dat$Distribution)
-#dists <- c("Heavy Tail", "Uniform")
 
 for (dist in dists) {
-  print(dist)
   print_relative_graphs(dist, "ThompsonSampling", "DynamicEpsilonGreedy", 0)
   print_relative_graphs(dist, "ThompsonSampling", "DynamicGreedy", 0)
   print_relative_graphs(dist, "DynamicEpsilonGreedy", "DynamicGreedy", 0)

### 2018-08-12 analysis_code/agents_table.Rmd
more updates

--- 2018-06-27 analysis_code/agents_table.Rmd
@@
 
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_tournament_raw_results.csv", sep=""))
-#dat_3 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_large_horizon_3_raw.csv", sep=""))
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_many_arms_haystack_raw.csv", sep=""))
-#dat_2 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_many_arms_2_raw.csv", sep=""))
-#dat_10 <- rbind(dat, dat_2)
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ws_small_erase_rep_raw.csv", sep=""))
-dat_3 <- dat
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_raw.csv", sep=""))
 
-p1_algs <- as.list(unique(dat_3['P1.Alg']))$P1.Alg
-p2_algs <- as.list(unique(dat_3['P2.Alg']))$P2.Alg
-agent_algs <- as.list(unique(dat_3['Agent.Alg']))$Agent.Alg
-time_horizons <-  as.list(unique(dat_3['Time.Horizon']))$Time.Horizon
-time_horizons <- c(5000)
-priors <- as.list(unique(dat_3['Prior']))$Prior
-warm_starts <-  sort(as.list(unique(dat_3['Warm.Start']))$Warm.Start, decreasing=FALSE)
+p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
+p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
+agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
+time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
+time_horizons <- c(2000)
+priors <- as.list(unique(dat['Prior']))$Prior
+warm_starts <-  sort(as.list(unique(dat['Warm.Start']))$Warm.Start, decreasing=FALSE)
 
@@
           } else {
-            filtered_dat <- filter(dat_10, P1.Alg == p1alg & Warm.Start == start & P2.Alg == p2alg & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+            filtered_dat <- filter(dat, P1.Alg == p1alg & Warm.Start == start & P2.Alg == p2alg & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
           }

### 2018-08-12 analysis_code/tournament_table.Rmd
more updates

--- 2018-08-09 analysis_code/tournament_table.Rmd
@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_testing_prelim_raw.csv", sep=""))
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_free_obs_10_arms_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg

### 2018-08-12 paper/competing_bandits.tex
more updates

@@
 We study the implications of our model via simulation.
-For computational tractability we focus our attention on Bernoulli bandits with beta priors.
+For computational tractability we focus our attention on Bernoulli bandit diBernoulli bandits with beta priors.
 We consider the following bandit ``priors" from which the bandit instances faced in simulation are drawn:
@@
 Run in isolation we expect that the performance of the algorithms (according to cumulative regret) should be \\
- Adaptive Exploration $>=$ Non-adaptive exploration $>=$ Greedy.
+ Adaptive Exploration $\geq$ Non-adaptive exploration $\geq$ Greedy.
 We verify that, for the chosen algorithms and chosen priors, this is indeed the case (with some caveats, discussed in the next section).
@@
 \For{Each prior $p$}
+	\For{$N$ simulations}
 		\State Generate true distribution from $p$
@@
 			\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
-			\For{$N$ simulations}
 				\State Give principal 2 $X$ free observations (the agents also get these observations)
@@
 In this section we present simulation results of the model discussed in Section \ref{S:2} where $X = 0$ so that both firms enter at the same time.
-We want to understand what algorithms the firms are incentivized to commit to in the competition game but the payoffs associated with each algorithm are too difficult to compete analytically and thus, we begin by simply imposing that the algorithm played by the firm is taken as given and study the consequences of playing them in the competition game.
+We want to understand what algorithms the firms are incentivized to commit to in the competition game but the payoffs associated with each algorithm are too difficult to compute analytically and thus, we begin by simply imposing that the algorithm played by the firm is taken as given and study the consequences of playing them in the competition game.
 Additionally, for this experiment we present results for the $HardMax$ agent decision rule and postpone discussion of alternative decision rules.
-Besides keeping track of market share, we also will keep track of the effective end of game in order to better understand what happens in the game.
+Besides keeping track of market share, we also will keep track of the ``effective end of game."
 
 \begin{definition}
-\textit{Effective End of Game (EEOG)} - the last round, $t$ in a simulation where there is a difference between the decision of the agent alive at $t-1$ and the agent alive at $t$ (todo: state this formally using notation of model)
+\textit{Effective End of Game (EEOG)} - the last round, $t$, in a simulation where there is a difference between the decision of the agent alive at $t-1$ and the agent alive at $t$ (todo: state this formally using notation of model)
 \end{definition}
@@
 \begin{table}[ht]
-\label{ht_incum}
 \centering
-\caption{Incumbent Experiment for Heavy Tail}
+\caption{Incumbent Experiment Heavy Tail X=200} 
 \begin{tabular}{rlll}
@@
   \hline
-TS & \makecell{\textbf{ 0.0067 } $\pm$ 0.0092 \\Variance: 0.007 \\ ES: 100 \%} & \makecell{\textbf{ 0.023 } $\pm$ 0.017 \\Variance: 0.02 \\ ES: 99 \%} & \makecell{\textbf{ 0.064 } $\pm$ 0.027 \\Variance: 0.05 \\ ES: 97 \%} \\
+TS & \makecell{\textbf{ 0.0029 } $\pm$ 0.0029 \\Variance: 0.002 \\ ES: 100 \%} & \makecell{\textbf{ 0.11 } $\pm$ 0.019 \\Variance: 0.09 \\ ES: 96 \%} & \makecell{\textbf{ 0.17 } $\pm$ 0.022 \\Variance: 0.1 \\ ES: 95 \%} \\
+  DEG & \makecell{\textbf{ 0.049 } $\pm$ 0.011 \\Variance: 0.03 \\ ES: 94 \%} & \makecell{\textbf{ 0.24 } $\pm$ 0.022 \\Variance: 0.1 \\ ES: 74 \%} & \makecell{\textbf{ 0.24 } $\pm$ 0.023 \\Variance: 0.1 \\ ES: 79 \%} \\
+   DG & \makecell{\textbf{ 0.12 } $\pm$ 0.017 \\Variance: 0.08 \\ ES: 87 \%} & \makecell{\textbf{ 0.35 } $\pm$ 0.026 \\Variance: 0.2 \\ ES: 77 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.023 \\Variance: 0.1 \\ ES: 65 \%} \\
    \hline
-  DEG & \makecell{\textbf{ 0.024 } $\pm$ 0.015 \\Variance: 0.02 \\ ES: 98 \%} & \makecell{\textbf{ 0.13 } $\pm$ 0.034 \\Variance: 0.09 \\ ES: 86 \%} & \makecell{\textbf{ 0.14 } $\pm$ 0.036 \\Variance: 0.1 \\ ES: 89 \%} \\
-\hline   
-   DG & \makecell{\textbf{ 0.063 } $\pm$ 0.024 \\Variance: 0.04 \\ ES: 93 \%} & \makecell{\textbf{ 0.19 } $\pm$ 0.041 \\Variance: 0.1 \\ ES: 91 \%} & \makecell{\textbf{ 0.15 } $\pm$ 0.032 \\Variance: 0.08 \\ ES: 77 \%} \\
 \end{tabular}
@@
 the free rounds to firm 2 do not count towards the share)}}
+\label{ht_incum}
 \end{table}
@@
 
-\section{Conclusion}
-\label{S:10}
-
-...
-
-\pagebreak
 \bibliography{refs}

### 2018-08-12 analysis_code/agents_table.R

@@ -2,18 +2,17 @@ library(dplyr)
 library(knitr)
 WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
-dat_3 <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_"))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_raw.csv", sep=""))
 
-p1_algs <- as.list(unique(dat_3['P1.Alg']))$P1.Alg
-p2_algs <- as.list(unique(dat_3['P2.Alg']))$P2.Alg
-agent_algs <- as.list(unique(dat_3['Agent.Alg']))$Agent.Alg
+p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
+p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
+agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
 agent_algs <- c("HardMax")
-time_horizons <-  as.list(unique(dat_3['Time.Horizon']))$Time.Horizon
-time_horizons <- c(5000)
-priors <- as.list(unique(dat_3['Prior']))$Prior
+time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
+time_horizons <- c(2000)
+priors <- as.list(unique(dat['Prior']))$Prior
 priors <- c("Heavy Tail", "Needle In Haystack - 0.5")
-warm_starts <- c(5)
+warm_starts <- c(20)
 
 ZERO_ONE_CUTOFF <- 0.1
 
@@ -59,7 +58,6 @@ for (prior in priors) {
     results <- matrix(nrow=length(agent_algs), ncol=(length(alg_pairs) / 2))
[…]

### 2018-08-12 analysis_code/compare_across.R

@@ -1,9 +1,9 @@
 library(dplyr)
 
-dat_1 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
-dat_2 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv")
+#dat_1 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
+#dat_2 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/tournament_experiment_full_sim_raw.csv")
 
-dist <- "Needle In Haystack"
+dist <- "Heavy Tail"
 ws_time <- 20
 iso_dat <- filter(dat_1, Distribution == dist)
 compet_dat <- filter(dat_2, Prior == dist & Warm.Start == 20)
@@ -18,7 +18,7 @@ losing_n <- c()
 for (n in n_vals) {
   cur_alg1 <- filter(alg_1,  t == ws_time & n == UQ(n))
   cur_alg2 <- filter(alg_2, t == ws_time & n == UQ(n))
-  if (cur_alg1$Realized.Reputation > cur_alg2$Realized.Reputation) {
+  if (cur_alg1$Realized.Reputation >= cur_alg2$Realized.Reputation) {
     winning_n <- c(winning_n, n)
   } else {
     losing_n <- c(losing_n, n)
@@ -26,7 +26,22 @@ for (n in n_vals) {
 }
 
 compet_dat_filtered <- filter(compet_dat, Time.Horizon == 2000 & Agent.Alg == "HardMax" & P1.Alg == alg1 & P2.Alg == alg2)
-winning_dat <- filter(compet_dat_filtered, row_number() %in% winning_n)
-losing_dat <- filter(compet_dat_filtered, row_number() %in% losing_n)
-cat("Winning dat", nrow(winning_dat), "Market share", mean(winning_dat$Market.Share.for.P1))
+winning_dat <- filter(compet_dat_filtered, N %in% winning_n)
+losing_dat <- filter(compet_dat_filtered, N %in% losing_n)
+cat("Winning dat", nrow(winning_dat), "Market share", mean(winning_dat$Market.Share.for.P1), "\n")
[…]

### 2018-08-12 analysis_code/relative_rep.jl

@@ -1,0 +1,44 @@
+using KernelDensity
+using CSV
+using Query
+using DataFrames
+using PyPlot
+
+dat = CSV.read("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
+distribution = "Uniform"
+
+function get_prob_rep_greater(data, alg1, alg2, t)
+  alg_1_dat = @from i in data begin
+      @where i.Algorithm == alg1 && i.t == t
+      @select i
+      @collect DataFrame
+    end
+
+  alg_2_dat = @from i in data begin
+      @where i.Algorithm == alg2 && i.t == t
+      @select i
+      @collect DataFrame
+    end
+
+  rep_diff = alg_1_dat[Symbol("Realized Reputation")] - alg_2_dat[Symbol("Realized Reputation")]
+  est = kde(rep_diff)
+  return quadgk(x -> pdf(est, x), 0, maximum(est.x))[1]
+end
+
+
+filtered_dat =  @from i in dat begin
+  @where i.Distribution == distribution
+  @select i
+  @collect DataFrame
+end
+
+t_vals = collect(range(100,100,20))
+probs = []
+alg1 = "ThompsonSampling"
+alg2 = "DynamicGreedy"
+for t in t_vals
+  push!(probs, get_prob_rep_greater(filtered_dat, alg1, alg2, t))
+end
+
+PyPlot.plot(t_vals, probs)
+

### 2018-08-12 analysis_code/relative_reputation_plots.R

@@ -29,7 +29,7 @@ print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
   for (t in t_vals) {
     cur_alg1 <- filter(alg_1,  t == UQ(t))
     cur_alg2 <- filter(alg_2, t == UQ(t))
-    count <- sum(cur_alg1$Realized.Reputation - cur_alg2$Realized.Reputation >= 0)
+    count <- sum(cur_alg1$Realized.Reputation - cur_alg2$Realized.Reputation > 0)
     p <- count / length(n_vals)
     counts <- c(counts, p)
     ses <- c(ses, sqrt(p * (1 - p) / length(n_vals)))
@@ -84,6 +84,7 @@ for (dist in dists) {
   print_relative_graphs(dist, "ThompsonSampling", "DynamicGreedy", 0)
   print_relative_graphs(dist, "DynamicEpsilonGreedy", "DynamicGreedy", 0)
 }
+
 # print mean plots
 for (dist in dists) {
   print_mean_graphs(dist, "ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy", 0)

### 2018-08-12 analysis_code/tournament_table.R

@@ -5,8 +5,7 @@ WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
 
 
-#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_parallel_test_raw.csv", sep=""))
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_with_realizations_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
 p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
@@ -15,6 +14,7 @@ time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
 time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior
 priors <- c("Needle In Haystack", "Heavy Tail")
+warm_start <- 20
 
 ZERO_ONE_CUTOFF <- 0.1
 algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
@@ -35,7 +35,7 @@ for (j in 1:length(time_horizons)) {
         for (l in 1:length(algs)) {
           p1alg = algs[k]
           p2alg = algs[l]
-          filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Warm.Start == 5 & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
+          filtered_dat <- filter(dat, P1.Alg == p1alg & P2.Alg == p2alg & Warm.Start == warm_start & Prior == prior & Agent.Alg == agent_alg & Time.Horizon == time)
           if (nrow(filtered_dat) == 0) {
             if (p1alg == p2alg) { print_breaks <- TRUE }
             next
@@ -51,7 +51,7 @@ for (j in 1:length(time_horizons)) {
           results[k, l] <- paste("\\makecell{\\textbf{", signif(mean(share), digits=2), "} $\\pm$", cin, "\\\\Variance:", signif(var(share), digits=1), "\\\\", "ES:", 100*signif(shares, digits=2), "\\%}")
         }
       }
[…]

### 2018-08-12 main.py

@@ -30,8 +30,10 @@ T = 2001
 NUM_SIMULATIONS = 1000
 
 FREE_OBS = True
+ERASE_REP = True
+ERASE_INFO = False
 FREE_OBS_NUM = 200
-exp_name = 'full_sim'
+exp_name = 'full_sim_no_rep'
 print('Exp name', exp_name)
 REALIZATIONS_NAME = 'preliminary' #if you want to pull in past realizations, fill this in with the realizations base name
 numCores = 12
@@ -78,7 +80,7 @@ raw_name = exp_base_name + '_raw.csv'
 realizations_name = exp_base_name + '_realizations.csv'
 dist_name = exp_base_name + '_dist.csv'
 
-INDIVIDUAL_FIELD_NAMES =['Prior', 'N', 'P1 Alg', 'EEOG', 'Instance Complexity', 'P2 Alg', 'Time Horizon', 'Agent Alg', 'Market Share for P1', 'P1 Regret', 'P2 Regret', 'P1 Reputation', 'P2 Reputation', 'Abs Delta Regret']
+INDIVIDUAL_FIELD_NAMES =['Prior', 'Erase Info', 'Erase Reputation', 'N', 'P1 Alg', 'EEOG', 'Instance Complexity', 'P2 Alg', 'Time Horizon', 'Agent Alg', 'Market Share for P1', 'P1 Regret', 'P2 Regret', 'P1 Reputation', 'P2 Reputation', 'Abs Delta Regret']
 
 # fetch distributions from previous run
 def fetch_distributions(filename, priorname):
@@ -154,7 +156,7 @@ def run_experiment(startSizes):
         for (principalAlg1, principalAlg2) in ALG_PAIRS:
           for startSize in startSizes:
             print('Running ' + agentAlg.__name__ + ' and principal 1 playing ' + principalAlg1.__name__ + ' and principal 2 playing ' + principalAlg2.__name__ + ' with warm start size ' + str(startSize) + ' with prior ' + banditDistrName)
-            simResults = Parallel(n_jobs=numCores)(delayed(simulate)(principalAlg1, principalAlg2, agentAlg, maxStart, K=K, T=T, warmStartNumObservations=startSize, realizations=realizations[i], freeObsForP2=FREE_OBS, freeObsNum=FREE_OBS_NUM, realDistributions=realDistributions[i], seed=i+1) for i in xrange(NUM_SIMULATIONS))
[…]

### 2018-08-12 simulate.py

@@ -73,6 +73,10 @@ Parameters
     Used for the incumbent experiment - does principal 2 (the incumbent) get free agents before the competition game?
   freeObsNum: Int
     For how many rounds is principal 2 the incumbent?
+  eraseReputation: Boolean
+    Whether to erase the reputation gain of the incumbent
+  eraseInformation: Boolean
+    Whether to erase the information gain of the incumbent
   principal1Priors / principal2Priors: Vector of Distributions
     Initial beliefs of the principals
   recordStatsAt: Vector of times
@@ -91,9 +95,10 @@ def simulate(principalAlg1, principalAlg2, agentAlg, maxWarmStart, K, T,
   warmStartNumObservations=DEFAULT_WARM_START_NUM_OBSERVATIONS,
   principal1Priors=None,
   principal2Priors=None,
+  eraseReputation=False,
+  eraseInformation=False,
   recordStatsAt=RECORD_STATS_AT,
-  seed=1.0,
-  eraseReputation = False
+  seed=1.0
 ):
   seed = int(seed)
   np.random.seed(seed)
@@ -127,7 +132,12 @@ def simulate(principalAlg1, principalAlg2, agentAlg, maxWarmStart, K, T,
   # now, reset the realizations to the warm start realizations so they're the same across firms and give each firm the warm start agents
   for principal in principals.values():
     principal.resetStats()
+    if eraseInformation:
+      principal.resetPriors()
     principal.setRealizations(warmStartRealizations)
+
+  if eraseReputation:
+    agents.resetInformationSet()
 
   for i in xrange(warmStartNumObservations):
     for (principalName, principal) in principals.iteritems():

### 2018-08-15 analysis_code/agents_table.Rmd
add rewrite of paper

@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ht_lhr_3_arms_raw.csv", sep=""))
 
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior

### 2018-08-15 analysis_code/tournament_table.Rmd
add rewrite of paper

--- 2018-08-10 analysis_code/tournament_table.Rmd
@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_testing_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ht_lhr_3_arms_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(1000)
+time_horizons <- c(7000)
 priors <- as.list(unique(dat['Prior']))$Prior

### 2018-08-15 paper/competing_bandits.tex
add rewrite of paper

@@
   \hline
-TS & \makecell{ \textbf{ 0.18 } $\pm$ 0.042 \\Var:  0.1 \\ ES: 97 \% } & \makecell{ \textbf{ 0.17 } $\pm$ 0.042 \\Var:  0.1 \\ ES: 98 \% } & \makecell{ \textbf{ 0.23 } $\pm$ 0.047 \\Var:  0.2 \\ ES: 98 \% } \\
-  DEG & \makecell{ \textbf{ 0.23 } $\pm$ 0.045 \\Var:  0.2 \\ ES: 92 \% } & \makecell{ \textbf{ 0.32 } $\pm$ 0.049 \\Var:  0.2 \\ ES: 86 \% } & \makecell{ \textbf{ 0.25 } $\pm$ 0.046 \\Var:  0.2 \\ ES: 90 \% } \\
-   DG & \makecell{ \textbf{ 0.28 } $\pm$ 0.047 \\Var:  0.2 \\ ES: 87 \% } & \makecell{ \textbf{ 0.33 } $\pm$ 0.051 \\Var:  0.2 \\ ES: 91 \% } & \makecell{ \textbf{ 0.3 } $\pm$ 0.047 \\Var:  0.2 \\ ES: 83 \% } \\
+TS & \makecell{\textbf{ 0.016 } $\pm$ 0.0075 \\Variance: 0.01 \\ ES: 100 \%} & \makecell{\textbf{ 0.13 } $\pm$ 0.02 \\Variance: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.2 } $\pm$ 0.024 \\Variance: 0.1 \\ ES: 96 \%} \\
+  DEG & \makecell{\textbf{ 0.068 } $\pm$ 0.013 \\Variance: 0.05 \\ ES: 93 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Variance: 0.2 \\ ES: 75 \%} & \makecell{\textbf{ 0.26 } $\pm$ 0.024 \\Variance: 0.2 \\ ES: 80 \%} \\
+   DG & \makecell{\textbf{ 0.15 } $\pm$ 0.019 \\Variance: 0.1 \\ ES: 87 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.028 \\Variance: 0.2 \\ ES: 80 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.024 \\Variance: 0.2 \\ ES: 67 \%} \\
    \hline
@@
 \centering
-\caption{Information Erased Experiment} 
+\caption{Information Erased Experiment Heavy Tail X=200} 
 \begin{tabular}{rlll}
@@
   \hline
-TS & \makecell{ \textbf{ 0.059 } $\pm$ 0.025 \\Var:  0.05 \\ ES: 97 \% } & \makecell{ \textbf{ 0.11 } $\pm$ 0.033 \\Var:  0.09 \\ ES: 94 \% } & \makecell{ \textbf{ 0.15 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 92 \% } \\
-  DEG & \makecell{ \textbf{ 0.11 } $\pm$ 0.031 \\Var:  0.08 \\ ES: 89 \% } & \makecell{ \textbf{ 0.19 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 82 \% } & \makecell{ \textbf{ 0.18 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 82 \% } \\
-   DG & \makecell{ \textbf{ 0.17 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 87 \% } & \makecell{ \textbf{ 0.22 } $\pm$ 0.042 \\Var:  0.1 \\ ES: 82 \% } & \makecell{ \textbf{ 0.26 } $\pm$ 0.043 \\Var:  0.1 \\ ES: 77 \% } \\
+TS & \makecell{\textbf{ 0.024 } $\pm$ 0.0094 \\Variance: 0.02 \\ ES: 100 \%} & \makecell{\textbf{ 0.16 } $\pm$ 0.022 \\Variance: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.22 } $\pm$ 0.025 \\Variance: 0.2 \\ ES: 95 \%} \\
+  DEG & \makecell{\textbf{ 0.24 } $\pm$ 0.025 \\Variance: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Variance: 0.1 \\ ES: 72 \%} & \makecell{\textbf{ 0.27 } $\pm$ 0.024 \\Variance: 0.1 \\ ES: 76 \%} \\
+   DG & \makecell{\textbf{ 0.33 } $\pm$ 0.028 \\Variance: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.026 \\Variance: 0.2 \\ ES: 74 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.023 \\Variance: 0.1 \\ ES: 58 \%} \\
    \hline

### 2018-08-15 paper_v2/competing_bandits.tex
add rewrite of paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}

% ready for submission
\usepackage{aaai}

% to compile a preprint version, e.g., for submission to arXiv, add
% add the [preprint] option:
% \usepackage[preprint]{nips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2018}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amssymb}
\usepackage{natbib}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{float}
\usepackage{sgame, tikz} % Game theory packages
\usepackage{caption} 
\usepackage{algorithm,algpseudocode}
\usepackage{makecell}
 \usepackage{multirow}
 \usepackage{graphicx}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\captionsetup{font=footnotesize}
\bibliographystyle{aaai}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\begin{document}

\title{Competing Bandits}

\author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
Zhiwei Steven Wu\textsuperscript{4} \\,
{\textsuperscript{1}Columbia University, Department of Economics}\\
{\textsuperscript{2}Columbia University, Department of Computer Science}\\
{\textsuperscript{3}Microsoft Research, New York, NY}\\
{\textsuperscript{4}University of Minnesota - Twin Cities, Department of Computer Science}
}
\maketitle

\begin{abstract}
...
\end{abstract}

\section{Introduction}
\label{intro}

Many modern online platforms simultaneously compete for users as well as learn from the users they manage to attract. For instance, Google Search and Bing compete for users in the search engine market yet at the same need to experiment with their search algorithms to learn what algorithms work the best. This creates a tradeoff between \textit{exploration} and \textit{competition} since firms need to experiment with potentially sub-optimal options for the benefit of gaining information to make better decisions tomorrow while at the same time firms need to incentivize consumers to select them over their competitors today.

From a social welfare perspective it is better for firms to adopt ``smarter" learning algorithms and our main high-level question is asking under what levels of competition are firms incentivized to adopt ``better" learning algorithms. Our primary contribution is that in our model, even in the case of a duopoly, firms are not incentivized to adopt better learning algorithms but that allowing one firm to be a temporary monopolist will incentivize this firm to play a better learning algorithm.

Unlike in the classic models in economics regarding the interplay between competition and innovation described in \citet{barro2004economic}, here we have no explicit R \& D cost. Rather, firms are not incentivized to experiment under duopoly because experimentation has an implicit relative \textit{reputational} cost compared to the myopic alternative and, in our model, consumers only care about the reputation of a firm. Thus, competition disincentivizes experimentation by firms precisely because competition forces firms to take actions in order to incentivize consumers to select them over their competitors and, as a result, by allowing a firm to be a temporary monopolist the firm can experiment without worrying about the reputational consequences. Our findings give an alternative intuition to the empirically documented inverse-U relationship between competition and innovation discussed in \citet{aghion2005competition}.

\textbf{Related Work} A major underlying component of our model is that firms face a multi-armed bandit (MAB) problem. Multi-armed bandits (MAB) are a tractable abstraction for the tradeoff between exploration and exploitation. MAB problems have been studied for many decades (see \citet{bubeck2012regret} for an overview). Most relevant to this paper is the thread that focuses on designing ``smart" and tractable algorithms that combine exploration and exploitation and ``naive" algorithms that separate exploration and exploitation (see \citet{slivkins2017bandits}).

The three-way tradeoff between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}. The strategic experimentation literature in economics, such as \citealt*{bolton1999strategic} and \citealt*{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.

The relationship between competition and innovation has been heavily studied in industrial organization \citep{tirole1988theory} and endogenous growth theory \citep{aghion2005competition, barro2004economic}, dating back to \citet{schumpeter2010capitalism}.

The most closely related work to this paper is \citealt*{mansour2017competing}. Our motivating questions are the same as \citet{mansour2017competing} though while \citet{mansour2017competing} use the rationality of consumers as their primary knob of competitiveness, we consider differences in the number of firms in the market as our primary knob. Additionally, while in \citet{mansour2017competing} agents select firms based on the Bayesian expected reward of each principal in our model the agents select firms based on a frequentist reputation score that is derived from the signals sent from past agents. This allows the model to become tractable for the purposes of evaluating it via simulation and this is the primary method of analysis that we employ.

\section{Model}
\label{model}

\textbf{Overview} There are two principals and $T+2k$ agents where $k$ is the warm start, or the agents that each principal gets for free at the beginning of the game. The timing of events is as follows:
\begin{enumerate}
\item At $t = 0$, the principals simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
\item Still at $t=0$ we suppose that each principal gets $k$ agents as a ``warm" start.
\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses among 2 principals, given a reputation score for both principals.
\item The principal that is chosen selects an action $a_{t} \in A$ from a set of fixed actions across principals and rounds.
\item Both the agent and the principal observe the reward $r_t \in [0, 1]$ from the action. The agent reports this reward to the principal and the future agents and the reputation score for the chosen principal is updated.
\item Repeat 2-4 for $T$ rounds.
\end{enumerate}

Generally, the rewards are i.i.d. with a common prior.  For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors. Each principal faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a ``fake" prior of $Beta(\alpha=1,\beta=1)$}. We assume that the principals commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.

\noindent \textbf{Agents} We suppose that agents are myopic and non-strategic. The utility function for the agents is simply to maximize their reward in the one period in which they are alive. We suppose that agents do not attempt to manipulate the strategy of the principal nor do they take the strategies of the principal into account when choosing between the principals. In our model, each agent uses the average reward of past agents as a proxy for their expected utility. For simplicity, the reputation score, $R_jt$ is defined as a sliding window average.
\begin{center}
$R_{jt} = \frac{1}{M} \sum\limits_{i=1}^{M} r_{t_j-i}$
\end{center}

Note that the reputation score is in terms of \textit{local} and not global time so that it is the sliding window average of the last $M$ times that principal $j$ was selected by the agents. The warm start of $k$ rounds allows this reputation score to be well-defined once the ``competition game" begins. In our model the agents deterministically choose the principal with the higher reputation score and ties are broken uniformly at random.

\noindent \textbf{Principals} We suppose that the principals simply care about maximizing market share (i.e. maximizing the number of $T$ agents who select them). We model the ``competition" game between principals as a simultaneous move game where both principals commit to a learning algorithm.

\noindent \textbf{MAB algorithms} We suppose that principals commit to a learning algorithm from $\mathcal{A}$. We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
\begin{enumerate}
\item ``Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation. We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is ``optimal" in the sense of having the highest mean reward \citep{agrawal2012analysis}.
\item ``Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation. We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability. For our experiments we keep $\epsilon = 0.05$ fixed.
\item Greedy / myopic algorithms that engage in no purposeful exploration and take the best short-sighted action. We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
\end{enumerate}

It is known that $TS > DEG > DG$ in the standard multi-armed bandit problem of minimizing total regret. The primary question that we want to understand is when, in competition, are the principals incentivized to adopt the ``better" algorithms?

\noindent \textbf{Incumbent} In some of our experiments we modify our model so that one principal enters the market $X$ number of rounds before the other. We refer to the principal that enters before as the ``incumbent" and the principal that enters after $X$ rounds as the ``entrant." The agents that arrive in these $X$ rounds are forced to select the only firm in the market as there is no outside option. We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$. We have that both the incumbent and entrant commit to a learning algorithm \textit{before} either firm receives any agent. After the $X$ rounds, both firms still receive the $k$ warm start agents.

\section{Simulation Details}
\label{sim_details}

We evaluate the consequences of our model via simulation.

\textbf{Bandit Priors} We look at bandit instances drawn from three different bandit ``priors" that are different types of learning problems. Recall that we only consider Bernoulli-distributed arms and so we draw the true means of the arms from these priors. For these experiments we fix the number of arms we consider, $K = 10$.
\begin{enumerate}
\item Needle In Haystack - $K-1$ arms with mean 0.5 and 1 arm with 0.7.
\item Uniform - the $K$ arms have means drawn uniformly at random from $[0.25, 0.75]$
\item "Heavy Tail" - the $K$ arms have means drawn from $Beta(\alpha=0.6, \beta = 0.6)$. With this prior it was likely to have means at the ``extremes" or means that were close to 0 as well as means that were close to 1.
\end{enumerate}

\noindent \textbf{Simulation of Competition Game} Unless otherwise noted, all of the reported results utilize the same set of randomly drawn bandit instances and realizations from the prior. Namely, for each bandit prior we draw $N = 1000$ bandit instances. For each of these instances we run simulations of our model for varying values of $k$ and $X$. We take the maximum values of $k$ and $X$ for the simulations we run, $k_{max}$ and $X_{max}$ respectively, and compute a realization table of dimension $(T+k_{max}+X_{max}) \times K$.

This realization table, as well as fixing the random seed for the same bandit instance and realization table across experiments, ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings. In the competition game we draw from the $T \times K$ portion of the table, so that if two different algorithms picks arm $a$ at time $t$, they get the same $[a, t]$ realization in the table. This setup also ensures that in the warm start period, increasing the warm start from $k$ to $k + 10$ results in the same behavior in the first $k$ rounds.

For the simulations we fix the sliding window size $M = 100$. Low values of $M$ induced too much random noise into the results and we found that increasing $M$ to be larger than $100$ did not make a substantial qualitative difference so we fix this value.

\section{Performance in Isolation}
\label{iso_section}

We evaluate the performance of the different learning algorithms on the bandit priors we consider in isolation. There are two reasons why this is important to consider. First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where $TS > DEG > DG$. Second, we want to better understand what statistics of the instance we can look at in isolation in order to help us predict and understand what to expect in the competition game.

\begin{figure}
\caption{Mean Reputation Trajectories in Isolation}
\includegraphics[scale=0.25]{"figures/nih_iso_mean"}
\label{prelim_means}
\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs. The shaded area display 95\% confidence intervals.}}
\end{figure}

Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior. The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the GitHub appendix.

In the ``competition" game the agents decide between the principals using the \textit{relative} reputation between the two principals. Namely, if principal $A$ has a higher reputation than principal $B$ agents will select principal $A$. A natural question to ask is whether looking at the mean performance is sufficient for understanding performance in the competition game. Figure \ref{rep_dist_nih} displays the reputation distribution at $t = 500$ for the same prior we reported the mean plot for previously. We see that the ``naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not. The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not. If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds so that the weak law of large numbers comes into play (our $M$ is set sufficiently high to ensure this most of the time). However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$. In these cases its reputation may be substantially worse but for the competition game the difference between the reputation does not matter but only the binary comparison between them \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, considering the SoftMax decision rule in \citet{mansour2017competing}}.

\begin{figure}
\caption{Reputation Distribution}
\includegraphics[scale=0.3]{figures/rep_distribution_nih}
\label{rep_dist_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 200$}}
\end{figure}

This motivates looking at the entire distribution of reputation difference between two algorithms. Figure \ref{ts_dg_rep_diff_nih} shows the distribution of reputation difference between $TS$ and $DG$ across $t$. This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right. Thus, the mean is not a representative statistic of the entire reputation difference distribution. As an alternative statistic for understanding the results of the competition game we introduce the \textit{relative reputation statistic} which looks at the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$. This corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a principal playing $A$ over a principal playing $B$ at time $t$ \footnote{Though we do not discuss it here, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion. In the GitHub appendix we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.}.

\begin{figure}
\caption{Reputation Difference Distribution}
\includegraphics[scale=0.3]{figures/ts_dg_rep_diff_nih}
\label{ts_dg_rep_diff_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $TS$ and $DG$ across $t$}}
\end{figure}

\begin{figure}
\caption{Relative Reputation Plots}
\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
\caption*{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$. The shaded area display 95\% confidence intervals.}
\label{relative_rep_plots}

\end{figure}

Figure \ref{relative_rep_plots} shows the relative reputation plots for $TS$ vs $DG$ on the Uniform and Needle In Haystack prior. For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $TS > DG$. The intuition behind this is that, especially since the principals start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information. However, this exploration leads to what we define as a \textit{relative reputation cost}. Namely, exploration leads to lower reputation relative to the myopic alternative. However, eventually the information acquired in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm. We define this as the \textit{relative reputation benefit} \footnote{We see similar behavior for the Heavy Tail prior}.

Is exploration always costly? Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $TS$ always does relatively better than $DG$. There are two contributing factors to this. First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration. Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same. Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm. However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.

\section{Monopoly vs Duopoly}

In this section we utilize the same instances and realizations as in the previous section and simulate the model described in section \ref{model}. We initially take the strategies of the firms as exogenous and simulate the model given these strategies in order to determine the payoffs associated with each pair of algorithms. Unless otherwise noted, all the results are reported at $t = 2000$.

\bibliography{refs}

\end{document}

### 2018-08-15 analysis_code/compare_across.R

@@ -1,15 +1,28 @@
 library(dplyr)
 
 #dat_1 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
-#dat_2 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/tournament_experiment_full_sim_raw.csv")
+#dat_2 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv")
 
-dist <- "Heavy Tail"
+
+concise_alg_rep <- function(alg) {
+  if (alg == "ThompsonSampling") {
+    return("TS")
+  } else if (alg == "DynamicEpsilonGreedy") {
+    return("DEG")
+  } else if (alg == "DynamicGreedy") {
+    return("DG")
+  }
+}
+
+dat_1$Algorithm<- sapply(dat_1$Algorithm, concise_alg_rep)
+tmp <- mutate(dat_1, Algorithm = concise_alg_rep(Algorithm))
+dist <- ".5/.7 Random Draw"
 ws_time <- 20
 iso_dat <- filter(dat_1, Distribution == dist)
 compet_dat <- filter(dat_2, Prior == dist & Warm.Start == 20)
 
-alg1 <- "ThompsonSampling"
-alg2 <- "DynamicGreedy"
+alg1 <- "TS"
+alg2 <- "DG"
 alg_1 <- filter(iso_dat, Algorithm == alg1)
 alg_2 <- filter(iso_dat, Algorithm == alg2)
 n_vals <- unique(iso_dat$n)
@@ -34,14 +47,41 @@ cat("Losing dat", nrow(losing_dat), "Market share", mean(losing_dat$Market.Share
 
 # look at density estimates of reputation
 
-dist <- "Heavy Tail"
+dist <- "Needle In Haystack"
 ws_time <- 20
 iso_dat <- filter(dat_1, Distribution == dist)
-iso_dat_t <- filter(iso_dat, t == 100)
-ggplot(iso_dat_t, aes(Realized.Reputation, colour=Algorithm)) + geom_density()
+iso_dat_t <- filter(iso_dat, t == 500)
+ggplot(iso_dat_t, aes(Realized.Reputation, colour=Algorithm)) +
+  geom_density() +
[…]

### 2018-08-15 analysis_code/relative_rep.jl

@@ -4,8 +4,8 @@ using Query
 using DataFrames
 using PyPlot
 
-dat = CSV.read("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
-distribution = "Uniform"
+dat = CSV.read("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/ht_long_horizon_3_arms.csv")
+distribution = "Heavy Tail"
 
 function get_prob_rep_greater(data, alg1, alg2, t)
   alg_1_dat = @from i in data begin
@@ -25,14 +25,13 @@ function get_prob_rep_greater(data, alg1, alg2, t)
   return quadgk(x -> pdf(est, x), 0, maximum(est.x))[1]
 end
 
-
 filtered_dat =  @from i in dat begin
   @where i.Distribution == distribution
   @select i
   @collect DataFrame
 end
 
-t_vals = collect(range(100,100,20))
+t_vals = collect(range(500,500,14))
 probs = []
 alg1 = "ThompsonSampling"
 alg2 = "DynamicGreedy"

### 2018-08-15 analysis_code/tournament_table.R

@@ -5,7 +5,7 @@ WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
 
 
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_no_info_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
 p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
@@ -51,7 +51,7 @@ for (j in 1:length(time_horizons)) {
           results[k, l] <- paste("\\makecell{\\textbf{", signif(mean(share), digits=2), "} $\\pm$", cin, "\\\\Variance:", signif(var(share), digits=1), "\\\\", "ES:", 100*signif(shares, digits=2), "\\%}")
         }
       }
-      tab <-xtable(results, caption=paste("Incumbent Experiment", prior, "X=200"))
+      tab <-xtable(results, caption=paste("Information Erased Experiment", prior, "X=200"))
       print(tab, type="latex")
     }
   }

### 2018-08-16 paper/competing_bandits.tex
finally finished a draft

--- 2018-08-12 paper/competing_bandits.tex
@@
 \theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
+\newtheorem{definition}{Definition}
 \captionsetup{font=footnotesize}
 \bibliographystyle{aaai}
-
-\title{Competing Bandits}
 
@@
 
-\author{-}
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-
 \begin{document}
 
+\title{Learning and Reputation under Monopoly and Duopoly: \\A Competing Bandits Approach}
+
+\author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
+Zhiwei Steven Wu\textsuperscript{4} \\,
+{\textsuperscript{1}Columbia University, Department of Economics}\\
+{\textsuperscript{2}Columbia University, Department of Computer Science}\\
+{\textsuperscript{3}Microsoft Research, New York, NY}\\
+{\textsuperscript{4}University of Minnesota - Twin Cities, Department of Computer Science}
+}
 \maketitle
@@
 \begin{abstract}
-...
+We empirically study the interplay between \textit{exploration} and \textit{competition} - how systems that learn from interactions with users tradeoff between making potentially suboptimal choices in order to acquire new information and the reputational consequences of this exploration which potentially has adverse competitive effects.
+Our model considers competition for myopic users between two firms deploying multi-armed bandit algorithms that face the same underlying multi-armed bandit instance.
+The agents select between the firms according to a reputation score, which is a function of the rewards past agents have experienced from this firm.
+
+We ask whether better algorithms are incentivized under varying degrees of competition.
+The environments we consider are a monopoly, a duopoly with one firm serving as an early entrant (or a temporary monopoly), and a duopoly.
+We show that, under certain conditions, monopoly and duopoly do not incentivize the adoption of better learning algorithms due to the reputational costs of exploration, but that a temporary monopoly may incentivize the adoption of better learning algorithms.
+Finally, we ask whether in this setting the data advantage or the reputational advantage of the early entrant serves as stronger barriers to entry.
 \end{abstract}
@@
 \section{Introduction}
-\label{S:1}
+\label{intro}
 
-...
+Many modern online platforms simultaneously compete for users as well as learn from the users they manage to attract.
+For instance, Google Search and Bing compete for users in the search engine market yet at the same need to experiment with their search algorithms to learn what algorithms work the best.
+This creates a tradeoff between \textit{exploration} and \textit{competition} since firms need to experiment with potentially sub-optimal options for the benefit of gaining information to make better decisions tomorrow while at the same time firms need to incentivize consumers to select them over their competitors today.
 
-\section{Related Work}
-\label{S:2}
+From a social welfare perspective it is better for firms to adopt ``smarter" learning algorithms and our main high-level question is asking under what levels of competition are firms incentivized to adopt ``better" learning algorithms.
+Our primary contribution is that in our model, even in the case of a duopoly, firms are not incentivized to adopt better learning algorithms but that allowing one firm to be a temporary monopolist will incentivize this firm to play a better learning algorithm.
 
-\citealt*{mansour2017competing}, \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}, etc.
+Unlike in the classic models in economics regarding the interplay between competition and innovation described in \citet{barro2004economic}, here we have no explicit R \& D cost.
+Rather, firms are not incentivized to experiment under duopoly because experimentation has an implicit relative \textit{reputational} cost compared to the myopic alternative and, in our model, consumers only care about the reputation of a firm.
+Thus, competition disincentivizes experimentation by firms precisely because competition forces firms to take actions in order to incentivize consumers to select them over their competitors and, as a result, by allowing a firm to be a temporary monopolist the firm can experiment without worrying about the reputational consequences.
+Our findings give an alternative intuition to the empirically documented inverse-U relationship between competition and innovation discussed in \citet{aghion2005competition}.
+
+\textbf{Related Work} A major underlying component of our model is that firms face a multi-armed bandit (MAB) problem.
+Multi-armed bandits (MAB) are a tractable abstraction for the tradeoff between exploration and exploitation.
+MAB problems have been studied for many decades (see \citet{bubeck2012regret} for an overview).
+Most relevant to this paper is the thread that focuses on designing ``smart" and tractable algorithms that combine exploration and exploitation and ``naive" algorithms that separate exploration and exploitation (see \citet{slivkins2017bandits}).
+
+The three-way tradeoff between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}.
+The strategic experimentation literature in economics, such as \citealt*{bolton1999strategic} and \citealt*{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.
+
+The relationship between competition and innovation has been heavily studied in industrial organization \citep{tirole1988theory} and endogenous growth theory \citep{aghion2005competition, barro2004economic}, dating back to \citet{schumpeter2010capitalism}.
+
+Our setting is also closely related to the ``dueling algorithms" framework described in \cite{immorlica2011dueling}, but this framework considers offline scenarios whereas we focus on online learning problems.
+
+The most closely related work to this paper is \citealt*{mansour2017competing}.
+Our motivating questions are the same as \citet{mansour2017competing} though while \citet{mansour2017competing} use the rationality of consumers as their primary knob of competitiveness, we consider differences in the number of firms in the market as our primary knob.
+Additionally, while in \citet{mansour2017competing} agents select firms based on the Bayesian expected reward of each principal in our model the agents select firms based on a frequentist reputation score that is derived from the signals sent from past agents.
+This allows the model to become tractable for the purposes of evaluating it via simulation and this is the primary method of analysis that we employ.
 
 \section{Model}
-\label{S:3}
+\label{model}
 
-...
-
-\section{Simulation Details}
-\label{S:4}
-
-\subsection{Bandit Priors}
-We study the implications of our model via simulation.
-For computational tractability we focus our attention on Bernoulli bandit diBernoulli bandits with beta priors.
-We consider the following bandit ``priors" from which the bandit instances faced in simulation are drawn:
+\textbf{Overview} There are two principals and $T+2k$ agents where $k$ is the warm start, or the agents that each principal gets for free at the beginning of the game and $T$ is the number of rounds in the game.
+The timing of events is as follows:
 \begin{enumerate}
-\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$).
-With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
-[We should rename this prior instance]
-\item Needle-in-haystack High - $K-1$ arms with mean 0.50, 1 arm with mean 0.70. 
+\item At $t = 0$, the principals simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
+\item Still at $t=0$ we suppose that each principal gets $k$ agents as a ``warm" start.
+\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses among 2 principals, given a reputation score for both principals.
+\item The principal that is chosen selects an action $a_{t} \in A$ from a set of fixed actions across principals and rounds.
+\item Both the agent and the principal observe the reward $r_t \in [0, 1]$ from the action.
+The agent reports this reward to the principal and the future agents and the reputation score for the chosen principal is updated.
+\item Repeat 2-4 for $T$ rounds.
 \end{enumerate}
 
-[Need some discussion of what makes these priors at all ``representative" - why can we just focus on them?]
+Generally, the rewards are i.i.d.
+with a common prior.
+For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors.
+Each principal faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a ``fake" prior of $Beta(\alpha=1,\beta=1)$}.
+We assume that the principals commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.
 
-[We can't feasibly include all the priors.
-I think Heavy Tail is interesting because of the relative reputation plots, but I'm not sure what other one to include that would be meaningful.
-For now I've just used NIH]
+\noindent \textbf{Agents} We suppose that agents are homogenous, myopic, and non-strategic.
+The utility function for the agents is simply to maximize their reward in the one period in which they are alive.
+We suppose that agents do not attempt to manipulate the strategy of the principal nor do they take the strategies of the principal into account when choosing between the principals.
+In our model, each agent uses the average reward of past agents as a proxy for their expected utility.
+For simplicity, the reputation score, $R_jt$ is defined as a sliding window average.
+\begin{center}
+$R_{jt} = \frac{1}{M} \sum\limits_{i=1}^{M} r_{t_j-i}$
+\end{center}
 
-In the appendix we consider the performance of additional priors, but qualitatively the results presented are unchanged.
+Note that the reputation score is in terms of \textit{local} and not global time so that it is the sliding window average of the last $M$ times that principal $j$ was selected by the agents.
+The warm start of $k$ rounds allows this reputation score to be well-defined once the ``competition game" begins.
+In our model the agents deterministically choose the principal with the higher reputation score and ties are broken uniformly at random.
 
-\subsection{Learning Algorithms}
+\noindent \textbf{Principals} We suppose that the principals simply care about maximizing their expected market share (i.e.
+maximizing the number of $T$ agents who select them).
+We model the ``competition" game between principals as a simultaneous move game where both principals commit to a learning algorithm.
 
-For our simulations we evaluate the performance of three algorithms, each endowed with a ``fake prior" of $Beta(1, 1)$.
-In general, there are three different classes of learning algorithms and we select one representative algorithm from each class:
+\noindent \textbf{MAB algorithms} We suppose that principals commit to a learning algorithm from $\mathcal{A}$.
+We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
 \begin{enumerate}
-\item Adaptive exploration algorithms that engage in purposeful exploration in a ``smart" way by adapting to the previous history of results.
-We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is ``optimal" in the sense of having the highest mean reward.
-\item Non-adaptive exploration algorithms that engage in purposeful exploration without considering the past results.
+\item ``Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
+We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is ``optimal" in the sense of having the highest mean reward \citep{agrawal2012analysis}.
+\item ``Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
 We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability.
 For our experiments we keep $\epsilon = 0.05$ fixed.
-\item Greedy / Myopic algorithms that engage in no purposeful exploration and take the best short-sighted action.
+\item Greedy / myopic algorithms that engage in no purposeful exploration and take the best short-sighted action.
 We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
@@
 
-Run in isolation we expect that the performance of the algorithms (according to cumulative regret) should be \\
- Adaptive Exploration $\geq$ Non-adaptive exploration $\geq$ Greedy.
-We verify that, for the chosen algorithms and chosen priors, this is indeed the case (with some caveats, discussed in the next section).
+It is known that $TS > DEG > DG$ in the standard multi-armed bandit problem of minimizing total regret.
+The primary question that we want to understand is when, in competition, are the principals incentivized to adopt the ``better" algorithms?
 
-The primary question that we are interested in is under what conditions in competition are the firms in our model incentivized to commit to adaptive exploration algorithms.
+\noindent \textbf{Incumbent} In some of our experiments we modify our model so that one principal enters the market $X$ number of rounds before the other.
+We refer to the principal that enters before as the ``incumbent" and the principal that enters after $X$ rounds as the ``entrant." The agents that arrive in these $X$ rounds are forced to select the only firm in the market as there is no outside option.
+We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$.
+We have that both the incumbent and entrant commit to a learning algorithm \textit{before} either firm receives any agent.
+After the $X$ rounds, both firms still receive the $k$ warm start agents.
 
-\subsection{Simulation of Competition Game}
-Algorithm \ref{alg_1} (FIX TO BE FIGURE) describes the simulation procedure utilized to evaluate the results of the competition game.
-$T$ is the finite time horizon of the competition game and $K$ is the number of arms in the bandit instance that is considered.
+\section{Simulation Details}
+\label{sim_details}
 
-\begin{algorithm}
-\caption{Simulation Pseudo-Code}
-\begin{algorithmic}[1]
-\For{Each prior $p$}
-	\For{$N$ simulations}
-		\State Generate true distribution from $p$
-		\State Generate $T \times K$ realizations for the arms 
-		\For{Each agent algorithm $agent alg$}
-			\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
-				\State Give principal 2 $X$ free observations (the agents also get these observations)
-				\State Give the agents $k$ observations from each principal (warm start)
-				\State Run simulation for $T$ periods
-			\EndFor
-		\EndFor
-	\EndFor
-\EndFor
-\end{algorithmic}
-\label{alg_1}
-\end{algorithm}
+We evaluate the consequences of our model via simulation.
+
+\textbf{Bandit Priors} We look at bandit instances drawn from three different bandit ``priors" that are different types of learning problems.
+Recall that we only consider Bernoulli-distributed arms and so we draw the true means of the arms from these priors.
+For these experiments we fix the number of arms we consider, $K = 10$.
+\begin{enumerate}
+\item Needle In Haystack - $K-1$ arms with mean 0.5 and 1 arm with 0.7.
+\item Uniform - the $K$ arms have means drawn uniformly at random from $[0.25, 0.75]$
+\item "Heavy Tail" - the $K$ arms have means drawn from $Beta(\alpha=0.6, \beta = 0.6)$.
+With this prior it was likely to have means at the ``extremes" or means that were close to 0 as well as means that were close to 1.
+\end{enumerate}
+
+\noindent \textbf{Simulation of Competition Game} Unless otherwise noted, all of the reported results utilize the same set of randomly drawn bandit instances and realizations from the prior.
+Namely, for each bandit prior we draw $N = 1000$ bandit instances.
+For each of these instances we run simulations of our model for varying values of $k$ and $X$.
+We take the maximum values of $k$ and $X$ for the simulations we run, $k_{max}$ and $X_{max}$ respectively, and compute a realization table of dimension $(T+k_{max}+X_{max}) \times K$.
+
+This realization table, as well as fixing the random seed for the same bandit instance and realization table across experiments, ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
+In the competition game we draw from the $T \times K$ portion of the table, so that if two different algorithms picks arm $a$ at time $t$, they get the same $[a, t]$ realization in the table.
+This setup also ensures that in the warm start period, increasing the warm start from $k$ to $k + 10$ results in the same behavior in the first $k$ rounds.
+
+For the simulations we fix the sliding window size $M = 100$.
+Low values of $M$ induced too much random noise into the results and we found that increasing $M$ to be larger than $100$ did not make a substantial qualitative difference so we fix this value.
 
 \section{Performance in Isolation}
-\label{S:5}
+\label{iso_section}
 
-We look at the performance of each of our learning algorithms in isolation, meaning that we simply run the learning algorithm for $T$ steps on a realization of the prior instances that we consider and record the realized rewards and the arm that is pulled.
-In our simulations the realizations of each of the arms were pre-drawn and the same across all algorithms and so the differences in algorithm performance come from the difference in the arms selected by the algorithms and not from differences in realizations.
-When looking at the results of each of the learning algorithms in isolation we want to confirm that, for the prior instances that we consider, ``better" algorithms win according to ``absolute" measures of performance (i.e.
-looking at mean reputation).
-Additionally, we want to understand what statistics we can look at in isolation to understand what wins under competition.
-
-In our model the reputation score is an input to the decision rule of the agents and thus it is important to understand how this score differs across algorithms when run outside of competition.
-Figure \ref{prelim_means} displays the mean average reputation score of each of the algorithms across the different simulations.
+We evaluate the performance of the different learning algorithms on the bandit priors we consider in isolation.
+There are two reasons why this is important to consider.
+First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where $TS > DEG > DG$.
+Second, we want to better understand what statistics of the instance we can look at in isolation in order to help us predict and understand what to expect in the competition game.
 
@@
 \caption{Mean Reputation Trajectories in Isolation}
-\includegraphics[scale=0.2]{"figures/heavy_tail_iso_mean"}
-\includegraphics[scale=0.2]{"figures/nih_iso_mean"}
+\includegraphics[scale=0.25]{"figures/nih_iso_mean"}
 \label{prelim_means}
-\caption*{\tiny{The plots contain the average reputation over $250$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
-TODO: get rid of smoothing on plot}}
+\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
+The shaded area display 95\% confidence intervals.}}
 \end{figure}
+
+Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior.
+The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the GitHub appendix.
+
+In the ``competition" game the agents decide between the principals using the \textit{relative} reputation between the two principals.
+Namely, if principal $A$ has a higher reputation than principal $B$ agents will select principal $A$.
+A natural question to ask is whether looking at the mean performance is sufficient for understanding performance in the competition game.
+Figure \ref{rep_dist_nih} displays the reputation distribution at $t = 500$ for the same prior we reported the mean plot for previously.
+We see that the ``naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not.
+The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not.
+If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds so that the weak law of large numbers comes into play (our $M$ is set sufficiently high to ensure this most of the time).
+However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$.
+In these cases its reputation may be substantially worse but for the competition game the difference between the reputation does not matter but only the binary comparison between them \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, considering the SoftMax decision rule in \citet{mansour2017competing}}.
+
+\begin{figure}
+\caption{Reputation Distribution}
+\includegraphics[scale=0.3]{figures/rep_distribution_nih}
+\label{rep_dist_nih}
+\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 200$}}
+\end{figure}
+
+This motivates looking at the entire distribution of reputation difference between two algorithms.
+Figure \ref{ts_dg_rep_diff_nih} shows the distribution of reputation difference between $TS$ and $DG$ across $t$.
+This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right.
+Thus, the mean is not a representative statistic of the entire reputation difference distribution.
+As an alternative statistic for understanding the results of the competition game we introduce the \textit{relative reputation statistic} which looks at the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$.
+This corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a principal playing $A$ over a principal playing $B$ at time $t$ \footnote{Though we do not discuss it here, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion.
+In the GitHub appendix we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.}.
+
+\begin{figure}
+\caption{Reputation Difference Distribution}
+\includegraphics[scale=0.3]{figures/ts_dg_rep_diff_nih}
+\label{ts_dg_rep_diff_nih}
+\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $TS$ and $DG$ across $t$}}
+\end{figure}
+
+Figure \ref{relative_rep_plots} shows the relative reputation plots for $TS$ vs $DG$ on the Uniform and Needle In Haystack prior.
+For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $TS > DG$.
+The intuition behind this is that, especially since the principals start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information.
+However, this exploration leads to what we define as a \textit{relative reputation cost}.
+Namely, exploration leads to lower reputation relative to the myopic alternative.
+However, eventually the information acquired in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm.
+We define this as the \textit{relative reputation benefit} \footnote{We see similar behavior for the Heavy Tail prior}.
 
@@
 \caption{Relative Reputation Plots}
-\includegraphics[scale=0.2]{"figures/deg_dg_ht_10_prelim"}
-\includegraphics[scale=0.2]{"figures/ts_dg_ht_10_prelim"} \\
-\includegraphics[scale=0.2]{"figures/deg_dg_nih_10_prelim"}
-\includegraphics[scale=0.2]{"figures/ts_dg_nih_10_prelim"} \\
-\caption*{\tiny{The plots contain the proportion of simulations over $250$ runs for a memory size of $100$ where, for a given $t$, the reputation of one algorithm is at least as high as the reputation of another on the same instance.}}
-\label{relative_rep}
+\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
+\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
+\caption*{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$.
+The shaded area display 95\% confidence intervals.}
+\label{relative_rep_plots}
+
 \end{figure}
 
-As expected we see that according to the mean reputation plots $TS > DEG > DG$ for sufficiently large values of $t$.
-While looking at the mean reputation is a natural statistic to consider when attempting to understand what algorithm wins under competition, we consider an additional statistic when attempting to understand what wins under competition.
-In the competition game the decision rule of the agents compares the relative reputation of the two firms so that if the reputation of firm $A$ is higher than that of firm $B$, the agents will (be more likely to) select firm $A$.
-Thus, another natural statistic to look at is to look at the two algorithms in isolation and consider the fraction of simulations where, for a fixed time $t$, the reputation of firm $A$ is at least as high as that of firm $B$.
-Figure \ref{prelim_comparison} displays the proportion of rounds where one algorithm has at least as high of a reputation as another.
+Is exploration always costly?
+Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $TS$ always does relatively better than $DG$.
+There are two contributing factors to this.
+First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration.
+Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
+Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
+However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
 
-Interestingly we observe that, on the Heavy Tail prior, $DG > DEG$ according to this metric even though $DEG > DG$ according to the mean reputation.
-The intuition for this is simple.
-$DG$ will either find the best arm or not.
-If it does then since $DEG$ engages in non-adaptive exploration $DG$ will, with $\epsilon$ probability, pick an arm uniformly at random which potentially hurts its reputation if the selected arm is not the best arm.
-$DEG$ eventually converges on the best arm with a sufficiently large time horizon, but $DG$ does not necessarily do so and so when $DG$ does not find the best arm it may potentially have substantially worse reputation which drags down its mean reputation.
-However, if it is able to find the best arm in sufficiently many simulations then it may have slightly higher reputation than $DEG$ in those simulations and thus do better according to the relative reputation proportion but, due to the simulations where it does not find the best arm it may then have worse mean reputation.
-In the next section we move to considering the competition game and revisit if this statistic is more useful for predicting who wins in competition.
+\section{Competition vs Better Algorithms: Inverted-U}
 
-\section{Simultaneous Entry Experiment}
-\label{S:6}
+In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \citet{aghion2005competition}.
+We consider three separate settings for our model: permanent monopoly, temporary monopoly, and permanent duopoly.
+In the permanent monopoly and permanent duopoly cases we show that, under certain conditions, firms are incentivized to commit to $DG$ whereas in the temporary monopoly case firms are incentivized to commit to $TS$.
+We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition.
+We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index).
+In our model we have no prices and no costs and so we interpret increases in ``market power" as decreases in the number of rounds where a firm has to compete for agents.
 
-In this section we present simulation results of the model discussed in Section \ref{S:2} where $X = 0$ so that both firms enter at the same time.
-We want to understand what algorithms the firms are incentivized to commit to in the competition game but the payoffs associated with each algorithm are too difficult to compute analytically and thus, we begin by simply imposing that the algorithm played by the firm is taken as given and study the consequences of playing them in the competition game.
-Additionally, for this experiment we present results for the $HardMax$ agent decision rule and postpone discussion of alternative decision rules.
-Besides keeping track of market share, we also will keep track of the ``effective end of game."
+We utilize the same instances and realizations as in the Section \ref{iso_section} and simulate the model described in section \ref{model}.
+We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the payoffs associated with each pair of algorithms.
+Unless otherwise noted, all the results are reported at $t = 2000$.
+
+\textbf{Permanent Monopoly} Under permanent monopoly only a single firm exists in the market for every period.
+Since, in our formulation, the firm only gets utility from a larger market share it is indifferent between each algorithm in $\mathcal{A}$ since, regardless of the algorithm it deploys, it will get the entire market.
+If we suppose that deploying the ``better" algorithms has even an $\epsilon > 0$ cost, then the firm would choose to deploy $DG$.
+
+\textbf{Permanent Duopoly} Permanent duopoly corresponds to the model described in section \ref{model} where $X = 0$ so that both firms enter the market simultaneously.
+What algorithms are firms incentivized to deploy?
+We report the average market share taken over the $N$ simulations for a set of exogenous values of $k$.
+Additionally, we report the mean and median of an additional quantity which we define as the effective end of game.
 
 \begin{definition}
-\textit{Effective End of Game (EEOG)} - the last round, $t$, in a simulation where there is a difference between the decision of the agent alive at $t-1$ and the agent alive at $t$ (todo: state this formally using notation of model)
+\textit{Effective End of Game (EEOG)} - the last round, $t$, in a simulation where the agent alive at $t-1$ and the agent alive at $t$ choose different firms.
 \end{definition}
+
+Since both firms enter at the same time the game is symmetric and we only need to compute the payoffs of three pairs of strategies.
+We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.
+Tables \ref{sim_nih}, \ref{sim_ht}, and \ref{sim_unif} report the results.
+\begin{enumerate}
+\item Under Needle In Haystack we have that $TS$ always does weakly better than $DEG$ and $DG$, regardless of the warm start
+\item Under Heavy Tail we have that, for low warm start, $DEG, DG > TS$ but that for the higher warm start values we consider $TS > DEG, DG$.
+Interestingly, even under the relatively high warm start we see $DEG > DG$.
+\item Under Uniform we have that $DEG, DG > TS$ for low warm start and $TS > DEG, DG$ for sufficiently high warm start.
+\end{enumerate}
+
+Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start value as fixing the starting point on the relative reputation plots.
+The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation statistic of $A \geq B$ at time $k$.
+Combined with the fact that the EEOG values being relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right, similar to the distribution of reputation differences}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.
+
+For the low values of the warm start we can see that $DG > TS$ for the Uniform prior but $TS > DG$ for the Needle In Haystack prior and this corresponds to what we see in the results of the competition game simulation.
+The intuition for this result aligns with the intuition given in the discussion of the relative reputation plots.
+Namely, for low warm start values $TS$ incurs a relative reputation cost compared to $DG$ since it engages in purposeful exploration.
+Moving to a sufficiently large warm start so that the better algorithm has recovered from the relative reputation cost incurred via exploration leads to better algorithms winning.
+However, it appears that the warm starts required for this are unreasonably high, especially for the Uniform prior.
+
+Focusing on reasonable warm starts, $DG > TS$ when it appears that better algorithms suffer an early relative reputational cost due to exploration and that both firms enter at the same time so that there is a ``permanent duopoly."
 
@@
    \hline
-   \label{hm_nih}
 \end{tabular}
+\label{sim_nih}
 \caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95 \% confidence band.
-For example, the cell in the top left indicates that TS gets on average 32\% of the market when played against DG.
+For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
 The next line contain the average and median effective end of game for this set of simulations.}}
 \end{table}
-
+% latex table generated in R 3.4.0 by xtable 1.8-2 package
+% Wed Aug 15 15:06:22 2018
 \begin{table}[ht]
@@
    \hline
-   \label{hm_ht}
 \end{tabular}
+\label{sim_ht}
+\end{table}
+% latex table generated in R 3.4.0 by xtable 1.8-2 package
+% Wed Aug 15 15:06:22 2018
+\begin{table}
+\centering
+\caption{Simultaneous Entry Uniform} 
+\begin{tabular}{rlll}
+  \hline
+ & k = 20 & k = 100 & k = 200 \\ 
+  \hline
+TS vs DG & \makecell{\textbf{0.43} $\pm$0.03\\ eeog \\ avg: 260\\ med: 0} & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 560\\ med: 290} & \makecell{\textbf{0.48} $\pm$0.03\\ eeog \\ avg: 770\\ med: 706.5} \\
+  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.37} $\pm$0.03\\ eeog \\ avg: 510\\ med: 222} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 810\\ med: 785.5} \\
+  DG vs DEG & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 530\\ med: 163.5} & \makecell{\textbf{0.5} $\pm$0.02\\ eeog \\ avg: 890\\ med: 922.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 990\\ med: 1071} \\
+   \hline
+\end{tabular}
+\label{sim_unif}
 \end{table}
 
-[IDEA: take a look at the distribution of reputation induced by each algorithm under each prior]
+\textbf{Temporary Monopoly} We now consider asymmetries in the timing of entry so that one firm enters the market before the other and serves as a monopolist in the periods until the other firm enters.
+In terms of our model, this corresponds to setting $X = 200$.
+For this section we report results fixing $k = 20$.
+Does this lead the incumbent to commit to $TS$?
+How about the entrant?
 
-Tables \ref{hm_nih} and \ref{hm_ht} display the results.
-Under Heavy Tail we see that for small warm start of $k = 20$, $DG$ gives the firm more of the market than $TS$ and $DEG$.
-As we increase the warm start to $k = 100, 200$ we observe that $TS$ does better than $DG$ and $DEG$.
-Note that, in particular for the simulations involving $TS$, the effective end of game had a median of 0.
-This means that the result of the simulation was \textit{completely determined by the warm start period}.
-Thus, when reasoning about what algorithm would win in the competition game under the HardMax decision rule we need to think about the performance of the algorithm on $k$ samples instead of $T$ samples, which may be small.
-For low warm start, it is thus not necessarily the case that $TS$ is the best algorithm.
-However, we see that under the Needle In Haystack prior $TS$ is always the best algorithm across all warm start values.
+Table \ref{ht_incum} shows the results of the simulations for this parameterization.
+We observe that, for the incumbent, $TS$ is the dominant strategy across all priors.
+In the appendix we include additional simulations where $X = 50$ and observe that $TS$ is not always a dominant strategy in this case.
+This result is robust across all of the priors that we considered so that for sufficiently large $X$, $TS$ is a dominant strategy for the incumbent where $X$ depends on the set of instances that we are considering.
 
-[Missing here is a discussion of the fact that with HardMax the market gets taken by one firm completely (in most simulations) as well as why this is the case.]
-
-First, we note that the relative reputation plots were better at predicting the result of the competition game than the ``absolute" mean reputation plots.
-For the Needle In Haystack prior, both plots showed that $TS > DEG > DG$ in isolation and this is also the result of the competition game.
-However, for the Heavy Tail prior, we see that $DG, DEG > TS$ for low warm start and that, in general, $DEG > DG$ \footnote{The result that $DEG > DG$ for this prior is robust across decision rules as well.}.
-While the ``absolute" reputation plots imply that $TS > DEG > DG$, we see that the relative reputation plot in Figure \ref{relative_rep} implies that $DEG > DG$ and in Figure \ref{rel_rep_ht_fine} we can see that in the early rounds, $DG$ does relatively better than $TS$ and that, around $t = 100$, $TS$ does relatively better than $DG$.
-Note that since the median effective end of game is $0$ in \ref{hm_ht} and the agents are using the $HardMax$ decision rule, this means that the relative reputation between the two firms after $k$ samples ends up determining who takes the market in the simulations and this is precisely what the relative reputation plots tell us.
-Thus, looking at the relative reputation plot at time $k$ gives us a good predictor of which algorithm will take more of the market in the competition game.
-
-Why is it that we see that $TS$ always wins under Needle in Haystack but, for lower warm start, $DG > TS$?
-We know that for a sufficiently long time horizon, $TS$ leads to a higher reputation than $DG$.
-However, in our model, firms need to incentivize the agents to select them over their competition.
-One possible intuition for the results here are that, in the early rounds, $TS$ engages in purposeful exploration whereas $DG$ is myopic.
-A consequence of exploration in our model is that it may lead to a reputational cost relative to the greedy algorithm since the sub-optimal arms pulled during exploration can harm reputation relative to the greedy algorithm.
-$TS$ thus incurs a relative reputational cost in the early rounds, but once it has explored enough and gained enough information it is able to recover the relative reputational loss and acquire a larger reputation than the greedy algorithm.
-A sufficiently large warm start allows $TS$ to explore in the early rounds and recover the reputational loss and thus win the competition game whereas a low warm start is not sufficient for this and thus $DG$ beats $TS$.
-Why do we only see this on Heavy Tail and not Needle in Haystack?
-One possible explanation for this is that learning on Heavy Tail is harder (use learning hardness metric here) and learning on Needle In Haystack is easier and thus TS needs more samples in the Heavy Tail instances compared to the Needle In Haystack instances in order to be able to recover the relative reputation loss [TODO: Explain this a bit more and with more formality.
-Can we have some experimental results that ``prove" this or generalize it?]
-
-[TODO: Can we say anything general about this using the ``learning hardness" metric that we have started looking at?]
-
-\section{Incumbent Experiment}
-\label{S:7}
-
-We now fix the same parameters as before, fixing a warm start of $k = 20$ [we should re-run this] but let $X = 200$ so that firm 2 has an ``incumbency" advantage.
-Recall that $X$ controls the number of rounds the firm is in the market by itself before the other firm enters and so in these rounds agents arrive and must choose firm 2.
-Importantly, we treat $X$ as an exogenous parameter (and not something optimally chosen by the entrant) and study the consequences of a fixed value of $X$.
-Thus, firm 2 is a monopolist for $X$ periods and, after $X$ rounds, firm 1 enters the market and each agent that enters in subsequent periods must select between the two firms.
-We want to understand what algorithms the incumbent (and the entrant) are incentivized to play.
-In particular, we are interested if introducing this asymmetry in entry induces firms who were previously incentivized to play $DG$ to play $TS$.
-As a result, we present the results of this experiment for $k = 20$ and on the Heavy Tail prior, where we saw that $DG$ was preferred to $TS$.
-
+% latex table generated in R 3.4.0 by xtable 1.8-2 package
+% Wed Aug 15 19:03:20 2018
 \begin{table}[ht]
 \centering
-\caption{Incumbent Experiment Heavy Tail X=200} 
-\begin{tabular}{rlll}
-\multicolumn{3}{c}{Incumbent Algorithm} \\
-\multirow{3}{*}{\rotatebox{90}{Entrant Algorithm}} \\
+\caption{Temporary Monopoly Experiment Heavy Tail} 
+\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
   \hline
@@
   \hline
-TS & \makecell{\textbf{ 0.0029 } $\pm$ 0.0029 \\Variance: 0.002 \\ ES: 100 \%} & \makecell{\textbf{ 0.11 } $\pm$ 0.019 \\Variance: 0.09 \\ ES: 96 \%} & \makecell{\textbf{ 0.17 } $\pm$ 0.022 \\Variance: 0.1 \\ ES: 95 \%} \\
-  DEG & \makecell{\textbf{ 0.049 } $\pm$ 0.011 \\Variance: 0.03 \\ ES: 94 \%} & \makecell{\textbf{ 0.24 } $\pm$ 0.022 \\Variance: 0.1 \\ ES: 74 \%} & \makecell{\textbf{ 0.24 } $\pm$ 0.023 \\Variance: 0.1 \\ ES: 79 \%} \\
-   DG & \makecell{\textbf{ 0.12 } $\pm$ 0.017 \\Variance: 0.08 \\ ES: 87 \%} & \makecell{\textbf{ 0.35 } $\pm$ 0.026 \\Variance: 0.2 \\ ES: 77 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.023 \\Variance: 0.1 \\ ES: 65 \%} \\
+TS & \makecell{\textbf{ 0.0029 } $\pm$ 0.003 \\Var: 0.002 \\ ES: 100 \%} & \makecell{\textbf{ 0.11 } $\pm$ 0.02 \\Var: 0.09 \\ ES: 96 \%} & \makecell{\textbf{ 0.17 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 95 \%} \\
+  DEG & \makecell{\textbf{ 0.049 } $\pm$ 0.01 \\Var: 0.03 \\ ES: 94 \%} & \makecell{\textbf{ 0.24 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 74 \%} & \makecell{\textbf{ 0.24 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 79 \%} \\
+   DG & \makecell{\textbf{ 0.12 } $\pm$ 0.02 \\Var: 0.08 \\ ES: 87 \%} & \makecell{\textbf{ 0.35 } $\pm$ 0.03 \\Var: 0.2 \\ ES: 77 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 65 \%} \\
    \hline
 \end{tabular}
-\caption*{\tiny{TODO: fix table names and general table alignment.
-The first line in each cell contains the average market share for the entrant over $N=250$ simulations as well as a 95\% confidence interval.
+\caption*{\tiny{The columns represent the strategy of the incumbent and the rows represent the strategy of the entrant.
+The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval.
 The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one principal getting $> 90\%$ of the market.
@@
 \end{table}
-
-Table \ref{ht_incum} shows the results of the simulations.
-We observe that, for the incumbent, $TS$ is a dominant strategy.
-In the appendix we include additional simulations where $X = 50$ and observe that $TS$ is not always a dominant strategy in this case.
-Qualitatively, these results are robust across the different priors we considered as well as across agent response models.
-Namely, for sufficiently large $X$, $TS$ is a dominant strategy for the incumbent.
-
-Why do we observe this reversal?
-The intuition for this is that competition forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
+Why do we observe that $TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, for this warm start, $DG$ was preferred to $TS$?
+The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
 This intuition is very similar to that observed in the previous section by increasing warm start.
@@
 As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
-Thus, counterintuitively, by allowing one firm to be a monopoly and dominate the market, we can incentivize them to play $TS$ and thus get higher social welfare than under competition.
-[add some more economic intuition here]
+Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $TS$.
 
-[Include some discussion about the entrant]
+What about for the entrant?
+We do not find that $TS$ is a dominant strategy, though $TS$ is not dominated by $DG$.
+Since the game we consider is a zero-sum game, there exists equililibria where $TS$ is played with some positive probability in equilibrium.
 
-\section{Data and Reputation \\ as Barriers to Entry}
-\label{S:8}
+\section{Data and Reputation as Barriers to Entry}
 
-As the simulation results of the incumbent experiment in Table \ref{ht_incum} and in the appendix show, the incumbent usually ends up with a substantial share of the market over the entrant.
-The rounds where the incumbent was a monopolist provided the incumbent with free data for learning (data advantage) as well as a chance to establish a reputational advantage over a potential entrant (reputational advantage).
+An alternative interpretation of the results in the previous section is that the ``temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant.
+This provides an alternative interpretation of the classic ``first mover advantage" that has been well-studied in economics and marketing \citep*{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry.
 Which plays a bigger role in preventing the entrant from being able to establish market share?
@@
 \centering
-\caption{Reputation Erased Experiment} 
-\begin{tabular}{rlll}
+\caption{Reputation Erased Experiment, Heavy Tail} 
+\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
   \hline
@@
   \hline
-TS & \makecell{ \textbf{ 0.18 } $\pm$ 0.042 \\Var:  0.1 \\ ES: 97 \% } & \makecell{ \textbf{ 0.17 } $\pm$ 0.042 \\Var:  0.1 \\ ES: 98 \% } & \makecell{ \textbf{ 0.23 } $\pm$ 0.047 \\Var:  0.2 \\ ES: 98 \% } \\
-  DEG & \makecell{ \textbf{ 0.23 } $\pm$ 0.045 \\Var:  0.2 \\ ES: 92 \% } & \makecell{ \textbf{ 0.32 } $\pm$ 0.049 \\Var:  0.2 \\ ES: 86 \% } & \makecell{ \textbf{ 0.25 } $\pm$ 0.046 \\Var:  0.2 \\ ES: 90 \% } \\
-   DG & \makecell{ \textbf{ 0.28 } $\pm$ 0.047 \\Var:  0.2 \\ ES: 87 \% } & \makecell{ \textbf{ 0.33 } $\pm$ 0.051 \\Var:  0.2 \\ ES: 91 \% } & \makecell{ \textbf{ 0.3 } $\pm$ 0.047 \\Var:  0.2 \\ ES: 83 \% } \\
+TS & \makecell{\textbf{ 0.016 } $\pm$ 0.0075 \\Var: 0.01 \\ ES: 100 \%} & \makecell{\textbf{ 0.13 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.2 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 96 \%} \\
+  DEG & \makecell{\textbf{ 0.068 } $\pm$ 0.013 \\Var: 0.05 \\ ES: 93 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 75 \%} & \makecell{\textbf{ 0.26 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 80 \%} \\
+   DG & \makecell{\textbf{ 0.15 } $\pm$ 0.019 \\Var: 0.1 \\ ES: 87 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.028 \\Var: 0.2 \\ ES: 80 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 67 \%} \\
    \hline
@@
 \centering
-\caption{Information Erased Experiment} 
-\begin{tabular}{rlll}
+\caption{Information Erased Experiment, Heavy Tail} 
+\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
   \hline
@@
   \hline
-TS & \makecell{ \textbf{ 0.059 } $\pm$ 0.025 \\Var:  0.05 \\ ES: 97 \% } & \makecell{ \textbf{ 0.11 } $\pm$ 0.033 \\Var:  0.09 \\ ES: 94 \% } & \makecell{ \textbf{ 0.15 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 92 \% } \\
-  DEG & \makecell{ \textbf{ 0.11 } $\pm$ 0.031 \\Var:  0.08 \\ ES: 89 \% } & \makecell{ \textbf{ 0.19 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 82 \% } & \makecell{ \textbf{ 0.18 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 82 \% } \\
-   DG & \makecell{ \textbf{ 0.17 } $\pm$ 0.038 \\Var:  0.1 \\ ES: 87 \% } & \makecell{ \textbf{ 0.22 } $\pm$ 0.042 \\Var:  0.1 \\ ES: 82 \% } & \makecell{ \textbf{ 0.26 } $\pm$ 0.043 \\Var:  0.1 \\ ES: 77 \% } \\
+TS & \makecell{\textbf{ 0.024 } $\pm$ 0.0094 \\Var: 0.02 \\ ES: 100 \%} & \makecell{\textbf{ 0.16 } $\pm$ 0.022 \\Var: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.22 } $\pm$ 0.025 \\Var: 0.2 \\ ES: 95 \%} \\
+  DEG & \makecell{\textbf{ 0.24 } $\pm$ 0.025 \\Var: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 72 \%} & \makecell{\textbf{ 0.27 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 76 \%} \\
+   DG & \makecell{\textbf{ 0.33 } $\pm$ 0.028 \\Var: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.026 \\Var: 0.2 \\ ES: 74 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.023 \\Var: 0.1 \\ ES: 58 \%} \\
    \hline
@@
 Tables \ref{rep_erase} and \ref{info_erase} display the results of these experiments.
-Even with the data advantage or reputation advantage alone the incumbent ends up getting a substantial portion of the market, regardless of the algorithm played.
-However, fixing the same algorithm played by the incumbent and the entrant, the market share for the incumbent is always higher when the incumbent retains her reputation instead of her information.
-The intuition for this is that the reputational advantage protects the incumbent from ``bad" decisions or bad luck on rewards that adversely affect reputation and thus, especially in the earlier rounds, agents in the earlier rounds will still choose the incumbent which will allow the incumbent to regain the information they lost.
-However, with reputation re-initialized the better information does not protect the incumbent from getting unlucky with her rewards.
-Looking at the variability in the shares in the two experiments indicates that not only does the incumbent get a smaller market share on average in the reputation erased experiment, but the variability in the shares is also higher.
+Maintaining the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
+Interestingly, when the incumbent plays $DEG$ or $DG$ reputation and information serve as substitutes (the market shares are almost identical), but when the incumbent plays $TS$ information serves as a larger barrier to entry than reputation.
+This is robust across the priors that we consider for this parametrization.
+One possible intuition for this is that, for Heavy Tail, the information acquired after the $X$ rounds under $TS$ is sufficient so that it becomes relatively easier for $TS$ to recover the reputation loss compared to the information loss, but since $DEG$ and $DG$ learn slower and they've acquired less information over the $X$ rounds relative to $TS$ there is less of an informational advantage.
 
-Though the setup is purely experimental, it is nonetheless interesting to look at if $TS$ still remains as a dominant strategy for the incumbent.
-In the information erased experiment, $TS$ is still a dominant strategy while in the reputation erased experiment it is not a dominant strategy.
-The intuition for this is that in the reputation erased experiment, the incumbent still has the data that she accumulated during her time as a monopolist and as a result, there is less value to an algorithm that engages in pure exploration as she wants to rebuild her reputation and thus might be better off exploiting the information she has in order to rebuild reputation instead of engaging in exploration and potentially damage her reputation.
-However, in the information erased experiment, she has a reputational cushion that allows for exploration so that she can regain the information that she lost and thus $TS$ remains a dominant strategy.
-
-In the appendix we include the results of the same experiment but with different agent response models. In these experiments we get that information and reputation serve as substitutes for each other
-
-[TODO: leave this, expound on this, or throw it out?]
-
-Thus, we see that both information and reputation advantages alone can serve as substantial barriers to entry in our model.
-However, under the HardMax agent response model, we see that reputation serves as more of a barrier to entry compared to information.
-
-\section{Varying the agent response model}
-\label{S:9}
-
-Thus far we have largely presented simulation results utilizing the deterministic HardMax rule.
-This is a natural decision rule where agents simply use the reputation score as a strict decision-making rule.
-In this section we consider the implications of adding randomness to the agent rule.
-What are the implications on what algorithm the firms are incentivized to play?
-How does the distribution of market share change?
-
-Under the HardMax decision rule, firms need to incentivize consumers to select them over their competition by having a higher reputation.
-First, we consider the $HardMaxWithRandom$ decision rule discussed in the Section \ref{S:3}.
-Similar to the result in \citet{mansour2017competing} we find that this induces firms to play the ``better" algorithm for sufficiently long time horizon (see appendix).
-The intuition for this is the same as before, where, even though the firm initially suffers reputation loss from exploration, eventually the firm gets sufficiently many agents that don't need to be incentivized to select the firm in order to get a higher reputation than a ``worse" algorithm.
-We have a similar observation for the $SoftMax$ decision rule discussed in Section \ref{S:3}.
-
-\begin{figure}
-\label{agent_variances}
-\end{figure}
-
-We see that better algorithms are incentivized when we consider random agent response rules, but there are differences in the overall market composition.
-As Figure \ref{agent_variances} shows, the variance of market shares across simulations varies greatly.
-Under $HardMax$, we see that, in a single simulation, the market is generally ``winner take all" where one firm generally gets a large share of the market and there is a high variance.
-Under $HardMaxWithRandom$ we see that, in a single simulation, the market is mostly taken by one firm (but to a lower extent than under $HardMax$) and there is a lower, though still high, variance.
-Under $SoftMax$ we see that the market variance is very low and market shares are generally close to 50/50.
+Though the setup is purely experimental, it is nonetheless interesting to look at if $TS$ still remains a dominant strategy for the incumbent.
+In the reputation erased experiment, $TS$ is still a dominant strategy while in the information erased experiment it is not a dominant strategy.
 

### 2018-08-16 paper/competing_bandits.tex
latest draft

@@
 Our model considers competition for myopic users between two firms deploying multi-armed bandit algorithms that face the same underlying multi-armed bandit instance.
-The agents select between the firms according to a reputation score, which is a function of the rewards past agents have experienced from this firm.
+The users select between the firms according to a reputation score, which is a function of the rewards past users have experienced from this firm.
 
 We ask whether better algorithms are incentivized under varying degrees of competition.
-The environments we consider are a monopoly, a duopoly with one firm serving as an early entrant (or a temporary monopoly), and a duopoly.
+The environments we consider are a monopoly, a duopoly with one firm serving as an early entrant (a temporary monopoly), and a duopoly.
 We show that, under certain conditions, monopoly and duopoly do not incentivize the adoption of better learning algorithms due to the reputational costs of exploration, but that a temporary monopoly may incentivize the adoption of better learning algorithms.
+Additionally, we interpret our results as providing an alternative intuition behind the classic first-mover advantage which gives the incumbent firm both a data advantage and a reputational advantange.
 Finally, we ask whether in this setting the data advantage or the reputational advantage of the early entrant serves as stronger barriers to entry.
@@
 
-From a social welfare perspective it is better for firms to adopt ``smarter" learning algorithms and our main high-level question is asking under what levels of competition are firms incentivized to adopt ``better" learning algorithms.
+From a social welfare perspective it is better for firms to adopt ``better" learning algorithms and our main high-level question is asking under what levels of competition are firms incentivized to adopt ``better" learning algorithms.
 Our primary contribution is that in our model, even in the case of a duopoly, firms are not incentivized to adopt better learning algorithms but that allowing one firm to be a temporary monopolist will incentivize this firm to play a better learning algorithm.
@@
 
-The most closely related work to this paper is \citealt*{mansour2017competing}.
-Our motivating questions are the same as \citet{mansour2017competing} though while \citet{mansour2017competing} use the rationality of consumers as their primary knob of competitiveness, we consider differences in the number of firms in the market as our primary knob.
-Additionally, while in \citet{mansour2017competing} agents select firms based on the Bayesian expected reward of each principal in our model the agents select firms based on a frequentist reputation score that is derived from the signals sent from past agents.
+The most closely related work to this paper is \citealt*{mansour2018competing}.
+Our motivating questions are the same as \citet{mansour2018competing} though while \citet{mansour2018competing} use the rationality of consumers as their primary knob of competitiveness, we consider differences in the number of firms in the market as our primary knob.
+Additionally, while in \citet{mansour2018competing} agents select firms based on which firm they expect to have a larger Bayesian expected reward, in our model the agents select firms based on a frequentist reputation score that is derived from the signals sent from past agents.
 This allows the model to become tractable for the purposes of evaluating it via simulation and this is the primary method of analysis that we employ.
@@
 
-\textbf{Overview} There are two principals and $T+2k$ agents where $k$ is the warm start, or the agents that each principal gets for free at the beginning of the game and $T$ is the number of rounds in the game.
+\textbf{Overview} There are two firms and $T+2k$ agents where $k$ is the warm start, or the agents that each firm gets for free at the beginning of the game and $T$ is the number of rounds in the game.
 The timing of events is as follows:
 \begin{enumerate}
-\item At $t = 0$, the principals simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
-\item Still at $t=0$ we suppose that each principal gets $k$ agents as a ``warm" start.
-\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses among 2 principals, given a reputation score for both principals.
-\item The principal that is chosen selects an action $a_{t} \in A$ from a set of fixed actions across principals and rounds.
-\item Both the agent and the principal observe the reward $r_t \in [0, 1]$ from the action.
-The agent reports this reward to the principal and the future agents and the reputation score for the chosen principal is updated.
+\item At $t = 0$, the firms simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
+\item Still at $t=0$ we suppose that each firm gets $k$ agents as a ``warm" start.
+\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses among 2 firms, given a reputation score for both firms.
+\item The firm that is chosen selects an action $a_{t} \in A$ from a set of actions that is fixed across firms and rounds.
+\item Both the agent and the firm observe the reward $r_t \in [0, 1]$ from the action.
+The agent reports this reward to the firm and the future agents and the reputation score for the chosen firm is updated.
 \item Repeat 2-4 for $T$ rounds.
@@
 For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors.
-Each principal faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a ``fake" prior of $Beta(\alpha=1,\beta=1)$}.
-We assume that the principals commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.
+Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a ``fake" prior of $Beta(\alpha=1,\beta=1)$}.
+We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.
 
@@
 The utility function for the agents is simply to maximize their reward in the one period in which they are alive.
-We suppose that agents do not attempt to manipulate the strategy of the principal nor do they take the strategies of the principal into account when choosing between the principals.
+We suppose that agents do not attempt to manipulate the strategy of the firm nor do they take the strategies of the firm into account when choosing between the firms.
 In our model, each agent uses the average reward of past agents as a proxy for their expected utility.
-For simplicity, the reputation score, $R_jt$ is defined as a sliding window average.
+For simplicity, the reputation score, $R_{jt}$ is defined as a sliding window average.
 \begin{center}
@@
 
-Note that the reputation score is in terms of \textit{local} and not global time so that it is the sliding window average of the last $M$ times that principal $j$ was selected by the agents.
+Note that the reputation score is in terms of \textit{local} and not global time so that it is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.
 The warm start of $k$ rounds allows this reputation score to be well-defined once the ``competition game" begins.
-In our model the agents deterministically choose the principal with the higher reputation score and ties are broken uniformly at random.
+In our model the agents deterministically choose the firm with the higher reputation score and ties are broken uniformly at random.
 
-\noindent \textbf{Principals} We suppose that the principals simply care about maximizing their expected market share (i.e.
+\noindent \textbf{firms} We suppose that the firms simply care about maximizing their expected market share (i.e.
 maximizing the number of $T$ agents who select them).
-We model the ``competition" game between principals as a simultaneous move game where both principals commit to a learning algorithm.
+We model the ``competition game" between firms as a simultaneous move game where both firms commit to a learning algorithm.
 
-\noindent \textbf{MAB algorithms} We suppose that principals commit to a learning algorithm from $\mathcal{A}$.
+\noindent \textbf{MAB algorithms} We suppose that firms commit to a learning algorithm from a fixed set of algorithms $\mathcal{A}$.
 We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
@@
 
-It is known that $TS > DEG > DG$ in the standard multi-armed bandit problem of minimizing total regret.
-The primary question that we want to understand is when, in competition, are the principals incentivized to adopt the ``better" algorithms?
+In the standard multi-armed bandit problem it is known that $TS > DEG > DG$ in terms of maximizing cumulative reward over a sufficiently large time horizon.
+The primary question that we want to understand is when, in competition, are the firms incentivized to adopt the ``better" algorithms?
 
-\noindent \textbf{Incumbent} In some of our experiments we modify our model so that one principal enters the market $X$ number of rounds before the other.
-We refer to the principal that enters before as the ``incumbent" and the principal that enters after $X$ rounds as the ``entrant." The agents that arrive in these $X$ rounds are forced to select the only firm in the market as there is no outside option.
+\noindent \textbf{Incumbent} In some of our experiments we modify our model so that one firm enters the market $X$ number of rounds before the other.
+We refer to the firm that enters before as the ``incumbent" and the firm that enters after $X$ rounds as the ``entrant." The agents that arrive in these $X$ rounds are forced to select the only firm in the market as there is no outside option.
 We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$.
@@
 There are two reasons why this is important to consider.
-First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where $TS > DEG > DG$.
+First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where $TS > DEG > DG$ for sufficiently large $t$.
 Second, we want to better understand what statistics of the instance we can look at in isolation in order to help us predict and understand what to expect in the competition game.
@@
 \caption{Mean Reputation Trajectories in Isolation}
-\includegraphics[scale=0.25]{"figures/nih_iso_mean"}
+\includegraphics[scale=0.35]{"figures/nih_iso_mean"}
 \label{prelim_means}
@@
 
-In the ``competition" game the agents decide between the principals using the \textit{relative} reputation between the two principals.
-Namely, if principal $A$ has a higher reputation than principal $B$ agents will select principal $A$.
+In the ``competition game" the agents decide between the firms using the \textit{relative} reputation between the two firms.
+Namely, if firm $A$ has a higher reputation than firm $B$ agents will select firm $A$.
 A natural question to ask is whether looking at the mean performance is sufficient for understanding performance in the competition game.
@@
 However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$.
-In these cases its reputation may be substantially worse but for the competition game the difference between the reputation does not matter but only the binary comparison between them \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, considering the SoftMax decision rule in \citet{mansour2017competing}}.
+In these cases its reputation may be substantially worse but for the competition game the difference between the reputation does not matter but only the binary comparison between them \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, considering the SoftMax decision rule in \citet{mansour2018competing}}.
 
@@
 \caption{Reputation Distribution}
-\includegraphics[scale=0.3]{figures/rep_distribution_nih}
+\includegraphics[scale=0.35]{figures/rep_distribution_nih}
 \label{rep_dist_nih}
@@
 Figure \ref{ts_dg_rep_diff_nih} shows the distribution of reputation difference between $TS$ and $DG$ across $t$.
-This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right.
+This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large.
 Thus, the mean is not a representative statistic of the entire reputation difference distribution.
-As an alternative statistic for understanding the results of the competition game we introduce the \textit{relative reputation statistic} which looks at the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$.
-This corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a principal playing $A$ over a principal playing $B$ at time $t$ \footnote{Though we do not discuss it here, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion.
+As an alternative statistic for understanding the results of the competition game we introduce the \textit{relative reputation proportion} which looks at the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$.
+This corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{Though we do not discuss it here, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion.
 In the GitHub appendix we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.}.
@@
 \caption{Reputation Difference Distribution}
-\includegraphics[scale=0.3]{figures/ts_dg_rep_diff_nih}
+\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
 \label{ts_dg_rep_diff_nih}
@@
 
-Figure \ref{relative_rep_plots} shows the relative reputation plots for $TS$ vs $DG$ on the Uniform and Needle In Haystack prior.
+Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $TS$ vs $DG$ on the Uniform and Needle In Haystack prior.
 For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $TS > DG$.
-The intuition behind this is that, especially since the principals start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information.
+The intuition behind this is that, especially since the firms start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information.
 However, this exploration leads to what we define as a \textit{relative reputation cost}.
@@
 In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \citet{aghion2005competition}.
-We consider three separate settings for our model: permanent monopoly, temporary monopoly, and permanent duopoly.
+We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly.
 In the permanent monopoly and permanent duopoly cases we show that, under certain conditions, firms are incentivized to commit to $DG$ whereas in the temporary monopoly case firms are incentivized to commit to $TS$.
@@
 \item Under Heavy Tail we have that, for low warm start, $DEG, DG > TS$ but that for the higher warm start values we consider $TS > DEG, DG$.
-Interestingly, even under the relatively high warm start we see $DEG > DG$.
+Interestingly, even under the relatively high warm start we see $DG > DEG$.
 \item Under Uniform we have that $DEG, DG > TS$ for low warm start and $TS > DEG, DG$ for sufficiently high warm start.
@@
 Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start value as fixing the starting point on the relative reputation plots.
-The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation statistic of $A \geq B$ at time $k$.
+The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$.
 Combined with the fact that the EEOG values being relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right, similar to the distribution of reputation differences}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.
 
-For the low values of the warm start we can see that $DG > TS$ for the Uniform prior but $TS > DG$ for the Needle In Haystack prior and this corresponds to what we see in the results of the competition game simulation.
+For the low values of the warm start we can see that $DG > TS$ for the Uniform prior but $TS > DG$ for the Needle In Haystack prior and this corresponds to what we see in the relative reputation proportion plots.
 The intuition for this result aligns with the intuition given in the discussion of the relative reputation plots.
@@
 \centering
-\caption{Simultaneous Entry Needle In Haystack} 
+\caption{Duopoly Experiment Needle In Haystack} 
 \begin{tabular}{rlll}
   \hline
- & k = 20 & k = 100 & k = 200 \\ 
+ & k = 20 & k = 250 & k = 500 \\ 
   \hline
-TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 210\\ med: 28} & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 260\\ med: 68} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 320\\ med: 0} \\
-  TS vs DEG & \makecell{\textbf{0.56} $\pm$0.03\\ eeog \\ avg: 150\\ med: 11} & \makecell{\textbf{0.54} $\pm$0.03\\ eeog \\ avg: 240\\ med: 24.5} & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 420\\ med: 36.5} \\
-  DG vs DEG & \makecell{\textbf{0.43} $\pm$0.03\\ eeog \\ avg: 330\\ med: 97.5} & \makecell{\textbf{0.42} $\pm$0.03\\ eeog \\ avg: 550\\ med: 229.5} & \makecell{\textbf{0.41} $\pm$0.02\\ eeog \\ avg: 650\\ med: 420} \\
+TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 200\\ med: 27} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 370\\ med: 0} & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 580\\ med: 121.5} \\
+  TS vs DEG & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 150\\ med: 14} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 460\\ med: 78.5} & \makecell{\textbf{0.56} $\pm$0.02\\ eeog \\ avg: 740\\ med: 627.5} \\
+  DG vs DEG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 340\\ med: 128.5} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 650\\ med: 408} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 690\\ med: 466.5} \\
    \hline
@@
 % latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Wed Aug 15 15:06:22 2018
+% Thu Aug 16 13:13:00 2018
 \begin{table}[ht]
 \centering
-\caption{Simultaneous Entry Heavy Tail} 
+\caption{Duopoly Experiment Heavy Tail} 
 \begin{tabular}{rlll}
   \hline
- & k = 20 & k = 100 & k = 200 \\ 
+ & k = 20 & k = 250 & k = 500 \\ 
   \hline
-TS vs DG & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 65\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 130\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 500\\ med: 0} \\
-  TS vs DEG & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 26\\ med: 0} & \makecell{\textbf{0.59} $\pm$0.03\\ eeog \\ avg: 160\\ med: 0} & \makecell{\textbf{0.86} $\pm$0.02\\ eeog \\ avg: 420\\ med: 0} \\
-  DG vs DEG & \makecell{\textbf{0.63} $\pm$0.03\\ eeog \\ avg: 390\\ med: 10} & \makecell{\textbf{0.62} $\pm$0.02\\ eeog \\ avg: 670\\ med: 500.5} & \makecell{\textbf{0.61} $\pm$0.02\\ eeog \\ avg: 770\\ med: 670.5} \\
+TS vs DG & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 55\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 570\\ med: 0} & \makecell{\textbf{0.76} $\pm$0.02\\ eeog \\ avg: 620\\ med: 98.5} \\
+  TS vs DEG & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 37\\ med: 0} & \makecell{\textbf{0.88} $\pm$0.01\\ eeog \\ avg: 480\\ med: 0} & \makecell{\textbf{0.9} $\pm$0.01\\ eeog \\ avg: 570\\ med: 113.5} \\
+  DG vs DEG & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 410\\ med: 7} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 790\\ med: 762} & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 730\\ med: 608} \\
    \hline
@@
 % latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Wed Aug 15 15:06:22 2018
-\begin{table}
+% Thu Aug 16 13:13:01 2018
+\begin{table}[ht]
 \centering
-\caption{Simultaneous Entry Uniform} 
+\caption{Duopoly Experiment Uniform} 
 \begin{tabular}{rlll}
   \hline
- & k = 20 & k = 100 & k = 200 \\ 
+ & k = 20 & k = 250 & k = 500 \\ 
   \hline
-TS vs DG & \makecell{\textbf{0.43} $\pm$0.03\\ eeog \\ avg: 260\\ med: 0} & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 560\\ med: 290} & \makecell{\textbf{0.48} $\pm$0.03\\ eeog \\ avg: 770\\ med: 706.5} \\
-  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.37} $\pm$0.03\\ eeog \\ avg: 510\\ med: 222} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 810\\ med: 785.5} \\
-  DG vs DEG & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 530\\ med: 163.5} & \makecell{\textbf{0.5} $\pm$0.02\\ eeog \\ avg: 890\\ med: 922.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 990\\ med: 1071} \\
+TS vs DG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.02\\ eeog \\ avg: 800\\ med: 754} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 910\\ med: 906.5} \\
+  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 180\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.02\\ eeog \\ avg: 810\\ med: 734} & \makecell{\textbf{0.55} $\pm$0.02\\ eeog \\ avg: 970\\ med: 987} \\
+  DG vs DEG & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 470\\ med: 57.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1088} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1142} \\
    \hline
@@
 The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval.
-The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one principal getting $> 90\%$ of the market.
+The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one firm getting $> 90\%$ of the market.
 Note that smaller values in the table are better for the incumbent.

### 2018-08-16 analysis_code/tournament_table.R

@@ -13,7 +13,7 @@ agent_algs - c("HardMax")
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
 time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior
-priors <- c("Needle In Haystack", "Heavy Tail")
+priors <- c("Heavy Tail")
 warm_start <- 20
 
 ZERO_ONE_CUTOFF <- 0.1
@@ -42,16 +42,16 @@ for (j in 1:length(time_horizons)) {
           }
           share <- filtered_dat$Market.Share.for.P1
           test <- t.test(share)
-          cin <- signif(test$conf.int[2] - mean(share), digits=2)
+          cin <- signif(test$conf.int[2] - mean(share), digits=1)
           shares <- sum(share >= (1 - ZERO_ONE_CUTOFF) | share <= ZERO_ONE_CUTOFF, na.rm=TRUE) / nrow(filtered_dat)
           cv_high <- qchisq(.975, df=length(share) - 1)
           cv_low <- qchisq(.025, df=length(share) - 1)
           lower_var <- signif((length(share) - 1) * var(share) / cv_high, digits=2)
           high_var <- signif((length(share) - 1) * var(share) / cv_low, digits=2)
-          results[k, l] <- paste("\\makecell{\\textbf{", signif(mean(share), digits=2), "} $\\pm$", cin, "\\\\Variance:", signif(var(share), digits=1), "\\\\", "ES:", 100*signif(shares, digits=2), "\\%}")
+          results[k, l] <- paste("\\makecell{\\textbf{", signif(mean(share), digits=2), "} $\\pm$", cin, "\\\\Var:", signif(var(share), digits=1), "\\\\", "ES:", 100*signif(shares, digits=2), "\\%}")
         }
       }
-      tab <-xtable(results, caption=paste("Information Erased Experiment", prior, "X=200"))
+      tab <-xtable(results, caption=paste("Temporary Monopoly Experiment", prior))
       print(tab, type="latex")
     }
   }

### 2018-08-16 paper/competing_bandits.tex
cosmetics

@@
 \includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
-\caption*{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$.
-The shaded area display 95\% confidence intervals.}
+\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$.
+The shaded area display 95\% confidence intervals.}}
 \label{relative_rep_plots}

### 2018-08-16 paper/competing_bandits.tex
more cosmetic changes
whoops

@@
 Many modern online platforms simultaneously compete for users as well as learn from the users they manage to attract.
-For instance, Google Search and Bing compete for users in the search engine market yet at the same need to experiment with their search algorithms to learn what algorithms work the best.
+For instance, Google Search and Bing compete for users in the search engine market yet at the same time need to experiment with their search algorithms to learn what algorithms work the best.
 This creates a tradeoff between \textit{exploration} and \textit{competition} since firms need to experiment with potentially sub-optimal options for the benefit of gaining information to make better decisions tomorrow while at the same time firms need to incentivize consumers to select them over their competitors today.
@@
 
-\noindent \textbf{firms} We suppose that the firms simply care about maximizing their expected market share (i.e.
+\noindent \textbf{Firms} We suppose that the firms simply care about maximizing their expected market share (i.e.
 maximizing the number of $T$ agents who select them).

### 2018-08-21 paper/competing_bandits.tex
updated paper

@@
 \author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
-Zhiwei Steven Wu\textsuperscript{4} \\,
+Zhiwei Steven Wu\textsuperscript{4} \\
 {\textsuperscript{1}Columbia University, Department of Economics}\\
@@
 \section{Introduction}
-\label{intro}
+\label{S:1}
 
@@
 
-\section{Model}
-\label{model}
+\section{Model}\label{S:2}
 
@@
 \item At $t = 0$, the firms simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
-\item Still at $t=0$ we suppose that each firm gets $k$ agents as a ``warm" start.
+\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a ``warm" start.
+The algorithm that the firm commits to makes $k$ rounds of progress and uses this information to initialize its information set.
+Additionally, the reputation score of the agents is initialized using the rewards from these $k$ rounds.
 \item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses among 2 firms, given a reputation score for both firms.
@@
 
-Note that the reputation score is in terms of \textit{local} and not global time so that it is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.
+Note $t_j$ is the \textit{local} time of the firm, not the global time, so the reputation score is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.
 The warm start of $k$ rounds allows this reputation score to be well-defined once the ``competition game" begins.
@@
 maximizing the number of $T$ agents who select them).
-We model the ``competition game" between firms as a simultaneous move game where both firms commit to a learning algorithm.
+We model the ``competition game" between firms as a simultaneous move game where both firms commit to a learning algorithm at $t = 0$.
 
@@
 \noindent \textbf{Incumbent} In some of our experiments we modify our model so that one firm enters the market $X$ number of rounds before the other.
-We refer to the firm that enters before as the ``incumbent" and the firm that enters after $X$ rounds as the ``entrant." The agents that arrive in these $X$ rounds are forced to select the only firm in the market as there is no outside option.
+We refer to the firm that enters before as the ``incumbent" and the firm that enters after $X$ rounds as the ``entrant." In the $X$ rounds before the ``entrant" enters the incumbent is a temporary monopolist as the agents that arrive in these $X$ rounds are forced to select the incumbent since it is the only firm in the market and there is no outside option.
 We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$.
@@
 
-\section{Simulation Details}
-\label{sim_details}
+\section{Simulation Details}\label{S:3}
 
@@
 \section{Performance in Isolation}
-\label{iso_section}
+\label{S:4}
 
@@
 This corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{Though we do not discuss it here, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion.
-In the GitHub appendix we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.}.
+In the GitHub appendix we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.
+Additionally, these results carry over to the competition game}.
 
@@
 \section{Competition vs Better Algorithms: Inverted-U}
+\label{S:5}
 
@@
 
-We utilize the same instances and realizations as in the Section \ref{iso_section} and simulate the model described in section \ref{model}.
-We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the payoffs associated with each pair of algorithms.
+We utilize the same instances and realizations as in the section~\ref{S:3} and simulate the model described in section~\ref{S:2}.
+We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the expected payoffs associated with each pair of algorithms.
 Unless otherwise noted, all the results are reported at $t = 2000$.
@@
 
-\textbf{Permanent Duopoly} Permanent duopoly corresponds to the model described in section \ref{model} where $X = 0$ so that both firms enter the market simultaneously.
+\textbf{Permanent Duopoly} Permanent duopoly corresponds to the model described in section~\ref{S:2} where $X = 0$ so that both firms enter the market simultaneously.
 What algorithms are firms incentivized to deploy?
@@
 \section{Data and Reputation as Barriers to Entry}
+\label{S:6}
 

### 2018-08-21 paper/competing_bandits.tex
reorg

@@
 
-\section{Introduction}
-\label{S:1}
+\section{Introduction}\label{section:1}
 
@@
 
-From a social welfare perspective it is better for firms to adopt ``better" learning algorithms and our main high-level question is asking under what levels of competition are firms incentivized to adopt ``better" learning algorithms.
+From a social welfare perspective it is better for firms to adopt "better" learning algorithms and our main high-level question is asking under what levels of competition are firms incentivized to adopt "better" learning algorithms.
 Our primary contribution is that in our model, even in the case of a duopoly, firms are not incentivized to adopt better learning algorithms but that allowing one firm to be a temporary monopolist will incentivize this firm to play a better learning algorithm.
@@
 MAB problems have been studied for many decades (see \citet{bubeck2012regret} for an overview).
-Most relevant to this paper is the thread that focuses on designing ``smart" and tractable algorithms that combine exploration and exploitation and ``naive" algorithms that separate exploration and exploitation (see \citet{slivkins2017bandits}).
+Most relevant to this paper is the thread that focuses on designing "smart" and tractable algorithms that combine exploration and exploitation and "naive" algorithms that separate exploration and exploitation (see \citet{slivkins2017bandits}).
 
@@
 
-Our setting is also closely related to the ``dueling algorithms" framework described in \cite{immorlica2011dueling}, but this framework considers offline scenarios whereas we focus on online learning problems.
+Our setting is also closely related to the "dueling algorithms" framework described in \cite{immorlica2011dueling}, but this framework considers offline scenarios whereas we focus on online learning problems.
 
@@
 
-\section{Model}\label{S:2}
+\section{Model}\label{section:2}
 
@@
 \item At $t = 0$, the firms simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
-\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a ``warm" start.
+\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a "warm" start.
 The algorithm that the firm commits to makes $k$ rounds of progress and uses this information to initialize its information set.
@@
 
-Generally, the rewards are i.i.d.
-with a common prior.
+Generally, the rewards are independent and identically distributed with a common prior.
 For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors.
-Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a ``fake" prior of $Beta(\alpha=1,\beta=1)$}.
+Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a "fake" prior of $Beta(\alpha=1,\beta=1)$}.
 We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.
@@
 Note $t_j$ is the \textit{local} time of the firm, not the global time, so the reputation score is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.
-The warm start of $k$ rounds allows this reputation score to be well-defined once the ``competition game" begins.
+The warm start of $k$ rounds allows this reputation score to be well-defined once the "competition game" begins.
 In our model the agents deterministically choose the firm with the higher reputation score and ties are broken uniformly at random.
@@
 maximizing the number of $T$ agents who select them).
-We model the ``competition game" between firms as a simultaneous move game where both firms commit to a learning algorithm at $t = 0$.
+We model the "competition game" between firms as a simultaneous move game where both firms commit to a learning algorithm at $t = 0$.
 
@@
 \begin{enumerate}
-\item ``Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
-We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is ``optimal" in the sense of having the highest mean reward \citep{agrawal2012analysis}.
-\item ``Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
+\item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
+We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \citep{agrawal2012analysis}.
+\item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
 We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability.
@@
 In the standard multi-armed bandit problem it is known that $TS > DEG > DG$ in terms of maximizing cumulative reward over a sufficiently large time horizon.
-The primary question that we want to understand is when, in competition, are the firms incentivized to adopt the ``better" algorithms?
+The primary question that we want to understand is when, in competition, are the firms incentivized to adopt the "better" algorithms?
 
 \noindent \textbf{Incumbent} In some of our experiments we modify our model so that one firm enters the market $X$ number of rounds before the other.
-We refer to the firm that enters before as the ``incumbent" and the firm that enters after $X$ rounds as the ``entrant." In the $X$ rounds before the ``entrant" enters the incumbent is a temporary monopolist as the agents that arrive in these $X$ rounds are forced to select the incumbent since it is the only firm in the market and there is no outside option.
+We refer to the firm that enters before as the "incumbent" and the firm that enters after $X$ rounds as the "entrant." In the $X$ rounds before the "entrant" enters the incumbent is a temporary monopolist as the agents that arrive in these $X$ rounds are forced to select the incumbent since it is the only firm in the market and there is no outside option.
 We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$.
@@
 
-\section{Simulation Details}\label{S:3}
+\section{Simulation Details}\label{section:3}
 
@@
 
-\textbf{Bandit Priors} We look at bandit instances drawn from three different bandit ``priors" that are different types of learning problems.
+\textbf{Bandit Priors} We look at bandit instances drawn from three different bandit "priors" that are different types of learning problems.
 Recall that we only consider Bernoulli-distributed arms and so we draw the true means of the arms from these priors.
@@
 \item "Heavy Tail" - the $K$ arms have means drawn from $Beta(\alpha=0.6, \beta = 0.6)$.
-With this prior it was likely to have means at the ``extremes" or means that were close to 0 as well as means that were close to 1.
+With this prior it was likely to have means at the "extremes" or means that were close to 0 as well as means that were close to 1.
 \end{enumerate}
@@
 
-\section{Performance in Isolation}
-\label{S:4}
+\section{Performance in Isolation}\label{section:4}
 
@@
 
-In the ``competition game" the agents decide between the firms using the \textit{relative} reputation between the two firms.
+In the "competition game" the agents decide between the firms using the \textit{relative} reputation between the two firms.
 Namely, if firm $A$ has a higher reputation than firm $B$ agents will select firm $A$.
@@
 Figure \ref{rep_dist_nih} displays the reputation distribution at $t = 500$ for the same prior we reported the mean plot for previously.
-We see that the ``naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not.
+We see that the "naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not.
 The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not.
-If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds so that the weak law of large numbers comes into play (our $M$ is set sufficiently high to ensure this most of the time).
+If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds.
 However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$.
@@
 Namely, exploration leads to lower reputation relative to the myopic alternative.
-However, eventually the information acquired in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm.
+However, eventually the information acquired in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is "hard enough" so that $DG$ cannot trivially find the best arm.
 We define this as the \textit{relative reputation benefit} \footnote{We see similar behavior for the Heavy Tail prior}.
@@
 First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration.
-Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
+Second, in the Needle In Haystack prior there are no "bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
 Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
@@
 
-\section{Competition vs Better Algorithms: Inverted-U}
-\label{S:5}
+\section{Competition vs Better Algorithms: Inverted-U}\label{section:5}
 
@@
 We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index).
-In our model we have no prices and no costs and so we interpret increases in ``market power" as decreases in the number of rounds where a firm has to compete for agents.
+In our model we have no prices and no costs and so we interpret increases in "market power" as decreases in the number of rounds where a firm has to compete for agents.
 
-We utilize the same instances and realizations as in the section~\ref{S:3} and simulate the model described in section~\ref{S:2}.
+We utilize the same instances and realizations as in the section 4 and simulate the model described in section 2.
 We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the expected payoffs associated with each pair of algorithms.
@@
 Since, in our formulation, the firm only gets utility from a larger market share it is indifferent between each algorithm in $\mathcal{A}$ since, regardless of the algorithm it deploys, it will get the entire market.
-If we suppose that deploying the ``better" algorithms has even an $\epsilon > 0$ cost, then the firm would choose to deploy $DG$.
+If we suppose that deploying the "better" algorithms has even an $\epsilon > 0$ cost, then the firm would choose to deploy $DG$.
 
-\textbf{Permanent Duopoly} Permanent duopoly corresponds to the model described in section~\ref{S:2} where $X = 0$ so that both firms enter the market simultaneously.
+\textbf{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously.
 What algorithms are firms incentivized to deploy?
@@
 
-Focusing on reasonable warm starts, $DG > TS$ when it appears that better algorithms suffer an early relative reputational cost due to exploration and that both firms enter at the same time so that there is a ``permanent duopoly."
+Focusing on reasonable warm starts, $DG > TS$ when it appears that better algorithms suffer an early relative reputational cost due to exploration and that both firms enter at the same time so that there is a "permanent duopoly."
 
@@
 This intuition is very similar to that observed in the previous section by increasing warm start.
-In some sense one can view allowing one firm to temporarily be a monopolist as temporarily relaxing the ``incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
+In some sense one can view allowing one firm to temporarily be a monopolist as temporarily relaxing the "incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
 The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.
@@
 
-What about for the entrant?
-We do not find that $TS$ is a dominant strategy, though $TS$ is not dominated by $DG$.
-Since the game we consider is a zero-sum game, there exists equililibria where $TS$ is played with some positive probability in equilibrium.
+What about for the entrant? The results are not consistent across priors and are summarized as follows:
+\begin{enumerate}
+\item Needle In Haystack - $TS > DG, DEG$
+\item Heavy Tail - $DG > TS, DEG$
+\item Uniform - $DG \sim DEG > TS$
+\end{enumerate}
+The raw results can be found in the GitHub appendix and though they are not consistent, it is interesting to note that on the priors where we see the relative reputation cost for $TS$ we have that $DG$ is preferred to $TS$ and for the priors where we see that exploration is not costly, we have that $TS$ is preferred to $DG$.
 
-\section{Data and Reputation as Barriers to Entry}
-\label{S:6}
+\section{Data and Reputation as Barriers to Entry}\label{section:6}
 
-An alternative interpretation of the results in the previous section is that the ``temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant.
-This provides an alternative interpretation of the classic ``first mover advantage" that has been well-studied in economics and marketing \citep*{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry.
+An alternative interpretation of the results in the previous section is that the "temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant.
+This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \citep*{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry.
 Which plays a bigger role in preventing the entrant from being able to establish market share?
@@
 Maintaining the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
-Interestingly, when the incumbent plays $DEG$ or $DG$ reputation and information serve as substitutes (the market shares are almost identical), but when the incumbent plays $TS$ information serves as a larger barrier to entry than reputation.
-This is robust across the priors that we consider for this parametrization.
+The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation.
 One possible intuition for this is that, for Heavy Tail, the information acquired after the $X$ rounds under $TS$ is sufficient so that it becomes relatively easier for $TS$ to recover the reputation loss compared to the information loss, but since $DEG$ and $DG$ learn slower and they've acquired less information over the $X$ rounds relative to $TS$ there is less of an informational advantage.
 
-Though the setup is purely experimental, it is nonetheless interesting to look at if $TS$ still remains a dominant strategy for the incumbent.
-In the reputation erased experiment, $TS$ is still a dominant strategy while in the information erased experiment it is not a dominant strategy.
+Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
+We find that the best responses remain the same with the exception of the Uniform prior where $TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.
 

### 2018-08-21 paper/competing_bandits.tex
add footnote

@@
 In our model, each agent uses the average reward of past agents as a proxy for their expected utility.
-For simplicity, the reputation score, $R_{jt}$ is defined as a sliding window average.
+For simplicity, the reputation score, $R_{jt}$ is defined as a sliding window average\footnote{Another natural formulation would be to have exponential discounting of the past rewards.
+We believe that the results should be qualitatively similar to those presented here as in both formulations the more recent past matters more than the distant past.}.
 \begin{center}

### 2018-08-23 analysis_code/agents_table.Rmd

@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ht_lhr_3_arms_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_longer_ws_raw.csv", sep=""))
 
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
+time_horizons <- c(1000, 2000)
 priors <- as.list(unique(dat['Prior']))$Prior

### 2018-08-23 analysis_code/tournament_table.Rmd

--- 2018-08-10 analysis_code/tournament_table.Rmd
@@
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_testing_raw.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ht_lhr_3_arms_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_no_rep_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(1000)
+time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior

### 2018-08-23 paper/competing_bandits.tex
updates to tex file

@@
 
-What about for the entrant? The results are not consistent across priors and are summarized as follows:
+What about for the entrant? The results are not the same across priors and are summarized as follows:
 \begin{enumerate}
@@
 \end{enumerate}
-The raw results can be found in the GitHub appendix and though they are not consistent, it is interesting to note that on the priors where we see the relative reputation cost for $TS$ we have that $DG$ is preferred to $TS$ and for the priors where we see that exploration is not costly, we have that $TS$ is preferred to $DG$.
+The raw results can be found in the GitHub appendix and though they are not same across priors, it is interesting to note that on the priors where we see the relative reputation cost for $TS$ we have that $DG$ is preferred to $TS$ and for the priors where we see that exploration is not costly, we have that $TS$ is preferred to $DG$.
 
@@
 Maintaining the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
-The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation.
+The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation when the incumbent plays $TS$.
 One possible intuition for this is that, for Heavy Tail, the information acquired after the $X$ rounds under $TS$ is sufficient so that it becomes relatively easier for $TS$ to recover the reputation loss compared to the information loss, but since $DEG$ and $DG$ learn slower and they've acquired less information over the $X$ rounds relative to $TS$ there is less of an informational advantage.

### 2018-08-26 paper/competing_bandits.tex
uncommented a package to make it compile.
updated paper and appendix

@@
 \usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
 \usepackage{hyperref}       % hyperlinks
@@
 \newtheorem{definition}{Definition}
+\newtheorem{finding}{Finding}
 \captionsetup{font=footnotesize}
 \bibliographystyle{aaai}
+\frenchspacing
 
@@
 Unlike in the classic models in economics regarding the interplay between competition and innovation described in \citet{barro2004economic}, here we have no explicit R \& D cost.
-Rather, firms are not incentivized to experiment under duopoly because experimentation has an implicit relative \textit{reputational} cost compared to the myopic alternative and, in our model, consumers only care about the reputation of a firm.
-Thus, competition disincentivizes experimentation by firms precisely because competition forces firms to take actions in order to incentivize consumers to select them over their competitors and, as a result, by allowing a firm to be a temporary monopolist the firm can experiment without worrying about the reputational consequences.
+Rather, firms are not incentivized to experiment under duopoly because exploration has an implicit relative \textit{reputational} cost compared to the greedy alternative and, in our model, consumers only care about the reputation of a firm.
+Thus, competition disincentivizes exploration by firms precisely because competition forces firms to take actions in order to incentivize consumers to select them over their competitors and, as a result, by allowing a firm to be a temporary monopolist the firm can experiment without worrying about the reputational consequences.
 Our findings give an alternative intuition to the empirically documented inverse-U relationship between competition and innovation discussed in \citet{aghion2005competition}.
@@
 \item At $t = 0$, the firms simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
-\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a "warm" start.
+\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a ``warm" start.
 The algorithm that the firm commits to makes $k$ rounds of progress and uses this information to initialize its information set.
-Additionally, the reputation score of the agents is initialized using the rewards from these $k$ rounds.
-\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses among 2 firms, given a reputation score for both firms.
+Additionally, the reputation score used by the agents is initialized using the rewards from these $k$ rounds.
+\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses the firm with the higher reputation at the time.
 \item The firm that is chosen selects an action $a_{t} \in A$ from a set of actions that is fixed across firms and rounds.
@@
 For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors.
-Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a "fake" prior of $Beta(\alpha=1,\beta=1)$}.
+Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require a prior to operate, such as Thompson Sampling, we use a "fake" prior of $Beta(\alpha=1,\beta=1)$}.
 We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.
@@
 The utility function for the agents is simply to maximize their reward in the one period in which they are alive.
-We suppose that agents do not attempt to manipulate the strategy of the firm nor do they take the strategies of the firm into account when choosing between the firms.
+We suppose that agents do not attempt to manipulate the strategy of the firm, do not take the strategies of the firm into account when choosing between the firms, and do not condition their behavior on calendar time.
 In our model, each agent uses the average reward of past agents as a proxy for their expected utility.
@@
 
-Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior.
-The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the GitHub appendix.
+Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior \footnote{The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the supplementary material.}.
 
-In the "competition game" the agents decide between the firms using the \textit{relative} reputation between the two firms.
-Namely, if firm $A$ has a higher reputation than firm $B$ agents will select firm $A$.
-A natural question to ask is whether looking at the mean performance is sufficient for understanding performance in the competition game.
-Figure \ref{rep_dist_nih} displays the reputation distribution at $t = 500$ for the same prior we reported the mean plot for previously.
-We see that the "naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not.
+\begin{finding}
+\textit{The mean performance of an algorithm is not a sufficient statistic for understanding the performance of an algorithm in competition.
+Looking at the distribution of reputation difference between algorithms we find that it tend is skewed to the right.
+Additionally we show that the mean performance of the algorithms is not always predictive of the competition game, even when both algorithms are given a large number of agents for free.}
+\end{finding}
+
+Is looking at the mean performance of between two algorithms sufficient for understanding what happens in competition?
+To answer this, we first look at the entire reputation distribution at $t = 500$ in Figure \ref{rep_dist_nih}.
+We see that the ``naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not.
 The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not.
@@
 However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$.
-In these cases its reputation may be substantially worse but for the competition game the difference between the reputation does not matter but only the binary comparison between them \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, considering the SoftMax decision rule in \citet{mansour2018competing}}.
+In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \citet{mansour2018competing}}.
+
+To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $TS$ and $DG$.
+This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large.
+Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution.
+There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the relative reputation proportion.
 
@@
 \label{rep_dist_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 200$}}
+\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
 \end{figure}
 
-This motivates looking at the entire distribution of reputation difference between two algorithms.
-Figure \ref{ts_dg_rep_diff_nih} shows the distribution of reputation difference between $TS$ and $DG$ across $t$.
-This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large.
-Thus, the mean is not a representative statistic of the entire reputation difference distribution.
-As an alternative statistic for understanding the results of the competition game we introduce the \textit{relative reputation proportion} which looks at the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$.
-This corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{Though we do not discuss it here, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion.
-In the GitHub appendix we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.
+We define the \textit{relative reputation proportion} in order to better understand what happens under competition.
+
+\begin{definition}
+\textit{Relative Reputation Proportion} - the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$
+\end{definition}
+
+\begin{finding}
+\textit{Algorithms that engage in purposeful exploration can lead to implicit relative reputation costs compared to the greedy alternative in the earlier rounds.
+We observe this under the Uniform and Heavy Tail priors.
+However, it is not observed under the Needle In Haystack prior}
+\end{finding}
+
+The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion.
+In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.
 Additionally, these results carry over to the competition game}.
@@
 The intuition behind this is that, especially since the firms start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information.
-However, this exploration leads to what we define as a \textit{relative reputation cost}.
-Namely, exploration leads to lower reputation relative to the myopic alternative.
-However, eventually the information acquired in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is "hard enough" so that $DG$ cannot trivially find the best arm.
-We define this as the \textit{relative reputation benefit} \footnote{We see similar behavior for the Heavy Tail prior}.
+However, eventually the information acquired from purposeful exploration in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm.
+The early exploration leads to what we define as a \textit{relative reputation cost} and the eventual gain in reputation leads to what we define as a \textit{relative reputation benefit}.
+
+\begin{definition}
+\textit{Relative reputation cost and benefit} - the relative reputation loss from an algorithm that engages in purposeful exploration compared to the greedy alternative.
+It is an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
+We call the priors where there is an early costly period followed by a later benefit period, relative reputation costly priors.
+\end{definition}
 
@@
 First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration.
-Second, in the Needle In Haystack prior there are no "bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
+Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
 Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
 However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
+Thus, only the Uniform and Heavy Tail priors are relative reputation costly priors.
 
@@
 We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly.
-In the permanent monopoly and permanent duopoly cases we show that, under certain conditions, firms are incentivized to commit to $DG$ whereas in the temporary monopoly case firms are incentivized to commit to $TS$.
+
+\begin{finding}
+\textit{Under the priors that are relative reputation costly and with low warm start, the optimal strategies in the competition game are:
+\begin{center}
+\textbf{Permanent Monopoly} - $DG$ \\
+\textbf{Temporary Monopoly} - $TS$ \\
+\textbf{Permanent Duopoly} - $DG$ \\
+\end{center}
+The conditions on $TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
+}
+\end{finding}
+
 We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition.
@@
 We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.
-Tables \ref{sim_nih}, \ref{sim_ht}, and \ref{sim_unif} report the results.
-\begin{enumerate}
-\item Under Needle In Haystack we have that $TS$ always does weakly better than $DEG$ and $DG$, regardless of the warm start
-\item Under Heavy Tail we have that, for low warm start, $DEG, DG > TS$ but that for the higher warm start values we consider $TS > DEG, DG$.
-Interestingly, even under the relatively high warm start we see $DG > DEG$.
-\item Under Uniform we have that $DEG, DG > TS$ for low warm start and $TS > DEG, DG$ for sufficiently high warm start.
-\end{enumerate}
 
+\begin{finding}
+\textit{For low warm starts, when exploration is relative reputation costly $DG >TS$ in competition, whereas when exploration is not relative reputation costly $TS > DG$.
+For sufficiently large warm starts, $TS$ is dominant across all priors.}
+\end{finding}
+
+Table \ref{sim_unif} shows the results under the Uniform prior.
+For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}.
 Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start value as fixing the starting point on the relative reputation plots.
 The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$.
-Combined with the fact that the EEOG values being relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right, similar to the distribution of reputation differences}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.
+The intuition for the result in competition aligns with the fact that for Uniform and Heavy Tail, $TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $TS$.
+However, if we move the warm start sufficiently high so that $TS$ moves to the relative reputation benefit regime, $TS > DG$.
 
-For the low values of the warm start we can see that $DG > TS$ for the Uniform prior but $TS > DG$ for the Needle In Haystack prior and this corresponds to what we see in the relative reputation proportion plots.
-The intuition for this result aligns with the intuition given in the discussion of the relative reputation plots.
-Namely, for low warm start values $TS$ incurs a relative reputation cost compared to $DG$ since it engages in purposeful exploration.
-Moving to a sufficiently large warm start so that the better algorithm has recovered from the relative reputation cost incurred via exploration leads to better algorithms winning.
-However, it appears that the warm starts required for this are unreasonably high, especially for the Uniform prior.
+However, we do not always see that $DG > TS$ under some warm starts.
+Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $TS > DG$.
+The intuition for this is precisely that, under the Needle In Haystack, $TS$ is not relative reputation costly compared to $DG$ and thus we see that $TS > DG$.
 
-Focusing on reasonable warm starts, $DG > TS$ when it appears that better algorithms suffer an early relative reputational cost due to exploration and that both firms enter at the same time so that there is a "permanent duopoly."
+Combining these results with the fact that the EEOG values are relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right, similar to the distribution of reputation differences}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.
+
+% latex table generated in R 3.4.0 by xtable 1.8-2 package
+% Thu Aug 16 13:13:01 2018
+\begin{table}[ht]
+\centering
+\caption{Duopoly Experiment Uniform} 
+\begin{tabular}{rlll}
+  \hline
+ & k = 20 & k = 250 & k = 500 \\ 
+  \hline
+TS vs DG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.02\\ eeog \\ avg: 800\\ med: 754} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 910\\ med: 906.5} \\
+  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 180\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.02\\ eeog \\ avg: 810\\ med: 734} & \makecell{\textbf{0.55} $\pm$0.02\\ eeog \\ avg: 970\\ med: 987} \\
+  DG vs DEG & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 470\\ med: 57.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1088} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1142} \\
+   \hline
+\end{tabular}
+\label{sim_unif}
+\end{table}
 
@@
 % Thu Aug 16 13:13:00 2018
-\begin{table}[ht]
-\centering
-\caption{Duopoly Experiment Heavy Tail} 
-\begin{tabular}{rlll}
-  \hline
- & k = 20 & k = 250 & k = 500 \\ 
-  \hline
-TS vs DG & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 55\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 570\\ med: 0} & \makecell{\textbf{0.76} $\pm$0.02\\ eeog \\ avg: 620\\ med: 98.5} \\
-  TS vs DEG & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 37\\ med: 0} & \makecell{\textbf{0.88} $\pm$0.01\\ eeog \\ avg: 480\\ med: 0} & \makecell{\textbf{0.9} $\pm$0.01\\ eeog \\ avg: 570\\ med: 113.5} \\
-  DG vs DEG & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 410\\ med: 7} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 790\\ med: 762} & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 730\\ med: 608} \\
-   \hline
-\end{tabular}
-\label{sim_ht}
-\end{table}
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Thu Aug 16 13:13:01 2018
-\begin{table}[ht]
-\centering
-\caption{Duopoly Experiment Uniform} 
-\begin{tabular}{rlll}
-  \hline
- & k = 20 & k = 250 & k = 500 \\ 
-  \hline
-TS vs DG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.02\\ eeog \\ avg: 800\\ med: 754} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 910\\ med: 906.5} \\
-  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 180\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.02\\ eeog \\ avg: 810\\ med: 734} & \makecell{\textbf{0.55} $\pm$0.02\\ eeog \\ avg: 970\\ med: 987} \\
-  DG vs DEG & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 470\\ med: 57.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1088} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1142} \\
-   \hline
-\end{tabular}
-\label{sim_unif}
-\end{table}
 
 \textbf{Temporary Monopoly} We now consider asymmetries in the timing of entry so that one firm enters the market before the other and serves as a monopolist in the periods until the other firm enters.
-In terms of our model, this corresponds to setting $X = 200$.
-For this section we report results fixing $k = 20$.
-Does this lead the incumbent to commit to $TS$?
-How about the entrant?
+In terms of our model, this corresponds to varying the value of $X$.
+For this section we report results fixing $k = 20$ and focus on the priors where we observed that, for this warm start, $DG > TS$ \footnote{However, we report the results for the Needle In Haystack prior in the supplementary material.
+The results are unsurprising, where $TS$ is the dominant strategy for both the entrant and incumbent.}.
+It seems clear that the temporary monopoly will get a larger market share regardless of what algorithm it plays, but is the incumbent incentivized to commit to $TS$?
 
-Table \ref{ht_incum} shows the results of the simulations for this parameterization.
-We observe that, for the incumbent, $TS$ is the dominant strategy across all priors.
-In the appendix we include additional simulations where $X = 50$ and observe that $TS$ is not always a dominant strategy in this case.
-This result is robust across all of the priors that we considered so that for sufficiently large $X$, $TS$ is a dominant strategy for the incumbent where $X$ depends on the set of instances that we are considering.
+\begin{finding}
+\textit{Counterintuitively, allowing one firm to be a temporary monopolist for a sufficiently long time will induce that firm to commit to the better learning algorithm, $TS$, even if, under the permanent duopoly, they would commit to $DG$}
+\end{finding}
+
+In table \ref{ht_incum} we can see that, for $X = 200$, $TS$ is a dominant strategy for the incumbent and that $DG$ is the dominant strategy for the entrant on the Heavy Tail prior \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $TS$.
+However, while $DG > TS$ for the entrant, $DG$ does roughly the same as $DEG$ on the Uniform prior.
+The results are in the supplementary material.}.
+In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $TS$ is the dominant strategy for the incumbent is robust across priors.
 
@@
 \centering
-\caption{Temporary Monopoly Experiment Heavy Tail} 
-\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
+\caption{Temporary Monopoly Heavy Tail X = 200} 
+\begin{tabular}{rlll}
   \hline
@@
   \hline
-TS & \makecell{\textbf{ 0.0029 } $\pm$ 0.003 \\Var: 0.002 \\ ES: 100 \%} & \makecell{\textbf{ 0.11 } $\pm$ 0.02 \\Var: 0.09 \\ ES: 96 \%} & \makecell{\textbf{ 0.17 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 95 \%} \\
-  DEG & \makecell{\textbf{ 0.049 } $\pm$ 0.01 \\Var: 0.03 \\ ES: 94 \%} & \makecell{\textbf{ 0.24 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 74 \%} & \makecell{\textbf{ 0.24 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 79 \%} \\
-   DG & \makecell{\textbf{ 0.12 } $\pm$ 0.02 \\Var: 0.08 \\ ES: 87 \%} & \makecell{\textbf{ 0.35 } $\pm$ 0.03 \\Var: 0.2 \\ ES: 77 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 65 \%} \\
+TS & \makecell{\textbf{0.003} $\pm$0.003\\Var:0.002\\ES:100\%} & \makecell{\textbf{0.083} $\pm$0.02\\Var:0.07\\ES:97\%} & \makecell{\textbf{0.17} $\pm$0.02\\Var:0.1\\ES:95\%} \\ 
+  DEG & \makecell{\textbf{0.045} $\pm$0.01\\Var:0.03\\ES:92\%} & \makecell{\textbf{0.25} $\pm$0.02\\Var:0.1\\ES:75\%} & \makecell{\textbf{0.23} $\pm$0.02\\Var:0.1\\ES:78\%} \\ 
+   DG & \makecell{\textbf{0.12} $\pm$0.02\\Var:0.08\\ES:88\%} & \makecell{\textbf{0.36} $\pm$0.03\\Var:0.2\\ES:76\%} & \makecell{\textbf{0.3} $\pm$0.02\\Var:0.1\\ES:64\%} \\ 
    \hline
 \end{tabular}
+\label{ht_incum}
 \caption*{\tiny{The columns represent the strategy of the incumbent and the rows represent the strategy of the entrant.
@@
 the free rounds to firm 2 do not count towards the share)}}
-\label{ht_incum}
 \end{table}
-Why do we observe that $TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, for this warm start, $DG$ was preferred to $TS$?
+
+Why do we observe that $TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same other conditions, $DG$ was preferred to $TS$?
 The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
 This intuition is very similar to that observed in the previous section by increasing warm start.
-In some sense one can view allowing one firm to temporarily be a monopolist as temporarily relaxing the "incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
+One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the "incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
 The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.
@@
 
-What about for the entrant? The results are not the same across priors and are summarized as follows:
-\begin{enumerate}
-\item Needle In Haystack - $TS > DG, DEG$
-\item Heavy Tail - $DG > TS, DEG$
-\item Uniform - $DG \sim DEG > TS$
-\end{enumerate}
-The raw results can be found in the GitHub appendix and though they are not same across priors, it is interesting to note that on the priors where we see the relative reputation cost for $TS$ we have that $DG$ is preferred to $TS$ and for the priors where we see that exploration is not costly, we have that $TS$ is preferred to $DG$.
-
 \section{Data and Reputation as Barriers to Entry}\label{section:6}
@@
 An alternative interpretation of the results in the previous section is that the "temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant.
-This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \citep*{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry.
+This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \citep*{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry simply due to the fact that they were in the market before the entrant.
 Which plays a bigger role in preventing the entrant from being able to establish market share?
-We run two additional experiments, modifying the previous incumbent experiment so that in one set of simulations the reputation of the incumbent is artificially erased and another in which the information gained by the incumbent is artificially erased so that the posterior is reset to the prior.
+We run two additional experiments, modifying the previous temporary monopoly experiment so that in one experiment the reputation of the incumbent is artificially erased (Reputation Erased Experiment) and another in which the information gained by the incumbent is artificially erased so that the posterior is reset to the prior (Information Erased Experiment).
 
@@
 
+\begin{finding}
+\textit{When the incumbent plays $TS$, the data advantage serves as a larger barrier to entry compared to reputation. However, when the incumbent plays $DEG$ or $DG$ it does not} 
+\end{finding}
+
 Tables \ref{rep_erase} and \ref{info_erase} display the results of these experiments.
-Maintaining the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
+Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
 The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation when the incumbent plays $TS$.
@@
 Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
-We find that the best responses remain the same with the exception of the Uniform prior where $TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.
+We find that the best responses remain the same for the Heavy Tail prior, but for the Uniform prior they change where $TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.
 

### 2018-08-26 paper/competing_bandits.tex
updated paper

@@
 \begin{finding}
-\textit{Algorithms that engage in purposeful exploration can lead to implicit relative reputation costs compared to the greedy alternative in the earlier rounds.
+\textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $TS$ doing worse than $DG$ for small time horizons.
 We observe this under the Uniform and Heavy Tail priors.
@@
 \begin{definition}
-\textit{Relative reputation cost and benefit} - the relative reputation loss from an algorithm that engages in purposeful exploration compared to the greedy alternative.
-It is an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
+\textit{Relative reputation cost and benefit} - the relative reputation loss an algorithm incurs from purposeful exploration compared to the greedy alternative.
+Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
 We call the priors where there is an early costly period followed by a later benefit period, relative reputation costly priors.

### 2018-08-27 paper/competing_bandits.tex
fixes to the paper

@@
 There are two reasons why this is important to consider.
-First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where $TS > DEG > DG$ for sufficiently large $t$.
-Second, we want to better understand what statistics of the instance we can look at in isolation in order to help us predict and understand what to expect in the competition game.
+First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $TS > DEG > DG$ for sufficiently large $t$.
+Second, we want to better understand what statistics of the priors we can look at in isolation in order to help us predict and understand what to expect in the competition game.
 
@@
 Looking at the distribution of reputation difference between algorithms we find that it tend is skewed to the right.
-Additionally we show that the mean performance of the algorithms is not always predictive of the competition game, even when both algorithms are given a large number of agents for free.}
+Additionally we provide a counterexample where the mean performance does not predict the result of the competition game.}
 \end{finding}
@@
 Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution.
-There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the relative reputation proportion.
+There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the\textit{relative reputation proportion}.
 
@@
 \end{figure}
-
-We define the \textit{relative reputation proportion} in order to better understand what happens under competition.
 
@@
 Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
-We call the priors where there is an early costly period followed by a later benefit period, relative reputation costly priors.
+We call the priors where there is an early costly period followed by a later benefit period relative reputation costly priors.
 \end{definition}
@@
 For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}.
-Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start value as fixing the starting point on the relative reputation plots.
+Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $k$ as fixing the starting point on the relative reputation plots.
 The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$.
-The intuition for the result in competition aligns with the fact that for Uniform and Heavy Tail, $TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $TS$.
+The intuition for the result in competition aligns with the fact that, for Uniform and Heavy Tail, $TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $TS$.
 However, if we move the warm start sufficiently high so that $TS$ moves to the relative reputation benefit regime, $TS > DG$.
@@
 Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $TS > DG$.
-The intuition for this is precisely that, under the Needle In Haystack, $TS$ is not relative reputation costly compared to $DG$ and thus we see that $TS > DG$.
+This is precisely due to the fact that, under the Needle In Haystack, $TS$ is not relative reputation costly compared to $DG$ and thus we see that $TS > DG$.
 
@@
 \begin{finding}
-\textit{When the incumbent plays $TS$, the data advantage serves as a larger barrier to entry compared to reputation. However, when the incumbent plays $DEG$ or $DG$ it does not} 
+\textit{When the incumbent plays $TS$, the data advantage serves as a larger barrier to entry compared to reputation.
+However, when the incumbent plays $DEG$ or $DG$ there is no substantial difference between the two.}
 \end{finding}

### 2018-08-27 paper/competing_bandits.tex
tweaks

@@
 Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution.
-There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the\textit{relative reputation proportion}.
+There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the \textit{relative reputation proportion}.
 
-\begin{figure}
+\begin{figure}[H]
 \caption{Reputation Distribution}
@@
 
-\begin{figure}
-\caption{Reputation Difference Distribution}
-\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
-\label{ts_dg_rep_diff_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $TS$ and $DG$ across $t$}}
+\begin{figure}[H]
+\caption{Relative Reputation Plots}
+\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
+\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
+\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$.
+The shaded area display 95\% confidence intervals.}}
+\label{relative_rep_plots}
+
 \end{figure}
@@
 \end{definition}
-
-\begin{figure}
-\caption{Relative Reputation Plots}
-\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
-\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
-\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$.
-The shaded area display 95\% confidence intervals.}}
-\label{relative_rep_plots}
-
-\end{figure}
-
 Is exploration always costly?

### 2018-08-27 paper/competing_bandits.tex
readd unintentionally omitted figure
Started revising abstract / intro.

@@
 \captionsetup{font=footnotesize}
-\bibliographystyle{aaai}
+
 \frenchspacing
@@
 
+\newcommand{\OMIT}[1]{}
+
 \begin{document}
@@
 
-\author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
-Zhiwei Steven Wu\textsuperscript{4} \\
-{\textsuperscript{1}Columbia University, Department of Economics}\\
-{\textsuperscript{2}Columbia University, Department of Computer Science}\\
-{\textsuperscript{3}Microsoft Research, New York, NY}\\
-{\textsuperscript{4}University of Minnesota - Twin Cities, Department of Computer Science}
-}
+\title{Competing Bandits: Exploration and Competition under Duopoly}
+
+%\author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
+%Zhiwei Steven Wu\textsuperscript{4} \\
+%{\textsuperscript{1}Columbia University, Department of Economics}\\
+%{\textsuperscript{2}Columbia University, Department of Computer Science}\\
+%{\textsuperscript{3}Microsoft Research, New York, NY}\\
+%{\textsuperscript{4}University of Minnesota - Twin Cities, Department of Computer Science}
+%}
 \maketitle
 
+\begin{abstract}
+We empirically study the interplay between \textit{exploration} and \textit{competition}. Systems that learn from interactions with users
+often engage in \emph{exploration}: making potentially suboptimal decisions in order to acquire new information for future decisions.
+However, exploration may hurt system's reputation in the near term, with adverse competitive effects.
+In particular, a system may enter a ``death spiral" when decreasing market share leaves the system with less users to learn from, which degrades system's performance relative to competition and further decreases the market share.
+
+We consider an idealized model in which two firms deploy multi-armed bandit algorithms and compete for myopic users.
+The users select between the firms based on a reputation score, which depends on the past users' rewards.
+We run extensive numerical experiments, asking whether better bandit algorithms are incentivized under varying degrees of competition, from monopoly to a temporary monopoly (a duopoly with an early entrant) to a duopoly.
+We find that duopoly (as well as monopoly) tends to favor the ``greedy algorithm" that does not explore, whereas a temporary monopoly may incentivize the adoption of better bandit algorithms.
+Our results provide new intuition on the ``first-mover advantage" phenomenon in the digital economy, where the incumbent firm enjoys both a ``data advantage" and a ``reputational advantange".
+We further investigate which of the two is a stronger barrier to entry.
+\end{abstract}
+
+\OMIT{ %%% Guy's original abstract
 \begin{abstract}
@@
 \end{abstract}
+} %%%%%%%
 
@@
 Many modern online platforms simultaneously compete for users as well as learn from the users they manage to attract.
-For instance, Google Search and Bing compete for users in the search engine market yet at the same time need to experiment with their search algorithms to learn what algorithms work the best.
-This creates a tradeoff between \textit{exploration} and \textit{competition} since firms need to experiment with potentially sub-optimal options for the benefit of gaining information to make better decisions tomorrow while at the same time firms need to incentivize consumers to select them over their competitors today.
+This creates a tradeoff between \textit{exploration} and \textit{competition}: firms experiment with potentially sub-optimal options for the sake of gaining information to make better decisions tomorrow, while they need to incentivize consumers to select them over their competitors today.
+For instance, Google Search and Bing compete for users in the search engine market yet at the same time need to experiment with their search and ranking algorithms to learn what works best.
 
-From a social welfare perspective it is better for firms to adopt "better" learning algorithms and our main high-level question is asking under what levels of competition are firms incentivized to adopt "better" learning algorithms.
+Platforms routinely deploy randomized controlled trials (a.k.a.
+A/B tests), and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
+While deploying ``better" learning algorithms for exploration would (deployment/maintenance costs aside) improve performance, this is not necessarily beneficial under competition.
+In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term, which would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+
+STOPPED HERE.
+
+From a social welfare perspective it is better for firms to adopt "better" learning algorithms for exploration.
+Our main high-level question is asking under what levels of competition are firms incentivized to adopt "better" learning algorithms.
 Our primary contribution is that in our model, even in the case of a duopoly, firms are not incentivized to adopt better learning algorithms but that allowing one firm to be a temporary monopolist will incentivize this firm to play a better learning algorithm.
@@
 
+\begin{figure}[H]
+\caption{Reputation Distribution}
+\includegraphics[scale=0.35]{figures/rep_distribution_nih}
+\label{rep_dist_nih}
+\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
+\end{figure}
+
 \begin{finding}
@@
 \begin{figure}[H]
-\caption{Reputation Distribution}
-\includegraphics[scale=0.35]{figures/rep_distribution_nih}
-\label{rep_dist_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
+\caption{Reputation Difference Distribution}
+\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
+\label{ts_dg_rep_diff_nih}
+\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $TS$ and $DG$ across $t$}}
 \end{figure}
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
 \caption{Relative Reputation Plots}
@@
 \bibliography{refs}
-
+\bibliographystyle{aaai}
 \end{document}

### 2018-08-28 paper/competing_bandits.tex
Half-way through the intro.

@@
 \newcommand{\OMIT}[1]{}
+\newcommand{\xhdr}[1]{\vspace{1mm} \noindent{\bf #1}}
+
+% a very useful package for edits and comments, from David Kempe (USC)
+\usepackage{color-edits}
+%\usepackage[suppress]{color-edits}  % use this to suppress the package
+\addauthor{as}{red}    % as for Alex
+\addauthor{ga}{blue}  % ga for Guy
+\addauthor{sw}{green} % sw for Steven
+% e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
 
@@
 
-We consider an idealized model in which two firms deploy multi-armed bandit algorithms and compete for myopic users.
-The users select between the firms based on a reputation score, which depends on the past users' rewards.
-We run extensive numerical experiments, asking whether better bandit algorithms are incentivized under varying degrees of competition, from monopoly to a temporary monopoly (a duopoly with an early entrant) to a duopoly.
-We find that duopoly (as well as monopoly) tends to favor the ``greedy algorithm" that does not explore, whereas a temporary monopoly may incentivize the adoption of better bandit algorithms.
-Our results provide new intuition on the ``first-mover advantage" phenomenon in the digital economy, where the incumbent firm enjoys both a ``data advantage" and a ``reputational advantange".
-We further investigate which of the two is a stronger barrier to entry.
+We ask whether better exploration algorithms are incentivized under competition.
+We run extensive numerical experiments in a stylized duopoly model in which two firms deploy multi-armed bandit algorithms and compete for myopic users.
+We find that duopoly (as well as monopoly) tends to favor a primitive ``greedy algorithm" that does not explore.
+However, a temporary monopoly (a duopoly with an early entrant) may incentivize the adoption of better bandit algorithms.
+Our findings provide new intuition on the ``first-mover advantage" phenomenon in the digital economy.
 \end{abstract}
@@
 \begin{abstract}
-We empirically study the interplay between \textit{exploration} and \textit{competition} - how systems that learn from interactions with users tradeoff between making potentially suboptimal choices in order to acquire new information and the reputational consequences of this exploration which potentially has adverse competitive effects.
+We empirically study the interplay between \textit{exploration} and \textit{competition} - how platforms that learn from interactions with users tradeoff between making potentially suboptimal choices in order to acquire new information and the reputational consequences of this exploration which potentially has adverse competitive effects.
 Our model considers competition for myopic users between two firms deploying multi-armed bandit algorithms that face the same underlying multi-armed bandit instance.
@@
 
-Platforms routinely deploy randomized controlled trials (a.k.a.
-A/B tests), and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
+Platforms routinely deploy randomized controlled trials (A/B tests), and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
 While deploying ``better" learning algorithms for exploration would (deployment/maintenance costs aside) improve performance, this is not necessarily beneficial under competition.
-In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term, which would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term.
+This would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+
+We ask whether competition incentivizes adoption of "better" algorithms for exploration.
+We investigate this issue via extensive numerical experiments in a stylized duopoly model.
+In our model, two firms compete for users, and simultaneously learn from them.
+Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm.
+Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window.
+Each firms' objective is the market share.
+
+We consider varying levels of competition, from monopoly to \emph{temporary monopoly} (a duopoly with an early entrant), to a true duopoly in which both firms start at the same time.
+We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly.
+This algorithm also prevails under monopoly simply because it tends to be easier to deploy.
+Whereas a temporary monopoly incentives more advanced exploration algorithms that perform better in the long run.
+These findings can be framed in terms of an ``inverted-U relationship" between competition and innovation.
+
+The disincentives to explore under duopoly arise entirely because of ``reputational cost", rather than R\&D costs (the latter are absent in our model).
+Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
+In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantange".
+We further investigate which of the two is a stronger barrier to entry.
+\ascomment{and find what?
+and what's the new intuition?}
+
+\subsection{Discussion and related work}
+
+Adoption of better exploration algorithms is a form of ``innovation", the interplay between competition and innovation being a well-studied topic in economics \citep[e.g.,][]{barro2004economic,aghion2005competition}, dating back to \citet{schumpeter2010capitalism}.
+A familiar theme is an ``inverted-U relationship" between competition and innovation -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
+While this literature focuses on R\&D costs of innovation, ``reputational costs" seem new and specific to exploration.
+
+Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information).
+MAB problems have been studied for many decades, e.g., see \citealt*{Bubeck-survey12} for background.
+We consider a simple and well-understood MAB model with i.i.d.
+rewards.
+We focus on a (well-known) distinction between the ``greedy" (exploitation-only) algorithm, "naive" exploration algorithms that separate exploration and exploitation, and more sophisticated algorithms that combine the two.
 
@@
 
-From a social welfare perspective it is better for firms to adopt "better" learning algorithms for exploration.
-Our main high-level question is asking under what levels of competition are firms incentivized to adopt "better" learning algorithms.
-Our primary contribution is that in our model, even in the case of a duopoly, firms are not incentivized to adopt better learning algorithms but that allowing one firm to be a temporary monopolist will incentivize this firm to play a better learning algorithm.
-
-Unlike in the classic models in economics regarding the interplay between competition and innovation described in \citet{barro2004economic}, here we have no explicit R \& D cost.
-Rather, firms are not incentivized to experiment under duopoly because exploration has an implicit relative \textit{reputational} cost compared to the greedy alternative and, in our model, consumers only care about the reputation of a firm.
-Thus, competition disincentivizes exploration by firms precisely because competition forces firms to take actions in order to incentivize consumers to select them over their competitors and, as a result, by allowing a firm to be a temporary monopolist the firm can experiment without worrying about the reputational consequences.
-Our findings give an alternative intuition to the empirically documented inverse-U relationship between competition and innovation discussed in \citet{aghion2005competition}.
-
-\textbf{Related Work} A major underlying component of our model is that firms face a multi-armed bandit (MAB) problem.
-Multi-armed bandits (MAB) are a tractable abstraction for the tradeoff between exploration and exploitation.
-MAB problems have been studied for many decades (see \citet{bubeck2012regret} for an overview).
-Most relevant to this paper is the thread that focuses on designing "smart" and tractable algorithms that combine exploration and exploitation and "naive" algorithms that separate exploration and exploitation (see \citet{slivkins2017bandits}).
-
-The three-way tradeoff between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}.
+The interplay between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}.
 The strategic experimentation literature in economics, such as \citealt*{bolton1999strategic} and \citealt*{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.
-
-The relationship between competition and innovation has been heavily studied in industrial organization \citep{tirole1988theory} and endogenous growth theory \citep{aghion2005competition, barro2004economic}, dating back to \citet{schumpeter2010capitalism}.
 
@@
 
-\section{Model}\label{section:2}
+\section{Model}\label{sec:model}
 
@@
 
-\bibliography{refs}
+\bibliography{refs,bib-bandits}
 \bibliographystyle{aaai}

### 2018-08-28 paper/competing_bandits.tex
Full pass on intro + related work.

@@
-\documentclass{article}
+\documentclass[letterpaper]{article}
+\usepackage{aaai}
+\usepackage{times}
+\usepackage{helvet}
+\usepackage{courier}
+\usepackage{url}
+\usepackage{graphicx}
+\frenchspacing
 
-% if you need to pass options to natbib, use, e.g.:
-% \PassOptionsToPackage{numbers, compress}{natbib}
-
-% ready for submission
-\usepackage{aaai}
-
-% to compile a preprint version, e.g., for submission to arXiv, add
-% add the [preprint] option:
-% \usepackage[preprint]{nips_2018}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2018}
-
-% to avoid loading the natbib package, add option nonatbib:
-% \usepackage[nonatbib]{nips_2018}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage{hyperref}       % hyperlinks
+%\usepackage[utf8]{inputenc} % allow utf-8 input -- cannot use with AAAI! A.S.
+%\usepackage{hyperref}       % hyperlinks
 \usepackage{url}            % simple URL typesetting
@@
 \usepackage{amssymb}
-\usepackage{natbib}
-%% The amsthm package provides extended theorem environments
-\usepackage{amsthm}
+%\usepackage{natbib} % cannot use with AAAI! A.S.
+
+\usepackage{amsthm} % The amsthm package provides extended theorem environments
 \usepackage{float}
@@
 
-\frenchspacing
-
 % The \author macro works with any number of authors. There are two
@@
 
+% Alex' macro
+\usepackage{xspace}
 \newcommand{\OMIT}[1]{}
 \newcommand{\xhdr}[1]{\vspace{1mm} \noindent{\bf #1}}
+\newcommand{\ie}{{\em i.e.,~\xspace}}
+\newcommand{\eg}{{\em e.g.,~\xspace}}
 
@@
 
-Adoption of better exploration algorithms is a form of ``innovation", the interplay between competition and innovation being a well-studied topic in economics \citep[e.g.,][]{barro2004economic,aghion2005competition}, dating back to \citet{schumpeter2010capitalism}.
+Adoption of better exploration algorithms is a form of ``innovation", the interplay between competition and innovation being a well-studied topic in economics, \eg \cite{barro2004economic,aghion2005competition}, dating back to \cite{schumpeter2010capitalism}.
 A familiar theme is an ``inverted-U relationship" between competition and innovation -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
@@
 Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information).
-MAB problems have been studied for many decades, e.g., see \citealt*{Bubeck-survey12} for background.
+MAB problems have been studied for many decades, \eg see \cite{Bubeck-survey12} for background.
 We consider a simple and well-understood MAB model with i.i.d.
@@
 
-STOPPED HERE.
+The study of competition vs.
+exploration has been initiated in \cite{CompetingBandits-itcs16}.
+Their model differs from ours in two key respects.
+First, users do not see any signal about firms' past performance, and instead choose between firms according to the Bayesian-expected reward.
+Second, they vary the strength of competition using assumptions about (ir)rational consumer behavior, whereas we use early entry.
+Their results are purely theoretical; their model is amenable to proofs but not to numerical experiments.
 
+The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system
+    \cite{Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17},
+dynamic auctions
+    \cite{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
+pay-per-click ad auctions 
+    \cite{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
+and human computation
+    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
+
+\OMIT{ %%%%%%
 The interplay between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}.
 The strategic experimentation literature in economics, such as \citealt*{bolton1999strategic} and \citealt*{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.
+} %%%%%%%%
 
-Our setting is also closely related to the "dueling algorithms" framework described in \cite{immorlica2011dueling}, but this framework considers offline scenarios whereas we focus on online learning problems.
-
-The most closely related work to this paper is \citealt*{mansour2018competing}.
-Our motivating questions are the same as \citet{mansour2018competing} though while \citet{mansour2018competing} use the rationality of consumers as their primary knob of competitiveness, we consider differences in the number of firms in the market as our primary knob.
-Additionally, while in \citet{mansour2018competing} agents select firms based on which firm they expect to have a larger Bayesian expected reward, in our model the agents select firms based on a frequentist reputation score that is derived from the signals sent from past agents.
-This allows the model to become tractable for the purposes of evaluating it via simulation and this is the primary method of analysis that we employ.
+Our setting is also closely related to the "dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
 
@@
 \item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
-We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \citep{agrawal2012analysis}.
+We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \cite{agrawal2012analysis}.
 \item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
@@
 However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$.
-In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \citet{mansour2018competing}}.
+In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \cite{mansour2018competing}}.
 
@@
 
-In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \citet{aghion2005competition}.
+In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \cite{aghion2005competition}.
 We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly.
@@
 An alternative interpretation of the results in the previous section is that the "temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant.
-This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \citep*{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry simply due to the fact that they were in the market before the entrant.
+This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \cite*{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry simply due to the fact that they were in the market before the entrant.
 Which plays a bigger role in preventing the entrant from being able to establish market share?
@@
 
-\bibliography{refs,bib-bandits}
+\bibliography{refs,bib-abbrv-short,bib-bandits,bib-AGT,bib-slivkins}
 \bibliographystyle{aaai}

### 2018-08-28 paper/competing_bandits.tex
edits on intro

@@
 \usepackage{graphicx}
+\usepackage{tikz} % SW: Added for plots
 \frenchspacing
@@
 \addauthor{ga}{blue}  % ga for Guy
-\addauthor{sw}{green} % sw for Steven
+\addauthor{sw}{orange} % sw for Steven
 % e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
@@
 Platforms routinely deploy randomized controlled trials (A/B tests), and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
-While deploying ``better" learning algorithms for exploration would (deployment/maintenance costs aside) improve performance, this is not necessarily beneficial under competition.
+While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition \swedit{(putting deployment/maintenance costs aside)}.
 In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term.
@@
 Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window.
-Each firms' objective is the market share.
+\swedit{The objective of both firms is to maximize the market share ---the fraction of users coming to them.}
 
@@
 Whereas a temporary monopoly incentives more advanced exploration algorithms that perform better in the long run.
-These findings can be framed in terms of an ``inverted-U relationship" between competition and innovation.
+These findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
+
+\tikzstyle{level 1}=[level distance=3.5cm, sibling distance=4.0cm]
+\tikzstyle{level 2}=[level distance=3.5cm, sibling distance=2cm]
+\tikzstyle{below} = [align=center]
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+      \draw[->] (-.5,0) -- (6,0) node[right] {};
+      \draw[->] (0,-.5) -- (0,2) node[above] {Better algorithms};
+      \draw[scale=0.6,domain=0.7:8.3,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{3.4 - 0.2 * (\x - 4.5)^2});
+     \node[below] at (0.7, -0.15) {\footnotesize monopoly};
+     \node[below] at (3, 0) {\footnotesize temporary \\ monopoly};
+     \node[below] at (5, -0.15) {\footnotesize duopoly};
+ \end{tikzpicture}
+ \caption{Inverted-U relationship between different regimes of
+   competition and the incentivized exploration algorithms.}
+\label{fig:inverted-U}
+\end{center}
+\end{figure}
 
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2018-08-28 paper/competing_bandits.tex
edits on related work

@@
 \subsection{Discussion and related work} 
-
-Adoption of better exploration algorithms is a form of ``innovation", the interplay between competition and innovation being a well-studied topic in economics, \eg \cite{barro2004economic,aghion2005competition}, dating back to \cite{schumpeter2010capitalism}.
+\swedit{We view the adoption of better exploration algorithms as a form of ``innovation", and thus our work is related to a longstanding economics literature on the interplay between competition and innovation,} \eg \cite{barro2004economic,aghion2005competition}, dating back to \cite{schumpeter2010capitalism}.
 A familiar theme is an ``inverted-U relationship" between competition and innovation -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.

### 2018-08-29 paper/competing_bandits.tex
small changes

@@
 \newcommand{\eg}{{\em e.g.,~\xspace}}
+\newcommand{\TS}{\mathrm{TS}}
+\newcommand{\DEG}{\mathrm{DEG}}
 
@@
 \item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
-We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \cite{agrawal2012analysis}.
+We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \cite{agrawal2012analysis}.
 \item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
-We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability.
+We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $\DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability.
 For our experiments we keep $\epsilon = 0.05$ fixed.
@@
 
-In the standard multi-armed bandit problem it is known that $TS > DEG > DG$ in terms of maximizing cumulative reward over a sufficiently large time horizon.
+In the standard multi-armed bandit problem it is known that $\TS > \DEG > DG$ in terms of maximizing cumulative reward over a sufficiently large time horizon.
 The primary question that we want to understand is when, in competition, are the firms incentivized to adopt the "better" algorithms?
@@
 There are two reasons why this is important to consider.
-First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $TS > DEG > DG$ for sufficiently large $t$.
+First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $\TS > \DEG > DG$ for sufficiently large $t$.
 Second, we want to better understand what statistics of the priors we can look at in isolation in order to help us predict and understand what to expect in the competition game.
@@
 To answer this, we first look at the entire reputation distribution at $t = 500$ in Figure \ref{rep_dist_nih}.
-We see that the ``naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not.
+We see that the ``naive" algorithms $DG$ and $\DEG$ have a bi-modal reputation distribution whereas $\TS$ does not.
 The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not.
 If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds.
-However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$.
+However, if it does not then it will get stuck on a bad arm and lose to $\TS$ or $\DEG$.
 In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \cite{mansour2018competing}}.
 
-To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $TS$ and $DG$.
+To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $\TS$ and $DG$.
 This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large.
@@
 \label{ts_dg_rep_diff_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $TS$ and $DG$ across $t$}}
+\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $DG$ across $t$}}
 \end{figure}
@@
 \begin{finding}
-\textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $TS$ doing worse than $DG$ for small time horizons.
+\textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $\TS$ doing worse than $DG$ for small time horizons.
 We observe this under the Uniform and Heavy Tail priors.
@@
 
-The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion.
-In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.
+The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $\DEG >DG$ or $\TS > DG$ according to the mean reputation but $DG > \DEG$ or $DG > TS$ according to the relative reputation proportion.
+In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $\DEG > DG$ according to the mean reputation but $DG > \DEG$ according to the relative reputation proportion.
 Additionally, these results carry over to the competition game}.
@@
 \includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
-\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$.
+\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq DG$.
 The shaded area display 95\% confidence intervals.}}
@@
 
-Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $TS$ vs $DG$ on the Uniform and Needle In Haystack prior.
-For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $TS > DG$.
-The intuition behind this is that, especially since the firms start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information.
-However, eventually the information acquired from purposeful exploration in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm.
+Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $DG$ on the Uniform and Needle In Haystack prior.
+For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $\TS > DG$.
+The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information.
+However, eventually the information acquired from purposeful exploration in the early rounds allows $\TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm.
 The early exploration leads to what we define as a \textit{relative reputation cost} and the eventual gain in reputation leads to what we define as a \textit{relative reputation benefit}.
@@
 Is exploration always costly?
-Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $TS$ always does relatively better than $DG$.
+Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $\TS$ always does relatively better than $DG$.
 There are two contributing factors to this.
-First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration.
+First, $\TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration.
 Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
-Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
+Thus, when $\TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
 However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
@@
 \textbf{Permanent Monopoly} - $DG$ \\
-\textbf{Temporary Monopoly} - $TS$ \\
+\textbf{Temporary Monopoly} - $\TS$ \\
 \textbf{Permanent Duopoly} - $DG$ \\
 \end{center}
-The conditions on $TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
+The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
 }
@@
 \begin{finding}
-\textit{For low warm starts, when exploration is relative reputation costly $DG >TS$ in competition, whereas when exploration is not relative reputation costly $TS > DG$.
-For sufficiently large warm starts, $TS$ is dominant across all priors.}
+\textit{For low warm starts, when exploration is relative reputation costly $DG >TS$ in competition, whereas when exploration is not relative reputation costly $\TS > DG$.
+For sufficiently large warm starts, $\TS$ is dominant across all priors.}
 \end{finding}
@@
 Table \ref{sim_unif} shows the results under the Uniform prior.
-For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}.
+For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $\TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}.
 Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $k$ as fixing the starting point on the relative reputation plots.
 The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$.
-The intuition for the result in competition aligns with the fact that, for Uniform and Heavy Tail, $TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $TS$.
-However, if we move the warm start sufficiently high so that $TS$ moves to the relative reputation benefit regime, $TS > DG$.
+The intuition for the result in competition aligns with the fact that, for Uniform and Heavy Tail, $\TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $\TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $\TS$.
+However, if we move the warm start sufficiently high so that $\TS$ moves to the relative reputation benefit regime, $\TS > DG$.
 
 However, we do not always see that $DG > TS$ under some warm starts.
-Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $TS > DG$.
-This is precisely due to the fact that, under the Needle In Haystack, $TS$ is not relative reputation costly compared to $DG$ and thus we see that $TS > DG$.
+Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $\TS > DG$.
+This is precisely due to the fact that, under the Needle In Haystack, $\TS$ is not relative reputation costly compared to $DG$ and thus we see that $\TS > DG$.
 
@@
 For this section we report results fixing $k = 20$ and focus on the priors where we observed that, for this warm start, $DG > TS$ \footnote{However, we report the results for the Needle In Haystack prior in the supplementary material.
-The results are unsurprising, where $TS$ is the dominant strategy for both the entrant and incumbent.}.
-It seems clear that the temporary monopoly will get a larger market share regardless of what algorithm it plays, but is the incumbent incentivized to commit to $TS$?
+The results are unsurprising, where $\TS$ is the dominant strategy for both the entrant and incumbent.}.
+It seems clear that the temporary monopoly will get a larger market share regardless of what algorithm it plays, but is the incumbent incentivized to commit to $\TS$?
 
 \begin{finding}
-\textit{Counterintuitively, allowing one firm to be a temporary monopolist for a sufficiently long time will induce that firm to commit to the better learning algorithm, $TS$, even if, under the permanent duopoly, they would commit to $DG$}
+\textit{Counterintuitively, allowing one firm to be a temporary monopolist for a sufficiently long time will induce that firm to commit to the better learning algorithm, $\TS$, even if, under the permanent duopoly, they would commit to $DG$}
 \end{finding}
 
-In table \ref{ht_incum} we can see that, for $X = 200$, $TS$ is a dominant strategy for the incumbent and that $DG$ is the dominant strategy for the entrant on the Heavy Tail prior \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $TS$.
-However, while $DG > TS$ for the entrant, $DG$ does roughly the same as $DEG$ on the Uniform prior.
+In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $DG$ is the dominant strategy for the entrant on the Heavy Tail prior \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.
+However, while $DG > TS$ for the entrant, $DG$ does roughly the same as $\DEG$ on the Uniform prior.
 The results are in the supplementary material.}.
-In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $TS$ is the dominant strategy for the incumbent is robust across priors.
+In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across priors.
 
@@
 
-Why do we observe that $TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same other conditions, $DG$ was preferred to $TS$?
+Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same other conditions, $DG$ was preferred to $\TS$?
 The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
@@
 As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
-Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $TS$.
+Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
 
@@
 \begin{finding}
-\textit{When the incumbent plays $TS$, the data advantage serves as a larger barrier to entry compared to reputation.
-However, when the incumbent plays $DEG$ or $DG$ there is no substantial difference between the two.}
+\textit{When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation.
+However, when the incumbent plays $\DEG$ or $DG$ there is no substantial difference between the two.}
 \end{finding}
@@
 Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
-The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation when the incumbent plays $TS$.
-One possible intuition for this is that, for Heavy Tail, the information acquired after the $X$ rounds under $TS$ is sufficient so that it becomes relatively easier for $TS$ to recover the reputation loss compared to the information loss, but since $DEG$ and $DG$ learn slower and they've acquired less information over the $X$ rounds relative to $TS$ there is less of an informational advantage.
+The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation when the incumbent plays $\TS$.
+One possible intuition for this is that, for Heavy Tail, the information acquired after the $X$ rounds under $\TS$ is sufficient so that it becomes relatively easier for $\TS$ to recover the reputation loss compared to the information loss, but since $\DEG$ and $DG$ learn slower and they've acquired less information over the $X$ rounds relative to $\TS$ there is less of an informational advantage.
 
 Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
-We find that the best responses remain the same for the Heavy Tail prior, but for the Uniform prior they change where $TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.
+We find that the best responses remain the same for the Heavy Tail prior, but for the Uniform prior they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.
 

### 2018-08-29 paper/competing_bandits.tex
fix weird bib issue and add additional citation

@@
 The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system
-    \cite{Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17},
+    \cite{Frazier-ec14,Kremer-JPE14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
 dynamic auctions
@@
 \OMIT{ %%%%%%
-The interplay between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}.
-The strategic experimentation literature in economics, such as \citealt*{bolton1999strategic} and \citealt*{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.
+The interplay between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \cite{che2017recommender}, \cite{kremer2014implementing}, \cite{mansour2015bayesian}.
+The strategic experimentation literature in economics, such as \cite{bolton1999strategic} and \cite{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.
 } %%%%%%%%
@@
 An alternative interpretation of the results in the previous section is that the "temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant.
-This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \cite*{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry simply due to the fact that they were in the market before the entrant.
+This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \cite{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry simply due to the fact that they were in the market before the entrant.
 Which plays a bigger role in preventing the entrant from being able to establish market share?

### 2018-08-29 paper/competing_bandits.tex
edits

@@
 Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require a prior to operate, such as Thompson Sampling, we use a "fake" prior of $Beta(\alpha=1,\beta=1)$}.
-We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.
+We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.\swcomment{this seems misleading and should not be here.
+just say that each action has bernouli rewards with unknown means}
 
@@
 
-Note $t_j$ is the \textit{local} time of the firm, not the global time, so the reputation score is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.
-The warm start of $k$ rounds allows this reputation score to be well-defined once the "competition game" begins.
+Note $t_j$ is the \textit{local} time of the firm, not the global time, so the reputation score is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.\swcomment{define ``calendar time'' and ``local time'' formally...
+don't think the current reputation defn makes sense} The warm start of $k$ rounds allows this reputation score to be well-defined once the "competition game" begins.
 In our model the agents deterministically choose the firm with the higher reputation score and ties are broken uniformly at random.
@@
 
-We evaluate the consequences of our model via simulation.
+We evaluate the consequences of our model via simulation.\swcomment{In the process of removing ``prior'' throughout}
 
-\textbf{Bandit Priors} We look at bandit instances drawn from three different bandit "priors" that are different types of learning problems.
-Recall that we only consider Bernoulli-distributed arms and so we draw the true means of the arms from these priors.
+\textbf{Bandit Instances} We look at bandit instances drawn from three different bandit \swedit{distributions} that capture different types of learning problems.
+Recall that we only consider Bernoulli rewards, which are fully parameterized by the means.
 For these experiments we fix the number of arms we consider, $K = 10$.

### 2018-08-29 paper/competing_bandits.tex
Another small pass on the intro.

@@
 We run extensive numerical experiments in a stylized duopoly model in which two firms deploy multi-armed bandit algorithms and compete for myopic users.
-We find that duopoly (as well as monopoly) tends to favor a primitive ``greedy algorithm" that does not explore.
-However, a temporary monopoly (a duopoly with an early entrant) may incentivize the adoption of better bandit algorithms.
-Our findings provide new intuition on the ``first-mover advantage" phenomenon in the digital economy.
+We find that duopoly \asedit{and monopoly} tend to favor a primitive ``greedy algorithm" that does not explore, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms.
+Our findings provide new intuition on the ``first-mover advantage" in the digital economy.
 \end{abstract}
@@
 
-Platforms routinely deploy randomized controlled trials (A/B tests), and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
-While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition \swedit{(putting deployment/maintenance costs aside)}.
+Platforms routinely deploy \asedit{A/B tests}, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
+While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition, \asedit{even putting aside the deployment/maintenance costs}.
 In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term.
@@
 Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window.
-\swedit{The objective of both firms is to maximize the market share ---the fraction of users coming to them.}
-
-We consider varying levels of competition, from monopoly to \emph{temporary monopoly} (a duopoly with an early entrant), to a true duopoly in which both firms start at the same time.
-We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly.
-This algorithm also prevails under monopoly simply because it tends to be easier to deploy.
-Whereas a temporary monopoly incentives more advanced exploration algorithms that perform better in the long run.
-These findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
+\asedit{Each firm's objective} is to maximize its market share ---the fraction of users coming to them.
 
@@
  \end{tikzpicture}
- \caption{Inverted-U relationship between different regimes of
-   competition and the incentivized exploration algorithms.}
+ \caption{A stylized ``inverted-U relationship" between strength of competition and ``level of innovation".}
 \label{fig:inverted-U}
@@
 
-The disincentives to explore under duopoly arise entirely because of ``reputational cost", rather than R\&D costs (the latter are absent in our model).
+We consider varying levels of competition, from monopoly to \emph{temporary monopoly} (a duopoly with an early entrant), to a true duopoly in which both firms start at the same time.
+We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly.
+This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
+Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run.
+The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model).
+
+\asedit{Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
+This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.}
+
 Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
@@
 \subsection{Discussion and related work}
-\swedit{We view the adoption of better exploration algorithms as a form of ``innovation", and thus our work is related to a longstanding economics literature on the interplay between competition and innovation,} \eg \cite{barro2004economic,aghion2005competition}, dating back to \cite{schumpeter2010capitalism}.
-A familiar theme is an ``inverted-U relationship" between competition and innovation -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
-While this literature focuses on R\&D costs of innovation, ``reputational costs" seem new and specific to exploration.
+Our work is related to a longstanding economics literature on the interplay between competition and innovation, \eg \cite{Schumpeter-42,barro2004economic,aghion2005competition}.
+While this literature focuses on R\&D costs of innovation, ``reputational costs" thereof seem new and specific to exploration.
 

### 2018-08-29 paper/competing_bandits.tex
Removed dup citations from ref.bib.

@@
 
-\title{Learning and Reputation under Monopoly and Duopoly: \\A Competing Bandits Approach}
+%\title{Learning and Reputation under Monopoly and Duopoly: \\A Competing Bandits Approach}
 
@@
 \subsection{Discussion and related work}
-Our work is related to a longstanding economics literature on the interplay between competition and innovation, \eg \cite{Schumpeter-42,barro2004economic,aghion2005competition}.
+Our work is related to a longstanding economics literature on competition vs.
+innovation, \eg \cite{Schumpeter-42,barro2004economic,Aghion-QJE05}.
 While this literature focuses on R\&D costs of innovation, ``reputational costs" thereof seem new and specific to exploration.
@@
 \item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
-We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \cite{agrawal2012analysis}.
+We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \cite{Shipra-colt12}.
 \item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
@@
 However, if it does not then it will get stuck on a bad arm and lose to $\TS$ or $\DEG$.
-In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \cite{mansour2018competing}}.
+In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \cite{CompetingBandits-itcs16}}.
 
@@
 
-In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \cite{aghion2005competition}.
+In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \cite{Aghion-QJE05}.
 We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly.

### 2018-08-30 paper/competing_bandits.tex
more edits in intro and related work.

@@
 
-\title{Competing Bandits: Exploration and Competition under Duopoly}
+%\title{Competing Bandits: Exploration and Competition under Duopoly}
+
+\title{Competing Bandits: The Perils of Exploration under Competition}
 
@@
 Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information).
-MAB problems have been studied for many decades, \eg see \cite{Bubeck-survey12} for background.
+MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background.
 We consider a simple and well-understood MAB model with i.i.d.
-rewards.
-We focus on a (well-known) distinction between the ``greedy" (exploitation-only) algorithm, "naive" exploration algorithms that separate exploration and exploitation, and more sophisticated algorithms that combine the two.
+rewards \cite{bandits-ucb1}.
+We focus on a well-known distinction between the ``greedy" (exploitation-only) algorithm, "naive" MAB  algorithms that separate exploration and exploitation, and better MAB algorithms that combine the two.
+As a token example of ``near-optimal" bandit algorithms,
+we use a classic algorithm called ``Thompson Sampling", see \cite{TS-survey-FTML18} for background.
 
@@
 dynamic auctions
-    \cite{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
-pay-per-click ad auctions
-    \cite{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
+    (see \cite{DynAuctions-survey10} for background),
+pay-per-click auctions
+    \cite{MechMAB-ec09,DevanurK09,Transform-ec10-jacm}
+%\cite{Amin-auctions-nips13,Jieming-ec18},
 and human computation
@@
 \item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
-We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \cite{Shipra-colt12}.
+We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward %\cite{Shipra-colt12}.
 \item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.

### 2018-08-30 paper/competing_bandits.tex
An edit in related work.

@@
 
-The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system
-    \cite{Frazier-ec14,Kremer-JPE14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
+The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
+    \eg \cite{Kremer-JPE14,Frazier-ec14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
 dynamic auctions
     (see \cite{DynAuctions-survey10} for background),
-pay-per-click auctions
-    \cite{MechMAB-ec09,DevanurK09,Transform-ec10-jacm}
-%\cite{Amin-auctions-nips13,Jieming-ec18},
+online ad auctions, \eg
+    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm,Amin-auctions-nips13},
 and human computation

### 2018-08-30 paper/competing_bandits.tex
break up the different sections into different files

@@
 \usepackage{url}
+\usepackage{subfiles}
 \usepackage{graphicx}
@@
 
-\section{Introduction}\label{section:1}
+\subfile{content/Introduction}
 
-Many modern online platforms simultaneously compete for users as well as learn from the users they manage to attract.
-This creates a tradeoff between \textit{exploration} and \textit{competition}: firms experiment with potentially sub-optimal options for the sake of gaining information to make better decisions tomorrow, while they need to incentivize consumers to select them over their competitors today.
-For instance, Google Search and Bing compete for users in the search engine market yet at the same time need to experiment with their search and ranking algorithms to learn what works best.
+\subfile{content/model}
 
-Platforms routinely deploy \asedit{A/B tests}, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
-While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition, \asedit{even putting aside the deployment/maintenance costs}.
-In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term.
-This would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+\subfile{content/sim_details}
 
-We ask whether competition incentivizes adoption of "better" algorithms for exploration.
-We investigate this issue via extensive numerical experiments in a stylized duopoly model.
-In our model, two firms compete for users, and simultaneously learn from them.
-Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm.
-Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window.
-\asedit{Each firm's objective} is to maximize its market share ---the fraction of users coming to them.
+\subfile{content/perf_in_iso}
 
-\tikzstyle{level 1}=[level distance=3.5cm, sibling distance=4.0cm]
-\tikzstyle{level 2}=[level distance=3.5cm, sibling distance=2cm]
-\tikzstyle{below} = [align=center]
+\subfile{content/inverted_u}
 
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-      \draw[->] (-.5,0) -- (6,0) node[right] {};
-      \draw[->] (0,-.5) -- (0,2) node[above] {Better algorithms};
-      \draw[scale=0.6,domain=0.7:8.3,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{3.4 - 0.2 * (\x - 4.5)^2});
-     \node[below] at (0.7, -0.15) {\footnotesize monopoly};
-     \node[below] at (3, 0) {\footnotesize temporary \\ monopoly};
-     \node[below] at (5, -0.15) {\footnotesize duopoly};
- \end{tikzpicture}
- \caption{A stylized ``inverted-U relationship" between strength of competition and ``level of innovation".}
-\label{fig:inverted-U}
-\end{center}
-\end{figure}
-
-We consider varying levels of competition, from monopoly to \emph{temporary monopoly} (a duopoly with an early entrant), to a true duopoly in which both firms start at the same time.
-We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly.
-This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
-Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run.
-The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model).
-
-\asedit{Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
-This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.}
-
-Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantange".
-We further investigate which of the two is a stronger barrier to entry.
-\ascomment{and find what?
-and what's the new intuition?}
-
-\subsection{Discussion and related work}
-Our work is related to a longstanding economics literature on competition vs.
-innovation, \eg \cite{Schumpeter-42,barro2004economic,Aghion-QJE05}.
-While this literature focuses on R\&D costs of innovation, ``reputational costs" thereof seem new and specific to exploration.
-
-Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information).
-MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background.
-We consider a simple and well-understood MAB model with i.i.d.
-rewards \cite{bandits-ucb1}.
-We focus on a well-known distinction between the ``greedy" (exploitation-only) algorithm, "naive" MAB  algorithms that separate exploration and exploitation, and better MAB algorithms that combine the two.
-As a token example of ``near-optimal" bandit algorithms,
-we use a classic algorithm called ``Thompson Sampling", see \cite{TS-survey-FTML18} for background.
-
-The study of competition vs.
-exploration has been initiated in \cite{CompetingBandits-itcs16}.
-Their model differs from ours in two key respects.
-First, users do not see any signal about firms' past performance, and instead choose between firms according to the Bayesian-expected reward.
-Second, they vary the strength of competition using assumptions about (ir)rational consumer behavior, whereas we use early entry.
-Their results are purely theoretical; their model is amenable to proofs but not to numerical experiments.
-
-The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
-    \eg \cite{Kremer-JPE14,Frazier-ec14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
-dynamic auctions
-    (see \cite{DynAuctions-survey10} for background),
-online ad auctions, \eg
-    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm,Amin-auctions-nips13},
-and human computation
-    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
-
-\OMIT{ %%%%%%
-The interplay between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \cite{che2017recommender}, \cite{kremer2014implementing}, \cite{mansour2015bayesian}.
-The strategic experimentation literature in economics, such as \cite{bolton1999strategic} and \cite{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.
-} %%%%%%%%
-
-Our setting is also closely related to the "dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
-
-\section{Model}\label{sec:model}
-
-\textbf{Overview} There are two firms and $T+2k$ agents where $k$ is the warm start, or the agents that each firm gets for free at the beginning of the game and $T$ is the number of rounds in the game.
-The timing of events is as follows:
-\begin{enumerate}
-\item At $t = 0$, the firms simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
-\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a ``warm" start.
-The algorithm that the firm commits to makes $k$ rounds of progress and uses this information to initialize its information set.
-Additionally, the reputation score used by the agents is initialized using the rewards from these $k$ rounds.
-\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses the firm with the higher reputation at the time.
-\item The firm that is chosen selects an action $a_{t} \in A$ from a set of actions that is fixed across firms and rounds.
-\item Both the agent and the firm observe the reward $r_t \in [0, 1]$ from the action.
-The agent reports this reward to the firm and the future agents and the reputation score for the chosen firm is updated.
-\item Repeat 2-4 for $T$ rounds.
-\end{enumerate}
-
-Generally, the rewards are independent and identically distributed with a common prior.
-For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors.
-Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require a prior to operate, such as Thompson Sampling, we use a "fake" prior of $Beta(\alpha=1,\beta=1)$}.
-We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.\swcomment{this seems misleading and should not be here.
-just say that each action has bernouli rewards with unknown means}
-
-\noindent \textbf{Agents} We suppose that agents are homogenous, myopic, and non-strategic.
-The utility function for the agents is simply to maximize their reward in the one period in which they are alive.
-We suppose that agents do not attempt to manipulate the strategy of the firm, do not take the strategies of the firm into account when choosing between the firms, and do not condition their behavior on calendar time.
-In our model, each agent uses the average reward of past agents as a proxy for their expected utility.
-For simplicity, the reputation score, $R_{jt}$ is defined as a sliding window average\footnote{Another natural formulation would be to have exponential discounting of the past rewards.
-We believe that the results should be qualitatively similar to those presented here as in both formulations the more recent past matters more than the distant past.}.
-\begin{center}
-$R_{jt} = \frac{1}{M} \sum\limits_{i=1}^{M} r_{t_j-i}$
-\end{center}
-
-Note $t_j$ is the \textit{local} time of the firm, not the global time, so the reputation score is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.\swcomment{define ``calendar time'' and ``local time'' formally...
-don't think the current reputation defn makes sense} The warm start of $k$ rounds allows this reputation score to be well-defined once the "competition game" begins.
-In our model the agents deterministically choose the firm with the higher reputation score and ties are broken uniformly at random.
-
-\noindent \textbf{Firms} We suppose that the firms simply care about maximizing their expected market share (i.e.
-maximizing the number of $T$ agents who select them).
-We model the "competition game" between firms as a simultaneous move game where both firms commit to a learning algorithm at $t = 0$.
-
-\noindent \textbf{MAB algorithms} We suppose that firms commit to a learning algorithm from a fixed set of algorithms $\mathcal{A}$.
-We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
-\begin{enumerate}
-\item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
-We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward %\cite{Shipra-colt12}.
-\item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
-We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $\DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability.
-For our experiments we keep $\epsilon = 0.05$ fixed.
-\item Greedy / myopic algorithms that engage in no purposeful exploration and take the best short-sighted action.
-We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
-\end{enumerate}
-
-In the standard multi-armed bandit problem it is known that $\TS > \DEG > DG$ in terms of maximizing cumulative reward over a sufficiently large time horizon.
-The primary question that we want to understand is when, in competition, are the firms incentivized to adopt the "better" algorithms?
-
-\noindent \textbf{Incumbent} In some of our experiments we modify our model so that one firm enters the market $X$ number of rounds before the other.
-We refer to the firm that enters before as the "incumbent" and the firm that enters after $X$ rounds as the "entrant." In the $X$ rounds before the "entrant" enters the incumbent is a temporary monopolist as the agents that arrive in these $X$ rounds are forced to select the incumbent since it is the only firm in the market and there is no outside option.
-We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$.
-We have that both the incumbent and entrant commit to a learning algorithm \textit{before} either firm receives any agent.
-After the $X$ rounds, both firms still receive the $k$ warm start agents.
-
-\section{Simulation Details}\label{section:3}
-
-We evaluate the consequences of our model via simulation.\swcomment{In the process of removing ``prior'' throughout}
-
-\textbf{Bandit Instances} We look at bandit instances drawn from three different bandit \swedit{distributions} that capture different types of learning problems.
-Recall that we only consider Bernoulli rewards, which are fully parameterized by the means.
-For these experiments we fix the number of arms we consider, $K = 10$.
-\begin{enumerate}
-\item Needle In Haystack - $K-1$ arms with mean 0.5 and 1 arm with 0.7.
-\item Uniform - the $K$ arms have means drawn uniformly at random from $[0.25, 0.75]$
-\item "Heavy Tail" - the $K$ arms have means drawn from $Beta(\alpha=0.6, \beta = 0.6)$.
-With this prior it was likely to have means at the "extremes" or means that were close to 0 as well as means that were close to 1.
-\end{enumerate}
-
-\noindent \textbf{Simulation of Competition Game} Unless otherwise noted, all of the reported results utilize the same set of randomly drawn bandit instances and realizations from the prior.
-Namely, for each bandit prior we draw $N = 1000$ bandit instances.
-For each of these instances we run simulations of our model for varying values of $k$ and $X$.
-We take the maximum values of $k$ and $X$ for the simulations we run, $k_{max}$ and $X_{max}$ respectively, and compute a realization table of dimension $(T+k_{max}+X_{max}) \times K$.
-
-This realization table, as well as fixing the random seed for the same bandit instance and realization table across experiments, ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
-In the competition game we draw from the $T \times K$ portion of the table, so that if two different algorithms picks arm $a$ at time $t$, they get the same $[a, t]$ realization in the table.
-This setup also ensures that in the warm start period, increasing the warm start from $k$ to $k + 10$ results in the same behavior in the first $k$ rounds.
-
-For the simulations we fix the sliding window size $M = 100$.
-Low values of $M$ induced too much random noise into the results and we found that increasing $M$ to be larger than $100$ did not make a substantial qualitative difference so we fix this value.
-
-\section{Performance in Isolation}\label{section:4}
-
-We evaluate the performance of the different learning algorithms on the bandit priors we consider in isolation.
-There are two reasons why this is important to consider.
-First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $\TS > \DEG > DG$ for sufficiently large $t$.
-Second, we want to better understand what statistics of the priors we can look at in isolation in order to help us predict and understand what to expect in the competition game.
-
-\begin{figure}
-\caption{Mean Reputation Trajectories in Isolation}
-\includegraphics[scale=0.35]{"figures/nih_iso_mean"}
-\label{prelim_means}
-\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
-The shaded area display 95\% confidence intervals.}}
-\end{figure}
-
-Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior \footnote{The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the supplementary material.}.
-
-\begin{figure}[H]
-\caption{Reputation Distribution}
-\includegraphics[scale=0.35]{figures/rep_distribution_nih}
-\label{rep_dist_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
-\end{figure}
-
-\begin{finding}
-\textit{The mean performance of an algorithm is not a sufficient statistic for understanding the performance of an algorithm in competition.
-Looking at the distribution of reputation difference between algorithms we find that it tend is skewed to the right.
-Additionally we provide a counterexample where the mean performance does not predict the result of the competition game.}
-\end{finding}
-
-Is looking at the mean performance of between two algorithms sufficient for understanding what happens in competition?
-To answer this, we first look at the entire reputation distribution at $t = 500$ in Figure \ref{rep_dist_nih}.
-We see that the ``naive" algorithms $DG$ and $\DEG$ have a bi-modal reputation distribution whereas $\TS$ does not.
-The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not.
-If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds.
-However, if it does not then it will get stuck on a bad arm and lose to $\TS$ or $\DEG$.
-In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \cite{CompetingBandits-itcs16}}.
-
-To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $\TS$ and $DG$.
-This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large.
-Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution.
-There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the \textit{relative reputation proportion}.
-
-\begin{figure}[H]
-\caption{Reputation Difference Distribution}
-\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
-\label{ts_dg_rep_diff_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $DG$ across $t$}}
-\end{figure}
-
-\begin{definition}
-\textit{Relative Reputation Proportion} - the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$
-\end{definition}
-
-\begin{finding}
-\textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $\TS$ doing worse than $DG$ for small time horizons.
-We observe this under the Uniform and Heavy Tail priors.
-However, it is not observed under the Needle In Haystack prior}
-\end{finding}
-
-The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $\DEG >DG$ or $\TS > DG$ according to the mean reputation but $DG > \DEG$ or $DG > TS$ according to the relative reputation proportion.
-In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $\DEG > DG$ according to the mean reputation but $DG > \DEG$ according to the relative reputation proportion.
-Additionally, these results carry over to the competition game}.
-
-\begin{figure}[ht]
-\caption{Relative Reputation Plots}
-\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
-\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
-\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq DG$.
-The shaded area display 95\% confidence intervals.}}
-\label{relative_rep_plots}
-
-\end{figure}
-
-Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $DG$ on the Uniform and Needle In Haystack prior.
-For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $\TS > DG$.
-The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information.
-However, eventually the information acquired from purposeful exploration in the early rounds allows $\TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm.
-The early exploration leads to what we define as a \textit{relative reputation cost} and the eventual gain in reputation leads to what we define as a \textit{relative reputation benefit}.
-
-\begin{definition}
-\textit{Relative reputation cost and benefit} - the relative reputation loss an algorithm incurs from purposeful exploration compared to the greedy alternative.
-Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
-We call the priors where there is an early costly period followed by a later benefit period relative reputation costly priors.
-\end{definition}
-Is exploration always costly?
-Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $\TS$ always does relatively better than $DG$.
-There are two contributing factors to this.
-First, $\TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration.
-Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
-Thus, when $\TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
-However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
-Thus, only the Uniform and Heavy Tail priors are relative reputation costly priors.
-
-\section{Competition vs Better Algorithms: Inverted-U}\label{section:5}
-
-In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \cite{Aghion-QJE05}.
-We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly.
-
-\begin{finding}
-\textit{Under the priors that are relative reputation costly and with low warm start, the optimal strategies in the competition game are:
-\begin{center}
-\textbf{Permanent Monopoly} - $DG$ \\
-\textbf{Temporary Monopoly} - $\TS$ \\
-\textbf{Permanent Duopoly} - $DG$ \\
-\end{center}
-The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
-}
-\end{finding}
-
-We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition.
-We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index).
-In our model we have no prices and no costs and so we interpret increases in "market power" as decreases in the number of rounds where a firm has to compete for agents.
-
-We utilize the same instances and realizations as in the section 4 and simulate the model described in section 2.
-We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the expected payoffs associated with each pair of algorithms.
-Unless otherwise noted, all the results are reported at $t = 2000$.
-
-\textbf{Permanent Monopoly} Under permanent monopoly only a single firm exists in the market for every period.
-Since, in our formulation, the firm only gets utility from a larger market share it is indifferent between each algorithm in $\mathcal{A}$ since, regardless of the algorithm it deploys, it will get the entire market.
-If we suppose that deploying the "better" algorithms has even an $\epsilon > 0$ cost, then the firm would choose to deploy $DG$.
-
-\textbf{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously.
-What algorithms are firms incentivized to deploy?
-We report the average market share taken over the $N$ simulations for a set of exogenous values of $k$.
-Additionally, we report the mean and median of an additional quantity which we define as the effective end of game.
-
-\begin{definition}
-\textit{Effective End of Game (EEOG)} - the last round, $t$, in a simulation where the agent alive at $t-1$ and the agent alive at $t$ choose different firms.
-\end{definition}
-
-Since both firms enter at the same time the game is symmetric and we only need to compute the payoffs of three pairs of strategies.
-We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.
-
-\begin{finding}
-\textit{For low warm starts, when exploration is relative reputation costly $DG >TS$ in competition, whereas when exploration is not relative reputation costly $\TS > DG$.
-For sufficiently large warm starts, $\TS$ is dominant across all priors.}
-\end{finding}
-
-Table \ref{sim_unif} shows the results under the Uniform prior.
-For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $\TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}.
-Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $k$ as fixing the starting point on the relative reputation plots.
-The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$.
-The intuition for the result in competition aligns with the fact that, for Uniform and Heavy Tail, $\TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $\TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $\TS$.
-However, if we move the warm start sufficiently high so that $\TS$ moves to the relative reputation benefit regime, $\TS > DG$.
-
-However, we do not always see that $DG > TS$ under some warm starts.
-Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $\TS > DG$.
-This is precisely due to the fact that, under the Needle In Haystack, $\TS$ is not relative reputation costly compared to $DG$ and thus we see that $\TS > DG$.
-
-Combining these results with the fact that the EEOG values are relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right, similar to the distribution of reputation differences}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.
-
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Thu Aug 16 13:13:01 2018
-\begin{table}[ht]
-\centering
-\caption{Duopoly Experiment Uniform}
-\begin{tabular}{rlll}
-  \hline
- & k = 20 & k = 250 & k = 500 \\
-  \hline
-TS vs DG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.02\\ eeog \\ avg: 800\\ med: 754} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 910\\ med: 906.5} \\
-  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 180\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.02\\ eeog \\ avg: 810\\ med: 734} & \makecell{\textbf{0.55} $\pm$0.02\\ eeog \\ avg: 970\\ med: 987} \\
-  DG vs DEG & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 470\\ med: 57.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1088} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1142} \\
-   \hline
-\end{tabular}
-\label{sim_unif}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\caption{Duopoly Experiment Needle In Haystack}
-\begin{tabular}{rlll}
-  \hline
- & k = 20 & k = 250 & k = 500 \\
-  \hline
-TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 200\\ med: 27} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 370\\ med: 0} & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 580\\ med: 121.5} \\
-  TS vs DEG & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 150\\ med: 14} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 460\\ med: 78.5} & \makecell{\textbf{0.56} $\pm$0.02\\ eeog \\ avg: 740\\ med: 627.5} \\
-  DG vs DEG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 340\\ med: 128.5} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 650\\ med: 408} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 690\\ med: 466.5} \\
-   \hline
-\end{tabular}
-\label{sim_nih}
-\caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95 \% confidence band.
-For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
-The next line contain the average and median effective end of game for this set of simulations.}}
-\end{table}
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Thu Aug 16 13:13:00 2018
-
-\textbf{Temporary Monopoly} We now consider asymmetries in the timing of entry so that one firm enters the market before the other and serves as a monopolist in the periods until the other firm enters.
-In terms of our model, this corresponds to varying the value of $X$.
-For this section we report results fixing $k = 20$ and focus on the priors where we observed that, for this warm start, $DG > TS$ \footnote{However, we report the results for the Needle In Haystack prior in the supplementary material.
-The results are unsurprising, where $\TS$ is the dominant strategy for both the entrant and incumbent.}.
-It seems clear that the temporary monopoly will get a larger market share regardless of what algorithm it plays, but is the incumbent incentivized to commit to $\TS$?
-
-\begin{finding}
-\textit{Counterintuitively, allowing one firm to be a temporary monopolist for a sufficiently long time will induce that firm to commit to the better learning algorithm, $\TS$, even if, under the permanent duopoly, they would commit to $DG$}
-\end{finding}
-
-In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $DG$ is the dominant strategy for the entrant on the Heavy Tail prior \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.
-However, while $DG > TS$ for the entrant, $DG$ does roughly the same as $\DEG$ on the Uniform prior.
-The results are in the supplementary material.}.
-In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across priors.
-
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Wed Aug 15 19:03:20 2018
-\begin{table}[ht]
-\centering
-\caption{Temporary Monopoly Heavy Tail X = 200}
-\begin{tabular}{rlll}
-  \hline
- & TS & DEG &  DG \\
-  \hline
-TS & \makecell{\textbf{0.003} $\pm$0.003\\Var:0.002\\ES:100\%} & \makecell{\textbf{0.083} $\pm$0.02\\Var:0.07\\ES:97\%} & \makecell{\textbf{0.17} $\pm$0.02\\Var:0.1\\ES:95\%} \\
-  DEG & \makecell{\textbf{0.045} $\pm$0.01\\Var:0.03\\ES:92\%} & \makecell{\textbf{0.25} $\pm$0.02\\Var:0.1\\ES:75\%} & \makecell{\textbf{0.23} $\pm$0.02\\Var:0.1\\ES:78\%} \\
-   DG & \makecell{\textbf{0.12} $\pm$0.02\\Var:0.08\\ES:88\%} & \makecell{\textbf{0.36} $\pm$0.03\\Var:0.2\\ES:76\%} & \makecell{\textbf{0.3} $\pm$0.02\\Var:0.1\\ES:64\%} \\
-   \hline
-\end{tabular}
-\label{ht_incum}
-\caption*{\tiny{The columns represent the strategy of the incumbent and the rows represent the strategy of the entrant.
-The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval.
-The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one firm getting $> 90\%$ of the market.
-Note that smaller values in the table are better for the incumbent.
-Market shares are calculated as the fraction of users selecting a particular firm \textit{after} the entrant has already entered (i.e.
-the free rounds to firm 2 do not count towards the share)}}
-\end{table}
-
-Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same other conditions, $DG$ was preferred to $\TS$?
-The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
-This intuition is very similar to that observed in the previous section by increasing warm start.
-One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the "incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
-The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.
-As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
-Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
-
-\section{Data and Reputation as Barriers to Entry}\label{section:6}
-
-An alternative interpretation of the results in the previous section is that the "temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant.
-This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \cite{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry simply due to the fact that they were in the market before the entrant.
-Which plays a bigger role in preventing the entrant from being able to establish market share?
-We run two additional experiments, modifying the previous temporary monopoly experiment so that in one experiment the reputation of the incumbent is artificially erased (Reputation Erased Experiment) and another in which the information gained by the incumbent is artificially erased so that the posterior is reset to the prior (Information Erased Experiment).
-
-\begin{table}[ht]
-\centering
-\caption{Reputation Erased Experiment, Heavy Tail}
-\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
-  \hline
- & TS & DEG &  DG \\
-  \hline
-TS & \makecell{\textbf{ 0.016 } $\pm$ 0.0075 \\Var: 0.01 \\ ES: 100 \%} & \makecell{\textbf{ 0.13 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.2 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 96 \%} \\
-  DEG & \makecell{\textbf{ 0.068 } $\pm$ 0.013 \\Var: 0.05 \\ ES: 93 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 75 \%} & \makecell{\textbf{ 0.26 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 80 \%} \\
-   DG & \makecell{\textbf{ 0.15 } $\pm$ 0.019 \\Var: 0.1 \\ ES: 87 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.028 \\Var: 0.2 \\ ES: 80 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 67 \%} \\
-   \hline
-   \label{rep_erase}
-\end{tabular}
-\end{table}
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Sun Jul 22 14:23:06 2018
-\begin{table}[ht]
-\centering
-\caption{Information Erased Experiment, Heavy Tail}
-\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
-  \hline
- & TS & DEG &  DG \\
-  \hline
-TS & \makecell{\textbf{ 0.024 } $\pm$ 0.0094 \\Var: 0.02 \\ ES: 100 \%} & \makecell{\textbf{ 0.16 } $\pm$ 0.022 \\Var: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.22 } $\pm$ 0.025 \\Var: 0.2 \\ ES: 95 \%} \\
-  DEG & \makecell{\textbf{ 0.24 } $\pm$ 0.025 \\Var: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 72 \%} & \makecell{\textbf{ 0.27 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 76 \%} \\
-   DG & \makecell{\textbf{ 0.33 } $\pm$ 0.028 \\Var: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.026 \\Var: 0.2 \\ ES: 74 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.023 \\Var: 0.1 \\ ES: 58 \%} \\
-   \hline
-   \label{info_erase}
-\end{tabular}
-\end{table}
-
-\begin{finding}
-\textit{When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation.
-However, when the incumbent plays $\DEG$ or $DG$ there is no substantial difference between the two.}
-\end{finding}
-
-Tables \ref{rep_erase} and \ref{info_erase} display the results of these experiments.
-Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
-The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation when the incumbent plays $\TS$.
-One possible intuition for this is that, for Heavy Tail, the information acquired after the $X$ rounds under $\TS$ is sufficient so that it becomes relatively easier for $\TS$ to recover the reputation loss compared to the information loss, but since $\DEG$ and $DG$ learn slower and they've acquired less information over the $X$ rounds relative to $\TS$ there is less of an informational advantage.
-
-Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
-We find that the best responses remain the same for the Heavy Tail prior, but for the Uniform prior they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.
+\subfile{content/barriers}
 

### 2018-08-30 paper/content/barriers.tex
break up the different sections into different files

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Data and Reputation as Barriers to Entry}\label{section:6}

An alternative interpretation of the results in the previous section is that the "temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant. This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \cite{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry simply due to the fact that they were in the market before the entrant. Which plays a bigger role in preventing the entrant from being able to establish market share? We run two additional experiments, modifying the previous temporary monopoly experiment so that in one experiment the reputation of the incumbent is artificially erased (Reputation Erased Experiment) and another in which the information gained by the incumbent is artificially erased so that the posterior is reset to the prior (Information Erased Experiment).

\begin{table}[ht]
\centering
\caption{Reputation Erased Experiment, Heavy Tail}
\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
  \hline
 & TS & DEG &  DG \\
  \hline
TS & \makecell{\textbf{ 0.016 } $\pm$ 0.0075 \\Var: 0.01 \\ ES: 100 \%} & \makecell{\textbf{ 0.13 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.2 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 96 \%} \\
  DEG & \makecell{\textbf{ 0.068 } $\pm$ 0.013 \\Var: 0.05 \\ ES: 93 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 75 \%} & \makecell{\textbf{ 0.26 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 80 \%} \\
   DG & \makecell{\textbf{ 0.15 } $\pm$ 0.019 \\Var: 0.1 \\ ES: 87 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.028 \\Var: 0.2 \\ ES: 80 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 67 \%} \\
   \hline
   \label{rep_erase}
\end{tabular}
\end{table}
% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Sun Jul 22 14:23:06 2018
\begin{table}[ht]
\centering
\caption{Information Erased Experiment, Heavy Tail}
\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
  \hline
 & TS & DEG &  DG \\
  \hline
TS & \makecell{\textbf{ 0.024 } $\pm$ 0.0094 \\Var: 0.02 \\ ES: 100 \%} & \makecell{\textbf{ 0.16 } $\pm$ 0.022 \\Var: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.22 } $\pm$ 0.025 \\Var: 0.2 \\ ES: 95 \%} \\
  DEG & \makecell{\textbf{ 0.24 } $\pm$ 0.025 \\Var: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 72 \%} & \makecell{\textbf{ 0.27 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 76 \%} \\
   DG & \makecell{\textbf{ 0.33 } $\pm$ 0.028 \\Var: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.026 \\Var: 0.2 \\ ES: 74 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.023 \\Var: 0.1 \\ ES: 58 \%} \\
   \hline
   \label{info_erase}
\end{tabular}
\end{table}

\begin{finding}
\textit{When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation. However, when the incumbent plays $\DEG$ or $DG$ there is no substantial difference between the two.}
\end{finding}

Tables \ref{rep_erase} and \ref{info_erase} display the results of these experiments. Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market. The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation when the incumbent plays $\TS$. One possible intuition for this is that, for Heavy Tail, the information acquired after the $X$ rounds under $\TS$ is sufficient so that it becomes relatively easier for $\TS$ to recover the reputation loss compared to the information loss, but since $\DEG$ and $DG$ learn slower and they've acquired less information over the $X$ rounds relative to $\TS$ there is less of an informational advantage.

Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained. We find that the best responses remain the same for the Heavy Tail prior, but for the Uniform prior they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.

\end{document}

### 2018-08-30 paper/content/introduction.tex
break up the different sections into different files

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Introduction}\label{section:1}

Many modern online platforms simultaneously compete for users as well as learn from the users they manage to attract. This creates a tradeoff between \textit{exploration} and \textit{competition}: firms experiment with potentially sub-optimal options for the sake of gaining information to make better decisions tomorrow, while they need to incentivize consumers to select them over their competitors today. For instance, Google Search and Bing compete for users in the search engine market yet at the same time need to experiment with their search and ranking algorithms to learn what works best.

Platforms routinely deploy \asedit{A/B tests}, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty. While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition, \asedit{even putting aside the deployment/maintenance costs}. In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term. This would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.

We ask whether competition incentivizes adoption of "better" algorithms for exploration. We investigate this issue via extensive numerical experiments in a stylized duopoly model. In our model, two firms compete for users, and simultaneously learn from them. Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm. Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window. \asedit{Each firm's objective} is to maximize its market share ---the fraction of users coming to them.

\tikzstyle{level 1}=[level distance=3.5cm, sibling distance=4.0cm]
\tikzstyle{level 2}=[level distance=3.5cm, sibling distance=2cm]
\tikzstyle{below} = [align=center]

\begin{figure}
\begin{center}
\begin{tikzpicture}
      \draw[->] (-.5,0) -- (6,0) node[right] {};
      \draw[->] (0,-.5) -- (0,2) node[above] {Better algorithms};
      \draw[scale=0.6,domain=0.7:8.3,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{3.4 - 0.2 * (\x - 4.5)^2});
     \node[below] at (0.7, -0.15) {\footnotesize monopoly};
     \node[below] at (3, 0) {\footnotesize temporary \\ monopoly};
     \node[below] at (5, -0.15) {\footnotesize duopoly};
 \end{tikzpicture}
 \caption{A stylized ``inverted-U relationship" between strength of competition and ``level of innovation".}
\label{fig:inverted-U}
\end{center}
\end{figure}

We consider varying levels of competition, from monopoly to \emph{temporary monopoly} (a duopoly with an early entrant), to a true duopoly in which both firms start at the same time. We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly. This algorithm also prevails under monopoly, simply because it tends to be easier to deploy. Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run. The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model).

\asedit{Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}). This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.}

Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantange". We further investigate which of the two is a stronger barrier to entry. \ascomment{and find what? and what's the new intuition?}

\subsection{Discussion and related work}
Our work is related to a longstanding economics literature on competition vs. innovation, \eg \cite{Schumpeter-42,barro2004economic,Aghion-QJE05}. While this literature focuses on R\&D costs of innovation, ``reputational costs" thereof seem new and specific to exploration.

Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information). MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background. We consider a simple and well-understood MAB model with i.i.d. rewards \cite{bandits-ucb1}. We focus on a well-known distinction between the ``greedy" (exploitation-only) algorithm, "naive" MAB  algorithms that separate exploration and exploitation, and better MAB algorithms that combine the two. As a token example of ``near-optimal" bandit algorithms, 
we use a classic algorithm called ``Thompson Sampling", see \cite{TS-survey-FTML18} for background.

The study of competition vs. exploration has been initiated in \cite{CompetingBandits-itcs16}. Their model differs from ours in two key respects. First, users do not see any signal about firms' past performance, and instead choose between firms according to the Bayesian-expected reward. Second, they vary the strength of competition using assumptions about (ir)rational consumer behavior, whereas we use early entry. Their results are purely theoretical; their model is amenable to proofs but not to numerical experiments.

The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
    \eg \cite{Kremer-JPE14,Frazier-ec14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
dynamic auctions
    (see \cite{DynAuctions-survey10} for background),
online ad auctions, \eg
    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm,Amin-auctions-nips13},
and human computation
    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.

\OMIT{ %%%%%%
The interplay between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \cite{che2017recommender}, \cite{kremer2014implementing}, \cite{mansour2015bayesian}. The strategic experimentation literature in economics, such as \cite{bolton1999strategic} and \cite{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.
} %%%%%%%%

Our setting is also closely related to the "dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.

\end{document}

### 2018-08-30 paper/content/inverted_u.tex
break up the different sections into different files

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Competition vs Better Algorithms: Inverted-U}\label{section:5}

In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \cite{Aghion-QJE05}. We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly.

\begin{finding}
\textit{Under the priors that are relative reputation costly and with low warm start, the optimal strategies in the competition game are:
\begin{center}
\textbf{Permanent Monopoly} - $DG$ \\
\textbf{Temporary Monopoly} - $\TS$ \\
\textbf{Permanent Duopoly} - $DG$ \\
\end{center}
The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
}
\end{finding}

We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition. We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index). In our model we have no prices and no costs and so we interpret increases in "market power" as decreases in the number of rounds where a firm has to compete for agents.

We utilize the same instances and realizations as in the section 4 and simulate the model described in section 2. We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the expected payoffs associated with each pair of algorithms. Unless otherwise noted, all the results are reported at $t = 2000$.

\textbf{Permanent Monopoly} Under permanent monopoly only a single firm exists in the market for every period. Since, in our formulation, the firm only gets utility from a larger market share it is indifferent between each algorithm in $\mathcal{A}$ since, regardless of the algorithm it deploys, it will get the entire market. If we suppose that deploying the "better" algorithms has even an $\epsilon > 0$ cost, then the firm would choose to deploy $DG$.

\textbf{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously. What algorithms are firms incentivized to deploy? We report the average market share taken over the $N$ simulations for a set of exogenous values of $k$. Additionally, we report the mean and median of an additional quantity which we define as the effective end of game.

\begin{definition}
\textit{Effective End of Game (EEOG)} - the last round, $t$, in a simulation where the agent alive at $t-1$ and the agent alive at $t$ choose different firms.
\end{definition}

Since both firms enter at the same time the game is symmetric and we only need to compute the payoffs of three pairs of strategies. We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.

\begin{finding}
\textit{For low warm starts, when exploration is relative reputation costly $DG >TS$ in competition, whereas when exploration is not relative reputation costly $\TS > DG$. For sufficiently large warm starts, $\TS$ is dominant across all priors.}
\end{finding}

Table \ref{sim_unif} shows the results under the Uniform prior. For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $\TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}. Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $k$ as fixing the starting point on the relative reputation plots. The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$. The intuition for the result in competition aligns with the fact that, for Uniform and Heavy Tail, $\TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $\TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $\TS$. However, if we move the warm start sufficiently high so that $\TS$ moves to the relative reputation benefit regime, $\TS > DG$.

However, we do not always see that $DG > TS$ under some warm starts. Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $\TS > DG$. This is precisely due to the fact that, under the Needle In Haystack, $\TS$ is not relative reputation costly compared to $DG$ and thus we see that $\TS > DG$.

Combining these results with the fact that the EEOG values are relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right, similar to the distribution of reputation differences}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Aug 16 13:13:01 2018
\begin{table}[ht]
\centering
\caption{Duopoly Experiment Uniform}
\begin{tabular}{rlll}
  \hline
 & k = 20 & k = 250 & k = 500 \\
  \hline
TS vs DG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.02\\ eeog \\ avg: 800\\ med: 754} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 910\\ med: 906.5} \\
  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 180\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.02\\ eeog \\ avg: 810\\ med: 734} & \makecell{\textbf{0.55} $\pm$0.02\\ eeog \\ avg: 970\\ med: 987} \\
  DG vs DEG & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 470\\ med: 57.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1088} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1142} \\
   \hline
\end{tabular}
\label{sim_unif}
\end{table}

\begin{table}[ht]
\centering
\caption{Duopoly Experiment Needle In Haystack}
\begin{tabular}{rlll}
  \hline
 & k = 20 & k = 250 & k = 500 \\
  \hline
TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 200\\ med: 27} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 370\\ med: 0} & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 580\\ med: 121.5} \\
  TS vs DEG & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 150\\ med: 14} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 460\\ med: 78.5} & \makecell{\textbf{0.56} $\pm$0.02\\ eeog \\ avg: 740\\ med: 627.5} \\
  DG vs DEG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 340\\ med: 128.5} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 650\\ med: 408} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 690\\ med: 466.5} \\
   \hline
\end{tabular}
\label{sim_nih}
\caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95 \% confidence band. For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG. The next line contain the average and median effective end of game for this set of simulations.}}
\end{table}
% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Aug 16 13:13:00 2018

\textbf{Temporary Monopoly} We now consider asymmetries in the timing of entry so that one firm enters the market before the other and serves as a monopolist in the periods until the other firm enters. In terms of our model, this corresponds to varying the value of $X$. For this section we report results fixing $k = 20$ and focus on the priors where we observed that, for this warm start, $DG > TS$ \footnote{However, we report the results for the Needle In Haystack prior in the supplementary material. The results are unsurprising, where $\TS$ is the dominant strategy for both the entrant and incumbent.}. It seems clear that the temporary monopoly will get a larger market share regardless of what algorithm it plays, but is the incumbent incentivized to commit to $\TS$?

\begin{finding}
\textit{Counterintuitively, allowing one firm to be a temporary monopolist for a sufficiently long time will induce that firm to commit to the better learning algorithm, $\TS$, even if, under the permanent duopoly, they would commit to $DG$}
\end{finding}

In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $DG$ is the dominant strategy for the entrant on the Heavy Tail prior \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$. However, while $DG > TS$ for the entrant, $DG$ does roughly the same as $\DEG$ on the Uniform prior. The results are in the supplementary material.}. In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across priors.

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Wed Aug 15 19:03:20 2018
\begin{table}[ht]
\centering
\caption{Temporary Monopoly Heavy Tail X = 200}
\begin{tabular}{rlll}
  \hline
 & TS & DEG &  DG \\
  \hline
TS & \makecell{\textbf{0.003} $\pm$0.003\\Var:0.002\\ES:100\%} & \makecell{\textbf{0.083} $\pm$0.02\\Var:0.07\\ES:97\%} & \makecell{\textbf{0.17} $\pm$0.02\\Var:0.1\\ES:95\%} \\
  DEG & \makecell{\textbf{0.045} $\pm$0.01\\Var:0.03\\ES:92\%} & \makecell{\textbf{0.25} $\pm$0.02\\Var:0.1\\ES:75\%} & \makecell{\textbf{0.23} $\pm$0.02\\Var:0.1\\ES:78\%} \\
   DG & \makecell{\textbf{0.12} $\pm$0.02\\Var:0.08\\ES:88\%} & \makecell{\textbf{0.36} $\pm$0.03\\Var:0.2\\ES:76\%} & \makecell{\textbf{0.3} $\pm$0.02\\Var:0.1\\ES:64\%} \\
   \hline
\end{tabular}
\label{ht_incum}
\caption*{\tiny{The columns represent the strategy of the incumbent and the rows represent the strategy of the entrant. The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval. The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one firm getting $> 90\%$ of the market. Note that smaller values in the table are better for the incumbent. Market shares are calculated as the fraction of users selecting a particular firm \textit{after} the entrant has already entered (i.e. the free rounds to firm 2 do not count towards the share)}}
\end{table}

Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same other conditions, $DG$ was preferred to $\TS$? The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds. This intuition is very similar to that observed in the previous section by increasing warm start. One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the "incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation. The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition. As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration. Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.

\end{document}

### 2018-08-30 paper/content/model.tex
break up the different sections into different files

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Model}\label{sec:model}

\textbf{Overview} There are two firms and $T+2k$ agents where $k$ is the warm start, or the agents that each firm gets for free at the beginning of the game and $T$ is the number of rounds in the game. The timing of events is as follows:
\begin{enumerate}
\item At $t = 0$, the firms simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a ``warm" start. The algorithm that the firm commits to makes $k$ rounds of progress and uses this information to initialize its information set. Additionally, the reputation score used by the agents is initialized using the rewards from these $k$ rounds.
\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses the firm with the higher reputation at the time.
\item The firm that is chosen selects an action $a_{t} \in A$ from a set of actions that is fixed across firms and rounds.
\item Both the agent and the firm observe the reward $r_t \in [0, 1]$ from the action. The agent reports this reward to the firm and the future agents and the reputation score for the chosen firm is updated.
\item Repeat 2-4 for $T$ rounds.
\end{enumerate}

Generally, the rewards are independent and identically distributed with a common prior.  For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors. Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require a prior to operate, such as Thompson Sampling, we use a "fake" prior of $Beta(\alpha=1,\beta=1)$}. We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.\swcomment{this seems misleading and should not be here. just say that each action has bernouli rewards with unknown means}

\noindent \textbf{Agents} We suppose that agents are homogenous, myopic, and non-strategic. The utility function for the agents is simply to maximize their reward in the one period in which they are alive. We suppose that agents do not attempt to manipulate the strategy of the firm, do not take the strategies of the firm into account when choosing between the firms, and do not condition their behavior on calendar time. In our model, each agent uses the average reward of past agents as a proxy for their expected utility. For simplicity, the reputation score, $R_{jt}$ is defined as a sliding window average\footnote{Another natural formulation would be to have exponential discounting of the past rewards. We believe that the results should be qualitatively similar to those presented here as in both formulations the more recent past matters more than the distant past.}.
\begin{center}
$R_{jt} = \frac{1}{M} \sum\limits_{i=1}^{M} r_{t_j-i}$
\end{center}

Note $t_j$ is the \textit{local} time of the firm, not the global time, so the reputation score is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.\swcomment{define ``calendar time'' and ``local time'' formally... don't think the current reputation defn makes sense} The warm start of $k$ rounds allows this reputation score to be well-defined once the "competition game" begins. In our model the agents deterministically choose the firm with the higher reputation score and ties are broken uniformly at random.

\noindent \textbf{Firms} We suppose that the firms simply care about maximizing their expected market share (i.e. maximizing the number of $T$ agents who select them). We model the "competition game" between firms as a simultaneous move game where both firms commit to a learning algorithm at $t = 0$.

\noindent \textbf{MAB algorithms} We suppose that firms commit to a learning algorithm from a fixed set of algorithms $\mathcal{A}$. We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
\begin{enumerate}
\item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation. We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \cite{Shipra-colt12}.
\item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation. We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $\DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability. For our experiments we keep $\epsilon = 0.05$ fixed.
\item Greedy / myopic algorithms that engage in no purposeful exploration and take the best short-sighted action. We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
\end{enumerate}

In the standard multi-armed bandit problem it is known that $\TS > \DEG > DG$ in terms of maximizing cumulative reward over a sufficiently large time horizon. The primary question that we want to understand is when, in competition, are the firms incentivized to adopt the "better" algorithms?

\noindent \textbf{Incumbent} In some of our experiments we modify our model so that one firm enters the market $X$ number of rounds before the other. We refer to the firm that enters before as the "incumbent" and the firm that enters after $X$ rounds as the "entrant." In the $X$ rounds before the "entrant" enters the incumbent is a temporary monopolist as the agents that arrive in these $X$ rounds are forced to select the incumbent since it is the only firm in the market and there is no outside option. We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$. We have that both the incumbent and entrant commit to a learning algorithm \textit{before} either firm receives any agent. After the $X$ rounds, both firms still receive the $k$ warm start agents.

\end{document}

### 2018-08-30 paper/content/perf_in_iso.tex
break up the different sections into different files

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Performance in Isolation}\label{section:4}

We evaluate the performance of the different learning algorithms on the bandit priors we consider in isolation. There are two reasons why this is important to consider. First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $\TS > \DEG > DG$ for sufficiently large $t$. Second, we want to better understand what statistics of the priors we can look at in isolation in order to help us predict and understand what to expect in the competition game.

\begin{figure}
\caption{Mean Reputation Trajectories in Isolation}
\includegraphics[scale=0.35]{"figures/nih_iso_mean"}
\label{prelim_means}
\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs. The shaded area display 95\% confidence intervals.}}
\end{figure}

Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior \footnote{The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the supplementary material.}.

\begin{figure}[H]
\caption{Reputation Distribution}
\includegraphics[scale=0.35]{figures/rep_distribution_nih}
\label{rep_dist_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
\end{figure}

\begin{finding}
\textit{The mean performance of an algorithm is not a sufficient statistic for understanding the performance of an algorithm in competition. Looking at the distribution of reputation difference between algorithms we find that it tend is skewed to the right. Additionally we provide a counterexample where the mean performance does not predict the result of the competition game.}
\end{finding}

Is looking at the mean performance of between two algorithms sufficient for understanding what happens in competition? To answer this, we first look at the entire reputation distribution at $t = 500$ in Figure \ref{rep_dist_nih}. We see that the ``naive" algorithms $DG$ and $\DEG$ have a bi-modal reputation distribution whereas $\TS$ does not. The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not. If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds. However, if it does not then it will get stuck on a bad arm and lose to $\TS$ or $\DEG$. In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \cite{CompetingBandits-itcs16}}.

To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $\TS$ and $DG$. This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large. Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution. There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the \textit{relative reputation proportion}.

\begin{figure}[H]
\caption{Reputation Difference Distribution}
\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
\label{ts_dg_rep_diff_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $DG$ across $t$}}
\end{figure}

\begin{definition}
\textit{Relative Reputation Proportion} - the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$
\end{definition}

\begin{finding}
\textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $\TS$ doing worse than $DG$ for small time horizons. We observe this under the Uniform and Heavy Tail priors. However, it is not observed under the Needle In Haystack prior}
\end{finding}

The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $\DEG >DG$ or $\TS > DG$ according to the mean reputation but $DG > \DEG$ or $DG > TS$ according to the relative reputation proportion. In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $\DEG > DG$ according to the mean reputation but $DG > \DEG$ according to the relative reputation proportion. Additionally, these results carry over to the competition game}.

\begin{figure}[ht]
\caption{Relative Reputation Plots}
\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq DG$. The shaded area display 95\% confidence intervals.}}
\label{relative_rep_plots}

\end{figure}

Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $DG$ on the Uniform and Needle In Haystack prior. For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $\TS > DG$. The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information. However, eventually the information acquired from purposeful exploration in the early rounds allows $\TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm. The early exploration leads to what we define as a \textit{relative reputation cost} and the eventual gain in reputation leads to what we define as a \textit{relative reputation benefit}.

\begin{definition}
\textit{Relative reputation cost and benefit} - the relative reputation loss an algorithm incurs from purposeful exploration compared to the greedy alternative. Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm. We call the priors where there is an early costly period followed by a later benefit period relative reputation costly priors.
\end{definition}
Is exploration always costly? Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $\TS$ always does relatively better than $DG$. There are two contributing factors to this. First, $\TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration. Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same. Thus, when $\TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm. However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option. Thus, only the Uniform and Heavy Tail priors are relative reputation costly priors.

\end{document}

### 2018-08-30 paper/content/sim_details.tex
break up the different sections into different files

--- 2018-08-15 paper_v2/competing_bandits.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-% \PassOptionsToPackage{numbers, compress}{natbib}
-
-% ready for submission
-\usepackage{aaai}
-
-% to compile a preprint version, e.g., for submission to arXiv, add
-% add the [preprint] option:
-% \usepackage[preprint]{nips_2018}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2018}
-
-% to avoid loading the natbib package, add option nonatbib:
-% \usepackage[nonatbib]{nips_2018}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{amssymb}
-\usepackage{natbib}
-%% The amsthm package provides extended theorem environments
-\usepackage{amsthm}
-\usepackage{float}
-\usepackage{sgame, tikz} % Game theory packages
-\usepackage{caption} 
-\usepackage{algorithm,algpseudocode}
-\usepackage{makecell}
- \usepackage{multirow}
- \usepackage{graphicx}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\captionsetup{font=footnotesize}
-\bibliographystyle{aaai}
-
-% The \author macro works with any number of authors. There are two
-% commands used to separate the names and addresses of multiple
-% authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to
-% break the lines. Using \AND forces a line break at that point. So,
-% if LaTeX puts 3 of 4 authors names on the first line, and the last
-% on the second line, try using \AND instead of \And before the third
-% author name.
-
+\documentclass[../competing_bandits.tex]{subfiles}
 \begin{document}
 
-\title{Competing Bandits}
+\section{Simulation Details}\label{section:3}
 
-\author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
-Zhiwei Steven Wu\textsuperscript{4} \\,
-{\textsuperscript{1}Columbia University, Department of Economics}\\
-{\textsuperscript{2}Columbia University, Department of Computer Science}\\
-{\textsuperscript{3}Microsoft Research, New York, NY}\\
-{\textsuperscript{4}University of Minnesota - Twin Cities, Department of Computer Science}
-}
-\maketitle
+We evaluate the consequences of our model via simulation.\swcomment{In the process of removing ``prior'' throughout}
 
-\begin{abstract}
-...
-\end{abstract}
-
-\section{Introduction}
-\label{intro}
-
-Many modern online platforms simultaneously compete for users as well as learn from the users they manage to attract.
-For instance, Google Search and Bing compete for users in the search engine market yet at the same need to experiment with their search algorithms to learn what algorithms work the best.
-This creates a tradeoff between \textit{exploration} and \textit{competition} since firms need to experiment with potentially sub-optimal options for the benefit of gaining information to make better decisions tomorrow while at the same time firms need to incentivize consumers to select them over their competitors today.
-
-From a social welfare perspective it is better for firms to adopt ``smarter" learning algorithms and our main high-level question is asking under what levels of competition are firms incentivized to adopt ``better" learning algorithms.
-Our primary contribution is that in our model, even in the case of a duopoly, firms are not incentivized to adopt better learning algorithms but that allowing one firm to be a temporary monopolist will incentivize this firm to play a better learning algorithm.
-
-Unlike in the classic models in economics regarding the interplay between competition and innovation described in \citet{barro2004economic}, here we have no explicit R \& D cost.
-Rather, firms are not incentivized to experiment under duopoly because experimentation has an implicit relative \textit{reputational} cost compared to the myopic alternative and, in our model, consumers only care about the reputation of a firm.
-Thus, competition disincentivizes experimentation by firms precisely because competition forces firms to take actions in order to incentivize consumers to select them over their competitors and, as a result, by allowing a firm to be a temporary monopolist the firm can experiment without worrying about the reputational consequences.
-Our findings give an alternative intuition to the empirically documented inverse-U relationship between competition and innovation discussed in \citet{aghion2005competition}.
-
-\textbf{Related Work} A major underlying component of our model is that firms face a multi-armed bandit (MAB) problem.
-Multi-armed bandits (MAB) are a tractable abstraction for the tradeoff between exploration and exploitation.
-MAB problems have been studied for many decades (see \citet{bubeck2012regret} for an overview).
-Most relevant to this paper is the thread that focuses on designing ``smart" and tractable algorithms that combine exploration and exploitation and ``naive" algorithms that separate exploration and exploitation (see \citet{slivkins2017bandits}).
-
-The three-way tradeoff between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}.
-The strategic experimentation literature in economics, such as \citealt*{bolton1999strategic} and \citealt*{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.
-
-The relationship between competition and innovation has been heavily studied in industrial organization \citep{tirole1988theory} and endogenous growth theory \citep{aghion2005competition, barro2004economic}, dating back to \citet{schumpeter2010capitalism}.
-
-The most closely related work to this paper is \citealt*{mansour2017competing}.
-Our motivating questions are the same as \citet{mansour2017competing} though while \citet{mansour2017competing} use the rationality of consumers as their primary knob of competitiveness, we consider differences in the number of firms in the market as our primary knob.
-Additionally, while in \citet{mansour2017competing} agents select firms based on the Bayesian expected reward of each principal in our model the agents select firms based on a frequentist reputation score that is derived from the signals sent from past agents.
-This allows the model to become tractable for the purposes of evaluating it via simulation and this is the primary method of analysis that we employ.
-
-\section{Model}
-\label{model}
-
-\textbf{Overview} There are two principals and $T+2k$ agents where $k$ is the warm start, or the agents that each principal gets for free at the beginning of the game.
-The timing of events is as follows:
-\begin{enumerate}
-\item At $t = 0$, the principals simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
-\item Still at $t=0$ we suppose that each principal gets $k$ agents as a ``warm" start.
-\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses among 2 principals, given a reputation score for both principals.
-\item The principal that is chosen selects an action $a_{t} \in A$ from a set of fixed actions across principals and rounds.
-\item Both the agent and the principal observe the reward $r_t \in [0, 1]$ from the action.
-The agent reports this reward to the principal and the future agents and the reputation score for the chosen principal is updated.
-\item Repeat 2-4 for $T$ rounds.
-\end{enumerate}
-
-Generally, the rewards are i.i.d.
-with a common prior.
-For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors.
-Each principal faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a ``fake" prior of $Beta(\alpha=1,\beta=1)$}.
-We assume that the principals commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.
-
-\noindent \textbf{Agents} We suppose that agents are myopic and non-strategic.
-The utility function for the agents is simply to maximize their reward in the one period in which they are alive.
-We suppose that agents do not attempt to manipulate the strategy of the principal nor do they take the strategies of the principal into account when choosing between the principals.
-In our model, each agent uses the average reward of past agents as a proxy for their expected utility.
-For simplicity, the reputation score, $R_jt$ is defined as a sliding window average.
-\begin{center}
-$R_{jt} = \frac{1}{M} \sum\limits_{i=1}^{M} r_{t_j-i}$
-\end{center}
-
-Note that the reputation score is in terms of \textit{local} and not global time so that it is the sliding window average of the last $M$ times that principal $j$ was selected by the agents.
-The warm start of $k$ rounds allows this reputation score to be well-defined once the ``competition game" begins.
-In our model the agents deterministically choose the principal with the higher reputation score and ties are broken uniformly at random.
-
-\noindent \textbf{Principals} We suppose that the principals simply care about maximizing market share (i.e.
-maximizing the number of $T$ agents who select them).
-We model the ``competition" game between principals as a simultaneous move game where both principals commit to a learning algorithm.
-
-\noindent \textbf{MAB algorithms} We suppose that principals commit to a learning algorithm from $\mathcal{A}$.
-We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
-\begin{enumerate}
-\item ``Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
-We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is ``optimal" in the sense of having the highest mean reward \citep{agrawal2012analysis}.
-\item ``Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
-We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability.
-For our experiments we keep $\epsilon = 0.05$ fixed.
-\item Greedy / myopic algorithms that engage in no purposeful exploration and take the best short-sighted action.
-We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
-\end{enumerate}
-
-It is known that $TS > DEG > DG$ in the standard multi-armed bandit problem of minimizing total regret.
-The primary question that we want to understand is when, in competition, are the principals incentivized to adopt the ``better" algorithms?
-
-\noindent \textbf{Incumbent} In some of our experiments we modify our model so that one principal enters the market $X$ number of rounds before the other.
-We refer to the principal that enters before as the ``incumbent" and the principal that enters after $X$ rounds as the ``entrant." The agents that arrive in these $X$ rounds are forced to select the only firm in the market as there is no outside option.
-We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$.
-We have that both the incumbent and entrant commit to a learning algorithm \textit{before} either firm receives any agent.
-After the $X$ rounds, both firms still receive the $k$ warm start agents.
-
-\section{Simulation Details}
-\label{sim_details}
-
-We evaluate the consequences of our model via simulation.
-
-\textbf{Bandit Priors} We look at bandit instances drawn from three different bandit ``priors" that are different types of learning problems.
-Recall that we only consider Bernoulli-distributed arms and so we draw the true means of the arms from these priors.
+\textbf{Bandit Instances} We look at bandit instances drawn from three different bandit \swedit{distributions} that capture different types of learning problems.
+Recall that we only consider Bernoulli rewards, which are fully parameterized by the means.
 For these experiments we fix the number of arms we consider, $K = 10$.
@@
 
-\section{Performance in Isolation}
-\label{iso_section}
-
-We evaluate the performance of the different learning algorithms on the bandit priors we consider in isolation.
-There are two reasons why this is important to consider.
-First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where $TS > DEG > DG$.
-Second, we want to better understand what statistics of the instance we can look at in isolation in order to help us predict and understand what to expect in the competition game.
-
-\begin{figure}
-\caption{Mean Reputation Trajectories in Isolation}
-\includegraphics[scale=0.25]{"figures/nih_iso_mean"}
-\label{prelim_means}
-\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
-The shaded area display 95\% confidence intervals.}}
-\end{figure}
-
-Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior.
-The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the GitHub appendix.
-
-In the ``competition" game the agents decide between the principals using the \textit{relative} reputation between the two principals.
-Namely, if principal $A$ has a higher reputation than principal $B$ agents will select principal $A$.
-A natural question to ask is whether looking at the mean performance is sufficient for understanding performance in the competition game.
-Figure \ref{rep_dist_nih} displays the reputation distribution at $t = 500$ for the same prior we reported the mean plot for previously.
-We see that the ``naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not.
-The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not.
-If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds so that the weak law of large numbers comes into play (our $M$ is set sufficiently high to ensure this most of the time).
-However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$.
-In these cases its reputation may be substantially worse but for the competition game the difference between the reputation does not matter but only the binary comparison between them \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, considering the SoftMax decision rule in \citet{mansour2017competing}}.
-
-\begin{figure}
-\caption{Reputation Distribution}
-\includegraphics[scale=0.3]{figures/rep_distribution_nih}
-\label{rep_dist_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 200$}}
-\end{figure}
-
-This motivates looking at the entire distribution of reputation difference between two algorithms.
-Figure \ref{ts_dg_rep_diff_nih} shows the distribution of reputation difference between $TS$ and $DG$ across $t$.
-This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right.
-Thus, the mean is not a representative statistic of the entire reputation difference distribution.
-As an alternative statistic for understanding the results of the competition game we introduce the \textit{relative reputation statistic} which looks at the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$.
-This corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a principal playing $A$ over a principal playing $B$ at time $t$ \footnote{Though we do not discuss it here, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion.
-In the GitHub appendix we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.}.
-
-\begin{figure}
-\caption{Reputation Difference Distribution}
-\includegraphics[scale=0.3]{figures/ts_dg_rep_diff_nih}
-\label{ts_dg_rep_diff_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $TS$ and $DG$ across $t$}}
-\end{figure}
-
-\begin{figure}
-\caption{Relative Reputation Plots}
-\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
-\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
-\caption*{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$.
-The shaded area display 95\% confidence intervals.}
-\label{relative_rep_plots}
-
-\end{figure}
-
-Figure \ref{relative_rep_plots} shows the relative reputation plots for $TS$ vs $DG$ on the Uniform and Needle In Haystack prior.
-For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $TS > DG$.
-The intuition behind this is that, especially since the principals start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information.
-However, this exploration leads to what we define as a \textit{relative reputation cost}.
-Namely, exploration leads to lower reputation relative to the myopic alternative.
-However, eventually the information acquired in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm.
-We define this as the \textit{relative reputation benefit} \footnote{We see similar behavior for the Heavy Tail prior}.
-
-Is exploration always costly?
-Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $TS$ always does relatively better than $DG$.
-There are two contributing factors to this.
-First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration.
-Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
-Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
-However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
-
-\section{Monopoly vs Duopoly}
-
-In this section we utilize the same instances and realizations as in the previous section and simulate the model described in section \ref{model}.
-We initially take the strategies of the firms as exogenous and simulate the model given these strategies in order to determine the payoffs associated with each pair of algorithms.
-Unless otherwise noted, all the results are reported at $t = 2000$.
-
-\bibliography{refs}
-
 \end{document}

### 2018-08-30 paper/content/barriers.tex
small edits about barriers

@@
 \begin{finding}
-\textit{When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation.
-However, when the incumbent plays $\DEG$ or $DG$ there is no substantial difference between the two.}
+\textit{Both the data advantage and reputation advantage alone serve as strong barriers to entry.
+When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $DG$ there is no substantial difference between the two.}
 \end{finding}

### 2018-08-30 paper/content/introduction.tex
small edits about barriers

@@
 Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantange".
+In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage".
 We further investigate which of the two is a stronger barrier to entry.
 \ascomment{and find what?
-and what's the new intuition?}
+and what's the new intuition?} \gaedit{We find that retaining either the reputation or the data advantage alone still serves as a strong barrier to entry, but that the data barrier to entry is only consistently stronger when the incumbent commits to a more advanced exploration algorithm.
+}
 

### 2018-08-30 paper/content/model.tex
add comment

@@
 Note $t_j$ is the \textit{local} time of the firm, not the global time, so the reputation score is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.\swcomment{define ``calendar time'' and ``local time'' formally...
-don't think the current reputation defn makes sense} The warm start of $k$ rounds allows this reputation score to be well-defined once the "competition game" begins.
+don't think the current reputation defn makes sense} \gacomment{yea, good point.
+Should define local time formally} The warm start of $k$ rounds allows this reputation score to be well-defined once the "competition game" begins.
 In our model the agents deterministically choose the firm with the higher reputation score and ties are broken uniformly at random.

### 2018-08-30 paper/content/introduction.tex
Minor edits.

--- 2018-08-30 paper/content/introduction.tex
@@
 
-Platforms routinely deploy \asedit{A/B tests}, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
-While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition, \asedit{even putting aside the deployment/maintenance costs}.
+Platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
+While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition, even putting aside the deployment/maintenance costs.
 In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term.
@@
 Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window.
-\asedit{Each firm's objective} is to maximize its market share ---the fraction of users coming to them.
+Each firm's objective is to maximize its market share ---the fraction of users coming to them.
 
@@
 
-\asedit{Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
-This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.}
+Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
+This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.
 
 Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantange".
+In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage".
 We further investigate which of the two is a stronger barrier to entry.
-\ascomment{and find what?
-and what's the new intuition?}
+We find that both are strong barriers on its own: removing either one substantially helps the late entrant.
+Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
 
-\subsection{Discussion and related work}
+\subsubsection{Discussion and related work.}
 Our work is related to a longstanding economics literature on competition vs.
@@
     \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
-
-\OMIT{ %%%%%%
-The interplay between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \cite{che2017recommender}, \cite{kremer2014implementing}, \cite{mansour2015bayesian}.
-The strategic experimentation literature in economics, such as \cite{bolton1999strategic} and \cite{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.
-} %%%%%%%%
-
 Our setting is also closely related to the "dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.

### 2018-08-30 paper/content/model.tex
Half-way into rewriting this thing.

--- 2018-08-30 paper/content/model.tex
@@
 
-\textbf{Overview} There are two firms and $T+2k$ agents where $k$ is the warm start, or the agents that each firm gets for free at the beginning of the game and $T$ is the number of rounds in the game.
-The timing of events is as follows:
+%\textbf{Overview} 
+We consider a game involving two firms and $T$ customers (henceforth, \emph{agents}).
+The game lasts for $T$ rounds.
+In each round, a new agent arrives, chooses among the two firms, interacts with the chosen firm, and leaves forever.
+
+Each interaction between a firm and an agent proceeds as follows.
+There is a set $A$ of $K$ actions, henceforth \emph{arms}, same for both firms and all rounds.
+The firm chooses an arm, and the agent experiences a numerical reward observed by the firm.
+Each arm corresponds to a different version of the experience that a firm can provide for an agent, and the reward corresponds to the agent's satisfaction level.
+The other firm does not observe anything about this interaction, not even the fact that this interaction has happened.
+
+From each firm's perspective, the interactions with agents follow the protocol of the multi-armed bandit problem (MAB).
+We focus on i.i.d.
+rewards: each reward is drawn independently from a distribution that depends on the chosen arm but not on the firm or the round.
+For concreteness, each reward is either $0$ or $1$.
+The mean reward of each action is the same for both firms (and each round), but initially unknown.
+
+Before the game starts, each firm commits to an MAB algorithm, and uses this algorithm to choose its actions.
+Each algorithm receives a ``warm start": additional $T_0$ agents that show up before the game starts, and interact with the firm in the same way as described above.
+
+\subsubsection{Agents.}
+Agents are myopic and non-strategic: they choose among the firms so as to maximize their expected reward, without attempting to influence the firms' learning algorithms or rewards of the future users.
+Agents are not well-informed: they only receive a rough signal about each firm's performance before they choose a firm, and no other information.
+
+Concretely, each firm has a \emph{reputation score}, and each agent's choice is driven by these two numbers.
+We posit a version of rational behavior: each agent chooses a firm with a maximal reputation score (breaking ties uniformly).
+The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
+
+\subsubsection{MAB algorithms.} We consider three classes of algorithms, ranging from more primitive to more sophisticated and better-performing.
+For concreteness, we fix one algorithm from each class.
+Our pilot experiments indicate that the results do not change substantially if other algorithms are chosen.
+
+STOPPED HERE.
+
 \begin{enumerate}
-\item At $t = 0$, the firms simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
-\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a ``warm" start.
-The algorithm that the firm commits to makes $k$ rounds of progress and uses this information to initialize its information set.
-Additionally, the reputation score used by the agents is initialized using the rewards from these $k$ rounds.
-\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses the firm with the higher reputation at the time.
-\item The firm that is chosen selects an action $a_{t} \in A$ from a set of actions that is fixed across firms and rounds.
-\item Both the agent and the firm observe the reward $r_t \in [0, 1]$ from the action.
-The agent reports this reward to the firm and the future agents and the reputation score for the chosen firm is updated.
-\item Repeat 2-4 for $T$ rounds.
-\end{enumerate}
+\item \emph{Greedy} algorithms that strive to take actions with maximal mean reward, based on the current information.
 
-Generally, the rewards are independent and identically distributed with a common prior.
-For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors.
-Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require a prior to operate, such as Thompson Sampling, we use a "fake" prior of $Beta(\alpha=1,\beta=1)$}.
-We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.\swcomment{this seems misleading and should not be here.
-just say that each action has bernouli rewards with unknown means}
+do not do explicit exploration.
+engage in no purposeful exploration and take the best short-sighted action.
+We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
 
-\noindent \textbf{Agents} We suppose that agents are homogenous, myopic, and non-strategic.
-The utility function for the agents is simply to maximize their reward in the one period in which they are alive.
-We suppose that agents do not attempt to manipulate the strategy of the firm, do not take the strategies of the firm into account when choosing between the firms, and do not condition their behavior on calendar time.
-In our model, each agent uses the average reward of past agents as a proxy for their expected utility.
-For simplicity, the reputation score, $R_{jt}$ is defined as a sliding window average\footnote{Another natural formulation would be to have exponential discounting of the past rewards.
-We believe that the results should be qualitatively similar to those presented here as in both formulations the more recent past matters more than the distant past.}.
-\begin{center}
-$R_{jt} = \frac{1}{M} \sum\limits_{i=1}^{M} r_{t_j-i}$
-\end{center}
-
-Note $t_j$ is the \textit{local} time of the firm, not the global time, so the reputation score is the sliding window average of the last $M$ times that firm $j$ was selected by the agents.\swcomment{define ``calendar time'' and ``local time'' formally...
-don't think the current reputation defn makes sense} The warm start of $k$ rounds allows this reputation score to be well-defined once the "competition game" begins.
-In our model the agents deterministically choose the firm with the higher reputation score and ties are broken uniformly at random.
-
-\noindent \textbf{Firms} We suppose that the firms simply care about maximizing their expected market share (i.e.
-maximizing the number of $T$ agents who select them).
-We model the "competition game" between firms as a simultaneous move game where both firms commit to a learning algorithm at $t = 0$.
-
-\noindent \textbf{MAB algorithms} We suppose that firms commit to a learning algorithm from a fixed set of algorithms $\mathcal{A}$.
-We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
-\begin{enumerate}
 \item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
-We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward \cite{Shipra-colt12}.
+We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward.
 \item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
@@
 For our experiments we keep $\epsilon = 0.05$ fixed.
-\item Greedy / myopic algorithms that engage in no purposeful exploration and take the best short-sighted action.
-We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
 \end{enumerate}
@@
 We have that both the incumbent and entrant commit to a learning algorithm \textit{before} either firm receives any agent.
-After the $X$ rounds, both firms still receive the $k$ warm start agents.
+After the $X$ rounds, both firms still receive the $T_0$ warm start agents.
 

### 2018-08-30 paper/content/introduction.tex
Minor edits.
Some other minor edit.

@@
 MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background.
-We consider a simple and well-understood MAB model with i.i.d.
-rewards \cite{bandits-ucb1}.
-We focus on a well-known distinction between the ``greedy" (exploitation-only) algorithm, "naive" MAB  algorithms that separate exploration and exploitation, and better MAB algorithms that combine the two.
-As a token example of ``near-optimal" bandit algorithms,
-we use a classic algorithm called ``Thompson Sampling", see \cite{TS-survey-FTML18} for background.
+We consider i.i.d.
+rewards, a well-studied and well-understood MAB model \cite{bandits-ucb1}.
+We focus on a well-known distinction between ``greedy", exploitation-only algorithms, "naive" algorithms that separate exploration and exploitation, and better algorithms that combine the two.
 

### 2018-08-30 paper/content/model.tex
Done with a full pass/rewrite.

@@
 
-\section{Model}\label{sec:model}
+\section{Model and Preliminaries}\label{sec:model}
 
@@
 
+In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``entrant").
+Formally, the incumbent enjoys additional $X$ rounds of the ``warm start".
+We treat $X$ as an exogenous element of the model, and study the consequences for a fixed $X$.
+
 \subsubsection{Agents.}
@@
 
-\subsubsection{MAB algorithms.} We consider three classes of algorithms, ranging from more primitive to more sophisticated and better-performing.
+\subsubsection{MAB algorithms.} We consider three classes of algorithms, ranging from more primitive to more sophisticated:
+
+\begin{enumerate}
+\item \emph{Greedy algorithms} that strive to take actions with maximal mean reward, based on the current information.
+
+\item \emph{Exploration-separating algorithms} that separate exploration and exploitation.
+The ``exploitation" choices strives to maximize mean reward in the next round, and the ``exploration" choices do not use the rewards observed so far.
+\item \emph{Adaptive exploration}: algorithms that combine exploration and exploitation, and sway the exploration choices towards more promising alternatives.
+\end{enumerate}
+
 For concreteness, we fix one algorithm from each class.
 Our pilot experiments indicate that the results do not change substantially if other algorithms are chosen.
-
-STOPPED HERE.
+For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge).
+We consider:
 
 \begin{enumerate}
-\item \emph{Greedy} algorithms that strive to take actions with maximal mean reward, based on the current information.
+\item a greedy algorithm that chooses an arm with largest posterior mean reward. We call it "Dynamic Greedy" (because the chosen arm may change over time), \DG in short.
 
-do not do explicit exploration.
-engage in no purposeful exploration and take the best short-sighted action.
-We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
+\item an exploration-separated algorithm that in each round, \emph{explores} with probability $\eps$: chooses an arm independently and uniformly at random, and with the remaining probability \emph{exploits} according to \DG.
+We call it ``dynamic epsilon-greedy", \DEG in short.
 
-\item "Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation.
-We consider Thompson Sampling (from hereon $\TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is "optimal" in the sense of having the highest mean reward.
-\item "Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation.
-We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $\DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability.
-For our experiments we keep $\epsilon = 0.05$ fixed.
+\item an adaptive-exploration algorithm called "Thompson Sampling" (\TS).
+In each round, this algorithm updates the posterior distribution for the mean reward of each arm $a$, draws an independent sample $s_a$ from this distribution, and chooses an arm with the largest $s_a$.
 \end{enumerate}
 
-In the standard multi-armed bandit problem it is known that $\TS > \DEG > DG$ in terms of maximizing cumulative reward over a sufficiently large time horizon.
-The primary question that we want to understand is when, in competition, are the firms incentivized to adopt the "better" algorithms?
+For ease of comparison, all three algorithms are parameterized with the same fake prior: namely, the mean reward of each arm is drawn independently from a Beta(1,1) distribution.
+Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.
 
-\noindent \textbf{Incumbent} In some of our experiments we modify our model so that one firm enters the market $X$ number of rounds before the other.
-We refer to the firm that enters before as the "incumbent" and the firm that enters after $X$ rounds as the "entrant." In the $X$ rounds before the "entrant" enters the incumbent is a temporary monopolist as the agents that arrive in these $X$ rounds are forced to select the incumbent since it is the only firm in the market and there is no outside option.
-We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$.
-We have that both the incumbent and entrant commit to a learning algorithm \textit{before} either firm receives any agent.
-After the $X$ rounds, both firms still receive the $T_0$ warm start agents.
+Both \DEG and \TS are classic MAB algorithms, see \cite{Bubeck-survey12,TS-survey-FTML18} for background.
+It is well-known that \TS is near-optimal in terms of the cumulative rewards, and \DEG is very suboptimal, but still much better than \DG.%
+\footnote{Precise formulations, namely upper/lower bound on \emph{regret} of the three algorithms, are not very relevant to this paper.} In a formula:
+    $ \TS \gg \DEG \gg \DG $
+as stand-alone MAB algorithms.
 

### 2018-08-30 paper/content/model.tex
Minor edits.

@@
 
-For ease of comparison, all three algorithms are parameterized with the same fake prior: namely, the mean reward of each arm is drawn independently from a Beta(1,1) distribution.
+For ease of comparison, all three algorithms are parameterized with the same fake prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution.
 Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.
@@
 It is well-known that \TS is near-optimal in terms of the cumulative rewards, and \DEG is very suboptimal, but still much better than \DG.%
-\footnote{Precise formulations, namely upper/lower bound on \emph{regret} of the three algorithms, are not very relevant to this paper.} In a formula:
+\footnote{Precise formulations, namely upper/lower bound on \emph{regret} of the three algorithms, are not very relevant to this paper.} In a stylized formula:
     $ \TS \gg \DEG \gg \DG $

### 2018-08-30 paper/content/introduction.tex
Renamed a subsection.

@@
 
-\subsubsection{Discussion and related work.}
+\subsubsection{Related work.}
 Our work is related to a longstanding economics literature on competition vs.
 innovation, \eg \cite{Schumpeter-42,barro2004economic,Aghion-QJE05}.
-While this literature focuses on R\&D costs of innovation, ``reputational costs" thereof seem new and specific to exploration.
+While this literature focuses on R\&D costs of innovation, ``reputational costs" seem new and specific to exploration.
 

### 2018-08-30 paper/content/sim_details.tex
Rewrote simulation details.

@@
 
-\section{Simulation Details}\label{section:3}
+%\subsubsection{Simulation details.}
+%\label{section:3}
 
-We evaluate the consequences of our model via simulation.\swcomment{In the process of removing ``prior'' throughout}
+\subsubsection{MAB instances.} We consider several representative instances of the MAB problem, hencenforth \emph{MAB instances}.
+In all experiments, there are $K=10$ arms.
+Since we focus on 0-1 rewards, a problem instance is specified by the mean reward $\mu(a)$ of each arm $a$.
+The mean rewards are chosen initially from some distribution, as specified below:
+\begin{enumerate}
+\item \emph{Needle In Haystack}: one arm (the ``needle") is chosen uniformly at random. This arm has mean reward $.7$, and the remaining ones have mean reward $.5$.
 
-\textbf{Bandit Instances} We look at bandit instances drawn from three different bandit \swedit{distributions} that capture different types of learning problems.
-Recall that we only consider Bernoulli rewards, which are fully parameterized by the means.
-For these experiments we fix the number of arms we consider, $K = 10$.
-\begin{enumerate}
-\item Needle In Haystack - $K-1$ arms with mean 0.5 and 1 arm with 0.7.
-\item Uniform - the $K$ arms have means drawn uniformly at random from $[0.25, 0.75]$
-\item "Heavy Tail" - the $K$ arms have means drawn from $Beta(\alpha=0.6, \beta = 0.6)$.
-With this prior it was likely to have means at the "extremes" or means that were close to 0 as well as means that were close to 1.
+\item \emph{Uniform instance}: the mean reward of each arm is drawn independently and uniformly from $[0.25, 0.75]$.
+\item \emph{Heavy-tail instance}: the mean reward of each arm is drawn independently from $\Beta(.6,.6)$ distribution.%
+    \footnote{This distribution has substantial ``tail probabilities".}
 \end{enumerate}
 
-\noindent \textbf{Simulation of Competition Game} Unless otherwise noted, all of the reported results utilize the same set of randomly drawn bandit instances and realizations from the prior.
-Namely, for each bandit prior we draw $N = 1000$ bandit instances.
-For each of these instances we run simulations of our model for varying values of $k$ and $X$.
-We take the maximum values of $k$ and $X$ for the simulations we run, $k_{max}$ and $X_{max}$ respectively, and compute a realization table of dimension $(T+k_{max}+X_{max}) \times K$.
+A particular realization of the mean rewards of the arms is called a \emph{realized} MAB instance. (Note that the actual rewards are still drawn stochastically.)
 
-This realization table, as well as fixing the random seed for the same bandit instance and realization table across experiments, ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
+\subsubsection{Simulation details.} For each MAB instance we draw $N = 1000$ realized MAB instances.
+For each of these $N$ instances we simulate our model for varying values of $T_0$ and $X$.
+Taking the maximum values of these parameters to be $T_0^{\max}$ and $X^{\max}$.
+resp., we compute the table of realized rewards (\emph{realization table}) of dimension $(T+T_0^{\max}+X^{\max}) \times K$.
+Rows of this table correspond to rounds, and columns correspond to arms.
+For each cell in column $a$, we draw the reward independently from Bernoulli distribution with expectation $\mu(a)$.
+We use the same table to specify realized rewards for all algorithms and all experiments on this realized MAB instance.
+This ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
+In the competition game, we use the last $T$ rows of the table, for consistency if we vary $T_0$ and $X$.
+
+\OMIT{This realization table, as well as fixing the random seed for the same bandit instance and realization table across experiments, ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
 In the competition game we draw from the $T \times K$ portion of the table, so that if two different algorithms picks arm $a$ at time $t$, they get the same $[a, t]$ realization in the table.
-This setup also ensures that in the warm start period, increasing the warm start from $k$ to $k + 10$ results in the same behavior in the first $k$ rounds.
+This setup also ensures that in the warm start period, increasing the warm start from $k$ to $k + 10$ results in the same behavior in the first $k$ rounds.}
 
-For the simulations we fix the sliding window size $M = 100$.
-Low values of $M$ induced too much random noise into the results and we found that increasing $M$ to be larger than $100$ did not make a substantial qualitative difference so we fix this value.
+We fix the sliding window size $M = 100$. We found that low values of $M$ induced too much random noise into the results, and further increasing $M$ did not make a substantial qualitative difference.
 

### 2018-08-30 paper/competing_bandits.tex
Added some macro.
Removed \asedit{}

@@
 \usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
 \usepackage{nicefrac}       % compact symbols for 1/2, etc.
 \usepackage{microtype}      % microtypography
-\usepackage{amssymb}
+\usepackage{amsmath,amssymb,amsfonts}
 %\usepackage{natbib} % cannot use with AAAI! A.S.
@@
 \newcommand{\eg}{{\em e.g.,~\xspace}}
-\newcommand{\TS}{\mathrm{TS}}
-\newcommand{\DEG}{\mathrm{DEG}}
+
+\newcommand{\term}[1]{\ensuremath{\mathtt{#1}}\xspace}
+\newcommand{\TS}{\term{TS}}
+\newcommand{\DEG}{\term{DEG}}
+\newcommand{\DG}{\term{DG}}
+\newcommand{\Beta}{\term{Beta}}
+
+\newcommand{\eps}{\varepsilon}
 
@@
 We run extensive numerical experiments in a stylized duopoly model in which two firms deploy multi-armed bandit algorithms and compete for myopic users.
-We find that duopoly \asedit{and monopoly} tend to favor a primitive ``greedy algorithm" that does not explore, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms.
+We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms.
 Our findings provide new intuition on the ``first-mover advantage" in the digital economy.

### 2018-08-30 paper/content/model.tex
Added a footnote with regret rates.

@@
 It is well-known that \TS is near-optimal in terms of the cumulative rewards, and \DEG is very suboptimal, but still much better than \DG.%
-\footnote{Precise formulations, namely upper/lower bound on \emph{regret} of the three algorithms, are not very relevant to this paper.} In a stylized formula:
+\footnote{Formally, \TS achieves regret 
+    $\tilde{O}(\sqrt{TK})$ and
+    $O(\tfrac{1}{\Delta} \log T)$,
+where $\Delta$ is the gap in mean rewards between the best and second-best arms.
+\DEG has regret $\tilde{\Theta}(T^{2/3} K^{1/3})$ in the worst case.
+And \DG can have regret as high as $\Omega(T)$.
+Deeper discussion of these distinctions is not very relevant to this paper.}
+In a stylized formula:
     $ \TS \gg \DEG \gg \DG $

### 2018-08-30 paper/content/sim_details.tex
Minor edits.

@@
 
-We fix the sliding window size $M = 100$. We found that low values of $M$ induced too much random noise into the results, and further increasing $M$ did not make a substantial qualitative difference.
+We fix the sliding window size $M = 100$. We found that low values of $M$ induced too much random noise in the results, and increasing $M$ further did not make a substantial qualitative difference.
 

### 2018-08-30 paper/content/perf_in_iso.tex
Barely started.

@@
 
-\section{Performance in Isolation}\label{section:4}
+\section{Algorithms' Performance in Isolation}\label{section:4}
 
-We evaluate the performance of the different learning algorithms on the bandit priors we consider in isolation.
-There are two reasons why this is important to consider.
+We start with a pilot experiment in which we investigate each algorithm's performance "in isolation": in a stand-alone bandit problem without the competition.
+
+STOPPED HERE.
+
 First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $\TS > \DEG > DG$ for sufficiently large $t$.

### 2018-08-31 paper/content/model.tex
Minor edit.
Changed terminology: realized MAB instance -> mean reward vector.

@@
 We focus on i.i.d.
-rewards: each reward is drawn independently from a distribution that depends on the chosen arm but not on the firm or the round.
-For concreteness, each reward is either $0$ or $1$.
-The mean reward of each action is the same for both firms (and each round), but initially unknown.
+Bernoulli rewards: the  reward of each arm $a$ is drawn from $\{0,1\}$ independently with expectation $\mu(a)$.
+The mean rewards $\mu(a)$ are the same for all rounds and both firms, but but initially unknown.
 
@@
 
-Both \DEG and \TS are classic MAB algorithms, see \cite{Bubeck-survey12,TS-survey-FTML18} for background.
+Both \DEG and \TS are classic and well-understood MAB algorithms, see \cite{Bubeck-survey12,TS-survey-FTML18} for background.
 It is well-known that \TS is near-optimal in terms of the cumulative rewards, and \DEG is very suboptimal, but still much better than \DG.%

### 2018-08-31 paper/content/perf_in_iso.tex
Changed terminology: realized MAB instance -> mean reward vector.

@@
 We start with a pilot experiment in which we investigate each algorithm's performance "in isolation": in a stand-alone bandit problem without the competition.
+We focus on reputation scores (\ie sliding window averages) generated by each algorithm.
 
-STOPPED HERE.
+First, we confirm that algorithms' performance is ordered as we'd expect:
+    $\TS \gg \DEG \gg \DG$.
+For a given algorithm and MAB instance
 
-First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $\TS > \DEG > DG$ for sufficiently large $t$.
+the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $\TS > \DEG > DG$ for sufficiently large $t$.
 Second, we want to better understand what statistics of the priors we can look at in isolation in order to help us predict and understand what to expect in the competition game.

### 2018-08-31 paper/content/sim_details.tex
Rewrote "simulation details".

@@
 
-\subsubsection{MAB instances.} We consider several representative instances of the MAB problem, hencenforth \emph{MAB instances}.
+\subsubsection{MAB instances.} 
 In all experiments, there are $K=10$ arms.
-Since we focus on 0-1 rewards, a problem instance is specified by the mean reward $\mu(a)$ of each arm $a$.
-The mean rewards are chosen initially from some distribution, as specified below:
+Since we focus on 0-1 rewards, an instance of the MAB problem is specified by the mean reward vector $(\mu(a):\; a\in A)$, henceforth \MRV.
+Initially the \MRV is drawn from some distribution, termed \emph{MAB instance}.
+We consider several representative MAB instances.
 \begin{enumerate}
@@
 
-A particular realization of the mean rewards of the arms is called a \emph{realized} MAB instance. (Note that the actual rewards are still drawn stochastically.)
-
-\subsubsection{Simulation details.} For each MAB instance we draw $N = 1000$ realized MAB instances.
-For each of these $N$ instances we simulate our model for varying values of $T_0$ and $X$.
-Taking the maximum values of these parameters to be $T_0^{\max}$ and $X^{\max}$.
-resp., we compute the table of realized rewards (\emph{realization table}) of dimension $(T+T_0^{\max}+X^{\max}) \times K$.
-Rows of this table correspond to rounds, and columns correspond to arms.
-For each cell in column $a$, we draw the reward independently from Bernoulli distribution with expectation $\mu(a)$.
-We use the same table to specify realized rewards for all algorithms and all experiments on this realized MAB instance.
+\subsubsection{Simulation details.} 
+For each MAB instance we draw $N = 1000$ \MRVs independently from the corresponding distribution.
+We use this same collection of \MRVs for all experiments with this MAB instance.
+For each \MRV we draw a table of realized rewards (\emph{realization table}), and use this same table for all experiments on this \MRV.
 This ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
-In the competition game, we use the last $T$ rows of the table, for consistency if we vary $T_0$ and $X$.
+More specifically, the realization table is a 0-1 matrix $W$ with $K$ columns which correspond to arms, and $T+T_{\max}$ rows, which correspond to rounds.
+Here $T_{\max}$ is the maximal duration of the ``warm start" in our experiments, \ie the maximal value of $X+T_0$.
+For each arm $a$, each value $W(\cdot,a)$ is drawn independently from Bernoulli distribution with expectation $\mu(a)$.
+Then in each experiment, the reward of this arm in round $t$ of the warm start is taken to be $W(t,a)$, and its reward in round $t$ of the game is $W(T_{\max}+t,a)$.
 
@@
 
+While we experiments with various MAB instances and parameter settings, in the main paper we only report plots for selected, representative experiments.
+Additional plots can be found in the supplement.
+Unless noted otherwise, our findings are based on and consistent with all these plots.
+
 \end{document}

### 2018-08-31 paper/content/perf_in_iso.tex
Stopping for lunch :).

@@
 
-We start with a pilot experiment in which we investigate each algorithm's performance "in isolation": in a stand-alone bandit problem without the competition.
-We focus on reputation scores (\ie sliding window averages) generated by each algorithm.
+We start with a pilot experiment in which we investigate each algorithm's performance "in isolation": in a stand-alone MAB problem without competition.
+We focus on reputation scores generated by each algorithm.
 
@@
     $\TS \gg \DEG \gg \DG$.
-For a given algorithm and MAB instance
+For each algorithm and each MAB instance, we compute the mean reputation score at each round, averaged over all \MRVs.
+We plot the \emph{mean reputation trajectory}: how this score evolves over time.
+Figure \ref{prelim_means} shows such plot for the Needle In Haystack instance (for other MAB instances the plots are similar).
 
-the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where, according to the mean reputation, $\TS > \DEG > DG$ for sufficiently large $t$.
-Second, we want to better understand what statistics of the priors we can look at in isolation in order to help us predict and understand what to expect in the competition game.
+\begin{finding}
+\textit{The mean reputation trajectories are arranged as predicted by prior work: 
+    $\TS > \DEG > \DG$.}
+\end{finding}
 
 \begin{figure}
-\caption{Mean Reputation Trajectories in Isolation}
+\caption{Mean Reputation Trajectories in Isolation. The shaded area displays 95\% confidence intervals.}
 \includegraphics[scale=0.35]{"figures/nih_iso_mean"}
 \label{prelim_means}
-\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
+%\caption*{\tiny{Mean reputation trajectory The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
 The shaded area display 95\% confidence intervals.}}
@@
 
-Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior \footnote{The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the supplementary material.}.
+Now we ask a more nuanced question: are the mean reputation trajectories the right statistic for our purposes? That is, how well do they predict the outcomes of the competition game?
 

### 2018-08-31 paper/content/introduction.tex
comment

@@
 In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage".
-We further investigate which of the two is a stronger barrier to entry.
-We find that both are strong barriers on its own: removing either one substantially helps the late entrant.
+We further investigate which of the two is a stronger barrier to entry.\swcomment{so which of the two?} We find that both are strong barriers on its own: removing either one substantially helps the late entrant.
 Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.

### 2018-08-31 paper/content/perf_in_iso.tex
Done for today ...

@@
 Figure \ref{prelim_means} shows such plot for the Needle In Haystack instance (for other MAB instances the plots are similar).
+We summarize this finding as follows:
 
@@
 \begin{figure}
-\caption{Mean Reputation Trajectories in Isolation. The shaded area displays 95\% confidence intervals.}
+\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area displays 95\% confidence intervals.}
 \includegraphics[scale=0.35]{"figures/nih_iso_mean"}
@@
 
-Now we ask a more nuanced question: are the mean reputation trajectories the right statistic for our purposes? That is, how well do they predict the outcomes of the competition game?
+The mean reputation trajectory is, arguably, the most natural way to represent algorithm's performance on a given problem instance.
+But is it the right statistic for our purposes?
+That is, how well do these trajectories predict the outcomes of the competition game?
+The following conjecture is natural:
 
-\begin{figure}[H]
-\caption{Reputation Distribution}
-\includegraphics[scale=0.35]{figures/rep_distribution_nih}
-\label{rep_dist_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
-\end{figure}
+\begin{conjecture}\label{conj:mean-trajectories}
+For a given MAB instance, if one algorithm's mean reputation trajectory lies above another, then the first algorithm prevails in the competition game.
+\end{conjecture}
 
+However, we find a more nuanced picture.
+
+\begin{finding}
+\textit{
+Conjecture~\ref{conj:mean-trajectories} is false. Mean reputation trajectories do not suffice to explain the outcomes under competition.}
+\end{finding}
+
+\OMIT{ %%%%%
 \begin{finding}
@@
 \end{finding}
+} %%%%%
 
-Is looking at the mean performance of between two algorithms sufficient for understanding what happens in competition?
-To answer this, we first look at the entire reputation distribution at $t = 500$ in Figure \ref{rep_dist_nih}.
-We see that the ``naive" algorithms $DG$ and $\DEG$ have a bi-modal reputation distribution whereas $\TS$ does not.
-The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not.
+\begin{figure}[H]
+\caption{Distribution of reputation scores for Needle-in-Haystack at $t=500$ (smoothed using a kernel density estimate)}
+\includegraphics[scale=0.35]{figures/rep_distribution_nih}
+\label{rep_dist_nih}
+%\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
+\end{figure}
+
+To see what could go wrong, consider how algorithm's reputation score is distributed at a particular time.
+That is, consider the empirical distribution of this score over different \MRVs.
+For concreteness, consider the needle-in-haystack instance at time $t=500$ (Figure \ref{rep_dist_nih}).
+We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
+The reason is that for this MAB instance, $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
+However, the mean reputation trajectory of an algorithm would summarize all this complexity the average over the \MRVs.
+This may be inadequate for explaining the outcome of the duopoly game, given that the latter is determined by a simple comparison between the firm's reputation scores.
+Further, consider the difference in reputation scores (\emph{reputation difference}) between \TS and \DG on a particular \MRV.
+Let's plot the empirical distribution of the reputation difference (over the \MRVs) at a particular time point.
+Figure~\ref{ts_dg_rep_diff_nih} shows such plots for several time points.
+We see that the distribution is skewed to the right, for which the mean is not a representative statistic.
+\ascomment{so what?}
+
+\OMIT{ %%%%%%%%
+ or not.
 If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds.
@@
 In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \cite{CompetingBandits-itcs16}}.
+} %%%%%%%%
 
-To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $\TS$ and $DG$.
-This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large.
-Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution.
-There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the \textit{relative reputation proportion}.
+\OMIT{There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(\TS) - reputation(\DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the \textit{relative reputation proportion}.}
 
 \begin{figure}[H]
-\caption{Reputation Difference Distribution}
+\caption{Distribution of reputation difference $\TS-\DG$ for the Needle-in-Haystack (smoothed via a kernel density estimate)}
 \includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
 \label{ts_dg_rep_diff_nih}
-\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $DG$ across $t$}}
+%\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $\DG$ across $t$}}
 \end{figure}
+
+Let us consider a statistic with a more direct connection to the competition game: for each time $t$, the proportion of \MRVs for which one algorithm has a higher reputation score than the other.
+
+\ascomment{Stopped here.}
 
@@
 \begin{finding}
-\textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $\TS$ doing worse than $DG$ for small time horizons.
+\textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $\TS$ doing worse than $\DG$ for small time horizons.
 We observe this under the Uniform and Heavy Tail priors.
@@
 
-The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $\DEG >DG$ or $\TS > DG$ according to the mean reputation but $DG > \DEG$ or $DG > TS$ according to the relative reputation proportion.
-In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $\DEG > DG$ according to the mean reputation but $DG > \DEG$ according to the relative reputation proportion.
+The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $\DEG >\DG$ or $\TS > \DG$ according to the mean reputation but $\DG > \DEG$ or $\DG > TS$ according to the relative reputation proportion.
+In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $\DEG > \DG$ according to the mean reputation but $\DG > \DEG$ according to the relative reputation proportion.
 Additionally, these results carry over to the competition game}.
@@
 \includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
-\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq DG$.
+\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.
 The shaded area display 95\% confidence intervals.}}
@@
 
-Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $DG$ on the Uniform and Needle In Haystack prior.
-For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $\TS > DG$.
+Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $\DG$ on the Uniform and Needle In Haystack prior.
+For the Uniform prior we see that, in the early rounds, $\DG > \TS$ for the majority of the simulations but that, eventually, $\TS > \DG$.
 The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information.
-However, eventually the information acquired from purposeful exploration in the early rounds allows $\TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm.
+However, eventually the information acquired from purposeful exploration in the early rounds allows $\TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $\DG$ cannot trivially find the best arm.
 The early exploration leads to what we define as a \textit{relative reputation cost} and the eventual gain in reputation leads to what we define as a \textit{relative reputation benefit}.
@@
 Is exploration always costly?
-Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $\TS$ always does relatively better than $DG$.
+Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $\TS$ always does relatively better than $\DG$.
 There are two contributing factors to this.

### 2018-09-01 paper/content/inverted_u.tex
restructure the paper a bit

@@
 
-\section{Competition vs Better Algorithms: Inverted-U}\label{section:5}
+\section{Competition vs Better Algorithms}\label{section:5}
 
-In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \cite{Aghion-QJE05}.
+In this section we document the results of the competition game for various parameterizations of our model.
 We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly.
 
+\OMIT{main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \cite{Aghion-QJE05}. 
+}
+
+\OMIT{
 \begin{finding}
-\textit{Under the priors that are relative reputation costly and with low warm start, the optimal strategies in the competition game are:
+\textit{For sufficiently low warm start, the optimal strategies in the competition game are:
 \begin{center}
-\textbf{Permanent Monopoly} - $DG$ \\
+\textbf{Permanent Monopoly} - $\DG$ \\
 \textbf{Temporary Monopoly} - $\TS$ \\
-\textbf{Permanent Duopoly} - $DG$ \\
+\textbf{Permanent Duopoly} - $\DG$ \\
 \end{center}
@@
 \end{finding}
+}
 
@@
 We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index).
-In our model we have no prices and no costs and so we interpret increases in "market power" as decreases in the number of rounds where a firm has to compete for agents.
+In our model we have no prices and no costs and so we interpret increases in ``market power" as decreases in the number of rounds where a firm has to compete for agents.
 
-We utilize the same instances and realizations as in the section 4 and simulate the model described in section 2.
+% couldn't get section references to work so hardcoding for now
+We utilize the same instances and realizations as in the section 3  and simulate the model described in section 2.
 We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the expected payoffs associated with each pair of algorithms.
@@
 \textbf{Permanent Monopoly} Under permanent monopoly only a single firm exists in the market for every period.
-Since, in our formulation, the firm only gets utility from a larger market share it is indifferent between each algorithm in $\mathcal{A}$ since, regardless of the algorithm it deploys, it will get the entire market.
-If we suppose that deploying the "better" algorithms has even an $\epsilon > 0$ cost, then the firm would choose to deploy $DG$.
+In our model the firm only gets utility from a larger market share so it is indifferent between each algorithm in $\mathcal{A}$ since, regardless of the algorithm it deploys, it will get the entire market.
+If we suppose that deploying the ``better" algorithms has even an $\epsilon > 0$ cost or, equivalently, that firms break indifference towards algorithms that are easier to deploy, then the firm would choose to deploy $\DG$.
 
@@
 What algorithms are firms incentivized to deploy?
-We report the average market share taken over the $N$ simulations for a set of exogenous values of $k$.
-Additionally, we report the mean and median of an additional quantity which we define as the effective end of game.
+We report the average market share taken over the $N$ simulations for a set of exogenous values of $T_0$.
+Additionally, we report the mean and median of an additional quantity which we define as the \textit{effective end of game}.
 
@@
 \begin{finding}
-\textit{For low warm starts, when exploration is relative reputation costly $DG >TS$ in competition, whereas when exploration is not relative reputation costly $\TS > DG$.
-For sufficiently large warm starts, $\TS$ is dominant across all priors.}
+\textit{
+\begin{enumerate}
+\item For low warm starts, under Heavy Tail and Uniform \\ $\DG > \TS$ in competition (see Table \ref{sim_ht}),
+\item Under Needle In Haystack, \\ $\TS > \DG$ (see Table \ref{sim_nih}). 
+\item For sufficiently large warm starts, $\TS$ is dominant across all the families of instances.
+\item Even for a large warm start, $\DG > \DEG$ under Heavy Tail
+\end{enumerate}
+}
 \end{finding}
 
-Table \ref{sim_unif} shows the results under the Uniform prior.
-For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $\TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}.
-Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $k$ as fixing the starting point on the relative reputation plots.
+\OMIT{
+Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $T_0$ as fixing the starting point on the relative reputation plots.
 The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$.
-The intuition for the result in competition aligns with the fact that, for Uniform and Heavy Tail, $\TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $\TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $\TS$.
-However, if we move the warm start sufficiently high so that $\TS$ moves to the relative reputation benefit regime, $\TS > DG$.
+The intuition for the result in competition aligns with the fact that, for Uniform and Heavy Tail, $\TS$ is relative reputation costly compared to $\DG$ and so, for low warm starts, $\TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $\DG$ beats $\TS$.
+However, if we move the warm start sufficiently high so that $\TS$ moves to the relative reputation benefit regime, $\TS > \DG$.
+}
 
-However, we do not always see that $DG > TS$ under some warm starts.
-Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $\TS > DG$.
-This is precisely due to the fact that, under the Needle In Haystack, $\TS$ is not relative reputation costly compared to $DG$ and thus we see that $\TS > DG$.
+ \OMIT{However, we do not always see that $\DG > \TS$ under some warm starts.
+Table \ref{sim_nih} shows the results for the Needle In Haystack family where we see that $\TS > \DG$.
+This is precisely due to the fact that, under the Needle In Haystack, $\TS$ is not relative reputation costly compared to $\DG$ and thus we see that $\TS > \DG$.}
 
-Combining these results with the fact that the EEOG values are relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right, similar to the distribution of reputation differences}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.
+Combining this finding with the fact that the EEOG values are relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.
 
@@
 \centering
-\caption{Duopoly Experiment Uniform}
+\caption{Duopoly Experiment Heavy Tail} 
 \begin{tabular}{rlll}
   \hline
- & k = 20 & k = 250 & k = 500 \\
+ & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 \\ 
   \hline
-TS vs DG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.02\\ eeog \\ avg: 800\\ med: 754} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 910\\ med: 906.5} \\
-  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 180\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.02\\ eeog \\ avg: 810\\ med: 734} & \makecell{\textbf{0.55} $\pm$0.02\\ eeog \\ avg: 970\\ med: 987} \\
-  DG vs DEG & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 470\\ med: 57.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1088} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1142} \\
+TS vs DG & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 55\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 570\\ med: 0} & \makecell{\textbf{0.76} $\pm$0.02\\ eeog \\ avg: 620\\ med: 98.5} \\
+  TS vs DEG & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 37\\ med: 0} & \makecell{\textbf{0.88} $\pm$0.01\\ eeog \\ avg: 480\\ med: 0} & \makecell{\textbf{0.9} $\pm$0.01\\ eeog \\ avg: 570\\ med: 113.5} \\
+  DG vs DEG & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 410\\ med: 7} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 790\\ med: 762} & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 730\\ med: 608} \\
    \hline
 \end{tabular}
-\label{sim_unif}
+\label{sim_ht}
 \end{table}
@@
   \hline
- & k = 20 & k = 250 & k = 500 \\
+ & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 \\
   \hline
@@
 In terms of our model, this corresponds to varying the value of $X$.
-For this section we report results fixing $k = 20$ and focus on the priors where we observed that, for this warm start, $DG > TS$ \footnote{However, we report the results for the Needle In Haystack prior in the supplementary material.
+For this section we report results fixing $T_0 = 20$ and focus on the family of instances where we observed that, for this warm start, $\DG > \TS$ \footnote{However, we report the results for the Needle In Haystack family in the supplementary material.
 The results are unsurprising, where $\TS$ is the dominant strategy for both the entrant and incumbent.}.
@@
 
-In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $DG$ is the dominant strategy for the entrant on the Heavy Tail prior \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.
-However, while $DG > TS$ for the entrant, $DG$ does roughly the same as $\DEG$ on the Uniform prior.
+In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the Heavy Tail family \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.
+However, while $\DG > \TS$ for the entrant, $\DG$ does roughly the same as $\DEG$ on the Uniform family.
 The results are in the supplementary material.}.
-In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across priors.
+In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the family of instances we consider.
 
@@
 
-Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same other conditions, $DG$ was preferred to $\TS$?
+Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was preferred to $\TS$?
 The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
-This intuition is very similar to that observed in the previous section by increasing warm start.
-One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the "incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
+One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the ``incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
 The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.
-As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
-Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
+\OMIT{As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.} Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
 

### 2018-09-01 paper/content/perf_in_iso.tex
restructure the paper a bit

@@
 
-We start with a pilot experiment in which we investigate each algorithm's performance "in isolation": in a stand-alone MAB problem without competition.
+We start with a pilot experiment in which we investigate each algorithm's performance ``in isolation": in a stand-alone MAB problem without competition.
 We focus on reputation scores generated by each algorithm.
 
-First, we confirm that algorithms' performance is ordered as we'd expect:
+First, we confirm that\gaedit{, for a sufficiently long time horizon,} algorithms' performance is ordered as we'd expect:
     $\TS \gg \DEG \gg \DG$.
@@
 
-The mean reputation trajectory is, arguably, the most natural way to represent algorithm's performance on a given problem instance.
+The mean reputation trajectory is, arguably, the most natural way to represent an algorithm's performance on a given problem instance.
 But is it the right statistic for our purposes?
@@
 \begin{conjecture}\label{conj:mean-trajectories}
-For a given MAB instance, if one algorithm's mean reputation trajectory lies above another, then the first algorithm prevails in the competition game.
+For a given MAB instance, if one algorithm's mean reputation trajectory lies above another at time $T_0$, then the first algorithm prevails in the competition game.
 \end{conjecture}
 
+We revisit this conjecture after documenting the results in the competition game.
+
+\OMIT{
 However, we find a more nuanced picture.
@@
 \end{finding}
+
+We will revisit this finding after presenting the results from the competition game.
 
@@
 
-To see what could go wrong, consider how algorithm's reputation score is distributed at a particular time.
+To see what could go wrong, consider how an algorithm's reputation score is distributed at a particular time.
 That is, consider the empirical distribution of this score over different \MRVs.
-For concreteness, consider the needle-in-haystack instance at time $t=500$ (Figure \ref{rep_dist_nih}).
+For concreteness, we consider the needle-in-haystack instance at time $t=500$ (Figure \ref{rep_dist_nih}).
 We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
 The reason is that for this MAB instance, $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
+This leads to two cases: $\DG$ either does slightly better than $\TS$ or where $\TS$ does substantially better than $\DG$.
+
 However, the mean reputation trajectory of an algorithm would summarize all this complexity the average over the \MRVs.
@@
 
+}
+
 \end{document}

### 2018-09-01 paper/content/revisited.tex
restructure the paper a bit

--- 2018-09-01 paper/content/perf_in_iso.tex
@@
 
-\section{Algorithms' Performance in Isolation}\label{section:4}
+\section{Algorithms' Performance in Isolation: Revisited}\label{section:revisited}
 
-We start with a pilot experiment in which we investigate each algorithm's performance ``in isolation": in a stand-alone MAB problem without competition.
-We focus on reputation scores generated by each algorithm.
+The results from the previous section present a few puzzles:
+\begin{enumerate}
+\item $\DG > \DEG$ on Heavy Tail even for large warm starts does not seem consistent with the ordering of the means.
+\item Why does $\TS$ beat $\DG$ for Needle In Haystack for low warm starts whereas on Uniform and Heavy Tail it does not?
+\end{enumerate}
 
-First, we confirm that\gaedit{, for a sufficiently long time horizon,} algorithms' performance is ordered as we'd expect:
-    $\TS \gg \DEG \gg \DG$.
-For each algorithm and each MAB instance, we compute the mean reputation score at each round, averaged over all \MRVs.
-We plot the \emph{mean reputation trajectory}: how this score evolves over time.
-Figure \ref{prelim_means} shows such plot for the Needle In Haystack instance (for other MAB instances the plots are similar).
-We summarize this finding as follows:
-
-\begin{finding}
-\textit{The mean reputation trajectories are arranged as predicted by prior work: 
-    $\TS > \DEG > \DG$.}
-\end{finding}
-
-\begin{figure}
-\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area displays 95\% confidence intervals.}
-\includegraphics[scale=0.35]{"figures/nih_iso_mean"}
-\label{prelim_means}
-%\caption*{\tiny{Mean reputation trajectory The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
-The shaded area display 95\% confidence intervals.}}
-\end{figure}
-
-The mean reputation trajectory is, arguably, the most natural way to represent an algorithm's performance on a given problem instance.
-But is it the right statistic for our purposes?
-That is, how well do these trajectories predict the outcomes of the competition game?
-The following conjecture is natural:
-
-\begin{conjecture}\label{conj:mean-trajectories}
-For a given MAB instance, if one algorithm's mean reputation trajectory lies above another at time $T_0$, then the first algorithm prevails in the competition game.
-\end{conjecture}
-
-We revisit this conjecture after documenting the results in the competition game.
-
-\OMIT{
-However, we find a more nuanced picture.
+We first attempt to explain the first puzzle. We revisit Conjecture~\ref{conj:mean-trajectories} and find a more nuanced picture.
 
@@
 
-We will revisit this finding after presenting the results from the competition game.
-
-\OMIT{ %%%%%
-\begin{finding}
-\textit{The mean performance of an algorithm is not a sufficient statistic for understanding the performance of an algorithm in competition.
-Looking at the distribution of reputation difference between algorithms we find that it tend is skewed to the right.
-Additionally we provide a counterexample where the mean performance does not predict the result of the competition game.}
-\end{finding}
-} %%%%%
-
-\begin{figure}[H]
+\begin{figure}[ht]
 \caption{Distribution of reputation scores for Needle-in-Haystack at $t=500$ (smoothed using a kernel density estimate)}
@@
 That is, consider the empirical distribution of this score over different \MRVs.
-For concreteness, we consider the needle-in-haystack instance at time $t=500$ (Figure \ref{rep_dist_nih}).
+For concreteness, we consider the needle-in-haystack instance at time $t=500$ (Figure \ref{rep_dist_nih}) since the intuition as to why the mean may be misleading is clearest on this instance (though the same result holds across the instances).
+
 We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
@@
 Figure~\ref{ts_dg_rep_diff_nih} shows such plots for several time points.
-We see that the distribution is skewed to the right, for which the mean is not a representative statistic.
-\ascomment{so what?}
 
-\OMIT{ %%%%%%%%
- or not.
-If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds.
-However, if it does not then it will get stuck on a bad arm and lose to $\TS$ or $\DEG$.
-In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \cite{CompetingBandits-itcs16}}.
-} %%%%%%%%
+We observe that the distribution is skewed to the right, precisely due to the fact that $\DG$ either does slightly better than $\TS$ or does substantially worse.
+This means that the mean is not a good measure of the central tendency, or typical value, of the reputation difference distribution and overstates how well an advanced exploration algorithm (that will eventually find the best arm) should do in competition compared to a naive algorithm that may never find the best arm.
 
-\OMIT{There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(\TS) - reputation(\DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the \textit{relative reputation proportion}.}
-
-\begin{figure}[H]
+\begin{figure}[ht]
 \caption{Distribution of reputation difference $\TS-\DG$ for the Needle-in-Haystack (smoothed via a kernel density estimate)}
@@
 
-\ascomment{Stopped here.}
-
 \begin{definition}
@@
 \textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $\TS$ doing worse than $\DG$ for small time horizons.
-We observe this under the Uniform and Heavy Tail priors.
-However, it is not observed under the Needle In Haystack prior}
+We observe this under the Uniform and Heavy Tail instances.
+However, it is not observed under the Needle In Haystack instances}
 \end{finding}
 
-The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $\DEG >\DG$ or $\TS > \DG$ according to the mean reputation but $\DG > \DEG$ or $\DG > TS$ according to the relative reputation proportion.
-In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $\DEG > \DG$ according to the mean reputation but $\DG > \DEG$ according to the relative reputation proportion.
-Additionally, these results carry over to the competition game}.
+The relative reputation proportion statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$.
 
@@
 
-Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $\DG$ on the Uniform and Needle In Haystack prior.
-For the Uniform prior we see that, in the early rounds, $\DG > \TS$ for the majority of the simulations but that, eventually, $\TS > \DG$.
+Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $\DG$ on the Uniform and Needle In Haystack instances.
+For the Uniform instance we see that, in the early rounds, $\DG > \TS$ for the majority of the simulations but that, eventually, $\TS > \DG$.
 The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information.
@@
 Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
-We call the priors where there is an early costly period followed by a later benefit period relative reputation costly priors.
+We call the priors where there is an early costly period followed by a later benefit period relative reputation costly instances.
 \end{definition}
 Is exploration always costly?
-Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $\TS$ always does relatively better than $\DG$.
+Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack instances, $\TS$ always does relatively better than $\DG$.
 There are two contributing factors to this.
@@
 However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
-Thus, only the Uniform and Heavy Tail priors are relative reputation costly priors.
+Thus, only the Uniform and Heavy Tail instances are relative reputation costly priors.
 
+This observation allows us to explain the second puzzle from the competition game with a more nuanced view of the inverted-U.
+
+\begin{finding}
+\textit{For sufficiently low warm start on relative reputation costly instances the optimal strategies in the competition game are:
+\begin{center}
+\textbf{Permanent Monopoly} - $\DG$ \\
+\textbf{Temporary Monopoly} - $\TS$ \\
+\textbf{Permanent Duopoly} - $\DG$ \\
+\end{center}
+The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
 }
+\end{finding}
+
+Thus, the reason we see that $\DG > \TS$ under Heavy Tail and Uniform but that $\TS$ is always preferred under Needle In Haystack is due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds leads it to suffer a relative reputational cost in the former case but not in the latter.
+Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $T_0$ as fixing the starting point on the relative reputation plots.
+The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $T_0$.
+As a result, if we begin with a warm start that falls in the region where the proportion $< 0.5$, we expect that $\DG > \TS$.
+
+This also allows us to give some intuition as to why allowing one firm to be a temporary monopoly incentivizes it to play $\TS$ even if, for the same warm start, the firm would be incentivized to play $\DG$ under permanent duopoly.
+By allowing it to be the only firm in the market for sufficiently long we allow it to recoup the ``reputation costs" by having it engage in most of its exploration while it does not have the pressure to incentivize agents to select them over their competition.
 

### 2018-09-01 analysis_code/compare_across.R

@@ -1,7 +1,7 @@
 library(dplyr)
 
-#dat_1 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
-#dat_2 <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv")
+dat_1 <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/longer_ws.csv")
+#dat_2 <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv")
 
 
 concise_alg_rep <- function(alg) {
@@ -70,7 +70,7 @@ ggplot(iso_deg, aes(rep_diff_dg)) + geom_density() +
 dist <- "Needle In Haystack"
 ws_time <- 20
 iso_dat <- filter(dat_1, Distribution == dist)
-t_vals <- c(500,1000, 2000)
+t_vals <- c(500,1000,2000)
 alg1 <- "TS"
 alg2 <- "DG"
 

### 2018-09-01 analysis_code/eeog.R

@@ -14,7 +14,7 @@ concise_alg_rep <- function(alg) {
   }
 }
 
-WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/"
+WORKING_PATH <- "/Users/garidor/Desktop/CompetingBandits/results/tournament_raw_results/"
 f <- "tournament_experiment_longer_ws_raw.csv"
 dat <- read.csv(file=paste(WORKING_PATH, f, sep=""))
 dat <- filter(dat, Time.Horizon == 2000)
@@ -27,7 +27,7 @@ algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
 for (prior in priors) {
   results <- matrix(nrow=3, ncol=length(warm_starts))
   rownames(results) <- c("TS vs DG", "TS vs DEG", "DG vs DEG")
-  colnames(results) <- lapply(warm_starts, function (start) { return(paste("k =", start))})
+  colnames(results) <- lapply(warm_starts, function (start) { return(paste("$T_0$ =", start))})
   for (i in 1:length(warm_starts)) {
     warm_start = warm_starts[i]
     for (k in 1:length(algs)) {

### 2018-09-01 analysis_code/relative_reputation_plots.R

@@ -3,7 +3,7 @@ library(ggplot2)
 library(reshape2)
 library(pBrackets)
 
-#dat <- read.csv("/Users/garidor/Desktop/bandits-rl-project/results/preliminary_raw_results/preliminary_plots_unified.csv")
+#dat <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/longer_ws.csv")
 
 
 bracketsGrob <- function(...){
@@ -79,17 +79,17 @@ print_mean_graphs <- function (dist, alg1, alg2, alg3, minComplexity) {
     cur_alg1 <- filter(alg_1, t == UQ(t))
     cur_alg2 <- filter(alg_2, t == UQ(t))
     cur_alg3 <- filter(alg_3, t == UQ(t))
-    df[nrow(df) + 1,] <-c(t, mean(cur_alg1$Realized.Reputation), 1.96 * sd(cur_alg1$Realized.Reputation) / sqrt(nrow(cur_alg1)), concise_alg_rep(alg1))
-    df[nrow(df) + 1,] <- c(t, mean(cur_alg2$Realized.Reputation), 1.96 * sd(cur_alg2$Realized.Reputation) / sqrt(nrow(cur_alg2)), concise_alg_rep(alg2))
-    df[nrow(df) + 1,] <-c(t, mean(cur_alg3$Realized.Reputation), 1.96 * sd(cur_alg3$Realized.Reputation) / sqrt(nrow(cur_alg3)), concise_alg_rep(alg3))
+    df[nrow(df) + 1,] <-c(t, mean(cur_alg1$Instantaneous.Realized.Reward.Mean), 1.96 * sd(cur_alg1$Instantaneous.Realized.Reward.Mean) / sqrt(nrow(cur_alg1)), concise_alg_rep(alg1))
+    df[nrow(df) + 1,] <- c(t, mean(cur_alg2$Instantaneous.Realized.Reward.Mean), 1.96 * sd(cur_alg2$Instantaneous.Realized.Reward.Mean) / sqrt(nrow(cur_alg2)), concise_alg_rep(alg2))
+    df[nrow(df) + 1,] <-c(t, mean(cur_alg3$Instantaneous.Realized.Reward.Mean), 1.96 * sd(cur_alg3$Instantaneous.Realized.Reward.Mean) / sqrt(nrow(cur_alg3)), concise_alg_rep(alg3))
   }
   df$t <- as.numeric(df$t)
   df$mean_rep <- as.numeric(df$mean_rep)
[…]

### 2018-09-01 analysis_code/tournament_table.R

@@ -1,15 +1,15 @@
 
 library(dplyr)
 library(knitr)
-WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
+WORKING_PATH <- "/Users/garidor/Desktop/CompetingBandits"
 
 
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_no_info_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_longer_ws_with_no_rep_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
 p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
-agent_algs - c("HardMax")
+agent_algs <- c("HardMax")
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
 time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior
@@ -51,7 +51,7 @@ for (j in 1:length(time_horizons)) {
           results[k, l] <- paste("\\makecell{\\textbf{", signif(mean(share), digits=2), "} $\\pm$", cin, "\\\\Var:", signif(var(share), digits=1), "\\\\", "ES:", 100*signif(shares, digits=2), "\\%}")
         }
       }
-      tab <-xtable(results, caption=paste("Temporary Monopoly Experiment", prior))
+      tab <-xtable(results, caption=paste("Information Erased Experiment", prior))
       print(tab, type="latex")
     }
   }

### 2018-09-01 paper/content/inverted_u.tex
editing...

@@
 
-In this section we document the results of the competition game for various parameterizations of our model.
-We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly.
+\swedit{Now we demonstrate the inverted-U relationship between the
+  incentivized algorithms and the competition levels in the game. In
+  the order of increasing competition levels, we study three different
+  settings: permanent monopoly, temporary monopoly, and permanent
+  duopoly. In more details:}
 
@@
 
-We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition.
+\swdelete{We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition.
 We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index).
@@
 Unless otherwise noted, all the results are reported at $t = 2000$.
+}
 
-\textbf{Permanent Monopoly} Under permanent monopoly only a single firm exists in the market for every period.
-In our model the firm only gets utility from a larger market share so it is indifferent between each algorithm in $\mathcal{A}$ since, regardless of the algorithm it deploys, it will get the entire market.
-If we suppose that deploying the ``better" algorithms has even an $\epsilon > 0$ cost or, equivalently, that firms break indifference towards algorithms that are easier to deploy, then the firm would choose to deploy $\DG$.
+\swedit{\xhdr{Permanent Monopoly.} There is only a single firm in the market
+for the entire period. In this case, the firm will can the entire
+market, regardless what algorithm it deploys. Assuming the deployment
+of any exploration algorithms will incur certain cost, then the firm
+would choose to deploy $\DG$.}
 
-\textbf{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously.
+\xhdr{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously.
 What algorithms are firms incentivized to deploy?
@@
 \end{document}
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../competing_bandits"
+%%% End:

### 2018-09-02 paper/content/inverted_u.tex

@@
 \section{Competition vs Better Algorithms}\label{section:5}
-
-\swedit{Now we demonstrate the inverted-U relationship between the
-  incentivized algorithms and the competition levels in the game. In
-  the order of increasing competition levels, we study three different
-  settings: permanent monopoly, temporary monopoly, and permanent
-  duopoly. In more details:}
+\swedit{how about just name the section ``inverted U''?}  \swedit{Now
+  we demonstrate the inverted-U relationship between the incentivized
+  algorithms and the competition levels. In particular, we investigate
+  the incentivized algorithms under three different regimes: permanent
+  monopoly, temporary monopoly, and permanent duopoly. First, the
+  story for permanent monopoly is relatively straightforward.}
 
@@
 
-\swedit{\xhdr{Permanent Monopoly.} There is only a single firm in the market
-for the entire period. In this case, the firm will can the entire
-market, regardless what algorithm it deploys. Assuming the deployment
-of any exploration algorithms will incur certain cost, then the firm
+\swedit{\xhdr{Permanent Monopoly.} Since there is only a single firm
+  in the market for the entire period, the firm can take the entire
+  market regardless what algorithm it deploys. Assuming the deployment
+  of any exploration algorithms will incur certain cost, the firm
   would choose to deploy $\DG$.}
+
+Next, to describe our results for temporary monopoly, we will
+introduce the following notion of \textit{effective end of game
+  (EEOG)} --- the last round $t$ such that the agent at round $(t-1)$
+and the next agent around at $t$ choose different firms.
 
@@
 
+\OMIT{
 \begin{definition}
@@
 \end{definition}
+}
 

### 2018-09-02 analysis_code/tournament_table.Rmd
update path for analysis code

--- 2018-08-15 analysis_code/tournament_table.Rmd
@@
 library(knitr)
-WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project"
+WORKING_PATH <- "/Users/garidor/Desktop/CompetingBandits"
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ht_lhr_3_arms_raw.csv", sep=""))
+#dat <- read.csv(file=paste(WORKING_PATH, "/results/tournament_raw_results/tournament_experiment_ht_lhr_3_arms_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_full_sim_no_info_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
@@
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
-time_horizons <- c(7000)
+time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior
+priors <- c("Needle In Haystack")
 

### 2018-09-02 analysis_code/relative_reputation_plots.R

@@ -3,7 +3,7 @@ library(ggplot2)
 library(reshape2)
 library(pBrackets)
 
-#dat <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/longer_ws.csv")
+dat <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/ht_3_arm_many_sim.csv")
 
 
 bracketsGrob <- function(...){
@@ -79,17 +79,17 @@ print_mean_graphs <- function (dist, alg1, alg2, alg3, minComplexity) {
     cur_alg1 <- filter(alg_1, t == UQ(t))
     cur_alg2 <- filter(alg_2, t == UQ(t))
     cur_alg3 <- filter(alg_3, t == UQ(t))
-    df[nrow(df) + 1,] <-c(t, mean(cur_alg1$Instantaneous.Realized.Reward.Mean), 1.96 * sd(cur_alg1$Instantaneous.Realized.Reward.Mean) / sqrt(nrow(cur_alg1)), concise_alg_rep(alg1))
-    df[nrow(df) + 1,] <- c(t, mean(cur_alg2$Instantaneous.Realized.Reward.Mean), 1.96 * sd(cur_alg2$Instantaneous.Realized.Reward.Mean) / sqrt(nrow(cur_alg2)), concise_alg_rep(alg2))
-    df[nrow(df) + 1,] <-c(t, mean(cur_alg3$Instantaneous.Realized.Reward.Mean), 1.96 * sd(cur_alg3$Instantaneous.Realized.Reward.Mean) / sqrt(nrow(cur_alg3)), concise_alg_rep(alg3))
+    df[nrow(df) + 1,] <-c(t, mean(cur_alg1$Realized.Reputation), 1.96 * sd(cur_alg1$Realized.Reputation) / sqrt(nrow(cur_alg1)), concise_alg_rep(alg1))
+    df[nrow(df) + 1,] <- c(t, mean(cur_alg2$Realized.Reputation), 1.96 * sd(cur_alg2$Realized.Reputation) / sqrt(nrow(cur_alg2)), concise_alg_rep(alg2))
+    df[nrow(df) + 1,] <-c(t, mean(cur_alg3$Realized.Reputation), 1.96 * sd(cur_alg3$Realized.Reputation) / sqrt(nrow(cur_alg3)), concise_alg_rep(alg3))
   }
   df$t <- as.numeric(df$t)
[…]

### 2018-09-02 analysis_code/tournament_table.R

@@ -5,7 +5,7 @@ WORKING_PATH <- "/Users/garidor/Desktop/CompetingBandits"
 
 
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_longer_ws_with_no_rep_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_longer_ws_no_info_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
 p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
@@ -13,7 +13,7 @@ agent_algs <- c("HardMax")
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
 time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior
-priors <- c("Heavy Tail")
+#priors <- c("Heavy Tail")
 warm_start <- 20
 
 ZERO_ONE_CUTOFF <- 0.1

### 2018-09-02 paper/competing_bandits.tex
Added \MRV macro.
Fixed the preamble as per the AAAI reqs, added section numbers.
restructure the paper a bit
attempt to make the figures look better

@@
 \usepackage{url}
-\usepackage{subfiles}
 \usepackage{graphicx}
-\usepackage{tikz} % SW: Added for plots
 \frenchspacing
+% Add additional packages here. The following
+% packages may NOT be used  (this list
+% is not exhaustive:
+% authblk, caption, CJK, float, fullpage, geometry,
+%hyperref, layout, nameref, natbib, savetrees,
+%setspace, titlesec, tocbibind, ulem
+%
+%US Lettersize Paper Is Required
+\setlength{\pdfpagewidth}{8.5in}
+\setlength{\pdfpageheight}{11in}
+%
+%
+% PDFINFO
+% You are required to complete the following
+% for pass-through to the PDF.
+% No LaTeX commands of any kind may be
+% entered. The parentheses and spaces
+% are an integral part of the
+% pdfinfo script and must not be removed.
+%
+\pdfinfo{
+/Title  (Competing Bandits: The Perils of Exploration under Competition)
+/Author (Guy Aridor, Kevin Liu, Aleksandrs Slivkins, Zhiwei Steven Wu)
+/Keywords (Input your keywords in this optional area)
+}
+%
+%Section Numbers
+% Uncomment if you want to use section numbers
+% and change the 0 to a 1 or 2
+ \setcounter{secnumdepth}{2} % section & subsection numbers
+
+% Title and Author Information Must Immediately Follow
+% the pdfinfo within the preamble
+%
 
@@
 %\usepackage{hyperref}       % hyperlinks
+\usepackage{tikz} % SW: Added for plots
+\usepackage{subfiles}
 \usepackage{url}            % simple URL typesetting
@@
 \newtheorem{definition}{Definition}
-\newtheorem{finding}{Finding}
+\newtheorem{finding}[definition]{Finding}
+\newtheorem{conjecture}[definition]{Conjecture}
 \captionsetup{font=footnotesize}
@@
 \newcommand{\eps}{\varepsilon}
+
+\newcommand{\MRV}{mean reward vector\xspace} % mean reward vector
+\newcommand{\MRVs}{mean reward vectors\xspace} % mean reward vector
 
@@
 
+\subfile{content/revisited}
+
 \subfile{content/barriers}
@@
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2018-09-02 paper/content/barriers.tex
attempt to make the figures look better

@@
 \centering
-\caption{Reputation Erased Experiment, Heavy Tail}
-\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
+\caption{Information Erased Experiment Heavy Tail} 
+\begin{tabular}{c@{\hspace{0.1\tabcolsep}}ccc}
+\hline
+\multicolumn{4}{c}{Incumbent Algorithm}\\
+\multirow{12}{0.6in}{\rotatebox{90}{Entrant Algorithm}} \\
   \hline
@@
   \hline
-TS & \makecell{\textbf{ 0.016 } $\pm$ 0.0075 \\Var: 0.01 \\ ES: 100 \%} & \makecell{\textbf{ 0.13 } $\pm$ 0.02 \\Var: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.2 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 96 \%} \\
-  DEG & \makecell{\textbf{ 0.068 } $\pm$ 0.013 \\Var: 0.05 \\ ES: 93 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 75 \%} & \makecell{\textbf{ 0.26 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 80 \%} \\
-   DG & \makecell{\textbf{ 0.15 } $\pm$ 0.019 \\Var: 0.1 \\ ES: 87 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.028 \\Var: 0.2 \\ ES: 80 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.024 \\Var: 0.2 \\ ES: 67 \%} \\
+TS & \makecell{\textbf{0.021} $\pm$0.009\\Var:0.02\\ES:100\%} & \makecell{\textbf{0.16} $\pm$0.02\\Var:0.1\\ES:97\%} & \makecell{\textbf{0.21} $\pm$0.02\\Var:0.2\\ES:95\%} \\ 
+  DEG & \makecell{\textbf{0.26} $\pm$0.03\\Var:0.2\\ES:95\%} & \makecell{\textbf{0.3} $\pm$0.02\\Var:0.2\\ES:74\%} & \makecell{\textbf{0.26} $\pm$0.02\\Var:0.1\\ES:76\%} \\ 
+   DG & \makecell{\textbf{0.34} $\pm$0.03\\Var:0.2\\ES:94\%} & \makecell{\textbf{0.4} $\pm$0.03\\Var:0.2\\ES:74\%} & \makecell{\textbf{0.33} $\pm$0.02\\Var:0.1\\ES:58\%} \\ 
    \hline
-   \label{rep_erase}
+  \label{info_erase}
 \end{tabular}
 \end{table}
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Sun Jul 22 14:23:06 2018
+
 \begin{table}[ht]
 \centering
-\caption{Information Erased Experiment, Heavy Tail}
-\begin{tabular}{c@{\hspace{1.0\tabcolsep}}ccc}
+\caption{Reputation Erased Experiment Heavy Tail} 
+\begin{tabular}{c@{\hspace{0.01\tabcolsep}}ccc}
+\hline
+\multicolumn{4}{c}{Incumbent Algorithm}\\
+\multirow{12}{0.6in}{\rotatebox{90}{Entrant Algorithm}} \\
   \hline
@@
   \hline
-TS & \makecell{\textbf{ 0.024 } $\pm$ 0.0094 \\Var: 0.02 \\ ES: 100 \%} & \makecell{\textbf{ 0.16 } $\pm$ 0.022 \\Var: 0.1 \\ ES: 97 \%} & \makecell{\textbf{ 0.22 } $\pm$ 0.025 \\Var: 0.2 \\ ES: 95 \%} \\
-  DEG & \makecell{\textbf{ 0.24 } $\pm$ 0.025 \\Var: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.29 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 72 \%} & \makecell{\textbf{ 0.27 } $\pm$ 0.024 \\Var: 0.1 \\ ES: 76 \%} \\
-   DG & \makecell{\textbf{ 0.33 } $\pm$ 0.028 \\Var: 0.2 \\ ES: 94 \%} & \makecell{\textbf{ 0.38 } $\pm$ 0.026 \\Var: 0.2 \\ ES: 74 \%} & \makecell{\textbf{ 0.33 } $\pm$ 0.023 \\Var: 0.1 \\ ES: 58 \%} \\
+TS & \makecell{\textbf{0.0096} $\pm$0.006\\Var:0.009\\ES:100\%} & \makecell{\textbf{0.11} $\pm$0.02\\Var:0.09\\ES:98\%} & \makecell{\textbf{0.18} $\pm$0.02\\Var:0.1\\ES:95\%} \\ 
+  DEG & \makecell{\textbf{0.073} $\pm$0.01\\Var:0.05\\ES:93\%} & \makecell{\textbf{0.29} $\pm$0.02\\Var:0.2\\ES:78\%} & \makecell{\textbf{0.25} $\pm$0.02\\Var:0.1\\ES:79\%} \\ 
+   DG & \makecell{\textbf{0.15} $\pm$0.02\\Var:0.1\\ES:89\%} & \makecell{\textbf{0.39} $\pm$0.03\\Var:0.2\\ES:78\%} & \makecell{\textbf{0.33} $\pm$0.02\\Var:0.2\\ES:66\%} \\ 
    \hline
-   \label{info_erase}
+    \label{rep_erase}
 \end{tabular}

### 2018-09-02 paper/content/inverted_u.tex
attempt to make the figures look better

--- 2018-09-01 paper/content/inverted_u.tex
@@
 \hline
+\multicolumn{4}{c}{Incumbent Algorithm}\\
+\multirow{12}{0.6in}{\rotatebox{90}{Entrant Algorithm}} \\
+  \hline
  & TS & DEG &  DG \\
@@
 \label{ht_incum}
-\caption*{\tiny{The columns represent the strategy of the incumbent and the rows represent the strategy of the entrant.
-The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval.
-The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one firm getting $> 90\%$ of the market.
+\caption*{\tiny{The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval.
+The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one firm getting an ``extreme" market share of $> 90\%$.
 Note that smaller values in the table are better for the incumbent.
 Market shares are calculated as the fraction of users selecting a particular firm \textit{after} the entrant has already entered (i.e.
-the free rounds to firm 2 do not count towards the share)}}
+the free rounds to the incumbent do not count towards the share)}}
 \end{table}

### 2018-09-02 paper/content/revisited.tex
attempt to make the figures look better

@@
 \begin{figure}[ht]
-\caption{Relative Reputation Plots}
+\caption{Relative Reputation Proportion Plots for $\TS$ vs $\DG$. Shaded area display 95\% confidence intervals.}
 \includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
-\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
-\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.
+\includegraphics[scale=0.35]{figures/relative_nih_ts_dg.png}
+%\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.
 The shaded area display 95\% confidence intervals.}}

### 2018-09-02 analysis_code/compare_across.R

@@ -1,6 +1,7 @@
 library(dplyr)
 
 dat_1 <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/longer_ws.csv")
+dat_1$Algorithm<- sapply(dat_1$Algorithm, concise_alg_rep)
 #dat_2 <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv")
 
 
@@ -14,9 +15,7 @@ concise_alg_rep <- function(alg) {
   }
 }
 
-dat_1$Algorithm<- sapply(dat_1$Algorithm, concise_alg_rep)
-tmp <- mutate(dat_1, Algorithm = concise_alg_rep(Algorithm))
-dist <- ".5/.7 Random Draw"
+dist <- "Uniform"
 ws_time <- 20
 iso_dat <- filter(dat_1, Distribution == dist)
 compet_dat <- filter(dat_2, Prior == dist & Warm.Start == 20)
@@ -54,7 +53,7 @@ iso_dat_t <- filter(iso_dat, t == 500)
 ggplot(iso_dat_t, aes(Realized.Reputation, colour=Algorithm)) +
   geom_density() +
   ggtitle("Reputation Distribution, Needle In Haystack") +
-  theme_bw(base_size = 12) +
+  theme_bw(base_size = 16) +
   xlab("Reputation") + 
   theme(plot.title = element_text(hjust = 0.5))
 
@@ -82,6 +81,6 @@ title <- paste(alg1, "-", alg2, "Reputation Distribution,", dist)
 ggplot(iso_alg1, aes(rep_diff)) + geom_density(aes(group=t, colour=t)) + 
   ggtitle(title) + 
   xlab("Reputation Difference") +
-  theme_bw(base_size = 12) +
+  theme_bw(base_size = 16) +
   theme(plot.title = element_text(hjust = 0.5)) +
   xlim(c(-0.25, 0.25))

### 2018-09-02 analysis_code/relative_reputation_plots.R

@@ -3,7 +3,7 @@ library(ggplot2)
 library(reshape2)
 library(pBrackets)
 
-dat <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/ht_3_arm_many_sim.csv")
+#dat <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/longer_ws.csv")
 
 
 bracketsGrob <- function(...){
@@ -53,14 +53,13 @@ print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
   q <- ggplot(df, aes(x=t, y=relative_rep)) + geom_line() + geom_point() +
     geom_errorbar(aes(ymin=relative_rep-1.96*se, ymax=relative_rep+1.96*se), width=.2) +
     ggtitle(plot_title) + xlab("time") + 
-    theme_grey(base_size = 12) +
     ylab(paste(concise_alg_rep(alg1), " >= ", concise_alg_rep(alg2), "(reputation)")) +
-    #annotation_custom(b1) +
-    #annotation_custom(b2) +
+    annotation_custom(b1) +
+    annotation_custom(b2) +
     ylim(c(0.2, 0.7)) +
-    #annotate("text", x = 225, y = 0.25, label = "Relative Reputation Cost") +
-    #annotate("text", x = 1000, y = 0.25, label = "Relative Reputation Benefit") +
-    theme_bw() +  
+    annotate("text", x = 225, y = 0.25, label = "Relative Reputation Cost") +
+    annotate("text", x = 1000, y = 0.25, label = "Relative Reputation Benefit") +
+    theme_bw(base_size = 16) +  
     theme(plot.title = element_text(hjust = 0.5))
   print(q)
 }
@@ -88,23 +87,23 @@ print_mean_graphs <- function (dist, alg1, alg2, alg3, minComplexity) {
   df$ci <- as.numeric(df$ci)
 
   plot_title <- paste("Mean Reputation -", dist)
-  q <- ggplot(data=df, aes(x=t, y=mean_rep, colour=alg)) + geom_smooth() +#geom_line() +
+  q <- ggplot(data=df, aes(x=t, y=mean_rep, colour=alg)) + geom_line() +
     geom_errorbar(aes(ymin=mean_rep-ci, ymax=mean_rep+ci), width=.2) +
     ggtitle(plot_title) + xlab("time") +
[…]

### 2018-09-02 analysis_code/tournament_table.R

@@ -5,7 +5,7 @@ WORKING_PATH <- "/Users/garidor/Desktop/CompetingBandits"
 
 
 
-dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_longer_ws_no_info_raw.csv", sep=""))
+dat <- read.csv(file=paste(WORKING_PATH, "/results/free_obs_raw_results/free_obs_experiment_longer_ws_no_rep_raw.csv", sep=""))
 p1_algs <- as.list(unique(dat['P1.Alg']))$P1.Alg
 p2_algs <- as.list(unique(dat['P2.Alg']))$P2.Alg
 agent_algs <- as.list(unique(dat['Agent.Alg']))$Agent.Alg
@@ -13,7 +13,7 @@ agent_algs <- c("HardMax")
 time_horizons <-  as.list(unique(dat['Time.Horizon']))$Time.Horizon
 time_horizons <- c(2000)
 priors <- as.list(unique(dat['Prior']))$Prior
-#priors <- c("Heavy Tail")
+priors <- c("Heavy Tail")
 warm_start <- 20
 
 ZERO_ONE_CUTOFF <- 0.1
@@ -48,10 +48,10 @@ for (j in 1:length(time_horizons)) {
           cv_low <- qchisq(.025, df=length(share) - 1)
           lower_var <- signif((length(share) - 1) * var(share) / cv_high, digits=2)
           high_var <- signif((length(share) - 1) * var(share) / cv_low, digits=2)
-          results[k, l] <- paste("\\makecell{\\textbf{", signif(mean(share), digits=2), "} $\\pm$", cin, "\\\\Var:", signif(var(share), digits=1), "\\\\", "ES:", 100*signif(shares, digits=2), "\\%}")
+          results[k, l] <- paste("\\makecell{\\textbf{", signif(mean(share), digits=2), "} $\\pm$", cin, "\\\\Var:", signif(var(share), digits=1), "\\\\", "ES:", 100*signif(shares, digits=2), "\\%}", sep="")
         }
       }
-      tab <-xtable(results, caption=paste("Information Erased Experiment", prior))
+      tab <-xtable(results, caption=paste("Reputation Erased Experiment", prior))
       print(tab, type="latex")
     }
   }

### 2018-09-02 paper/content/inverted_u.tex
add back steven's changes

--- 2018-09-02 paper/content/inverted_u.tex
@@
 \begin{document}
-
 \section{Competition vs Better Algorithms}\label{section:5}
@@
   story for permanent monopoly is relatively straightforward.}
-
-\OMIT{main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \cite{Aghion-QJE05}. 
-}
-
-\OMIT{
-\begin{finding}
-\textit{For sufficiently low warm start, the optimal strategies in the competition game are:
-\begin{center}
-\textbf{Permanent Monopoly} - $\DG$ \\
-\textbf{Temporary Monopoly} - $\TS$ \\
-\textbf{Permanent Duopoly} - $\DG$ \\
-\end{center}
-The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
-}
-\end{finding}
-}
 
@@
 
+\swdelete{We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition.
+We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index).
+In our model we have no prices and no costs and so we interpret increases in ``market power" as decreases in the number of rounds where a firm has to compete for agents.
+
+% couldn't get section references to work so hardcoding for now
+We utilize the same instances and realizations as in the section 3  and simulate the model described in section 2.
+We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the expected payoffs associated with each pair of algorithms.
+Unless otherwise noted, all the results are reported at $t = 2000$.
+
+\xhdr{Permanent Monopoly} Under permanent monopoly only a single firm exists in the market for every period.
+In our model the firm only gets utility from a larger market share so it is indifferent between each algorithm in $\mathcal{A}$ since, regardless of the algorithm it deploys, it will get the entire market.
+If we suppose that deploying the ``better" algorithms has even an $\epsilon > 0$ cost or, equivalently, that firms break indifference towards algorithms that are easier to deploy, then the firm would choose to deploy $\DG$.
+}
+
 \xhdr{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously.
@@
 }
-
 Since both firms enter at the same time the game is symmetric and we only need to compute the payoffs of three pairs of strategies.
@@
 
-\textbf{Temporary Monopoly} We now consider asymmetries in the timing of entry so that one firm enters the market before the other and serves as a monopolist in the periods until the other firm enters.
+\xhdr{Temporary Monopoly} We now consider asymmetries in the timing of entry so that one firm enters the market before the other and serves as a monopolist in the periods until the other firm enters.
 In terms of our model, this corresponds to varying the value of $X$.
@@
 \end{document}
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "../competing_bandits"
-%%% End:

### 2018-09-02 paper/content/barriers.tex
basically rewrite barriers section, tweaks to inverted U and sim details

@@
 
-An alternative interpretation of the results in the previous section is that the "temporary monopoly" provides a first mover advantage to the incumbent firm and that this first mover advantage allows the firm to both get a data advantage as well as a reputational advantage over the entrant.
-This provides an alternative interpretation of the classic "first mover advantage" that has been well-studied in economics and marketing \cite{kerin1992first} where, in our model, an incumbent can use data and reputation as a barrier to entry simply due to the fact that they were in the market before the entrant.
-Which plays a bigger role in preventing the entrant from being able to establish market share?
-We run two additional experiments, modifying the previous temporary monopoly experiment so that in one experiment the reputation of the incumbent is artificially erased (Reputation Erased Experiment) and another in which the information gained by the incumbent is artificially erased so that the posterior is reset to the prior (Information Erased Experiment).
+An alternative interpretation of the results in the previous section is that the temporary monopoly provides a first mover advantage to the incumbent firm.
+The early entry allows the incumbent to acquire data on potentially suboptimal actions without incurring an immediate reputational consequence as well as the ability to eventually build up a high reputation before the entrant enters the market.
+Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant.
+This gives new intuition for the classic ``first mover advantage" \cite{kerin1992first} in the digital economy where data can serve as a barrier to entry in many markets.
+
+While combined these two advantages serve as strong barriers to entry a natural question to ask is would having only the data advantage or only the reputation advantage still serve as such strong barriers to entry?
+Which of the two serves as a stronger barrier to entry and under what conditions?
+In order to answer these questions we run two additional experiments, artificially modifying the previous temporary monopoly experiment.
+In one experiment the reputation of the incumbent is artificially erased upon the entry of the entrant (Reputation Erased Experiment) and in the other the information gained by the incumbent is artificially erased so that the posterior is reset to the prior (Information Erased Experiment).
 
@@
 \begin{finding}
-\textit{Both the data advantage and reputation advantage alone serve as strong barriers to entry.
-When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $DG$ there is no substantial difference between the two.}
+\textit{Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
+\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two.}
 \end{finding}
 
-Tables \ref{rep_erase} and \ref{info_erase} display the results of these experiments.
-Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
-The main interesting finding here, which is robust across priors for this parameterization, is that information serves as a larger barrier to entry than reputation when the incumbent plays $\TS$.
-One possible intuition for this is that, for Heavy Tail, the information acquired after the $X$ rounds under $\TS$ is sufficient so that it becomes relatively easier for $\TS$ to recover the reputation loss compared to the information loss, but since $\DEG$ and $DG$ learn slower and they've acquired less information over the $X$ rounds relative to $\TS$ there is less of an informational advantage.
+Tables \ref{info_erase} and \ref{rep_erase} display the results of these experiments.
+The reason for the finding is that, under $\TS$, the early entry allows the incumbent to do exploration without immediate reputational consequence in the early rounds and so, even when reputation is erased, the incumbent has already acquired sufficient information that simply exploiting that information allows her to quickly recover her reputation.
+However, while the reputational advantage gives the incumbent some cushion for doing exploration the incumbent is still vulnerable to getting unlucky while doing exploration and suffering the reputational consequences.
+This only happens under $\TS$ and not under $\DEG$ or $\DG$ since $\TS$, by construction, is more intelligent about the data that it acquires compared to $\DEG$ and $\DG$.
 
 Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
-We find that the best responses remain the same for the Heavy Tail prior, but for the Uniform prior they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.
+We find that the best responses remain the same for the Heavy Tail instance, but for the Uniform instance they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.
 

### 2018-09-02 paper/content/inverted_u.tex
basically rewrite barriers section, tweaks to inverted U and sim details

@@
   the incentivized algorithms under three different regimes: permanent
-  monopoly, temporary monopoly, and permanent duopoly. First, the
-  story for permanent monopoly is relatively straightforward.}
-
-\swdelete{We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition.
-We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index).
-In our model we have no prices and no costs and so we interpret increases in ``market power" as decreases in the number of rounds where a firm has to compete for agents.
-
-% couldn't get section references to work so hardcoding for now
-We utilize the same instances and realizations as in the section 3  and simulate the model described in section 2.
-We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the expected payoffs associated with each pair of algorithms.
-Unless otherwise noted, all the results are reported at $t = 2000$.
-}
+  monopoly, temporary monopoly, and permanent duopoly.
+\gaedit{We break indifference between exploration algorithms and non-exploration algorithms towards non-exploration algorithms.} First, the story for permanent monopoly is relatively straightforward.}
 
@@
 We report the average market share taken over the $N$ simulations for a set of exogenous values of $T_0$.
-Additionally, we report the mean and median of an additional quantity which we define as the \textit{effective end of game}.
+Additionally, we report the mean and median of the effective end of game.
 
@@
 \item Under Needle In Haystack, \\ $\TS > \DG$ (see Table \ref{sim_nih}). 
-\item For sufficiently large warm starts, $\TS$ is dominant across all the families of instances.
+\item For sufficiently large warm starts, $\TS$ is dominant across all instances.
 \item Even for a large warm start, $\DG > \DEG$ under Heavy Tail
@@
 
-In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the Heavy Tail family \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.
-However, while $\DG > \TS$ for the entrant, $\DG$ does roughly the same as $\DEG$ on the Uniform family.
-The results are in the supplementary material.}.
+In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the Heavy Tail family \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.}.
 In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the family of instances we consider.
@@
 The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.
+
 \OMIT{As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.} Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.

### 2018-09-02 paper/content/inverted_u.tex
edits

@@
 \section{Competition vs Better Algorithms}\label{section:5}
-\swedit{how about just name the section ``inverted U''?}  \swedit{Now
+\swcomment{how about just name the section ``inverted U''?}  \swedit{Now
   we demonstrate the inverted-U relationship between the incentivized

### 2018-09-02 paper/content/inverted_u.tex
small change

@@
 Next, to describe our results for temporary monopoly, we will
-introduce the following notion of \textit{effective end of game
-  (EEOG)} --- the last round $t$ such that the agent at round $(t-1)$
-and the next agent around at $t$ choose different firms.
+introduce the notion of \textit{effective end of game (EEOG)} --- the
+last round $t$ such that the agent at round $(t-1)$ and the next agent
+around at $t$ choose different firms.
 

### 2018-09-02 paper/content/barriers.tex
edits

@@
 
-While combined these two advantages serve as strong barriers to entry a natural question to ask is would having only the data advantage or only the reputation advantage still serve as such strong barriers to entry?
-Which of the two serves as a stronger barrier to entry and under what conditions?
-In order to answer these questions we run two additional experiments, artificially modifying the previous temporary monopoly experiment.
-In one experiment the reputation of the incumbent is artificially erased upon the entry of the entrant (Reputation Erased Experiment) and in the other the information gained by the incumbent is artificially erased so that the posterior is reset to the prior (Information Erased Experiment).
+While combined these two advantages serve as strong barriers to entry,
+a natural question to ask is would having only the data advantage or
+only the reputation advantage still serve as such strong barriers to
+entry? Which of the two serves as a stronger barrier to entry and
+under what conditions?
+
+\swedit{To answer these questions, we isolate the effects from the two
+  advantages and run two additional experiments. In the first
+  experiment (called \emph{reputation erased experiment}), the
+  reputation of the incumbent is artificially erased upon the entry of
+  the entrant. In the second experiment (called \emph{information
+    erased experiment}), the information gained by the incumbent is
+  erased in the sense that its posterior is reset to the initial
+  (fake) prior upon the entry of the entrant.}
+
+\swcomment{Guy, could you make it clear what the first ``erase'' mean? Like reset to default?}
 
@@
 \textit{Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
-\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two.}
+\swcomment{too vague.
+what does significant mean?}\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two.}
 \end{finding}

### 2018-09-02 paper/content/inverted_u.tex

@@
 \begin{document}
-\section{Competition vs Better Algorithms}\label{section:5}
-\swcomment{how about just name the section ``inverted U''?}  \swedit{Now
-  we demonstrate the inverted-U relationship between the incentivized
+\section{Competition vs Better Algorithms: Inverted-U}\label{section:5}
+ Now we demonstrate the inverted-U relationship between the incentivized
   algorithms and the competition levels. In particular, we investigate
@@
   monopoly, temporary monopoly, and permanent duopoly.
-\gaedit{We break indifference between exploration algorithms and non-exploration algorithms towards non-exploration algorithms.} First, the story for permanent monopoly is relatively straightforward.}
+\gaedit{We break indifference between exploration algorithms and non-exploration algorithms towards non-exploration algorithms due to deployment costs.} \gadelete{First, the story for permanent monopoly is relatively straightforward.}
+ We summarize the relationship between the different competition levels and incentivized algorithms as follows:
 
-\swedit{\xhdr{Permanent Monopoly.} Since there is only a single firm
+\begin{conjecture}\label{conj:inverted-u}
+\textit{For sufficiently low warm start,
+ the optimal strategies in the competition game are:
+\begin{center}
+\textbf{Permanent Monopoly} - $\DG$ \\
+\textbf{Temporary Monopoly} - $\TS$ \\
+\textbf{Permanent Duopoly} - $\DG$ \\
+\end{center}
+The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
+}
+\end{conjecture}
+
+\xhdr{Permanent Monopoly.} Since there is only a single firm
   in the market for the entire period, the firm can take the entire
-  market regardless what algorithm it deploys. Assuming the deployment
-  of any exploration algorithms will incur certain cost, the firm
-  would choose to deploy $\DG$.}
+  market regardless what algorithm it deploys.
+\gadelete{Assuming the deployment of any exploration algorithms will incur certain cost} \gaedit{Since we break indifferences by supposing the firm prefers easier to deploy algorithms}, the firm would choose to deploy $\DG$.
 
@@
 
-\swdelete{We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition.
-We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index).
-In our model we have no prices and no costs and so we interpret increases in ``market power" as decreases in the number of rounds where a firm has to compete for agents.
-
-% couldn't get section references to work so hardcoding for now
-We utilize the same instances and realizations as in the section 3  and simulate the model described in section 2.
-We initially take the strategies of the firms as exogenous (chosen from $\mathcal{A}$) and simulate the model given these strategies in order to determine the expected payoffs associated with each pair of algorithms.
-Unless otherwise noted, all the results are reported at $t = 2000$.
-
-\xhdr{Permanent Monopoly} Under permanent monopoly only a single firm exists in the market for every period.
-In our model the firm only gets utility from a larger market share so it is indifferent between each algorithm in $\mathcal{A}$ since, regardless of the algorithm it deploys, it will get the entire market.
-If we suppose that deploying the ``better" algorithms has even an $\epsilon > 0$ cost or, equivalently, that firms break indifference towards algorithms that are easier to deploy, then the firm would choose to deploy $\DG$.
-}
-
 \xhdr{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously.
@@
 
+Since both firms enter at the same time the game is symmetric and the competition game is zero-sum so we only need to compute the expected payoffs of three pairs of strategies.
+We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.
+Since our results are from simulations, we interpret algorithm $A$ being better than algorithm $B$ if the resulting market share of $A$ is statistically greater than $0.5$.
+Otherwise, the firm is indifferent between $A$ and $B$ with ties broken towards algorithms that are easier to deploy.
+
+Tables \ref{sim_ht} and \ref{sim_nih} show the results of the simulations.
+For the most part Conjecture~\ref{conj:inverted-u} is validated where, for the Heavy Tail and Uniform instances, we see that $\DG$ does better than $\TS$ and is a weakly dominant strategy in the game.
+However, there are two results that are at odds with the conjecture:
+
+\begin{puzzle}\label{puz:dg_ht}
+Even for large warm start, $\DG$ is better than $\DEG$ under Heavy Tail
+\end{puzzle}
+
+\begin{puzzle}\label{puz:nih}
+Under Needle In Haystack, $\TS$ is weakly dominant across all warm starts
+\end{puzzle}
+
 \OMIT{
-\begin{definition}
-\textit{Effective End of Game (EEOG)} - the last round, $t$, in a simulation where the agent alive at $t-1$ and the agent alive at $t$ choose different firms.
-\end{definition}
-}
-Since both firms enter at the same time the game is symmetric and we only need to compute the payoffs of three pairs of strategies.
-We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.
-
 \begin{finding}
@@
 \begin{enumerate}
-\item For low warm starts, under Heavy Tail and Uniform \\ $\DG > \TS$ in competition (see Table \ref{sim_ht}),
-\item Under Needle In Haystack, \\ $\TS > \DG$ (see Table \ref{sim_nih}). 
+\item For low warm starts, under Heavy Tail and Uniform \\ $\DG$ is a weakly dominant strategy in competition (see Table \ref{sim_ht}),
+\item  (see Table \ref{sim_nih}). 
 \item For sufficiently large warm starts, $\TS$ is dominant across all instances.
-\item Even for a large warm start, $\DG > \DEG$ under Heavy Tail
+\item Even for a large warm start, $\DG$ leads to a higher expected market share than $\DEG$ under Heavy Tail
 \end{enumerate}
@@
 
-Combining this finding with the fact that the EEOG values are relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.
+We defer the explanation for these puzzles until the next section.
+Looking at the EEOG values we note that, even though the game continues until $t=2000$, the game effectively ends much earlier than that.
+This seems to imply that when attempting to understand the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotic performance.
+This is due to the fact that the losing firm may enter into a ``death spiral" where the winner keeps learning and improving from its customers thus increasing its reputation while the losing firm gets starved of customers and progressively falls farther and farther behind its competitors.
 
@@
 \label{sim_nih}
-\caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95 \% confidence band.
+\caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95\% confidence band.
 For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
@@
 In terms of our model, this corresponds to varying the value of $X$.
-For this section we report results fixing $T_0 = 20$ and focus on the family of instances where we observed that, for this warm start, $\DG > \TS$ \footnote{However, we report the results for the Needle In Haystack family in the supplementary material.
+For this section we report results fixing $T_0 = 20$ and focus on the family of instances where we observed that, for this warm start, $\DG$ was the optimal strategy \footnote{However, we report the results for the Needle In Haystack instances in the supplementary material.
 The results are unsurprising, where $\TS$ is the dominant strategy for both the entrant and incumbent.}.
@@
 
-In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the Heavy Tail family \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.}.
-In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the family of instances we consider.
+In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the Heavy Tail instances \footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.}.
+In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the instances we consider.
 

### 2018-09-02 paper/content/revisited.tex

@@
 
-The results from the previous section present a few puzzles:
-\begin{enumerate}
-\item $\DG > \DEG$ on Heavy Tail even for large warm starts does not seem consistent with the ordering of the means.
-\item Why does $\TS$ beat $\DG$ for Needle In Haystack for low warm starts whereas on Uniform and Heavy Tail it does not?
-\end{enumerate}
+Conjecture~\ref{conj:inverted-u} for the most part is validated by our simulation results, but Puzzles~\ref{puz:dg_ht} and \ref{puz:nih} seem to be problematic for it.
 
-We first attempt to explain the first puzzle. We revisit Conjecture~\ref{conj:mean-trajectories} and find a more nuanced picture.
+We first attempt to explain Puzzle~\ref{puz:dg_ht} by revisiting Conjecture~\ref{conj:mean-trajectories}.
 
@@
 Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $\DG$ on the Uniform and Needle In Haystack instances.
-For the Uniform instance we see that, in the early rounds, $\DG > \TS$ for the majority of the simulations but that, eventually, $\TS > \DG$.
+For the Uniform instance we see that, in the early rounds, the relative reputation proportion for $\TS$  vs.
+$\DG$ is less than 0.5 meaning that $\DG$ has a higher reputation in a majority of the simulations  but that, eventually, the proportion $\TS$ does better than $\DG$ in a majority of the simulations.
 The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information.
@@
 \end{definition}
+
 Is exploration always costly?
@@
 
-This observation allows us to explain the second puzzle from the competition game with a more nuanced view of the inverted-U.
+This observation allows us to explain the Puzzle~\ref{puz:nih} from the competition game and establish a more nuanced view of the inverted-U.
 
 \begin{finding}
-\textit{For sufficiently low warm start on relative reputation costly instances the optimal strategies in the competition game are:
+\textit{For sufficiently low warm start \emph{under the instances that are relative reputation costly}, the optimal strategies in the competition game are:
 \begin{center}
-\textbf{Permanent Monopoly} - $\DG$ \\
+\textbf{Permanent Monopoly} - $DG$ \\
 \textbf{Temporary Monopoly} - $\TS$ \\
-\textbf{Permanent Duopoly} - $\DG$ \\
+\textbf{Permanent Duopoly} - $DG$ \\
 \end{center}

### 2018-09-02 paper/content/inverted_u.tex
fixups

@@
 We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.
-Since our results are from simulations, we interpret algorithm $A$ being better than algorithm $B$ if the resulting market share of $A$ is statistically greater than $0.5$.
+Since our results are from simulations, we interpret algorithm $A$ being better than algorithm $B$ if the resulting average market share of $A$ is statistically greater than $0.5$.
 Otherwise, the firm is indifferent between $A$ and $B$ with ties broken towards algorithms that are easier to deploy.
@@
 \begin{puzzle}\label{puz:dg_ht}
-Even for large warm start, $\DG$ is better than $\DEG$ under Heavy Tail
+Even for large warm start, $\DG$ is better than $\DEG$ under Heavy Tail \footnote{We demonstrate this holds even for $K = 3$ in the supplementary material}
 \end{puzzle}

### 2018-09-02 paper/content/revisited.tex
fixups

@@
 For the Uniform instance we see that, in the early rounds, the relative reputation proportion for $\TS$  vs.
-$\DG$ is less than 0.5 meaning that $\DG$ has a higher reputation in a majority of the simulations  but that, eventually, the proportion $\TS$ does better than $\DG$ in a majority of the simulations.
-The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information.
+$\DG$ is less than 0.5 meaning that $\DG$ has a higher reputation in a majority of the simulations  but that, eventually, $\TS$ does better than $\DG$ in a majority of the simulations.
+The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information which leads to a relatively lower reputation compared to $\DG$.
 However, eventually the information acquired from purposeful exploration in the early rounds allows $\TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $\DG$ cannot trivially find the best arm.
@@
 Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
-We call the priors where there is an early costly period followed by a later benefit period relative reputation costly instances.
+We call the instances where there is an early costly period followed by a later benefit period relative reputation costly instances.
 \end{definition}
@@
 There are two contributing factors to this.
-First, $\TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration.
-Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
+First, $\TS$ identifies the best arm faster in the Needle In Haystack instances than the Uniform instances so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration.
+Second, in the Needle In Haystack instances there are no ``bad" arms as there may be in the Uniform instances since by construction all the arms except one in Needle In Haystack are the same.
 Thus, when $\TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
-However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
-Thus, only the Uniform and Heavy Tail instances are relative reputation costly priors.
+However, with the Uniform instances, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
+Thus, only the Uniform and Heavy Tail instances are relative reputation costly instances.
 
@@
 
-Thus, the reason we see that $\DG > \TS$ under Heavy Tail and Uniform but that $\TS$ is always preferred under Needle In Haystack is due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds leads it to suffer a relative reputational cost in the former case but not in the latter.
+Thus, the reason we see that $\DG$ is better than $\TS$ under Heavy Tail and Uniform but that $\TS$ is always preferred under Needle In Haystack is due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds leads it to suffer a relative reputational cost in the former case but not in the latter.
 Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $T_0$ as fixing the starting point on the relative reputation plots.
 The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $T_0$.
-As a result, if we begin with a warm start that falls in the region where the proportion $< 0.5$, we expect that $\DG > \TS$.
+As a result, if we begin with a warm start that falls in the region where the proportion $< 0.5$, we expect that $\DG$ does better than $\TS$.
 
 This also allows us to give some intuition as to why allowing one firm to be a temporary monopoly incentivizes it to play $\TS$ even if, for the same warm start, the firm would be incentivized to play $\DG$ under permanent duopoly.
-By allowing it to be the only firm in the market for sufficiently long we allow it to recoup the ``reputation costs" by having it engage in most of its exploration while it does not have the pressure to incentivize agents to select them over their competition.
+By allowing it to be the only firm in the market for sufficiently long we allow it to recoup the reputation costs by having it engage in most of its exploration while it does not have the pressure to incentivize agents to select them over their competition.
 

### 2018-09-02 paper/content/revisited.tex
actually let me bold this

@@
 \begin{finding}
-\textit{For sufficiently low warm start \emph{under the instances that are relative reputation costly}, the optimal strategies in the competition game are:
+\textit{For sufficiently low warm start \textbf{under the instances that are relative reputation costly}, the optimal strategies in the competition game are:
 \begin{center}

### 2018-09-03 paper/content/barriers.tex
edits

@@
 \swcomment{too vague.
-what does significant mean?}\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two.}
+what does significant mean?}\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}
 \end{finding}

### 2018-09-03 paper/content/inverted_u.tex
edits

@@
 \section{Competition vs Better Algorithms: Inverted-U}\label{section:5}
- Now we demonstrate the inverted-U relationship between the incentivized
-  algorithms and the competition levels. In particular, we investigate
-  the incentivized algorithms under three different regimes: permanent
-  monopoly, temporary monopoly, and permanent duopoly.
-\gaedit{We break indifference between exploration algorithms and non-exploration algorithms towards non-exploration algorithms due to deployment costs.} \gadelete{First, the story for permanent monopoly is relatively straightforward.}
- We summarize the relationship between the different competition levels and incentivized algorithms as follows:
-
-\begin{conjecture}\label{conj:inverted-u}
-\textit{For sufficiently low warm start,
- the optimal strategies in the competition game are:
+Now we demonstrate the inverted-U relationship between the
+incentivized algorithms and the competition levels. In particular, we
+investigate the incentivized algorithms under three different regimes:
+permanent monopoly, temporary monopoly, and permanent duopoly:
+\swdelete{We break indifference between exploration algorithms and
+  non-exploration algorithms towards non-exploration algorithms due to
+  deployment costs.}  \swdelete{We summarize the relationship between
+  the different competition levels and incentivized algorithms as
+  follows:} \swcomment{deleted the discussion of tie-breaking since it
+  feels out of place.}
+\begin{finding}\label{conj:inverted-u}
+  \textit{For sufficiently low warm start, the optimal strategies
+    \swcomment{for whom??} in the competition game are:
 \begin{center}
@@
 \end{center}
-The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
+The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.\swcomment{vague here. }
 }
-\end{conjecture}
+\end{finding}
 
-\xhdr{Permanent Monopoly.} Since there is only a single firm
-  in the market for the entire period, the firm can take the entire
-  market regardless what algorithm it deploys.
-\gadelete{Assuming the deployment of any exploration algorithms will incur certain cost} \gaedit{Since we break indifferences by supposing the firm prefers easier to deploy algorithms}, the firm would choose to deploy $\DG$.
+\xhdr{Permanent Monopoly.} Since there is only a single firm in the
+market for the entire period, the firm can take the entire market
+regardless what algorithm it deploys. \gadelete{Assuming the
+  deployment of any exploration algorithms will incur certain cost}
+\swedit{Since we assume that exploration algorithms ($\DEG$ and $\TS$)
+  would incur deployment cost}, the firm would choose to deploy $\DG$.
 
@@
 
-\xhdr{Permanent Duopoly} Permanent duopoly corresponds to the model described in section 2 where $X = 0$ so that both firms enter the market simultaneously.
-What algorithms are firms incentivized to deploy?
-We report the average market share taken over the $N$ simulations for a set of exogenous values of $T_0$.
-Additionally, we report the mean and median of the effective end of game.
+\xhdr{Permanent Duopoly} \swdelete{Permanent duopoly corresponds to the model
+described in section 2 where $X = 0$ so that both firms enter the
+market simultaneously. What algorithms are firms incentivized to
+deploy? We report the average market share taken over the $N$
+simulations for a set of exogenous values of $T_0$. Additionally, we
+report the mean and median of the effective end of game.}
+\swedit{
 
-Since both firms enter at the same time the game is symmetric and the competition game is zero-sum so we only need to compute the expected payoffs of three pairs of strategies.
+  Since in this case both firms enter at the same time, they will both
+  get an expected market share of $1/2$ whenever they deploy the same
+  algorithms.
+
+To see what algorithms are the firms incentivized to deploy, we look at the average market
+share taken over the $N$ simulations for different values of warm-up length $T_0$.
+
+}
+
+\swdelete{Since both firms enter at the same time the game is symmetric and the competition game is zero-sum so we only need to compute the expected payoffs of three pairs of strategies.
 We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.
 Since our results are from simulations, we interpret algorithm $A$ being better than algorithm $B$ if the resulting average market share of $A$ is statistically greater than $0.5$.
-Otherwise, the firm is indifferent between $A$ and $B$ with ties broken towards algorithms that are easier to deploy.
+Otherwise, the firm is indifferent between $A$ and $B$ with ties broken towards algorithms that are easier to deploy.}
 

### 2018-09-03 paper/content/inverted_u.tex
edits and comments

@@
   the different competition levels and incentivized algorithms as
-  follows:} \swcomment{deleted the discussion of tie-breaking since it
-  feels out of place.}
+  follows:} \swcomment{think the tie-breaking should be discussed in the prelim?}
 \begin{finding}\label{conj:inverted-u}
@@
 \begin{center}
-\textbf{Permanent Monopoly} - $\DG$ \\
-\textbf{Temporary Monopoly} - $\TS$ \\
-\textbf{Permanent Duopoly} - $\DG$ \\
+{Permanent Monopoly} - $\DG$ \\
+{Temporary Monopoly} - $\TS$ \\
+{Permanent Duopoly} - $\DG$ \\
 \end{center}
-The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.\swcomment{vague here. }
+The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.\swcomment{can we quantify?}
 }
@@
 
-\xhdr{Permanent Duopoly} \swdelete{Permanent duopoly corresponds to the model
-described in section 2 where $X = 0$ so that both firms enter the
-market simultaneously. What algorithms are firms incentivized to
-deploy? We report the average market share taken over the $N$
-simulations for a set of exogenous values of $T_0$. Additionally, we
-report the mean and median of the effective end of game.}
-\swedit{
+\xhdr{Permanent Duopoly} \swdelete{Permanent duopoly corresponds to
+  the model described in section 2 where $X = 0$ so that both firms
+  enter the market simultaneously. What algorithms are firms
+  incentivized to deploy? We report the average market share taken
+  over the $N$ simulations for a set of exogenous values of
+  $T_0$. Additionally, we report the mean and median of the effective
+  end of game.}  \swedit{Since both firms enter at the same time, they
+  will both get an expected market share of $1/2$ whenever they deploy
+  the same algorithms. We will mainly investigate the market shares
+  when the firms play different algorithms.  In Tables \ref{sim_ht}
+  and \ref{sim_nih}, we report the average market share taken over the
+  $N$ simulations for different values of warm-up length $T_0$, and
+  also the mean and median of the EEOG, under heavy tail and needle in
+  haystack instances.
 
-  Since in this case both firms enter at the same time, they will both
-  get an expected market share of $1/2$ whenever they deploy the same
-  algorithms.
-
-To see what algorithms are the firms incentivized to deploy, we look at the average market
-share taken over the $N$ simulations for different values of warm-up length $T_0$.
-
-}
+  For the heavy tail and uniform instances,\footnote{We defer the
+    table for uniform instances to the appendix.} we see that $\DG$ is
+  a weakly dominant strategy for the competition game, and in
+  particular weakly dominates $\TS$. This establishes our claim that
+  $\DG$ is the incentivized algorithm.  However, there are two results
+  that are at odds with the conjecture: }
 
@@
 Since our results are from simulations, we interpret algorithm $A$ being better than algorithm $B$ if the resulting average market share of $A$ is statistically greater than $0.5$.
-Otherwise, the firm is indifferent between $A$ and $B$ with ties broken towards algorithms that are easier to deploy.}
+Otherwise, the firm is indifferent between $A$ and $B$ with ties broken towards algorithms that are easier to deploy.
 
@@
 However, there are two results that are at odds with the conjecture:
+}
+
+\swcomment{Guy, please check above! for what instances are $\DG$ a weakly dominant strategy again? Also, what is $N$?}
 
 \begin{puzzle}\label{puz:dg_ht}
-Even for large warm start, $\DG$ is better than $\DEG$ under Heavy Tail \footnote{We demonstrate this holds even for $K = 3$ in the supplementary material}
+Even for large warm start, $\DG$ is better than $\DEG$ under heavy tail instances.\footnote{We demonstrate this holds even for $K = 3$ in the supplementary material}
 \end{puzzle}
@@
 \begin{puzzle}\label{puz:nih}
-Under Needle In Haystack, $\TS$ is weakly dominant across all warm starts
+Under Needle In Haystack, $\TS$ is weakly dominant across all warm starts. 
 \end{puzzle}
@@
 \centering
-\caption{Duopoly Experiment Heavy Tail}
+\caption{Duopoly experiment for heavy tail instances}
 \begin{tabular}{rlll}
@@
 \centering
-\caption{Duopoly Experiment Needle In Haystack}
+\caption{Duopoly experiment for needle in haystack instances}
 \begin{tabular}{rlll}
@@
 \label{sim_nih}
-\caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95\% confidence band.
+\caption*{{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95\% confidence band.
 For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
@@
 
-\xhdr{Temporary Monopoly} We now consider asymmetries in the timing of entry so that one firm enters the market before the other and serves as a monopolist in the periods until the other firm enters.
-In terms of our model, this corresponds to varying the value of $X$.
-For this section we report results fixing $T_0 = 20$ and focus on the family of instances where we observed that, for this warm start, $\DG$ was the optimal strategy \footnote{However, we report the results for the Needle In Haystack instances in the supplementary material.
-The results are unsurprising, where $\TS$ is the dominant strategy for both the entrant and incumbent.}.
-It seems clear that the temporary monopoly will get a larger market share regardless of what algorithm it plays, but is the incumbent incentivized to commit to $\TS$?
+\swcomment{Guy: could you add lines and make the table 3 $\times$ 3 grid? Now it's hard to read which number belongs to which cell...}
+
+\xhdr{Temporary Monopoly.} Recall that the incumbent firm enters the
+market and serves as a monopolist until the entrant firm enters at
+round $X$. For this section we report results fixing $T_0 = 20$ and
+focus on the family of instances where we observed that, for this warm
+start, $\DG$ was the optimal strategy.\footnote{However, we report the
+  results for the Needle In Haystack instances in the supplementary
+  material. The results are unsurprising, where $\TS$ is the dominant
+  strategy for both the entrant and incumbent.}\swcomment{I don't
+  understand this sentence. what's the point?} \swdelete{It seems clear that the
+  temporary monopoly will get a larger market share regardless of what
+  algorithm it plays.}
 
 \begin{finding}
-\textit{Counterintuitively, allowing one firm to be a temporary monopolist for a sufficiently long time will induce that firm to commit to the better learning algorithm, $\TS$, even if, under the permanent duopoly, they would commit to $DG$}
+  \textit{If the incumbent firm can stay as a temporary monopolist for
+    a sufficiently long time, it will deploy the better learning
+    algorithm, $\TS$, even though, under the permanent duopoly, they
+    would commit to $\DG$.}
 \end{finding}
 
-In table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the Heavy Tail instances \footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.}.
-In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the instances we consider.
+In Table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the instances we consider.
 
@@
 \label{ht_incum}
-\caption*{\tiny{The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval.
+\caption*{{The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval.
 The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one firm getting an ``extreme" market share of $> 90\%$.

### 2018-09-03 paper/content/barriers.tex
clear up some of hte comments

@@
   experiment (called \emph{reputation erased experiment}), the
-  reputation of the incumbent is artificially erased upon the entry of
-  the entrant. In the second experiment (called \emph{information
+  reputation of the incumbent is artificially reset upon the entry of
+  the entrant \gaedit{so that the incumbent's reputation score is only set by the warm start periods and not any of the incumbency periods}. In the second experiment (called \emph{information
     erased experiment}), the information gained by the incumbent is
@@
 \begin{finding}
-\textit{Maintaining either the data or reputation advantage alone still allows the incumbent to retain a significant portion of the market.
+\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a large share of the market compared to the share they get under permanent duopoly}.
 \swcomment{too vague.
@@
 
-Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
-We find that the best responses remain the same for the Heavy Tail instance, but for the Uniform instance they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.
+\gadelete{Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
+We find that the best responses remain the same for the Heavy Tail instance, but for the Uniform instance they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.}
 

### 2018-09-03 paper/content/inverted_u.tex
clear up some of hte comments

@@
 \caption{Duopoly experiment for heavy tail instances}
-\begin{tabular}{rlll}
+\begin{tabular}{|c|c|c|c|}
   \hline
@@
 TS vs DG & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 55\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 570\\ med: 0} & \makecell{\textbf{0.76} $\pm$0.02\\ eeog \\ avg: 620\\ med: 98.5} \\
+\hline  
   TS vs DEG & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 37\\ med: 0} & \makecell{\textbf{0.88} $\pm$0.01\\ eeog \\ avg: 480\\ med: 0} & \makecell{\textbf{0.9} $\pm$0.01\\ eeog \\ avg: 570\\ med: 113.5} \\
+\hline  
   DG vs DEG & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 410\\ med: 7} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 790\\ med: 762} & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 730\\ med: 608} \\
@@
 \caption{Duopoly experiment for needle in haystack instances}
-\begin{tabular}{rlll}
+\begin{tabular}{|c|c|c|c|}
   \hline
@@
 TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 200\\ med: 27} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 370\\ med: 0} & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 580\\ med: 121.5} \\
+\hline  
   TS vs DEG & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 150\\ med: 14} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 460\\ med: 78.5} & \makecell{\textbf{0.56} $\pm$0.02\\ eeog \\ avg: 740\\ med: 627.5} \\
+  \hline
   DG vs DEG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 340\\ med: 128.5} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 650\\ med: 408} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 690\\ med: 466.5} \\
@@
 
-\swcomment{Guy: could you add lines and make the table 3 $\times$ 3 grid? Now it's hard to read which number belongs to which cell...}
-
 \xhdr{Temporary Monopoly.} Recall that the incumbent firm enters the
 market and serves as a monopolist until the entrant firm enters at
-round $X$. For this section we report results fixing $T_0 = 20$ and
-focus on the family of instances where we observed that, for this warm
-start, $\DG$ was the optimal strategy.\footnote{However, we report the
+round $X$.
+\gaedit{For this section we focus on the warm start values (concretely, $T_0 = 20$) and on the instances where we observed that, under permanent duopoly, $\DG$ was the weakly dominant strategy} \footnote{However, we report the
   results for the Needle In Haystack instances in the supplementary
   material. The results are unsurprising, where $\TS$ is the dominant
-  strategy for both the entrant and incumbent.}\swcomment{I don't
-  understand this sentence. what's the point?} \swdelete{It seems clear that the
-  temporary monopoly will get a larger market share regardless of what
+  strategy for both the entrant and incumbent.}.
+
+\gadelete{we report results fixing $T_0 = 20$ and
+focus on the family of instances where we observed that, for this warm
+start, $\DG$ was the optimal strategy.}\swcomment{I don't
+  understand this sentence. what's the point?} \swdelete{It seems clear that the temporary monopoly will get a larger market share regardless of what
   algorithm it plays.}

### 2018-09-03 paper/content/model.tex
edits and comments
edits

@@
 We treat $X$ as an exogenous element of the model, and study the consequences for a fixed $X$.
+\swedit{Each firm aims to maximize its market share---the number of agents coming to them.}
 
@@
 
+\swedit{For any fixed firm, we say that algorithm 1 \emph{weakly
+    dominates} algorithm 2 if algorithm 1 never provides a lower
+  market share than than 2 against all opposing strategies. An
+  algorithm is a weakly dominant strategy for a firm if the algorithm
+  weakly dominates all other algorithms.}
+
+\swcomment{please use the above terminology}
+
 \end{document} 
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../competing_bandits"
+%%% End:

### 2018-09-03 paper/content/inverted_u.tex
commit updated section 3 and 4

@@
 \section{Competition vs Better Algorithms: Inverted-U}\label{section:5}
+
 Now we demonstrate the inverted-U relationship between the
@@
 investigate the incentivized algorithms under three different regimes:
-permanent monopoly, temporary monopoly, and permanent duopoly:
-\swdelete{We break indifference between exploration algorithms and
-  non-exploration algorithms towards non-exploration algorithms due to
-  deployment costs.}  \swdelete{We summarize the relationship between
-  the different competition levels and incentivized algorithms as
-  follows:} \swcomment{think the tie-breaking should be discussed in the prelim?}
-\begin{finding}\label{conj:inverted-u}
-  \textit{For sufficiently low warm start, the optimal strategies
-    \swcomment{for whom??} in the competition game are:
+permanent monopoly, temporary monopoly, and permanent duopoly: \swcomment{think the tie-breaking should be discussed in the prelim?}
+
+\begin{finding}
+\textit{For sufficiently low warm start under the instances that are relative reputation costly, the optimal strategies in the competition game are:
 \begin{center}
-{Permanent Monopoly} - $\DG$ \\
-{Temporary Monopoly} - $\TS$ \\
-{Permanent Duopoly} - $\DG$ \\
+\textbf{Permanent Monopoly} - $\DG$ \\
+\textbf{Temporary Monopoly} (incumbent)- $\TS$ \\
+\textbf{Permanent Duopoly} - $\DG$ \\
 \end{center}
-The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.\swcomment{can we quantify?}
+The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
 }
@@
 market for the entire period, the firm can take the entire market
-regardless what algorithm it deploys. \gadelete{Assuming the
+regardless of what algorithm it deploys. \gadelete{Assuming the
   deployment of any exploration algorithms will incur certain cost}
@@
   also the mean and median of the EEOG, under heavy tail and needle in
-  haystack instances.
+  haystack instances.}
 
-  For the heavy tail and uniform instances,\footnote{We defer the
-    table for uniform instances to the appendix.} we see that $\DG$ is
-  a weakly dominant strategy for the competition game, and in
-  particular weakly dominates $\TS$. This establishes our claim that
-  $\DG$ is the incentivized algorithm.  However, there are two results
-  that are at odds with the conjecture: }
+For the relative reputation costly instances,\footnote{We defer the table for uniform instances to the appendix.} we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$.
+This establishes our claim that $\DG$ is the incentivized algorithm.
 
-\swdelete{Since both firms enter at the same time the game is symmetric and the competition game is zero-sum so we only need to compute the expected payoffs of three pairs of strategies.
-We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.
-Since our results are from simulations, we interpret algorithm $A$ being better than algorithm $B$ if the resulting average market share of $A$ is statistically greater than $0.5$.
-Otherwise, the firm is indifferent between $A$ and $B$ with ties broken towards algorithms that are easier to deploy.
-
-Tables \ref{sim_ht} and \ref{sim_nih} show the results of the simulations.
-For the most part Conjecture~\ref{conj:inverted-u} is validated where, for the Heavy Tail and Uniform instances, we see that $\DG$ does better than $\TS$ and is a weakly dominant strategy in the game.
-However, there are two results that are at odds with the conjecture:
-}
-
-\swcomment{Guy, please check above! for what instances are $\DG$ a weakly dominant strategy again? Also, what is $N$?}
-
+\gadelete{
 \begin{puzzle}\label{puz:dg_ht}
@@
 \end{puzzle}
-
-\OMIT{
-\begin{finding}
-\textit{
-\begin{enumerate}
-\item For low warm starts, under Heavy Tail and Uniform \\ $\DG$ is a weakly dominant strategy in competition (see Table \ref{sim_ht}),
-\item  (see Table \ref{sim_nih}). 
-\item For sufficiently large warm starts, $\TS$ is dominant across all instances.
-\item Even for a large warm start, $\DG$ leads to a higher expected market share than $\DEG$ under Heavy Tail
-\end{enumerate}
-}
-\end{finding}
 }
 
-\OMIT{
+The reason we see that $\DG$ weakly dominates under the Heavy Tail and Uniform instances but that $\TS$ is weakly dominant under Needle In Haystack is precisely due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds leads it to suffer a relative reputational cost in the former case but not in the latter.
 Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $T_0$ as fixing the starting point on the relative reputation plots.
-The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$.
-The intuition for the result in competition aligns with the fact that, for Uniform and Heavy Tail, $\TS$ is relative reputation costly compared to $\DG$ and so, for low warm starts, $\TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $\DG$ beats $\TS$.
-However, if we move the warm start sufficiently high so that $\TS$ moves to the relative reputation benefit regime, $\TS > \DG$.
-}
+The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $T_0$.
+As a result, we see that for warm start $T_0 = 20$ on the relative reputation costly instances, $\DG$ will win the first agent in a larger proportion of the simulations than $\TS$.
+However, on the Needle In Haystack instances, we see that $\TS$ wins a larger proportion of the first agents compared to $\DG$.
 
- \OMIT{However, we do not always see that $\DG > \TS$ under some warm starts.
-Table \ref{sim_nih} shows the results for the Needle In Haystack family where we see that $\TS > \DG$.
-This is precisely due to the fact that, under the Needle In Haystack, $\TS$ is not relative reputation costly compared to $\DG$ and thus we see that $\TS > \DG$.}
-
-We defer the explanation for these puzzles until the next section.
+What happens after the first agent?
 Looking at the EEOG values we note that, even though the game continues until $t=2000$, the game effectively ends much earlier than that.
@@
 This is due to the fact that the losing firm may enter into a ``death spiral" where the winner keeps learning and improving from its customers thus increasing its reputation while the losing firm gets starved of customers and progressively falls farther and farther behind its competitors.
+As a result, even though if $\TS$ could get sufficiently many agents it would eventually have a higher reputation than $\DG$, since it starts the game by losing agents to $\DG$ due to its lowered reputation from exploration in the warm start it ends up doing worse in the competition game.
 
@@
 round $X$.
-\gaedit{For this section we focus on the warm start values (concretely, $T_0 = 20$) and on the instances where we observed that, under permanent duopoly, $\DG$ was the weakly dominant strategy} \footnote{However, we report the
-  results for the Needle In Haystack instances in the supplementary
+\gaedit{For this section we focus on the warm start values and on the relative reputation costly instances where we observed that, under permanent duopoly, $\DG$ was the weakly dominant strategy.
+Concretely, our presented results are for $T_0 = 20$.} \footnote{However, we report the results for the Needle In Haystack instances in the supplementary
   material. The results are unsurprising, where $\TS$ is the dominant
@@
 
-\gadelete{we report results fixing $T_0 = 20$ and
-focus on the family of instances where we observed that, for this warm
-start, $\DG$ was the optimal strategy.}\swcomment{I don't
-  understand this sentence. what's the point?} \swdelete{It seems clear that the temporary monopoly will get a larger market share regardless of what
-  algorithm it plays.}
-
 \begin{finding}
-  \textit{If the incumbent firm can stay as a temporary monopolist for
+  \textit{If the incumbent firm is a temporary monopolist for
     a sufficiently long time, it will deploy the better learning
@@
 % Wed Aug 15 19:03:20 2018
+
 \begin{table}[ht]
@@
 \caption{Temporary Monopoly Heavy Tail X = 200}
-\begin{tabular}{rlll}
-\hline
-\multicolumn{4}{c}{Incumbent Algorithm}\\
-\multirow{12}{0.6in}{\rotatebox{90}{Entrant Algorithm}} \\
-  \hline
- & TS & DEG &  DG \\
-  \hline
-TS & \makecell{\textbf{0.003} $\pm$0.003\\Var:0.002\\ES:100\%} & \makecell{\textbf{0.083} $\pm$0.02\\Var:0.07\\ES:97\%} & \makecell{\textbf{0.17} $\pm$0.02\\Var:0.1\\ES:95\%} \\
-  DEG & \makecell{\textbf{0.045} $\pm$0.01\\Var:0.03\\ES:92\%} & \makecell{\textbf{0.25} $\pm$0.02\\Var:0.1\\ES:75\%} & \makecell{\textbf{0.23} $\pm$0.02\\Var:0.1\\ES:78\%} \\
-   DG & \makecell{\textbf{0.12} $\pm$0.02\\Var:0.08\\ES:88\%} & \makecell{\textbf{0.36} $\pm$0.03\\Var:0.2\\ES:76\%} & \makecell{\textbf{0.3} $\pm$0.02\\Var:0.1\\ES:64\%} \\
-   \hline
+\begin{tabular}{cc|c|c|c|}
+  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
+  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
+            & $\TS$ & \makecell{\textbf{0.003} \\$\pm$0.003} & \makecell{\textbf{0.083} \\ $\pm$0.02} & \makecell{\textbf{0.17} \\$\pm$0.02} \\ \cline{3-5}
+Entrant  & $\DEG$ & \makecell{\textbf{0.045} \\$\pm$0.01} & \makecell{\textbf{0.25} \\$\pm$0.02} & \makecell{\textbf{0.23} \\$\pm$0.02} \\\cline{3-5}
+            & $\DG$ & \makecell{\textbf{0.12} \\ $\pm$0.02} & \makecell{\textbf{0.36} \\$\pm$0.03} & \makecell{\textbf{0.3} \\$\pm$0.02} \\\cline{3-5}
 \end{tabular}
 \label{ht_incum}
-\caption*{{The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval.
-The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one firm getting an ``extreme" market share of $> 90\%$.
+\caption*{{The table contains the average market share for the entrant as well as 95\% confidence intervals.
 Note that smaller values in the table are better for the incumbent.
@@
 
-Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was preferred to $\TS$?
+Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was the weakly dominant strategy?
 The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
@@
 
-\OMIT{As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.} Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
+As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same immediate relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
+Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
 

### 2018-09-03 paper/content/perf_in_iso.tex
attempt to make the figures look better
commit updated section 3 and 4

@@
 We plot the \emph{mean reputation trajectory}: how this score evolves over time.
-Figure \ref{prelim_means} shows such plot for the Needle In Haystack instance (for other MAB instances the plots are similar).
+Figure \ref{prelim_means} shows such \gaedit{a} plot for the Needle In Haystack instance (for other MAB instances the plots are similar).
 We summarize this finding as follows:
@@
 \begin{finding}
-\textit{The mean reputation trajectories are arranged as predicted by prior work: 
+\textit{The mean reputation trajectories\gaedit{, for sufficiently long time horizon,} are arranged as predicted by prior work:
     $\TS > \DEG > \DG$.}
@@
 \caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area displays 95\% confidence intervals.}
-\includegraphics[scale=0.35]{"figures/nih_iso_mean"}
+\includegraphics[scale=0.35]{figures/nih_iso_mean}
 \label{prelim_means}
@@
 That is, how well do these trajectories predict the outcomes of the competition game?
-The following conjecture is natural:
+
+Our experiments led us to consider a statistic with a more direct connection to the competition game: for each time $t$, the proportion of \MRVs for which one algorithm has a higher reputation score than the other which we define as the \textit{relative reputation proportion}.
+
+\begin{figure}[ht]
+\caption{Relative Reputation Proportion Plots for $\TS$ vs $\DG$. Shaded area display 95\% confidence intervals.}
+\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
+\includegraphics[scale=0.35]{figures/relative_nih_ts_dg.png}
+%\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.
+The shaded area display 95\% confidence intervals.}}
+\label{relative_rep_plots}
+
+\end{figure}
+
+Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $\DG$ on the Uniform and Needle In Haystack instances.
+For the Uniform instance we see that, in the early rounds, the relative reputation proportion for $\TS$  vs.
+$\DG$ is less than 0.5 meaning that $\DG$ has a higher reputation in a majority of the simulations but that, eventually, $\TS$ has a higher reputation than $\DG$ in a majority of the simulations.
+The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information which leads to a relatively lower reputation compared to $\DG$.
+However, eventually the information acquired from purposeful exploration in the early rounds allows $\TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $\DG$ cannot trivially find the best arm.
+
+\begin{definition}
+\textit When purposeful exploration leads to a relative reputation loss compared to the greedy alternative we define this as a \textit{relative reputation cost}.
+The eventual gain in reputation from the early acquired information leads to what we define as a \textit{relative reputation benefit}.
+Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
+We call the instances where there is an early costly period followed by a later benefit period \textit{relative reputation costly instances}.
+\end{definition}
+
+\begin{finding}
+\textit{Purposeful exploration can lead to relative reputational costs and, as a result, the relative reputation proportion of $\TS$ vs $\DG$ favors $\DG$ for small time horizons.
+We observe this under the Uniform and Heavy Tail instances.
+However, it is not observed under the Needle In Haystack instances}
+\end{finding}
+
+Is exploration always costly?
+Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack instances, $\TS$ always does relatively better than $\DG$.
+There are two contributing factors to this.
+First, $\TS$ identifies the best arm faster in the Needle In Haystack instances than the Uniform instances so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration.
+Second, in the Needle In Haystack instances there are no ``bad" arms as there may be in the Uniform instances since by construction all the arms except one in Needle In Haystack are the same.
+Thus, when $\TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
+However, with the Uniform instances, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
+Thus, only the Uniform and Heavy Tail instances are relative reputation costly instances.
+
+\OMIT{The following conjecture is natural:
 
@@
 We revisit this conjecture after documenting the results in the competition game.
+}
 
@@
 
-Let us consider a statistic with a more direct connection to the competition game: for each time $t$, the proportion of \MRVs for which one algorithm has a higher reputation score than the other.
+.
 

### 2018-09-03 paper/content/inverted_u.tex
cosmetic changes

@@
   \hline
-TS vs DG & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 55\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 570\\ med: 0} & \makecell{\textbf{0.76} $\pm$0.02\\ eeog \\ avg: 620\\ med: 98.5} \\
+$\TS$ vs $\DG$ & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 55\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 570\\ med: 0} & \makecell{\textbf{0.76} $\pm$0.02\\ eeog \\ avg: 620\\ med: 98.5} \\
 \hline  
-  TS vs DEG & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 37\\ med: 0} & \makecell{\textbf{0.88} $\pm$0.01\\ eeog \\ avg: 480\\ med: 0} & \makecell{\textbf{0.9} $\pm$0.01\\ eeog \\ avg: 570\\ med: 113.5} \\
+  $\TS$ vs $\DEG$ & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 37\\ med: 0} & \makecell{\textbf{0.88} $\pm$0.01\\ eeog \\ avg: 480\\ med: 0} & \makecell{\textbf{0.9} $\pm$0.01\\ eeog \\ avg: 570\\ med: 113.5} \\
 \hline  
-  DG vs DEG & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 410\\ med: 7} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 790\\ med: 762} & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 730\\ med: 608} \\
+  $\DG$ vs $\DEG$ & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 410\\ med: 7} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 790\\ med: 762} & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 730\\ med: 608} \\
    \hline
@@
   \hline
-TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 200\\ med: 27} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 370\\ med: 0} & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 580\\ med: 121.5} \\
+$\TS$ vs $\DG$ & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 200\\ med: 27} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 370\\ med: 0} & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 580\\ med: 121.5} \\
 \hline  
-  TS vs DEG & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 150\\ med: 14} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 460\\ med: 78.5} & \makecell{\textbf{0.56} $\pm$0.02\\ eeog \\ avg: 740\\ med: 627.5} \\
+  $\TS$ vs $\DEG$ & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 150\\ med: 14} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 460\\ med: 78.5} & \makecell{\textbf{0.56} $\pm$0.02\\ eeog \\ avg: 740\\ med: 627.5} \\
   \hline
-  DG vs DEG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 340\\ med: 128.5} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 650\\ med: 408} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 690\\ med: 466.5} \\
+  $\DG$ vs $\DEG$ & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 340\\ med: 128.5} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 650\\ med: 408} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 690\\ med: 466.5} \\
    \hline

### 2018-09-03 paper/content/barriers.tex
updates

@@
   reputation of the incumbent is artificially reset upon the entry of
-  the entrant \gaedit{so that the incumbent's reputation score is only set by the warm start periods and not any of the incumbency periods}. In the second experiment (called \emph{information
+  the entrant \gaedit{so that the incumbent's reputation score is not influenced by the rewards during the incumbency periods and only determined by the rewards generated in the warm start periods}.
+In the second experiment (called \emph{information
     erased experiment}), the information gained by the incumbent is
@@
 \caption{Information Erased Experiment Heavy Tail}
-\begin{tabular}{c@{\hspace{0.1\tabcolsep}}ccc}
-\hline
-\multicolumn{4}{c}{Incumbent Algorithm}\\
-\multirow{12}{0.6in}{\rotatebox{90}{Entrant Algorithm}} \\
-  \hline
- & TS & DEG &  DG \\ 
-  \hline
-TS & \makecell{\textbf{0.021} $\pm$0.009\\Var:0.02\\ES:100\%} & \makecell{\textbf{0.16} $\pm$0.02\\Var:0.1\\ES:97\%} & \makecell{\textbf{0.21} $\pm$0.02\\Var:0.2\\ES:95\%} \\ 
-  DEG & \makecell{\textbf{0.26} $\pm$0.03\\Var:0.2\\ES:95\%} & \makecell{\textbf{0.3} $\pm$0.02\\Var:0.2\\ES:74\%} & \makecell{\textbf{0.26} $\pm$0.02\\Var:0.1\\ES:76\%} \\ 
-   DG & \makecell{\textbf{0.34} $\pm$0.03\\Var:0.2\\ES:94\%} & \makecell{\textbf{0.4} $\pm$0.03\\Var:0.2\\ES:74\%} & \makecell{\textbf{0.33} $\pm$0.02\\Var:0.1\\ES:58\%} \\ 
-   \hline
+\begin{tabular}{cc|c|c|c|}
+  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
+  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
+            & $\TS$ & \makecell{\textbf{0.021} \\$\pm$0.009}& 		\makecell{\textbf{0.16} \\$\pm$0.02} & 
+            \makecell{\textbf{0.21} \\ $\pm$0.02} \\ \cline{3-5}
+Entrant  & $\DEG$ & \makecell{\textbf{0.26} \\$\pm$0.03} & \makecell{\textbf{0.3} \\$\pm$0.02} & 
+\makecell{\textbf{0.26} \\ $\pm$0.02} \\ \cline{3-5}
+            & $\DG$ & \makecell{\textbf{0.34} \\ $\pm$0.03} & \makecell{\textbf{0.4} \\$\pm$0.03} & 
+            \makecell{\textbf{0.33} \\$\pm$0.02} \\\cline{3-5}
+\end{tabular}
 \label{info_erase}
-\end{tabular}
 \end{table}
@@
 \caption{Reputation Erased Experiment Heavy Tail}
-\begin{tabular}{c@{\hspace{0.01\tabcolsep}}ccc}
-\hline
-\multicolumn{4}{c}{Incumbent Algorithm}\\
-\multirow{12}{0.6in}{\rotatebox{90}{Entrant Algorithm}} \\
-  \hline
- & TS & DEG &  DG \\ 
-  \hline
-TS & \makecell{\textbf{0.0096} $\pm$0.006\\Var:0.009\\ES:100\%} & \makecell{\textbf{0.11} $\pm$0.02\\Var:0.09\\ES:98\%} & \makecell{\textbf{0.18} $\pm$0.02\\Var:0.1\\ES:95\%} \\ 
-  DEG & \makecell{\textbf{0.073} $\pm$0.01\\Var:0.05\\ES:93\%} & \makecell{\textbf{0.29} $\pm$0.02\\Var:0.2\\ES:78\%} & \makecell{\textbf{0.25} $\pm$0.02\\Var:0.1\\ES:79\%} \\ 
-   DG & \makecell{\textbf{0.15} $\pm$0.02\\Var:0.1\\ES:89\%} & \makecell{\textbf{0.39} $\pm$0.03\\Var:0.2\\ES:78\%} & \makecell{\textbf{0.33} $\pm$0.02\\Var:0.2\\ES:66\%} \\ 
-   \hline
+\begin{tabular}{cc|c|c|c|}
+  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
+  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
+            & $\TS$ & \makecell{\textbf{0.0096} \\$\pm$0.006}& 		\makecell{\textbf{0.11} \\$\pm$0.02} & 
+            \makecell{\textbf{0.18} \\$\pm$0.02} \\ \cline{3-5}
+Entrant  & $\DEG$ & \makecell{\textbf{0.073} \\ $\pm$0.01} & \makecell{\textbf{0.29} \\ $\pm$0.02} & 
+\makecell{\textbf{0.25} \\$\pm$0.02} \\ \cline{3-5}
+            & $\DG$ &\makecell{\textbf{0.15} \\$\pm$0.02} & \makecell{\textbf{0.39} \\$\pm$0.03} & 
+           \makecell{\textbf{0.33} \\ $\pm$0.02} \\\cline{3-5}
+\end{tabular}
 \label{rep_erase}
-\end{tabular}
 \end{table}
@@
 \begin{finding}
-\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a large share of the market compared to the share they get under permanent duopoly}.
-\swcomment{too vague.
-what does significant mean?}\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}
+\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a large share of the market compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}
 \end{finding}
@@
 
-\gadelete{Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
+\OMIT{Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
 We find that the best responses remain the same for the Heavy Tail instance, but for the Uniform instance they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.}

### 2018-09-03 paper/content/revisited.tex
updates

--- 2018-09-02 paper/content/revisited.tex
@@
 
-Conjecture~\ref{conj:inverted-u} for the most part is validated by our simulation results, but Puzzles~\ref{puz:dg_ht} and \ref{puz:nih} seem to be problematic for it.
+A natural question to ask is why we had to introduce the relative reputation proportion instead of simply using the mean to understand the results.
+For instance, the following would have been a natural conjecture:
+\begin{conjecture}\label{conj:mean-trajectories}
+\begin{enumerate}
+\item To understand which algorithm would win in the competition game we could compare the mean reputation score of the two algorithms at $T_0$.
+\item Given a sufficiently large $T_0$, if one algorithm's mean reputation trajectory lies above another, then the first algorithm prevails in the competition game.
+\end{enumerate}
+\end{conjecture}
 
-We first attempt to explain Puzzle~\ref{puz:dg_ht} by revisiting Conjecture~\ref{conj:mean-trajectories}.
+However, we find a more nuanced picture.
 
@@
 \textit{
-Conjecture~\ref{conj:mean-trajectories} is false. Mean reputation trajectories do not suffice to explain the outcomes under competition.}
+Conjecture~\ref{conj:mean-trajectories} is false.
+Mean reputation trajectories do not suffice to explain the outcomes under competition.
+We observe in Figure \ref{sim_ht} that $DG$ attains a larger market share than $DEG$ even for large warm starts.\footnote{In the appendix we show that this holds for $K = 3$ and longer time horizons than we consider here as well}}
 \end{finding}
@@
 %\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
+\end{figure}
+
+\begin{figure}[ht]
+\caption{Distribution of reputation difference $\TS-\DG$ for the Needle-in-Haystack (smoothed via a kernel density estimate)}
+\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
+\label{ts_dg_rep_diff_nih}
+%\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $\DG$ across $t$}}
 \end{figure}
@@
 
-\begin{figure}[ht]
-\caption{Distribution of reputation difference $\TS-\DG$ for the Needle-in-Haystack (smoothed via a kernel density estimate)}
-\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
-\label{ts_dg_rep_diff_nih}
-%\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $\DG$ across $t$}}
-\end{figure}
-
-Let us consider a statistic with a more direct connection to the competition game: for each time $t$, the proportion of \MRVs for which one algorithm has a higher reputation score than the other.
-
-\begin{definition}
-\textit{Relative Reputation Proportion} - the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$
-\end{definition}
-
-\begin{finding}
-\textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $\TS$ doing worse than $\DG$ for small time horizons.
-We observe this under the Uniform and Heavy Tail instances.
-However, it is not observed under the Needle In Haystack instances}
-\end{finding}
-
-The relative reputation proportion statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$.
-
-\begin{figure}[ht]
-\caption{Relative Reputation Proportion Plots for $\TS$ vs $\DG$. Shaded area display 95\% confidence intervals.}
-\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
-\includegraphics[scale=0.35]{figures/relative_nih_ts_dg.png}
-%\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.
-The shaded area display 95\% confidence intervals.}}
-\label{relative_rep_plots}
-
-\end{figure}
-
-Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $\DG$ on the Uniform and Needle In Haystack instances.
-For the Uniform instance we see that, in the early rounds, the relative reputation proportion for $\TS$  vs.
-$\DG$ is less than 0.5 meaning that $\DG$ has a higher reputation in a majority of the simulations  but that, eventually, the proportion $\TS$ does better than $\DG$ in a majority of the simulations.
-The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information.
-However, eventually the information acquired from purposeful exploration in the early rounds allows $\TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $\DG$ cannot trivially find the best arm.
-The early exploration leads to what we define as a \textit{relative reputation cost} and the eventual gain in reputation leads to what we define as a \textit{relative reputation benefit}.
-
-\begin{definition}
-\textit{Relative reputation cost and benefit} - the relative reputation loss an algorithm incurs from purposeful exploration compared to the greedy alternative.
-Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
-We call the priors where there is an early costly period followed by a later benefit period relative reputation costly instances.
-\end{definition}
-
-Is exploration always costly?
-Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack instances, $\TS$ always does relatively better than $\DG$.
-There are two contributing factors to this.
-First, $\TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration.
-Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
-Thus, when $\TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
-However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
-Thus, only the Uniform and Heavy Tail instances are relative reputation costly priors.
-
-This observation allows us to explain the Puzzle~\ref{puz:nih} from the competition game and establish a more nuanced view of the inverted-U.
-
-\begin{finding}
-\textit{For sufficiently low warm start \emph{under the instances that are relative reputation costly}, the optimal strategies in the competition game are:
-\begin{center}
-\textbf{Permanent Monopoly} - $DG$ \\
-\textbf{Temporary Monopoly} - $\TS$ \\
-\textbf{Permanent Duopoly} - $DG$ \\
-\end{center}
-The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
-}
-\end{finding}
-
-Thus, the reason we see that $\DG > \TS$ under Heavy Tail and Uniform but that $\TS$ is always preferred under Needle In Haystack is due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds leads it to suffer a relative reputational cost in the former case but not in the latter.
-Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $T_0$ as fixing the starting point on the relative reputation plots.
-The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $T_0$.
-As a result, if we begin with a warm start that falls in the region where the proportion $< 0.5$, we expect that $\DG > \TS$.
-
-This also allows us to give some intuition as to why allowing one firm to be a temporary monopoly incentivizes it to play $\TS$ even if, for the same warm start, the firm would be incentivized to play $\DG$ under permanent duopoly.
-By allowing it to be the only firm in the market for sufficiently long we allow it to recoup the ``reputation costs" by having it engage in most of its exploration while it does not have the pressure to incentivize agents to select them over their competition.
-
 \end{document}

### 2018-09-03 paper/content/revisited.tex
cosmetic change

@@
 Mean reputation trajectories do not suffice to explain the outcomes under competition.
-We observe in Figure \ref{sim_ht} that $DG$ attains a larger market share than $DEG$ even for large warm starts.\footnote{In the appendix we show that this holds for $K = 3$ and longer time horizons than we consider here as well}}
+We observe in Figure \ref{sim_ht} that $\DG$ attains a larger market share than $\DEG$ even for large warm starts.\footnote{In the appendix we show that this holds for $K = 3$ and longer time horizons than we consider here as well}}
 \end{finding}

### 2018-09-03 paper/content/model.tex

@@
 
-\swedit{For any fixed firm, we say that algorithm 1 \emph{weakly
+\swedit{For any single firm, we say that algorithm 1 \emph{weakly
     dominates} algorithm 2 if algorithm 1 never provides a lower
@@
   weakly dominates all other algorithms.}
+\swcomment{please use the above terminology}
 
-\swcomment{please use the above terminology}
+\swedit{todo: tie-breaking}
 

### 2018-09-04 paper/content/introduction.tex
Added a sentence of previous paper's result vs ours.
fixups to paper

--- 2018-08-30 paper/content/introduction.tex
@@
 rewards, a well-studied and well-understood MAB model \cite{bandits-ucb1}.
-We focus on a well-known distinction between ``greedy", exploitation-only algorithms, "naive" algorithms that separate exploration and exploitation, and better algorithms that combine the two.
+We focus on a well-known distinction between ``greedy", exploitation-only algorithms, ``naive" algorithms that separate exploration and exploitation, and better algorithms that combine the two.
 
@@
 Their results are purely theoretical; their model is amenable to proofs but not to numerical experiments.
+\asedit{Their high-level conclusion is an inverted-U relationship between competition and innovation that is similar to ours.}
 
@@
     \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
-Our setting is also closely related to the "dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
+Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
 

### 2018-09-04 paper/content/sim_details.tex
Minor edit.
basically rewrite barriers section, tweaks to inverted U and sim details
fixups to paper

@@
 \begin{enumerate}
-\item \emph{Needle In Haystack}: one arm (the ``needle") is chosen uniformly at random. This arm has mean reward $.7$, and the remaining ones have mean reward $.5$.
+\item \emph{Needle-In-Haystack}: one arm (the ``needle") is chosen uniformly at random. This arm has mean reward $.7$, and the remaining ones have mean reward $.5$.
 
@@
 This ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
+
 More specifically, the realization table is a 0-1 matrix $W$ with $K$ columns which correspond to arms, and $T+T_{\max}$ rows, which correspond to rounds.
@@
 
-\OMIT{This realization table, as well as fixing the random seed for the same bandit instance and realization table across experiments, ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
-In the competition game we draw from the $T \times K$ portion of the table, so that if two different algorithms picks arm $a$ at time $t$, they get the same $[a, t]$ realization in the table.
-This setup also ensures that in the warm start period, increasing the warm start from $k$ to $k + 10$ results in the same behavior in the first $k$ rounds.}
-
-We fix the sliding window size $M = 100$. We found that low values of $M$ induced too much random noise in the results, and increasing $M$ further did not make a substantial qualitative difference.
+We fix the sliding window size $M = 100$.
+We found that low values of $M$ induced too much random noise in the results, and increasing $M$ further did not make a substantial qualitative difference.
+Unless otherwise noted, all results are reported at $t = 2000$.
+\gaedit{The simulations are computationally intensive with each experiment, after extensive optimization, usually taking around half a day parallelized over a cluster of 12 CPU cores.}
 

### 2018-09-04 paper/content/sim_details.tex
tweak

@@
 Unless otherwise noted, all results are reported at $t = 2000$.
-\gaedit{The simulations are computationally intensive with each experiment, after extensive optimization, usually taking around half a day parallelized over a cluster of 12 CPU cores.}
+\gaedit{The simulations are computationally intensive with each experiment usually taking around half a day parallelized over a cluster of 12 CPU cores.}
 

### 2018-09-04 paper/content/inverted_u.tex
fixups to paper
use more game theoretic terminology

@@
 investigate the incentivized algorithms under three different regimes:
-permanent monopoly, temporary monopoly, and permanent duopoly: \swcomment{think the tie-breaking should be discussed in the prelim?}
+permanent monopoly, temporary monopoly, and permanent duopoly:
 
 \begin{finding}
-\textit{For sufficiently low warm start under the instances that are relative reputation costly, the optimal strategies in the competition game are:
+\textit{For sufficiently low warm start under the instances that are relative reputation costly, the unique pure Nash equilibrium strategies in the competition game are:\footnote{the uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
 \begin{center}
@@
 market for the entire period, the firm can take the entire market
-regardless of what algorithm it deploys. \gadelete{Assuming the
-  deployment of any exploration algorithms will incur certain cost}
-\swedit{Since we assume that exploration algorithms ($\DEG$ and $\TS$)
-  would incur deployment cost}, the firm would choose to deploy $\DG$.
+regardless of what algorithm it deploys.
+\swedit{Since we assume that exploration algorithms ($\DEG$ and $\TS$) would incur deployment cost and firms break indifferences towards easier to deploy algorithms}, the firm would choose to deploy $\DG$.
 
@@
 
-\xhdr{Permanent Duopoly} \swdelete{Permanent duopoly corresponds to
-  the model described in section 2 where $X = 0$ so that both firms
-  enter the market simultaneously. What algorithms are firms
-  incentivized to deploy? We report the average market share taken
-  over the $N$ simulations for a set of exogenous values of
-  $T_0$. Additionally, we report the mean and median of the effective
-  end of game.}  \swedit{Since both firms enter at the same time, they
-  will both get an expected market share of $1/2$ whenever they deploy
+\xhdr{Permanent Duopoly} \swedit{Since both firms enter at the same time, they will both get an expected market share of $1/2$ whenever they deploy
   the same algorithms. We will mainly investigate the market shares
@@
 
-For the relative reputation costly instances,\footnote{We defer the table for uniform instances to the appendix.} we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$.
+For the relative reputation costly instances,\footnote{We defer the table for uniform instances to the appendix.
+Summarizing the results, we see that, for low warm start, \DG yields more than half the market against \TS but is indifferent between \DG and \DEG.
+Since we break indifference towards easier to deploy algorithms, we also find that \DG is the incentivized algorithm in equilibrium} we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$.
 This establishes our claim that $\DG$ is the incentivized algorithm.
-
-\gadelete{
-\begin{puzzle}\label{puz:dg_ht}
-Even for large warm start, $\DG$ is better than $\DEG$ under heavy tail instances.\footnote{We demonstrate this holds even for $K = 3$ in the supplementary material}
-\end{puzzle}
-
-\begin{puzzle}\label{puz:nih}
-Under Needle In Haystack, $\TS$ is weakly dominant across all warm starts. 
-\end{puzzle}
-}
 
@@
   material. The results are unsurprising, where $\TS$ is the dominant
-  strategy for both the entrant and incumbent.}.
+  strategy for both the entrant and incumbent.}
 

### 2018-09-04 paper/content/model.tex
fixups to paper
use more game theoretic terminology

--- 2018-09-03 paper/content/model.tex
@@
 \item an exploration-separated algorithm that in each round, \emph{explores} with probability $\eps$: chooses an arm independently and uniformly at random, and with the remaining probability \emph{exploits} according to \DG.
-We call it ``dynamic epsilon-greedy", \DEG in short.
+We call it ``dynamic epsilon-greedy", \DEG in short.\gaedit{\footnote{For our experiments, we fix $\eps = 0.05$.
+Our pilot experiments showed that different $\eps$ did not qualitatively change the results.}}
 
@@
 
-\swedit{For any fixed firm, we say that algorithm 1 \emph{weakly
+\swedit{For any single firm, we say that algorithm 1 \emph{weakly
     dominates} algorithm 2 if algorithm 1 never provides a lower
-  market share than than 2 against all opposing strategies. An
+  market share than 2 against all opposing strategies. An
   algorithm is a weakly dominant strategy for a firm if the algorithm
   weakly dominates all other algorithms.}
-
+  \gaedit{Since the algorithms that involve exploration have larger deployment costs, we break any indifference towards algorithms that do not involve exploration (i.e.
+if \DG and \DEG provide the same market share, the firm will play \DG).}
 \swcomment{please use the above terminology}

### 2018-09-04 paper/content/sim_details.tex
This file is now obsolete!

@@
 \begin{document}
+
+OBSOLETE!
 
@@
 
+\subsubsection{Terminology.}
+\asedit{Following a standard game-theoretic terminology, algorithm Alg1 \emph{(weakly) dominates} algorithm Alg2 for a given firm if Alg1 provides a larger (or equal) market share than Alg2 at the end of the game.
+An algorithm is a (weakly) dominant strategy for the firm if it (weakly) dominates all other algorithms.
+This is for a particular MAB instance and a particular selection of the game parameters.}
+
+\OMIT{\gaedit{Since the algorithms that involve exploration have larger deployment costs, we break any indifference towards algorithms that do not involve exploration (i.e.
+if \DG and \DEG provide the same market share, the firm will play \DG).}}
+
 \subsubsection{Simulation details.}
@@
 We fix the sliding window size $M = 100$.
-We found that low values of $M$ induced too much random noise in the results, and increasing $M$ further did not make a substantial qualitative difference.
-Unless otherwise noted, all results are reported at $t = 2000$.
-\gaedit{The simulations are computationally intensive with each experiment usually taking around half a day parallelized over a cluster of 12 CPU cores.}
+We found that lower values induced too much random noise in the results, and increasing $M$ further did not make a qualitative difference.
+Unless otherwise noted, \asedit{we used $T = 2000$}.
+
+\asedit{The simulations are computationally intensive.
+An experiment on a particular MAB instance comprised multiple runs of the competition game: $N$ mean reward vectors times $9$ pairs of algorithms times three values for the warm start.
+We used a parallel implementation over a cluster of 12 XX Hz CPU cores, with 8 GB RAM per core.
+With this implementation, each experiment took about XX hours.}
 

### 2018-09-04 paper/content/model.tex
Moved content from "simulation details", edited "terminology" and "why so hard" para.

@@
 Before the game starts, each firm commits to an MAB algorithm, and uses this algorithm to choose its actions.
-Each algorithm receives a ``warm start": additional $T_0$ agents that show up before the game starts, and interact with the firm in the same way as described above.
+Each algorithm receives a ``warm start": additional $T_0$ agents that arrive before the game starts, and interact with the firm as described above.
+\asedit{Each firm's objective is to maximize its market share: the fraction of users who chose this firm.}
 
@@
 We treat $X$ as an exogenous element of the model, and study the consequences for a fixed $X$.
-\swedit{Each firm aims to maximize its market share---the number of agents coming to them.}
 
@@
 
-\swedit{For any single firm, we say that algorithm 1 \emph{weakly
-    dominates} algorithm 2 if algorithm 1 never provides a lower
-  market share than 2 against all opposing strategies. An
-  algorithm is a weakly dominant strategy for a firm if the algorithm
-  weakly dominates all other algorithms.}
-  \gaedit{Since the algorithms that involve exploration have larger deployment costs, we break any indifference towards algorithms that do not involve exploration (i.e.
-if \DG and \DEG provide the same market share, the firm will play \DG).}
-\swcomment{please use the above terminology}
+\subsubsection{MAB instances.}
+In all experiments, there are $K=10$ arms.
+Since we focus on 0-1 rewards, an instance of the MAB problem is specified by the mean reward vector $(\mu(a):\; a\in A)$, henceforth \MRV.
+Initially the \MRV is drawn from some distribution, termed \emph{MAB instance}.
+We consider several representative MAB instances.
+\begin{enumerate}
+\item \emph{Needle-In-Haystack}: one arm (the ``needle") is chosen uniformly at random. This arm has mean reward $.7$, and the remaining ones have mean reward $.5$.
+
+\item \emph{Uniform instance}: the mean reward of each arm is drawn independently and uniformly from $[0.25, 0.75]$.
+\item \emph{Heavy-tail instance}: the mean reward of each arm is drawn independently from $\Beta(.6,.6)$ distribution.%
+    \footnote{This distribution has substantial ``tail probabilities".}
+\end{enumerate}
+
+\subsubsection{Terminology.}
+\asedit{Following a standard game-theoretic terminology, algorithm Alg1 \emph{(weakly) dominates} algorithm Alg2 for a given firm if Alg1 provides a larger (or equal) market share than Alg2 at the end of the game.
+An algorithm is a (weakly) dominant strategy for the firm if it (weakly) dominates all other algorithms.
+This is for a particular MAB instance and a particular selection of the game parameters.}
+
+\OMIT{\gaedit{Since the algorithms that involve exploration have larger deployment costs, we break any indifference towards algorithms that do not involve exploration (i.e.
+if \DG and \DEG provide the same market share, the firm will play \DG).}}
+
+\subsubsection{Simulation details.}
+For each MAB instance we draw $N = 1000$ \MRVs independently from the corresponding distribution.
+We use this same collection of \MRVs for all experiments with this MAB instance.
+For each \MRV we draw a table of realized rewards (\emph{realization table}), and use this same table for all experiments on this \MRV.
+This ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
+
+More specifically, the realization table is a 0-1 matrix $W$ with $K$ columns which correspond to arms, and $T+T_{\max}$ rows, which correspond to rounds.
+Here $T_{\max}$ is the maximal duration of the ``warm start" in our experiments, \ie the maximal value of $X+T_0$.
+For each arm $a$, each value $W(\cdot,a)$ is drawn independently from Bernoulli distribution with expectation $\mu(a)$.
+Then in each experiment, the reward of this arm in round $t$ of the warm start is taken to be $W(t,a)$, and its reward in round $t$ of the game is $W(T_{\max}+t,a)$.
+
+We fix the sliding window size $M = 100$.
+We found that lower values induced too much random noise in the results, and increasing $M$ further did not make a qualitative difference.
+Unless otherwise noted, \asedit{we used $T = 2000$}.
+
+\asedit{The simulations are computationally intensive.
+An experiment on a particular MAB instance comprised multiple runs of the competition game: $N$ mean reward vectors times $9$ pairs of algorithms times three values for the warm start.
+We used a parallel implementation over a cluster of 12 XX Hz CPU cores, with 8 GB RAM per core.
+With this implementation, each experiment took about XX hours.}
+
+While we experiments with various MAB instances and parameter settings, in the main paper we only report plots for selected, representative experiments.
+Additional plots can be found in the supplement.
+Unless noted otherwise, our findings are based on and consistent with all these plots.
 

### 2018-09-04 paper/content/introduction.tex
Commented out "simulation details" file which is now obsolete.

@@
 Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window.
-Each firm's objective is to maximize its market share ---the fraction of users coming to them.
+Each firm's objective is to maximize its market share (the fraction of users choosing this firm).
 
@@
 rewards, a well-studied and well-understood MAB model \cite{bandits-ucb1}.
-We focus on a well-known distinction between ``greedy", exploitation-only algorithms, ``naive" algorithms that separate exploration and exploitation, and better algorithms that combine the two.
+We focus on a well-known distinction between ``greedy" (exploitation-only) algorithms, ``naive" algorithms that separate exploration and exploitation, and ``smart" algorithms that combine them.
+\asedit{Switching from ``greedy" to ``naive" to ``smart" comes with substantial adoption costs in infrastructure and personnel training \cite{MWT-WhitePaper-2016,DS-arxiv}.}
 
@@
 Their results are purely theoretical; their model is amenable to proofs but not to numerical experiments.
-\asedit{Their high-level conclusion is an inverted-U relationship between competition and innovation that is similar to ours.}
+Their high-level conclusion is an inverted-U relationship between competition and innovation that is similar to ours.
 

### 2018-09-04 paper/content/revisited.tex
comments

--- 2018-09-03 paper/content/revisited.tex
@@
 \begin{enumerate}
-\item To understand which algorithm would win in the competition game we could compare the mean reputation score of the two algorithms at $T_0$.
+\item To understand which algorithm would win in the competition game we could compare the mean reputation score of the two algorithms at $T_0$.\swcomment{should leave a pointer to the trajectories plot}
 \item Given a sufficiently large $T_0$, if one algorithm's mean reputation trajectory lies above another, then the first algorithm prevails in the competition game.
@@
 \end{conjecture}
+\swcomment{why enumerate? sounds like two conjectures, but they are not. first part seems redundant}
 
@@
 Mean reputation trajectories do not suffice to explain the outcomes under competition.
-We observe in Figure \ref{sim_ht} that $DG$ attains a larger market share than $DEG$ even for large warm starts.\footnote{In the appendix we show that this holds for $K = 3$ and longer time horizons than we consider here as well}}
+We observe in Figure \ref{sim_ht} that $\DG$ attains a larger market share than $\DEG$ even for large warm starts.\footnote{In the appendix we show that this holds for $K = 3$ and longer time horizons than we consider here as well.}}
 \end{finding}
@@
 To see what could go wrong, consider how an algorithm's reputation score is distributed at a particular time.
-That is, consider the empirical distribution of this score over different \MRVs.
-For concreteness, we consider the needle-in-haystack instance at time $t=500$ (Figure \ref{rep_dist_nih}) since the intuition as to why the mean may be misleading is clearest on this instance (though the same result holds across the instances).
+That is, consider the empirical distribution of this score over different \MRVs.\swcomment{not sure what this means.
+over the randomness of \MRVs only?
+you mean marginal distribution?} For concreteness, we consider the needle-in-haystack instance at time $t=500$ (Figure \ref{rep_dist_nih}) since the intuition as to why the mean may be misleading is clearest on this instance (though the same result holds across the instances).
 
 We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
-The reason is that for this MAB instance, $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
+The reason is that for this MAB instance,\swcomment{what is ``this''?} $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
 This leads to two cases: $\DG$ either does slightly better than $\TS$ or where $\TS$ does substantially better than $\DG$.
 
-However, the mean reputation trajectory of an algorithm would summarize all this complexity the average over the \MRVs.
+However, the mean reputation trajectory \swedit{may fail to capture} this complexity since it simply takes average over different realizations of \MRVs.
 This may be inadequate for explaining the outcome of the duopoly game, given that the latter is determined by a simple comparison between the firm's reputation scores.
@@
 Figure~\ref{ts_dg_rep_diff_nih} shows such plots for several time points.
-
+\swcomment{I don't quite understand Figure~\ref{ts_dg_rep_diff_nih}. there are three curves, so I assume two black ones are for the two algos? If so, could you explain}
 We observe that the distribution is skewed to the right, precisely due to the fact that $\DG$ either does slightly better than $\TS$ or does substantially worse.
@@
 \end{document} 
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../competing_bandits"
+%%% End:

### 2018-09-04 paper/content/revisited.tex
revisited revision

@@
 \begin{conjecture}\label{conj:mean-trajectories}
-\begin{enumerate}
-\item To understand which algorithm would win in the competition game we could compare the mean reputation score of the two algorithms at $T_0$.\swcomment{should leave a pointer to the trajectories plot}
-\item Given a sufficiently large $T_0$, if one algorithm's mean reputation trajectory lies above another, then the first algorithm prevails in the competition game.
-\end{enumerate}
+\item Given a sufficiently large $T_0$, if one algorithm's mean reputation trajectory lies above another (for instance, Figure  \ref{prelim_means}), then the first algorithm prevails in the competition game.
 \end{conjecture}
-\swcomment{why enumerate? sounds like two conjectures, but they are not. first part seems redundant}
 
@@
 \begin{figure}[ht]
+\includegraphics[scale=0.35]{figures/rep_distribution_nih}
 \caption{Distribution of reputation scores for Needle-in-Haystack at $t=500$ (smoothed using a kernel density estimate)}
-\includegraphics[scale=0.35]{figures/rep_distribution_nih}
 \label{rep_dist_nih}
@@
 \begin{figure}[ht]
-\caption{Distribution of reputation difference $\TS-\DG$ for the Needle-in-Haystack (smoothed via a kernel density estimate)}
 \includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
+\caption{Distribution of reputation difference $\TS-\DG$ for the Needle-in-Haystack at $t = 500, 1000, 2000$(smoothed via a kernel density estimate)}
 \label{ts_dg_rep_diff_nih}
@@
 We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
-The reason is that for this MAB instance,\swcomment{what is ``this''?} $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
+The reason is that \gaedit{for the Needle In Haystack instance}, $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
 This leads to two cases: $\DG$ either does slightly better than $\TS$ or where $\TS$ does substantially better than $\DG$.
@@
 Figure~\ref{ts_dg_rep_diff_nih} shows such plots for several time points.
-\swcomment{I don't quite understand Figure~\ref{ts_dg_rep_diff_nih}. there are three curves, so I assume two black ones are for the two algos? If so, could you explain}
 We observe that the distribution is skewed to the right, precisely due to the fact that $\DG$ either does slightly better than $\TS$ or does substantially worse.

### 2018-09-04 paper/content/revisited.tex
fixups

@@
 \begin{conjecture}\label{conj:mean-trajectories}
-\item Given a sufficiently large $T_0$, if one algorithm's mean reputation trajectory lies above another (for instance, Figure  \ref{prelim_means}), then the first algorithm prevails in the competition game.
+Given a sufficiently large $T_0$, if one algorithm's mean reputation trajectory lies above another (for instance, Figure  \ref{prelim_means}), then the first algorithm prevails in the competition game.
 \end{conjecture}

### 2018-09-04 paper/content/perf_in_iso.tex
Rewrote most of the section.

--- 2018-09-01 paper/content/perf_in_iso.tex
@@
 We focus on reputation scores generated by each algorithm.
-
-First, we confirm that\gaedit{, for a sufficiently long time horizon,} algorithms' performance is ordered as we'd expect:
-    $\TS \gg \DEG \gg \DG$.
+We confirm that algorithms' performance is ordered as we'd expect:
+    $\TS > \DEG > \DG$
+for a sufficiently long time horizon.
 For each algorithm and each MAB instance, we compute the mean reputation score at each round, averaged over all \MRVs.
 We plot the \emph{mean reputation trajectory}: how this score evolves over time.
-Figure \ref{prelim_means} shows such plot for the Needle In Haystack instance (for other MAB instances the plots are similar).
+Figure \ref{prelim_means} shows such a plot for the Needle In Haystack instance; for other MAB instances the plots are similar.
 We summarize this finding as follows:
@@
 \textit{The mean reputation trajectories are arranged as predicted by prior work:
-    $\TS > \DEG > \DG$.}
+    $\TS > \DEG > \DG$ for a sufficiently long time horizon.}
 \end{finding}
@@
 \begin{figure}
-\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area displays 95\% confidence intervals.}
-\includegraphics[scale=0.35]{"figures/nih_iso_mean"}
-\label{prelim_means}
+\includegraphics[scale=0.35]{figures/nih_iso_mean}
 %\caption*{\tiny{Mean reputation trajectory The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
 The shaded area display 95\% confidence intervals.}}
+\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area displays 95\% confidence intervals.}
+\label{prelim_means}
 \end{figure}
 
-The mean reputation trajectory is, arguably, the most natural way to represent an algorithm's performance on a given problem instance.
-But is it the right statistic for our purposes?
-That is, how well do these trajectories predict the outcomes of the competition game?
-The following conjecture is natural:
+\ascomment{Rewrote from here down.}
+
+The mean reputation trajectory is probably the most natural way to represent an algorithm's performance on a given MAB instance.
+However, we found that the outcomes of the competition game are better explained with a different ``performance-in-isolation" statistic that is more directly connected to the game.
+Consider the  performance of two algorithms, Alg1 and Alg2, ``in isolation" on a particular MAB instance.
+The \emph{relative reputation} of Alg1 (vs.
+Alg2) at a given time $t$ is the fraction of \MRVs / realization tables for which Alg1 has a higher reputation score than Alg2.
+The intuition is that agent's selection in our model depends only on the comparison between the reputation scores.
+
+\begin{figure}[ht]
+\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
+\includegraphics[scale=0.35]{figures/relative_nih_ts_dg.png}
+%\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.
+The shaded area display 95\% confidence intervals.}}
+\caption{Relative Reputation Proportion Plots for $\TS$ vs $\DG$. MAB instances: Uniform (top) and Needle-in-Haystack (bottom). Shaded area display 95\% confidence intervals.}
+\label{relative_rep_plots}
+\end{figure}
+
+This angle allows a more nuanced analysis of reputation costs vs.
+benefits under competition.
+Figure \ref{relative_rep_plots} (top) shows the relative reputation trajectory for $\TS$ vs $\DG$ for the Uniform instance.
+The relative reputation is less than $\tfrac12$ in the early rounds, meaning that $\DG$ has a higher reputation score in a majority of the simulations, and more than $\tfrac12$ later on.
+The reason is the exploration in \TS leads to worse decisions initially, but allows for better decisions later.
+The time period when relative reputation vs.
+\DG dips below $\tfrac12$ can be seen as an explanation for the competitive disadvantage of exploration.
+However, such period does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
+\footnote{We see two explanations for this: $\TS$ identifies the best arm faster for Needle-in-Haystack than for Uniform instance, and there are no ``very bad" arms which make exploration very expensive.}
+
+\begin{finding}\label{find:period}
+\textit{Exploration can lead to relative reputation vs. $\DG$ going below $\tfrac12$ for some initial time period. This happens for some MAB instances but not for some others.}
+\end{finding}
+
+\begin{definition}
+For a particular MAB algorithm, a time period when relative reputation vs.
+\DG goes below $\tfrac12$ is called {\em exploration disadvantage period}.
+An MAB instance is called \emph{exploration-disadvantaged} if such period exists.
+\end{definition}
+
+\OMIT{ %%%%%%%
+\begin{definition}
+When purposeful exploration leads to a relative reputation loss compared to the greedy alternative we define this as a \textit{relative reputation cost}.
+The eventual gain in reputation from the early acquired information leads to what we define as a \textit{relative reputation benefit}.
+Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
+We call the instances where there is an early costly period followed by a later benefit period \textit{relative reputation costly instances}.
+\end{definition}
+} %%%%%
+
+\OMIT{ %%%%% 
+$\TS$ identifies the best arm faster for Needle-in-Haystack than for Uniform instance, so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration.
+Second, in the Needle In Haystack instances there are no ``bad" arms as there may be in the Uniform instances since by construction all the arms except one in Needle In Haystack are the same.
+Thus, when $\TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
+However, with the Uniform instances, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
+Thus, only the Uniform and Heavy Tail instances are relative reputation costly instances.
+} %%%%
+
+\OMIT{The following conjecture is natural:
 
@@
 \begin{figure}[H]
+\includegraphics[scale=0.35]{figures/rep_distribution_nih}
+%\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
 \caption{Distribution of reputation scores for Needle-in-Haystack at $t=500$ (smoothed using a kernel density estimate)}
-\includegraphics[scale=0.35]{figures/rep_distribution_nih}
 \label{rep_dist_nih}
-%\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
 \end{figure}

### 2018-09-04 paper/content/barriers.tex
fixups
typo

@@
 \begin{finding}
-\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a large share of the market compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}
+\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger share of the market compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}
 \end{finding}

### 2018-09-04 paper/content/perf_in_iso.tex
Minor edit.

@@
 The shaded area display 95\% confidence intervals.}}
-\caption{Relative Reputation Proportion Plots for $\TS$ vs $\DG$. MAB instances: Uniform (top) and Needle-in-Haystack (bottom). Shaded area display 95\% confidence intervals.}
+\caption{Relative reputation trajectory for $\TS$ vs $\DG$, on Uniform instance (top) and Needle-in-Haystack instance (bottom). Shaded area display 95\% confidence intervals.}
 \label{relative_rep_plots}
@@
 However, such period does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
-\footnote{We see two explanations for this: $\TS$ identifies the best arm faster for Needle-in-Haystack than for Uniform instance, and there are no ``very bad" arms which make exploration very expensive.}
+\footnote{We see two explanations for this: $\TS$ identifies the best arm faster than for the Uniform instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
 

### 2018-09-04 paper/content/introduction.tex
Minor edits.

@@
 
-We consider varying levels of competition, from monopoly to \emph{temporary monopoly} (a duopoly with an early entrant), to a true duopoly in which both firms start at the same time.
+We consider varying levels of competition, from permanent monopoly to \emph{temporary monopoly} (a duopoly with an early entrant), to a duopoly in which both firms start at the same time.
 We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly.

### 2018-09-04 paper/content/model.tex
Minor edits.

@@
 
-In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``entrant").
+In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``entrant"), and therefore enjoys a \emph{temporary monopoly}.
 Formally, the incumbent enjoys additional $X$ rounds of the ``warm start".
@@
 In all experiments, there are $K=10$ arms.
-Since we focus on 0-1 rewards, an instance of the MAB problem is specified by the mean reward vector $(\mu(a):\; a\in A)$, henceforth \MRV.
+Since we focus on 0-1 rewards, an instance of the MAB problem is specified by the \emph{\MRV} $(\mu(a):\; a\in A)$.
 Initially the \MRV is drawn from some distribution, termed \emph{MAB instance}.

### 2018-09-04 paper/content/inverted_u.tex
Revised the tables.

@@
 \begin{document}
-\section{Competition vs Better Algorithms: Inverted-U}\label{section:5}
+\section{Competition and Better MAB Algorithms}\label{section:5}
 
-Now we demonstrate the inverted-U relationship between the
-incentivized algorithms and the competition levels. In particular, we
-investigate the incentivized algorithms under three different regimes:
-permanent monopoly, temporary monopoly, and permanent duopoly:
+We describe our main experiments with the duopoly game.
+On a high level, we demonstrate the inverted-U relationship between the algorithms incentivized by the competition algorithms and the intensity of the competition, from
+permanent monopoly to temporary monopoly to permanent duopoly.
 
@@
 % Thu Aug 16 13:13:01 2018
-\begin{table}[ht]
+\begin{table*}[h]
 \centering
-\caption{Duopoly experiment for heavy tail instances}
-\begin{tabular}{|c|c|c|c|}
+\begin{tabular}{|c|c|c|c||c|c|c|}
   \hline
+  & \multicolumn{3}{c}{Heavy-Tail} 
+  & \multicolumn{3}{c}{Needle-in-Haystack}\\
+  \hline 
+  & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 
   & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 \\
   \hline
-$\TS$ vs $\DG$ & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 55\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 570\\ med: 0} & \makecell{\textbf{0.76} $\pm$0.02\\ eeog \\ avg: 620\\ med: 98.5} \\
+\TS vs \DG 
+  & \makecell{\textbf{0.29} $\pm$0.03\\ \Eeog 55 (0)} 
+  & \makecell{\textbf{0.72} $\pm$0.02\\ \Eeog 570 (0)} 
+  & \makecell{\textbf{0.76} $\pm$0.02\\ \Eeog 620 (99)} 
+  %%
+  & \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 200 (27)}
+  & \makecell{\textbf{0.6} $\pm$0.03\\ \Eeog 370 (0}
+  & \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 580 (122)}
+  \\
 \hline
-  $\TS$ vs $\DEG$ & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 37\\ med: 0} & \makecell{\textbf{0.88} $\pm$0.01\\ eeog \\ avg: 480\\ med: 0} & \makecell{\textbf{0.9} $\pm$0.01\\ eeog \\ avg: 570\\ med: 113.5} \\
+  $\TS$ vs $\DEG$ 
+  & \makecell{\textbf{0.3} $\pm$0.03\\ \Eeog 37 (0)} 
+  & \makecell{\textbf{0.88} $\pm$0.01\\ \Eeog 480 (0)}
+  & \makecell{\textbf{0.9} $\pm$0.01\\ \Eeog 570 (114)} 
+  %%
+  & \makecell{\textbf{0.57} $\pm$0.03\\ \Eeog 150 (14}
+  & \makecell{\textbf{0.52} $\pm$0.03\\ \Eeog 460 (79)}
+  & \makecell{\textbf{0.56} $\pm$0.02\\ \Eeog 740 (628)} 
+  \\
 \hline
-  $\DG$ vs $\DEG$ & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 410\\ med: 7} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 790\\ med: 762} & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 730\\ med: 608} \\
+  $\DG$ vs $\DEG$ 
+  & \makecell{\textbf{0.62} $\pm$0.03\\ \Eeog 410 (7)}
+  & \makecell{\textbf{0.6} $\pm$0.02\\ \Eeog 790 (762)}
+  & \makecell{\textbf{0.57} $\pm$0.03\\ \Eeog 730 (608)} 
+  %%
+  & \makecell{\textbf{0.46} $\pm$0.03\\ \Eeog 340 (129)}
+  & \makecell{\textbf{0.42} $\pm$0.02\\ \Eeog 650 (408)}
+  & \makecell{\textbf{0.42} $\pm$0.02\\ \Eeog 690 (467)} 
+  \\
    \hline
 \end{tabular}
+\caption{The duopoly experiment for Uniform and Needle-in-Haystack instances.
+Each cell describes a game between two algorithms, call them Alg1 vs.
+Alg2, for a particular value of the warm start $T_0$.
+The first line in the cell is the market share of Alg 1: the average (in bold) and the 95\% confidence band.
+%For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG. 
+The second line specifies the ``effective end of game" (\Eeog): the average and the median (in brackets).
+\ascomment{Can't add vertical lines in the top row - pls help!}}
 \label{sim_ht}
-\end{table}
+\end{table*}
 
+\OMIT{ %%%%%%
 \begin{table}[ht]
@@
   \hline
-$\TS$ vs $\DG$ & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 200\\ med: 27} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 370\\ med: 0} & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 580\\ med: 121.5} \\
+$\TS$ vs $\DG$ 
+& \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 200 (27)} 
+& \makecell{\textbf{0.6} $\pm$0.03\\ \Eeog 370 (0} 
+& \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 580 (122)} \\
 \hline
-  $\TS$ vs $\DEG$ & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 150\\ med: 14} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 460\\ med: 78.5} & \makecell{\textbf{0.56} $\pm$0.02\\ eeog \\ avg: 740\\ med: 627.5} \\
+  $\TS$ vs $\DEG$ 
+  & \makecell{\textbf{0.57} $\pm$0.03\\ \Eeog 150 (14} 
+  & \makecell{\textbf{0.52} $\pm$0.03\\ \Eeog 460 (79)} 
+  & \makecell{\textbf{0.56} $\pm$0.02\\ \Eeog 740 (628)} \\
   \hline
-  $\DG$ vs $\DEG$ & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 340\\ med: 128.5} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 650\\ med: 408} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 690\\ med: 466.5} \\
+  $\DG$ vs $\DEG$ 
+  & \makecell{\textbf{0.46} $\pm$0.03\\ \Eeog 340 (129)} 
+  & \makecell{\textbf{0.42} $\pm$0.02\\ \Eeog 650 (408)} 
+  & \makecell{\textbf{0.42} $\pm$0.02\\ \Eeog 690 (467)} \\
    \hline
@@
 % Thu Aug 16 13:13:00 2018
+} %%%%%%%
 
@@
 
-In Table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the instances we consider.
+In Table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the
+instances we consider.
 
+\begin{table}[ht]
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+   & $\TS$  & $\DEG$  & $\DG$ \\ \hline
+$\TS$ 
+    & \makecell{\textbf{0.003}$\pm$0.003} 
+    & \makecell{\textbf{0.083}$\pm$0.02} 
+    & \makecell{\textbf{0.17}$\pm$0.02} \\\hline
+$\DEG$ 
+    & \makecell{\textbf{0.045}$\pm$0.01} 
+    & \makecell{\textbf{0.25}$\pm$0.02} 
+    & \makecell{\textbf{0.23}$\pm$0.02} \\\hline
+$\DG$ 
+    & \makecell{\textbf{0.12}$\pm$0.02} 
+    & \makecell{\textbf{0.36}$\pm$0.03} 
+    & \makecell{\textbf{0.3}$\pm$0.02} \\\hline
+\end{tabular}
+\label{ht_incum}
+\caption{The temporary monopoly experiment with $X=200$ (and $T_0=\asedit{???}$), for the Heavy-Tail MAB instance.
+Each cell describes the duopoly game between the entrant's algorithm (the row) and the incumbent's algorithm (the column).
+The cell specifies the entrant's market share for the rounds in which he was present.
+We give the average (in bold) and the 95\% confidence interval.
+NB: smaller average is better for the incumbent.}
+\end{table}
+
+\OMIT{ %%%%%%%
 % latex table generated in R 3.4.0 by xtable 1.8-2 package
 % Wed Aug 15 19:03:20 2018
-
 \begin{table}[ht]
@@
 \end{table}
+} %%%%%%%
 

### 2018-09-04 paper/content/inverted_u.tex
Revised top para.

@@
 
+\ascomment{stopped here.}
+
 \begin{finding}
-\textit{For sufficiently low warm start under the instances that are relative reputation costly, the unique pure Nash equilibrium strategies in the competition game are:\footnote{the uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
+\textit{For sufficiently low warm start under the instances that are exploration-disadvantaged, the unique pure Nash equilibrium strategies in the competition game are:\footnote{the uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
 \begin{center}
@@
 
-For the relative reputation costly instances,\footnote{We defer the table for uniform instances to the appendix.
+For the exploration-disadvantaged instances,\footnote{We defer the table for uniform instances to the appendix.
 Summarizing the results, we see that, for low warm start, \DG yields more than half the market against \TS but is indifferent between \DG and \DEG.
@@
 The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $T_0$.
-As a result, we see that for warm start $T_0 = 20$ on the relative reputation costly instances, $\DG$ will win the first agent in a larger proportion of the simulations than $\TS$.
+As a result, we see that for warm start $T_0 = 20$ on the exploration-disadvantaged instances, $\DG$ will win the first agent in a larger proportion of the simulations than $\TS$.
 However, on the Needle In Haystack instances, we see that $\TS$ wins a larger proportion of the first agents compared to $\DG$.
@@
 round $X$.
-\gaedit{For this section we focus on the warm start values and on the relative reputation costly instances where we observed that, under permanent duopoly, $\DG$ was the weakly dominant strategy.
+\gaedit{For this section we focus on the warm start values and on the exploration-disadvantaged instances where we observed that, under permanent duopoly, $\DG$ was the weakly dominant strategy.
 Concretely, our presented results are for $T_0 = 20$.} \footnote{However, we report the results for the Needle In Haystack instances in the supplementary

### 2018-09-04 paper/content/inverted_u.tex
Minor edit.

@@
 \caption{The temporary monopoly experiment with $X=200$ (and $T_0=\asedit{???}$), for the Heavy-Tail MAB instance.
-Each cell describes the duopoly game between the entrant's algorithm (the row) and the incumbent's algorithm (the column).
-The cell specifies the entrant's market share for the rounds in which he was present.
+Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}).
+The cell specifies the entrant's market share for this game, over the rounds in which he was present.
 We give the average (in bold) and the 95\% confidence interval.

### 2018-09-04 paper/content/barriers.tex

@@
 
-\begin{table}[ht]
+\begin{table*}[h]
 \centering
-\caption{Information Erased Experiment Heavy Tail}
-\begin{tabular}{cc|c|c|c|}
-  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
-  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
-            & $\TS$ & \makecell{\textbf{0.021} \\$\pm$0.009}& 		\makecell{\textbf{0.16} \\$\pm$0.02} & 
-            \makecell{\textbf{0.21} \\ $\pm$0.02} \\ \cline{3-5}
-Entrant  & $\DEG$ & \makecell{\textbf{0.26} \\$\pm$0.03} & \makecell{\textbf{0.3} \\$\pm$0.02} & 
-\makecell{\textbf{0.26} \\ $\pm$0.02} \\ \cline{3-5}
-            & $\DG$ & \makecell{\textbf{0.34} \\ $\pm$0.03} & \makecell{\textbf{0.4} \\$\pm$0.03} & 
-            \makecell{\textbf{0.33} \\$\pm$0.02} \\\cline{3-5}
+\begin{tabular}{|c|c|c|c||c|c|c|}
+\hline
+  & \multicolumn{3}{c}{Information erased}
+  & \multicolumn{3}{c}{Reputation erased}\\
+\hline
+& $\TS$  & $\DEG$  & $\DG$ 
+& $\TS$  & $\DEG$  & $\DG$
+\\\hline
+$\TS$ 
+    & \makecell{\textbf{0.021}$\pm$0.009}
+    & \makecell{\textbf{0.16}$\pm$0.02} 
+    & \makecell{\textbf{0.21} $\pm$0.02} 
+    %%
+    & \makecell{\textbf{0.0096}$\pm$0.006}
+    & \makecell{\textbf{0.11}$\pm$0.02}
+    & \makecell{\textbf{0.18}$\pm$0.02} 
+    \\ \hline
+$\DEG$ 
+    & \makecell{\textbf{0.26}$\pm$0.03} 
+    & \makecell{\textbf{0.3}$\pm$0.02} 
+    & \makecell{\textbf{0.26}$\pm$0.02} 
+    %%
+    & \makecell{\textbf{0.073}$\pm$0.01}
+    & \makecell{\textbf{0.29}$\pm$0.02}
+    & \makecell{\textbf{0.25}$\pm$0.02} 
+    \\ \hline
+$\DG$ 
+    & \makecell{\textbf{0.34}$\pm$0.03} 
+    & \makecell{\textbf{0.4}$\pm$0.03} 
+    & \makecell{\textbf{0.33}$\pm$0.02} 
+    %%
+    & \makecell{\textbf{0.15}$\pm$0.02}
+    & \makecell{\textbf{0.39}$\pm$0.03}
+    & \makecell{\textbf{0.33}$\pm$0.02}
+    \\\hline
 \end{tabular}
+\caption{Information-erased vs.
+reputation-erased experiment, on Heavy-Tail MAB instance.
+Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}).
+The cell specifies the entrant's market share in this game: the average (in bold) and the 95\% confidence interval.
+NB: smaller average is better for the incumbent.}
 \label{info_erase}
-\end{table}
+\end{table*}
 
-\begin{table}[ht]
-\centering
-\caption{Reputation Erased Experiment Heavy Tail}
-\begin{tabular}{cc|c|c|c|}
-  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
-  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
-            & $\TS$ & \makecell{\textbf{0.0096} \\$\pm$0.006}& 		\makecell{\textbf{0.11} \\$\pm$0.02} & 
-            \makecell{\textbf{0.18} \\$\pm$0.02} \\ \cline{3-5}
-Entrant  & $\DEG$ & \makecell{\textbf{0.073} \\ $\pm$0.01} & \makecell{\textbf{0.29} \\ $\pm$0.02} & 
-\makecell{\textbf{0.25} \\$\pm$0.02} \\ \cline{3-5}
-            & $\DG$ &\makecell{\textbf{0.15} \\$\pm$0.02} & \makecell{\textbf{0.39} \\$\pm$0.03} & 
-           \makecell{\textbf{0.33} \\ $\pm$0.02} \\\cline{3-5}
-\end{tabular}
-\label{rep_erase}
-\end{table}
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{cc|c|c|c|}
+%  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
+%  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
+%            & $\TS$ & \makecell{\textbf{0.021} \\$\pm$0.009}& 		\makecell{\textbf{0.16} \\$\pm$0.02} &
+%            \makecell{\textbf{0.21} \\ $\pm$0.02} \\ \cline{3-5}
+%Entrant  & $\DEG$ & \makecell{\textbf{0.26} \\$\pm$0.03} & \makecell{\textbf{0.3} \\$\pm$0.02} &
+%\makecell{\textbf{0.26} \\ $\pm$0.02} \\ \cline{3-5}
+%            & $\DG$ & \makecell{\textbf{0.34} \\ $\pm$0.03} & \makecell{\textbf{0.4} \\$\pm$0.03} &
+%            \makecell{\textbf{0.33} \\$\pm$0.02} \\\cline{3-5}
+%\end{tabular}
+%\caption{Information Erased Experiment Heavy Tail}
+%\label{info_erase}
+%\end{table}
+%
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{cc|c|c|c|}
+%  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
+%  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
+%            & $\TS$ & \makecell{\textbf{0.0096} \\$\pm$0.006}& 		\makecell{\textbf{0.11} \\$\pm$0.02} &
+%            \makecell{\textbf{0.18} \\$\pm$0.02} \\ \cline{3-5}
+%Entrant  & $\DEG$ & \makecell{\textbf{0.073} \\ $\pm$0.01} & \makecell{\textbf{0.29} \\ $\pm$0.02} &
+%\makecell{\textbf{0.25} \\$\pm$0.02} \\ \cline{3-5}
+%            & $\DG$ &\makecell{\textbf{0.15} \\$\pm$0.02} & \makecell{\textbf{0.39} \\$\pm$0.03} &
+%           \makecell{\textbf{0.33} \\ $\pm$0.02} \\\cline{3-5}
+%\end{tabular}
+%\caption{Reputation Erased Experiment Heavy Tail}
+%\label{rep_erase}
+%\end{table}
 

### 2018-09-04 paper/content/revisited.tex
Just changed the section title to fit one line.

@@
 
-\section{Algorithms' Performance in Isolation: Revisited}\label{section:revisited}
+\section{Performance in Isolation, Revisited}\label{section:revisited}
 
@@
 \includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
-\caption{Distribution of reputation difference $\TS-\DG$ for the Needle-in-Haystack at $t = 500, 1000, 2000$(smoothed via a kernel density estimate)}
+\caption{Distribution of reputation difference $\TS-\DG$ for  Needle-in-Haystack 
+% at $t = 500, 1000, 2000$ AS: this is redundant, and it saves a line :).
+(smoothed via a kernel density estimate)}
 \label{ts_dg_rep_diff_nih}

### 2018-09-04 paper/content/inverted_u.tex
fix table lines
more table edits'
fixups to table

--- 2018-09-04 paper/content/inverted_u.tex
@@
   \hline
-  & \multicolumn{3}{c}{Heavy-Tail} 
-  & \multicolumn{3}{c}{Needle-in-Haystack}\\
+  & \multicolumn{3}{c|}{Heavy-Tail} 
+  & \multicolumn{3}{c|}{Needle-in-Haystack}\\
   \hline 
@@
 %For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG. 
-The second line specifies the ``effective end of game" (\Eeog): the average and the median (in brackets).
-\ascomment{Can't add vertical lines in the top row - pls help!}}
+The second line specifies the ``effective end of game" (\Eeog): the average and the median (in brackets).}
 \label{sim_ht}
@@
 \label{ht_incum}
-\caption{The temporary monopoly experiment with $X=200$ (and $T_0=\asedit{???}$), for the Heavy-Tail MAB instance.
+\caption{The temporary monopoly experiment with $X=200$ (and $T_0=20$), for the Heavy-Tail MAB instance.
 Each cell describes the duopoly game between the entrant's algorithm (the row) and the incumbent's algorithm (the column).

### 2018-09-04 paper/content/inverted_u.tex
updates

@@
 \begin{finding}
-\textit{For sufficiently low warm start under the instances that are exploration-disadvantaged, the unique pure Nash equilibrium strategies in the competition game are:\footnote{the uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
+\textit{For sufficiently low warm start under the instances that are exploration-disadvantaged, the unique incentivized strategies in the competition game are:\footnote{The uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
 \begin{center}
@@
 \textbf{Temporary Monopoly} (incumbent)- $\TS$ \\
-\textbf{Permanent Duopoly} - $\DG$ \\
+\textbf{Permanent Duopoly} - $(\DG, \DG)$ \\
 \end{center}
@@
 
-For the exploration-disadvantaged instances,\footnote{We defer the table for uniform instances to the appendix.
+For the exploration-disadvantaged instances, we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$.
+This establishes our claim that $\DG$ is the incentivized algorithm.\footnote{We defer the table for uniform instances to the appendix.
 Summarizing the results, we see that, for low warm start, \DG yields more than half the market against \TS but is indifferent between \DG and \DEG.
-Since we break indifference towards easier to deploy algorithms, we also find that \DG is the incentivized algorithm in equilibrium} we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$.
-This establishes our claim that $\DG$ is the incentivized algorithm.
+Since we break indifference towards easier to deploy algorithms, we also find that \DG is the incentivized algorithm in equilibrium}
 
-The reason we see that $\DG$ weakly dominates under the Heavy Tail and Uniform instances but that $\TS$ is weakly dominant under Needle In Haystack is precisely due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds leads it to suffer a relative reputational cost in the former case but not in the latter.
+The reason we see that $\DG$ is incentivized under the Heavy Tail and Uniform instances but that $\TS$ is weakly dominant under Needle In Haystack is precisely due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds puts it in a disadvantage in the former case but not in the latter.
 Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $T_0$ as fixing the starting point on the relative reputation plots.
@@
 \gaedit{For this section we focus on the warm start values and on the exploration-disadvantaged instances where we observed that, under permanent duopoly, $\DG$ was the weakly dominant strategy.
-Concretely, our presented results are for $T_0 = 20$.} \footnote{However, we report the results for the Needle In Haystack instances in the supplementary
-  material. The results are unsurprising, where $\TS$ is the dominant
+Concretely, our presented results are for $T_0 = 20$.} \footnote{However, we report the results for the Needle In Haystack instances in the supplementary material.
+The results are unsurprising, where $\TS$ is the dominant
   strategy for both the entrant and incumbent.}
@@
 
-In Table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the
+In Table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$ and $\DG$ is only weakly dominant.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the
 instances we consider.
@@
 
-Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was the weakly dominant strategy?
+Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was the incentivized algorithm?
 The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.

### 2018-09-04 paper/competing_bandits.tex
first pass at another restructure
updates
Commented out "perf-in-iso" which is now obsolete.
Added a macro \Eeod.
Changed \Eeog macro.
more small fixups

@@
 /Title  (Competing Bandits: The Perils of Exploration under Competition)
-/Author (Guy Aridor, Kevin Liu, Aleksandrs Slivkins, Zhiwei Steven Wu)
+/Author (...)
 /Keywords (Input your keywords in this optional area)
@@
 \newtheorem{definition}{Definition}
-\newtheorem{finding}[definition]{Finding}
-\newtheorem{conjecture}[definition]{Conjecture}
+\newtheorem{finding}{Finding}
+\newtheorem{conjecture}{Conjecture}
+\newtheorem{puzzle}{Puzzle}
 \captionsetup{font=footnotesize}
@@
 \newcommand{\DG}{\term{DG}}
-\newcommand{\Beta}{\term{Beta}}
+\newcommand{\Beta}{\term{Beta}} % for Beta distribution
+\newcommand{\Eeog}{\term{EoG}} % shorthand for "effective end of game"
 
@@
 % e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
+\setlength{\tabcolsep}{8pt} % Default value: 6pt
+\renewcommand{\arraystretch}{1.5} % Default value: 1
 
@@
 
-\subfile{content/sim_details}
+%\subfile{content/sim_details}
 
@@
 
+\subfile{content/barriers}
+
 \subfile{content/revisited}
-
-\subfile{content/barriers}
 

### 2018-09-04 paper/content/inverted_u.tex
more small fixups

@@
 around at $t$ choose different firms.
-
-\xhdr{Permanent Duopoly} \swedit{Since both firms enter at the same time, they will both get an expected market share of $1/2$ whenever they deploy
-  the same algorithms. We will mainly investigate the market shares
-  when the firms play different algorithms.  In Tables \ref{sim_ht}
-  and \ref{sim_nih}, we report the average market share taken over the
-  $N$ simulations for different values of warm-up length $T_0$, and
-  also the mean and median of the EEOG, under heavy tail and needle in
-  haystack instances.}
-
-For the exploration-disadvantaged instances, we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$.
-This establishes our claim that $\DG$ is the incentivized algorithm.\footnote{We defer the table for uniform instances to the appendix.
-Summarizing the results, we see that, for low warm start, \DG yields more than half the market against \TS but is indifferent between \DG and \DEG.
-Since we break indifference towards easier to deploy algorithms, we also find that \DG is the incentivized algorithm in equilibrium}
-
-The reason we see that $\DG$ is incentivized under the Heavy Tail and Uniform instances but that $\TS$ is weakly dominant under Needle In Haystack is precisely due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds puts it in a disadvantage in the former case but not in the latter.
-Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $T_0$ as fixing the starting point on the relative reputation plots.
-The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $T_0$.
-As a result, we see that for warm start $T_0 = 20$ on the exploration-disadvantaged instances, $\DG$ will win the first agent in a larger proportion of the simulations than $\TS$.
-However, on the Needle In Haystack instances, we see that $\TS$ wins a larger proportion of the first agents compared to $\DG$.
-
-What happens after the first agent?
-Looking at the EEOG values we note that, even though the game continues until $t=2000$, the game effectively ends much earlier than that.
-This seems to imply that when attempting to understand the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotic performance.
-This is due to the fact that the losing firm may enter into a ``death spiral" where the winner keeps learning and improving from its customers thus increasing its reputation while the losing firm gets starved of customers and progressively falls farther and farther behind its competitors.
-As a result, even though if $\TS$ could get sufficiently many agents it would eventually have a higher reputation than $\DG$, since it starts the game by losing agents to $\DG$ due to its lowered reputation from exploration in the warm start it ends up doing worse in the competition game.
 
@@
 The second line specifies the ``effective end of game" (\Eeog): the average and the median (in brackets).}
-\label{sim_ht}
+\label{sim_table}
 \end{table*}
+\xhdr{Permanent Duopoly} \swedit{Since both firms enter at the same time, they will both get an expected market share of $1/2$ whenever they deploy
+  the same algorithms. We will mainly investigate the market shares
+  when the firms play different algorithms.  In Table \ref{sim_table}, we report the average market share taken over the
+  $N$ simulations for different values of warm-up length $T_0$, and
+  also the mean and median of the EEOG, under heavy tail and needle in
+  haystack instances.}
+
+For the exploration-disadvantaged instances, we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$.
+This establishes our claim that $\DG$ is the incentivized algorithm.\footnote{We defer the table for uniform instances to the appendix.
+Summarizing the results, we see that, for low warm start, \DG yields more than half the market against \TS but is indifferent between \DG and \DEG.
+Since we break indifference towards easier to deploy algorithms, we also find that \DG is the incentivized algorithm in equilibrium}
+
+The reason we see that $\DG$ is incentivized under the Heavy Tail and Uniform instances but that $\TS$ is weakly dominant under Needle In Haystack is precisely due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds puts it in a disadvantage in the former case but not in the latter.
+Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $T_0$ as fixing the starting point on the relative reputation plots.
+The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $T_0$.
+As a result, we see that for warm start $T_0 = 20$ on the exploration-disadvantaged instances, $\DG$ will win the first agent in a larger proportion of the simulations than $\TS$.
+However, on the Needle In Haystack instances, we see that $\TS$ wins a larger proportion of the first agents compared to $\DG$.
+
+What happens after the first agent?
+Looking at the EEOG values we note that, even though the game continues until $t=2000$, the game effectively ends much earlier than that.
+This seems to imply that when attempting to understand the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotic performance.
+This is due to the fact that the losing firm may enter into a ``death spiral" where the winner keeps learning and improving from its customers thus increasing its reputation while the losing firm gets starved of customers and progressively falls farther and farther behind its competitors.
+As a result, even though if $\TS$ could get sufficiently many agents it would eventually have a higher reputation than $\DG$, since it starts the game by losing agents to $\DG$ due to its lowered reputation from exploration in the warm start it ends up doing worse in the competition game.
 

### 2018-09-04 paper/content/perf_in_iso.tex
typo

@@
 However, such period does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
-\footnote{We see two explanations for this: $\TS$ identifies the best arm faster than for the Uniform instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
+\footnote{We see two explanations for this: $\TS$ identifies the best arm faster  for the Uniform instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
 

### 2018-09-05 paper/content/introduction.tex
Minor edit.

@@
 Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information).
-MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background.
+MAB problems have been studied for many decades, see \cite{Gittins-book11,Bubeck-survey12} for background.
 We consider i.i.d.
@@
 We focus on a well-known distinction between ``greedy" (exploitation-only) algorithms, ``naive" algorithms that separate exploration and exploitation, and ``smart" algorithms that combine them.
-\asedit{Switching from ``greedy" to ``naive" to ``smart" comes with substantial adoption costs in infrastructure and personnel training \cite{MWT-WhitePaper-2016,DS-arxiv}.}
+\asedit{Switching from ``greedy" to ``naive" to ``smart" MAB algorithms involves substantial adoption costs in infrastructure and personnel training \cite{MWT-WhitePaper-2016,DS-arxiv}.}
 

### 2018-09-05 paper/content/model.tex
Added a para on how our MAB instances are representative.

@@
 Each algorithm receives a ``warm start": additional $T_0$ agents that arrive before the game starts, and interact with the firm as described above.
-\asedit{Each firm's objective is to maximize its market share: the fraction of users who chose this firm.}
+Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
 
@@
 Initially the \MRV is drawn from some distribution, termed \emph{MAB instance}.
-We consider several representative MAB instances.
+We consider three MAB instances:
 \begin{enumerate}
@@
 \end{enumerate}
+\asedit{We argue that these MAB instances are (somewhat) representative.
+Consider the ``gap" between the best and the second-best arm, an essential parameter in the literature on MAB.
+The ``gap" is fixed in Needle-in-Haystack, spread over a wide spectrum of values under the Uniform instance, and is spread but  focused on the large values under the Heavy-Tail instance.}
 
 \subsubsection{Terminology.}
-\asedit{Following a standard game-theoretic terminology, algorithm Alg1 \emph{(weakly) dominates} algorithm Alg2 for a given firm if Alg1 provides a larger (or equal) market share than Alg2 at the end of the game.
+Following a standard game-theoretic terminology, algorithm Alg1 \emph{(weakly) dominates} algorithm Alg2 for a given firm if Alg1 provides a larger (or equal) market share than Alg2 at the end of the game.
 An algorithm is a (weakly) dominant strategy for the firm if it (weakly) dominates all other algorithms.
-This is for a particular MAB instance and a particular selection of the game parameters.}
+This is for a particular MAB instance and a particular selection of the game parameters.
 
@@
 
-\asedit{The simulations are computationally intensive.
+The simulations are computationally intensive.
 An experiment on a particular MAB instance comprised multiple runs of the competition game: $N$ mean reward vectors times $9$ pairs of algorithms times three values for the warm start.
-We used a parallel implementation over a cluster of 12 XX Hz CPU cores, with 8 GB RAM per core.
-With this implementation, each experiment took about XX hours.}
+We used a parallel implementation over a cluster of 12 \asedit{XX-Hz} CPU cores, with 8 GB RAM per core.
+With this implementation, each experiment took about \asedit{XX} hours.
 

### 2018-09-05 paper/content/perf_in_iso.tex
Added which instances are exploration-disadvantaged.

@@
 \DG dips below $\tfrac12$ can be seen as an explanation for the competitive disadvantage of exploration.
-However, such period does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
+\asedit{Such period also exists for the Heavy-tail MAB instance.} However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
 \footnote{We see two explanations for this: $\TS$ identifies the best arm faster  for the Uniform instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
@@
 \end{definition}
+
+\asedit{Thus, Uniform and Heavy-tail instance are exploration-disadvantaged, but Needle-in-Haystack is not.}
 

### 2018-09-05 paper/content/model.tex
add timing results

@@
 We found that lower values induced too much random noise in the results, and increasing $M$ further did not make a qualitative difference.
-Unless otherwise noted, \asedit{we used $T = 2000$}.
+Unless otherwise noted, we used $T = 2000$.
 
@@
 An experiment on a particular MAB instance comprised multiple runs of the competition game: $N$ mean reward vectors times $9$ pairs of algorithms times three values for the warm start.
-We used a parallel implementation over a cluster of 12 \asedit{XX-Hz} CPU cores, with 8 GB RAM per core.
-With this implementation, each experiment took about \asedit{XX} hours.
+We used a parallel implementation over a cluster of 12 2200Hz CPU cores, with 8 GB RAM per core.
+With this implementation, each experiment took about 10.5 hours.
 

### 2018-09-05 paper/content/barriers.tex
more small fixups
stuff

@@
 \section{Data and Reputation as Barriers to Entry}\label{section:6}
+In the temporary monopoly setting, the incumbent firm enjoys a first
+mover advantage, which allows the firm to acquire data on potentially
+suboptimal actions without incurring an immediate reputational
+consequence, and to build up a high reputation before the entrant
+enters the market. Thus, the early entry gives the incumbent both a
+\textit{data} advantage and a \textit{reputational} advantage over the
+entrant. This provides a quantitative perspective on the classic
+``first mover advantage" \cite{kerin1992first} in the digital economy
+where data can serve as a barrier to entry in many markets.
 
-An alternative interpretation of the results in the previous section is that the temporary monopoly provides a first mover advantage to the incumbent firm.
-The early entry allows the incumbent to acquire data on potentially suboptimal actions without incurring an immediate reputational consequence as well as the ability to eventually build up a high reputation before the entrant enters the market.
-Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant.
-This gives new intuition for the classic ``first mover advantage" \cite{kerin1992first} in the digital economy where data can serve as a barrier to entry in many markets.
+Together the two advantages serve as strong barriers to entry, and a
+natural question is whether either one of the advantages alone suffice
+as a strong barrier to entry. Also, which of the two is the stronger
+barrier?
 
-While combined these two advantages serve as strong barriers to entry,
-a natural question to ask is would having only the data advantage or
-only the reputation advantage still serve as such strong barriers to
-entry? Which of the two serves as a stronger barrier to entry and
-under what conditions?
-
-\swedit{To answer these questions, we isolate the effects from the two
+{To answer these questions, we isolate the effect from each of the two
   advantages and run two additional experiments. In the first
@@
   reputation of the incumbent is artificially reset upon the entry of
-  the entrant \gaedit{so that the incumbent's reputation score is not influenced by the rewards during the incumbency periods and only determined by the rewards generated in the warm start periods}.
-In the second experiment (called \emph{information
-    erased experiment}), the information gained by the incumbent is
-  erased in the sense that its posterior is reset to the initial
-  (fake) prior upon the entry of the entrant.}
-
-\swcomment{Guy, could you make it clear what the first ``erase'' mean? Like reset to default?}
+  the entrant {so that the incumbent's reputation score is only based
+    on the rewards generated in the warm start period}. In the second
+  experiment (called \emph{information erased experiment}), the
+  information gained by the incumbent is erased upon the entry of the
+  entrant in the sense that its posterior is reset to the initial
+  (fake) prior.}
 
@@
 \hline
-  & \multicolumn{3}{c}{Information erased}
-  & \multicolumn{3}{c}{Reputation erased}\\
+  & \multicolumn{3}{c|}{Information erased}
+  & \multicolumn{3}{c|}{Reputation erased}\\
 \hline
@@
 NB: smaller average is better for the incumbent.}
-\label{info_erase}
+\label{barrier_exp}
 \end{table*}
@@
 
-Tables \ref{info_erase} and \ref{rep_erase} display the results of these experiments.
+Table \ref{barrier_exp} display the results of these experiments.
 The reason for the finding is that, under $\TS$, the early entry allows the incumbent to do exploration without immediate reputational consequence in the early rounds and so, even when reputation is erased, the incumbent has already acquired sufficient information that simply exploiting that information allows her to quickly recover her reputation.

### 2018-09-05 paper/content/introduction.tex

--- 2018-09-04 paper/content/introduction.tex
@@
 We focus on a well-known distinction between ``greedy" (exploitation-only) algorithms, ``naive" algorithms that separate exploration and exploitation, and ``smart" algorithms that combine them.
-\asedit{Switching from ``greedy" to ``naive" to ``smart" comes with substantial adoption costs in infrastructure and personnel training \cite{MWT-WhitePaper-2016,DS-arxiv}.}
+Switching from ``greedy" to ``naive" to ``smart" algorithms involves substantial adoption costs in infrastructure and personnel training \cite{MWT-WhitePaper-2016,DS-arxiv}.
 

### 2018-09-05 paper/content/inverted_u.tex
A partial pass.

@@
 
-We describe our main experiments with the duopoly game.
-On a high level, we demonstrate the inverted-U relationship between the algorithms incentivized by the competition algorithms and the intensity of the competition, from
-permanent monopoly to temporary monopoly to permanent duopoly.
+Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
+As the ``intensity of competition" varies from permanent monopoly to temporary monopoly to permanent duopoly, we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
 
-\ascomment{stopped here.}
+More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
+We do this for each "intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
 
-\begin{finding}
-\textit{For sufficiently low warm start under the instances that are exploration-disadvantaged, the unique incentivized strategies in the competition game are:\footnote{The uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
-\begin{center}
-\textbf{Permanent Monopoly} - $\DG$ \\
-\textbf{Temporary Monopoly} (incumbent)- $\TS$ \\
-\textbf{Permanent Duopoly} - $(\DG, \DG)$ \\
-\end{center}
-The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
+For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{MWT-WhitePaper-2016,DS-arxiv}).
+Then \DG is trivially the dominant strategy under permanent monopoly.
+
+%(1) Permanent monopoly - indifferent between all algorithms but break indifference towards DG cause of deployment costs
+%(2) Temporary monopoly - TS is the dominant strategy for the incumbent
+%(3) Permanent duopoly -
+%   (a) Under NIH, TS is dominant
+%   (b) Under Heavy Tail, DG is weakly dominant
+%   (c) Under Uniform, DG "beats" TS, but DEG "beats" TS by more than DG and DEG vs DG leads to ~50% market share (it's slightly in favor of DG but 50/50 is within the confidence band).
+If we break indifference towards DG cause of deployment costs, then we have that the unique PSNE is (DG, DG).
+If we don't add the indifference breaking, then we have four PSNE, (DG, DEG), (DEG, DG), (DG, DG), (DEG, DEG)
+
+For temporary monopoly, we find that \TS dominates:
+
+\begin{finding}\label{find:temp-monopoly}
+\textit{Under temporary monopoly, TS is the dominant strategy for the incumbent, across all MAB instances, if the temporary monopoly is sufficiently long-lasting.
 }
@@
 
-\xhdr{Permanent Monopoly.} Since there is only a single firm in the
-market for the entire period, the firm can take the entire market
-regardless of what algorithm it deploys.
+For permanent duopoly, it is a crucial whether an MAB instance is exploration-disadvantaged:
+
+\begin{finding}\label{find:duopoly}
+\textit{Under permanent duopoly: 
+\begin{itemize}
+\item[(a)] (\DG,\DG) is the unique pure-strategy Nash equilibrium for exploration-disadvantaged MAB instances with a sufficiently small ``warm start".
+\item[(b)] This is not necessarily the case for MAB instances that are not exploration-disadvantaged. In particular, \TS is a weakly dominating strategy for Needle-in-Haystack.
+\end{itemize}
+}
+\end{finding}
+
+%\begin{finding}
+%\textit{For sufficiently low warm start under the instances that are exploration-disadvantaged, the unique incentivized strategies in the competition game are:\footnote{The uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
+%\begin{center}
+%\textbf{Permanent Monopoly} - $\DG$ \\
+%\textbf{Temporary Monopoly} (incumbent)- $\TS$ \\
+%\textbf{Permanent Duopoly} - $(\DG, \DG)$ \\
+%\end{center}
+%The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.}
+%\end{finding}
+
+%\xhdr{Permanent Monopoly.} Since there is only a single firm in the
+%market for the entire period, the firm can take the entire market
+%regardless of what algorithm it deploys.
 \swedit{Since we assume that exploration algorithms ($\DEG$ and $\TS$) would incur deployment cost and firms break indifferences towards easier to deploy algorithms}, the firm would choose to deploy $\DG$.
-
-Next, to describe our results for temporary monopoly, we will
-introduce the notion of \textit{effective end of game (EEOG)} --- the
-last round $t$ such that the agent at round $(t-1)$ and the next agent
-around at $t$ choose different firms.
 
@@
   \hline
-  & \multicolumn{3}{c|}{Heavy-Tail} 
+  & \multicolumn{3}{c||}{Heavy-Tail}
   & \multicolumn{3}{c|}{Needle-in-Haystack}\\
@@
 \end{tabular}
-\caption{The duopoly experiment for Uniform and Needle-in-Haystack instances.
+\caption{Permanent duopoly, for Heavy-Tail and Needle-in-Haystack instances.
 Each cell describes a game between two algorithms, call them Alg1 vs.
 Alg2, for a particular value of the warm start $T_0$.
-The first line in the cell is the market share of Alg 1: the average (in bold) and the 95\% confidence band.
+Line 1 in the cell is the market share of Alg 1: the average (in bold) and the 95\% confidence band.
 %For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
-The second line specifies the ``effective end of game" (\Eeog): the average and the median (in brackets).}
+Line 2 specifies the ``effective end of game" (\Eeog): the average and the median (in brackets). The time horizon is $T=2000$.}
 \label{sim_table}
 \end{table*}
-\xhdr{Permanent Duopoly} \swedit{Since both firms enter at the same time, they will both get an expected market share of $1/2$ whenever they deploy
-  the same algorithms. We will mainly investigate the market shares
-  when the firms play different algorithms.  In Table \ref{sim_table}, we report the average market share taken over the
-  $N$ simulations for different values of warm-up length $T_0$, and
-  also the mean and median of the EEOG, under heavy tail and needle in
-  haystack instances.}
 
-For the exploration-disadvantaged instances, we see that, for low warm start, $\DG$ is a weakly dominant strategy for the competition game, and in particular weakly dominates $\TS$.
-This establishes our claim that $\DG$ is the incentivized algorithm.\footnote{We defer the table for uniform instances to the appendix.
+\subsubsection{Permanent Duopoly (Finding~\ref{find:duopoly}).} 
+We investigate the firms' market shares when they choose different algorithms (otherwise, by symmetry both firms get half of the agents).
+We report the market shares for Heavy-Tail and Needle-in-Haystack instances in Table \ref{sim_table}  (see the first line in each cell), for a range of values of the warm start $T_0$.
+
+We find that $\DG$ is a weakly dominant strategy for the Heavy-tail instance (with a sufficiently small $T_0$). However, \TS is a weakly dominating strategy for the Needle-in-Haystack instance.
+
+A similar table for the Uniform MAB instance is deferred to the appendix.
+We find that for a sufficiently small $T_0$, \DG yields more than half the market against \TS,  but achieves similar market share vs.
+\DG and \DEG.
+By our tie-breaking rule, (\DG,\DG) is the only pure-strategy equilibrium.
+
+\ascomment{stopped here.}
+
+\OMIT{This establishes our claim that $\DG$ is the incentivized algorithm.\footnote{We defer the table for uniform instances to the appendix.
 Summarizing the results, we see that, for low warm start, \DG yields more than half the market against \TS but is indifferent between \DG and \DEG.
-Since we break indifference towards easier to deploy algorithms, we also find that \DG is the incentivized algorithm in equilibrium}
+Since we break indifference towards easier to deploy algorithms, we also find that \DG is the incentivized algorithm in equilibrium}}
 
@@
 However, on the Needle In Haystack instances, we see that $\TS$ wins a larger proportion of the first agents compared to $\DG$.
+
+We introduce the notion of \textit{effective end of game (\Eeog)} the
+last round $t$ such that the agents at this and previous round choose different firms.
 
@@
 
-\xhdr{Temporary Monopoly.} Recall that the incumbent firm enters the
+\subsubsection{Temporary Monopoly (Finding~\ref{find:temp-monopoly}).} 
+Recall that the incumbent firm enters the
 market and serves as a monopolist until the entrant firm enters at
@@
 \label{ht_incum}
-\caption{The temporary monopoly experiment with $X=200$ (and $T_0=20$), for the Heavy-Tail MAB instance.
+\caption{Temporary monopoly, with $X=200$ (and $T_0=20$), for the Heavy-Tail MAB instance.
 Each cell describes the duopoly game between the entrant's algorithm (the row) and the incumbent's algorithm (the column).
-The cell specifies the entrant's market share for the rounds in which he was present.
+The cell specifies the entrant's market share (fraction of rounds in which it was chosen) for the rounds in which he was present.
 We give the average (in bold) and the 95\% confidence interval.

### 2018-09-05 paper/content/barriers.tex
edits

@@
 
-Together the two advantages serve as strong barriers to entry, and a
-natural question is whether either one of the advantages alone suffice
-as a strong barrier to entry. Also, which of the two is the stronger
-barrier?
+While the two advantages together serve as strong barriers to entry, a
+natural question to ask is whether either one of the advantages alone
+suffices as a strong barrier to entry. If both of them are barriers of
+entry, which of the two is the stronger?
 
-{To answer these questions, we isolate the effect from each of the two
-  advantages and run two additional experiments. In the first
-  experiment (called \emph{reputation erased experiment}), the
-  reputation of the incumbent is artificially reset upon the entry of
-  the entrant {so that the incumbent's reputation score is only based
-    on the rewards generated in the warm start period}. In the second
-  experiment (called \emph{information erased experiment}), the
-  information gained by the incumbent is erased upon the entry of the
-  entrant in the sense that its posterior is reset to the initial
-  (fake) prior.}
+To answer these questions, we isolate the effect from each of the two
+advantages and run two additional experiments. In the first experiment
+(called \emph{reputation erased experiment}), the reputation of the
+incumbent is artificially reset upon the entry of the entrant {so that
+  the incumbent's reputation score is only based on the rewards
+  generated in the warm start period}. In the second experiment
+(called \emph{information erased experiment}), the information gained
+by the incumbent is erased upon the entry of the entrant in the sense
+that its posterior is reset to the initial prior. The results of the
+experiments are presented in Table~\ref{barrier_exp}.
 
@@
 Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}).
-The cell specifies the entrant's market share in this game: the average (in bold) and the 95\% confidence interval.
+The cell specifies the entrant's market share for the rounds in which hit was present: the average (in bold) and the 95\% confidence interval.
 NB: smaller average is better for the incumbent.}
@@
 \begin{finding}
-\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger share of the market compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}
+\swedit{During the rounds in which the entrant is present, the incumbent
+  with either data or reputational advantage alone gets a larger share
+  than the share of an incumbent without any advantage.  When the
+  incumbent plays $\TS$, the data advantage serves as a larger barrier
+  to entry compared to reputation, but when the incumbent plays $\DEG$
+  or $\DG$ there is no substantial difference between the two
+  barriers.}
+\OMIT{\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger market share compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}}
 \end{finding}
 
-Table \ref{barrier_exp} display the results of these experiments.
-The reason for the finding is that, under $\TS$, the early entry allows the incumbent to do exploration without immediate reputational consequence in the early rounds and so, even when reputation is erased, the incumbent has already acquired sufficient information that simply exploiting that information allows her to quickly recover her reputation.
-However, while the reputational advantage gives the incumbent some cushion for doing exploration the incumbent is still vulnerable to getting unlucky while doing exploration and suffering the reputational consequences.
-This only happens under $\TS$ and not under $\DEG$ or $\DG$ since $\TS$, by construction, is more intelligent about the data that it acquires compared to $\DEG$ and $\DG$.
+The reason for the finding is that, under $\TS$,\swcomment{also true
+  for other algos?} the early entry allows the incumbent to do
+exploration without immediate reputational consequence in the early
+rounds and so, even when reputation is erased, the incumbent has
+already acquired sufficient information that simply exploiting that
+information allows her to quickly recover her
+reputation.\swcomment{which part are you explaining?} However, while
+the reputational advantage gives the incumbent some cushion for doing
+exploration the incumbent is still vulnerable to getting unlucky while
+doing exploration and suffering the reputational consequences. This
+only happens under $\TS$ since $\TS$, by construction, is more
+intelligent about the data that it acquires compared to $\DEG$ and
+$\DG$.
 

### 2018-09-05 paper/content/barriers.tex
edits

@@
 
+\swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}
+
 The reason for the finding is that, under $\TS$,\swcomment{also true
@@
 intelligent about the data that it acquires compared to $\DEG$ and
-$\DG$.
+$\DG$.\swcomment{more intelligent about the data means?}
 

### 2018-09-05 paper/content/barriers.tex
updated version of barriers

@@
   with either data or reputational advantage alone gets a larger share
-  than the share of an incumbent without any advantage.  When the
-  incumbent plays $\TS$, the data advantage serves as a larger barrier
-  to entry compared to reputation, but when the incumbent plays $\DEG$
-  or $\DG$ there is no substantial difference between the two
+  than the share \gaedit{in the permanent duopoly, regardless of the algorithm she plays}.
+When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two
   barriers.}
@@
 
-The reason for the finding is that, under $\TS$,\swcomment{also true
-  for other algos?} the early entry allows the incumbent to do
+\gaedit{While the data and reputation advantages seem to serve as substitutes for \DEG and \DG, it is interesting that under \TS the data advantage is larger.
+The reason for this is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}.
+Doing purposeful exploration in competition may have immediate adverse reputational consequences because exploration sometimes involves picking bad arms.
+However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms.
+As a result, the data advantage is larger for \TS precisely because the advantage that data gives is not just about the quantity of data but also about its distribution across actions.
+
+Why is it that data and reputation alone provide such large barriers to entry?
+Having more data allows the incumbent to select better actions which allows her to quickly get a high reputation compared to an entrant that both needs to learn and work on building up a reputation.
+Starting with a large reputation score gives the incumbent a cushion for doing exploration and learning without facing as stark reputational consequences.
+It is interesting that, for \DEG and \DG, these two effects seem to serve nearly as perfectly substitutes.
+}
+
+\gadelete{
+The reason \gadelete{for the finding} \gaedit{that we see that the data advantage is greater
+under $\TS$} is that the early entry allows the incumbent to do
 exploration without immediate reputational consequence in the early
@@
 information allows her to quickly recover her
-reputation.\swcomment{which part are you explaining?} However, while
+reputation. \swcomment{which part are you explaining?} However, while
 the reputational advantage gives the incumbent some cushion for doing
 exploration the incumbent is still vulnerable to getting unlucky while
-doing exploration and suffering the reputational consequences. This
+doing exploration and suffering the reputational consequences. Under all of the incumbent algorithms,
+the incumbent obtains a data advantage by the mere fact that it is around in the market and gets data from consumers.
+However, we see that the data advantage is stronger under \TS since \TS acquires data that is hard to acquire under competition.
+We only observe this under $\TS$ since
+\TS not only acquiresThis
 only happens under $\TS$ since $\TS$, by construction, is more
@@
 $\DG$.\swcomment{more intelligent about the data means?}
+}
 

### 2018-09-05 paper/content/barriers.tex
stuff

@@
 that its posterior is reset to the initial prior. The results of the
-experiments are presented in Table~\ref{barrier_exp}.
+experiments are presented in Table~\ref{barrier_exp}, which lead to
+our Finding~\ref{barrier-find}.
 
-\begin{table*}[h]
+\begin{table*}[h*]
 \centering
@@
 
-\begin{finding}
+\begin{finding}\label{barrier-find}
 \swedit{During the rounds in which the entrant is present, the incumbent
@@
 
-\swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}
+% \swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}
 
@@
 The reason for this is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}.
-Doing purposeful exploration in competition may have immediate adverse reputational consequences because exploration sometimes involves picking bad arms.
+In contrast, purposeful exploration in competition may have immediate adverse reputational consequences.
 However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms.

### 2018-09-05 paper/content/inverted_u.tex
Minor edit in the intro para.
Small edit.
Almost done with the revision.

@@
 As the ``intensity of competition" varies from permanent monopoly to temporary monopoly to permanent duopoly, we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
-
 More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
 We do this for each "intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
-
 For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{MWT-WhitePaper-2016,DS-arxiv}).
-Then \DG is trivially the dominant strategy under permanent monopoly.
+Note that \DG is trivially the dominant strategy under permanent monopoly.
 
@@
 
-For temporary monopoly, we find that \TS dominates:
-
-\begin{finding}\label{find:temp-monopoly}
-\textit{Under temporary monopoly, TS is the dominant strategy for the incumbent, across all MAB instances, if the temporary monopoly is sufficiently long-lasting.
-}
-\end{finding}
-
-For permanent duopoly, it is a crucial whether an MAB instance is exploration-disadvantaged:
+\subsubsection{Permanent duopoly.}
+A crucial distinction in our findings is whether an MAB instance is exploration-disadvantaged:
 
@@
 % Thu Aug 16 13:13:01 2018
-\begin{table*}[h]
+\begin{table*}[t]
 \centering
@@
   & \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 200 (27)}
-  & \makecell{\textbf{0.6} $\pm$0.03\\ \Eeog 370 (0}
+  & \makecell{\textbf{0.6} $\pm$0.03\\ \Eeog 370 (0)}
   & \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 580 (122)}
@@
   %%
-  & \makecell{\textbf{0.57} $\pm$0.03\\ \Eeog 150 (14}
+  & \makecell{\textbf{0.57} $\pm$0.03\\ \Eeog 150 (14)}
   & \makecell{\textbf{0.52} $\pm$0.03\\ \Eeog 460 (79)}
@@
 \end{tabular}
-\caption{Permanent duopoly, for Heavy-Tail and Needle-in-Haystack instances.
+\caption{{\bf Permanent duopoly}, for Heavy-Tail and Needle-in-Haystack instances.
 Each cell describes a game between two algorithms, call them Alg1 vs.
@@
 
-\subsubsection{Permanent Duopoly (Finding~\ref{find:duopoly}).} 
 We investigate the firms' market shares when they choose different algorithms (otherwise, by symmetry both firms get half of the agents).
 We report the market shares for Heavy-Tail and Needle-in-Haystack instances in Table \ref{sim_table}  (see the first line in each cell), for a range of values of the warm start $T_0$.
-
-We find that $\DG$ is a weakly dominant strategy for the Heavy-tail instance (with a sufficiently small $T_0$). However, \TS is a weakly dominating strategy for the Needle-in-Haystack instance.
-
+We find that $\DG$ is a weakly dominant strategy for the Heavy-tail instance, as long as $T_0$ is sufficiently small.
+However, \TS is a weakly dominating strategy for the Needle-in-Haystack instance.
 A similar table for the Uniform MAB instance is deferred to the appendix.
@@
 
-\ascomment{stopped here.}
-
 \OMIT{This establishes our claim that $\DG$ is the incentivized algorithm.\footnote{We defer the table for uniform instances to the appendix.
@@
 
+We attribute the prevalence of \DG on exploration-disadvantaged MAB instances to its prevalence on the initial ``exploration disadvantage period", as described in Section~\ref{sec:isolation}.
+Increasing the warm start length $T_0$ makes this period shorter: indeed, considering relative reputation trajectory in Figure~\ref{relative_rep_plots} (top), increasing $T_0$ effectively shifts the starting time point to the right.
+This is why it helps \DG if $T_0$ is small.
+
+\OMIT{
 The reason we see that $\DG$ is incentivized under the Heavy Tail and Uniform instances but that $\TS$ is weakly dominant under Needle In Haystack is precisely due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds puts it in a disadvantage in the former case but not in the latter.
@@
 However, on the Needle In Haystack instances, we see that $\TS$ wins a larger proportion of the first agents compared to $\DG$.
+}
 
-We introduce the notion of \textit{effective end of game (\Eeog)} the
-last round $t$ such that the agents at this and previous round choose different firms.
+\subsubsection{Temporary Monopoly.}
+Recall that the incumbent firm enters the market and serves as a monopolist until the entrant firm enters at round $X$.
+We make $X$ large enough, but still much smaller than the time horizon $T$.
+We find that the incumbent is incentivized to choose \TS, in a strong sense:
 
+\begin{finding}\label{find:temp-monopoly}
+\textit{Under temporary monopoly, \TS is the dominant strategy for the incumbent. This holds across all MAB instances, if $X$ is large enough.
+}
+\end{finding}
+
+The simulation results for the Heavy-Tail MAB instance are reported in Table~\ref{tab:ht-incum}, for a particular $X=200$.
+We see that \TS is a dominant strategy for the incumbent.
+Similar tables for the other MAB instances and other values of $X$ are reported in the supplement, with the same conclusion.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+   & $\TS$  & $\DEG$  & $\DG$ \\ \hline
+$\TS$
+    & \makecell{\textbf{0.003}$\pm$0.003}
+    & \makecell{\textbf{0.083}$\pm$0.02}
+    & \makecell{\textbf{0.17}$\pm$0.02} \\\hline
+$\DEG$
+    & \makecell{\textbf{0.045}$\pm$0.01}
+    & \makecell{\textbf{0.25}$\pm$0.02}
+    & \makecell{\textbf{0.23}$\pm$0.02} \\\hline
+$\DG$
+    & \makecell{\textbf{0.12}$\pm$0.02}
+    & \makecell{\textbf{0.36}$\pm$0.03}
+    & \makecell{\textbf{0.3}$\pm$0.02} \\\hline
+\end{tabular}
+\caption{{\bf Temporary monopoly}, with $X=200$ (and $T_0=20$), for the Heavy-Tail MAB instance.
+Each cell describes the duopoly game between the entrant's algorithm (the row) and the incumbent's algorithm (the column).
+The cell specifies the entrant's market share (fraction of rounds in which it was chosen) for the rounds in which he was present.
+We give the average (in bold) and the 95\% confidence interval.
+NB: smaller average is better for the incumbent.}
+\label{tab:ht-incum}
+\end{table}
+
+\DG is a weakly dominant strategy for the entrant, for Heavy-Tail instance in Table~\ref{tab:ht-incum} and the Uniform instance, but not for the Needle-in-Haystack instance.
+We attribute this finding to exploration-disadvantaged property of these two MAB instances, for the same reasons as discussed above.
+
+\begin{finding}\label{find:temp-monopoly-entrant}
+\textit{Under temporary monopoly with a sufficiently large $X$, \DG is a weakly dominant strategy for the incumbent for exploration-disadvantaged instances.
+}
+\end{finding}
+
+\OMIT{
+In Table~\ref{tab:ht-incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$ and $\DG$ is only weakly dominant.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the instances we consider.}
+
+\subsubsection{Inverted-U relationship.}
+
+\ascomment{Describe the inverted-U in the findings above.}
+\ascomment{stopped here}
+
+Why permanent duopoly creates different 
+Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was the incentivized algorithm?
+The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
+One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the ``incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
+The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.
+
+As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same immediate relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
+Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
+
+\subsubsection{Death spiral effect.}
+Further, we investigate the ``death spiral" effect mentioned in the Introduction.
+Restated in terms of our model, the effect is that one firm attracts new customers at a lower rate than the other, and falls behind in terms of performance because the other firm has more customer to learn from, and this gets worse over time until (almost) all new customers go to the other firm.
+With this intuition in mind, we define  \textit{effective end of game (\Eeog)} for a particular \MRV and realization table, as the last round $t$ such that the agents at this and previous round choose different firms.
+Indeed, the game, effectively, ends after this round.
+We interpret low \Eeog as a strong evidence of the ``death spiral" effect.
+Focusing on the permanent duopoly scanario, we specify the \Eeog values in Table \ref{sim_table} (the second line of each cell).
+We find that the \Eeog values are indeed small:
+
+\begin{finding}
+\textit{
+Under permanent duopoly, \Eeog values tend to be much smaller than the time horizon $T$, but tend to increase as the warm start $T_0$ increases.
+}
+\end{finding}
+
+\ascomment{Can we cook up an explanation for why \Eeog grows with $T_0$?}
+
+\OMIT{ %%%
+Enter into a ``death spiral" where the winner keeps learning and improving from its customers thus increasing its reputation while the losing firm gets starved of customers and progressively falls farther and farther behind its competitors.
 What happens after the first agent?
@@
 As a result, even though if $\TS$ could get sufficiently many agents it would eventually have a higher reputation than $\DG$, since it starts the game by losing agents to $\DG$ due to its lowered reputation from exploration in the warm start it ends up doing worse in the competition game.
+} %%%%%
 
@@
 
-\subsubsection{Temporary Monopoly (Finding~\ref{find:temp-monopoly}).} 
-Recall that the incumbent firm enters the
-market and serves as a monopolist until the entrant firm enters at
-round $X$.
-\gaedit{For this section we focus on the warm start values and on the exploration-disadvantaged instances where we observed that, under permanent duopoly, $\DG$ was the weakly dominant strategy.
-Concretely, our presented results are for $T_0 = 20$.} \footnote{However, we report the results for the Needle In Haystack instances in the supplementary material.
-The results are unsurprising, where $\TS$ is the dominant
-  strategy for both the entrant and incumbent.}
-
-\begin{finding}
-  \textit{If the incumbent firm is a temporary monopolist for
-    a sufficiently long time, it will deploy the better learning
-    algorithm, $\TS$, even though, under the permanent duopoly, they
-    would commit to $\DG$.}
-\end{finding}
-
-In Table \ref{ht_incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$ and $\DG$ is only weakly dominant.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the
-instances we consider.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|c|c|c|c|}
-\hline
-   & $\TS$  & $\DEG$  & $\DG$ \\ \hline
-$\TS$
-    & \makecell{\textbf{0.003}$\pm$0.003}
-    & \makecell{\textbf{0.083}$\pm$0.02}
-    & \makecell{\textbf{0.17}$\pm$0.02} \\\hline
-$\DEG$
-    & \makecell{\textbf{0.045}$\pm$0.01}
-    & \makecell{\textbf{0.25}$\pm$0.02}
-    & \makecell{\textbf{0.23}$\pm$0.02} \\\hline
-$\DG$
-    & \makecell{\textbf{0.12}$\pm$0.02}
-    & \makecell{\textbf{0.36}$\pm$0.03}
-    & \makecell{\textbf{0.3}$\pm$0.02} \\\hline
-\end{tabular}
-\label{ht_incum}
-\caption{Temporary monopoly, with $X=200$ (and $T_0=20$), for the Heavy-Tail MAB instance.
-Each cell describes the duopoly game between the entrant's algorithm (the row) and the incumbent's algorithm (the column).
-The cell specifies the entrant's market share (fraction of rounds in which it was chosen) for the rounds in which he was present.
-We give the average (in bold) and the 95\% confidence interval.
-NB: smaller average is better for the incumbent.}
-\end{table}
-
 \OMIT{ %%%%%%%
@@
 
-Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was the incentivized algorithm?
-The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
-One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the ``incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
-The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.
-
-As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same immediate relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
-Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
-
 \end{document}

### 2018-09-05 paper/content/revisited.tex
fix broken reference

@@
 Mean reputation trajectories do not suffice to explain the outcomes under competition.
-We observe in Figure \ref{sim_ht} that $\DG$ attains a larger market share than $\DEG$ even for large warm starts.\footnote{In the appendix we show that this holds for $K = 3$ and longer time horizons than we consider here as well.}}
+We observe in Figure \ref{sim_table} that $\DG$ attains a larger market share than $\DEG$ even for large warm starts.\footnote{In the appendix we show that this holds for $K = 3$ and longer time horizons than we consider here as well.}}
 \end{finding}

### 2018-09-05 paper/content/model.tex
more cosmetic fixes

@@
 
-\item an adaptive-exploration algorithm called "Thompson Sampling" (\TS).
+\item an adaptive-exploration algorithm called ``Thompson Sampling" (\TS).
 In each round, this algorithm updates the posterior distribution for the mean reward of each arm $a$, draws an independent sample $s_a$ from this distribution, and chooses an arm with the largest $s_a$.
@@
 \end{enumerate}
-\asedit{We argue that these MAB instances are (somewhat) representative.
+We argue that these MAB instances are (somewhat) representative.
 Consider the ``gap" between the best and the second-best arm, an essential parameter in the literature on MAB.
-The ``gap" is fixed in Needle-in-Haystack, spread over a wide spectrum of values under the Uniform instance, and is spread but  focused on the large values under the Heavy-Tail instance.}
+The ``gap" is fixed in Needle-in-Haystack, spread over a wide spectrum of values under the Uniform instance, and is spread but  focused on the large values under the Heavy-Tail instance.
 

### 2018-09-05 paper/content/inverted_u.tex
Done with the revision (with one small caveat).

@@
 Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from permanent monopoly to temporary monopoly to permanent duopoly, we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
+As the ``intensity of competition" varies from permanent monopoly to ``early entry" to permanent duopoly to ``late entry", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
 More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
@@
 \subsubsection{Permanent duopoly.}
-A crucial distinction in our findings is whether an MAB instance is exploration-disadvantaged:
+The basic scenario is when both firms are competing from round $1$. A crucial distinction is whether an MAB instance is exploration-disadvantaged:
 
@@
 \subsubsection{Temporary Monopoly.}
+We turn our attention to the temporary monopoly scenario.
 Recall that the incumbent firm enters the market and serves as a monopolist until the entrant firm enters at round $X$.
@@
 \DG is a weakly dominant strategy for the entrant, for Heavy-Tail instance in Table~\ref{tab:ht-incum} and the Uniform instance, but not for the Needle-in-Haystack instance.
-We attribute this finding to exploration-disadvantaged property of these two MAB instances, for the same reasons as discussed above.
+We attribute this finding to exploration-disadvantaged property of these two MAB instance, for the same reasons as discussed above.
 
@@
 \subsubsection{Inverted-U relationship.}
+We interpret our findings through the lens of the inverted-U relationship between the ``intensity of competition" and the ``quality of technology".
+The lowest level of competition is monopoly, when \DG wins out for the trivial reason of tie-breaking.
+The highest levels are permanent duopoly and ``late entry" (temporary monopoly from the entrant's perspective).
+We see that \DG is incentivized for exploration-disadvantaged MAB instances.
+In fact, incentives for \DG get stronger when the model transitions from permanent duopoly to ``late entry".%
+\footnote{For the Heavy-Tail instance, \DG goes from a weakly dominant startegy to a strictly dominant one.
+For the Uniform instance, \DG goes from a Nash equilibrium strategy to a weakly dominant one.}
+Finally, the middle level of competition, ``early entry" (temporary monopoly for the entrant) creates strong incentives for \TS.
+In  stylized form, this relationship is captured in Figure~\ref{fig:inverted-U}.
 
-\ascomment{Describe the inverted-U in the findings above.}
-\ascomment{stopped here}
+% Transition Duopoly -> LateStart creates stronger incentives for DG.
+% In Heavy Tail, DG is strictly dominant (before it was only weakly).
+% In Uniform, DG is weakly dominant (before it was only PSNE).
 
-Why permanent duopoly creates different 
+Our intuition for why ``early entry" creates more incentives for exploration is as follows.
+During the temporary monopoly period, reputation costs of exploration vanish.
+Instead, the firm wants to improve its performance as much as possible by the time competition starts.
+Essentially, the firm only faces a classical explore-exploit tradeoff, and is incentivized to choose algorithms that are best at optimizing this tradeoff.
+
+\OMIT{ %%%%%%%%%
 Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was the incentivized algorithm?
@@
 Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
+} %%%%%%%%
 

### 2018-09-05 paper/content/barriers.tex
Fixed a compile bug: \begin{table*}[h*] -> \begin{table*}[h]
updatd tables

@@
 \section{Data and Reputation as Barriers to Entry}\label{section:6}
-In the temporary monopoly setting, the incumbent firm enjoys a first
-mover advantage, which allows the firm to acquire data on potentially
-suboptimal actions without incurring an immediate reputational
-consequence, and to build up a high reputation before the entrant
-enters the market. Thus, the early entry gives the incumbent both a
-\textit{data} advantage and a \textit{reputational} advantage over the
-entrant. This provides a quantitative perspective on the classic
-``first mover advantage" \cite{kerin1992first} in the digital economy
-where data can serve as a barrier to entry in many markets.
 
-While the two advantages together serve as strong barriers to entry, a
-natural question to ask is whether either one of the advantages alone
-suffices as a strong barrier to entry. If both of them are barriers of
-entry, which of the two is the stronger?
-
-To answer these questions, we isolate the effect from each of the two
-advantages and run two additional experiments. In the first experiment
-(called \emph{reputation erased experiment}), the reputation of the
-incumbent is artificially reset upon the entry of the entrant {so that
-  the incumbent's reputation score is only based on the rewards
-  generated in the warm start period}. In the second experiment
-(called \emph{information erased experiment}), the information gained
-by the incumbent is erased upon the entry of the entrant in the sense
-that its posterior is reset to the initial prior. The results of the
-experiments are presented in Table~\ref{barrier_exp}, which lead to
-our Finding~\ref{barrier-find}.
-
-\begin{table*}[h*]
+\begin{table*}[h]
 \centering
@@
 \hline
-  & \multicolumn{3}{c|}{Information erased}
+  & \multicolumn{3}{c||}{Information erased}
   & \multicolumn{3}{c|}{Reputation erased}\\
@@
 \end{table*}
+
+In the temporary monopoly setting, the incumbent firm enjoys a first
+mover advantage, which allows the firm to acquire data on potentially
+suboptimal actions without incurring an immediate reputational
+consequence, and to build up a high reputation before the entrant
+enters the market. Thus, the early entry gives the incumbent both a
+\textit{data} advantage and a \textit{reputational} advantage over the
+entrant. This provides a quantitative perspective on the classic
+``first mover advantage" \cite{kerin1992first} in the digital economy
+where data can serve as a barrier to entry in many markets.
+
+While the two advantages together serve as strong barriers to entry, a
+natural question to ask is whether either one of the advantages alone
+suffices as a strong barrier to entry. If both of them are barriers of
+entry, which of the two is the stronger?
+
+To answer these questions, we isolate the effect from each of the two
+advantages and run two additional experiments. In the first experiment
+(called \emph{reputation erased experiment}), the reputation of the
+incumbent is artificially reset upon the entry of the entrant {so that
+  the incumbent's reputation score is only based on the rewards
+  generated in the warm start period}. In the second experiment
+(called \emph{information erased experiment}), the information gained
+by the incumbent is erased upon the entry of the entrant in the sense
+that its posterior is reset to the initial prior. The results of the
+experiments are presented in Table~\ref{barrier_exp}, which lead to
+our Finding~\ref{barrier-find}.
 

### 2018-09-05 paper/content/barriers.tex
more table placements

@@
 \section{Data and Reputation as Barriers to Entry}\label{section:6}
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|c|c|c|c||c|c|c|}
-\hline
-  & \multicolumn{3}{c||}{Information erased}
-  & \multicolumn{3}{c|}{Reputation erased}\\
-\hline
-& $\TS$  & $\DEG$  & $\DG$
-& $\TS$  & $\DEG$  & $\DG$
-\\\hline
-$\TS$
-    & \makecell{\textbf{0.021}$\pm$0.009}
-    & \makecell{\textbf{0.16}$\pm$0.02}
-    & \makecell{\textbf{0.21} $\pm$0.02}
-    %%
-    & \makecell{\textbf{0.0096}$\pm$0.006}
-    & \makecell{\textbf{0.11}$\pm$0.02}
-    & \makecell{\textbf{0.18}$\pm$0.02}
-    \\ \hline
-$\DEG$
-    & \makecell{\textbf{0.26}$\pm$0.03}
-    & \makecell{\textbf{0.3}$\pm$0.02}
-    & \makecell{\textbf{0.26}$\pm$0.02}
-    %%
-    & \makecell{\textbf{0.073}$\pm$0.01}
-    & \makecell{\textbf{0.29}$\pm$0.02}
-    & \makecell{\textbf{0.25}$\pm$0.02}
-    \\ \hline
-$\DG$
-    & \makecell{\textbf{0.34}$\pm$0.03}
-    & \makecell{\textbf{0.4}$\pm$0.03}
-    & \makecell{\textbf{0.33}$\pm$0.02}
-    %%
-    & \makecell{\textbf{0.15}$\pm$0.02}
-    & \makecell{\textbf{0.39}$\pm$0.03}
-    & \makecell{\textbf{0.33}$\pm$0.02}
-    \\\hline
-\end{tabular}
-\caption{Information-erased vs.
-reputation-erased experiment, on Heavy-Tail MAB instance.
-Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}).
-The cell specifies the entrant's market share for the rounds in which hit was present: the average (in bold) and the 95\% confidence interval.
-NB: smaller average is better for the incumbent.}
-\label{barrier_exp}
-\end{table*}
 

### 2018-09-05 paper/content/inverted_u.tex
updatd tables
more table placements

@@
 \section{Competition and Better MAB Algorithms}\label{section:5}
-
-Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from permanent monopoly to ``early entry" to permanent duopoly to ``late entry", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
-More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
-We do this for each "intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
-For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{MWT-WhitePaper-2016,DS-arxiv}).
-Note that \DG is trivially the dominant strategy under permanent monopoly.
-
-%(1) Permanent monopoly - indifferent between all algorithms but break indifference towards DG cause of deployment costs
-%(2) Temporary monopoly - TS is the dominant strategy for the incumbent
-%(3) Permanent duopoly -
-%   (a) Under NIH, TS is dominant
-%   (b) Under Heavy Tail, DG is weakly dominant
-%   (c) Under Uniform, DG "beats" TS, but DEG "beats" TS by more than DG and DEG vs DG leads to ~50% market share (it's slightly in favor of DG but 50/50 is within the confidence band).
-If we break indifference towards DG cause of deployment costs, then we have that the unique PSNE is (DG, DG).
-If we don't add the indifference breaking, then we have four PSNE, (DG, DEG), (DEG, DG), (DG, DG), (DEG, DEG)
-
-\subsubsection{Permanent duopoly.}
-The basic scenario is when both firms are competing from round $1$. A crucial distinction is whether an MAB instance is exploration-disadvantaged:
-
-\begin{finding}\label{find:duopoly}
-\textit{Under permanent duopoly:
-\begin{itemize}
-\item[(a)] (\DG,\DG) is the unique pure-strategy Nash equilibrium for exploration-disadvantaged MAB instances with a sufficiently small ``warm start".
-\item[(b)] This is not necessarily the case for MAB instances that are not exploration-disadvantaged. In particular, \TS is a weakly dominating strategy for Needle-in-Haystack.
-\end{itemize}
-}
-\end{finding}
-
-%\begin{finding}
-%\textit{For sufficiently low warm start under the instances that are exploration-disadvantaged, the unique incentivized strategies in the competition game are:\footnote{The uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
-%\begin{center}
-%\textbf{Permanent Monopoly} - $\DG$ \\
-%\textbf{Temporary Monopoly} (incumbent)- $\TS$ \\
-%\textbf{Permanent Duopoly} - $(\DG, \DG)$ \\
-%\end{center}
-%The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.}
-%\end{finding}
-
-%\xhdr{Permanent Monopoly.} Since there is only a single firm in the
-%market for the entire period, the firm can take the entire market
-%regardless of what algorithm it deploys.
-\swedit{Since we assume that exploration algorithms ($\DEG$ and $\TS$) would incur deployment cost and firms break indifferences towards easier to deploy algorithms}, the firm would choose to deploy $\DG$.
 
@@
 \end{table*}
+
+Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
+As the ``intensity of competition" varies from permanent monopoly to ``early entry" to permanent duopoly to ``late entry", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
+More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
+We do this for each "intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
+For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{MWT-WhitePaper-2016,DS-arxiv}).
+Note that \DG is trivially the dominant strategy under permanent monopoly.
+
+%(1) Permanent monopoly - indifferent between all algorithms but break indifference towards DG cause of deployment costs
+%(2) Temporary monopoly - TS is the dominant strategy for the incumbent
+%(3) Permanent duopoly -
+%   (a) Under NIH, TS is dominant
+%   (b) Under Heavy Tail, DG is weakly dominant
+%   (c) Under Uniform, DG "beats" TS, but DEG "beats" TS by more than DG and DEG vs DG leads to ~50% market share (it's slightly in favor of DG but 50/50 is within the confidence band).
+If we break indifference towards DG cause of deployment costs, then we have that the unique PSNE is (DG, DG).
+If we don't add the indifference breaking, then we have four PSNE, (DG, DEG), (DEG, DG), (DG, DG), (DEG, DEG)
+
+\subsubsection{Permanent duopoly.}
+The basic scenario is when both firms are competing from round $1$. A crucial distinction is whether an MAB instance is exploration-disadvantaged:
+
+\begin{finding}\label{find:duopoly}
+\textit{Under permanent duopoly:
+\begin{itemize}
+\item[(a)] (\DG,\DG) is the unique pure-strategy Nash equilibrium for exploration-disadvantaged MAB instances with a sufficiently small ``warm start".
+\item[(b)] This is not necessarily the case for MAB instances that are not exploration-disadvantaged. In particular, \TS is a weakly dominating strategy for Needle-in-Haystack.
+\end{itemize}
+}
+\end{finding}
+
+%\begin{finding}
+%\textit{For sufficiently low warm start under the instances that are exploration-disadvantaged, the unique incentivized strategies in the competition game are:\footnote{The uniqueness comes from the fact that we break indifferences towards easier to deploy algorithms}
+%\begin{center}
+%\textbf{Permanent Monopoly} - $\DG$ \\
+%\textbf{Temporary Monopoly} (incumbent)- $\TS$ \\
+%\textbf{Permanent Duopoly} - $(\DG, \DG)$ \\
+%\end{center}
+%The conditions on $\TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.}
+%\end{finding}
+
+%\xhdr{Permanent Monopoly.} Since there is only a single firm in the
+%market for the entire period, the firm can take the entire market
+%regardless of what algorithm it deploys.
+\swedit{Since we assume that exploration algorithms ($\DEG$ and $\TS$) would incur deployment cost and firms break indifferences towards easier to deploy algorithms}, the firm would choose to deploy $\DG$.
 
@@
 In Table~\ref{tab:ht-incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$ and $\DG$ is only weakly dominant.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the instances we consider.}
+
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|c|c|c|c||c|c|c|}
+\hline
+  & \multicolumn{3}{c||}{Information erased}
+  & \multicolumn{3}{c|}{Reputation erased}\\
+\hline
+& $\TS$  & $\DEG$  & $\DG$
+& $\TS$  & $\DEG$  & $\DG$
+\\\hline
+$\TS$
+    & \makecell{\textbf{0.021}$\pm$0.009}
+    & \makecell{\textbf{0.16}$\pm$0.02}
+    & \makecell{\textbf{0.21} $\pm$0.02}
+    %%
+    & \makecell{\textbf{0.0096}$\pm$0.006}
+    & \makecell{\textbf{0.11}$\pm$0.02}
+    & \makecell{\textbf{0.18}$\pm$0.02}
+    \\ \hline
+$\DEG$
+    & \makecell{\textbf{0.26}$\pm$0.03}
+    & \makecell{\textbf{0.3}$\pm$0.02}
+    & \makecell{\textbf{0.26}$\pm$0.02}
+    %%
+    & \makecell{\textbf{0.073}$\pm$0.01}
+    & \makecell{\textbf{0.29}$\pm$0.02}
+    & \makecell{\textbf{0.25}$\pm$0.02}
+    \\ \hline
+$\DG$
+    & \makecell{\textbf{0.34}$\pm$0.03}
+    & \makecell{\textbf{0.4}$\pm$0.03}
+    & \makecell{\textbf{0.33}$\pm$0.02}
+    %%
+    & \makecell{\textbf{0.15}$\pm$0.02}
+    & \makecell{\textbf{0.39}$\pm$0.03}
+    & \makecell{\textbf{0.33}$\pm$0.02}
+    \\\hline
+\end{tabular}
+\caption{Information-erased vs.
+reputation-erased experiment, on Heavy-Tail MAB instance.
+Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}).
+The cell specifies the entrant's market share for the rounds in which hit was present: the average (in bold) and the 95\% confidence interval.
+NB: smaller average is better for the incumbent.}
+\label{barrier_exp}
+\end{table*}
 

### 2018-09-05 paper/content/inverted_u.tex
cosmetic fixups

@@
 More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
-We do this for each "intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
+We do this for each ``intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
 For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{MWT-WhitePaper-2016,DS-arxiv}).
@@
 \item[(a)] (\DG,\DG) is the unique pure-strategy Nash equilibrium for exploration-disadvantaged MAB instances with a sufficiently small ``warm start".
-\item[(b)] This is not necessarily the case for MAB instances that are not exploration-disadvantaged. In particular, \TS is a weakly dominating strategy for Needle-in-Haystack.
+\item[(b)] This is not necessarily the case for MAB instances that are not exploration-disadvantaged. In particular, \TS is a weakly dominant strategy for Needle-in-Haystack.
 \end{itemize}
@@
 We find that $\DG$ is a weakly dominant strategy for the Heavy-tail instance, as long as $T_0$ is sufficiently small.
-However, \TS is a weakly dominating strategy for the Needle-in-Haystack instance.
+However, \TS is a weakly dominant strategy for the Needle-in-Haystack instance.
 A similar table for the Uniform MAB instance is deferred to the appendix.
@@
 Further, we investigate the ``death spiral" effect mentioned in the Introduction.
-Restated in terms of our model, the effect is that one firm attracts new customers at a lower rate than the other, and falls behind in terms of performance because the other firm has more customer to learn from, and this gets worse over time until (almost) all new customers go to the other firm.
+Restated in terms of our model, the effect is that one firm attracts new customers at a lower rate than the other, and falls behind in terms of performance because the other firm has more customers to learn from, and this gets worse over time until (almost) all new customers go to the other firm.
 With this intuition in mind, we define  \textit{effective end of game (\Eeog)} for a particular \MRV and realization table, as the last round $t$ such that the agents at this and previous round choose different firms.

### 2018-09-05 paper/content/perf_in_iso.tex
Changed section label.
cosmetic fixups

@@
 
-\section{Algorithms' Performance in Isolation}\label{section:4}
+\section{Algorithms' Performance in Isolation}\label{sec:isolation}
 
@@
 
-\ascomment{Rewrote from here down.}
-
 The mean reputation trajectory is probably the most natural way to represent an algorithm's performance on a given MAB instance.
@@
 The \emph{relative reputation} of Alg1 (vs.
-Alg2) at a given time $t$ is the fraction of \MRVs / realization tables for which Alg1 has a higher reputation score than Alg2.
+Alg2) at a given time $t$ is the fraction of \MRVs /realization tables for which Alg1 has a higher reputation score than Alg2.
 The intuition is that agent's selection in our model depends only on the comparison between the reputation scores.

### 2018-09-05 paper/content/introduction.tex
more cosmetic fixes
update intro

@@
 
-We ask whether competition incentivizes adoption of "better" algorithms for exploration.
+We ask whether competition incentivizes adoption of ``better" algorithms for exploration.
 We investigate this issue via extensive numerical experiments in a stylized duopoly model.
@@
 Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage".
+\gadelete{In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage".
 We further investigate which of the two is a stronger barrier to entry.
 We find that both are strong barriers on its own: removing either one substantially helps the late entrant.
+Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.}
+
+\gaedit{Being first in the market gives firms free data to learn from (a ``data advantage") as well a larger reputation compared to an entrant (a ``reputation advantage").
+We investigate which of the two is a stronger barrier to entry.
+We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed.
 Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
+This occurs because the lack of initial competition allows the incumbent to gather data on actions that may be harder to gather in competition (due to reputational consequences) but that are useful for learning and maximizing long-run reputation.
+}
 

### 2018-09-05 paper/content/introduction.tex
Edits in inverted-U and "barriers" paragraphs.

@@
 
-We consider varying levels of competition, from permanent monopoly to \emph{temporary monopoly} (a duopoly with an early entrant), to a duopoly in which both firms start at the same time.
-We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly.
-This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
+\asedit{We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with an early entrant.
+Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``early entry" (in the temporary monopoly) to permanent duopoly to ``late entry" (temporary monopoly from the incumbent's perspective).
+We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly, and even more so under ``late entry".} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
 Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run.
-The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model).
+The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model, aside from the tie-breaking).
 
@@
 Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
+Being first in the market gives firms free data to learn from (a ``data advantage") as well as \asedit{a more definite (and possibly better)} reputation compared to an entrant (a ``reputation advantage").
+We investigate which of the two is a stronger barrier to entry.
+We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed.
+Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
+
+\OMIT{This occurs because the lack of initial competition allows the incumbent to gather data on actions that may be harder to gather in competition (due to reputational consequences) but that are useful for learning and maximizing long-run reputation.}
+
+\OMIT{ %%%%%
 \gadelete{In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage".
@@
 Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.}
-
-\gaedit{Being first in the market gives firms free data to learn from (a ``data advantage") as well a larger reputation compared to an entrant (a ``reputation advantage").
-We investigate which of the two is a stronger barrier to entry.
-We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed.
-Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
-This occurs because the lack of initial competition allows the incumbent to gather data on actions that may be harder to gather in competition (due to reputational consequences) but that are useful for learning and maximizing long-run reputation.
-}
+} %%%%%%
 

### 2018-09-05 paper/content/model.tex
new plot...also fix computation details

@@
 The ``gap" is fixed in Needle-in-Haystack, spread over a wide spectrum of values under the Uniform instance, and is spread but  focused on the large values under the Heavy-Tail instance.
+\asedit{We also ran smaller experiments with versions of these instances, and achieved similar qualitative results.}
 
@@
 An experiment on a particular MAB instance comprised multiple runs of the competition game: $N$ mean reward vectors times $9$ pairs of algorithms times three values for the warm start.
-We used a parallel implementation over a cluster of 12 2200Hz CPU cores, with 8 GB RAM per core.
-With this implementation, each experiment took about 10.5 hours.
+We used a parallel implementation over a cluster of 12 2.2 GHz CPU cores, with 8 GB RAM per core.
+With this implementation, each experiment took about $10$ hours.
 

### 2018-09-05 paper/content/inverted_u.tex
Added a para re why EoG values increase with T0.

@@
 \textit{
-Under permanent duopoly, \Eeog values tend to be much smaller than the time horizon $T$, but tend to increase as the warm start $T_0$ increases.
+Under permanent duopoly, \Eeog values tend to be much smaller than the time horizon $T$. 
 }
@@
 
-\ascomment{Can we cook up an explanation for why \Eeog grows with $T_0$?}
+\asedit{We also see that the \Eeog values tend to increase as the warm start $T_0$ increases.
+We conjecture this is because larger $T_0$ tends to help a better algorithm.
+Indeed, we know that the ``effective end of game" in this scenario typically occurs when a better algorithm loses, and helping it delays the loss.}
 

### 2018-09-05 paper/content/barriers.tex
Half-way done with a pass.

@@
 
-In the temporary monopoly setting, the incumbent firm enjoys a first
-mover advantage, which allows the firm to acquire data on potentially
-suboptimal actions without incurring an immediate reputational
-consequence, and to build up a high reputation before the entrant
-enters the market. Thus, the early entry gives the incumbent both a
-\textit{data} advantage and a \textit{reputational} advantage over the
-entrant. This provides a quantitative perspective on the classic
-``first mover advantage" \cite{kerin1992first} in the digital economy
-where data can serve as a barrier to entry in many markets.
+Under temporary monopoly, the incumbent can explore without incurring immediate reputational costs, and build up a high reputation before the entrant appears.
+Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant.
+While both advantages create a strong barrier to entry, we investigate whether either one is a strong barrier alone, and which one is stronger.
+Our findings provide a quantitative insight into the classic ``first mover advantage" phenomenon in the digital economy.
 
-While the two advantages together serve as strong barriers to entry, a
-natural question to ask is whether either one of the advantages alone
-suffices as a strong barrier to entry. If both of them are barriers of
-entry, which of the two is the stronger?
+For a more succinct terminology, recall that the incumbent enjoys an extended warm start of $X+T_0$ rounds.
+Call the first $X$ of these rounds the \emph{monopoly period} (and the rest is the proper ``warm start").
+The rounds when both firms are competing for customers are called \emph{competition period.}
 
-To answer these questions, we isolate the effect from each of the two
-advantages and run two additional experiments. In the first experiment
-(called \emph{reputation erased experiment}), the reputation of the
-incumbent is artificially reset upon the entry of the entrant {so that
-  the incumbent's reputation score is only based on the rewards
-  generated in the warm start period}. In the second experiment
-(called \emph{information erased experiment}), the information gained
-by the incumbent is erased upon the entry of the entrant in the sense
-that its posterior is reset to the initial prior. The results of the
-experiments are presented in Table~\ref{barrier_exp}, which lead to
-our Finding~\ref{barrier-find}.
+We run two additional experiments to isolate the effects of the two
+advantages mentioned above.
+The \emph{data-advantage experiment} focuses on the data advantage by, essentially, erasing the reputation advantage.
+Namely, the data from the monopoly period is not used in the computation of the incumbent's reputation score.
+Likewise, the \emph{reputation-advantage experiment} erases the data advantage and focuses on the reputation advantage: namely, the incumbent's algorithm `forgets' the data gathered during the monopoly period.
 
@@
 
+We find that either data or reputational advantage alone gives a substantial boost to the incumbent, compared to permanent duopoly.
+The results for the Heavy-Tail instance are presented in Table~\ref{barrier_exp}, in the same structure as Table~\ref{tab:ht-incum}.
+For the other two instances, the results are qualitatively similar.
+
+We can quantitatively define the data (resp., reputation) advantage as the incumbent's market share in the competition period in the data-advantage (resp., reputation advantage) experiment, minus the said share under permanent duopoly, for the same pair of algorithms and the same problem instance.
+In this language, our findings are as follows.
+
 \begin{finding}\label{barrier-find}
-\swedit{During the rounds in which the entrant is present, the incumbent
-  with either data or reputational advantage alone gets a larger share
-  than the share \gaedit{in the permanent duopoly, regardless of the algorithm she plays}.
-When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two
-  barriers.}
+Data advantage and reputation advantage alone are substantially large, across all algorithms and all MAB instances.
+The data advantage is larger than the reputation advantage when the incumbent chooses \TS; the two advantages are similar when the incumbent chooses \DEG or \DG.
+\end{finding}
+
+\ascomment{stopped here.}
+
+\OMIT{ %%%%%
+\begin{finding}\label{barrier-find}
+The incumbent with either data or reputational advantage alone gets a substantially larger share of agents in the competition period, compared to the said share under permanent duopoly.
+This holds across all algorithms and all MAB instances.
+When the incumbent chooses $\TS$, the data advantage serves as a larger barrier to entry compared to the reputation advantage, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.
 \OMIT{\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger market share compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}}
 \end{finding}
-
+} %%%%%
 % \swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}
 
-\gaedit{While the data and reputation advantages seem to serve as substitutes for \DEG and \DG, it is interesting that under \TS the data advantage is larger.
-The reason for this is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}.
+Our intuition for why the data advantage is larger  under \TS is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}.
 In contrast, purposeful exploration in competition may have immediate adverse reputational consequences.
@@
 It is interesting that, for \DEG and \DG, these two effects seem to serve nearly as perfectly substitutes.
-}
 

### 2018-09-05 paper/content/introduction.tex
new plot...also fix computation details
Added a citation for first mover advantage.
new figure + clean up some edits

@@
 \begin{tikzpicture}
-      \draw[->] (-.5,0) -- (6,0) node[right] {};
-      \draw[->] (0,-.5) -- (0,2) node[above] {Better algorithms};
-      \draw[scale=0.6,domain=0.7:8.3,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{3.4 - 0.2 * (\x - 4.5)^2});
-     \node[below] at (0.7, -0.15) {\footnotesize monopoly};
-     \node[below] at (3, 0) {\footnotesize temporary \\ monopoly};
-     \node[below] at (5, -0.15) {\footnotesize duopoly};
+      \draw[->] (-.5,0) -- (7,0) node[right] {};
+      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
+      \draw[scale=0.6,domain=0.7:9.8,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - 0.18 * (\x - 5.25)^2});
+     \node[below] at (0.7, -0.2) {\footnotesize monopoly};
+     \node[below] at (3, 0) {\footnotesize early \\ entry};
+     \node[below] at (4.5, -0.2) {\footnotesize duopoly};
+     \node[below] at (6, 0) {\footnotesize late \\entry};
  \end{tikzpicture}
@@
 
-\asedit{We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with an early entrant.
+We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with an early entrant.
 Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``early entry" (in the temporary monopoly) to permanent duopoly to ``late entry" (temporary monopoly from the incumbent's perspective).
-We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly, and even more so under ``late entry".} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
+We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly, and even more so under ``late entry".
+This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
 Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run.
@@
 Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
-
-\OMIT{This occurs because the lack of initial competition allows the incumbent to gather data on actions that may be harder to gather in competition (due to reputational consequences) but that are useful for learning and maximizing long-run reputation.}
-
-\OMIT{ %%%%%
-\gadelete{In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage".
-We further investigate which of the two is a stronger barrier to entry.
-We find that both are strong barriers on its own: removing either one substantially helps the late entrant.
-Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.}
-} %%%%%%
 

### 2018-09-05 paper/content/perf_in_iso.tex
new figure + clean up some edits

@@
 \DG dips below $\tfrac12$ can be seen as an explanation for the competitive disadvantage of exploration.
-\asedit{Such period also exists for the Heavy-tail MAB instance.} However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
+Such period also exists for the Heavy-tail MAB instance.
+However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
 \footnote{We see two explanations for this: $\TS$ identifies the best arm faster  for the Uniform instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
@@
 
-\asedit{Thus, Uniform and Heavy-tail instance are exploration-disadvantaged, but Needle-in-Haystack is not.}
-
-\OMIT{ %%%%%%%
-\begin{definition}
-When purposeful exploration leads to a relative reputation loss compared to the greedy alternative we define this as a \textit{relative reputation cost}.
-The eventual gain in reputation from the early acquired information leads to what we define as a \textit{relative reputation benefit}.
-Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
-We call the instances where there is an early costly period followed by a later benefit period \textit{relative reputation costly instances}.
-\end{definition}
-} %%%%%
-
-\OMIT{ %%%%%
-$\TS$ identifies the best arm faster for Needle-in-Haystack than for Uniform instance, so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration.
-Second, in the Needle In Haystack instances there are no ``bad" arms as there may be in the Uniform instances since by construction all the arms except one in Needle In Haystack are the same.
-Thus, when $\TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
-However, with the Uniform instances, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
-Thus, only the Uniform and Heavy Tail instances are relative reputation costly instances.
-} %%%%
-
-\OMIT{The following conjecture is natural:
-
-\begin{conjecture}\label{conj:mean-trajectories}
-For a given MAB instance, if one algorithm's mean reputation trajectory lies above another at time $T_0$, then the first algorithm prevails in the competition game.
-\end{conjecture}
-
-We revisit this conjecture after documenting the results in the competition game.
-}
-
-\OMIT{
-However, we find a more nuanced picture.
-
-\begin{finding}
-\textit{
-Conjecture~\ref{conj:mean-trajectories} is false. Mean reputation trajectories do not suffice to explain the outcomes under competition.}
-\end{finding}
-
-We will revisit this finding after presenting the results from the competition game.
-
-\OMIT{ %%%%%
-\begin{finding}
-\textit{The mean performance of an algorithm is not a sufficient statistic for understanding the performance of an algorithm in competition.
-Looking at the distribution of reputation difference between algorithms we find that it tend is skewed to the right.
-Additionally we provide a counterexample where the mean performance does not predict the result of the competition game.}
-\end{finding}
-} %%%%%
-
-\begin{figure}[H]
-\includegraphics[scale=0.35]{figures/rep_distribution_nih}
-%\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
-\caption{Distribution of reputation scores for Needle-in-Haystack at $t=500$ (smoothed using a kernel density estimate)}
-\label{rep_dist_nih}
-\end{figure}
-
-To see what could go wrong, consider how an algorithm's reputation score is distributed at a particular time.
-That is, consider the empirical distribution of this score over different \MRVs.
-For concreteness, we consider the needle-in-haystack instance at time $t=500$ (Figure \ref{rep_dist_nih}).
-We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
-The reason is that for this MAB instance, $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
-This leads to two cases: $\DG$ either does slightly better than $\TS$ or where $\TS$ does substantially better than $\DG$.
-
-However, the mean reputation trajectory of an algorithm would summarize all this complexity the average over the \MRVs.
-This may be inadequate for explaining the outcome of the duopoly game, given that the latter is determined by a simple comparison between the firm's reputation scores.
-Further, consider the difference in reputation scores (\emph{reputation difference}) between \TS and \DG on a particular \MRV.
-Let's plot the empirical distribution of the reputation difference (over the \MRVs) at a particular time point.
-Figure~\ref{ts_dg_rep_diff_nih} shows such plots for several time points.
-We see that the distribution is skewed to the right, for which the mean is not a representative statistic.
-\ascomment{so what?}
-
-\OMIT{ %%%%%%%%
- or not.
-If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds.
-However, if it does not then it will get stuck on a bad arm and lose to $\TS$ or $\DEG$.
-In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \cite{CompetingBandits-itcs16}}.
-} %%%%%%%%
-
-\OMIT{There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(\TS) - reputation(\DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the \textit{relative reputation proportion}.}
-
-\begin{figure}[H]
-\caption{Distribution of reputation difference $\TS-\DG$ for the Needle-in-Haystack (smoothed via a kernel density estimate)}
-\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
-\label{ts_dg_rep_diff_nih}
-%\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $\DG$ across $t$}}
-\end{figure}
-
-.
-
-\ascomment{Stopped here.}
-
-\begin{definition}
-\textit{Relative Reputation Proportion} - the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$
-\end{definition}
-
-\begin{finding}
-\textit{Purposeful exploration can lead to relative reputational costs compared to the greedy alternative and this leads to $\TS$ doing worse than $\DG$ for small time horizons.
-We observe this under the Uniform and Heavy Tail priors.
-However, it is not observed under the Needle In Haystack prior}
-\end{finding}
-
-The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $\DEG >\DG$ or $\TS > \DG$ according to the mean reputation but $\DG > \DEG$ or $\DG > TS$ according to the relative reputation proportion.
-In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $\DEG > \DG$ according to the mean reputation but $\DG > \DEG$ according to the relative reputation proportion.
-Additionally, these results carry over to the competition game}.
-
-\begin{figure}[ht]
-\caption{Relative Reputation Plots}
-\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
-\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
-\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.
-The shaded area display 95\% confidence intervals.}}
-\label{relative_rep_plots}
-
-\end{figure}
-
-Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $\TS$ vs $\DG$ on the Uniform and Needle In Haystack prior.
-For the Uniform prior we see that, in the early rounds, $\DG > \TS$ for the majority of the simulations but that, eventually, $\TS > \DG$.
-The intuition behind this is that, especially since the firms start with no substantive initial information, $\TS$ does purposeful exploration in the early rounds in order to acquire information.
-However, eventually the information acquired from purposeful exploration in the early rounds allows $\TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $\DG$ cannot trivially find the best arm.
-The early exploration leads to what we define as a \textit{relative reputation cost} and the eventual gain in reputation leads to what we define as a \textit{relative reputation benefit}.
-
-\begin{definition}
-\textit{Relative reputation cost and benefit} - the relative reputation loss an algorithm incurs from purposeful exploration compared to the greedy alternative.
-Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm.
-We call the priors where there is an early costly period followed by a later benefit period relative reputation costly priors.
-\end{definition}
-Is exploration always costly?
-Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $\TS$ always does relatively better than $\DG$.
-There are two contributing factors to this.
-First, $\TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $\TS$ needs to engage in purposeful exploration.
-Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same.
-Thus, when $\TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm.
-However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.
-Thus, only the Uniform and Heavy Tail priors are relative reputation costly priors.
-
-}
+Thus, Uniform and Heavy-tail instance are exploration-disadvantaged, but Needle-in-Haystack is not.
 

### 2018-09-06 paper/content/barriers.tex
Done with my pass, I think.

@@
 \begin{finding}\label{barrier-find}
-Data advantage and reputation advantage alone are substantially large, across all algorithms and all MAB instances.
-The data advantage is larger than the reputation advantage when the incumbent chooses \TS; the two advantages are similar when the incumbent chooses \DEG or \DG.
+(a) Data advantage and reputation advantage alone are substantially large, across all algorithms and all MAB instances.
+(b) The data advantage is larger than the reputation advantage when the incumbent chooses \TS.
+(c) The two advantages are similar in magnitude when the incumbent chooses \DEG or \DG.
 \end{finding}
-
-\ascomment{stopped here.}
-
 \OMIT{ %%%%%
@@
 
-Our intuition for why the data advantage is larger  under \TS is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}.
-In contrast, purposeful exploration in competition may have immediate adverse reputational consequences.
+Our intuition for Finding~\ref{barrier-find}(b) is as follows.
+Suppose the incumbent switches from \DG to \TS.
+This switch allows the incumbent to explore actions more efficiently, and therefore should benefit the data advantage.
+However, the switch increases the reputation cost of exploration in the short run, which could weaken the reputation advantage.
+
+\OMIT{ %%%%%
+\textit{that may be very costly to explore in competition}.
+In contrast, purposeful exploration in \TS may damage reputation in the short run.
+in competition may have immediate adverse reputational consequences.
 However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms.
 As a result, the data advantage is larger for \TS precisely because the advantage that data gives is not just about the quantity of data but also about its distribution across actions.
+} %%%%%%
 
+\OMIT{ %%%%%
 Why is it that data and reputation alone provide such large barriers to entry?
@@
 It is interesting that, for \DEG and \DG, these two effects seem to serve nearly as perfectly substitutes.
+} %%%%%%
 
+\OMIT{ %%%%%
 \gadelete{
@@
 }
+} %%%%%%%
 

### 2018-09-06 paper/content/perf_in_iso.tex
this should be NIH

@@
 However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
-\footnote{We see two explanations for this: $\TS$ identifies the best arm faster  for the Uniform instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
+\footnote{We see two explanations for this: $\TS$ identifies the best arm faster for the Needle In Haystack instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
 

### 2018-09-06 paper/content/revisited.tex
Done with my pass, except for the very last sentence.

--- 2018-09-04 paper/content/revisited.tex
@@
 
-A natural question to ask is why we had to introduce the relative reputation proportion instead of simply using the mean to understand the results.
-For instance, the following would have been a natural conjecture:
+We saw in Section~\ref{sec:competition} that mean reputation trajectories do not suffice to explain the outcomes under competition. Let us provide more evidence and intuition for this.
+
+Mean reputation trajectories are so natural that one is tempted to conjecture that they determine the outcomes under competition. More specifically:
 \begin{conjecture}\label{conj:mean-trajectories}
-Given a sufficiently large $T_0$, if one algorithm's mean reputation trajectory lies above another (for instance, Figure  \ref{prelim_means}), then the first algorithm prevails in the competition game.
+If one algorithm's mean reputation trajectory lies above another, perhaps after some initial time interval (\eg as in Figure~\ref{prelim_means}), then the first algorithm prevails under competition, for a sufficiently large warm start $T_0$.
 \end{conjecture}
@@
 However, we find a more nuanced picture.
+For example, in Figure \ref{sim_table} we see that $\DG$ attains a larger market share than $\DEG$ even for large warm starts.
+We find that this also holds for $K = 3$ arms and longer time horizons, see the supplement for more details.
+We conclude:
 
@@
 \textit{
-Conjecture~\ref{conj:mean-trajectories} is false.
-Mean reputation trajectories do not suffice to explain the outcomes under competition.
-We observe in Figure \ref{sim_ht} that $\DG$ attains a larger market share than $\DEG$ even for large warm starts.\footnote{In the appendix we show that this holds for $K = 3$ and longer time horizons than we consider here as well.}}
+Conjecture~\ref{conj:mean-trajectories} is false: mean reputation trajectories do not suffice to explain the outcomes under competition.}
 \end{finding}
+
+To see what could go wrong with Conjecture~\ref{conj:mean-trajectories}, consider how an algorithm's reputation score is distributed at a particular time.
+That is, consider the empirical distribution of this score over different \MRVs.%
+\footnote{Recall that each \MRV in our experimental setup comes with one specific realization table.} For concreteness, consider the needle-in-haystack instance at time $t=500$, plotted in Figure \ref{rep_dist_nih}.
+(The other MAB instances lead to a similar intuition.)
 
@@
 
-To see what could go wrong, consider how an algorithm's reputation score is distributed at a particular time.
-That is, consider the empirical distribution of this score over different \MRVs.\swcomment{not sure what this means.
-over the randomness of \MRVs only?
-you mean marginal distribution?} For concreteness, we consider the needle-in-haystack instance at time $t=500$ (Figure \ref{rep_dist_nih}) since the intuition as to why the mean may be misleading is clearest on this instance (though the same result holds across the instances).
+We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
+The reason is that for this MAB instance, $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
+In the former case \DG does slightly better than $\TS$, and in the latter case it does substantially worse.
+However, the mean reputation trajectory may fail to capture this complexity since it simply takes average over different \MRVs.
+This may be inadequate for explaining the outcome of the duopoly game, given that the latter is determined by a simple comparison between the firm's reputation scores.
 
-We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
-The reason is that \gaedit{for the Needle In Haystack instance}, $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
-This leads to two cases: $\DG$ either does slightly better than $\TS$ or where $\TS$ does substantially better than $\DG$.
-
-However, the mean reputation trajectory \swedit{may fail to capture} this complexity since it simply takes average over different realizations of \MRVs.
-This may be inadequate for explaining the outcome of the duopoly game, given that the latter is determined by a simple comparison between the firm's reputation scores.
-Further, consider the difference in reputation scores (\emph{reputation difference}) between \TS and \DG on a particular \MRV.
+To further this intuition, consider the difference in reputation scores (\emph{reputation difference}) between \TS and \DG on a particular \MRV.
 Let's plot the empirical distribution of the reputation difference (over the \MRVs) at a particular time point.
@@
 We observe that the distribution is skewed to the right, precisely due to the fact that $\DG$ either does slightly better than $\TS$ or does substantially worse.
-This means that the mean is not a good measure of the central tendency, or typical value, of the reputation difference distribution and overstates how well an advanced exploration algorithm (that will eventually find the best arm) should do in competition compared to a naive algorithm that may never find the best arm.
+Therefore, the mean is not a good measure of the central tendency, or typical value, of the reputation difference distribution.
+In particular, the mean reputation difference tends to overstate how well an advanced exploration algorithm (that will eventually find the best arm) should do in competition compared to a naive algorithm that may never find the best arm.
 

### 2018-09-06 paper/content/revisited.tex
Done with my pass.

@@
 
-\begin{figure}[ht]
-\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
-\caption{Distribution of reputation difference $\TS-\DG$ for  Needle-in-Haystack
-% at $t = 500, 1000, 2000$ AS: this is redundant, and it saves a line :).
-(smoothed via a kernel density estimate)}
-\label{ts_dg_rep_diff_nih}
-%\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $\DG$ across $t$}}
-\end{figure}
-
 We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
@@
 We observe that the distribution is skewed to the right, precisely due to the fact that $\DG$ either does slightly better than $\TS$ or does substantially worse.
-Therefore, the mean is not a good measure of the central tendency, or typical value, of the reputation difference distribution.
-In particular, the mean reputation difference tends to overstate how well an advanced exploration algorithm (that will eventually find the best arm) should do in competition compared to a naive algorithm that may never find the best arm.
+Therefore, the mean is not a good measure of the central tendency, or typical value, of this distribution.
+
+\OMIT{In particular, the mean reputation difference tends to overstate how well an advanced exploration algorithm (that will eventually find the best arm) should do in competition compared to a naive algorithm that may never find the best arm.}
+
+\begin{figure}[ht]
+\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
+\caption{Distribution of reputation difference $\TS-\DG$ for  Needle-in-Haystack
+% at $t = 500, 1000, 2000$ AS: this is redundant, and it saves a line :).
+(smoothed via a kernel density estimate)}
+\label{ts_dg_rep_diff_nih}
+%\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $\DG$ across $t$}}
+\end{figure}
 

### 2018-09-06 paper/content/introduction.tex
Minor edit.

@@
 Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run.
-The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model, aside from the tie-breaking).
+The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model).
 

### 2018-09-06 paper/content/inverted_u.tex
Changed terminology for the "data vs. reputation" table.
Changed section label.
Minor edit, and added a table for duopoly / uniform.

@@
 \begin{document}
-\section{Competition and Better MAB Algorithms}\label{section:5}
+\section{Competition and Better MAB Algorithms}\label{sec:competition}
 
@@
 We report the market shares for Heavy-Tail and Needle-in-Haystack instances in Table \ref{sim_table}  (see the first line in each cell), for a range of values of the warm start $T_0$.
-We find that $\DG$ is a weakly dominant strategy for the Heavy-tail instance, as long as $T_0$ is sufficiently small.
+Table~\ref{tab:duopoly_unif} reports similarly on the Uniform instance.
+We find that $\DG$ is a weakly dominant strategy for the Heavy-tail and Uniform instances, as long as $T_0$ is sufficiently small.
 However, \TS is a weakly dominant strategy for the Needle-in-Haystack instance.
-A similar table for the Uniform MAB instance is deferred to the appendix.
 We find that for a sufficiently small $T_0$, \DG yields more than half the market against \TS,  but achieves similar market share vs.
@@
 By our tie-breaking rule, (\DG,\DG) is the only pure-strategy equilibrium.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|c|c|c|c|}
+  \hline
+  & \multicolumn{3}{c|}{Uniform} \\
+\hline
+   & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 \\ \hline
+\TS vs \DG
+  & \makecell{\textbf{0.46} $\pm$0.03}
+    & \makecell{\textbf{0.52} $\pm$0.02}
+    & \makecell{\textbf{0.6} $\pm$0.02} \\ \hline
+\TS vs \DEG
+    & \makecell{\textbf{0.41} $\pm$0.03}
+    & \makecell{\textbf{0.51} $\pm$0.02}
+    & \makecell{\textbf{0.55} $\pm$0.02} \\ \hline
+\DG vs \DEG
+    & \makecell{\textbf{0.51} $\pm$0.03}
+    & \makecell{\textbf{0.48} $\pm$0.02}
+    & \makecell{\textbf{0.45} $\pm$0.02} \\\hline
+\end{tabular}
+\caption{{\bf Permanent duopoly}, for the Uniform MAB instance. Semantics are the same as in Table \ref{sim_table}.}
+\label{tab:duopoly_unif}
+\end{table}
 
@@
 \hline
-  & \multicolumn{3}{c||}{Information erased}
-  & \multicolumn{3}{c|}{Reputation erased}\\
+  & \multicolumn{3}{c||}{Reputation advantage}
+  & \multicolumn{3}{c|}{Data advantage}\\
 \hline
@@
 \end{tabular}
-\caption{Information-erased vs.
-reputation-erased experiment, on Heavy-Tail MAB instance.
+\caption{Data advantage vs.
+reputation advantage experiment, on Heavy-Tail MAB instance.
 Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}).
@@
 
-\asedit{We also see that the \Eeog values tend to increase as the warm start $T_0$ increases.
-We conjecture this is because larger $T_0$ tends to help a better algorithm.
-Indeed, we know that the ``effective end of game" in this scenario typically occurs when a better algorithm loses, and helping it delays the loss.}
+We also see that the \Eeog values tend to increase as the warm start $T_0$ increases.
+We conjecture this is because larger $T_0$ tends to be more beneficial for a better algorithm (as it tends to follow a better learning curve).
+Indeed, we know that the ``effective end of game" in this scenario typically occurs when a better algorithm loses, and helping it delays the loss.
 

### 2018-09-06 paper/content/introduction.tex
Removed an \asedit{}.

@@
 Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-Being first in the market gives firms free data to learn from (a ``data advantage") as well as \asedit{a more definite (and possibly better)} reputation compared to an entrant (a ``reputation advantage").
+Being first in the market gives firms free data to learn from (a ``data advantage") as well as a more definite (and possibly better) reputation compared to an entrant (a ``reputation advantage").
 We investigate which of the two is a stronger barrier to entry.

### 2018-09-06 paper/content/introduction.tex
Added a para on mean reputation.

@@
 
+\asedit{We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition.
+In particular, we find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor.
+We suggest a more refined performance measure, and use it to explain (and frame) some of the competition outcomes.}
+
 \subsubsection{Related work.}

### 2018-09-06 paper/content/barriers.tex
update appendix
Clarified in "barriers".

@@
 Suppose the incumbent switches from \DG to \TS.
-This switch allows the incumbent to explore actions more efficiently, and therefore should benefit the data advantage.
-However, the switch increases the reputation cost of exploration in the short run, which could weaken the reputation advantage.
+This switch allows the incumbent to explore actions more efficiently -- collect better data, not just more data -- and therefore should benefit the data advantage.
+However, the same switch increases the reputation cost of exploration in the short run, which could weaken the reputation advantage.
 

### 2018-09-06 paper/content/model.tex
new figure + clean up some edits
Minor edit to save two lines.

@@
 \subsubsection{MAB instances.}
-In all experiments, there are $K=10$ arms.
+We consider instances with $K=10$ arms.
 Since we focus on 0-1 rewards, an instance of the MAB problem is specified by the \emph{\MRV} $(\mu(a):\; a\in A)$.
-Initially the \MRV is drawn from some distribution, termed \emph{MAB instance}.
+Initially this vector is drawn from some distribution, termed \emph{MAB instance}.
 We consider three MAB instances:
@@
 The ``gap" is fixed in Needle-in-Haystack, spread over a wide spectrum of values under the Uniform instance, and is spread but  focused on the large values under the Heavy-Tail instance.
-\asedit{We also ran smaller experiments with versions of these instances, and achieved similar qualitative results.}
+We also ran smaller experiments with versions of these instances, and achieved similar qualitative results.
 

### 2018-09-06 paper/content/barriers.tex
Another rewording in the last para.

@@
 Suppose the incumbent switches from \DG to \TS.
-This switch allows the incumbent to explore actions more efficiently -- collect better data, not just more data -- and therefore should benefit the data advantage.
+This switch allows the incumbent to explore actions more efficiently -- collect better data in the same number of rounds -- and therefore should benefit the data advantage.
 However, the same switch increases the reputation cost of exploration in the short run, which could weaken the reputation advantage.
+One take-away from this finding is that the data advantage is about not just quantity but quality of data.
 

### 2018-09-06 paper/content/introduction.tex
Added a sentence on barriers / quality vs quantity.

@@
 Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-Being first in the market gives firms free data to learn from (a ``data advantage") as well as a more definite (and possibly better) reputation compared to an entrant (a ``reputation advantage").
+Being first in the market gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
 We investigate which of the two is a stronger barrier to entry.
 We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed.
-Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
+However, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
+\asedit{One take-away is that the data advantage is about not just quantity but quality of data.}
 

### 2018-09-06 paper/content/barriers.tex
Moved the "barriers" table back to the "barriers" section.

@@
 For the other two instances, the results are qualitatively similar.
+
+\begin{table*}[t]
+\centering
+\begin{tabular}{|c|c|c|c||c|c|c|}
+\hline
+  & \multicolumn{3}{c||}{Reputation advantage}
+  & \multicolumn{3}{c|}{Data advantage}\\
+\hline
+& $\TS$  & $\DEG$  & $\DG$
+& $\TS$  & $\DEG$  & $\DG$
+\\\hline
+$\TS$
+    & \makecell{\textbf{0.021}$\pm$0.009}
+    & \makecell{\textbf{0.16}$\pm$0.02}
+    & \makecell{\textbf{0.21} $\pm$0.02}
+    %%
+    & \makecell{\textbf{0.0096}$\pm$0.006}
+    & \makecell{\textbf{0.11}$\pm$0.02}
+    & \makecell{\textbf{0.18}$\pm$0.02}
+    \\ \hline
+$\DEG$
+    & \makecell{\textbf{0.26}$\pm$0.03}
+    & \makecell{\textbf{0.3}$\pm$0.02}
+    & \makecell{\textbf{0.26}$\pm$0.02}
+    %%
+    & \makecell{\textbf{0.073}$\pm$0.01}
+    & \makecell{\textbf{0.29}$\pm$0.02}
+    & \makecell{\textbf{0.25}$\pm$0.02}
+    \\ \hline
+$\DG$
+    & \makecell{\textbf{0.34}$\pm$0.03}
+    & \makecell{\textbf{0.4}$\pm$0.03}
+    & \makecell{\textbf{0.33}$\pm$0.02}
+    %%
+    & \makecell{\textbf{0.15}$\pm$0.02}
+    & \makecell{\textbf{0.39}$\pm$0.03}
+    & \makecell{\textbf{0.33}$\pm$0.02}
+    \\\hline
+\end{tabular}
+\caption{Data advantage vs.
+reputation advantage experiment, on Heavy-Tail MAB instance.
+Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}).
+The cell specifies the entrant's market share for the rounds in which hit was present: the average (in bold) and the 95\% confidence interval.
+NB: smaller average is better for the incumbent.}
+\label{barrier_exp}
+\end{table*}
 

### 2018-09-06 paper/content/inverted_u.tex
Removed a redundant line in a table.
Moved the "barriers" table back to the "barriers" section.

@@
 \begin{tabular}{|c|c|c|c|}
-  \hline
-  & \multicolumn{3}{c|}{Uniform} \\
+ % \hline
+ % & \multicolumn{3}{c|}{Uniform} \\
 \hline
@@
 In Table~\ref{tab:ht-incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$ and $\DG$ is only weakly dominant.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the instances we consider.}
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|c|c|c|c||c|c|c|}
-\hline
-  & \multicolumn{3}{c||}{Reputation advantage}
-  & \multicolumn{3}{c|}{Data advantage}\\
-\hline
-& $\TS$  & $\DEG$  & $\DG$
-& $\TS$  & $\DEG$  & $\DG$
-\\\hline
-$\TS$
-    & \makecell{\textbf{0.021}$\pm$0.009}
-    & \makecell{\textbf{0.16}$\pm$0.02}
-    & \makecell{\textbf{0.21} $\pm$0.02}
-    %%
-    & \makecell{\textbf{0.0096}$\pm$0.006}
-    & \makecell{\textbf{0.11}$\pm$0.02}
-    & \makecell{\textbf{0.18}$\pm$0.02}
-    \\ \hline
-$\DEG$
-    & \makecell{\textbf{0.26}$\pm$0.03}
-    & \makecell{\textbf{0.3}$\pm$0.02}
-    & \makecell{\textbf{0.26}$\pm$0.02}
-    %%
-    & \makecell{\textbf{0.073}$\pm$0.01}
-    & \makecell{\textbf{0.29}$\pm$0.02}
-    & \makecell{\textbf{0.25}$\pm$0.02}
-    \\ \hline
-$\DG$
-    & \makecell{\textbf{0.34}$\pm$0.03}
-    & \makecell{\textbf{0.4}$\pm$0.03}
-    & \makecell{\textbf{0.33}$\pm$0.02}
-    %%
-    & \makecell{\textbf{0.15}$\pm$0.02}
-    & \makecell{\textbf{0.39}$\pm$0.03}
-    & \makecell{\textbf{0.33}$\pm$0.02}
-    \\\hline
-\end{tabular}
-\caption{Data advantage vs.
-reputation advantage experiment, on Heavy-Tail MAB instance.
-Each cell describes the duopoly game between the entrant's algorithm (the {\bf row}) and the incumbent's algorithm (the {\bf column}).
-The cell specifies the entrant's market share for the rounds in which hit was present: the average (in bold) and the 95\% confidence interval.
-NB: smaller average is better for the incumbent.}
-\label{barrier_exp}
-\end{table*}
 

### 2018-09-06 paper/content/barriers.tex
edits

@@
 \begin{finding}\label{barrier-find}
+\textit{
 (a) Data advantage and reputation advantage alone are substantially large, across all algorithms and all MAB instances.
@@
 (c) The two advantages are similar in magnitude when the incumbent chooses \DEG or \DG.
+}
 \end{finding}
@@
 However, the same switch increases the reputation cost of exploration in the short run, which could weaken the reputation advantage.
-One take-away from this finding is that the data advantage is about not just quantity but quality of data.
+One take-away is that the data advantage is not just about data quantity but also data quality.
 

### 2018-09-06 paper/content/introduction.tex
also edits to intro

@@
 However, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
-\asedit{One take-away is that the data advantage is about not just quantity but quality of data.}
+One take-away is that the data advantage is not just about data quantity but also data quality.
 
-\asedit{We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition.
+We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition.
 In particular, we find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor.
-We suggest a more refined performance measure, and use it to explain (and frame) some of the competition outcomes.}
+We suggest a more refined performance measure, and use it to explain (and frame) some of the competition outcomes.
 

### 2018-10-15 paper/competing_bandits.tex
A minor edit.
A minor rewording in "barriers" sentence.
aaai -> www

--- 2018-08-30 paper/competing_bandits.tex
@@
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{url}
+\documentclass[sigconf, anonymous, review,natbib=false]{acmart}
+
+\usepackage{tikz} % SW: Added for plots
 \usepackage{subfiles}
-\usepackage{graphicx}
-\usepackage{tikz} % SW: Added for plots
-\frenchspacing
-
-%\usepackage[utf8]{inputenc} % allow utf-8 input -- cannot use with AAAI! A.S.
-%\usepackage{hyperref}       % hyperlinks
 \usepackage{url}            % simple URL typesetting
@@
 \theoremstyle{definition}
-\newtheorem{definition}{Definition}
+%\newtheorem{definition}{Definition}
 \newtheorem{finding}{Finding}
+%\newtheorem{conjecture}{Conjecture}
+\newtheorem{puzzle}{Puzzle}
 \captionsetup{font=footnotesize}
-
-% The \author macro works with any number of authors. There are two
-% commands used to separate the names and addresses of multiple
-% authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to
-% break the lines. Using \AND forces a line break at that point. So,
-% if LaTeX puts 3 of 4 authors names on the first line, and the last
-% on the second line, try using \AND instead of \And before the third
-% author name.
 
@@
 \newcommand{\DG}{\term{DG}}
-\newcommand{\Beta}{\term{Beta}}
+\newcommand{\Beta}{\term{Beta}} % for Beta distribution
+\newcommand{\Eeog}{\term{EoG}} % shorthand for "effective end of game"
 
 \newcommand{\eps}{\varepsilon}
+
+\newcommand{\MRV}{mean reward vector\xspace} % mean reward vector
+\newcommand{\MRVs}{mean reward vectors\xspace} % mean reward vector
 
@@
 % e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
+\setlength{\tabcolsep}{8pt} % Default value: 6pt
+\renewcommand{\arraystretch}{1.5} % Default value: 1
+
+%drt24 hacks
+% Letter paper
+\setlength{\paperheight}{11in}
+\setlength{\paperwidth}{8.5in}
+
+% Hack to try to make acmart work with biblatex: https://tex.stackexchange.com/questions/37076/is-it-possible-to-load-biblatex-with-a-class-that-has-already-loaded-natbib
+\let\citename\relax
+%\RequirePackage[abbreviate=true, dateabbrev=true, isbn=true, doi=true, urldate=comp, url=true, maxbibnames=9, backref=false, backend=biber, style=ACM-Reference-Format, language=american]{biblatex}
+
+\usepackage{booktabs} % For formal tables
+
+% Copyright
+%\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+
+% DOI
+\acmDOI{10.475/123_4}
+
+% ISBN
+\acmISBN{123-4567-24-567/08/06}
+
+%Conference
+\acmConference[WWW'19]{ACM WWW conference}{May 2019}{San Francisco, CA 2019}
+\acmYear{2019}
+\copyrightyear{2019}
+
+\acmPrice{15.00}
 
 \begin{document}
-
-%\title{Learning and Reputation under Monopoly and Duopoly: \\A Competing Bandits Approach}
-
-%\title{Competing Bandits: Exploration and Competition under Duopoly}
-
-\title{Competing Bandits: The Perils of Exploration under Competition}
-
-%\author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
-%Zhiwei Steven Wu\textsuperscript{4} \\
-%{\textsuperscript{1}Columbia University, Department of Economics}\\
-%{\textsuperscript{2}Columbia University, Department of Computer Science}\\
-%{\textsuperscript{3}Microsoft Research, New York, NY}\\
-%{\textsuperscript{4}University of Minnesota - Twin Cities, Department of Computer Science}
-%}
-\maketitle
+\title{Competing Bandits: \\ The Perils of Exploration under Competition}
 
@@
 We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms.
-Our findings provide new intuition on the ``first-mover advantage" in the digital economy.
+Our findings shed light on the ``first-mover advantage" in the digital economy.
 \end{abstract}
 
-\OMIT{ %%% Guy's original abstract
-\begin{abstract}
-We empirically study the interplay between \textit{exploration} and \textit{competition} - how platforms that learn from interactions with users tradeoff between making potentially suboptimal choices in order to acquire new information and the reputational consequences of this exploration which potentially has adverse competitive effects.
-Our model considers competition for myopic users between two firms deploying multi-armed bandit algorithms that face the same underlying multi-armed bandit instance.
-The users select between the firms according to a reputation score, which is a function of the rewards past users have experienced from this firm.
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below.
+%
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10010147.10010257.10010282.10010284</concept_id>
+<concept_desc>Computing methodologies~Online learning settings</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10010147.10010257.10010258.10010261.10010272</concept_id>
+<concept_desc>Computing methodologies~Sequential decision making</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10010405.10010455.10010460</concept_id>
+<concept_desc>Applied computing~Economics</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
 
-We ask whether better algorithms are incentivized under varying degrees of competition.
-The environments we consider are a monopoly, a duopoly with one firm serving as an early entrant (a temporary monopoly), and a duopoly.
-We show that, under certain conditions, monopoly and duopoly do not incentivize the adoption of better learning algorithms due to the reputational costs of exploration, but that a temporary monopoly may incentivize the adoption of better learning algorithms.
-Additionally, we interpret our results as providing an alternative intuition behind the classic first-mover advantage which gives the incumbent firm both a data advantage and a reputational advantange.
-Finally, we ask whether in this setting the data advantage or the reputational advantage of the early entrant serves as stronger barriers to entry.
-\end{abstract}
-} %%%%%%%
+\ccsdesc[500]{Computing methodologies~Online learning settings}
+\ccsdesc[300]{Computing methodologies~Sequential decision making}
+\ccsdesc[500]{Applied computing~Economics}
+
+\keywords{Bandits, Competition}
+
+\maketitle
 
@@
 \subfile{content/model}
-
-\subfile{content/sim_details}
 
@@
 
+\subfile{content/revisited}
+
+\bibliographystyle{acm}
 \bibliography{refs,bib-abbrv-short,bib-bandits,bib-AGT,bib-slivkins}
-\bibliographystyle{aaai}
+
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2018-10-15 paper/competing_bandits_with_appendix.tex
aaai -> www

--- 2018-09-04 paper/competing_bandits.tex
@@
 
+\usepackage[title]{appendix}
 %\usepackage[utf8]{inputenc} % allow utf-8 input -- cannot use with AAAI! A.S.
@@
 
-%\author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
-%Zhiwei Steven Wu\textsuperscript{4} \\
-%{\textsuperscript{1}Columbia University, Department of Economics}\\
-%{\textsuperscript{2}Columbia University, Department of Computer Science}\\
-%{\textsuperscript{3}Microsoft Research, New York, NY}\\
-%{\textsuperscript{4}University of Minnesota - Twin Cities, Department of Computer Science}
-%}
+\author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
+Zhiwei Steven Wu\textsuperscript{4} \\
+{\textsuperscript{1}Columbia University, Department of Economics}\\
+{\textsuperscript{2}Columbia University, Department of Computer Science}\\
+{\textsuperscript{3}Microsoft Research, New York, NY}\\
+{\textsuperscript{4}University of Minnesota - Twin Cities, Department of Computer Science}
+}
 \maketitle
@@
 We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms.
-Our findings provide new intuition on the ``first-mover advantage" in the digital economy.
+Our findings shed light on the ``first-mover advantage" in the digital economy.
 \end{abstract}
@@
 \bibliographystyle{aaai}
+
+\newpage
+\subfile{content/appendix_for_one_version}
 \end{document}

### 2018-10-15 paper/content/inverted_u.tex
edits
Minor edit.
aaai -> www

@@
 \end{table}
+\OMIT{
+\begin{table}[h]
+\centering
+\caption{Heavy Tail}
+\begin{tabular}{|c|c|c|c|}
+\hline
+   & $\TS$  & $\DEG$  & $\DG$ \\ \hline
+$\TS$
+    & \makecell{\textbf{0.50, 0.50}}
+    & \makecell{\textbf{0.3, 0.7}}
+    & \makecell{\textbf{0.29, 0.71}} \\\hline
+$\DEG$
+    & \makecell{\textbf{0.7, 0.3}}
+    & \makecell{\textbf{0.50, 0.50}}
+    & \makecell{\textbf{0.38, 0.62}} \\\hline
+$\DG$
+    & \makecell{\textbf{0.71, 0.29}}
+    & \makecell{\textbf{0.62, 0.38}}
+    & \makecell{\textbf{0.50, 0.50}} \\\hline
+\end{tabular}
+\end{table}
+
+\begin{table}[h]
+\centering
+\caption{Needle In Haystack}
+\begin{tabular}{|c|c|c|c|}
+\hline
+   & $\TS$  & $\DEG$  & $\DG$ \\ \hline
+$\TS$
+    & \makecell{\textbf{0.50, 0.50}}
+    & \makecell{\textbf{0.57, 0.43}}
+    & \makecell{\textbf{0.64, 0.36}} \\\hline
+$\DEG$
+    & \makecell{\textbf{0.43, 0.57}}
+    & \makecell{\textbf{0.50, 0.50}}
+    & \makecell{\textbf{0.54, 0.46}} \\\hline
+$\DG$
+    & \makecell{\textbf{0.36, 0.64}}
+    & \makecell{\textbf{0.43, 0.57}}
+    & \makecell{\textbf{0.50, 0.50}} \\\hline
+\end{tabular}
+\end{table}
+}
 
@@
 \begin{finding}\label{find:temp-monopoly-entrant}
-\textit{Under temporary monopoly with a sufficiently large $X$, \DG is a weakly dominant strategy for the incumbent for exploration-disadvantaged instances.
+\textit{Under temporary monopoly, \DG is a weakly dominant strategy for the entrant for exploration-disadvantaged MAB instances.
 }
 \end{finding}
-
-\OMIT{
-In Table~\ref{tab:ht-incum} we can see that, for $X = 200$, $\TS$ is a dominant strategy for the incumbent and that $\DG$ is the dominant strategy for the entrant on the heavy tail instances.\footnote{The result for the Uniform instances is the same, though it takes a higher $X$ to induce the incumbent to play $\TS$ and $\DG$ is only weakly dominant.} In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $\TS$ is the dominant strategy for the incumbent is robust across the instances we consider.}
 
@@
 
-\OMIT{ %%%%%%%%%
-Why do we observe that $\TS$ is the dominant strategy for the incumbent whereas in the permanent duopoly experiment we saw that, under the same parameters, $\DG$ was the incentivized algorithm?
-The intuition for this is that competition in the duopoly forces the firms to worry about their reputation which dissuades them from committing to algorithms that involve pure exploration in the early rounds.
-One can view allowing one firm to temporarily be a monopolist as temporarily relaxing the ``incentive" component of exploration, exploitation, and incentives so that the incumbent firm only faces the classic tradeoff between exploration and exploitation.
-The incumbent only needs to worry about her reputation after $X$ periods when the entrant comes into the market and again needs to worry about incentivizing agents to select them over their competition.
-
-As a result, the incumbent is incentivized to commit to an algorithm that does exploration in the early rounds since she no longer suffers the same immediate relative reputational cost that she would suffer under competition as long as the $X$ is sufficiently large that she can begin to recover the reputational costs of exploration.
-Thus, counterintuitively, by having one firm be a monopoly and dominate the market, we can incentivize them to play $\TS$.
-} %%%%%%%%
-
 \subsubsection{Death spiral effect.}
@@
 
-\OMIT{ %%%
-Enter into a ``death spiral" where the winner keeps learning and improving from its customers thus increasing its reputation while the losing firm gets starved of customers and progressively falls farther and farther behind its competitors.
-What happens after the first agent?
-Looking at the EEOG values we note that, even though the game continues until $t=2000$, the game effectively ends much earlier than that.
-This seems to imply that when attempting to understand the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotic performance.
-This is due to the fact that the losing firm may enter into a ``death spiral" where the winner keeps learning and improving from its customers thus increasing its reputation while the losing firm gets starved of customers and progressively falls farther and farther behind its competitors.
-As a result, even though if $\TS$ could get sufficiently many agents it would eventually have a higher reputation than $\DG$, since it starts the game by losing agents to $\DG$ due to its lowered reputation from exploration in the warm start it ends up doing worse in the competition game.
-} %%%%%
-
-\OMIT{ %%%%%%
-\begin{table}[ht]
-\centering
-\caption{Duopoly experiment for needle in haystack instances}
-\begin{tabular}{|c|c|c|c|}
-  \hline
- & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 \\
-  \hline
-$\TS$ vs $\DG$
-& \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 200 (27)}
-& \makecell{\textbf{0.6} $\pm$0.03\\ \Eeog 370 (0}
-& \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 580 (122)} \\
-\hline
-  $\TS$ vs $\DEG$
-  & \makecell{\textbf{0.57} $\pm$0.03\\ \Eeog 150 (14}
-  & \makecell{\textbf{0.52} $\pm$0.03\\ \Eeog 460 (79)}
-  & \makecell{\textbf{0.56} $\pm$0.02\\ \Eeog 740 (628)} \\
-  \hline
-  $\DG$ vs $\DEG$
-  & \makecell{\textbf{0.46} $\pm$0.03\\ \Eeog 340 (129)}
-  & \makecell{\textbf{0.42} $\pm$0.02\\ \Eeog 650 (408)}
-  & \makecell{\textbf{0.42} $\pm$0.02\\ \Eeog 690 (467)} \\
-   \hline
-\end{tabular}
-\label{sim_nih}
-\caption*{{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95\% confidence band.
-For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
-The next line contain the average and median effective end of game for this set of simulations.}}
-\end{table}
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Thu Aug 16 13:13:00 2018
-} %%%%%%%
-
-\OMIT{ %%%%%%%
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Wed Aug 15 19:03:20 2018
-\begin{table}[ht]
-\centering
-\caption{Temporary Monopoly Heavy Tail X = 200}
-\begin{tabular}{cc|c|c|c|}
-  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
-  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
-            & $\TS$ & \makecell{\textbf{0.003} \\$\pm$0.003} & \makecell{\textbf{0.083} \\ $\pm$0.02} & \makecell{\textbf{0.17} \\$\pm$0.02} \\ \cline{3-5}
-Entrant  & $\DEG$ & \makecell{\textbf{0.045} \\$\pm$0.01} & \makecell{\textbf{0.25} \\$\pm$0.02} & \makecell{\textbf{0.23} \\$\pm$0.02} \\\cline{3-5}
-            & $\DG$ & \makecell{\textbf{0.12} \\ $\pm$0.02} & \makecell{\textbf{0.36} \\$\pm$0.03} & \makecell{\textbf{0.3} \\$\pm$0.02} \\\cline{3-5}
-\end{tabular}
-\label{ht_incum}
-\caption*{{The table contains the average market share for the entrant as well as 95\% confidence intervals.
-Note that smaller values in the table are better for the incumbent.
-Market shares are calculated as the fraction of users selecting a particular firm \textit{after} the entrant has already entered (i.e.
-the free rounds to the incumbent do not count towards the share)}}
-\end{table}
-} %%%%%%%
-
 \end{document}

### 2018-10-18 paper/content/barriers.tex
somewhat major revision

@@
 
-\section{Data and Reputation as Barriers to Entry}\label{section:6}
+\section{Data as a Barrier to Entry}\label{section:6}
+
+\gaedit{Even though welfare is highest in the temporary monopoly case there are other reasons one may worry about a single firm having a substantial market share.
+In this section we explore what drives the substantial market share of the temporary monopolist and discuss how this relates to thinking about data as a barrier to entry.}
 
@@
 Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant.
-While both advantages create a strong barrier to entry, we investigate whether either one is a strong barrier alone, and which one is stronger.
-Our findings provide a quantitative insight into the classic ``first mover advantage" phenomenon in the digital economy.
+\gaedit{First, our findings provide a quantitative insight into the role of the classic ``first mover advantage" phenomenon in the digital economy.
+Second, our findings provide a viewpoint in terms of thinking about the role that data can play as a competitive advantage.}
 
@@
 Likewise, the \emph{reputation-advantage experiment} erases the data advantage and focuses on the reputation advantage: namely, the incumbent's algorithm `forgets' the data gathered during the monopoly period.
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{cc|c|c|c|}
-%  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
-%  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
-%            & $\TS$ & \makecell{\textbf{0.021} \\$\pm$0.009}& 		\makecell{\textbf{0.16} \\$\pm$0.02} &
-%            \makecell{\textbf{0.21} \\ $\pm$0.02} \\ \cline{3-5}
-%Entrant  & $\DEG$ & \makecell{\textbf{0.26} \\$\pm$0.03} & \makecell{\textbf{0.3} \\$\pm$0.02} &
-%\makecell{\textbf{0.26} \\ $\pm$0.02} \\ \cline{3-5}
-%            & $\DG$ & \makecell{\textbf{0.34} \\ $\pm$0.03} & \makecell{\textbf{0.4} \\$\pm$0.03} &
-%            \makecell{\textbf{0.33} \\$\pm$0.02} \\\cline{3-5}
-%\end{tabular}
-%\caption{Information Erased Experiment Heavy Tail}
-%\label{info_erase}
-%\end{table}
-%
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{cc|c|c|c|}
-%  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
-%  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
-%            & $\TS$ & \makecell{\textbf{0.0096} \\$\pm$0.006}& 		\makecell{\textbf{0.11} \\$\pm$0.02} &
-%            \makecell{\textbf{0.18} \\$\pm$0.02} \\ \cline{3-5}
-%Entrant  & $\DEG$ & \makecell{\textbf{0.073} \\ $\pm$0.01} & \makecell{\textbf{0.29} \\ $\pm$0.02} &
-%\makecell{\textbf{0.25} \\$\pm$0.02} \\ \cline{3-5}
-%            & $\DG$ &\makecell{\textbf{0.15} \\$\pm$0.02} & \makecell{\textbf{0.39} \\$\pm$0.03} &
-%           \makecell{\textbf{0.33} \\ $\pm$0.02} \\\cline{3-5}
-%\end{tabular}
-%\caption{Reputation Erased Experiment Heavy Tail}
-%\label{rep_erase}
-%\end{table}
 
@@
 \begin{finding}\label{barrier-find}
-\textit{
+\textit{\\
 (a) Data advantage and reputation advantage alone are substantially large, across all algorithms and all MAB instances.
-(b) The data advantage is larger than the reputation advantage when the incumbent chooses \TS.
-(c) The two advantages are similar in magnitude when the incumbent chooses \DEG or \DG.
+\\(b) The data advantage is larger than the reputation advantage when the incumbent chooses \TS.
+\\(c) The two advantages are similar in magnitude when the incumbent chooses \DEG or \DG.
 }
 \end{finding}
-\OMIT{ %%%%%
-\begin{finding}\label{barrier-find}
-The incumbent with either data or reputational advantage alone gets a substantially larger share of agents in the competition period, compared to the said share under permanent duopoly.
-This holds across all algorithms and all MAB instances.
-When the incumbent chooses $\TS$, the data advantage serves as a larger barrier to entry compared to the reputation advantage, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.
-\OMIT{\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger market share compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}}
-\end{finding}
-} %%%%%
-% \swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}
 
@@
 However, the same switch increases the reputation cost of exploration in the short run, which could weaken the reputation advantage.
-One take-away is that the data advantage is not just about data quantity but also data quality.
 
-\OMIT{ %%%%%
-\textit{that may be very costly to explore in competition}.
-In contrast, purposeful exploration in \TS may damage reputation in the short run.
-in competition may have immediate adverse reputational consequences.
-However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms.
-As a result, the data advantage is larger for \TS precisely because the advantage that data gives is not just about the quantity of data but also about its distribution across actions.
-} %%%%%%
-
-\OMIT{ %%%%%
-Why is it that data and reputation alone provide such large barriers to entry?
-Having more data allows the incumbent to select better actions which allows her to quickly get a high reputation compared to an entrant that both needs to learn and work on building up a reputation.
-Starting with a large reputation score gives the incumbent a cushion for doing exploration and learning without facing as stark reputational consequences.
-It is interesting that, for \DEG and \DG, these two effects seem to serve nearly as perfectly substitutes.
-} %%%%%%
-
-\OMIT{ %%%%%
-\gadelete{
-The reason \gadelete{for the finding} \gaedit{that we see that the data advantage is greater
-under $\TS$} is that the early entry allows the incumbent to do
-exploration without immediate reputational consequence in the early
-rounds and so, even when reputation is erased, the incumbent has
-already acquired sufficient information that simply exploiting that
-information allows her to quickly recover her
-reputation. \swcomment{which part are you explaining?} However, while
-the reputational advantage gives the incumbent some cushion for doing
-exploration the incumbent is still vulnerable to getting unlucky while
-doing exploration and suffering the reputational consequences. Under all of the incumbent algorithms,
-the incumbent obtains a data advantage by the mere fact that it is around in the market and gets data from consumers.
-However, we see that the data advantage is stronger under \TS since \TS acquires data that is hard to acquire under competition.
-We only observe this under $\TS$ since
-\TS not only acquiresThis
-only happens under $\TS$ since $\TS$, by construction, is more
-intelligent about the data that it acquires compared to $\DEG$ and
-$\DG$.\swcomment{more intelligent about the data means?}
-}
-} %%%%%%%
-
-\OMIT{Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained.
-We find that the best responses remain the same for the Heavy Tail instance, but for the Uniform instance they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.}
+\gaedit{How does this help us in thinking about data as a barrier to entry?
+The result that the data advantage is stronger under \TS than \DEG and \DG gives evidence that data quality, and not just quantity, matters in thinking about the competitive advantage of data.
+In particular, in the incumbency period \TS gathers data that may be hard for the entrant to gather during competition due to its potential reputational consequences (i.e.
+engaging in exploration), but which are needed in order to learn.
+If the entrant had some way of acquiring this data without incurring reputational consequences then the data advantage would be substantially weaker.
+This insight is what differentiates our results from the static viewpoint of \cite{varian2018artificial, bajari2018impact}.
+While it remains true that, for a given action, increased data on that action has diminishing returns to scale, competitive pressures may make it harder for smaller firms to be able to engage in the exploration necessary to accumulate sufficient data on actions to learn if they are good or bad.
+Otherwise, they may be more likely to fall into the death-spiral effect discussed earlier.
+This effect only shows itself when thinking about the issue from the perspective of competition dynamics and endogenous data acquisition.}
 

### 2018-10-18 paper/content/conclusion.tex
somewhat major revision

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Conclusion}\label{sec:conclusion}

\gaedit{In this paper we considered a setting where firms simultaneously learn from and compete for users. We showed that competition may not always induce firms to commit to algorithms that engage in exploration and that this leads to welfare losses for consumers. The primary mechanism behind this was that exploration may have short-term reputational consequences that lead to naive algorithms winning in competition. Allowing one firm to have a head-start, or a first-mover advantage, leads to the first-mover having an incentive to play traditionally ``good" bandit algorithms and having significantly better welfare for consumers. Finally, we attempted to isolate the effect that the data advantage of an incumbent has on its ability to garner a considerable market share. Contrary to previous studies and discussions which do not consider competition dynamics or endogenizing data collection, we find that the data advantage is substantial in generating significant market power. Additionally we show that this effect is strongest when an incumbent performs exploration in the incumbency period and generates a high-quality dataset on actions that may be difficult for an entrant to generate due to the pressure of competition.}

\gacomment{Not sure what else to put in the conclusion?}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../competing_bandits"
%%% End:

### 2018-10-18 paper/content/introduction.tex
aaai -> www
somewhat major revision

@@
 This would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+\gaedit{This ``data feedback" loop has been discussed in policy circles and in the business literature, but, to our knowledge, our paper is the first to formally provide a model that allows us to understand this effect.}
 
-We ask whether competition incentivizes adoption of ``better" algorithms for exploration.
-We investigate this issue via extensive numerical experiments in a stylized duopoly model.
-In our model, two firms compete for users, and simultaneously learn from them.
+\gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model.
+In our model, two firms compete for users \gaedit{on a single dimension, quality}, and simultaneously learn from them.
 Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm.
 Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window.
-Each firm's objective is to maximize its market share (the fraction of users choosing this firm).
+Each firm's objective is to maximize its \gaedit{expected} market share (the fraction of users choosing this firm).
 
@@
       \draw[scale=0.6,domain=0.7:9.8,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - 0.18 * (\x - 5.25)^2});
-     \node[below] at (0.7, -0.2) {\footnotesize monopoly};
-     \node[below] at (3, 0) {\footnotesize early \\ entry};
-     \node[below] at (4.5, -0.2) {\footnotesize duopoly};
-     \node[below] at (6, 0) {\footnotesize late \\entry};
+     \node[below] at (0.7, -0.22) {\footnotesize monopoly};
+     \node[below] at (3, -0.2) {\footnotesize incumbent};
+     \node[below] at (4.5, -0.22) {\footnotesize duopoly};
+     \node[below] at (6, -0.2) {\footnotesize entrant};
  \end{tikzpicture}
@@
 We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with an early entrant.
-Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``early entry" (in the temporary monopoly) to permanent duopoly to ``late entry" (temporary monopoly from the incumbent's perspective).
-We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly, and even more so under ``late entry".
-This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
+Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (first-mover in duopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late arriver in duopoly)}.
+\gaedit{We find that, even in a duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late entrant in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
 Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run.
-The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model).
+\gaedit{We find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}
 
@@
 This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.
+\gaedit{The mechanism that leads to this behavior in our model is different from the existing literature in economics since the disincentives to ``innovate" arise entirely from the reputational consequences of exploration in competition and not from R\&D costs (which are absent from our model).}
 
-Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
+\gaedit{Our findings show that the data feedback loop arises either when a firm deploys a ``naive" greedy learning algorithm (under simultaneous entry) or has an initial temporary monopoly in the market.
+The findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
+Being first in the market in the digital economy gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
+
+Even though, under our model, consumers are best off when there is a ``temporary" monopoly, a recent subject of debate in policy circles is whether the data advantage alone poses a sufficiently powerful barrier to entry to warrant anti-trust interest.
+While \cite{varian2018artificial} and \cite{bajari2018impact} argue that, from a statistical perspective, data has decreasing returns to scale and therefore does not serve as a significant competitive advantage, they focus on the problem in isolation.
+We argue that, to reason about this question, one needs to consider what the impact of better data has in terms of competition dynamics and when data collection is endogenous and has consequences, especially when firms need to learn while competing.
+Therefore, we run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
+We find that the data advantage alone still leads to a significant incumbency advantage.
+Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm.
+This highlights that, when thinking about data as a barrier to entry, data quality and not just quantity matters.
+In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}
+
+\OMIT{Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
 Being first in the market gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
@@
 However, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
-One take-away is that the data advantage is not just about data quantity but also data quality.
+One take-away is that the data advantage is not just about data quantity but also data quality.}
 
@@
 Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
+\gaedit{There is a superficial similarity (in name only) between this paper and the line of work on ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}).
+The latter is not about competing
+bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".
+
+The study of whether data can give a significant competitive advantage has been discussed theoretically in \cite{varian2018artificial, lambrecht2015can} and empirically in \cite{bajari2018impact}.
+These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role, whereas we ask how additional data helps while firms simultaneously solve a machine learning problem and competing for a market of consumers.
+In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm.
+Additionally, the first-mover advantage has been well-studied in economics and marketing \cite{kerin1992first}.
+}
 

### 2018-10-18 paper/content/inverted_u.tex
somewhat major revision

@@
 Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from permanent monopoly to ``early entry" to permanent duopoly to ``late entry", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
+As the ``intensity of competition" varies from permanent monopoly to ``incumbent" to permanent duopoly to ``entrant", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
 More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
@@
 The lowest level of competition is monopoly, when \DG wins out for the trivial reason of tie-breaking.
-The highest levels are permanent duopoly and ``late entry" (temporary monopoly from the entrant's perspective).
+The highest levels are permanent duopoly and ``entrant" (temporary monopoly from the entrant's perspective).
 We see that \DG is incentivized for exploration-disadvantaged MAB instances.
-In fact, incentives for \DG get stronger when the model transitions from permanent duopoly to ``late entry".%
+In fact, incentives for \DG get stronger when the model transitions from permanent duopoly to ``entrant".%
 \footnote{For the Heavy-Tail instance, \DG goes from a weakly dominant startegy to a strictly dominant one.
 For the Uniform instance, \DG goes from a Nash equilibrium strategy to a weakly dominant one.}
-Finally, the middle level of competition, ``early entry" (temporary monopoly for the entrant) creates strong incentives for \TS.
+Finally, the middle level of competition, ``incumbent" (\gaedit{duopoly with an initial temporary monopoly}) creates strong incentives for \TS.
 In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U}.
@@
 
-Our intuition for why ``early entry" creates more incentives for exploration is as follows.
+Our intuition for why incumbency creates more incentives for exploration is as follows.
 During the temporary monopoly period, reputation costs of exploration vanish.
@@
 
+\gaedit{
+\subsubsection{Welfare}
+
+What do our results mean for the welfare of consumers?
+In our model we define consumer welfare in terms of pseudo-regret.
+Following the bandits literature we define pseudo-regret as being the difference between the true mean of the best action and the true mean of the action that is selected.
+We equate consumer welfare with \textit{market} pseudo-regret where market pseudo-regret is the sum of the pseudo-regret of the actions that are experienced in the competition game\footnote{Note that smaller pseudo-regret is better for consumers}.
+
+\begin{figure}
+\includegraphics[scale=0.3]{figures/unif_eq_welfare}
+\includegraphics[scale=0.3]{figures/ht_eq_welfare}
+\caption{Smoothed welfare plots resulting from equilibrium strategies in the different market structures.
+Note that welfare at $t = 0$ incorporates the regret incurred during the incumbent and warm start periods.}
+\label{eq_regret}
+\end{figure}
+
+\begin{finding}
+\textit{ \\ (a) Consumer welfare is highest under temporary monopoly. \\
+(b) Increasing the number of firms in the market playing \DG weakly decreases consumer welfare.
+}
+\end{finding}
+
+Figure \ref{eq_regret} displays the simulated welfare when the equilibrium strategies are played.
+Consumer welfare is \textit{better off} in the temporary monopoly case than in the duopoly case.
+However, this was expected from the fact that the incumbent is incentivized to play \TS.
+The welfare in the temporary monopoly case is close to the welfare to having a single firm playing \TS.
+
+What is surprising is that welfare under monopoly and duopoly are very similar, with monopoly being marginally better than duopoly.
+Recall that in equilibrium in these market structures all firms play \DG.
+One would think that welfare would be weakly increasing with the number of firms playing \DG since, from the perspective of the market, if adding a firm playing \DG results in that firm, by chance, identifying a better arm than the existing firms then that firm ought to attract more consumers and increase market welfare.
+If the firm identifies a worse arm than the rest then one ought to expect consumers to not select this firm and market welfare would be unchanged.
+Indeed, adding sufficiently many firms one would expect that eventually one firm by chance would discover the best arm quickly and the market welfare could even beat that of a stand-alone firm playing \TS.
+
+\begin{figure}
+\includegraphics[scale=0.3]{figures/eeog_vs_num_ht}
+\includegraphics[scale=0.3]{figures/ht_many_firm_welfare}\\
+\caption{Average welfare and EEOG as we increase the number of firms playing \DG}
+\label{many_firm_welfare}
+\end{figure}
+
+To test if it is possible that increasing competition past a duopoly leads to better welfare than under temporary monopoly we run simulations where we increase the number of firms playing \DG.
+Figure \ref{many_firm_welfare} reports the average welfare across these simulations.
+Welfare not only does not get better, \textit{but is even worse} as we increase the number of firms.
+Correspondingly, we track the average \Eeog in each of the simulations and notice that it is \textit{increasing} as the number of firms increases.
+As we increase the number of firms, a single individual firm gets fewer consumers for learning as consumers switch between firms more often.
+Additionally, bad luck with rewards or arm selection is punished even more harshly as we increase the number of firms.
+This further strengthens the intuition from our previous result, which is that learning may be hard under competition since accidental exploration and mistakes are punished harshly but are necessary for learning.}
+
 \end{document}

### 2018-10-18 paper/content/model.tex
somewhat major revision

@@
 \subsubsection{Agents.}
-Agents are myopic and non-strategic: they choose among the firms so as to maximize their expected reward, without attempting to influence the firms' learning algorithms or rewards of the future users.
+\gaedit{For simplicity we suppose that firms compete for agents on a single dimension, quality, that is summarized via the reward experienced by the consumer\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}}.
+Agents are myopic and non-strategic: they choose among the firms so as to maximize their expected reward \gaedit{(i.e.
+select the firm with the highest quality)}, without attempting to influence the firms' learning algorithms or rewards of the future users.
 Agents are not well-informed: they only receive a rough signal about each firm's performance before they choose a firm, and no other information.
@@
 We posit a version of rational behavior: each agent chooses a firm with a maximal reputation score (breaking ties uniformly).
-The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
+The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm\gaedit{\footnote{One can think of this as similar to ratings aggregations on many online platforms such as Yelp, Tripadvisor, etc.}}.
 
@@
 Our pilot experiments indicate that the results do not change substantially if other algorithms are chosen.
-For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge).
+\gaedit{We make this restriction since we are mainly interested in if there is a qualitative differences between what we know is good in isolation and what is incentivized in competition.} For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge).
 We consider:

### 2018-10-18 analysis_code/compare_across.R

@@ -1,6 +1,6 @@
 library(dplyr)
 
-dat_1 <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/longer_ws.csv")
+dat_1 <- read.csv("/Volumes/Mac/final_bandit_results/results/preliminary_raw_results/longer_ws.csv")
 dat_1$Algorithm<- sapply(dat_1$Algorithm, concise_alg_rep)
 #dat_2 <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv")
 
@@ -49,12 +49,12 @@ cat("Losing dat", nrow(losing_dat), "Market share", mean(losing_dat$Market.Share
 dist <- "Needle In Haystack"
 ws_time <- 20
 iso_dat <- filter(dat_1, Distribution == dist)
-iso_dat_t <- filter(iso_dat, t == 500)
+iso_dat_t <- filter(iso_dat, t == 1000)
 ggplot(iso_dat_t, aes(Realized.Reputation, colour=Algorithm)) +
   geom_density() +
-  ggtitle("Reputation Distribution, Needle In Haystack") +
-  theme_bw(base_size = 16) +
-  xlab("Reputation") + 
+  ggtitle("Quality Distribution, Needle In Haystack, t=1000") +
+  theme_bw(base_size = 12) +
+  xlab("Quality") + 
   theme(plot.title = element_text(hjust = 0.5))
 
 iso_deg <- filter(iso_dat, t %in% t_vals & Algorithm == "ThompsonSampling") 

### 2018-10-18 analysis_code/relative_reputation_plots.R

@@ -3,7 +3,7 @@ library(ggplot2)
 library(reshape2)
 library(pBrackets)
 
-#dat <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/longer_ws.csv")
+dat <- read.csv("/Volumes/Mac/final_bandit_results/results/preliminary_raw_results/longer_ws.csv")
 
 
 bracketsGrob <- function(...){
@@ -15,7 +15,7 @@ bracketsGrob <- function(...){
   }, e)
 }
 
-b1 <- bracketsGrob(0.34, 0.15, 0, 0.15, h=0.1, lwd=2, col="red")
+b1 <- bracketsGrob(0.18, 0.15, 0, 0.15, h=0.1, lwd=2, col="red")
 b2 <- bracketsGrob(1, 0.15, 0.35, 0.15, h=0.1,  lwd=2, col="red")
 concise_alg_rep <- function(alg) {
   if (alg == "ThompsonSampling") {
@@ -31,7 +31,7 @@ print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
   alg_1 <- filter(dist_dat, Algorithm == alg1)
   alg_2 <- filter(dist_dat, Algorithm == alg2)
   n_vals <- unique(dist_dat$n)
-  t_vals <- seq(10, 1500, 10)
+  t_vals <- seq(10, 1000, 10)
   df <- as.data.frame(matrix(nrow=length(t_vals), ncol=3))
   colnames(df) <- c("t", "relative_rep")
   df$t <- t_vals
@@ -49,17 +49,17 @@ print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
   }
   df$relative_rep <- counts
   df$se <- ses
-  plot_title <- paste("Relative Reputation -", dist)
+  plot_title <- paste("Relative Quality -", dist)
   q <- ggplot(df, aes(x=t, y=relative_rep)) + geom_line() + geom_point() +
     geom_errorbar(aes(ymin=relative_rep-1.96*se, ymax=relative_rep+1.96*se), width=.2) +
     ggtitle(plot_title) + xlab("time") + 
-    ylab(paste(concise_alg_rep(alg1), " >= ", concise_alg_rep(alg2), "(reputation)")) +
+    ylab(paste(concise_alg_rep(alg1), " >= ", concise_alg_rep(alg2), "(quality)")) +
     annotation_custom(b1) +
-    annotation_custom(b2) +
-    ylim(c(0.2, 0.7)) +
[…]

### 2018-10-18 analysis_code/welfare.R

@@ -1,0 +1,58 @@
+library(tidyverse)
+library(ggplot2)
+
+simultaneous_welfare <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/tournament_raw_results/tournament_experiment_full_welfare_calc_raw.csv")
+free_obs_welfare <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/free_obs_raw_results/free_obs_experiment_welfare_calc_raw.csv")
+#prelim <- read.csv("/Volumes/Mac/final_bandit_results/results/preliminary_raw_results/longer_ws.csv")
+prelim <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/regret_calc.csv")
+sim_eq_welfare <- filter(simultaneous_welfare, ((Prior == "Heavy Tail" | Prior == "Uniform") & (P1.Alg == "DynamicGreedy" & P2.Alg == "DynamicGreedy")) | (Prior == "Needle In Haystack" & P1.Alg == "ThompsonSampling" & P2.Alg == "ThompsonSampling"))
+free_obs_eq_welfare <- filter(free_obs_welfare, ((Prior == "Heavy Tail" | Prior == "Uniform") & (P1.Alg == "DynamicGreedy" & P2.Alg == "ThompsonSampling")))
+
+one_welfare <- combine(sim_eq_welfare, free_obs_eq_welfare)
+one_welfare <- one_welfare %>% mutate(label=ifelse(source == "free_obs_eq_welfare", "Temporary Monopoly", "Duopoly"))
+
+prelim <- prelim %>% mutate(Prior = Distribution, Warm.Start = 20, label = "Monopoly", Total.Regret = Cumulative.Regret)
+prelim <- prelim %>% mutate(Time.Horizon = t - 20)
+
+ht_greedy <- filter(prelim, Prior == "Heavy Tail" & Algorithm == "DynamicGreedy")
+unif_greedy <- filter(prelim, Prior == "Uniform" & Algorithm == "DynamicGreedy")
+ht_ts <- filter(prelim, Prior == "Heavy Tail" & Algorithm == "ThompsonSampling")
+ht_ts <- ht_ts %>% mutate(label = "Thompson Sampling")
+unif_ts <- filter(prelim, Prior == "Uniform" & Algorithm == "ThompsonSampling")
+unif_ts <- unif_ts %>% mutate(label = "Thompson Sampling")
+
[…]

### 2018-10-31 paper/competing_bandits.tex
somewhat major revision
updates to paper

@@
 \usepackage{subfiles}
+\usepackage[toc,page]{appendix}
 \usepackage{url}            % simple URL typesetting
@@
 \begin{abstract}
-We empirically study the interplay between \textit{exploration} and \textit{competition}. Systems that learn from interactions with users
-often engage in \emph{exploration}: making potentially suboptimal decisions in order to acquire new information for future decisions.
+We empirically study the interplay between \textit{exploration} and \textit{competition}.
+Systems that learn from interactions with users often engage in \emph{exploration}: making potentially suboptimal decisions in order to acquire new information for future decisions.
 However, exploration may hurt system's reputation in the near term, with adverse competitive effects.
@@
 We run extensive numerical experiments in a stylized duopoly model in which two firms deploy multi-armed bandit algorithms and compete for myopic users.
-We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms.
-Our findings shed light on the ``first-mover advantage" in the digital economy.
+We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore \gaedit{and leads to low consumer welfare}, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms \gaedit{and lead to higher consumer welfare.
+Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.
+We show that a data advantage may serve as a significant barrier to entry when exploration has significant reputational consequences for entrants.}
 \end{abstract}
@@
 
-\subfile{content/Introduction}
+\subfile{content/introduction}
 
@@
 
+\subfile{content/conclusion}
+
+\subfile{content/appendix_for_one_version}
+
 \bibliographystyle{acm}

### 2018-10-31 paper/content/conclusion.tex
updates to paper

@@
 \gaedit{In this paper we considered a setting where firms simultaneously learn from and compete for users.
-We showed that competition may not always induce firms to commit to algorithms that engage in exploration and that this leads to welfare losses for consumers.
-The primary mechanism behind this was that exploration may have short-term reputational consequences that lead to naive algorithms winning in competition.
+We showed that competition may not always induce firms to commit to algorithms that engage in exploration resulting in welfare losses for consumers.
+The primary mechanism behind this was that exploration may have short-term reputational consequences that lead to naive algorithms winning in competition that are bad in the long-run.
 Allowing one firm to have a head-start, or a first-mover advantage, leads to the first-mover having an incentive to play traditionally ``good" bandit algorithms and having significantly better welfare for consumers.
 Finally, we attempted to isolate the effect that the data advantage of an incumbent has on its ability to garner a considerable market share.
-Contrary to previous studies and discussions which do not consider competition dynamics or endogenizing data collection, we find that the data advantage is substantial in generating significant market power.
+Contrary to previous studies and discussions which do not consider competition dynamics or endogenous data collection, we find that the data advantage is substantial in generating significant market power.
 Additionally we show that this effect is strongest when an incumbent performs exploration in the incumbency period and generates a high-quality dataset on actions that may be difficult for an entrant to generate due to the pressure of competition.}
-
-\gacomment{Not sure what else to put in the conclusion?}
 

### 2018-10-31 paper/content/introduction.tex
remove this line
updates to paper

@@
-\documentclass[../competing_bandits.tex]{subfiles}
+\documentclass[../competing_bandits_with_appendix.tex]{subfiles}
 \begin{document}
@@
 This would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
-\gaedit{This ``data feedback" loop has been discussed in policy circles and in the business literature, but, to our knowledge, our paper is the first to formally provide a model that allows us to understand this effect.}
 
@@
 
-We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with an early entrant.
-Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (first-mover in duopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late arriver in duopoly)}.
-\gaedit{We find that, even in a duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late entrant in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
-Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run.
-\gaedit{We find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}
+We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a \gaedit{first-mover}.
+Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (the first-mover in temporary monopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late-arriver in temporary monopoly)}.
+\gaedit{We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
+\gaedit{Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that perform better in the long run.
+As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}
 
@@
 
-\gaedit{Our findings show that the data feedback loop arises either when a firm deploys a ``naive" greedy learning algorithm (under simultaneous entry) or has an initial temporary monopoly in the market.
-The findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
+\gaedit{The findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
 Being first in the market in the digital economy gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
@@
 While \cite{varian2018artificial} and \cite{bajari2018impact} argue that, from a statistical perspective, data has decreasing returns to scale and therefore does not serve as a significant competitive advantage, they focus on the problem in isolation.
-We argue that, to reason about this question, one needs to consider what the impact of better data has in terms of competition dynamics and when data collection is endogenous and has consequences, especially when firms need to learn while competing.
+We argue that, to reason about this question, one needs to consider what the impact of better data is when data collection has competitive consequences, especially when firms need to learn while competing.
 Therefore, we run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
@@
 The study of whether data can give a significant competitive advantage has been discussed theoretically in \cite{varian2018artificial, lambrecht2015can} and empirically in \cite{bajari2018impact}.
-These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role, whereas we ask how additional data helps while firms simultaneously solve a machine learning problem and competing for a market of consumers.
+These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.
+We ask how additional data helps while firms simultaneously solve a machine learning problem and compete for consumers.
 In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm.

### 2018-10-31 paper/content/inverted_u.tex
fix typos
updates to paper

@@
 Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from permanent monopoly to ``incumbent" to permanent duopoly to ``entrant", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
+As the ``intensity of competition" varies from permanent monopoly to ``incumbent" to permanent duopoly to ``late entrant", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
 More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
@@
 The lowest level of competition is monopoly, when \DG wins out for the trivial reason of tie-breaking.
-The highest levels are permanent duopoly and ``entrant" (temporary monopoly from the entrant's perspective).
+The highest levels are permanent duopoly and ``late entrant".
 We see that \DG is incentivized for exploration-disadvantaged MAB instances.
-In fact, incentives for \DG get stronger when the model transitions from permanent duopoly to ``entrant".%
+In fact, incentives for \DG get stronger when the model transitions from permanent duopoly to ``late entrant".%
 \footnote{For the Heavy-Tail instance, \DG goes from a weakly dominant startegy to a strictly dominant one.
 For the Uniform instance, \DG goes from a Nash equilibrium strategy to a weakly dominant one.}
-Finally, the middle level of competition, ``incumbent" (\gaedit{duopoly with an initial temporary monopoly}) creates strong incentives for \TS.
-In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U}.
+Finally, the middle level of competition, ``incumbent" in the temporary monopoly creates strong incentives for \TS. In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U}.
 
@@
 Restated in terms of our model, the effect is that one firm attracts new customers at a lower rate than the other, and falls behind in terms of performance because the other firm has more customers to learn from, and this gets worse over time until (almost) all new customers go to the other firm.
-With this intuition in mind, we define  \textit{effective end of game (\Eeog)} for a particular \MRV and realization table, as the last round $t$ such that the agents at this and previous round choose different firms.
+With this intuition in mind, we define  \textit{effective end of game} (\Eeog) for a particular \MRV and realization table, as the last round $t$ such that the agents at this and previous round choose different firms.
 Indeed, the game, effectively, ends after this round.
 We interpret low \Eeog as a strong evidence of the ``death spiral" effect.
-Focusing on the permanent duopoly scanario, we specify the \Eeog values in Table \ref{sim_table} (the second line of each cell).
+Focusing on the permanent duopoly scenario, we specify the \Eeog values in Table \ref{sim_table} (the second line of each cell).
 We find that the \Eeog values are indeed small:
@@
 Following the bandits literature we define pseudo-regret as being the difference between the true mean of the best action and the true mean of the action that is selected.
-We equate consumer welfare with \textit{market} pseudo-regret where market pseudo-regret is the sum of the pseudo-regret of the actions that are experienced in the competition game\footnote{Note that smaller pseudo-regret is better for consumers}.
+We equate consumer welfare with \textit{market} pseudo-regret where market pseudo-regret is the sum of the pseudo-regret of the actions that are experienced in the competition game.\footnote{Note that smaller pseudo-regret is better for consumers}
 
@@
 
-Figure \ref{eq_regret} displays the simulated welfare when the equilibrium strategies are played.
+Figure \ref{eq_regret} displays the average welfare when the equilibrium strategies are played.
 Consumer welfare is \textit{better off} in the temporary monopoly case than in the duopoly case.
@@
 Figure \ref{many_firm_welfare} reports the average welfare across these simulations.
-Welfare not only does not get better, \textit{but is even worse} as we increase the number of firms.
+Welfare not only does not get better, \textit{but is weakly worse} as we increase the number of firms.
 Correspondingly, we track the average \Eeog in each of the simulations and notice that it is \textit{increasing} as the number of firms increases.
@@
 Additionally, bad luck with rewards or arm selection is punished even more harshly as we increase the number of firms.
-This further strengthens the intuition from our previous result, which is that learning may be hard under competition since accidental exploration and mistakes are punished harshly but are necessary for learning.}
+This further strengthens the intuition from our previous result, which is that learning may be hard under competition since  exploration and mistakes are punished harshly but are necessary for learning.}
 

### 2018-10-31 paper/content/model.tex
fix typos
updates to paper

--- 2018-09-06 paper/content/model.tex
@@
 Bernoulli rewards: the  reward of each arm $a$ is drawn from $\{0,1\}$ independently with expectation $\mu(a)$.
-The mean rewards $\mu(a)$ are the same for all rounds and both firms, but but initially unknown.
+The mean rewards $\mu(a)$ are the same for all rounds and both firms, but initially unknown.
 
@@
 Each algorithm receives a ``warm start": additional $T_0$ agents that arrive before the game starts, and interact with the firm as described above.
-Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
+\gaedit{The warm start serves to vary the initial information the firm has before beginning to compete for users as well as to initialize the agents' perception of the firms.} Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
 
-In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``entrant"), and therefore enjoys a \emph{temporary monopoly}.
+In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``late entrant"), and therefore enjoys a \emph{temporary monopoly}.
 Formally, the incumbent enjoys additional $X$ rounds of the ``warm start".
@@
 \subsubsection{Agents.}
-Agents are myopic and non-strategic: they choose among the firms so as to maximize their expected reward, without attempting to influence the firms' learning algorithms or rewards of the future users.
+\gaedit{For simplicity we suppose that firms compete for agents on a single dimension, quality, that is summarized via the rewards experienced by past consumers.\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}} Agents are myopic and non-strategic: they choose among the firms so as to maximize their expected reward \gaedit{(i.e.
+select the firm with the highest quality)}, without attempting to influence the firms' learning algorithms or rewards of the future users.
 Agents are not well-informed: they only receive a rough signal about each firm's performance before they choose a firm, and no other information.
@@
 We posit a version of rational behavior: each agent chooses a firm with a maximal reputation score (breaking ties uniformly).
-The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
+The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.\gaedit{\footnote{One can think of this as similar to ratings aggregations on many online platforms such as Yelp, Tripadvisor, etc.}}
 
@@
 Our pilot experiments indicate that the results do not change substantially if other algorithms are chosen.
-For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge).
+\gaedit{We make this restriction since we are mainly interested in qualitative differences between standard ``good" multi-armed bandit algorithms and what algorithms are incentivized in competition.} For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge).
 We consider:

### 2018-11-04 paper/content/introduction.tex
comment

@@
 \gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model.
-In our model, two firms compete for users \gaedit{on a single dimension, quality}, and simultaneously learn from them.
+In our model, two firms compete for users \gaedit{on a single dimension, quality},\swcomment{not sure what this means} and simultaneously learn from them.
 Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm.
@@
 We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a \gaedit{first-mover}.
-Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (the first-mover in temporary monopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late-arriver in temporary monopoly)}.
+\swcomment{confused about how they correspond to ``competition intensity''}Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (the first-mover in temporary monopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late-arriver in temporary monopoly)}.
 \gaedit{We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.

### 2018-11-04 paper/content/introduction.tex
intro

@@
 This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.
-\gaedit{The mechanism that leads to this behavior in our model is different from the existing literature in economics since the disincentives to ``innovate" arise entirely from the reputational consequences of exploration in competition and not from R\&D costs (which are absent from our model).}
+\swedit{However, the ``inverted-U relationship'' is driven by different aspects in our model than the ones in the existing literature in economics.
+In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, as opposed to the R\&D costs (which is the more standard cause in prior work)}.
 

### 2018-11-04 paper/content/introduction.tex
intro

@@
 This highlights that, when thinking about data as a barrier to entry, data quality and not just quantity matters.
-In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}
+In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}\swcomment{this is quite confusing.
+let's talk}
 

### 2018-11-04 paper/content/introduction.tex
intro

@@
     \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
-Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
+Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.\swcomment{not sure we need to talk about dueling bandits}
 \gaedit{There is a superficial similarity (in name only) between this paper and the line of work on ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}).

### 2018-11-04 paper/content/introduction.tex
intro

@@
 \gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model.
-In our model, two firms compete for users \gaedit{on a single dimension, quality},\swcomment{not sure what this means} and simultaneously learn from them.
+In our model, two firms compete for users \swdelete{on a single dimension, quality},\swcomment{not sure what this means} and simultaneously learn from them.
 Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm.
@@
 
-Even though, under our model, consumers are best off when there is a ``temporary" monopoly, a recent subject of debate in policy circles is whether the data advantage alone poses a sufficiently powerful barrier to entry to warrant anti-trust interest.
+\swdelete{Even though, under our model, consumers are best off when there is a ``temporary" monopoly, a recent subject of debate in policy circles is whether the data advantage alone poses a sufficiently powerful barrier to entry to warrant anti-trust interest.}
+\swedit{Beyond the welfare implications, we also investigate the barrier of entry from the incumbant's data advantage.}
  While \cite{varian2018artificial} and \cite{bajari2018impact} argue that, from a statistical perspective, data has decreasing returns to scale and therefore does not serve as a significant competitive advantage, they focus on the problem in isolation.
@@
 This highlights that, when thinking about data as a barrier to entry, data quality and not just quantity matters.
-In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}\swcomment{this is quite confusing.
-let's talk}
+In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}\swcomment{guy, could you make another pass over above}
 
@@
 
-The study of whether data can give a significant competitive advantage has been discussed theoretically in \cite{varian2018artificial, lambrecht2015can} and empirically in \cite{bajari2018impact}.
-These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.
-We ask how additional data helps while firms simultaneously solve a machine learning problem and compete for consumers.
+The study of whether data can give a significant \swdelete{competitive} advantage
+has been discussed theoretically in \cite{varian2018artificial,
+  lambrecht2015can} and empirically in
+\cite{bajari2018impact}. \swedit{While they argue that small amounts
+  of additional data do not provide significant improvement in
+  training a more predictive model, this set of results focus on learning in isolation, which overlooks certain aspects in a competition.}
+\swdelete{
+These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.} We ask how additional data helps while firms simultaneously solve a machine learning problem and compete for consumers.
 In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm.

### 2018-11-04 paper/content/inverted_u.tex
editing

@@
 
-What do our results mean for the welfare of consumers?
-In our model we define consumer welfare in terms of pseudo-regret.
-Following the bandits literature we define pseudo-regret as being the difference between the true mean of the best action and the true mean of the action that is selected.
-We equate consumer welfare with \textit{market} pseudo-regret where market pseudo-regret is the sum of the pseudo-regret of the actions that are experienced in the competition game.\footnote{Note that smaller pseudo-regret is better for consumers}
+\swedit{We also study the effects the competition has on the consumer
+  welfare---the cumulative rewards received by the users over time.}
 
@@
 
-Figure \ref{eq_regret} displays the average welfare when the equilibrium strategies are played.
-Consumer welfare is \textit{better off} in the temporary monopoly case than in the duopoly case.
-However, this was expected from the fact that the incumbent is incentivized to play \TS.
-The welfare in the temporary monopoly case is close to the welfare to having a single firm playing \TS.
+\swedit{For our purpose, it will be clearer to look at \emph{market
+    regret}: the difference between the cumulative mean rewards
+  achieved by taking the best action over all rounds and the
+  cumulative mean rewards received by the agents. (Hence smaller
+  regret means higher welfare.)}
 
-What is surprising is that welfare under monopoly and duopoly are very similar, with monopoly being marginally better than duopoly.
-Recall that in equilibrium in these market structures all firms play \DG.
-One would think that welfare would be weakly increasing with the number of firms playing \DG since, from the perspective of the market, if adding a firm playing \DG results in that firm, by chance, identifying a better arm than the existing firms then that firm ought to attract more consumers and increase market welfare.
-If the firm identifies a worse arm than the rest then one ought to expect consumers to not select this firm and market welfare would be unchanged.
-Indeed, adding sufficiently many firms one would expect that eventually one firm by chance would discover the best arm quickly and the market welfare could even beat that of a stand-alone firm playing \TS.
+Figure \ref{eq_regret} displays the \swedit{market regret (average
+  over multiple runs)} under different levels of competitions,
+assuming both firms playing their equilibrium strategies. Consumer
+welfare is \textit{better off} in the temporary monopoly case than in
+the duopoly case. However, this was expected from the fact that the
+incumbent is incentivized to play \TS. The welfare in the temporary
+monopoly case is close to the welfare to having a single firm playing
+\TS.
 
+\swedit{We also observe that monopoly and duopoly achieve similar
+  welare, with monopoly being marginally better than duopoly. This is
+  surprising since one might conjecture that the welfare should
+  increase with the number of firms playing \DG since more $\DG$ runs
+  should result in faster concentration on playing the better arms. In
+  particular,, more firms playing \DG should result in faster
+  discovery of a better arm, and the firm with such discovery would be
+  able to attract more consumers and increase market welfare.}
 \begin{figure}
@@
 
-To test if it is possible that increasing competition past a duopoly leads to better welfare than under temporary monopoly we run simulations where we increase the number of firms playing \DG.
-Figure \ref{many_firm_welfare} reports the average welfare across these simulations.
-Welfare not only does not get better, \textit{but is weakly worse} as we increase the number of firms.
-Correspondingly, we track the average \Eeog in each of the simulations and notice that it is \textit{increasing} as the number of firms increases.
-As we increase the number of firms, a single individual firm gets fewer consumers for learning as consumers switch between firms more often.
-Additionally, bad luck with rewards or arm selection is punished even more harshly as we increase the number of firms.
-This further strengthens the intuition from our previous result, which is that learning may be hard under competition since  exploration and mistakes are punished harshly but are necessary for learning.}
+\swedit{To study this phenomenon further, we go beyond the two-firm
+  setting and simulate the game an increasing number of firms playing
+  \DG.} Figure \ref{many_firm_welfare} reports the average welfare
+across these simulations. Welfare not only does not get better,
+\textit{but is weakly worse} as we increase the number of
+firms. Correspondingly, we track the average \Eeog in each of the
+simulations and notice that it is \textit{increasing} as the number of
+firms increases. As we increase the number of firms, a single
+individual firm gets fewer consumers for learning as consumers switch
+between firms more often. Additionally, bad luck with rewards or arm
+selection is punished even more harshly as we increase the number of
+firms. This further strengthens the intuition from our previous
+result, which is that learning may be hard under competition since
+exploration and mistakes are punished harshly but are necessary for
+learning.}
 

### 2018-11-04 paper/content/introduction.tex
updated intro

@@
 \gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model.
-In our model, two firms compete for users \swdelete{on a single dimension, quality},\swcomment{not sure what this means} and simultaneously learn from them.
+In our model, two firms compete for users and simultaneously learn from them.
 Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm.
@@
 
-\swdelete{Even though, under our model, consumers are best off when there is a ``temporary" monopoly, a recent subject of debate in policy circles is whether the data advantage alone poses a sufficiently powerful barrier to entry to warrant anti-trust interest.}
-\swedit{Beyond the welfare implications, we also investigate the barrier of entry from the incumbant's data advantage.}
- While \cite{varian2018artificial} and \cite{bajari2018impact} argue that, from a statistical perspective, data has decreasing returns to scale and therefore does not serve as a significant competitive advantage, they focus on the problem in isolation.
-We argue that, to reason about this question, one needs to consider what the impact of better data is when data collection has competitive consequences, especially when firms need to learn while competing.
-Therefore, we run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
+\swedit{Beyond the welfare implications, we also investigate the barrier to entry that follows from the incumbent's data advantage.}  We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
 We find that the data advantage alone still leads to a significant incumbency advantage.
 Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm.
-This highlights that, when thinking about data as a barrier to entry, data quality and not just quantity matters.
-In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}\swcomment{guy, could you make another pass over above}
+Our results highlight the importance of considering competition dynamics when thinking about the anti-trust question of whether data can serve as a significant barrier to entry in an industry.
+While data has decreasing returns to scale from a statistical perspective, in a competitive environment the relative differences in the data possessed by the firms may lead to a large competitive benefit due to the feedback loops discussed previously.
+In our model this effect is amplified when the incumbent commits to a more advanced bandit algorithm since, in the incumbency period, the firm gathers data on possible actions more efficiently leading to the relative gap upon entry to be larger.
+This is important when exploration can have larger reputational consequences.}\swcomment{guy, could you make another pass over above}
 

### 2018-11-04 paper/content/barriers.tex
updates to paper
more tweaks

@@
 If the entrant had some way of acquiring this data without incurring reputational consequences then the data advantage would be substantially weaker.
-This insight is what differentiates our results from the static viewpoint of \cite{varian2018artificial, bajari2018impact}.
+This insight is what differentiates our results from the static viewpoint of the previous literature.
 While it remains true that, for a given action, increased data on that action has diminishing returns to scale, competitive pressures may make it harder for smaller firms to be able to engage in the exploration necessary to accumulate sufficient data on actions to learn if they are good or bad.
 Otherwise, they may be more likely to fall into the death-spiral effect discussed earlier.
-This effect only shows itself when thinking about the issue from the perspective of competition dynamics and endogenous data acquisition.}
+This effect arises when thinking about the issue from the perspective of competition dynamics.}
 

### 2018-11-04 paper/content/conclusion.tex
more tweaks

@@
 The primary mechanism behind this was that exploration may have short-term reputational consequences that lead to naive algorithms winning in competition that are bad in the long-run.
-Allowing one firm to have a head-start, or a first-mover advantage, leads to the first-mover having an incentive to play traditionally ``good" bandit algorithms and having significantly better welfare for consumers.
+Allowing one firm to have a head-start, or a first-mover advantage, leads to the first-mover having an incentive to play traditionally ``good" bandit algorithms and resulting in significantly better welfare for consumers.
 Finally, we attempted to isolate the effect that the data advantage of an incumbent has on its ability to garner a considerable market share.
-Contrary to previous studies and discussions which do not consider competition dynamics or endogenous data collection, we find that the data advantage is substantial in generating significant market power.
+Contrary to previous studies and discussions which do not consider competition dynamics, we find that the data advantage is substantial in generating significant market power.
 Additionally we show that this effect is strongest when an incumbent performs exploration in the incumbency period and generates a high-quality dataset on actions that may be difficult for an entrant to generate due to the pressure of competition.}

### 2018-11-05 paper/content/inverted_u.tex
small update
better eeog fig

@@
   over multiple runs)} under different levels of competitions,
-assuming both firms playing their equilibrium strategies. Consumer
-welfare is \textit{better off} in the temporary monopoly case than in
-the duopoly case. However, this was expected from the fact that the
-incumbent is incentivized to play \TS. The welfare in the temporary
-monopoly case is close to the welfare to having a single firm playing
+assuming both firms playing their equilibrium strategies. Consumers are \textit{better off} in the temporary monopoly case than in
+the duopoly case.
+However, this was expected from the fact that the incumbent is incentivized to play \TS.
+The welfare in the temporary monopoly case is close to the welfare to having a single firm playing
 \TS.
@@
 \swedit{We also observe that monopoly and duopoly achieve similar
-  welare, with monopoly being marginally better than duopoly. This is
+  welfare, with monopoly being marginally better than duopoly. This is
   surprising since one might conjecture that the welfare should

### 2018-11-05 paper/competing_bandits.tex
Inserted a short/running title (without a line break).

@@
 \begin{document}
-\title{Competing Bandits: \\ The Perils of Exploration under Competition}
+\title[Competing Bandits: The Perils of Exploration under Competition]
+{Competing Bandits:\\The Perils of Exploration under Competition}
 
@@
 Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.
-We show that a data advantage may serve as a significant barrier to entry when exploration has significant reputational consequences for entrants.}
+We show that a data advantage may serve as a significant barrier to entry when exploration has reputational consequences for entrants.}
 \end{abstract}

### 2018-11-05 paper/content/introduction.tex
edits

@@
 Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm.
-Our results highlight the importance of considering competition dynamics when thinking about the anti-trust question of whether data can serve as a significant barrier to entry in an industry.
+Our results highlight the importance of considering competition dynamics when thinking about the anti-trust question of whether data can become a significant barrier to entry in an industry.
 While data has decreasing returns to scale from a statistical perspective, in a competitive environment the relative differences in the data possessed by the firms may lead to a large competitive benefit due to the feedback loops discussed previously.
 In our model this effect is amplified when the incumbent commits to a more advanced bandit algorithm since, in the incumbency period, the firm gathers data on possible actions more efficiently leading to the relative gap upon entry to be larger.
-This is important when exploration can have larger reputational consequences.}\swcomment{guy, could you make another pass over above}
+This is important when exploration can have larger reputational consequences.}
 
@@
     \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
-Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.\swcomment{not sure we need to talk about dueling bandits}
-\gaedit{There is a superficial similarity (in name only) between this paper and the line of work on ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}).
-The latter is not about competing
-bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".
+Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
+\swedit{There is also a line of work on ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}) that also studies a form of competition with MAB problems.
+However, their framework is fundamentally different from ours in that they study learning from setting ``duels'' between pairs of arms, and we study how learning algorithms compete with each other for data.
+}
 
@@
   of additional data do not provide significant improvement in
-  training a more predictive model, this set of results focus on learning in isolation, which overlooks certain aspects in a competition.}
+  training a more predictive model, this set of results focus on learning in isolation, which overlooks certain aspects in a competition.
 \swdelete{

### 2018-11-05 paper/content/introduction.tex
edit

@@
 \swedit{There is also a line of work on ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}) that also studies a form of competition with MAB problems.
-However, their framework is fundamentally different from ours in that they study learning from setting ``duels'' between pairs of arms, and we study how learning algorithms compete with each other for data.
+However, their framework is fundamentally different from ours in that they study learning from setting up ``duels'' between pairs of arms, and we study how learning algorithms compete with each other for data.
 }

### 2018-11-05 paper/content/introduction.tex
intro

@@
 
-\swedit{Beyond the welfare implications, we also investigate the barrier to entry that follows from the incumbent's data advantage.}  We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
-We find that the data advantage alone still leads to a significant incumbency advantage.
-Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm.
-Our results highlight the importance of considering competition dynamics when thinking about the anti-trust question of whether data can become a significant barrier to entry in an industry.
-While data has decreasing returns to scale from a statistical perspective, in a competitive environment the relative differences in the data possessed by the firms may lead to a large competitive benefit due to the feedback loops discussed previously.
-In our model this effect is amplified when the incumbent commits to a more advanced bandit algorithm since, in the incumbency period, the firm gathers data on possible actions more efficiently leading to the relative gap upon entry to be larger.
-This is important when exploration can have larger reputational consequences.}
+\swedit{Beyond the welfare implications, we also investigate the barrier to entry that follows from the incumbent's data advantage.
+We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
+We find that the data advantage from the incumbent's ``small head start'' can lead to a significant incumbency advantage, especially when the incumbent commits to a more advanced bandit algorithm.
+Our results show that even a small degree of data assymmetry can be amplified to be a larger market advantage through the competition dynamics.
+This is in contrast to the data advantage in a static environment \cite{varian2018artificial,bajari2018impact}, in which a small amount of additional data can only provide limited improvement in the predictive accuracy.
+Our results shed lights on the importance of considering competition dynamics when reasoning about the anti-trust question of whether data can become a significant barrier to entry in an industry. 
+% in a competitive environment the relative differences in the data
+% possessed by the firms may lead to a large competitive benefit due
+% to the feedback loops discussed previously. In our model this effect
+% is amplified when the incumbent commits to a more advanced bandit
+% algorithm since, in the incumbency period, the firm gathers data on
+% possible actions more efficiently leading to the relative gap upon
+% entry to be larger. This is important when exploration can have
+% larger reputational consequences.
+}}\swcomment{still not super happy about above.}
 

### 2018-11-05 paper/content/barriers.tex
compete

@@
 \gaedit{Even though welfare is highest in the temporary monopoly case there are other reasons one may worry about a single firm having a substantial market share.
-In this section we explore what drives the substantial market share of the temporary monopolist and discuss how this relates to thinking about data as a barrier to entry.}
+In this section we explore what drives the substantial market share of the temporary monopolist and discuss how this relates to thinking about data as a barrier to entry.}\swcomment{maybe articulate the ``other'' reasons?}
 
@@
 
-\gaedit{How does this help us in thinking about data as a barrier to entry?
+\swedit{Our finding also highlights that the amplification of data
+  advantage through the competition dynamics. In particular, in the
+  incumbency period \TS gathers data that may be hard for the entrant
+  to gather during competition due to its potential reputational
+  consequences (by engaging in exploration). This small ``head start''
+  goes a long way---the incumbent continues to attract data
+  efficiently without sacrificing much of its exploration, while the
+  entrant tends to struggle to gather data even with limited
+  exploration.
+}
+
+\swdelete{\gaedit{How does this help us in thinking about data as a barrier to entry?
 The result that the data advantage is stronger under \TS than \DEG and \DG gives evidence that data quality, and not just quantity, matters in thinking about the competitive advantage of data.
@@
 Otherwise, they may be more likely to fall into the death-spiral effect discussed earlier.
-This effect arises when thinking about the issue from the perspective of competition dynamics.}
+This effect arises when thinking about the issue from the perspective of competition dynamics.}}
 

### 2018-11-05 paper/content/conclusion.tex
compete

@@
 The primary mechanism behind this was that exploration may have short-term reputational consequences that lead to naive algorithms winning in competition that are bad in the long-run.
-Allowing one firm to have a head-start, or a first-mover advantage, leads to the first-mover having an incentive to play traditionally ``good" bandit algorithms and resulting in significantly better welfare for consumers.
+Allowing one firm to have a head start, or a first-mover advantage, leads to the first-mover having an incentive to play traditionally ``good" bandit algorithms and resulting in significantly better welfare for consumers.
 Finally, we attempted to isolate the effect that the data advantage of an incumbent has on its ability to garner a considerable market share.
 Contrary to previous studies and discussions which do not consider competition dynamics, we find that the data advantage is substantial in generating significant market power.
-Additionally we show that this effect is strongest when an incumbent performs exploration in the incumbency period and generates a high-quality dataset on actions that may be difficult for an entrant to generate due to the pressure of competition.}
+Additionally we show that this effect is strongest when an incumbent \swedit{performs efficient exploration in the incumbency period without the pressure from a competitor.}}
 

### 2018-11-05 paper/content/barriers.tex
updates

--- 2018-11-04 paper/content/barriers.tex
@@
 
+\OMIT{
 \gaedit{Even though welfare is highest in the temporary monopoly case there are other reasons one may worry about a single firm having a substantial market share.
-In this section we explore what drives the substantial market share of the temporary monopolist and discuss how this relates to thinking about data as a barrier to entry.}
+In this section we explore what drives the substantial market share of the temporary monopolist and discuss how this relates to thinking about data as a barrier to entry.}\swcomment{maybe articulate the ``other'' reasons?}
+}
+\gaedit{In this section we explore what factors drive the large market share for the incumbent in the temporary monopoly.}
 
@@
 
-\gaedit{How does this help us in thinking about data as a barrier to entry?
+\gadelete{
+\swedit{Our finding also highlights the amplification of the data
+  advantage through competition dynamics. In particular, in the
+  incumbency period \TS gathers data \gaedit{ through exploration } that may be hard for the entrant to gather during competition due to its potential reputational consequences \gadelete{(by engaging in exploration)}.
+This small ``head start'' goes a long way---the incumbent continues to \gadelete{attract data} \gaedit{attract users} efficiently without sacrificing much of its exploration, while the entrant tends to struggle to gather data even with limited exploration.
+}
+}
+
+\gaedit{Our findings imply two aspects of how and when data can serve as a barrier to entry.
+
+The first is that competition dynamics can lead to small asymmetries in data between firms amplifying into large effects through the same mechanism as the death spiral effect discussed in the previous section.
+
+The second is that data can serve as a larger barrier to entry when exploration is reputationally costly.
+This is due to the fact that the incumbent can explore actions in the incumbency period without immediate reputational consequences, whereas in order for the entrant to acquire the same data it needs to incur the immediate reputational consequences.
+As a result, when the incumbent does exploration in the incumbency period the value of the data accumulated is higher and it is harder for an entrant to compete.
+}
+
+\swdelete{\gaedit{How does this help us in thinking about data as a barrier to entry?
 The result that the data advantage is stronger under \TS than \DEG and \DG gives evidence that data quality, and not just quantity, matters in thinking about the competitive advantage of data.

### 2018-11-05 paper/content/introduction.tex
updates

@@
 We find that the data advantage from the incumbent's ``small head start'' can lead to a significant incumbency advantage, especially when the incumbent commits to a more advanced bandit algorithm.
-Our results show that even a small degree of data assymmetry can be amplified to be a larger market advantage through the competition dynamics.
-This is in contrast to the data advantage in a static environment \cite{varian2018artificial,bajari2018impact}, in which a small amount of additional data can only provide limited improvement in the predictive accuracy.
+Our results show that even a small degree of data asymmetry can be amplified to be a larger market advantage through competition dynamics.
+This is in contrast to the data advantage in a static environment such as discussed in \cite{varian2018artificial,bajari2018impact}, in which a small amount of additional data can only provide limited improvement in the predictive accuracy.
 Our results shed lights on the importance of considering competition dynamics when reasoning about the anti-trust question of whether data can become a significant barrier to entry in an industry. 

### 2018-11-05 paper/content/inverted_u.tex
Some space-saving.
updates

@@
 \begin{document}
-\section{Competition and Better MAB Algorithms}\label{sec:competition}
+\section{Competition vs. Better Algorithms}\label{sec:competition}
 
@@
 
-\subsubsection{Permanent duopoly.}
+\xhdr{Permanent duopoly.}
 The basic scenario is when both firms are competing from round $1$. A crucial distinction is whether an MAB instance is exploration-disadvantaged:
@@
 
-\subsubsection{Temporary Monopoly.}
+\xhdr{Temporary Monopoly.}
 We turn our attention to the temporary monopoly scenario.
@@
 
-\subsubsection{Inverted-U relationship.}
+\xhdr{Inverted-U relationship.}
 We interpret our findings through the lens of the inverted-U relationship between the ``intensity of competition" and the ``quality of technology".
@@
 
-\subsubsection{Death spiral effect.}
+\xhdr{Death spiral effect.}
 Further, we investigate the ``death spiral" effect mentioned in the Introduction.
@@
 \gaedit{
-\subsubsection{Welfare}
+\xhdr{Welfare}
 
@@
 \includegraphics[scale=0.3]{figures/ht_many_firm_welfare}\\
-\caption{Average welfare and EEOG as we increase the number of firms playing \DG}
+\caption{Average welfare and EEOG as we increase \#firms playing \DG}
 \label{many_firm_welfare}

### 2018-11-05 paper/content/model.tex
updates

@@
 
-\subsubsection{Agents.}
+\xhdr{Agents.}
 \gaedit{For simplicity we suppose that firms compete for agents on a single dimension, quality, that is summarized via the rewards experienced by past consumers.\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}} Agents are myopic and non-strategic: they choose among the firms so as to maximize their expected reward \gaedit{(i.e.
@@
 
-\subsubsection{MAB algorithms.} We consider three classes of algorithms, ranging from more primitive to more sophisticated:
+\xhdr{MAB algorithms.} We consider three classes of algorithms, ranging from more primitive to more sophisticated:
 
@@
 
-\subsubsection{MAB instances.}
+\xhdr{MAB instances.}
 We consider instances with $K=10$ arms.
@@
 
-\subsubsection{Terminology.}
+\xhdr{Terminology.}
 Following a standard game-theoretic terminology, algorithm Alg1 \emph{(weakly) dominates} algorithm Alg2 for a given firm if Alg1 provides a larger (or equal) market share than Alg2 at the end of the game.
@@
 
-\subsubsection{Simulation details.}
+\xhdr{Simulation details.}
 For each MAB instance we draw $N = 1000$ \MRVs independently from the corresponding distribution.

### 2018-11-05 paper/content/barriers.tex
more tweaks

@@
 
-The first is that competition dynamics can lead to small asymmetries in data between firms amplifying into large effects through the same mechanism as the death spiral effect discussed in the previous section.
+The first is that small asymmetries in data between firms can amplify into large effects through the same mechanism as the death spiral effect discussed in the previous section.
 
@@
 This is due to the fact that the incumbent can explore actions in the incumbency period without immediate reputational consequences, whereas in order for the entrant to acquire the same data it needs to incur the immediate reputational consequences.
-As a result, when the incumbent does exploration in the incumbency period the value of the data accumulated is higher and it is harder for an entrant to compete.
+As a result, when the incumbent does exploration in the incumbency period the value of the data accumulated is higher and it is harder for an entrant to compete in the market.
 }

### 2018-11-05 paper/content/introduction.tex
superficial tweaks to intro

@@
 \gaedit{We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
-\gaedit{Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that perform better in the long run.
+\gaedit{Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that performs better in the long run.
 As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}
@@
 In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm.
-Additionally, the first-mover advantage has been well-studied in economics and marketing \cite{kerin1992first}.
+Additionally, the first-mover advantage has been well-studied in other settings in economics and marketing \cite{kerin1992first}.
 }

### 2018-11-06 paper/content/introduction.tex
Mostly, edits in related work.

@@
 
-\gaedit{We ask whether algorithms that are traditionally considered ``better" in the standard multi-armed bandit problem are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model.
+\gaedit{We ask whether algorithms that are traditionally considered ``better" in the literature on multi-armed bandits are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model.
 In our model, two firms compete for users and simultaneously learn from them.
@@
 
-\subsubsection{Related work.}
-Our work is related to a longstanding economics literature on competition vs.
-innovation, \eg \cite{Schumpeter-42,barro2004economic,Aghion-QJE05}.
-While this literature focuses on R\&D costs of innovation, ``reputational costs" seem new and specific to exploration.
+\subsection{Related work}
 
-Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information).
+\xhdr{Machine learning.} Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information).
 MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background.
@@
     \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
+
 Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
-\swedit{There is also a line of work on ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}) that also studies a form of competition with MAB problems.
-However, their framework is fundamentally different from ours in that they study learning from setting up ``duels'' between pairs of arms, and we study how learning algorithms compete with each other for data.
-}
 
-The study of whether data can give a significant \swdelete{competitive} advantage
-has been discussed theoretically in \cite{varian2018artificial,
+\asedit{In ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}), an algorithm sets up a ``duel" between a pair of arms in each round, and only learns which arm has ``won".
+While this setting features a form of competition inside an MAB problem, it is very different from ours.}
+
+\xhdr{Economics.}
+Our work is related to a longstanding economics literature on competition vs.
+innovation, \eg \cite{Schumpeter-42,barro2004economic,Aghion-QJE05}.
+While this literature focuses on R\&D costs of innovation, ``reputational costs" seem new and specific to exploration.
+
+\asedit{Whether data gives competitive advantage
+has been studied theoretically in \cite{varian2018artificial,
   lambrecht2015can} and empirically in
-\cite{bajari2018impact}. \swedit{While they argue that small amounts
-  of additional data do not provide significant improvement in
-  training a more predictive model, this set of results focus on learning in isolation, which overlooks certain aspects in a competition.
-\swdelete{
-These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.} We ask how additional data helps while firms simultaneously solve a machine learning problem and compete for consumers.
-In our setting, the competition dynamics from our model lead to different conclusions about the competitive advantage that a data advantage can give a firm.
-Additionally, the first-mover advantage has been well-studied in other settings in economics and marketing \cite{kerin1992first}.
-}
+\cite{bajari2018impact}. 
+While these papers find that small amounts
+  of additional data do not provide significant improvement,
+ % in training a more predictive model, 
+  they focus on learning in isolation.
+  %, which overlooks certain aspects in a competition.
+%\swdelete{These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.} 
+We ask how additional data helps when firms simultaneously solve a machine learning problem and compete for consumers.
+Due to the competition dynamics in our model, we find that even a small amount of additional data tends to help a lot.}
+
+The first-mover advantage has been well-studied in other settings in economics and marketing, see survey \cite{kerin1992first}.
 

### 2018-11-06 paper/content/introduction.tex
Another edit in related work.
Added a few more refs.

@@
 
+Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
+
+\asedit{In ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}), an algorithm sets up a ``duel" between a pair of arms in each round, and only learns which arm has ``won".
+While this setting features a form of competition inside an MAB problem, it is very different from ours.}
+
 The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
@@
 online ad auctions, \eg
-    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm,Amin-auctions-nips13},
-and human computation
-    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
-
-Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.
-
-\asedit{In ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}), an algorithm sets up a ``duel" between a pair of arms in each round, and only learns which arm has ``won".
-While this setting features a form of competition inside an MAB problem, it is very different from ours.}
+    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm},
+human computation
+    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13},
+and repeated actions, \eg
+    \cite{Amin-auctions-nips13,Amin-auctions-nips14,Jieming-ec18}.
 

### 2018-11-06 paper/content/model.tex
Some edits in the model.

@@
 \xhdr{Agents.}
-\gaedit{For simplicity we suppose that firms compete for agents on a single dimension, quality, that is summarized via the rewards experienced by past consumers.\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}} Agents are myopic and non-strategic: they choose among the firms so as to maximize their expected reward \gaedit{(i.e.
+\gaedit{For simplicity we suppose that firms compete for agents on a single dimension, quality, that is summarized via the rewards experienced by past consumers.
+\OMIT{\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}}}
+Agents are myopic and non-strategic: they \asedit{would like} to choose among the firms so as to maximize their expected reward \gaedit{(i.e.
 select the firm with the highest quality)}, without attempting to influence the firms' learning algorithms or rewards of the future users.
@@
 
-Concretely, each firm has a \emph{reputation score}, and each agent's choice is driven by these two numbers.
+Concretely, each of the two firms has a \emph{reputation score}, and each agent's choice is driven by these two numbers.
 We posit a version of rational behavior: each agent chooses a firm with a maximal reputation score (breaking ties uniformly).
-The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.\gaedit{\footnote{One can think of this as similar to ratings aggregations on many online platforms such as Yelp, Tripadvisor, etc.}}
+The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
+\asedit{This is similar to ratings aggregators on many online platforms such as Yelp, Tripadvisor, etc.).}
 
@@
 
-For concreteness, we fix one algorithm from each class.
-Our pilot experiments indicate that the results do not change substantially if other algorithms are chosen.
-\gaedit{We make this restriction since we are mainly interested in qualitative differences between standard ``good" multi-armed bandit algorithms and what algorithms are incentivized in competition.} For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge).
+\asedit{We are mainly interested in qualitative differences between the three classes.} For concreteness, we fix one algorithm from each class.
+Our pilot experiments indicate that our findings do not change substantially if other algorithms are chosen.
+For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge).
 We consider:

### 2018-11-06 paper/content/inverted_u.tex
Minor edits.

@@
 
-\gaedit{
-\xhdr{Welfare}
-
-\swedit{We also study the effects the competition has on the consumer
-  welfare---the cumulative rewards received by the users over time.}
+\xhdr{Welfare.}
+We also study the effects of the competition on the consumer welfare---the total reward received by the users over time.
 
@@
 
-\swedit{For our purpose, it will be clearer to look at \emph{market
-    regret}: the difference between the cumulative mean rewards
-  achieved by taking the best action over all rounds and the
-  cumulative mean rewards received by the agents. (Hence smaller
-  regret means higher welfare.)}
+\asedit{Rather than the welfare directly, we find it more illuminating to consider} \emph{market regret}: the difference between the cumulative mean rewards achieved by taking the best action over all rounds and the cumulative mean rewards received by the agents.
+(Hence smaller regret means higher welfare.)
 
@@
   should result in faster concentration on playing the better arms. In
-  particular,, more firms playing \DG should result in faster
+  particular, more firms playing \DG should result in faster
   discovery of a better arm, and the firm with such discovery would be
@@
 exploration and mistakes are punished harshly but are necessary for
-learning.}
+learning.
 

### 2018-11-06 paper/content/inverted_u.tex
restate findings

--- 2018-11-05 paper/content/inverted_u.tex
@@
 
-\swedit{We also study the effects the competition has on the consumer
-  welfare---the cumulative rewards received by the users over time.}
+\swedit{We also study the effects the competition has on the consumer welfare---the cumulative rewards received by the users over time.}
 
@@
 \end{figure}
-
-\begin{finding}
-\textit{ \\ (a) Consumer welfare is highest under temporary monopoly. \\
-(b) Increasing the number of firms in the market playing \DG weakly decreases consumer welfare.
-}
-\end{finding}
 
@@
 The welfare in the temporary monopoly case is close to the welfare to having a single firm playing
-\TS.
+\TS. \gaedit{We also observe that monopoly and duopoly achieve similar welfare, with monopoly being marginally better than duopoly.}
 
-\swedit{We also observe that monopoly and duopoly achieve similar
-  welfare, with monopoly being marginally better than duopoly. This is
-  surprising since one might conjecture that the welfare should
-  increase with the number of firms playing \DG since more $\DG$ runs
+\begin{finding}
+\textit{ In equilibrium, consumer welfare is highest under temporary monopoly. Monopoly and duopoly achieve similar welfare with monopoly leading to marginally better welfare than duopoly.
+}
+\end{finding}
+
+\gaedit{The result that monopoly leads to marginally better welfare than duopoly is surprising since, in equilibrium, all firms play \DG in both settings.
+One might conjecture that the welfare should increase with the number of firms playing \DG since more \DG runs
   should result in faster concentration on playing the better arms. In
-  particular,, more firms playing \DG should result in faster
+  particular, more firms playing \DG should result in faster
   discovery of a better arm, and the firm with such discovery would be
@@
 
+\begin{finding}
+\textit{When all firms in the market deploy \DG, welfare is weakly decreasing as the number of firms in the market decreases.}
+\end{finding}
 \swedit{To study this phenomenon further, we go beyond the two-firm

### 2018-11-06 paper/content/introduction.tex
tweaks to barriers intro

@@
 \gaedit{The findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-Being first in the market in the digital economy gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
+Being first in the market in the digital economy gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").}
 
+\gaedit{Beyond the welfare implications, we also investigate the degree to which the incumbent's data advantage serves as a barrier to entry.
+We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
+We find that the data advantage from the incumbent's ``small head start'' can lead to a significant incumbency advantage, especially when the incumbent commits to a more advanced bandit algorithm.
+Our results show that even a small degree of data asymmetry can be amplified to be a larger market advantage through competition dynamics.
+This effect is more pronounced when the incumbent commits to a more advanced bandit algorithm since the incumbent does not have to suffer from the reputational consequences of exploration while the entrant does.
+Our results shed lights on the importance of considering competition dynamics when reasoning about the anti-trust question of whether data can become a significant barrier to entry in an industry.}
+\gadelete{
 \swedit{Beyond the welfare implications, we also investigate the barrier to entry that follows from the incumbent's data advantage.
@@
 Our results show that even a small degree of data asymmetry can be amplified to be a larger market advantage through competition dynamics.
-This is in contrast to the data advantage in a static environment such as discussed in \cite{varian2018artificial,bajari2018impact}, in which a small amount of additional data can only provide limited improvement in the predictive accuracy.
-Our results shed lights on the importance of considering competition dynamics when reasoning about the anti-trust question of whether data can become a significant barrier to entry in an industry.
+\gadelete{This is in contrast to the data advantage in a static environment such as discussed in \cite{varian2018artificial,bajari2018impact}, in which a small amount of additional data can only provide limited improvement in the predictive accuracy.} The effect is stronger when the incumbent commits to a more advanced bandit algorithm due to
+
 % in a competitive environment the relative differences in the data
@@
 % larger reputational consequences.
-}}\swcomment{still not super happy about above.}
+}\swcomment{still not super happy about above.} }
 

### 2018-11-06 paper/content/inverted_u.tex
Edits in Sec4.

@@
 
-\gaedit{
-\xhdr{Welfare}
-
-\swedit{We also study the effects the competition has on the consumer welfare---the cumulative rewards received by the users over time.}
+\xhdr{Welfare.}
+\asedit{We study the effects of competition on consumer welfare: the total reward collected by the users over time. Rather than welfare directly, we find it more lucid to consider
+\emph{market regret}: 
+\[ \textstyle T\, \max_a \mu(a) - \sum_{t\in [T]} \mu(a_t), \]
+where $a_t$ is the arm chosen by agent $t$. This is a standard performance measure in the literature on multi-armed bandits. Note that smaller regret means higher welfare.}
 
@@
 
-\swedit{For our purpose, it will be clearer to look at \emph{market
-    regret}: the difference between the cumulative mean rewards
-  achieved by taking the best action over all rounds and the
-  cumulative mean rewards received by the agents. (Hence smaller
-  regret means higher welfare.)}
+\asedit{We assume that both firms play their respective equilibrium strategies for the corresponding competition level. As discussed previously, these are:}
+\ascomment{Guy, pls fill in, and refer to the corresponding findings.}
 
 Figure \ref{eq_regret} displays the \swedit{market regret (average
-  over multiple runs)} under different levels of competitions,
-assuming both firms playing their equilibrium strategies. Consumers are \textit{better off} in the temporary monopoly case than in
+  over multiple runs)} under different levels of competition.
+Consumers are \textit{better off} in the temporary monopoly case than in
 the duopoly case.

### 2018-11-06 paper/content/inverted_u.tex
update eq strategies

@@
 
-\asedit{We assume that both firms play their respective equilibrium strategies for the corresponding competition level. As discussed previously, these are:}
-\ascomment{Guy, pls fill in, and refer to the corresponding findings.}
+\asedit{We assume that both firms play their respective equilibrium strategies for the corresponding competition level} \gaedit{for $T_0 = 20$.
+As discussed previously, these are: (1) \DG in the monopoly, (2) \DG for both firms in duopoly from Finding \ref{find:duopoly}, (3) \TS for the incumbent (Finding \ref{find:temp-monopoly}) and \DG for the entrant in temporary monopoly (Finding \ref{find:temp-monopoly-entrant}).}
 

### 2018-11-06 paper/content/barriers.tex
merge barriers paragraph into one paragraph
Minor edit.

@@
 }
-\gaedit{In this section we explore what factors drive the large market share for the incumbent in the temporary monopoly.}
+
+\asdelete{We explore what factors drive the large market share for the incumbent in the temporary monopoly model.}
 
@@
 Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant.
-\gaedit{First, our findings provide a quantitative insight into the role of the classic ``first mover advantage" phenomenon in the digital economy.
+\asedit{We explore which of the two factors is more significant.}  \gaedit{First, our findings provide a quantitative insight into the role of the classic ``first mover advantage" phenomenon in the digital economy.
 Second, our findings provide a viewpoint in terms of thinking about the role that data can play as a competitive advantage.}
@@
 \gaedit{Our findings imply two aspects of how and when data can serve as a barrier to entry.
-
 The first is that small asymmetries in data between firms can amplify into large effects through the same mechanism as the death spiral effect discussed in the previous section.
-
 The second is that data can serve as a larger barrier to entry when exploration is reputationally costly.

### 2018-11-06 paper/content/inverted_u.tex
Revision in "welfare".

--- 2018-11-06 paper/content/inverted_u.tex
@@
 
-\asedit{We assume that both firms play their respective equilibrium strategies for the corresponding competition level. As discussed previously, these are:}
-\ascomment{Guy, pls fill in, and refer to the corresponding findings.}
+\asedit{We assume that both firms play their respective equilibrium strategies for the corresponding competition level. As discussed previously, these are: 
+\begin{itemize}
+\item \DG in the monopoly, 
+\item \DG for both firms in duopoly (Finding \ref{find:duopoly}),
+\item \TS for the incumbent (Finding \ref{find:temp-monopoly}) and \DG for the entrant in temporary monopoly (Finding \ref{find:temp-monopoly-entrant}).
+\end{itemize}
+} %%%%%%%
 
-Figure \ref{eq_regret} displays the \swedit{market regret (average
-  over multiple runs)} under different levels of competition.
+Figure \ref{eq_regret} displays the market regret (averaged
+  over multiple runs) under different levels of competition.
 Consumers are \textit{better off} in the temporary monopoly case than in
 the duopoly case.
-However, this was expected from the fact that the incumbent is incentivized to play \TS.
-The welfare in the temporary monopoly case is close to the welfare to having a single firm playing
-\TS. \gaedit{We also observe that monopoly and duopoly achieve similar welfare, with monopoly being marginally better than duopoly.}
+\asedit{Recall that under temporary monopoly, the incumbent is incentivized to play \TS.
+Moreover, we find that the welfare is close to that of having a single firm for all agents and running \TS.
+We also observe that monopoly and duopoly achieve similar welfare.}
+%with monopoly being marginally better than duopoly.
 
-\begin{finding}
-\textit{ In equilibrium, consumer welfare is highest under temporary monopoly. Monopoly and duopoly achieve similar welfare with monopoly leading to marginally better welfare than duopoly.
+\begin{finding}\label{find:welfare}
+\textit{In equilibrium, consumer welfare is (a) highest under temporary monopoly, (b) similar for monopoly and duopoly.
+%monopoly being marginally better leading to marginally better welfare than duopoly.
 }
@@
 
-\gaedit{The result that monopoly leads to marginally better welfare than duopoly is surprising since, in equilibrium, all firms play \DG in both settings.
-One might conjecture that the welfare should increase with the number of firms playing \DG since more \DG runs
-  should result in faster concentration on playing the better arms. In
-  particular, more firms playing \DG should result in faster
-  discovery of a better arm, and the firm with such discovery would be
-  able to attract more consumers and increase market welfare.}
+\asedit{Finding~\ref{find:welfare}(b) is interesting because, in equilibrium, both firms play \DG in both settings, and one might conjecture that the welfare should increase with the number of firms playing \DG.
+Indeed, one run of \DG may get stuck on a bad arm.
+However, two firms independently playing \DG are less likely to get stuck simultaneously.
+If one firm gets stuck and the other does not, then the latter should attract most agents, leading to improved welfare.}
 \begin{figure}
@@
 
+To study this phenomenon further, we go beyond the duopoly setting to more than two firms playing \DG (and starting at the same time). Figure \ref{many_firm_welfare} reports the average welfare
+across these simulations. Welfare not only does not get better, \textit{but is weakly worse} as we increase the number of firms.
+
 \begin{finding}
-\textit{When all firms in the market deploy \DG, welfare is weakly decreasing as the number of firms in the market decreases.}
+\textit{When all firms deploy \DG, and start at the same time, welfare is weakly decreasing as the number of firms decreases.}
 \end{finding}
-\swedit{To study this phenomenon further, we go beyond the two-firm
-  setting and simulate the game an increasing number of firms playing
-  \DG.} Figure \ref{many_firm_welfare} reports the average welfare
-across these simulations. Welfare not only does not get better,
-\textit{but is weakly worse} as we increase the number of
-firms. Correspondingly, we track the average \Eeog in each of the
+
+We track the average \Eeog in each of the
 simulations and notice that it is \textit{increasing} as the number of

### 2018-11-06 paper/content/model.tex
Minor edit in the model.

@@
 We used a parallel implementation over a cluster of 12 2.2 GHz CPU cores, with 8 GB RAM per core.
-With this implementation, each experiment took about $10$ hours.
+Each experiment took about $10$ hours.
 
+\xhdr{Consistency.}
 While we experiments with various MAB instances and parameter settings, in the main paper we only report plots for selected, representative experiments.

### 2018-11-06 paper/content/barriers.tex
More edits.

@@
 Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant.
-\asedit{We explore which of the two factors is more significant.}  \gaedit{First, our findings provide a quantitative insight into the role of the classic ``first mover advantage" phenomenon in the digital economy.
-Second, our findings provide a viewpoint in terms of thinking about the role that data can play as a competitive advantage.}
+\asedit{We explore which of the two factors is more significant.}  Our findings provide a quantitative insight into the role of the classic ``first mover advantage" phenomenon in the digital economy.
+
+%Second, our findings provide a viewpoint in terms of thinking about the role that data can play as a competitive advantage.
 
@@
 
-\gaedit{Our findings imply two aspects of how and when data can serve as a barrier to entry.
+\OMIT{\gaedit{Our findings imply two aspects of how and when data can serve as a barrier to entry.
 The first is that small asymmetries in data between firms can amplify into large effects through the same mechanism as the death spiral effect discussed in the previous section.
@@
 As a result, when the incumbent does exploration in the incumbency period the value of the data accumulated is higher and it is harder for an entrant to compete in the market.
-}
+}}
 

### 2018-11-06 paper/content/model.tex
Minor edit.

@@
 
-\xhdr{Consistency.}
+\asedit{\xhdr{Consistency.}}
 While we experiments with various MAB instances and parameter settings, in the main paper we only report plots for selected, representative experiments.
-Additional plots can be found in the supplement.
+Additional plots will be published in the full version.
 Unless noted otherwise, our findings are based on and consistent with all these plots.

### 2018-11-06 paper/content/inverted_u.tex
More edits.
Revised last para in Sec4.

@@
 \includegraphics[scale=0.3]{figures/ht_many_firm_welfare}\\
-\caption{Average welfare and EEOG as we increase \#firms playing \DG}
+\caption{Average welfare and \Eeog as we increase \#firms playing \DG}
 \label{many_firm_welfare}
@@
 
-We track the average \Eeog in each of the
-simulations and notice that it is \textit{increasing} as the number of
-firms increases. As we increase the number of firms, a single
+\asedit{We track the average \Eeog in each of the
+simulations and notice that it \textit{increases} with the number of firms.
+This observation also runs counter of the intuition that with more firms running \DG, one of them is more likely to ``get lucky" and take over the market (which would cause \Eeog to \emph{decrease} with the number of firms).}
+
+\OMIT{As we increase the number of firms, a single
 individual firm gets fewer consumers for learning as consumers switch
@@
 exploration and mistakes are punished harshly but are necessary for
-learning.
+learning.}
 

### 2018-11-06 paper/competing_bandits.tex
updates
Added a short title.

--- 2018-10-31 paper/competing_bandits.tex
@@
 % a very useful package for edits and comments, from David Kempe (USC)
-\usepackage{color-edits}
-%\usepackage[suppress]{color-edits}  % use this to suppress the package
+%\usepackage{color-edits}
+\usepackage[suppress]{color-edits}  % use this to suppress the package
 \addauthor{as}{red}    % as for Alex
@@
 \begin{document}
-\title{Competing Bandits: \\ The Perils of Exploration under Competition}
+\title[Competing Bandits: The Perils of Exploration under Competition]
+{Competing Bandits: \\ The Perils of Exploration under Competition}
 
@@
 We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore \gaedit{and leads to low consumer welfare}, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms \gaedit{and lead to higher consumer welfare.
-Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.
-We show that a data advantage may serve as a significant barrier to entry when exploration has significant reputational consequences for entrants.}
+Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.}
 \end{abstract}
@@
 
-\keywords{Bandits, Competition}
+\keywords{Multi-armed bandits,Exploration,Competition,Competition vs Innovation}
 

### 2019-01-27 paper/competing_bandits.tex
convert to EC format
fix formatting mistakes

@@
-\documentclass[sigconf, anonymous, review,natbib=false]{acmart}
+\documentclass[format=acmsmall, review=false]{acmart}
 
+\usepackage{acm-ec-19}
 \usepackage{tikz} % SW: Added for plots
@@
 % Letter paper
-\setlength{\paperheight}{11in}
-\setlength{\paperwidth}{8.5in}
 
@@
 %\setcopyright{cagovmixed}
-
-% DOI
-\acmDOI{10.475/123_4}
-
-% ISBN
-\acmISBN{123-4567-24-567/08/06}
-
-%Conference
-\acmConference[WWW'19]{ACM WWW conference}{May 2019}{San Francisco, CA 2019}
-\acmYear{2019}
-\copyrightyear{2019}
-
-\acmPrice{15.00}
 
@@
 %
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10010147.10010257.10010282.10010284</concept_id>
-<concept_desc>Computing methodologies~Online learning settings</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10010147.10010257.10010258.10010261.10010272</concept_id>
-<concept_desc>Computing methodologies~Sequential decision making</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10010405.10010455.10010460</concept_id>
-<concept_desc>Applied computing~Economics</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Computing methodologies~Online learning settings}
-\ccsdesc[300]{Computing methodologies~Sequential decision making}
-\ccsdesc[500]{Applied computing~Economics}
 
@@
 \maketitle
-
 \subfile{content/introduction}

### 2019-01-29 paper/content/model.tex
updates
tweaks

@@
 The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
-\asedit{This is similar to ratings aggregators on many online platforms such as Yelp, Tripadvisor, etc.).}
+\asedit{This is similar to ratings aggregators on many online platforms such as Yelp, Tripadvisor, etc.}
 
@@
 \asedit{\xhdr{Consistency.}}
-While we experiments with various MAB instances and parameter settings, in the main paper we only report plots for selected, representative experiments.
+While we experiment with various MAB instances and parameter settings, in the main paper we only report plots for selected, representative experiments.
 Additional plots will be published in the full version.

### 2019-01-29 paper/content/perf_in_iso.tex
Some space-saving.
tweaks

@@
 
-\section{Algorithms' Performance in Isolation}\label{sec:isolation}
+\section{Performance in Isolation}\label{sec:isolation}
 
@@
     $\TS > \DEG > \DG$
-for a sufficiently long time horizon.
+for a sufficiently long time horizon.\footnote{For most of our experiments we report results for $T = 2000$ and this claim holds at this $T$}, 
 For each algorithm and each MAB instance, we compute the mean reputation score at each round, averaged over all \MRVs.

### 2019-02-07 paper/content/introduction.tex
tweaks
Minor edits in related work.

@@
 
-\xhdr{Machine learning.} Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information).
+\xhdr{Machine learning.} Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions).
 MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background.
@@
 
-\asedit{In ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}), an algorithm sets up a ``duel" between a pair of arms in each round, and only learns which arm has ``won".
-While this setting features a form of competition inside an MAB problem, it is very different from ours.}
+In ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}), an algorithm sets up a ``duel" between a pair of arms in each round, and only learns which arm has ``won".
+While this setting features a form of competition inside an MAB problem, it is very different from ours.
 
@@
     \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13},
-and repeated actions, \eg
+and repeated auctions, \eg
     \cite{Amin-auctions-nips13,Amin-auctions-nips14,Jieming-ec18}.
@@
 
-\asedit{Whether data gives competitive advantage
+Whether data gives competitive advantage
 has been studied theoretically in \cite{varian2018artificial,
@@
 We ask how additional data helps when firms simultaneously solve a machine learning problem and compete for consumers.
-Due to the competition dynamics in our model, we find that even a small amount of additional data tends to help a lot.}
-
+Due to the competition dynamics in our model, we find that even a small amount of additional data tends to help a lot.
 The first-mover advantage has been well-studied in other settings in economics and marketing, see survey \cite{kerin1992first}.

### 2019-02-07 paper/content/introduction.tex
Minor edits.

@@
 \xhdr{Machine learning.} Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions).
-MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background.
+MAB problems have been studied for many decades, see \cite{Bubeck-survey12,LS19bandit-book} for background on this immense body of work; below we only comment on the most related developments.
 We consider i.i.d.
@@
 The study of competition vs.
-exploration has been initiated in \cite{CompetingBandits-itcs16}.
+exploration has been initiated in \cite{CompetingBandits-itcs18}.
 Their model differs from ours in two key respects.

### 2019-02-07 paper/content/perf_in_iso.tex
Minor edits.

@@
 The shaded area display 95\% confidence intervals.}}
-\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area displays 95\% confidence intervals.}
+\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area shows 95\% confidence intervals.}
 \label{prelim_means}
@@
 
-Thus, Uniform and Heavy-tail instance are exploration-disadvantaged, but Needle-in-Haystack is not.
+\noindent Uniform and Heavy-tail instance are exploration-disadvantaged, but Needle-in-Haystack is not.
 

### 2019-02-07 paper/content/introduction.tex
Minor edits.

@@
 
-\gaedit{We ask whether algorithms that are traditionally considered ``better" in the literature on multi-armed bandits are incentivized to be deployed under competition.} We investigate this issue via extensive numerical experiments in a stylized duopoly model.
-In our model, two firms compete for users and simultaneously learn from them.
+We ask whether algorithms that are traditionally considered ``better" in the literature on multi-armed bandits are incentivized to be deployed under competition.
+We investigate this issue via extensive numerical experiments in a stylized duopoly model.
+
+\xhdr{Our model.} Two firms compete for users and simultaneously learn from them.
 Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm.
@@
 
-We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a \gaedit{first-mover}.
-\swcomment{confused about how they correspond to ``competition intensity''}Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``\gadelete{early entry}\gaedit{incumbent}" (the first-mover in temporary monopoly) to permanent duopoly to ``\gadelete{late entry}\gaedit{entrant" (late-arriver in temporary monopoly)}.
-\gaedit{We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
-\gaedit{Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that performs better in the long run.
-As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.}
+We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a first-mover.
+Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``incumbent" (the first-mover in temporary monopoly) to permanent duopoly to ``entrant" (late-arriver in temporary monopoly).
+
+\xhdr{Main findings.}
+We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.
+This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
+Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that performs better in the long run.
+As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.
 
@@
 This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.
-\swedit{However, the ``inverted-U relationship'' is driven by different aspects in our model than the ones in the existing literature in economics.
-In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, as opposed to the R\&D costs (which is the more standard cause in prior work)}.
+However, the ``inverted-U relationship'' is driven by different aspects in our model than the ones in the existing literature in economics.
+In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, as opposed to the R\&D costs (which is the more standard cause in prior work).
 
-\gaedit{The findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-Being first in the market in the digital economy gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").}
+\xhdr{Additional findings.}
+The findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
+Being first in the market in the digital economy gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
 
-\gaedit{Beyond the welfare implications, we also investigate the degree to which the incumbent's data advantage serves as a barrier to entry.
+Beyond the welfare implications, we also investigate the degree to which the incumbent's data advantage serves as a barrier to entry.
 We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
@@
 This effect is more pronounced when the incumbent commits to a more advanced bandit algorithm since the incumbent does not have to suffer from the reputational consequences of exploration while the entrant does.
-Our results shed lights on the importance of considering competition dynamics when reasoning about the anti-trust question of whether data can become a significant barrier to entry in an industry.}
+Our results shed lights on the importance of considering competition dynamics when reasoning about the anti-trust question of whether data can become a significant barrier to entry in an industry.
+
 \gadelete{

### 2019-02-10 paper/content/barriers.tex
add non-greedy choice model section and one line clarification in intro

@@
 
-\gadelete{
-\swedit{Our finding also highlights the amplification of the data
-  advantage through competition dynamics. In particular, in the
-  incumbency period \TS gathers data \gaedit{ through exploration } that may be hard for the entrant to gather during competition due to its potential reputational consequences \gadelete{(by engaging in exploration)}.
-This small ``head start'' goes a long way---the incumbent continues to \gadelete{attract data} \gaedit{attract users} efficiently without sacrificing much of its exploration, while the entrant tends to struggle to gather data even with limited exploration.
-}
-}
-
-\OMIT{\gaedit{Our findings imply two aspects of how and when data can serve as a barrier to entry.
-The first is that small asymmetries in data between firms can amplify into large effects through the same mechanism as the death spiral effect discussed in the previous section.
-The second is that data can serve as a larger barrier to entry when exploration is reputationally costly.
-This is due to the fact that the incumbent can explore actions in the incumbency period without immediate reputational consequences, whereas in order for the entrant to acquire the same data it needs to incur the immediate reputational consequences.
-As a result, when the incumbent does exploration in the incumbency period the value of the data accumulated is higher and it is harder for an entrant to compete in the market.
-}}
-
-\swdelete{\gaedit{How does this help us in thinking about data as a barrier to entry?
-The result that the data advantage is stronger under \TS than \DEG and \DG gives evidence that data quality, and not just quantity, matters in thinking about the competitive advantage of data.
-In particular, in the incumbency period \TS gathers data that may be hard for the entrant to gather during competition due to its potential reputational consequences (i.e.
-engaging in exploration), but which are needed in order to learn.
-If the entrant had some way of acquiring this data without incurring reputational consequences then the data advantage would be substantially weaker.
-This insight is what differentiates our results from the static viewpoint of the previous literature.
-While it remains true that, for a given action, increased data on that action has diminishing returns to scale, competitive pressures may make it harder for smaller firms to be able to engage in the exploration necessary to accumulate sufficient data on actions to learn if they are good or bad.
-Otherwise, they may be more likely to fall into the death-spiral effect discussed earlier.
-This effect arises when thinking about the issue from the perspective of competition dynamics.}}
-
 \end{document}

### 2019-02-10 paper/content/introduction.tex
add non-greedy choice model section and one line clarification in intro

@@
 For instance, Google Search and Bing compete for users in the search engine market yet at the same time need to experiment with their search and ranking algorithms to learn what works best.
+\gaedit{Bing experimenting with a ranking algorithm that ends up with a poor user experience could lead to negative short-run consequences in consumers' perceived quality of Bing and lead to more consumers choosing Google over Bing.
+However, this experimentation can help Bing eventually find a better ranking algorithm that would lead to positive long-run consequences in consumers' perceived quality of Bing and lead to more consumers choosing Bing over Google.}
 

### 2019-02-10 paper/content/inverted_u.tex
tweaks
add non-greedy choice model section and one line clarification in intro

@@
 % Thu Aug 16 13:13:01 2018
+\footnotesize
 \begin{table*}[t]
@@
 \end{tabular}
+\normalsize
 \caption{{\bf Permanent duopoly}, for Heavy-Tail and Needle-in-Haystack instances.
@@
 
+\normalsize
 Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
@@
 \begin{figure}
-\includegraphics[scale=0.3]{figures/unif_eq_welfare}
-\includegraphics[scale=0.3]{figures/ht_eq_welfare}
+\includegraphics[scale=0.25]{figures/unif_eq_welfare}
+\includegraphics[scale=0.25]{figures/ht_eq_welfare}
 \caption{Smoothed welfare plots resulting from equilibrium strategies in the different market structures.
@@
 \begin{figure}
-\includegraphics[scale=0.3]{figures/eeog_vs_num_ht}
-\includegraphics[scale=0.3]{figures/ht_many_firm_welfare}\\
+\includegraphics[scale=0.25]{figures/eeog_vs_num_ht}
+\includegraphics[scale=0.25]{figures/ht_many_firm_welfare}\\
 \caption{Average welfare and \Eeog as we increase \#firms playing \DG}

### 2019-02-10 paper/content/non_greedy_choice.tex
add non-greedy choice model section and one line clarification in intro

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Non-Greedy Choice}\label{section:non_greedy}

In this section we consider an extension of our base model where the agents' choice model is no longer greedy. We study a small deviation from greedy where the agent's choice rule is now that the agent selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the greedy choice with probability $1 - \epsilon$. A variant of this choice model is also studied as the HardMaxWithRandom choice model in \cite{mansour2018competing}.

In Section \ref{sec:competition} we saw that in the permanent duopoly case it was possible for exploration to lead to a death spiral which eventually starved the firm of agents. However, giving one firm a small head start or enough free agents via the warm start incentivized it to play \TS since it could recover the reputation costs it incurred from early exploration. An interpretation of the non-greedy choice model is that instead of concentrating the free agents as coming in the beginning of the game instead they are dispersed throughout the game.

\footnotesize
\begin{table*}[t]
\centering
\begin{tabular}{|c|c|c|c||c|c|c|}
  \hline
  & \multicolumn{3}{c||}{Heavy-Tail (Non-Greedy)}
  & \multicolumn{3}{c|}{Heavy-Tail (Greedy)}\\
  \hline
  & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG 
 & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG  \\
  \hline
$t = 2000$
 & \makecell{ \textbf{0.43} $\pm$ 0.02 \\Var: 0.15 } 
  & \makecell{ \textbf{0.44} $\pm$ 0.02 \\Var: 0.15 } 
  & \makecell{ \textbf{0.6} $\pm$ 0.02 \\Var: 0.1 }
  %%
 &  \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
  & \makecell{ \textbf{0.28} $\pm$ 0.03 \\Var: 0.19 } 
  & \makecell{ \textbf{0.63} $\pm$ 0.03 \\Var: 0.18 }
    \\
\hline
  $t = 5000$ 
   & \makecell{ \textbf{0.66} $\pm$ 0.01 \\Var: 0.056 } 
  & \makecell{ \textbf{0.59} $\pm$ 0.02 \\Var: 0.092 } 
  & \makecell{ \textbf{0.56} $\pm$ 0.02 \\Var: 0.098 } 
  %%
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
 & \makecell{ \textbf{0.62} $\pm$ 0.03 \\Var: 0.19 }
 \\ 
  \hline
  $t = 10000$
  & \makecell{ \textbf{0.76} $\pm$ 0.01 \\Var: 0.026 } 
 & \makecell{ \textbf{0.67} $\pm$ 0.02 \\Var: 0.067 } 
 & \makecell{ \textbf{0.52} $\pm$ 0.02 \\Var: 0.11 }
 %%
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.21 } 
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.2 } 
  & \makecell{ \textbf{0.6} $\pm$ 0.03 \\Var: 0.2 }
  %%
  \\
   \hline
\end{tabular}
\normalsize
\caption{Performance Comparison between Greedy and Non-Greedy Choice on the Heavy-Tail MAB instance. Each cell describes the market shares in a game between two algorithms, call them Alg1 vs. Alg2, at a particular value of $t$. Line 1 in the cell is the market share of Alg 1: the average (in bold) and the 95\% confidence band.
%For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\epsilon = 0.1$ in the Non-Greedy choice model and $T_0 = 20$.}
\label{tab:non_greedy_table}
\end{table*}

\normalsize

\begin{finding}\label{find:non_greedy_choice}
\textit{Under the non-greedy choice model, \TS is incentivized for sufficiently large $T$}
\end{finding}

Table \ref{tab:non_greedy_table} shows the average market shares under the Greedy vs Non-Greedy choice model. As $t$ gets sufficiently large, \TS becomes incentivized for the Non-Greedy choice model whereas, as expected, it does not for the Greedy choice model. These results hold across all problem instances.\footnote{The results here are pulled using different \MRV and realizations than the results pulled previously (due to the larger $T$). However, they are drawn from the same prior instances and so qualitatively are the same but the quantitative results are not directly comparable to those from the previous section.} The intuition for this is simple. The existence of the random agents ensures that no algorithm ever gets completely starved of agents. As a result, with sufficiently large $T$, the algorithms with better asymptotic performance in isolation according to relative reputation will win since they will eventually get enough random agents to achieve a higher reputation than the other algorithms.

While we report the results for $\epsilon = 0.1$ the results qualitatively should not depend on the parameterization of $\epsilon$. The $\epsilon$ parameter determines the rate at which a firm will get agents for free, which should primarily affect how quickly the algorithm converges to its asymptotic performance. As a result, we expect that the value of $T$ where Finding \ref{find:non_greedy_choice} holds is decreasing in $\epsilon$.

Finding \ref{find:non_greedy_choice} implies that we can re-interpret the inverted-U findings from before in terms of the number of agents that a firm receives without having to worry about incentives. In the extreme when the firm gets all agents for free as in the monopoly case then it is incentivized to play \DG. When it only gets some of the agents for free, either via a large warm start, a temporary monopoly, or non-greedy consumer choice, then \TS is incentivized. However, if the number of free agents gets small enough then \DG is incentivized as in the permanent duopoly analysis from before.

Table \ref{tab:non_greedy_table} also shows that another difference between the Greedy and Non-Greedy choice model is that the Non-Greedy choice model leads to lower variance in market shares across simulations.

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../competing_bandits"
%%% End:

### 2019-02-10 paper/content/non_greedy_choice.tex
fixup in terminology

@@
   \hline
-$t = 2000$
+$T = 2000$
  & \makecell{ \textbf{0.43} $\pm$ 0.02 \\Var: 0.15 } 
@@
 \hline
-  $t = 5000$ 
+  $T= 5000$ 
    & \makecell{ \textbf{0.66} $\pm$ 0.01 \\Var: 0.056 } 
@@
   \hline
-  $t = 10000$
+  $T = 10000$
   & \makecell{ \textbf{0.76} $\pm$ 0.01 \\Var: 0.026 } 
@@
 \begin{finding}\label{find:non_greedy_choice}
-\textit{Under the non-greedy choice model, \TS is incentivized for sufficiently large $T$}
+\textit{Under the non-greedy choice model, \TS is weakly dominant as long as $T$ is sufficiently large}
 \end{finding}
@@
 Table \ref{tab:non_greedy_table} shows the average market shares under the Greedy vs Non-Greedy choice model.
-As $t$ gets sufficiently large, \TS becomes incentivized for the Non-Greedy choice model whereas, as expected, it does not for the Greedy choice model.
+As $T$ gets sufficiently large, \TS becomes weakly dominant for the Non-Greedy choice model whereas, as expected, it does not for the Greedy choice model.
 These results hold across all problem instances.\footnote{The results here are pulled using different \MRV and realizations than the results pulled previously (due to the larger $T$).

### 2019-02-14 paper/content/non_greedy_choice.tex
update section labels and fixup appendix
updates

@@
 
-\section{Non-Greedy Choice}\label{section:non_greedy}
+\section{Non-deterministic Choice Models}\label{sec:non_greedy}
 
-In this section we consider an extension of our base model where the agents' choice model is no longer greedy.
-We study a small deviation from greedy where the agent's choice rule is now that the agent selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the greedy choice with probability $1 - \epsilon$.
-A variant of this choice model is also studied as the HardMaxWithRandom choice model in \cite{mansour2018competing}.
+In this section we consider an extension of our base model where the agents' choice model is no longer deterministic.
+Recall that in our baseline model agents deterministically choose the firm with the highest reputation score.
+We define this choice model as \textit{HardMax}.
+We study a small deviation from this choice model where the agent now selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the firm with the highest reputation score with probability $1 - \epsilon$.
+Following \cite{mansour2018competing} we define this choice model as \textit{HardMaxWithRandom}.
 
@@
 However, giving one firm a small head start or enough free agents via the warm start incentivized it to play \TS since it could recover the reputation costs it incurred from early exploration.
-An interpretation of the non-greedy choice model is that instead of concentrating the free agents as coming in the beginning of the game instead they are dispersed throughout the game.
+An interpretation of HardMaxWithRandom is that instead of concentrating the free agents as arriving in the beginning of the game instead they are dispersed throughout the game.
 
@@
   \hline
-  & \multicolumn{3}{c||}{Heavy-Tail (Non-Greedy)}
-  & \multicolumn{3}{c|}{Heavy-Tail (Greedy)}\\
+  & \multicolumn{3}{c||}{Heavy-Tail (HardMaxWithRandom)}
+  & \multicolumn{3}{c|}{Heavy-Tail (HardMax)}\\
   \hline
@@
 \normalsize
-\caption{Performance Comparison between Greedy and Non-Greedy Choice on the Heavy-Tail MAB instance.
+\caption{Performance Comparison between HardMax and HardMaxWithRandom Choice on the Heavy-Tail MAB instance.
 Each cell describes the market shares in a game between two algorithms, call them Alg1 vs.
@@
 %For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
-Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\epsilon = 0.1$ in the Non-Greedy choice model and $T_0 = 20$.}
+Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\epsilon = 0.1$ in the HardMaxWithRandom choice model and $T_0 = 20$.}
 \label{tab:non_greedy_table}
@@
 
-Table \ref{tab:non_greedy_table} shows the average market shares under the Greedy vs Non-Greedy choice model.
-As $T$ gets sufficiently large, \TS becomes weakly dominant for the Non-Greedy choice model whereas, as expected, it does not for the Greedy choice model.
+Table \ref{tab:non_greedy_table} shows the average market shares under the HardMax vs HardMaxWithRandom choice model.
+As $T$ gets sufficiently large, \TS becomes weakly dominant for the HardMaxWithRandom model whereas, as expected, it does not for the HardMax model.
 These results hold across all problem instances.\footnote{The results here are pulled using different \MRV and realizations than the results pulled previously (due to the larger $T$).
-However, they are drawn from the same prior instances and so qualitatively are the same but the quantitative results are not directly comparable to those from the previous section.} The intuition for this is simple.
+However, they are drawn from the same prior instances and so qualitatively are the same but the quantitative results are not directly comparable to those from the previous section.} The intuition for why \TS should become the dominant strategy eventually is simple.
 The existence of the random agents ensures that no algorithm ever gets completely starved of agents.
-As a result, with sufficiently large $T$, the algorithms with better asymptotic performance in isolation according to relative reputation will win since they will eventually get enough random agents to achieve a higher reputation than the other algorithms.
+As a result, each algorithm should eventually converge to its asymptotic performance in isolation.
 
-While we report the results for $\epsilon = 0.1$ the results qualitatively should not depend on the parameterization of $\epsilon$.
-The $\epsilon$ parameter determines the rate at which a firm will get agents for free, which should primarily affect how quickly the algorithm converges to its asymptotic performance.
-As a result, we expect that the value of $T$ where Finding \ref{find:non_greedy_choice} holds is decreasing in $\epsilon$.
+However, it takes a significant amount of randomness and a relatively large time horizon for this convergence to occur.
+Even with $T = 10000$ and $\epsilon = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance.
+Additionally Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that HardMaxWithRandom leads to lower variance in market shares across simulations compared to HardMax.
 
@@
 In the extreme when the firm gets all agents for free as in the monopoly case then it is incentivized to play \DG.
-When it only gets some of the agents for free, either via a large warm start, a temporary monopoly, or non-greedy consumer choice, then \TS is incentivized.
+When it only gets some of the agents for free, either via a large warm start, a temporary monopoly, or non-deterministic choice, then \TS is incentivized.
 However, if the number of free agents gets small enough then \DG is incentivized as in the permanent duopoly analysis from before.
-
-Table \ref{tab:non_greedy_table} also shows that another difference between the Greedy and Non-Greedy choice model is that the Non-Greedy choice model leads to lower variance in market shares across simulations.
 

### 2019-02-14 paper/content/non_greedy_choice.tex
fixup the finding
tweak

@@
 \begin{finding}\label{find:non_greedy_choice}
-\textit{Under the non-greedy choice model, \TS is weakly dominant as long as $T$ is sufficiently large}
+\textit{Under the HardMaxWithRandom choice model \TS is weakly dominant as long as $T$ is sufficiently large. However, both $T$ and $\epsilon$ have to be relatively large for this to occur.}
 \end{finding}
@@
 Even with $T = 10000$ and $\epsilon = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance.
-Additionally Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that HardMaxWithRandom leads to lower variance in market shares across simulations compared to HardMax.
+Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that HardMaxWithRandom leads to lower variance in market shares across simulations compared to HardMax.
 

### 2019-02-14 paper/content/non_greedy_choice.tex
macros

@@
 In this section we consider an extension of our base model where the agents' choice model is no longer deterministic.
-Recall that in our baseline model agents deterministically choose the firm with the highest reputation score.
+Recall that in our base model agents deterministically choose the firm with the higher reputation score.
 We define this choice model as \textit{HardMax}.
-We study a small deviation from this choice model where the agent now selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the firm with the highest reputation score with probability $1 - \epsilon$.
-Following \cite{mansour2018competing} we define this choice model as \textit{HardMaxWithRandom}.
+We study a small deviation from this choice model where the agent now selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the firm with the higher reputation score with probability $1 - \epsilon$.
+Following \cite{mansour2018competing} we define this choice model as \textit{HardMax with random} (\HMR).
 
@@
 However, giving one firm a small head start or enough free agents via the warm start incentivized it to play \TS since it could recover the reputation costs it incurred from early exploration.
-An interpretation of HardMaxWithRandom is that instead of concentrating the free agents as arriving in the beginning of the game instead they are dispersed throughout the game.
+An interpretation of \HMR is that instead of concentrating the free agents as arriving in the beginning of the game instead they are dispersed throughout the game.
 
@@
   \hline
-  & \multicolumn{3}{c||}{Heavy-Tail (HardMaxWithRandom)}
+  & \multicolumn{3}{c||}{Heavy-Tail (\HMR)}
   & \multicolumn{3}{c|}{Heavy-Tail (HardMax)}\\
@@
 \normalsize
-\caption{Performance Comparison between HardMax and HardMaxWithRandom Choice on the Heavy-Tail MAB instance.
+\caption{Performance Comparison between HardMax and \HMR Choice on the Heavy-Tail MAB instance.
 Each cell describes the market shares in a game between two algorithms, call them Alg1 vs.
@@
 %For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
-Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\epsilon = 0.1$ in the HardMaxWithRandom choice model and $T_0 = 20$.}
+Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\epsilon = 0.1$ in the \HMR choice model and $T_0 = 20$.}
 \label{tab:non_greedy_table}
@@
 \begin{finding}\label{find:non_greedy_choice}
-\textit{Under the HardMaxWithRandom choice model \TS is weakly dominant as long as $T$ is sufficiently large. However, both $T$ and $\epsilon$ have to be relatively large for this to occur.}
+\textit{Under the \HMR choice model \TS is weakly dominant as long as $T$ is sufficiently large. However, both $T$ and $\epsilon$ have to be relatively large for this to occur.}
 \end{finding}
 
-Table \ref{tab:non_greedy_table} shows the average market shares under the HardMax vs HardMaxWithRandom choice model.
-As $T$ gets sufficiently large, \TS becomes weakly dominant for the HardMaxWithRandom model whereas, as expected, it does not for the HardMax model.
+Table \ref{tab:non_greedy_table} shows the average market shares under the HardMax vs \HMR choice model.
+As $T$ gets sufficiently large, \TS becomes weakly dominant for the \HMR model whereas, as expected, it does not for the HardMax model.
 These results hold across all problem instances.\footnote{The results here are pulled using different \MRV and realizations than the results pulled previously (due to the larger $T$).
@@
 Even with $T = 10000$ and $\epsilon = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance.
-Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that HardMaxWithRandom leads to lower variance in market shares across simulations compared to HardMax.
+Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that \HMR leads to lower variance in market shares across simulations compared to HardMax.
 

### 2019-02-14 paper/content/non_greedy_choice.tex
add HardMax macro

@@
 Recall that in our base model agents deterministically choose the firm with the higher reputation score.
-We define this choice model as \textit{HardMax}.
+We define this choice model as \textit{HardMax} (\HM).
 We study a small deviation from this choice model where the agent now selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the firm with the higher reputation score with probability $1 - \epsilon$.
@@
   & \multicolumn{3}{c||}{Heavy-Tail (\HMR)}
-  & \multicolumn{3}{c|}{Heavy-Tail (HardMax)}\\
+  & \multicolumn{3}{c|}{Heavy-Tail (\HM)}\\
   \hline
@@
 \normalsize
-\caption{Performance Comparison between HardMax and \HMR Choice on the Heavy-Tail MAB instance.
+\caption{Performance Comparison between \HM and \HMR Choice on the Heavy-Tail MAB instance.
 Each cell describes the market shares in a game between two algorithms, call them Alg1 vs.
@@
 
-Table \ref{tab:non_greedy_table} shows the average market shares under the HardMax vs \HMR choice model.
-As $T$ gets sufficiently large, \TS becomes weakly dominant for the \HMR model whereas, as expected, it does not for the HardMax model.
+Table \ref{tab:non_greedy_table} shows the average market shares under the \HM vs \HMR choice model.
+As $T$ gets sufficiently large, \TS becomes weakly dominant for the \HMR model whereas, as expected, it does not for the \HM model.
 These results hold across all problem instances.\footnote{The results here are pulled using different \MRV and realizations than the results pulled previously (due to the larger $T$).
@@
 Even with $T = 10000$ and $\epsilon = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance.
-Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that \HMR leads to lower variance in market shares across simulations compared to HardMax.
+Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that \HMR leads to lower variance in market shares across simulations compared to \HM.
 

### 2019-02-14 paper/content/introduction.tex
update section labels and fixup appendix
Full pass on the intro.

--- 2019-02-07 paper/content/introduction.tex
@@
 
-\section{Introduction}\label{section:1}
+\section{Introduction}\label{sec:intro}
 
 Many modern online platforms simultaneously compete for users as well as learn from the users they manage to attract.
-This creates a tradeoff between \textit{exploration} and \textit{competition}: firms experiment with potentially sub-optimal options for the sake of gaining information to make better decisions tomorrow, while they need to incentivize consumers to select them over their competitors today.
+This creates a tension between \textit{exploration} and \textit{competition}: firms experiment with potentially sub-optimal options for the sake of gaining information to make better decisions tomorrow, while they need to incentivize consumers to select them over their competitors today.
 For instance, Google Search and Bing compete for users in the search engine market yet at the same time need to experiment with their search and ranking algorithms to learn what works best.
+Similar exploration vs.
+competition tension arises in other application domains: recommendation systems, news and entertainment websites, online commerce, and so forth.
 
-Platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
+\OMIT{\gaedit{Bing experimenting with a ranking algorithm that ends up with a poor user experience could lead to negative short-run consequences in consumers' perceived quality of Bing and lead to more consumers choosing Google over Bing.
+However, this experimentation can help Bing eventually find a better ranking algorithm that would lead to positive long-run consequences in consumers' perceived quality of Bing and lead to more consumers choosing Bing over Google.}}
+
+Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for making decisions under uncertainty and trading off exploration and exploitation (making good near-term decisions).
 While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition, even putting aside the deployment/maintenance costs.
@@
 
-We ask whether algorithms that are traditionally considered ``better" in the literature on multi-armed bandits are incentivized to be deployed under competition.
-We investigate this issue via extensive numerical experiments in a stylized duopoly model.
+In this paper, we ask how the interplay of exploration and competition affects platforms' incentives.
+While some bandit algorithms are traditionally considered ``better" than others in the literature, {\bf\em does competition incentivize the adoption of the better algorithms}?
+How is this affected by the intensity of competition?
+We investigate these issues via extensive numerical experiments in a stylized duopoly model.
 
-\xhdr{Our model.} Two firms compete for users and simultaneously learn from them.
+\xhdr{Our model.} We consider two firms that compete for users and simultaneously learn from them.
 Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm.
 Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window.
-Each firm's objective is to maximize its \gaedit{expected} market share (the fraction of users choosing this firm).
+Each firm's objective is to maximize its market share (the fraction of users choosing this firm).
 
@@
 
+We focus on three classes of bandit algorithms, ranging from more primitive to more sophisticated: \emph{greedy algorithms} that do not explicitly explore, algorithms that separate exploration and exploitation, and algorithms that combine the two.
+We know from prior work that in the absence of competition,  ``better" algorithms are better in the long run, but could be worse initially.
+
 \xhdr{Main findings.}
-We find that, even in the permanent duopoly, in equilibrium competition incentivizes firms to play a ``greedy algorithm" that does not explicitly explore and even more so if the firm is a late arriver in a market.
+We find that in the permanent duopoly, competition incentivizes firms to choose the ``greedy algorithm", and even more so if the firm is a late arriver in a market.
 This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
@@
 
-Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of an ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
+Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
 This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.
@@
 \xhdr{Additional findings.}
-The findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-Being first in the market in the digital economy gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
+We investigate the ``first-mover advantage" phenomenon in more detail.
+Being first in the market gives free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
+We run additional experiments so as to isolate these two effects.
+We find that the data advantage alone can lead to a significant advantage under competition.
+In fact, even a small ``head start'' can get amplified into a large incumbency advantage under competition.
+This effect is most pronounced when the incumbent commits to a more advanced bandit algorithm.
+These findings suggest that competition dynamics are pertinent to the anti-trust considerations of data as a barrier to entry.
 
-Beyond the welfare implications, we also investigate the degree to which the incumbent's data advantage serves as a barrier to entry.
-We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
-We find that the data advantage from the incumbent's ``small head start'' can lead to a significant incumbency advantage, especially when the incumbent commits to a more advanced bandit algorithm.
-Our results show that even a small degree of data asymmetry can be amplified to be a larger market advantage through competition dynamics.
-This effect is more pronounced when the incumbent commits to a more advanced bandit algorithm since the incumbent does not have to suffer from the reputational consequences of exploration while the entrant does.
-Our results shed lights on the importance of considering competition dynamics when reasoning about the anti-trust question of whether data can become a significant barrier to entry in an industry.
-
-\gadelete{
+\OMIT{ %%%%%%
 \swedit{Beyond the welfare implications, we also investigate the barrier to entry that follows from the incumbent's data advantage.
@@
 \gadelete{This is in contrast to the data advantage in a static environment such as discussed in \cite{varian2018artificial,bajari2018impact}, in which a small amount of additional data can only provide limited improvement in the predictive accuracy.} The effect is stronger when the incumbent commits to a more advanced bandit algorithm due to
-
+}} %%%%%%
 % in a competitive environment the relative differences in the data
@@
 % larger reputational consequences.
-}\swcomment{still not super happy about above.} }
 
@@
 We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition.
-In particular, we find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor.
-We suggest a more refined performance measure, and use it to explain (and frame) some of the competition outcomes.
+We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor.
+We suggest a
+more refined performance measure, and use it to explain some of the competition outcomes.
+
+While reputation scores already introduce some noise and randomness into users' choices, we consider an alternative choice model in which the noise/randomness is explicit: a small fraction of users choose a firm uniformly at random.
+We confirm the theoretical intuition that better algorithms prevail if the parameters -- the ``random fraction" and the time horizon -- are sufficiently large.
+However, we find that this effect is negligible for some smaller but still ``relevant" parameter values.
+
+\xhdr{Discussion.}
+The present paper is an experimental counterpart to our prior work \cite{CompetingBandits-itcs18}, which considered a similar duopoly model and obtained a number of theoretical results with ``asymptotic" flavor.
+For the sake of analytical tractability, that paper makes a somewhat unrealistic simplification that users do not observe any signals about firms' ongoing performance (instead, users choose between firms according to the firms' Bayesian-expected rewards).
+The high-level conclusion in \cite{CompetingBandits-itcs18} is an inverted-U relationship similar to ours, but the strength of competition is varied in a different way, using assumptions about (ir)rational user behavior.
+
+The present paper provides a more nuanced and ``non-asymptotic" perspective.
+In essence, we look for substantial effects within relevant time scales.
+Indeed, we start our investigation by determining what time scales are relevant in the context of our model.
+The reputation-based choice model accounts for competition in a more direct way, allows to separate reputation vs.
+data advantage, and makes our model amenable to numerical simulations (unlike the model in \cite{CompetingBandits-itcs18}).
+
+To elucidate the interplay of competition and exploration, our model is stylized in several important  respects, some of which we discuss below.
+Firms compete only on the quality of service, rather than, say, pricing or the range of products.
+Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores", which further simplified to average performance over a time window.
+On the machine learning side, our setup is geared to distinguish between ``simple" vs.
+``better" vs.
+``smart" bandit algorithms; we are not interested in state-of-art algorithms for very realistic bandit settings.
 
@@
 
-\xhdr{Machine learning.} Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions).
-MAB problems have been studied for many decades, see \cite{Bubeck-survey12,LS19bandit-book} for background on this immense body of work; below we only comment on the most related developments.
-We consider i.i.d.
+\xhdr{Machine learning.} Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and exploitation.
+MAB problems have been studied for many decades, see \cite{Bubeck-survey12,LS19bandit-book} for background on this immense body of work; we only comment on the most related aspects.
+
+We consider MAB with i.i.d.
 rewards, a well-studied and well-understood MAB model \cite{bandits-ucb1}.
@@
 
-The study of competition vs.
-exploration has been initiated in \cite{CompetingBandits-itcs18}.
-Their model differs from ours in two key respects.
+The study of competition vs. exploration has been initiated in \cite{CompetingBandits-itcs18}, as discussed above. 
+\OMIT{Their model differs from ours in two key respects.
 First, users do not see any signal about firms' past performance, and instead choose between firms according to the Bayesian-expected reward.
@@
 Their results are purely theoretical; their model is amenable to proofs but not to numerical experiments.
-Their high-level conclusion is an inverted-U relationship between competition and innovation that is similar to ours.
-
+Their high-level conclusion is an inverted-U relationship between competition and innovation that is similar to ours.}
 Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.

### 2019-02-14 paper/content/model.tex
Minor edits in prelims.

@@
 Each algorithm receives a ``warm start": additional $T_0$ agents that arrive before the game starts, and interact with the firm as described above.
-\gaedit{The warm start serves to vary the initial information the firm has before beginning to compete for users as well as to initialize the agents' perception of the firms.} Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
+\asedit{The warm ensures that each firm has a meaningful reputation when competition starts.} Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
 
@@
 \xhdr{Agents.}
-\gaedit{For simplicity we suppose that firms compete for agents on a single dimension, quality, that is summarized via the rewards experienced by past consumers.
-\OMIT{\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}}}
+\asedit{Firms compete on a single dimension, quality of service, as expressed by agents' rewards.}
+\OMIT{\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}}
 Agents are myopic and non-strategic: they \asedit{would like} to choose among the firms so as to maximize their expected reward \gaedit{(i.e.
@@
 The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
-\asedit{This is similar to ratings aggregators on many online platforms such as Yelp, Tripadvisor, etc.}
+
+%\asedit{This is similar to ratings aggregators on many online platforms such as Yelp, Tripadvisor, etc.}
 

### 2019-02-14 paper/content/introduction.tex
Removed "our prior work".

@@
 \xhdr{Discussion.}
-The present paper is an experimental counterpart to our prior work \cite{CompetingBandits-itcs18}, which considered a similar duopoly model and obtained a number of theoretical results with ``asymptotic" flavor.
+The present paper is an experimental counterpart to \cite{CompetingBandits-itcs18}, which considered a similar duopoly model and obtained a number of theoretical results with ``asymptotic" flavor.
 For the sake of analytical tractability, that paper makes a somewhat unrealistic simplification that users do not observe any signals about firms' ongoing performance (instead, users choose between firms according to the firms' Bayesian-expected rewards).

### 2019-02-14 paper/content/introduction.tex
In intro/discussion, added a para on challenges.

@@
 
+Even with a stylized model, numerical investigation is quite challenging.
+An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given instance of a multi-armed bandit problem.%
+\footnote{Each such experiment is run many times to reduce variance.}
+Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit instances\}, and we are looking for findings that are consistent across this entire space.
+It is essential to keep each of the three dimensions small yet representative.
+In particular, we need to capture a huge variety of bandit instances with only a few representative examples.
+Further, one needs succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).
+
 \subsection{Related work}

### 2019-02-14 paper/content/perf_in_iso.tex
In "isol", added a sentence re choice of T.

--- 2018-09-06 paper/content/perf_in_iso.tex
@@
 The shaded area display 95\% confidence intervals.}}
-\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area displays 95\% confidence intervals.}
+\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area shows 95\% confidence intervals.}
 \label{prelim_means}
 \end{figure}
+
+\asedit{We also use Figure~\ref{prelim_means} to choose a reasonable time-horizon for the subsequent experiments, as $T=2000$.
+The idea is, we want $T$ to be large enough so that algorithms performance starts to plateau, but small enough such that algorithms are still learning.}
 

### 2019-02-14 paper/content/non_greedy_choice.tex
commit

--- 2019-02-14 paper/content/non_greedy_choice.tex
@@
 
-In Section \ref{sec:competition} we saw that in the permanent duopoly case it was possible for exploration to lead to a death spiral which eventually starved the firm of agents.
-However, giving one firm a small head start or enough free agents via the warm start incentivized it to play \TS since it could recover the reputation costs it incurred from early exploration.
-An interpretation of \HMR is that instead of concentrating the free agents as arriving in the beginning of the game instead they are dispersed throughout the game.
+In Section \ref{sec:competition} we show that in the permanent duopoly case exploration can lead to a death spiral which eventually starved the firm of agents.
+However, giving one firm a small head start or enough free agents via the warm start incentivizes it to play \TS since it could recover the reputation costs it incurred from early exploration.
+An interpretation of \HMR~ is that instead of concentrating the free agents as arriving in the beginning of the game instead they are dispersed throughout the game.
 
@@
 \begin{finding}\label{find:non_greedy_choice}
-\textit{Under the \HMR choice model \TS is weakly dominant as long as $T$ is sufficiently large. However, both $T$ and $\epsilon$ have to be relatively large for this to occur.}
+\textit{Under the \HMR~ choice model \TS is weakly dominant as long as $T$ is sufficiently large. \swdelete{However, both $T$ and $\epsilon$ have to be relatively large for this to occur.}}
 \end{finding}
 
-Table \ref{tab:non_greedy_table} shows the average market shares under the HardMax vs \HMR choice model.
+Table \ref{tab:non_greedy_table} shows the average market shares under the \HM~ vs \HMR~ choice model.
 As $T$ gets sufficiently large, \TS becomes weakly dominant for the \HMR model whereas, as expected, it does not for the HardMax model.

### 2019-02-14 paper/content/model.tex

@@
 
-\asedit{\xhdr{Consistency.}}
-While we experiment with various MAB instances and parameter settings, in the main paper we only report plots for selected, representative experiments.
-Additional plots will be published in the full version.
-Unless noted otherwise, our findings are based on and consistent with all these plots.
+\xhdr{Consistency.}
+While we experiment with various MAB instances and parameter settings, we only report on selected, representative experiments in the body of the paper.
+Additional plots and tables are provided in the appendix.
+Unless noted otherwise, our findings are based on and consistent with all these experiments.
 

### 2019-02-14 paper/content/non_greedy_choice.tex
stuff

@@
 
-In Section \ref{sec:competition} we show that in the permanent duopoly case exploration can lead to a death spiral which eventually starved the firm of agents.
-However, giving one firm a small head start or enough free agents via the warm start incentivizes it to play \TS since it could recover the reputation costs it incurred from early exploration.
-An interpretation of \HMR~ is that instead of concentrating the free agents as arriving in the beginning of the game instead they are dispersed throughout the game.
+In Section \ref{sec:competition} we show that in the permanent duopoly
+case exploration can lead to a death spiral which eventually starves
+the firm of agents. However, giving one firm a small head start or
+enough free agents via the warm start incentivizes it to play \TS
+since it could recover the reputation costs it incurred from early
+exploration. An interpretation of \HMR~ is that instead of
+concentrating the free agents as arriving in the beginning of the game
+instead they are dispersed throughout the game.
 
@@
 
-Table \ref{tab:non_greedy_table} shows the average market shares under the \HM~ vs \HMR~ choice model.
-As $T$ gets sufficiently large, \TS becomes weakly dominant for the \HMR model whereas, as expected, it does not for the HardMax model.
-These results hold across all problem instances.\footnote{The results here are pulled using different \MRV and realizations than the results pulled previously (due to the larger $T$).
-However, they are drawn from the same prior instances and so qualitatively are the same but the quantitative results are not directly comparable to those from the previous section.} The intuition for why \TS should become the dominant strategy eventually is simple.
-The existence of the random agents ensures that no algorithm ever gets completely starved of agents.
-As a result, each algorithm should eventually converge to its asymptotic performance in isolation.
+Table \ref{tab:non_greedy_table} shows the average market shares under
+the \HM~ vs \HMR~ choice model. \swedit{In contrast to the \HM~ model,
+  \TS becomes weakly dominant under the \HMR model, as $T$ gets
+  sufficiently large.} These results hold across all problem
+instances.\footnote{The results here are pulled using different \MRV
+  and realizations than the results pulled previously (due to the
+  larger $T$). However, they are drawn from the same prior instances
+  and so qualitatively are the same but the quantitative results are
+  not directly comparable to those from the previous
+  section.\swcomment{not sure we need this}} The intuition for why \TS
+should become the dominant strategy eventually is simple. The
+consistent stream of random agents ensures that each principal is
+chosen at least $\Omega(\epsilon t)$ times at every time step $t$. As
+a result, each algorithm should eventually converge to its asymptotic
+performance in isolation.
 
@@
 When it only gets some of the agents for free, either via a large warm start, a temporary monopoly, or non-deterministic choice, then \TS is incentivized.
-However, if the number of free agents gets small enough then \DG is incentivized as in the permanent duopoly analysis from before.
+However, if the number of free agents gets small enough then \DG is incentivized as in the permanent duopoly analysis from before.\swcomment{I don't understand this paragraph; I think it's too vague.
+Perhaps remove?}
 

### 2019-02-14 paper/competing_bandits.tex
tweaks
add non-greedy choice model section and one line clarification in intro
update section labels and fixup appendix
updates
add HardMax macro
commit
fix
stuff
edits in the abtract

@@
 
+\newcommand{\HMR}{\term{HMR}}
+\newcommand{\HM}{\term{HM}}
 % a very useful package for edits and comments, from David Kempe (USC)
 %\usepackage{color-edits}
-\usepackage[suppress]{color-edits}  % use this to suppress the package
+\usepackage{color-edits}  % use this to suppress the package
 \addauthor{as}{red}    % as for Alex
@@
 \begin{abstract}
-We empirically study the interplay between \textit{exploration} and \textit{competition}.
-Systems that learn from interactions with users often engage in \emph{exploration}: making potentially suboptimal decisions in order to acquire new information for future decisions.
-However, exploration may hurt system's reputation in the near term, with adverse competitive effects.
-In particular, a system may enter a ``death spiral" when decreasing market share leaves the system with less users to learn from, which degrades system's performance relative to competition and further decreases the market share.
+  We empirically study the interplay between \textit{exploration} and
+  \textit{competition}. Systems that learn from interactions with
+  users often engage in \emph{exploration}: making potentially
+  suboptimal decisions in order to acquire new information for future
+  decisions. However, \swedit{when multiple systems are competing for
+    the same market of users together, exploration may hurt a system's
+    reputation in the near term, resulting in potential adverse
+    ``death spiral'' effects: the short-term reputation cost decreases
+    the number of users for the system to learn from, which degrades
+    system's performance relative to competition and further decreases
+    the market share.}
 
@@
 %
-
+\maketitle
 \keywords{Multi-armed bandits,Exploration,Competition,Competition vs Innovation}
 
-\maketitle
+\newpage
+
 \subfile{content/introduction}
@@
 
+\subfile{content/non_greedy_choice}
+
 \subfile{content/conclusion}
-
-\subfile{content/appendix_for_one_version}
 
@@
 
+\subfile{content/appendix_for_one_version}
+
 \end{document} 
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2019-02-14 paper/content/introduction.tex
edits in the abtract
light edits on intro

@@
 
+We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a first-mover.
+Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``incumbent" (the first-mover in temporary monopoly) to permanent duopoly to ``entrant" (late-arriver in temporary monopoly).
+
+We focus on three classes of bandit algorithms, ranging from more primitive to more sophisticated: \emph{greedy algorithms} that do not explicitly explore, algorithms that separate exploration and exploitation, and algorithms that combine the two.
+We know from prior work that in the absence of competition,  ``better" algorithms are better in the long run, but could be worse initially.
+
+\xhdr{Main findings.}
+We find that in the permanent duopoly, competition incentivizes firms to choose the ``greedy algorithm", and even more so if the firm is a late arriver in a market.
+This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
+Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that performs better in the long run.
+As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.\swcomment{why ``even''?}
+
+Interpreting the adoption of better algorithms as ``innovation", our
+findings can be framed in terms of the ``inverted-U relationship"
+between competition and innovation (see Figure~\ref{fig:inverted-U}),
+\swedit{a well-established concept in the economics literature, dating
+  back to \cite{Schumpeter-42}.} \swedit{In this relationship, too
+  little or too much competition is bad for innovation, but
+  intermediate levels of competition tend to be better.} However, the
+``inverted-U relationship'' is driven by different aspects in our
+model than the ones in the existing literature in economics. In our
+case, the barriers for innovations arise entirely from the
+reputational consequences of exploration in competition, as opposed to
+the R\&D costs (which is the more standard cause in prior work).
+
 \tikzstyle{level 1}=[level distance=3.5cm, sibling distance=4.0cm]
@@
 
-\begin{figure}
+\begin{figure}[t]
 \begin{center}
@@
 
-We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a first-mover.
-Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``incumbent" (the first-mover in temporary monopoly) to permanent duopoly to ``entrant" (late-arriver in temporary monopoly).
-
-We focus on three classes of bandit algorithms, ranging from more primitive to more sophisticated: \emph{greedy algorithms} that do not explicitly explore, algorithms that separate exploration and exploitation, and algorithms that combine the two.
-We know from prior work that in the absence of competition,  ``better" algorithms are better in the long run, but could be worse initially.
-
-\xhdr{Main findings.}
-We find that in the permanent duopoly, competition incentivizes firms to choose the ``greedy algorithm", and even more so if the firm is a late arriver in a market.
-This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
-Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm that performs better in the long run.
-As a result, we find that consumer welfare is highest under temporary monopoly and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.
-
-Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
-This relationship -- too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better -- is a familiar theme in the economics literature, dating back to \cite{Schumpeter-42}.
-However, the ``inverted-U relationship'' is driven by different aspects in our model than the ones in the existing literature in economics.
-In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, as opposed to the R\&D costs (which is the more standard cause in prior work).
-
 \xhdr{Additional findings.}
@@
 These findings suggest that competition dynamics are pertinent to the anti-trust considerations of data as a barrier to entry.
-
+\swcomment{should we say ``reputation advantage'' alone does not give much advantage?}
 \OMIT{ %%%%%%

### 2019-02-14 paper/content/non_greedy_choice.tex
removed the last paragraph

@@
 
-Finding \ref{find:non_greedy_choice} implies that we can re-interpret the inverted-U findings from before in terms of the number of agents that a firm receives without having to worry about incentives.
+\OMIT{Finding \ref{find:non_greedy_choice} implies that we can re-interpret the inverted-U findings from before in terms of the number of agents that a firm receives without having to worry about incentives.
 In the extreme when the firm gets all agents for free as in the monopoly case then it is incentivized to play \DG.
@@
 However, if the number of free agents gets small enough then \DG is incentivized as in the permanent duopoly analysis from before.\swcomment{I don't understand this paragraph; I think it's too vague.
-Perhaps remove?}
+Perhaps remove?}}
 

### 2019-02-14 paper/content/inverted_u.tex
Remove old \asedit commands.

--- 2018-11-06 paper/content/inverted_u.tex
@@
 
-\xhdr{Welfare.}
-\asedit{We study the effects of competition on consumer welfare: the total reward collected by the users over time. Rather than welfare directly, we find it more lucid to consider
+\xhdr{Welfare implications.}
+We study the effects of competition on consumer welfare: the total reward collected by the users over time. Rather than welfare directly, we find it more lucid to consider
 \emph{market regret}:
 \[ \textstyle T\, \max_a \mu(a) - \sum_{t\in [T]} \mu(a_t), \]
-where $a_t$ is the arm chosen by agent $t$. This is a standard performance measure in the literature on multi-armed bandits. Note that smaller regret means higher welfare.}
+where $a_t$ is the arm chosen by agent $t$. This is a standard performance measure in the literature on multi-armed bandits. Note that smaller regret means higher welfare.
 
@@
 
-\asedit{We assume that both firms play their respective equilibrium strategies for the corresponding competition level. As discussed previously, these are: 
+We assume that both firms play their respective equilibrium strategies for the corresponding competition level. As discussed previously, these are:
 \begin{itemize}
@@
 \end{itemize}
-} %%%%%%%
 
@@
 the duopoly case.
-\asedit{Recall that under temporary monopoly, the incumbent is incentivized to play \TS.
+Recall that under temporary monopoly, the incumbent is incentivized to play \TS.
 Moreover, we find that the welfare is close to that of having a single firm for all agents and running \TS.
-We also observe that monopoly and duopoly achieve similar welfare.}
+We also observe that monopoly and duopoly achieve similar welfare.
 %with monopoly being marginally better than duopoly.
@@
 
-\asedit{Finding~\ref{find:welfare}(b) is interesting because, in equilibrium, both firms play \DG in both settings, and one might conjecture that the welfare should increase with the number of firms playing \DG.
+Finding~\ref{find:welfare}(b) is interesting because, in equilibrium, both firms play \DG in both settings, and one might conjecture that the welfare should increase with the number of firms playing \DG.
 Indeed, one run of \DG may get stuck on a bad arm.
 However, two firms independently playing \DG are less likely to get stuck simultaneously.
-If one firm gets stuck and the other does not, then the latter should attract most agents, leading to improved welfare.}
+If one firm gets stuck and the other does not, then the latter should attract most agents, leading to improved welfare.
 \begin{figure}
@@
 
-\asedit{We track the average \Eeog in each of the
+We track the average \Eeog in each of the
 simulations and notice that it \textit{increases} with the number of firms.
-This observation also runs counter of the intuition that with more firms running \DG, one of them is more likely to ``get lucky" and take over the market (which would cause \Eeog to \emph{decrease} with the number of firms).}
+This observation also runs counter of the intuition that with more firms running \DG, one of them is more likely to ``get lucky" and take over the market (which would cause \Eeog to \emph{decrease} with the number of firms).
 

### 2019-02-14 paper/content/barriers.tex
update section labels and fixup appendix
Removed an old color-edit macro.

@@
 
-\section{Data as a Barrier to Entry}\label{section:6}
+\section{Data as a Barrier to Entry}\label{sec:barriers}
 
@@
 
-\asdelete{We explore what factors drive the large market share for the incumbent in the temporary monopoly model.}
+%We explore what factors drive the large market share for the incumbent in the temporary monopoly model.
 
@@
 Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant.
-\asedit{We explore which of the two factors is more significant.}  Our findings provide a quantitative insight into the role of the classic ``first mover advantage" phenomenon in the digital economy.
+We explore which of the two factors is more significant.
+Our findings provide a quantitative insight into the role of the classic ``first mover advantage" phenomenon in the digital economy.
 

### 2019-02-14 paper/content/introduction.tex
Intro: some edits in "results".

--- 2019-02-14 paper/content/introduction.tex
@@
 In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term.
-This would leave the learning algorithms with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+This would leave the learning algorithm with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+\asedit{Taken to the extreme, such dynamics may create a ``death spiral" effect when the vast majority of customers eventually switch to the competitors.}
 
@@
 
+We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a first-mover.
+Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``incumbent" (the first-mover in temporary monopoly) to permanent duopoly to ``entrant" (late-arriver in temporary monopoly).
+
+We focus on three classes of bandit algorithms, ranging from more primitive to more sophisticated: \emph{greedy algorithms} that do not explicitly explore, algorithms that separate exploration and exploitation, and algorithms that combine the two.
+We know from prior work that in the absence of competition,  ``better" algorithms are better in the long run, but could be worse initially.
+
+\xhdr{Main findings.}
+We find that in the permanent duopoly, competition incentivizes firms to choose the ``greedy algorithm", and even more so if the firm is a late arriver in a market.
+This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
+Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm.
+As a result, consumer welfare is highest under temporary monopoly.
+\asedit{We find strong evidence of the ``death spiral" effect mentioned earlier; this effect is strongest under permanent duopoly.}
+
+%and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.\swcomment{why ``even''?}
+
+Interpreting the adoption of better algorithms as ``innovation", our
+findings can be framed in terms of the ``inverted-U relationship"
+between competition and innovation (see Figure~\ref{fig:inverted-U}).
+This is a well-established concept in the economics literature, dating
+  back to \cite{Schumpeter-42}, whereby too
+  little or too much competition is bad for innovation, but
+  intermediate levels of competition tend to be better. However, the
+``inverted-U relationship'' is driven by different aspects in our
+model than the ones in the existing literature in economics. In our
+case, the barriers for innovations arise entirely from the
+reputational consequences of exploration in competition, as opposed to
+the R\&D costs (which is the more standard cause in prior work).
+
 \tikzstyle{level 1}=[level distance=3.5cm, sibling distance=4.0cm]
@@
 Being first in the market gives free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
-We run additional experiments so as to isolate these two effects.
-We find that the data advantage alone can lead to a significant advantage under competition.
-In fact, even a small ``head start'' can get amplified into a large incumbency advantage under competition.
-This effect is most pronounced when the incumbent commits to a more advanced bandit algorithm.
-These findings suggest that competition dynamics are pertinent to the anti-trust considerations of data as a barrier to entry.
+We run additional experiments so as to isolate and compare these two effects.
+We find that either effects alone leads to a significant advantage under competition.
+The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.
+
+\asedit{Data advantage is significant from the anti-trust perspective, as a possible barrier to entry.
+We find that even a small amount ``data advantage" gets amplified under competition, causing a large difference in eventual market shares.
+This observation runs contrary to prior work  \cite{varian2018artificial,lambrecht2015can,bajari2018impact}, which studied learning without competition, and found that small amounts of additional data do not provide significant improvement in eventual outcomes.
+We conclude that competition dynamics -- that firms compete as they learn over time -- are pertinent to these anti-trust considerations.}
 

### 2019-02-14 paper/content/introduction.tex
Deleted smth in intro/related work/econ.

@@
 %\swdelete{These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.}
-We ask how additional data helps when firms simultaneously solve a machine learning problem and compete for consumers.
-Due to the competition dynamics in our model, we find that even a small amount of additional data tends to help a lot.
+\asdelete{We ask how additional data helps when firms simultaneously solve a machine learning problem and compete for consumers.
+Due to the competition dynamics in our model, we find that even a small amount of additional data tends to help a lot.}
 The first-mover advantage has been well-studied in other settings in economics and marketing, see survey \cite{kerin1992first}.

### 2019-02-14 paper/content/model.tex
Minor edit.

@@
 \item an exploration-separated algorithm that in each round, \emph{explores} with probability $\eps$: chooses an arm independently and uniformly at random, and with the remaining probability \emph{exploits} according to \DG.
-We call it ``dynamic epsilon-greedy", \DEG in short.\gaedit{\footnote{For our experiments, we fix $\eps = 0.05$.
+We call it ``dynamic epsilon-greedy", \DEG in short.\gaedit{\footnote{Throughout, we fix $\eps = 0.05$.
 Our pilot experiments showed that different $\eps$ did not qualitatively change the results.}}
@@
 
-\item \emph{Uniform instance}: the mean reward of each arm is drawn independently and uniformly from $[0.25, 0.75]$.
-\item \emph{Heavy-tail instance}: the mean reward of each arm is drawn independently from $\Beta(.6,.6)$ distribution.%
-    \footnote{This distribution has substantial ``tail probabilities".}
+\item \emph{Uniform instance}: the mean reward of each arm is drawn independently and uniformly from $[\nicefrac{1}{4}, \nicefrac{3}{4}]$.
+\item \emph{Heavy-tail instance}: the mean reward of each arm is drawn independently from $\Beta(.6,.6)$ distribution (which is known to have substantial ``tail probabilities").
 \end{enumerate}

### 2019-02-14 paper/competing_bandits.tex
Minor edit in abstract.

@@
   suboptimal decisions in order to acquire new information for future
-  decisions. However, \swedit{when multiple systems are competing for
-    the same market of users together, exploration may hurt a system's
-    reputation in the near term, resulting in potential adverse
-    ``death spiral'' effects: the short-term reputation cost decreases
+  decisions. However, when multiple systems are competing for
+    the same market of users, exploration may hurt a system's
+    reputation in the near term, with adverse competitive effects. In particular, a system may enter a ``death spiral", when the short-term reputation cost decreases
     the number of users for the system to learn from, which degrades
     system's performance relative to competition and further decreases
-    the market share.}
+    the market share.
 
@@
 We run extensive numerical experiments in a stylized duopoly model in which two firms deploy multi-armed bandit algorithms and compete for myopic users.
-We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore \gaedit{and leads to low consumer welfare}, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms \gaedit{and lead to higher consumer welfare.
-Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.}
+We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore and leads to low consumer welfare, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms and lead to higher consumer welfare.
+Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.
 \end{abstract}

### 2019-02-14 paper/content/conclusion.tex
Revised conclusions.

@@
 
-\gaedit{In this paper we considered a setting where firms simultaneously learn from and compete for users.
-We showed that competition may not always induce firms to commit to algorithms that engage in exploration resulting in welfare losses for consumers.
-The primary mechanism behind this was that exploration may have short-term reputational consequences that lead to naive algorithms winning in competition that are bad in the long-run.
-Allowing one firm to have a head start, or a first-mover advantage, leads to the first-mover having an incentive to play traditionally ``good" bandit algorithms and resulting in significantly better welfare for consumers.
-Finally, we attempted to isolate the effect that the data advantage of an incumbent has on its ability to garner a considerable market share.
-Contrary to previous studies and discussions which do not consider competition dynamics, we find that the data advantage is substantial in generating significant market power.
-Additionally we show that this effect is strongest when an incumbent \swedit{performs efficient exploration in the incumbency period without the pressure from a competitor.}}
+\asedit{We consider a stylized duopoly setting where firms simultaneously learn from and compete for users.
+We showed that competition may not always induce firms to commit to better exploration algorithms, resulting in welfare losses for consumers.
+The primary reason is that exploration may have short-term reputational consequences that lead to more naive algorithms winning in a long-term competition.
+Allowing one firm to have a head start, a.k.a.
+the first-mover advantage, incentivizes the first-mover to deploy ``better" algorithms, which in turn leads to better welfare for consumers.
+Finally, we isolate the component of the first-mover advantage that is due to having more initial data, and find that even a small amount of this ``data advantage" leads to substantial long-term market power.}
 

### 2019-02-14 paper/content/introduction.tex
Added an asedit command.

@@
 We run additional experiments so as to isolate and compare these two effects.
-We find that either effects alone leads to a significant advantage under competition.
-The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.
+\asedit{We find that either effects alone leads to a significant advantage under competition.
+The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.}
 

### 2019-02-14 paper/content/model.tex
fix typos

@@
 Each algorithm receives a ``warm start": additional $T_0$ agents that arrive before the game starts, and interact with the firm as described above.
-\asedit{The warm ensures that each firm has a meaningful reputation when competition starts.} Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
+\asedit{The warm start ensures that each firm has a meaningful reputation when competition starts.} Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
 

### 2019-02-14 paper/content/revisited.tex
updates to paper
update section labels and fixup appendix
Minor edit.

@@
-\documentclass[../competing_bandits.tex]{subfiles}
+\documentclass[../competing_bandits_with_appendix.tex]{subfiles}
 \begin{document}
 
-\section{Performance in Isolation, Revisited}\label{section:revisited}
+\section{Performance in Isolation, Revisited}\label{sec:revisited}
 
@@
 \includegraphics[scale=0.35]{figures/rep_distribution_nih}
-\caption{Distribution of reputation scores for Needle-in-Haystack at $t=500$ (smoothed using a kernel density estimate)}
+\caption{Reputation scores for Needle-in-Haystack at $t=500$ (smoothed using a kernel density estimate)}
 \label{rep_dist_nih}
@@
 \includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
-\caption{Distribution of reputation difference $\TS-\DG$ for  Needle-in-Haystack
+\caption{Reputation difference $\TS-\DG$ for  Needle-in-Haystack
 % at $t = 500, 1000, 2000$ AS: this is redundant, and it saves a line :).
-(smoothed via a kernel density estimate)}
+(smoothed using a kernel density estimate)}
 \label{ts_dg_rep_diff_nih}

### 2019-02-14 paper/content/introduction.tex
fix typos
Minor edit.

@@
 We run additional experiments so as to isolate and compare these two effects.
-\asedit{We find that either effects alone leads to a significant advantage under competition.
+\asedit{We find that either effect alone leads to a significant advantage under competition.
 The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.}
@@
 
-While reputation scores already introduce some noise and randomness into users' choices, we consider an alternative choice model in which the noise/randomness is explicit: a small fraction of users choose a firm uniformly at random.
-We confirm the theoretical intuition that better algorithms prevail if the parameters -- the ``random fraction" and the time horizon -- are sufficiently large.
+While reputation scores already introduce some noise and randomness into users' choices, we consider an alternative choice rule in which the noise/randomness is explicit: a small fraction of users choose a firm uniformly at random.
+We confirm the theoretical intuition that better algorithms prevail if the parameter -- \asedit{essentially, the expected number ``random users"} -- is sufficiently large.
 However, we find that this effect is negligible for some smaller but still ``relevant" parameter values.

### 2019-02-14 paper/content/non_greedy_choice.tex
Some edits.

@@
 
-In this section we consider an extension of our base model where the agents' choice model is no longer deterministic.
-Recall that in our base model agents deterministically choose the firm with the higher reputation score.
-We define this choice model as \textit{HardMax} (\HM).
-We study a small deviation from this choice model where the agent now selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the firm with the higher reputation score with probability $1 - \epsilon$.
-Following \cite{mansour2018competing} we define this choice model as \textit{HardMax with random} (\HMR).
+Let us consider an extension in which the agents' choice is no longer deterministic.
+Recall that in our main model agents deterministically choose the firm with the higher reputation score; call this choice rule \emph{HardMax} (\HM).
+Now, we introduce some randomness: each agent selects between the firms uniformly with probability $\eps\in (0,1)$, and takes the firm with the higher reputation score with the remaining probability; call this choice rule \emph{HardMax with randomness} (\HMR).
 
-In Section \ref{sec:competition} we show that in the permanent duopoly
+One can view \HMR as a version of ``warm start", where a firm receives some customers without competition, but these customers are dispersed throughout the game.
+The expected duration of this ``dispersed warm start" is $\eps T$.
+If this quantity is large enough, we expect better algorithms to reach their long-term performance and prevail in competition.
+We confirm this intuition; we also find that this effect is negligible for smaller (but still relevant) values of $\eps$ or $T$.
+
+\OMIT{In Section \ref{sec:competition} we show that in the permanent duopoly
 case exploration can lead to a death spiral which eventually starves
@@
 concentrating the free agents as arriving in the beginning of the game
-instead they are dispersed throughout the game.
+instead they are dispersed throughout the game.}
 
@@
 %For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
-Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\epsilon = 0.1$ in the \HMR choice model and $T_0 = 20$.}
+Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\eps = 0.1$ in the \HMR choice rule and $T_0 = 20$.}
 \label{tab:non_greedy_table}
@@
 \begin{finding}\label{find:non_greedy_choice}
-\textit{Under the \HMR~ choice model \TS is weakly dominant as long as $T$ is sufficiently large. \swdelete{However, both $T$ and $\epsilon$ have to be relatively large for this to occur.}}
+\textit{\TS is weakly dominant under the \HMR~ choice rule, if and only if $\eps T$ is sufficiently large.}
 \end{finding}
@@
 Table \ref{tab:non_greedy_table} shows the average market shares under
-the \HM~ vs \HMR~ choice model. \swedit{In contrast to the \HM~ model,
+the \HM~ vs \HMR~ choice rule. In contrast to the \HM~ model,
   \TS becomes weakly dominant under the \HMR model, as $T$ gets
-  sufficiently large.} These results hold across all problem
+  sufficiently large. These findings hold across all problem
 instances.\footnote{The results here are pulled using different \MRV
@@
   not directly comparable to those from the previous
-  section.\swcomment{not sure we need this}} The intuition for why \TS
+  section.\swcomment{not sure we need this}}
+
+\OMIT{The intuition for why \TS
 should become the dominant strategy eventually is simple. The
 consistent stream of random agents ensures that each principal is
-chosen at least $\Omega(\epsilon t)$ times at every time step $t$. As
+chosen at least $\Omega(\eps t)$ times at every time step $t$. As
 a result, each algorithm should eventually converge to its asymptotic
-performance in isolation.
+performance in isolation.}
 
-However, it takes a significant amount of randomness and a relatively large time horizon for this convergence to occur.
-Even with $T = 10000$ and $\epsilon = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance.
-Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that \HMR leads to lower variance in market shares across simulations compared to \HM.
+However, it takes a significant amount of randomness and a relatively large time horizon for this effect to take place.
+Even with $T = 10000$ and $\eps = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance.
+Table \ref{tab:non_greedy_table} also shows that another difference between the two choice rules is that \HMR leads to lower variance in market shares across simulations compared to \HM.
 

### 2019-02-14 paper/content/non_greedy_choice.tex
Commented out the footnote that Steven didn't like.

@@
   sufficiently large. These findings hold across all problem
-instances.\footnote{The results here are pulled using different \MRV
+instances.
+
+\OMIT{\footnote{The results here are pulled using different \MRV
   and realizations than the results pulled previously (due to the
@@
   not directly comparable to those from the previous
-  section.\swcomment{not sure we need this}}
+  section.\swcomment{not sure we need this}}}
 

### 2019-02-14 paper/content/non_greedy_choice.tex
add new table
Some more edits.

@@
   \hline
-  & \multicolumn{3}{c||}{Heavy-Tail (\HMR)}
+  & \multicolumn{3}{c||}{Heavy-Tail (\HMR with $\eps=.1$)}
   & \multicolumn{3}{c|}{Heavy-Tail (\HM)}\\
@@
 \normalsize
-\caption{Performance Comparison between \HM and \HMR Choice on the Heavy-Tail MAB instance.
+\caption{\HM and \HMR choice models on the Heavy-Tail MAB instance.
 Each cell describes the market shares in a game between two algorithms, call them Alg1 vs.
@@
 %For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
-Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\eps = 0.1$ in the \HMR choice rule and $T_0 = 20$.}
+Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $T_0 = 20$.}
 \label{tab:non_greedy_table}
@@
 \begin{finding}\label{find:non_greedy_choice}
-\textit{\TS is weakly dominant under the \HMR~ choice rule, if and only if $\eps T$ is sufficiently large.}
+\textit{\TS is weakly dominant under the \HMR~ choice rule, if and only if $\eps T$ is sufficiently large. \asedit{\HMR leads to lower variance in market shares, compared to \HM.}}
 \end{finding}
@@
   sufficiently large. These findings hold across all problem
-instances.
+instances, see Table \ref{tab:additional_results} (with the same semantics as in Table \ref{tab:non_greedy_table}).
 
@@
 
+\normalsize
 \OMIT{The intuition for why \TS
@@
 
-However, it takes a significant amount of randomness and a relatively large time horizon for this effect to take place.
-Even with $T = 10000$ and $\eps = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance.
-Table \ref{tab:non_greedy_table} also shows that another difference between the two choice rules is that \HMR leads to lower variance in market shares across simulations compared to \HM.
-
 \OMIT{Finding \ref{find:non_greedy_choice} implies that we can re-interpret the inverted-U findings from before in terms of the number of agents that a firm receives without having to worry about incentives.
@@
 
+\footnotesize
+\begin{table*}[ht]
+\centering
+\begin{tabular}{|c|c|c|c||c|c|c|}
+  \hline
+  & \multicolumn{3}{c||}{Uniform (\HMR  with $\eps=.1$)}
+  & \multicolumn{3}{c|}{Needle-In-Haystack (\HMR  with $\eps=.1$)}\\
+  \hline
+  & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG
+ & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG  \\
+ \hline
+$T = 2000$
+ & \makecell{ \textbf{0.42} $\pm$ 0.02 \\Var: 0.13 }
+ & \makecell{ \textbf{0.45} $\pm$ 0.02 \\Var: 0.13 }
+ & \makecell{ \textbf{0.49} $\pm$ 0.02 \\Var: 0.093 }
+  %%
+  & \makecell{  \textbf{0.55} $\pm$ 0.02 \\Var: 0.15 }
+  & \makecell{  \textbf{0.61} $\pm$ 0.02 \\Var: 0.13 }
+  & \makecell{  \textbf{0.46} $\pm$ 0.02 \\Var: 0.12 }
+    \\
+\hline
+  $T= 5000$
+ & \makecell{ \textbf{0.48} $\pm$ 0.02 \\Var: 0.089 }
+ & \makecell{ \textbf{0.53} $\pm$ 0.02 \\Var: 0.098 }
+ & \makecell{ \textbf{0.46} $\pm$ 0.02 \\Var: 0.072 }
+  %%
+ & \makecell{  \textbf{0.56} $\pm$ 0.02 \\Var: 0.13 }
+ & \makecell{  \textbf{0.63} $\pm$ 0.02 \\Var: 0.12 }
+ & \makecell{  \textbf{0.43} $\pm$ 0.02 \\Var: 0.11 }
+ \\
+  \hline
+  $T = 10000$
+& \makecell{ \textbf{0.54} $\pm$ 0.01 \\Var: 0.055 }
+& \makecell{  \textbf{0.6} $\pm$ 0.02 \\Var: 0.073 }
+& \makecell{  \textbf{0.44} $\pm$ 0.02 \\Var: 0.064 }
+ %%
+  & \makecell{ \textbf{0.58} $\pm$ 0.02 \\Var: 0.083 }
+  & \makecell{ \textbf{0.65} $\pm$ 0.02 \\Var: 0.096 }
+  & \makecell{ \textbf{0.4} $\pm$ 0.02 \\Var: 0.1 }
+  %%
+  \\
+   \hline
+\end{tabular}
+\caption{\HMR choice model for Uniform and Needle-In-Haystack MAB instances.}
+%Same semantics as in Table \ref{tab:non_greedy_table}.
+\label{tab:additional_results}
+\end{table*}
+\normalsize
+
+However, it takes a significant amount of randomness and a relatively large time horizon for this effect to take place.
+Even with $T = 10000$ and $\eps = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance.
+
+\OMIT{Table \ref{tab:non_greedy_table} also shows that another difference between the two choice rules is that \HMR leads to lower variance in market shares across simulations compared to \HM.}
+
 \end{document}

### 2019-02-14 paper/content/model.tex
ensured consistency of instance naming

@@
 \item \emph{Uniform instance}: the mean reward of each arm is drawn independently and uniformly from $[\nicefrac{1}{4}, \nicefrac{3}{4}]$.
-\item \emph{Heavy-tail instance}: the mean reward of each arm is drawn independently from $\Beta(.6,.6)$ distribution (which is known to have substantial ``tail probabilities").
+\item \emph{Heavy-Tail instance}: the mean reward of each arm is drawn independently from $\Beta(.6,.6)$ distribution (which is known to have substantial ``tail probabilities").
 \end{enumerate}

### 2019-02-14 paper/content/perf_in_iso.tex
Heavy-tail -> Heavy-Tail
ensured consistency of instance naming

@@
 We plot the \emph{mean reputation trajectory}: how this score evolves over time.
-Figure \ref{prelim_means} shows such a plot for the Needle In Haystack instance; for other MAB instances the plots are similar.
+Figure \ref{prelim_means} shows such a plot for the Needle-in-Haystack instance; for other MAB instances the plots are similar.
 We summarize this finding as follows:
@@
 \DG dips below $\tfrac12$ can be seen as an explanation for the competitive disadvantage of exploration.
-Such period also exists for the Heavy-tail MAB instance.
+Such period also exists for the Heavy-Tail MAB instance.
 However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
-\footnote{We see two explanations for this: $\TS$ identifies the best arm faster for the Needle In Haystack instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
+\footnote{We see two explanations for this: $\TS$ identifies the best arm faster for the Needle-in-Haystack instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
 

### 2019-02-14 paper/content/revisited.tex
ensured consistency of instance naming

@@
 That is, consider the empirical distribution of this score over different \MRVs.%
-\footnote{Recall that each \MRV in our experimental setup comes with one specific realization table.} For concreteness, consider the needle-in-haystack instance at time $t=500$, plotted in Figure \ref{rep_dist_nih}.
+\footnote{Recall that each \MRV in our experimental setup comes with one specific realization table.} For concreteness, consider the Needle-in-Haystack instance at time $t=500$, plotted in Figure \ref{rep_dist_nih}.
 (The other MAB instances lead to a similar intuition.)

### 2019-02-14 paper/competing_bandits.tex
Added a newline break.
fixes to intro and abstract

@@
-\documentclass[format=acmsmall, review=false]{acmart}
+\documentclass[format=acmsmall, review=true]{acmart}
 
@@
 %\usepackage{color-edits}
-\usepackage{color-edits}  % use this to suppress the package
+\usepackage[suppress]{color-edits}  % use this to suppress the package
 \addauthor{as}{red}    % as for Alex
@@
 
+\author{Submission 316}
+
 \begin{abstract}
@@
     reputation in the near term, with adverse competitive effects. In particular, a system may enter a ``death spiral", when the short-term reputation cost decreases
-    the number of users for the system to learn from, which degrades
+    the number of users for the system to learn from, which degrades the
     system's performance relative to competition and further decreases
@@
 \subfile{content/conclusion}
-
+\newpage 
 \bibliographystyle{acm}

### 2019-02-14 paper/content/introduction.tex
fixes to intro and abstract

@@
 While deploying ``better" learning algorithms for exploration would improve performance, this is not necessarily beneficial under competition, even putting aside the deployment/maintenance costs.
-In particular, excessive experimentation may hurt platform's reputation and decrease market share in the near term.
-This would leave the learning algorithm with less users to learn from, which may further degrade platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
-\asedit{Taken to the extreme, such dynamics may create a ``death spiral" effect when the vast majority of customers eventually switch to the competitors.}
+In particular, excessive experimentation may hurt a platform's reputation and decrease its market share in the near term.
+This would leave the learning algorithm with less users to learn from, which may further degrade the platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+\asedit{Taken to the extreme, such dynamics may create a ``death spiral" effect when the vast majority of customers eventually switch to competitors.}
 

### 2019-04-03 paper/ec19-rebuttal.txt
Full draft!

Apologies for brevity! We'll discuss all issues in revision.

Several important differences compared to [MSW] [26], as discussed on page3:

In [MSW], users don't observe anything about firms' actual performance, and instead choose based on complicated Bayesian reasoning. We consider a much more realistic model with "reputation signals". This model is amenable to numerical simulations (whereas simulating the priors in [MSW] is computationally very difficult). Moreover, we can separate data vs. reputational advantage (Section5).

[MSW] changes "competitiveness" via behavioral model ("HardMax" -> "HardMaxWithRandomness" -> "SoftMax"). We focus on different "channel": early -> simultaneous -> late entry (and investigate the "first-mover advantage").

[MSW] is theory-only, with strong but "asymptotic" results ("after sufficiently many rounds"). Our findings are more quantitative: whether effects are sufficiently strong in realistic time.

(Finding11 is a good example: as we emphasize right after, the effect is not observed until fairly large T. Put differently, no effect when/if such T is unrealistic, even though this effect is predicted by [MSW].)

Thus, theorems from [MSW] have no direct bearing on our simulations.

Re lack of proofs:

We believe theorems wouldn't not add much value. Indeed, we are in realm of  "asymptotic" theorems about stylized Economic models, which provide some "hard" intuition but no direct bearing on the world. But [MSW] already has an elaborate set of such results, with similar high-level conclusions. So, we'd need to go beyond the first-order conclusions to add value.

While amenable to simulations, our model appears difficult to analyze:

    - intricate feedback loop (performance -> reputations -> users -> performance),

    - mean reputation, most connected to our intuition, is sometimes bad predictor in competition (Sections 3,6),

    - mathematical tools from regret-minimization only produce "asymptotic" results.

We proved *something*: if incumbent starts so early that it has permanent "first-order stochastic dominance" over entrant as standalone algorithm, then it has some advantage under competition. This result is quite weak, yet not easy to prove, and we didn't see much added value. Perhaps we should have included it, just to illustrate the difficulty of analysis.

Re significance:

General adage, we don't know which intuitions/effects are significant until we test. In particular, our intuitions are all about mean reputations, which is sometimes wrong thing to track (Sections 3,6). Also, some effects matter only after very long time (ex: Finding 11).

Welfare implications are at least somewhat surprising.

We didn't have much/any intuition about "data advantage" vs. "reputational advantage"

Re monopoly [Review3]: No findings here, just assumption: monopolist chooses
Dynamic Greedy (DG) because it is easier to deploy in practice (page 2). Implicitly, users have no "outside option": the service provided is an improvement over not having it. This is plausible with free ad-supported platforms such as Yelp or Google. Pricing-for-quality modifies firms' utilities in all scenarios, not just under monopoly. Also a valid model, worth investigating.

[Review3] Re changing competitiveness: "channels" from literature (Lerner Index [3] or Herfindahl-Hirschman Index) just aren't applicable,  we and [MSW] invent channels that are (more details don't fit here).

[Review3] Lower "memory size"  induces noise which we cannot easily characterize. Such noise is also unrealistic, as reputation signals usually average over many datapoints. We remove "noise" in main experiments (but study it in Section7).

[Review1] Agreed, "adaptive strategies" (ex: DG then TS) are interesting! However, "fixed strategies" are arguably more realistic: detecting competition, committing to new algorithms, and adapting/deploying them takes very long time in practice (background: https://arxiv.org/abs/1606.03966). Much larger strategy space, obviously. We did some pilot experiments, where we found that changing the algorithm did not matter.

### 2019-04-03 paper/ec19-rebuttal.txt
light edits

@@
 We consider a much more realistic model with "reputation signals".
-This model is amenable to numerical simulations (whereas simulating the priors in [MSW] is computationally very difficult).
+This model is amenable to numerical simulations (whereas simulating the posteriors in [MSW] is computationally very difficult).
 Moreover, we can separate data vs.
@@
 
-We believe theorems wouldn't not add much value.
+We believe theorems wouldn't add much value.
 Indeed, we are in realm of  "asymptotic" theorems about stylized Economic models, which provide some "hard" intuition but no direct bearing on the world.
@@
 General adage, we don't know which intuitions/effects are significant until we test.
-In particular, our intuitions are all about mean reputations, which is sometimes wrong thing to track (Sections 3,6).
-Also, some effects matter only after very long time (ex: Finding 11).
+In particular, much of our intuitions are about mean reputations, which is sometimes wrong thing to track (Sections 3,6).
+Also, some effects matter only after very long time (see Finding 11).
 
@@
 
-[Review3] Lower "memory size"  induces noise which we cannot easily characterize.
+[Review3] Lower "memory size"  induces noise that we cannot easily characterize.
 Such noise is also unrealistic, as reputation signals usually average over many datapoints.
@@
 
-[Review1] Agreed, "adaptive strategies" (ex: DG then TS) are interesting!
+[Review1] Agreed, "adaptive strategies" are interesting!
 However, "fixed strategies" are arguably more realistic: detecting competition, committing to new algorithms, and adapting/deploying them takes very long time in practice (background: https://arxiv.org/abs/1606.03966).

### 2019-04-03 paper/ec19-rebuttal.txt
Minor edits.
Minor edit.

@@
 
-[Review3] Re changing competitiveness: "channels" from literature (Lerner Index [3] or Herfindahl-Hirschman Index) just aren't applicable,  we and [MSW] invent channels that are (more details don't fit here).
+[Review3] Re changing competitiveness: "channels" from literature (Lerner Index [3] or Herfindahl-Hirschman Index) just aren't applicable in our setting, whereas we and [MSW] identify channels that are.
+(More details don't fit here).
 
@@
 
-[Review1] Agreed, "adaptive strategies" are interesting!
+[Review1] Agreed, "dynamic strategies" (ex: DG then TS) are interesting!
 However, "fixed strategies" are arguably more realistic: detecting competition, committing to new algorithms, and adapting/deploying them takes very long time in practice (background: https://arxiv.org/abs/1606.03966).

### 2019-04-29 paper/content/conclusion.tex
Some post-rebuttal edits. Also, removed old color-edits.

@@
 
-\asedit{We consider a stylized duopoly setting where firms simultaneously learn from and compete for users.
+We consider a stylized duopoly setting where firms simultaneously learn from and compete for users.
 We showed that competition may not always induce firms to commit to better exploration algorithms, resulting in welfare losses for consumers.
@@
 the first-mover advantage, incentivizes the first-mover to deploy ``better" algorithms, which in turn leads to better welfare for consumers.
-Finally, we isolate the component of the first-mover advantage that is due to having more initial data, and find that even a small amount of this ``data advantage" leads to substantial long-term market power.}
+Finally, we isolate the component of the first-mover advantage that is due to having more initial data, and find that even a small amount of this ``data advantage" leads to substantial long-term market power.
 

### 2019-04-29 paper/content/introduction.tex
Some post-rebuttal edits. Also, removed old color-edits.

@@
 This would leave the learning algorithm with less users to learn from, which may further degrade the platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
-\asedit{Taken to the extreme, such dynamics may create a ``death spiral" effect when the vast majority of customers eventually switch to competitors.}
+Taken to the extreme, such dynamics may create a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
 
@@
 We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a first-mover.
-Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``incumbent" (the first-mover in temporary monopoly) to permanent duopoly to ``entrant" (late-arriver in temporary monopoly).
+Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``incumbent" (the first-mover in temporary monopoly) to permanent duopoly to ``entrant" (late-arriver in temporary monopoly).%
+\footnote{\asedit{We consider the ``permanent monopoly" scenario for comparison only, without presenting any findings. We just assume that a monopolist chooses
+the greedy algorithm, because it is easier to deploy in practice.
+Implicitly, users have no ``outside option": the service provided is an improvement over not having it (and therefore the monopolist is not incentivized to deploy better learning algorithms).
+This is plausible with free ad-supported platforms such as Yelp or Google.}}
 
@@
 As a result, consumer welfare is highest under temporary monopoly.
-\asedit{We find strong evidence of the ``death spiral" effect mentioned earlier; this effect is strongest under permanent duopoly.}
+We find strong evidence of the ``death spiral" effect mentioned earlier; this effect is strongest under permanent duopoly.
 
@@
 We run additional experiments so as to isolate and compare these two effects.
-\asedit{We find that either effect alone leads to a significant advantage under competition.
-The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.}
+We find that either effect alone leads to a significant advantage under competition.
+The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.
 
-\asedit{Data advantage is significant from the anti-trust perspective, as a possible barrier to entry.
+Data advantage is significant from the anti-trust perspective, as a possible barrier to entry.
 We find that even a small amount ``data advantage" gets amplified under competition, causing a large difference in eventual market shares.
 This observation runs contrary to prior work  \cite{varian2018artificial,lambrecht2015can,bajari2018impact}, which studied learning without competition, and found that small amounts of additional data do not provide significant improvement in eventual outcomes.
-We conclude that competition dynamics -- that firms compete as they learn over time -- are pertinent to these anti-trust considerations.}
+We conclude that competition dynamics -- that firms compete as they learn over time -- are pertinent to these anti-trust considerations.
 
@@
 
-While reputation scores already introduce some noise and randomness into users' choices, we consider an alternative choice rule in which the noise/randomness is explicit: a small fraction of users choose a firm uniformly at random.
-We confirm the theoretical intuition that better algorithms prevail if the parameter -- \asedit{essentially, the expected number ``random users"} -- is sufficiently large.
+We also consider an alternative choice rule with explicit noise/randomness: a small fraction of users choose a firm uniformly at random.%
+\footnote{\asedit{Reputation scores already introduce some noise into users' choices.
+However, the amount of noise due to this channel is typically small, both in our simulations and in practice, because reputation signals average over many datapoints.}}
+
+ We confirm the theoretical intuition that better algorithms prevail if the parameter -- essentially, the expected number ``random users" -- is sufficiently large.
 However, we find that this effect is negligible for some smaller but still ``relevant" parameter values.
 
-\xhdr{Discussion.}
+\subsection{Discussion}
+
 The present paper is an experimental counterpart to \cite{CompetingBandits-itcs18}, which considered a similar duopoly model and obtained a number of theoretical results with ``asymptotic" flavor.
-For the sake of analytical tractability, that paper makes a somewhat unrealistic simplification that users do not observe any signals about firms' ongoing performance (instead, users choose between firms according to the firms' Bayesian-expected rewards).
-The high-level conclusion in \cite{CompetingBandits-itcs18} is an inverted-U relationship similar to ours, but the strength of competition is varied in a different way, using assumptions about (ir)rational user behavior.
+For the sake of analytical tractability, that paper makes a somewhat unrealistic simplification that users do not observe any signals about firms' ongoing performance.
+Instead, users choose between firms according to the firms' Bayesian-expected rewards.
+The strength of competition is varied in a different way, using assumptions about (ir)rational user behavior.
+\asedit{For these reasons, the theorems from \cite{CompetingBandits-itcs18} have no direct bearing on our simulations.} However, their high-level conclusion in is an inverted-U relationship similar to ours.
 
@@
 Further, one needs succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).
+
+\asedit{While amenable to simulations, our model appears difficult to analyze. This is for several reasons:
+%
+intricate feedback loop from performance to reputations to users to performance;
+%
+mean reputation, most connected to our intuition, is sometimes bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited});
+%
+mathematical tools from regret-minimization would only produce ``asymptotic" results, which does not seem to suffice.
+We believe that resolving first-order theoretical questions would not add much value to this paper, because of \cite{CompetingBandits-itcs18}.
+We are in realm of ``asymptotic" theorems about stylized Economic models, which provide some mathematical intuition but no direct bearing on the world.
+But \cite{CompetingBandits-itcs18} already has an elaborate set of such results, with similar high-level conclusions.
+So, one would need to go beyond the first-order conclusions to add value.}
 
@@
 %\swdelete{These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.}
-\asdelete{We ask how additional data helps when firms simultaneously solve a machine learning problem and compete for consumers.
-Due to the competition dynamics in our model, we find that even a small amount of additional data tends to help a lot.}
 The first-mover advantage has been well-studied in other settings in economics and marketing, see survey \cite{kerin1992first}.

### 2019-04-29 paper/content/model.tex
Some post-rebuttal edits. Also, removed old color-edits.

@@
 Each algorithm receives a ``warm start": additional $T_0$ agents that arrive before the game starts, and interact with the firm as described above.
-\asedit{The warm start ensures that each firm has a meaningful reputation when competition starts.} Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
+The warm start ensures that each firm has a meaningful reputation when competition starts.
+Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
 
@@
 \xhdr{Agents.}
-\asedit{Firms compete on a single dimension, quality of service, as expressed by agents' rewards.}
+Firms compete on a single dimension, quality of service, as expressed by agents' rewards. 
 \OMIT{\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}}
-Agents are myopic and non-strategic: they \asedit{would like} to choose among the firms so as to maximize their expected reward \gaedit{(i.e.
-select the firm with the highest quality)}, without attempting to influence the firms' learning algorithms or rewards of the future users.
+Agents are myopic and non-strategic: they would like to choose among the firms so as to maximize their expected reward (i.e.
+select the firm with the highest quality), without attempting to influence the firms' learning algorithms or rewards of the future users.
 Agents are not well-informed: they only receive a rough signal about each firm's performance before they choose a firm, and no other information.
@@
 
-\asedit{We are mainly interested in qualitative differences between the three classes.} For concreteness, we fix one algorithm from each class.
+We are mainly interested in qualitative differences between the three classes.
+For concreteness, we fix one algorithm from each class.
 Our pilot experiments indicate that our findings do not change substantially if other algorithms are chosen.

### 2019-04-29 paper/content/perf_in_iso.tex
Some post-rebuttal edits. Also, removed old color-edits.

@@
 
-\asedit{We also use Figure~\ref{prelim_means} to choose a reasonable time-horizon for the subsequent experiments, as $T=2000$.
-The idea is, we want $T$ to be large enough so that algorithms performance starts to plateau, but small enough such that algorithms are still learning.}
+We also use Figure~\ref{prelim_means} to choose a reasonable time-horizon for the subsequent experiments, as $T=2000$.
+The idea is, we want $T$ to be large enough so that algorithms performance starts to plateau, but small enough such that algorithms are still learning.
 

### 2019-04-29 paper/ec19 workshop/cover-straregic-learning.tex
Full draft.

--- 2019-02-14 paper/competing_bandits.tex
@@
-\documentclass[format=acmsmall, review=true]{acmart}
+\documentclass[11pt,letterpaper]{article}
 
-\usepackage{acm-ec-19}
-\usepackage{tikz} % SW: Added for plots
-\usepackage{subfiles}
-\usepackage[toc,page]{appendix}
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{amsmath,amssymb,amsfonts}
-%\usepackage{natbib} % cannot use with AAAI! A.S.
+\usepackage{fullpage,times}
 
-\usepackage{amsthm} % The amsthm package provides extended theorem environments
-\usepackage{float}
-\usepackage{sgame, tikz} % Game theory packages
-\usepackage{caption}
-\usepackage{algorithm,algpseudocode}
-\usepackage{makecell}
- \usepackage{multirow}
- \usepackage{graphicx}
-\theoremstyle{definition}
-%\newtheorem{definition}{Definition}
-\newtheorem{finding}{Finding}
-%\newtheorem{conjecture}{Conjecture}
-\newtheorem{puzzle}{Puzzle}
-\captionsetup{font=footnotesize}
+\usepackage[round]{natbib}
 
-% Alex' macro
-\usepackage{xspace}
-\newcommand{\OMIT}[1]{}
-\newcommand{\xhdr}[1]{\vspace{1mm} \noindent{\bf #1}}
-\newcommand{\ie}{{\em i.e.,~\xspace}}
-\newcommand{\eg}{{\em e.g.,~\xspace}}
-
-\newcommand{\term}[1]{\ensuremath{\mathtt{#1}}\xspace}
-\newcommand{\TS}{\term{TS}}
-\newcommand{\DEG}{\term{DEG}}
-\newcommand{\DG}{\term{DG}}
-\newcommand{\Beta}{\term{Beta}} % for Beta distribution
-\newcommand{\Eeog}{\term{EoG}} % shorthand for "effective end of game"
-
-\newcommand{\eps}{\varepsilon}
-
-\newcommand{\MRV}{mean reward vector\xspace} % mean reward vector
-\newcommand{\MRVs}{mean reward vectors\xspace} % mean reward vector
-
-\newcommand{\HMR}{\term{HMR}}
-\newcommand{\HM}{\term{HM}}
 % a very useful package for edits and comments, from David Kempe (USC)
-%\usepackage{color-edits}
-\usepackage[suppress]{color-edits}  % use this to suppress the package
+\usepackage{color-edits}
+%\usepackage[suppress]{color-edits}  % use this to suppress the package
+%\addauthor{ab}{green}   % ab for Ashwin
+%\addauthor{bk}{magenta} % bk for Bobby
 \addauthor{as}{red}      % as for Alex
-\addauthor{ga}{blue}  % ga for Guy
-\addauthor{sw}{orange} % sw for Steven
+%\addauthor{mb}{blue}    % mb for Moshe
 % e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
-\setlength{\tabcolsep}{8pt} % Default value: 6pt
-\renewcommand{\arraystretch}{1.5} % Default value: 1
-
-%drt24 hacks
-% Letter paper
-
-% Hack to try to make acmart work with biblatex: https://tex.stackexchange.com/questions/37076/is-it-possible-to-load-biblatex-with-a-class-that-has-already-loaded-natbib
-\let\citename\relax
-%\RequirePackage[abbreviate=true, dateabbrev=true, isbn=true, doi=true, urldate=comp, url=true, maxbibnames=9, backref=false, backend=biber, style=ACM-Reference-Format, language=american]{biblatex}
-
-\usepackage{booktabs} % For formal tables
-
-% Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
 
 \begin{document}
-\title[Competing Bandits: The Perils of Exploration under Competition]
-{Competing Bandits: \\ The Perils of Exploration under Competition}
+%\vspace{-30mm}
 
-\author{Submission 316}
+\section*{Submission: Competing Bandits: The perils of Competition\\
+{\large (by Guy Aridor, Aleksandrs Slivkins and Zhiwei Steven Wu)}\\
+{\large Workshop: Learning in the Presence of Strategic Behavior}
+}
 
-\begin{abstract}
 We empirically study the interplay between \textit{exploration} and
@@
 Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.
-\end{abstract}
 
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below.
-%
-\maketitle
-\keywords{Multi-armed bandits,Exploration,Competition,Competition vs Innovation}
+\vspace{2mm}
 
-\newpage
+This paper continues the literature on ``exploration and incentives", which in turn belongs to the area of ``learning in the presence of strategic behavior".
+It is an experimental counterpart to \citet{CompetingBandits-itcs18}, which has initiated the study of ``exploration and competition" and obtained obtained a number of theoretical results with ``asymptotic" flavor.
+While they considered a similar duopoly model and arrived at similar high-level conclusions, their details and assumptions are substantially different from ours for the sake analytical tractability, and their theorems have no direct bearing on our simulations.
+In comparison, we provide a more nuanced and ``non-asymptotic" perspective, looking for substantial effects within relevant time scales.
+We plan a merged journal submission to a venue in economics or operations research.
 
-\subfile{content/introduction}
-
-\subfile{content/model}
-
-\subfile{content/perf_in_iso}
-
-\subfile{content/inverted_u}
-
-\subfile{content/barriers}
-
-\subfile{content/revisited}
-
-\subfile{content/non_greedy_choice}
-
-\subfile{content/conclusion}
-\newpage 
-\bibliographystyle{acm}
-\bibliography{refs,bib-abbrv-short,bib-bandits,bib-AGT,bib-slivkins}
-
-\subfile{content/appendix_for_one_version}
+\begin{small}
+\bibliographystyle{plainnat}
+\bibliography{bib-abbrv,bib-slivkins}
+\end{small}
 
 \end{document}
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2019-04-30 paper/competing_bandits.tex
Allowed color-edits.
some edits

@@
-\documentclass[format=acmsmall, review=true]{acmart}
+\documentclass[format=acmsmall, review=false]{acmart}
 
@@
 % a very useful package for edits and comments, from David Kempe (USC)
-%\usepackage{color-edits}
-\usepackage[suppress]{color-edits}  % use this to suppress the package
+\usepackage{color-edits}
+%\usepackage[suppress]{color-edits}  % use this to suppress the package
 \addauthor{as}{red}    % as for Alex
@@
 \begin{document}
-\title[Competing Bandits: The Perils of Exploration under Competition]
-{Competing Bandits: \\ The Perils of Exploration under Competition}
+\title[The Perils of Exploration under Competition: A Computational Modeling Approach]
+{The Perils of Exploration under Competition: \\ A Computational Modeling Approach}
 
-\author{Submission 316}
+ \author{Guy Aridor}
+ \affiliation{ \institution{Columbia University}}
+ \author{Kevin Liu}
+ \affiliation{\institution{Columbia University}}
+ \author{Aleksandrs Slivkins}
+ \affiliation{\institution{Microsoft Research, New York City}}
+ \author{Zhiwei Steven Wu}
+ \affiliation{\institution{University of Minnesota - Twin Cities}}
+
+ \authorsaddresses{%
+  Authors' addresses: Guy Aridor, Columbia University, New York, NY, g.aridor@columbia.edu; Kevin Liu, Columbia University, New York, NY, kevin.liu@columbia.edu; Aleksandrs Slivkins, Microsoft Research, New York, NY, slivkins@microsoft.com; Zhiwei Steven Wu, University of Minnesota - Twin Cities, Minneapolis, MN, zsw@umn.edu}
 
@@
 
-\newpage
-
 \subfile{content/introduction}

### 2019-04-30 paper/content/introduction.tex
some edits

@@
 competition tension arises in other application domains: recommendation systems, news and entertainment websites, online commerce, and so forth.
-
-\OMIT{\gaedit{Bing experimenting with a ranking algorithm that ends up with a poor user experience could lead to negative short-run consequences in consumers' perceived quality of Bing and lead to more consumers choosing Google over Bing.
-However, this experimentation can help Bing eventually find a better ranking algorithm that would lead to positive long-run consequences in consumers' perceived quality of Bing and lead to more consumers choosing Bing over Google.}}
 
@@
 Implicitly, users have no ``outside option": the service provided is an improvement over not having it (and therefore the monopolist is not incentivized to deploy better learning algorithms).
-This is plausible with free ad-supported platforms such as Yelp or Google.}}
+This is plausible with free ad-supported platforms such as Yelp or Google.}}\footnote{\gaedit{The most common measures of market ``competitiveness" such as the Lerner Index or the Herfindahl-Hirschman Index of a market rely on ex-post observable attributes of a market such as prices or market shares \cite{tirole1988theory}.
+However, since we have no prices and market shares are an endogenous element of our model, we simply define ``competitiveness" in terms of the number of firms and timing of entry into the market.}}
 
@@
 We find strong evidence of the ``death spiral" effect mentioned earlier; this effect is strongest under permanent duopoly.
-
-%and that even increasing the number of firms in the market past a duopoly makes consumers weakly worse off.\swcomment{why ``even''?}
 
@@
 
-\OMIT{ %%%%%%
-\swedit{Beyond the welfare implications, we also investigate the barrier to entry that follows from the incumbent's data advantage.
-We run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage.
-We find that the data advantage from the incumbent's ``small head start'' can lead to a significant incumbency advantage, especially when the incumbent commits to a more advanced bandit algorithm.
-Our results show that even a small degree of data asymmetry can be amplified to be a larger market advantage through competition dynamics.
-\gadelete{This is in contrast to the data advantage in a static environment such as discussed in \cite{varian2018artificial,bajari2018impact}, in which a small amount of additional data can only provide limited improvement in the predictive accuracy.} The effect is stronger when the incumbent commits to a more advanced bandit algorithm due to
-}} %%%%%%
-% in a competitive environment the relative differences in the data
-% possessed by the firms may lead to a large competitive benefit due
-% to the feedback loops discussed previously. In our model this effect
-% is amplified when the incumbent commits to a more advanced bandit
-% algorithm since, in the incumbency period, the firm gathers data on
-% possible actions more efficiently leading to the relative gap upon
-% entry to be larger. This is important when exploration can have
-% larger reputational consequences.
-
-\OMIT{Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy.
-Being first in the market gives firms free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
-We investigate which of the two is a stronger barrier to entry.
-We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed.
-However, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.
-One take-away is that the data advantage is not just about data quantity but also data quality.}
-
 We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition.
@@
 \footnote{\asedit{Reputation scores already introduce some noise into users' choices.
-However, the amount of noise due to this channel is typically small, both in our simulations and in practice, because reputation signals average over many datapoints.}}
-
- We confirm the theoretical intuition that better algorithms prevail if the parameter -- essentially, the expected number ``random users" -- is sufficiently large.
-However, we find that this effect is negligible for some smaller but still ``relevant" parameter values.
+However, the amount of noise due to this channel is typically small, both in our simulations and in practice, because reputation signals average over many datapoints.}} We confirm the theoretical intuition that better algorithms prevail if the \gaedit{expected number of ``random" users is sufficiently large.}\gadelete{parameter -- essentially, the expected number ``random users" --is sufficiently large.} However, we find that this effect is negligible for some smaller but still ``relevant" parameter values.
 
@@
 %
-mean reputation, most connected to our intuition, is sometimes bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited});
+mean reputation, most connected to our intuition, is sometimes \gaedit{a} bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited});
 %
-mathematical tools from regret-minimization would only produce ``asymptotic" results, which does not seem to suffice.
-We believe that resolving first-order theoretical questions would not add much value to this paper, because of \cite{CompetingBandits-itcs18}.
-We are in realm of ``asymptotic" theorems about stylized Economic models, which provide some mathematical intuition but no direct bearing on the world.
+mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice.
+We believe that resolving first-order theoretical questions would not add much value to this paper, because of \cite{CompetingBandits-itcs18}.\gadelete{We are in \gaedit{the} realm of ``asymptotic" theorems about stylized economic models, which provide some mathematical intuition but no direct bearing on the world.
 But \cite{CompetingBandits-itcs18} already has an elaborate set of such results, with similar high-level conclusions.
-So, one would need to go beyond the first-order conclusions to add value.}
+So, one would need to go beyond the first-order conclusions to add value.}}
 
@@
 The study of competition vs. exploration has been initiated in \cite{CompetingBandits-itcs18}, as discussed above.
-\OMIT{Their model differs from ours in two key respects.
-First, users do not see any signal about firms' past performance, and instead choose between firms according to the Bayesian-expected reward.
-Second, they vary the strength of competition using assumptions about (ir)rational consumer behavior, whereas we use early entry.
-Their results are purely theoretical; their model is amenable to proofs but not to numerical experiments.
-Their high-level conclusion is an inverted-U relationship between competition and innovation that is similar to ours.}
 Our setting is also closely related to the ``dueling algorithms" framework \cite{DuelingAlgs-stoc11}, but this framework considers offline / full feedback scenarios whereas we focus on online machine learning problems.

### 2019-04-30 paper/content/inverted_u.tex
Heavy-tail -> Heavy-Tail
some edits

@@
 Table~\ref{tab:duopoly_unif} reports similarly on the Uniform instance.
-We find that $\DG$ is a weakly dominant strategy for the Heavy-tail and Uniform instances, as long as $T_0$ is sufficiently small.
+We find that $\DG$ is a weakly dominant strategy for the Heavy-Tail and Uniform instances, as long as $T_0$ is sufficiently small.
 However, \TS is a weakly dominant strategy for the Needle-in-Haystack instance.
@@
 \begin{finding}
-\textit{When all firms deploy \DG, and start at the same time, welfare is weakly decreasing as the number of firms decreases.}
+\textit{When all firms deploy \DG, and start at the same time, welfare is weakly decreasing as the number of firms \gaedit{increases}}
 \end{finding}

### 2019-05-04 paper/ec19-abstract.tex
Full draft for the extended abstract.

--- 2019-04-30 paper/competing_bandits.tex
@@
 
-\begin{abstract}
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below.
+%
+\maketitle
+\keywords{Multi-armed bandits,Exploration,Competition,Competition vs Innovation}
+
 We empirically study the interplay between \textit{exploration} and
@@
 We ask whether better exploration algorithms are incentivized under competition.
-We run extensive numerical experiments in a stylized duopoly model in which two firms deploy multi-armed bandit algorithms and compete for myopic users.
-We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore and leads to low consumer welfare, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms and lead to higher consumer welfare.
-Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.
-\end{abstract}
+We run extensive numerical experiments in a stylized duopoly model, where two firms compete for users and simultaneously learn from them.
+Each firm commits to a multi-armed bandit algorithm, and \emph{explores} according to this algorithm.
+Users select between the two firms based on the current reputation score: rewards from the firm's algorithm, averaged over a recent time window.
+Each firm's objective is to maximize its market share (the fraction of users choosing this firm).
+We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with a first-mover.
+We focus on three classes of bandit algorithms, ranging from more primitive to more sophisticated: \emph{greedy algorithms} that do not explicitly explore, algorithms that separate exploration and exploitation, and algorithms that combine the two.
+We know from prior work that in the absence of competition,  ``better" algorithms are typcially better in the long run, but could be worse initially.
 
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below.
-%
-\maketitle
-\keywords{Multi-armed bandits,Exploration,Competition,Competition vs Innovation}
+\xhdr{Main findings.}
+We find that in the permanent duopoly, competition incentivizes firms to choose the ``greedy algorithm", and even more so if the firm is a late arriver in a market.
+This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
+Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm.
+As a result, consumer welfare is highest under temporary monopoly.
+We find strong evidence of the ``death spiral" effect mentioned earlier; this effect is strongest under permanent duopoly.
 
-\subfile{content/introduction}
+Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
+This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
+However, the ``inverted-U relationship'' is driven by different aspects in our model than the ones in the existing literature in economics:  the barriers for innovations arise due to the reputational consequences of exploration rather than the R\&D costs.
 
-\subfile{content/model}
+\tikzstyle{level 1}=[level distance=3.5cm, sibling distance=4.0cm]
+\tikzstyle{level 2}=[level distance=3.5cm, sibling distance=2cm]
+\tikzstyle{below} = [align=center]
 
-\subfile{content/perf_in_iso}
+\begin{figure}[t]
+\begin{center}
+\begin{tikzpicture}
+      \draw[->] (-.5,0) -- (7,0) node[right] {};
+      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
+      \draw[scale=0.6,domain=0.7:9.8,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - 0.18 * (\x - 5.25)^2});
+     \node[below] at (0.7, -0.22) {\footnotesize monopoly};
+     \node[below] at (3, -0.2) {\footnotesize incumbent};
+     \node[below] at (4.5, -0.22) {\footnotesize duopoly};
+     \node[below] at (6, -0.2) {\footnotesize entrant};
+ \end{tikzpicture}
+ \caption{A stylized ``inverted-U relationship" between strength of competition and ``level of innovation".}
+\label{fig:inverted-U}
+\end{center}
+\end{figure}
 
-\subfile{content/inverted_u}
+\xhdr{Additional findings.}
+We investigate the ``first-mover advantage" phenomenon in more detail.
+Being first in the market gives free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
+We run additional experiments so as to isolate and compare these two effects.
+We find that either effect alone leads to a significant advantage under competition.
+The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.
 
-\subfile{content/barriers}
+Data advantage is significant from the anti-trust perspective, as a possible barrier to entry.
+We find that even a small amount ``data advantage" gets amplified under competition, causing a large difference in eventual market shares.
+This observation runs contrary to prior work,  %\cite{varian2018artificial,lambrecht2015can,bajari2018impact},
+which studied learning without competition, and found that small amounts of additional data do not provide significant improvement in eventual outcomes.
+We conclude that competition dynamics -- that firms compete as they learn over time -- are pertinent to these anti-trust considerations.
 
-\subfile{content/revisited}
+We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition.
+We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor.
+We suggest a
+more refined performance measure, and use it to explain some of the competition outcomes.
 
-\subfile{content/non_greedy_choice}
+We also consider an alternative choice rule with explicit noise/randomness: a small fraction of users choose a firm uniformly at random.
+We confirm the theoretical intuition that better algorithms prevail if the expected number of ``random" users is sufficiently large.
+However, we find that this effect is negligible for some smaller but still ``relevant" parameter values.
 
-\subfile{content/conclusion}
-\newpage
+\xhdr{Discussion.}
+Our paper is an experimental counterpart to \cite{CompetingBandits-itcs18}, which considered a similar duopoly model and obtained a number of theoretical results with ``asymptotic" flavor.
+For analytical tractability, \cite{CompetingBandits-itcs18} makes a somewhat unrealistic simplification that users do not observe any signals about firms' ongoing performance.
+The strength of competition is varied using assumptions about (ir)rational user behavior.
+For these reasons, the theorems from \cite{CompetingBandits-itcs18} have no direct bearing on our simulations.
+However, their high-level conclusion in is an inverted-U relationship similar to ours.
+
+The present paper provides a more nuanced and ``non-asymptotic" perspective.
+The reputation-based choice model accounts for competition in a more direct way, allows to separate reputation vs.
+data advantage, and makes our model amenable to numerical simulations (unlike the model in \cite{CompetingBandits-itcs18}).
+
 \bibliographystyle{acm}
 \bibliography{refs,bib-abbrv-short,bib-bandits,bib-AGT,bib-slivkins}
-
-\subfile{content/appendix_for_one_version}
 

### 2019-05-05 paper/ec19-abstract.tex
Added pointer to the arxiv.
update abstract

@@
-\documentclass[format=acmsmall, review=false]{acmart}
+\documentclass[acmsmall]{ec19acm}
 
-\usepackage{acm-ec-19}
+\settopmatter{printacmref=true}
+\fancyhead{}
+\copyrightyear{2019} 
+\acmYear{2019} 
+\setcopyright{rightsretained} 
+\acmConference[EC '19]{ACM EC '19: ACM Conference on Economics and Computation}{June 24--28, 2019}{Phoenix, AZ, USA}
+\acmBooktitle{ACM EC '19: ACM Conference on Economics and Computation (EC '19), June 24--28, 2019, Phoenix, AZ, USA}
+\acmDOI{10.1145/3328526.3329635}
+\acmISBN{978-1-4503-6792-9/19/06}
+
 \usepackage{tikz} % SW: Added for plots
@@
 %\usepackage{color-edits}
-\usepackage[suppress]{color-edits}  % use this to suppress the package
+\usepackage{color-edits}  % use this to suppress the package
 \addauthor{as}{red}    % as for Alex
@@
 
-% Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
 \begin{document}
@@
 {The Perils of Exploration under Competition: \\ A Computational Modeling Approach}
+
+\titlenote{Extended abstract. Full version is available as a technical report at \url{https://arxiv.org/abs/1902.05590}~.}
 
@@
 
+  \begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10010405.10010455.10010460</concept_id>
+<concept_desc>Applied computing~Economics</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Applied computing~Economics}
+
+\fancyhead{} 
 %
@@
 %
+\keywords{Multi-armed bandits;Exploration;Competition;Competition vs Innovation}
 \maketitle
-\keywords{Multi-armed bandits,Exploration,Competition,Competition vs Innovation}
 
@@
 
+\gadelete{
 \xhdr{Discussion.}
@@
 
+}
+
 \bibliographystyle{acm}

### 2019-05-05 paper/ec19-abstract.tex
updated abstract

@@
 \usepackage{subfiles}
+\usepackage{wrapfig}
 \usepackage[toc,page]{appendix}
@@
 
+\tikzstyle{level 1}=[level distance=2.0cm, sibling distance=4.0cm]
+\tikzstyle{level 2}=[level distance=2.0cm, sibling distance=2cm]
+\tikzstyle{below} = [align=center]
+
 \xhdr{Main findings.}
@@
 
-\tikzstyle{level 1}=[level distance=3.5cm, sibling distance=4.0cm]
-\tikzstyle{level 2}=[level distance=3.5cm, sibling distance=2cm]
-\tikzstyle{below} = [align=center]
-
-\begin{figure}[t]
+\begin{wrapfigure}{r}{0.6\textwidth}
 \begin{center}
 \begin{tikzpicture}
-      \draw[->] (-.5,0) -- (7,0) node[right] {};
+      \draw[->] (-.5,0) -- (6,0) node[right] {};
       \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
@@
 \end{center}
-\end{figure}
+\end{wrapfigure}
 
@@
 
-\gadelete{
 \xhdr{Discussion.}
@@
 For these reasons, the theorems from \cite{CompetingBandits-itcs18} have no direct bearing on our simulations.
-However, their high-level conclusion in is an inverted-U relationship similar to ours.
+However, their high-level conclusion \gadelete{in} is an inverted-U relationship similar to ours.
 
@@
 data advantage, and makes our model amenable to numerical simulations (unlike the model in \cite{CompetingBandits-itcs18}).
-
-}
 

### 2019-05-05 paper/ec19-abstract.tex
A few edits for space-saving.

@@
 This algorithm also prevails under monopoly, simply because it tends to be easier to deploy.
-Whereas the incumbent in the temporary monopoly is incentivized to deploy a more advanced exploration algorithm.
-As a result, consumer welfare is highest under temporary monopoly.
-We find strong evidence of the ``death spiral" effect mentioned earlier; this effect is strongest under permanent duopoly.
-
-Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
-This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
-However, the ``inverted-U relationship'' is driven by different aspects in our model than the ones in the existing literature in economics:  the barriers for innovations arise due to the reputational consequences of exploration rather than the R\&D costs.
-
-\begin{wrapfigure}{r}{0.6\textwidth}
+Whereas the incumbent in
+\begin{wrapfigure}{r}{0.5\textwidth}
+\vspace{-2mm}
 \begin{center}
 \begin{tikzpicture}
-      \draw[->] (-.5,0) -- (6,0) node[right] {};
-      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
+      \draw[->] (0,0) -- (6,0) node[right] {};
+      \draw[->] (0,0) -- (0,3) node[right] {Better algorithms};
       \draw[scale=0.6,domain=0.7:9.8,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - 0.18 * (\x - 5.25)^2});
@@
      \node[below] at (4.5, -0.22) {\footnotesize duopoly};
-     \node[below] at (6, -0.2) {\footnotesize entrant};
+     \node[below] at (5.7, -0.2) {\footnotesize entrant};
  \end{tikzpicture}
@@
 \end{wrapfigure}
+the temporary monopoly is incentivized to deploy a more advanced exploration algorithm.
+As a result, consumer welfare is highest under temporary monopoly.
+We find strong evidence of the ``death spiral" effect mentioned earlier; it is strongest under permanent duopoly.
+
+Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
+This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
+However, the ``inverted-U relationship'' is driven by different aspects in our model than the ones in the existing literature in economics:  the barriers for innovations arise due to the reputational consequences of exploration rather than the R\&D costs.
 

### 2019-05-05 paper/ec19-abstract.tex
More edits, fits two pages!

@@
 </concept>
+<concept>
+<concept_id>10010147.10010257.10010258.10010261.10010272</concept_id>
+<concept_desc>Computing methodologies~Sequential decision making</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10010147.10010257.10010282.10010284</concept_id>
+<concept_desc>Computing methodologies~Online learning settings</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
 </ccs2012>
@@
 \ccsdesc[500]{Applied computing~Economics}
+\ccsdesc[300]{Computing methodologies~Sequential decision making}
+\ccsdesc[300]{Computing methodologies~Online learning settings}
 
@@
 %
-\keywords{Multi-armed bandits;Exploration;Competition;Competition vs Innovation}
+\keywords{Multi-armed bandits; Exploration;Competition vs. Innovation}
 \maketitle
@@
 We focus on three classes of bandit algorithms, ranging from more primitive to more sophisticated: \emph{greedy algorithms} that do not explicitly explore, algorithms that separate exploration and exploitation, and algorithms that combine the two.
-We know from prior work that in the absence of competition,  ``better" algorithms are typcially better in the long run, but could be worse initially.
+In the absence of competition,  ``better" algorithms are are typically better in the long run, but could be worse initially.
 
@@
 Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the ``inverted-U relationship" between competition and innovation (see Figure~\ref{fig:inverted-U}).
-This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
-However, the ``inverted-U relationship'' is driven by different aspects in our model than the ones in the existing literature in economics:  the barriers for innovations arise due to the reputational consequences of exploration rather than the R\&D costs.
+This is a well-established concept in the economics literature,
+%, dating back to \cite{Schumpeter-42}, 
+whereby too little or too much competition is bad for innovation, but the middle range tends to be better.
+However, the ``inverted-U relationship'' is driven by different aspects in our model compared to the literature in economics:  the barriers for innovations arise due to the reputational consequences of exploration rather than the R\&D costs.
 
@@
 Data advantage is significant from the anti-trust perspective, as a possible barrier to entry.
-We find that even a small amount ``data advantage" gets amplified under competition, causing a large difference in eventual market shares.
+We find that even a small ``data advantage" gets amplified under competition, causing a large difference in eventual market shares.
 This observation runs contrary to prior work,  %\cite{varian2018artificial,lambrecht2015can,bajari2018impact},
-which studied learning without competition, and found that small amounts of additional data do not provide significant improvement in eventual outcomes.
+which studied learning \emph{without} competition, and found that small amounts of additional data do not provide significant improvement.
 We conclude that competition dynamics -- that firms compete as they learn over time -- are pertinent to these anti-trust considerations.
@@
 \xhdr{Discussion.}
-Our paper is an experimental counterpart to \cite{CompetingBandits-itcs18}, which considered a similar duopoly model and obtained a number of theoretical results with ``asymptotic" flavor.
-For analytical tractability, \cite{CompetingBandits-itcs18} makes a somewhat unrealistic simplification that users do not observe any signals about firms' ongoing performance.
+Our paper is an experimental counterpart to a theoretical paper \cite{CompetingBandits-itcs18}, which considered a similar duopoly model under a somewhat unrealistic simplification that users do not observe any signals about firms' ongoing performance.
 The strength of competition is varied using assumptions about (ir)rational user behavior.
-For these reasons, the theorems from \cite{CompetingBandits-itcs18} have no direct bearing on our simulations.
-However, their high-level conclusion \gadelete{in} is an inverted-U relationship similar to ours.
-
+While the theorems from \cite{CompetingBandits-itcs18} have no direct bearing on our simulations, their high-level conclusion is an inverted-U relationship similar to ours.
 The present paper provides a more nuanced and ``non-asymptotic" perspective.

### 2019-06-05 journal paper/itcs18 paper/app-perturb.tex
copied the files from itcs18

We show that Assumption~\eqref{eq:assn-distinct} holds almost surely under a small random perturbation of the prior. We focus on problem instances with 0-1 rewards, and assume that the prior $\priorMu$ is independent across arms and has a finite support.%
\footnote{The assumption of 0-1 rewards is for clarity. Our results hold under a more general assumption that for each arm $a$, rewards can only take finitely many values, and each of these values is possible (with positive probability) for every possible value of the mean reward $\mu_a$.}
 Consider the probability vector in the prior for arm $a$:
\[ \vec{p}_a = \left(\; \Pr[\mu_a=\nu]:\; \nu\in \support(\mu_a)\; \right).\]
We apply a small random perturbation independently to each such vector:
\begin{align}\label{eq:app:perturb:noise}
\vec{p}_a \leftarrow \vec{p}_a + \vec{q}_a,
    \quad \text{where}\quad \vec{q}_a\sim  \mN_a.
\end{align}
Here $\mN_a$ is the noise distribution for arm $a$: a distribution over real-valued, zero-sum vectors of dimension $d_a = |\support(\mu_a)|$. We need the noise distribution to satisfy the following property:
\begin{align}\label{eq:app:perturb:noise-prop}
\forall x\in [-1,1]^{d_a}\setminus \{0\}\qquad
\Pr_{q\sim \mN_a}\left[ x\cdot(\vec{p}_a+ q) \neq 0 \right] =1.
\end{align}

\begin{theorem}\label{thm:perturb}
Consider an instance of MAB with 0-1 rewards. Assume that the prior $\priorMu$ is independent across arms, and each mean reward $\mu_a$ has a finite support that does not include $0$ or $1$. Assume that noise distributions $\mN_a$ satisfy property \eqref{eq:app:perturb:noise-prop}. If random perturbation~\eqref{eq:app:perturb:noise} is applied independently to each arm $a$, then \refeq{eq:assn-distinct} holds almost surely for each history $h$.
\end{theorem}

\begin{remark}
As a generic example of a noise distribution which satisfies
  Property \eqref{eq:app:perturb:noise-prop}, consider the uniform
  distribution $\mN$ over the bounded convex set
  \[ Q = \left\{q \in \R^{d_a} \mid q \cdot \vec{1} = 0 \mbox{ and }
      \|q\|_2 \leq \eps\right\}, \] where $\vec{1}$ denotes the all-1
  vector. If $x = a \vec{1}$ for some non-zero value of $a$, then
  \eqref{eq:app:perturb:noise-prop} holds because
  $$x \cdot (p+q) = x \cdot p = a\neq 0.$$ Otherwise, denote
  $p=\vec{p}_a$ and observe that $x\cdot({p}+ {q}) = 0$ only if
  $x \cdot q = c \triangleq x \cdot (-p)$. Since $x\neq \vec{1}$, the
  intersection $Q\cap\{ x\cdot q = c \}$ either is empty or has
  measure 0 in $Q$, which implies
  $\Pr_{{q}}\left[ x\cdot({p}+ {q}) \neq 0 \right] =1$.
\end{remark}

To prove Theorem~\ref{thm:perturb}, it suffices to focus on two arms, and perturb one of them. Since realized rewards have finite support, there are only finitely many possible histories. Therefore, it suffices to focus on a fixed history $h$.

\begin{lemma}\label{lm:perturb}
Consider an instance of MAB with 0-1 rewards. Assume that the prior $\priorMu$ is independent across arms, and that $\support(\mu_1)$ is finite and does not include $0$ or $1$. Fix history $h$. Suppose random perturbation~\eqref{eq:app:perturb:noise} is applied to arm $1$, with noise distribution $\mN_1$ that satisfies \eqref{eq:app:perturb:noise-prop}. Then
    $\E[\mu_1\mid h] \neq \E[\mu_2\mid h] $
almost surely.
\end{lemma}

\begin{proof}
Note that $\E[\mu_a\mid h]$ does not depend on the algorithm which produced this history. Therefore, for the sake of the analysis, we can assume w.l.o.g. that this history has been generated by a particular algorithm, as long as this algorithm can can produce this history with non-zero probability. Let us consider  the algorithm that deterministically chooses same actions as $h$.

Let $S = \support(\mu_1)$. Then:
\begin{align*}
\E[\mu_1\mid h]
    &= \sum_{\nu\in S} \nu \cdot \Pr[\mu_1 =\nu \mid h]
    = \sum_{\nu\in S} \nu \cdot \Pr[h \mid \mu_1 =\nu] \cdot \Pr[\mu_1=\nu]\;/\;\Pr[h], \\
\Pr[h]
    &= \sum_{\nu\in S} \Pr[h \mid \mu_1 =\nu] \cdot \Pr[\mu_1=\nu].
\end{align*}
Therefore,
    $\E[\mu_1\mid h] = \E[\mu_2\mid h] $
if and only if
\begin{align*}
\sum_{\nu\in S} (\nu-C) \cdot \Pr[h \mid \mu_1 =\nu] \cdot \Pr[\mu_1=\nu] = 0,
\quad\text{ where }\quad C=\E[\mu_2\mid h].
\end{align*}
Since $\E[\mu_2\mid h]$ and $\Pr[h \mid \mu_1 =\nu]$ do not depend on the probability vector $\vec{p}_1$, we conclude that
\begin{align*}
  \E[\mu_1\mid h] = \E[\mu_2\mid h]
\quad\Leftrightarrow\quad x\cdot \vec{p}_1 =0,
\end{align*}
where vector
    \[ x := \left(\; (\nu-C) \cdot \Pr[h \mid \mu_1 =\nu]:\; \nu\in S\; \right) \in [-1,1]^{d_1}\]
does not depend on $\vec{p}_1$.

Thus, it suffices to prove that $x\cdot \vec{p}_1 \neq 0$ almost surely under the perturbation. In a formula:
\begin{align}\label{eq:lm:perturb-pf}
    \Pr_{q\sim \mN_1}\left[ x\cdot( \vec{p}_1+q) \neq 0 \right] =1
\end{align}

Note that $\Pr[h \mid \mu_1 =\nu]>0$ for all $\nu\in S$, because $0,1\not\in S$. It follows that at most one coordinate of $x$ can be zero. So  \eqref{eq:lm:perturb-pf} follows from property \eqref{eq:app:perturb:noise-prop}.
\end{proof}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main.tex"
%%% End:

### 2019-06-05 journal paper/itcs18 paper/sec-welfare.tex
copied the files from itcs18

\asedit{We frame our contributions in terms of the relationship between \competitiveness and \rationality on one side, and adoption of better algorithms on the other. Recall that both \competitiveness (of the game between the two principals) and \rationality (of the agents) are controlled by the response function $\respF$.}

\OMIT{ %%%%%%
We frame our contributions in terms of the relationship between \competition and \innovation, \ie between the extent to which the game between the two principals is competitive, and the degree of innovation --- adoption of better that these models incentivize. \Competition is controlled via the response function $\respF$, and \innovation refers to the quality of the technology (MAB algorithms) adopted by the principals. The \competition vs. \innovation relationship is well-studied in the economics literature, and is commonly known to often follow an inverted-U shape, as in \reffig{fig:inverted-U} (see Section~\ref{sec:related-work} for citations). \Competition in our models is closely correlated with \rationality: the extent to which agents make rational decisions, and indeed \rationality is what $\respF$ controls directly.
} %%%%%%%%

\xhdr{Main story.}
Our main story concerns the restricted competition game between the two principals where one allowed algorithm \alg is ``better" than the others. \asedit{We track whether and when \alg is chosen in an equilibrium.} We vary \competitiveness/\rationality by changing the response function from \HardMax (full rationality, very competitive environment) to \HardMaxRandom to  \SoftMaxRandom (less rationality and competition). Our conclusions are as follows:
\begin{OneLiners}
\item Under \HardMax, no innovation: \DynGreedy is chosen over \alg.
\item Under \HardMaxRandom, some innovation:  \alg is chosen as long as it BIR-dominates.
\item Under \SoftMaxRandom, more innovation: \alg is chosen as long as it weakly-BIR-dominates.%
\footnote{This is a weaker condition, the better algorithm is chosen in a broader range of scenarios.}
\end{OneLiners}
These conclusions follow, respectively, from Corollaries~\ref{cor:DG-dominance}, \ref{cor:random} and \ref{cor:SoftMax}. Further, \asedit{we consider the uniform choice between the principals. It corresponds to the least amount of rationality and competition, and (when principals' utility is the number of agents) uniform choice provides no incentives to innovate.}%
\footnote{On the other hand, if principals' utility is somewhat aligned with agents' welfare, as in \eqref{eq:general-utility}, then a monopolist principal is incentivized to choose the best possible MAB algorithm (namely, to minimize cumulative Bayesian regret $\BReg(T)$). Accordingly, monopoly would result in better social welfare than competition, as the latter is likely to split the market and cause each principal to learn more slowly. This is a very generic and well-known effect regarding economies of scale.}
Thus, we have an inverted-U relationship, see \reffig{fig:inverted-U2}.

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=1]
      \draw[->] (-.5,0) -- (9.5,0) node[above] 
        {\qquad\qquad Competitiveness/Rationality};
      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithm in equilibrium};
      \draw[scale=0.8,domain=0.5:9.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{3.5 - 0.15*(\x - 5)^2});
     \node[below] at (1, 0) {\footnotesize \Uniform};
     \node[below] at (3.9, 0) {\footnotesize \SoftMaxRandom};
     \node[below] at (6, 0) {\footnotesize \HardMaxRandom};
     \node[below] at (8, 0) {\footnotesize \HardMax};
      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
 \end{tikzpicture}

\caption{The stylized inverted-U relationship in the ``main story".}
\label{fig:inverted-U2}
\end{center}
\end{figure}

\xhdr{Secondary story.}
Let us zoom in on the symmetric  \HardMaxRandom model. \asedit{Competitiveness and rationality within this model are controlled by the baseline probability $\eps_0 = \respF(- 1)$, which goes smoothly between the two extremes of \HardMax ($\eps_0=0$) and the uniform choice ($\eps_0=\tfrac12$). Smaller $\eps_0$ corresponds to increased rationality and increased competitiveness.} For clarity, we assume that principal's utility is the number of agents.

We consider the marginal utility of switching to a better algorithm. Suppose initially both principals use some algorithm \alg, and principal 1 ponders switching to another algorithm \alg' which BIR-dominates \alg. \asedit{We are interested in the marginal utility of this switch. Then:}

\begin{itemize}
\item $\eps_0 = 0$ (\HardMax):~~~~ the marginal utility can be negative if \alg is \DynGreedy.

\item $\eps_0$ near $0$:~~~~ only a small marginal utility can be guaranteed, as it may take a long time for $\alg'$ to ``catch up" with \alg, and hence less time to reap the benefits.

\item ``medium-range" $\eps_0$:~~~~ large marginal utility, as $\alg'$ learns fast and gets most agents.

\item $\eps_0$ near $\tfrac12$:~~~~ small marginal utility, as principal 1 gets most agents for free no matter what.
\end{itemize}
The familiar inverted-U shape is depicted in Figure~\ref{fig:inverted-U3}.

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=1]
      \draw[->] (-.5,0) -- (9.5,0) node[above]  {$\eps_0$};
      \draw[->] (0,-.5) -- (0,3) node[above] {marginal utility};
      \draw[scale=0.8,domain=0.5:9.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{3.5 - 0.15*(\x - 5)^2});
     \node[below] at (.6, 0) {\footnotesize \Uniform};
     \node[above] at (.5, 0) {\footnotesize 0};
     % \node[below] at (3.9, 0) {\footnotesize \SoftMaxRandom};
     % \node[below] at (6, 0) {\footnotesize \HardMaxRandom};
     \node[below] at (7.5,0) {\footnotesize \HardMax};
     \node[above] at (7.5, 0) {\footnotesize 1/2};
      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
 \end{tikzpicture}

\caption{The stylized inverted-U relationship from the ``secondary story"}
\label{fig:inverted-U3}
\end{center}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2019-06-17 journal paper/itcs18 paper/abstract.tex

Most modern systems strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information. We initiate a study of the interplay between \emph{exploration and competition}---how such systems balance the exploration for learning and the competition for users. Here the users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing systems.

In our model, we consider competition between two multi-armed bandit algorithms faced with the same bandit instance. Users arrive one by one and choose among the two algorithms, so that each algorithm makes progress if and only if it is chosen.
%We ask whether better algorithms perform better in such competition,
% and whether the competition incentivizes better algorithms
% and improves social welfare.
We ask whether and to what extent competition incentivizes \asedit{the adoption of better bandit algorithms}. We investigate this issue for several models of user response, as we vary the degree of rationality and competitiveness in the model. \asedit{Our findings are closely related to} the ``competition vs. innovation" relationship, a well-studied theme in economics.

### 2019-06-17 journal paper/itcs18 paper/sec-random.tex

This section is dedicated to the \HardMaxRandom response model, where each principal is always chosen with some positive baseline probability. The main technical result for this model states that a principal with asymptotically better \BIR wins by a large margin: after a ``learning phase" of constant duration, all agents choose this principal with maximal possible probability $\respF(1)$. For example, a principal with $\BIR(n)\leq \tilde{O}(n^{-1/2})$ wins over a principal with $\BIR(n)\geq \Omega(n^{-1/3})$. However, this positive result comes with a significant caveat detailed in Section~\ref{sec:random-greedy}.

We formulate and prove a cleaner version of the result, followed by a more general formulation developed in a subsequent Remark~\ref{rem:random-messy}. We need to express a property that \alg[1] eventually catches up and surpasses \alg[2], even if initially it receives only a fraction of traffic. For the cleaner version, we assume that both algorithms are well-defined for an infinite time horizon, so that their \BIR does not depend on the time horizon $T$ of the game.  Then this property can be formalized as:
\begin{align}\label{eq:random-better-clean}
(\forall \eps>0)\qquad
\BIR_1(\eps n)/\BIR_2(n) \to 0.
\end{align}
In fact, a weaker version of \eqref{eq:random-better-clean} suffices:
denoting $\eps_0 = \respF(-1)$, for some constant $n_0$ we have
\begin{align}\label{eq:random-better-weaker}
(\forall n\geq n_0) \qquad
\BIR_1(\eps_0 n/2)/\BIR_2(n) <\tfrac12.
\end{align}
We also need a very mild technical assumption on the ``bad" algorithm:
\begin{align}\label{eq:random-assn}
 (\forall n\geq n_0) \qquad
  \BIR_2(n) > 4\,e^{-\eps_0 n/12}.
\end{align}

\begin{theorem}\label{thm:random-clean}
Assume \HardMaxRandom response function. Suppose both algorithms are monotone and well-defined for an infinite time horizon, and satisfy~\eqref{eq:random-better-weaker} and~\eqref{eq:random-assn}. Then each agent $t\geq n_0$ chooses principal $1$ with maximal possible probability $\respF(1) = 1- \eps_0$.
\end{theorem}

\begin{proof}
Consider global round $t\geq n_0$. Recall that each agent chooses principal $1$ with probability at least
    $\respF(-1)>0$.

Then
    $\E[n_1(t+1)] \geq 2\eps_0\,t $.
By Chernoff Bounds (Theorem~\ref{thm:chernoff}), we have that
    $n_1(t+1)\geq \eps_0 t$
holds with probability at least $1-q$,
where $q = \exp(-\eps_0 t/12)$.

We need to prove that
    $\PMR_1(t) - \PMR_2(t)>0$.
For any $m_1$ and $m_2$, consider the quantity
\[ \Delta(m_1,m_2) := \BIR_2(m_2+1) - \BIR_1(m_1+1).\]
Whenever $m_1 \geq \eps_0 t/2 -1$ and $m_2<t$, it holds that
\begin{align*}
\Delta(m_1,m_2) \geq \Delta(\eps_0 t / 2, t)
    \geq \BIR_2(t)/2.
\end{align*}
The above inequalities follow, resp., from algorithms' monotonicity and \eqref{eq:random-better-weaker}. Now,
\begin{align*}
\PMR_1(t) - \PMR_2(t)
    &= \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}{\Delta(m_1,m_2)} \\
    &\geq -q
        + \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}
            {\Delta(m_1,m_2)\mid m_1 \geq \eps_0 t/2-1} \\
    &\geq \BIR_2(t)/2-q \\
    &> \BIR_2(t)/4 > 0  \EqComment{by \eqref{eq:random-assn}}. \qedhere
\end{align*}
\end{proof}

\begin{remark}\label{rem:random-messy}
  Many standard MAB algorithms in the literature are parameterized by
  the time horizon $T$. Regret bounds for such algorithms usually
  include a polylogarithmic dependence on $T$. In particular, a
  typical upper bound for \BIR has the following form:
  \OMIT{\footnote{We
    provide upper bounds on \BIR for several standard MAB algorithms
    to illustrate these dependencies in the appendix.\sw{added}}}
\begin{align}
    \BIR(n\mid T)\leq \polylog(T)\cdot n^{-\gamma}
    \quad \text{for some $\gamma\in(0,\tfrac12]$}.
\end{align}
Here we write $\BIR(n\mid T)$ to emphasize the dependence on $T$.

We generalize \eqref{eq:random-better-weaker} to handle the dependence
on $T$: there exists a number $T_0$ and a function $n_0(T)\in \polylog(T)$ 
such that 
\begin{align}\label{eq:random-better-messy}
(\forall T\geq T_0,\; n\geq n_0(T)) \quad
\frac{\BIR_1(\eps_0 n /2\mid T)}{\BIR_2(n\mid T)} <\frac12.
\end{align}
If this holds, we say that \alg[1] \emph{BIR-dominates} \alg[2].

We provide a version of Theorem~\ref{thm:random-clean} in which algorithms are parameterized with time horizon $T$ and condition \eqref{eq:random-better-weaker} is replaced with \eqref{eq:random-better-messy}; its proof is very similar and is omitted.
\end{remark}

%\begin{theorem}\label{thm:random-messy}
%Assume \HardMaxRandom response function, and algorithms that satisfy \eqref{eq:random-better-messy} and~\eqref{eq:random-assn}. Then each agent
%    $t\geq n_0(T)$ chooses principal $1$ with probability $\respF(1)$.
%\end{theorem}

%Let \alg be a monotone MAB algorithm and $\mA$ be a finite set of monotone MAB algorithms such that each algorithm in $\mA$ satisfies \eqref{eq:random-assn} and is ``dominated" by \alg in the sense of \eqref{eq:random-better-messy}.
%Assume principals can only choose algorithms from $\mA\cup \{\alg\}$.

To state a game-theoretic corollary of Theorem~\ref{thm:random-clean}, we consider a version of the competition game between the two principals in which they can only choose from a finite set $\mA$ of monotone MAB algorithms. One of these algorithms is ``better" than all others; we call it the \emph{special} algorithm. Unless specified otherwise, it BIR-dominates all other allowed algorithms. The other algorithms satisfy \eqref{eq:random-assn}. We call this game the \emph{restricted competition game}.

\begin{corollary}\label{cor:random}
Assume \HardMaxRandom response function. Consider the restricted competition game with special algorithm \alg. Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium: both principals choose \alg.
\end{corollary}

\subsection{A little greedy goes a long way}
\label{sec:random-greedy}

Given any monotone MAB algorithm other than \DynGreedy, we design a modified algorithm which learns at a slower rate, yet ``wins the game" in the sense of Theorem~\ref{thm:random-clean}. As a corollary, the competition game with unrestricted choice of algorithms typically does not have a Nash equilibrium.

Given an algorithm \alg[1] that deviates from \DynGreedy starting from
step $n_0$ and a ``mixing'' parameter $p$, we will construct a
modified algorithm as follows.
\begin{enumerate}
\item The modified algorithm coincides with \alg[1] (and \DynGreedy)
for the first $n_0-1$ steps;
\item In each step $n\geq n_0$, \alg[1] is invoked with probability
  $1-p$, and with the remaining probability $p$ does the ``greedy
  choice": chooses an action with the largest posterior mean reward
  given the current information collected by \alg[1].
\end{enumerate}
%The modified algorithm is called \emph{greedy deviation}, with probability parameter $p$.
For a cleaner comparison between the two algorithms, the modified algorithm does not record rewards received in steps with the ``greedy choice". Parameter $p>0$ is the same for all steps.

\begin{theorem}\label{thm:random-greedy}
Assume symmetric \HardMaxRandom response function. Let $\eps_0 = \respF(-1)$ be the baseline probability. Suppose \alg[1] deviates from \DynGreedy starting from some step $n_0$. Let \alg[2] be the modified algorithm, as described above, with mixing parameter $p$ such that
    $(1-\eps_0)(1-p)>\eps_0$.
Then each agent $t\geq n_0$ chooses principal $2$ with maximal possible probability $1-\eps_0$.
\end{theorem}

\begin{corollary}\label{cor:random-greedy}
  Suppose that both principals can choose any monotone MAB algorithm, and assume the symmetric \HardMaxRandom response
  function. Then for any time
  horizon $T$, the only possible \emph{pure} Nash equilibrium is one
  where both principals choose \DynGreedy. Moreover, no pure Nash
  equilibrium exists when some algorithm ``dominates" \DynGreedy in
  the sense of \eqref{eq:random-better-messy} and the time horizon $T$
  is sufficiently large.
\end{corollary}

\begin{remark}
The modified algorithm performs exploration % ingests information
at a slower rate. Let us argue how this may translate  into a larger \BIR compared to the original algorithm. Let  $\BIR'_1(n)$ be the \BIR of the ``greedy choice" after after $n-1$ steps of \alg[1]. Then
\begin{align}\label{eq:random-BIR}
\BIR_2(n)
    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIR'_1(m)}.
\end{align}
In this expression, $m$ is the number of times \alg[1] is invoked in the first $n$ steps of the modified algorithm. Note that
    $\E[m] = n_0-1 + (n-n_0+1)(1-p) \geq (1-p)n$.

Suppose $\BIR_1(n)= \beta n^{-\gamma}$ for some constants $\beta,\gamma>0$. Further, assume 
    $\BIR'_1(n)\geq  c\; \BIR_1(n)$,
for some $c>1-\gamma$.
Then for all $n\geq n_0$ and small enough $p>0$ it holds that:
\begin{align*}
 \BIR_2(n) 
    &\geq  (1-p+pc)\; \E[\; \BIR_1(m) \;] \\
\E[\; \BIR_1(m) \;]
    &\geq \BIR_1(\; \E[m] \;) &\qquad\text{(By Jensen's inequality)} \\
    &\geq \BIR_1(\; (1-p)n \;) &\qquad\text{(since $\E[m]\geq n(1-p)$)}  \\
    &\geq \beta\cdot n^{-\gamma} \cdot (1-p)^{-\gamma} 
        &\qquad\text{(plugging in $\BIR_1(n)=\beta n^{-\gamma}$)}  \\
    &> \BIR_1(n)\;\; (1-p\gamma)^{-1}
        &\qquad\text{(since $(1-p)^\gamma < 1-p\gamma$)}.\\
\BIR_2(n) 
    &>\alpha\cdot \BIR_1(n),
    &\text{where} \quad 
    \alpha = \tfrac{1-p+pc}{1-p\gamma}>1.
\end{align*}
(In the above equations, all expectations are over $m$ distributed as in \eqref{eq:random-BIR}.)
\end{remark}

\begin{proof}[Proof of Theorem~\ref{thm:random-greedy}]
  Let $\rew'_1(n)$ denote the Bayesian-expected reward of the ``greedy
  choice'' after after $n-1$ steps of \alg[1]. Note that
  $\rew_1(\cdot)$ and $\rew'_1(\cdot)$ are non-decreasing: the former
  because \alg[1] is monotone and the latter because the ``greedy
  choice'' is only improved with an increasing set of
  observations. Therefore, the modified algorithm \alg[2] is monotone
  by \eqref{eq:random-BIR}.

  By definition of the ``greedy choice,'' $\rew_1(n)\leq \rew'_1(n)$
  for all steps $n$. Moreover, by Lemma~\ref{lm:DG-rew},
  \alg[1] has a strictly smaller $\rew(n_0)$ compared to \DynGreedy;
  so, $\rew_1(n_0)<\rew_2(n_0)$.\sw{fixed inequalities; but this paragraph is still not perfectly clear}

Let $\alg$ denote a copy of \alg[1] that is running ``inside" the modified algorithm \alg[2]. Let $m_2(t)$ be the number of global rounds before $t$ in which the agent chooses principal $2$ \emph{and} \alg is invoked; in other words, it is the number of agents seen by \alg before global round $t$. Let $\mM_{2,t}$ be the agents' posterior distribution for $m_2(t)$.

We claim that in each global round $t\geq n_0$, distribution $\mM_{2,t}$ stochastically dominates distribution $\posteriorN{1}{t}$, and $\PMR_1(t)<\PMR_2(t)$. We use induction on $t$. The base case $t=n_0$ holds because $\mM_{2,t} = \posteriorN{1}{t}$ (because the two algorithms coincide on the first $n_0-1$ steps), and $\PMR_1(n_0)<\PMR_2(n_0)$ is proved as in \eqref{eq:pf:thm:DG-dominance}, using the fact that $\rew_1(n_0)<\rew_2(n_0)$.

The induction step is proved as follows. The induction hypothesis for global round $t-1$ implies that agent $t-1$ is seen by \alg with probability $(1-\eps_0)(1-p)$, which is strictly larger than $\eps_0$, the probability with which this agent is seen by \alg[2]. Therefore, $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$.
\begin{align}
\PMR_1(t)
  &= \Ex{n\sim \posteriorN{1}{t}} {\rew_1(n+1)} \nonumber \\
  &\leq \Ex{m\sim \mM_{2,t}} {\rew_1(m+1)}
    \label{eq:pf:thm:random-greedy-1}\\
  &< \Ex{m\sim \mM_{2,t}} {(1-p)\cdot \rew_1(m+1) + p\cdot \rew'_1(m+1)}
    \label{eq:pf:thm:random-greedy-2} \\
  &= \PMR_2(t). \nonumber
\end{align}
Here inequality \eqref{eq:pf:thm:random-greedy-1} holds because $\rew_1(\cdot)$ is monotone and $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$, and inequality \eqref{eq:pf:thm:random-greedy-2} holds because $\rew_1(n_0)<\rew_2(n_0)$ and $\mM_{2,t}(n_0)>0$.%
\footnote{If $\rew_1(\cdot)$ is strictly increasing, then inequality \eqref{eq:pf:thm:random-greedy-1} is strict, too; this is because $\mM_{2,t}(t-1)>\posteriorN{1}{t}(t-1)$.  }
\end{proof}

%%% Local Variables:
%%% TeX-master: "main.tex"
%%% End:

### 2019-06-17 journal paper/itcs18 paper/sec-related-work.tex

Multi-armed bandits (\emph{MAB}) is a particularly elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information. MAB problems have been studied in Economics, Operations Research and Computer Science for many decades; see \citep{Bubeck-survey12,Gittins-book11,slivkins-MABbook} for background on regret-minimizing and Bayesian formulations, respectively. A discussion of industrial applications of MAB can be found in \citet{MWT-WhitePaper-2016}.

The literature on MAB is vast and multi-threaded. The most related
thread concerns regret-minimizing MAB formulations with IID rewards
\citep{Lai-Robbins-85,bandits-ucb1}. This thread includes ``smart" MAB
algorithms that combine exploration and exploitation, such as UCB1
\citep{bandits-ucb1} and Successive Elimination
\citep{EvenDar-icml06}, and ``naive'' MAB algorithms that separate
exploration and exploitation, including explore-first and
$\eps$-Greedy \citep[\eg see][]{slivkins-MABbook}.

The three-way tradeoff between exploration, exploitation and incentives has been studied in several other settings:
incentivizing exploration in a recommendation system
    \citep{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,ICexplorationGames-ec16-working},
dynamic auctions
    \cite[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
pay-per-click ad auctions with unknown click probabilities
    \cite[\eg][]{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
coordinating search and matching by self-interested agents
    \citep{Bobby-Glen-ec16},
as well as human computation
    \cite[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.

\citet{Bolton-econometrica99,Keller-econometrica05,Johari-ec12} studied models with self-interested agents jointly performing exploration, with no principal to coordinate them.

There is a superficial similarity (in name only) between this paper and the line of work on ``dueling bandits"
    \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
The latter is not about competing bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".

Our setting is closely related to the ``dueling algorithms" framework \citep{DuelingAlgs-stoc11} which studies competition between two principals, each running an algorithm for the same problem. However, this work considers algorithms for offline / full input scenarios, whereas we focus on online machine learning and the explore-exploit-incentives tradeoff therein. Also, this work specifically assumes binary payoffs (\ie win or lose) for the principals.

\xhdr{Other related work in economics.} The competition vs. innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{Aghion-QJE05,Vives-08}. Production costs aside, this literature treats innovation as a priori beneficial for the firm. Our setting is very different, as innovation in exploration algorithms may potentially hurt the firm.

A line of work on \emph{platform competition}, starting with \cite{Rysman09}, concerns competition between firms (\emph{platforms}) that improve as they attract more users (\emph{network effect}); see \citet{Weyl-White-14} for a recent survey. This literature is not concerned with \innovation, and typically models network effects exogenously, whereas in our model network effects are endogenous: they are created by MAB algorithms, an essential part of the model.

Relaxed versions of rationality similar to ours are found in several notable lines of work. For example, ``random agents" (a.k.a. noise traders) can side-step the ``no-trade theorem'' \citep{Milgrom-Stokey-82}, a famous impossibility result in financial economics. The \SoftMaxRandom model is closely related to the literature on \emph{product differentiation}, starting from \cite{Hotelling-29}, see \cite{Perloff-Salop-85} for a notable later paper.

There is a large literature on non-existence of equilibria due to small deviations   (which is related to the corresponding result for \HardMaxRandom), starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets. Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom.

% moved this thought to the intro.
\OMIT{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}

%%% Local Variables:
%%% TeX-master: "main.tex"
%%% End:

### 2019-06-17 journal paper/itcs18 paper/sec-soft.tex

This section is devoted to the \SoftMaxRandom model. We recover a positive result under the assumptions from Theorem~\ref{thm:random-clean} (albeit with a weaker conclusion), and then proceed to a much more challenging result under weaker assumptions. We start with a formal definition:

\begin{definition}\label{def:SoftMax}
A response function $\respF$ is \SoftMaxRandom if the following conditions hold:
\begin{OneLiners}
\item  $\respF(\cdot)$ is bounded away from $0$ and $1$:
    $\respF(\cdot)\in [\eps,1-\eps]$ for some $\eps\in (0,\tfrac12)$,
\item  the response function
 $\respF(\cdot)$ is ``smooth" around $0$:
 \begin{align}\label{eq:SoftMax-smooth}
 \exists\, \text{constants $\delta_0,c_0,c'_0>0$}
    \qquad \forall x\in [-\delta_0,\delta_0] \qquad
    c_0 \leq \respF'(x) \leq c'_0.
 \end{align}
\item fair tie-breaking: $\respF(0)=\tfrac12$.
\end{OneLiners}
\end{definition}

\begin{remark}
\asedit{This definition is fruitful when parameters $c_0$ and $c_0'$ are close to $\tfrac12$. Throughout, we assume that \alg[1] is better than \alg[2], and obtain results parameterized by $c_0$. By symmetry, one could assume that \alg[2] is better than \alg[1], and obtain similar results parameterized by $c_0'$.}
\end{remark}

Our first result is a version of Theorem~\ref{thm:random-clean}, with the same assumptions about the algorithms and essentially the same proof. The conclusion is much weaker: we can only guarantee that each agent $t\geq n_0$ chooses principal 1 with probability slightly larger than $\tfrac12$. This is essentially unavoidable in a typical case when both algorithms satisfy $\BIR(n)\to 0$, by Definition~\ref{def:SoftMax}.

\begin{theorem}\label{thm:SoftMax-weak}
  Assume \SoftMaxRandom response function. Suppose \alg[1] has better
  \BIR in the sense of \eqref{eq:random-better-weaker}, and \alg[2]
  satisfies the condition \eqref{eq:random-assn}. Then each agent
  $t\geq n_0$ chooses principal $1$ with probability
\begin{align}\label{eq:thm:SoftMax-weak}
     \Pr[i_t = 1]\geq \tfrac12 +  \tfrac{c_0}{4}\; \BIR_2(t).
\end{align}
\end{theorem}

\begin{proof}[Proof Sketch]
We follow the steps in the proof of Theorem~\ref{thm:random-clean} to derive \begin{align*}
\PMR_1(t) - \PMR_2(t)
    &\geq \BIR_2(t)/2 -q,
    \quad \text{where $q = \exp(-\eps_0 t/12)$.}
\end{align*}
This is at least $\BIR_2(t)/4$ by \eqref{eq:random-assn}. Then \eqref{eq:thm:SoftMax-weak} follows by the smoothness condition \eqref{eq:SoftMax-smooth}.
\end{proof}

\newcommand{\BReg}{\term{BReg}}

We recover a version of Corollary~\ref{cor:random}, if each principal's utility is the number of users (rather than the more general model in \eqref{eq:general-utility}). We also need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity. \BReg is a standard notion from the literature (along with \BIR):
\begin{align}\label{eq:SoftMax-BReg}
\BReg(n) := n\cdot \E_{\mu\sim\priorMu}
    \left[ \max_{a\in A} \mu_a\right] - \sum_{n=1}^n \rew(n')
    = \sum_{n'=1}^n \BIR(n').
\end{align}

\begin{corollary}\label{cor:SoftMax}
Assume that the response function is \SoftMaxRandom, and each principal's  utility is the number of users.
%
%Suppose principals can only choose algorithms from $\mA\cup \{\alg\}$, where $\mA\cup \{\alg\}$ is a finite set of monotone MAB algorithms such that each algorithm in $\mA$ satisfies \eqref{eq:random-assn} and $\BReg(n)\to \infty$, and is ``dominated" by \alg in the sense of \eqref{eq:random-better-messy}.
%
Consider the restricted competition game with special algorithm \alg, and assume that all other allowed algorithms satisfy $\BReg(n)\to \infty$. Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium: both principals choose \alg.
\end{corollary}

Further, we prove a much more challenging result in which the
condition \eqref{eq:random-better-weaker} is replaced with a much
weaker ``BIR-dominance'' condition. For clarity, we will again assume
that both algorithms are well-defined for an infinite time
horizon. The \emph{weak BIR dominance} condition says there exist
constants $\beta_0, \alpha_0\in (0, 1/2)$ and $n_0$ such that
 \begin{align}\label{eq:SoftMax-better}
   (\forall n\geq n_0) \quad
   \frac{\BIR_1((1-\beta_0)\, n)}{\BIR_2(n)} <1-\alpha_0.
 \end{align}
% for some $n_0(T)\in \polylog(T)$ and constants
% $\beta_0,\alpha_0\in (0, 1/2)$,
% \begin{align}\label{eq:SoftMax-better}
% (\forall n\geq n_0(T)) \quad
% \frac{\BIR_1((1-\beta_0)\, n\mid T)}{\BIR_2(n\mid T)} <1-\alpha_0.
% \end{align}
 If this holds, we say that \alg[1] \emph{weakly BIR-dominates}
 \alg[2]. Note that the condition \eqref{eq:random-better-messy}
 involves sufficiently small multiplicative factors (resp., $\eps_0/2$
 and $\tfrac12$), the new condition replaces them with factors that
 can be arbitrarily close to $1$.

 We make a mild assumption on \alg[1] that its $\BIR_1(n)$ tends to
 0. Formally, for any $\eps > 0$, there exists some $n(\eps)$ such
 that
 \begin{align}\label{eq:so-mild}
   (\forall n\geq n(\eps)) \qquad \BIR_1(n) \leq \eps.
 \end{align}
 We also require a slightly stronger version of the technical
 assumption~\eqref{eq:random-assn}:%  for any $\eps > 0$, there exists
 % $n(\eps)$ such that for
for some $n_0$,
\begin{align}\label{eq:SoftMax-assn-strong}
 % (\forall n\geq n(\eps)) \qquad
 %  \BIR_2(n) > e^{-\eps n}. %used to be 4exp( -\eps n/6)
(\forall n\geq n_0) \qquad 
\BIR_2(n) \geq \frac{4}{\alpha_0} \exp \left( \frac{-\min\{\eps_0, 1/8\} n}{12}\right)
\end{align}

\begin{theorem}\label{thm:SoftMax-strong}
  Assume the \SoftMaxRandom response function. Suppose \alg[1]
  weakly-BIR-dominates \alg[2], \alg[1] satisfies \eqref{eq:so-mild},
  and \alg[2] satisfies \eqref{eq:SoftMax-assn-strong}. Then there
  exists some $t_0$ such that each agent $t\geq t_0$ chooses principal
  $1$ with probability
\begin{align}\label{eq:thm:SoftMax-strong}
     \Pr[i_t = 1]\geq \tfrac12 +  \tfrac{c_0\alpha_0}{4}\; \BIR_2(t).
\end{align}
\end{theorem}

The main idea behind our proof is that even though \alg[1] may have a
slower rate of learning in the beginning, it will gradually catch up
and surpass \alg[2]. We will describe this process in two phases. In
the first phase, \alg[1] receives a random agent with probability at
least $\respF(-1) = \eps_0$ in each round. Since $\BIR_1$ tends to 0,
the difference in \BIR{s} between the two algorithms is also
diminishing. Due to the \SoftMaxRandom response function, \alg[1]
attracts each agent with probability at least $1/2 - O(\beta_0)$ after
a sufficient number of rounds. Then the game enters the second phase:
both algorithms receive agents at a rate close to $\tfrac12$, and the
fractions of agents received by both algorithms --- $n_1(t)/t$ and
$n_2(t)/t$ --- also converge to $\tfrac12$. At the end of the second
phase and in each global round afterwards, the counts $n_1(t)$ and
$n_2(t)$ satisfy the weak BIR-dominance condition, in the sense that
they both are larger than $n_0$ and $n_1(t)\geq (1-\beta_0)\; n_2(t)$.
At this point, \alg[1] actually has smaller $\BIR$, which reflected in the {\PMR}s eventually. Accordingly, from then on \alg[1]
attracts agents at a rate slightly larger than $\tfrac12$. We prove
that the ``bump'' over $\tfrac12$ is at least on the order of
$\BIR_2(t)$.

\begin{proof}[Proof of Theorem~\ref{thm:SoftMax-strong}]
  Let $\beta_1 = \min\{c_0'\delta_0, \beta_0/20\}$ with $\delta_0$
  defined in~\eqref{eq:SoftMax-smooth}.  Recall each agent chooses
  \alg[1] with probability at least $\respF(-1)= \eps_0$.  By
  condition \eqref{eq:so-mild} and \eqref{eq:SoftMax-assn-strong},
  there exists some sufficiently large $T_1$ such that for any
  $t\geq T_1$, $\BIR_1(\eps_0 T_1/2) \leq \beta_1/c_0'$ and
  $\BIR_2(t) > e^{-\eps_0 t/12}$. Moreover, for any $t\geq T_1$, we
  know $\E[n_1(t+1)] \geq \eps_0\,t $, and by the Chernoff Bounds
  (Theorem~\ref{thm:chernoff}), we have $n_1(t+1) \geq \eps_0 t/2$
  holds with probability at least $1 - q_1(t)$ with
  $q_1(t) = \exp(-\eps_0 t/12) < \BIR_2(t)$. It follows that for any $t\geq T_1$,
\begin{align*}
  \PMR_2(t) - \PMR_1(t) &= \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}{\BIR_1(m_1+ 1) - \BIR_2(m_2+1)} \\
                        &\leq q_1(t)  + \Ex{m_1\sim \posteriorN{1}{t}}{\BIR_1(m_1+ 1)\mid m_1 \geq \eps_0 t/2 - 1 } - \BIR_2(t)\\
                        &\leq \BIR_1(\eps_0 T_1/2) \leq \beta_1/c_0'
\end{align*}
% where the second inequality follows from~\eqref{eq:random-assn}.
Since the response function $\respF$ is $c_0'$-Lipschitz in the
neighborhood of $[-\delta_0, \delta_0]$, each agent after round $T_1$
will choose \alg[1] with probability at least
\[
  p_t \geq \frac{1}{2} - c_0'\left(\PMR_2(t) - \PMR_1(t)\right) \geq
  \frac{1}{2} - \beta_1.
\]

Next, we will show that there exists a sufficiently large $T_2$ such
that for any $t\geq T_1 + T_2$, with high probability
$n_1(t) > \max\{n_0, (1 - \beta_0)n_2(t)\}$, where $n_0$ is defined
in~\eqref{eq:SoftMax-better}. %  Let us first lower bound the number of
% agents received by \alg[1] after some number of rounds $t = T_1 + T'$
% for any $T' \geq T_1$. 
Fix any $t \geq T_1 + T_2$. 
Since each agent chooses \alg[1] with probability at least
$1/2 - \beta_1$, by Chernoff Bounds (Theorem~\ref{thm:chernoff}) we
have with probability at least $1 - q_2(t)$ that the number of agents
that choose \alg[1] is at least $\beta_0(1/2 - \beta_1)t/5$, where the
function
$$
q_2(x) = \exp\left( \frac{-(1/2 - \beta_1)(1 - \beta_0/5)^2x}{3} \right).
$$
Note that the number of agents received by \alg[2] is at most
$T_1 + (1/2 + \beta_1)t + (1/2 - \beta_1)(1 - \beta_0/5)t$.

Then as long as
% $T_2 \geq \max\left\{\frac{3T_1}{(1 - \beta_0)}, 8 n_0\right\}$, we
% can guarantee that for any $t \geq T_1 + T_2$,
% $n_1(t) > n_2(t) (1 - \beta_0)$ and $n_1(t) > n_0$ with probability at
% least $1 - q_2(t)$.
$T_2 \geq \frac{5T_1}{\beta_0}$, we can guarantee that
$n_1(t) > n_2(t) (1 - \beta_0)$ and $n_1(t) > n_0$ with probability at
least $1 - q_2(t)$ for any $t \geq T_1 + T_2$.
% for any $t\geq T_1 + T_2$, as long as $T_2 \geq .$ Moreover, to
% guarantee that , we will just need $T_2 \geq$.  Finally, we will
% argue that in each round $t \geq T_1 + T_2$, we recover the
% guarantee in \eqref{eq:thm:SoftMax-strong}.
% % \[
% %   \Pr[i_t = 1] \geq \frac{1}{2} + \frac{c_0\alpha_0\BIR_2(t)}{4}.
% % \]
Note that the weak BIR-dominance condition
in~\eqref{eq:SoftMax-better} implies that for any $t\geq T_1 + T_2$
with probability at least $1 - q_2(t)$,
\[
  \BIR_1(n_1(t)) < (1- \alpha_0)\BIR_2(n_2(t)).
\]
It follows that for any $t\geq T_1 + T_2$,
\begin{align*}
  \PMR_1(t) - \PMR_2(t) &= \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}{\BIR_2(m_2+ 1) - \BIR_1(m_1+1)} \\
                        &\geq (1 - q_2(t))\alpha_0 \BIR_2(t) - q_2(t)\\
                        &\geq \alpha_0 \BIR_2(t)/4
\end{align*}
where the last inequality holds as long as
$q_2(t) \leq \alpha_0\BIR_2(t)/4$, and is implied by the condition
in~\eqref{eq:SoftMax-assn-strong} as long as $T_2$ is sufficiently
large. Hence, by the definition of our \SoftMaxRandom response
function and assumption in~\eqref{eq:SoftMax-smooth}, we have
\[
  \Pr[i_t = 1] \geq \frac{1}{2} + \frac{c_0\alpha_0\BIR_2(t)}{4}. \qedhere
\]
\end{proof}

Similar to the condition \eqref{eq:random-better-weaker}, we can also
generalize the weak BIR-dominance condition \eqref{eq:SoftMax-better}
to handle the dependence on $T$: there exist some $T_0$, a function
$n_0(T)\in \polylog(T)$, and constants $\beta_0,\alpha_0\in (0, 1/2)$, such that 
\begin{align}\label{eq:SoftMax-better-weaker}
(\forall T\geq T_0,  n\geq n_0(T)) \quad
\frac{\BIR_1((1-\beta_0)\, n\mid T)}{\BIR_2(n\mid T)} <1-\alpha_0.
\end{align}

We also provide a version of Theorem~\ref{thm:SoftMax-weak} under this
more general weak BIR-dominance condition; its proof is very similar
and is omitted. The following is just a direct consequence of
Theorem~\ref{thm:SoftMax-weak} with this general condition.\sw{added}

\begin{corollary}\label{cor:SoftMax-strong}
Assume that the response function is \SoftMaxRandom, and each principal's  utility is the number of users. Consider the restricted competition game in which the special algorithm \alg weakly-BIR-dominates the other allowed algorithms, and the latter satisfy $\BReg(n)\to \infty$. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.
\end{corollary}

%If both $\BIR_1()$ and $\BIR_2()$ are of the form $\tilde{\Theta}(n^{-\gamma})$, $\gamma>0$, then the old condition requires $\BIR_1(\cdot)$ to be better by constant multiplicative factor $C$, with $C$ sufficiently large, whereas the new condition allows any $C>1$.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2019-08-14 journal paper/ec19paper/content/introduction.tex
the latex compiles

--- 2019-04-30 paper/content/introduction.tex
@@
 Implicitly, users have no ``outside option": the service provided is an improvement over not having it (and therefore the monopolist is not incentivized to deploy better learning algorithms).
-This is plausible with free ad-supported platforms such as Yelp or Google.}}\footnote{\gaedit{The most common measures of market ``competitiveness" such as the Lerner Index or the Herfindahl-Hirschman Index of a market rely on ex-post observable attributes of a market such as prices or market shares \cite{tirole1988theory}.
-However, since we have no prices and market shares are an endogenous element of our model, we simply define ``competitiveness" in terms of the number of firms and timing of entry into the market.}}
+This is plausible with free ad-supported platforms such as Yelp or Google.}}
 
@@
 \footnote{\asedit{Reputation scores already introduce some noise into users' choices.
-However, the amount of noise due to this channel is typically small, both in our simulations and in practice, because reputation signals average over many datapoints.}} We confirm the theoretical intuition that better algorithms prevail if the \gaedit{expected number of ``random" users is sufficiently large.}\gadelete{parameter -- essentially, the expected number ``random users" --is sufficiently large.} However, we find that this effect is negligible for some smaller but still ``relevant" parameter values.
+However, the amount of noise due to this channel is typically small, both in our simulations and in practice, because reputation signals average over many datapoints.}} We confirm the theoretical intuition that better algorithms prevail if the expected number of ``random" users is sufficiently large.
+However, we find that this effect is negligible for some smaller but still ``relevant" parameter values.
 
@@
 
-\asedit{While amenable to simulations, our model appears difficult to analyze. This is for several reasons:
+While amenable to simulations, our model appears difficult to analyze. This is for several reasons:
 %
@@
 %
-mean reputation, most connected to our intuition, is sometimes \gaedit{a} bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited});
+mean reputation, most connected to our intuition, is sometimes a bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited});
 %
 mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice.
-We believe that resolving first-order theoretical questions would not add much value to this paper, because of \cite{CompetingBandits-itcs18}.\gadelete{We are in \gaedit{the} realm of ``asymptotic" theorems about stylized economic models, which provide some mathematical intuition but no direct bearing on the world.
-But \cite{CompetingBandits-itcs18} already has an elaborate set of such results, with similar high-level conclusions.
-So, one would need to go beyond the first-order conclusions to add value.}}
+Further, we believe that resolving first-order theoretical questions about our model would not add much value to this paper.
+\asedit{Indeed, we are in the realm of stylized economic models that provide mathematical intuition about the world, and \cite{CompetingBandits-itcs18} already has an elaborate analysis with similar high-level conclusions.}
 
@@
 
+\asedit{The most common measures of market ``competitiveness" such as the Lerner Index or the Herfindahl-Hirschman Index of a market rely on ex-post observable attributes of a market such as prices or market shares \cite{tirole1988theory}.
+However, neither is applicable to our setting: in our model, there are no prices, and market shares are endogenous.}
+
 \end{document}

### 2019-08-14 journal paper/ec19paper/content/inverted_u.tex
the latex compiles

--- 2019-04-30 paper/content/inverted_u.tex
@@
 \end{tabular}
-\caption{{\bf Temporary monopoly}, with $X=200$ (and $T_0=20$), for the Heavy-Tail MAB instance.
+\captionsetup{labelformat=empty}
+\caption{User share of row player (entrant), 200 round head-start, Heavy-Tail Instance}
+%\caption{{\bf Temporary monopoly}, with $X=200$ (and $T_0=20$), for the Heavy-Tail MAB instance.
 Each cell describes the duopoly game between the entrant's algorithm (the row) and the incumbent's algorithm (the column).
@@
 \begin{figure}
-\includegraphics[scale=0.3]{figures/unif_eq_welfare}
-\includegraphics[scale=0.3]{figures/ht_eq_welfare}
+\includegraphics[scale=0.3]{ec19paper/figures/unif_eq_welfare}
+\includegraphics[scale=0.3]{ec19paper/figures/ht_eq_welfare}
 \caption{Smoothed welfare plots resulting from equilibrium strategies in the different market structures.
@@
 \begin{figure}
-\includegraphics[scale=0.3]{figures/eeog_vs_num_ht}
-\includegraphics[scale=0.3]{figures/ht_many_firm_welfare}\\
+\includegraphics[scale=0.3]{ec19paper/figures/eeog_vs_num_ht}
+\includegraphics[scale=0.3]{ec19paper/figures/ht_many_firm_welfare}\\
 \caption{Average welfare and \Eeog as we increase \#firms playing \DG}

### 2019-08-14 journal paper/ec19paper/content/perf_in_iso.tex
the latex compiles

--- 2019-04-29 paper/content/perf_in_iso.tex
@@
 \begin{figure}
-\includegraphics[scale=0.35]{figures/nih_iso_mean}
+\includegraphics[scale=0.35]{ec19paper/figures/nih_iso_mean}
 %\caption*{\tiny{Mean reputation trajectory The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
@@
 \begin{figure}[ht]
-\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
-\includegraphics[scale=0.35]{figures/relative_nih_ts_dg.png}
+\includegraphics[scale=0.35]{ec19paper/figures/relative_uniform_annotated_plot}
+\includegraphics[scale=0.35]{ec19paper/figures/relative_nih_ts_dg.png}
 %\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.

### 2019-08-14 journal paper/itcs18paper/app-examples.tex
the latex compiles

\newcommand{\ExplorExploit}{\term{ExplorExploit}}
\newcommand{\PhasedExplorExploit}{\term{PhasedExplorExploit}}
\newcommand{\RandomDynGreedy}{\term{RandomDynGreedy}}
\newcommand{\SuccesiveElimination}{\term{SuccesiveElimination}}
\newcommand{\SuccesiveEliminationReset}{\term{SuccesiveEliminationReset}}

This appendix provides some pertinent background on multi-armed
bandits (\emph{MAB}). We discuss \BIR and monotonicity of several MAB algorithms, touching upon: \DynGreedy and \StaticGreedy (Section~\ref{sec:MAB-greedy}), ``naive" MAB algorithms that separate exploration and exploitation (Section~\ref{sec:MAB-naive}), and ``smart" MAB algorithms that combine exploration and exploitation (Section~\ref{sec:MAB-smart}).

As we do throughout the paper, we focus on MAB with i.i.d. rewards and a Bayesian prior; we call it \emph{Bayesian MAB} for brevity.

% For a given mean reward vector $\mu$, the $n$-th step instantaneous regret is
%\begin{align}
%\regret(n\mid\mu) &:= \max_{a\in A} \mu_a - \rew(n\mid\mu),\\
%\regretWC(n)    &:=  \sup_{\text{mean reward vectors $\mu$}} \; \BIR(n\mid \mu).
%\end{align}

\subsection{\DynGreedy and \StaticGreedy}
\label{sec:MAB-greedy}

We provide an example when \DynGreedy and \StaticGreedy have
constant \BIR, and prove monotonicity of \DynGreedy. For the
example, it suffices to consider \emph{deterministic rewards} (for
each action $a$, the realized reward is always equal to the mean
$\mu_a$) and \emph{independent priors} (according to the prior
$\priorMu$, random variables $\mu_1 \LDOTS \mu_K$ are mutually
independent) each of {\em full support}.

%\ascomment{This lemma is just one example. Can we argue that constant \BIR is typical/reasonable?}
%
%
%\begin{lemma}
%There is a problem instance of Bayesian MAB such that \StaticGreedy
%and \DynGreedy have (at least) a constant \BIR for all steps. The
%problem instance has two actions, independent priors, and
%deterministic rewards.
%\end{lemma}
%
%\begin{proof}
%To specify the problem instance, it remains to specify the marginal
%distributions mean rewards $\mu_1$ and $\mu_2$ according to the
%prior. We posit that $\mu_1$ is uniform on the interval $[0, 1]$,
%and $\mu_2$ is uniform on the interval $[\tfrac13,1]$.
%
%Note that
%    $\OPT := \E[\max_{a\in A} \mu_a] =(1/3)(2/3)+(2/3)(7/9)=20/27$.
%%\ascomment{The numbers seem wrong, given the below?}
%This holds since with
%probability $1/3$ we have $\mu_1\leq 1/3$ in which case clearly the
%better arm is 2 with $\E[\mu_2]=2/3$ and with probability $2/3$ we
%have both $\mu_1,\mu_2\in[1/3,1]$ and distributed uniformly each.
%
%\StaticGreedy always recommends action $2$. Since
%$\E[\mu_2]=\tfrac23$, \BIR is $2/27$ in all rounds.
%
%\DynGreedy recommends the first agent action $2$. If $\mu_2 > 2/3$
%For all the following agents it recommends action $2$. Otherwise, it
%recommends agent $2$ action $1$ and for all the following agents it
%recommends the better action. The expectation for agents $t\geq 3$
%is $(1/2)(5/6)+(1/2)((1/2)(1/2)+(1/2)(5/9))=49/72$ and has
%instantaneous regret $13/216$.
%\end{proof}

The following claim is immediate from the definition of the CDF
function
\begin{claim}
Assume independent priors. Let $F_i$ be the CDF of the mean reward
$\mu_i$ of action $a_i\in A$. Then, for any numbers
$z_2>z_1>\E[\mu_2]$ we have
    $\Pr[\text{$\mu_1\leq z_1$ and $\mu_2\geq z_2$}] = F_1(z_1)(1-F_2(z_2)) $.
\end{claim}

%\begin{lemma}
%Let $\priorMu$ be the prior of the means, and assume that it is
%independent for each action, where for arm $i$ it has a mean $
%\E[\mu_i]$ and CDF $F_i$. Then, for any $\E[\mu_2] < z_1<z_2$ we
%have that with probability $F_1(z_1)(1-F_2(z_2))$ that both
%$\mu_1\leq z_1$ and $z_2\leq \mu_2$.
%\end{lemma}

%\begin{corollary}
%Any problem instance of Bayesian MAB for two actions and independent
%priors which are full support and deterministic rewards, with
%constant probability \DynGreedy has a constant \BIR for all steps.
%\end{corollary}

We can now draw an immediate corollary of the above claim

\begin{corollary}
Consider any problem instance of Bayesian MAB with two actions and independent
priors which are full support. Then:
\begin{OneLiners}
\item[(a)] With constant probability, \StaticGreedy  has a constant \BIR for all steps.
\item[(b)] Assuming deterministic rewards, with
constant probability \DynGreedy has a constant \BIR for all steps.
\end{OneLiners}
\end{corollary}

\begin{remark}
A similar result holds for  rewards which are distributed as
Bernoulli random variables. In this case we consider accumulative
reward of an action as a random walk, and use a high probability
variation of the law of iterated logarithms. (Details omitted.)
\end{remark}

Next, we show that \DynGreedy is monotone.

\begin{lemma}\label{dgmono}
\DynGreedy is monotone, in the sense that $\rew(n)$ is non-decreasing.
Further, $\rew(n)$ is strictly increasing for every time step $n$ with $\Pr[a_n\neq a_{n+1}]>0$.
\end{lemma}

\begin{proof}
We prove by induction on $n$ that $\rew(n)\leq \rew(n+1)$ for
\DynGreedy. Let $a_n$ be the random variable recommended at time
$t$, then $\E[\mu_{a_n}| \mI_n ]=\rew(n)$. We can rewrite this as:
\[
\rew(n)=\E_{\mI_n}[\E_{r_n}[\mu_{a_n}|r_n,\mI_n]] =
\E_{\mI_{n+1}}[\mu_{a_n}|\mI_{n+1}]
\]
since $\mI_{n+1}=(\mI_n,r_n)$. At time $n+1$ \DynGreedy will select
an action $a_{n+1}$ such that:
\[
\rew(n+1)=\E[\mu_{a_{n+1}}|\mI_{n+1}]\geq \E[\mu_{a_n}
|\mI_n]=\rew(n)
\]
%
%Now after executing action $a_n$ and observing its reward $r_n$
%there are two possibilities. If $a_n$ is still the best action, then
%its expected reward given the information at time $n$ has not
%changed. On the other hand, if $a_n$ is not the best action anymore,
%then it implies that we recommend a better action. In such a case,
%we increase the expected reward. \ymcomment{need to make this
%formal}
which proves the monotonicity. In cases that $\Pr[a_n\neq
a_{n+1}]>0]$ we have a strict inequality, since with some
probability we select a better action then the realization of $a_n$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%

\subsection{``Naive" MAB algorithms that separate exploration and exploitation}
\label{sec:MAB-naive}

%\ascomment{Nicer to have $\BIR(n)\leq n^{-1/3}$.
%Explore-then-exploit does not have that! So, let's use phases of
%exponentially increasing duration.}

MAB algorithm \ExplorExploit$(m)$ initially explores each action
with $m$ agents and for the remaining $T-|A|m$ agents recommends the
action with the highest observed average. In the explore phase it
assigns a random permutation of the $mK$ recommendations.

\begin{lemma}
%Assume that rewards in the range $[0,1]$.
The \ExplorExploit$(T^{2/3}\log |A|/\delta)$ algorithm has, with
probability $1-\delta$, for any  $n\geq |A|T^{2/3}$ we have
\BIR$(n)=O(T^{-1/3})$. In addition, \ExplorExploit$(m)$ is monotone.
\end{lemma}

\begin{proof}
In the explore phase we we approximate for each action $a\in A$, the
value of $\mu_a$ by $\hat{\mu}_a$. Using the standard Chernoff
bounds we have that with probability $1-\delta$, for every action
$a\in A$ we have $|\mu_a -\hat{\mu}_a| \leq T^{-1/3}$.

Let $a^* = \arg\max_a \mu_a$ and $a^{ee}$ the action that
\ExplorExploit selects in the explore phase after the first
$|A|T^{2/3}$ agents. Since $\hat{\mu}_{a^*} \leq
\hat{\mu}_{a^{ee}}$, this implies that $\mu_{a^*} -
\mu_{a^{ee}}=O(T^{-1/3})$.

To show that \ExplorExploit$(m)$ is monotone, we need to show only
that $\rew(mK) \leq \rew(mK+1)$. This follows since for any $t< mK$
we have $\rew(t)=\rew(t+1)$, since the recommended action is
uniformly distributed for each time $t$. Also, for any $t\geq mK+1$
we have $\rew(t)=\rew(t+1)$ since we are recommending the same
exploration action. The proof that $\rew(mK) \leq \rew(mK+1)$ is the
same as for \DynGreedy in Lemma~\ref{dgmono}.
\end{proof}

We can also have a a phased version which we call
\PhasedExplorExploit$(m_t)$, where time is partition in to phases.
In phase $t$ we have $m_t$ agents and a random subset of $K$ explore
the actions (each action explored by a single agent) and the other
agents exploit. (This implies that we need that $m_t\geq K$ for all
$t$. We also assume that $m_t$ is monotone in $t$.)

\begin{lemma}
Consider the case that $K=2$ and the rewards of the actions are
Bernoulli r.v. with parameter $\mu_i$ and $\Delta=\mu_1-\mu_2$.
Algorithm \PhasedExplorExploit$(m_t)$ is monotone and for $m_t =
\sqrt{t}$ it has $\BIR(n)=O(n^{-1/3}+e^{-O(\Delta^2 n^{2/3})}))$. 
\end{lemma}

\begin{proof}
We first show that it is monotone. Recall that $\mu_1>\mu_2$. Let
$S_i=\sum_{j=1}^t r_{i,j}$ be the sum of the rewards of action $i$
up to phase $t$. We need to show that $\Pr[S_1>S_2]+ (1/2)
\Pr[S_1=S_2]$ is monotonically increasing in $t$. Consider the
random variable $Z=S_1-S_2$. At each phase it increases by $+1$ with
probability $\mu_1(1-\mu_2)$, decreases by $-1$ with probability
$(1-\mu_1)\mu_2$ and otherwise does not change.

Consider the values of $Z$ up to phase $t$. We really care only
about the probability that is shifted from positive to negative and
vice versa.

First, consider the probability that $Z=0$. We can partition it to
$S_1=S_2=r$ events, and let $p(r,r)$ be the probability of this
event. For each such event, we have $p(r,r)\mu_1$ moved to $Z=+1$
and $p(r,r)\mu_2$ moved to $Z=-1$. Since $\mu_1>\mu_2$ we have that
$p(r,r)\mu_1\geq p(r,r)\mu_2$ (note that $p(r,r)$ might be zero, so
we do not have a strict inequality).

Second, consider the probability that $Z=+1$ or $Z=-1$. We can
partition it to $S_1=r+1;S_2=r$ and $S_1=r;S_2=r+1$ events, and let
$p(r+1,r)$ and $p(r,r+1)$ be the probabilities of those events.
%
It is not hard to see that $p(r+1,r)\mu_2=p(r,r+1)\mu_1$.
%
This implies that the probability mass moved from $Z=+1$ to $Z=0$ is
identical to that moved from $Z=-1$ to $Z=0$.

We have showed that $\Pr[S_1>S_2]+ (1/2) \Pr[S_1=S_2]$ and therefore
the expected valued of the exploit action is non-decreasing. Since
we have that the size of the phases are increasing, the $\BIR$ is
strictly increasing between phases and identical within each phase.

We now analyze the $\BIR$ regret. Note that agent $n$ is in phase
$O(n^{2/3})$ and the length of his phase is $O(n^{1/3})$. The $\BIR$
has two parts. The first is due to the exploration, which is at most
$O(n^{-1/3})$. The second is due to the probability that we exploit
the wrong action. This happens with probability $\Pr[S_1<S_2]+ (1/2)
\Pr[S_1=S_2]$ which we can bound using a Chernoff bound by
$e^{-O(\Delta^2n^{2/3})}$, since we explored each action
$O(n^{2/3})$ times.
\end{proof}

\begin{remark}
Actually we have a tradeoff depending on the parameter $m_t$ between
the regret due to exploration and exploitation. (Note that the
monotonicity is always guarantee assuming $m_t$ is monotone.) If we
can set that $m_t = 2^t$ then at time $n$ we have $2/ n$ probability
of an exploit action. For the explore action we are in phase $\log
n$ so the probability of a sub-optimal explore action is
$n^{-O(\Delta^{-2})}$. This should give us
$\BIR(n)=O(n^{-O(\Delta^{-2})})$.
\end{remark}

%MAB algorithm \RandomDynGreedy$(q_t)$for agent $t$ with probability
%$q_t$ selects a random action from $A$ and with probability $1-q_t$
%recommends the best action according to the current posterior
%(similar to \DynGreedy).
%\begin{lemma}
%%Assume that rewards in the range $[0,1]$.
%The algorithm \RandomDynGreedy$(q_t)$, where $q_t=t^{-1/3}$, has,
%with probability $1-\delta$, \BIR$(n)=O(K^{1/2}n^{-1/3}\log
%(K/\delta))$.
%\end{lemma}
%
%\begin{proof}
%At time $n$ we have that with probability $1-\delta/(2K)$, each
%action $a\in A$ was selected at least $n_a=\Theta(n^{2/3}/K)$ times.
%%
%This implies that with probability $1-\delta/(2K)$, for each $a$ we
%have $|\hat{\mu}_a-\mu_a|\leq K^{1/2}n^{-1/3}\log((2K)/\delta)$.
%%
%Therefore, the \BIR is $O(K^{1/2}n^{-1/3}\log((K)/\delta))$
%\end{proof}

%\ascomment{Add: can we modify Phased Explore-then-exploit to make it
%monotone? Perhaps choose exploration rounds u.a.r. within the
%phase?}

%
%\begin{lemma}
%\ExplorExploit$(m)$ is monotone
%\end{lemma}
%
%\begin{proof}
%We need to consider only time $m'=|A|m+1$. For any $n<m'$, we
%clearly have $\rew(n)=\rew(1)$.
%
%Consider two actions $a_1,a_2\in A$, such that $\mu_{a_1} \geq
%\mu_{a_2}$. We claim that the probability that $a_1$ is selected at
%time $m$ to be the exploit action is larger than the probability
%that $a_2$ is selected. This follows since $\hat{\mu}_{a_1}$
%stochastically dominates $\hat{\mu}_{a_2}$, which implies that for
%any threshold $\theta$ we have $\Pr[\hat{\mu}_{a_1}\geq\theta]\geq
%\Pr[\hat{\mu}_{a_2}\geq\theta]$.
%
%After the elimination we consider the expected reward of the
%selected action $\sum_{i\in A} \mu_i q_i$, where $q_i$ is the
%probability that action $i$ was selected for the exploit phase. We
%have that $q_i \geq q_{i+1}$, from the stochastic dominance.
%
%The sum $\sum_{i\in A} \mu_i q_i$ with $q_i \geq q_{i+1}$ and
%$\sum_i q_i=1$ is minimized by setting $q_i=1/|A|$. (We can see that
%if there are $q_i\neq 1/|A|$, then there are two $q_{i}> q_{i+1}$,
%and one can see that setting both to $(q_{i}+ q_{i+1})/2$ decreases
%the value.) Therefore we have that the $\rew(m')\geq\rew(m'-1)$.
%\end{proof}

\subsection{``Smart" MAB algorithms that combine exploration and exploitation}
\label{sec:MAB-smart}

%MAB algorithm \SuccesiveElimination holds a set of action
%$A_s\subset A$ which are called {\em surviving actions} and for
%agent $t$ selects a random action from $A_s$. Let $n_{i,t}$ be the
%number of times action $i$ has been selected up to time $t$, and
%$\hat{\mu}_{i,t}$ be the average of the rewards of action $i$ up to
%time $t$ and $\hat{\mu}^*=\max_i \hat{\mu}_{i,t}$. We eliminate
%action $i$ at time $t$, i.e., delete it from $A_s$, if
%$\hat{\mu}_t^*-\hat{\mu}_{i,t} > \log(T/\delta)/\sqrt{n/K}$.
%
%\begin{lemma}
%Assume that the prior is independent and if $\mu_i\geq \mu_j$ then
%$r_i$ stochastically dominates $r_j$.
%%
%The algorithm \SuccesiveElimination, has, with probability
%$1-\delta$, \BIR$(n)=O(\log(T/\delta)/\sqrt{n/K})$.
%\end{lemma}

%\begin{proof}
%Let the best action be $a^*=\arg\max_a \mu_a$. With probability
%$1-\delta$ at any time $n$ we have that for any action $i\in A_s$
%that $|\hat{\mu}_i -\mu_i|\leq \log(T/\delta)/\sqrt{n/K}$, and
%$a^*\in A_s$. This implies that any action $a$ such
%$\mu_{a^*}-\mu_{a}> 3\log(T/\delta)/\sqrt{n/K}$ is eliminated.
%Therefore, any action in $A_s$ has \BIR$(n)$ of at most
%$O(\log(T/\delta)/\sqrt{n/K})$. \ymcomment{should we get the
%constant right?}
%\end{proof}

%\ascomment{Add: Successive Elimination is monotone (or a simple
%modification thereof).}
%
%\begin{lemma}
%\SuccesiveElimination is monotone
%\end{lemma}
%
%\begin{proof}
%We will first show that for any two actions $a_1,a_2\in A$ that if
%$\mu_{a_1} \geq \mu_{a_2}$ then at any time $t$ we have that the
%probability that $a_1$ is eliminated is smaller than the probability
%that $a_2$ is eliminated, i.e., $\Pr[a_1 \not\in A_{s,t+1}| a_1\in
%A_{s,t} ]\geq \Pr[ a_2 \not\in A_{s,t+1}| a_2\in A_{s,t}]$.
%
%We first show that given this statement, the lemma holds and
%\SuccesiveElimination is monotone. Consider $\rew(t)$ versus
%$\rew(t+1)$. If the elimination probabilities would have been
%identical then we would have $\rew(t)=\rew(t+1)$. Consider the top
%two actions, and consider the elimination process in two phases,
%firs we eliminate both with the same probability, and hence the
%reward is unchanged, and then we eliminate only the second, which
%implies that we increase the the reward. The claim about the
%elimination probabilities rules out the case that the elimination
%probability of the best action is higher than then of the the second
%best. Therefore, overall the reward can only increase.
%
%
%We now need to prove the claim. Since $\mu_{a_1}\geq \mu_{a_2}$ and
%they use the same parametric family $\psi_a(\cdot)$ then
%$\hat{\mu}_{a_1}$ stochastically dominates $\hat{\mu}_{a_2}$. We
%actually need to show that $\hat{\mu}_{a_1}$ stochastically
%dominates $\hat{\mu}_{a_2}$ given that $a_1,a_2\in A_{s,t}$. If this
%holds, then the claim about the elimination would be immediate,
%since for any threshold $\theta$ we have $\Pr[\hat{\mu}_{a_1,t} <
%\theta |a_1\in A_{s,t}] \leq \Pr[\hat{\mu}_{a_2,t} < \theta |a_2\in
%A_{s,t}] $. By setting $\theta=\hat{\mu}^*-
%\log(T/\delta)/\sqrt{n/K}$ we derive the lemma.
%
%We now need to prove the claim. We first show that
%$\E[\hat{\mu}_{a_1}-\hat{\mu}_{a_2} | a_1,a_2\in A_s]\geq 0$ which
%is equivalent to $\E[\hat{\mu}_{a_1}-\hat{\mu}_{a_2} |
%\hat{\mu}_{a_1},\hat{\mu}_{a_2}\geq \hat{\mu}^*-\theta]\geq 0$ where
%$\theta$ is the threshold used in \SuccesiveElimination. Since
%action $a_1$ has a high expectation than $a_2$ it holds (at least
%for Bernoulli r.v.).
%
%
%[[OLD]]
%
% Clearly, for any actions $a\in A$ we have that $\Pr[a \in
%A_s]$ decreases with time. In addition, we have that with
%probability $1-\delta$ for every time $t$ we have $a^*\in A_s$.
%
%If $\mu_{a_1} \geq \mu_{a_2}$ then at any time $t$ we have
%$\Pr[a_1\in A_s]\geq \Pr[a_2\in A_s]$.
%
%
%We probably need to show that if $\mu_{a_1} \geq \mu_{a_2}$ then at
%any time $t$ we have that the probability that $a_1$ is eliminated
%is smaller than the probability that $a_2$ is eliminated, i.e.,
%$\Pr[a_1\in A_{s,t}, a_1 \not\in A_{s,t+1}]\geq \Pr[a_2\in A_s, a_2
%\not\in A_{s,t+1}]$.
%\end{proof}

%\ymcomment{Here is a simple modification of Successive-Elimination}

MAB algorithm \SuccesiveEliminationReset works as follows. It keeps
a set of surviving actions $A_s\subseteq A$, where initially
$A_s=A$. The agents are partition into phases, where each phase is a
random permutation of the non-eliminated actions.
% Initially, we select actions at random. Let $n_{i,t}$
%be the number of times action $i$ has been selected up to time $t$,
Let $\hat{\mu}_{i,t}$ be the average of the rewards of action $i$ up
to phase $t$ and $\hat{\mu}^*=\max_i \hat{\mu}_{i,t}$. We eliminate
action $i$ at the end of phase $t$, i.e., delete it from $A_s$, if
$\hat{\mu}_t^*-\hat{\mu}_{i,t} > \log(T/\delta)/\sqrt{t}$.
%So far identical to \SuccesiveElimination, the difference is how we
%continue after elimination.
In \SuccesiveEliminationReset we simply reset the algorithm with
$A=A_s-A_{e,t}$, where $A_{e,t}$ is the set of eliminated actions
after phase $t$. Namely, we restart $\hat{\mu}_{i,t}$ and ignore the
old rewards before the elimination.

\begin{lemma}
%
The algorithm \SuccesiveEliminationReset, has, with probability
$1-\delta$, \BIR$(n)=O(\log(T/\delta)/\sqrt{n/K})$.
\end{lemma}

\begin{proof}
Let the best action be $a^*=\arg\max_a \mu_a$. With probability
$1-\delta$ at any time $n$ we have that for any action $i\in A_s$
that $|\hat{\mu}_i -\mu_i|\leq \log(T/\delta)/\sqrt{n/K}$, and
$a^*\in A_s$. This implies that any action $a$ such
$\mu_{a^*}-\mu_{a}> 3\log(T/\delta)/\sqrt{n/K}$ is eliminated.
Therefore, any action in $A_s$ has \BIR$(n)$ of at most
$6\log(T/\delta)/\sqrt{n/K}$.
\end{proof}

\begin{lemma}
Assume that if $\mu_i\geq \mu_j$ then the rewards $r_i$
stochastically dominates the rewards $r_j$.
%
Then, \SuccesiveEliminationReset is monotone
\end{lemma}

\begin{proof}
Consider the first time $T$ an action is eliminated, and let
$T=\tau$ be a realized value of $T$. Then, clearly for $n<\tau$ we
have $\rew(n)=\rew(1)$ .

Consider two actions $a_1,a_2\in A$, such that $\mu_{a_1} \geq
\mu_{a_2}$. At time $T=\tau$, the probability that  $a_1$ is
eliminated is smaller than the probability that $a_2$ is eliminated.
This follows since $\hat{\mu}_{a_1}$ stochastically dominates
$\hat{\mu}_{a_2}$, which implies that for any threshold $\theta$ we
have $\Pr[\hat{\mu}_{a_1}\geq\theta]\geq
\Pr[\hat{\mu}_{a_2}\geq\theta]$.

After the elimination we consider the expected reward of the
eliminated action $\sum_{i\in A} \mu_i q_i$, where $q_i$ is the
probability that action $i$ was eliminated in time $T=\tau$. We have
that $q_i \leq q_{i+1}$, from the probabilities of elimination.

The sum $\sum_{i\in A} \mu_i q_i$ with $q_i \leq q_{i+1}$ and
$\sum_i q_i=1$ is maximized by setting $q_i=1/|A|$. (We can see that
if there are $q_i\neq 1/|A|$, then there are two $q_{i}< q_{i+1}$,
and one can see that setting both to $(q_{i}+ q_{i+1})/2$ increases
the value.) Therefore we have that the $\rew(\tau)\geq\rew(\tau-1)$.

Now we can continue by induction. For the induction, we can show the
property for {\em any} remaining set of at most $k-1$ actions. The
main issue is that \SuccesiveEliminationReset restarts from scratch,
so we can use induction.
\end{proof}

### 2019-08-14 journal paper/itcs18paper/sec-intro.tex
the latex compiles

Learning from interactions with users is ubiquitous in modern customer-facing systems, from product recommendations to web search to spam detection to content selection to fine-tuning the interface. Many systems purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information. Randomized controlled trials, a.k.a. A/B testing, are an industry standard, with a number of companies such as \emph{Optimizely} offering tools and platforms to facilitate them. Many companies use more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known theoretical framework for exploration and making decisions under uncertainty.

%~\cite{KohaviAB-2015,KohaviLSH09}

Systems that engage in exploration typically need to compete against one another; most importantly, they compete for users. This creates an interesting tension between \exploration and \competition. In a nutshell, while exploring may be essential for improving the service tomorrow, it may degrade quality and make users leave \emph{today}, in which case there will be no users to learn from! Thus, users play three distinct roles: they are customers that generate revenue, they generate data for the systems to learn from, and they are self-interested agents which choose among the competing systems.

We initiate a study of the interplay between \exploration and \competition. The main high-level question is: {\bf whether and to what extent competition incentivizes adoption of better exploration algorithms}. This translates into a number of more concrete questions. While it is commonly assumed that better learning technology always helps, is this so for our setting? In other words, would a better learning algorithm result in higher utility for a principal? Would it be used in an equilibrium of the ``competition game"? Also, does competition lead to better social welfare compared to a monopoly? We investigate these questions for several models, as we vary the capacity of users to make rational decisions (\rationality ) and the severity of competition between the learning systems (\competitiveness). The two are controlled by the same ``knob" in our models; \asedit{such coupling is not unusual in the literature, \eg see \citet{Gabaix-16}}.

\asedit{On a high level, our contributions can be framed in terms of the ``inverted-U relationship" between rationality/competitiveness and the quality of adopted algorithms (see \reffig{fig:inverted-U}).}

\OMIT{The relationship between the severity of competition among firms and the quality of technology adopted as a result of this competition is a familiar  theme in the economics literature, known as ``competition vs. innovation".%
\footnote{In this context, \innovation usually refers to adoption of a better technology as a substantial R\&D expense to a given firm. It is not as salient whether similar ideas and/or technologies already exist outside the firm. }
We frame our contributions in terms of the ``inverted-U relationship", a conventional wisdom regarding ``competition vs. innovation" (see \reffig{fig:inverted-U}).}

\begin{figure}
\begin{center}
\begin{tikzpicture}
      \draw[->] (-.5,0) -- (3,0) node[right] {Rationality/competitiveness};
      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
 \end{tikzpicture}

\caption{Inverted-U relationship between rationality/competitiveness and algorithms.}
\label{fig:inverted-U}
\end{center}
\end{figure}

% the extent to which the game between the two principals is competitive
% degree of innovation that these models incentivize.
% the extent to which agents make rational decisions

%\subsection{Our model}
%\label{sec:intro-model}

\xhdr{Our model.} We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}). These two processes are interlinked, as exploration decisions are experienced by users and informed by their feedback. We need to specify several conceptual pieces: how the principals and agents interact, what is the machine learning problem faced by each principal, and what is the information structure. Each piece can get rather complicated in isolation, let alone jointly, so we strive for simplicity. Thus, the basic model is as follows:

\begin{itemize}

\item A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals. The principal chooses an action (\eg a list of web search results to show to the agent), the user experiences this action, and reports a reward. \asedit{All agents have the same ``decision rule" for choosing among the principals given the available information.}

\item Each principal faces a very basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a. \emph{arms}) and receives a reward drawn independently from a fixed distribution specific to this action.

\item Principals simultaneously announce their learning algorithms \asedit{before round $1$}, and cannot change them afterwards. There is a common Bayesian prior on the rewards (but the realized reward distributions are not observed by the principals or the agents). Agents do not receive \asedit{any other information}. Each principal only observes agents that chose him.
\end{itemize}

\xhdr{Technical results.}
Our results depend crucially on agents' ``decision rule" for choosing among the principals. The simplest and perhaps the most obvious rule is to select the principal which maximizes their expected utility; we refer to it as \HardMax. We find that \HardMax is not conducive to \asedit{adopting better algorithms}. In fact, each principal's dominant strategy is to do no purposeful exploration whatsoever, and instead always choose an action that maximizes expected reward given the current information; we call this algorithm \DynGreedy. While this algorithm may potentially try out different actions over time and acquire useful information, it is known to be dramatically bad in many important cases of multi-armed bandits --- precisely because it does not explore on purpose, and may therefore fail to discover best/better actions. Further, we show that \HardMax is very sensitive to tie-breaking when both principals have exactly the same expected utility according to agents' beliefs. If tie-breaking is probabilistically biased --- say, principal 1 is always chosen with probability strictly larger than $\tfrac12$ --- then this principal has a simple ``winning strategy" no matter what the other principal does.

We relax \HardMax to allow each principal to be chosen with some fixed baseline probability. One intuitive interpretation is that there are ``random agents" who choose a principal uniformly at random, and each arriving agent is either \HardMax or ``random" with some fixed probability. We call this model \HardMaxRandom. We find that \asedit{better algorithms} help in a big way: a sufficiently better algorithm is guaranteed to win all \asedit{non-random} agents after an initial learning phase. While the precise notion of ``sufficiently better algorithm" is rather subtle, we note that commonly known ``smart" bandit algorithms typically defeat the commonly known ``naive" ones, and the latter typically defeat \DynGreedy. However, there is a substantial caveat: one can defeat any algorithm by interleaving it with \DynGreedy. This has two undesirable corollaries: a better algorithm may sometimes lose, and a pure Nash equilibrium typically does not exist.

We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' expected rewards; we call it \SoftMaxRandom. For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm. This is the most technical result of the paper. The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm may attract slightly more).

All results extend to a much more general version of the multi-armed bandit problem in which the principal may observe additional feedback before and/or after each decision, as long as the feedback distribution does not change over time. In most results, principal's utility may depend on both the market share and agents' rewards.

\xhdr{Economic interpretation.}
\asedit{The inverted-U relationship between the severity of competition among firms and the quality of technologies that they adopt is a familiar theme in the economics literature \citep[\eg][]{Aghion-QJE05,Vives-08}.%
\footnote{The literature frames this relationship as one between ``competition" and ``innovation". In this context, ``innovation" refers to adoption of a better technology, at a substantial R\&D expense to a given firm. It is not salient whether similar ideas and/or technologies already exist outside the firm. It is worth noting that adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature; see \citet{MWT-WhitePaper-2016} for an example of such R\&D effort.}
We find it illuminating to frame our contributions in a similar manner, as illustrated in \reffig{fig:inverted-U}.

Our models differ in terms of rationality in agents' decision-making: from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom. The same distinctions also control the severity of competition between the principals: from cut-throat competition with \HardMax to a more relaxed competition with \HardMaxRandom, to an even more relaxed competition with \SoftMaxRandom. Indeed, with \HardMax you lose all customers as soon as you fall behind in performance, with \HardMaxRandom you get some small market share no matter what, and with \SoftMaxRandom you are further guaranteed a market share close to $\tfrac12$ as long as your performance is not much worse than the competition. The uniform choice among principals corresponds to no rationality and no competition.

We identify the inverted-U relationship in the spirit of \reffig{fig:inverted-U} that is driven by the rationality/competitiveness distinctions outlined above: from \HardMax to \HardMaxRandom to \SoftMaxRandom to \Uniform. We also find another, technically different inverted-U relationship which zeroes in on the \HardMaxRandom model. We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.

These inverted-U relationships arise for a fundamentally different reason, compared to the existing literature on ``competition vs. innovation.'' In the literature, better technology always helps in a competitive environment, other things being equal. Thus, the tradeoff is between the costs of improving the technology and the benefits that the improved technology provides in the competition. Meanwhile, we find that a better exploration algorithm may sometimes perform much worse under competition, even in the absence of R\&D costs.}

\ascomment{Yishay's edits, slightly edited by Alex}

\xhdr{Discussion.} We capture some pertinent features of reality while ignoring some others for the sake of tractability. Most notably, we assume that agents do not receive any information about other agents' rewards after the game starts. In the final analysis, this assumption makes agents' behavior independent of a particular realization of the Bayesian prior, and therefore enables us to summarize each learning algorithm via its Bayesian-expected rewards (as opposed to detailed performance on the particular realizations of the prior). Such summarization is essential for formulating lucid and general analytic results, let alone proving them. It is a major open question whether one can incorporate signals about other agents' rewards and obtain a tractable model.

We also make a standard assumption that agents are myopic: they do not worry about how their actions impact their future utility. In particular, they do not attempt to learn over time, to second-guess or game future agents, or to manipulate principal's learning algorithm. We believe this is a typical case in practice, in part because agent's influence tend to be small compared to the overall system. We model this simply by assuming that each agent only arrives once.

Much of the challenge in this paper, both conceptual and technical, was in setting up the right model and the matching theorems, and not only in proving the theorems. Apart from making the modeling choices described above, it was crucial to interpret the results and intuitions from the literature on multi-armed bandits so as to formulate meaningful assumptions on bandit algorithms and Bayesian priors which are productive in our setting.

\xhdr{Open questions.}
\asedit{How to incorporate signals about the other agents' rewards? One needs to reason about how exact or coarse these signals are, and how the agents update their beliefs after receiving them. Also, one may need to allow principals' learning algorithms to respond to updates about the other principal's performance. (Or not, since this is not how learning algorithms are usually designed!) A clean, albeit idealized, model would be that (i) each agent learns her exact expected reward from each principal before she needs to choose which principal to go to, but (ii) these updates are invisible to the principals. Even then, one needs to argue about the competition on particular realizations of the Bayesian prior, which appears very challenging.

Another promising extension is to heterogeneous agents. Then the agents' choices are impacted by their idiosyncratic signals/beliefs, instead of being entirely determined by priors and/or signals about the average performance. It would be particularly interesting to investigate the emergence of \emph{specialization}: whether/when an algorithm learns to target specific population segments in order to compete against a more powerful ``incumbent".
} %\asedit

\xhdr{Map of the paper.}
We survey related work (Section~\ref{sec:related-work}), lay out the model and preliminaries (Section~\ref{sec:model}), and proceed to analyze the three main models, \HardMax, \HardMaxRandom and \SoftMaxRandom (in Sections~\ref{sec:rational},~\ref{sec:random}, and ~\ref{sec:soft}, respectively). We discuss economic implications in Section~\ref{sec:welfare}. Appendix~\ref{app:examples} provides some pertinent background on multi-armed bandits. Appendix~\ref{app:perturb} gives a broad example to support an  assumption in our model.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2019-08-14 journal paper/itcs18paper/sec-model.tex
the latex compiles

\xhdr{Principals and agents.} There are two principals and $T$ agents. The game proceeds in rounds (we will sometimes refer to them as \emph{global rounds}). In each round $t\in [T]$, the following  interaction takes place. A new agent arrives and chooses one of the two principals. The principal chooses a recommendation: an action $a_t\in A$, where $A$ is a fixed set of actions (same for both principals and all rounds). The agent follows this recommendation, receives a reward $r_t\in [0,1]$, and reports it back to the principal.

The rewards are i.i.d. with a common prior. More formally, for each action $a\in A$ there is a parametric family $\psi_a(\cdot)$ of
reward distributions, parameterized by the mean reward $\mu_a$. (The paradigmatic case is 0-1 rewards with a given expectation.) The
mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from prior distribution $\priorMu$ before round $1$. Whenever a given action $a\in A$ is chosen, the reward is drawn independently from distribution $\psi_a(\mu_a)$. The prior $\priorMu$ and the distributions $(\psi_a(\cdot)\colon a\in A)$ constitute the (full) Bayesian prior on rewards, denoted $\prior$.

Each principal commits to a learning algorithm for making recommendations. This algorithm follows a protocol of \emph{multi-armed bandits} (\emph{MAB}). Namely, the algorithm proceeds in time-steps:%
\footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before. We will omit the local vs. local distinction when clear from the context.} each time it is called, it outputs a chosen action $a\in A$ and then inputs the reward for this action. The algorithm is called only in global rounds when the corresponding principal is chosen.

The information structure is as follows. The prior $\prior$ is known to everyone. The mean rewards $\mu_a$ are not revealed to anybody. Each agent knows both principals' algorithms, and the global round when (s)he arrives, \emph{but not} the rewards of the previous agents. Each principal is completely unaware of the rounds when the other is chosen.

\xhdr{Some terminology.} The two principals are called ``Principal
1" and ``Principal 2".
%Principal $i\in \{1,2\}$ is also denoted \prin[i].
The algorithm of principal $i\in \{1,2\}$ is called ``algorithm $i$" and denoted
\alg[i]. The agent in global round $t$ is called ``agent $t$"; the
chosen principal is denoted $i_t$.

Throughout, $\E[\cdot]$ denotes expectation over all applicable randomness.

\xhdr{Bayesian-expected rewards.}
Consider the performance of a given algorithm \alg[i], $i\in \{1,2\}$, when it is run in isolation (\ie without competition, just as a bandit algorithm). Let $\rew_i(n)$ denote its Bayesian-expected reward for the $n$-th step.

Now, going back to our game, fix global round $t$ and let $n_i(t)$ denote the number of global rounds before $t$ in which this principal is chosen. Then:
\begin{align*}
 \E[r_t \mid \text{principal $i$ is chosen in round $t$ and $n_i(t)=n$} ]
    = \rew_i(n+1) \quad (\forall n\in\N).
\end{align*}

\xhdr{Agents' response.}
Each agent $t$ chooses principal $i_t$ as as follows: it chooses a distribution over the principals, and then draws independently from this distribution. Let $p_t$ be the probability of choosing principal $1$ according to this distribution. Below we specify $p_t$; we need to be careful so as to avoid a circular definition.

Let $\mI_t$ be the information available to agent $t$ before the
round. Assume $\mI_t$ suffices to form posteriors for quantities
$n_i(t)$, $i\in \{1,2\}$, denote them by $\posteriorN{i}{t}$. Note
that the Bayesian expected reward of each principal $i$ is a function
only of the number rounds he was chosen by the agents, so the
posterior mean reward for each principal $i$ can be written as
\begin{align*}
 \PMR_i(t) := \E[r_t \mid \text{$\mI_t$ and $i_t=i$} ]
    = \E[\rew_i(n_i(t)+1) \mid \mI_t]
    = \E_{n\sim \posteriorN{i}{t}}[\rew_i(n+1)].
\end{align*}
This quantity represents the posterior mean reward for principal $i$
at round $t$, according to information $\mI_t$; hence the notation \PMR. In general, probability $p_t$ is defined by the
posterior mean rewards $\PMR_i(t)$ for both principals. We assume a
somewhat more specific shape:
\begin{align}
p_t = \respF\left(\; \PMR_1(t) - \PMR_2(t) \;\right).
%p_t = \respF(\Delta_t), \quad \text{where } \Delta_t := \PMR_1(t) - \PMR_2(t).
\end{align}
Here $\respF:[-1,1]\to [0,1]$ is the \emph{response function}, which is the same for all agents. We assume that the response function is known to all agents.

To make the model well-defined, it remains to argue that information $\mI_t$ is indeed sufficient to form posteriors on $n_1(t)$ and $n_2(t)$. This can be easily seen using induction on $t$.

Since all agents arrive with identical information (other than knowing which global round they arrive in), it follows that all agents have identical posteriors for $n_{i,t}$ (for a given principal $i$ and a given global round $t$). This posterior is denoted $\posteriorN{i}{t}$.

\xhdr{Response functions.}
We use the response function $\respF$ to characterize the amount of rationality and competitiveness in our model. We assume that $\respF$ is monotonically non-decreasing, is larger than $\tfrac12$ on the interval $(0,1]$, and smaller than $\tfrac12$ on the interval $[-1,0)$. Beyond that, we consider three specific models, listed in the order of decreasing rationality and competitiveness (see \reffig{fig:response-functions}):

\begin{figure}
\begin{center}
  \begin{tikzpicture}[scale=2.0]
    \draw[->] (-1.1,0) -- (1.1,0) node[right] {$\Delta_t := \PMR_1(t) - \PMR_2(t)$};
    \draw[->] (0,-0.1) -- (0,1.1) node[above]
        {$p_t = \text{prob. of choosing principal 1}$};
    \draw[scale=1.0,domain=-1:0,smooth,variable=\q,blue, line width=0.50mm] plot ({\q},{0});
    \draw[scale=1.0,domain=0:1,smooth,variable=\q,blue,line width=0.50mm] plot ({\q},{1});
    \draw[scale=1.0,domain=-1:0,smooth,variable=\y,red]  plot ({\y},{0.1});
    \draw[scale=1.0,domain=0:1,smooth,variable=\y,red]  plot ({\y},{0.9});
    \draw[scale=1.0,domain=-1:1,smooth,variable=\y,cyan, line width=0.45mm, dash pattern=on 3pt off 2pt]  plot ({\y},{1/(1 + 1/(9^\y))});
    % \node[above, blue] at (0.5, 0.5) {\footnotesize $2 q (1 - q)$};
     \node[left] at (0, 0.5) {\footnotesize $1/2$};
     \node[left] at (0, 1) {\footnotesize $1$};
     \node[below left] at (0, 0) {\footnotesize $0$};
     \node[below ] at (1, 0) {\footnotesize $1$};
     \node[below ] at (-1, 0) {\footnotesize $-1$};
  \end{tikzpicture}
\end{center}
\caption{The three models for agents' response function: \HardMax is thick blue, \HardMaxRandom is slim red, and \SoftMaxRandom is the dashed curve.}
\label{fig:response-functions}
\end{figure}

\begin{OneLiners}
\item \HardMax: $\respF$ equals $0$ on the interval $[-1,0)$ and $1$
  on the interval $(0,1]$. In other words, the agents will
  deterministically choose the principal with the higher posterior
  mean reward.

\item \HardMaxRandom:
    % $\respF$ equals $\eps$ on the interval $[-1,0)$ and $1-\eps'$ on the interval $(0,1]$, where $\eps,\eps'\in (0,\tfrac12)$ are some positive constants. In words, each agent is a \HardMax agent with probability $1-\eps-\eps'$, and with the remaining probability she makes a random choice.
    $\respF$ equals $\eps_0$ on the interval $[-1,0)$ and $1-\eps_0$ on the interval $(0,1]$, where $\eps_0\in (0,\tfrac12)$ are some positive constants. In words, each agent is a \HardMax agent with probability $1-2\eps_0$, and with the remaining probability she makes a random choice.

\item \SoftMaxRandom: $\respF(\cdot)$ lies in the interval $[\eps_0,1-\eps_0]$, $\eps_0>0$, and is ``smooth" around $0$ (in the sense defined precisely in Section~\ref{sec:soft}).
\end{OneLiners}
% \asedit{(We assume $\respF(-1)+\respF(1)=1$ in \HardMaxRandom and \SoftMaxRandom only to simplify notation.)}

We say that $\respF$ is \emph{symmetric} if $\respF(-x)+\respF(x)=1$
for any $x\in [0,1]$. This implies \emph{fair tie-breaking}:
$\respF(0)=\tfrac12$.% , and $\eps=\eps'$ in the
% definitions above.

\xhdr{MAB algorithms.}
We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
\begin{align}
\BIR(n) := \E_{\mu\sim\priorMu}
    \left[ \max_{a\in A} \mu_a\right] - \rew(n),
\end{align}
where $\rew(n)$ is the Bayesian-expected reward of the algorithm for the $n$-th step, when the algorithm is run in isolation. We are primarily interested in how \BIR scales with $n$; we treat $K$, the number of arms, as a constant unless specified otherwise.

We will emphasize several specific algorithms or classes thereof:

\begin{itemize}
\item ``smart" MAB algorithms that combine exploration and exploitation, such as UCB1 \cite{bandits-ucb1} and Successive Elimination \cite{EvenDar-icml06}. These algorithms achieve
        $\BIR(n) \leq  \tilde{O}(n^{-1/2})$
    for all priors and all (or all but a very few) steps $n$. This bound is known to be tight for any fixed $n$.
    \footnote{This follows from the lower-bound analysis in \cite{bandits-exp3}.}

\item ``naive" MAB algorithms that separate exploration and exploitation, such as Explore-then-Exploit and $\eps$-Greedy. These algorithms have dedicated rounds in which they explore by choosing an action uniformly at random. When these rounds are known in advance, the algorithm suffers constant \BIR in such rounds. When the ``exploration rounds" are instead randomly chosen by the algorithm, one can usually guarantee an inverse-polynomial upper bound \BIR, but not as good as the one above: namely,
        $ \BIR(n) \leq \tilde{O}(n^{-1/3})$.
    This is the best possible upper bound on \BIR for the two algorithms mentioned above.

\item \DynGreedy: at each step, recommends the best action according to the current posterior: an action $a$ with the highest posterior expected reward
        $\E[\mu_a \mid \mI]$,
     where $\mI$ is the information available to the algorithm so far.
     \DynGreedy has (at least) a constant \BIR for some reasonable priors, \ie
        $\BIR(n)>\Omega(1)$.

\item \StaticGreedy: always recommends the prior best action,\ie an action $a$ with the highest prior mean reward  $\E_{\mu\sim \priorMu}[\mu_a]$. This algorithm typically has constant \BIR.
\end{itemize}

We focus on MAB algorithms such that $\BIR(n)$ is non-increasing; we call such algorithms \emph{monotone}. While some reasonable MAB algorithms may occasionally violate monotonicity, they can usually be easily modified so that monotonicity violations either vanish altogether, or only occur at very specific rounds (so that agents are extremely unlikely to exploit them in practice).

More background and examples can be found in Appendix~\ref{app:examples}. In particular, we prove that \DynGreedy is monotone.

\xhdr{Competition game between principals.}
Some of our results explicitly study the game between the two principals. We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm. Thus, choosing a pure strategy in this game corresponds to choosing an MAB algorithm (and, implicitly, announcing this algorithm to the agents).

Principal's utility is primarily defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral, in the sense that they optimize their expected utility.

\xhdr{Assumptions on the prior.} We make some technical assumptions for the sake of simplicity. First, each action $a$ has a positive probability of being the best action according to the prior:
\begin{align}\label{eq:assn-prob}
\forall a\in A:\;\;\;  \Pr_{\mu\sim \priorMu}[\mu_a  > \mu_{a'}
\;\forall a'\in A]
> 0.
\end{align}

Second, posterior mean rewards of actions are pairwise distinct almost
surely. That is, the history $h$ at any step of an MAB algorithm%
\footnote{The \emph{history} of an MAB algorithm at a given step
  comprises the chosen actions and the observed rewards in all
  previous steps in the execution of this algorithm.}
 satisfies
\begin{align}\label{eq:assn-distinct}
    \E[\mu_a \mid h] \neq \E[\mu_{a'}\mid h] \quad \forall a,a'\in A,
\end{align}
\asedit{except at a set of histories of probability $0$}.  In
particular, prior mean rewards of actions are pairwise distinct:
$\E[\mu_a] \neq \E[\mu_a']$ for any $a,a'\in A$.

We provide two examples for which property \eqref{eq:assn-distinct} is `generic', in the sense that it can be enforced almost surely by a
  small random perturbation of the prior. Both examples focus on 0-1 rewards and priors $\priorMu$ that are independent across arms. The first example
  assumes Beta priors on the mean rewards, and is very easy.%
  \footnote{Suppose the rewards are Bernouli r.v. and the mean reward
    $\mu_a$ for each arm $a$ is drawn from some Beta distribution
    $\text{Beta}(\alpha_a, \beta_a)$. Given any history that contains
    $h_a$ number of heads and $t_a$ number of tails from arm $a$, the
    posterior mean reward is
    $\frac{\alpha_a + h_a}{\alpha_a + h_a + \beta_a + t_a}$. Note that
    $h_a$ and $t_a$ take integer values. Therefore, perturbing the
    parameters $\alpha_a$ and $\beta_a$ independently with any
    continuous noise will induce a prior with property
    \eqref{eq:assn-distinct} with probability 1.  }  The second
  example assumes that mean rewards have a
  finite support, see Appendix~\ref{app:perturb} for details.

\xhdr{Some more notation.} Without loss of generality, we label actions as $A=[K]$ and sort them according to their prior mean rewards, so that
    $ \E[\mu_1] > \E[\mu_2] > \ldots > \E[\mu_K]$.

Fix principal $i\in \{1,2\}$ and (local) step $n$. The arm chosen by algorithm \alg[i] at this step is denoted $a_{i,n}$, and the corresponding \BIR is denoted $\BIR_i(n)$. History of \alg[i] up to this step is denoted $H_{i,n}$.
%The posterior distribution of agent $t$ for quantity $n_t(t)$ is denoted $\posteriorN{i}{t}$.

Write
    $\PMR(a\mid E) = \E[\mu_a \mid E]$
for posterior mean reward of action $a$ given event $E$.

%\end{assumption}

\subsection{Generalizations}
\label{sec:model-extensions}

Our results can be extended compared to the basic model described above.

%First, the two principals may have different action sets, as long as
%    $\E\left[ \max_{a\in A} \mu_a\right]$,
%the Bayesian-expected reward of the best action, is the same for both principals.

First, unless specified otherwise, our results allow a more general notion of principal's utility that can depend on both the market share and agents' rewards. Namely, principal $i$ collects $U_i(r_t)$ units of utility in each global round $t$ when she is chosen (and $0$ otherwise), where $U_i(\cdot)$ is some fixed non-decreasing function with $U_i(0)>0$. In a formula,
\begin{align}\label{eq:general-utility}
\textstyle U_i := \sum_{t=1}^T\; \indicator{i_t=i}\cdot U_i(r_r).
\end{align}

Second, our results carry over, with little or no modification of the proofs, to much more general versions of MAB, as long as it satisfies the i.i.d. property. In each round, an algorithm can see a \emph{context} before choosing an action (as in \emph{contextual bandits}) and/or additional feedback other than the reward after the reward is chosen (as in, \eg \emph{semi-bandits}), as long as the contexts are drawn from a fixed distribution, and the (reward, feedback) pair is drawn from a fixed distribution that depends only on the context and the chosen action. The Bayesian prior $\prior$ needs to be a more complicated object, to make sure that \PMR and \BIR are well-defined. Mean rewards may also have a known structure, such as Lipschitzness, convexity, or linearity; such structure can be incorporated via $\prior$. All these extensions have been studied extensively in the literature on MAB, and account for a substantial segment thereof; see \cite{Bubeck-survey12} for background and details.

\subsection{Chernoff Bounds}

We use an elementary concentration inequality known as {\em Chernoff Bounds}, in a formulation from~\cite{MitzUpfal-book05}.
\begin{theorem}[Chernoff Bounds]
\label{thm:chernoff}
Consider $n$ i.i.d. random variables $X_1 \ldots X_n$ with values in $[0,1]$. Let
    $X = \tfrac{1}{n} \sum_{i=1}^n X_i$ be their average, and let $\nu = \E[X]$. Then:
\[ \min\left(\; \Pr[ X-\nu > \delta \nu ],\quad
                \Pr[ \nu-X > \delta \nu ]
    \; \right)
    < e^{-\nu n \delta^2/3}
    \quad \text{for any $\delta\in (0,1)$.}
\]
\end{theorem}

%, and the choice of algorithm are commonly known
%to the agents. Given any finite class of MAB learning algorithms
%$\Pi$, a pure strategy for the principal $j$ in the competition game
%is an algorithm $\pi_j$ in the class $\Pi$ and a mixed strategy
%$\sigma_j$ is a probability distribution over $\Pi$. We will write
%$u_j(\sigma)$ and $u(\pi)$ to denote the expected utility for the
%principal $j$ in mixed strategy profile $\sigma$ and pure strategy
%profile $\pi$.  The principal gets a unit reward for being selected by
%an agent, so the utility function can be written as
%\[
%u_j(\pi) = \Ex{}{\sum_{t=1}^T\mathbf{1}[P^t = j]}
%\]
%A variant of the utility function we have is the number of agents
%weighted by the rewards for the agents
%\[
%u_j(\pi) = \Ex{}{\sum_{t=1}^T \mathbf{1}[P^t = j] r^t}
%\]
%\iffalse
%We consider a few models for the utility of the principals.  For
%example, the principal gets the reward realized by the agents it
%advised.
%\fi
%
%We will analyze the equilibrium strategies of the game under different
%environments.
%
%\begin{definition}[Equilibrium in the competition game]
%  A pure strategy profile $\pi$ is a pure strategy Nash equilibrium (PSNE) if
%  for any principal $j$ in the game and any algorithm $a\in \Pi$
%\[
%  u_j(\pi) \geq u_j(a, \pi_{-j}).
%\]
%A mixed strategy profile $\sigma$ is a (mixed strategy) Nash
%equilibrium if for any principal $j$ and any algorithm $a\in\Pi$,
%\[
%  \Ex{\sigma}{u_j(\sigma)} \geq u_j(a, \sigma_{-j}).
%\]
%\end{definition}
%
%We will measure the quality of an equilibrium by studying the welfare
%of the agents with respect to the system regret.

%%% Local Variables:
%%% TeX-master: "main.tex"
%%% End:

### 2019-08-14 journal paper/itcs18paper/sec-rational.tex
the latex compiles

In this section, we will consider the version in which
the agents are fully rational, in the sense that their response function is \HardMax. We show that principals are not incentivized to \emph{explore}--- \ie to deviate from \DynGreedy. The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates.

To make this more precise, let us say that two MAB algorithms \emph{deviate} at (local) step $n$ if there is an action $a\in A$ and \asedit{a set of step-$n$ histories of positive probability such that any history $h$ in this set} is feasible for both algorithms, and under this history the two algorithms choose action $a$ with different probability.

\iffalse
\begin{definition}
  Suppose that an algorithm $\pi$ is not dynamic greedy. Then
  {\em the first exploration} for $\pi$ is the first time step where
  there is a non-zero probability that the recommended action has a
  lower expectation than exploitation option.\sw{define...}
\end{definition}
\fi

\begin{theorem}\label{thm:DG-dominance}
Assume \HardMax response function with fair tie-breaking. Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$. Then all agents in global rounds $t\geq n_0$ select principal $1$.
\end{theorem}

\begin{corollary}\label{cor:DG-dominance}
The competition game between principals has a unique Nash equilibirium: both principals choose \DynGreedy.
\end{corollary}

\begin{remark}
This corollary holds under a more general model which allows time-discounting: namely, the utility of each principal $i$ in each global round $t$ is $U_{i,t}(r_t)$ if this principal is chosen, and $0$ otherwise, where $U_{i,t}(\cdot)$ is an arbitrary non-decreasing function with $U_{i,t}(0)>0$.
\end{remark}

\subsection{Proof of Theorem~\ref{thm:DG-dominance}}

The proof starts with two auxiliary lemmas: that deviating from \DynGreedy implies a strictly smaller Bayesian-expected reward, and that \HardMax implies a ``sudden-death" property: if one agent chooses principal $1$ with certainty, then so do all subsequent agents do. \asedit{We re-use both lemmas in later sections, so we state them in sufficient generality.}

\begin{lemma}\label{lm:DG-rew}
\asedit{Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$. Then $\rew_1(n_0)>\rew_2(n_0)$. This holds for any response function $\respF$.}
\end{lemma}

\asedit{Lemma~\ref{lm:DG-rew} does not rely on any particular shape of the response function because it only considers the performance of each algorithm without competition.}

\begin{proof}[Proof of Lemma~\ref{lm:DG-rew}]
Since the two algorithms coincide on the first $n_0-1$ steps, it follows by symmetry that histories $H_{1,n_0}$ and $H_{2,n_0}$ have the same distribution. We use a \emph{coupling argument}: w.l.o.g., we assume the two histories coincide,
    $H_{1,n_0} = H_{2,n_0} = H$.

At local step $n_0$, \DynGreedy chooses an action $a_{1,n_0} = a_{1,n_0}(H)$
which maximizes the posterior mean reward given history $H$: for any realized history $h\in \support(H)$ and any action $a\in A$
\begin{align}\label{eq:lm:DG-rew-1}
 \PMR(a_{1,n_0} \mid H = h) \geq \PMR(a \mid H=h).
\end{align}

\ascomment{Rewrote the rest of the proof to account for positive-prob set of histories.}

By assumption \eqref{eq:assn-distinct}, it follows that 
\begin{align}\label{eq:lm:DG-rew-2}
 \PMR(a_{1,n_0} \mid H = h) > \PMR(a \mid H=h)
 \quad \text{for any $h\in \support(H)$ and $a\neq a_{1,n_0}(h)$.}
\end{align}

Since the two algorithms deviate at step $n_0$, there is a set $S\subset \support(H)$ of step-$n_0$ histories such that $\Pr[S]>0$ and any history $h\in S$ satisfies
    $\Pr[a_{2,n_0}\neq a_{1,n_0} \mid H=h]>0$.
Combining this with \eqref{eq:lm:DG-rew-2}, we deduce that 
\begin{align}\label{eq:lm:DG-rew-3}
 \PMR(a_{1,n_0} \mid H = h) > \E\left[ \mu_{a_{2,n_0}}\mid H=h \right]
 \quad\text{for each history $h\in S$}.
\end{align}
Using \eqref{eq:lm:DG-rew-1} and \eqref{eq:lm:DG-rew-3} and integrating over realized histories $h$, we obtain
    $\rew_1(n_0) > \rew_2(n_0)$.
\end{proof}

\begin{lemma}\label{lm:DG-sudden}
\asedit{Consider \HardMax response function with $\respF(0)\geq\tfrac12$.} 
Suppose \alg[1] is monotone, and $\PMR_1(t_0)>\PMR_2(t_0)$ for some global round $t_0$. Then $\PMR_1(t)>\PMR_2(t)$ for all subsequent rounds $t$.
\end{lemma}

\begin{proof}
Let us use induction on round $t\geq t_0$, with the base case $t=t_0$. Let $\mN=\mN_{1,t_0}$ be the agents' posterior distribution for $n_{1,t_0}$, the number of global rounds before $t_0$ in which principal $1$ is chosen. By induction, all agents from $t_0$ to $t-1$ chose principal $1$, so $\PMR_2(t_0)= \PMR_2(t)$. Therefore,
\[ \PMR_1(t)
    = \Ex{n\sim \mN}{\rew_1(n+1+t-t_0)}
    \geq \Ex{n\sim \mN}{\rew_1(n+1)}
    =\PMR_1(t_0) > \PMR_2(t_0)= \PMR_2(t), \]
where the first inequality holds because \alg[1] is monotone, and the second one is the base case.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:DG-dominance}]
Since the two algorithms coincide on the first $n_0-1$ steps, it follows by symmetry that $\rew_1(n) = \rew_2(n)$ for any $n< n_0$.
By Lemma~\ref{lm:DG-rew},
    $\rew_1(n_0) > \rew_2(n_0)$.

Recall that $n_i(t)$ is the number of global rounds $s<t$ in which principal $i$ is chosen, and $\posteriorN{i}{t}$ is the agents' posterior distribution for this quantity. By symmetry, each agent $t<n_0$ chooses a principal uniformly at random. It follows that
    $\posteriorN{1}{n_0} = \posteriorN{2}{n_0}$
(denote both distributions by $\mN$ for brevity), and $\mN(n_0-1)>0$.
Therefore:
\begin{align}
\PMR_1(n_0)
  &= \Ex{n\sim \mN} {\rew_1(n+ 1)}
   = \sum_{n = 0}^{n_0-1} \mN(n) \cdot{\rew_1(n + 1)} \nonumber \\
  & > \mN(n_0-1)\cdot {\rew_2(n_0)} + \sum_{n = 0}^{n_0-2}  \mN(n)\cdot{\rew_2(n + 1)}
    \nonumber \\
  &= \Ex{n\sim \mN}{\rew_2(n + 1)} = \PMR_2(n_0) \label{eq:pf:thm:DG-dominance}
\end{align}
So, agent $n_0$ chooses principal $1$. By Lemma~\ref{lm:DG-sudden} \asedit{(noting that \DynGreedy is monotone)}, all subsequent agents choose principal $1$, too.
\end{proof}

%\begin{theorem}
%  Suppose that the agents break ties uniformly at random, that is
%  $q=1/2$, then both principals playing \DynGreedy is the unique
%  equilibrium.
%\end{theorem}
%
%\begin{proof}
%  Suppose that both principals play dynamic greedy. The result
%  of Theorem~\ref{thm:DG-dominance} shows that there is no beneficial deviation
%  for both principals, so this forms a pure strategy equilibrium.
%
%  Now we will show that no other strategy profile can form equilibrium
%  in this game.  Suppose that both principals play some mixed
%  strategies that places non-zero probability on algorithms other than
%  dynamic greedy. We know one of the principals $p_j$ is getting no
%  more than $T/2$ agents in expectation. If $p_j$ switch to playing
%  only dynamic greedy, he can get strictly more than $T/2$ agents
%  by Theorem~\ref{thm:DG-dominance}. Thus, such a strategy profile will not form
%  an equilibrium.  Finally, suppose that $p_1$ plays dynamic greedy
%  and $p_2$ plays algorithms other than dynamic greedy with non-zero
%  probability. By Theorem~\ref{thm:DG-dominance}, we know that $p_2$ gets less
%  than $T/2$ agents in expectation. This means switching to dynamic
%  greedy will strictly improve the expected utility to $T/2$, so such
%  strategy profile cannot form an equilibrium either.
%\end{proof}

\subsection{\HardMax with biased tie-breaking}
\label{sec:HardMax-biased}

The \HardMax model is very sensitive to the tie-breaking rule. For starters, if ties are  broken deterministically in favor of principal $1$, then principal 1 can get all agents no matter what the other principal does, simply by using \StaticGreedy.

\begin{theorem}\label{thm:HardMax-hardTies}
Assume \HardMax response function with $\respF(0)=1$ (ties are always broken in favor of principal $1$). If \alg[1] is \StaticGreedy, then all agents choose principal $1$.
\end{theorem}

\begin{proof}
Agent $1$ chooses principal $1$ because of the tie-breaking rule. Since \StaticGreedy is trivially monotone, all the subsequent agents choose principal $1$ by an induction argument similar to the one in the proof of Lemma~\ref{lm:DG-sudden}.
\end{proof}

A more challenging scenario is when the tie-breaking is biased in favor of principal 1, but not deterministically so: $\respF(0)>\tfrac12$. Then this principal also has a ``winning strategy" no matter what the other principal does. Specifically, principal 1 can get all but the first few agents, under a mild technical assumption that \DynGreedy deviates from \StaticGreedy. Principal 1 can use \DynGreedy, or any other monotone MAB algorithm that coincides with \DynGreedy in the first few steps.

%We can generalize the theorem top the case of $q >1/2$ if the
%principal $p_1$ can guarantee better than the a priori best action to
%all the agents following the second.

\begin{theorem}\label{thm:HardMax-biased}
Assume \HardMax response function with $\respF(0)>\tfrac12$ (\ie tie-breaking is biased in favor of principal $1$). Assume the prior $\mP$ is such that \DynGreedy deviates from \StaticGreedy starting from some step $n_0$. Suppose that principal $1$ runs a monotone MAB algorithm that coincides with \DynGreedy in the first $n_0$ steps. Then all agents $t\geq n_0$ choose principal $1$.
\end{theorem}

\begin{proof}
The proof re-uses Lemmas~\ref{lm:DG-rew} and~\ref{lm:DG-sudden}, which do not rely on fair tie-breaking.

Because of the biased tie-breaking, for each global round $t$ we have:
\begin{align}\label{eq:thm:HardMax-biased-PMRtoPr}
\text{if $\PMR_1(t)\geq \PMR_2(t)$ then $\Pr[i_t=1]>\tfrac12$.}
\end{align}
Recall that $i_t$ is the principal chosen in global round $t$.

Let $m_0$ be the first step when \alg[2] deviates from \DynGreedy, or \DynGreedy deviates from \StaticGreedy, whichever comes sooner. \asedit{Then \alg[2], \DynGreedy and \StaticGreedy coincide on the first $m_0-1$ steps. Moreover, $m_0\leq n_0$ (since \DynGreedy deviates from \StaticGreedy at step $n_0$), so \alg[1] coincides with \DynGreedy on the first $m_0$ steps.

So, $\rew_1(n)=\rew_2(n)$ for each step $n<m_0$, because \alg[1] and \alg[2] coincide on the first $m_0-1$ steps. Moreover, if \alg[2] deviates from \DynGreedy at step $m_0$ then
    $\rew_1(m_0) > \rew_2(m_0)$ 
by Lemma~\ref{lm:DG-rew}; else, we trivially have
    $\rew_1(m_0) = \rew_2(m_0)$.} To summarize:
\begin{align}\label{eq:thm:HardMax-biased-rew}
    \rew_1(n)\geq\rew_2(n)\quad \text{for all steps $n\leq m_0$}.
\end{align}

We claim that $\Pr[i_t=1]>\tfrac12$ for all global rounds $t\leq m_0$. We prove this claim using induction on $t$. The base case $t=1$ holds by \eqref{eq:thm:HardMax-biased-PMRtoPr} and the fact that in step 1, \DynGreedy chooses the arm with the highest prior mean reward. For the induction step, we assume that $\Pr[i_t=1]>\tfrac12$ for all global rounds $t<t_0$, for some $t_0\leq  m_0$. It follows that distribution $\posteriorN{1}{t_0}$ stochastically dominates distribution $\posteriorN{2}{t_0}$.%
\footnote{For random variables $X,Y$ on \R, we say that $X$ \emph{stochastically dominates} $Y$ if $\Pr[X\geq x] \geq \Pr[Y\geq x]$ for any $x\in \R$.}
Observe that
\begin{align}\label{eq:thm:HardMax-biased-PMR-aux}
\PMR_1(t_0)
  = \Ex{n\sim \posteriorN{1}{t_0}} {\rew_1(n+1)}
  \geq \Ex{n\sim \posteriorN{2}{t_0}} {\rew_2(n+1)}
  = \PMR_2(t_0).
\end{align}
So the induction step follows by \eqref{eq:thm:HardMax-biased-PMRtoPr}. Claim proved.

Now let us focus on global round $m_0$, and denote $\mN_i = \posteriorN{i}{m_0}$.  By the above claim,
\begin{align}\label{eq:thm:HardMax-biased-mN}
\text{$\mN_1$ stochastically dominates $\mN_2$, and moreover
    $\mN_i(m_0-1)>\mN_i(m_0-1)$.}
\end{align}

By definition of $m_0$, either (i) \alg[2] deviates from \DynGreedy starting from local step $m_0$, which implies $\rew_1(m_0)> \rew_2(m_0)$ by Lemma~\ref{lm:DG-rew}, or (ii) \DynGreedy deviates from \StaticGreedy starting from local step $m_0$, which implies $\rew_1(m_0)>\rew_1(m_0-1)$ by Lemma~\ref{dgmono}. In both cases, using \eqref{eq:thm:HardMax-biased-rew} and \eqref{eq:thm:HardMax-biased-mN}, it follows that the inequality in \eqref{eq:thm:HardMax-biased-PMR-aux} is strict for $t_0=m_0$.

Therefore, agent $m_0$ chooses principal $1$, and by Lemma~\ref{lm:DG-sudden} so do all subsequent agents.
\end{proof}

% OLD PROOF (assuming alg1 coincides with DG on first two rounds, etc.
% We consider two cases, depending on the arm $a_{2,1}$ that principal $2$ chooses in the same step. The interesting case is $a_{2,1}=1$.
%Then agent $1$ selects principal $1$ with probability $q:=\respF(0)>\tfrac12$. Therefore, agent $2$ forms the following posterior mean rewards:
%\begin{align*}
%  \PMR_1(2) &= (1-q)\cdot \E[\mu_1] + q\cdot \rew_1(2)
%    \EqComment{for principal 1}\\
%  \PMR_2(2) &= q\cdot \E[\mu_1] + (1-q)\cdot \rew_2(2)
%    \EqComment{for principal 2}.
%\end{align*}
%We know that
%    $\rew_1(2) > \E[\mu_1]$
%(\DynGreedy is strictly monotone in the first $2$ rounds)
%from Lemma~\ref{dgmono}, and
%    $\rew_1(2) \geq \rew_2(2)$
%by Lemma~\ref{lm:DG-rew}. It follows that
%    $\PMR_1(2)>\PMR_2(2)$,
%and consequently agent $2$ chooses principal $1$. The subsequent agents choose principal $1$ by Lemma~\ref{lm:DG-sudden}.
%
%The remaining case is $a_{2,1}\neq 1$: principal $p_2$ does not recommend arm $1$ in the first step. Since all other arms have a strictly lower prior mean reward, agent $1$ chooses principal $1$. The subsequent agents choose principal $1$ by Lemma~\ref{lm:DG-sudden}.

% \begin{lemma}
%   Let $\mu^2_{DG}$ the expected reward of action of \DynGreedy
%   in the second step. Then $\mu^2_{DG} > \mu_1$.
% \end{lemma}

% \begin{proof}
%   \sw{todo. This requires an assumption that the second arm has
%     non-zero probability of beating the first one.}
% \end{proof}

% Steven's proof for a more general version with alg1 = BIC algo
%Agent $1$ chooses principal $1$ because of the tie-breaking rule,
%  Since principal $p_1$ in the first step recommends arm 1, the agent
%  selects principal $p_1$ (due to the tie breaking rule).  Agent $2$
%  knows that agent $1$ selected $p_1$. Because of BIC, she knows $p_1$
%  will offer reward no lower than $p_2$ in expectation.  By induction,
%  all the agents select principal $p_1$.

%\sw{We also have an observation saying if one principal starts with a
%  better dynamic greedy will just dominate the game. It could do some
%  amount of exploration while having better Bayesian per-round regret
%  than the other principal. This means the other principal will never
%  have any agents arriving for learning.}

%%% Local Variables:
%%% TeX-master: "main.tex"
%%% End:

### 2019-08-14 journal paper/journal_main.tex
the latex compiles

--- 2019-04-30 paper/competing_bandits.tex
@@
 
-\usepackage{acm-ec-19}
+\usepackage{ec19paper/acm-ec-19}
+\usepackage{itcs18paper/slivkins-setup,itcs18paper/slivkins-theorems}
 \usepackage{tikz} % SW: Added for plots
@@
 
-% Alex' macro
-\usepackage{xspace}
-\newcommand{\OMIT}[1]{}
-\newcommand{\xhdr}[1]{\vspace{1mm} \noindent{\bf #1}}
-\newcommand{\ie}{{\em i.e.,~\xspace}}
-\newcommand{\eg}{{\em e.g.,~\xspace}}
-
+% Commands from EC 19 paper
 \newcommand{\term}[1]{\ensuremath{\mathtt{#1}}\xspace}
@@
 \newcommand{\DG}{\term{DG}}
+\newcommand{\AD}{\term{AD}}
+\newcommand{\EGR}{\term{EGR}}
+\newcommand{\GR}{\term{GR}}
 \newcommand{\Beta}{\term{Beta}} % for Beta distribution
 \newcommand{\Eeog}{\term{EoG}} % shorthand for "effective end of game"
-
-\newcommand{\eps}{\varepsilon}
 
@@
 \newcommand{\HM}{\term{HM}}
+
+% commands from ITCS 18 paper
+
+\DeclareMathOperator*{\Expectation}{\mathbb{E}}
+\newcommand{\Ex}[2]{\Expectation_{#1}\left[#2\right]}
+
+% notation
+%%% advanced notation
+\newcommand{\OPT}{\term{OPT}}
+\newcommand{\rew}{\term{rew}}  % Bayesian-expected reward after n local rounds
+\newcommand{\PMR}{\term{PMR}} % posterior mean reward
+\newcommand{\support}{\term{support}}
+
+\newcommand{\BIR}{\term{BIR}} % Bayesian Instantaneous Regret
+\newcommand{\regret}{R^{\term{inst}}} % regret
+\newcommand{\regretWC}{\regret_{\term{wc}}} % regret
+
+% response function
+\newcommand{\respF}{f_{\term{resp}}}
+\newcommand{\respEps}{\eps_\term{resp}}
+
+\newcommand{\BReg}{\term{BReg}}
+\newcommand{\HardMax}{\term{HardMax}}
+\newcommand{\HardMaxRandom}{\term{HardMax\&Random}}
+\newcommand{\SoftMaxRandom}{\term{SoftMax}}
+\newcommand{\Uniform}{\term{Uniform}}
+\newcommand{\Random}{\term{Random}}
+
+\newcommand{\StaticGreedy}{\term{StaticGreedy}}
+\newcommand{\DynGreedy}{\term{DynamicGreedy}}
+\newcommand{\DynamicGreedy}{\term{DynamicGreedy}}
+
+\newcommand{\termSub}[2]{\ensuremath{\mathtt{#1}_{#2}}\xspace}
+\newcommand{\alg}[1][]{\termSub{alg}{#1}}
+%\newcommand{\prin}[1][]{\termSub{prin}{#1}}  % principal
+\newcommand{\agent}[1][]{\termSub{agent}{#1}}
+
+% priors and posteriors
+\newcommand{\prior}{\ensuremath{\mP}\xspace}
+\newcommand{\priorMu}{\ensuremath{\prior_\mathtt{mean}}\xspace}
+\newcommand{\posteriorN}[2]{\mN_{#1,#2}}  % \posteriorN{principal}{round}
+
+% rationality / innovation / competition
+\newcommand{\termTXT}[1]{{\em {#1}}\xspace}
+
+\newcommand{\rationality}{\termTXT{rationality}}
+\newcommand{\Rationality}{\termTXT{Rationality}}
+\newcommand{\innovation}{\termTXT{innovation}}
+\newcommand{\Innovation}{\termTXT{Innovation}}
+\newcommand{\competition}{\termTXT{competition}}
+\newcommand{\Competition}{\termTXT{Competition}}
+\newcommand{\competitiveness}{\termTXT{competitiveness}}
+\newcommand{\exploration}{\termTXT{exploration}}
+\newcommand{\Exploration}{\termTXT{Exploration}}
 % a very useful package for edits and comments, from David Kempe (USC)
-\usepackage{color-edits}
+\usepackage{ec19paper/color-edits}
 %\usepackage[suppress]{color-edits}  % use this to suppress the package
@@
 \begin{document}
-\title[The Perils of Exploration under Competition: A Computational Modeling Approach]
-{The Perils of Exploration under Competition: \\ A Computational Modeling Approach}
+\title[Competing Bandits]
+{Competing Bandits}
 
@@
  \affiliation{ \institution{Columbia University}}
- \author{Kevin Liu}
- \affiliation{\institution{Columbia University}}
  \author{Aleksandrs Slivkins}
@@
  \authorsaddresses{%
-  Authors' addresses: Guy Aridor, Columbia University, New York, NY, g.aridor@columbia.edu; Kevin Liu, Columbia University, New York, NY, kevin.liu@columbia.edu; Aleksandrs Slivkins, Microsoft Research, New York, NY, slivkins@microsoft.com; Zhiwei Steven Wu, University of Minnesota - Twin Cities, Minneapolis, MN, zsw@umn.edu}
+  Authors' addresses: Guy Aridor, Columbia University, New York, NY, g.aridor@columbia.edu; Aleksandrs Slivkins, Microsoft Research, New York, NY, slivkins@microsoft.com; Zhiwei Steven Wu, University of Minnesota - Twin Cities, Minneapolis, MN, zsw@umn.edu}
 
@@
 
-\subfile{content/introduction}
+\section{Introduction}
+\label{sec:intro}
+\subfile{itcs18paper/sec-intro}
 
-\subfile{content/model}
+\section{Related work}
+\label{sec:related-work}
+\subfile{itcs18paper/sec-related-work}
 
-\subfile{content/perf_in_iso}
+\section{Our model and preliminaries}
+\label{sec:model}
+\subfile{itcs18paper/sec-model}
 
-\subfile{content/inverted_u}
+\section{Full rationality (HardMax)}
+\label{sec:rational}
+\subfile{itcs18paper/sec-rational}
 
-\subfile{content/barriers}
+\section{Relaxed rationality: HardMax \& Random}
+\label{sec:random}
+\subfile{itcs18paper/sec-random}
 
-\subfile{content/revisited}
+\section{SoftMax response function}
+\label{sec:soft}
+\subfile{itcs18paper/sec-soft}
 
-\subfile{content/non_greedy_choice}
+\section{Economic implications}
+\label{sec:welfare}
+\subfile{itcs18paper/sec-welfare}
 
-\subfile{content/conclusion}
+\subfile{ec19paper/content/introduction}
+
+\subfile{ec19paper/content/model}
+
+\subfile{ec19paper/content/perf_in_iso}
+
+\subfile{ec19paper/content/inverted_u}
+
+\subfile{ec19paper/content/barriers}
+
+\subfile{ec19paper/content/revisited}
+
+\subfile{ec19paper/content/non_greedy_choice}
+
+\subfile{ec19paper/content/conclusion}
+
 \newpage
 \bibliographystyle{acm}
-\bibliography{refs,bib-abbrv-short,bib-bandits,bib-AGT,bib-slivkins}
+\bibliography{bib-ML,refs,bib-abbrv-short,bib-bandits,bib-AGT,bib-slivkins}
 
-\subfile{content/appendix_for_one_version}
+\subfile{ec19paper/content/appendix_for_one_version}
 

### 2019-08-18 journal paper/content/sec-intro.tex
first pass through intro and model section

Learning from interactions with users is ubiquitous in modern customer-facing systems, from product recommendations to web search to spam detection to content selection to fine-tuning the interface. Many systems purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information. Randomized controlled trials, a.k.a. A/B testing, are an industry standard, with a number of companies such as \emph{Optimizely} offering tools and platforms to facilitate them. Many companies use more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known theoretical framework for exploration and making decisions under uncertainty.

%~\cite{KohaviAB-2015,KohaviLSH09}

Systems that engage in exploration typically need to compete against one another; most importantly, they compete for users. This creates an interesting tension between \exploration and \competition. In a nutshell, while exploring may be essential for improving the service tomorrow, it may degrade quality and make users leave \emph{today}, in which case there will be no users to learn from! Thus, users play three distinct roles: they are customers that generate revenue, they generate data for the systems to learn from, and they are self-interested agents which choose among the competing systems.

We initiate a study of the interplay between \exploration and \competition. The main high-level question is: {\bf whether and to what extent competition incentivizes adoption of better exploration algorithms}. \gaedit{This translates into a number of more concrete questions. While it is commonly assumed that better learning technology always helps, is this so for our setting? Does increasing competition lead to higher consumer welfare? To what extent does one firm possessing more information or having a better perception amongst consumers than its competitors serve as a barrier to entry in this environment?

We investigate these questions for two separate but connected models. In the first we suppose that consumers are well-informed about the firm's exploration strategies but do not observe any information about past user experiences with the firms. In the second, we suppose that consumers \textit{only} consider the average of past user experiences with each of the firms. In the first model we analytically characterize the equilibrium exploration strategies of two firms competing for the same stream of consumers as we vary the degree of randomness in consumer decisions.\footnote{The randomness in consumer decisions can be interpreted both as the degree of rationality as well as a reduced-form way of capturing the degree to which consumers themselves engage in exploration} In the second model we account for competition in a more direct way and utilize numerical simulations to consider asymmetries in the timing of entry, increasing competition beyond two firms, and allow us to isolate the effect that role that data can play as a barrier to entry.

On a high level, our contributions can be framed in terms of the ``inverted-U relationship" between competition and the quality of adopted algorithms (see \reffig{fig:inverted-U}).

}

\begin{figure}
\begin{center}
\begin{tikzpicture}
      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
 \end{tikzpicture}

\caption{Inverted-U relationship between rationality/competitiveness and algorithms.}
\label{fig:inverted-U}
\end{center}
\end{figure}

% the extent to which the game between the two principals is competitive
% degree of innovation that these models incentivize.
% the extent to which agents make rational decisions

%\subsection{Our model}
%\label{sec:intro-model}

\xhdr{Our model.} We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}). These two processes are interlinked, as exploration decisions are experienced by users and informed by their feedback. We need to specify several conceptual pieces: how the principals and agents interact, what is the machine learning problem faced by each principal, and what is the information structure. Each piece can get rather complicated in isolation, let alone jointly, so we strive for simplicity. Thus, the basic model is as follows:

\begin{itemize}

\item A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals. The principal chooses an action (\eg a list of web search results to show to the agent), the user experiences this action, and reports a reward. All agents have the same ``decision rule" for choosing among the principals given the available information.

\item Each principal faces a very basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a. \emph{arms}) and receives a reward drawn independently from a fixed distribution specific to this action.

\item Principals simultaneously announce their learning algorithms before round $1$, and cannot change them afterwards. There is a common Bayesian prior on the rewards (but the realized reward distributions are not observed by the principals or the agents).  Each principal only observes agents that chose him. \gaedit{The information set for the agent differs between the two variants of the model. \begin{enumerate}
\item In the first, agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms
\item In the second, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal
\end{enumerate}
}
\end{itemize}

\gaedit{
\xhdr{Main Findings.}
In both variants of our model, we find that in a simultaneous entry duopoly a greedy algorithm that does no purposeful exploration is incentivized in equilibrium. In the first variant of the model, once a firm does any exploration it is immediately starved of consumers when its opponent plays a greedy algorithm since we show that the expected reward for every subsequent consumer is higher for the firm that plays the greedy algorithm. In the second variant of the model, the same mechanism drives the greedy algorithm to be incentivized in equilibrium. In this variant, exploration hurts a firm's reputation and decrease its market share in the near term. This leaves the firm with less users to learn from, which may further degrade the firm's performance relative to competitors who keep learning and improving from \emph{their} customers, and so forth. Taken to the extreme, such dynamics lead to a ``death spiral" effect when the vast majority of customers eventually switch to competitors. In both cases, consumer welfare is low since the greedy algorithm is known to be dramatically bad in many important cases of multi-armed bandits.

The primary mechanism that generates this stark result is the same in both variants of the model: consumers need to be incentivized to select a firm over its competitors, leading firms that engage in exploration to be starved of consumers before they make enough progress on their learning problem. In order to incentivize ``better" exploration strategies the key intuition is that the firm needs to have some ``free" consumers to visit them without having to incentivize them. This allows the firm to eventually get out of the hole of exploration.

In the first variant of the model, we relax the decision rule of the consumers to allow for each firm to be chosen with some fixed baseline probability so that firms get a constant stream of ``free" consumers. This can be interpreted as a reduced-form relaxation of the assumption that consumers are myopic and themselves may be engaging in exploration or, alternatively, as deviations from rational behavior. We find that, in this setting, better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase. While the precise notion of ``sufficiently better algorithm" is rather subtle, we note that commonly known ``smart" bandit algorithms typically defeat the commonly known ``naive" ones, and the latter typically defeat the greedy algorithm. However, there is a substantial caveat: one can defeat any algorithm by interleaving it with the greedy algorithm. This has two undesirable corollaries: a better algorithm may sometimes lose, and a pure Nash equilibrium typically does not exist.

In the second variant of the model, we keep the decision rule of the consumers but instead vary the timing of entry and allow one firm to have a first-mover advantage so that this firm (denoted as the incumbent) gets all the agents in the market before the other firm enters. We find that, if the first-mover period is sufficiently large, then the incumbent has a dominant strategy to play ``smart" bandit algorithms and that consumer welfare is substantially higher than in the case of simultaneous entry. The intuition is simply that a sufficiently long incumbency period allows the firm to make sufficient progress on its learning problem that it can overcome the original drop in its reputation from exploration in the beginning rounds.

\xhdr{Additional Findings} In the first variant of the model, we further relax the decision rule of the agents so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' expected rewards; we call it \SoftMaxRandom. For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm. This is the most technical result of the paper. The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm may attract slightly more).

In the second variant of the model, we investigate the ``first-mover advantage" phenomenon in more detail. Being first in the market gives free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage"). We run additional experiments so as to isolate and compare these two effects. We find that either effect alone leads to a significant advantage under competition. The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.

Data advantage is significant from an anti-trust perspective, as a possible barrier to entry. We find that even a small amount ``data advantage" gets amplified under competition, causing a large difference in eventual market shares. This observation runs contrary to prior work  \cite{lambrecht2015can,bajari2018impact}, which studied learning without competition, and found that small amounts of additional data do not provide significant improvement in eventual outcomes. We conclude that competition dynamics -- that firms compete as they learn over time -- are pertinent to these anti-trust considerations.

We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition. We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor. We suggest a
more refined performance measure, and use it to explain some of the competition outcomes.

\OMIT{All results extend to a much more general version of the multi-armed bandit problem in which the principal may observe additional feedback before and/or after each decision, as long as the feedback distribution does not change over time. In most results, principal's utility may depend on both the market share and agents' rewards.
}
\xhdr{Economic interpretation.}
The inverted-U relationship between the severity of competition among firms and the quality of technologies that they adopt is a familiar theme in the economics literature \citep[\eg][]{Aghion-QJE05,Vives-08}.%
\footnote{The literature frames this relationship as one between ``competition" and ``innovation". In this context, ``innovation" refers to adoption of a better technology, at a substantial R\&D expense to a given firm. It is not salient whether similar ideas and/or technologies already exist outside the firm. It is worth noting that the adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature; see \citet{MWT-WhitePaper-2016} for an example of such R\&D effort.}
We find it illuminating to frame our contributions in a similar manner, as illustrated in \reffig{fig:inverted-U}.

In our model, competition varies based on the number of ``free" consumers that a firm receives. In the first variant of the model, this comes from introducing ``random" consumers where we vary from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom. Indeed, with \HardMax you lose all customers as soon as you fall behind in performance, with \HardMaxRandom you get some small market share no matter what, and with \SoftMaxRandom you are further guaranteed a market share close to $\tfrac12$ as long as your performance is not much worse than the competition. The uniform choice among principals corresponds to no rationality and no competition. In the second variant of the model, this comes from the number of firms and the timing of entry in the market. Competition in this case ranges from monopoly to ``incumbent" (first-mover in duopoly) to simultaneous entry duopoly to ``late entrant" (last mover in duopoly). The same distinctions in both cases also control the severity of competition between the principals.

In both variants we identify an inverted-U relationship between competition and innovation in the spirit of \reffig{fig:inverted-U}. These inverted-U relationships arise for a fundamentally different reason, compared to the existing literature on ``competition vs. innovation.'' In the literature, better technology always helps in a competitive environment, other things being equal. Thus, the trade-off is between the costs of improving the technology and the benefits that the improved technology provides in the competition. Meanwhile, we find that a better exploration algorithm may sometimes perform much worse under competition, even in the absence of R\&D costs. This stems from the nature of exploration technologies in online markets which rely on learning from interactions with users. This leads to an implicit cost from exploration in the form of a reduced rate of users that a firm attracts and can learn from. However, interestingly, the economic mechanism in our model for incentivizing firms to engage in R\&D has a qualitative similarity to the role that patents play in incentivizing innovation in textbook R\&D models. In these models, patents temporarily relax competition for the innovating firm by giving them exclusive access to their innovation for a limited period of time in order to incentivize them to invest in the better technology. In our model, temporarily relaxing competition in the form of giving firms free periods to learn incentivizes the firm to invest in the better technology.

\xhdr{Discussion}
The first variant of our model is tractable enough to allow us to obtain theoretical results with ``asymptotic" flavor. However, for the sake of analytical tractability, we make the unrealistic simplification that users do not observe any signals about firms' ongoing performance.

The second variant of our model relaxes this simplification and accounts for competition in a more direct way, but becomes considerably harder to analyze analytically. This is for several reasons: intricate feedback loop from performance to reputations to users to performance;
%
mean reputation, most connected to our intuition, is sometimes a bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited});
%
mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice. We therefore analyze our model using numerical simulation, which has several benefits. It allows us to analyze our model from a ``non-asymptotic" perspective, looking for substantial effects within relevant time scales. Indeed, we start our investigation by determining what time scales are relevant in the context of this variant of the model. Further, it allows us to investigate important economic mechanisms that arise in environments where exploration and competition tensions are at play, such as the effect of incumbency and the extent to which data can serve as a barrier to entry.

\xhdr{Map of the paper.}
We survey related work (Section~\ref{sec:related-work}), lay out the model and preliminaries (Section~\ref{sec:model}), and proceed to analyze the different models. In Sections ~\ref{sec:rational},~\ref{sec:random}, and ~\ref{sec:soft} we analyze the first main variant of the model and characterize the equilibrium behavior under three different agent decision rules. Then, we turn to analyze the second variant of the model. Section [prelim] discusses the details of the numerical simulations. Sections ~\ref{sec:isolation} and ~\ref{sec:revisited} overview results from running different bandit algorithms in isolation (i.e. without competition). Sections ~\ref{sec:competition} and ~\ref{sec:barriers} overview the results of the second variant of the model. Appendix~\ref{app:examples} provides some pertinent background on multi-armed bandits. Appendix~\ref{app:perturb} gives a broad example to support an assumption in our model.

}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2019-08-18 journal paper/content/sec-model.tex
first pass through intro and model section

--- 2019-08-14 journal paper/itcs18paper/sec-model.tex
@@
 
+\gaedit{
 The information structure is as follows.
 The prior $\prior$ is known to everyone.
-The mean rewards $\mu_a$ are not revealed to anybody.
-Each agent knows both principals' algorithms, and the global round when (s)he arrives, \emph{but not} the rewards of the previous agents.
-Each principal is completely unaware of the rounds when the other is chosen.
+The mean rewards $\mu_a$ are not revealed to anybody and each principal is completely unaware of the rounds when the other is chosen.
+We consider two variants of our model with different information structures.
+In the first, each agent knows both principals' algorithms, and the global round when (s)he arrives, \emph{but not} the rewards of the previous agents.
+In the second, agents only make decisions based on th rewards of previous agents.
+Concretely, each of the two principals has a \emph{reputation score}, and each agent's choice is driven by these two numbers.
+The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
 
+}
 \xhdr{Some terminology.} The two principals are called ``Principal
@@
 \xhdr{Agents' response.}
-Each agent $t$ chooses principal $i_t$ as as follows: it chooses a distribution over the principals, and then draws independently from this distribution.
+
+Each agent $t$ chooses principal $i_t$ as follows: it chooses a distribution over the principals, and then draws independently from this distribution.
 Let $p_t$ be the probability of choosing principal $1$ according to this distribution.
 Below we specify $p_t$; we need to be careful so as to avoid a circular definition.
+
+\gaedit{
+
+The form of $p_t$ depends on the information structure that we consider.
+In the first, where the agents only know the global round and the principals' algorithms, it is defined as follows.
+
+}
 
@@
 
+\gaedit{
+In the second information structure we consider, at a given time $t$ each principal $i \in \{1, 2 \}$ has a reputation score denoted as $\REP_i (t)$.
+In this case we have that the agent's responses take an analogous form as in the first information structure:
+\begin{align}
+p_t = \respF\left(\; \REP_1(t) - \REP_2(t) \;\right).
+\end{align}
+}
 \xhdr{Response functions.}
-We use the response function $\respF$ to characterize the amount of rationality and competitiveness in our model.
+We use the response function $\respF$ \gaedit{to characterize the decision rule of the agents in our model}.
 We assume that $\respF$ is monotonically non-decreasing, is larger than $\tfrac12$ on the interval $(0,1]$, and smaller than $\tfrac12$ on the interval $[-1,0)$.
-Beyond that, we consider three specific models, listed in the order of decreasing rationality and competitiveness (see \reffig{fig:response-functions}):
+Beyond that, we consider three specific models (see \reffig{fig:response-functions}):
 
@@
 
-We will emphasize several specific algorithms or classes thereof:
+We will emphasize several specific algorithms or classes thereof. In the second half of the paper we will :
 
 \begin{itemize}
-\item ``smart" MAB algorithms that combine exploration and exploitation, such as UCB1 \cite{bandits-ucb1} and Successive Elimination \cite{EvenDar-icml06}. These algorithms achieve
+\item ``smart" MAB algorithms that combine exploration and exploitation, such as UCB1 \cite{bandits-ucb1} and \gaedit{Thompson Sampling [citation for TS]}. These algorithms achieve
         $\BIR(n) \leq  \tilde{O}(n^{-1/2})$
@@
 
-We focus on MAB algorithms such that $\BIR(n)$ is non-increasing; we call such algorithms \emph{monotone}.
+When we consider the analytical model, we focus on MAB algorithms such that $\BIR(n)$ is non-increasing; we call such algorithms \emph{monotone}.
 While some reasonable MAB algorithms may occasionally violate monotonicity, they can usually be easily modified so that monotonicity violations either vanish altogether, or only occur at very specific rounds (so that agents are extremely unlikely to exploit them in practice).
@@
 More background and examples can be found in Appendix~\ref{app:examples}. In particular, we prove that \DynGreedy is monotone.
+
+\gaedit{When we move to the variant of the model where we consider numerical simulations, we will pick a representative algorithm from each class of algorithms discussed here as our main interest is in qualitative differences between the different classes.
+We will utilize Thompson Sampling (\TS) from the ``smart" MAB algorithms, $\eps$-greedy (\DEG)\footnote{Throughout, we fix $\eps = 0.05$.
+Our pilot experiments showed that different $\eps$ did not qualitatively change the results.} from the ``naive" algorithms, and \DynGreedy (\DG).
+For ease of comparison, all three algorithms are parameterized with the same fake prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution.
+Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.
+}
 
@@
 for posterior mean reward of action $a$ given event $E$.
+
+\gaedit{
+\xhdr{Varying Competition}
+
+In both variants of the model that we consider, we have several levers for varying the nature of competition. In the strictest version of the model,
+
+In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``late entrant"), and therefore enjoys a \emph{temporary monopoly}.
+Formally, the incumbent enjoys additional $X$ rounds of the ``warm start".
+We treat $X$ as an exogenous element of the model, and study the consequences for a fixed $X$.
+
+}
 

### 2019-08-18 journal paper/journal_main.tex
first pass through intro and model section

@@
 \newcommand{\rew}{\term{rew}}  % Bayesian-expected reward after n local rounds
+\newcommand{\REP}{\term{REP}} % posterior mean reward
 \newcommand{\PMR}{\term{PMR}} % posterior mean reward
@@
 \begin{document}
-\title[Competing Bandits]
-{Competing Bandits}
+\title[Competing Bandits: \\ The Perils of Exploration under Competition]
+{Competing Bandits: \\ The Perils of Exploration under Competition}
 
@@
 \begin{abstract}
-  We empirically study the interplay between \textit{exploration} and
-  \textit{competition}. Systems that learn from interactions with
-  users often engage in \emph{exploration}: making potentially
-  suboptimal decisions in order to acquire new information for future
-  decisions. However, when multiple systems are competing for
-    the same market of users, exploration may hurt a system's
-    reputation in the near term, with adverse competitive effects. In particular, a system may enter a ``death spiral", when the short-term reputation cost decreases
-    the number of users for the system to learn from, which degrades the
-    system's performance relative to competition and further decreases
-    the market share.
-
-We ask whether better exploration algorithms are incentivized under competition.
-We run extensive numerical experiments in a stylized duopoly model in which two firms deploy multi-armed bandit algorithms and compete for myopic users.
-We find that duopoly and monopoly tend to favor a primitive ``greedy algorithm" that does not explore and leads to low consumer welfare, whereas a temporary monopoly (a duopoly with an early entrant) may incentivize better bandit algorithms and lead to higher consumer welfare.
-Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.
+ Abstract goes here
 \end{abstract}
@@
 \label{sec:intro}
-\subfile{itcs18paper/sec-intro}
+\subfile{content/sec-intro}
 
@@
 \label{sec:model}
-\subfile{itcs18paper/sec-model}
+\subfile{content/sec-model}
 
@@
 
-\section{Economic implications}
-\label{sec:welfare}
-\subfile{itcs18paper/sec-welfare}
-
-\subfile{ec19paper/content/introduction}
-
-\subfile{ec19paper/content/model}
+%\section{Economic implications}
+%\label{sec:welfare}
+%\subfile{itcs18paper/sec-welfare}
 
@@
 
-\subfile{ec19paper/content/non_greedy_choice}
-
-\subfile{ec19paper/content/conclusion}
+%\subfile{ec19paper/content/non_greedy_choice}
 

### 2019-08-18 analysis_code/eeog.Rmd
finish first pass at journal version

@@
 }
-
-WORKING_PATH <- "/Users/garidor/Desktop/bandits-rl-project/results/tournament_raw_results/"
-f <- "tournament_experiment_test_raw.csv"
+WORKING_PATH <- "/Volumes/My Passport/bandits/tournament_raw_results/"
+#WORKING_PATH <- "/Volumes/My Passport/final_bandit_results/results/tournament_raw_results/"
+#f <- "tournament_experiment_full_sim_with_realizations_raw.csv"
+f <- "tournament_experiment_full_welfare_calc_raw.csv"
 dat <- read.csv(file=paste(WORKING_PATH, f, sep=""))
@@
 dat <- filter(dat, Time.Horizon == 2000)
-#priors <- as.list(unique(dat['Prior']))$Prior
-priors <- c("Heavy Tail")
+priors <- as.list(unique(dat['Prior']))$Prior
+#priors <- c("Heavy Tail")
 #warm_starts <- as.list(unique(dat['Warm.Start']))$Warm.Start

### 2019-08-18 journal paper/content/conclusion.tex
finish first pass at journal version

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Conclusion}\label{sec:conclusion}

We considered a stylized duopoly setting where firms simultaneously learn from and compete for users. We analyzed two separate but connected variants of this model, one where consumers chose between firms based on their Bayesian expected reward and another where consumers chose between them based on their past performance. We showed in both cases that competition may not always induce firms to commit to better exploration algorithms, resulting in welfare losses for consumers. The primary mechanism in both was that consumers need to be incentivized to select a firm over its competitors, leading firms that engage in exploration to be starved of consumers before they make enough progress on their learning problem. We found that in order to incentivize ``better" exploration strategies the key intuition is that a firm needs to have some ``free" consumers that visit them without having to be incentivized to do so. We explored several economic mechanisms that can generate this. The first was by considering different variants of random utility models where stochastic choice gives firms enough free consumers to incentivize them to commit to ``better" algorithms and we explored how this varied based on the nature of the random utility model. The second was by considering giving one firm a first-mover advantage where all consumers in the market would go to this firm without having to be incentivized to do so. While this first-mover advantage lead to the incumbent firm being incentivized to commit to a ``better" algorithm it lead to a substantial market share going to this firm. We further analyzed to what extent this advantage came from having a more defined reputation or more data from the incumbency period. We found that even a small amount of ``data advantage" leads to substantial long-term market power due to the dynamics of competition.

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../competing_bandits"
%%% End:

### 2019-08-18 journal paper/content/related_work.tex
finish first pass at journal version

--- 2019-06-17 journal paper/itcs18 paper/sec-related-work.tex
@@
 This literature is not concerned with \innovation, and typically models network effects exogenously, whereas in our model network effects are endogenous: they are created by MAB algorithms, an essential part of the model.
+\gaedit{The ``death spiral" effect that we find is similar to the increasing dominance dominance notion found in the industrial organization literature (see e.g.
+\cite{cabral1994learning, athey2001investment}).
+Further, there is also a line of work that investigates how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers \cite{schmalensee1982product, bagwell1990informational}.
+In the reputation choice model, we observe a similar effect when we investigate the role that reputation can serve as a barrier to entry.
+However, in our model this effect is further strengthened by the fact that the firms have to learn while competing adding that the incumbent may not only have a reputational advantage but additionally the information it acquires serves as a further barrier to entry.
+The idea that data can serve as a barrier to entry, especially in online markets, has been studied theoretically in \cite{ lambrecht2015can} and empirically in
+\cite{bajari2018impact}. While these papers find that small amounts of additional data do not provide significant improvement, they focus on learning in isolation. }
 

### 2019-08-18 journal paper/content/sec-intro.tex
finish first pass at journal version

@@
-Learning from interactions with users is ubiquitous in modern customer-facing systems, from product recommendations to web search to spam detection to content selection to fine-tuning the interface.
+Learning from interactions with users is ubiquitous in modern customer-facing systems, from product recommendations to web search to spam detection to content selection to fine-tuning user interfaces.
 Many systems purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
@@
 
-We investigate these questions for two separate but connected models.
-In the first we suppose that consumers are well-informed about the firm's exploration strategies but do not observe any information about past user experiences with the firms.
-In the second, we suppose that consumers \textit{only} consider the average of past user experiences with each of the firms.
-In the first model we analytically characterize the equilibrium exploration strategies of two firms competing for the same stream of consumers as we vary the degree of randomness in consumer decisions.\footnote{The randomness in consumer decisions can be interpreted both as the degree of rationality as well as a reduced-form way of capturing the degree to which consumers themselves engage in exploration} In the second model we account for competition in a more direct way and utilize numerical simulations to consider asymmetries in the timing of entry, increasing competition beyond two firms, and allow us to isolate the effect that role that data can play as a barrier to entry.
+We investigate these questions for two separate but connected stylized duopoly models where two firms commit to exploration strategies and compete for a stream of consumers.
+In the first, which we denote as the \textit{expectation choice model}, we suppose that consumers are well-informed about the firm's exploration strategies but do not observe any information about past user experiences with the firms.
+In the second, which we denote as the \textit{reputation choice model}, we suppose that consumers \textit{only} consider the average of past user experiences with each of the firms.
+In the expectation choice model we analytically characterize the equilibrium exploration strategies as we vary the degree of randomness in consumer decisions.
+In the reputation choice model we account for competition in a more direct way and utilize numerical simulations to consider asymmetries in the timing of entry, increasing competition beyond two firms, and allow us to isolate the effect that role that informational advantages can play as a barrier to entry in markets where exploration and competition tensions arise.
 
@@
 
-\caption{Inverted-U relationship between rationality/competitiveness and algorithms.}
+\caption{Inverted-U relationship between competitiveness and algorithms.}
 \label{fig:inverted-U}
@@
 \begin{enumerate}
-\item In the first, agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms
-\item In the second, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal
+\item In the expectation choice model, agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms
+\item In the reputation choice model, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal
 \end{enumerate}
@@
 In both variants of our model, we find that in a simultaneous entry duopoly a greedy algorithm that does no purposeful exploration is incentivized in equilibrium.
-In the first variant of the model, once a firm does any exploration it is immediately starved of consumers when its opponent plays a greedy algorithm since we show that the expected reward for every subsequent consumer is higher for the firm that plays the greedy algorithm.
-In the second variant of the model, the same mechanism drives the greedy algorithm to be incentivized in equilibrium.
+In the expectation choice model, once a firm does any exploration it is immediately starved of consumers when its opponent plays a greedy algorithm since we show that the expected reward for every subsequent consumer is higher for the firm that plays the greedy algorithm.
+In the reputation choice model, the same mechanism drives the greedy algorithm to be incentivized in equilibrium.
 In this variant, exploration hurts a firm's reputation and decrease its market share in the near term.
-This leaves the firm with less users to learn from, which may further degrade the firm's performance relative to competitors who keep learning and improving from \emph{their} customers, and so forth.
+This leaves the firm with less users to learn from, which further degrades the firm's performance relative to competitors who keep learning and improving from \emph{their} customers, and so forth.
 Taken to the extreme, such dynamics lead to a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
@@
 The primary mechanism that generates this stark result is the same in both variants of the model: consumers need to be incentivized to select a firm over its competitors, leading firms that engage in exploration to be starved of consumers before they make enough progress on their learning problem.
-In order to incentivize ``better" exploration strategies the key intuition is that the firm needs to have some ``free" consumers to visit them without having to incentivize them.
-This allows the firm to eventually get out of the hole of exploration.
+In order to incentivize ``better" exploration strategies the key intuition is that the firm needs to have some ``free" consumers that visit them without the firm having to incentivize them to do so.
+This allows the firm to eventually overcome the initial losses in consumer perception from exploration.
 
-In the first variant of the model, we relax the decision rule of the consumers to allow for each firm to be chosen with some fixed baseline probability so that firms get a constant stream of ``free" consumers.
-This can be interpreted as a reduced-form relaxation of the assumption that consumers are myopic and themselves may be engaging in exploration or, alternatively, as deviations from rational behavior.
+In the expectation choice model, we relax the decision rule of the consumers to allow for each firm to be chosen with some fixed baseline probability so that firms get a constant stream of ``free" consumers; we call this choice rule \HardMaxRandom.
 We find that, in this setting, better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
@@
 
-In the second variant of the model, we keep the decision rule of the consumers but instead vary the timing of entry and allow one firm to have a first-mover advantage so that this firm (denoted as the incumbent) gets all the agents in the market before the other firm enters.
+In the reputation choice model, we maintain the same decision rule for the consumers but instead vary the timing of entry and allow one firm to have a first-mover advantage so that this firm (denoted as the incumbent) gets all the agents in the market before the other firm enters.
 We find that, if the first-mover period is sufficiently large, then the incumbent has a dominant strategy to play ``smart" bandit algorithms and that consumer welfare is substantially higher than in the case of simultaneous entry.
@@
 
-\xhdr{Additional Findings} In the first variant of the model, we further relax the decision rule of the agents so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' expected rewards; we call it \SoftMaxRandom.
+\xhdr{Additional Findings} In the expectation choice model, we further relax the decision rule of the agents so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' expected rewards; we call it \SoftMaxRandom.
 For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
@@
 
-In the second variant of the model, we investigate the ``first-mover advantage" phenomenon in more detail.
+In the reputation choice model, we investigate the ``first-mover advantage" phenomenon in more detail.
 Being first in the market gives free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
@@
 In our model, competition varies based on the number of ``free" consumers that a firm receives.
-In the first variant of the model, this comes from introducing ``random" consumers where we vary from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
+In the expectation choice model, this comes from introducing ``random" consumers where we vary from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
 Indeed, with \HardMax you lose all customers as soon as you fall behind in performance, with \HardMaxRandom you get some small market share no matter what, and with \SoftMaxRandom you are further guaranteed a market share close to $\tfrac12$ as long as your performance is not much worse than the competition.
 The uniform choice among principals corresponds to no rationality and no competition.
-In the second variant of the model, this comes from the number of firms and the timing of entry in the market.
+In the reputation choice model, this comes from the number of firms and the timing of entry in the market.
 Competition in this case ranges from monopoly to ``incumbent" (first-mover in duopoly) to simultaneous entry duopoly to ``late entrant" (last mover in duopoly).
@@
 This leads to an implicit cost from exploration in the form of a reduced rate of users that a firm attracts and can learn from.
-However, interestingly, the economic mechanism in our model for incentivizing firms to engage in R\&D has a qualitative similarity to the role that patents play in incentivizing innovation in textbook R\&D models.
+However, interestingly, the economic mechanism in our model for incentivizing firms to engage in R\&D has a qualitative similarity to the role that patents play in incentivizing innovation in standard R\&D models.
 In these models, patents temporarily relax competition for the innovating firm by giving them exclusive access to their innovation for a limited period of time in order to incentivize them to invest in the better technology.
@@
 \xhdr{Discussion}
-The first variant of our model is tractable enough to allow us to obtain theoretical results with ``asymptotic" flavor.
-However, for the sake of analytical tractability, we make the unrealistic simplification that users do not observe any signals about firms' ongoing performance.
+We consider two separate variants of the model in order to provide a more thorough investigation of the tension between exploration and competition.
+The expectation choice model is tractable enough to allow us to obtain theoretical results with ``asymptotic" flavor.
+However, for the sake of analytical tractability, we make the unrealistic simplification that users do not observe any signals about firms' ongoing performance and it is difficult to analyze important economic mechanisms such as asymmetries in the timing of entry.
 
-The second variant of our model relaxes this simplification and accounts for competition in a more direct way, but becomes considerably harder to analyze analytically.
+The reputation choice model relaxes this simplification and accounts for competition in a more direct way as well as allows us to explore other relevant economic mechanisms in understanding the tension between exploration and competition.
+However, it now becomes considerably harder to analyze analytically.
 This is for several reasons: intricate feedback loop from performance to reputations to users to performance;
@@
 Indeed, we start our investigation by determining what time scales are relevant in the context of this variant of the model.
-Further, it allows us to investigate important economic mechanisms that arise in environments where exploration and competition tensions are at play, such as the effect of incumbency and the extent to which data can serve as a barrier to entry.
+Further, it allows us to investigate important economic mechanisms that arise in environments where exploration and competition tensions are at play, such as the effect of incumbency, increasing the number of firms, and the extent to which data can serve as a barrier to entry.
 
@@
 We survey related work (Section~\ref{sec:related-work}), lay out the model and preliminaries (Section~\ref{sec:model}), and proceed to analyze the different models.
+We start by analyzing the expectation choice model analytically.
 In Sections ~\ref{sec:rational},~\ref{sec:random}, and ~\ref{sec:soft} we analyze the first main variant of the model and characterize the equilibrium behavior under three different agent decision rules.
-Then, we turn to analyze the second variant of the model.
-Section [prelim] discusses the details of the numerical simulations.
+Then, we turn to analyze the reputation choice variant of the model with details of the analysis described in Section ~\ref{sec:sim_details}.
 Sections ~\ref{sec:isolation} and ~\ref{sec:revisited} overview results from running different bandit algorithms in isolation (i.e.
 without competition).
-Sections ~\ref{sec:competition} and ~\ref{sec:barriers} overview the results of the second variant of the model.
-Appendix~\ref{app:examples} provides some pertinent background on multi-armed bandits.
-Appendix~\ref{app:perturb} gives a broad example to support an assumption in our model.
-
+Sections ~\ref{sec:competition} and ~\ref{sec:barriers} overview the results of the second variant of the mode.
+Section ~\ref{sec:conclusion} concludes.
 }

### 2019-08-18 journal paper/content/sec-model.tex
finish first pass at journal version

--- 2019-08-14 journal paper/itcs18paper/sec-model.tex
@@
 
+\gaedit{
 The information structure is as follows.
 The prior $\prior$ is known to everyone.
-The mean rewards $\mu_a$ are not revealed to anybody.
-Each agent knows both principals' algorithms, and the global round when (s)he arrives, \emph{but not} the rewards of the previous agents.
-Each principal is completely unaware of the rounds when the other is chosen.
+The mean rewards $\mu_a$ are not revealed to anybody and each principal is completely unaware of the rounds when the other is chosen.
+We consider two variants of our model with different information structures.
+In the first, the  \textit{expectation choice} model, each agent knows both principals' algorithms, and the global round when (s)he arrives, \emph{but not} the rewards of the previous agents.
+In the second, the \textit{reputation choice} model, agents only make decisions based on the rewards of previous agents.
+Concretely, each of the two principals has a \emph{reputation score}, and each agent's choice is driven by these two numbers.
+The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
 
+}
 \xhdr{Some terminology.} The two principals are called ``Principal
@@
 \xhdr{Agents' response.}
-Each agent $t$ chooses principal $i_t$ as as follows: it chooses a distribution over the principals, and then draws independently from this distribution.
+Each agent $t$ chooses principal $i_t$ as follows: it chooses a distribution over the principals, and then draws independently from this distribution.
 Let $p_t$ be the probability of choosing principal $1$ according to this distribution.
 Below we specify $p_t$; we need to be careful so as to avoid a circular definition.
+
+\gaedit{
+
+The form of $p_t$ depends on the information structure that we consider.
+In the expectation choice model, where the agents only know the global round and the principals' algorithms, it is defined as follows.
+
+}
 
@@
 
+\gaedit{
+In the reputation choice model we consider, at a given time $t$ each principal $i \in \{1, 2 \}$ has a reputation score denoted as $\REP_i (t)$.
+In this case we have that the agent's responses take an analogous form as in the first case:
+\begin{align}
+p_t = \respF\left(\; \REP_1(t) - \REP_2(t) \;\right).
+\end{align}
+}
 \xhdr{Response functions.}
-We use the response function $\respF$ to characterize the amount of rationality and competitiveness in our model.
+We use the response function $\respF$ \gaedit{to characterize the decision rule of the agents in our model}.
 We assume that $\respF$ is monotonically non-decreasing, is larger than $\tfrac12$ on the interval $(0,1]$, and smaller than $\tfrac12$ on the interval $[-1,0)$.
-Beyond that, we consider three specific models, listed in the order of decreasing rationality and competitiveness (see \reffig{fig:response-functions}):
+Beyond that, we consider three specific models (see \reffig{fig:response-functions}):
 
@@
 
-We will emphasize several specific algorithms or classes thereof:
+We will emphasize several specific algorithms or classes thereof. In the second half of the paper we will :
 
 \begin{itemize}
-\item ``smart" MAB algorithms that combine exploration and exploitation, such as UCB1 \cite{bandits-ucb1} and Successive Elimination \cite{EvenDar-icml06}. These algorithms achieve
+\item ``smart" MAB algorithms that combine exploration and exploitation, such as UCB1 \cite{bandits-ucb1} and \gaedit{Thompson Sampling \cite{russo2018tutorial}}. These algorithms achieve
         $\BIR(n) \leq  \tilde{O}(n^{-1/2})$

### 2019-08-18 journal paper/content/sim_details.tex
finish first pass at journal version

--- 2019-04-29 paper/content/model.tex
@@
 
-\section{Model and Preliminaries}\label{sec:model}
+\section{Reputation Model Details}\label{sec:sim_details}
 
-%\textbf{Overview}
-We consider a game involving two firms and $T$ customers (henceforth, \emph{agents}).
-The game lasts for $T$ rounds.
-In each round, a new agent arrives, chooses among the two firms, interacts with the chosen firm, and leaves forever.
+In this section we move to the reputation choice model.
+Recall that in this variant of the model, instead of making choices between firms based on the Bayesian expected reward, each agent chooses the firm with a maximal reputation score (breaking ties uniformly).
+The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
+We focus on numerical investigation of this model instead of analytical characterizations of equilibrium strategies.
 
-Each interaction between a firm and an agent proceeds as follows.
-There is a set $A$ of $K$ actions, henceforth \emph{arms}, same for both firms and all rounds.
-The firm chooses an arm, and the agent experiences a numerical reward observed by the firm.
-Each arm corresponds to a different version of the experience that a firm can provide for an agent, and the reward corresponds to the agent's satisfaction level.
-The other firm does not observe anything about this interaction, not even the fact that this interaction has happened.
-
-From each firm's perspective, the interactions with agents follow the protocol of the multi-armed bandit problem (MAB).
+\xhdr{Key Model Differences}
+The timing of the model is the same as before where each firm commits to a MAB algorithm before the game starts and uses this algorithm to choose its actions.
 We focus on i.i.d.
@@
 The mean rewards $\mu(a)$ are the same for all rounds and both firms, but initially unknown.
+However, instead of starting from some prior, we suppose that each firm has a uniform, ``fake", prior and that the initial information set of the firm is given by a ``warm start".
+Each algorithm receives a ``warm start": additional $T_0$ agents that arrive before the game starts, and interact with the firm as described above.
+The warm start ensures that each firm has a meaningful reputation and initial prior when competition starts..
 
-Before the game starts, each firm commits to an MAB algorithm, and uses this algorithm to choose its actions.
-Each algorithm receives a ``warm start": additional $T_0$ agents that arrive before the game starts, and interact with the firm as described above.
-The warm start ensures that each firm has a meaningful reputation when competition starts.
-Each firm's objective is to maximize its market share: the fraction of users who chose this firm.
-
-In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``late entrant"), and therefore enjoys a \emph{temporary monopoly}.
+In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``late entrant"), and therefore enjoys a \emph{first-mover advantage}.
 Formally, the incumbent enjoys additional $X$ rounds of the ``warm start".
@@
 
-\xhdr{Agents.}
-Firms compete on a single dimension, quality of service, as expressed by agents' rewards. 
-\OMIT{\footnote{One can think that there are two quality types - a reward of 1 is a high type and a reward of 0 is a low type}}
-Agents are myopic and non-strategic: they would like to choose among the firms so as to maximize their expected reward (i.e.
-select the firm with the highest quality), without attempting to influence the firms' learning algorithms or rewards of the future users.
-Agents are not well-informed: they only receive a rough signal about each firm's performance before they choose a firm, and no other information.
-
-Concretely, each of the two firms has a \emph{reputation score}, and each agent's choice is driven by these two numbers.
-We posit a version of rational behavior: each agent chooses a firm with a maximal reputation score (breaking ties uniformly).
-The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
-
-%\asedit{This is similar to ratings aggregators on many online platforms such as Yelp, Tripadvisor, etc.}
-
-\xhdr{MAB algorithms.} We consider three classes of algorithms, ranging from more primitive to more sophisticated:
-
-\begin{enumerate}
-\item \emph{Greedy algorithms} that strive to take actions with maximal mean reward, based on the current information.
-
-\item \emph{Exploration-separating algorithms} that separate exploration and exploitation.
-The ``exploitation" choices strives to maximize mean reward in the next round, and the ``exploration" choices do not use the rewards observed so far.
-\item \emph{Adaptive exploration}: algorithms that combine exploration and exploitation, and sway the exploration choices towards more promising alternatives.
-\end{enumerate}
-
-We are mainly interested in qualitative differences between the three classes.
-For concreteness, we fix one algorithm from each class.
-Our pilot experiments indicate that our findings do not change substantially if other algorithms are chosen.
-For technical reasons, we consider Bayesian versions initialized with a ``fake" prior (\ie not based on actual knowledge).
-We consider:
-
-\begin{enumerate}
-\item a greedy algorithm that chooses an arm with largest posterior mean reward. We call it "Dynamic Greedy" (because the chosen arm may change over time), \DG in short.
-
-\item an exploration-separated algorithm that in each round, \emph{explores} with probability $\eps$: chooses an arm independently and uniformly at random, and with the remaining probability \emph{exploits} according to \DG.
-We call it ``dynamic epsilon-greedy", \DEG in short.\gaedit{\footnote{Throughout, we fix $\eps = 0.05$.
-Our pilot experiments showed that different $\eps$ did not qualitatively change the results.}}
-
-\item an adaptive-exploration algorithm called ``Thompson Sampling" (\TS).
-In each round, this algorithm updates the posterior distribution for the mean reward of each arm $a$, draws an independent sample $s_a$ from this distribution, and chooses an arm with the largest $s_a$.
-\end{enumerate}
-
+\xhdr{MAB algorithms.} We consider the same three classes of bandit algorithms discussed in Section ~\ref{sec:model}, but take a representative algorithm from each class and look for qualitative differences between the different classes.
+We will utilize Thompson Sampling (\TS) from the ``smart" MAB algorithms, $\eps$-greedy (\DEG)\footnote{Throughout, we fix $\eps = 0.05$.
+Our pilot experiments showed that different $\eps$ did not qualitatively change the results.} from the ``naive" algorithms, and \DynGreedy (\DG).
 For ease of comparison, all three algorithms are parameterized with the same fake prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution.
 Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.
-
-Both \DEG and \TS are classic and well-understood MAB algorithms, see \cite{Bubeck-survey12,TS-survey-FTML18} for background.
-It is well-known that \TS is near-optimal in terms of the cumulative rewards, and \DEG is very suboptimal, but still much better than \DG.%
-\footnote{Formally, \TS achieves regret
-    $\tilde{O}(\sqrt{TK})$ and
-    $O(\tfrac{1}{\Delta} \log T)$,
-where $\Delta$ is the gap in mean rewards between the best and second-best arms.
-\DEG has regret $\tilde{\Theta}(T^{2/3} K^{1/3})$ in the worst case.
-And \DG can have regret as high as $\Omega(T)$.
-Deeper discussion of these distinctions is not very relevant to this paper.}
-In a stylized formula:
-    $ \TS \gg \DEG \gg \DG $
-as stand-alone MAB algorithms.
 
@@
 
-\xhdr{Terminology.}
-Following a standard game-theoretic terminology, algorithm Alg1 \emph{(weakly) dominates} algorithm Alg2 for a given firm if Alg1 provides a larger (or equal) market share than Alg2 at the end of the game.
-An algorithm is a (weakly) dominant strategy for the firm if it (weakly) dominates all other algorithms.
-This is for a particular MAB instance and a particular selection of the game parameters.
-
-\OMIT{\gaedit{Since the algorithms that involve exploration have larger deployment costs, we break any indifference towards algorithms that do not involve exploration (i.e.
-if \DG and \DEG provide the same market share, the firm will play \DG).}}
-
 \xhdr{Simulation details.}
@@
 
+\xhdr{Terminology.}
+Following a standard game-theoretic terminology, algorithm Alg1 \emph{(weakly) dominates} algorithm Alg2 for a given firm if Alg1 provides a larger (or equal) market share than Alg2 at the end of the game.
+An algorithm is a (weakly) dominant strategy for the firm if it (weakly) dominates all other algorithms.
+This is for a particular MAB instance and a particular selection of the game parameters.
+
+\xhdr{Discussion}
+While we experiment with various MAB instances and parameter settings, we only report on selected, representative experiments in the body of the paper.
+Additional plots and tables are provided in the appendix.
+Unless noted otherwise, our findings are based on and consistent with all these experiments.
+
+Even with a stylized model, numerical investigation is quite challenging.
+An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given instance of a multi-armed bandit problem.%
+\footnote{Each such experiment is run many times to reduce variance.}
+Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit instances\}, and we are looking for findings that are consistent across this entire space.
+It is essential to keep each of the three dimensions small yet representative.
+In particular, we need to capture a huge variety of bandit instances with only a few representative examples.
+Further, one needs succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).
+
 The simulations are computationally intensive.
@@
 Each experiment took about $10$ hours.
-
-\xhdr{Consistency.}
-While we experiment with various MAB instances and parameter settings, we only report on selected, representative experiments in the body of the paper.
-Additional plots and tables are provided in the appendix.
-Unless noted otherwise, our findings are based on and consistent with all these experiments.
 

### 2019-08-18 journal paper/journal_main.tex
finish first pass at journal version

@@
 % a very useful package for edits and comments, from David Kempe (USC)
-\usepackage{ec19paper/color-edits}
+\usepackage[suppress]{ec19paper/color-edits}
 %\usepackage[suppress]{color-edits}  % use this to suppress the package
@@
 \begin{abstract}
- Abstract goes here
+\gaedit{Most modern systems strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
+We initiate a study of the interplay between \emph{exploration and competition}---how such systems balance the exploration for learning and the competition for users.
+Here the users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing systems.
+
+In our model, we consider competition between two multi-armed bandit algorithms faced with the same bandit instance.
+Users arrive one by one and choose among the two algorithms, so that each algorithm makes progress if and only if it is chosen.
+We ask whether and to what extent competition incentivizes the adoption of better bandit algorithms.
+We investigate this issue analytically for several models of user response, as we vary the degree of competitiveness in the model.
+We further investigate this issue using numerical simulation to understand the effect of the first-mover advantage in our model.
+Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets and are closely related to the ``competition vs.
+innovation" relationship, a well-studied theme in economics.}
 \end{abstract}
@@
 \label{sec:related-work}
-\subfile{itcs18paper/sec-related-work}
+\subfile{content/related_work}
 
@@
 
+\gaedit{
+\subfile{content/sim_details}
+}
+
 \subfile{ec19paper/content/perf_in_iso}
@@
 
+\gaedit{
+\subfile{content/conclusion}
+}
 %\subfile{ec19paper/content/non_greedy_choice}
@@
 
+\begin{appendices}
+
+\section{Background on multi-armed bandits}
+\label{app:examples}
+\subfile{itcs18paper/app-examples}
+
+\section{Non-degeneracy via a random perturbation}
+\label{app:perturb}
+\subfile{itcs18paper/app-perturb}
 \subfile{ec19paper/content/appendix_for_one_version}
 
+\end{appendices}
 \end{document}

### 2019-08-18 analysis_code/welfare.R

@@ -1,14 +1,15 @@
 library(tidyverse)
 library(ggplot2)
+library(gdata)
 
-simultaneous_welfare <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/tournament_raw_results/tournament_experiment_full_welfare_calc_raw.csv")
-free_obs_welfare <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/free_obs_raw_results/free_obs_experiment_welfare_calc_raw.csv")
+simultaneous_welfare <- read.csv("/Volumes/My Passport/bandits/tournament_raw_results/tournament_experiment_full_welfare_calc_raw.csv")
+free_obs_welfare <- read.csv("/Volumes/My Passport/bandits/free_obs_raw_results/free_obs_experiment_welfare_calc_raw.csv")
 #prelim <- read.csv("/Volumes/Mac/final_bandit_results/results/preliminary_raw_results/longer_ws.csv")
-prelim <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/preliminary_raw_results/regret_calc.csv")
+prelim <- read.csv("/Volumes/My Passport/bandits/preliminary_raw_results/regret_calc.csv")
 sim_eq_welfare <- filter(simultaneous_welfare, ((Prior == "Heavy Tail" | Prior == "Uniform") & (P1.Alg == "DynamicGreedy" & P2.Alg == "DynamicGreedy")) | (Prior == "Needle In Haystack" & P1.Alg == "ThompsonSampling" & P2.Alg == "ThompsonSampling"))
 free_obs_eq_welfare <- filter(free_obs_welfare, ((Prior == "Heavy Tail" | Prior == "Uniform") & (P1.Alg == "DynamicGreedy" & P2.Alg == "ThompsonSampling")))
-
 one_welfare <- combine(sim_eq_welfare, free_obs_eq_welfare)
+
 one_welfare <- one_welfare %>% mutate(label=ifelse(source == "free_obs_eq_welfare", "Temporary Monopoly", "Duopoly"))
 
 prelim <- prelim %>% mutate(Prior = Distribution, Warm.Start = 20, label = "Monopoly", Total.Regret = Cumulative.Regret)
@@ -17,15 +18,15 @@ prelim <- prelim %>% mutate(Time.Horizon = t - 20)
[…]

### 2019-10-10 journal paper/journal_main.tex
tweaks
Added Yishay to authors.

@@
 % a very useful package for edits and comments, from David Kempe (USC)
-\usepackage[suppress]{ec19paper/color-edits}
+\usepackage{ec19paper/color-edits}
 %\usepackage[suppress]{color-edits}  % use this to suppress the package
@@
  \affiliation{ \institution{Columbia University}}
+ \author{Yishay Mansour}
+ \affiliation{\institution{Tel Aviv University and Google}}
  \author{Aleksandrs Slivkins}
- \affiliation{\institution{Microsoft Research, New York City}}
+ \affiliation{\institution{Microsoft Research New York City}}
  \author{Zhiwei Steven Wu}
@@
  \authorsaddresses{%
-  Authors' addresses: Guy Aridor, Columbia University, New York, NY, g.aridor@columbia.edu; Aleksandrs Slivkins, Microsoft Research, New York, NY, slivkins@microsoft.com; Zhiwei Steven Wu, University of Minnesota - Twin Cities, Minneapolis, MN, zsw@umn.edu}
+  Authors' addresses: 
+  Guy Aridor, Columbia University, New York, NY, USA. g.aridor@columbia.edu; 
+  Yishay Mansour, Tel Aviv University, Tel Aviv, Israel. mansour.yishay@gmail.com;
+  Aleksandrs Slivkins, Microsoft Research, New York, NY, USA. slivkins@microsoft.com; 
+  Zhiwei Steven Wu, University of Minnesota - Twin Cities, Minneapolis, MN, USA. zsw@umn.edu}
 

### 2019-10-22 journal paper/content/related_work.tex
updates

@@
 This literature is not concerned with \innovation, and typically models network effects exogenously, whereas in our model network effects are endogenous: they are created by MAB algorithms, an essential part of the model.
-\gaedit{The ``death spiral" effect that we find is similar to the increasing dominance dominance notion found in the industrial organization literature (see e.g.
+\gaedit{The ``death spiral" effect that we find is similar to the increasing dominance notion found in the industrial organization literature (see e.g.
 \cite{cabral1994learning, athey2001investment}).

### 2019-10-22 journal paper/content/sec-intro.tex
updates

@@
-Learning from interactions with users is ubiquitous in modern customer-facing systems, from product recommendations to web search to spam detection to content selection to fine-tuning user interfaces.
-Many systems purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-Randomized controlled trials, a.k.a.
-A/B testing, are an industry standard, with a number of companies such as \emph{Optimizely} offering tools and platforms to facilitate them.
-Many companies use more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known theoretical framework for exploration and making decisions under uncertainty.
+Learning from interactions with consumers is ubiquitous in modern customer-facing systems, from product recommendations to web search to content selection to fine-tuning user interfaces.
+These interactions are essential for the development of high-quality products, but, in order to learn what products are high quality, companies need to implement systems that engage in \emph{exploration}: deploying potentially suboptimal products in order to acquire new information to make more informed product decisions tomorrow.
 
-%~\cite{KohaviAB-2015,KohaviLSH09}
+In this paper, we initiate a study of the interplay between \exploration and \competition.
+While exploration may be beneficial for both firms and consumers in the long-run, it may lead to bad experiences for consumers today and result in subsequent consumers going to the firm's competitors tomorrow.
+In particular, a firm can only make progress on its learning problem if it is able to incentivize consumers to visit it over its competitors and excessive exploration may lead to low reputation which leads consumers to increasingly visit the firm's competitors.
+Consumers therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing firms.
 
-Systems that engage in exploration typically need to compete against one another; most importantly, they compete for users.
-This creates an interesting tension between \exploration and \competition.
-In a nutshell, while exploring may be essential for improving the service tomorrow, it may degrade quality and make users leave \emph{today}, in which case there will be no users to learn from!
-Thus, users play three distinct roles: they are customers that generate revenue, they generate data for the systems to learn from, and they are self-interested agents which choose among the competing systems.
+Given the prevalence of these technologies and the increased importance of data in the digital economy, our main high-level question is: {\bf does competition lead to the adoption of better exploration algorithms?}.
+This translates into a number of more concrete questions.
+While it is commonly assumed that ``better" technology always helps, is this so for our setting?
+Does increasing competition lead to higher consumer welfare?
+To what extent is there a ``data feedback loop" where one firm having more data leads to that firm attracting more users which leads to that firm having more data, etc.?
 
-We initiate a study of the interplay between \exploration and \competition.
-The main high-level question is: {\bf whether and to what extent competition incentivizes adoption of better exploration algorithms}.
-\gaedit{This translates into a number of more concrete questions.
-While it is commonly assumed that better learning technology always helps, is this so for our setting?
-Does increasing competition lead to higher consumer welfare?
-To what extent does one firm possessing more information or having a better perception amongst consumers than its competitors serve as a barrier to entry in this environment?
-
-We investigate these questions for two separate but connected stylized duopoly models where two firms commit to exploration strategies and compete for a stream of consumers.
-In the first, which we denote as the \textit{expectation choice model}, we suppose that consumers are well-informed about the firm's exploration strategies but do not observe any information about past user experiences with the firms.
-In the second, which we denote as the \textit{reputation choice model}, we suppose that consumers \textit{only} consider the average of past user experiences with each of the firms.
-In the expectation choice model we analytically characterize the equilibrium exploration strategies as we vary the degree of randomness in consumer decisions.
-In the reputation choice model we account for competition in a more direct way and utilize numerical simulations to consider asymmetries in the timing of entry, increasing competition beyond two firms, and allow us to isolate the effect that role that informational advantages can play as a barrier to entry in markets where exploration and competition tensions arise.
+We investigate these questions with a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
+We first look at a version of this model that allows us to make analytical characterizations of the optimal algorithms firms should deploy in equilibrium.
+In this variant of the model, which we denote as the \textit{expectation choice variant}, we suppose that consumers are well-informed about the firm's exploration strategies but do not observe any information about past user experiences with the firms.
+In this variant we analytically characterize the equilibrium exploration strategies as we vary the degree of randomness in consumer decisions.
+We then look at a second variant of the model that accounts for competition in a more direct way, which we denote as the \textit{reputation choice variant}, where we suppose that consumers \textit{only} consider the average of past user experiences with each of the firms.
+This variant allows us to utilize numerical simulations to consider a number of additional issues such as asymmetries in the timing of entry, increasing competition beyond two firms, and allowing us to isolate the role that informational advantages can play as a barrier to entry in markets where exploration and competition tensions arise.
 
 On a high level, our contributions can be framed in terms of the ``inverted-U relationship" between competition and the quality of adopted algorithms (see \reffig{fig:inverted-U}).
-
-}
 
@@
 Each principal only observes agents that chose him.
-\gaedit{The information set for the agent differs between the two variants of the model.
+The information set for the agent differs between the two variants of the model.
 \begin{enumerate}
-\item In the expectation choice model, agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms
-\item In the reputation choice model, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal
+\item In the expectation choice variant, agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms
+\item In the reputation choice variant, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal
 \end{enumerate}
-}
 \end{itemize}
 
-\gaedit{
 \xhdr{Main Findings.}
-In both variants of our model, we find that in a simultaneous entry duopoly a greedy algorithm that does no purposeful exploration is incentivized in equilibrium.
-In the expectation choice model, once a firm does any exploration it is immediately starved of consumers when its opponent plays a greedy algorithm since we show that the expected reward for every subsequent consumer is higher for the firm that plays the greedy algorithm.
-In the reputation choice model, the same mechanism drives the greedy algorithm to be incentivized in equilibrium.
+We find that in a simultaneous entry duopoly a greedy (myopic) algorithm that does no purposeful exploration is incentivized in equilibrium.
+In the expectation choice variant, once a firm does any exploration it is immediately starved of consumers when its opponent plays a greedy algorithm since we show that the expected reward for every subsequent consumer is higher for the firm that plays the greedy algorithm.
+In the reputation choice variant, the same mechanism drives the greedy algorithm to be incentivized in equilibrium.
 In this variant, exploration hurts a firm's reputation and decrease its market share in the near term.
@@
 Taken to the extreme, such dynamics lead to a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
+As a result, our model provides an example of the mechanisms that can lead to a ``data feedback loop" that has been hypothesized in policy discussions.
 In both cases, consumer welfare is low since the greedy algorithm is known to be dramatically bad in many important cases of multi-armed bandits.
 
-The primary mechanism that generates this stark result is the same in both variants of the model: consumers need to be incentivized to select a firm over its competitors, leading firms that engage in exploration to be starved of consumers before they make enough progress on their learning problem.
+The primary mechanism that generates this stark result is that consumers need to be incentivized to select a firm over its competitors, leading firms that engage in exploration to be starved of consumers before they make enough progress on their learning problem.
 In order to incentivize ``better" exploration strategies the key intuition is that the firm needs to have some ``free" consumers that visit them without the firm having to incentivize them to do so.
@@
 
-In the expectation choice model, we relax the decision rule of the consumers to allow for each firm to be chosen with some fixed baseline probability so that firms get a constant stream of ``free" consumers; we call this choice rule \HardMaxRandom.
+We relax the decision rule of the consumers to allow for each firm to be chosen with some fixed baseline probability so that firms get a constant stream of ``free" consumers; we call this choice rule \HardMaxRandom.
 We find that, in this setting, better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
@@
 
-In the reputation choice model, we maintain the same decision rule for the consumers but instead vary the timing of entry and allow one firm to have a first-mover advantage so that this firm (denoted as the incumbent) gets all the agents in the market before the other firm enters.
+We use the reputation choice variant to analyze the effect of varying the timing of entry between the firms and allow one firm to have a first-mover advantage so that this firm (denoted as the incumbent) gets all the agents in the market before the other firm enters.
 We find that, if the first-mover period is sufficiently large, then the incumbent has a dominant strategy to play ``smart" bandit algorithms and that consumer welfare is substantially higher than in the case of simultaneous entry.
@@
 
-\xhdr{Additional Findings} In the expectation choice model, we further relax the decision rule of the agents so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' expected rewards; we call it \SoftMaxRandom.
-For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
+\xhdr{\underline{Additional Findings}}
+
+\xhdr{Noise in consumer choice}: We further relax the decision rule of the agents so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' expected rewards; we call it \SoftMaxRandom.
+For this decision rule, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
 This is the most technical result of the paper.
@@
 
-In the reputation choice model, we investigate the ``first-mover advantage" phenomenon in more detail.
+We use the reputation choice variant to explore numerically the extent to which better algorithms under the \HardMaxRandom decision rule are incentivized in realistic time-scales.
+We show that for small but still ``relevant" parameter values, this effect does not show up for unless the time horizon is very long.
+\\
+\xhdr{Reputation and Data First-Mover Advantage}
+In the reputation choice variant, we investigate the ``first-mover advantage" phenomenon in more detail.
 Being first in the market gives free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
@@
 
-We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition.
+\xhdr{Predicting Outcomes in Competition} We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition in the reputation choice variant.
 We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor.
-We suggest a
-more refined performance measure, and use it to explain some of the competition outcomes.
+We suggest a more refined performance measure, based on a comparison between the reputation of the two firms, and use it to explain some of the competition outcomes.
+\textbf{Add some discussion about comparison with ``better" algorithms in analytical part}
 
@@
 In our model, competition varies based on the number of ``free" consumers that a firm receives.
-In the expectation choice model, this comes from introducing ``random" consumers where we vary from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
+One economic mechanism that can generate this comes from introducing ``random" consumers where we vary from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
 Indeed, with \HardMax you lose all customers as soon as you fall behind in performance, with \HardMaxRandom you get some small market share no matter what, and with \SoftMaxRandom you are further guaranteed a market share close to $\tfrac12$ as long as your performance is not much worse than the competition.
 The uniform choice among principals corresponds to no rationality and no competition.
-In the reputation choice model, this comes from the number of firms and the timing of entry in the market.
+Another economic mechanism that can generate this comes from the number of firms and the timing of entry in the market.
 Competition in this case ranges from monopoly to ``incumbent" (first-mover in duopoly) to simultaneous entry duopoly to ``late entrant" (last mover in duopoly).
@@
 We start by analyzing the expectation choice model analytically.
-In Sections ~\ref{sec:rational},~\ref{sec:random}, and ~\ref{sec:soft} we analyze the first main variant of the model and characterize the equilibrium behavior under three different agent decision rules.
+In Sections ~\ref{sec:rational},~\ref{sec:random}, and ~\ref{sec:soft} we analyze the expectation choice variant and characterize the equilibrium behavior under three different agent decision rules.
 Then, we turn to analyze the reputation choice variant of the model with details of the analysis described in Section ~\ref{sec:sim_details}.
@@
 without competition).
-Sections ~\ref{sec:competition} and ~\ref{sec:barriers} overview the results of the second variant of the mode.
+Sections ~\ref{sec:competition} and ~\ref{sec:barriers} overview the results of the reputation choice variant.
+Section ~\ref{sec:non_greedy} presents results of varying the consumer choice rule in the reputation choice variant.
 Section ~\ref{sec:conclusion} concludes.
-}
 

### 2019-11-08 journal paper/content/sec-intro.tex
updates

@@
 Learning from interactions with consumers is ubiquitous in modern customer-facing systems, from product recommendations to web search to content selection to fine-tuning user interfaces.
-These interactions are essential for the development of high-quality products, but, in order to learn what products are high quality, companies need to implement systems that engage in \emph{exploration}: deploying potentially suboptimal products in order to acquire new information to make more informed product decisions tomorrow.
+These interactions are essential for the development of high quality products, but, in order to learn what products are high quality, firms need to implement systems that engage in \emph{exploration}: deploying potentially suboptimal products in order to acquire new information to make more informed product decisions tomorrow.
 
 In this paper, we initiate a study of the interplay between \exploration and \competition.
-While exploration may be beneficial for both firms and consumers in the long-run, it may lead to bad experiences for consumers today and result in subsequent consumers going to the firm's competitors tomorrow.
-In particular, a firm can only make progress on its learning problem if it is able to incentivize consumers to visit it over its competitors and excessive exploration may lead to low reputation which leads consumers to increasingly visit the firm's competitors.
+Competition induces firms to want to produce high quality goods, but they need to learn how to do so via exploration.
+However, while exploration may provide firms with the necessary information to allow them to produce high quality products in the long-run, it may lead to bad experiences for consumers today and adversely affect a firm's reputation.
+This may incentivize subsequent consumers to go to the firm's competitors tomorrow and leave the firm with a lower market share and less consumers to learn from.
 Consumers therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing firms.
 
-Given the prevalence of these technologies and the increased importance of data in the digital economy, our main high-level question is: {\bf does competition lead to the adoption of better exploration algorithms?}.
-This translates into a number of more concrete questions.
+Given the prevalence of these technologies and the increased importance of data in the digital economy, our main high-level question is: {\bf does competition lead to the adoption of better exploration algorithms?} This translates into a number of more concrete questions.
 While it is commonly assumed that ``better" technology always helps, is this so for our setting?
@@
 To what extent is there a ``data feedback loop" where one firm having more data leads to that firm attracting more users which leads to that firm having more data, etc.?
-
-We investigate these questions with a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
-We first look at a version of this model that allows us to make analytical characterizations of the optimal algorithms firms should deploy in equilibrium.
-In this variant of the model, which we denote as the \textit{expectation choice variant}, we suppose that consumers are well-informed about the firm's exploration strategies but do not observe any information about past user experiences with the firms.
-In this variant we analytically characterize the equilibrium exploration strategies as we vary the degree of randomness in consumer decisions.
-We then look at a second variant of the model that accounts for competition in a more direct way, which we denote as the \textit{reputation choice variant}, where we suppose that consumers \textit{only} consider the average of past user experiences with each of the firms.
-This variant allows us to utilize numerical simulations to consider a number of additional issues such as asymmetries in the timing of entry, increasing competition beyond two firms, and allowing us to isolate the role that informational advantages can play as a barrier to entry in markets where exploration and competition tensions arise.
-
-On a high level, our contributions can be framed in terms of the ``inverted-U relationship" between competition and the quality of adopted algorithms (see \reffig{fig:inverted-U}).
-
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
-      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
-      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
-      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
- \end{tikzpicture}
-
-\caption{Inverted-U relationship between competitiveness and algorithms.}
-\label{fig:inverted-U}
-\end{center}
-\end{figure}
 
@@
 
-\xhdr{Our model.} We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
+\xhdr{Our model.} We investigate these questions with a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
+We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
 These two processes are interlinked, as exploration decisions are experienced by users and informed by their feedback.
@@
 Each principal only observes agents that chose him.
-The information set for the agent differs between the two variants of the model.
+We consider two variants of the baseline model and the information set for the agent differs between the two:
 \begin{enumerate}
-\item In the expectation choice variant, agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms
-\item In the reputation choice variant, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal
+\item In the \textit{expectation choice variant}, agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms.
+\item In the \textit{reputation choice variant}, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal.
 \end{enumerate}
@@
 }
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
+      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
+      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
+      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
+ \end{tikzpicture}
+
+\caption{Inverted-U relationship between competitiveness and algorithms.}
+\label{fig:inverted-U}
+\end{center}
+\end{figure}
 \xhdr{Economic interpretation.}

### 2019-11-08 journal paper/content/sec-model.tex
updates

@@
 We consider two variants of our model with different information structures.
-In the first, the  \textit{expectation choice} model, each agent knows both principals' algorithms, and the global round when (s)he arrives, \emph{but not} the rewards of the previous agents.
-In the second, the \textit{reputation choice} model, agents only make decisions based on the rewards of previous agents.
+In the first, the  \textit{expectation choice} variant, each agent knows both principals' algorithms, and the global round when (s)he arrives, \emph{but not} the rewards of the previous agents.
+In the second, the \textit{reputation choice} variant, agents only make decisions based on the rewards of previous agents.
 Concretely, each of the two principals has a \emph{reputation score}, and each agent's choice is driven by these two numbers.
@@
 \gaedit{
-In the reputation choice model we consider, at a given time $t$ each principal $i \in \{1, 2 \}$ has a reputation score denoted as $\REP_i (t)$.
+In the reputation choice variant we consider, at a given time $t$ each principal $i \in \{1, 2 \}$ has a reputation score denoted as $\REP_i (t)$.
 In this case we have that the agent's responses take an analogous form as in the first case:

### 2019-11-08 journal paper/ec19paper/content/barriers.tex
the latex compiles
updates

--- 2019-02-14 paper/content/barriers.tex
@@
 
-We find that either data or reputational advantage alone gives a substantial boost to the incumbent, compared to permanent duopoly.
+We find that either data or reputational advantage alone gives a substantial boost to the incumbent, compared to simultaneous entry duopoly.
 The results for the Heavy-Tail instance are presented in Table~\ref{barrier_exp}, in the same structure as Table~\ref{tab:ht-incum}.
@@
 
-\begin{table*}[t]
+\begin{table*}[h]
 \centering
@@
 \hline
-  & \multicolumn{3}{c||}{Reputation advantage}
-  & \multicolumn{3}{c|}{Data advantage}\\
+  & \multicolumn{3}{c||}{Reputation advantage (only)}
+  & \multicolumn{3}{c|}{Data advantage (only)}\\
 \hline
@@
 
-We can quantitatively define the data (resp., reputation) advantage as the incumbent's market share in the competition period in the data-advantage (resp., reputation advantage) experiment, minus the said share under permanent duopoly, for the same pair of algorithms and the same problem instance.
+We can quantitatively define the data (resp., reputation) advantage as the incumbent's market share in the competition period in the data-advantage (resp., reputation advantage) experiment, minus the said share under simultaneous entry duopoly, for the same pair of algorithms and the same problem instance.
 In this language, our findings are as follows.

### 2019-11-08 journal paper/ec19paper/content/inverted_u.tex
updates

@@
 \normalsize
-\caption{{\bf Permanent duopoly}, for Heavy-Tail and Needle-in-Haystack instances.
+\caption{{\bf Simultaneous Entry Duopoly}, for Heavy-Tail and Needle-in-Haystack instances.
 Each cell describes a game between two algorithms, call them Alg1 vs.
@@
 Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from permanent monopoly to ``incumbent" to permanent duopoly to ``late entrant", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
+As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
 More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
@@
 For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{MWT-WhitePaper-2016,DS-arxiv}).
-Note that \DG is trivially the dominant strategy under permanent monopoly.
+Note that \DG is trivially the dominant strategy under monopoly.
 
@@
 
-\xhdr{Permanent duopoly.}
+\xhdr{Simultaneous entry duopoly.}
 The basic scenario is when both firms are competing from round $1$. A crucial distinction is whether an MAB instance is exploration-disadvantaged:
@@
 \begin{finding}\label{find:duopoly}
-\textit{Under permanent duopoly:
+\textit{Under simultaneous entry duopoly:
 \begin{itemize}
@@
 \end{tabular}
-\caption{{\bf Permanent duopoly}, for the Uniform MAB instance. Semantics are the same as in Table \ref{sim_table}.}
+\caption{{\bf Simultaneous entry duopoly}, for the Uniform MAB instance. Semantics are the same as in Table \ref{sim_table}.}
 \label{tab:duopoly_unif}
@@
 The lowest level of competition is monopoly, when \DG wins out for the trivial reason of tie-breaking.
-The highest levels are permanent duopoly and ``late entrant".
+The highest levels are simultaneous entry duopoly and ``late entrant".
 We see that \DG is incentivized for exploration-disadvantaged MAB instances.
-In fact, incentives for \DG get stronger when the model transitions from permanent duopoly to ``late entrant".%
-\footnote{For the Heavy-Tail instance, \DG goes from a weakly dominant startegy to a strictly dominant one.
+In fact, incentives for \DG get stronger when the model transitions from simultaneous entry duopoly to ``late entrant".%
+\footnote{For the Heavy-Tail instance, \DG goes from a weakly dominant strategy to a strictly dominant one.
 For the Uniform instance, \DG goes from a Nash equilibrium strategy to a weakly dominant one.}
@@
 Our intuition for why incumbency creates more incentives for exploration is as follows.
-During the temporary monopoly period, reputation costs of exploration vanish.
+During the temporary monopoly period, reputation consequences of exploration vanish.
 Instead, the firm wants to improve its performance as much as possible by the time competition starts.
-Essentially, the firm only faces a classical explore-exploit tradeoff, and is incentivized to choose algorithms that are best at optimizing this tradeoff.
+Essentially, the firm only faces a classical explore-exploit trade-off, and is incentivized to choose algorithms that are best at optimizing this trade-off.
 
@@
 We interpret low \Eeog as a strong evidence of the ``death spiral" effect.
-Focusing on the permanent duopoly scenario, we specify the \Eeog values in Table \ref{sim_table} (the second line of each cell).
+Focusing on the simultaneous entry duopoly scenario, we specify the \Eeog values in Table \ref{sim_table} (the second line of each cell).
 We find that the \Eeog values are indeed small:
@@
 \textit{
-Under permanent duopoly, \Eeog values tend to be much smaller than the time horizon $T$.
+Under simultaneous entry duopoly, \Eeog values tend to be much smaller than the time horizon $T$.
 }

### 2019-11-08 journal paper/journal_main.tex
updates

@@
 \begin{abstract}
-\gaedit{Most modern systems strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-We initiate a study of the interplay between \emph{exploration and competition}---how such systems balance the exploration for learning and the competition for users.
-Here the users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing systems.
+Most modern firms strive to learn from interactions with consumers, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
+We initiate a study of the interplay between \emph{exploration} and \emph{competition}—how such firms balance the exploration for learning and the competition for consumers.
+Here consumers play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing firms.
 
-In our model, we consider competition between two multi-armed bandit algorithms faced with the same bandit instance.
-Users arrive one by one and choose among the two algorithms, so that each algorithm makes progress if and only if it is chosen.
-We ask whether and to what extent competition incentivizes the adoption of better bandit algorithms.
-We investigate this issue analytically for several models of user response, as we vary the degree of competitiveness in the model.
-We further investigate this issue using numerical simulation to understand the effect of the first-mover advantage in our model.
+In our model, we consider competition between two firms facing the same multi-armed bandit instance who simultaneously choose a bandit algorithm.
+Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance if and only if it is chosen.
+We study to what extent competition leads to welfare increases for consumers and what algorithms firms adopt under competition.
+We find that stark competition induces firms to commit to a ``greedy" algorithm that leads to low consumer welfare.
+However, we find that in order to incentivize ``better" exploration strategies a firm needs to have some ``free" consumers that visit them without having to be incentivized to do so.
+We investigate this effect via noise in consumer decision-making and asymmetries in timing of entry of firms into the market.
 Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets and are closely related to the ``competition vs.
-innovation" relationship, a well-studied theme in economics.}
+innovation" relationship, a well-studied theme in economics.
 \end{abstract}
@@
 
+\subfile{ec19paper/content/non_greedy_choice}
+
 \gaedit{
@@
 }
-%\subfile{ec19paper/content/non_greedy_choice}
 
@@
 \bibliographystyle{acm}
-\bibliography{bib-ML,refs,bib-abbrv-short,bib-bandits,bib-AGT,bib-slivkins}
+\bibliography{bib-ML,refs,bib-abbrv-short,bib-bandits,bib-AGT,bib-slivkins, bib-random}
 

### 2019-11-30 journal paper/content/sec-intro.tex
add updated poster and edits in journal sec

@@
 
-Given the prevalence of these technologies and the increased importance of data in the digital economy, our main high-level question is: {\bf does competition lead to the adoption of better exploration algorithms?} This translates into a number of more concrete questions.
+Given the prevalence of these technologies in the digital economy, our main high-level question is: {\bf does competition lead to the adoption of better exploration algorithms?} This translates into a number of more concrete questions.
 While it is commonly assumed that ``better" technology always helps, is this so for our setting?

### 2019-12-10 journal paper/content/sec-intro.tex
tweaks to paper

@@
 
-Given the prevalence of these technologies in the digital economy, our main high-level question is: {\bf does competition lead to the adoption of better exploration algorithms?} This translates into a number of more concrete questions.
+The main high-level question that we focus on in this paper is: {\bf does competition lead to the adoption of better exploration algorithms?} This translates into a number of more concrete questions.
 While it is commonly assumed that ``better" technology always helps, is this so for our setting?

### 2019-12-10 journal paper/content/sec-model.tex
fixup
tweaks to paper

@@
 \footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before.
-We will omit the local vs.
+We will omit the global vs.
 local distinction when clear from the context.} each time it is called, it outputs a chosen action $a\in A$ and then inputs the reward for this action.
@@
 
-We will emphasize several specific algorithms or classes thereof. In the second half of the paper we will :
+We will emphasize several specific algorithms or classes thereof:
 
@@
         $\BIR(n) \leq  \tilde{O}(n^{-1/2})$
-    for all priors and all (or all but a very few) steps $n$. This bound is known to be tight for any fixed $n$.
-    \footnote{This follows from the lower-bound analysis in \cite{bandits-exp3}.}
+    for all priors and all (or all but a very few) steps $n$. This bound is known to be tight for any fixed $n$.\footnote{This follows from the lower-bound analysis in \cite{bandits-exp3}.}
 
@@
 
-Write
+We denote
     $\PMR(a\mid E) = \E[\mu_a \mid E]$

### 2019-12-10 journal paper/ec19paper/content/non_greedy_choice.tex
the latex compiles
tweaks to paper

--- 2019-02-14 paper/content/non_greedy_choice.tex
@@
 Let us consider an extension in which the agents' choice is no longer deterministic.
-Recall that in our main model agents deterministically choose the firm with the higher reputation score; call this choice rule \emph{HardMax} (\HM).
-Now, we introduce some randomness: each agent selects between the firms uniformly with probability $\eps\in (0,1)$, and takes the firm with the higher reputation score with the remaining probability; call this choice rule \emph{HardMax with randomness} (\HMR).
+Recall that in the reputation choice variant agents deterministically choose the firm with the higher reputation score, which, as before, we denote as \emph{HardMax} (\HM).
+Now, we introduce some randomness: each agent selects between the firms uniformly with probability $\eps\in (0,1)$, and takes the firm with the higher reputation score with the remaining probability; as before, we denote this rule as \emph{HardMax with randomness} (\HMR).
 
@@
 instead they are dispersed throughout the game.}
+
+\begin{finding}\label{find:non_greedy_choice}
+\textit{\TS is weakly dominant under the \HMR~ choice rule, if and only if $\eps T$ is sufficiently large. \HMR leads to lower variance in market shares, compared to \HM.}
+\end{finding}
 
@@
 
-\begin{finding}\label{find:non_greedy_choice}
-\textit{\TS is weakly dominant under the \HMR~ choice rule, if and only if $\eps T$ is sufficiently large. \asedit{\HMR leads to lower variance in market shares, compared to \HM.}}
-\end{finding}
-
-Table \ref{tab:non_greedy_table} shows the average market shares under
-the \HM~ vs \HMR~ choice rule. In contrast to the \HM~ model,
-  \TS becomes weakly dominant under the \HMR model, as $T$ gets
-  sufficiently large. These findings hold across all problem
-instances, see Table \ref{tab:additional_results} (with the same semantics as in Table \ref{tab:non_greedy_table}).
-
-\OMIT{\footnote{The results here are pulled using different \MRV
-  and realizations than the results pulled previously (due to the
-  larger $T$). However, they are drawn from the same prior instances
-  and so qualitatively are the same but the quantitative results are
-  not directly comparable to those from the previous
-  section.\swcomment{not sure we need this}}}
-
-\normalsize
-\OMIT{The intuition for why \TS
-should become the dominant strategy eventually is simple. The
-consistent stream of random agents ensures that each principal is
-chosen at least $\Omega(\eps t)$ times at every time step $t$. As
-a result, each algorithm should eventually converge to its asymptotic
-performance in isolation.}
-
-\OMIT{Finding \ref{find:non_greedy_choice} implies that we can re-interpret the inverted-U findings from before in terms of the number of agents that a firm receives without having to worry about incentives.
-In the extreme when the firm gets all agents for free as in the monopoly case then it is incentivized to play \DG.
-When it only gets some of the agents for free, either via a large warm start, a temporary monopoly, or non-deterministic choice, then \TS is incentivized.
-However, if the number of free agents gets small enough then \DG is incentivized as in the permanent duopoly analysis from before.\swcomment{I don't understand this paragraph; I think it's too vague.
-Perhaps remove?}}
-
 \footnotesize
-\begin{table*}[ht]
+\begin{table*}[t]
 \centering
@@
 
+Table \ref{tab:non_greedy_table} shows the average market shares under
+the \HM~ vs \HMR~ choice rule. In contrast to the \HM~ model,
+  \TS becomes weakly dominant under the \HMR model, as $T$ gets
+  sufficiently large. These findings hold across all problem
+instances, see Table \ref{tab:additional_results} (with the same semantics as in Table \ref{tab:non_greedy_table}).
+
+\OMIT{\footnote{The results here are pulled using different \MRV
+  and realizations than the results pulled previously (due to the
+  larger $T$). However, they are drawn from the same prior instances
+  and so qualitatively are the same but the quantitative results are
+  not directly comparable to those from the previous
+  section.\swcomment{not sure we need this}}}
+
+\normalsize
+\OMIT{The intuition for why \TS
+should become the dominant strategy eventually is simple. The
+consistent stream of random agents ensures that each principal is
+chosen at least $\Omega(\eps t)$ times at every time step $t$. As
+a result, each algorithm should eventually converge to its asymptotic
+performance in isolation.}
+
+\OMIT{Finding \ref{find:non_greedy_choice} implies that we can re-interpret the inverted-U findings from before in terms of the number of agents that a firm receives without having to worry about incentives.
+In the extreme when the firm gets all agents for free as in the monopoly case then it is incentivized to play \DG.
+When it only gets some of the agents for free, either via a large warm start, a temporary monopoly, or non-deterministic choice, then \TS is incentivized.
+However, if the number of free agents gets small enough then \DG is incentivized as in the permanent duopoly analysis from before.\swcomment{I don't understand this paragraph; I think it's too vague.
+Perhaps remove?}}
+
 However, it takes a significant amount of randomness and a relatively large time horizon for this effect to take place.

### 2020-06-28 journal paper/content/omitted_proofs.tex
updates

\subsection{Proof of Theorem~\ref{thm:HardMax-hardTies}}

\begin{proof}
The proof re-uses Lemmas~\ref{lm:DG-rew} and~\ref{lm:DG-sudden}, which do not rely on fair tie-breaking.

Because of the biased tie-breaking, for each global round $t$ we have:
\begin{align}\label{eq:thm:HardMax-biased-PMRtoPr}
\text{if $\PMR_1(t)\geq \PMR_2(t)$ then $\Pr[i_t=1]>\tfrac12$.}
\end{align}
Recall that $i_t$ is the principal chosen in global round $t$.

Let $m_0$ be the first step when \alg[2] deviates from \DynGreedy, or \DynGreedy deviates from \StaticGreedy, whichever comes sooner. \asedit{Then \alg[2], \DynGreedy and \StaticGreedy coincide on the first $m_0-1$ steps. Moreover, $m_0\leq n_0$ (since \DynGreedy deviates from \StaticGreedy at step $n_0$), so \alg[1] coincides with \DynGreedy on the first $m_0$ steps.

So, $\rew_1(n)=\rew_2(n)$ for each step $n<m_0$, because \alg[1] and \alg[2] coincide on the first $m_0-1$ steps. Moreover, if \alg[2] deviates from \DynGreedy at step $m_0$ then
    $\rew_1(m_0) > \rew_2(m_0)$ 
by Lemma~\ref{lm:DG-rew}; else, we trivially have
    $\rew_1(m_0) = \rew_2(m_0)$.} To summarize:
\begin{align}\label{eq:thm:HardMax-biased-rew}
    \rew_1(n)\geq\rew_2(n)\quad \text{for all steps $n\leq m_0$}.
\end{align}

We claim that $\Pr[i_t=1]>\tfrac12$ for all global rounds $t\leq m_0$. We prove this claim using induction on $t$. The base case $t=1$ holds by \eqref{eq:thm:HardMax-biased-PMRtoPr} and the fact that in step 1, \DynGreedy chooses the arm with the highest prior mean reward. For the induction step, we assume that $\Pr[i_t=1]>\tfrac12$ for all global rounds $t<t_0$, for some $t_0\leq  m_0$. It follows that distribution $\posteriorN{1}{t_0}$ stochastically dominates distribution $\posteriorN{2}{t_0}$.%
\footnote{For random variables $X,Y$ on \R, we say that $X$ \emph{stochastically dominates} $Y$ if $\Pr[X\geq x] \geq \Pr[Y\geq x]$ for any $x\in \R$.}
Observe that
\begin{align}\label{eq:thm:HardMax-biased-PMR-aux}
\PMR_1(t_0)
  = \Ex{n\sim \posteriorN{1}{t_0}} {\rew_1(n+1)}
  \geq \Ex{n\sim \posteriorN{2}{t_0}} {\rew_2(n+1)}
  = \PMR_2(t_0).
\end{align}
So the induction step follows by \eqref{eq:thm:HardMax-biased-PMRtoPr}. Claim proved.

Now let us focus on global round $m_0$, and denote $\mN_i = \posteriorN{i}{m_0}$.  By the above claim,
\begin{align}\label{eq:thm:HardMax-biased-mN}
\text{$\mN_1$ stochastically dominates $\mN_2$, and moreover
    $\mN_i(m_0-1)>\mN_i(m_0-1)$.}
\end{align}

By definition of $m_0$, either (i) \alg[2] deviates from \DynGreedy starting from local step $m_0$, which implies $\rew_1(m_0)> \rew_2(m_0)$ by Lemma~\ref{lm:DG-rew}, or (ii) \DynGreedy deviates from \StaticGreedy starting from local step $m_0$, which implies $\rew_1(m_0)>\rew_1(m_0-1)$ by Lemma~\ref{dgmono}. In both cases, using \eqref{eq:thm:HardMax-biased-rew} and \eqref{eq:thm:HardMax-biased-mN}, it follows that the inequality in \eqref{eq:thm:HardMax-biased-PMR-aux} is strict for $t_0=m_0$.

Therefore, agent $m_0$ chooses principal $1$, and by Lemma~\ref{lm:DG-sudden} so do all subsequent agents.
\end{proof}

\subsection{Proof of Theorem~\ref{thm:random-greedy}}
\begin{proof}
  Let $\rew'_1(n)$ denote the Bayesian-expected reward of the ``greedy
  choice'' after after $n-1$ steps of \alg[1]. Note that
  $\rew_1(\cdot)$ and $\rew'_1(\cdot)$ are non-decreasing: the former
  because \alg[1] is monotone and the latter because the ``greedy
  choice'' is only improved with an increasing set of
  observations. Therefore, the modified algorithm \alg[2] is monotone
  by \eqref{eq:random-BIR}.

  By definition of the ``greedy choice,'' $\rew_1(n)\leq \rew'_1(n)$
  for all steps $n$. Moreover, by Lemma~\ref{lm:DG-rew},
  \alg[1] has a strictly smaller $\rew(n_0)$ compared to \DynGreedy;
  so, $\rew_1(n_0)<\rew_2(n_0)$.

Let $\alg$ denote a copy of \alg[1] that is running ``inside" the modified algorithm \alg[2]. Let $m_2(t)$ be the number of global rounds before $t$ in which the agent chooses principal $2$ \emph{and} \alg is invoked; in other words, it is the number of agents seen by \alg before global round $t$. Let $\mM_{2,t}$ be the agents' posterior distribution for $m_2(t)$.

We claim that in each global round $t\geq n_0$, distribution $\mM_{2,t}$ stochastically dominates distribution $\posteriorN{1}{t}$, and $\PMR_1(t)<\PMR_2(t)$. We use induction on $t$. The base case $t=n_0$ holds because $\mM_{2,t} = \posteriorN{1}{t}$ (because the two algorithms coincide on the first $n_0-1$ steps), and $\PMR_1(n_0)<\PMR_2(n_0)$ is proved as in \eqref{eq:pf:thm:DG-dominance}, using the fact that $\rew_1(n_0)<\rew_2(n_0)$.

The induction step is proved as follows. The induction hypothesis for global round $t-1$ implies that agent $t-1$ is seen by \alg with probability $(1-\eps_0)(1-p)$, which is strictly larger than $\eps_0$, the probability with which this agent is seen by \alg[2]. Therefore, $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$.
\begin{align}
\PMR_1(t)
  &= \Ex{n\sim \posteriorN{1}{t}} {\rew_1(n+1)} \nonumber \\
  &\leq \Ex{m\sim \mM_{2,t}} {\rew_1(m+1)}
    \label{eq:pf:thm:random-greedy-1}\\
  &< \Ex{m\sim \mM_{2,t}} {(1-p)\cdot \rew_1(m+1) + p\cdot \rew'_1(m+1)}
    \label{eq:pf:thm:random-greedy-2} \\
  &= \PMR_2(t). \nonumber
\end{align}
Here inequality \eqref{eq:pf:thm:random-greedy-1} holds because $\rew_1(\cdot)$ is monotone and $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$, and inequality \eqref{eq:pf:thm:random-greedy-2} holds because $\rew_1(n_0)<\rew_2(n_0)$ and $\mM_{2,t}(n_0)>0$.%
\footnote{If $\rew_1(\cdot)$ is strictly increasing, then inequality \eqref{eq:pf:thm:random-greedy-1} is strict, too; this is because $\mM_{2,t}(t-1)>\posteriorN{1}{t}(t-1)$.  }
\end{proof}

\subsection{Proof of Theorem~\ref{thm:SoftMax-strong}}
\begin{proof}
  Let $\beta_1 = \min\{c_0'\delta_0, \beta_0/20\}$ with $\delta_0$
  defined in~\eqref{eq:SoftMax-smooth}.  Recall each agent chooses
  \alg[1] with probability at least $\respF(-1)= \eps_0$.  By
  condition \eqref{eq:so-mild} and \eqref{eq:SoftMax-assn-strong},
  there exists some sufficiently large $T_1$ such that for any
  $t\geq T_1$, $\BIR_1(\eps_0 T_1/2) \leq \beta_1/c_0'$ and
  $\BIR_2(t) > e^{-\eps_0 t/12}$. Moreover, for any $t\geq T_1$, we
  know $\E[n_1(t+1)] \geq \eps_0\,t $, and by the Chernoff Bounds
  (Theorem~\ref{thm:chernoff}), we have $n_1(t+1) \geq \eps_0 t/2$
  holds with probability at least $1 - q_1(t)$ with
  $q_1(t) = \exp(-\eps_0 t/12) < \BIR_2(t)$. It follows that for any $t\geq T_1$,
\begin{align*}
  \PMR_2(t) - \PMR_1(t) &= \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}{\BIR_1(m_1+ 1) - \BIR_2(m_2+1)} \\
                        &\leq q_1(t)  + \Ex{m_1\sim \posteriorN{1}{t}}{\BIR_1(m_1+ 1)\mid m_1 \geq \eps_0 t/2 - 1 } - \BIR_2(t)\\
                        &\leq \BIR_1(\eps_0 T_1/2) \leq \beta_1/c_0'
\end{align*}
% where the second inequality follows from~\eqref{eq:random-assn}.
Since the response function $\respF$ is $c_0'$-Lipschitz in the
neighborhood of $[-\delta_0, \delta_0]$, each agent after round $T_1$
will choose \alg[1] with probability at least
\[
  p_t \geq \frac{1}{2} - c_0'\left(\PMR_2(t) - \PMR_1(t)\right) \geq
  \frac{1}{2} - \beta_1.
\]

Next, we will show that there exists a sufficiently large $T_2$ such
that for any $t\geq T_1 + T_2$, with high probability
$n_1(t) > \max\{n_0, (1 - \beta_0)n_2(t)\}$, where $n_0$ is defined
in~\eqref{eq:SoftMax-better}. %  Let us first lower bound the number of
% agents received by \alg[1] after some number of rounds $t = T_1 + T'$
% for any $T' \geq T_1$. 
Fix any $t \geq T_1 + T_2$. 
Since each agent chooses \alg[1] with probability at least
$1/2 - \beta_1$, by Chernoff Bounds (Theorem~\ref{thm:chernoff}) we
have with probability at least $1 - q_2(t)$ that the number of agents
that choose \alg[1] is at least $\beta_0(1/2 - \beta_1)t/5$, where the
function
$$
q_2(x) = \exp\left( \frac{-(1/2 - \beta_1)(1 - \beta_0/5)^2x}{3} \right).
$$
Note that the number of agents received by \alg[2] is at most
$T_1 + (1/2 + \beta_1)t + (1/2 - \beta_1)(1 - \beta_0/5)t$.

Then as long as
% $T_2 \geq \max\left\{\frac{3T_1}{(1 - \beta_0)}, 8 n_0\right\}$, we
% can guarantee that for any $t \geq T_1 + T_2$,
% $n_1(t) > n_2(t) (1 - \beta_0)$ and $n_1(t) > n_0$ with probability at
% least $1 - q_2(t)$.
$T_2 \geq \frac{5T_1}{\beta_0}$, we can guarantee that
$n_1(t) > n_2(t) (1 - \beta_0)$ and $n_1(t) > n_0$ with probability at
least $1 - q_2(t)$ for any $t \geq T_1 + T_2$.
% for any $t\geq T_1 + T_2$, as long as $T_2 \geq .$ Moreover, to
% guarantee that , we will just need $T_2 \geq$.  Finally, we will
% argue that in each round $t \geq T_1 + T_2$, we recover the
% guarantee in \eqref{eq:thm:SoftMax-strong}.
% % \[
% %   \Pr[i_t = 1] \geq \frac{1}{2} + \frac{c_0\alpha_0\BIR_2(t)}{4}.
% % \]
Note that the weak BIR-dominance condition
in~\eqref{eq:SoftMax-better} implies that for any $t\geq T_1 + T_2$
with probability at least $1 - q_2(t)$,
\[
  \BIR_1(n_1(t)) < (1- \alpha_0)\BIR_2(n_2(t)).
\]
It follows that for any $t\geq T_1 + T_2$,
\begin{align*}
  \PMR_1(t) - \PMR_2(t) &= \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}{\BIR_2(m_2+ 1) - \BIR_1(m_1+1)} \\
                        &\geq (1 - q_2(t))\alpha_0 \BIR_2(t) - q_2(t)\\
                        &\geq \alpha_0 \BIR_2(t)/4
\end{align*}
where the last inequality holds as long as
$q_2(t) \leq \alpha_0\BIR_2(t)/4$, and is implied by the condition
in~\eqref{eq:SoftMax-assn-strong} as long as $T_2$ is sufficiently
large. Hence, by the definition of our \SoftMaxRandom response
function and assumption in~\eqref{eq:SoftMax-smooth}, we have
\[
  \Pr[i_t = 1] \geq \frac{1}{2} + \frac{c_0\alpha_0\BIR_2(t)}{4}. \qedhere
\]
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main.tex"
%%% End:

### 2020-06-28 journal paper/content/sec-intro.tex
updates

@@
-Learning from interactions with consumers is ubiquitous in modern customer-facing systems, from product recommendations to web search to content selection to fine-tuning user interfaces.
+Learning from interactions with users is ubiquitous in modern customer-facing systems, from product recommendations to web search to content selection to fine-tuning user interfaces.
 These interactions are essential for the development of high quality products, but, in order to learn what products are high quality, firms need to implement systems that engage in \emph{exploration}: deploying potentially suboptimal products in order to acquire new information to make more informed product decisions tomorrow.
@@
 In this paper, we initiate a study of the interplay between \exploration and \competition.
-Competition induces firms to want to produce high quality goods, but they need to learn how to do so via exploration.
-However, while exploration may provide firms with the necessary information to allow them to produce high quality products in the long-run, it may lead to bad experiences for consumers today and adversely affect a firm's reputation.
-This may incentivize subsequent consumers to go to the firm's competitors tomorrow and leave the firm with a lower market share and less consumers to learn from.
-Consumers therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing firms.
+The fundamental tension is that while exploration allows for the deployment of high quality products in the long-run, it leads to poor user experiences in the short-run.
+Exploration can therefore harm the perception users have of a system, which may lead them to instead utilize a competing system.
+This would leave the system with lower market share and therefore less users to learn from relative to its competitors and allow its competitors to make further progress on their learning problem.
+Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.
 
-The main high-level question that we focus on in this paper is: {\bf does competition lead to the adoption of better exploration algorithms?} This translates into a number of more concrete questions.
+The main high-level question that we focus on in this paper is whether competition between systems incentives the adoption of better exploration algorithms.
+This translates into a number of more concrete questions.
 While it is commonly assumed that ``better" technology always helps, is this so for our setting?
 Does increasing competition lead to higher consumer welfare?
-To what extent is there a ``data feedback loop" where one firm having more data leads to that firm attracting more users which leads to that firm having more data, etc.?
+To what extent is there a ``data feedback loop" where one firm having more data leads to that firm attracting more users which leads to that firm having more data, etc.?\footnote{This is a fundamental question which is part of a larger policy discussion around whether data can serve as an indirect network effect and lead to similar ``market tipping" results as is standard in the literature on competition in markets with network effects (see \cite{jullien2019economics} for a policy oriented discussion of this).}
 
@@
 \xhdr{Main Findings.}
-We find that in a simultaneous entry duopoly a greedy (myopic) algorithm that does no purposeful exploration is incentivized in equilibrium.
-In the expectation choice variant, once a firm does any exploration it is immediately starved of consumers when its opponent plays a greedy algorithm since we show that the expected reward for every subsequent consumer is higher for the firm that plays the greedy algorithm.
-In the reputation choice variant, the same mechanism drives the greedy algorithm to be incentivized in equilibrium.
-In this variant, exploration hurts a firm's reputation and decrease its market share in the near term.
-This leaves the firm with less users to learn from, which further degrades the firm's performance relative to competitors who keep learning and improving from \emph{their} customers, and so forth.
-Taken to the extreme, such dynamics lead to a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
-As a result, our model provides an example of the mechanisms that can lead to a ``data feedback loop" that has been hypothesized in policy discussions.
-In both cases, consumer welfare is low since the greedy algorithm is known to be dramatically bad in many important cases of multi-armed bandits.
+We find that competition induces firms to commit to a greedy (myopic) algorithm in equilibrium that does no purposeful exploration.
+This leads to low consumer welfare since the greedy algorithm is known to be dramatically bad in many important cases of multi-armed bandits.
+The primary mechanism that generates this stark result is that consumers need to be incentivized to select a firm over its competitors, leading a firm that engages in exploration to be starved of consumers before it makes enough progress on its learning problem.
+In order to incentivize ``better" exploration strategies in equilibrium, the key intuition is that the firm needs to have some ``free" consumers that visit them without the firm having to incentivize them to do so.
+This allows the firm to eventually overcome the initial losses in consumer perception from exploration.
+We explore two economic mechanisms that can generate such an effect.
 
-The primary mechanism that generates this stark result is that consumers need to be incentivized to select a firm over its competitors, leading firms that engage in exploration to be starved of consumers before they make enough progress on their learning problem.
-In order to incentivize ``better" exploration strategies the key intuition is that the firm needs to have some ``free" consumers that visit them without the firm having to incentivize them to do so.
-This allows the firm to eventually overcome the initial losses in consumer perception from exploration.
+The first is the presence of users who randomly choose between the firms which gives each firm a constant stream of ``free" users and thus ensures it never gets fully starved.
+We find that, even when there is a small fraction of such consumers, better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.\footnote{While the precise notion of ``sufficiently better algorithm" is rather subtle, we note that commonly known ``smart" bandit algorithms typically defeat the commonly known ``naive" ones, and the latter typically defeat the greedy algorithm.
+However, there is a substantial caveat: one can defeat any algorithm by interleaving it with the greedy algorithm.
+This has two undesirable corollaries: a better algorithm may sometimes lose, and a pure Nash equilibrium typically does not exist.} However, while this holds asymptotically, our numerical experiments show that this requires either unreasonably large time scales or a large fraction of such users.
 
-We relax the decision rule of the consumers to allow for each firm to be chosen with some fixed baseline probability so that firms get a constant stream of ``free" consumers; we call this choice rule \HardMaxRandom.
-We find that, in this setting, better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
-While the precise notion of ``sufficiently better algorithm" is rather subtle, we note that commonly known ``smart" bandit algorithms typically defeat the commonly known ``naive" ones, and the latter typically defeat the greedy algorithm.
-However, there is a substantial caveat: one can defeat any algorithm by interleaving it with the greedy algorithm.
-This has two undesirable corollaries: a better algorithm may sometimes lose, and a pure Nash equilibrium typically does not exist.
-
-We use the reputation choice variant to analyze the effect of varying the timing of entry between the firms and allow one firm to have a first-mover advantage so that this firm (denoted as the incumbent) gets all the agents in the market before the other firm enters.
-We find that, if the first-mover period is sufficiently large, then the incumbent has a dominant strategy to play ``smart" bandit algorithms and that consumer welfare is substantially higher than in the case of simultaneous entry.
-The intuition is simply that a sufficiently long incumbency period allows the firm to make sufficient progress on its learning problem that it can overcome the original drop in its reputation from exploration in the beginning rounds.
+The second is that if one firm has a first-mover advantage then this firm gets free users before the other firm enters into the market.
+We show that if this incumbency period is sufficiently long, then it allows the firm to make sufficient progress on its learning problem and incentivizes it to commit to better algorithms.
+While this leads to the incumbent getting almost all of the market, it leads to higher consumer welfare than under competition when the firms enter at the same time.
 
@@
 
-\xhdr{Noise in consumer choice}: We further relax the decision rule of the agents so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' expected rewards; we call it \SoftMaxRandom.
+\xhdr{Reputation and Data Advantage}: In the reputation choice variant we investigate the first-mover advantage phenomenon in more detail.
+Being first in the market gives the incumbent free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
+We run additional experiments so as to isolate and compare these two effects.
+We find that either effect alone leads to a significant advantage under competition.
+The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.
+This result shows that even a small amount ``data advantage" gets amplified under competition, causing a large difference in eventual market shares.
+
+\xhdr{Noise in consumer choice}: We further relax the decision rule of the agents so that the probability of choosing a given firm varies smoothly as a function of the difference between  principals' expected rewards.
 For this decision rule, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
@@
 The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm may attract slightly more).
-
-We use the reputation choice variant to explore numerically the extent to which better algorithms under the \HardMaxRandom decision rule are incentivized in realistic time-scales.
-We show that for small but still ``relevant" parameter values, this effect does not show up for unless the time horizon is very long.
-\\
-\xhdr{Reputation and Data First-Mover Advantage}
-In the reputation choice variant, we investigate the ``first-mover advantage" phenomenon in more detail.
-Being first in the market gives free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
-We run additional experiments so as to isolate and compare these two effects.
-We find that either effect alone leads to a significant advantage under competition.
-The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.
-
-Data advantage is significant from an anti-trust perspective, as a possible barrier to entry.
-We find that even a small amount ``data advantage" gets amplified under competition, causing a large difference in eventual market shares.
-This observation runs contrary to prior work  \cite{lambrecht2015can,bajari2018impact}, which studied learning without competition, and found that small amounts of additional data do not provide significant improvement in eventual outcomes.
-We conclude that competition dynamics -- that firms compete as they learn over time -- are pertinent to these anti-trust considerations.
 
@@
 \end{figure}
-\xhdr{Economic interpretation.}
-The inverted-U relationship between the severity of competition among firms and the quality of technologies that they adopt is a familiar theme in the economics literature \citep[\eg][]{Aghion-QJE05,Vives-08}.%
-\footnote{The literature frames this relationship as one between ``competition" and ``innovation".
-In this context, ``innovation" refers to adoption of a better technology, at a substantial R\&D expense to a given firm.
+\xhdr{Economic interpretation.} Our model speaks to two distinct strands of the economics literature.
+
+The first is the literature that studies the relationship between competition and innovation where, in our context,  the choice between different algorithms is interpreted as firms determining whether to invest in a ``better" technology as the degree of competition varies.\footnote{The choice of algorithms has a clear analogue to ``innovation" in our context as the literature on multi-armed bandits has identified distinct ``classes" of algorithms according to asymptotic regret bounds.
+Thus, choosing algorithms in ``better" classes corresponds with innovation on the part of the firm in our interpretation of the model.
 It is not salient whether similar ideas and/or technologies already exist outside the firm.
-It is worth noting that the adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature; see \citet{MWT-WhitePaper-2016} for an example of such R\&D effort.}
+It is worth noting that the adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature; see \citet{MWT-WhitePaper-2016} for an example of such R\&D effort.} A classic result in this literature is that there is an inverted-U relationship between the severity of competition among firms and the quality of technologies that they adopt is a familiar theme in the economics literature \citep[\eg][]{Aghion-QJE05,Vives-08}.
+
 We find it illuminating to frame our contributions in a similar manner, as illustrated in \reffig{fig:inverted-U}.
-
-In our model, competition varies based on the number of ``free" consumers that a firm receives.
-One economic mechanism that can generate this comes from introducing ``random" consumers where we vary from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
-Indeed, with \HardMax you lose all customers as soon as you fall behind in performance, with \HardMaxRandom you get some small market share no matter what, and with \SoftMaxRandom you are further guaranteed a market share close to $\tfrac12$ as long as your performance is not much worse than the competition.
-The uniform choice among principals corresponds to no rationality and no competition.
-Another economic mechanism that can generate this comes from the number of firms and the timing of entry in the market.
-Competition in this case ranges from monopoly to ``incumbent" (first-mover in duopoly) to simultaneous entry duopoly to ``late entrant" (last mover in duopoly).
-The same distinctions in both cases also control the severity of competition between the principals.
-
-In both variants we identify an inverted-U relationship between competition and innovation in the spirit of \reffig{fig:inverted-U}.
-These inverted-U relationships arise for a fundamentally different reason, compared to the existing literature on ``competition vs.
+In our model we consider varying the intensity of competition as varying the degree to which a firm needs to incentivize users to visit them.
+Competition is less intense when there are more ``random" users or one firm has a sufficiently long first-mover advantage.
+The inverted-U relationship in our model arise for a fundamentally different reason, compared to the existing literature on ``competition vs.
 innovation.'' In the literature, better technology always helps in a competitive environment, other things being equal.
@@
 This leads to an implicit cost from exploration in the form of a reduced rate of users that a firm attracts and can learn from.
-However, interestingly, the economic mechanism in our model for incentivizing firms to engage in R\&D has a qualitative similarity to the role that patents play in incentivizing innovation in standard R\&D models.
+
+Interestingly, the economic mechanism in our model for incentivizing firms to engage in R\&D has a qualitative similarity to the role that patents play in incentivizing innovation in standard R\&D models.
 In these models, patents temporarily relax competition for the innovating firm by giving them exclusive access to their innovation for a limited period of time in order to incentivize them to invest in the better technology.
 In our model, temporarily relaxing competition in the form of giving firms free periods to learn incentivizes the firm to invest in the better technology.
+
+The second is the nascent literature which studies the economics of data and how data can serve as a barrier to entry in the digital economy \cite{de2020data, hagiu2020data}.
+Our model directly speaks to this literature as we endogenize data-driven network effects by explicitly considering a model where firms are simultaneously solving a machine learning problem while competing against each other for users.
+The results of our model point to the fact that competition dynamics -- that firms compete as they learn over time and that this competition endogenously determines the data observed by firms -- are pertinent to understanding whether or not data can serve as a barrier to entry.
+Indeed, our results are similar to the market tipping results in the literature on indirect network effects and show a channel through which differences in data ownership can result in market tipping.
+However, it is important to note that the channel through which this arises in our model is not simply from differences in the quantity of data possessed by the firms, but also from the quality of the data.
 
@@
 \xhdr{Map of the paper.}
-We survey related work (Section~\ref{sec:related-work}), lay out the model and preliminaries (Section~\ref{sec:model}), and proceed to analyze the different models.
+We survey related work (Section~\ref{sec:related-work}), lay out the model and preliminaries (Section~\ref{sec:model}).
 We start by analyzing the expectation choice model analytically.

### 2020-06-28 journal paper/itcs18paper/sec-random.tex
the latex compiles
updates

--- 2019-06-17 journal paper/itcs18 paper/sec-random.tex
@@
 %The modified algorithm is called \emph{greedy deviation}, with probability parameter $p$.
-For a cleaner comparison between the two algorithms, the modified algorithm does not record rewards received in steps with the ``greedy choice". Parameter $p>0$ is the same for all steps.
+For a cleaner comparison between the two algorithms, the modified algorithm does not record rewards received in steps with the ``greedy choice".
+Parameter $p>0$ is the same for all steps.
+We state the main result and defer the detailed proof to the appendix.
 
@@
 
-\begin{proof}[Proof of Theorem~\ref{thm:random-greedy}]
-  Let $\rew'_1(n)$ denote the Bayesian-expected reward of the ``greedy
-  choice'' after after $n-1$ steps of \alg[1]. Note that
-  $\rew_1(\cdot)$ and $\rew'_1(\cdot)$ are non-decreasing: the former
-  because \alg[1] is monotone and the latter because the ``greedy
-  choice'' is only improved with an increasing set of
-  observations. Therefore, the modified algorithm \alg[2] is monotone
-  by \eqref{eq:random-BIR}.
-
-  By definition of the ``greedy choice,'' $\rew_1(n)\leq \rew'_1(n)$
-  for all steps $n$. Moreover, by Lemma~\ref{lm:DG-rew},
-  \alg[1] has a strictly smaller $\rew(n_0)$ compared to \DynGreedy;
-  so, $\rew_1(n_0)<\rew_2(n_0)$.\sw{fixed inequalities; but this paragraph is still not perfectly clear}
-
-Let $\alg$ denote a copy of \alg[1] that is running ``inside" the modified algorithm \alg[2].
-Let $m_2(t)$ be the number of global rounds before $t$ in which the agent chooses principal $2$ \emph{and} \alg is invoked; in other words, it is the number of agents seen by \alg before global round $t$.
-Let $\mM_{2,t}$ be the agents' posterior distribution for $m_2(t)$.
-
-We claim that in each global round $t\geq n_0$, distribution $\mM_{2,t}$ stochastically dominates distribution $\posteriorN{1}{t}$, and $\PMR_1(t)<\PMR_2(t)$.
-We use induction on $t$.
-The base case $t=n_0$ holds because $\mM_{2,t} = \posteriorN{1}{t}$ (because the two algorithms coincide on the first $n_0-1$ steps), and $\PMR_1(n_0)<\PMR_2(n_0)$ is proved as in \eqref{eq:pf:thm:DG-dominance}, using the fact that $\rew_1(n_0)<\rew_2(n_0)$.
-
-The induction step is proved as follows.
-The induction hypothesis for global round $t-1$ implies that agent $t-1$ is seen by \alg with probability $(1-\eps_0)(1-p)$, which is strictly larger than $\eps_0$, the probability with which this agent is seen by \alg[2].
-Therefore, $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$.
-\begin{align}
-\PMR_1(t)
-  &= \Ex{n\sim \posteriorN{1}{t}} {\rew_1(n+1)} \nonumber \\
-  &\leq \Ex{m\sim \mM_{2,t}} {\rew_1(m+1)}
-    \label{eq:pf:thm:random-greedy-1}\\
-  &< \Ex{m\sim \mM_{2,t}} {(1-p)\cdot \rew_1(m+1) + p\cdot \rew'_1(m+1)}
-    \label{eq:pf:thm:random-greedy-2} \\
-  &= \PMR_2(t). \nonumber
-\end{align}
-Here inequality \eqref{eq:pf:thm:random-greedy-1} holds because $\rew_1(\cdot)$ is monotone and $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$, and inequality \eqref{eq:pf:thm:random-greedy-2} holds because $\rew_1(n_0)<\rew_2(n_0)$ and $\mM_{2,t}(n_0)>0$.%
-\footnote{If $\rew_1(\cdot)$ is strictly increasing, then inequality \eqref{eq:pf:thm:random-greedy-1} is strict, too; this is because $\mM_{2,t}(t-1)>\posteriorN{1}{t}(t-1)$.  }
-\end{proof}
-
 %%% Local Variables:

### 2020-06-28 journal paper/itcs18paper/sec-rational.tex
updates

@@
 Principal 1 can use \DynGreedy, or any other monotone MAB algorithm that coincides with \DynGreedy in the first few steps.
+We state the result and defer the proof to the appendix.
 
@@
 \end{theorem}
-
-\begin{proof}
-The proof re-uses Lemmas~\ref{lm:DG-rew} and~\ref{lm:DG-sudden}, which do not rely on fair tie-breaking.
-
-Because of the biased tie-breaking, for each global round $t$ we have:
-\begin{align}\label{eq:thm:HardMax-biased-PMRtoPr}
-\text{if $\PMR_1(t)\geq \PMR_2(t)$ then $\Pr[i_t=1]>\tfrac12$.}
-\end{align}
-Recall that $i_t$ is the principal chosen in global round $t$.
-
-Let $m_0$ be the first step when \alg[2] deviates from \DynGreedy, or \DynGreedy deviates from \StaticGreedy, whichever comes sooner.
-\asedit{Then \alg[2], \DynGreedy and \StaticGreedy coincide on the first $m_0-1$ steps.
-Moreover, $m_0\leq n_0$ (since \DynGreedy deviates from \StaticGreedy at step $n_0$), so \alg[1] coincides with \DynGreedy on the first $m_0$ steps.
-
-So, $\rew_1(n)=\rew_2(n)$ for each step $n<m_0$, because \alg[1] and \alg[2] coincide on the first $m_0-1$ steps. Moreover, if \alg[2] deviates from \DynGreedy at step $m_0$ then
-    $\rew_1(m_0) > \rew_2(m_0)$ 
-by Lemma~\ref{lm:DG-rew}; else, we trivially have
-    $\rew_1(m_0) = \rew_2(m_0)$.} To summarize:
-\begin{align}\label{eq:thm:HardMax-biased-rew}
-    \rew_1(n)\geq\rew_2(n)\quad \text{for all steps $n\leq m_0$}.
-\end{align}
-
-We claim that $\Pr[i_t=1]>\tfrac12$ for all global rounds $t\leq m_0$.
-We prove this claim using induction on $t$.
-The base case $t=1$ holds by \eqref{eq:thm:HardMax-biased-PMRtoPr} and the fact that in step 1, \DynGreedy chooses the arm with the highest prior mean reward.
-For the induction step, we assume that $\Pr[i_t=1]>\tfrac12$ for all global rounds $t<t_0$, for some $t_0\leq  m_0$.
-It follows that distribution $\posteriorN{1}{t_0}$ stochastically dominates distribution $\posteriorN{2}{t_0}$.%
-\footnote{For random variables $X,Y$ on \R, we say that $X$ \emph{stochastically dominates} $Y$ if $\Pr[X\geq x] \geq \Pr[Y\geq x]$ for any $x\in \R$.}
-Observe that
-\begin{align}\label{eq:thm:HardMax-biased-PMR-aux}
-\PMR_1(t_0)
-  = \Ex{n\sim \posteriorN{1}{t_0}} {\rew_1(n+1)}
-  \geq \Ex{n\sim \posteriorN{2}{t_0}} {\rew_2(n+1)}
-  = \PMR_2(t_0).
-\end{align}
-So the induction step follows by \eqref{eq:thm:HardMax-biased-PMRtoPr}. Claim proved.
-
-Now let us focus on global round $m_0$, and denote $\mN_i = \posteriorN{i}{m_0}$.  By the above claim,
-\begin{align}\label{eq:thm:HardMax-biased-mN}
-\text{$\mN_1$ stochastically dominates $\mN_2$, and moreover
-    $\mN_i(m_0-1)>\mN_i(m_0-1)$.}
-\end{align}
-
-By definition of $m_0$, either (i) \alg[2] deviates from \DynGreedy starting from local step $m_0$, which implies $\rew_1(m_0)> \rew_2(m_0)$ by Lemma~\ref{lm:DG-rew}, or (ii) \DynGreedy deviates from \StaticGreedy starting from local step $m_0$, which implies $\rew_1(m_0)>\rew_1(m_0-1)$ by Lemma~\ref{dgmono}.
-In both cases, using \eqref{eq:thm:HardMax-biased-rew} and \eqref{eq:thm:HardMax-biased-mN}, it follows that the inequality in \eqref{eq:thm:HardMax-biased-PMR-aux} is strict for $t_0=m_0$.
-
-Therefore, agent $m_0$ chooses principal $1$, and by Lemma~\ref{lm:DG-sudden} so do all subsequent agents.
-\end{proof}
 

### 2020-06-28 journal paper/itcs18paper/sec-soft.tex
the latex compiles
updates

--- 2019-06-17 journal paper/itcs18 paper/sec-soft.tex
@@
 \end{proof}
-
-\newcommand{\BReg}{\term{BReg}}
 
@@
 
-The main idea behind our proof is that even though \alg[1] may have a
-slower rate of learning in the beginning, it will gradually catch up
+The main idea behind our proof is that even though \alg[1] may have a slower rate of learning in the beginning, it will gradually catch up
 and surpass \alg[2]. We will describe this process in two phases. In
@@
 that the ``bump'' over $\tfrac12$ is at least on the order of
-$\BIR_2(t)$.
-
-\begin{proof}[Proof of Theorem~\ref{thm:SoftMax-strong}]
-  Let $\beta_1 = \min\{c_0'\delta_0, \beta_0/20\}$ with $\delta_0$
-  defined in~\eqref{eq:SoftMax-smooth}.  Recall each agent chooses
-  \alg[1] with probability at least $\respF(-1)= \eps_0$.  By
-  condition \eqref{eq:so-mild} and \eqref{eq:SoftMax-assn-strong},
-  there exists some sufficiently large $T_1$ such that for any
-  $t\geq T_1$, $\BIR_1(\eps_0 T_1/2) \leq \beta_1/c_0'$ and
-  $\BIR_2(t) > e^{-\eps_0 t/12}$. Moreover, for any $t\geq T_1$, we
-  know $\E[n_1(t+1)] \geq \eps_0\,t $, and by the Chernoff Bounds
-  (Theorem~\ref{thm:chernoff}), we have $n_1(t+1) \geq \eps_0 t/2$
-  holds with probability at least $1 - q_1(t)$ with
-  $q_1(t) = \exp(-\eps_0 t/12) < \BIR_2(t)$. It follows that for any $t\geq T_1$,
-\begin{align*}
-  \PMR_2(t) - \PMR_1(t) &= \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}{\BIR_1(m_1+ 1) - \BIR_2(m_2+1)} \\
-                        &\leq q_1(t)  + \Ex{m_1\sim \posteriorN{1}{t}}{\BIR_1(m_1+ 1)\mid m_1 \geq \eps_0 t/2 - 1 } - \BIR_2(t)\\
-                        &\leq \BIR_1(\eps_0 T_1/2) \leq \beta_1/c_0'
-\end{align*}
-% where the second inequality follows from~\eqref{eq:random-assn}.
-Since the response function $\respF$ is $c_0'$-Lipschitz in the
-neighborhood of $[-\delta_0, \delta_0]$, each agent after round $T_1$
-will choose \alg[1] with probability at least
-\[
-  p_t \geq \frac{1}{2} - c_0'\left(\PMR_2(t) - \PMR_1(t)\right) \geq
-  \frac{1}{2} - \beta_1.
-\]
-
-Next, we will show that there exists a sufficiently large $T_2$ such
-that for any $t\geq T_1 + T_2$, with high probability
-$n_1(t) > \max\{n_0, (1 - \beta_0)n_2(t)\}$, where $n_0$ is defined
-in~\eqref{eq:SoftMax-better}. %  Let us first lower bound the number of
-% agents received by \alg[1] after some number of rounds $t = T_1 + T'$
-% for any $T' \geq T_1$. 
-Fix any $t \geq T_1 + T_2$. 
-Since each agent chooses \alg[1] with probability at least
-$1/2 - \beta_1$, by Chernoff Bounds (Theorem~\ref{thm:chernoff}) we
-have with probability at least $1 - q_2(t)$ that the number of agents
-that choose \alg[1] is at least $\beta_0(1/2 - \beta_1)t/5$, where the
-function
-$$
-q_2(x) = \exp\left( \frac{-(1/2 - \beta_1)(1 - \beta_0/5)^2x}{3} \right).
-$$
-Note that the number of agents received by \alg[2] is at most
-$T_1 + (1/2 + \beta_1)t + (1/2 - \beta_1)(1 - \beta_0/5)t$.
-
-Then as long as
-% $T_2 \geq \max\left\{\frac{3T_1}{(1 - \beta_0)}, 8 n_0\right\}$, we
-% can guarantee that for any $t \geq T_1 + T_2$,
-% $n_1(t) > n_2(t) (1 - \beta_0)$ and $n_1(t) > n_0$ with probability at
-% least $1 - q_2(t)$.
-$T_2 \geq \frac{5T_1}{\beta_0}$, we can guarantee that
-$n_1(t) > n_2(t) (1 - \beta_0)$ and $n_1(t) > n_0$ with probability at
-least $1 - q_2(t)$ for any $t \geq T_1 + T_2$.
-% for any $t\geq T_1 + T_2$, as long as $T_2 \geq .$ Moreover, to
-% guarantee that , we will just need $T_2 \geq$.  Finally, we will
-% argue that in each round $t \geq T_1 + T_2$, we recover the
-% guarantee in \eqref{eq:thm:SoftMax-strong}.
-% % \[
-% %   \Pr[i_t = 1] \geq \frac{1}{2} + \frac{c_0\alpha_0\BIR_2(t)}{4}.
-% % \]
-Note that the weak BIR-dominance condition
-in~\eqref{eq:SoftMax-better} implies that for any $t\geq T_1 + T_2$
-with probability at least $1 - q_2(t)$,
-\[
-  \BIR_1(n_1(t)) < (1- \alpha_0)\BIR_2(n_2(t)).
-\]
-It follows that for any $t\geq T_1 + T_2$,
-\begin{align*}
-  \PMR_1(t) - \PMR_2(t) &= \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}{\BIR_2(m_2+ 1) - \BIR_1(m_1+1)} \\
-                        &\geq (1 - q_2(t))\alpha_0 \BIR_2(t) - q_2(t)\\
-                        &\geq \alpha_0 \BIR_2(t)/4
-\end{align*}
-where the last inequality holds as long as
-$q_2(t) \leq \alpha_0\BIR_2(t)/4$, and is implied by the condition
-in~\eqref{eq:SoftMax-assn-strong} as long as $T_2$ is sufficiently
-large. Hence, by the definition of our \SoftMaxRandom response
-function and assumption in~\eqref{eq:SoftMax-smooth}, we have
-\[
-  \Pr[i_t = 1] \geq \frac{1}{2} + \frac{c_0\alpha_0\BIR_2(t)}{4}. \qedhere
-\]
-\end{proof}
+$\BIR_2(t)$. The detailed proof is deferred to the appendix.
 
@@
 and is omitted. The following is just a direct consequence of
-Theorem~\ref{thm:SoftMax-weak} with this general condition.\sw{added}
+Theorem~\ref{thm:SoftMax-weak} with this general condition
 

### 2020-06-28 journal paper/journal_main.tex
add updated poster and edits in journal sec
updates

@@
 
+In the appendix, we provide background on multi-armed bandits as well as several discussions and proofs omitted from the main text.
+Furthermore, we provide plots and tables for our experiments, which were omitted from the main text.
+In all cases, the plots and tables here are in line with those in the main text, and lead to similar qualitative conclusions.
+
 \section{Background on multi-armed bandits}
@@
 \subfile{itcs18paper/app-perturb}
+
+\section{Omitted Proofs}
+\label{app:omitted_proofs}
+\subfile{content/omitted_proofs}
+
 \subfile{ec19paper/content/appendix_for_one_version}

### 2020-06-30 journal paper/content/sec-intro.tex
Started going through the intro.

@@
 Learning from interactions with users is ubiquitous in modern customer-facing systems, from product recommendations to web search to content selection to fine-tuning user interfaces.
-These interactions are essential for the development of high quality products, but, in order to learn what products are high quality, firms need to implement systems that engage in \emph{exploration}: deploying potentially suboptimal products in order to acquire new information to make more informed product decisions tomorrow.
+Many systems purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
+Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
+
+%~\cite{KohaviAB-2015,KohaviLSH09}
 
 In this paper, we initiate a study of the interplay between \exploration and \competition.
-The fundamental tension is that while exploration allows for the deployment of high quality products in the long-run, it leads to poor user experiences in the short-run.
-Exploration can therefore harm the perception users have of a system, which may lead them to instead utilize a competing system.
-This would leave the system with lower market share and therefore less users to learn from relative to its competitors and allow its competitors to make further progress on their learning problem.
+Systems that engage in exploration typically need to compete against one another; most importantly, they compete for users.
+This creates a tension:
+%between \exploration and \competition. 
+while exploration may be essential for improving the service tomorrow, it may degrade quality and make users leave \emph{today}, in which case there will be less users to learn from.
+This may further degrade the platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+Taken to the extreme, such dynamics may create a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
 Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.
+
+\ascomment{stopped here.}
 

### 2020-07-01 journal paper/journal_main.tex
Switched to long conf names.
Switched to style files in the main folder.
un-suppressed color-edits.
update format + add ToC

@@
-\documentclass[format=acmsmall, review=false]{acmart}
+\documentclass[12pt]{article}
 
-\usepackage{ec19paper/acm-ec-19}
-\usepackage{itcs18paper/slivkins-setup,itcs18paper/slivkins-theorems}
+\usepackage{amssymb,amsmath,amsfonts,
+float,geometry,ulem,graphicx,caption,color,
+setspace,sectsty,comment,footmisc,caption,natbib,
+pdflscape,subcaption,array,hyperref, subfiles,adjustbox,threeparttable
+}
 \usepackage{tikz} % SW: Added for plots
@@
 \usepackage{amsmath,amssymb,amsfonts}
-%\usepackage{natbib} % cannot use with AAAI! A.S.
+\usepackage{natbib} % cannot use with AAAI! A.S.
 
@@
 \usepackage{sgame, tikz} % Game theory packages
-\usepackage{caption}
 \usepackage{algorithm,algpseudocode}
@@
 \captionsetup{font=footnotesize}
+
+\usepackage{slivkins-setup,slivkins-theorems}
 
@@
 \newcommand{\Exploration}{\termTXT{Exploration}}
+
 % a very useful package for edits and comments, from David Kempe (USC)
-\usepackage[suppress]{ec19paper/color-edits}
+\usepackage{color-edits}
 %\usepackage[suppress]{color-edits}  % use this to suppress the package
@@
 
+\normalem
+\raggedbottom
+% \usepackage[hidelinks]{hyperref}
+\onehalfspacing
+
+\geometry{left=1.0in,right=1.0in,top=1.0in,bottom=1.0in}
+
 \usepackage{booktabs} % For formal tables
 
-% Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
+\begin{document}
 
-\begin{document}
-\title[Competing Bandits: \\ The Perils of Exploration under Competition]
-{Competing Bandits: \\ The Perils of Exploration under Competition}
-
- \author{Guy Aridor}
- \affiliation{ \institution{Columbia University}}
- \author{Yishay Mansour}
- \affiliation{\institution{Tel Aviv University and Google}}
- \author{Aleksandrs Slivkins}
- \affiliation{\institution{Microsoft Research New York City}}
- \author{Zhiwei Steven Wu}
- \affiliation{\institution{University of Minnesota - Twin Cities}}
-
- \authorsaddresses{%
-  Authors' addresses: 
-  Guy Aridor, Columbia University, New York, NY, USA. g.aridor@columbia.edu; 
-  Yishay Mansour, Tel Aviv University, Tel Aviv, Israel. mansour.yishay@gmail.com;
-  Aleksandrs Slivkins, Microsoft Research, New York, NY, USA. slivkins@microsoft.com; 
-  Zhiwei Steven Wu, University of Minnesota - Twin Cities, Minneapolis, MN, USA. zsw@umn.edu}
-
+\begin{titlepage}
+\title{Competing Bandits: \\ The Perils of Exploration under Competition\thanks{This paper is a merged version of \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
+\author{Guy Aridor\footnote{Columbia University, Department of Economics.
+Email: g.aridor@columbia.edu} \rule{0.2in}{0pt} Yishay Mansour\footnote{Google and Tel Aviv University, Department of Computer Science.
+Email: mansour.yishay@gmail.com;} \rule{0.2in}{0pt}   Aleksandrs Slivkins\footnote{Microsoft Research - New York City.
+Email: slivkins@microsoft.com} \rule{0.2in}{0pt}   Zhiwei Steven Wu\footnote{University of Minnesota - Twin Cities, Department of Computer Science.
+Email: zsw@umn.edu.}}
+\date{\today}
+\maketitle
 \begin{abstract}
@@
 
-In our model, we consider competition between two firms facing the same multi-armed bandit instance who simultaneously choose a bandit algorithm.
+We consider competition between two firms facing the same multi-armed bandit instance who simultaneously choose a bandit algorithm.
 Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance if and only if it is chosen.
@@
 We find that stark competition induces firms to commit to a ``greedy" algorithm that leads to low consumer welfare.
-However, we find that in order to incentivize ``better" exploration strategies a firm needs to have some ``free" consumers that visit them without having to be incentivized to do so.
-We investigate this effect via noise in consumer decision-making and asymmetries in timing of entry of firms into the market.
+However, we find that weakening that lead to higher consumer welfare.
+competition by providing firms with some ``free" consumers incentivizes better exploration strategies.
 Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets and are closely related to the ``competition vs.
 innovation" relationship, a well-studied theme in economics.
+
+\vspace{0.2in}
+
+\bigskip
 \end{abstract}
+\setcounter{page}{0}
+\thispagestyle{empty}
+\end{titlepage}
+\pagebreak \newpage
+
+\tableofcontents
 
@@
 %
-\maketitle
-\keywords{Multi-armed bandits,Exploration,Competition,Competition vs Innovation}
-
 \section{Introduction}
@@
 \newpage
-\bibliographystyle{acm}
-\bibliography{bib-ML,refs,bib-abbrv-short,bib-bandits,bib-AGT,bib-slivkins, bib-random}
+\bibliographystyle{plain}
+\bibliography{bib-abbrv,bib-ML,refs,bib-bandits,bib-AGT,bib-slivkins, bib-random}
 

### 2020-07-01 journal paper/journal_main.tex
fix incomplete sentence in abstract

@@
 We find that stark competition induces firms to commit to a ``greedy" algorithm that leads to low consumer welfare.
-However, we find that weakening that lead to higher consumer welfare.
-competition by providing firms with some ``free" consumers incentivizes better exploration strategies.
+However, we find that weakening competition by providing firms with some ``free" consumers incentivizes better exploration strategies and increases consumer welfare.
 Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets and are closely related to the ``competition vs.

### 2020-07-01 journal paper/journal_main.tex
change font.
Revised formatting a bit.

@@
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 
-\usepackage{amssymb,amsmath,amsfonts,
-float,geometry,ulem,graphicx,caption,color,
-setspace,sectsty,comment,footmisc,caption,natbib,
-pdflscape,subcaption,array,hyperref, subfiles,adjustbox,threeparttable
+\usepackage{ulem,color,
+setspace,sectsty,comment,footmisc,caption,
+pdflscape,subcaption,array,hyperref,adjustbox,threeparttable
 }
-\usepackage{tikz} % SW: Added for plots
+
+\usepackage{fullpage}
+%\usepackage{geometry}
+%\geometry{left=1.0in,right=1.0in,top=1.0in,bottom=1.0in}
+
 \usepackage{subfiles}
+\usepackage{libertine}
 \usepackage[toc,page]{appendix}
@@
 \usepackage{microtype}      % microtypography
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{natbib} % cannot use with AAAI! A.S.
+\usepackage{amsmath,amssymb,amsfonts,amsthm}
+\usepackage[round]{natbib} % cannot use with AAAI! A.S.
 
-\usepackage{amsthm} % The amsthm package provides extended theorem environments
 \usepackage{float}
@@
  \usepackage{graphicx}
-\theoremstyle{definition}
+
+\usepackage{booktabs} % For formal tables
+
+%\theoremstyle{definition}
 %\newtheorem{definition}{Definition}
@@
 % e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
+
 \setlength{\tabcolsep}{8pt} % Default value: 6pt
@@
 
-%drt24 hacks
-% Letter paper
-
-% Hack to try to make acmart work with biblatex: https://tex.stackexchange.com/questions/37076/is-it-possible-to-load-biblatex-with-a-class-that-has-already-loaded-natbib
-\let\citename\relax
-%\RequirePackage[abbreviate=true, dateabbrev=true, isbn=true, doi=true, urldate=comp, url=true, maxbibnames=9, backref=false, backend=biber, style=ACM-Reference-Format, language=american]{biblatex}
-
-\normalem
-\raggedbottom
-% \usepackage[hidelinks]{hyperref}
-\onehalfspacing
-
-\geometry{left=1.0in,right=1.0in,top=1.0in,bottom=1.0in}
-
-\usepackage{booktabs} % For formal tables
-
 \begin{document}
 
-\begin{titlepage}
-\title{Competing Bandits: \\ The Perils of Exploration under Competition\thanks{This paper is a merged version of \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
-\author{Guy Aridor\footnote{Columbia University, Department of Economics.
-Email: g.aridor@columbia.edu} \rule{0.2in}{0pt} Yishay Mansour\footnote{Google and Tel Aviv University, Department of Computer Science.
-Email: mansour.yishay@gmail.com;} \rule{0.2in}{0pt}   Aleksandrs Slivkins\footnote{Microsoft Research - New York City.
-Email: slivkins@microsoft.com} \rule{0.2in}{0pt}   Zhiwei Steven Wu\footnote{University of Minnesota - Twin Cities, Department of Computer Science.
-Email: zsw@umn.edu.}}
-\date{\today}
+%\begin{titlepage}
+
+\title{Competing Bandits: 
+\\ The Perils of Exploration under Competition
+\thanks{This is a merge of two conference papers,
+\citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, 
+with a unified presentation.}}
+% \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
+
+\author{Guy Aridor
+\footnote{Columbia University, Department of Economics. Email: g.aridor@columbia.edu} 
+\and
+%\rule{0.2in}{0pt}
+Yishay Mansour
+\footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com;} 
+\and
+%\rule{0.2in}{0pt}   
+Aleksandrs Slivkins\footnote{Microsoft Research New York City. Email: slivkins@microsoft.com} 
+\and
+%\rule{0.2in}{0pt}   
+Zhiwei Steven Wu
+\footnote{University of Minnesota - Twin Cities, Department of Computer Science. Email: zsw@umn.edu.\newline
+The research has been done when Z.S. Wu has been an intern and a postdoc at Microsoft Research NYC.
+}}
+\date{First version: July 2020}
 \maketitle
@@
 \end{abstract}
-\setcounter{page}{0}
-\thispagestyle{empty}
-\end{titlepage}
-\pagebreak \newpage
+%\setcounter{page}{0}
+%\thispagestyle{empty}
+%\end{titlepage}
 
+\newpage
+\begin{small}
 \tableofcontents
+\end{small}
+\newpage
 
@@
 \newpage
-\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 \bibliography{bib-abbrv,bib-ML,refs,bib-bandits,bib-AGT,bib-slivkins, bib-random}

### 2020-07-01 journal paper/journal_main.tex
A few more edits re formatting.
Restored \\ to the title.
Edit in the title note.

@@
 
-\title{Competing Bandits: 
-\\ The Perils of Exploration under Competition
-\thanks{This is a merge of two conference papers,
+\title{Competing Bandits:\\ 
+The Perils of Exploration under Competition%
+\thanks{This is a merged and final version of two conference papers,
 \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, 
-with a unified presentation.}}
+with a unified and streamlined presentation.}}
 % \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
 
-\author{Guy Aridor
+\author{Guy Aridor%
 \footnote{Columbia University, Department of Economics. Email: g.aridor@columbia.edu} 
@@
 %\rule{0.2in}{0pt}
-Yishay Mansour
+Yishay Mansour%
 \footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com;} 
@@
 %\rule{0.2in}{0pt}   
-Aleksandrs Slivkins\footnote{Microsoft Research New York City. Email: slivkins@microsoft.com} 
+Aleksandrs Slivkins%
+\footnote{Microsoft Research New York City. Email: slivkins@microsoft.com} 
 \and
 %\rule{0.2in}{0pt}   
-Zhiwei Steven Wu
+Zhiwei Steven Wu%
 \footnote{University of Minnesota - Twin Cities, Department of Computer Science. Email: zsw@umn.edu.\newline
@@
 
-\newpage
+\clearpage
+%\addcontentsline{toc}{section}{References}
 \bibliographystyle{plainnat}
@@
 
+\clearpage
 \begin{appendices}

### 2020-07-02 journal paper/content/sec-intro.tex
Revised through the end of "our model".

@@
-Learning from interactions with users is ubiquitous in modern customer-facing systems, from product recommendations to web search to content selection to fine-tuning user interfaces.
-Many systems purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
+Learning from interactions with users is ubiquitous in modern customer-facing platforms, from product recommendations to web search to content selection to fine-tuning user interfaces.
+Many platforms purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
 Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
@@
 In this paper, we initiate a study of the interplay between \exploration and \competition.
-Systems that engage in exploration typically need to compete against one another; most importantly, they compete for users.
+Platforms that engage in exploration typically need to compete against one another; most importantly, they compete for users.
 This creates a tension:
@@
 
-\ascomment{stopped here.}
+The main high-level question that we focus on in this paper is: 
+\begin{align*}
+\textbf{How does competition incentivize the adoption of better exploration algorithms?}
+\end{align*}
 
-The main high-level question that we focus on in this paper is whether competition between systems incentives the adoption of better exploration algorithms.
-This translates into a number of more concrete questions.
-While it is commonly assumed that ``better" technology always helps, is this so for our setting?
-Does increasing competition lead to higher consumer welfare?
-To what extent is there a ``data feedback loop" where one firm having more data leads to that firm attracting more users which leads to that firm having more data, etc.?\footnote{This is a fundamental question which is part of a larger policy discussion around whether data can serve as an indirect network effect and lead to similar ``market tipping" results as is standard in the literature on competition in markets with network effects (see \cite{jullien2019economics} for a policy oriented discussion of this).}
+\noindent This translates into a number of more concrete questions.
+While it is commonly assumed that better technology always helps, is this so under competition?
+Does increased competition lead to higher consumer welfare?
+How significant are ``data feedback loops", where one firm having more data leads to that firm attracting more users, which leads to that firm having more data, and so forth?%
+\footnote{This question is part of a larger policy discussion on whether data serves as an indirect ``network effect" \citep[\eg see][]{jullien2019economics}.
+In particular, can data lead to ``market tipping", as in the literature on competition with network effects.}
+Finding formalizations that admit meaningful answers is a major part of the overall challenge.
+
+To answer these questions, we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
+The choice of a particular technology (exploration algorithm) is no longer an abstract, static choice with a predetermined outcome for the platform.
+Instead, we model the algorithms explicitly, and investigate how they play out in competition over an extended period of time.
+
+%\footnote{This is a fundamental question which is part of a larger policy discussion around whether data can serve as an indirect network effect and lead to similar ``market tipping" results as is standard in the literature on competition in markets with network effects (see \cite{jullien2019economics} for a policy oriented discussion of this).}
 
@@
 
-\xhdr{Our model.} We investigate these questions with a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
+\subsection{Our model}
+\label{sec:intro-model}
+
+%We investigate these questions with 
+\xhdr{Competition game.} We consider a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
 We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
@@
 
-\item Each principal faces a very basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
+\item Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
 \emph{arms}) and receives a reward drawn independently from a fixed distribution specific to this action.
+The reward distributions are initially unknown (but can be estimated over time from the data).
 
-\item Principals simultaneously announce their learning algorithms before round $1$, and cannot change them afterwards.
-There is a common Bayesian prior on the rewards (but the realized reward distributions are not observed by the principals or the agents).
-Each principal only observes agents that chose him.
-We consider two variants of the baseline model and the information set for the agent differs between the two:
-\begin{enumerate}
-\item In the \textit{expectation choice variant}, agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms.
-\item In the \textit{reputation choice variant}, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal.
-\end{enumerate}
+\item Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
+    Each principal only observes agents that chose this principal.
+
+\item Agents do not directly observe other agents' rewards, but may have access to each platforms' reputation score (more on this in    Section~\ref{sec:intro-discussion}).
 \end{itemize}
 
-\xhdr{Main Findings.}
+\xhdr{Technology: multi-armed bandit algorithms.}
+While some bandit algorithms are better than others, these comparisons are rather subtle.
+Indeed, some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+
+We distinguish three classes of bandit algorithms, based on 
+the prevalent intuition in the area.
+The distinction concerns the way in which they resolve the fundamental tradeoff between  exploration and \emph{exploitation} (making optimal decisions using the available data).
+Going from more primitive to more sophisticated:
+
+\begin{itemize}
+\item \emph{Greedy algorithms} that strive to maximize the reward for the next round given the available information. Thus, they always ``exploit" and never explicitly ``explore".
+
+\item \emph{Exploration-separating algorithms}
+ that separate exploration and exploitation: each round is dedicated to one and completely ignores the other.
+
+\item \emph{Adaptive exploration} algorithms that combine exploration and exploitation, and sway the exploration choices towards more promising alternatives.
+\end{itemize}
+
+\noindent We are mainly interested in qualitative differences between the three classes.
+For numerical simulations, we pick one representative algorithm from each class.
+For theoretical results, we allow arbitrary algorithms and focus on asymptotic differences in the algorithms' performance.
+
+\ascomment{stopped here.}
+
+\subsection{Our Results}
+\label{sec:intro-results}
+
 We find that competition induces firms to commit to a greedy (myopic) algorithm in equilibrium that does no purposeful exploration.
@@
 
+\subsection{Discussion}
+\label{sec:intro-discussion}
+
 \begin{figure}
@@
 \end{figure}
+
 \xhdr{Economic interpretation.} Our model speaks to two distinct strands of the economics literature.
@@
 
-\xhdr{Discussion}
+\xhdr{Model variants.}
+We consider two model variants, which determine agents' decision rule and their information sets: 
+\ascomment{copied-and-pasted, will edit later.}
+
+\begin{description}
+\item[\TheoryModel] 
+There is a common Bayesian prior on the reward distributions.
+Agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms.
+\item[\ExptsModel] agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal.
+\end{description}
+
 We consider two separate variants of the model in order to provide a more thorough investigation of the tension between exploration and competition.

### 2020-07-03 journal paper/content/sec-intro.tex
Some more edits.

@@
 Each piece can get rather complicated in isolation, let alone jointly, so we strive for simplicity.
-Thus, the basic model is as follows:
+Thus, the key features of our model are as follows:
 
@@
 the prevalent intuition in the area.
-The distinction concerns the way in which they resolve the fundamental tradeoff between  exploration and \emph{exploitation} (making optimal decisions using the available data).
-Going from more primitive to more sophisticated:
+The distinction concerns the way in which they resolve the fundamental tradeoff between  exploration and \emph{exploitation} (making optimal myopic decisions using the available data).
+Going from more primitive to more sophisticated, these three classes are as follows:
 
@@
 \item \emph{Exploration-separating algorithms}
- that separate exploration and exploitation: each round is dedicated to one and completely ignores the other.
+ that separate exploration and exploitation: essentially, each round is dedicated to one and completely ignores the other.
 
-\item \emph{Adaptive exploration} algorithms that combine exploration and exploitation, and sway the exploration choices towards more promising alternatives.
+\item \emph{Adaptive-exploration} algorithms that combine exploration and exploitation, and gradually sway the exploration choices towards more promising alternatives.
 \end{itemize}
 
-\noindent We are mainly interested in qualitative differences between the three classes.
-For numerical simulations, we pick one representative algorithm from each class.
-For theoretical results, we allow arbitrary algorithms and focus on asymptotic differences in the algorithms' performance.
+\noindent Greedy algorithms are terrible for a wide variety of problem instances, precisely because they never explore.
+Adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances.
+Exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances, and fail to improve for the ``easy" ones.
 
@@
 
-We find that competition induces firms to commit to a greedy (myopic) algorithm in equilibrium that does no purposeful exploration.
+We offer a mix of theoretical results and numerical simulations.
+We are mainly interested in qualitative differences between the three classes of algorithms in Section~\ref{sec:intro-model}.
+For numerical simulations, we pick one representative algorithm from each class.
+For theoretical results, we allow arbitrary algorithms and focus on asymptotic differences in the algorithms' performance.
+
+\xhdr{Main results.}
+We start with a basic version in which the principals are symmetric (\ie do not have an inherent advantage over one another) and the agents are utility-maximizers.
+We find that competition induces firms to choose the greedy algorithm.
+
+commit to a greedy (myopic) algorithm in equilibrium that does no purposeful exploration.
 This leads to low consumer welfare since the greedy algorithm is known to be dramatically bad in many important cases of multi-armed bandits.

### 2020-07-04 journal paper/content/sec-intro.tex
Full draft up to and including "theory results".

@@
 The main high-level question that we focus on in this paper is: 
-\begin{align*}
+\begin{align}\label{eq:main-Q}
 \textbf{How does competition incentivize the adoption of better exploration algorithms?}
-\end{align*}
+\end{align}
 
@@
 
-\subsection{Our model}
+\subsection{Our model: competing bandits}
 \label{sec:intro-model}
@@
 \item Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
-    Each principal only observes agents that chose this principal.
-
-\item Agents do not directly observe other agents' rewards, but may have access to each platforms' reputation score (more on this in    Section~\ref{sec:intro-discussion}).
+    Each principal only observes agents that chose this principal. Agents have little or no information about other agents' choices and rewards.
+%, but may have access to each platforms' reputation score (more on this in    Section~\ref{sec:intro-discussion}).
 \end{itemize}
 
+\noindent We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
+
+%We consider two model variants, which determine agents' decision rule and their information sets.
+In \emTheoryModel, there is a common Bayesian prior on the reward distributions.
+Agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms.
+In \emExptsModel, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal.
+The former variant used for the theoretical results, the latter for simulations.
+
 \xhdr{Technology: multi-armed bandit algorithms.}
-While some bandit algorithms are better than others, these comparisons are rather subtle.
-Indeed, some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+To flesh out \eqref{eq:main-Q}, let us elaborate what we mean by `better' bandit algorithms.
+In general, comparisons between bandit algorithms are rather subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
 
@@
 \noindent Greedy algorithms are terrible for a wide variety of problem instances, precisely because they never explore.
+Exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances.
 Adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances.
-Exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances, and fail to improve for the ``easy" ones.
-
-\ascomment{stopped here.}
 
@@
 For theoretical results, we allow arbitrary algorithms and focus on asymptotic differences in the algorithms' performance.
+
+\xhdr{Theoretical results.} 
+We endow agents with Bayesian rationality, arguably the most natural modeling approach for a theoretical investigation.
+We consider a `bare-bones' variant of Bayesian rationality: agents have a common Bayesian prior on reward distributions, and know the principals' algorithms.
+For simplicity, agents do not receive any information about previous agents' choices and rewards.
+Each agent knows the round (s)he arrives in, and computes principals' Bayesian-expected rewards for this round.
+
+Our results depend crucially on agents' decision rule to choose among the two principals.
+The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
+We find that \HardMax is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
+Further, we show that \HardMax is very sensitive to tie-breaking: if the tie-breaking is probabilistically biased in favor of one principal, then this principal has a simple ``winning strategy" no matter what the other principal does.
+
+Next, we dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
+We call this model \HardMaxRandom.
+We find that better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
+However, there is a substantial caveat: one can defeat any algorithm by interleaving it with the greedy algorithm.
+This has two undesirable corollaries: a better algorithm may sometimes lose in competition, and a pure Nash equilibrium typically does not exist.
+
+We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
+For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
+This is the most technical result of the paper.
+The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm would attract slightly more).
+
+All results extend to a much more general version of the multi-armed bandit problem in which the principal may observe additional feedback before and/or after each decision, as long as the feedback distribution does not change over time.
+In most results, principal's utility may depend on both the market share and agents' rewards.
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
+      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
+      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
+      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
+ \end{tikzpicture}
+
+\caption{Inverted-U relationship between competitiveness and algorithms.}
+\label{fig:inverted-U}
+\end{center}
+\end{figure}
+
+\xhdr{Economic interpretation: the inverted-U relationship.}
+Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the inverted-U relationship between competition and innovation, as in Figure~\ref{fig:inverted-U}.%
+\footnote{Here, ``innovation" refers to adoption of a better technology at a substantial R\&D expense to a given firm.
+It is not salient whether similar ideas and/or technologies already exist elsewhere.
+Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
+This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
+
+Our decision rules differ in terms of rationality: from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
+The same distinctions also control the severity of competition between the principals: from cut-throat competition with \HardMax to a more relaxed competition with \HardMaxRandom, to an even more relaxed competition with \SoftMaxRandom.
+Indeed, with \HardMax you lose all customers as soon as you fall behind in performance, with \HardMaxRandom you get some small market share no matter what, and with \SoftMaxRandom you are further guaranteed a market share close to $\tfrac12$ as long as your performance is not much worse than the competition.
+The uniform choice among principals corresponds to no rationality and no competition.
+
+We identify the inverted-U relationship driven by the rationality/competitiveness distinctions outlined above: from \HardMax to \HardMaxRandom to \SoftMaxRandom to \Uniform.
+We also find another, technically different inverted-U relationship which zeroes in on the \HardMaxRandom model.
+We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
+
+These inverted-U relationships are driven by different aspects in our model than the ones in the existing literature in economics.
+The latter focuses on the tradeoff between the R\&D costs and the benefits that the improved technology provides in the competition.
+In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
+
+\asedit{Stopped here.}
 
@@
 
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
-      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
-      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
-      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
- \end{tikzpicture}
-
-\caption{Inverted-U relationship between competitiveness and algorithms.}
-\label{fig:inverted-U}
-\end{center}
-\end{figure}
-
 \xhdr{Economic interpretation.} Our model speaks to two distinct strands of the economics literature.
@@
 \xhdr{Model variants.}
-We consider two model variants, which determine agents' decision rule and their information sets: 
-\ascomment{copied-and-pasted, will edit later.}
-
-\begin{description}
-\item[\TheoryModel] 
-There is a common Bayesian prior on the reward distributions.
-Agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms.
-\item[\ExptsModel] agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal.
-\end{description}
-
 We consider two separate variants of the model in order to provide a more thorough investigation of the tension between exploration and competition.

### 2020-07-06 journal paper/content/sec-intro.tex
Done with "our results".

@@
 Does increased competition lead to higher consumer welfare?
-How significant are ``data feedback loops", where one firm having more data leads to that firm attracting more users, which leads to that firm having more data, and so forth?%
-\footnote{This question is part of a larger policy discussion on whether data serves as an indirect ``network effect" \citep[\eg see][]{jullien2019economics}.
-In particular, can data lead to ``market tipping", as in the literature on competition with network effects.}
+How significant are ``data feedback loops" --- when more data leads to more users, which leads to even more data, etc.
+--- and how they relate to the anti-trust considerations?
 Finding formalizations that admit meaningful answers is a major part of the overall challenge.
@@
 \item Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
-    Each principal only observes agents that chose this principal. Agents have little or no information about other agents' choices and rewards.
+Each principal's objective is to maximize its market share (the fraction of users choosing this principal).
+Each principal only observes agents that chose this principal.
 %, but may have access to each platforms' reputation score (more on this in    Section~\ref{sec:intro-discussion}).
@@
 \noindent We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
+In all variants, agents have little or no information about other agents' choices and rewards.
 
@@
 
-\noindent Greedy algorithms are terrible for a wide variety of problem instances, precisely because they never explore.
+In isolation, \ie in the absence of competition, these classes are fairly well-understood.
+Greedy algorithms are terrible for a wide variety of problem instances, precisely because they never explore.
 Exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances.
 Adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances.
+Generally,  ``better" algorithms are better in the long run, but could be worse initially.
 
@@
 \xhdr{Theoretical results.}
-We endow agents with Bayesian rationality, arguably the most natural modeling approach for a theoretical investigation.
-We consider a `bare-bones' variant of Bayesian rationality: agents have a common Bayesian prior on reward distributions, and know the principals' algorithms.
-For simplicity, agents do not receive any information about previous agents' choices and rewards.
-Each agent knows the round (s)he arrives in, and computes principals' Bayesian-expected rewards for this round.
+%We endow agents with Bayesian rationality, a common modeling approach for a theoretical investigation. 
+We consider a basic Bayesian model.
+We posit that agents have a common Bayesian prior on reward distributions, and know the principals' algorithms.
+For simplicity, agents do not receive any information about the previous agents' choices and rewards.
+Each agent knows the round (s)he arrives in, computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
+We refer to this variant as \emTheoryModel.
 
-Our results depend crucially on agents' decision rule to choose among the two principals.
-The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
+Our results depend crucially on the agents' decision rule. 
+\begin{itemize}
+
+\item The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
 We find that \HardMax is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
@@
 
-Next, we dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
+\item We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
 We call this model \HardMaxRandom.
@@
 
-We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
+\item We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
 For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
@@
 The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm would attract slightly more).
+\end{itemize}
 
@@
 Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
-This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
+This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{Aghion-QJE05,Vives-08}.
 
@@
 
-\asedit{Stopped here.}
+\xhdr{Numerical simulations.}
+We consider a basic frequentist model.
+We posit that the agents observe some performance signals, and the agents base their decisions on the performance signals alone, without invoking any prior knowledge or beliefs.
+The performance signals are abstracted and aggregated as a scalar \emph{reputation score} for each principal, modeled as a sliding window average of its rewards.
+Thus, agents' decision rule depends only on the two reputation scores.
+We refer to this variant as \emExptsModel.%
+%\footnote{In comparison, the theoretical results focus on another extreme, with Bayesian rationality and no performance signals.}
+
+We refine and expand the theoretical results in several ways.
+
+\begin{itemize}
+\item We compare \HardMax and \HardMaxRandom decision rules.
+We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.
+As predicted by the theory, better algorithms prevail if the expected number of ``random" users is sufficiently large.
+However, this effect is negligible for smaller parameter values.
+
+%\footnote{\asedit{Reputation scores already introduce some noise into users' choices.
+However, the amount of noise due to this channel is typically small, both in our simulations and in practice, because reputation signals average over many datapoints.}}
+
+\item We investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
+(We focus on the \HardMax decision rule.) We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous arrival).
+Consumer welfare is higher under early/late arrival than under simultaneous entry.
+We frame these results in terms inverted-U relationships.
+
+%\footnote{\asedit{We consider the ``permanent monopoly" scenario for comparison only, without presenting any findings.
+We just assume that a monopolist chooses the greedy algorithm, because it is easier to deploy in practice.
+Implicitly, users have no ``outside option": the service provided is an improvement over not having it (and therefore the monopolist is not incentivized to deploy better learning algorithms).
+This is plausible with free ad-supported platforms such as Yelp or Google.}}
+
+\item We investigate the algorithms' performance ``in isolation", \ie without competition.
+We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.
+We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes \emph{not} a good predictor for the outcomes under competition.
+
+\item We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}).
+We run additional experiments so as to isolate and compare these two effects.
+We find that either effect alone leads to a significant advantage under competition.
+The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.
+Finally, we find an ``amplification effect" of the data advantage: even a small amount thereof gets amplified under competition, causing a large difference in eventual market shares.
+
+\end{itemize}
+
+\ascomment{One downside of the story above: we may need to explain why we didn't simulate \SoftMaxRandom.}
+
+\ascomment{Guy: "anti-trust xhdr" synthesises several points from various places, pls see if I got it right.}
+
+\xhdr{Economic interpretation: anti-trust perspective.}
+Data advantage is significant as a possible barrier to entry in the digital economy 
+\citep[\eg][]{jullien2019economics,de2020data, hagiu2020data}. 
+Our model directly speaks to this literature as we endogenize data-driven network effects by explicitly considering the dynamics of learning under competition.
+Our results are similar to those on ``market tipping" because of indirect network effects \ascomment{some citations??}, and show channels through which differences in data ownership can result in market tipping.
+We observe two important differences compared to prior work:
+
+\begin{itemize}
+\item What matters is not just the quantity of data but its quality: how informative it is for the learning problem at hand. Data collected by better bandit algorithms tends to be more informative.
+
+    \ascomment{Guy: cite prior work that only talks about data quantity?
+Are we really the first paper on anti-trust that talks about data quality?
+Seems like a big claim, we better be right on this.
+}
+
+\item The ``amplification effect" of the data advantage runs contrary to prior work on learning without competition \citep{varian2018artificial,lambrecht2015can,bajari2018impact}, which found that small amounts of additional data do not provide significant improvement in eventual outcomes.
+\end{itemize}
+
+\noindent We conclude that competition dynamics are pertinent for this discussion.
+
+\ascomment{We could, in principle, move both "Econ interpretation" xhdrs to a separate subsection.
+But the current order of presentation seems better to me: we can explain the inverted-U right after "theory", before we put a whole bunch of simulation results on the stack.
+And the anti-trust stuff goes nicely right after the numerical results, without a big subsection break.}
+
+\newpage
+\ascomment{Left the previous version for now (from Guy).}
 

### 2020-07-06 journal paper/content/sec-intro.tex
Full draft of the intro!

@@
 Each agent knows the round (s)he arrives in, computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
-We refer to this variant as \emTheoryModel.
+We refer to this variant as the \emph{\TheoryModel}.
 
@@
 We consider a basic frequentist model.
-We posit that the agents observe some performance signals, and the agents base their decisions on the performance signals alone, without invoking any prior knowledge or beliefs.
+We posit that the agents observe signals about the principals' past performance, and base their decisions on these signals alone, without invoking any prior knowledge or beliefs.
 The performance signals are abstracted and aggregated as a scalar \emph{reputation score} for each principal, modeled as a sliding window average of its rewards.
 Thus, agents' decision rule depends only on the two reputation scores.
-We refer to this variant as \emExptsModel.%
+We refer to this variant as the \emph{\ExptsModel}.%
 %\footnote{In comparison, the theoretical results focus on another extreme, with Bayesian rationality and no performance signals.}
@@
 And the anti-trust stuff goes nicely right after the numerical results, without a big subsection break.}
+
+\subsection{Further discussion}
+\label{sec:discussion}
+
+\xhdr{Significance.}
+Our theory takes a basic Bayesian approach, a standard perspective in economic theory, and discovers several strong asymptotic results.
+The numerical simulations provide a more nuanced and ``non-asymptotic" perspective.
+In essence, we look for substantial effects within relevant time scales.
+(In fact, we start our investigation by determining what time scales are relevant in the context of our model.)
+
+\ascomment{The next para makes a point that seems important (to me). Is this point clear?}
+
+Our study has a dual purpose: (i) shed light on real-world implications of some typical scenarios, and (ii) investigate the space of models for describing the real world.
+As an example to clarify the latter point, consider the \HardMax model with simultaneous entry.
+It is is not necessarily the most realistic model, but arguably the most natural one to study \emph{a priori}.
+However, our results elucidate the need for more refined models with ``free exploration" (\eg via random agents or early entry).
+
+\xhdr{Assumptions.}
+Our models are stylized in several important  respects, some of which we discuss below.
+Firms compete only on the quality of service, rather than, say, pricing or the range of products.
+Agents are myopic: they do not worry about how their actions impact their future utility.
+\footnote{In particular, agents do not attempt to learn over time, to second-guess or game future agents, or to manipulate the principals' learning algorithms.
+This is arguably typical in practice, in part because one agent's influence tends to be relatively small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores", and further simplified to average performance over a time window.
+On the machine learning side, our setup is geared to distinguish between ``simple" vs.
+``better" vs.
+``smart" bandit algorithms; we are not interested in state-of-art algorithms for very realistic bandit settings.
+
+We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs.
+For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior.
+Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
+Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
+It is unclear how to incorporate performance signals in a theoretically tractable model.
+For the numerical results, the \ExptsModel accounts for competition in a more direct way, allows to separate reputation vs.
+data advantage, and makes our model amenable to numerical simulations.
+Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.
+
+\xhdr{Challenges.}
+Much of the challenge in the theoretical results, both conceptual and technical, was in setting up the model and the theorems.
+Apart from zeroing in on the \TheoryModel, it was crucial to interpret the results and intuitions from the literature on multi-armed bandits so as to formulate meaningful and productive assumptions on bandit algorithms and Bayesian priors.
+
+The numerical investigation is quite challenging even with a stylized model such as ours.
+An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given instance of a multi-armed bandit problem (and each such experiment is run many times to reduce variance).
+Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit instances\}, and we are looking for findings that are consistent across this entire space.
+It is essential to keep each of the three dimensions small yet representative.
+In particular, we need to capture a huge variety of bandit instances with only a few representative examples.
+Further, we need a succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).
+
+While amenable to simulations, the \ExptsModel appears difficult to analyze. This is for several reasons:
+%
+(i) intricate feedback loop from performance to reputations to users to performance;
+%
+(ii) mean reputation, most connected to our intuition, is sometimes a bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited}); and
+%
+(iii)
+mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice.
+Given the theoretical results on the \TheoryModel, and the fact that we are in the realm of stylized economic models that provide mathematical intuition about the world, we believe that resolving first-order theoretical questions about the \ExptsModel would not add much value to this paper.
+
+\ascomment{Intro ends here.}
 

### 2020-07-06 journal paper/content/sec-intro.tex
Minor formatting thing.

@@
 mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice.
-Given the theoretical results on the \TheoryModel, and the fact that we are in the realm of stylized economic models that provide mathematical intuition about the world, we believe that resolving first-order theoretical questions about the \ExptsModel would not add much value to this paper.
+Given the theoretical results on the \TheoryModel, and the fact that we are in the realm of stylized economic models, we believe that resolving first-order theoretical questions about the \ExptsModel would not add much value to this paper.
 
@@
 \newpage
+\section{STUFF FROM THE PREVIOUS VERSION}
+
 \ascomment{Left the previous version for now (from Guy).}

### 2020-07-07 journal paper/content/sec-model.tex
Full draft for the "theory section".

@@
 
-\xhdr{Assumptions on the prior.} We make some technical assumptions for the sake of simplicity. First, each action $a$ has a positive probability of being the best action according to the prior:
+\xhdr{Assumptions on the prior.} We make some technical assumptions on the Bayesian prior. First, each action $a$ has a positive probability of being the best action according to the prior:
 \begin{align}\label{eq:assn-prob}
@@
 \end{align}
-\asedit{except at a set of histories of probability $0$}.  In
+except at a set of histories of probability $0$.  In
 particular, prior mean rewards of actions are pairwise distinct:
-$\E[\mu_a] \neq \E[\mu_a']$ for any $a,a'\in A$.
+\[ \E[\mu_a] \neq \E[\mu_a'] \quad \text{for any $a,a'\in A$}. \]
 
@@
 %\end{assumption}
-
-\subsection{Generalizations}
-\label{sec:model-extensions}
-
-Our results can be extended compared to the basic model described above.
-
-%First, the two principals may have different action sets, as long as
-%    $\E\left[ \max_{a\in A} \mu_a\right]$,
-%the Bayesian-expected reward of the best action, is the same for both principals.
-
-First, unless specified otherwise, our results allow a more general notion of principal's utility that can depend on both the market share and agents' rewards.
-Namely, principal $i$ collects $U_i(r_t)$ units of utility in each global round $t$ when she is chosen (and $0$ otherwise), where $U_i(\cdot)$ is some fixed non-decreasing function with $U_i(0)>0$.
-In a formula,
-\begin{align}\label{eq:general-utility}
-\textstyle U_i := \sum_{t=1}^T\; \indicator{i_t=i}\cdot U_i(r_r).
-\end{align}
-
-Second, our results carry over, with little or no modification of the proofs, to much more general versions of MAB, as long as it satisfies the i.i.d.
-property.
-In each round, an algorithm can see a \emph{context} before choosing an action (as in \emph{contextual bandits}) and/or additional feedback other than the reward after the reward is chosen (as in, \eg \emph{semi-bandits}), as long as the contexts are drawn from a fixed distribution, and the (reward, feedback) pair is drawn from a fixed distribution that depends only on the context and the chosen action.
-The Bayesian prior $\prior$ needs to be a more complicated object, to make sure that \PMR and \BIR are well-defined.
-Mean rewards may also have a known structure, such as Lipschitzness, convexity, or linearity; such structure can be incorporated via $\prior$.
-All these extensions have been studied extensively in the literature on MAB, and account for a substantial segment thereof; see \cite{Bubeck-survey12} for background and details.
 

### 2020-07-07 journal paper/content/sec-theory-proofs.tex
Full draft for the "theory section".

--- 2020-06-28 journal paper/content/omitted_proofs.tex
@@
-\subsection{Proof of Theorem~\ref{thm:HardMax-hardTies}}
+\subsubsection{Main result on \HardMax: Proof of Theorem~\ref{thm:DG-dominance}}
+\label{sec:proofs-HM-main}
+
+The proof starts with two auxiliary lemmas: that deviating from \DynGreedy implies a strictly smaller Bayesian-expected reward, and that \HardMax implies a ``sudden-death" property: if one agent chooses principal $1$ with certainty, then so do all subsequent agents do.
+{We re-use both lemmas in later sections, so we state them in sufficient generality.}
+
+\begin{lemma}\label{lm:DG-rew}
+{Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$. Then $\rew_1(n_0)>\rew_2(n_0)$. This holds for any response function $\respF$.}
+\end{lemma}
+
+{Lemma~\ref{lm:DG-rew} does not rely on any particular shape of the response function because it only considers the performance of each algorithm without competition.}
+
+\begin{proof}[Proof of Lemma~\ref{lm:DG-rew}]
+Since the two algorithms coincide on the first $n_0-1$ steps, it follows by symmetry that histories $H_{1,n_0}$ and $H_{2,n_0}$ have the same distribution.
+We use a \emph{coupling argument}: w.l.o.g., we assume the two histories coincide,
+    $H_{1,n_0} = H_{2,n_0} = H$.
+
+At local step $n_0$, \DynGreedy chooses an action $a_{1,n_0} = a_{1,n_0}(H)$
+which maximizes the posterior mean reward given history $H$: for any realized history $h\in \support(H)$ and any action $a\in A$
+\begin{align}\label{eq:lm:DG-rew-1}
+ \PMR(a_{1,n_0} \mid H = h) \geq \PMR(a \mid H=h).
+\end{align}
+
+By assumption \eqref{eq:assn-distinct}, it follows that
+\begin{align}\label{eq:lm:DG-rew-2}
+ \PMR(a_{1,n_0} \mid H = h) > \PMR(a \mid H=h)
+ \quad \text{for any $h\in \support(H)$ and $a\neq a_{1,n_0}(h)$.}
+\end{align}
+
+Since the two algorithms deviate at step $n_0$, there is a set $S\subset \support(H)$ of step-$n_0$ histories such that $\Pr[S]>0$ and any history $h\in S$ satisfies
+    $\Pr[a_{2,n_0}\neq a_{1,n_0} \mid H=h]>0$.
+Combining this with \eqref{eq:lm:DG-rew-2}, we deduce that
+\begin{align}\label{eq:lm:DG-rew-3}
+ \PMR(a_{1,n_0} \mid H = h) > \E\left[ \mu_{a_{2,n_0}}\mid H=h \right]
+ \quad\text{for each history $h\in S$}.
+\end{align}
+Using \eqref{eq:lm:DG-rew-1} and \eqref{eq:lm:DG-rew-3} and integrating over realized histories $h$, we obtain
+    $\rew_1(n_0) > \rew_2(n_0)$.
+\end{proof}
+
+\begin{lemma}\label{lm:DG-sudden}
+{Consider \HardMax response function with $\respF(0)\geq\tfrac12$.}
+Suppose \alg[1] is monotone, and $\PMR_1(t_0)>\PMR_2(t_0)$ for some global round $t_0$. Then $\PMR_1(t)>\PMR_2(t)$ for all subsequent rounds $t$.
+\end{lemma}
 
 \begin{proof}
+Let us use induction on round $t\geq t_0$, with the base case $t=t_0$.
+Let $\mN=\mN_{1,t_0}$ be the agents' posterior distribution for $n_{1,t_0}$, the number of global rounds before $t_0$ in which principal $1$ is chosen.
+By induction, all agents from $t_0$ to $t-1$ chose principal $1$, so $\PMR_2(t_0)= \PMR_2(t)$.
+Therefore,
+\[ \PMR_1(t)
+    = \Ex{n\sim \mN}{\rew_1(n+1+t-t_0)}
+    \geq \Ex{n\sim \mN}{\rew_1(n+1)}
+    =\PMR_1(t_0) > \PMR_2(t_0)= \PMR_2(t), \]
+where the first inequality holds because \alg[1] is monotone, and the second one is the base case.
+\end{proof}
+
+\begin{proof}[Proof of Theorem~\ref{thm:DG-dominance}]
+Since the two algorithms coincide on the first $n_0-1$ steps, it follows by symmetry that $\rew_1(n) = \rew_2(n)$ for any $n< n_0$.
+By Lemma~\ref{lm:DG-rew},
+    $\rew_1(n_0) > \rew_2(n_0)$.
+
+Recall that $n_i(t)$ is the number of global rounds $s<t$ in which principal $i$ is chosen, and $\posteriorN{i}{t}$ is the agents' posterior distribution for this quantity.
+By symmetry, each agent $t<n_0$ chooses a principal uniformly at random.
+It follows that
+    $\posteriorN{1}{n_0} = \posteriorN{2}{n_0}$
+(denote both distributions by $\mN$ for brevity), and $\mN(n_0-1)>0$.
+Therefore:
+\begin{align}
+\PMR_1(n_0)
+  &= \Ex{n\sim \mN} {\rew_1(n+ 1)}
+   = \sum_{n = 0}^{n_0-1} \mN(n) \cdot{\rew_1(n + 1)} \nonumber \\
+  & > \mN(n_0-1)\cdot {\rew_2(n_0)} + \sum_{n = 0}^{n_0-2}  \mN(n)\cdot{\rew_2(n + 1)}
+    \nonumber \\
+  &= \Ex{n\sim \mN}{\rew_2(n + 1)} = \PMR_2(n_0) \label{eq:pf:thm:DG-dominance}
+\end{align}
+So, agent $n_0$ chooses principal $1$. By Lemma~\ref{lm:DG-sudden} {(noting that \DynGreedy is monotone)}, all subsequent agents choose principal $1$, too.
+\end{proof}
+
+%%%%%%%%%%%%
+\subsubsection{\HardMax with biased tie-breaking}
+\label{sec:proofs-HM-main}
+
+\begin{proof}[Proof of Theorem~\ref{thm:HardMax-hardTies}]
+Agent $1$ chooses principal $1$ because of the tie-breaking rule.
+Since \StaticGreedy is trivially monotone, all the subsequent agents choose principal $1$ by an induction argument similar to the one in the proof of Lemma~\ref{lm:DG-sudden}.
+\end{proof}
+
+\begin{proof}[Proof of Theorem~\ref{thm:HardMax-biased}]
 The proof re-uses Lemmas~\ref{lm:DG-rew} and~\ref{lm:DG-sudden}, which do not rely on fair tie-breaking.
@@
 Let $m_0$ be the first step when \alg[2] deviates from \DynGreedy, or \DynGreedy deviates from \StaticGreedy, whichever comes sooner.
-\asedit{Then \alg[2], \DynGreedy and \StaticGreedy coincide on the first $m_0-1$ steps.
+{Then \alg[2], \DynGreedy and \StaticGreedy coincide on the first $m_0-1$ steps.
 Moreover, $m_0\leq n_0$ (since \DynGreedy deviates from \StaticGreedy at step $n_0$), so \alg[1] coincides with \DynGreedy on the first $m_0$ steps.
@@
 
-\subsection{Proof of Theorem~\ref{thm:random-greedy}}
-\begin{proof}
+%%%%%%%%%%%%
+\subsubsection{The main result for \HardMaxRandom: 
+Proof of Theorem~\ref{thm:random-clean}}
+\label{sec:proofs-HMR-main}
+
+Consider global round $t\geq n_0$. Recall that each agent chooses principal $1$ with probability at least
+    $\respF(-1)>0$.
+
+Then
+    $\E[n_1(t+1)] \geq 2\eps_0\,t $.
+By Chernoff Bounds (Theorem~\ref{thm:chernoff}), we have that
+    $n_1(t+1)\geq \eps_0 t$
+holds with probability at least $1-q$,
+where $q = \exp(-\eps_0 t/12)$.
+
+We need to prove that
+    $\PMR_1(t) - \PMR_2(t)>0$.
+For any $m_1$ and $m_2$, consider the quantity
+\[ \Delta(m_1,m_2) := \BIR_2(m_2+1) - \BIR_1(m_1+1).\]
+Whenever $m_1 \geq \eps_0 t/2 -1$ and $m_2<t$, it holds that
+\begin{align*}
+\Delta(m_1,m_2) \geq \Delta(\eps_0 t / 2, t)
+    \geq \BIR_2(t)/2.
+\end{align*}
+The above inequalities follow, resp., from algorithms' monotonicity and \eqref{eq:random-better-weaker}. Now,
+\begin{align*}
+\PMR_1(t) - \PMR_2(t)
+    &= \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}{\Delta(m_1,m_2)} \\
+    &\geq -q
+        + \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}
+            {\Delta(m_1,m_2)\mid m_1 \geq \eps_0 t/2-1} \\
+    &\geq \BIR_2(t)/2-q \\
+    &> \BIR_2(t)/4 > 0  \EqComment{by \eqref{eq:random-assn}}. \qedhere
+\end{align*}
+
+%%%%%%%%%%%%
+\subsubsection{For \HardMaxRandom, a little greedy goes a long way
+(Proof of Theorem~\ref{thm:random-greedy})}
+\label{sec:proofs-HMR-main}
+
+%\begin{proof}[Proof of Theorem~\ref{thm:random-greedy}]
   Let $\rew'_1(n)$ denote the Bayesian-expected reward of the ``greedy
@@
 \footnote{If $\rew_1(\cdot)$ is strictly increasing, then inequality \eqref{eq:pf:thm:random-greedy-1} is strict, too; this is because $\mM_{2,t}(t-1)>\posteriorN{1}{t}(t-1)$.  }
-\end{proof}
+%\end{proof}
 
-\subsection{Proof of Theorem~\ref{thm:SoftMax-strong}}
-\begin{proof}
+%%%%%%%%%%%%
+\subsubsection{\SoftMaxRandom: proof of Theorem~\ref{thm:SoftMax-strong}}
+\label{sec:proofs-SoftMax}
+
+\xhdr{Proof Sketch.}
+The main idea is that even though \alg[1] may have a
+slower rate of learning in the beginning, it will gradually catch up
+and surpass \alg[2]. We will describe this process in two phases. In
+the first phase, \alg[1] receives a random agent with probability at
+least $\respF(-1) = \eps_0$ in each round. Since $\BIR_1$ tends to 0,
+the difference in \BIR{s} between the two algorithms is also
+diminishing. Due to the \SoftMaxRandom response function, \alg[1]
+attracts each agent with probability at least $1/2 - O(\beta_0)$ after
+a sufficient number of rounds. Then the game enters the second phase:
+both algorithms receive agents at a rate close to $\tfrac12$, and the
+fractions of agents received by both algorithms --- $n_1(t)/t$ and
+$n_2(t)/t$ --- also converge to $\tfrac12$. At the end of the second
+phase and in each global round afterwards, the counts $n_1(t)$ and
+$n_2(t)$ satisfy the weak BIR-dominance condition, in the sense that
+they both are larger than $n_0$ and $n_1(t)\geq (1-\beta_0)\; n_2(t)$.
+At this point, \alg[1] actually has smaller $\BIR$, which reflected in the {\PMR}s eventually. Accordingly, from then on \alg[1]
+attracts agents at a rate slightly larger than $\tfrac12$. We prove
+that the ``bump'' over $\tfrac12$ is at least on the order of
+$\BIR_2(t)$.
+
+%\begin{proof}[Proof of Theorem~\ref{thm:SoftMax-strong}]
+\xhdr{Full proof.}
   Let $\beta_1 = \min\{c_0'\delta_0, \beta_0/20\}$ with $\delta_0$
@@
   \alg[1] with probability at least $\respF(-1)= \eps_0$.  By
-  condition \eqref{eq:so-mild} and \eqref{eq:SoftMax-assn-strong},
+By condition \eqref{eq:SoftMax-assn-strong} and the fact that
+$\BIR_1(n) \to 0$,
   there exists some sufficiently large $T_1$ such that for any
@@
 \]
-\end{proof}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "main.tex"
-%%% End:
+%\end{proof}

### 2020-07-07 journal paper/content/sec-theory.tex
Full draft for the "theory section".

\ascomment{some intro para goes here.}

\subsection{Full rationality: \HardMax}
\label{sec:theory-HM}

We consider the version in which the agents are fully rational, in the sense that their response function is \HardMax. We show that principals are not incentivized to \emph{explore}--- \ie to deviate from \DynGreedy. The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates.

To make this more precise, let us say that two MAB algorithms \emph{deviate} at (local) step $n$ if there is an action $a\in A$ and a set of step-$n$ histories of positive probability such that any history $h$ in this set is feasible for both algorithms, and under this history the two algorithms choose action $a$ with different probability.

\begin{theorem}\label{thm:DG-dominance}
Assume \HardMax response function with fair tie-breaking. Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$. Then all agents in global rounds $t\geq n_0$ select principal $1$.
\end{theorem}

\begin{corollary}\label{cor:DG-dominance}
The competition game between principals has a unique Nash equilibirium: both principals choose \DynGreedy.
\end{corollary}

\begin{remark}
This corollary holds under a more general model which allows time-discounting: namely, the utility of each principal $i$ in each global round $t$ is $U_{i,t}(r_t)$ if this principal is chosen, and $0$ otherwise, where $U_{i,t}(\cdot)$ is an arbitrary non-decreasing function with $U_{i,t}(0)>0$.
\end{remark}

%\subsection{\HardMax with biased tie-breaking}
%\label{sec:HardMax-biased}

\xhdr{Biased tie-breaking.}
The \HardMax model is very sensitive to the tie-breaking rule. For starters, if ties are  broken deterministically in favor of principal $1$, then principal 1 can get all agents no matter what the other principal does, simply by using \StaticGreedy.

\begin{theorem}\label{thm:HardMax-hardTies}
Assume \HardMax response function with $\respF(0)=1$ (ties are always broken in favor of principal $1$). If \alg[1] is \StaticGreedy, then all agents choose principal $1$.
\end{theorem}

A more challenging scenario is when the tie-breaking is biased in favor of principal 1, but not deterministically so: $\respF(0)>\tfrac12$. Then this principal also has a ``winning strategy" no matter what the other principal does. Specifically, principal 1 can get all but the first few agents, under a mild technical assumption that \DynGreedy deviates from \StaticGreedy. Principal 1 can use \DynGreedy, or any other monotone MAB algorithm that coincides with \DynGreedy in the first few steps.

%We can generalize the theorem top the case of $q >1/2$ if the
%principal $p_1$ can guarantee better than the a priori best action to
%all the agents following the second.

\begin{theorem}\label{thm:HardMax-biased}
Assume \HardMax response function with $\respF(0)>\tfrac12$ (\ie tie-breaking is biased in favor of principal $1$). Assume the prior $\mP$ is such that \DynGreedy deviates from \StaticGreedy starting from some step $n_0$. Suppose that principal $1$ runs a monotone MAB algorithm that coincides with \DynGreedy in the first $n_0$ steps. Then all agents $t\geq n_0$ choose principal $1$.
\end{theorem}

%%%%%%%%%%
\subsection{Relaxed rationality: \HardMaxRandom}
\label{sec:theory-HMR}

Consider the \HardMaxRandom response model, where each principal is chosen with some positive baseline probability. The main technical result for this model states that a principal with asymptotically better \BIR wins by a large margin: after a ``learning phase" of constant duration, all agents choose this principal with maximal possible probability $\respF(1)$. For example, a principal with $\BIR(n)\leq \tilde{O}(n^{-1/2})$ wins over a principal with $\BIR(n)\geq \Omega(n^{-1/3})$.

%However, this positive result comes with a significant caveat detailed in Section~\ref{sec:random-greedy}.

%We formulate and prove a cleaner version of the result, followed by a more general formulation developed in a subsequent Remark~\ref{rem:random-messy}.

We need to express a property that \alg[1] eventually catches up and surpasses \alg[2], even if initially it receives only a fraction of traffic. We assume that both algorithms run indefinitely and do not depend on the time horizon $T$; call such algorithms \emph{$T$-oblivious}. In particular, their \BIR does not depend on the time horizon $T$ of the game.  Then this property can be formalized as:
\begin{align}\label{eq:random-better-clean}
(\forall \eps>0)\qquad
\frac{\BIR_1(\eps n)}{\BIR_2(n)} \to 0.
\end{align}
In fact, a weaker version of \eqref{eq:random-better-clean} suffices:
denoting $\eps_0 = \respF(-1)$, for some constant $n_0$ we have
\begin{align}\label{eq:random-better-weaker}
(\forall n\geq n_0) \qquad
\frac{\BIR_1(\eps_0\, n/2)}{\BIR_2(n)} <\nicefrac12.
\end{align}
If this holds, we say that \alg[1] \emph{BIR-dominates} \alg[2].

\begin{theorem}\label{thm:random-clean}
Assume \HardMaxRandom response function. Suppose algorithms \alg[1], \alg[2] are monotone and $T$-oblivious. Assume that \alg[1] \emph{BIR-dominates} \alg[2], and the latter satisfies a mild technical assumption:
\begin{align}\label{eq:random-assn}
 (\forall n\geq n_0) \qquad
  \BIR_2(n) > 4\,e^{-\eps_0\ n/12}.
\end{align}
Then each agent $t\geq n_0$ chooses principal $1$ with maximal possible probability $\respF(1) = 1- \eps_0$.
\end{theorem}

%Let \alg be a monotone MAB algorithm and $\mA$ be a finite set of monotone MAB algorithms such that each algorithm in $\mA$ satisfies \eqref{eq:random-assn} and is ``dominated" by \alg in the sense of \eqref{eq:random-better-messy}.
%Assume principals can only choose algorithms from $\mA\cup \{\alg\}$.

To state a game-theoretic corollary of Theorem~\ref{thm:random-clean}, we restrict the set of feasible MAB algorithms:

\begin{definition}\label{def:restricted-competition}
A \emph{restricted competition game} is a version of the competition game between the two principals in which they can only choose from a finite set of MAB algorithms, which are called \emph{feasible}. All feasible algorithms are assumed to be
monotone and $T$-oblivious, and to satisfy \eqref{eq:random-assn}. 
\end{definition}

\begin{corollary}\label{cor:random}
Assume \HardMaxRandom response function. Consider the restricted competition game with a feasible algorithm \alg which BIR-dominates all other feasible algorithms. Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium: both principals choose \alg.
\end{corollary}

%\subsection{A little greedy goes a long way}
%\label{sec:random-greedy}

\xhdr{A little greedy goes a long way.}
Given any monotone MAB algorithm other than \DynGreedy, we design a modified algorithm which learns at a slower rate, yet ``wins the game" in the sense of Theorem~\ref{thm:random-clean}. So, the competition game with unrestricted choice of algorithms typically does not have a Nash equilibrium.

Given an algorithm \alg[1] that deviates from \DynGreedy starting from
step $n_0$ and a ``mixing'' parameter $p$, we will construct a
modified algorithm as follows.
\begin{enumerate}
\item The modified algorithm coincides with \alg[1] (and \DynGreedy)
for the first $n_0-1$ steps;
\item In each step $n\geq n_0$, \alg[1] is invoked with probability
  $1-p$, and with the remaining probability $p$ does the ``greedy
  choice": chooses an action with the largest posterior mean reward
  given the current information collected by \alg[1].
\end{enumerate}
%The modified algorithm is called \emph{greedy deviation}, with probability parameter $p$.
For a cleaner comparison between the two algorithms, the modified algorithm does not record rewards received in steps with the ``greedy choice". Parameter $p>0$ is the same for all steps.

\begin{theorem}\label{thm:random-greedy}
Assume symmetric \HardMaxRandom response function. Let $\eps_0 = \respF(-1)$ be the baseline probability. Suppose \alg[1] deviates from \DynGreedy starting from some step $n_0$. Let \alg[2] be the modified algorithm, as described above, with mixing parameter $p$ such that
    $(1-\eps_0)(1-p)>\eps_0$.
Then each agent $t\geq n_0$ chooses principal $2$ with maximal possible probability $1-\eps_0$.
\end{theorem}

\begin{corollary}\label{cor:random-greedy}
  Suppose both principals can choose any monotone MAB algorithm, and assume the symmetric \HardMaxRandom response
  function. Then for any time
  horizon $T$, the only possible \emph{pure} Nash equilibrium is one
  where both principals choose \DynGreedy. Moreover, no pure Nash
  equilibrium exists when some algorithm ``dominates" \DynGreedy in
  the sense of \eqref{eq:random-better-messy} and the time horizon $T$
  is sufficiently large.
\end{corollary}

\begin{remark}
The modified algorithm performs exploration % ingests information
at a slower rate. Let us argue how this may translate  into a larger \BIR compared to the original algorithm. Let  $\BIR'_1(n)$ be the \BIR of the ``greedy choice" after after $n-1$ steps of \alg[1]. Then
\begin{align}\label{eq:random-BIR}
\BIR_2(n)
    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIR'_1(m)}.
\end{align}
In this expression, $m$ is the number of times \alg[1] is invoked in the first $n$ steps of the modified algorithm. Note that
    $\E[m] = n_0-1 + (n-n_0+1)(1-p) \geq (1-p)n$.

Suppose $\BIR_1(n)= \beta n^{-\gamma}$ for some constants $\beta,\gamma>0$. Further, assume
    $\BIR'_1(n)\geq  c\; \BIR_1(n)$,
for some $c>1-\gamma$.
Then for all $n\geq n_0$ and small enough $p>0$ it holds that:
\begin{align*}
 \BIR_2(n)
    &\geq  (1-p+pc)\; \E[\; \BIR_1(m) \;] \\
\E[\; \BIR_1(m) \;]
    &\geq \BIR_1(\; \E[m] \;) &\qquad\text{(By Jensen's inequality)} \\
    &\geq \BIR_1(\; (1-p)n \;) &\qquad\text{(since $\E[m]\geq n(1-p)$)}  \\
    &\geq \beta\cdot n^{-\gamma} \cdot (1-p)^{-\gamma}
        &\qquad\text{(plugging in $\BIR_1(n)=\beta n^{-\gamma}$)}  \\
    &> \BIR_1(n)\;\; (1-p\gamma)^{-1}
        &\qquad\text{(since $(1-p)^\gamma < 1-p\gamma$)}.\\
\BIR_2(n)
    &>\alpha\cdot \BIR_1(n),
    &\text{where} \quad
    \alpha = \tfrac{1-p+pc}{1-p\gamma}>1.
\end{align*}
(In the above equations, all expectations are over $m$ distributed as in \eqref{eq:random-BIR}.)
\end{remark}

%%%%%%%%%
\subsection{\SoftMaxRandom response function}
\label{sec:theory-SoftMax}

This subsection is devoted to the \SoftMaxRandom model. We derive a ``better algorithm wins" result under a much weaker version of BIR-dominance. This is the most technical result of the paper.

We start with a formal definition of \SoftMaxRandom:

\begin{definition}\label{def:SoftMax}
A response function $\respF$ is \SoftMaxRandom if the following conditions hold:
\begin{OneLiners}
\item  $\respF(\cdot)$ is bounded away from $0$ and $1$:
    $\respF(\cdot)\in [\eps,1-\eps]$ for some $\eps\in (0,\tfrac12)$,
\item  the response function
 $\respF(\cdot)$ is ``smooth" around $0$:
 \begin{align}\label{eq:SoftMax-smooth}
 \exists\, \text{constants $\delta_0,c_0,c'_0>0$}
    \qquad \forall x\in [-\delta_0,\delta_0] \qquad
    c_0 \leq \respF'(x) \leq c'_0.
 \end{align}
\item fair tie-breaking: $\respF(0)=\tfrac12$.
\end{OneLiners}
\end{definition}

\begin{remark}
This definition is fruitful when parameters $c_0$ and $c_0'$ are close to $\tfrac12$. Throughout, we assume that \alg[1] is better than \alg[2], and obtain results parameterized by $c_0$. By symmetry, one could assume that \alg[2] is better than \alg[1], and obtain similar results parameterized by $c_0'$.
\end{remark}

For the sake of intuition, let us derive a version of Theorem~\ref{thm:random-clean}, with the same assumptions about the algorithms and essentially the same proof. The conclusion is much weaker: we can only guarantee that each agent $t\geq n_0$ chooses principal 1 with probability slightly larger than $\tfrac12$. This is essentially unavoidable in a typical case when both algorithms satisfy $\BIR(n)\to 0$, by Definition~\ref{def:SoftMax}.

\begin{theorem}\label{thm:SoftMax-weak}
Assume \SoftMaxRandom response function.  Suppose algorithms \alg[1], \alg[2] satisfy the assumptions in Theorem~\ref{thm:random-clean}. Then each agent
  $t\geq n_0$ chooses principal $1$ with probability
\begin{align}\label{eq:thm:SoftMax-weak}
     \Pr[i_t = 1]\geq \tfrac12 +  \tfrac{c_0}{4}\; \BIR_2(t).
\end{align}
\end{theorem}

\begin{proof}[Proof Sketch]
We follow the steps in the proof of Theorem~\ref{thm:random-clean} to derive \begin{align*}
\PMR_1(t) - \PMR_2(t)
    &\geq \BIR_2(t)/2 -q,
    \quad \text{where $q = \exp(-\eps_0 t/12)$.}
\end{align*}
This is at least $\BIR_2(t)/4$ by \eqref{eq:random-assn}. Then \eqref{eq:thm:SoftMax-weak} follows by the smoothness condition \eqref{eq:SoftMax-smooth}.
\end{proof}

\OMIT {%%%%%%%%
We recover a version of Corollary~\ref{cor:random}, if each principal's utility is the number of users (rather than the more general model in \eqref{eq:general-utility}).

\begin{corollary}\label{cor:SoftMax}
Assume that the response function is \SoftMaxRandom, and each principal's  utility is the number of users.
%
%Suppose principals can only choose algorithms from $\mA\cup \{\alg\}$, where $\mA\cup \{\alg\}$ is a finite set of monotone MAB algorithms such that each algorithm in $\mA$ satisfies \eqref{eq:random-assn} and $\BReg(n)\to \infty$, and is ``dominated" by \alg in the sense of \eqref{eq:random-better-messy}.
%
Consider the restricted competition game with special algorithm \alg, and assume that all other allowed algorithms satisfy $\BReg(n)\to \infty$. Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium: both principals choose \alg.
\end{corollary}
} %%%% \OMIT{

Let us relax the notion of BIR-dominance so that the constant multiplicative factors in \eqref{eq:random-better-weaker}, namely
 $\eps_0/2$ and $\tfrac12$, are replaced by constants that can be arbitrarily close to $1$.

\begin{definition}
Suppose MAB algorithms \alg[1],  \alg[2] are $T$-oblivious. Say that
\alg[1] \emph{weakly BIR-dominates} \alg[2] if there exist
absolute  constants $\beta_0, \alpha_0\in (0, 1/2)$ and $n_0\in\N$ such that
 \begin{align}\label{eq:SoftMax-better}
   (\forall n\geq n_0) \quad
   \frac{\BIR_1((1-\beta_0)\, n)}{\BIR_2(n)} <1-\alpha_0.
 \end{align}
 \end{definition}

Now we are ready to state the main result for \SoftMaxRandom:

\begin{theorem}\label{thm:SoftMax-strong}
Assume the \SoftMaxRandom response function. Suppose algorithms \alg[1], \alg[2] are monotone and $T$-oblivious, and \alg[1] weakly-BIR-dominates \alg[2]. Posit mild technical assumptions: 
  $\BIR_1(n) \to 0$ and 
\begin{align}\label{eq:SoftMax-assn-strong}
 % (\forall n\geq n(\eps)) \qquad
 %  \BIR_2(n) > e^{-\eps n}. %used to be 4exp( -\eps n/6)
(\exists n_0\; \forall n\geq n_0) \qquad
\BIR_2(n) \geq \frac{4}{\alpha_0}
\exp \left( \frac{-n\,\min\{\eps_0, 1/8\}}{12}\right).
\end{align}
Then there
  exists some $t_0$ such that each agent $t\geq t_0$ chooses principal
  $1$ with probability
\begin{align}\label{eq:thm:SoftMax-strong}
     \Pr[i_t = 1]\geq \tfrac12 +  \tfrac14 \,c_0\,\alpha_0\; \BIR_2(t).
\end{align}
\end{theorem}

As in the previous subsection, we can state a game-theoretic corollary for the restricted competition game (as in Definition~\ref{def:restricted-competition}). We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity. \BReg is a standard notion from the literature (along with \BIR):
\begin{align}\label{eq:SoftMax-BReg}
\BReg(n) := n\cdot \E_{\mu\sim\priorMu}
    \left[ \max_{a\in A} \mu_a\right] - \sum_{n=1}^n \rew(n')
    = \sum_{n'=1}^n \BIR(n').
\end{align}

\begin{corollary}\label{cor:SoftMax-strong}
Assume \SoftMaxRandom response function. Consider the restricted competition game in which all feasible algorithms satisfy $\BReg(n)\to \infty$.%
\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition.
\ascomment{explain why, cite smth?}} 
Suppose a feasible algorithm \alg BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.
\end{corollary}

%If both $\BIR_1()$ and $\BIR_2()$ are of the form $\tilde{\Theta}(n^{-\gamma})$, $\gamma>0$, then the old condition requires $\BIR_1(\cdot)$ to be better by constant multiplicative factor $C$, with $C$ sufficiently large, whereas the new condition allows any $C>1$.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

%%%%%%%%%%%%%
\subsection{Economic implications}
\label{sec:theory-welfare}

{We frame our contributions in terms of the relationship between \competitiveness and \rationality on one side, and adoption of better algorithms on the other. Recall that both \competitiveness (of the game between the two principals) and \rationality (of the agents) are controlled by the response function $\respF$.}

\OMIT{ %%%%%%
We frame our contributions in terms of the relationship between \competition and \innovation, \ie between the extent to which the game between the two principals is competitive, and the degree of innovation --- adoption of better that these models incentivize. \Competition is controlled via the response function $\respF$, and \innovation refers to the quality of the technology (MAB algorithms) adopted by the principals. The \competition vs. \innovation relationship is well-studied in the economics literature, and is commonly known to often follow an inverted-U shape, as in \reffig{fig:inverted-U} (see Section~\ref{sec:related-work} for citations). \Competition in our models is closely correlated with \rationality: the extent to which agents make rational decisions, and indeed \rationality is what $\respF$ controls directly.
} %%%%%%%%

\xhdr{Main story.}
Our main story concerns the restricted competition game between the two principals where one allowed algorithm \alg is ``better" than the others. {We track whether and when \alg is chosen in an equilibrium.} We vary \competitiveness/\rationality by changing the response function from \HardMax (full rationality, very competitive environment) to \HardMaxRandom to  \SoftMaxRandom (less rationality and competition). Our conclusions are as follows:
\begin{OneLiners}
\item Under \HardMax, no innovation: \DynGreedy is chosen over \alg.
\item Under \HardMaxRandom, some innovation:  \alg is chosen as long as it BIR-dominates.
\item Under \SoftMaxRandom, more innovation: \alg is chosen as long as it weakly-BIR-dominates.%
\footnote{This is a weaker condition, the better algorithm is chosen in a broader range of scenarios.}
\end{OneLiners}
These conclusions follow, respectively, from Corollaries~\ref{cor:DG-dominance}, \ref{cor:random} and \ref{cor:SoftMax}. Further, {we consider the uniform choice between the principals. It corresponds to the least amount of rationality and competition, and (when principals' utility is the number of agents) uniform choice provides no incentives to innovate.}%
\footnote{On the other hand, if principals' utility is somewhat aligned with agents' welfare, as in \eqref{eq:general-utility}, then a monopolist principal is incentivized to choose the best possible MAB algorithm (namely, to minimize cumulative Bayesian regret $\BReg(T)$). Accordingly, monopoly would result in better social welfare than competition, as the latter is likely to split the market and cause each principal to learn more slowly. This is a very generic and well-known effect regarding economies of scale.}
Thus, we have an inverted-U relationship, see \reffig{fig:inverted-U2}.

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=1]
      \draw[->] (-.5,0) -- (9.5,0) node[above]
        {\qquad\qquad Competitiveness/Rationality};
      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithm in equilibrium};
      \draw[scale=0.8,domain=0.5:9.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{3.5 - 0.15*(\x - 5)^2});
     \node[below] at (1, 0) {\footnotesize \Uniform};
     \node[below] at (3.9, 0) {\footnotesize \SoftMaxRandom};
     \node[below] at (6, 0) {\footnotesize \HardMaxRandom};
     \node[below] at (8, 0) {\footnotesize \HardMax};
      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
 \end{tikzpicture}

\caption{The stylized inverted-U relationship in the ``main story".}
\label{fig:inverted-U2}
\end{center}
\end{figure}

\xhdr{Secondary story.}
Let us zoom in on the symmetric  \HardMaxRandom model. {Competitiveness and rationality within this model are controlled by the baseline probability $\eps_0 = \respF(- 1)$, which goes smoothly between the two extremes of \HardMax ($\eps_0=0$) and the uniform choice ($\eps_0=\tfrac12$). Smaller $\eps_0$ corresponds to increased rationality and increased competitiveness.} For clarity, we assume that principal's utility is the number of agents.

We consider the marginal utility of switching to a better algorithm. Suppose initially both principals use some algorithm \alg, and principal 1 ponders switching to another algorithm \alg' which BIR-dominates \alg. {We are interested in the marginal utility of this switch. Then:}

\begin{itemize}
\item $\eps_0 = 0$ (\HardMax):~~~~ the marginal utility can be negative if \alg is \DynGreedy.

\item $\eps_0$ near $0$:~~~~ only a small marginal utility can be guaranteed, as it may take a long time for $\alg'$ to ``catch up" with \alg, and hence less time to reap the benefits.

\item ``medium-range" $\eps_0$:~~~~ large marginal utility, as $\alg'$ learns fast and gets most agents.

\item $\eps_0$ near $\tfrac12$:~~~~ small marginal utility, as principal 1 gets most agents for free no matter what.
\end{itemize}
The familiar inverted-U shape is depicted in Figure~\ref{fig:inverted-U3}.

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=1]
      \draw[->] (-.5,0) -- (9.5,0) node[above]  {$\eps_0$};
      \draw[->] (0,-.5) -- (0,3) node[above] {marginal utility};
      \draw[scale=0.8,domain=0.5:9.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{3.5 - 0.15*(\x - 5)^2});
     \node[below] at (.6, 0) {\footnotesize \Uniform};
     \node[above] at (.5, 0) {\footnotesize 0};
     % \node[below] at (3.9, 0) {\footnotesize \SoftMaxRandom};
     % \node[below] at (6, 0) {\footnotesize \HardMaxRandom};
     \node[below] at (7.5,0) {\footnotesize \HardMax};
     \node[above] at (7.5, 0) {\footnotesize 1/2};
      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
 \end{tikzpicture}

\caption{The stylized inverted-U relationship from the ``secondary story"}
\label{fig:inverted-U3}
\end{center}
\end{figure}

%%%%%%%%%%%%%
\subsection{Generalizations}
\label{sec:theory-extensions}

\ascomment{need to polish this subsection}

Our results can be extended compared to the basic model described above.

%First, the two principals may have different action sets, as long as
%    $\E\left[ \max_{a\in A} \mu_a\right]$,
%the Bayesian-expected reward of the best action, is the same for both principals.

First, unless specified otherwise, our results allow a more general notion of principal's utility that can depend on both the market share and agents' rewards. Namely, principal $i$ collects $U_i(r_t)$ units of utility in each global round $t$ when she is chosen (and $0$ otherwise), where $U_i(\cdot)$ is some fixed non-decreasing function with $U_i(0)>0$. In a formula,
\begin{align}\label{eq:general-utility}
\textstyle U_i := \sum_{t=1}^T\; \indicator{i_t=i}\cdot U_i(r_r).
\end{align}

\ascomment{Edit in: this extension does not apply to Corollary~\ref{cor:SoftMax-strong}.}

Second, our results carry over, with little or no modification of the proofs, to much more general versions of MAB, as long as it satisfies the i.i.d. property. In each round, an algorithm can see a \emph{context} before choosing an action (as in \emph{contextual bandits}) and/or additional feedback other than the reward after the reward is chosen (as in, \eg \emph{semi-bandits}), as long as the contexts are drawn from a fixed distribution, and the (reward, feedback) pair is drawn from a fixed distribution that depends only on the context and the chosen action. The Bayesian prior $\prior$ needs to be a more complicated object, to make sure that \PMR and \BIR are well-defined. Mean rewards may also have a known structure, such as Lipschitzness, convexity, or linearity; such structure can be incorporated via $\prior$. All these extensions have been studied extensively in the literature on MAB, and account for a substantial segment thereof; see \cite{Bubeck-survey12} for background and details.

\xhdr{Dependence on the time horizon.}
%\begin{remark}\label{rem:random-messy}
  Many standard MAB algorithms in the literature are parameterized by
  the time horizon $T$. Regret bounds for such algorithms usually
  include a polylogarithmic dependence on $T$. In particular, a
  typical upper bound for \BIR has the following form:
  \OMIT{\footnote{We
    provide upper bounds on \BIR for several standard MAB algorithms
    to illustrate these dependencies in the appendix.}}
\begin{align}
    \BIR(n\mid T)\leq \polylog(T)\cdot n^{-\gamma}
    \quad \text{for some $\gamma\in(0,\tfrac12]$}.
\end{align}
Here we write $\BIR(n\mid T)$ to emphasize the dependence on $T$.

Theorem~\ref{thm:random-clean} easily extends if condition \eqref{eq:random-better-weaker} is replaced with the following:
there exists a number $T_0$ and a function $n_0(T)\in \polylog(T)$
such that
\begin{align}\label{eq:random-better-messy}
(\forall T\geq T_0,\; n\geq n_0(T)) \quad
\frac{\BIR_1(\eps_0 n /2\mid T)}{\BIR_2(n\mid T)} <\frac12.
\end{align}
%If this holds, we say that \alg[1] \emph{BIR-dominates} \alg[2].
%The proof of this extension is very similar and is omitted.

%\begin{theorem}\label{thm:random-messy}
%Assume \HardMaxRandom response function, and algorithms that satisfy \eqref{eq:random-better-messy} and~\eqref{eq:random-assn}. Then each agent
%    $t\geq n_0(T)$ chooses principal $1$ with probability $\respF(1)$.
%\end{theorem}

Similar to the condition \eqref{eq:random-better-weaker}, we can also
generalize the weak BIR-dominance condition \eqref{eq:SoftMax-better}
to handle the dependence on $T$: there exist some $T_0$, a function
$n_0(T)\in \polylog(T)$, and constants $\beta_0,\alpha_0\in (0, 1/2)$, such that
\begin{align}\label{eq:SoftMax-better-weaker}
(\forall T\geq T_0,  n\geq n_0(T)) \quad
\frac{\BIR_1((1-\beta_0)\, n\mid T)}{\BIR_2(n\mid T)} <1-\alpha_0.
\end{align}

We also provide a version of Theorem~\ref{thm:SoftMax-weak} under this
more general weak BIR-dominance condition; its proof is very similar
and is omitted.

### 2020-07-07 journal paper/ec19paper/content/revisited.tex
the latex compiles
tweaks
Reformatted so that: - all expts in the body are in one section. - all proofs in the app are in one section - all expts in the app are in one section.

--- 2019-02-14 paper/content/revisited.tex
@@
 
-\section{Performance in Isolation, Revisited}\label{sec:revisited}
+\subsection{Performance in Isolation, Revisited}\label{sec:revisited}
 
@@
 
-\begin{figure}[ht]
-\includegraphics[scale=0.35]{figures/rep_distribution_nih}
+\begin{figure}[H]
+\includegraphics[scale=0.35]{ec19paper/figures/rep_distribution_nih}
 \caption{Reputation scores for Needle-in-Haystack at $t=500$ (smoothed using a kernel density estimate)}
@@
 
-\begin{figure}[ht]
-\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
+\begin{figure}[H]
+\includegraphics[scale=0.35]{ec19paper/figures/ts_dg_rep_diff_nih}
 \caption{Reputation difference $\TS-\DG$ for  Needle-in-Haystack

### 2020-07-07 journal paper/content/sec-intro.tex
Full draft for the "theory section".
Added a footnote re discrepancy for HardMax.

@@
 \begin{itemize}
-\item We compare \HardMax and \HardMaxRandom decision rules.
-We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.
-As predicted by the theory, better algorithms prevail if the expected number of ``random" users is sufficiently large.
-However, this effect is negligible for smaller parameter values.
+\item We compare \HardMax and \HardMaxRandom decision rules. We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.%
+\footnote{The greedy algorithm wins often but not always, whereas it
+always wins under \TheoryModel. The reason for this discrepancy is that the greedy algorithm in the theoretical results is aware of the Bayesian prior.
+}
+ As predicted by the theory, better algorithms prevail if the expected number of ``random" users is sufficiently large. However, this effect is negligible for smaller parameter values.
 
@@
 \newpage
-\section{STUFF FROM THE PREVIOUS VERSION}
+\section*{STUFF FROM THE PREVIOUS VERSION}
 

### 2020-07-07 journal paper/content/old-intro.tex
Moved the stuff from Guy's intro to a separate file.

--- 2020-07-06 journal paper/content/sec-intro.tex
@@
-Learning from interactions with users is ubiquitous in modern customer-facing platforms, from product recommendations to web search to content selection to fine-tuning user interfaces.
-Many platforms purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
-
-%~\cite{KohaviAB-2015,KohaviLSH09}
-
-In this paper, we initiate a study of the interplay between \exploration and \competition.
-Platforms that engage in exploration typically need to compete against one another; most importantly, they compete for users.
-This creates a tension:
-%between \exploration and \competition.
-while exploration may be essential for improving the service tomorrow, it may degrade quality and make users leave \emph{today}, in which case there will be less users to learn from.
-This may further degrade the platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
-Taken to the extreme, such dynamics may create a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
-Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.
-
-The main high-level question that we focus on in this paper is:
-\begin{align}\label{eq:main-Q}
-\textbf{How does competition incentivize the adoption of better exploration algorithms?}
-\end{align}
-
-\noindent This translates into a number of more concrete questions.
-While it is commonly assumed that better technology always helps, is this so under competition?
-Does increased competition lead to higher consumer welfare?
-How significant are ``data feedback loops" --- when more data leads to more users, which leads to even more data, etc.
---- and how they relate to the anti-trust considerations?
-Finding formalizations that admit meaningful answers is a major part of the overall challenge.
-
-To answer these questions, we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
-The choice of a particular technology (exploration algorithm) is no longer an abstract, static choice with a predetermined outcome for the platform.
-Instead, we model the algorithms explicitly, and investigate how they play out in competition over an extended period of time.
-
-%\footnote{This is a fundamental question which is part of a larger policy discussion around whether data can serve as an indirect network effect and lead to similar ``market tipping" results as is standard in the literature on competition in markets with network effects (see \cite{jullien2019economics} for a policy oriented discussion of this).}
-
-% the extent to which the game between the two principals is competitive
-% degree of innovation that these models incentivize.
-% the extent to which agents make rational decisions
-
-%\subsection{Our model}
-%\label{sec:intro-model}
-
-\subsection{Our model: competing bandits}
-\label{sec:intro-model}
-
-%We investigate these questions with
-\xhdr{Competition game.} We consider a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
-We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
-These two processes are interlinked, as exploration decisions are experienced by users and informed by their feedback.
-We need to specify several conceptual pieces: how the principals and agents interact, what is the machine learning problem faced by each principal, and what is the information structure.
-Each piece can get rather complicated in isolation, let alone jointly, so we strive for simplicity.
-Thus, the key features of our model are as follows:
-
-\begin{itemize}
-
-\item A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals.
-The principal chooses an action (\eg a list of web search results to show to the agent), the user experiences this action, and reports a reward.
-All agents have the same ``decision rule" for choosing among the principals given the available information.
-
-\item Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
-\emph{arms}) and receives a reward drawn independently from a fixed distribution specific to this action.
-The reward distributions are initially unknown (but can be estimated over time from the data).
-
-\item Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
-Each principal's objective is to maximize its market share (the fraction of users choosing this principal).
-Each principal only observes agents that chose this principal.
-%, but may have access to each platforms' reputation score (more on this in    Section~\ref{sec:intro-discussion}).
-\end{itemize}
-
-\noindent We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
-In all variants, agents have little or no information about other agents' choices and rewards.
-
-%We consider two model variants, which determine agents' decision rule and their information sets.
-In \emTheoryModel, there is a common Bayesian prior on the reward distributions.
-Agents do not receive any other information and choose between the principals using their knowledge of $t$ and the principals' algorithms.
-In \emExptsModel, agents have access to a reputation score for each principal, which is a sliding window average of the rewards experienced by previous agents that have visited this principal.
-The former variant used for the theoretical results, the latter for simulations.
-
-\xhdr{Technology: multi-armed bandit algorithms.}
-To flesh out \eqref{eq:main-Q}, let us elaborate what we mean by `better' bandit algorithms.
-In general, comparisons between bandit algorithms are rather subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
-
-We distinguish three classes of bandit algorithms, based on
-the prevalent intuition in the area.
-The distinction concerns the way in which they resolve the fundamental tradeoff between  exploration and \emph{exploitation} (making optimal myopic decisions using the available data).
-Going from more primitive to more sophisticated, these three classes are as follows:
-
-\begin{itemize}
-\item \emph{Greedy algorithms} that strive to maximize the reward for the next round given the available information. Thus, they always ``exploit" and never explicitly ``explore".
-
-\item \emph{Exploration-separating algorithms}
- that separate exploration and exploitation: essentially, each round is dedicated to one and completely ignores the other.
-
-\item \emph{Adaptive-exploration} algorithms that combine exploration and exploitation, and gradually sway the exploration choices towards more promising alternatives.
-\end{itemize}
-
-In isolation, \ie in the absence of competition, these classes are fairly well-understood.
-Greedy algorithms are terrible for a wide variety of problem instances, precisely because they never explore.
-Exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances.
-Adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances.
-Generally,  ``better" algorithms are better in the long run, but could be worse initially.
-
-\subsection{Our Results}
-\label{sec:intro-results}
-
-We offer a mix of theoretical results and numerical simulations.
-We are mainly interested in qualitative differences between the three classes of algorithms in Section~\ref{sec:intro-model}.
-For numerical simulations, we pick one representative algorithm from each class.
-For theoretical results, we allow arbitrary algorithms and focus on asymptotic differences in the algorithms' performance.
-
-\xhdr{Theoretical results.}
-%We endow agents with Bayesian rationality, a common modeling approach for a theoretical investigation.
-We consider a basic Bayesian model.
-We posit that agents have a common Bayesian prior on reward distributions, and know the principals' algorithms.
-For simplicity, agents do not receive any information about the previous agents' choices and rewards.
-Each agent knows the round (s)he arrives in, computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
-We refer to this variant as the \emph{\TheoryModel}.
-
-Our results depend crucially on the agents' decision rule.
-\begin{itemize}
-
-\item The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
-We find that \HardMax is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
-Further, we show that \HardMax is very sensitive to tie-breaking: if the tie-breaking is probabilistically biased in favor of one principal, then this principal has a simple ``winning strategy" no matter what the other principal does.
-
-\item We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
-We call this model \HardMaxRandom.
-We find that better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
-However, there is a substantial caveat: one can defeat any algorithm by interleaving it with the greedy algorithm.
-This has two undesirable corollaries: a better algorithm may sometimes lose in competition, and a pure Nash equilibrium typically does not exist.
-
-\item We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
-For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
-This is the most technical result of the paper.
-The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm would attract slightly more).
-\end{itemize}
-
-All results extend to a much more general version of the multi-armed bandit problem in which the principal may observe additional feedback before and/or after each decision, as long as the feedback distribution does not change over time.
-In most results, principal's utility may depend on both the market share and agents' rewards.
-
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
-      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
-      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
-      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
- \end{tikzpicture}
-
-\caption{Inverted-U relationship between competitiveness and algorithms.}
-\label{fig:inverted-U}
-\end{center}
-\end{figure}
-
-\xhdr{Economic interpretation: the inverted-U relationship.}
-Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the inverted-U relationship between competition and innovation, as in Figure~\ref{fig:inverted-U}.%
-\footnote{Here, ``innovation" refers to adoption of a better technology at a substantial R\&D expense to a given firm.
-It is not salient whether similar ideas and/or technologies already exist elsewhere.
-Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
-This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{Aghion-QJE05,Vives-08}.
-
-Our decision rules differ in terms of rationality: from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
-The same distinctions also control the severity of competition between the principals: from cut-throat competition with \HardMax to a more relaxed competition with \HardMaxRandom, to an even more relaxed competition with \SoftMaxRandom.
-Indeed, with \HardMax you lose all customers as soon as you fall behind in performance, with \HardMaxRandom you get some small market share no matter what, and with \SoftMaxRandom you are further guaranteed a market share close to $\tfrac12$ as long as your performance is not much worse than the competition.
-The uniform choice among principals corresponds to no rationality and no competition.
-
-We identify the inverted-U relationship driven by the rationality/competitiveness distinctions outlined above: from \HardMax to \HardMaxRandom to \SoftMaxRandom to \Uniform.
-We also find another, technically different inverted-U relationship which zeroes in on the \HardMaxRandom model.
-We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
-
-These inverted-U relationships are driven by different aspects in our model than the ones in the existing literature in economics.
-The latter focuses on the tradeoff between the R\&D costs and the benefits that the improved technology provides in the competition.
-In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
-
-\xhdr{Numerical simulations.}
-We consider a basic frequentist model.
-We posit that the agents observe signals about the principals' past performance, and base their decisions on these signals alone, without invoking any prior knowledge or beliefs.
-The performance signals are abstracted and aggregated as a scalar \emph{reputation score} for each principal, modeled as a sliding window average of its rewards.
-Thus, agents' decision rule depends only on the two reputation scores.
-We refer to this variant as the \emph{\ExptsModel}.%
-%\footnote{In comparison, the theoretical results focus on another extreme, with Bayesian rationality and no performance signals.}
-
-We refine and expand the theoretical results in several ways.
-
-\begin{itemize}
-\item We compare \HardMax and \HardMaxRandom decision rules.
-We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.
-As predicted by the theory, better algorithms prevail if the expected number of ``random" users is sufficiently large.
-However, this effect is negligible for smaller parameter values.
-
-%\footnote{\asedit{Reputation scores already introduce some noise into users' choices.
-However, the amount of noise due to this channel is typically small, both in our simulations and in practice, because reputation signals average over many datapoints.}}
-
-\item We investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
-(We focus on the \HardMax decision rule.) We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous arrival).
-Consumer welfare is higher under early/late arrival than under simultaneous entry.
-We frame these results in terms inverted-U relationships.
-
-%\footnote{\asedit{We consider the ``permanent monopoly" scenario for comparison only, without presenting any findings.
-We just assume that a monopolist chooses the greedy algorithm, because it is easier to deploy in practice.
-Implicitly, users have no ``outside option": the service provided is an improvement over not having it (and therefore the monopolist is not incentivized to deploy better learning algorithms).
-This is plausible with free ad-supported platforms such as Yelp or Google.}}
-
-\item We investigate the algorithms' performance ``in isolation", \ie without competition.
-We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.
-We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes \emph{not} a good predictor for the outcomes under competition.
-
-\item We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}).
-We run additional experiments so as to isolate and compare these two effects.
-We find that either effect alone leads to a significant advantage under competition.
-The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.
-Finally, we find an ``amplification effect" of the data advantage: even a small amount thereof gets amplified under competition, causing a large difference in eventual market shares.
-
-\end{itemize}
-
-\ascomment{One downside of the story above: we may need to explain why we didn't simulate \SoftMaxRandom.}
-
-\ascomment{Guy: "anti-trust xhdr" synthesises several points from various places, pls see if I got it right.}
-
-\xhdr{Economic interpretation: anti-trust perspective.}
-Data advantage is significant as a possible barrier to entry in the digital economy
-\citep[\eg][]{jullien2019economics,de2020data, hagiu2020data}.
-Our model directly speaks to this literature as we endogenize data-driven network effects by explicitly considering the dynamics of learning under competition.
-Our results are similar to those on ``market tipping" because of indirect network effects \ascomment{some citations??}, and show channels through which differences in data ownership can result in market tipping.
-We observe two important differences compared to prior work:
-
-\begin{itemize}
-\item What matters is not just the quantity of data but its quality: how informative it is for the learning problem at hand. Data collected by better bandit algorithms tends to be more informative.
-
-    \ascomment{Guy: cite prior work that only talks about data quantity?
-Are we really the first paper on anti-trust that talks about data quality?
-Seems like a big claim, we better be right on this.
-}
-
-\item The ``amplification effect" of the data advantage runs contrary to prior work on learning without competition \citep{varian2018artificial,lambrecht2015can,bajari2018impact}, which found that small amounts of additional data do not provide significant improvement in eventual outcomes.
-\end{itemize}
-
-\noindent We conclude that competition dynamics are pertinent for this discussion.
-
-\ascomment{We could, in principle, move both "Econ interpretation" xhdrs to a separate subsection.
-But the current order of presentation seems better to me: we can explain the inverted-U right after "theory", before we put a whole bunch of simulation results on the stack.
-And the anti-trust stuff goes nicely right after the numerical results, without a big subsection break.}
-
-\subsection{Further discussion}
-\label{sec:discussion}
-
-\xhdr{Significance.}
-Our theory takes a basic Bayesian approach, a standard perspective in economic theory, and discovers several strong asymptotic results.
-The numerical simulations provide a more nuanced and ``non-asymptotic" perspective.
-In essence, we look for substantial effects within relevant time scales.
-(In fact, we start our investigation by determining what time scales are relevant in the context of our model.)
-
-\ascomment{The next para makes a point that seems important (to me). Is this point clear?}
-
-Our study has a dual purpose: (i) shed light on real-world implications of some typical scenarios, and (ii) investigate the space of models for describing the real world.
-As an example to clarify the latter point, consider the \HardMax model with simultaneous entry.
-It is is not necessarily the most realistic model, but arguably the most natural one to study \emph{a priori}.
-However, our results elucidate the need for more refined models with ``free exploration" (\eg via random agents or early entry).
-
-\xhdr{Assumptions.}
-Our models are stylized in several important  respects, some of which we discuss below.
-Firms compete only on the quality of service, rather than, say, pricing or the range of products.
-Agents are myopic: they do not worry about how their actions impact their future utility.
-\footnote{In particular, agents do not attempt to learn over time, to second-guess or game future agents, or to manipulate the principals' learning algorithms.
-This is arguably typical in practice, in part because one agent's influence tends to be relatively small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores", and further simplified to average performance over a time window.
-On the machine learning side, our setup is geared to distinguish between ``simple" vs.
-``better" vs.
-``smart" bandit algorithms; we are not interested in state-of-art algorithms for very realistic bandit settings.
-
-We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs.
-For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior.
-Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
-Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
-It is unclear how to incorporate performance signals in a theoretically tractable model.
-For the numerical results, the \ExptsModel accounts for competition in a more direct way, allows to separate reputation vs.
-data advantage, and makes our model amenable to numerical simulations.
-Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.
-
-\xhdr{Challenges.}
-Much of the challenge in the theoretical results, both conceptual and technical, was in setting up the model and the theorems.
-Apart from zeroing in on the \TheoryModel, it was crucial to interpret the results and intuitions from the literature on multi-armed bandits so as to formulate meaningful and productive assumptions on bandit algorithms and Bayesian priors.
-
-The numerical investigation is quite challenging even with a stylized model such as ours.
-An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given instance of a multi-armed bandit problem (and each such experiment is run many times to reduce variance).
-Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit instances\}, and we are looking for findings that are consistent across this entire space.
-It is essential to keep each of the three dimensions small yet representative.
-In particular, we need to capture a huge variety of bandit instances with only a few representative examples.
-Further, we need a succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).
-
-While amenable to simulations, the \ExptsModel appears difficult to analyze. This is for several reasons:
-%
-(i) intricate feedback loop from performance to reputations to users to performance;
-%
-(ii) mean reputation, most connected to our intuition, is sometimes a bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited}); and
-%
-(iii)
-mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice.
-Given the theoretical results on the \TheoryModel, and the fact that we are in the realm of stylized economic models, we believe that resolving first-order theoretical questions about the \ExptsModel would not add much value to this paper.
-
-\ascomment{Intro ends here.}
-
-\newpage
-\section{STUFF FROM THE PREVIOUS VERSION}
+\section{STUFF FROM THE PREVIOUS VERSION OF THE INTRO}
 
@@
 
-\subsection{Discussion}
+\subsection*{Discussion}
 \label{sec:intro-discussion}
@@
 Section ~\ref{sec:conclusion} concludes.
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "main"
-%%% End:

### 2020-07-07 journal paper/content/sec-intro.tex
Moved the stuff from Guy's intro to a separate file.

@@
 
-\ascomment{Intro ends here.}
-
-\newpage
-\section*{STUFF FROM THE PREVIOUS VERSION}
-
-\ascomment{Left the previous version for now (from Guy).}
-
-\xhdr{Main results.}
-We start with a basic version in which the principals are symmetric (\ie do not have an inherent advantage over one another) and the agents are utility-maximizers.
-We find that competition induces firms to choose the greedy algorithm.
-
-commit to a greedy (myopic) algorithm in equilibrium that does no purposeful exploration.
-This leads to low consumer welfare since the greedy algorithm is known to be dramatically bad in many important cases of multi-armed bandits.
-The primary mechanism that generates this stark result is that consumers need to be incentivized to select a firm over its competitors, leading a firm that engages in exploration to be starved of consumers before it makes enough progress on its learning problem.
-In order to incentivize ``better" exploration strategies in equilibrium, the key intuition is that the firm needs to have some ``free" consumers that visit them without the firm having to incentivize them to do so.
-This allows the firm to eventually overcome the initial losses in consumer perception from exploration.
-We explore two economic mechanisms that can generate such an effect.
-
-The first is the presence of users who randomly choose between the firms which gives each firm a constant stream of ``free" users and thus ensures it never gets fully starved.
-We find that, even when there is a small fraction of such consumers, better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.\footnote{While the precise notion of ``sufficiently better algorithm" is rather subtle, we note that commonly known ``smart" bandit algorithms typically defeat the commonly known ``naive" ones, and the latter typically defeat the greedy algorithm.
-However, there is a substantial caveat: one can defeat any algorithm by interleaving it with the greedy algorithm.
-This has two undesirable corollaries: a better algorithm may sometimes lose, and a pure Nash equilibrium typically does not exist.} However, while this holds asymptotically, our numerical experiments show that this requires either unreasonably large time scales or a large fraction of such users.
-
-The second is that if one firm has a first-mover advantage then this firm gets free users before the other firm enters into the market.
-We show that if this incumbency period is sufficiently long, then it allows the firm to make sufficient progress on its learning problem and incentivizes it to commit to better algorithms.
-While this leads to the incumbent getting almost all of the market, it leads to higher consumer welfare than under competition when the firms enter at the same time.
-
-\xhdr{\underline{Additional Findings}}
-
-\xhdr{Reputation and Data Advantage}: In the reputation choice variant we investigate the first-mover advantage phenomenon in more detail.
-Being first in the market gives the incumbent free data to learn from (a ``data advantage") as well as a more definite, and possibly better reputation compared to an entrant (a ``reputation advantage").
-We run additional experiments so as to isolate and compare these two effects.
-We find that either effect alone leads to a significant advantage under competition.
-The data advantage is larger than reputation advantage when the incumbent commits to a more advanced bandit algorithm.
-This result shows that even a small amount ``data advantage" gets amplified under competition, causing a large difference in eventual market shares.
-
-\xhdr{Noise in consumer choice}: We further relax the decision rule of the agents so that the probability of choosing a given firm varies smoothly as a function of the difference between  principals' expected rewards.
-For this decision rule, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
-This is the most technical result of the paper.
-The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm may attract slightly more).
-
-\xhdr{Predicting Outcomes in Competition} We also investigate how algorithms' performance ``in isolation" (without competition) is predictive of the outcomes under competition in the reputation choice variant.
-We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes not a good predictor.
-We suggest a more refined performance measure, based on a comparison between the reputation of the two firms, and use it to explain some of the competition outcomes.
-\textbf{Add some discussion about comparison with ``better" algorithms in analytical part}
-
-\OMIT{All results extend to a much more general version of the multi-armed bandit problem in which the principal may observe additional feedback before and/or after each decision, as long as the feedback distribution does not change over time.
-In most results, principal's utility may depend on both the market share and agents' rewards.
-}
-
-\subsection{Discussion}
-\label{sec:intro-discussion}
-
-\xhdr{Economic interpretation.} Our model speaks to two distinct strands of the economics literature.
-
-The first is the literature that studies the relationship between competition and innovation where, in our context,  the choice between different algorithms is interpreted as firms determining whether to invest in a ``better" technology as the degree of competition varies.\footnote{The choice of algorithms has a clear analogue to ``innovation" in our context as the literature on multi-armed bandits has identified distinct ``classes" of algorithms according to asymptotic regret bounds.
-Thus, choosing algorithms in ``better" classes corresponds with innovation on the part of the firm in our interpretation of the model.
-It is not salient whether similar ideas and/or technologies already exist outside the firm.
-It is worth noting that the adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature; see \citet{MWT-WhitePaper-2016} for an example of such R\&D effort.} A classic result in this literature is that there is an inverted-U relationship between the severity of competition among firms and the quality of technologies that they adopt is a familiar theme in the economics literature \citep[\eg][]{Aghion-QJE05,Vives-08}.
-
-We find it illuminating to frame our contributions in a similar manner, as illustrated in \reffig{fig:inverted-U}.
-In our model we consider varying the intensity of competition as varying the degree to which a firm needs to incentivize users to visit them.
-Competition is less intense when there are more ``random" users or one firm has a sufficiently long first-mover advantage.
-The inverted-U relationship in our model arise for a fundamentally different reason, compared to the existing literature on ``competition vs.
-innovation.'' In the literature, better technology always helps in a competitive environment, other things being equal.
-Thus, the trade-off is between the costs of improving the technology and the benefits that the improved technology provides in the competition.
-Meanwhile, we find that a better exploration algorithm may sometimes perform much worse under competition, even in the absence of R\&D costs.
-This stems from the nature of exploration technologies in online markets which rely on learning from interactions with users.
-This leads to an implicit cost from exploration in the form of a reduced rate of users that a firm attracts and can learn from.
-
-Interestingly, the economic mechanism in our model for incentivizing firms to engage in R\&D has a qualitative similarity to the role that patents play in incentivizing innovation in standard R\&D models.
-In these models, patents temporarily relax competition for the innovating firm by giving them exclusive access to their innovation for a limited period of time in order to incentivize them to invest in the better technology.
-In our model, temporarily relaxing competition in the form of giving firms free periods to learn incentivizes the firm to invest in the better technology.
-
-The second is the nascent literature which studies the economics of data and how data can serve as a barrier to entry in the digital economy \cite{de2020data, hagiu2020data}.
-Our model directly speaks to this literature as we endogenize data-driven network effects by explicitly considering a model where firms are simultaneously solving a machine learning problem while competing against each other for users.
-The results of our model point to the fact that competition dynamics -- that firms compete as they learn over time and that this competition endogenously determines the data observed by firms -- are pertinent to understanding whether or not data can serve as a barrier to entry.
-Indeed, our results are similar to the market tipping results in the literature on indirect network effects and show a channel through which differences in data ownership can result in market tipping.
-However, it is important to note that the channel through which this arises in our model is not simply from differences in the quantity of data possessed by the firms, but also from the quality of the data.
-
-\xhdr{Model variants.}
-We consider two separate variants of the model in order to provide a more thorough investigation of the tension between exploration and competition.
-The expectation choice model is tractable enough to allow us to obtain theoretical results with ``asymptotic" flavor.
-However, for the sake of analytical tractability, we make the unrealistic simplification that users do not observe any signals about firms' ongoing performance and it is difficult to analyze important economic mechanisms such as asymmetries in the timing of entry.
-
-The reputation choice model relaxes this simplification and accounts for competition in a more direct way as well as allows us to explore other relevant economic mechanisms in understanding the tension between exploration and competition.
-However, it now becomes considerably harder to analyze analytically.
-This is for several reasons: intricate feedback loop from performance to reputations to users to performance;
-%
-mean reputation, most connected to our intuition, is sometimes a bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited});
-%
-mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice.
-We therefore analyze our model using numerical simulation, which has several benefits.
-It allows us to analyze our model from a ``non-asymptotic" perspective, looking for substantial effects within relevant time scales.
-Indeed, we start our investigation by determining what time scales are relevant in the context of this variant of the model.
-Further, it allows us to investigate important economic mechanisms that arise in environments where exploration and competition tensions are at play, such as the effect of incumbency, increasing the number of firms, and the extent to which data can serve as a barrier to entry.
-
-\xhdr{Map of the paper.}
-We survey related work (Section~\ref{sec:related-work}), lay out the model and preliminaries (Section~\ref{sec:model}).
-We start by analyzing the expectation choice model analytically.
-In Sections ~\ref{sec:rational},~\ref{sec:random}, and ~\ref{sec:soft} we analyze the expectation choice variant and characterize the equilibrium behavior under three different agent decision rules.
-Then, we turn to analyze the reputation choice variant of the model with details of the analysis described in Section ~\ref{sec:sim_details}.
-Sections ~\ref{sec:isolation} and ~\ref{sec:revisited} overview results from running different bandit algorithms in isolation (i.e.
-without competition).
-Sections ~\ref{sec:competition} and ~\ref{sec:barriers} overview the results of the reputation choice variant.
-Section ~\ref{sec:non_greedy} presents results of varying the consumer choice rule in the reputation choice variant.
-Section ~\ref{sec:conclusion} concludes.
-
 %%% Local Variables:

### 2020-07-07 journal paper/content/sec-intro.tex
Moved/reworded the note re discrepancy on HardMax.

@@
 \begin{itemize}
-\item We compare \HardMax and \HardMaxRandom decision rules. We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.%
-\footnote{The greedy algorithm wins often but not always, whereas it
-always wins under \TheoryModel. The reason for this discrepancy is that the greedy algorithm in the theoretical results is aware of the Bayesian prior.
-}
+\item We compare \HardMax and \HardMaxRandom decision rules. We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier. 
  As predicted by the theory, better algorithms prevail if the expected number of ``random" users is sufficiently large. However, this effect is negligible for smaller parameter values.
@@
 \item We investigate the algorithms' performance ``in isolation", \ie without competition.
-We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.
+We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.%
+    \footnote{Note the discrepancy with our theoretical results on \HardMax: there, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations the prior is not available).}
      We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes \emph{not} a good predictor for the outcomes under competition.

### 2020-07-07 journal paper/content/sec-theory-proofs.tex
Another quick pass through the theory section.

@@
 \label{sec:proofs-HM-main}
-
-The proof starts with two auxiliary lemmas: that deviating from \DynGreedy implies a strictly smaller Bayesian-expected reward, and that \HardMax implies a ``sudden-death" property: if one agent chooses principal $1$ with certainty, then so do all subsequent agents do.
-{We re-use both lemmas in later sections, so we state them in sufficient generality.}
-
-\begin{lemma}\label{lm:DG-rew}
-{Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$. Then $\rew_1(n_0)>\rew_2(n_0)$. This holds for any response function $\respF$.}
-\end{lemma}
-
-{Lemma~\ref{lm:DG-rew} does not rely on any particular shape of the response function because it only considers the performance of each algorithm without competition.}
 
@@
 
-\begin{lemma}\label{lm:DG-sudden}
-{Consider \HardMax response function with $\respF(0)\geq\tfrac12$.}
-Suppose \alg[1] is monotone, and $\PMR_1(t_0)>\PMR_2(t_0)$ for some global round $t_0$. Then $\PMR_1(t)>\PMR_2(t)$ for all subsequent rounds $t$.
-\end{lemma}
-
-\begin{proof}
+\begin{proof}[Proof of Lemma~\ref{lm:DG-sudden}]
 Let us use induction on round $t\geq t_0$, with the base case $t=t_0$.
@@
 %%%%%%%%%%%%
-\subsubsection{\HardMax with biased tie-breaking}
+\subsubsection{\HardMax with biased tie-breaking: 
+Proof of Theorem~\ref{thm:HardMax-biased}}
 \label{sec:proofs-HM-main}
 
-\begin{proof}[Proof of Theorem~\ref{thm:HardMax-hardTies}]
-Agent $1$ chooses principal $1$ because of the tie-breaking rule.
-Since \StaticGreedy is trivially monotone, all the subsequent agents choose principal $1$ by an induction argument similar to the one in the proof of Lemma~\ref{lm:DG-sudden}.
-\end{proof}
-
-\begin{proof}[Proof of Theorem~\ref{thm:HardMax-biased}]
 The proof re-uses Lemmas~\ref{lm:DG-rew} and~\ref{lm:DG-sudden}, which do not rely on fair tie-breaking.
@@
 Therefore, agent $m_0$ chooses principal $1$, and by Lemma~\ref{lm:DG-sudden} so do all subsequent agents.
-\end{proof}
 
@@
 
-\xhdr{Proof Sketch.}
-The main idea is that even though \alg[1] may have a
-slower rate of learning in the beginning, it will gradually catch up
-and surpass \alg[2]. We will describe this process in two phases. In
-the first phase, \alg[1] receives a random agent with probability at
-least $\respF(-1) = \eps_0$ in each round. Since $\BIR_1$ tends to 0,
-the difference in \BIR{s} between the two algorithms is also
-diminishing. Due to the \SoftMaxRandom response function, \alg[1]
-attracts each agent with probability at least $1/2 - O(\beta_0)$ after
-a sufficient number of rounds. Then the game enters the second phase:
-both algorithms receive agents at a rate close to $\tfrac12$, and the
-fractions of agents received by both algorithms --- $n_1(t)/t$ and
-$n_2(t)/t$ --- also converge to $\tfrac12$. At the end of the second
-phase and in each global round afterwards, the counts $n_1(t)$ and
-$n_2(t)$ satisfy the weak BIR-dominance condition, in the sense that
-they both are larger than $n_0$ and $n_1(t)\geq (1-\beta_0)\; n_2(t)$.
-At this point, \alg[1] actually has smaller $\BIR$, which reflected in the {\PMR}s eventually. Accordingly, from then on \alg[1]
-attracts agents at a rate slightly larger than $\tfrac12$. We prove
-that the ``bump'' over $\tfrac12$ is at least on the order of
-$\BIR_2(t)$.
-
 %\begin{proof}[Proof of Theorem~\ref{thm:SoftMax-strong}]
-\xhdr{Full proof.}
   Let $\beta_1 = \min\{c_0'\delta_0, \beta_0/20\}$ with $\delta_0$

### 2020-07-07 journal paper/content/sec-theory.tex
Another quick pass through the theory section.

@@
-\ascomment{some intro para goes here.}
+In this section, we present our theoretical results for the \TheoryModel. While we provide intuition and (some) proof sketches, the detailed proofs are deferred to Section~\ref{sec:theory-proofs}.
 
@@
 
-We consider the version in which the agents are fully rational, in the sense that their response function is \HardMax.
+We posit fully rational agents, in the sense that their response function is \HardMax.
 We show that principals are not incentivized to \emph{explore}--- \ie to deviate from \DynGreedy.
 The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates.
+To make this more precise, let us say that two MAB algorithms \emph{deviate} at (local) step $n$ if
+\begin{align}
+\Pr\sbr{ \alg[1](n)\neq \alg[2](n)} >0.
+\end{align}
 
-To make this more precise, let us say that two MAB algorithms \emph{deviate} at (local) step $n$ if there is an action $a\in A$ and a set of step-$n$ histories of positive probability such that any history $h$ in this set is feasible for both algorithms, and under this history the two algorithms choose action $a$ with different probability.
+\ascomment{Steven: we used to have a much more cumbersome definition (commented out), not sure why!}
+
+\ascomment{Steven: in the thm, can we just say "deviates at $n_0$" instead of "deviates starting from $n_0$"? What do we mean by the latter, anyway?}
+
+%there is an action $a\in A$ and a set of step-$n$ histories of positive probability such that any history $h$ in this set is feasible for both algorithms, and under this history the two algorithms choose action $a$ with different probability.
 
@@
 \begin{corollary}\label{cor:DG-dominance}
-The competition game between principals has a unique Nash equilibirium: both principals choose \DynGreedy.
+The competition game has a unique Nash equilibirium: both principals choose \DynGreedy.
 \end{corollary}
 
-\begin{remark}
-This corollary holds under a more general model which allows time-discounting: namely, the utility of each principal $i$ in each global round $t$ is $U_{i,t}(r_t)$ if this principal is chosen, and $0$ otherwise, where $U_{i,t}(\cdot)$ is an arbitrary non-decreasing function with $U_{i,t}(0)>0$.
-\end{remark}
+The proof of Theorem~\ref{thm:DG-dominance} relies on two key lemmas: that deviating from \DynGreedy implies a strictly smaller Bayesian-expected reward, and that \HardMax implies a ``sudden-death" property: if one agent chooses principal $1$ with certainty, then so do all subsequent agents.
+We re-use both lemmas in later results, so we state them in sufficient generality.
+In particular, Lemma~\ref{lm:DG-rew} works for any response function because it only considers the performance of each algorithm without competition.
+
+\begin{lemma}\label{lm:DG-rew}
+Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$. Then $\rew_1(n_0)>\rew_2(n_0)$. This holds for any response function $\respF$.
+\end{lemma}
+
+\begin{lemma}\label{lm:DG-sudden}
+Consider \HardMax response function with $\respF(0)\geq\tfrac12$.
+Suppose \alg[1] is monotone, and $\PMR_1(t_0)>\PMR_2(t_0)$ for some global round $t_0$. Then $\PMR_1(t)>\PMR_2(t)$ for all subsequent rounds $t$.
+\end{lemma}
+
+The remainder of the proof of Theorem~\ref{thm:DG-dominance} uses the conclusion of one lemma to derive the precondition for another, \ie goes from $\rew_1(n_0)>\rew_2(n_0)$ to $\PMR_1(n_0)>\PMR_2(n_0)$.
+The subtlety one needs to deal with is that the principal's ``local" round corresponding to a given ``global" round is a random quantity due to the random tie-breaking.
 
@@
 \end{theorem}
+
+\begin{proof}[Proof Sketch]
+Agent $1$ chooses principal $1$ because of the tie-breaking rule.
+Since \StaticGreedy is trivially monotone, all the subsequent agents choose principal $1$ by an induction argument similar to the one in the proof of Lemma~\ref{lm:DG-sudden}.
+\end{proof}
 
@@
 
+The proof re-uses Lemmas~\ref{lm:DG-rew} and~\ref{lm:DG-sudden}, which do not rely on fair tie-breaking.
+
+\ascomment{Steven: can you come up with a short proof sketch / intuition why this thm works?}
+
 %%%%%%%%%%
@@
 
-We need to express a property that \alg[1] eventually catches up and surpasses \alg[2], even if initially it receives only a fraction of traffic.
+To state this result, we need to express a property that \alg[1] eventually catches up and surpasses \alg[2], even if initially it receives only a fraction of traffic.
 We assume that both algorithms run indefinitely and do not depend on the time horizon $T$; call such algorithms \emph{$T$-oblivious}.
@@
 
-To state a game-theoretic corollary of Theorem~\ref{thm:random-clean}, we restrict the set of feasible MAB algorithms:
+\ascomment{Steven: can you come up with a short proof sketch / intuition why this thm works?}
+
+To state the game-theoretic implications of Theorem~\ref{thm:random-clean}, we restrict the set of feasible MAB algorithms:
 
@@
 step $n_0$ and a ``mixing'' parameter $p$, we will construct a
-modified algorithm as follows.
+modified algorithm by ``mixing in" some greedy choices:
 \begin{enumerate}
@@
   given the current information collected by \alg[1].
+
+\item The modified algorithm does not record rewards received in steps with the ``greedy choice".%
+    \footnote{This is probably not essential, but makes for a cleaner comparison between the two algorithms.}
+
 \end{enumerate}
 %The modified algorithm is called \emph{greedy deviation}, with probability parameter $p$.
-For a cleaner comparison between the two algorithms, the modified algorithm does not record rewards received in steps with the ``greedy choice". Parameter $p>0$ is the same for all steps.
+%Parameter $p>0$ is the same for all steps.
+
+\ascomment{Steven: why can't we start `mixing' at $n=1$?}
+
+Note that the modified algorithm is another pure strategy in the competition game between the two principals, not a mixture of \alg[1] and the greedy algorithm.
 
@@
   where both principals choose \DynGreedy. Moreover, no pure Nash
-  equilibrium exists when some algorithm ``dominates" \DynGreedy in
-  the sense of \eqref{eq:random-better-messy} and the time horizon $T$
-  is sufficiently large.
+  equilibrium exists when some algorithm BIR-dominates \DynGreedy and the time horizon $T$ is sufficiently large.
 \end{corollary}
@@
 
+\ascomment{Steven: in this thm and the next, do we need the algorithms to be monotone? We didn't assume this in the conf paper, which I assume was a minor bug.}
+
 \begin{proof}[Proof Sketch]
@@
 
+\ascomment{Steven: in the conf paper, we spell out the defn of $\BIR_1(n) \to 0$. We don't need to, do we?}
+
+\begin{proof}[Proof Sketch]
+The main idea is that even though \alg[1] may have a
+slower rate of learning in the beginning, it will gradually catch up
+and surpass \alg[2]. We distinguish two phases. In
+the first phase, \alg[1] receives a random agent with probability at
+least $\respF(-1) = \eps_0$ in each round. Since $\BIR_1$ tends to 0,
+the difference in \BIR{s} between the two algorithms is also
+diminishing. Due to the \SoftMaxRandom response function, \alg[1]
+attracts each agent with probability at least $1/2 - O(\beta_0)$ after
+a sufficient number of rounds. Then the game enters the second phase:
+both algorithms receive agents at a rate close to $\tfrac12$, and the
+fractions of agents received by both algorithms --- $n_1(t)/t$ and
+$n_2(t)/t$ --- also converge to $\tfrac12$. At the end of the second
+phase and in each global round afterwards, the counts $n_1(t)$ and
+$n_2(t)$ satisfy the weak BIR-dominance condition, in the sense that
+they both are larger than $n_0$ and $n_1(t)\geq (1-\beta_0)\; n_2(t)$.
+At this point, \alg[1] actually has smaller $\BIR$, which reflected in the {\PMR}s eventually. Accordingly, from then on \alg[1]
+attracts agents at a rate slightly larger than $\tfrac12$. We prove
+that the ``bump'' over $\tfrac12$ is at least on the order of
+$\BIR_2(t)$.
+\end{proof}
+
 As in the previous subsection, we can state a game-theoretic corollary for the restricted competition game (as in Definition~\ref{def:restricted-competition}).
@@
 \footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition.
-\ascomment{explain why, cite smth?}} 
+\ascomment{Steven, Yishay: how do we explain why? Pls help ...}}
 Suppose a feasible algorithm \alg BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.
@@
 
+Corollary~\ref{cor:DG-dominance} holds under a more general model which allows time-discounting: namely, the utility of each principal $i$ in each global round $t$ is $U_{i,t}(r_t)$ if this principal is chosen, and $0$ otherwise, where $U_{i,t}(\cdot)$ is an arbitrary non-decreasing function with $U_{i,t}(0)>0$.
+
 \xhdr{Dependence on the time horizon.}

### 2020-07-07 journal paper/content/sec-theory.tex
Expanded the footnote re "a little greedy".

@@
 
-\item The modified algorithm does not record rewards received in steps with the ``greedy choice".%
-    \footnote{This is probably not essential, but makes for a cleaner comparison between the two algorithms.}
+\item The modified algorithm does not record the ``greedy choice" steps in its data set.%
+    \footnote{In other words: in the subsequent rounds, as far as the modified algorithm is concerned, the ``greedy choice" steps have never happened.
+While it is usually more efficient to consider all available data, this simplification enables a cleaner comparison between the two algorithms.
+Also, it makes for  a cleaner setup: otherwise, just to make the modified algorithm well-defined, we'd need to assume that \alg[1] is a mapping from step-$n$ histories to actions, for each step $n$.
+This is usually OK -- most bandit algorithms have this shape anyway -- but it prohibits \alg[1] to correlate its internal randomness across steps.}
 

### 2020-07-07 journal paper/content/sec-intro.tex
another pass at economic interpretation

@@
 
-\xhdr{Economic interpretation: anti-trust perspective.}
-Data advantage is significant as a possible barrier to entry in the digital economy
-\citep[\eg][]{jullien2019economics,de2020data, hagiu2020data}.
-Our model directly speaks to this literature as we endogenize data-driven network effects by explicitly considering the dynamics of learning under competition.
-Our results are similar to those on ``market tipping" because of indirect network effects \ascomment{some citations??}, and show channels through which differences in data ownership can result in market tipping.
-We observe two important differences compared to prior work:
+\xhdr{Economic interpretation: data network effects.}
+\gaedit{Our model speaks to the ongoing debate of the role that data plays as a competitive advantage in the digital economy.
+One fundamental question in this debate is whether data serves as a network effect, whereby competition for the market is more important than competition in the market due to the winner-take-all nature of such markets \citep{farrell2007coordination, jullien2019economics}.
+We find a similar effect in our setting where slight asymmetries between the firms leads to a winner-take-all outcome.
+Indeed, we find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market.
 
-\begin{itemize}
-\item What matters is not just the quantity of data but its quality: how informative it is for the learning problem at hand. Data collected by better bandit algorithms tends to be more informative.
-
-    \ascomment{Guy: cite prior work that only talks about data quantity?
-Are we really the first paper on anti-trust that talks about data quality?
-Seems like a big claim, we better be right on this.
+A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them as deploying algorithms for solving such problems.
+Our results highlight that understanding the performance of learning algorithms in isolation does not necessary translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
+Thus, approaches such as \cite{bajari2018impact} that argue that the diminishing returns to scale and scope of data in isolation render such data feedback loops as non-existent ignore the differences induced by learning in isolation versus under competition.
+Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not just by data quantity, but also the increased data quality gathered by better learning algorithms.
 }
-
-\item The ``amplification effect" of the data advantage runs contrary to prior work on learning without competition \citep{varian2018artificial,lambrecht2015can,bajari2018impact}, which found that small amounts of additional data do not provide significant improvement in eventual outcomes.
-\end{itemize}
-
-\noindent We conclude that competition dynamics are pertinent for this discussion.
 

### 2020-07-08 journal paper/content/related_work.tex
tweaks to paper
updated related work + intro

--- 2019-06-17 journal paper/itcs18 paper/sec-related-work.tex
@@
 algorithms that combine exploration and exploitation, such as UCB1
-\citep{bandits-ucb1} and Successive Elimination
+\citep{bandits-ucb1}, Thompson Sampling \citep{russo2018tutorial}, and Successive Elimination
 \citep{EvenDar-icml06}, and ``naive'' MAB algorithms that separate
@@
 Also, this work specifically assumes binary payoffs (\ie win or lose) for the principals.
+\gaedit{\cite{ben2017best, ben2019regression} similarly study competing learning algorithms, though they consider competition between offline learning algorithms, whereas our focus here is on online learning algorithms.}
 
@@
 
-A line of work on \emph{platform competition}, starting with \cite{Rysman09}, concerns competition between firms (\emph{platforms}) that improve as they attract more users (\emph{network effect}); see \citet{Weyl-White-14} for a recent survey.
+\gaedit{There is a literature that studies the interaction between learning-by-doing and competition \citep{fudenberg1983learning, dasgupta1988learning, cabral1994learning}.
+In these models, firms learn while competing against each other where a firm attracting more consumers reduces its per-unit production costs.
+However, our model focuses on firms learning product quality as opposed to reducing production costs and the impact that this learning has on attracting consumers.}
+
+\gaedit{A line of work on \emph{platform competition}, starting with \cite{Rysman09}, concerns competition between firms (\emph{platforms}) that improve as they attract more users (\emph{network effect}); see \citet{Weyl-White-14} for a recent survey.
 This literature is not concerned with \innovation, and typically models network effects exogenously, whereas in our model network effects are endogenous: they are created by MAB algorithms, an essential part of the model.
+A nascent literature studies whether and when network effects manifest themselves in data-intensive markets \citep{prufer2017competing, hagiu2020data}, but typically model learning as a reduced-form function of past consumer history and focus on the role of prices, as opposed to the reputational consequences of learning.
 
-Relaxed versions of rationality similar to ours are found in several notable lines of work.
+\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers.
+We observe a similar effect when we investigate the role that reputation can serve as a barrier to entry.
+However, in our model this effect is further strengthened by the fact that the firms have to learn while competing adding that the incumbent may not only have a reputational advantage, but additionally has a further advantage due to the data it acquires in the incumbency period.
+Thus, our model also highlights the role that data can serve as a barrier to entry in online markets which has similarly been noted in \cite{lambrecht2015can, de2020data}.}
+
+Finally, relaxed versions of rationality similar to ours are found in several notable lines of work.
 For example, ``random agents" (a.k.a.
@@
 The \SoftMaxRandom model is closely related to the literature on \emph{product differentiation}, starting from \cite{Hotelling-29}, see \cite{Perloff-Salop-85} for a notable later paper.
-
 There is a large literature on non-existence of equilibria due to small deviations   (which is related to the corresponding result for \HardMaxRandom), starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets.

### 2020-07-08 journal paper/content/sec-intro.tex
updated related work + intro

--- 2020-07-07 journal paper/content/sec-intro.tex
@@
 
-\xhdr{Economic interpretation: anti-trust perspective.}
-Data advantage is significant as a possible barrier to entry in the digital economy
-\citep[\eg][]{jullien2019economics,de2020data, hagiu2020data}.
-Our model directly speaks to this literature as we endogenize data-driven network effects by explicitly considering the dynamics of learning under competition.
-Our results are similar to those on ``market tipping" because of indirect network effects \ascomment{some citations??}, and show channels through which differences in data ownership can result in market tipping.
-We observe two important differences compared to prior work:
+\xhdr{Economic interpretation: data network effects.}
+\gaedit{Our model speaks to the ongoing debate on the role that data plays in the digital economy.
+One fundamental question in this debate is whether data can serve a similar role as traditional network effects, whereby competition for the market is more important than competition in the market due to the winner-take-all nature of such markets \citep{Rysman09, jullien2019economics}.
+The ``death spiral" effect that we find has a similar flavor as this as once a firm engages in exploration it gets completely starved of users.
+Indeed, we further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market.
+A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them as deploying algorithms for solving this problem.
+Thus, we do not explicitly model such network effects, but rather they arise endogenously from our setup.
 
-\begin{itemize}
-\item What matters is not just the quantity of data but its quality: how informative it is for the learning problem at hand. Data collected by better bandit algorithms tends to be more informative.
-
-    \ascomment{Guy: cite prior work that only talks about data quantity?
-Are we really the first paper on anti-trust that talks about data quality?
-Seems like a big claim, we better be right on this.
+Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
+Approaches such as \cite{bajari2018impact} that argue that the diminishing returns to scale and scope of data in isolation render such data feedback loops as non-existent ignore the differences induced by learning in isolation versus under competition.
+Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.
 }
-
-\item The ``amplification effect" of the data advantage runs contrary to prior work on learning without competition \citep{varian2018artificial,lambrecht2015can,bajari2018impact}, which found that small amounts of additional data do not provide significant improvement in eventual outcomes.
-\end{itemize}
-
-\noindent We conclude that competition dynamics are pertinent for this discussion.
 

### 2020-07-08 journal paper/content/sec-model.tex
Minor edit.
Daily save -- half-done.

@@
-\xhdr{Principals and agents.} There are two principals and $T$ agents.
-The game proceeds in rounds (we will sometimes refer to them as \emph{global rounds}).
+We define the basic model; some extensions thereof are specified in  Section~\ref{sec:theory-extensions}.
+
+\xhdr{Principals and agents.} There are two principals and $T$ agents,
+denoted, resp., principal $i\in \{1,2\}$ and agent $t\in [T]$.
+The game proceeds in (global) rounds.
 In each round $t\in [T]$, the following  interaction takes place.
-A new agent arrives and chooses one of the two principals.
-The principal chooses a recommendation: an action $a_t\in A$, where $A$ is a fixed set of actions (same for both principals and all rounds).
-The agent follows this recommendation, receives a reward $r_t\in [0,1]$, and reports it back to the principal.
+Agent $t$ arrives and chooses a principal $i_t\in \{1,2\}$.
+The principal chooses action $a_t\in A$, where $A$ is a fixed set of actions (same for both principals and all rounds).
+The agent experiences this action, receives a reward $r_t\in \{ 0,1\}$, and reports it back to the principal.
+We posit \emph{stochastic rewards}: whenever a given action $a\in A$ is chosen, the reward is an independent draw from Bernoulli distribution with mean $\mu_a$.
+The mean rewards $\mu_a$ are initially not known to anybody.
+Each principal is completely unaware of the rounds when the other principal is chosen.
 
-The rewards are i.i.d. with a common prior. More formally, for each action $a\in A$ there is a parametric family $\psi_a(\cdot)$ of
+Each principal $i$ commits to a learning algorithm \alg[i] before round $1$, and uses this algorithm throughout.
+The algorithm follows the protocol of \emph{multi-armed bandits} (\emph{MAB}).
+Namely, it proceeds in time-steps:%
+\footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before.
+We will omit the global vs.
+local distinction when clear from the context.} each time it is called, it outputs an action from $A$, and inputs a reward for this action.
+\alg[i] is called only in global rounds when principal $i$ is chosen.
+
+\newcommand{\est}{\mathtt{est}}
+
+\xhdr{Agents' response.} Each agent $t$ forms a reward estimate $\est_i(t)$ for each principal $i$, as specified below, and chooses the principal based on these reward estimates.
+Specifically, principal $1$ is chosen with probability
+\begin{align}
+p_t = \respF\rbr{ \est_1(t) - \est_2(t) },
+\end{align}
+where $\respF:[-1,1]\to [0,1]$ is the \emph{response function}, same for all agents and known to everyone.
+
+We assume that $\respF$ is monotonically non-decreasing, is larger than $\nicefrac12$ on the interval $(0,1]$, and smaller than $\nicefrac12$ on the interval $[-1,0)$.
+We consider three specific models (see \reffig{fig:response-functions}):
+
+\begin{figure}
+\begin{center}
+  \begin{tikzpicture}[scale=2.0]
+    \draw[->] (-1.1,0) -- (1.1,0) node[right]
+    {$\Delta_t := \est_1(t) - \est_2(t)$};
+    \draw[->] (0,-0.1) -- (0,1.1) node[above]
+        {$p_t = \text{prob. of choosing principal 1}$};
+    \draw[scale=1.0,domain=-1:0,smooth,variable=\q,blue, line width=0.50mm] plot ({\q},{0});
+    \draw[scale=1.0,domain=0:1,smooth,variable=\q,blue,line width=0.50mm] plot ({\q},{1});
+    \draw[scale=1.0,domain=-1:0,smooth,variable=\y,red]  plot ({\y},{0.1});
+    \draw[scale=1.0,domain=0:1,smooth,variable=\y,red]  plot ({\y},{0.9});
+    \draw[scale=1.0,domain=-1:1,smooth,variable=\y,cyan, line width=0.45mm, dash pattern=on 3pt off 2pt]  plot ({\y},{1/(1 + 1/(9^\y))});
+    % \node[above, blue] at (0.5, 0.5) {\footnotesize $2 q (1 - q)$};
+     \node[left] at (0, 0.5) {\footnotesize $1/2$};
+     \node[left] at (0, 1) {\footnotesize $1$};
+     \node[below left] at (0, 0) {\footnotesize $0$};
+     \node[below ] at (1, 0) {\footnotesize $1$};
+     \node[below ] at (-1, 0) {\footnotesize $-1$};
+  \end{tikzpicture}
+\end{center}
+\caption{The three models for $\respF$: \HardMax is thick blue, \HardMaxRandom is slim red, and \SoftMaxRandom is the dashed curve.}
+\label{fig:response-functions}
+\end{figure}
+
+\begin{itemize}
+\item \HardMax: $\respF$ equals $0$ on the interval $[-1,0)$ and $1$
+  on the interval $(0,1]$. In words, a \HardMax agent
+  deterministically chooses a principal with a higher reward estimate.
+
+\item \HardMaxRandom:
+    % $\respF$ equals $\eps$ on the interval $[-1,0)$ and $1-\eps'$ on the interval $(0,1]$, where $\eps,\eps'\in (0,\tfrac12)$ are some positive constants.
+In words, each agent is a \HardMax agent with probability $1-\eps-\eps'$, and with the remaining probability she makes a random choice.
+    $\respF$ equals $\eps_0$ on the interval $[-1,0)$ and $1-\eps_0$ on the interval $(0,1]$, for some constant $\eps_0\in (0,\tfrac12)$.
+In words, each agent is a \HardMax agent with probability $1-2\eps_0$, and makes a random choice otherwise.
+
+\item \SoftMaxRandom: $\respF(\cdot)$ is bounded away from $0$ and $1$, and ``smooth" around $0$ (see Section~\ref{sec:theory-SoftMax}).
+\end{itemize}
+% \asedit{(We assume $\respF(-1)+\respF(1)=1$ in \HardMaxRandom and \SoftMaxRandom only to simplify notation.)}
+
+%We say that $\respF$ is \emph{symmetric} if $\respF(-x)+\respF(x)=1$ for any $x\in [0,1]$. This implies \emph{fair tie-breaking}: $\respF(0)=\tfrac12$.
+
+\xhdr{The \TheoryModel.}
+In this model variant, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$ before round $1$.
+Agents know the principals' algorithms.
+Their reward estimates are defined as Bayesian-expected rewards: for each agent $t$ and principal $i$,
+\[ \est_i(t) := \E\sbr{ r_t\mid i_t = i }.\]
+
+\ascomment{stopped here.}
+
+with a common prior. More formally, for each action $a\in A$ there is a parametric family $\psi_a(\cdot)$ of
 reward distributions, parameterized by the mean reward $\mu_a$. (The paradigmatic case is 0-1 rewards with a given expectation.) The
@@
 
-Each principal commits to a learning algorithm for making recommendations. This algorithm follows a protocol of \emph{multi-armed bandits} (\emph{MAB}). Namely, the algorithm proceeds in time-steps:%
-\footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before.
-We will omit the global vs.
-local distinction when clear from the context.} each time it is called, it outputs a chosen action $a\in A$ and then inputs the reward for this action.
-The algorithm is called only in global rounds when the corresponding principal is chosen.
-
-\gaedit{
 The information structure is as follows.
@@
 The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
-
-}
-\xhdr{Some terminology.} The two principals are called ``Principal
-1" and ``Principal 2".
-%Principal $i\in \{1,2\}$ is also denoted \prin[i].
-The algorithm of principal $i\in \{1,2\}$ is called ``algorithm $i$" and denoted
-\alg[i]. The agent in global round $t$ is called ``agent $t$"; the
-chosen principal is denoted $i_t$.
 
@@
 }
-\xhdr{Response functions.}
-We use the response function $\respF$ \gaedit{to characterize the decision rule of the agents in our model}.
-We assume that $\respF$ is monotonically non-decreasing, is larger than $\tfrac12$ on the interval $(0,1]$, and smaller than $\tfrac12$ on the interval $[-1,0)$.
-Beyond that, we consider three specific models (see \reffig{fig:response-functions}):
-
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[scale=2.0]
-    \draw[->] (-1.1,0) -- (1.1,0) node[right] {$\Delta_t := \PMR_1(t) - \PMR_2(t)$};
-    \draw[->] (0,-0.1) -- (0,1.1) node[above]
-        {$p_t = \text{prob. of choosing principal 1}$};
-    \draw[scale=1.0,domain=-1:0,smooth,variable=\q,blue, line width=0.50mm] plot ({\q},{0});
-    \draw[scale=1.0,domain=0:1,smooth,variable=\q,blue,line width=0.50mm] plot ({\q},{1});
-    \draw[scale=1.0,domain=-1:0,smooth,variable=\y,red]  plot ({\y},{0.1});
-    \draw[scale=1.0,domain=0:1,smooth,variable=\y,red]  plot ({\y},{0.9});
-    \draw[scale=1.0,domain=-1:1,smooth,variable=\y,cyan, line width=0.45mm, dash pattern=on 3pt off 2pt]  plot ({\y},{1/(1 + 1/(9^\y))});
-    % \node[above, blue] at (0.5, 0.5) {\footnotesize $2 q (1 - q)$};
-     \node[left] at (0, 0.5) {\footnotesize $1/2$};
-     \node[left] at (0, 1) {\footnotesize $1$};
-     \node[below left] at (0, 0) {\footnotesize $0$};
-     \node[below ] at (1, 0) {\footnotesize $1$};
-     \node[below ] at (-1, 0) {\footnotesize $-1$};
-  \end{tikzpicture}
-\end{center}
-\caption{The three models for agents' response function: \HardMax is thick blue, \HardMaxRandom is slim red, and \SoftMaxRandom is the dashed curve.}
-\label{fig:response-functions}
-\end{figure}
-
-\begin{OneLiners}
-\item \HardMax: $\respF$ equals $0$ on the interval $[-1,0)$ and $1$
-  on the interval $(0,1]$. In other words, the agents will
-  deterministically choose the principal with the higher posterior
-  mean reward.
-
-\item \HardMaxRandom:
-    % $\respF$ equals $\eps$ on the interval $[-1,0)$ and $1-\eps'$ on the interval $(0,1]$, where $\eps,\eps'\in (0,\tfrac12)$ are some positive constants.
-In words, each agent is a \HardMax agent with probability $1-\eps-\eps'$, and with the remaining probability she makes a random choice.
-    $\respF$ equals $\eps_0$ on the interval $[-1,0)$ and $1-\eps_0$ on the interval $(0,1]$, where $\eps_0\in (0,\tfrac12)$ are some positive constants.
-In words, each agent is a \HardMax agent with probability $1-2\eps_0$, and with the remaining probability she makes a random choice.
-
-\item \SoftMaxRandom: $\respF(\cdot)$ lies in the interval $[\eps_0,1-\eps_0]$, $\eps_0>0$, and is ``smooth" around $0$ (in the sense defined precisely in Section~\ref{sec:soft}).
-\end{OneLiners}
-% \asedit{(We assume $\respF(-1)+\respF(1)=1$ in \HardMaxRandom and \SoftMaxRandom only to simplify notation.)}
-
-We say that $\respF$ is \emph{symmetric} if $\respF(-x)+\respF(x)=1$
-for any $x\in [0,1]$. This implies \emph{fair tie-breaking}:
-$\respF(0)=\tfrac12$.% , and $\eps=\eps'$ in the
-% definitions above.
 
@@
 Fix principal $i\in \{1,2\}$ and (local) step $n$.
-The arm chosen by algorithm \alg[i] at this step is denoted $a_{i,n}$, and the corresponding \BIR is denoted $\BIR_i(n)$.
+The arm chosen by algorithm \alg[i] at this step is denoted $a_{i,n} = \alg[i](n)$, and the corresponding \BIR is denoted $\BIR_i(n)$.
 History of \alg[i] up to this step is denoted $H_{i,n}$.

### 2020-07-08 journal paper/content/sec-intro.tex
formatting tweaks

@@
 \item We investigate the algorithms' performance ``in isolation", \ie without competition.
-We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.%
-    \footnote{Note the discrepancy with our theoretical results on \HardMax: there, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations the prior is not available).}
+We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.\footnote{Note the discrepancy with our theoretical results on \HardMax: there, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations the prior is not available).}
      We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes \emph{not} a good predictor for the outcomes under competition.
@@
 Firms compete only on the quality of service, rather than, say, pricing or the range of products.
-Agents are myopic: they do not worry about how their actions impact their future utility.
-\footnote{In particular, agents do not attempt to learn over time, to second-guess or game future agents, or to manipulate the principals' learning algorithms.
+Agents are myopic: they do not worry about how their actions impact their future utility.\footnote{In particular, agents do not attempt to learn over time, to second-guess or game future agents, or to manipulate the principals' learning algorithms.
 This is arguably typical in practice, in part because one agent's influence tends to be relatively small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores", and further simplified to average performance over a time window.

### 2020-07-08 journal paper/content/sim_details.tex
tweaks to paper
Full draft for the "theory section".
formatting tweaks

@@
 
-\section{Reputation Model Details}\label{sec:sim_details}
-
-In this section we move to the reputation choice model.
+In this section we move to the reputation choice variant.
 Recall that in this variant of the model, instead of making choices between firms based on the Bayesian expected reward, each agent chooses the firm with a maximal reputation score (breaking ties uniformly).
@@
 
-Even with a stylized model, numerical investigation is quite challenging.
+\OMIT{Even with a stylized model, numerical investigation is quite challenging.
 An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given instance of a multi-armed bandit problem.%
@@
 In particular, we need to capture a huge variety of bandit instances with only a few representative examples.
-Further, one needs succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).
+Further, one needs succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).}
 

### 2020-07-08 journal paper/ec19paper/content/perf_in_iso.tex
Reformatted so that: - all expts in the body are in one section. - all proofs in the app are in one section - all expts in the app are in one section.
formatting tweaks

@@
 
-\section{Performance in Isolation}\label{sec:isolation}
+\subsection{Performance in Isolation}\label{sec:isolation}
 
@@
 \begin{figure}
-\includegraphics[scale=0.35]{ec19paper/figures/nih_iso_mean}
+\centering
+\includegraphics[scale=0.3]{ec19paper/figures/nih_iso_mean}
+\includegraphics[scale=0.4]{ec19paper/appendix_figures/uniform_mean}
 %\caption*{\tiny{Mean reputation trajectory The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
 The shaded area display 95\% confidence intervals.}}
-\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area shows 95\% confidence intervals.}
+\caption{Mean reputation trajectories for Needle-in-Haystack and Uniform. The shaded area shows 95\% confidence intervals.}
 \label{prelim_means}
@@
 \begin{figure}[ht]
+\centering
 \includegraphics[scale=0.35]{ec19paper/figures/relative_uniform_annotated_plot}

### 2020-07-08 journal paper/journal_main.tex
fix em issue
Added some macro.
Some macro.
Changed some macro.
Changed the same macro, again ... :(
Full draft for the "theory section".
Reformatted so that: - all expts in the body are in one section. - all proofs in the app are in one section - all expts in the app are in one section.
Moved the stuff from Guy's intro to a separate file.
formatting fix

@@
 
-\usepackage{ulem,color,
+\usepackage{color,
 setspace,sectsty,comment,footmisc,caption,
@@
 \usepackage{slivkins-setup,slivkins-theorems}
+
+% model variants
+\newcommand{\TheoryModel}{Bayesian-choice model\xspace}
+\newcommand{\ExptsModel}{reputation-choice model\xspace}
 
@@
 Yishay Mansour%
-\footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com;} 
+\footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com}
 \and
@@
 Zhiwei Steven Wu%
-\footnote{University of Minnesota - Twin Cities, Department of Computer Science. Email: zsw@umn.edu.\newline
+\footnote{University of Minnesota - Twin Cities, Department of Computer Science. Email: zsw@umn.edu\newline
 The research has been done when Z.S. Wu has been an intern and a postdoc at Microsoft Research NYC.
@@
 \begin{small}
+\setcounter{tocdepth}{2}
 \tableofcontents
@@
 
-\section{Full rationality (HardMax)}
-\label{sec:rational}
-\subfile{itcs18paper/sec-rational}
+\section{Theoretical results: the \TheoryModel}
+\label{sec:theory}
+\subfile{content/sec-theory}
 
-\section{Relaxed rationality: HardMax \& Random}
-\label{sec:random}
-\subfile{itcs18paper/sec-random}
+\subsection{Proofs}
+\label{sec:theory-proofs}
+\subfile{content/sec-theory-proofs}
 
-\section{SoftMax response function}
-\label{sec:soft}
-\subfile{itcs18paper/sec-soft}
+%\section{Full rationality (HardMax)}
+%\label{sec:rational}
+%\subfile{itcs18paper/sec-rational}
+
+%\section{Relaxed rationality: HardMax \& Random}
+%\label{sec:random}
+%\subfile{itcs18paper/sec-random}
+
+%\section{SoftMax response function}
+%\label{sec:soft}
+%\subfile{itcs18paper/sec-soft}
 
@@
 
-\gaedit{
+\section{Numerical simulations: the \ExptsModel}
+\label{sec:sim}
+%\label{sec:sim_details}
+
 \subfile{content/sim_details}
-}
 
@@
 \clearpage
-%\addcontentsline{toc}{section}{References}
+\addcontentsline{toc}{section}{References}
 \bibliographystyle{plainnat}
@@
 
-\section{Background on multi-armed bandits}
+\section{Supplementary theoretical results}
 \label{app:examples}
@@
 
-\section{Non-degeneracy via a random perturbation}
+\subsection{Non-degeneracy via a random perturbation}
 \label{app:perturb}
@@
 
-\section{Omitted Proofs}
-\label{app:omitted_proofs}
-\subfile{content/omitted_proofs}
+\subfile{ec19paper/content/appendix_for_one_version}
 
-\subfile{ec19paper/content/appendix_for_one_version}
+\subfile{content/old-intro}
 

### 2020-07-08 journal paper/content/sec-theory.tex
Minor edit.
Added "prelims" subsections.

@@
 In this section, we present our theoretical results for the \TheoryModel. While we provide intuition and (some) proof sketches, the detailed proofs are deferred to Section~\ref{sec:theory-proofs}.
+
+\subsection{Preliminaries}
+\label{sec:theory-prelims}
+
+\ascomment{to be added / copied from the ITCS paper}
 
@@
 
-\ascomment{need to polish this subsection}
+\ascomment{TODO: revise this subsection.}
 
@@
 and is omitted.
+
+\ascomment{TODO: arbitrary reward distributions.}
+For each action $a\in A$ there is a parametric family $\psi_a(\cdot)$ of reward distributions, parameterized by the mean reward $\mu_a$.
+(The paradigmatic case is 0-1 rewards with a given expectation.) Whenever a given action $a\in A$ is chosen, the reward is drawn independently from distribution $\psi_a(\mu_a)$.
+The prior $\priorMu$ and the distributions $(\psi_a(\cdot)\colon a\in A)$ constitute the (full) Bayesian prior on rewards, denoted $\prior$.

### 2020-07-08 journal paper/content/sec-model.tex
Full draft of the "model" section.

@@
-We define the basic model; some extensions thereof are specified in  Section~\ref{sec:theory-extensions}.
+Let us define the basic model; our theoretical results admit some extensions, detailed in Section~\ref{sec:theory-extensions}.
 
@@
 We posit \emph{stochastic rewards}: whenever a given action $a\in A$ is chosen, the reward is an independent draw from Bernoulli distribution with mean $\mu_a$.
-The mean rewards $\mu_a$ are initially not known to anybody.
+The mean rewards $\mu_a$ are fixed over time, and initially not known to anybody.
 Each principal is completely unaware of the rounds when the other principal is chosen.
@@
 
-\newcommand{\est}{\mathtt{est}}
+\newcommand{\est}{\mathtt{EST}}
 
-\xhdr{Agents' response.} Each agent $t$ forms a reward estimate $\est_i(t)$ for each principal $i$, as specified below, and chooses the principal based on these reward estimates.
+\xhdr{Agents' response.} Each agent $t$ forms a reward estimate $\est_i(t)\in [0,1]$ for each principal $i$, and chooses the principal based on these reward estimates.
 Specifically, principal $1$ is chosen with probability
@@
 
-\xhdr{The \TheoryModel.}
-In this model variant, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$ before round $1$.
-Agents know the principals' algorithms.
-Their reward estimates are defined as Bayesian-expected rewards: for each agent $t$ and principal $i$,
-\[ \est_i(t) := \E\sbr{ r_t\mid i_t = i }.\]
-
-\ascomment{stopped here.}
-
-with a common prior. More formally, for each action $a\in A$ there is a parametric family $\psi_a(\cdot)$ of
-reward distributions, parameterized by the mean reward $\mu_a$. (The paradigmatic case is 0-1 rewards with a given expectation.) The
-mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from prior distribution $\priorMu$ before round $1$.
-Whenever a given action $a\in A$ is chosen, the reward is drawn independently from distribution $\psi_a(\mu_a)$.
-The prior $\priorMu$ and the distributions $(\psi_a(\cdot)\colon a\in A)$ constitute the (full) Bayesian prior on rewards, denoted $\prior$.
-
-The information structure is as follows.
-The prior $\prior$ is known to everyone.
-The mean rewards $\mu_a$ are not revealed to anybody and each principal is completely unaware of the rounds when the other is chosen.
-We consider two variants of our model with different information structures.
-In the first, the  \textit{expectation choice} variant, each agent knows both principals' algorithms, and the global round when (s)he arrives, \emph{but not} the rewards of the previous agents.
-In the second, the \textit{reputation choice} variant, agents only make decisions based on the rewards of previous agents.
-Concretely, each of the two principals has a \emph{reputation score}, and each agent's choice is driven by these two numbers.
-The reputation score is simply a sliding window average: an average reward of the last $M$ agents that chose this firm.
-
-Throughout, $\E[\cdot]$ denotes expectation over all applicable randomness.
-
-\xhdr{Bayesian-expected rewards.}
-Consider the performance of a given algorithm \alg[i], $i\in \{1,2\}$, when it is run in isolation (\ie without competition, just as a bandit algorithm).
-Let $\rew_i(n)$ denote its Bayesian-expected reward for the $n$-th step.
-
-Now, going back to our game, fix global round $t$ and let $n_i(t)$ denote the number of global rounds before $t$ in which this principal is chosen. Then:
-\begin{align*}
- \E[r_t \mid \text{principal $i$ is chosen in round $t$ and $n_i(t)=n$} ]
-    = \rew_i(n+1) \quad (\forall n\in\N).
-\end{align*}
-
-\xhdr{Agents' response.}
-Each agent $t$ chooses principal $i_t$ as follows: it chooses a distribution over the principals, and then draws independently from this distribution.
-Let $p_t$ be the probability of choosing principal $1$ according to this distribution.
-Below we specify $p_t$; we need to be careful so as to avoid a circular definition.
-
-\gaedit{
-
-The form of $p_t$ depends on the information structure that we consider.
-In the expectation choice model, where the agents only know the global round and the principals' algorithms, it is defined as follows.
-
-}
-
-Let $\mI_t$ be the information available to agent $t$ before the
-round. Assume $\mI_t$ suffices to form posteriors for quantities
-$n_i(t)$, $i\in \{1,2\}$, denote them by $\posteriorN{i}{t}$. Note
-that the Bayesian expected reward of each principal $i$ is a function
-only of the number rounds he was chosen by the agents, so the
-posterior mean reward for each principal $i$ can be written as
-\begin{align*}
- \PMR_i(t) := \E[r_t \mid \text{$\mI_t$ and $i_t=i$} ]
-    = \E[\rew_i(n_i(t)+1) \mid \mI_t]
-    = \E_{n\sim \posteriorN{i}{t}}[\rew_i(n+1)].
-\end{align*}
-This quantity represents the posterior mean reward for principal $i$
-at round $t$, according to information $\mI_t$; hence the notation \PMR. In general, probability $p_t$ is defined by the
-posterior mean rewards $\PMR_i(t)$ for both principals. We assume a
-somewhat more specific shape:
-\begin{align}
-p_t = \respF\left(\; \PMR_1(t) - \PMR_2(t) \;\right).
-%p_t = \respF(\Delta_t), \quad \text{where } \Delta_t := \PMR_1(t) - \PMR_2(t).
-\end{align}
-Here $\respF:[-1,1]\to [0,1]$ is the \emph{response function}, which is the same for all agents. We assume that the response function is known to all agents.
-
-To make the model well-defined, it remains to argue that information $\mI_t$ is indeed sufficient to form posteriors on $n_1(t)$ and $n_2(t)$. This can be easily seen using induction on $t$.
-
-Since all agents arrive with identical information (other than knowing which global round they arrive in), it follows that all agents have identical posteriors for $n_{i,t}$ (for a given principal $i$ and a given global round $t$).
-This posterior is denoted $\posteriorN{i}{t}$.
-
-\gaedit{
-In the reputation choice variant we consider, at a given time $t$ each principal $i \in \{1, 2 \}$ has a reputation score denoted as $\REP_i (t)$.
-In this case we have that the agent's responses take an analogous form as in the first case:
-\begin{align}
-p_t = \respF\left(\; \REP_1(t) - \REP_2(t) \;\right).
-\end{align}
-}
-
-\xhdr{MAB algorithms.}
-We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
-\begin{align}
-\BIR(n) := \E_{\mu\sim\priorMu}
-    \left[ \max_{a\in A} \mu_a\right] - \rew(n),
-\end{align}
-where $\rew(n)$ is the Bayesian-expected reward of the algorithm for the $n$-th step, when the algorithm is run in isolation.
-We are primarily interested in how \BIR scales with $n$; we treat $K$, the number of arms, as a constant unless specified otherwise.
-
-We will emphasize several specific algorithms or classes thereof:
+\xhdr{Bayesian vs. frequentist.}
+We consider two model variants, Bayesian and frequentist. The main difference between the two concerns the agents' reward estimates $\est_i(t)$.
 
 \begin{itemize}
-\item ``smart" MAB algorithms that combine exploration and exploitation, such as UCB1 \cite{bandits-ucb1} and \gaedit{Thompson Sampling \cite{russo2018tutorial}}. These algorithms achieve
-        $\BIR(n) \leq  \tilde{O}(n^{-1/2})$
-    for all priors and all (or all but a very few) steps $n$. This bound is known to be tight for any fixed $n$.\footnote{This follows from the lower-bound analysis in \cite{bandits-exp3}.}
+\item In the \emph{\TheoryModel}, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$.
+Agents know the principals' algorithms.
+Their reward estimates are defined as Bayesian-expected rewards:
+        $\est_i(t) := \E\sbr{ r_t\mid i_t = i }$
+    for each agent $t$ and principal $i$.
 
-\item ``naive" MAB algorithms that separate exploration and exploitation, such as Explore-then-Exploit and $\eps$-Greedy.
-These algorithms have dedicated rounds in which they explore by choosing an action uniformly at random.
-When these rounds are known in advance, the algorithm suffers constant \BIR in such rounds.
-When the ``exploration rounds" are instead randomly chosen by the algorithm, one can usually guarantee an inverse-polynomial upper bound \BIR, but not as good as the one above: namely,
-        $ \BIR(n) \leq \tilde{O}(n^{-1/3})$.
-    This is the best possible upper bound on \BIR for the two algorithms mentioned above.
+\item In the \emph{\ExptsModel}, agents' reward estimate for a given principal is the average reward of the last $M$ agents that chose this principal (interpreted as a \emph{reputation score}).
+To make the reputation score meaningful initially, each principal enjoys a ``warm start": additional $T_0$ agents arrive before the game starts, and interact with the principal as described above.
 
-\item \DynGreedy: at each step, recommends the best action according to the current posterior: an action $a$ with the highest posterior expected reward
-        $\E[\mu_a \mid \mI]$,
-     where $\mI$ is the information available to the algorithm so far.
-     \DynGreedy has (at least) a constant \BIR for some reasonable priors, \ie
-        $\BIR(n)>\Omega(1)$.
-
-\item \StaticGreedy: always recommends the prior best action,\ie an action $a$ with the highest prior mean reward  $\E_{\mu\sim \priorMu}[\mu_a]$. This algorithm typically has constant \BIR.
 \end{itemize}
 
-When we consider the analytical model, we focus on MAB algorithms such that $\BIR(n)$ is non-increasing; we call such algorithms \emph{monotone}.
-While some reasonable MAB algorithms may occasionally violate monotonicity, they can usually be easily modified so that monotonicity violations either vanish altogether, or only occur at very specific rounds (so that agents are extremely unlikely to exploit them in practice).
-
-More background and examples can be found in Appendix~\ref{app:examples}. In particular, we prove that \DynGreedy is monotone.
-
-\xhdr{Competition game between principals.}
-Some of our results explicitly study the game between the two principals.
+\xhdr{Competition game.}
+Some of our results explicitly study the game between the two principals, termed the \emph{competition game}.
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-Thus, choosing a pure strategy in this game corresponds to choosing an MAB algorithm (and, implicitly, announcing this algorithm to the agents).
-
-Principal's utility is primarily defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral, in the sense that they optimize their expected utility.
-
-\xhdr{Assumptions on the prior.} We make some technical assumptions on the Bayesian prior. First, each action $a$ has a positive probability of being the best action according to the prior:
-\begin{align}\label{eq:assn-prob}
-\forall a\in A:\;\;\;  \Pr_{\mu\sim \priorMu}[\mu_a  > \mu_{a'}
-\;\forall a'\in A]
-> 0.
-\end{align}
-
-Second, posterior mean rewards of actions are pairwise distinct almost
-surely. That is, the history $h$ at any step of an MAB algorithm%
-\footnote{The \emph{history} of an MAB algorithm at a given step
-  comprises the chosen actions and the observed rewards in all
-  previous steps in the execution of this algorithm.}
- satisfies
-\begin{align}\label{eq:assn-distinct}
-    \E[\mu_a \mid h] \neq \E[\mu_{a'}\mid h] \quad \forall a,a'\in A,
-\end{align}
-except at a set of histories of probability $0$.  In
-particular, prior mean rewards of actions are pairwise distinct:
-\[ \E[\mu_a] \neq \E[\mu_a'] \quad \text{for any $a,a'\in A$}. \]
-
-We provide two examples for which property \eqref{eq:assn-distinct} is `generic', in the sense that it can be enforced almost surely by a
-  small random perturbation of the prior. Both examples focus on 0-1 rewards and priors $\priorMu$ that are independent across arms. The first example
-  assumes Beta priors on the mean rewards, and is very easy.%
-  \footnote{Suppose the rewards are Bernouli r.v. and the mean reward
-    $\mu_a$ for each arm $a$ is drawn from some Beta distribution
-    $\text{Beta}(\alpha_a, \beta_a)$. Given any history that contains
-    $h_a$ number of heads and $t_a$ number of tails from arm $a$, the
-    posterior mean reward is
-    $\frac{\alpha_a + h_a}{\alpha_a + h_a + \beta_a + t_a}$. Note that
-    $h_a$ and $t_a$ take integer values. Therefore, perturbing the
-    parameters $\alpha_a$ and $\beta_a$ independently with any
-    continuous noise will induce a prior with property
-    \eqref{eq:assn-distinct} with probability 1.  }  The second
-  example assumes that mean rewards have a
-  finite support, see Appendix~\ref{app:perturb} for details.
-
-\xhdr{Some more notation.} Without loss of generality, we label actions as $A=[K]$ and sort them according to their prior mean rewards, so that
-    $ \E[\mu_1] > \E[\mu_2] > \ldots > \E[\mu_K]$.
-
-Fix principal $i\in \{1,2\}$ and (local) step $n$.
-The arm chosen by algorithm \alg[i] at this step is denoted $a_{i,n} = \alg[i](n)$, and the corresponding \BIR is denoted $\BIR_i(n)$.
-History of \alg[i] up to this step is denoted $H_{i,n}$.
-%The posterior distribution of agent $t$ for quantity $n_t(t)$ is denoted $\posteriorN{i}{t}$.
-
-We denote
-    $\PMR(a\mid E) = \E[\mu_a \mid E]$
-for posterior mean reward of action $a$ given event $E$.
-
-%\end{assumption}
-
-\subsection{Chernoff Bounds}
-
-We use an elementary concentration inequality known as {\em Chernoff Bounds}, in a formulation from~\cite{MitzUpfal-book05}.
-\begin{theorem}[Chernoff Bounds]
-\label{thm:chernoff}
-Consider $n$ i.i.d. random variables $X_1 \ldots X_n$ with values in $[0,1]$. Let
-    $X = \tfrac{1}{n} \sum_{i=1}^n X_i$ be their average, and let $\nu = \E[X]$. Then:
-\[ \min\left(\; \Pr[ X-\nu > \delta \nu ],\quad
-                \Pr[ \nu-X > \delta \nu ]
-    \; \right)
-    < e^{-\nu n \delta^2/3}
-    \quad \text{for any $\delta\in (0,1)$.}
-\]
-\end{theorem}
-
-%, and the choice of algorithm are commonly known
-%to the agents. Given any finite class of MAB learning algorithms
-%$\Pi$, a pure strategy for the principal $j$ in the competition game
-%is an algorithm $\pi_j$ in the class $\Pi$ and a mixed strategy
-%$\sigma_j$ is a probability distribution over $\Pi$. We will write
-%$u_j(\sigma)$ and $u(\pi)$ to denote the expected utility for the
-%principal $j$ in mixed strategy profile $\sigma$ and pure strategy
-%profile $\pi$.  The principal gets a unit reward for being selected by
-%an agent, so the utility function can be written as
-%\[
-%u_j(\pi) = \Ex{}{\sum_{t=1}^T\mathbf{1}[P^t = j]}
-%\]
-%A variant of the utility function we have is the number of agents
-%weighted by the rewards for the agents
-%\[
-%u_j(\pi) = \Ex{}{\sum_{t=1}^T \mathbf{1}[P^t = j] r^t}
-%\]
-%\iffalse
-%We consider a few models for the utility of the principals.  For
-%example, the principal gets the reward realized by the agents it
-%advised.
-%\fi
-%
-%We will analyze the equilibrium strategies of the game under different
-%environments.
-%
-%\begin{definition}[Equilibrium in the competition game]
-%  A pure strategy profile $\pi$ is a pure strategy Nash equilibrium (PSNE) if
-%  for any principal $j$ in the game and any algorithm $a\in \Pi$
-%\[
-%  u_j(\pi) \geq u_j(a, \pi_{-j}).
-%\]
-%A mixed strategy profile $\sigma$ is a (mixed strategy) Nash
-%equilibrium if for any principal $j$ and any algorithm $a\in\Pi$,
-%\[
-%  \Ex{\sigma}{u_j(\sigma)} \geq u_j(a, \sigma_{-j}).
-%\]
-%\end{definition}
-%
-%We will measure the quality of an equilibrium by studying the welfare
-%of the agents with respect to the system regret.
-
-%%% Local Variables:
-%%% TeX-master: "main.tex"
-%%% End:
+Thus, choosing a pure strategy in this game corresponds to choosing an MAB algorithm.
+Principal's utility is defined as the market share, \ie the number of agents that chose this principal.
+Principals are risk-neutral, in the sense that they optimize their expected utility.

### 2020-07-08 journal paper/content/sec-theory-proofs.tex
Moved proofs to the appendix.

@@
-\subsubsection{Main result on \HardMax: Proof of Theorem~\ref{thm:DG-dominance}}
+\subsection{Main result on \HardMax: Proof of Theorem~\ref{thm:DG-dominance}}
 \label{sec:proofs-HM-main}
@@
 %%%%%%%%%%%%
-\subsubsection{\HardMax with biased tie-breaking: 
+\subsection{\HardMax with biased tie-breaking:
 Proof of Theorem~\ref{thm:HardMax-biased}}
@@
 %%%%%%%%%%%%
-\subsubsection{The main result for \HardMaxRandom:
+\subsection{The main result for \HardMaxRandom:
 Proof of Theorem~\ref{thm:random-clean}}
@@
 %%%%%%%%%%%%
-\subsubsection{For \HardMaxRandom, a little greedy goes a long way
+\subsection{For \HardMaxRandom, a little greedy goes a long way
 (Proof of Theorem~\ref{thm:random-greedy})}
@@
 %%%%%%%%%%%%
-\subsubsection{\SoftMaxRandom: proof of Theorem~\ref{thm:SoftMax-strong}}
+\subsection{\SoftMaxRandom: proof of Theorem~\ref{thm:SoftMax-strong}}
 \label{sec:proofs-SoftMax}

### 2020-07-09 journal paper/content/sec-theory.tex
Minor edit.
Added prelims; minor edits elsewhere.

--- 2020-07-07 journal paper/content/sec-theory.tex
@@
-In this section, we present our theoretical results for the \TheoryModel. While we provide intuition and (some) proof sketches, the detailed proofs are deferred to Section~\ref{sec:theory-proofs}.
+In this section, we present our theoretical results for the \TheoryModel. While we provide intuition and (some) proof sketches, the detailed proofs are deferred to Appendix~\ref{sec:theory-proofs}.
+
+\subsection{Preliminaries}
+\label{sec:theory-prelims}
+
+\xhdr{Notation.} Let $\rew_i(n)$ denote the Bayesian-expected reward of principal $i$ at local step $n$.
+For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen.
+Note that
+    \[ \PMR_i(t):=  \E\sbr{ r_t\mid i_t = i } 
+        = \E\sbr{ \rew_i(n_i(t)+1) }. \]
+
+\xhdr{Assumptions.} We make two mild assumptions on the prior. First, each arm $a$ can be the best:
+\begin{align}\label{eq:assn-prob}
+\forall a\in A:\;\;\;  
+\Pr_{\mu\sim \priorMu}\sbr{\mu_a  > \mu_{a'}
+\;\forall a'\in A}
+> 0.
+\end{align}
+
+\noindent Second, actions' posterior mean rewards are pairwise distinct given any feasible history $h$:%
+\footnote{The \emph{history} of an MAB algorithm at a given step
+  comprises the chosen actions and the observed rewards in all
+  previous steps. The history is \emph{feasible} if for each arm-reward pair $(a,r)$ in the history, $r$ is in the support of the reward distribution for $a$.}
+\begin{align}\label{eq:assn-distinct}
+    \E[\mu_a \mid h] \neq \E[\mu_{a'}\mid h] \quad \forall a,a'\in A.
+\end{align}
+In particular, \emph{prior} mean rewards of actions are pairwise distinct:
+$\E[\mu_a] \neq \E[\mu_a']$ for any $a,a'\in A$.
+
+In Appendix~\ref{app:perturb}, we provide two examples for which property \eqref{eq:assn-distinct} is `generic', in the sense that it can be enforced almost surely by a small random perturbation of the prior.
+The two examples concern, resp., Beta priors and priors with a finite support, and focus on priors $\priorMu$ that are independent across arms.
+
+\xhdr{MAB algorithms.}
+We consider two (Bayesian) greedy algorithms.
+The first one, called  \DynGreedy, chooses an arm $a$ with the largest posterior mean reward $\E[\mu_a \mid \cdot ]$ given all information currently available to the algorithm.
+The second algorithm, called \StaticGreedy, chooses an arm $a$ with the largest prior mean reward $\E[\mu_a]$, and uses this arm in all rounds.
+In both algorithms, ties are broken arbitrarily.
+
+We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
+\begin{align}
+\BIR(n) := \E_{\mu\sim\priorMu}
+    \left[ \max_{a\in A} \mu_a\right] - \rew(n),
+\end{align}
+where $\rew(n)$ is the Bayesian-expected reward of the algorithm for the $n$-th step, when the algorithm is run in isolation.
+We are primarily interested in how \BIR scales with $n$: typically, how fast \BIR decreases with $n$.
+We treat $K$, the number of arms, as a constant unless specified otherwise.
+Let $\BIR_i(n) = \BIR(n)$ for the algorithm of principal $i$.
+
+Let us connect \BIR with the three classes of algorithms listed in Section~\ref{sec:intro-model}.
+\DynGreedy has at least a constant \BIR for many reasonable priors (where the constant can depend on $K$ and the prior, but not on $n$).
+\StaticGreedy typically has at least a constant \BIR, in the same sense.
+Exploration-separating algorithms can have
+    $\BIR(n) \sim n^{-1/3}$
+for all priors, whereas adaptive-exploration algorithms can have an even better rate
+    $\BIR(n) \sim n^{-1/3}$
+for all priors. Both rates are up to $\log(n)$ factors, and are the best possible \BIR rates that can be achieved for all priors. More details on this can be found in Appendix~\ref{app:examples}.
 
@@
 \begin{align}
-\Pr\sbr{ \alg[1](n)\neq \alg[2](n)} >0.
+\Pr\sbr{ \alg[1](n)\neq \alg[2](n)} >0,
 \end{align}
+where $\alg[i](n)$ is the arm chosen by \alg[i] at local step $n$.
 
@@
 
+\ascomment{This whole remark is supposed to add intuition, but it seems pretty hard to parse ... :(}
+
 \begin{remark}
@@
 at a slower rate.
-Let us argue how this may translate  into a larger \BIR compared to the original algorithm.
+For the sake of intuition, let us argue how this may translate  into a larger \BIR compared to the original algorithm.
 Let  $\BIR'_1(n)$ be the \BIR of the ``greedy choice" after after $n-1$ steps of \alg[1].
@@
     $\E[m] = n_0-1 + (n-n_0+1)(1-p) \geq (1-p)n$.
+
+\ascomment{Steven: why can't we take $c=1$ below??}
 
@@
 \end{OneLiners}
-These conclusions follow, respectively, from Corollaries~\ref{cor:DG-dominance}, \ref{cor:random} and \ref{cor:SoftMax}.
+These conclusions follow, respectively, from Corollaries~\ref{cor:DG-dominance}, \ref{cor:random} and \ref{cor:SoftMax-strong}.
 Further, {we consider the uniform choice between the principals.
@@
 %%%%%%%%%%%%%
-\subsection{Generalizations}
+\subsection{Extensions}
 \label{sec:theory-extensions}
 
-\ascomment{need to polish this subsection}
+Our theoretical results can be extended beyond the basic model in Section~\ref{sec:model}.
 
-Our results can be extended compared to the basic model described above.
-
-%First, the two principals may have different action sets, as long as
-%    $\E\left[ \max_{a\in A} \mu_a\right]$,
-%the Bayesian-expected reward of the best action, is the same for both principals.
-
-First, unless specified otherwise, our results allow a more general notion of principal's utility that can depend on both the market share and agents' rewards.
+\xhdr{Reward-dependent utility.}
+Except for Corollary~\ref{cor:SoftMax-strong}, our results allow a more general notion of principal's utility that can depend on both the market share and agents' rewards.
 Namely, principal $i$ collects $U_i(r_t)$ units of utility in each global round $t$ when she is chosen (and $0$ otherwise), where $U_i(\cdot)$ is some fixed non-decreasing function with $U_i(0)>0$.
@@
 
-\ascomment{Edit in: this extension does not apply to Corollary~\ref{cor:SoftMax-strong}.}
+\xhdr{Time-discounted utility.}
+Corollary~\ref{cor:DG-dominance} holds under a more general model which allows time-discounting: namely, the utility of each principal $i$ in each global round $t$ is $U_{i,t}(r_t)$ if this principal is chosen, and $0$ otherwise, where $U_{i,t}(\cdot)$ is an arbitrary non-decreasing function with $U_{i,t}(0)>0$.
 
-Second, our results carry over, with little or no modification of the proofs, to much more general versions of MAB, as long as it satisfies the i.i.d.
+\xhdr{Arbitrary reward distributions.}
+While Section~\ref{sec:model} posits Bernoulli rewards, we allow arbitrary reward distributions.
+Formally, for each arm $a\in A$ there is a parametric family $\psi_a(\cdot)$ of reward distributions, parameterized by the mean reward.
+Whenever arm $a$ is chosen, the reward is drawn independently from distribution $\psi_a(\mu_a)$.
+The prior $\priorMu$ and the distributions $(\psi_a(\cdot)\colon a\in A)$ constitute the (full) Bayesian prior on rewards, denoted $\prior$.
+
+\xhdr{MAB extensions.} Our results carry over, with little or no modification of the proofs, to much more general versions of MAB, as long as it satisfies the i.i.d.
 property.
@@
 Mean rewards may also have a known structure, such as Lipschitzness, convexity, or linearity; such structure can be incorporated via $\prior$.
-All these extensions have been studied extensively in the literature on MAB, and account for a substantial segment thereof; see \cite{Bubeck-survey12} for background and details.
+All these extensions have been studied extensively in the literature on MAB, and account for a substantial segment thereof.
+Background can be found in any of the recent books on MAB \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
 
-Corollary~\ref{cor:DG-dominance} holds under a more general model which allows time-discounting: namely, the utility of each principal $i$ in each global round $t$ is $U_{i,t}(r_t)$ if this principal is chosen, and $0$ otherwise, where $U_{i,t}(\cdot)$ is an arbitrary non-decreasing function with $U_{i,t}(0)>0$.
-
-\xhdr{Dependence on the time horizon.}
-%\begin{remark}\label{rem:random-messy}
-  Many standard MAB algorithms in the literature are parameterized by
-  the time horizon $T$. Regret bounds for such algorithms usually
-  include a polylogarithmic dependence on $T$. In particular, a
-  typical upper bound for \BIR has the following form:
-  \OMIT{\footnote{We
-    provide upper bounds on \BIR for several standard MAB algorithms
-    to illustrate these dependencies in the appendix.}}
+\xhdr{\BIR can depend on $T$.}
+Many/most MAB algorithms are parameterized by the time horizon $T$, and their regret bounds usually include $\polylog(T)$. In particular,  a regret bound for \BIR is
 \begin{align}
@@
 
-Theorem~\ref{thm:random-clean} easily extends if condition \eqref{eq:random-better-weaker} is replaced with the following:
-there exists a number $T_0$ and a function $n_0(T)\in \polylog(T)$
+Accordingly, BIR-dominance can be redefined as follows: there exists a number $T_0$ and a function $n_0(T)\in \polylog(T)$
 such that
@@
 
-Similar to the condition \eqref{eq:random-better-weaker}, we can also
-generalize the weak BIR-dominance condition \eqref{eq:SoftMax-better}
-to handle the dependence on $T$: there exist some $T_0$, a function
+Similarly, weak BIR-dominance can be redefined as follows: there exist some $T_0$, a function
 $n_0(T)\in \polylog(T)$, and constants $\beta_0,\alpha_0\in (0, 1/2)$, such that
@@
 
-We also provide a version of Theorem~\ref{thm:SoftMax-weak} under this
-more general weak BIR-dominance condition; its proof is very similar
-and is omitted.
+Theorem~\ref{thm:random-clean} and Theorem~\ref{thm:SoftMax-weak} extend to these generalized versions; the easy modifications are omitted.

### 2020-07-09 journal paper/content/sec-theory-proofs.tex
Minor edit.
Added some notation in the "preamble".

@@
+\xhdr{Some notation.} Without loss of generality, we label actions as $A=[K]$ and sort them according to their prior mean rewards, so that
+    $ \E[\mu_1] > \E[\mu_2] > \ldots > \E[\mu_K]$.
+
+Fix principal $i\in \{1,2\}$ and (local) step $n$.
+The arm chosen by algorithm \alg[i] at this step is denoted $a_{i,n}$, and the corresponding \BIR is denoted $\BIR_i(n)$.
+History of \alg[i] up to this step is denoted $H_{i,n}$.
+
+Fix agent $t$. Recall that $n_i(t)$ denotes the number of global rounds before $t$ in which principal $i$ is chosen. Let $\posteriorN{i}{t}$ denote the distribution of $n_i(t)$.
+
+\ascomment{Steven: in the ITCS paper, we defined $\posteriorN{i}{t}$ as the posterior distribution given all info available to agent $t$. But this is the same thing, isn't it?}
+
+Write
+    $\PMR(a\mid E) = \E[\mu_a \mid E]$
+for posterior mean reward of action $a$ given event $E$.
+
+\xhdr{Chernoff Bounds.}
+We use an elementary concentration inequality known as {\em Chernoff Bounds}, in a formulation from~\cite{MitzUpfal-book05}.
+\begin{theorem}[Chernoff Bounds]
+\label{thm:chernoff}
+Consider $n$ i.i.d. random variables $X_1 \ldots X_n$ with values in $[0,1]$. Let
+    $X = \tfrac{1}{n} \sum_{i=1}^n X_i$ be their average, and let $\nu = \E[X]$. Then:
+\[ \min\left(\; \Pr[ X-\nu > \delta \nu ],\quad
+                \Pr[ \nu-X > \delta \nu ]
+    \; \right)
+    < e^{-\nu n \delta^2/3}
+    \quad \text{for any $\delta\in (0,1)$.}
+\]
+\end{theorem}
+
 \subsection{Main result on \HardMax: Proof of Theorem~\ref{thm:DG-dominance}}
@@
     &\geq \BIR_2(t)/2-q \\
-    &> \BIR_2(t)/4 > 0  \EqComment{by \eqref{eq:random-assn}}. \qedhere
+    &> \BIR_2(t)/4 > 0
+    &\EqComment{by \refeq{eq:random-assn}}. \qedhere
 \end{align*}

### 2020-07-09 journal paper/itcs18paper/app-perturb.tex
Copied the beta-priors example from the body; revised accordingly.

--- 2019-06-05 journal paper/itcs18 paper/app-perturb.tex
@@
-We show that Assumption~\eqref{eq:assn-distinct} holds almost surely under a small random perturbation of the prior.
-We focus on problem instances with 0-1 rewards, and assume that the prior $\priorMu$ is independent across arms and has a finite support.%
-\footnote{The assumption of 0-1 rewards is for clarity.
+We provide two examples when Assumption~\eqref{eq:assn-distinct} holds almost surely under a small random perturbation of the prior.
+We focus on problem instances with Bernoulli rewards, and assume that the prior $\priorMu$ is independent across arms.
+
+% and has a finite support.%
+%\footnote{The assumption of 0-1 rewards is for clarity.
 Our results hold under a more general assumption that for each arm $a$, rewards can only take finitely many values, and each of these values is possible (with positive probability) for every possible value of the mean reward $\mu_a$.}
- Consider the probability vector in the prior for arm $a$:
+
+\xhdr{Beta priors.} Suppose the mean reward $\mu_a$ for each arm $a$ is drawn from some Beta distribution $\text{Beta}(\alpha_a, \beta_a)$.
+Given any history $H$ that contains $h_a$ number of heads and $t_a$ number of tails from arm $a$, the posterior mean reward is
+    $\E[\mu_a \mid H] =  \frac{\alpha_a + h_a}{\alpha_a + h_a + \beta_a + t_a}$.
+Therefore, perturbing the
+    parameters $\alpha_a$ and $\beta_a$ independently with any
+    continuous noise will induce a prior with property
+    \eqref{eq:assn-distinct} with probability 1.
+
+\xhdr{A prior with a finite support.} Consider the probability vector in the prior for arm $a$:
 \[ \vec{p}_a = \left(\; \Pr[\mu_a=\nu]:\; \nu\in \support(\mu_a)\; \right).\]

### 2020-07-09 journal paper/content/sec-model.tex
Minor edits.

@@
 Agent $t$ arrives and chooses a principal $i_t\in \{1,2\}$.
-The principal chooses action $a_t\in A$, where $A$ is a fixed set of actions (same for both principals and all rounds).
+The principal chooses action $a_t\in A$, where $A$ is a fixed set of actions (same for both principals and all rounds).%
+\footnote{Throughout, we use `action' and `arm' interchangeably, as common in the literature on multi-armed bandits.}
  The agent experiences this action, receives a reward $r_t\in \{ 0,1\}$, and reports it back to the principal.
@@
 Agents know the principals' algorithms.
-Their reward estimates are defined as Bayesian-expected rewards:
-        $\est_i(t) := \E\sbr{ r_t\mid i_t = i }$
+Their reward estimates are defined as posterior mean rewards:
+        $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i }$
     for each agent $t$ and principal $i$.

### 2020-07-09 journal paper/content/sim_details.tex
Added "prelims" subsections.
A quick pass on experiments.

@@
 
-\xhdr{Key Model Differences}
-The timing of the model is the same as before where each firm commits to a MAB algorithm before the game starts and uses this algorithm to choose its actions.
+\subsection{Experiment setup}
+\label{expts-prelims}
+
+%The timing of the model is the same as before where each firm commits to a MAB algorithm before the game starts and uses this algorithm to choose its actions.
 We focus on i.i.d.
@@
 
-In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``late entrant"), and therefore enjoys a \emph{first-mover advantage}.
+\xhdr{Agent model.} All experiments use \HardMax response function (without mentioning it), except Section~\ref{sec:non_greedy} where we use \HardMaxRandom agents.
+
+\xhdr{First-movers.}
+Recall that in some of our experiments, one firm is the ``incumbent" who enters the market before the other (``late entrant"), and therefore enjoys a \emph{first-mover advantage}.
 Formally, the incumbent enjoys additional $X$ rounds of the ``warm start".
@@
 
-\xhdr{MAB algorithms.} We consider the same three classes of bandit algorithms discussed in Section ~\ref{sec:model}, but take a representative algorithm from each class and look for qualitative differences between the different classes.
-We will utilize Thompson Sampling (\TS) from the ``smart" MAB algorithms, $\eps$-greedy (\DEG)\footnote{Throughout, we fix $\eps = 0.05$.
-Our pilot experiments showed that different $\eps$ did not qualitatively change the results.} from the ``naive" algorithms, and \DynGreedy (\DG).
-For ease of comparison, all three algorithms are parameterized with the same fake prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution.
+\xhdr{MAB algorithms.} We consider the same three classes of bandit algorithms discussed in Section~\ref{sec:intro-model}, but take a representative algorithm from each class and look for qualitative differences between the different classes.
+We will utilize Thompson Sampling (\TS) from the ``adaptive-exploration" MAB algorithms, the dynamic $\eps$-greedy (\DEG)\footnote{Throughout, we fix $\eps = 0.05$.
+Our pilot experiments showed that different $\eps$ did not qualitatively change the results.} from the ``exploration-separating" algorithms, and the ``dynamic greedy" algorithm (\DG) as in Section~\ref{sec:theory-prelims}.
+For ease of comparison, all three algorithms are parameterized with the same ``fake" Bayesian prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution.
 Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.
+
+\ascomment{TODO: add a bit more detail -- \eg define \eps-greedy.}
 

### 2020-07-09 journal paper/ec19paper/content/non_greedy_choice.tex
Reformatted so that: - all expts in the body are in one section. - all proofs in the app are in one section - all expts in the app are in one section.
A quick pass on experiments.

--- 2019-02-14 paper/content/non_greedy_choice.tex
@@
 
-\section{Non-deterministic Choice Models}\label{sec:non_greedy}
+\subsection{Non-deterministic choice Model (HardMaxRandom)}\label{sec:non_greedy}
 
-Let us consider an extension in which the agents' choice is no longer deterministic.
-Recall that in our main model agents deterministically choose the firm with the higher reputation score; call this choice rule \emph{HardMax} (\HM).
-Now, we introduce some randomness: each agent selects between the firms uniformly with probability $\eps\in (0,1)$, and takes the firm with the higher reputation score with the remaining probability; call this choice rule \emph{HardMax with randomness} (\HMR).
+Let us consider an extension in which the agents' response function \eqref{eq:model-f} is no longer deterministic.
+We focus on \HardMaxRandom model, where each agent selects between the firms uniformly with probability $\eps\in (0,1)$, and takes the firm with the higher reputation score with the remaining probability.
 
-One can view \HMR as a version of ``warm start", where a firm receives some customers without competition, but these customers are dispersed throughout the game.
+One can view \HardMaxRandom as a version of ``warm start", where a firm receives some customers without competition, but these customers are dispersed throughout the game.
 The expected duration of this ``dispersed warm start" is $\eps T$.
@@
 instead they are dispersed throughout the game.}
+
+\begin{finding}\label{find:non_greedy_choice}
+\textit{\TS is weakly dominant under \HardMaxRandom, if and only if $\eps T$ is sufficiently large. Moreover, \HardMaxRandom leads to lower variance in market share, compared to \HardMax.}
+\end{finding}
 
@@
 \normalsize
-\caption{\HM and \HMR choice models on the Heavy-Tail MAB instance.
+\caption{\HardMax (\HM) and \HardMaxRandom (\HMR) choice models on the Heavy-Tail MAB instance.
 Each cell describes the market shares in a game between two algorithms, call them Alg1 vs.
@@
 \end{tabular}
-\caption{\HMR choice model for Uniform and Needle-In-Haystack MAB instances.}
+\caption{\HardMaxRandom (\HMR) choice model for Uniform and Needle-In-Haystack MAB instances.}
 %Same semantics as in Table \ref{tab:non_greedy_table}.
@@
 \normalsize
+
+Table \ref{tab:non_greedy_table} shows the average market shares under
+\HardMax vs \HardMaxRandom. In contrast to what happens under \HardMax,
+  \TS becomes weakly dominant under \HardMaxRandom, as $T$ gets
+  sufficiently large. These findings hold across all problem
+instances, see Table \ref{tab:additional_results} (with the same semantics as in Table \ref{tab:non_greedy_table}).
+
+\OMIT{\footnote{The results here are pulled using different \MRV
+  and realizations than the results pulled previously (due to the
+  larger $T$). However, they are drawn from the same prior instances
+  and so qualitatively are the same but the quantitative results are
+  not directly comparable to those from the previous
+  section.\swcomment{not sure we need this}}}
+
+\normalsize
+\OMIT{The intuition for why \TS
+should become the dominant strategy eventually is simple. The
+consistent stream of random agents ensures that each principal is
+chosen at least $\Omega(\eps t)$ times at every time step $t$. As
+a result, each algorithm should eventually converge to its asymptotic
+performance in isolation.}
+
+\OMIT{Finding \ref{find:non_greedy_choice} implies that we can re-interpret the inverted-U findings from before in terms of the number of agents that a firm receives without having to worry about incentives.
+In the extreme when the firm gets all agents for free as in the monopoly case then it is incentivized to play \DG.
+When it only gets some of the agents for free, either via a large warm start, a temporary monopoly, or non-deterministic choice, then \TS is incentivized.
+However, if the number of free agents gets small enough then \DG is incentivized as in the permanent duopoly analysis from before.\swcomment{I don't understand this paragraph; I think it's too vague.
+Perhaps remove?}}
 

### 2020-07-09 journal paper/journal_main.tex
Renamed the "model" section.
Moved proofs to the appendix.
Renamed a section.
Swapped two subsections.
Revised the abstract.

@@
 \begin{abstract}
-Most modern firms strive to learn from interactions with consumers, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-We initiate a study of the interplay between \emph{exploration} and \emph{competition}—how such firms balance the exploration for learning and the competition for consumers.
-Here consumers play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing firms.
+Most online platforms strive to learn from interactions with consumers, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
+We initiate a study of the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for consumers.
+Here consumers play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
 
-We consider competition between two firms facing the same multi-armed bandit instance who simultaneously choose a bandit algorithm.
-Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance if and only if it is chosen.
-We study to what extent competition leads to welfare increases for consumers and what algorithms firms adopt under competition.
-We find that stark competition induces firms to commit to a ``greedy" algorithm that leads to low consumer welfare.
+We consider a stylized duopoly model in which two firms face the same instance of multi-armed bandits.
+Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen.
+We study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for consumers.
+We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low consumer welfare.
 However, we find that weakening competition by providing firms with some ``free" consumers incentivizes better exploration strategies and increases consumer welfare.
-Our findings shed light on the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets and are closely related to the ``competition vs.
+We investigate two channels for weakening the competition: relaxing the rationality of consumers and giving one firm a first-mover advantage.
+We provide a mix of theoretical results and numerical simulations.
+
+Our findings are closely related to the ``competition vs.
 innovation" relationship, a well-studied theme in economics.
+They also elucidate the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.
 
@@
 
+\newpage
 \section{Related work}
@@
 
-\section{Our model and preliminaries}
+\section{Our model in detail}
 \label{sec:model}
@@
 \subfile{content/sec-theory}
-
-\subsection{Proofs}
-\label{sec:theory-proofs}
-\subfile{content/sec-theory-proofs}
 
@@
 
+\subfile{ec19paper/content/non_greedy_choice}
+
 \subfile{ec19paper/content/revisited}
-
-\subfile{ec19paper/content/non_greedy_choice}
 
@@
 
-\section{Supplementary theoretical results}
+\section{Background on multi-armed bandits}
 \label{app:examples}
@@
 
+\section{Proofs for Section~\ref{sec:theory}}
+\label{sec:theory-proofs}
+\subfile{content/sec-theory-proofs}
+
 \subfile{ec19paper/content/appendix_for_one_version}
 
-\subfile{content/old-intro}
+%\subfile{content/old-intro}
 

### 2020-07-09 journal paper/content/related_work.tex
A quick pass on related work.

@@
-Multi-armed bandits (\emph{MAB}) is a particularly elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information.
-MAB problems have been studied in Economics, Operations Research and Computer Science for many decades; see \citep{Bubeck-survey12,Gittins-book11,slivkins-MABbook} for background on regret-minimizing and Bayesian formulations, respectively.
-A discussion of industrial applications of MAB can be found in \citet{MWT-WhitePaper-2016}.
+\xhdr{Bandits and exploration.} Multi-armed bandits (\emph{MAB}) is a particularly elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information.
+MAB problems have been studied in Economics, Operations Research and Computer Science for many decades, \asedit{and are covered in many books and surveys.
+Particularly, see
+\citet{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} for background on regret-minimizing formulations of bandits, \citet{Gittins-book11} for Bayesian and Markovian formulations, and \citet{CesaBL-book,slivkins-MABbook} for connections to economics and game theory.} A discussion of industrial applications of MAB can be found in \citep{DS-arxiv} and \citep[Ch.
+8]{slivkins-MABbook}.
 
@@
 thread concerns regret-minimizing MAB formulations with IID rewards
-\citep{Lai-Robbins-85,bandits-ucb1}. This thread includes ``smart" MAB
-algorithms that combine exploration and exploitation, such as UCB1
-\citep{bandits-ucb1}, Thompson Sampling \citep{russo2018tutorial}, and Successive Elimination
-\citep{EvenDar-icml06}, and ``naive'' MAB algorithms that separate
-exploration and exploitation, including explore-first and
-$\eps$-Greedy \citep[\eg see][]{slivkins-MABbook}.
+\citep{Lai-Robbins-85,bandits-ucb1}. This thread includes ``smart" adaptive-exploration
+algorithms such as UCB1
+\citep{bandits-ucb1}, Thompson Sampling \citep{TS-survey-FTML18}, and Successive Elimination
+\citep{EvenDar-icml06}, and ``naive'' exploration-separating algorithms, including explore-first and
+Epsilon-Greedy 
+\asedit{\citep[\eg see Ch. 1 in ][for discussion and literature review]{slivkins-MABbook}.}
 
@@
 incentivizing exploration in a recommendation system
-    \citep{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,ICexplorationGames-ec16-working},
+    \citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},
 dynamic auctions
@@
     \cite[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
-
 \citet{Bolton-econometrica99,Keller-econometrica05,Johari-ec12} studied models with self-interested agents jointly performing exploration, with no principal to coordinate them.
@@
 Also, this work specifically assumes binary payoffs (\ie win or lose) for the principals.
-\gaedit{\cite{ben2017best, ben2019regression} similarly study competing learning algorithms, though they consider competition between offline learning algorithms, whereas our focus here is on online learning algorithms.}
+\gaedit{\citet{ben2017best, ben2019regression} similarly study competing learning algorithms, though they consider competition between offline learning algorithms, whereas our focus here is on online learning algorithms.}
 
@@
 In these models, firms learn while competing against each other where a firm attracting more consumers reduces its per-unit production costs.
-However, our model focuses on firms learning product quality as opposed to reducing production costs and the impact that this learning has on attracting consumers.}
+However, our model focuses on firms learning product quality (as opposed to reducing production costs) and the impact that this learning has on attracting consumers.}
 
-\gaedit{A line of work on \emph{platform competition}, starting with \cite{Rysman09}, concerns competition between firms (\emph{platforms}) that improve as they attract more users (\emph{network effect}); see \citet{Weyl-White-14} for a recent survey.
+A line of work on \emph{platform competition}, starting with \cite{Rysman09}, concerns competition between firms (\emph{platforms}) that improve as they attract more users (\emph{network effect}); see \citet{Weyl-White-14} for a recent survey.
 This literature is not concerned with \innovation, and typically models network effects exogenously, whereas in our model network effects are endogenous: they are created by MAB algorithms, an essential part of the model.
-A nascent literature studies whether and when network effects manifest themselves in data-intensive markets \citep{prufer2017competing, hagiu2020data}, but typically model learning as a reduced-form function of past consumer history and focus on the role of prices, as opposed to the reputational consequences of learning.
+\gaedit{A nascent literature studies whether and when network effects manifest themselves in data-intensive markets \citep{prufer2017competing, hagiu2020data}, but typically models learning as a reduced-form function of past consumer history and focus on the role of prices.}
+%as opposed to the reputational consequences of learning.
 
-\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers.
+\gaedit{\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers.
 We observe a similar effect when we investigate the role that reputation can serve as a barrier to entry.

### 2020-07-09 journal paper/ec19paper/content/inverted_u.tex
Reformatted so that: - all expts in the body are in one section. - all proofs in the app are in one section - all expts in the app are in one section.
formatting tweaks
Inserted an inverted-U figure.

@@
 \begin{document}
-\section{Competition vs. Better Algorithms}\label{sec:competition}
+\subsection{Competition vs. Better Algorithms}\label{sec:competition}
 
@@
 
+\tikzstyle{level 1}=[level distance=3.5cm, sibling distance=4.0cm]
+\tikzstyle{level 2}=[level distance=3.5cm, sibling distance=2cm]
+\tikzstyle{below} = [align=center]
+
+\begin{figure}[t]
+\begin{center}
+\begin{tikzpicture}
+      \draw[->] (-.5,0) -- (7,0) node[right] {};
+      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
+      \draw[scale=0.6,domain=0.7:9.8,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - 0.18 * (\x - 5.25)^2});
+     \node[below] at (0.7, -0.22) {\footnotesize monopoly};
+     \node[below] at (3, -0.2) {\footnotesize incumbent};
+     \node[below] at (4.5, -0.22) {\footnotesize duopoly};
+     \node[below] at (6, -0.2) {\footnotesize entrant};
+ \end{tikzpicture}
+ \caption{A stylized ``inverted-U relationship" between strength of competition and ``level of innovation".}
+\label{fig:inverted-U-expts}
+\end{center}
+\end{figure}
+
 \xhdr{Inverted-U relationship.}
@@
 For the Uniform instance, \DG goes from a Nash equilibrium strategy to a weakly dominant one.}
-Finally, the middle level of competition, ``incumbent" in the temporary monopoly creates strong incentives for \TS. In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U}.
+Finally, the middle level of competition, ``incumbent" in the temporary monopoly creates strong incentives for \TS.
+In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U-expts}.
 
@@
 \begin{figure}
-\includegraphics[scale=0.3]{ec19paper/figures/unif_eq_welfare}
-\includegraphics[scale=0.3]{ec19paper/figures/ht_eq_welfare}
+\centering
+\includegraphics[scale=0.35]{ec19paper/figures/unif_eq_welfare}
+\includegraphics[scale=0.35]{ec19paper/figures/ht_eq_welfare}
 \caption{Smoothed welfare plots resulting from equilibrium strategies in the different market structures.
@@
 \begin{figure}
+\centering
 \includegraphics[scale=0.3]{ec19paper/figures/eeog_vs_num_ht}

### 2020-07-09 journal paper/ec19paper/content/inverted_u.tex
Added a footnote re monopoly.

@@
 Finally, the middle level of competition, ``incumbent" in the temporary monopoly creates strong incentives for \TS.
-In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U-expts}.
+In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U-expts}.%
+\footnote{\asedit{We consider the ``permanent monopoly" scenario for comparison only, without presenting any findings. We just assume that a monopolist chooses
+the greedy algorithm, because it is easier to deploy in practice.
+Implicitly, users have no ``outside option": the service provided is an improvement over not having it (and therefore the monopolist is not incentivized to deploy better learning algorithms).
+This is plausible with free ad-supported platforms such as Yelp or Google.}}
 

### 2020-07-09 journal paper/content/sec-intro.tex
Added a sentence.
Minor edits in "econ interpretation: network effects".

@@
 The uniform choice among principals corresponds to no rationality and no competition.
+\asedit{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
 
@@
 Consumer welfare is higher under early/late arrival than under simultaneous entry.
-We frame these results in terms inverted-U relationships.
+We frame these results in terms of an inverted-U relationship.
 
@@
 
-\ascomment{Guy: "anti-trust xhdr" synthesises several points from various places, pls see if I got it right.}
+\ascomment{The xhdr below is basically from Guy, with minor edits by Alex.}
 
-\xhdr{Economic interpretation: data network effects.}
-\gaedit{Our model speaks to the ongoing debate on the role that data plays in the digital economy.
-One fundamental question in this debate is whether data can serve a similar role as traditional network effects, whereby competition for the market is more important than competition in the market due to the winner-take-all nature of such markets \citep{Rysman09, jullien2019economics}.
-The ``death spiral" effect that we find has a similar flavor as this as once a firm engages in exploration it gets completely starved of users.
-Indeed, we further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market.
-A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them as deploying algorithms for solving this problem.
-Thus, we do not explicitly model such network effects, but rather they arise endogenously from our setup.
+\xhdr{Economic interpretation: network effects \asedit{of data}.}
+Our model speaks to the ongoing debate on the role that data plays in the digital economy.
+One fundamental question in this debate is whether data can serve a similar role as traditional network effects, whereby competition \emph{for} the market is more important than competition \emph{in} the market, due to the winner-take-all nature of \asedit{the digital economy} \citep{Rysman09, jullien2019economics}.
+The ``death spiral" effect that we find has a similar flavor: 
+\asdelete{as this as}
+once a firm engages in exploration it gets completely starved of users.
+We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market.
+A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.
+Thus, we do not explicitly model \asedit{the} network effects, but they arise endogenously from our setup.
+
+\ascomment{Guy: could you pls clarify the diff between "for the market" and "in the market".
+Is this the distinction between "competing for market share" and "competing for profit"?
+If so, I understand what you mean by "for" the market; but why "in" the market?
+Perhaps a small edit would make it much clearer?}
 
 Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
-Approaches such as \cite{bajari2018impact} that argue that the diminishing returns to scale and scope of data in isolation render such data feedback loops as non-existent ignore the differences induced by learning in isolation versus under competition.
+Approaches such as \citet{bajari2018impact} argue that the diminishing returns to scale and scope of data in isolation \asedit{mitigate} such data feedback loops,
+%as non-existent
+\asedit{but} ignore the differences induced by learning in isolation versus under competition.
 Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.
-}
+
+\ascomment{Guy: Previously you also cited \cite{varian2018artificial,lambrecht2015can} for the same thing as \citet{bajari2018impact}.
+Why not now?
+Also, previously you explicitly mentioned "barriers to entry" and "market tipping" and "anti-trust implications".
+Why not now?
+I don't really know any of this literature, so I defer to your judgement on this ...}
 
@@
 It is is not necessarily the most realistic model, but arguably the most natural one to study \emph{a priori}.
-However, our results elucidate the need for more refined models with ``free exploration" (\eg via random agents or early entry).
+However, our results elucidate the need for more refined models that allow for ``free exploration" (\eg via random agents or early entry).
 

### 2020-07-09 journal paper/content/related_work.tex
Added a comment.
Added a para on "greedy".

@@
 Particularly, see
-\citet{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} for background on regret-minimizing formulations of bandits, \citet{Gittins-book11} for Bayesian and Markovian formulations, and \citet{CesaBL-book,slivkins-MABbook} for connections to economics and game theory.} A discussion of industrial applications of MAB can be found in \citep{DS-arxiv} and \citep[Ch.
-8]{slivkins-MABbook}.
+\citet{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} for background on regret-minimizing formulations of bandits, \citet{Gittins-book11} for Bayesian and Markovian formulations, and \citet{CesaBL-book,slivkins-MABbook} for connections to economics and game theory.}
 
@@
 \asedit{\citep[\eg see Ch. 1 in ][for discussion and literature review]{slivkins-MABbook}.}
+
+\asedit{The greedy algorithm is terrible on a wide variety of problem instances, in the sense that it fails to try the best arm and therefore suffers regret linear in the time horizon  
+\citep[see Chapter 11.2 in][]{slivkins-MABbook}. 
+On the other hand, a recent line of work \citep{kannan2018smoothed,bastani2017exploiting,externalities-colt18}
+finds that the greedy algorithm performs well (in theory), in a model of linear contextual bandits under substantial assumptions on smoothing or heterogeneity of the context vectors.}
+
+A discussion of industrial applications of MAB can be found in \citep{DS-arxiv} and \citep[Chapter 8]{slivkins-MABbook}.
+\asedit{In particular, switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training.}
 
@@
 
+\ascomment{Guy: In EC20 you cited \cite{kerin1992first} for a survey on first-mover advantage. Why not now? Likewise, you also had a para on Lerner Inder and such (last para in related work in EC20).}
+
 % moved this thought to the intro.

### 2020-07-09 journal paper/ec19paper/content/perf_in_iso.tex
updated performance in isolation figures to be PDF + slightly nicer

@@
 \centering
-\includegraphics[scale=0.3]{ec19paper/figures/nih_iso_mean}
-\includegraphics[scale=0.4]{ec19paper/appendix_figures/uniform_mean}
+\includegraphics[scale=0.35]{ec19paper/figures/nih_iso_mean}
+\includegraphics[scale=0.35]{ec19paper/figures/uniform_mean}
 %\caption*{\tiny{Mean reputation trajectory The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
@@
 \includegraphics[scale=0.35]{ec19paper/figures/relative_uniform_annotated_plot}
-\includegraphics[scale=0.35]{ec19paper/figures/relative_nih_ts_dg.png}
+\includegraphics[scale=0.35]{ec19paper/figures/relative_nih_ts_dg}
 %\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.

### 2020-07-09 analysis_code/compare_across.R

@@ -1,6 +1,6 @@
 library(dplyr)
 
-dat_1 <- read.csv("/Volumes/Mac/final_bandit_results/results/preliminary_raw_results/longer_ws.csv")
+dat_1 <- read.csv("/Users/guyaridor/Dropbox/Competing Bandits/Isolation Results/longer_ws.csv", stringsAsFactors=F)
 dat_1$Algorithm<- sapply(dat_1$Algorithm, concise_alg_rep)
 #dat_2 <- read.csv("/Users/garidor/Desktop/CompetingBandits/results/tournament_raw_results/tournament_experiment_full_sim_with_realizations_raw.csv")
 
@@ -52,7 +52,7 @@ iso_dat <- filter(dat_1, Distribution == dist)
 iso_dat_t <- filter(iso_dat, t == 1000)
 ggplot(iso_dat_t, aes(Realized.Reputation, colour=Algorithm)) +
   geom_density() +
-  ggtitle("Quality Distribution, Needle In Haystack, t=1000") +
+  ggtitle("Reputation Distribution, Needle In Haystack, t=1000") +
   theme_bw(base_size = 12) +
   xlab("Quality") + 
   theme(plot.title = element_text(hjust = 0.5))

### 2020-07-09 analysis_code/relative_reputation_plots.R

@@ -2,8 +2,10 @@ library(dplyr)
 library(ggplot2)
 library(reshape2)
 library(pBrackets)
+WORKING_PATH <- "/Users/guyaridor/Dropbox/Competing Bandits/Isolation Results"
+setwd(WORKING_PATH)
 
-dat <- read.csv("/Volumes/Mac/final_bandit_results/results/preliminary_raw_results/longer_ws.csv")
+dat <- read.csv("longer_ws.csv", stringsAsFactors=F)
 
 
 bracketsGrob <- function(...){
@@ -15,9 +17,10 @@ bracketsGrob <- function(...){
   }, e)
 }
 
-b1 <- bracketsGrob(0.18, 0.15, 0, 0.15, h=0.1, lwd=2, col="red")
+b1 <- bracketsGrob(0.33, 0.15, 0, 0.15, h=0.1, lwd=2, col="red")
 b2 <- bracketsGrob(1, 0.15, 0.35, 0.15, h=0.1,  lwd=2, col="red")
 concise_alg_rep <- function(alg) {
+  alg <- as.character(alg)
   if (alg == "ThompsonSampling") {
     return("TS")
   } else if (alg == "DynamicEpsilonGreedy") {
@@ -26,12 +29,14 @@ concise_alg_rep <- function(alg) {
     return("DG")
   }
 }
+
+dat$alg<- sapply(dat$Algorithm, concise_alg_rep)
 print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
   dist_dat <- filter(dat, Distribution == dist & Realized.Complexity >= minComplexity)
   alg_1 <- filter(dist_dat, Algorithm == alg1)
   alg_2 <- filter(dist_dat, Algorithm == alg2)
   n_vals <- unique(dist_dat$n)
-  t_vals <- seq(10, 1000, 10)
+  t_vals <- seq(10, 1500, 10)
   df <- as.data.frame(matrix(nrow=length(t_vals), ncol=3))
   colnames(df) <- c("t", "relative_rep")
   df$t <- t_vals
@@ -49,22 +54,22 @@ print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
   }
   df$relative_rep <- counts
   df$se <- ses
-  plot_title <- paste("Relative Quality -", dist)
+  plot_title <- paste("Relative Reputation -", dist)
   q <- ggplot(df, aes(x=t, y=relative_rep)) + geom_line() + geom_point() +
[…]

### 2020-07-10 journal paper/ec19paper/content/inverted_u.tex
finish pass throug h figures

@@
 \caption{Smoothed welfare plots resulting from equilibrium strategies in the different market structures.
-Note that welfare at $t = 0$ incorporates the regret incurred during the incumbent and warm start periods.}
+Note that welfare at $t = 0$ incorporates the regret incurred during the incumbent and warm start periods.
+The Thompson Sampling trajectory displays the regret incurred by running Thompson Sampling in isolation on the given instances.}
 \label{eq_regret}
@@
 \centering
-\includegraphics[scale=0.3]{ec19paper/figures/eeog_vs_num_ht}
+\includegraphics[scale=0.3]{ec19paper/figures/eeog_vs_num_firms_ht}
 \includegraphics[scale=0.3]{ec19paper/figures/ht_many_firm_welfare}\\
-\caption{Average welfare and \Eeog as we increase \#firms playing \DG}
+\caption{Average welfare and \Eeog as we increase the number of firms playing \DG}
 \label{many_firm_welfare}

### 2020-07-10 analysis_code/welfare.R

@@ -2,15 +2,16 @@ library(tidyverse)
 library(ggplot2)
 library(gdata)
 
-simultaneous_welfare <- read.csv("/Volumes/My Passport/bandits/tournament_raw_results/tournament_experiment_full_welfare_calc_raw.csv")
-free_obs_welfare <- read.csv("/Volumes/My Passport/bandits/free_obs_raw_results/free_obs_experiment_welfare_calc_raw.csv")
+setwd("/Users/guyaridor/Dropbox/bandits_results")
+simultaneous_welfare <- read.csv("tournament_raw_results/tournament_experiment_full_welfare_calc_raw.csv")
+free_obs_welfare <- read.csv("free_obs_raw_results/free_obs_experiment_welfare_calc_raw.csv")
 #prelim <- read.csv("/Volumes/Mac/final_bandit_results/results/preliminary_raw_results/longer_ws.csv")
-prelim <- read.csv("/Volumes/My Passport/bandits/preliminary_raw_results/regret_calc.csv")
+prelim <- read.csv("preliminary_raw_results/regret_calc.csv")
 sim_eq_welfare <- filter(simultaneous_welfare, ((Prior == "Heavy Tail" | Prior == "Uniform") & (P1.Alg == "DynamicGreedy" & P2.Alg == "DynamicGreedy")) | (Prior == "Needle In Haystack" & P1.Alg == "ThompsonSampling" & P2.Alg == "ThompsonSampling"))
 free_obs_eq_welfare <- filter(free_obs_welfare, ((Prior == "Heavy Tail" | Prior == "Uniform") & (P1.Alg == "DynamicGreedy" & P2.Alg == "ThompsonSampling")))
 one_welfare <- combine(sim_eq_welfare, free_obs_eq_welfare)
 
-one_welfare <- one_welfare %>% mutate(label=ifelse(source == "free_obs_eq_welfare", "Temporary Monopoly", "Duopoly"))
+one_welfare <- one_welfare %>% mutate(label=ifelse(source == "free_obs_eq_welfare", "First-Mover", "Duopoly"))
 
 prelim <- prelim %>% mutate(Prior = Distribution, Warm.Start = 20, label = "Monopoly", Total.Regret = Cumulative.Regret)
 prelim <- prelim %>% mutate(Time.Horizon = t - 20)
[…]

### 2020-07-10 journal paper/content/sec-theory.tex
Updated some comments after talking with Steven.

@@
 
-\ascomment{Steven: we used to have a much more cumbersome definition (commented out), not sure why!}
+\ascomment{TODO: we used to have a much more cumbersome definition (commented out), and we need it back! because, what if we have two copies of the same randomized algorithm, this eqn will hold!}
 
-\ascomment{Steven: in the thm, can we just say "deviates at $n_0$" instead of "deviates starting from $n_0$"? What do we mean by the latter, anyway?}
+\ascomment{TODO: define "deviate starting from" as just "this is the first $n_0$ when the deviation happens.}
 
@@
 
-\ascomment{Steven: why can't we start `mixing' at $n=1$?}
-
 Note that the modified algorithm is another pure strategy in the competition game between the two principals, not a mixture of \alg[1] and the greedy algorithm.
@@
 For the sake of intuition, let us argue how this may translate  into a larger \BIR compared to the original algorithm.
-Let  $\BIR'_1(n)$ be the \BIR of the ``greedy choice" after after $n-1$ steps of \alg[1].
+Let  $\BIR'_1(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1].
 Then
@@
 
-\ascomment{Steven: why can't we take $c=1$ below??}
-
 Suppose $\BIR_1(n)= \beta n^{-\gamma}$ for some constants $\beta,\gamma>0$. Further, assume
@@
 for some $c>1-\gamma$.
+\ascomment{Steven,Alex: not clear why this holds ...}
+
 Then for all $n\geq n_0$ and small enough $p>0$ it holds that:
@@
 
-\ascomment{Steven: in this thm and the next, do we need the algorithms to be monotone? We didn't assume this in the conf paper, which I assume was a minor bug.}
-
 \begin{proof}[Proof Sketch]
@@
 
-\ascomment{Steven: in the conf paper, we spell out the defn of $\BIR_1(n) \to 0$. We don't need to, do we?}
-
 \begin{proof}[Proof Sketch]
@@
 \footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition.
-\ascomment{Steven, Yishay: how do we explain why? Pls help ...}}
+\ascomment{TODO: can we actually prove it? Assume Bernoulli priors, discrete prior, reduce to Lai-Robbins.}}
 Suppose a feasible algorithm \alg BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.

### 2020-07-10 journal paper/content/related_work.tex
tweaks to related work

@@
 
-\gaedit{\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers.
+\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers.
 We observe a similar effect when we investigate the role that reputation can serve as a barrier to entry.
 However, in our model this effect is further strengthened by the fact that the firms have to learn while competing adding that the incumbent may not only have a reputational advantage, but additionally has a further advantage due to the data it acquires in the incumbency period.
-Thus, our model also highlights the role that data can serve as a barrier to entry in online markets which has similarly been noted in \cite{lambrecht2015can, de2020data}.}
+Thus, our model also highlights the role that data can serve as a barrier to entry in online markets which has similarly been noted in \cite{de2020data}.
+\gaedit{For an extensive overview of the other channels through which first-mover advantages can lead to a competitive advantage, see \cite{kerin1992first}.}
 
@@
 Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom.
+\gaedit{We utilize first-mover advantage and relaxed versions of rationality to model varying competition, instead of classic market ``competitiveness" measures such as the Lerner Index or the Herfindahl-Hirschman Index, since these measures rely on ex-post observable attributes of a market such as prices or market shares \cite{tirole1988theory}.
+However, neither is applicable to our setting: in our model, there are no prices, and market shares are endogenous.}
 
 \ascomment{Guy: In EC20 you cited \cite{kerin1992first} for a survey on first-mover advantage. Why not now? Likewise, you also had a para on Lerner Inder and such (last para in related work in EC20).}
+\gacomment{Good catch, I re-added them back in. I had removed \cite{kerin1992first} since I added more directly relevant papers to the first-mover advantages in our model.}
 

### 2020-07-10 analysis_code/welfare.R

@@ -65,7 +65,7 @@ ggsave(paste("ht_many_firm_welfare.pdf", sep=""), device="pdf", plot=g, width=8.
 end_time <- ht_competition %>% filter(Time.Horizon == 2000)
 end_time <- end_time %>% group_by(NumFirms) %>% summarise(mean_eeog = mean(EEOG), sd_eeog = sd(EEOG), N=n())
 g <- ggplot(end_time, aes(x=as.numeric(NumFirms), y=mean_eeog)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=mean_eeog - 1.96 * sd_eeog / N, ymax=mean_eeog + 1.96 * sd_eeog / N)) + 
-  xlab("Number of Firms") + ylab("Effective End of Game") + ggtitle("EEOG vs Number of Firms, Heavy Tail") + 
+  xlab("Number of Firms") + ylab("Effective End of Game") + ggtitle("EoG vs Number of Firms, Heavy Tail") + 
   theme_bw(base_size = 20) + theme(plot.title = element_text(hjust = 0.5))
 print(g)
 ggsave(paste("eeog_vs_num_firms_ht.pdf", sep=""), device="pdf", plot=g, width=8.14, height=7.6)
@@ -74,7 +74,7 @@ ggsave(paste("eeog_vs_num_firms_ht.pdf", sep=""), device="pdf", plot=g, width=8.
 end_time <- unif_competition %>% filter(Time.Horizon == 2000)
 end_time <- end_time %>% group_by(NumFirms) %>% summarise(mean_eeog = mean(EEOG), sd_eeog = sd(EEOG), N=n())
 g <- ggplot(end_time, aes(x=as.numeric(NumFirms), y=mean_eeog)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=mean_eeog - 1.96 * sd_eeog / N, ymax=mean_eeog + 1.96 * sd_eeog / N)) + 
-  xlab("Number of Firms") + ylab("Effective End of Game") + ggtitle("EEOG vs Number of Firms, Uniform") + 
+  xlab("Number of Firms") + ylab("Effective End of Game") + ggtitle("EoG vs Number of Firms, Uniform") + 
   theme_bw(base_size = 20) + theme(plot.title = element_text(hjust = 0.5))
 print(g)
 ggsave(paste("eeog_vs_num_firms_unif.pdf", sep=""), device="pdf", plot=g, width=8.14, height=7.6)

### 2020-07-10 journal paper/content/sec-intro.tex
updates to intro + numerical sections

@@
 Our model speaks to the ongoing debate on the role that data plays in the digital economy.
-One fundamental question in this debate is whether data can serve a similar role as traditional network effects, whereby competition \emph{for} the market is more important than competition \emph{in} the market, due to the winner-take-all nature of \asedit{the digital economy} \citep{Rysman09, jullien2019economics}.
+One fundamental question in this debate is whether data can serve a similar role as traditional network effects, \gaedit{whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.}
 The ``death spiral" effect that we find has a similar flavor: 
@@
 once a firm engages in exploration it gets completely starved of users.
-We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market.
-A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.
+\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.} A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.
 Thus, we do not explicitly model \asedit{the} network effects, but they arise endogenously from our setup.
@@
 Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
-Approaches such as \citet{bajari2018impact} argue that the diminishing returns to scale and scope of data in isolation \asedit{mitigate} such data feedback loops,
+Approaches such as \citet{lambrecht2015can, bajari2018impact} argue that the diminishing returns to scale and scope of data in isolation \asedit{mitigate} such data feedback loops,
 %as non-existent
@@
 I don't really know any of this literature, so I defer to your judgement on this ...}
+
+\gacomment{I added back the \cite{lambrecht2015can} reference, but \cite{varian2018artificla} is a more informal discussion of these issues and not essential to include.
+However, even the \cite{lambrecht2015can} is relatively informal discussion of the same issues.
+Part of the problem that makes this tricky is that the literature on network effects is a well-developed and mature literature, but the question of whether data plays a similar role primarily started as more of a policy discussion, of which there's only starting to be some serious theoretical and empirical work on the issue.}
 

### 2020-07-10 journal paper/ec19paper/content/barriers.tex
Reformatted so that: - all expts in the body are in one section. - all proofs in the app are in one section - all expts in the app are in one section.
updates to intro + numerical sections

@@
 
-\section{Data as a Barrier to Entry}\label{sec:barriers}
+\subsection{Data as a Barrier to Entry}\label{sec:barriers}
 
@@
 
-Under temporary monopoly, the incumbent can explore without incurring immediate reputational costs, and build up a high reputation before the entrant appears.
+In the first-mover regime, the incumbent can explore without incurring immediate reputational costs, and build up a high reputation before the entrant appears.
 Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant.

### 2020-07-10 journal paper/ec19paper/content/inverted_u.tex
updates to intro + numerical sections

@@
 \normalsize
-\caption{{\bf Simultaneous Entry Duopoly}, for Heavy-Tail and Needle-in-Haystack instances.
+\caption{{\bf Simultaneous Entry}, for Heavy-Tail and Needle-in-Haystack instances.
 Each cell describes a game between two algorithms, call them Alg1 vs.
@@
 
-\xhdr{Simultaneous entry duopoly.}
+\xhdr{Simultaneous entry.}
 The basic scenario is when both firms are competing from round $1$. A crucial distinction is whether an MAB instance is exploration-disadvantaged:
@@
 \begin{finding}\label{find:duopoly}
-\textit{Under simultaneous entry duopoly:
+\textit{Under simultaneous entry:
 \begin{itemize}
@@
 \end{tabular}
-\caption{{\bf Simultaneous entry duopoly}, for the Uniform MAB instance. Semantics are the same as in Table \ref{sim_table}.}
+\caption{{\bf Simultaneous entry}, for the Uniform MAB instance. Semantics are the same as in Table \ref{sim_table}.}
 \label{tab:duopoly_unif}
@@
 We attribute the prevalence of \DG on exploration-disadvantaged MAB instances to its prevalence on the initial ``exploration disadvantage period", as described in Section~\ref{sec:isolation}.
-Increasing the warm start length $T_0$ makes this period shorter: indeed, considering relative reputation trajectory in Figure~\ref{relative_rep_plots} (top), increasing $T_0$ effectively shifts the starting time point to the right.
+Increasing the warm start length $T_0$ makes this period shorter: indeed, considering the relative reputation trajectory in Figure~\ref{relative_rep_plots} (left), increasing $T_0$ effectively shifts the starting time point to the right.
 This is why it helps \DG if $T_0$ is small.
@@
 
-\xhdr{Temporary Monopoly.}
-We turn our attention to the temporary monopoly scenario.
+\xhdr{First-Mover.}
+We turn our attention to the first-mover scenario.
 Recall that the incumbent firm enters the market and serves as a monopolist until the entrant firm enters at round $X$.
@@
 \begin{finding}\label{find:temp-monopoly}
-\textit{Under temporary monopoly, \TS is the dominant strategy for the incumbent. This holds across all MAB instances, if $X$ is large enough.
+\textit{Underfirst-mover, \TS is the dominant strategy for the incumbent. This holds across all MAB instances, if $X$ is large enough.
 }
@@
 \begin{finding}\label{find:temp-monopoly-entrant}
-\textit{Under temporary monopoly, \DG is a weakly dominant strategy for the entrant for exploration-disadvantaged MAB instances.
+\textit{Under first-mover, \DG is a weakly dominant strategy for the entrant for exploration-disadvantaged MAB instances.
 }
@@
      \node[below] at (0.7, -0.22) {\footnotesize monopoly};
-     \node[below] at (3, -0.2) {\footnotesize incumbent};
-     \node[below] at (4.5, -0.22) {\footnotesize duopoly};
-     \node[below] at (6, -0.2) {\footnotesize entrant};
+     \node[below] at (2.8, -0.2) {\footnotesize incumbent};
+     \node[below] at (4.8, -0.22) {\footnotesize sim. entry};
+     \node[below] at (6.2, -0.2) {\footnotesize entrant};
  \end{tikzpicture}
@@
 The lowest level of competition is monopoly, when \DG wins out for the trivial reason of tie-breaking.
-The highest levels are simultaneous entry duopoly and ``late entrant".
+The highest levels are simultaneous entry and ``late entrant".
 We see that \DG is incentivized for exploration-disadvantaged MAB instances.
-In fact, incentives for \DG get stronger when the model transitions from simultaneous entry duopoly to ``late entrant".%
+In fact, incentives for \DG get stronger when the model transitions from simultaneous entry to ``late entrant".%
 \footnote{For the Heavy-Tail instance, \DG goes from a weakly dominant strategy to a strictly dominant one.
 For the Uniform instance, \DG goes from a Nash equilibrium strategy to a weakly dominant one.}
-Finally, the middle level of competition, ``incumbent" in the temporary monopoly creates strong incentives for \TS.
+Finally, the middle level of competition, ``incumbent" in the first-mover regime creates strong incentives for \TS.
 In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U-expts}.%
-\footnote{\asedit{We consider the ``permanent monopoly" scenario for comparison only, without presenting any findings. We just assume that a monopolist chooses
+\footnote{\asedit{We consider the monopoly scenario for comparison only, without presenting any findings. We just assume that a monopolist chooses
 the greedy algorithm, because it is easier to deploy in practice.
@@
 Our intuition for why incumbency creates more incentives for exploration is as follows.
-During the temporary monopoly period, reputation consequences of exploration vanish.
+\gaedit{During the period in which the incumbent is the only firm in the market}, reputation consequences of exploration vanish.
 Instead, the firm wants to improve its performance as much as possible by the time competition starts.
@@
 We interpret low \Eeog as a strong evidence of the ``death spiral" effect.
-Focusing on the simultaneous entry duopoly scenario, we specify the \Eeog values in Table \ref{sim_table} (the second line of each cell).
+Focusing on the simultaneous entry scenario, we specify the \Eeog values in Table \ref{sim_table} (the second line of each cell).
 We find that the \Eeog values are indeed small:
@@
 \textit{
-Under simultaneous entry duopoly, \Eeog values tend to be much smaller than the time horizon $T$.
+Under simultaneous entry, \Eeog values tend to be much smaller than the time horizon $T$.
 }
@@
 \item \DG in the monopoly,
-\item \DG for both firms in duopoly (Finding \ref{find:duopoly}),
-\item \TS for the incumbent (Finding \ref{find:temp-monopoly}) and \DG for the entrant in temporary monopoly (Finding \ref{find:temp-monopoly-entrant}).
+\item \DG for both firms in simultaneous entry (Finding \ref{find:duopoly}),
+\item \TS for the incumbent (Finding \ref{find:temp-monopoly}) and \DG for the entrant in first-mover (Finding \ref{find:temp-monopoly-entrant}).
 \end{itemize}
@@
   over multiple runs) under different levels of competition.
-Consumers are \textit{better off} in the temporary monopoly case than in
-the duopoly case.
-Recall that under temporary monopoly, the incumbent is incentivized to play \TS.
+Consumers are \textit{better off} in the first-mover case than in
+the simultaneous entry case.
+Recall that under first-mover, the incumbent is incentivized to play \TS.
 Moreover, we find that the welfare is close to that of having a single firm for all agents and running \TS.
-We also observe that monopoly and duopoly achieve similar welfare.
+We also observe that monopoly and simultaneous entry achieve similar welfare.
 %with monopoly being marginally better than duopoly.
@@
 \begin{finding}\label{find:welfare}
-\textit{In equilibrium, consumer welfare is (a) highest under temporary monopoly, (b) similar for monopoly and duopoly.
+\textit{In equilibrium, consumer welfare is (a) highest under first-mover, (b) similar for monopoly and simultaneous entry.
 %monopoly being marginally better leading to marginally better welfare than duopoly.
@@
 \begin{finding}
-\textit{When all firms deploy \DG, and start at the same time, welfare is weakly decreasing as the number of firms \gaedit{increases}}
+\textit{When all firms deploy \DG, and start at the same time, welfare is weakly decreasing as the number of firms increases.}
 \end{finding}

### 2020-07-10 journal paper/ec19paper/content/perf_in_iso.tex
updates to intro + numerical sections

@@
 The shaded area display 95\% confidence intervals.}}
-\caption{Relative reputation trajectory for $\TS$ vs $\DG$, on Uniform instance (top) and Needle-in-Haystack instance (bottom). Shaded area display 95\% confidence intervals.}
+\caption{Relative reputation trajectory for $\TS$ vs $\DG$, on Uniform instance (left) and Needle-in-Haystack instance (right). Shaded area display 95\% confidence intervals.}
 \label{relative_rep_plots}
@@
 benefits under competition.
-Figure \ref{relative_rep_plots} (top) shows the relative reputation trajectory for $\TS$ vs $\DG$ for the Uniform instance.
+Figure \ref{relative_rep_plots} (left) shows the relative reputation trajectory for $\TS$ vs $\DG$ for the Uniform instance.
 The relative reputation is less than $\tfrac12$ in the early rounds, meaning that $\DG$ has a higher reputation score in a majority of the simulations, and more than $\tfrac12$ later on.
@@
 Such period also exists for the Heavy-Tail MAB instance.
-However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
+However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (right).%
 \footnote{We see two explanations for this: $\TS$ identifies the best arm faster for the Needle-in-Haystack instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}

### 2020-07-10 journal paper/content/conclusion.tex
Minor edit.
new conclusion

@@
 
-We considered a stylized duopoly setting where firms simultaneously learn from and compete for users.
-We analyzed two separate but connected variants of this model, one where consumers chose between firms based on their Bayesian expected reward and another where consumers chose between them based on their past performance.
-We showed in both cases that competition may not always induce firms to commit to better exploration algorithms, resulting in welfare losses for consumers.
-The primary mechanism in both was that consumers need to be incentivized to select a firm over its competitors, leading firms that engage in exploration to be starved of consumers before they make enough progress on their learning problem.
-We found that in order to incentivize ``better" exploration strategies the key intuition is that a firm needs to have some ``free" consumers that visit them without having to be incentivized to do so.
-We explored several economic mechanisms that can generate this.
-The first was by considering different variants of random utility models where stochastic choice gives firms enough free consumers to incentivize them to commit to ``better" algorithms and we explored how this varied based on the nature of the random utility model.
-The second was by considering giving one firm a first-mover advantage where all consumers in the market would go to this firm without having to be incentivized to do so.
-While this first-mover advantage lead to the incumbent firm being incentivized to commit to a ``better" algorithm it lead to a substantial market share going to this firm.
-We further analyzed to what extent this advantage came from having a more defined reputation or more data from the incumbency period.
-We found that even a small amount of ``data advantage" leads to substantial long-term market power due to the dynamics of competition.
+In this paper we studied the tension between exploration and competition.
+We considered a stylized duopoly model in which two firms face an identical multi-armed bandit problem and compete for a stream of users.
+The firms make progress on their learning problem if and only if they are able to attract a user to their platform over their competitors.
+We investigated two variants of this problem.
+The first was an analytical variant, \TheoryModel, where the users choose between the firms according to their Bayesian expected reward of visiting the firm's platform.
+The second we evaluated via numerical simulation, \ExptsModel, where users choose between the firms based on a reputation score determined by the average reward of the recent users who have utilized that firm.
+
+We find that, in both variants, firms are incentivized to adopt a greedy algorithm which does no purposeful exploration and leads to welfare losses for users.
+We then consider two relaxations of competition.
+In the first, we relax the rationality of the users and, in the second, we give one of the firms a first-mover advantage.
+Both of these relaxations of competition induce firms to adopt ``better" bandit algorithms and increase user welfare.
+
+Our model and its implications have two economic interpretations.
+The first is that, interpreting the adoption of ``better" bandit algorithms as firm innovation, our results can be framed in terms of the classic inverted-U relationship between innovation and competition.
+However, unlike other models in the literature, the reason that firms do not ``innovate" is not due to the direct costs of innovation.
+Rather, it is since ``better" algorithms engage in exploration which has negative short-term consequences on user perception of the firms' quality and thus dis-incentivizes the firm to adopt such technologies under stark competition.
+
+The second economic interpretation is that our model speaks to the role that data plays in the digital economy.
+We find a ``death spiral" effect where exploration by a firm degrades its performance relative to competitors who keep learning and improving from \emph{their} users, and so forth until the platform is completely starved of users.
+Furthermore, we find that even a small amount of initial data advantage for a firm gets amplified due to the dynamics of competition and leads to that firm acquiring a large market share.
+Thus, our results highlight that competition endogenously generates the data observed by firms, which naturally leads to results similar to those in markets with network effects and underscores that data can serve as a barrier to entry in the digital economy.
+
+There are several interesting directions to explore for future work.
+The model we consider in this paper is well-suited as a model for competition between ad-supported platforms who compete primarily on quality and whose services are provided without cost to users.
+However, it would be interesting to consider the role that giving the firms to ability to set prices would play as this could potentially enable the firms to compensate users in the earlier periods while it is engaging in exploration and prevent such ``death spiral" effects.
+The second interesting extension would be to consider a setting with horizontally differentiated user preferences\footnote{This would be similar to the extension in Section \ref{sec:theory-extensions} that considers contextual bandits, though it would be interesting to study how this would change based on different user models that had context-specific user decision rules.} as such a model would be better suited for understanding how competition may induce firms to specialize.
+In general, there are many further directions to undertake in order to better understand how the intertwined relationship between competition and data generation can help us better understand the digital economy.
 

### 2020-07-11 journal paper/content/sec-intro.tex
Some edits / comments in "network effects" xhdr.

--- 2020-07-09 journal paper/content/sec-intro.tex
@@
 
-\xhdr{Economic interpretation: network effects \asedit{of data}.}
+\xhdr{Economic interpretation: network effects of data.}
 Our model speaks to the ongoing debate on the role that data plays in the digital economy.
-One fundamental question in this debate is whether data can serve a similar role as traditional network effects, whereby competition \emph{for} the market is more important than competition \emph{in} the market, due to the winner-take-all nature of \asedit{the digital economy} \citep{Rysman09, jullien2019economics}.
-The ``death spiral" effect that we find has a similar flavor: 
-\asdelete{as this as}
-once a firm engages in exploration it gets completely starved of users.
-We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market.
+One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
+%whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
+\asedit{The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved out of users (resp., take over most of the market).}
+%\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}
 A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.
-Thus, we do not explicitly model \asedit{the} network effects, but they arise endogenously from our setup.
-
-\ascomment{Guy: could you pls clarify the diff between "for the market" and "in the market".
-Is this the distinction between "competing for market share" and "competing for profit"?
-If so, I understand what you mean by "for" the market; but why "in" the market?
-Perhaps a small edit would make it much clearer?}
+Thus, we do not explicitly model the network effects, but they arise endogenously from our setup.
 
 Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
-Approaches such as \citet{bajari2018impact} argue that the diminishing returns to scale and scope of data in isolation \asedit{mitigate} such data feedback loops,
+Approaches such as \citet{lambrecht2015can, bajari2018impact} argue that the diminishing returns to scale and scope of data in isolation mitigate such data feedback loops,
 %as non-existent
-\asedit{but} ignore the differences induced by learning in isolation versus under competition.
+but ignore the differences induced by learning in isolation versus under competition.
 Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.
@@
 I don't really know any of this literature, so I defer to your judgement on this ...}
+
+\gacomment{I added back the \cite{lambrecht2015can} reference, but \cite{varian2018artificial} is a more informal discussion of these issues and not essential to include.
+However, even the \cite{lambrecht2015can} is relatively informal discussion of the same issues.
+Part of the problem that makes this tricky is that the literature on network effects is a well-developed and mature literature, but the question of whether data plays a similar role primarily started as more of a policy discussion, of which there's only starting to be some serious theoretical and empirical work on the issue.}
 

### 2020-07-11 journal paper/content/related_work.tex
Slightly reworded the last para.

@@
 
-Finally, relaxed versions of rationality similar to ours are found in several notable lines of work.
+Relaxed versions of rationality similar to ours are found in several notable lines of work.
 For example, ``random agents" (a.k.a.
@@
 Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom.
-\gaedit{We utilize first-mover advantage and relaxed versions of rationality to model varying competition, instead of classic market ``competitiveness" measures such as the Lerner Index or the Herfindahl-Hirschman Index, since these measures rely on ex-post observable attributes of a market such as prices or market shares \cite{tirole1988theory}.
-However, neither is applicable to our setting: in our model, there are no prices, and market shares are endogenous.}
 
-\ascomment{Guy: In EC20 you cited \cite{kerin1992first} for a survey on first-mover advantage. Why not now? Likewise, you also had a para on Lerner Inder and such (last para in related work in EC20).}
-\gacomment{Good catch, I re-added them back in. I had removed \cite{kerin1992first} since I added more directly relevant papers to the first-mover advantages in our model.}
+We use first-mover advantage and relaxed versions of rationality to model varying competition, instead of classic ``market competitiveness" measures such as the Lerner Index or the Herfindahl-Hirschman Index
+\citep{tirole1988theory}.
+The latter measures rely on ex-post observable attributes of a market such as prices or market shares.
+Neither is applicable to our setting, since there are no prices, and market shares are endogenous.
+
+\ascomment{Guy: slightly reworded the previous para.}
 

### 2020-07-11 journal paper/content/sim_details.tex
Added a sentence.
Revised the intro para.

@@
 
-In this section we move to the reputation choice variant.
+\asedit{In this section we present our numerical simulations.
+As discussed in the Introduction, we focus on the \ExptsModel, whereby each agent chooses the firm with a maximal reputation score, modeled as a sliding window average of its rewards.
+Given the volume of experimental results, we (only) present the typical / representative ones in the body of the paper; the full results can be found in Appendix~\ref{app:expts}.}
+
+%In this section we move to the reputation choice variant.
 Recall that in this variant of the model, instead of making choices between firms based on the Bayesian expected reward, each agent chooses the firm with a maximal reputation score (breaking ties uniformly).
@@
 \xhdr{MAB algorithms.} We consider the same three classes of bandit algorithms discussed in Section~\ref{sec:intro-model}, but take a representative algorithm from each class and look for qualitative differences between the different classes.
+\asedit{Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.}
+
 We will utilize Thompson Sampling (\TS) from the ``adaptive-exploration" MAB algorithms, the dynamic $\eps$-greedy (\DEG)\footnote{Throughout, we fix $\eps = 0.05$.

### 2020-07-11 journal paper/content/related_work.tex
Minor edits.

@@
 \xhdr{Bandits and exploration.} Multi-armed bandits (\emph{MAB}) is a particularly elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information.
-MAB problems have been studied in Economics, Operations Research and Computer Science for many decades, \asedit{and are covered in many books and surveys.
+MAB problems have been studied in Economics, Operations Research and Computer Science for many decades, and are covered in many books and surveys.
 Particularly, see
-\citet{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} for background on regret-minimizing formulations of bandits, \citet{Gittins-book11} for Bayesian and Markovian formulations, and \citet{CesaBL-book,slivkins-MABbook} for connections to economics and game theory.}
+\citet{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} for background on regret-minimizing formulations, \citet{Gittins-book11} for Bayesian and Markovian formulations, and \citet{CesaBL-book,slivkins-MABbook} for connections to economics and game theory.
 
@@
 thread concerns regret-minimizing MAB formulations with IID rewards
-\citep{Lai-Robbins-85,bandits-ucb1}. This thread includes ``smart" adaptive-exploration
-algorithms such as UCB1
-\citep{bandits-ucb1}, Thompson Sampling \citep{TS-survey-FTML18}, and Successive Elimination
-\citep{EvenDar-icml06}, and ``naive'' exploration-separating algorithms, including explore-first and
-Epsilon-Greedy
-\asedit{\citep[\eg see Ch. 1 in ][for discussion and literature review]{slivkins-MABbook}.}
+\citep{Lai-Robbins-85,bandits-ucb1}. This thread includes adaptive-exploration
+algorithms such as 
+Thompson Sampling \citep{Thompson-1933,TS-survey-FTML18},
+UCB1 \citep{bandits-ucb1}, 
+and Successive Elimination \citep{EvenDar-icml06}, 
+as well as more ``naive'', exploration-separating algorithms such as Explore-first and Epsilon-Greedy.
+%\citep[\eg see Ch. 1 in ][for discussion and literature review]{slivkins-MABbook}.
 
@@
 
-A discussion of industrial applications of MAB can be found in \citep{DS-arxiv} and \citep[Chapter 8]{slivkins-MABbook}.
+A discussion of industrial applications of MAB can be found in \citep{DS-arxiv}. 
+%and \citep[Chapter 8]{slivkins-MABbook}. 
 \asedit{In particular, switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training.}
 
-The three-way tradeoff between exploration, exploitation and incentives has been studied in several other settings:
+The three-way tradeoff between exploration, exploitation and incentives has been studied in several settings other than ours:
 incentivizing exploration in a recommendation system
@@
 
-There is a superficial similarity (in name only) between this paper and the line of work on ``dueling bandits"
+There is a superficial similarity --- in name only --- between this paper and the line of work on ``dueling bandits"
     \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
@@
 
-Our setting is closely related to the ``dueling algorithms" framework \citep{DuelingAlgs-stoc11} which studies competition between two principals, each running an algorithm for the same problem.
-However, this work considers algorithms for offline / full input scenarios, whereas we focus on online machine learning and the explore-exploit-incentives tradeoff therein.
-Also, this work specifically assumes binary payoffs (\ie win or lose) for the principals.
-\gaedit{\citet{ben2017best, ben2019regression} similarly study competing learning algorithms, though they consider competition between offline learning algorithms, whereas our focus here is on online learning algorithms.}
+Our setting is closely related to the ``dueling algorithms" framework of \citet{DuelingAlgs-stoc11}, which studies competition between two principals, each running an algorithm for the same problem.
+However, \citet{DuelingAlgs-stoc11} consider algorithms for offline / full input scenarios, \asedit{and posit binary (win/lose) payoffs for the principals.
+Similarly, \citet{ben2017best, ben2019regression} study competition between offline learning algorithms.
+Whereas we focus on bandit algorithms and competition for users.}
 

### 2020-07-12 journal paper/content/related_work.tex
Writing "background". Daily save.

@@
-\xhdr{Bandits and exploration.} Multi-armed bandits (\emph{MAB}) is a particularly elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information.
-MAB problems have been studied in Economics, Operations Research and Computer Science for many decades, and are covered in many books and surveys.
+\xhdr{Bandits and exploration.} Multi-armed bandits (\emph{MAB}) is an elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information.
+MAB problems have been studied for many decades \asedit{by researchers from computer science, operations research, statistics and economics.
+The literature on MAB is extremely vast and multi-threaded.
+The most relevant thread studies a basic model of MAB with IID rewards and no auxiliary structure, see Section~\ref{sec:related-classes}.
+The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.} This literature is covered in many books and surveys.
 Particularly, see
 \citet{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} for background on regret-minimizing formulations, \citet{Gittins-book11} for Bayesian and Markovian formulations, and \citet{CesaBL-book,slivkins-MABbook} for connections to economics and game theory.
-
-The literature on MAB is vast and multi-threaded. The most related
-thread concerns regret-minimizing MAB formulations with IID rewards
-\citep{Lai-Robbins-85,bandits-ucb1}. This thread includes adaptive-exploration
-algorithms such as 
-Thompson Sampling \citep{Thompson-1933,TS-survey-FTML18},
-UCB1 \citep{bandits-ucb1}, 
-and Successive Elimination \citep{EvenDar-icml06}, 
-as well as more ``naive'', exploration-separating algorithms such as Explore-first and Epsilon-Greedy.
-%\citep[\eg see Ch. 1 in ][for discussion and literature review]{slivkins-MABbook}.
-
-\asedit{The greedy algorithm is terrible on a wide variety of problem instances, in the sense that it fails to try the best arm and therefore suffers regret linear in the time horizon
-\citep[see Chapter 11.2 in][]{slivkins-MABbook}.
-On the other hand, a recent line of work \citep{kannan2018smoothed,bastani2017exploiting,externalities-colt18}
-finds that the greedy algorithm performs well (in theory), in a model of linear contextual bandits under substantial assumptions on smoothing or heterogeneity of the context vectors.}
-
 A discussion of industrial applications of MAB can be found in \citep{DS-arxiv}.
-%and \citep[Chapter 8]{slivkins-MABbook}. 
-\asedit{In particular, switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training.}
 
@@
 
+\subsection{Background: the three classes of algorithms [IN PROGRESS, don't read]}
+\label{sec:related-classes}
+
+We present sufficient background on the three classes of MAB algorithms listed in Section~\ref{sec:intro-model}: from adaptive-exploration to exploration-separating to greedy.
+
+We are concerned with the following problem.
+There are $T$ rounds and $K$ \emph{arms} to choose from.
+In each round $t\in [T]$, the algorithm chooses an arm and receives a reward $r_t\in[0,1]$ for this arm, drawn from a fixed but unknown distribution.
+Algorithm's goal is to maximize the total reward.
+A standard performance measure is \emph{regret}, defined as the difference in the total expected reward between the algorithm and the best fixed arm.%
+\footnote{In a formula, regret is
+    $\E\sbr{ T\cdot \max_{\text{arms $a$}}\mu_a -  \sum_{t\in[T]} r_t }$,
+where $\mu_a$ is the mean reward of arm $a$.}
+The primary concern is the asymptotic growth rate of regret as a function of $T$.
+
+The three classes of algorithms perform very differently in terms of regret. Adaptive-exploration algorithms achieve optimal regret rates:
+    $\tildeO(\sqrt{KT})$
+for all problem instances, and an improved regret rate of 
+    $O(\tfrac{K}{\Delta}\log T)$
+for problem instances with $\text{gap}\geq \Delta$
+\citep{Lai-Robbins-85,bandits-ucb1,bandits-exp3}.%
+\footnote{The \emph{gap} is the difference in mean reward between the best arm and the second-best arm.
+As usual in computer science, the $O(\cdot)$ and $\tildeO(\cdot)$ notation suppresses the constant and
+$\polylog(T)$ factors, respectively.}
+Exploration-separating algorithms can only achieve regret $\tildeO(T^{2/3})$ across all problem instances.
+Finally, the greedy algorithm is terrible on a wide variety of problem instances, in the sense that it fails to try the best arm and therefore suffers regret linear in $T$
+\citep[see Chapter 11.2 in][]{slivkins-MABbook}.
+
+The optimal regret rates are achieved by many adaptive-exploration  algorithms, of which the most known are 
+Thompson Sampling \citep{Thompson-1933,TS-survey-FTML18},%
+\footnote{While Thompson Sampling is the first and arguably best-known bandit algorithm, the corresponding regret bounds have been proved relatively recently \citep{Shipra-colt12,Kaufmann-alt12,Shipra-aistats13}.}
+UCB1 \citep{bandits-ucb1},
+and Successive Elimination \citep{EvenDar-icml06}. 
+Further work on refining these regret rates is not as relevant.
+
+Exploration-separating algorithms completely separate exploration and exploitation.
+The simplest approach, called \emph{Explore-First}, is to explore uniformly for a predetermined number of rounds, then choose one arm and ``exploit".
+A slightly more refined approach, called \emph{Epsilon-Greedy}, explores uniformly in each round $t$ with predetermined probability $\eps_t$, and exploits with the remaining probability.
+
+\ascomment{stopped here.}
+
+%A formal definition is that the distribution over arms in each round does not depend on rewards observed in the previous rounds.
+
+A recent line of work \citep{kannan2018smoothed,bastani2017exploiting,externalities-colt18}
+finds that the greedy algorithm performs well (in theory), in a model of linear contextual bandits under substantial assumptions on smoothing or heterogeneity of the context vectors.
+
+Switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training.
+
 %%% Local Variables:

### 2020-07-12 journal paper/content/related_work.tex
Minor edits.

@@
 A slightly more refined approach, called \emph{Epsilon-Greedy}, explores uniformly in each round $t$ with predetermined probability $\eps_t$, and exploits with the remaining probability.
+A formal definition is that each round is either selected exploration, in which case  the distribution over arms does not depend on data, or it is assigned to exploitation, in which case the data from this round is discarded.
+The $T^{2/3}$ lower bound on regret follows from a somewhat more general lower bound in \citet{MechMAB-ec09}.
 
 \ascomment{stopped here.}
-
-%A formal definition is that the distribution over arms in each round does not depend on rewards observed in the previous rounds.
 

### 2020-07-12 journal paper/content/related_work.tex
More edits.

@@
 
-\subsection{Background: the three classes of algorithms [IN PROGRESS, don't read]}
+\subsection{Background for non-specialists: the three classes of algorithms [IN PROGRESS]}
 \label{sec:related-classes}
@@
 
+\xhdr{Fundamentals.}
 We are concerned with the following problem.
@@
 Algorithm's goal is to maximize the total reward.
-A standard performance measure is \emph{regret}, defined as the difference in the total expected reward between the algorithm and the best fixed arm.%
+
+A standard performance measure is \emph{regret}, defined as the difference in the total expected reward between the algorithm and the best arm.%
 \footnote{In a formula, regret is
-    $\E\sbr{ T\cdot \max_{\text{arms $a$}}\mu_a -  \sum_{t\in[T]} r_t }$,
+    $T\cdot \max_{\text{arms $a$}}\mu_a 
+    -  \E\sbr{ \sum_{t\in[T]} r_t }$,
 where $\mu_a$ is the mean reward of arm $a$.}
-The primary concern is the asymptotic growth rate of regret as a function of $T$.
+Since regret is normalized by the best arm, it allows to compare algorithms across different problem instances. 
+The primary concern is the asymptotic growth rate of regret as a function of the time horizon $T$.
 
@@
 
+\xhdr{Advanced aspects.}
+Switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training \citep{DS-arxiv}.
+
 A recent line of work \citep{kannan2018smoothed,bastani2017exploiting,externalities-colt18}
@@
 
-Switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training.
+%\xhdr{}
 

### 2020-07-12 journal paper/content/sec-intro.tex
Added a sentence, reworded some comments.

@@
 Generally,  ``better" algorithms are better in the long run, but could be worse initially.
+
+\asedit{While this material builds on a vast and active research area, we present sufficient background in Section~\ref{sec:related-classes}, so as to keep this paper accessible to non-specialists.}
 
@@
 
-\ascomment{Guy: Previously you also cited \cite{varian2018artificial,lambrecht2015can} for the same thing as \citet{bajari2018impact}.
-Why not now?
-Also, previously you explicitly mentioned "barriers to entry" and "market tipping" and "anti-trust implications".
-Why not now?
-I don't really know any of this literature, so I defer to your judgement on this ...}
+\ascomment{Guy: citing \citet{varian2018artificial} might serve us, even if it is a more informal discussion. The existence of the policy discussion is worth mentioning, too (with citations).}
 
-\gacomment{I added back the \cite{lambrecht2015can} reference, but \cite{varian2018artificial} is a more informal discussion of these issues and not essential to include.
-However, even the \cite{lambrecht2015can} is relatively informal discussion of the same issues.
-Part of the problem that makes this tricky is that the literature on network effects is a well-developed and mature literature, but the question of whether data plays a similar role primarily started as more of a policy discussion, of which there's only starting to be some serious theoretical and empirical work on the issue.}
-
-\ascomment{We could, in principle, move both "Econ interpretation" xhdrs to a separate subsection.
-But the current order of presentation seems better to me: we can explain the inverted-U right after "theory", before we put a whole bunch of simulation results on the stack.
-And the anti-trust stuff goes nicely right after the numerical results, without a big subsection break.}
+\ascomment{"Econ interpretation" xhdrs fit incely in "our results", as opposed to a separate subsection.
+Because: we can explain the "inverted-U" stuff right after "theory", before we put the simulation results on the stack.
+And the "network effects" stuff goes nicely right after the numerical results, without a big subsection break.}
 

### 2020-07-12 journal paper/content/related_work.tex
More edits.

@@
 
-\ascomment{stopped here.}
-
 \xhdr{Advanced aspects.}
 Switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training \citep{DS-arxiv}.
+Inserting exploration in a complex decision pipeline necessitates a substantial awareness of the technology and a certain change in a mindset, as well as an infrastructure to collect and analyze the data.
+Adaptive exploration requires the said infrastructure to propagate the data analysis back to the ``front-end" where the decisions are made, and do it on a sufficiently fast and regular cadence.
+Framing the problem in terms of the appropriate modeling assumptions and action features, and debugging the machine learning code, tend to be quite subtle, too.
 
-A recent line of work \citep{kannan2018smoothed,bastani2017exploiting,externalities-colt18}
-finds that the greedy algorithm performs well (in theory), in a model of linear contextual bandits under substantial assumptions on smoothing or heterogeneity of the context vectors.
+The greedy algorithm can perform well \emph{sometimes} in a more general model of \emph{contextual bandits}, where auxiliary payoff-relevant signals, a.k.a.
+contexts, are observed before each round.
+This phenomenon has been observed in practice
+\citep{practicalCB-arxiv18}, and in theory \citep{kannan2018smoothed,bastani2017exploiting,externalities-colt18} under (very) substantial assumptions.
+The prevalent intuition is that the diversity of contexts can --- under some conditions and to a limited extent --- substitute for explicit exploration.
+
+\ascomment{stopped here.}
 

### 2020-07-13 journal paper/content/sec-intro.tex
updates

@@
 \xhdr{Economic interpretation: network effects of data.}
-Our model speaks to the ongoing debate on the role that data plays in the digital economy.
-One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
+Our model speaks to the ongoing debate on the role that data plays in the digital economy, \gaedit{which has been at the forefront of policy discussions on how to regulate major, data-intensive, digital platforms \citep{competition2019digital, scott2019committee}.} One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
 %whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
@@
 Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
-Approaches such as \citet{lambrecht2015can, bajari2018impact} argue that the diminishing returns to scale and scope of data in isolation mitigate such data feedback loops,
+Approaches such as \citet{lambrecht2015can, bajari2018impact, varian2018artificial} argue that the diminishing returns to scale and scope of data in isolation mitigate such data feedback loops,
 %as non-existent

### 2020-07-13 journal paper/ec19paper/content/perf_in_iso.tex
updates

@@
 \centering
-\includegraphics[scale=0.35]{ec19paper/figures/nih_iso_mean}
-\includegraphics[scale=0.35]{ec19paper/figures/uniform_mean}
+\includegraphics[scale=0.35]{ec19paper/figures/nih_mean}
+\includegraphics[scale=0.35]{ec19paper/figures/mean_inst_reward_nih}
 %\caption*{\tiny{Mean reputation trajectory The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs.
 The shaded area display 95\% confidence intervals.}}
-\caption{Mean reputation trajectories for Needle-in-Haystack and Uniform. The shaded area shows 95\% confidence intervals.}
+\caption{Mean reputation trajectory (left) and mean instantaneous reward trajectory (right) for Needle-in-Haystack. The shaded area shows 95\% confidence intervals.}
 \label{prelim_means}

### 2020-07-13 journal paper/content/sec-intro.tex
Reworded in "network effects".

@@
 
-\ascomment{The xhdr below is basically from Guy, with minor edits by Alex.}
+\ascomment{The xhdr below is basically from Guy, with some rewordings by Alex.}
 
 \xhdr{Economic interpretation: network effects of data.}
-Our model speaks to the ongoing debate on the role that data plays in the digital economy, \gaedit{which has been at the forefront of policy discussions on how to regulate major, data-intensive, digital platforms \citep{competition2019digital, scott2019committee}.} One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
+\asedit{Our model speaks to policy discussions on regulating data-intensive digital platforms \citep{competition2019digital, scott2019committee}, and particularly to the ongoing debate on the role of data in the digital economy.} One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
 %whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
@@
 Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.
-
-\ascomment{Guy: citing \citet{varian2018artificial} might serve us, even if it is a more informal discussion. The existence of the policy discussion is worth mentioning, too (with citations).}
 

### 2020-07-13 journal paper/content/related_work.tex
Moved "background" to Appendix A.

--- 2020-07-12 journal paper/content/related_work.tex
@@
 The literature on MAB is extremely vast and multi-threaded.
-The most relevant thread studies a basic model of MAB with IID rewards and no auxiliary structure, see Section~\ref{sec:related-classes}.
+The most relevant thread studies a basic model of MAB with IID rewards and no auxiliary structure, see Appendix~\ref{sec:related-classes}.
 The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.} This literature is covered in many books and surveys.
@@
 
-\ascomment{Guy: slightly reworded the previous para.}
-
 % moved this thought to the intro.
 \OMIT{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
-
-\subsection{Background: the three classes of algorithms [IN PROGRESS, don't read]}
-\label{sec:related-classes}
-
-We present sufficient background on the three classes of MAB algorithms listed in Section~\ref{sec:intro-model}: from adaptive-exploration to exploration-separating to greedy.
-
-We are concerned with the following problem.
-There are $T$ rounds and $K$ \emph{arms} to choose from.
-In each round $t\in [T]$, the algorithm chooses an arm and receives a reward $r_t\in[0,1]$ for this arm, drawn from a fixed but unknown distribution.
-Algorithm's goal is to maximize the total reward.
-A standard performance measure is \emph{regret}, defined as the difference in the total expected reward between the algorithm and the best fixed arm.%
-\footnote{In a formula, regret is
-    $\E\sbr{ T\cdot \max_{\text{arms $a$}}\mu_a -  \sum_{t\in[T]} r_t }$,
-where $\mu_a$ is the mean reward of arm $a$.}
-The primary concern is the asymptotic growth rate of regret as a function of $T$.
-
-The three classes of algorithms perform very differently in terms of regret. Adaptive-exploration algorithms achieve optimal regret rates:
-    $\tildeO(\sqrt{KT})$
-for all problem instances, and an improved regret rate of 
-    $O(\tfrac{K}{\Delta}\log T)$
-for problem instances with $\text{gap}\geq \Delta$
-\citep{Lai-Robbins-85,bandits-ucb1,bandits-exp3}.%
-\footnote{The \emph{gap} is the difference in mean reward between the best arm and the second-best arm.
-As usual in computer science, the $O(\cdot)$ and $\tildeO(\cdot)$ notation suppresses the constant and
-$\polylog(T)$ factors, respectively.}
-Exploration-separating algorithms can only achieve regret $\tildeO(T^{2/3})$ across all problem instances.
-Finally, the greedy algorithm is terrible on a wide variety of problem instances, in the sense that it fails to try the best arm and therefore suffers regret linear in $T$
-\citep[see Chapter 11.2 in][]{slivkins-MABbook}.
-
-The optimal regret rates are achieved by many adaptive-exploration  algorithms, of which the most known are 
-Thompson Sampling \citep{Thompson-1933,TS-survey-FTML18},%
-\footnote{While Thompson Sampling is the first and arguably best-known bandit algorithm, the corresponding regret bounds have been proved relatively recently \citep{Shipra-colt12,Kaufmann-alt12,Shipra-aistats13}.}
-UCB1 \citep{bandits-ucb1},
-and Successive Elimination \citep{EvenDar-icml06}. 
-Further work on refining these regret rates is not as relevant.
-
-Exploration-separating algorithms completely separate exploration and exploitation.
-The simplest approach, called \emph{Explore-First}, is to explore uniformly for a predetermined number of rounds, then choose one arm and ``exploit".
-A slightly more refined approach, called \emph{Epsilon-Greedy}, explores uniformly in each round $t$ with predetermined probability $\eps_t$, and exploits with the remaining probability.
-A formal definition is that each round is either selected exploration, in which case  the distribution over arms does not depend on data, or it is assigned to exploitation, in which case the data from this round is discarded.
-The $T^{2/3}$ lower bound on regret follows from a somewhat more general lower bound in \citet{MechMAB-ec09}.
-
-\ascomment{stopped here.}
-
-A recent line of work \citep{kannan2018smoothed,bastani2017exploiting,externalities-colt18}
-finds that the greedy algorithm performs well (in theory), in a model of linear contextual bandits under substantial assumptions on smoothing or heterogeneity of the context vectors.
-
-Switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training.
 

### 2020-07-13 journal paper/content/sec-bg.tex
Moved "background" to Appendix A.

We present sufficient background on the three classes of MAB algorithms listed in Section~\ref{sec:intro-model}: from adaptive-exploration to exploration-separating to greedy.

\xhdr{Fundamentals.}
We are concerned with the following problem. There are $T$ rounds and $K$ \emph{arms} to choose from. In each round $t\in [T]$, the algorithm chooses an arm and receives a reward $r_t\in[0,1]$ for this arm, drawn from a fixed but unknown distribution. Algorithm's goal is to maximize the total reward.

A standard performance measure is \emph{regret}, defined as the difference in the total expected reward between the algorithm and the best arm.%
\footnote{In a formula, regret is
    $T\cdot \max_{\text{arms $a$}}\mu_a
    -  \E\sbr{ \sum_{t\in[T]} r_t }$,
where $\mu_a$ is the mean reward of arm $a$.}
Since regret is normalized by the best arm, it allows to compare algorithms across different problem instances.
The primary concern is the asymptotic growth rate of regret as a function of the time horizon $T$.

The three classes of algorithms perform very differently in terms of regret. Adaptive-exploration algorithms achieve optimal regret rates:
    $\tildeO(\sqrt{KT})$
for all problem instances, and an improved regret rate of
    $O(\tfrac{K}{\Delta}\log T)$
for problem instances with $\text{gap}\geq \Delta$
\citep{Lai-Robbins-85,bandits-ucb1,bandits-exp3}.%
\footnote{The \emph{gap} is the difference in mean reward between the best arm and the second-best arm. As usual in computer science, the $O(\cdot)$ and $\tildeO(\cdot)$ notation suppresses the constant and
$\polylog(T)$ factors, respectively.}
Exploration-separating algorithms can only achieve regret $\tildeO(T^{2/3})$ across all problem instances. Finally, the greedy algorithm is terrible on a wide variety of problem instances, in the sense that it fails to try the best arm and therefore suffers regret linear in $T$
\citep[see Chapter 11.2 in][]{slivkins-MABbook}.

The optimal regret rates are achieved by many adaptive-exploration  algorithms, of which the most known are
Thompson Sampling \citep{Thompson-1933,TS-survey-FTML18},%
\footnote{While Thompson Sampling is the first and arguably best-known bandit algorithm, the corresponding regret bounds have been proved relatively recently \citep{Shipra-colt12,Kaufmann-alt12,Shipra-aistats13}.}
UCB1 \citep{bandits-ucb1},
and Successive Elimination \citep{EvenDar-icml06}.
Further work on refining these regret rates is not as relevant.

Exploration-separating algorithms completely separate exploration and exploitation. The simplest approach, called \emph{Explore-First}, is to explore uniformly for a predetermined number of rounds, then choose one arm and ``exploit". A slightly more refined approach, called \emph{Epsilon-Greedy}, explores uniformly in each round $t$ with predetermined probability $\eps_t$, and exploits with the remaining probability. A formal definition is that each round is either selected exploration, in which case  the distribution over arms does not depend on data, or it is assigned to exploitation, in which case the data from this round is discarded. The $T^{2/3}$ lower bound on regret follows from a somewhat more general lower bound in \citet{MechMAB-ec09}.

\xhdr{Advanced aspects.}
Switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training \citep{DS-arxiv}. Inserting exploration in a complex decision pipeline necessitates a substantial awareness of the technology and a certain change in a mindset, as well as an infrastructure to collect and analyze the data. Adaptive exploration requires the said infrastructure to propagate the data analysis back to the ``front-end" where the decisions are made, and do it on a sufficiently fast and regular cadence. Framing the problem in terms of the appropriate modeling assumptions and action features, and debugging the machine learning code, tend to be quite subtle, too.

The greedy algorithm can perform well \emph{sometimes} in a more general model of \emph{contextual bandits}, where auxiliary payoff-relevant signals, a.k.a. contexts, are observed before each round. This phenomenon has been observed in practice
\citep{practicalCB-arxiv18}, and in theory \citep{kannan2018smoothed,bastani2017exploiting,externalities-colt18} under (very) substantial assumptions. The prevalent intuition is that the diversity of contexts can --- under some conditions and to a limited extent --- substitute for explicit exploration.

\xhdr{Instantaneous regret.}

\ascomment{stopped here.}

%\xhdr{}

### 2020-07-13 journal paper/content/sec-intro.tex
Moved "background" to Appendix A.

@@
 
-\asedit{While this material builds on a vast and active research area, we present sufficient background in Section~\ref{sec:related-classes}, so as to keep this paper accessible to non-specialists.}
+\asedit{While this material builds on a vast and active research area, we present sufficient background in Appendix~\ref{sec:related-classes}, so as to keep this paper accessible to non-specialists.}
 

### 2020-07-13 journal paper/content/sec-theory.tex
Minor issue-fixes.

@@
 The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates.
-To make this more precise, let us say that two MAB algorithms \emph{deviate} at (local) step $n$ if
-\begin{align}
-\Pr\sbr{ \alg[1](n)\neq \alg[2](n)} >0,
-\end{align}
-where $\alg[i](n)$ is the arm chosen by \alg[i] at local step $n$.
 
-\ascomment{TODO: we used to have a much more cumbersome definition (commented out), and we need it back! because, what if we have two copies of the same randomized algorithm, this eqn will hold!}
+To make this more precise, let us define what it means for MAB algorithms to deviate from each other.
 
-\ascomment{TODO: define "deviate starting from" as just "this is the first $n_0$ when the deviation happens.}
-
-%there is an action $a\in A$ and a set of step-$n$ histories of positive probability such that any history $h$ in this set is feasible for both algorithms, and under this history the two algorithms choose action $a$ with different probability.
+\begin{definition}
+Two MAB algorithms \emph{deviate} at (local) step $n$ if there is a set $H$ of histories over the previous local steps such that both algorithms lead to $H$ with positive probability, and choose different distributions over arms given any history $h\in H$.
+The two algorithms deviate \emph{starting from} step $n_0$ if $n=n_0$ is the smallest step when they deviate.
+\end{definition}
 
@@
 
-\ascomment{This whole remark is supposed to add intuition, but it seems pretty hard to parse ... :(}
+\newcommand{\BIRgr}{\BIR^{\texttt{gr}}}
 
-\begin{remark}
-The modified algorithm performs exploration % ingests information
-at a slower rate.
-For the sake of intuition, let us argue how this may translate  into a larger \BIR compared to the original algorithm.
-Let  $\BIR'_1(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1].
-Then
+For the sake of intuition, let us give a concrete example how the modified algorithm \alg[2] may have a larger \BIR compared to the original algorithm \alg[1], even though it performs exploration at a slower rate.
+Let $\BIRgr_1(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1].
+Assume
+\begin{align}\label{eq:cor:random-greedy-ex-assumption}
+ \tfrac{1}{c}\; \BIRgr_1(n) \geq \BIR_1(n)= \beta\cdot n^{-\gamma}  
+ \qquad \text{for all local steps $n$},
+\end{align}
+for some absolute constants $\beta,\gamma>0$ and $c>1-\gamma$. We will derive
+\begin{align}\label{eq:cor:random-greedy-ex-conclusion}
+\BIR_2(n)
+    >\alpha\cdot \BIR_1(n),
+    \text{where} \quad
+    \alpha = \frac{1-p+pc}{1-p\gamma}>1.
+\end{align}
+
+Indeed, by definition of the modified algorithm we have that
 \begin{align}\label{eq:random-BIR}
 \BIR_2(n)
-    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIR'_1(m)}.
+    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIRgr_1(m)}.
 \end{align}
@@
     $\E[m] = n_0-1 + (n-n_0+1)(1-p) \geq (1-p)n$.
-
-Suppose $\BIR_1(n)= \beta n^{-\gamma}$ for some constants $\beta,\gamma>0$. Further, assume
-    $\BIR'_1(n)\geq  c\; \BIR_1(n)$,
-for some $c>1-\gamma$.
-\ascomment{Steven,Alex: not clear why this holds ...}
 
@@
 \BIR_2(n)
-    &>\alpha\cdot \BIR_1(n),
-    &\text{where} \quad
-    \alpha = \tfrac{1-p+pc}{1-p\gamma}>1.
+    &>\alpha\cdot \BIR_1(n).
 \end{align*}
 (In the above equations, all expectations are over $m$ distributed as in \eqref{eq:random-BIR}.)
-\end{remark}
 

### 2020-07-13 journal paper/content/sec-intro.tex
cite furman report instead of ACCC report

@@
 \xhdr{Economic interpretation: network effects of data.}
-\asedit{Our model speaks to policy discussions on regulating data-intensive digital platforms \citep{competition2019digital, scott2019committee}, and particularly to the ongoing debate on the role of data in the digital economy.} One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
+\asedit{Our model speaks to policy discussions on regulating data-intensive digital platforms \citep{furman2019unlocking, scott2019committee}, and particularly to the ongoing debate on the role of data in the digital economy.} One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
 %whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.

### 2020-07-13 journal paper/journal_main.tex
updates to intro + numerical sections
Inserted some \newpage commands.
Removed page number from ToC.
Added \newpage in the appendix.
Removed "first version" from \date.
Inserted a \newpage.
Moved "background" to Appendix A.
Revised some appendix titles.
Changed ToC to small font.
correcting ack

@@
 \footnote{University of Minnesota - Twin Cities, Department of Computer Science. Email: zsw@umn.edu\newline
-The research has been done when Z.S. Wu has been an intern and a postdoc at Microsoft Research NYC.
+Part of the research was done when Z.S. Wu was an intern and a postdoc at Microsoft Research NYC.
 }}
-\date{First version: July 2020}
+\date{July 2020}
 \maketitle
@@
 \end{abstract}
-%\setcounter{page}{0}
-%\thispagestyle{empty}
+\setcounter{page}{0}
+\thispagestyle{empty}
 %\end{titlepage}
@@
 \end{small}
+\setcounter{page}{0}
+\thispagestyle{empty}
 \newpage
@@
 \label{sec:intro}
+
 \subfile{content/sec-intro}
@@
 
+\newpage
 \section{Theoretical results: the \TheoryModel}
@@
 
+\newpage
 \gaedit{
@@
 
-\clearpage
+\newpage
 \addcontentsline{toc}{section}{References}
+
 \bibliographystyle{plainnat}
+\begin{small}
 \bibliography{bib-abbrv,bib-ML,refs,bib-bandits,bib-AGT,bib-slivkins, bib-random}
+\end{small}
 
@@
 
-\section{Background on multi-armed bandits}
+\section{Background for non-specialists: the three classes of algorithms [IN PROGRESS]}
+\label{sec:related-classes}
+\subfile{content/sec-bg}
+
+\section{Auxiliary theoretical results}
 \label{app:examples}
@@
 
-\section{Proofs for Section~\ref{sec:theory}}
+\newpage
+\section{Full proofs for Section~\ref{sec:theory}}
 \label{sec:theory-proofs}
 \subfile{content/sec-theory-proofs}
+
+\newpage
+\section{Full experimental results}
+\label{app:expts}
 

### 2020-07-13 journal paper/content/sec-theory.tex
Another pass on theory.

@@
 The competition game has a unique Nash equilibirium: both principals choose \DynGreedy.
+\asedit{This holds even if the principals can only choose from restricted strategy sets which include \DynGreedy.}
 \end{corollary}
@@
 
-Consider the \HardMaxRandom response model, where each principal is chosen with some positive baseline probability.
-The main technical result for this model states that a principal with asymptotically better \BIR wins by a large margin: after a ``learning phase" of constant duration, all agents choose this principal with maximal possible probability $\respF(1)$.
-For example, a principal with $\BIR(n)\leq \tilde{O}(n^{-1/2})$ wins over a principal with $\BIR(n)\geq \Omega(n^{-1/3})$.
+Consider the \HardMaxRandom response model, where each principal is chosen with some positive baseline probability. Informally, we find that:
+\begin{align}\label{eq:theory-HMR-informal}
+\text{A much better algorithm wins big.}
+\end{align}
+In more detail, a principal with asymptotically better \BIR wins by a large margin: after a ``learning phase" of constant duration, all agents choose this principal with maximal possible probability $\respF(1)$.
+For example, a principal with $\BIR(n)\leq \tilde{O}(n^{-1/2})$ prevails over a principal with $\BIR(n)\geq \Omega(n^{-1/3})$.
 
@@
 \end{align}
-If this holds, we say that \alg[1] \emph{BIR-dominates} \alg[2].
+If this holds, we say that \alg[1] \emph{BIR-dominates} \alg[2] starting from (local) step $n_0$.
+
+We also need a mild technical assumption that $\BIR_2(\cdot)$ is not extremely small: for some constant $m_0$,
+\begin{align}\label{eq:random-assn}
+ (\forall n\geq m_0) \qquad
+  \BIR_2(n) > 4\,e^{-\eps_0\ n/12}.
+\end{align}
+
+Thus, the main result is stated as follows:
 
@@
 Assume \HardMaxRandom response function.
-Suppose algorithms \alg[1], \alg[2] are monotone and $T$-oblivious.
-Assume that \alg[1] \emph{BIR-dominates} \alg[2], and the latter satisfies a mild technical assumption:
-\begin{align}\label{eq:random-assn}
- (\forall n\geq n_0) \qquad
-  \BIR_2(n) > 4\,e^{-\eps_0\ n/12}.
-\end{align}
-Then each agent $t\geq n_0$ chooses principal $1$ with maximal possible probability $\respF(1) = 1- \eps_0$.
+Suppose algorithms \alg[1], \alg[2] are monotone and $T$-oblivious, and \eqref{eq:random-assn} holds.
+If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible).
 \end{theorem}
@@
 
-To state the game-theoretic implications of Theorem~\ref{thm:random-clean}, we restrict the set of feasible MAB algorithms:
+\asedit{We'd like to use Theorem~\ref{thm:random-clean} to conclude that a (much) better algorithm prevails in equilibrium.
+Indeed, this holds if the principals can only choose from a finite set of MAB algorithms.}
 
 \begin{definition}\label{def:restricted-competition}
-A \emph{restricted competition game} is a version of the competition game between the two principals in which they can only choose from a finite set of MAB algorithms, which are called \emph{feasible}.
+A \emph{\FiniteGame} is a version of the competition game between the two principals in which they can only choose from a finite set of MAB algorithms, which are called \emph{feasible}.
 All feasible algorithms are assumed to be
@@
 Assume \HardMaxRandom response function.
-Consider the restricted competition game with a feasible algorithm \alg which BIR-dominates all other feasible algorithms.
+Consider the \FiniteGame with a feasible algorithm \alg which BIR-dominates all other feasible algorithms.
 Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium: both principals choose \alg.
@@
 
-\xhdr{A little greedy goes a long way.}
-Given any monotone MAB algorithm other than \DynGreedy, we design a modified algorithm which learns at a slower rate, yet ``wins the game" in the sense of Theorem~\ref{thm:random-clean}.
-So, the competition game with unrestricted choice of algorithms typically does not have a Nash equilibrium.
+\xhdr{Counterpoint: A little greedy goes a long way.}
+Given any monotone MAB algorithm \alg other than \DynGreedy, we design a modified algorithm which ``mixes in" some greedy choices (and consequently learns at a slower rate), yet prevails over \alg in the competition game.
+Thus, we have a counterpoint to \eqref{eq:theory-HMR-informal}: even under \HardMaxRandom, a better algorithm may lose in competition.
+The corresponding counterpoint to Corollary~\ref{cor:random} states that non-greedy algorithms are \emph{not} chosen in equilibrium.
 
-Given an algorithm \alg[1] that deviates from \DynGreedy starting from
-step $n_0$ and a ``mixing'' parameter $p$, we will construct a
-modified algorithm by ``mixing in" some greedy choices:
+The modified algorithm, called the \emph{greedy modification} of \alg  with \emph{mixing parameter} $p\in (0,1)$, is defined as follows.
+Suppose \alg deviates from \DynGreedy starting from some (local) step $n_0$.
+Then
 \begin{enumerate}
-\item The modified algorithm coincides with \alg[1] (and \DynGreedy)
+\item The modified algorithm coincides with \DynGreedy
 for the first $n_0-1$ steps;
-\item In each step $n\geq n_0$, \alg[1] is invoked with probability
+\item In each step $n\geq n_0$, \alg is invoked with probability
   $1-p$, and with the remaining probability $p$ does the ``greedy
   choice": chooses an action with the largest posterior mean reward
-  given the current information collected by \alg[1].
+  given the current information collected by \alg.
 
@@
 
-Note that the modified algorithm is another pure strategy in the competition game between the two principals, not a mixture of \alg[1] and the greedy algorithm.
+\noindent Note that the modified algorithm is another pure strategy in the competition game between the two principals, not a mixed strategy that randomizes between \alg and the greedy algorithm.
+
+We find that the greedy modification prevails in competition if the mixing parameter is small enough.
 
 \begin{theorem}\label{thm:random-greedy}
-Assume symmetric \HardMaxRandom response function.
-Let $\eps_0 = \respF(-1)$ be the baseline probability.
-Suppose \alg[1] deviates from \DynGreedy starting from some step $n_0$.
-Let \alg[2] be the modified algorithm, as described above, with mixing parameter $p$ such that
+Consider a symmetric \HardMaxRandom response function $\respF$ with baseline probability $\eps_0 = \respF(-1)$.
+Suppose \alg[1] is monotone, and deviates from \DynGreedy starting from some step $n_0$.
+Let \alg[2] be the greedy modification of \alg[1] with mixing parameter $p>0$ such that
     $(1-\eps_0)(1-p)>\eps_0$.
-Then each agent $t\geq n_0$ chooses principal $2$ with maximal possible probability $1-\eps_0$.
+Then each agent $t\geq n_0$ chooses principal $2$ with probability $1-\eps_0$ (which is the largest possible).
 \end{theorem}
 
+%It is easy to see that \alg[2] is monotone. Indeed,
+
 \begin{corollary}\label{cor:random-greedy}
-  Suppose both principals can choose any monotone MAB algorithm, and assume the symmetric \HardMaxRandom response
-  function. Then for any time
-  horizon $T$, the only possible \emph{pure} Nash equilibrium is one
-  where both principals choose \DynGreedy. Moreover, no pure Nash
-  equilibrium exists when some algorithm BIR-dominates \DynGreedy and the time horizon $T$ is sufficiently large.
+Posit a symmetric \HardMaxRandom response function.
+Suppose both principals can only choose from monotone MAB algorithms.
+Then for any time horizon $T$, the only possible \emph{pure} Nash equilibrium is one where both principals choose \DynGreedy.
+Moreover, no pure Nash equilibrium exists when some algorithm BIR-dominates \DynGreedy and $T$ is sufficiently large.
+\asedit{These conclusions hold even if the principals' strategy sets are any collections of monotone MAB algorithms that are closed under greedy mixing.}%
+\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) monotone MAB algorithms, and contains all greedy mixtures of all algorithms in this set.}
 \end{corollary}
@@
 
-For the sake of intuition, let us give a concrete example how the modified algorithm \alg[2] may have a larger \BIR compared to the original algorithm \alg[1], even though it performs exploration at a slower rate.
-Let $\BIRgr_1(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1].
-Assume
+\ascomment{Steven: I revised this example to clarify what it is supposed to mean.
+But the conclusion can't be right: "greedy modification" can only improve the BIR at step $n_0$!
+Perhaps what we derive is that
+$\BIR_2(n)>\BIR_1(n)$ starting from $n=n_0+1$? }
+
+For the sake of intuition, let us show how \alg[1] can be a better algorithm than its greedy modification, in the sense of having a strictly smaller \BIR.
+
+\begin{example}
+Consider \alg[1] and \alg[2] as in Theorem~\ref{thm:random-greedy}. 
+Let $\BIRgr_1(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1]. Assume, for some absolute constants $\beta,\gamma>0$ and $c>1-\gamma$, that
 \begin{align}\label{eq:cor:random-greedy-ex-assumption}
  \tfrac{1}{c}\; \BIRgr_1(n) \geq \BIR_1(n)= \beta\cdot n^{-\gamma}
- \qquad \text{for all local steps $n$},
+ \qquad \text{for all local steps $n$}.
 \end{align}
-for some absolute constants $\beta,\gamma>0$ and $c>1-\gamma$. We will derive
-\begin{align}\label{eq:cor:random-greedy-ex-conclusion}
-\BIR_2(n)
-    >\alpha\cdot \BIR_1(n),
-    \text{where} \quad
-    \alpha = \frac{1-p+pc}{1-p\gamma}>1.
-\end{align}
+Then, for a sufficiently small $p>0$, we have $\BIR_2(n)>\BIR_1(n)$ for all local steps $n\geq n_0$.
+\end{example}
 
-Indeed, by definition of the modified algorithm we have that
+\begin{proof}
+By definition of the modified algorithm we have that
 \begin{align}\label{eq:random-BIR}
@@
 \BIR_2(n)
-    &>\alpha\cdot \BIR_1(n).
+    &>\alpha\cdot \BIR_1(n),
+    &\text{where} \quad
+    \alpha = \frac{1-p(1-c)}{1-p\gamma}>1.
 \end{align*}
 (In the above equations, all expectations are over $m$ distributed as in \eqref{eq:random-BIR}.)
+\end{proof}
 
@@
 %
-Consider the restricted competition game with special algorithm \alg, and assume that all other allowed algorithms satisfy $\BReg(n)\to \infty$.
+Consider the \FiniteGame with special algorithm \alg, and assume that all other allowed algorithms satisfy $\BReg(n)\to \infty$.
 Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium: both principals choose \alg.
@@
 
-As in the previous subsection, we can state a game-theoretic corollary for the restricted competition game (as in Definition~\ref{def:restricted-competition}).
+As in the previous subsection, we can state a game-theoretic corollary for the \FiniteGame (as in Definition~\ref{def:restricted-competition}).
 We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity.
@@
 \begin{corollary}\label{cor:SoftMax-strong}
-Assume \SoftMaxRandom response function. Consider the restricted competition game in which all feasible algorithms satisfy $\BReg(n)\to \infty$.%
+Assume \SoftMaxRandom response function. Consider the \FiniteGame in which all feasible algorithms satisfy $\BReg(n)\to \infty$.%
 \footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition.
@@
 \xhdr{Main story.}
-Our main story concerns the restricted competition game between the two principals where one allowed algorithm \alg is ``better" than the others.
+Our main story concerns the \FiniteGame between the two principals where one allowed algorithm \alg is ``better" than the others.
 {We track whether and when \alg is chosen in an equilibrium.} We vary \competitiveness/\rationality by changing the response function from \HardMax (full rationality, very competitive environment) to \HardMaxRandom to  \SoftMaxRandom (less rationality and competition).

### 2020-07-13 journal paper/content/sec-theory.tex
Added "monotonicity of alg2" proof.

@@
 
-\xhdr{Notation.} Let $\rew_i(n)$ denote the Bayesian-expected reward of principal $i$ at local step $n$.
+\xhdr{Notation.} Let $\rew_i(n)$ denote the realized reward of principal $i$ at local step $n$.
 For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen.
@@
 We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
-\begin{align}
-\BIR(n) := \E_{\mu\sim\priorMu}
-    \left[ \max_{a\in A} \mu_a\right] - \rew(n),
+\begin{align}\label{eq:theory-BIR-defn}
+\BIR_i(n) := \E\sbr{ n\cdot \max_{a\in A} \mu_a - \rew_i(n)}.
 \end{align}
-where $\rew(n)$ is the Bayesian-expected reward of the algorithm for the $n$-th step, when the algorithm is run in isolation.
-We are primarily interested in how \BIR scales with $n$: typically, how fast \BIR decreases with $n$.
-We treat $K$, the number of arms, as a constant unless specified otherwise.
-Let $\BIR_i(n) = \BIR(n)$ for the algorithm of principal $i$.
+We are primarily interested in how \BIR scales with $n$: typically, how fast \BIR decreases with $n$. We treat $K$, the number of arms, as a constant unless specified otherwise. 
+%Let $\BIR_i(n) = \BIR(n)$ for the algorithm of principal $i$.
 
@@
 
-%It is easy to see that \alg[2] is monotone. Indeed,
+\newcommand{\BIRgr}{\BIR^{\texttt{gr}}}
+
+Next, let us argue that \alg[2] is monotone.
+We spell out the proof for the sake of intuition.
+For each $n\in\N$, let $M_n$ be the number of times \alg[1] is invoked in the first $n$ steps of \alg[2], and let $\BIRgr_1(n)$ be the $\BIR$ of the ``greedy choice" after $n-1$ steps of \alg[1].
+Let
+    $\mathtt{reg}_2(n) = n\cdot \max_a \mu_a - \rew_2(n)$ 
+be the (frequentist) instantaneous regret of \alg[2]. Then by definition of the greedy modification it holds that 
+\begin{align}\label{eq:random-BIR}
+\E\sbr{ \mathtt{reg}_2(n) \mid M_n = m}
+    &= (1-p) \cdot \BIR_1(m) + p \cdot \BIRgr_1(m) .\\
+\BIR_2(n)
+    &= \E\sbr{ (1-p) \cdot \BIR_1(M_n) + p \cdot \BIRgr_1(M_n) }.
+\end{align}
+Now, monotonicity of \alg[2] follows from monotonicity of $\BIR_1(\cdot)$ and $\BIRgr_1(\cdot)$.
 
@@
 \end{corollary}
-
-\newcommand{\BIRgr}{\BIR^{\texttt{gr}}}
 
@@
 \BReg(n) := n\cdot \E_{\mu\sim\priorMu}
-    \left[ \max_{a\in A} \mu_a\right] - \sum_{n=1}^n \rew(n')
-    = \sum_{n'=1}^n \BIR(n').
+    \left[ \max_{a\in A} \mu_a\right] - 
+    \E\sbr{ \sum_{m=1}^n \rew(m)}
+    = \sum_{m=1}^n \BIR(m).
 \end{align}

### 2020-07-13 journal paper/content/sec-theory.tex
A few more sentences in HMR.

@@
 \end{align}
-Now, monotonicity of \alg[2] follows from monotonicity of $\BIR_1(\cdot)$ and $\BIRgr_1(\cdot)$.
+Now, monotonicity of $\BIR_2$ follows from monotonicity of $\BIR_1(\cdot)$ and $\BIRgr_1(\cdot)$.
+The former is assumed, and the latter is the monotonicity-in-information property of the ``greedy step" (see \ascomment{pointer} for a precise formulation and proof).
+
+Thus, the greedy modification is beneficial in competition \emph{and} keeps us inside the set of monotone MAB algorithms.
+Consequently, if the principals are restricted to choosing monotone MAB algorithms, then non-greedy algorithms cannot be chosen in equilibrium.
 

### 2020-07-14 journal paper/content/sec-intro.tex
small fixes

@@
 We identify the inverted-U relationship driven by the rationality/competitiveness distinctions outlined above: from \HardMax to \HardMaxRandom to \SoftMaxRandom to \Uniform.
-We also find another, technically different inverted-U relationship which zeroes in on the \HardMaxRandom model.
+We also find another, technically different, inverted-U relationship which zeroes in on the \HardMaxRandom model.
 We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
@@
 In essence, we look for substantial effects within relevant time scales.
-(In fact, we start our investigation by determining what time scales are relevant in the context of our model.)
+In fact, we start our investigation by determining what time scales are relevant in the context of our model.
 

### 2020-07-14 journal paper/content/sec-intro.tex
Removed some color-editing.

@@
 
-\asedit{While this material builds on a vast and active research area, we present sufficient background in Appendix~\ref{sec:related-classes}, so as to keep this paper accessible to non-specialists.}
+While we build on a vast and active research area, we present sufficient background in Appendix~\ref{sec:related-classes} so as to keep this paper accessible to non-specialists.
 
@@
 The uniform choice among principals corresponds to no rationality and no competition.
-\asedit{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
+While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.
 
@@
 \xhdr{Economic interpretation: network effects of data.}
-\asedit{Our model speaks to policy discussions on regulating data-intensive digital platforms \citep{furman2019unlocking, scott2019committee}, and particularly to the ongoing debate on the role of data in the digital economy.} One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", \asedit{creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.}
+Our model speaks to policy discussions on regulating data-intensive digital platforms \citep{furman2019unlocking, scott2019committee}, and particularly to the ongoing debate on the role of data in the digital economy.
+One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.
 %whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
-\asedit{The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved out of users (resp., take over most of the market).}
+The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved out of users (resp., take over most of the market).
 %\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}

### 2020-07-14 journal paper/content/sec-theory.tex
Revised HMR subsection.

@@
 
-\newcommand{\BIRgr}{\BIR^{\texttt{gr}}}
+\newcommand{\gr}{\texttt{gr}}
+\newcommand{\BIRgr}{\BIR^{\gr}}
+\newcommand{\alggr}{\alg[\gr]}
 
-Next, let us argue that \alg[2] is monotone.
-We spell out the proof for the sake of intuition.
-For each $n\in\N$, let $M_n$ be the number of times \alg[1] is invoked in the first $n$ steps of \alg[2], and let $\BIRgr_1(n)$ be the $\BIR$ of the ``greedy choice" after $n-1$ steps of \alg[1].
+Let us analyze the \BIR of the greedy modification (we will use it for what follows).
+Use \alg[1] and \alg[2] as in the theorem.
+For each $n\in\N$, let $M_n$ be the number of times \alg[1] is invoked in the first $n$ steps of \alg[2].
+Let \alggr be a hypothetical algorithm which at each step $n$ outputs the ``Bayesian-greedy choice" based on the data collected by \alg[1] in the first $n-1$ steps.
+Let $\BIRgr(n)$ be the \BIR of this algorithm.
 Let
     $\mathtt{reg}_2(n) = n\cdot \max_a \mu_a - \rew_2(n)$
-be the (frequentist) instantaneous regret of \alg[2]. Then by definition of the greedy modification it holds that 
-\begin{align}\label{eq:random-BIR}
+be the (frequentist) instantaneous regret of \alg[2]. Then 
+\begin{align}
 \E\sbr{ \mathtt{reg}_2(n) \mid M_n = m}
-    &= (1-p) \cdot \BIR_1(m) + p \cdot \BIRgr_1(m) .\\
+    &= (1-p) \cdot \BIR_1(m) + p \cdot \BIRgr(m). \nonumber \\
 \BIR_2(n)
-    &= \E\sbr{ (1-p) \cdot \BIR_1(M_n) + p \cdot \BIRgr_1(M_n) }.
+    &= \E\sbr{ (1-p) \cdot \BIR_1(M_n) + p \cdot \BIRgr(M_n) }.
+    \label{eq:BIR-modification}
 \end{align}
-Now, monotonicity of $\BIR_2$ follows from monotonicity of $\BIR_1(\cdot)$ and $\BIRgr_1(\cdot)$.
-The former is assumed, and the latter is the monotonicity-in-information property of the ``greedy step" (see \ascomment{pointer} for a precise formulation and proof).
+
+\ascomment{Steven: do you buy this example? }
+
+Let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2].
+We need to assume that \alg[1] is (much) better than the greedy step \alggr if the latter is computed on a fraction of the data.
+Formally, we posit a convex function $f(\cdot)$ and $m_0>0$ such that
+\[ \BIRgr(n) \geq f(n) 
+    \quad\text{and}\quad 
+    f\rbr{\E[M_n]} \geq 2\, \BIR_1(n) \quad \forall n\geq m_0 .  \]
+Then by \eqref{eq:BIR-modification} and Jensen's inequality, for each step $n\geq m_0$ we have
+\[\BIR_2(n) 
+    \geq  \E\sbr{ \BIRgr(M_n) }
+    \geq \E\sbr{ f(M_n) } 
+    \geq f\rbr{ \E[M_n] }
+    \geq 2\,\BIR_1(n). \]
+
+Second, let us argue that \alg[2] is monotone.
+This follows from \eqref{eq:BIR-modification} since both \alg[1] and \alggr are monotone.
+The latter follows from a ``monotonicity-in-information" property of the ``greedy step": essentially, it can only get better with more information
+(see \ascomment{add a pointer}).
 
@@
 Moreover, no pure Nash equilibrium exists when some algorithm BIR-dominates \DynGreedy and $T$ is sufficiently large.
-\asedit{These conclusions hold even if the principals' strategy sets are any collections of monotone MAB algorithms that are closed under greedy mixing.}%
-\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) monotone MAB algorithms, and contains all greedy mixtures of all algorithms in this set.}
+\asedit{These conclusions hold even if each principal's strategy set is a subset of monotone MAB algorithms that is closed under the greedy modification.}%
+\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) monotone MAB algorithms, and contains the greedy modification for each algorithms in this set and each mixing parameter $p$.}
 \end{corollary}
 
-\ascomment{Steven: I revised this example to clarify what it is supposed to mean.
-But the conclusion can't be right: "greedy modification" can only improve the BIR at step $n_0$!
-Perhaps what we derive is that
+\ascomment{Steven: OLD EXAMPLE. I revised it to clarify what it is supposed to mean. But my new example is better, isn't it?}
+
+\ascomment{The conclusion can't be right: "greedy modification" can only improve the BIR at step $n_0$! Perhaps what we derive is that
 $\BIR_2(n)>\BIR_1(n)$ starting from $n=n_0+1$? }
-
-For the sake of intuition, let us show how \alg[1] can be a better algorithm than its greedy modification, in the sense of having a strictly smaller \BIR.
 
@@
 Consider \alg[1] and \alg[2] as in Theorem~\ref{thm:random-greedy}.
-Let $\BIRgr_1(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1]. Assume, for some absolute constants $\beta,\gamma>0$ and $c>1-\gamma$, that
+Let $\BIRgr(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1]. Assume, for some absolute constants $\beta,\gamma>0$ and $c>1-\gamma$, that
 \begin{align}\label{eq:cor:random-greedy-ex-assumption}
- \tfrac{1}{c}\; \BIRgr_1(n) \geq \BIR_1(n)= \beta\cdot n^{-\gamma}
+ \tfrac{1}{c}\; \BIRgr(n) \geq \BIR_1(n)= \beta\cdot n^{-\gamma}
  \qquad \text{for all local steps $n$}.
@@
 \BIR_2(n)
-    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIRgr_1(m)}.
+    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIRgr(m)}.
 \end{align}

### 2020-07-14 journal paper/content/sec-theory.tex
Added a comment.
Minor edits in SoftMax.

@@
 
+\ascomment{This whole blurb before the corollary is new.}
+
 Let us analyze the \BIR of the greedy modification (we will use it for what follows).
@@
 
-This subsection is devoted to the \SoftMaxRandom model. We derive a ``better algorithm wins" result under a much weaker version of BIR-dominance. This is the most technical result of the paper.
-
-We start with a formal definition of \SoftMaxRandom:
+For the \SoftMaxRandom model, we derive a ``better algorithm wins" result under a much weaker version of BIR-dominance. We start with a formal definition of \SoftMaxRandom:
 
@@
 For the sake of intuition, let us derive a version of Theorem~\ref{thm:random-clean}, with the same assumptions about the algorithms and essentially the same proof.
-The conclusion is much weaker: we can only guarantee that each agent $t\geq n_0$ chooses principal 1 with probability slightly larger than $\tfrac12$.
+The conclusion is much weaker, though: we can only guarantee that each agent $t\geq n_0$ chooses principal 1 with probability slightly larger than $\tfrac12$.
 This is essentially unavoidable in a typical case when both algorithms satisfy $\BIR(n)\to 0$, by Definition~\ref{def:SoftMax}.

### 2020-07-14 journal paper/content/sec-bg.tex
Revised some appendix titles.
Copied stuff from body.

@@
 
+Let us connect \BIR with the three classes of algorithms listed in Section~\ref{sec:intro-model}.
+\DynGreedy has at least a constant \BIR for many reasonable priors (where the constant can depend on $K$ and the prior, but not on $n$).
+\StaticGreedy typically has at least a constant \BIR, in the same sense.
+Exploration-separating algorithms can have
+    $\BIR(n) \sim n^{-1/3}$
+for all priors, whereas adaptive-exploration algorithms can have an even better rate
+    $\BIR(n) \sim n^{-1/3}$
+for all priors. Both rates are up to $\log(n)$ factors, and are the best possible \BIR rates that can be achieved for all priors. More details on this can be found in Appendix~\ref{app:examples}.
+
 \ascomment{stopped here.}

### 2020-07-14 journal paper/content/sec-theory.tex
Minor edits throughout, almost done.

@@
 \end{align}
-We are primarily interested in how \BIR scales with $n$: typically, how fast \BIR decreases with $n$. We treat $K$, the number of arms, as a constant unless specified otherwise.
+We are primarily interested in how \BIR scales with $n$: typically, how fast \BIR decreases with $n$.
+We treat $K$, the number of arms, as a constant unless specified otherwise.
+Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:examples} for more background.
+
 %Let $\BIR_i(n) = \BIR(n)$ for the algorithm of principal $i$.
-
-Let us connect \BIR with the three classes of algorithms listed in Section~\ref{sec:intro-model}.
-\DynGreedy has at least a constant \BIR for many reasonable priors (where the constant can depend on $K$ and the prior, but not on $n$).
-\StaticGreedy typically has at least a constant \BIR, in the same sense.
-Exploration-separating algorithms can have
-    $\BIR(n) \sim n^{-1/3}$
-for all priors, whereas adaptive-exploration algorithms can have an even better rate
-    $\BIR(n) \sim n^{-1/3}$
-for all priors. Both rates are up to $\log(n)$ factors, and are the best possible \BIR rates that can be achieved for all priors. More details on this can be found in Appendix~\ref{app:examples}.
 
@@
 
-\ascomment{Steven: do you buy this example? }
-
 Let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2].
@@
 
-\ascomment{Steven: OLD EXAMPLE. I revised it to clarify what it is supposed to mean. But my new example is better, isn't it?}
-
-\ascomment{The conclusion can't be right: "greedy modification" can only improve the BIR at step $n_0$! Perhaps what we derive is that
-$\BIR_2(n)>\BIR_1(n)$ starting from $n=n_0+1$? }
-
-\begin{example}
-Consider \alg[1] and \alg[2] as in Theorem~\ref{thm:random-greedy}.
-Let $\BIRgr(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1]. Assume, for some absolute constants $\beta,\gamma>0$ and $c>1-\gamma$, that
-\begin{align}\label{eq:cor:random-greedy-ex-assumption}
- \tfrac{1}{c}\; \BIRgr(n) \geq \BIR_1(n)= \beta\cdot n^{-\gamma}
- \qquad \text{for all local steps $n$}.
-\end{align}
-Then, for a sufficiently small $p>0$, we have $\BIR_2(n)>\BIR_1(n)$ for all local steps $n\geq n_0$.
-\end{example}
-
-\begin{proof}
-By definition of the modified algorithm we have that
-\begin{align}\label{eq:random-BIR}
-\BIR_2(n)
-    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIRgr(m)}.
-\end{align}
-In this expression, $m$ is the number of times \alg[1] is invoked in the first $n$ steps of the modified algorithm. Note that
-    $\E[m] = n_0-1 + (n-n_0+1)(1-p) \geq (1-p)n$.
-
-Then for all $n\geq n_0$ and small enough $p>0$ it holds that:
-\begin{align*}
- \BIR_2(n)
-    &\geq  (1-p+pc)\; \E[\; \BIR_1(m) \;] \\
-\E[\; \BIR_1(m) \;]
-    &\geq \BIR_1(\; \E[m] \;) &\qquad\text{(By Jensen's inequality)} \\
-    &\geq \BIR_1(\; (1-p)n \;) &\qquad\text{(since $\E[m]\geq n(1-p)$)}  \\
-    &\geq \beta\cdot n^{-\gamma} \cdot (1-p)^{-\gamma}
-        &\qquad\text{(plugging in $\BIR_1(n)=\beta n^{-\gamma}$)}  \\
-    &> \BIR_1(n)\;\; (1-p\gamma)^{-1}
-        &\qquad\text{(since $(1-p)^\gamma < 1-p\gamma$)}.\\
-\BIR_2(n)
-    &>\alpha\cdot \BIR_1(n),
-    &\text{where} \quad
-    \alpha = \frac{1-p(1-c)}{1-p\gamma}>1.
-\end{align*}
-(In the above equations, all expectations are over $m$ distributed as in \eqref{eq:random-BIR}.)
-\end{proof}
+%\ascomment{Steven: OLD EXAMPLE. I revised it to clarify what it is supposed to mean. But my new example is better, isn't it?}
+%
+%\ascomment{The conclusion can't be right: "greedy modification" can only improve the BIR at step $n_0$! Perhaps what we derive is that
+%$\BIR_2(n)>\BIR_1(n)$ starting from $n=n_0+1$? }
+%
+%
+%\begin{example}
+%Consider \alg[1] and \alg[2] as in Theorem~\ref{thm:random-greedy}.
+%Let $\BIRgr(n)$ be the \BIR of the ``greedy choice" after $n-1$ steps of \alg[1]. Assume, for some absolute constants $\beta,\gamma>0$ and $c>1-\gamma$, that
+%\begin{align}\label{eq:cor:random-greedy-ex-assumption}
+% \tfrac{1}{c}\; \BIRgr(n) \geq \BIR_1(n)= \beta\cdot n^{-\gamma}
+% \qquad \text{for all local steps $n$}.
+%\end{align}
+%Then, for a sufficiently small $p>0$, we have $\BIR_2(n)>\BIR_1(n)$ for all local steps $n\geq n_0$.
+%\end{example}
+%
+%\begin{proof}
+%By definition of the modified algorithm we have that
+%\begin{align}\label{eq:random-BIR}
+%\BIR_2(n)
+%    &= \Ex{m\sim (n_0-1)+\term{Binomial}(n-n_0+1,1-p)}{(1-p) \cdot \BIR_1(m) + p \cdot \BIRgr(m)}.
+%\end{align}
+%In this expression, $m$ is the number of times \alg[1] is invoked in the first $n$ steps of the modified algorithm. Note that
+%    $\E[m] = n_0-1 + (n-n_0+1)(1-p) \geq (1-p)n$.
+%
+%
+%Then for all $n\geq n_0$ and small enough $p>0$ it holds that:
+%\begin{align*}
+% \BIR_2(n)
+%    &\geq  (1-p+pc)\; \E[\; \BIR_1(m) \;] \\
+%\E[\; \BIR_1(m) \;]
+%    &\geq \BIR_1(\; \E[m] \;) &\qquad\text{(By Jensen's inequality)} \\
+%    &\geq \BIR_1(\; (1-p)n \;) &\qquad\text{(since $\E[m]\geq n(1-p)$)}  \\
+%    &\geq \beta\cdot n^{-\gamma} \cdot (1-p)^{-\gamma}
+%        &\qquad\text{(plugging in $\BIR_1(n)=\beta n^{-\gamma}$)}  \\
+%    &> \BIR_1(n)\;\; (1-p\gamma)^{-1}
+%        &\qquad\text{(since $(1-p)^\gamma < 1-p\gamma$)}.\\
+%\BIR_2(n)
+%    &>\alpha\cdot \BIR_1(n),
+%    &\text{where} \quad
+%    \alpha = \frac{1-p(1-c)}{1-p\gamma}>1.
+%\end{align*}
+%(In the above equations, all expectations are over $m$ distributed as in \eqref{eq:random-BIR}.)
+%\end{proof}
 
@@
 Suppose MAB algorithms \alg[1],  \alg[2] are $T$-oblivious. Say that
-\alg[1] \emph{weakly BIR-dominates} \alg[2] if there exist
-absolute  constants $\beta_0, \alpha_0\in (0, 1/2)$ and $n_0\in\N$ such that
+\alg[1] \emph{weakly BIR-dominates} \alg[2] if there exist absolute  constants $\beta_0, \alpha_0\in (0, 1/2)$ and $n_0\in\N$ such that
  \begin{align}\label{eq:SoftMax-better}
@@
 Assume the \SoftMaxRandom response function. Suppose algorithms \alg[1], \alg[2] are monotone and $T$-oblivious, and \alg[1] weakly-BIR-dominates \alg[2]. Posit mild technical assumptions:
-  $\BIR_1(n) \to 0$ and
+  $\BIR_1(n) \to 0$ and that $\BIR_2$ cannot be extremely small, namely: 
 \begin{align}\label{eq:SoftMax-assn-strong}
@@
  %  \BIR_2(n) > e^{-\eps n}. %used to be 4exp( -\eps n/6)
-(\exists n_0\; \forall n\geq n_0) \qquad
+(\exists m_0\; \forall n\geq m_0) \qquad
 \BIR_2(n) \geq \frac{4}{\alpha_0}
-\exp \left( \frac{-n\,\min\{\eps_0, 1/8\}}{12}\right).
+\exp \left( \frac{-n\,\min\{\eps_0, \nicefrac{1}{8}\}}{12}\right).
 \end{align}
@@
 Assume \SoftMaxRandom response function. Consider the \FiniteGame in which all feasible algorithms satisfy $\BReg(n)\to \infty$.%
-\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition.
-\ascomment{TODO: can we actually prove it? Assume Bernoulli priors, discrete prior, reduce to Lai-Robbins.}}
+\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition, see 
+\ascomment{insert pointer}.}
 Suppose a feasible algorithm \alg BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.

### 2020-07-14 journal paper/content/sim_details.tex
Minor edits for "in isolation" stuff.

@@
 
+\ascomment{TODO: revise!}
 We will utilize Thompson Sampling (\TS) from the ``adaptive-exploration" MAB algorithms, the dynamic $\eps$-greedy (\DEG)\footnote{Throughout, we fix $\eps = 0.05$.
@@
 Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.
-
-\ascomment{TODO: add a bit more detail -- \eg define \eps-greedy.}
 

### 2020-07-14 journal paper/ec19paper/content/inverted_u.tex
Removed the reference to the DS white paper (instead, we cite the arxiv paper).
Minor edits for "in isolation" stuff.

@@
 We do this for each ``intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
-For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{MWT-WhitePaper-2016,DS-arxiv}).
+For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{DS-arxiv}).
 Note that \DG is trivially the dominant strategy under monopoly.
@@
 \begin{finding}\label{find:temp-monopoly}
-\textit{Underfirst-mover, \TS is the dominant strategy for the incumbent. This holds across all MAB instances, if $X$ is large enough.
+\textit{Under first-mover, \TS is the dominant strategy for the incumbent. This holds across all MAB instances, if $X$ is large enough.
 }

### 2020-07-14 journal paper/ec19paper/content/perf_in_iso.tex
Minor edits for "in isolation" stuff.

@@
 We plot the \emph{mean reputation trajectory}: how this score evolves over time.
-Figure \ref{prelim_means} shows such a plot for the Needle-in-Haystack instance; for other MAB instances the plots are similar.
+\asedit{We also plot the trajectory for instantaneous rewards (\emph{not} averaged over the previous time-periods), which provides a better view into algorithm's performance at a given time.}%
+\footnote{For ``instantaneous reward" at a given time $t$, we report the average (over all \MRVs) of the mean rewards at this time, instead of the average of the \emph{realized} rewards, so as to decrease the noise.}
+Figure \ref{prelim_means} shows these trajectories for the Needle-in-Haystack instance; for other MAB instances the plots are similar.
+
 We summarize this finding as follows:
@@
 \begin{finding}
-\textit{The mean reputation trajectories are arranged as predicted by prior work:
+\textit{The mean reputation trajectories \asedit{and the instantaneous reward trajectories} are arranged as predicted by prior work:
     $\TS > \DEG > \DG$ for a sufficiently long time horizon.}

### 2020-07-14 journal paper/journal_main.tex
Added a macro.
Revised macro for the algorithms
Revised the title note.

@@
 \newcommand{\term}[1]{\ensuremath{\mathtt{#1}}\xspace}
-\newcommand{\TS}{\term{TS}}
-\newcommand{\DEG}{\term{DEG}}
-\newcommand{\DG}{\term{DG}}
-\newcommand{\AD}{\term{AD}}
-\newcommand{\EGR}{\term{EGR}}
-\newcommand{\GR}{\term{GR}}
+
+% shorthand for algorithms
+\newcommand{\TS}{\term{TS}}    % Thompson Sampling
+\newcommand{\DEG}{\term{BEG}}  % "Dynamic eps-greedy"
+\newcommand{\DG}{\term{BG}}    % "Dynamic Greedy"
+
+% shorthand for models
+\newcommand{\HMR}{\term{HMR}} % HardMaxRandom
+\newcommand{\HM}{\term{HM}}    % HardMax
+
 \newcommand{\Beta}{\term{Beta}} % for Beta distribution
@@
 
-\newcommand{\HMR}{\term{HMR}}
-\newcommand{\HM}{\term{HM}}
-
-% commands from ITCS 18 paper
+% commands from "theory"
 
@@
 \newcommand{\Ex}[2]{\Expectation_{#1}\left[#2\right]}
+\newcommand{\FiniteGame}{finite competition game\xspace}
 
@@
 
+% algorithms
 \newcommand{\StaticGreedy}{\term{StaticGreedy}}
-\newcommand{\DynGreedy}{\term{DynamicGreedy}}
-\newcommand{\DynamicGreedy}{\term{DynamicGreedy}}
+\newcommand{\DynGreedy}{\term{BayesianGreedy}}
+\newcommand{\DynamicGreedy}{\DynGreedy}
+\newcommand{\Thompson}{\term{ThompsonSampling}}
 
@@
 \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19},
-with a unified and streamlined presentation.}}
+with a unified and streamlined presentation and expanded background materials.
+All theoretical results are from \citet{CompetingBandits-itcs18}, and all experiments are from \citet{CompetingBandits-ec19}.}}
 % \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}

### 2020-07-14 journal paper/content/sec-theory-proofs.tex
Minor edit.
Minor edits for consisntency.

@@
 Fix agent $t$. Recall that $n_i(t)$ denotes the number of global rounds before $t$ in which principal $i$ is chosen. Let $\posteriorN{i}{t}$ denote the distribution of $n_i(t)$.
-
 \ascomment{Steven: in the ITCS paper, we defined $\posteriorN{i}{t}$ as the posterior distribution given all info available to agent $t$. But this is the same thing, isn't it?}
@@
     $\Pr[a_{2,n_0}\neq a_{1,n_0} \mid H=h]>0$.
-Combining this with \eqref{eq:lm:DG-rew-2}, we deduce that
+Combining this with \eqref{eq:lm:DG-rew-2}, 
 \begin{align}\label{eq:lm:DG-rew-3}
@@
 
-Because of the biased tie-breaking, for each global round $t$ we have:
+Recall that $i_t$ is the principal chosen in a given global round $t$.
+Because of the biased tie-breaking, 
 \begin{align}\label{eq:thm:HardMax-biased-PMRtoPr}
@@
 \end{align}
-Recall that $i_t$ is the principal chosen in global round $t$.
 
@@
 
+Without loss of generality, assume $m_0 = n_0$.
 Consider global round $t\geq n_0$. Recall that each agent chooses principal $1$ with probability at least
@@
 %\begin{proof}[Proof of Theorem~\ref{thm:random-greedy}]
-  Let $\rew'_1(n)$ denote the Bayesian-expected reward of the ``greedy
+  Let $\rewgr(n)$ denote the Bayesian-expected reward of the ``greedy
   choice'' after after $n-1$ steps of \alg[1]. Note that
-  $\rew_1(\cdot)$ and $\rew'_1(\cdot)$ are non-decreasing: the former
+  $\rew_1(\cdot)$ and $\rewgr(\cdot)$ are non-decreasing: the former
   because \alg[1] is monotone and the latter because the ``greedy
   choice'' is only improved with an increasing set of
-  observations. Therefore, the modified algorithm \alg[2] is monotone
-  by \eqref{eq:random-BIR}.
+  observations. \ascomment{insert a pointer} 
+Using \eqref{eq:BIR-modification}, we conclude that
+the greedy modification \alg[2] is monotone.
 
-  By definition of the ``greedy choice,'' $\rew_1(n)\leq \rew'_1(n)$
+  By definition of the ``greedy choice,'' $\rew_1(n)\leq \rewgr(n)$
   for all steps $n$. Moreover, by Lemma~\ref{lm:DG-rew},
@@
 
-Let $\alg$ denote a copy of \alg[1] that is running ``inside" the modified algorithm \alg[2].
-Let $m_2(t)$ be the number of global rounds before $t$ in which the agent chooses principal $2$ \emph{and} \alg is invoked; in other words, it is the number of agents seen by \alg before global round $t$.
+Let $\alg$ denote a copy of \alg[1] that is running ``inside" \alg[2].
+Let $m_2(t)$ be the number of global rounds before $t$ in which the agent chooses principal $2$ \emph{and} \alg is invoked; \ie it is the number of agents seen by \alg before global round $t$.
 Let $\mM_{2,t}$ be the agents' posterior distribution for $m_2(t)$.
@@
     \label{eq:pf:thm:random-greedy-1}\\
-  &< \Ex{m\sim \mM_{2,t}} {(1-p)\cdot \rew_1(m+1) + p\cdot \rew'_1(m+1)}
+  &< \Ex{m\sim \mM_{2,t}} {(1-p)\cdot \rew_1(m+1) + p\cdot \rewgr(m+1)}
     \label{eq:pf:thm:random-greedy-2} \\
@@
 \[
-  p_t \geq \frac{1}{2} - c_0'\left(\PMR_2(t) - \PMR_1(t)\right) \geq
-  \frac{1}{2} - \beta_1.
+  p_t \geq \tfrac{1}{2} - c_0'\left(\PMR_2(t) - \PMR_1(t)\right) \geq
+  \tfrac{1}{2} - \beta_1.
 \]
@@
 have with probability at least $1 - q_2(t)$ that the number of agents
-that choose \alg[1] is at least $\beta_0(1/2 - \beta_1)t/5$, where the
-function
+that choose \alg[1] is at least 
+    $\beta_0(\nicefrac{1}{2} - \beta_1)t/5$, 
+where %the function
 $$
-q_2(x) = \exp\left( \frac{-(1/2 - \beta_1)(1 - \beta_0/5)^2x}{3} \right).
+q_2(x) = \exp\rbr{ -\nicefrac{1}{3}\; 
+    (\nicefrac{1}{2} - \beta_1)(1 - \beta_0/5)^2x} .
 $$
@@
 in~\eqref{eq:SoftMax-better} implies that for any $t\geq T_1 + T_2$
-with probability at least $1 - q_2(t)$,
-\[
-  \BIR_1(n_1(t)) < (1- \alpha_0)\BIR_2(n_2(t)).
-\]
+with probability at least $1 - q_2(t)$, we have 
+$ \BIR_1(n_1(t)) < (1- \alpha_0)\,\BIR_2(n_2(t))$.
+
 It follows that for any $t\geq T_1 + T_2$,
@@
   \PMR_1(t) - \PMR_2(t) &= \Ex{m_1\sim \posteriorN{1}{t},\;m_2\sim \posteriorN{2}{t}}{\BIR_2(m_2+ 1) - \BIR_1(m_1+1)} \\
-                        &\geq (1 - q_2(t))\alpha_0 \BIR_2(t) - q_2(t)\\
-                        &\geq \alpha_0 \BIR_2(t)/4
+                        &\geq (1 - q_2(t))\,\alpha_0\, \BIR_2(t) - q_2(t)
+                        \geq \alpha_0\, \BIR_2(t)/4,
 \end{align*}
@@
 function and assumption in~\eqref{eq:SoftMax-smooth}, we have
-\[
-  \Pr[i_t = 1] \geq \frac{1}{2} + \frac{c_0\alpha_0\BIR_2(t)}{4}. \qedhere
-\]
+$  \Pr[i_t = 1] \geq \nicefrac{1}{2} + 
+    \nicefrac{1}{4}\; c_0\,\alpha_0\,\BIR_2(t)$.
 %\end{proof}

### 2020-07-14 journal paper/content/sec-theory.tex
Moved some macro to the main file.
Minor edits in "theory/extensions".

@@
 \end{theorem}
-
-\newcommand{\gr}{\texttt{gr}}
-\newcommand{\BIRgr}{\BIR^{\gr}}
-\newcommand{\alggr}{\alg[\gr]}
 
@@
 It corresponds to the least amount of rationality and competition, and (when principals' utility is the number of agents) uniform choice provides no incentives to innovate.}%
-\footnote{On the other hand, if principals' utility is somewhat aligned with agents' welfare, as in \eqref{eq:general-utility}, then a monopolist principal is incentivized to choose the best possible MAB algorithm (namely, to minimize cumulative Bayesian regret $\BReg(T)$).
+\footnote{On the other hand, if principals' utility is aligned with agents' welfare, then a monopolist principal is incentivized to choose the best possible MAB algorithm (namely, to minimize cumulative Bayesian regret $\BReg(T)$).
 Accordingly, monopoly would result in better social welfare than competition, as the latter is likely to split the market and cause each principal to learn more slowly.
@@
 
-\begin{itemize}
+\begin{OneLiners}
 \item $\eps_0 = 0$ (\HardMax):~~~~ the marginal utility can be negative if \alg is \DynGreedy.
@@
 \item $\eps_0$ near $\tfrac12$:~~~~ small marginal utility, as principal 1 gets most agents for free no matter what.
-\end{itemize}
+\end{OneLiners}
 The familiar inverted-U shape is depicted in Figure~\ref{fig:inverted-U3}.
 
-\begin{figure}
+\begin{figure}[t]
 \begin{center}
@@
 In a formula,
-\begin{align}\label{eq:general-utility}
-\textstyle U_i := \sum_{t=1}^T\; \indicator{i_t=i}\cdot U_i(r_r).
-\end{align}
+%\begin{align}\label{eq:general-utility}
+%\textstyle 
+$U_i := \sum_{t=1}^T\; \indicator{i_t=i}\cdot U_i(r_r)$.
+%\end{align}
 
@@
 \xhdr{\BIR can depend on $T$.}
-Many/most MAB algorithms are parameterized by the time horizon $T$, and their regret bounds usually include $\polylog(T)$. In particular,  a regret bound for \BIR is
+Many MAB algorithms are parameterized by the time horizon $T$, and their regret bounds usually include $\polylog(T)$. In particular,  a typical regret bound for \BIR is
 \begin{align}
@@
 \end{align}
-Here we write $\BIR(n\mid T)$ to emphasize the dependence on $T$.
-
-Accordingly, BIR-dominance can be redefined as follows: there exists a number $T_0$ and a function $n_0(T)\in \polylog(T)$
+Here we write $\BIR(n\mid T)$ to emphasize the dependence on $T$. Accordingly, BIR-dominance can be redefined as follows: there exists a number $T_0$ and a function $n_0(T)\in \polylog(T)$
 such that
@@
 
+\noindent Weak BIR-dominance can be redefined similarly.
+Theorem~\ref{thm:random-clean} and Theorem~\ref{thm:SoftMax-weak} extend to these generalized versions; the easy modifications are omitted.
+
 %\begin{theorem}\label{thm:random-messy}
@@
 
-Similarly, weak BIR-dominance can be redefined as follows: there exist some $T_0$, a function
-$n_0(T)\in \polylog(T)$, and constants $\beta_0,\alpha_0\in (0, 1/2)$, such that
-\begin{align}\label{eq:SoftMax-better-weaker}
-(\forall T\geq T_0,  n\geq n_0(T)) \quad
-\frac{\BIR_1((1-\beta_0)\, n\mid T)}{\BIR_2(n\mid T)} <1-\alpha_0.
-\end{align}
+%Similarly, weak BIR-dominance can be redefined as follows: there exist some $T_0$, a function
+%$n_0(T)\in \polylog(T)$, and constants $\beta_0,\alpha_0\in (0, 1/2)$, such that
+%\begin{align}\label{eq:SoftMax-better-weaker}
+%(\forall T\geq T_0,  n\geq n_0(T)) \quad
+%\frac{\BIR_1((1-\beta_0)\, n\mid T)}{\BIR_2(n\mid T)} <1-\alpha_0.
+%\end{align}
 
-Theorem~\ref{thm:random-clean} and Theorem~\ref{thm:SoftMax-weak} extend to these generalized versions; the easy modifications are omitted.
+%Theorem~\ref{thm:random-clean} and Theorem~\ref{thm:SoftMax-weak} extend to these generalized versions; the easy modifications are omitted.

### 2020-07-14 journal paper/content/sim_details.tex
Minor edits throughout.

@@
 
-\asedit{In this section we present our numerical simulations.
+In this section we present our numerical simulations.
 As discussed in the Introduction, we focus on the \ExptsModel, whereby each agent chooses the firm with a maximal reputation score, modeled as a sliding window average of its rewards.
-Given the volume of experimental results, we (only) present the typical / representative ones in the body of the paper; the full results can be found in Appendix~\ref{app:expts}.}
+While we experiment with various MAB instances and parameter settings, we only report on selected, representative experiments.
+Additional plots and tables are provided in Appendix~\ref{app:expts}.
+Unless noted otherwise, our findings are based on and consistent with all these experiments.
 
@@
 
-\xhdr{Agent model.} All experiments use \HardMax response function (without mentioning it), except Section~\ref{sec:non_greedy} where we use \HardMaxRandom agents.
-
-\xhdr{First-movers.}
-Recall that in some of our experiments, one firm is the ``incumbent" who enters the market before the other (``late entrant"), and therefore enjoys a \emph{first-mover advantage}.
+\xhdr{Competition model.} All experiments use \HardMax response function (without mentioning it), except Section~\ref{sec:non_greedy} where we use \HardMaxRandom agents.
+In some of our experiments, one firm is the ``incumbent" who enters the market before the other (``late entrant"), and therefore enjoys a \emph{first-mover advantage}.
 Formally, the incumbent enjoys additional $X$ rounds of the ``warm start".
@@
 
-\xhdr{MAB algorithms.} We consider the same three classes of bandit algorithms discussed in Section~\ref{sec:intro-model}, but take a representative algorithm from each class and look for qualitative differences between the different classes.
+\xhdr{MAB algorithms.} We consider the three classes of bandit algorithms discussed in Section~\ref{sec:intro-model}, and look for qualitative differences between them.
+We take a representative algorithm from each class.
 \asedit{Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.}
-
-\ascomment{TODO: revise!}
-We will utilize Thompson Sampling (\TS) from the ``adaptive-exploration" MAB algorithms, the dynamic $\eps$-greedy (\DEG)\footnote{Throughout, we fix $\eps = 0.05$.
-Our pilot experiments showed that different $\eps$ did not qualitatively change the results.} from the ``exploration-separating" algorithms, and the ``dynamic greedy" algorithm (\DG) as in Section~\ref{sec:theory-prelims}.
+We use Thompson Sampling (\TS) from the ``adaptive-exploration" MAB algorithms, Bayesian Epsilon-Greeedy (\DEG) with exploration probability $\eps=5\%$ from the ``exploration-separating" algorithms,%
+\footnote{Our pilot experiments show that different $\eps$ did not qualitatively change the results.}  and \DynGreedy algorithm (\DG) as in Section~\ref{sec:theory-prelims}.
 For ease of comparison, all three algorithms are parameterized with the same ``fake" Bayesian prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution.
 Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.
+Self-contained background can be found in Appendix~\ref{sec:related-classes}.
 
@@
 
-We fix the sliding window size $M = 100$.
+For the reputation scores, we fix the sliding window size $M = 100$.
 We found that lower values induced too much random noise in the results, and increasing $M$ further did not make a qualitative difference.
@@
 This is for a particular MAB instance and a particular selection of the game parameters.
-
-\xhdr{Discussion}
-While we experiment with various MAB instances and parameter settings, we only report on selected, representative experiments in the body of the paper.
-Additional plots and tables are provided in the appendix.
-Unless noted otherwise, our findings are based on and consistent with all these experiments.
 
@@
 
+\xhdr{Running time.}
 The simulations are computationally intensive.

### 2020-07-14 journal paper/content/sim_details.tex
Removed some color-edits.

@@
 We take a representative algorithm from each class.
-\asedit{Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.}
+Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.
 We use Thompson Sampling (\TS) from the ``adaptive-exploration" MAB algorithms, Bayesian Epsilon-Greeedy (\DEG) with exploration probability $\eps=5\%$ from the ``exploration-separating" algorithms,%

### 2020-07-14 journal paper/content/sim_details.tex
Revised a footnote re BEG.

@@
 Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.
-We use Thompson Sampling (\TS) from the ``adaptive-exploration" MAB algorithms, Bayesian Epsilon-Greeedy (\DEG) with exploration probability $\eps=5\%$ from the ``exploration-separating" algorithms,%
-\footnote{Our pilot experiments show that different $\eps$ did not qualitatively change the results.}  and \DynGreedy algorithm (\DG) as in Section~\ref{sec:theory-prelims}.
+We use Thompson Sampling (\TS) from the ``adaptive-exploration" algorithms, Bayesian Epsilon-Greeedy (\DEG) with exploration probability $\eps=5\%$ from the ``exploration-separating" algorithms,%
+\footnote{In each round, \DEG algorithm explores uniformly with a predetermined probability $\eps$, and ``exploits" with the remaining probability, choosing an arm with maximal posterior mean reward given the current data.
+Our pilot experiments show that choosing a different exploration probability $\eps$ does not qualitatively change the results.}  and \DynGreedy (\DG) algorithm  as in Section~\ref{sec:theory-prelims}.
 For ease of comparison, all three algorithms are parameterized with the same ``fake" Bayesian prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution.

### 2020-07-14 journal paper/content/sec-theory.tex
Minor fix.
Minor edits.

@@
 \begin{align}\label{eq:theory-BIR-defn}
-\BIR_i(n) := \E\sbr{ n\cdot \max_{a\in A} \mu_a - \rew_i(n)}.
+\BIR_i(n) := \E\sbr{ \max_{a\in A} \mu_a - \rew_i(n)}.
 \end{align}
-We are primarily interested in how \BIR scales with $n$: typically, how fast \BIR decreases with $n$.
-We treat $K$, the number of arms, as a constant unless specified otherwise.
-Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:examples} for more background.
+We are primarily interested in how fast \BIR decreases with $n$.
+(We treat the number of arms as a constant.) Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:examples} for background.
 

### 2020-07-14 journal paper/ec19paper/content/revisited.tex
formatting tweaks
updates to intro + numerical sections
update figure placement and use BEG, BG instead of DEG, DG

@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.35]{ec19paper/figures/rep_distribution_nih}
+\includegraphics[scale=0.35]{ec19paper/figures/ts_dg_rep_diff_nih}
+\caption{Reputation scores for Needle-in-Haystack at $t=500$ (left), Reputation difference $\TS - \DG$ for Needle-in-Haystack (right).Both are smoothed using a kernel density estimate.}
+\label{rep_dist_nih}
+%\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
+\end{figure}
+
 \begin{finding}
@@
 That is, consider the empirical distribution of this score over different \MRVs.%
-\footnote{Recall that each \MRV in our experimental setup comes with one specific realization table.} For concreteness, consider the Needle-in-Haystack instance at time $t=500$, plotted in Figure \ref{rep_dist_nih}.
+\footnote{Recall that each \MRV in our experimental setup comes with one specific realization table.} For concreteness, consider the Needle-in-Haystack instance at time $t=500$, plotted in Figure \ref{rep_dist_nih} (left).
 (The other MAB instances lead to a similar intuition.)
-
-\begin{figure}[H]
-\includegraphics[scale=0.35]{ec19paper/figures/rep_distribution_nih}
-\caption{Reputation scores for Needle-in-Haystack at $t=500$ (smoothed using a kernel density estimate)}
-\label{rep_dist_nih}
-%\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
-\end{figure}
 
@@
 Let's plot the empirical distribution of the reputation difference (over the \MRVs) at a particular time point.
-Figure~\ref{ts_dg_rep_diff_nih} shows such plots for several time points.
+Figure~\ref{ts_dg_rep_diff_nih} (right) shows such plots for several time points.
 We observe that the distribution is skewed to the right, precisely due to the fact that $\DG$ either does slightly better than $\TS$ or does substantially worse.
@@
 \OMIT{In particular, the mean reputation difference tends to overstate how well an advanced exploration algorithm (that will eventually find the best arm) should do in competition compared to a naive algorithm that may never find the best arm.}
-
-\begin{figure}[H]
-\includegraphics[scale=0.35]{ec19paper/figures/ts_dg_rep_diff_nih}
-\caption{Reputation difference $\TS-\DG$ for  Needle-in-Haystack
-% at $t = 500, 1000, 2000$ AS: this is redundant, and it saves a line :).
-(smoothed using a kernel density estimate)}
-\label{ts_dg_rep_diff_nih}
-%\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $\TS$ and $\DG$ across $t$}}
-\end{figure}
 

### 2020-07-14 analysis_code/relative_reputation_plots.R

@@ -24,14 +24,14 @@ concise_alg_rep <- function(alg) {
   if (alg == "ThompsonSampling") {
     return("TS")
   } else if (alg == "DynamicEpsilonGreedy") {
-    return("DEG")
+    return("BEG")
   } else if (alg == "DynamicGreedy") {
-    return("DG")
+    return("BG")
   }
 }
 
 dat$alg<- sapply(dat$Algorithm, concise_alg_rep)
-print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
+print_relative_graphs <- function (dat, dist, alg1, alg2, minComplexity, withAnnotation) {
   dist_dat <- filter(dat, Distribution == dist & Realized.Complexity >= minComplexity)
   alg_1 <- filter(dist_dat, Algorithm == alg1)
   alg_2 <- filter(dist_dat, Algorithm == alg2)
@@ -55,21 +55,29 @@ print_relative_graphs <- function (dist, alg1, alg2, minComplexity) {
   df$relative_rep <- counts
   df$se <- ses
   plot_title <- paste("Relative Reputation -", dist)
-  q <- ggplot(df, aes(x=t, y=relative_rep)) + geom_line() + geom_point() +
-    geom_errorbar(aes(ymin=relative_rep-1.96*se, ymax=relative_rep+1.96*se), width=.2) +
-    ggtitle(plot_title) + xlab("time") + 
-    ylab(paste(concise_alg_rep(alg1), " >= ", concise_alg_rep(alg2), "(reputation)")) +
-    #annotation_custom(b1) +
-    #annotation_custom(b2) +
-    #annotate("text", x = 275, y = 0.28, label = "Exploration Disadvantage Period", size=5) +
-    #annotate("text", x = 1000, y = 0.19, label = "Relative Reputation Benefit") +
-    theme_bw(base_size = 20) +  ylim(c(0.2, 0.7)) +
-    theme(plot.title = element_text(hjust = 0.5))
+  q <- NA
+  if (withAnnotation) {
+    q <- ggplot(df, aes(x=t, y=relative_rep)) + geom_line() + geom_point() +
+      geom_errorbar(aes(ymin=relative_rep-1.96*se, ymax=relative_rep+1.96*se), width=.2) +
+      ggtitle(plot_title) + xlab("time") + 
[…]

### 2020-07-15 journal paper/content/sim_details.tex
DEG -> BEG, DG -> BG

@@
 Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.
-We use Thompson Sampling (\TS) from the ``adaptive-exploration" algorithms, Bayesian Epsilon-Greeedy (\DEG) with exploration probability $\eps=5\%$ from the ``exploration-separating" algorithms,%
-\footnote{In each round, \DEG algorithm explores uniformly with a predetermined probability $\eps$, and ``exploits" with the remaining probability, choosing an arm with maximal posterior mean reward given the current data.
+We use \Thompson (\TS) from the ``adaptive-exploration" algorithms, \DynamicEpsGreedy (\DEG) with exploration probability $\eps=5\%$ from the ``exploration-separating" algorithms,%
+\footnote{In each round, \DynamicEpsGreedy algorithm explores uniformly with a predetermined probability $\eps$, and ``exploits" with the remaining probability, choosing an arm with maximal posterior mean reward given the current data.
 Our pilot experiments show that choosing a different exploration probability $\eps$ does not qualitatively change the results.}  and \DynGreedy (\DG) algorithm  as in Section~\ref{sec:theory-prelims}.

### 2020-07-15 journal paper/ec19paper/content/barriers.tex
Minor edits throughout.
DEG -> BEG, DG -> BG

@@
 \textit{\\
-(a) Data advantage and reputation advantage alone are substantially large, across all algorithms and all MAB instances.
-\\(b) The data advantage is larger than the reputation advantage when the incumbent chooses \TS.
-\\(c) The two advantages are similar in magnitude when the incumbent chooses \DEG or \DG.
+(a) Data advantage and reputation advantage alone are large, across all algorithms and MAB instances.
+\\(b) The data advantage is larger than the reputation advantage when the incumbent chooses \Thompson.
+\\(c) The two advantages are similar in magnitude when the incumbent chooses \DynamicEpsGreedy or \DynamicGreedy.
 }
@@
 Our intuition for Finding~\ref{barrier-find}(b) is as follows.
-Suppose the incumbent switches from \DG to \TS.
+Suppose the incumbent switches from \DynamicGreedy to \Thompson.
 This switch allows the incumbent to explore actions more efficiently -- collect better data in the same number of rounds -- and therefore should benefit the data advantage.

### 2020-07-15 journal paper/ec19paper/content/inverted_u.tex
Minor edits.
DEG -> BEG, DG -> BG

@@
 As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
-More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DG, \DEG and \TS.
+More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DynamicGreedy, \DynamicEpsGreedy and \Thompson.
 We do this for each ``intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
 For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{DS-arxiv}).
-Note that \DG is trivially the dominant strategy under monopoly.
+Note that \DynamicGreedy is trivially the dominant strategy under monopoly.
 
@@
 \begin{itemize}
-\item[(a)] (\DG,\DG) is the unique pure-strategy Nash equilibrium for exploration-disadvantaged MAB instances with a sufficiently small ``warm start".
-\item[(b)] This is not necessarily the case for MAB instances that are not exploration-disadvantaged. In particular, \TS is a weakly dominant strategy for Needle-in-Haystack.
+\item[(a)] (\DynamicGreedy,\DynamicGreedy) is the unique pure-strategy Nash equilibrium for exploration-disadvantaged MAB instances with a sufficiently small ``warm start".
+\item[(b)] This is not necessarily the case for MAB instances that are not exploration-disadvantaged. In particular, \Thompson is a weakly dominant strategy for Needle-in-Haystack.
 \end{itemize}
@@
 We investigate the firms' market shares when they choose different algorithms (otherwise, by symmetry both firms get half of the agents).
-We report the market shares for Heavy-Tail and Needle-in-Haystack instances in Table \ref{sim_table}  (see the first line in each cell), for a range of values of the warm start $T_0$.
+We report the market shares for Heavy-Tail and Needle-in-Haystack instances in Table~\ref{sim_table}  (see the first line in each cell), for a range of values of the warm start $T_0$.
 Table~\ref{tab:duopoly_unif} reports similarly on the Uniform instance.
 We find that $\DG$ is a weakly dominant strategy for the Heavy-Tail and Uniform instances, as long as $T_0$ is sufficiently small.
-However, \TS is a weakly dominant strategy for the Needle-in-Haystack instance.
-We find that for a sufficiently small $T_0$, \DG yields more than half the market against \TS,  but achieves similar market share vs.
-\DG and \DEG.
-By our tie-breaking rule, (\DG,\DG) is the only pure-strategy equilibrium.
+However, \Thompson is a weakly dominant strategy for the Needle-in-Haystack instance.
+We find that for a sufficiently small $T_0$, \DynamicGreedy yields more than half the market against \Thompson,  but achieves similar market share vs.
+\DynamicGreedy and \DynamicEpsGreedy.
+By our tie-breaking rule, (\DynamicGreedy,\DynamicGreedy) is the only pure-strategy equilibrium.
 
@@
 
-We attribute the prevalence of \DG on exploration-disadvantaged MAB instances to its prevalence on the initial ``exploration disadvantage period", as described in Section~\ref{sec:isolation}.
+We attribute the prevalence of \DynamicGreedy on exploration-disadvantaged MAB instances to its prevalence on the initial ``exploration disadvantage period", as described in Section~\ref{sec:isolation}.
 Increasing the warm start length $T_0$ makes this period shorter: indeed, considering the relative reputation trajectory in Figure~\ref{relative_rep_plots} (left), increasing $T_0$ effectively shifts the starting time point to the right.
-This is why it helps \DG if $T_0$ is small.
+This is why it helps \DynamicGreedy if $T_0$ is small.
 
 \OMIT{
-The reason we see that $\DG$ is incentivized under the Heavy Tail and Uniform instances but that $\TS$ is weakly dominant under Needle In Haystack is precisely due to the fact that the purposeful exploration engaged by $\TS$ in the early rounds puts it in a disadvantage in the former case but not in the latter.
+The reason we see that $\DynamicGreedy$ is incentivized under the Heavy Tail and Uniform instances but that $\Thompson$ is weakly dominant under Needle In Haystack is precisely due to the fact that the purposeful exploration engaged by $\Thompson$ in the early rounds puts it in a disadvantage in the former case but not in the latter.
 Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start $T_0$ as fixing the starting point on the relative reputation plots.
 The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $T_0$.
-As a result, we see that for warm start $T_0 = 20$ on the exploration-disadvantaged instances, $\DG$ will win the first agent in a larger proportion of the simulations than $\TS$.
-However, on the Needle In Haystack instances, we see that $\TS$ wins a larger proportion of the first agents compared to $\DG$.
+As a result, we see that for warm start $T_0 = 20$ on the exploration-disadvantaged instances, $\DynamicGreedy$ will win the first agent in a larger proportion of the simulations than $\Thompson$.
+However, on the Needle In Haystack instances, we see that $\Thompson$ wins a larger proportion of the first agents compared to $\DynamicGreedy$.
 }
@@
 We make $X$ large enough, but still much smaller than the time horizon $T$.
-We find that the incumbent is incentivized to choose \TS, in a strong sense:
+We find that the incumbent is incentivized to choose \Thompson, in a strong sense:
 
 \begin{finding}\label{find:temp-monopoly}
-\textit{Under first-mover, \TS is the dominant strategy for the incumbent. This holds across all MAB instances, if $X$ is large enough.
+\textit{Under first-mover, \Thompson is the dominant strategy for the incumbent. This holds across all MAB instances, if $X$ is large enough.
 }
@@
 The simulation results for the Heavy-Tail MAB instance are reported in Table~\ref{tab:ht-incum}, for a particular $X=200$.
-We see that \TS is a dominant strategy for the incumbent.
+We see that \Thompson is a dominant strategy for the incumbent.
 Similar tables for the other MAB instances and other values of $X$ are reported in the supplement, with the same conclusion.
@@
 \captionsetup{labelformat=empty}
-\caption{User share of row player (entrant), 200 round head-start, Heavy-Tail Instance}
+\caption{Market share of row player (entrant), 200 round head-start, Heavy-Tail Instance}
 %\caption{{\bf Temporary monopoly}, with $X=200$ (and $T_0=20$), for the Heavy-Tail MAB instance.
@@
 
-\DG is a weakly dominant strategy for the entrant, for Heavy-Tail instance in Table~\ref{tab:ht-incum} and the Uniform instance, but not for the Needle-in-Haystack instance.
+\DynamicGreedy is a weakly dominant strategy for the entrant, for Heavy-Tail instance in Table~\ref{tab:ht-incum} and the Uniform instance, but not for the Needle-in-Haystack instance.
 We attribute this finding to exploration-disadvantaged property of these two MAB instance, for the same reasons as discussed above.
@@
 \begin{finding}\label{find:temp-monopoly-entrant}
-\textit{Under first-mover, \DG is a weakly dominant strategy for the entrant for exploration-disadvantaged MAB instances.
+\textit{Under first-mover, \DynamicGreedy is a weakly dominant strategy for the entrant for exploration-disadvantaged MAB instances.
 }
@@
 We interpret our findings through the lens of the inverted-U relationship between the ``intensity of competition" and the ``quality of technology".
-The lowest level of competition is monopoly, when \DG wins out for the trivial reason of tie-breaking.
+The lowest level of competition is monopoly, when \DynamicGreedy wins out for the trivial reason of tie-breaking.
 The highest levels are simultaneous entry and ``late entrant".
-We see that \DG is incentivized for exploration-disadvantaged MAB instances.
-In fact, incentives for \DG get stronger when the model transitions from simultaneous entry to ``late entrant".%
-\footnote{For the Heavy-Tail instance, \DG goes from a weakly dominant strategy to a strictly dominant one.
-For the Uniform instance, \DG goes from a Nash equilibrium strategy to a weakly dominant one.}
-Finally, the middle level of competition, ``incumbent" in the first-mover regime creates strong incentives for \TS.
+We see that \DynamicGreedy is incentivized for exploration-disadvantaged MAB instances.
+In fact, incentives for \DynamicGreedy get stronger when the model transitions from simultaneous entry to ``late entrant".%
+\footnote{For the Heavy-Tail instance, \DynamicGreedy goes from a weakly dominant strategy to a strictly dominant one.
+For the Uniform instance, \DynamicGreedy goes from a Nash equilibrium strategy to a weakly dominant one.}
+Finally, the middle level of competition, ``incumbent" in the first-mover regime creates strong incentives for \Thompson.
 In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U-expts}.%
@@
 We interpret low \Eeog as a strong evidence of the ``death spiral" effect.
-Focusing on the simultaneous entry scenario, we specify the \Eeog values in Table \ref{sim_table} (the second line of each cell).
+Focusing on the simultaneous entry scenario, we specify the \Eeog values in Table~\ref{sim_table} (the second line of each cell).
 We find that the \Eeog values are indeed small:
@@
 We assume that both firms play their respective equilibrium strategies for the corresponding competition level. As discussed previously, these are:
-\begin{itemize}
-\item \DG in the monopoly,
-\item \DG for both firms in simultaneous entry (Finding \ref{find:duopoly}),
-\item \TS for the incumbent (Finding \ref{find:temp-monopoly}) and \DG for the entrant in first-mover (Finding \ref{find:temp-monopoly-entrant}).
-\end{itemize}
+\begin{OneLiners}
+\item \DynamicGreedy in the monopoly,
+\item \DynamicGreedy for both firms in simultaneous entry (Finding \ref{find:duopoly}),
+\item \Thompson for the incumbent (Finding \ref{find:temp-monopoly}) and \DynamicGreedy for the entrant in first-mover (Finding \ref{find:temp-monopoly-entrant}).
+\end{OneLiners}
 
@@
 the simultaneous entry case.
-Recall that under first-mover, the incumbent is incentivized to play \TS.
-Moreover, we find that the welfare is close to that of having a single firm for all agents and running \TS.
+Recall that under first-mover, the incumbent is incentivized to play \Thompson.
+Moreover, we find that the welfare is close to that of having a single firm for all agents and running \Thompson.
 We also observe that monopoly and simultaneous entry achieve similar welfare.
@@
 
-Finding~\ref{find:welfare}(b) is interesting because, in equilibrium, both firms play \DG in both settings, and one might conjecture that the welfare should increase with the number of firms playing \DG.
-Indeed, one run of \DG may get stuck on a bad arm.
-However, two firms independently playing \DG are less likely to get stuck simultaneously.
+Finding~\ref{find:welfare}(b) is interesting because, in equilibrium, both firms play \DynamicGreedy in both settings, and one might conjecture that the welfare should increase with the number of firms playing \DynamicGreedy.
+Indeed, one run of \DynamicGreedy may get stuck on a bad arm.
+However, two firms independently playing \DynamicGreedy are less likely to get stuck simultaneously.
 If one firm gets stuck and the other does not, then the latter should attract most agents, leading to improved welfare.
@@
 \includegraphics[scale=0.3]{ec19paper/figures/ht_many_firm_welfare}\\
-\caption{Average welfare and \Eeog as we increase the number of firms playing \DG}
+\caption{Average welfare and \Eeog as we increase the number of firms playing \DynamicGreedy}
 \label{many_firm_welfare}
@@
 
-To study this phenomenon further, we go beyond the duopoly setting to more than two firms playing \DG (and starting at the same time). Figure \ref{many_firm_welfare} reports the average welfare
+To study this phenomenon further, we go beyond the duopoly setting to more than two firms playing \DynamicGreedy (and starting at the same time).
+Figure~\ref{many_firm_welfare} reports the average welfare
 across these simulations. Welfare not only does not get better, \textit{but is weakly worse} as we increase the number of firms.
@@
 \begin{finding}
-\textit{When all firms deploy \DG, and start at the same time, welfare is weakly decreasing as the number of firms increases.}
+\textit{When all firms deploy \DynamicGreedy, and start at the same time, welfare is weakly decreasing as the number of firms increases.}
 \end{finding}
@@
 simulations and notice that it \textit{increases} with the number of firms.
-This observation also runs counter of the intuition that with more firms running \DG, one of them is more likely to ``get lucky" and take over the market (which would cause \Eeog to \emph{decrease} with the number of firms).
+This observation also runs counter of the intuition that with more firms running \DynamicGreedy, one of them is more likely to ``get lucky" and take over the market (which would cause \Eeog to \emph{decrease} with the number of firms).
 

### 2020-07-15 journal paper/ec19paper/content/perf_in_iso.tex
Minor edits throughout.
DEG -> BEG, DG -> BG

@@
 We confirm that algorithms' performance is ordered as we'd expect:
-    $\TS > \DEG > \DG$
+    $\Thompson > \DynamicEpsGreedy > \DynamicGreedy$
 for a sufficiently long time horizon.
@@
 \textit{The mean reputation trajectories \asedit{and the instantaneous reward trajectories} are arranged as predicted by prior work:
-    $\TS > \DEG > \DG$ for a sufficiently long time horizon.}
+    $\Thompson > \DynamicEpsGreedy > \DynamicGreedy$ for a sufficiently long time horizon.}
 \end{finding}
@@
 \includegraphics[scale=0.35]{ec19paper/figures/relative_nih_ts_dg}
-%\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$.
+%\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\Thompson \geq \DynamicGreedy$.
 The shaded area display 95\% confidence intervals.}}
-\caption{Relative reputation trajectory for $\TS$ vs $\DG$, on Uniform instance (left) and Needle-in-Haystack instance (right). Shaded area display 95\% confidence intervals.}
+\caption{Relative reputation trajectory for $\Thompson$ vs $\DynamicGreedy$, on Uniform instance (left) and Needle-in-Haystack instance (right). Shaded area display 95\% confidence intervals.}
 \label{relative_rep_plots}
@@
 benefits under competition.
-Figure \ref{relative_rep_plots} (left) shows the relative reputation trajectory for $\TS$ vs $\DG$ for the Uniform instance.
-The relative reputation is less than $\tfrac12$ in the early rounds, meaning that $\DG$ has a higher reputation score in a majority of the simulations, and more than $\tfrac12$ later on.
-The reason is the exploration in \TS leads to worse decisions initially, but allows for better decisions later.
+Figure \ref{relative_rep_plots} (left) shows the relative reputation trajectory for $\Thompson$ vs $\DynamicGreedy$ for the Uniform instance.
+The relative reputation is less than $\tfrac12$ in the early rounds, meaning that $\DynamicGreedy$ has a higher reputation score in a majority of the simulations, and more than $\tfrac12$ later on.
+The reason is the exploration in \Thompson leads to worse decisions initially, but allows for better decisions later.
 The time period when relative reputation vs.
-\DG dips below $\tfrac12$ can be seen as an explanation for the competitive disadvantage of exploration.
+\DynamicGreedy dips below $\tfrac12$ can be seen as an explanation for the competitive disadvantage of exploration.
 Such period also exists for the Heavy-Tail MAB instance.
-However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (right).%
-\footnote{We see two explanations for this: $\TS$ identifies the best arm faster for the Needle-in-Haystack instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
+However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots}.%
+\footnote{We see two explanations for this: $\Thompson$ identifies the best arm faster for the Needle-in-Haystack instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
 
 \begin{finding}\label{find:period}
-\textit{Exploration can lead to relative reputation vs. $\DG$ going below $\tfrac12$ for some initial time period. This happens for some MAB instances but not for some others.}
+\textit{Exploration can lead to relative reputation vs. $\DynamicGreedy$ going below $\tfrac12$ for some initial time period. This happens for some MAB instances but not for some others.}
 \end{finding}
@@
 For a particular MAB algorithm, a time period when relative reputation vs.
-\DG goes below $\tfrac12$ is called {\em exploration disadvantage period}.
+\DynamicGreedy goes below $\tfrac12$ is called {\em exploration disadvantage period}.
 An MAB instance is called \emph{exploration-disadvantaged} if such period exists.

### 2020-07-15 journal paper/ec19paper/content/revisited.tex
DEG -> BEG, DG -> BG

@@
 However, we find a more nuanced picture.
-For example, in Figure \ref{sim_table} we see that $\DG$ attains a larger market share than $\DEG$ even for large warm starts.
+For example, in Figure \ref{sim_table} we see that $\DynamicGreedy$ attains a larger market share than $\DynamicEpsGreedy$ even for large warm starts.
 We find that this also holds for $K = 3$ arms and longer time horizons, see the supplement for more details.
@@
 \includegraphics[scale=0.35]{ec19paper/figures/ts_dg_rep_diff_nih}
-\caption{Reputation scores for Needle-in-Haystack at $t=500$ (left), Reputation difference $\TS - \DG$ for Needle-in-Haystack (right).Both are smoothed using a kernel density estimate.}
-\label{rep_dist_nih}
+\caption{Reputation scores for Needle-in-Haystack at $t=500$ (left), Reputation difference $\Thompson - \DynamicGreedy$ for Needle-in-Haystack (right).Both are smoothed using a kernel density estimate.}
+\label{fig:rep_dist_nih}
 %\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
@@
 That is, consider the empirical distribution of this score over different \MRVs.%
-\footnote{Recall that each \MRV in our experimental setup comes with one specific realization table.} For concreteness, consider the Needle-in-Haystack instance at time $t=500$, plotted in Figure \ref{rep_dist_nih} (left).
+\footnote{Recall that each \MRV in our experimental setup comes with one specific realization table.} For concreteness, consider the Needle-in-Haystack instance at time $t=500$, plotted in Figure~\ref{fig:rep_dist_nih} (left).
 (The other MAB instances lead to a similar intuition.)
 
-We see that the ``naive" algorithms $\DG$ and $\DEG$ have a bi-modal reputation distribution, whereas $\TS$ does not.
-The reason is that for this MAB instance, $\DG$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
-In the former case \DG does slightly better than $\TS$, and in the latter case it does substantially worse.
+We see that the ``naive" algorithms $\DynamicGreedy$ and $\DynamicEpsGreedy$ have a bi-modal reputation distribution, whereas $\Thompson$ does not.
+The reason is that for this MAB instance, $\DynamicGreedy$ either finds the best arm and sticks to it, or gets stuck on the bad arms.
+In the former case \DynamicGreedy does slightly better than $\Thompson$, and in the latter case it does substantially worse.
 However, the mean reputation trajectory may fail to capture this complexity since it simply takes average over different \MRVs.
@@
 
-To further this intuition, consider the difference in reputation scores (\emph{reputation difference}) between \TS and \DG on a particular \MRV.
+To further this intuition, consider the difference in reputation scores (\emph{reputation difference}) between \Thompson and \DynamicGreedy on a particular \MRV.
 Let's plot the empirical distribution of the reputation difference (over the \MRVs) at a particular time point.
-Figure~\ref{ts_dg_rep_diff_nih} (right) shows such plots for several time points.
-We observe that the distribution is skewed to the right, precisely due to the fact that $\DG$ either does slightly better than $\TS$ or does substantially worse.
+Figure~\ref{fig:rep_dist_nih} (right) shows such plots for several time points.
+We observe that the distribution is skewed to the right, precisely due to the fact that $\DynamicGreedy$ either does slightly better than $\Thompson$ or does substantially worse.
 Therefore, the mean is not a good measure of the central tendency, or typical value, of this distribution.

### 2020-07-15 journal paper/content/sec-bg.tex
Minor edits.

--- 2020-07-13 journal paper/content/sec-bg.tex
@@
-We present sufficient background on the three classes of MAB algorithms listed in Section~\ref{sec:intro-model}: from adaptive-exploration to exploration-separating to greedy.
+We present self-contained background on multi-armed bandits (MAB), to make the paper accessible to non-specialists.
+We focus on the three classes of MAB algorithms listed in Section~\ref{sec:intro-model}: from adaptive-exploration to exploration-separating to greedy.
 
@@
 
-A standard performance measure is \emph{regret}, defined as the difference in the total expected reward between the algorithm and the best arm.%
-\footnote{In a formula, regret is
+A standard performance measure is \emph{regret}, defined as the difference in the total expected reward between the algorithm and the best arm. In a formula, regret is
     $T\cdot \max_{\text{arms $a$}}\mu_a
     -  \E\sbr{ \sum_{t\in[T]} r_t }$,
-where $\mu_a$ is the mean reward of arm $a$.}
-Since regret is normalized by the best arm, it allows to compare algorithms across different problem instances.
-The primary concern is the asymptotic growth rate of regret as a function of the time horizon $T$.
+where $\mu_a$ is the mean reward of arm $a$.
+Normalized by the best arm, regret allows to compare algorithms across different problem instances.
+The primary concern is the asymptotic growth rate of regret as a function of $T$.
 
-The three classes of algorithms perform very differently in terms of regret. Adaptive-exploration algorithms achieve optimal regret rates:
+The three classes of algorithms perform very differently in terms of regret: adaptive-exploration algorithms are by far the best, greedy algorithms are by far the worst, and exploration-separating ones are in the middle.
+The precise regret rates are not essential to this paper; we present them for illustration only.
+Adaptive-exploration algorithms achieve optimal regret rates:
     $\tildeO(\sqrt{KT})$
@@
 Exploration-separating algorithms completely separate exploration and exploitation.
+Each round is either selected for exploration, in which case  the distribution over arms does not depend on data, or it is assigned to exploitation, in which case the data from this round is discarded.
 The simplest approach, called \emph{Explore-First}, is to explore uniformly for a predetermined number of rounds, then choose one arm and ``exploit".
-A slightly more refined approach, called \emph{Epsilon-Greedy}, explores uniformly in each round $t$ with predetermined probability $\eps_t$, and exploits with the remaining probability.
-A formal definition is that each round is either selected exploration, in which case  the distribution over arms does not depend on data, or it is assigned to exploitation, in which case the data from this round is discarded.
+A slightly more refined approach, called \emph{Epsilon-Greedy}, explores uniformly in each round with a predetermined probability, and ``exploits" with the remaining probability.
 The $T^{2/3}$ lower bound on regret follows from a somewhat more general lower bound in \citet{MechMAB-ec09}.
 
-\xhdr{Advanced aspects.}
+\xhdr{Some advanced aspects.}
 Switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training \citep{DS-arxiv}.
@@
 \xhdr{Instantaneous regret.}
+Cumulative performance measures such as regret are not quite appropriate for our setting, as we need to characterize interactions in particular rounds.
+Instead, our theoretical results focus on \emph{Bayesian instantaneous regret} (\BIR), as defined in Section~\ref{sec:theory-prelims}.
+Recall that we posit a Bayesian prior on the \MRVs.
+In the notation of this appendix, the \BIR is simply:
+\begin{align*}%\label{eq:bg-BIR-defn}
+\BIR(t) := \E_{\text{prior}}\sbr{ \max_{a\in A} \mu_a - r_t}.
+\end{align*}
+\noindent Note that Bayesian regret (\ie regret in expectation over the prior) is precisely
+\begin{align}\label{eq:bg-BReg}
+  \BReg(T):= 
+    \E_{\text{prior}}\sbr{ T\cdot \max_{\text{arms $a$}}\mu_a - \sum_{t=1}^T r_t }
+    = \sum_{t=1}^T \BIR(t).
+\end{align}
+We are primarily interested in how fast \BIR decreases with $t$.%
+\footnote{We treat $K$, the number of arms, as a constant in what follows.}
+
+The three classes of algorithms are well-separated in terms of \BIR, much like they are in terms of regret.
+
+\begin{itemize}
+\item \DynGreedy has at least a constant \BIR for many reasonable priors (where the constant can depend on $K$ and the prior, but not on $t$).
+The reason and the proof are exactly the same as for regret.
+
+\item Exploration-separating algorithms can achieve
+    $\BIR(t) = \tildeO(t^{-1/3})$
+for all priors, \eg by using Epsilon-Greedy with exploration probability $\eps_t = t^{-1/3}$ in each round $t$. 
+The $t^{-1/3}$ rate is the best possible, in the sense that
 
@@
 
+\item Adaptive-exploration algorithms can have an even better rate
+    $\BIR(n) \sim n^{-1/3}$
+for all priors. 
+\end{itemize}
+
+Both rates are up to $\log(n)$ factors, and are the best possible \BIR rates that can be achieved for all priors.
+\ascomment{stopped here.}
+
 %\xhdr{}

### 2020-07-15 journal paper/content/sim_details.tex
Minor edits.

@@
 Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.
-We use \Thompson (\TS) from the ``adaptive-exploration" algorithms, \DynamicEpsGreedy (\DEG) with exploration probability $\eps=5\%$ from the ``exploration-separating" algorithms,%
+We use \Thompson (\TS) from the ``adaptive-exploration" algorithms, \DynamicEpsGreedy (\DEG) from the ``exploration-separating" algorithms,%
 \footnote{In each round, \DynamicEpsGreedy algorithm explores uniformly with a predetermined probability $\eps$, and ``exploits" with the remaining probability, choosing an arm with maximal posterior mean reward given the current data.
-Our pilot experiments show that choosing a different exploration probability $\eps$ does not qualitatively change the results.}  and \DynGreedy (\DG) algorithm  as in Section~\ref{sec:theory-prelims}.
+We use $\eps=5\%$ throughout.
+Our pilot experiments show that choosing a different $\eps$ does not qualitatively change the results.}  and \DynGreedy (\DG) algorithm  as in Section~\ref{sec:theory-prelims}.
 For ease of comparison, all three algorithms are parameterized with the same ``fake" Bayesian prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution.

### 2020-07-15 journal paper/ec19paper/content/inverted_u.tex
Minor edits.

@@
 In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U-expts}.%
-\footnote{\asedit{We consider the monopoly scenario for comparison only, without presenting any findings. We just assume that a monopolist chooses
+\footnote{We consider the monopoly scenario for comparison only, without presenting any findings. We just assume that a monopolist chooses
 the greedy algorithm, because it is easier to deploy in practice.
 Implicitly, users have no ``outside option": the service provided is an improvement over not having it (and therefore the monopolist is not incentivized to deploy better learning algorithms).
-This is plausible with free ad-supported platforms such as Yelp or Google.}}
+This is plausible with free ad-supported platforms such as Yelp or Google.}
 

### 2020-07-15 journal paper/ec19paper/content/perf_in_iso.tex
Minor edits.

@@
 \textit{The mean reputation trajectories \asedit{and the instantaneous reward trajectories} are arranged as predicted by prior work:
-    $\Thompson > \DynamicEpsGreedy > \DynamicGreedy$ for a sufficiently long time horizon.}
+    $\Thompson > \DynamicEpsGreedy > \DynamicGreedy$ for a sufficiently long time horizon $T$.}
 \end{finding}

### 2020-07-15 journal paper/content/sec-theory.tex
Minor edits.

@@
 Assume \SoftMaxRandom response function. Consider the \FiniteGame in which all feasible algorithms satisfy $\BReg(n)\to \infty$.%
-\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition, see
-\ascomment{insert pointer}.}
+\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition due to the standard lower bounds \cite{Lai-Robbins-85,bandits-exp3}. \ascomment{Think more.}}
 Suppose a feasible algorithm \alg BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.

### 2020-07-15 journal paper/content/sec-bg.tex
Full draft of the bg.

@@
 
-The three classes of algorithms are well-separated in terms of \BIR, much like they are in terms of regret.
-
+The three classes of algorithms are well-separated in terms of \BIR, much like they are in terms of regret. (The precise regret rates are listed below only to illustrate this point.)
 \begin{itemize}
@@
 \item Exploration-separating algorithms can achieve
-    $\BIR(t) = \tildeO(t^{-1/3})$
+    $\BIR(t) = \tildeO\rbr{t^{-1/3}}$
 for all priors, \eg by using Epsilon-Greedy with exploration probability $\eps_t = t^{-1/3}$ in each round $t$.
-The $t^{-1/3}$ rate is the best possible, in the sense that
+The $t^{-1/3}$ rate is the best possible in the worst case, in the following sense: 
+    $\BIR(t) = \tildeO\rbr{t^{-\gamma}}$ 
+for all $t$ and some $\gamma>0$,
+over all problem instances, then $\gamma\geq \nicefrac{1}{3}$.
+This is by \eqref{eq:bg-BReg} and the fact that $\BReg(T)\geq \Omega(T^{2/3})$ in the worst case.
+(The latter follows from the `frequentist" lower bound in \citet{MechMAB-ec09}.)
 
-\ascomment{stopped here.}
+\item Adaptive-exploration algorithms \emph{can} have an even better regret rate: $\BIR(t) = \tildeO\rbr{ t^{-1/2} } $, \eg this holds for Successive Elimination \citep{EvenDar-icml06}.
+Any optimal MAB algorithm such as Thompson Sampling and UCB1 enjoys this regret rate ``on average", because it satisfies
+        $\BReg(T)\leq \tildeO(sqrt{T})$.
+    Put differently, if such algorithm satisfies 
+        $\BIR(t) = \tildeO\rbr{t^{-\gamma}}$
+    for all $t$ and some $\gamma>0$, then $\gamma \leq \nicefrac{1}{2}$.
+    That said, a uniform upper bound on $\BIR$ of Thompson Sampling or UCB1 is not known, to the best of our knowledge.
 
-\item Adaptive-exploration algorithms can have an even better rate
-    $\BIR(n) \sim n^{-1/3}$
-for all priors. 
 \end{itemize}
 
-Both rates are up to $\log(n)$ factors, and are the best possible \BIR rates that can be achieved for all priors.
-\ascomment{stopped here.}
+This theoretical intuition is supported by plots in 
+Figure~\ref{relative_rep_plots} and Appendix~\ref{app:isol}.
 

### 2020-07-15 journal paper/itcs18paper/app-examples.tex
Lots of edits and comments in the "theory" appendix.

@@
 \newcommand{\PhasedExplorExploit}{\term{PhasedExplorExploit}}
-\newcommand{\RandomDynGreedy}{\term{RandomDynGreedy}}
-\newcommand{\SuccesiveElimination}{\term{SuccesiveElimination}}
+%\newcommand{\RandomDynGreedy}{\term{RandomDynGreedy}}
+%\newcommand{\SuccesiveElimination}{\term{SuccesiveElimination}}
 \newcommand{\SuccesiveEliminationReset}{\term{SuccesiveEliminationReset}}
 
-This appendix provides some pertinent background on multi-armed
-bandits (\emph{MAB}).
+\newcommand{\IReg}{R^{\term{inst}}} % instantaneous regret
+
+%This appendix provides some pertinent background on multi-armed
+%bandits (\emph{MAB}).
 We discuss \BIR and monotonicity of several MAB algorithms, touching upon: \DynGreedy and \StaticGreedy (Section~\ref{sec:MAB-greedy}), ``naive" MAB algorithms that separate exploration and exploitation (Section~\ref{sec:MAB-naive}), and ``smart" MAB algorithms that combine exploration and exploitation (Section~\ref{sec:MAB-smart}).
-
-As we do throughout the paper, we focus on MAB with i.i.d. rewards and a Bayesian prior; we call it \emph{Bayesian MAB} for brevity.
+As we do throughout the paper, we focus on MAB with i.i.d.
+rewards and a Bayesian prior; we call it \emph{Bayesian MAB} for brevity.
 
@@
 
-\subsection{\DynGreedy and \StaticGreedy}
+\ascomment{TODO: insert an intro para.}
+
+Throughout this section, we consider a standard MAB setting.
+There are $T$ rounds and $K$ arms.
+In each round $t\in [T]$, the algorithm chooses an arm $a_t\in A$ and receives a reward $r_t\in[0,1]$ for this arm, drawn from a fixed but unknown distribution.
+The set of all arms is $A$; mean reward of arm $a$ is denoted $\mu_a$.
+
+Throughout, we let
+    $\rew(t) = \mu_{a_t}$
+denote the instantaneous mean reward of the algorithm.
+
+Instantaneous regret is denoted
+    $\IReg(t) := \max_{a\in A} \mu_a - \mu_{a_t}$,
+so that 
+    $\BIR(t) = \E[\IReg(t)]$.
+
+\subsection{Monotonicity-in-information for the greedy step}
 \label{sec:MAB-greedy}
 
-We provide an example when \DynGreedy and \StaticGreedy have
-constant \BIR, and prove monotonicity of \DynGreedy. For the
-example, it suffices to consider \emph{deterministic rewards} (for
-each action $a$, the realized reward is always equal to the mean
-$\mu_a$) and \emph{independent priors} (according to the prior
-$\priorMu$, random variables $\mu_1 \LDOTS \mu_K$ are mutually
-independent) each of {\em full support}.
+\ascomment{TODO: insert an intro sentence.}
 
-%\ascomment{This lemma is just one example. Can we argue that constant \BIR is typical/reasonable?}
-%
-%
-%\begin{lemma}
-%There is a problem instance of Bayesian MAB such that \StaticGreedy
-%and \DynGreedy have (at least) a constant \BIR for all steps. The
-%problem instance has two actions, independent priors, and
-%deterministic rewards.
-%\end{lemma}
-%
-%\begin{proof}
-%To specify the problem instance, it remains to specify the marginal
-%distributions mean rewards $\mu_1$ and $\mu_2$ according to the
-%prior. We posit that $\mu_1$ is uniform on the interval $[0, 1]$,
-%and $\mu_2$ is uniform on the interval $[\tfrac13,1]$.
-%
-%Note that
-%    $\OPT := \E[\max_{a\in A} \mu_a] =(1/3)(2/3)+(2/3)(7/9)=20/27$.
-%%\ascomment{The numbers seem wrong, given the below?}
-%This holds since with
-%probability $1/3$ we have $\mu_1\leq 1/3$ in which case clearly the
-%better arm is 2 with $\E[\mu_2]=2/3$ and with probability $2/3$ we
-%have both $\mu_1,\mu_2\in[1/3,1]$ and distributed uniformly each.
-%
-%\StaticGreedy always recommends action $2$. Since
-%$\E[\mu_2]=\tfrac23$, \BIR is $2/27$ in all rounds.
-%
-%\DynGreedy recommends the first agent action $2$. If $\mu_2 > 2/3$
-%For all the following agents it recommends action $2$. Otherwise, it
-%recommends agent $2$ action $1$ and for all the following agents it
-%recommends the better action. The expectation for agents $t\geq 3$
-%is $(1/2)(5/6)+(1/2)((1/2)(1/2)+(1/2)(5/9))=49/72$ and has
-%instantaneous regret $13/216$.
-%\end{proof}
+\ascomment{Steven: NB: we need monotonicity-in-info to argue about "a little greedy goes a long way".}
 
-The following claim is immediate from the definition of the CDF
-function
-\begin{claim}
-Assume independent priors. Let $F_i$ be the CDF of the mean reward
-$\mu_i$ of action $a_i\in A$. Then, for any numbers
-$z_2>z_1>\E[\mu_2]$ we have
-    $\Pr[\text{$\mu_1\leq z_1$ and $\mu_2\geq z_2$}] = F_1(z_1)(1-F_2(z_2)) $.
-\end{claim}
+The $n$-step history is the random sequence 
+    $H_n = \rbr{ (a_t,r_t):\, t\in[n]}$.
+Realizations of $H_n$ are called \emph{realized histories}. 
+Let $\mH_n$ be the set of all possible values of $H_n$.
 
-%\begin{lemma}
-%Let $\priorMu$ be the prior of the means, and assume that it is
-%independent for each action, where for arm $i$ it has a mean $
-%\E[\mu_i]$ and CDF $F_i$. Then, for any $\E[\mu_2] < z_1<z_2$ we
-%have that with probability $F_1(z_1)(1-F_2(z_2))$ that both
-%$\mu_1\leq z_1$ and $z_2\leq \mu_2$.
-%\end{lemma}
+The \emph{Bayesian-greedy step} given an $n$-step history $h\in \mH_n$ is defined as
+\[ \DG(h) := \argmax_{a\in A} \E\sbr{ \mu_a\mid H_n = h },
+\quad\text{ties broken uniformly}.\]
 
-%\begin{corollary}
-%Any problem instance of Bayesian MAB for two actions and independent
-%priors which are full support and deterministic rewards, with
-%constant probability \DynGreedy has a constant \BIR for all steps.
-%\end{corollary}
-
-We can now draw an immediate corollary of the above claim
-
+\begin{lemma}\label{lm:MII}
+Let $h,h'$ be two realized histories such that $h$ is a prefix of $h'$. Then
+\[ \E\sbr{ \mu_{\DG(h)} } \leq \E\sbr{ \mu_{\DG(h')} }. \]
+\end{lemma}
 \begin{corollary}
-Consider any problem instance of Bayesian MAB with two actions and independent
-priors which are full support. Then:
-\begin{OneLiners}
-\item[(a)] With constant probability, \StaticGreedy  has a constant \BIR for all steps.
-\item[(b)] Assuming deterministic rewards, with
-constant probability \DynGreedy has a constant \BIR for all steps.
-\end{OneLiners}
+\DynGreedy is monotone.
 \end{corollary}
 
-\begin{remark}
-A similar result holds for  rewards which are distributed as
-Bernoulli random variables. In this case we consider accumulative
-reward of an action as a random walk, and use a high probability
-variation of the law of iterated logarithms. (Details omitted.)
-\end{remark}
+\ascomment{TODO: find a citation for Lemma~\ref{lm:MII}.}
 
-Next, we show that \DynGreedy is monotone.
+\ascomment{OLD TEXT:}
 
 \begin{lemma}\label{dgmono}
-\DynGreedy is monotone, in the sense that $\rew(n)$ is non-decreasing.
-Further, $\rew(n)$ is strictly increasing for every time step $n$ with $\Pr[a_n\neq a_{n+1}]>0$.
+\DynGreedy is monotone, in the sense that $\E[\rew(n)]$ is non-decreasing.
+\ascomment{Steven: do we ever need the second sentence here?}
+Further, $\E[\rew(n)]$ is strictly increasing for every time step $n$ with $\Pr[a_n\neq a_{n+1}]>0$.
 \end{lemma}
@@
 
-\subsection{``Naive" MAB algorithms that separate exploration and exploitation}
-\label{sec:MAB-naive}
+\subsection{Monotone MAB algorithms}
+\label{app:MAB-monotone}
+
+\ascomment{Steven: NB: this is the point of this subsection...}
+This subsection argues that ``monotonicity" is a mild assumption in our context.
+To this end, we present versions of some standard MAB algorithms that are monotone and satisfy ``standard" regret bounds.
+
+\subsubsection{Monotone, exploration-separating MAB algorithms}
+\label{app:MAB-monotone-naive}
 
@@
 
-MAB algorithm \ExplorExploit$(m)$ initially explores each action
-with $m$ agents and for the remaining $T-|A|m$ agents recommends the
-action with the highest observed average. In the explore phase it
-assigns a random permutation of the $mK$ recommendations.
+We provide two exploration-separating algorithms that are monotone and achieve $\IReg(t)=\tildeO\rbr{t^{-1/3}}$.
+(As we saw in Appendix~\ref{sec:related-classes}, this is the ``right" regret rate for exploration-separating algorithms.)
+
+\ascomment{Steven: can't we prove that Epsilon-Greedy is monotone, too?}
+
+We start with a simple \emph{Explore-First} algorithm:
 
 \begin{lemma}
-%Assume that rewards in the range $[0,1]$.
-The \ExplorExploit$(T^{2/3}\log |A|/\delta)$ algorithm has, with
-probability $1-\delta$, for any  $n\geq |A|T^{2/3}$ we have
-\BIR$(n)=O(T^{-1/3})$. In addition, \ExplorExploit$(m)$ is monotone.
+Consider an algorithm which initially explores each action $m$ times (where $m$ is a parameter), chooses an action  with the highest observed average reward, and uses it for the remaining steps.
+In the initial exploration phase, actions are randomly permuted.
+\begin{itemize}
+\item[(a)] This algorithm is monotone, for every \MRV.
+\ascomment{Steven: right?}
+
+\item[(b)] Fix $\delta>0$ and 
+    $m = T^{2/3}\log(|A|/\delta)$.
+Then, with probability $1-\delta$, it holds that  
+    $\IReg(t)=\tildeO\rbr{T^{-1/3}}$
+for each step $n> m\,|A|$.
+\ascomment{Steven: can you derive a statement on regret, too?}
+\end{itemize}
 \end{lemma}
@@
 
-We can also have a a phased version which we call
-\PhasedExplorExploit$(m_t)$, where time is partition in to phases.
-In phase $t$ we have $m_t$ agents and a random subset of $K$ explore
-the actions (each action explored by a single agent) and the other
-agents exploit. (This implies that we need that $m_t\geq K$ for all
-$t$. We also assume that $m_t$ is monotone in $t$.)
+We consider a slightly more advanced version, called \PhasedExplorExploit, which enjoys a similar regret bound uniformly over all steps.
+
+The algorithm is defined as follows.
+The time is partition in to phases $i = 1,2, \ldots $ of $m_i\geq K$ rounds each, where $m_i$ is chosen in advance and increase with $i$.
+In each phase $i$, $K$ rounds are chosen at random for exploration, and are assigned randomly to one arm each.
+The remaining rounds are for exploitation.
+\ascomment{Steven: exploit how?
+Bayesian or frequentist?
+Do we use the data from the same phase?}
 
 \begin{lemma}
-Consider the case that $K=2$ and the rewards of the actions are
-Bernoulli r.v. with parameter $\mu_i$ and $\Delta=\mu_1-\mu_2$.
-Algorithm \PhasedExplorExploit$(m_t)$ is monotone and for $m_t =
-\sqrt{t}$ it has $\BIR(n)=O(n^{-1/3}+e^{-O(\Delta^2 n^{2/3})}))$. 
+Posit $K=2$ arms and Bernoulli rewards. 
+\begin{itemize}
+\item[(a)] Algorithm \PhasedExplorExploit is monotone, for every \MRV.
+    \ascomment{Steven: right?}
+
+\item[(b)] For phase durations $m_i =\sqrt{i}$, it has 
+    $\IReg(n)=O(n^{-1/3}+e^{-O(\Delta^2 n^{2/3})}))$,
+where $\Delta = |\mu_1-\mu_2|$.
+\ascomment{Steven: can you derive $O(T^{2/3})$ regret, too?}
+\end{itemize}
 \end{lemma}
@@
 \begin{proof}
-We first show that it is monotone. Recall that $\mu_1>\mu_2$. Let
+We first show that it is monotone. Recall that $\mu_1>\mu_2$. 
+\ascomment{Steven: we only have it for prior means, \ie 
+$\E[\mu_1]>\E[\mu_2]$}
+
+Let
 $S_i=\sum_{j=1}^t r_{i,j}$ be the sum of the rewards of action $i$
@@
 
-\subsection{``Smart" MAB algorithms that combine exploration and exploitation}
-\label{sec:MAB-smart}
+\subsubsection{Monotone, adaptive-exploration MAB algorithms}
+\label{app:MAB-monotone-smart}
 
+We provide a simple modification of Successive Elimination algorithm from \citep{EvenDar-icml06} that is monotone and (still) enjoys the same bounds on regret and BIR. 
 %MAB algorithm \SuccesiveElimination holds a set of action
@@
 
-MAB algorithm \SuccesiveEliminationReset works as follows. It keeps
-a set of surviving actions $A_s\subseteq A$, where initially
-$A_s=A$. The agents are partition into phases, where each phase is a
-random permutation of the non-eliminated actions.
-% Initially, we select actions at random. Let $n_{i,t}$
-%be the number of times action $i$ has been selected up to time $t$,
-Let $\hat{\mu}_{i,t}$ be the average of the rewards of action $i$ up
+The algorithm, called \SuccesiveEliminationReset, proceeds in phases.
+In each phase $j$, the algorithm is given a set of ``surviving" arms $A_j\subseteq A$, with $A_1=A$.
+The phase lasts $|A_j|$ rounds, randomly permuted across the arms in $A_j$ (so that each arm in $A_j$ is chosen exactly once).
+
+\ascomment{Stopped here.}
+
+\ascomment{Steven: I don't understand the algorithm! It resets after each phase, right? But then, the history is just one sample per each arm? WTF?}
+
+Let $\hat{\mu}_{i,t}$ be the average reward of action $i$ up
 to phase $t$ and $\hat{\mu}^*=\max_i \hat{\mu}_{i,t}$. We eliminate
@@
 \begin{lemma}
-%
-The algorithm \SuccesiveEliminationReset, has, with probability
-$1-\delta$, \BIR$(n)=O(\log(T/\delta)/\sqrt{n/K})$.
+Consider algorithm \SuccesiveEliminationReset.
+\begin{itemize}
+\item[(a)]
+With probability $1-\delta$, $\IReg(n)=O(\log(T/\delta)/\sqrt{n/K})$.
+
+\item[(b)]
+Assuming Bernoulli rewards, the algorithm is monotone.
+\ascomment{Steven: for each \MRV, or only in expectation?}
+\ascomment{Steven: can you derive optimal regret?}
+\end{itemize}
 \end{lemma}
@@
 \begin{proof}
+\xhdr{Part (a).}
 Let the best action be $a^*=\arg\max_a \mu_a$. With probability
@@
 $6\log(T/\delta)/\sqrt{n/K}$.
-\end{proof}
 
-\begin{lemma}
-Assume that if $\mu_i\geq \mu_j$ then the rewards $r_i$
-stochastically dominates the rewards $r_j$.
-%
-Then, \SuccesiveEliminationReset is monotone
-\end{lemma}
-
-\begin{proof}
+\xhdr{Part (b).}
 Consider the first time $T$ an action is eliminated, and let

### 2020-07-15 journal paper/content/sec-theory-proofs.tex
Inserted pointers re monotonicity-in-info lemma.

@@
   choice'' is only improved with an increasing set of
-  observations. \ascomment{insert a pointer} 
+  observations, see Lemma~\ref{lm:MII}. 
 Using \eqref{eq:BIR-modification}, we conclude that

### 2020-07-15 journal paper/content/sec-theory.tex
Added a note re monotone algs in theory/prelims.
Inserted pointers re monotonicity-in-info lemma.

@@
 (We treat the number of arms as a constant.) Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:examples} for background.
+
+An algorithm is called \emph{monotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing (and therefore $\BIR(\cdot)$ is non-decreasing). We argue this is a mild assumption, see
+Appendix~\ref{app:MAB-monotone}.
 
@@
 The latter follows from a ``monotonicity-in-information" property of the ``greedy step": essentially, it can only get better with more information
-(see \ascomment{add a pointer}).
+(see Lemma~\ref{lm:MII}).
 

### 2020-07-15 journal paper/journal_main.tex
Moved some macro to the main file.
Minor edits in "theory/extensions".
Changed an appendix title.
DEG -> BEG, DG -> BG
Minor edits.
Removed "in progress" from a section title.
Revised some macro.
fix wording in abstract

@@
 
+% in connection with "a little greedy goes a long way"
+%     stuff for the "greedy modification".
+\newcommand{\gr}{\texttt{gr}}
+\newcommand{\BIRgr}{\BIR^{\gr}}
+\newcommand{\alggr}{\alg[\gr]}
+\newcommand{\rewgr}{\rew_{\gr}}
+
 \DeclareMathOperator*{\Expectation}{\mathbb{E}}
@@
 \newcommand{\rew}{\term{rew}}  % Bayesian-expected reward after n local rounds
-\newcommand{\REP}{\term{REP}} % posterior mean reward
 \newcommand{\PMR}{\term{PMR}} % posterior mean reward
@@
 \newcommand{\BIR}{\term{BIR}} % Bayesian Instantaneous Regret
-\newcommand{\regret}{R^{\term{inst}}} % regret
-\newcommand{\regretWC}{\regret_{\term{wc}}} % regret
 
@@
 \newcommand{\DynamicGreedy}{\DynGreedy}
+\newcommand{\DynamicEpsGreedy}{\term{BayesianEpsilonGreedy}}
 \newcommand{\Thompson}{\term{ThompsonSampling}}
@@
 
-We consider a stylized duopoly model in which two firms face the same instance of multi-armed bandits.
-Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen.
+We consider a stylized duopoly model in which two firms face the same \gaedit{multi-armed bandit instance.} Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen.
 We study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for consumers.
@@
 \subfile{content/sec-theory}
+\newpage
 
@@
 
-\section{Background for non-specialists: the three classes of algorithms [IN PROGRESS]}
+\section{Background for non-specialists: multi-armed bandits}
 \label{sec:related-classes}

### 2020-07-15 journal paper/ec19paper/content/revisited.tex
fix typos

@@
 \includegraphics[scale=0.35]{ec19paper/figures/ts_dg_rep_diff_nih}
-\caption{Reputation scores for Needle-in-Haystack at $t=500$ (left), Reputation difference $\Thompson - \DynamicGreedy$ for Needle-in-Haystack (right).Both are smoothed using a kernel density estimate.}
+\caption{Reputation scores for Needle-in-Haystack at $t=500$ (left), Reputation difference $\Thompson - \DynamicGreedy$ for Needle-in-Haystack (right).
+Both are smoothed using a kernel density estimate.}
 \label{fig:rep_dist_nih}

### 2020-07-15 journal paper/content/conclusion.tex
tweaks to final sentence

@@
 The second interesting extension would be to consider a setting with horizontally differentiated user preferences\footnote{This would be similar to the extension in Section \ref{sec:theory-extensions} that considers contextual bandits, though it would be interesting to study how this would change based on different user models that had context-specific user decision rules.} as such a model would be better suited for understanding how competition may induce firms to specialize.
-In general, there are many further directions to undertake in order to better understand how the intertwined relationship between competition and data generation can help us better understand the digital economy.
+In general, there are several further directions to undertake in order to better characterize the intertwined relationship between competition and data generation that can help us better understand competition in the digital economy.
 

### 2020-07-16 journal paper/content/sec-bg.tex
fix typo

@@
 Any optimal MAB algorithm such as Thompson Sampling and UCB1 enjoys this regret rate ``on average", because it satisfies
-        $\BReg(T)\leq \tildeO(sqrt{T})$.
+        $\BReg(T)\leq \tildeO(\sqrt{T})$.
     Put differently, if such algorithm satisfies 

### 2020-07-17 journal paper/journal_main.tex
Added a color-edit macro for Yishay

@@
 \addauthor{sw}{orange} % sw for Steven
+\addauthor{ym}{madenta} % sw for Steven
 % e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
@@
 
-In the appendix, we provide background on multi-armed bandits as well as several discussions and proofs omitted from the main text.
+%In the appendix, we provide background on multi-armed bandits as well as several discussions and proofs omitted from the main text.
 Furthermore, we provide plots and tables for our experiments, which were omitted from the main text.

### 2020-07-17 journal paper/content/sec-bg.tex
Another pass, full draft.

--- 2020-07-15 journal paper/content/sec-bg.tex
@@
-We present self-contained background on multi-armed bandits (MAB), to make the paper accessible to non-specialists.
-We focus on the three classes of MAB algorithms listed in Section~\ref{sec:intro-model}: from adaptive-exploration to exploration-separating to greedy.
+We present self-contained background on multi-armed bandits (MAB), to make the paper accessible to researchers who are not experts on MAB. 
+We focus on the three classes of MAB algorithms listed in Section~\ref{sec:intro-model}: adaptive-exploration algorithms, exploration-separating algorithms, and greedy algorithms.
+We list the precise upper and lower bounds on the regret rates only to illustrate how the three algorithm classes are separated from one another; the exact results are not essential for the paper.
+
+We use standard conventions: $O(f(t))$ and $\Omega(f(t))$ means at most (resp., at least) $f(n)$, up to constant factors, starting from large enough $t$.
+Likewise, $\tildeO(f(t))$ notation suppresses the $\polylog(t)$ factors.
 
@@
 The three classes of algorithms perform very differently in terms of regret: adaptive-exploration algorithms are by far the best, greedy algorithms are by far the worst, and exploration-separating ones are in the middle.
-The precise regret rates are not essential to this paper; we present them for illustration only.
 Adaptive-exploration algorithms achieve optimal regret rates:
     $\tildeO(\sqrt{KT})$
-for all problem instances, and an improved regret rate of
+for all problem instances, and simultaneously a vastly improved regret rate of
     $O(\tfrac{K}{\Delta}\log T)$
-for problem instances with $\text{gap}\geq \Delta$
+for all problem instances with $\text{gap}\geq \Delta$, without knowing the $\Delta$ in advance
 \citep{Lai-Robbins-85,bandits-ucb1,bandits-exp3}.%
 \footnote{The \emph{gap} is the difference in mean reward between the best arm and the second-best arm.
-As usual in computer science, the $O(\cdot)$ and $\tildeO(\cdot)$ notation suppresses the constant and
-$\polylog(T)$ factors, respectively.}
+It is a standard notion of difficulty for MAB instances: instances with large gap are deemed  ``easy".}
 Exploration-separating algorithms can only achieve regret $\tildeO(T^{2/3})$ across all problem instances.
-Finally, the greedy algorithm is terrible on a wide variety of problem instances, in the sense that it fails to try the best arm and therefore suffers regret linear in $T$
+They can achieve the ``gap-dependent" regret rate stated above, but \emph{only} if they know the $\Delta$ in advance, and at the cost of a terrible regret $\Omega(\Delta T)$ for some other problem instances \citep{MechMAB-ec09}.
+Finally, the greedy algorithm is terrible on a wide variety of problem instances, in the sense that with constant probability it fails to try the best arm even once, and therefore suffers regret $\Omega(T)$
 \citep[see Chapter 11.2 in][]{slivkins-MABbook}.
 
-The optimal regret rates are achieved by many adaptive-exploration  algorithms, of which the most known are
+The optimal regret rates are achieved by several adaptive-exploration  algorithms, of which the most known are
 Thompson Sampling \citep{Thompson-1933,TS-survey-FTML18},%
@@
 UCB1 \citep{bandits-ucb1},
-and Successive Elimination \citep{EvenDar-icml06}.
-Further work on refining these regret rates is not as relevant.
+and Successive Elimination \citep{EvenDar-icml06}. There is a substantial amount of follow-up work on refining these regret rates, but it is not as relevant to this paper.
 
 Exploration-separating algorithms completely separate exploration and exploitation.
-Each round is either selected for exploration, in which case  the distribution over arms does not depend on data, or it is assigned to exploitation, in which case the data from this round is discarded.
-The simplest approach, called \emph{Explore-First}, is to explore uniformly for a predetermined number of rounds, then choose one arm and ``exploit".
-A slightly more refined approach, called \emph{Epsilon-Greedy}, explores uniformly in each round with a predetermined probability, and ``exploits" with the remaining probability.
-The $T^{2/3}$ lower bound on regret follows from a somewhat more general lower bound in \citet{MechMAB-ec09}.
+Ahead of time, each round is either selected for exploration, in which case  the distribution over arms does not depend on the observed data, or it is assigned to exploitation, in which case the data from this round is discarded.
+The simplest approach, called \emph{Explore-First}, explores uniformly for a predetermined number of rounds, then chooses one arm for ``exploitation" and uses it from then on.
+A more refined approach, called \emph{Epsilon-Greedy}, explores uniformly in each round with a predetermined probability, and ``exploits" with the remaining probability.
+Both algorithms, and the associated $\tildeO(T^{2/3})$ regret bounds, have been ``folklore knowledge" for decades.
+The general definition of ``exploration-separating algorithms" and the associated lower bounds trace back to \citet{MechMAB-ec09}.%
+\footnote{\citet{MechMAB-ec09} considers a closely related, but technically different setting, which can be easily ``translated" into ours (either as a corollary or as another application of the same proof technique).
+Detailing this is beyond the scope of this appendix.}
 
-\xhdr{Some advanced aspects.}
+% add: this is all more general
+
+\xhdr{Advanced aspects.}
 Switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training \citep{DS-arxiv}.
-Inserting exploration in a complex decision pipeline necessitates a substantial awareness of the technology and a certain change in a mindset, as well as an infrastructure to collect and analyze the data.
+Inserting exploration into a complex decision-making pipeline necessitates a substantial awareness of the technology and a certain change in a mindset for the personnel, as well as an infrastructure to collect and analyze the data.
 Adaptive exploration requires the said infrastructure to propagate the data analysis back to the ``front-end" where the decisions are made, and do it on a sufficiently fast and regular cadence.
-Framing the problem in terms of the appropriate modeling assumptions and action features, and debugging the machine learning code, tend to be quite subtle, too.
+Framing the problem (in terms of the appropriate modeling assumptions and action features) and debugging the machine learning code tend to be quite subtle, too.
+
+The lower bounds mentioned above are fairly typical, in the sense that they hold for a wide variety of problem instances, even though they are usually (and most cleanly) presented as worst-case.
+The $\Omega(\sqrt{T})$ lower bound from \citet{bandits-exp3} can be extended to hold for most problem instances, in the following sense: for each instance $\mI$ there exists a ``decoy instance" $\mI'$ such that any algorithm incurs regret $\Omega(\sqrt{T})$ on at least one of them.
+The ``gap-dependent" lower bound of
+ $\Omega(\tfrac{K}{\Delta}\log T)$
+in fact holds for all problem instances and all algorithms that are not \emph{terrible} on the large-gap instances \citep{Lai-Robbins-85}.
+The $\Omega(T^{2/3})$ lower bound for exploration-separating algorithms in fact applies to all problem instances, as long as the algorithm achieves $\tildeO(T^{2/3})$ regret rate in the worst case \citep{MechMAB-ec09}.
+Moreover, there is a tradeoff between the worst-case upper bound on the regret rate and a lower bound that applies for all problem instances \citep[Theorem 4.3 in][using the same proof technique]{MechMAB-ec09}.
+
+The intuition on (the separation between) the three algorithm classes applies more generally, far beyond the basic MAB model discussed above.
+In particular, all algorithms that we explicitly mentioned are in fact general algorithmic techniques that are known to extend to a variety of more general MAB scenarios, typically with a similarly stark separation in regret bounds.
 
@@
 \end{align}
-We are primarily interested in how fast \BIR decreases with $t$.%
-\footnote{We treat $K$, the number of arms, as a constant in what follows.}
+We are primarily interested in how fast \BIR decreases with $t$, treating $K$ as a constant.
 
-The three classes of algorithms are well-separated in terms of \BIR, much like they are in terms of regret. (The precise regret rates are listed below only to illustrate this point.)
+The three classes of algorithms are well-separated in terms of \BIR, much like they are in terms of regret. 
+%(The precise regret rates are listed below only to illustrate this point.)
 \begin{itemize}
@@
 
-\item Adaptive-exploration algorithms \emph{can} have an even better regret rate: $\BIR(t) = \tildeO\rbr{ t^{-1/2} } $, \eg this holds for Successive Elimination \citep{EvenDar-icml06}.
-Any optimal MAB algorithm such as Thompson Sampling and UCB1 enjoys this regret rate ``on average", because it satisfies
-        $\BReg(T)\leq \tildeO(sqrt{T})$.
-    Put differently, if such algorithm satisfies 
+\item Adaptive-exploration algorithms \emph{can} have an even better regret rate: $\BIR(t) = \tildeO\rbr{ t^{-1/2} } $.
+This holds for Successive Elimination \citep{EvenDar-icml06} and for Thompson Sampling (see Appendix~\ref{app:MAB-monotone-smart}).%
+    \footnote{However, such result is not known for UCB1 algorithm, to the best of our knowledge.}
+    Any optimal MAB algorithm enjoys this regret rate ``on average" by \eqref{eq:bg-BReg}, since 
+        $\BReg(T)\leq \tildeO(\sqrt{T})$.
+    In particular, if such algorithm satisfies
         $\BIR(t) = \tildeO\rbr{t^{-\gamma}}$
     for all $t$ and some $\gamma>0$, then $\gamma \leq \nicefrac{1}{2}$.
-    That said, a uniform upper bound on $\BIR$ of Thompson Sampling or UCB1 is not known, to the best of our knowledge.
 
@@
 
-This theoretical intuition is supported by plots in 
+This theoretical intuition is supported by our numerical simulations: see 
 Figure~\ref{relative_rep_plots} and Appendix~\ref{app:isol}.

### 2020-07-17 journal paper/content/sec-intro.tex
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.

@@
 
-While we build on a vast and active research area, we present sufficient background in Appendix~\ref{sec:related-classes} so as to keep this paper accessible to non-specialists.
+While we build on a vast and active research area, we present sufficient background in Appendix~\ref{app:bg} so as to keep this paper accessible to non-specialists.
 

### 2020-07-17 journal paper/content/sec-theory.tex
fix typos
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.

@@
 We are primarily interested in how fast \BIR decreases with $n$.
-(We treat the number of arms as a constant.) Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:examples} for background.
+(We treat the number of arms as a constant.) Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:bg} for background.
 
 An algorithm is called \emph{monotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing (and therefore $\BIR(\cdot)$ is non-decreasing). We argue this is a mild assumption, see
-Appendix~\ref{app:MAB-monotone}.
+Appendix~\ref{app:MAB-mono}.
 
@@
 \begin{corollary}\label{cor:DG-dominance}
-The competition game has a unique Nash equilibirium: both principals choose \DynGreedy.
+The competition game has a unique Nash equilibrium: both principals choose \DynGreedy.
 \asedit{This holds even if the principals can only choose from restricted strategy sets which include \DynGreedy.}

### 2020-07-17 journal paper/content/sim_details.tex
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.

@@
 
-\xhdr{MAB algorithms.} We consider the three classes of bandit algorithms discussed in Section~\ref{sec:intro-model}, and look for qualitative differences between them.
+\xhdr{MAB algorithms.} We consider the three classes of bandit algorithms discussed in Section~\ref{sec:intro-model} (and in more detail in Appendix~\ref{app:bg}), and look for qualitative differences between them.
 We take a representative algorithm from each class.
@@
 Recall that Beta priors with 0-1 rewards form a conjugate family, which allows for simple posterior updates.
-Self-contained background can be found in Appendix~\ref{sec:related-classes}.
+
+%Self-contained background can be found in Appendix~\ref{sec:related-classes}.
 

### 2020-07-17 journal paper/itcs18paper/app-examples.tex
Minor edits.
Added comments.
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.

@@
 
-\subsection{Monotonicity-in-information for the greedy step}
-\label{sec:MAB-greedy}
+\subsection{Monotone algorithms}
+\label{app:MAB-mono}
 
-\ascomment{TODO: insert an intro sentence.}
+%This subsection argues that ``monotonicity" is a mild assumption in our context.
+To this end, we present versions of some standard MAB algorithms that are monotone and satisfy ``standard" regret bounds.
+
+\ascomment{TODO: insert an intro sentence, cite the ec15 paper.}
 
@@
 \end{corollary}
-
-\ascomment{TODO: find a citation for Lemma~\ref{lm:MII}.}
 
@@
 
-\subsection{Monotone MAB algorithms}
-\label{app:MAB-monotone}
-
-\ascomment{Steven: NB: this is the point of this subsection...}
-This subsection argues that ``monotonicity" is a mild assumption in our context.
-To this end, we present versions of some standard MAB algorithms that are monotone and satisfy ``standard" regret bounds.
+%\subsection{Monotone MAB algorithms}
+%\label{app:MAB-monotone}
 
@@
 \begin{itemize}
-\item[(a)] This algorithm is monotone, for every \MRV.
-\ascomment{Steven: right?}
+\item[(a)] This algorithm is monotone.
+\ascomment{Steven: for every \MRV or only in expectation?}
 
@@
 for each step $n> m\,|A|$.
-\ascomment{Steven: can you derive a statement on regret, too?}
+\ascomment{Steven: can you derive $O(T^{2/3})$ regret, too?}
 \end{itemize}
@@
 \begin{itemize}
-\item[(a)] Algorithm \PhasedExplorExploit is monotone, for every \MRV.
-    \ascomment{Steven: right?}
+\item[(a)] Algorithm \PhasedExplorExploit is monotone.
+\ascomment{Steven: for every \MRV or only in expectation?}
 

### 2020-07-17 journal paper/content/sec-theory-proofs.tex
Changed "monotone" to \bmonotone.

@@
     =\PMR_1(t_0) > \PMR_2(t_0)= \PMR_2(t), \]
-where the first inequality holds because \alg[1] is monotone, and the second one is the base case.
+where the first inequality holds because \alg[1] is \bmonotone, and the second one is the base case.
 \end{proof}
@@
 \end{align}
-So, agent $n_0$ chooses principal $1$. By Lemma~\ref{lm:DG-sudden} {(noting that \DynGreedy is monotone)}, all subsequent agents choose principal $1$, too.
+So, agent $n_0$ chooses principal $1$. By Lemma~\ref{lm:DG-sudden} {(noting that \DynGreedy is \bmonotone)}, all subsequent agents choose principal $1$, too.
 \end{proof}
@@
   $\rew_1(\cdot)$ and $\rewgr(\cdot)$ are non-decreasing: the former
-  because \alg[1] is monotone and the latter because the ``greedy
+  because \alg[1] is \bmonotone and the latter because the ``greedy
   choice'' is only improved with an increasing set of
@@
 Using \eqref{eq:BIR-modification}, we conclude that
-the greedy modification \alg[2] is monotone.
+the greedy modification \alg[2] is \bmonotone.
 
@@
 \end{align}
-Here inequality \eqref{eq:pf:thm:random-greedy-1} holds because $\rew_1(\cdot)$ is monotone and $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$, and inequality \eqref{eq:pf:thm:random-greedy-2} holds because $\rew_1(n_0)<\rew_2(n_0)$ and $\mM_{2,t}(n_0)>0$.%
+Here inequality \eqref{eq:pf:thm:random-greedy-1} holds because $\rew_1(\cdot)$ is \bmonotone and $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$, and inequality \eqref{eq:pf:thm:random-greedy-2} holds because $\rew_1(n_0)<\rew_2(n_0)$ and $\mM_{2,t}(n_0)>0$.%
 \footnote{If $\rew_1(\cdot)$ is strictly increasing, then inequality \eqref{eq:pf:thm:random-greedy-1} is strict, too; this is because $\mM_{2,t}(t-1)>\posteriorN{1}{t}(t-1)$.  }

### 2020-07-17 journal paper/content/sec-theory.tex
Changed "monotone" to \bmonotone.

@@
 
-An algorithm is called \emph{monotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing (and therefore $\BIR(\cdot)$ is non-decreasing). We argue this is a mild assumption, see
+An algorithm is called \emph{\bmonotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing (and therefore $\BIR(\cdot)$ is non-decreasing). We argue this is a mild assumption, see
 Appendix~\ref{app:MAB-mono}.
@@
 Consider \HardMax response function with $\respF(0)\geq\tfrac12$.
-Suppose \alg[1] is monotone, and $\PMR_1(t_0)>\PMR_2(t_0)$ for some global round $t_0$. Then $\PMR_1(t)>\PMR_2(t)$ for all subsequent rounds $t$.
+Suppose \alg[1] is \bmonotone, and $\PMR_1(t_0)>\PMR_2(t_0)$ for some global round $t_0$. Then $\PMR_1(t)>\PMR_2(t)$ for all subsequent rounds $t$.
 \end{lemma}
@@
 Agent $1$ chooses principal $1$ because of the tie-breaking rule.
-Since \StaticGreedy is trivially monotone, all the subsequent agents choose principal $1$ by an induction argument similar to the one in the proof of Lemma~\ref{lm:DG-sudden}.
+Since \StaticGreedy is trivially \bmonotone, all the subsequent agents choose principal $1$ by an induction argument similar to the one in the proof of Lemma~\ref{lm:DG-sudden}.
 \end{proof}
@@
 Specifically, principal 1 can get all but the first few agents, under a mild technical assumption that \DynGreedy deviates from \StaticGreedy.
-Principal 1 can use \DynGreedy, or any other monotone MAB algorithm that coincides with \DynGreedy in the first few steps.
+Principal 1 can use \DynGreedy, or any other \bmonotone MAB algorithm that coincides with \DynGreedy in the first few steps.
 
@@
 Assume the prior $\mP$ is such that \DynGreedy deviates from \StaticGreedy starting from some step $n_0$.
-Suppose that principal $1$ runs a monotone MAB algorithm that coincides with \DynGreedy in the first $n_0$ steps.
+Suppose that principal $1$ runs a \bmonotone MAB algorithm that coincides with \DynGreedy in the first $n_0$ steps.
 Then all agents $t\geq n_0$ choose principal $1$.
@@
 Assume \HardMaxRandom response function.
-Suppose algorithms \alg[1], \alg[2] are monotone and $T$-oblivious, and \eqref{eq:random-assn} holds.
+Suppose algorithms \alg[1], \alg[2] are \bmonotone and $T$-oblivious, and \eqref{eq:random-assn} holds.
 If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible).
@@
 
-%Let \alg be a monotone MAB algorithm and $\mA$ be a finite set of monotone MAB algorithms such that each algorithm in $\mA$ satisfies \eqref{eq:random-assn} and is ``dominated" by \alg in the sense of \eqref{eq:random-better-messy}.
+%Let \alg be a \bmonotone MAB algorithm and $\mA$ be a finite set of \bmonotone MAB algorithms such that each algorithm in $\mA$ satisfies \eqref{eq:random-assn} and is ``dominated" by \alg in the sense of \eqref{eq:random-better-messy}.
 %Assume principals can only choose algorithms from $\mA\cup \{\alg\}$.
@@
 All feasible algorithms are assumed to be
-monotone and $T$-oblivious, and to satisfy \eqref{eq:random-assn}.
+\bmonotone and $T$-oblivious, and to satisfy \eqref{eq:random-assn}.
 \end{definition}
@@
 \xhdr{Counterpoint: A little greedy goes a long way.}
-Given any monotone MAB algorithm \alg other than \DynGreedy, we design a modified algorithm which ``mixes in" some greedy choices (and consequently learns at a slower rate), yet prevails over \alg in the competition game.
+Given any \bmonotone MAB algorithm \alg other than \DynGreedy, we design a modified algorithm which ``mixes in" some greedy choices (and consequently learns at a slower rate), yet prevails over \alg in the competition game.
 Thus, we have a counterpoint to \eqref{eq:theory-HMR-informal}: even under \HardMaxRandom, a better algorithm may lose in competition.
@@
 Consider a symmetric \HardMaxRandom response function $\respF$ with baseline probability $\eps_0 = \respF(-1)$.
-Suppose \alg[1] is monotone, and deviates from \DynGreedy starting from some step $n_0$.
+Suppose \alg[1] is \bmonotone, and deviates from \DynGreedy starting from some step $n_0$.
 Let \alg[2] be the greedy modification of \alg[1] with mixing parameter $p>0$ such that
@@
 
-Second, let us argue that \alg[2] is monotone.
-This follows from \eqref{eq:BIR-modification} since both \alg[1] and \alggr are monotone.
+Second, let us argue that \alg[2] is \bmonotone.
+This follows from \eqref{eq:BIR-modification} since both \alg[1] and \alggr are \bmonotone.
 The latter follows from a ``monotonicity-in-information" property of the ``greedy step": essentially, it can only get better with more information
@@
 
-Thus, the greedy modification is beneficial in competition \emph{and} keeps us inside the set of monotone MAB algorithms.
-Consequently, if the principals are restricted to choosing monotone MAB algorithms, then non-greedy algorithms cannot be chosen in equilibrium.
+Thus, the greedy modification is beneficial in competition \emph{and} keeps us inside the set of \bmonotone MAB algorithms.
+Consequently, if the principals are restricted to choosing \bmonotone MAB algorithms, then non-greedy algorithms cannot be chosen in equilibrium.
 
@@
 Posit a symmetric \HardMaxRandom response function.
-Suppose both principals can only choose from monotone MAB algorithms.
+Suppose both principals can only choose from \bmonotone MAB algorithms.
 Then for any time horizon $T$, the only possible \emph{pure} Nash equilibrium is one where both principals choose \DynGreedy.
 Moreover, no pure Nash equilibrium exists when some algorithm BIR-dominates \DynGreedy and $T$ is sufficiently large.
-\asedit{These conclusions hold even if each principal's strategy set is a subset of monotone MAB algorithms that is closed under the greedy modification.}%
-\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) monotone MAB algorithms, and contains the greedy modification for each algorithms in this set and each mixing parameter $p$.}
+\asedit{These conclusions hold even if each principal's strategy set is a subset of \bmonotone MAB algorithms that is closed under the greedy modification.}%
+\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) \bmonotone MAB algorithms, and contains the greedy modification for each algorithms in this set and each mixing parameter $p$.}
 \end{corollary}
@@
 %
-%Suppose principals can only choose algorithms from $\mA\cup \{\alg\}$, where $\mA\cup \{\alg\}$ is a finite set of monotone MAB algorithms such that each algorithm in $\mA$ satisfies \eqref{eq:random-assn} and $\BReg(n)\to \infty$, and is ``dominated" by \alg in the sense of \eqref{eq:random-better-messy}.
+%Suppose principals can only choose algorithms from $\mA\cup \{\alg\}$, where $\mA\cup \{\alg\}$ is a finite set of \bmonotone MAB algorithms such that each algorithm in $\mA$ satisfies \eqref{eq:random-assn} and $\BReg(n)\to \infty$, and is ``dominated" by \alg in the sense of \eqref{eq:random-better-messy}.
 %
@@
 \begin{theorem}\label{thm:SoftMax-strong}
-Assume the \SoftMaxRandom response function. Suppose algorithms \alg[1], \alg[2] are monotone and $T$-oblivious, and \alg[1] weakly-BIR-dominates \alg[2]. Posit mild technical assumptions:
+Assume the \SoftMaxRandom response function. Suppose algorithms \alg[1], \alg[2] are \bmonotone and $T$-oblivious, and \alg[1] weakly-BIR-dominates \alg[2]. Posit mild technical assumptions:
   $\BIR_1(n) \to 0$ and that $\BIR_2$ cannot be extremely small, namely:
@@
 Assume \SoftMaxRandom response function. Consider the \FiniteGame in which all feasible algorithms satisfy $\BReg(n)\to \infty$.%
-\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition due to the standard lower bounds \cite{Lai-Robbins-85,bandits-exp3}. \ascomment{Think more.}}
+\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition, see Appendix~\ref{app:bg} for background.} 
 Suppose a feasible algorithm \alg BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.

### 2020-07-17 journal paper/content/sec-bg.tex
Minor edits.

@@
     $\BIR(t) = \tildeO\rbr{t^{-1/3}}$
-for all priors, \eg by using Epsilon-Greedy with exploration probability $\eps_t = t^{-1/3}$ in each round $t$.
-The $t^{-1/3}$ rate is the best possible in the worst case, in the following sense:
+for all priors, \eg by using Epsilon-Greedy algorithm with exploration probability $\eps_t = t^{-1/3}$ in each round $t$.
+In the typical scenario when
+ $\BReg(t)\geq \Omega(t^{2/3})$,
+the \BIR rate of $t^{-1/3}$  cannot be improved, in the following sense:
+if
     $\BIR(t) = \tildeO\rbr{t^{-\gamma}}$
-for all $t$ and some $\gamma>0$,
-over all problem instances, then $\gamma\geq \nicefrac{1}{3}$.
-This is by \eqref{eq:bg-BReg} and the fact that $\BReg(T)\geq \Omega(T^{2/3})$ in the worst case.
-(The latter follows from the `frequentist" lower bound in \citet{MechMAB-ec09}.)
+for all $t$, then $\gamma\geq \nicefrac{1}{3}$.
 
 \item Adaptive-exploration algorithms \emph{can} have an even better regret rate: $\BIR(t) = \tildeO\rbr{ t^{-1/2} } $.
-This holds for Successive Elimination \citep{EvenDar-icml06} and for Thompson Sampling (see Appendix~\ref{app:MAB-monotone-smart}).%
+This holds for Successive Elimination \citep{EvenDar-icml06} and for Thompson Sampling (see Appendix~\ref{app:examples}).%
     \footnote{However, such result is not known for UCB1 algorithm, to the best of our knowledge.}
@@
         $\BIR(t) = \tildeO\rbr{t^{-\gamma}}$
-    for all $t$ and some $\gamma>0$, then $\gamma \leq \nicefrac{1}{2}$.
+    for all $t$, then $\gamma \leq \nicefrac{1}{2}$.
 

### 2020-07-17 journal paper/itcs18paper/app-examples.tex
Full draft!

@@
 
-\ascomment{TODO: insert an intro para.}
+This appendix provides some auxiliary results on \bmonotonicity of some standard algorithms: \DynGreedy, \DynamicEpsGreedy and Thompson Sampling.
+The former is needed in Section~\ref{sec:theory}, the last two merely add motivation for our theoretical story.
+The result on Thompson Sampling also implies that $\BIR(t)\leq \tildeO(t^{-1/2})$.
+Recall that an algorithm is called \bmonotone if its Bayesian-expected reward is non-decreasing in time.
 
-Throughout this section, we consider a standard MAB setting.
+Throughout, we consider a standard MAB setting.
 There are $T$ rounds and $K$ arms.
@@
 
-Throughout, we let
+We let
     $\rew(t) = \mu_{a_t}$
@@
 
-\subsection{Monotone algorithms}
-\label{app:MAB-mono}
+%\subsection{Monotone algorithms}
+%\label{app:MAB-mono}
 
@@
 
-\ascomment{TODO: insert an intro sentence, cite the ec15 paper.}
+\xhdr{Monotonicity for the greedy algorithm.}
+We state the monotonicity-in-information result for the ``Bayesian-greedy step": informally, exploitation can only get better with more data.
+We invoke this result directly in Section~\ref{sec:theory}, and use it to derive monotonicity of \DynGreedy and \DynamicEpsGreedy.
 
-\ascomment{Steven: NB: we need monotonicity-in-info to argue about "a little greedy goes a long way".}
-
-The $n$-step history is the random sequence
+A formal statement needs some scaffolding. The $n$-step history is the random sequence
     $H_n = \rbr{ (a_t,r_t):\, t\in[n]}$.
 Realizations of $H_n$ are called \emph{realized histories}.
-Let $\mH_n$ be the set of all possible values of $H_n$.
+Let $\mH_n$ be the set of all possible values of $H_n$. The \emph{Bayesian-greedy step} given an $n$-step history $h\in \mH_n$ is defined as
+\[ \DG(h) := \argmax_{a\in A} \E\sbr{ \mu_a\mid H_n = h },
+\quad\text{ties broken arbitrarily}.\]
+(However, recall that such ties are ruled out by Assumption~\ref{eq:assn-distinct}.) Now, the result is as follows:
 
-The \emph{Bayesian-greedy step} given an $n$-step history $h\in \mH_n$ is defined as
-\[ \DG(h) := \argmax_{a\in A} \E\sbr{ \mu_a\mid H_n = h },
-\quad\text{ties broken uniformly}.\]
-
-\begin{lemma}\label{lm:MII}
+\begin{lemma}[\citet{ICexplorationGames-ec16}]\label{lm:MII}
 Let $h,h'$ be two realized histories such that $h$ is a prefix of $h'$. Then
@@
 \end{lemma}
-\begin{corollary}
-\DynGreedy is monotone.
+
+\begin{corollary}\label{dgmono}
+\DynGreedy is \bmonotone. Moreover,
+$\E[\rew(n)]$ strictly increases in each time step $n$ with $\Pr[a_n\neq a_{n+1}]>0$.
 \end{corollary}
 
-\ascomment{OLD TEXT:}
-
-\begin{lemma}\label{dgmono}
-\DynGreedy is monotone, in the sense that $\E[\rew(n)]$ is non-decreasing.
-\ascomment{Steven: do we ever need the second sentence here?}
-Further, $\E[\rew(n)]$ is strictly increasing for every time step $n$ with $\Pr[a_n\neq a_{n+1}]>0$.
-\end{lemma}
-
 \begin{proof}
-We prove by induction on $n$ that $\rew(n)\leq \rew(n+1)$ for
-\DynGreedy. Let $a_n$ be the random variable recommended at time
-$t$, then $\E[\mu_{a_n}| \mI_n ]=\rew(n)$. We can rewrite this as:
-\[
-\rew(n)=\E_{\mI_n}[\E_{r_n}[\mu_{a_n}|r_n,\mI_n]] =
-\E_{\mI_{n+1}}[\mu_{a_n}|\mI_{n+1}]
-\]
-since $\mI_{n+1}=(\mI_n,r_n)$. At time $n+1$ \DynGreedy will select
-an action $a_{n+1}$ such that:
-\[
-\rew(n+1)=\E[\mu_{a_{n+1}}|\mI_{n+1}]\geq \E[\mu_{a_n}
-|\mI_n]=\rew(n)
-\]
-%
-%Now after executing action $a_n$ and observing its reward $r_n$
-%there are two possibilities. If $a_n$ is still the best action, then
-%its expected reward given the information at time $n$ has not
-%changed. On the other hand, if $a_n$ is not the best action anymore,
-%then it implies that we recommend a better action. In such a case,
-%we increase the expected reward. \ymcomment{need to make this
-%formal}
-which proves the monotonicity. In cases that $\Pr[a_n\neq
-a_{n+1}]>0]$ we have a strict inequality, since with some
-probability we select a better action then the realization of $a_n$.
+Bayesian-monotonicity follows directly. The ``strictly increases" statement holds because the arm chosen in a given round has a strictly largest Bayesian-expected reward for that round.
 \end{proof}
 
-%%%%%%%%%%%%%%%%%%%%%%%
+\xhdr{Monotonicity for Epsilon-Greedy.} Lemma~\ref{lm:MII} immediately implies monotonicity of \DynamicEpsGreedy, for a generic choice of exploration probabilities.
+Recall that in each round $t$, \DynamicEpsGreedy algorithm explores uniformly with a predetermined probability $\eps_t$, and ``exploits" with the remaining probability using the Bayesian-greedy step:
+    $a_t = \DG(\text{current data})$.
 
-%\subsection{Monotone MAB algorithms}
-%\label{app:MAB-monotone}
+\begin{corollary}\label{cor:mono-eps-greedy}
+\DynamicEpsGreedy is \bmonotone whenever probabilities $\eps_t$ are non-increasing.
+\end{corollary}
 
-\subsubsection{Monotone, exploration-separating MAB algorithms}
-\label{app:MAB-monotone-naive}
+\xhdr{Monotonicity and \BIR of Thompson Sampling.} 
+Monotonicity follows from \citet{Selke-BE-2020}; the \BIR result is an immediate corollary given that
+    $\BReg(t) \leq \tilde{O}(\sqrt{t})$
+for all steps $t$ \citep{Russo-MathOR-14}. Neither result has appeared in print, to the best of our knowledge.
 
-%\ascomment{Nicer to have $\BIR(n)\leq n^{-1/3}$.
-%Explore-then-exploit does not have that! So, let's use phases of
-%exponentially increasing duration.}
+\begin{theorem}[\citet{Selke-BE-2020}]\label{thm:TS-mono}
+Thompson Sampling is \bmonotone.
+\end{theorem}
 
-We provide two exploration-separating algorithms that are monotone and achieve $\IReg(t)=\tildeO\rbr{t^{-1/3}}$.
-(As we saw in Appendix~\ref{sec:related-classes}, this is the ``right" regret rate for exploration-separating algorithms.)
+\begin{corollary}
+Thompson Sampling satisfies $\BIR(t)\leq \tildeO(t^{-1/2})$.
+\end{corollary}
 
-\ascomment{Steven: can't we prove that Epsilon-Greedy is monotone, too?}
+The theorem follows easily from two observations in  \citet{Selke-BE-2020}, we provide the proof for completeness.
+First, let us state the algorithm.
+Let $\Pr^t$ and $\E^t$ denote, resp., Bayesian posterior probability and expectation conditional on the history observed at time $t$.
+Let $a^*$ be the best arm: $a^*\in \argmax_a \mu_a$.
+In each round $t$, Thompson Sampling samples arm $a_t$ as follows:
+\begin{align}\label{eq:TS-defn}
+    \Pr[a_t=a] := {\Pr}^t[a^*=a]
+    \quad\text{for each arm $a$}.
+\end{align}
 
-We start with a simple \emph{Explore-First} algorithm:
+\begin{proof}[Proof of Theorem~\ref{thm:TS-mono}]
+The two observations from \citet{Selke-BE-2020}  are as follows: for each arm $a$,
+\begin{OneLiners}
+\item[(i)]
+$\E\sbr{ \mu_a \mid a_t = a} \;\Pr[a_t=a] = \E\sbr{H_{t,a}}$,
+where
+    $H_{t,a} := \E^t[\mu_a]\; \Pr^t[a_t = a]$.
 
-\begin{lemma}
-Consider an algorithm which initially explores each action $m$ times (where $m$ is a parameter), chooses an action  with the highest observed average reward, and uses it for the remaining steps.
-In the initial exploration phase, actions are randomly permuted.
-\begin{itemize}
-\item[(a)] This algorithm is monotone.
-\ascomment{Steven: for every \MRV or only in expectation?}
+\item[(ii)]
+The process $(H_{t,a}:\;t\in\N)$ 
+is a submartingale.
+\end{OneLiners}
 
-\item[(b)] Fix $\delta>0$ and
-    $m = T^{2/3}\log(|A|/\delta)$.
-Then, with probability $1-\delta$, it holds that
-    $\IReg(t)=\tildeO\rbr{T^{-1/3}}$
-for each step $n> m\,|A|$.
-\ascomment{Steven: can you derive $O(T^{2/3})$ regret, too?}
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-In the explore phase we we approximate for each action $a\in A$, the
-value of $\mu_a$ by $\hat{\mu}_a$. Using the standard Chernoff
-bounds we have that with probability $1-\delta$, for every action
-$a\in A$ we have $|\mu_a -\hat{\mu}_a| \leq T^{-1/3}$.
-
-Let $a^* = \arg\max_a \mu_a$ and $a^{ee}$ the action that
-\ExplorExploit selects in the explore phase after the first
-$|A|T^{2/3}$ agents. Since $\hat{\mu}_{a^*} \leq
-\hat{\mu}_{a^{ee}}$, this implies that $\mu_{a^*} -
-\mu_{a^{ee}}=O(T^{-1/3})$.
-
-To show that \ExplorExploit$(m)$ is monotone, we need to show only
-that $\rew(mK) \leq \rew(mK+1)$. This follows since for any $t< mK$
-we have $\rew(t)=\rew(t+1)$, since the recommended action is
-uniformly distributed for each time $t$. Also, for any $t\geq mK+1$
-we have $\rew(t)=\rew(t+1)$ since we are recommending the same
-exploration action. The proof that $\rew(mK) \leq \rew(mK+1)$ is the
-same as for \DynGreedy in Lemma~\ref{dgmono}.
+We use these observations as follows:
+\begin{align*}
+\E[\rew(t)]
+    &= \sum_{\text{arms $a$}} 
+        \E\sbr{ \mu_a \mid a_t = a} \;\Pr[a_t=a] 
+    =\sum_{\text{arms $a$}}  \E\sbr{ H_{t,a} }
+    \qquad\EqComment{by (i)}\\
+\E[\rew(t+1)]
+    &= \sum_{\text{arms $a$}}  \E\sbr{ H_{t+1,\;a} }
+    \geq \sum_{\text{arms $a$}}  \E\sbr{ H_{t,a} }
+    =\E[\rew(t)]
+    \qquad\EqComment{by (ii)} \qedhere
+\end{align*}
 \end{proof}
-
-We consider a slightly more advanced version, called \PhasedExplorExploit, which enjoys a similar regret bound uniformly over all steps.
-
-The algorithm is defined as follows.
-The time is partition in to phases $i = 1,2, \ldots $ of $m_i\geq K$ rounds each, where $m_i$ is chosen in advance and increase with $i$.
-In each phase $i$, $K$ rounds are chosen at random for exploration, and are assigned randomly to one arm each.
-The remaining rounds are for exploitation.
-\ascomment{Steven: exploit how?
-Bayesian or frequentist?
-Do we use the data from the same phase?}
-
-\begin{lemma}
-Posit $K=2$ arms and Bernoulli rewards.
-\begin{itemize}
-\item[(a)] Algorithm \PhasedExplorExploit is monotone.
-\ascomment{Steven: for every \MRV or only in expectation?}
-
-\item[(b)] For phase durations $m_i =\sqrt{i}$, it has
-    $\IReg(n)=O(n^{-1/3}+e^{-O(\Delta^2 n^{2/3})}))$,
-where $\Delta = |\mu_1-\mu_2|$.
-\ascomment{Steven: can you derive $O(T^{2/3})$ regret, too?}
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-We first show that it is monotone. Recall that $\mu_1>\mu_2$.
-\ascomment{Steven: we only have it for prior means, \ie
-$\E[\mu_1]>\E[\mu_2]$}
-
-Let
-$S_i=\sum_{j=1}^t r_{i,j}$ be the sum of the rewards of action $i$
-up to phase $t$. We need to show that $\Pr[S_1>S_2]+ (1/2)
-\Pr[S_1=S_2]$ is monotonically increasing in $t$. Consider the
-random variable $Z=S_1-S_2$. At each phase it increases by $+1$ with
-probability $\mu_1(1-\mu_2)$, decreases by $-1$ with probability
-$(1-\mu_1)\mu_2$ and otherwise does not change.
-
-Consider the values of $Z$ up to phase $t$. We really care only
-about the probability that is shifted from positive to negative and
-vice versa.
-
-First, consider the probability that $Z=0$. We can partition it to
-$S_1=S_2=r$ events, and let $p(r,r)$ be the probability of this
-event. For each such event, we have $p(r,r)\mu_1$ moved to $Z=+1$
-and $p(r,r)\mu_2$ moved to $Z=-1$. Since $\mu_1>\mu_2$ we have that
-$p(r,r)\mu_1\geq p(r,r)\mu_2$ (note that $p(r,r)$ might be zero, so
-we do not have a strict inequality).
-
-Second, consider the probability that $Z=+1$ or $Z=-1$. We can
-partition it to $S_1=r+1;S_2=r$ and $S_1=r;S_2=r+1$ events, and let
-$p(r+1,r)$ and $p(r,r+1)$ be the probabilities of those events.
-%
-It is not hard to see that $p(r+1,r)\mu_2=p(r,r+1)\mu_1$.
-%
-This implies that the probability mass moved from $Z=+1$ to $Z=0$ is
-identical to that moved from $Z=-1$ to $Z=0$.
-
-We have showed that $\Pr[S_1>S_2]+ (1/2) \Pr[S_1=S_2]$ and therefore
-the expected valued of the exploit action is non-decreasing. Since
-we have that the size of the phases are increasing, the $\BIR$ is
-strictly increasing between phases and identical within each phase.
-
-We now analyze the $\BIR$ regret. Note that agent $n$ is in phase
-$O(n^{2/3})$ and the length of his phase is $O(n^{1/3})$. The $\BIR$
-has two parts. The first is due to the exploration, which is at most
-$O(n^{-1/3})$. The second is due to the probability that we exploit
-the wrong action. This happens with probability $\Pr[S_1<S_2]+ (1/2)
-\Pr[S_1=S_2]$ which we can bound using a Chernoff bound by
-$e^{-O(\Delta^2n^{2/3})}$, since we explored each action
-$O(n^{2/3})$ times.
-\end{proof}
-
-\begin{remark}
-Actually we have a tradeoff depending on the parameter $m_t$ between
-the regret due to exploration and exploitation. (Note that the
-monotonicity is always guarantee assuming $m_t$ is monotone.) If we
-can set that $m_t = 2^t$ then at time $n$ we have $2/ n$ probability
-of an exploit action. For the explore action we are in phase $\log
-n$ so the probability of a sub-optimal explore action is
-$n^{-O(\Delta^{-2})}$. This should give us
-$\BIR(n)=O(n^{-O(\Delta^{-2})})$.
-\end{remark}
-
-%MAB algorithm \RandomDynGreedy$(q_t)$for agent $t$ with probability
-%$q_t$ selects a random action from $A$ and with probability $1-q_t$
-%recommends the best action according to the current posterior
-%(similar to \DynGreedy).
-%\begin{lemma}
-%%Assume that rewards in the range $[0,1]$.
-%The algorithm \RandomDynGreedy$(q_t)$, where $q_t=t^{-1/3}$, has,
-%with probability $1-\delta$, \BIR$(n)=O(K^{1/2}n^{-1/3}\log
-%(K/\delta))$.
-%\end{lemma}
-%
-%\begin{proof}
-%At time $n$ we have that with probability $1-\delta/(2K)$, each
-%action $a\in A$ was selected at least $n_a=\Theta(n^{2/3}/K)$ times.
-%%
-%This implies that with probability $1-\delta/(2K)$, for each $a$ we
-%have $|\hat{\mu}_a-\mu_a|\leq K^{1/2}n^{-1/3}\log((2K)/\delta)$.
-%%
-%Therefore, the \BIR is $O(K^{1/2}n^{-1/3}\log((K)/\delta))$
-%\end{proof}
-
-%\ascomment{Add: can we modify Phased Explore-then-exploit to make it
-%monotone? Perhaps choose exploration rounds u.a.r. within the
-%phase?}
-
-%
-%\begin{lemma}
-%\ExplorExploit$(m)$ is monotone
-%\end{lemma}
-%
-%\begin{proof}
-%We need to consider only time $m'=|A|m+1$. For any $n<m'$, we
-%clearly have $\rew(n)=\rew(1)$.
-%
-%Consider two actions $a_1,a_2\in A$, such that $\mu_{a_1} \geq
-%\mu_{a_2}$. We claim that the probability that $a_1$ is selected at
-%time $m$ to be the exploit action is larger than the probability
-%that $a_2$ is selected. This follows since $\hat{\mu}_{a_1}$
-%stochastically dominates $\hat{\mu}_{a_2}$, which implies that for
-%any threshold $\theta$ we have $\Pr[\hat{\mu}_{a_1}\geq\theta]\geq
-%\Pr[\hat{\mu}_{a_2}\geq\theta]$.
-%
-%After the elimination we consider the expected reward of the
-%selected action $\sum_{i\in A} \mu_i q_i$, where $q_i$ is the
-%probability that action $i$ was selected for the exploit phase. We
-%have that $q_i \geq q_{i+1}$, from the stochastic dominance.
-%
-%The sum $\sum_{i\in A} \mu_i q_i$ with $q_i \geq q_{i+1}$ and
-%$\sum_i q_i=1$ is minimized by setting $q_i=1/|A|$. (We can see that
-%if there are $q_i\neq 1/|A|$, then there are two $q_{i}> q_{i+1}$,
-%and one can see that setting both to $(q_{i}+ q_{i+1})/2$ decreases
-%the value.) Therefore we have that the $\rew(m')\geq\rew(m'-1)$.
-%\end{proof}
-
-\subsubsection{Monotone, adaptive-exploration MAB algorithms}
-\label{app:MAB-monotone-smart}
-
-We provide a simple modification of Successive Elimination algorithm from \citep{EvenDar-icml06} that is monotone and (still) enjoys the same bounds on regret and BIR.
-%MAB algorithm \SuccesiveElimination holds a set of action
-%$A_s\subset A$ which are called {\em surviving actions} and for
-%agent $t$ selects a random action from $A_s$. Let $n_{i,t}$ be the
-%number of times action $i$ has been selected up to time $t$, and
-%$\hat{\mu}_{i,t}$ be the average of the rewards of action $i$ up to
-%time $t$ and $\hat{\mu}^*=\max_i \hat{\mu}_{i,t}$. We eliminate
-%action $i$ at time $t$, i.e., delete it from $A_s$, if
-%$\hat{\mu}_t^*-\hat{\mu}_{i,t} > \log(T/\delta)/\sqrt{n/K}$.
-%
-%\begin{lemma}
-%Assume that the prior is independent and if $\mu_i\geq \mu_j$ then
-%$r_i$ stochastically dominates $r_j$.
-%%
-%The algorithm \SuccesiveElimination, has, with probability
-%$1-\delta$, \BIR$(n)=O(\log(T/\delta)/\sqrt{n/K})$.
-%\end{lemma}
-
-%\begin{proof}
-%Let the best action be $a^*=\arg\max_a \mu_a$. With probability
-%$1-\delta$ at any time $n$ we have that for any action $i\in A_s$
-%that $|\hat{\mu}_i -\mu_i|\leq \log(T/\delta)/\sqrt{n/K}$, and
-%$a^*\in A_s$. This implies that any action $a$ such
-%$\mu_{a^*}-\mu_{a}> 3\log(T/\delta)/\sqrt{n/K}$ is eliminated.
-%Therefore, any action in $A_s$ has \BIR$(n)$ of at most
-%$O(\log(T/\delta)/\sqrt{n/K})$. \ymcomment{should we get the
-%constant right?}
-%\end{proof}
-
-%\ascomment{Add: Successive Elimination is monotone (or a simple
-%modification thereof).}
-%
-%\begin{lemma}
-%\SuccesiveElimination is monotone
-%\end{lemma}
-%
-%\begin{proof}
-%We will first show that for any two actions $a_1,a_2\in A$ that if
-%$\mu_{a_1} \geq \mu_{a_2}$ then at any time $t$ we have that the
-%probability that $a_1$ is eliminated is smaller than the probability
-%that $a_2$ is eliminated, i.e., $\Pr[a_1 \not\in A_{s,t+1}| a_1\in
-%A_{s,t} ]\geq \Pr[ a_2 \not\in A_{s,t+1}| a_2\in A_{s,t}]$.
-%
-%We first show that given this statement, the lemma holds and
-%\SuccesiveElimination is monotone. Consider $\rew(t)$ versus
-%$\rew(t+1)$. If the elimination probabilities would have been
-%identical then we would have $\rew(t)=\rew(t+1)$. Consider the top
-%two actions, and consider the elimination process in two phases,
-%firs we eliminate both with the same probability, and hence the
-%reward is unchanged, and then we eliminate only the second, which
-%implies that we increase the the reward. The claim about the
-%elimination probabilities rules out the case that the elimination
-%probability of the best action is higher than then of the the second
-%best. Therefore, overall the reward can only increase.
-%
-%
-%We now need to prove the claim. Since $\mu_{a_1}\geq \mu_{a_2}$ and
-%they use the same parametric family $\psi_a(\cdot)$ then
-%$\hat{\mu}_{a_1}$ stochastically dominates $\hat{\mu}_{a_2}$. We
-%actually need to show that $\hat{\mu}_{a_1}$ stochastically
-%dominates $\hat{\mu}_{a_2}$ given that $a_1,a_2\in A_{s,t}$. If this
-%holds, then the claim about the elimination would be immediate,
-%since for any threshold $\theta$ we have $\Pr[\hat{\mu}_{a_1,t} <
-%\theta |a_1\in A_{s,t}] \leq \Pr[\hat{\mu}_{a_2,t} < \theta |a_2\in
-%A_{s,t}] $. By setting $\theta=\hat{\mu}^*-
-%\log(T/\delta)/\sqrt{n/K}$ we derive the lemma.
-%
-%We now need to prove the claim. We first show that
-%$\E[\hat{\mu}_{a_1}-\hat{\mu}_{a_2} | a_1,a_2\in A_s]\geq 0$ which
-%is equivalent to $\E[\hat{\mu}_{a_1}-\hat{\mu}_{a_2} |
-%\hat{\mu}_{a_1},\hat{\mu}_{a_2}\geq \hat{\mu}^*-\theta]\geq 0$ where
-%$\theta$ is the threshold used in \SuccesiveElimination. Since
-%action $a_1$ has a high expectation than $a_2$ it holds (at least
-%for Bernoulli r.v.).
-%
-%
-%[[OLD]]
-%
-% Clearly, for any actions $a\in A$ we have that $\Pr[a \in
-%A_s]$ decreases with time. In addition, we have that with
-%probability $1-\delta$ for every time $t$ we have $a^*\in A_s$.
-%
-%If $\mu_{a_1} \geq \mu_{a_2}$ then at any time $t$ we have
-%$\Pr[a_1\in A_s]\geq \Pr[a_2\in A_s]$.
-%
-%
-%We probably need to show that if $\mu_{a_1} \geq \mu_{a_2}$ then at
-%any time $t$ we have that the probability that $a_1$ is eliminated
-%is smaller than the probability that $a_2$ is eliminated, i.e.,
-%$\Pr[a_1\in A_{s,t}, a_1 \not\in A_{s,t+1}]\geq \Pr[a_2\in A_s, a_2
-%\not\in A_{s,t+1}]$.
-%\end{proof}
-
-%\ymcomment{Here is a simple modification of Successive-Elimination}
-
-The algorithm, called \SuccesiveEliminationReset, proceeds in phases.
-In each phase $j$, the algorithm is given a set of ``surviving" arms $A_j\subseteq A$, with $A_1=A$.
-The phase lasts $|A_j|$ rounds, randomly permuted across the arms in $A_j$ (so that each arm in $A_j$ is chosen exactly once).
-
-\ascomment{Stopped here.}
-
-\ascomment{Steven: I don't understand the algorithm! It resets after each phase, right? But then, the history is just one sample per each arm? WTF?}
-
-Let $\hat{\mu}_{i,t}$ be the average reward of action $i$ up
-to phase $t$ and $\hat{\mu}^*=\max_i \hat{\mu}_{i,t}$. We eliminate
-action $i$ at the end of phase $t$, i.e., delete it from $A_s$, if
-$\hat{\mu}_t^*-\hat{\mu}_{i,t} > \log(T/\delta)/\sqrt{t}$.
-%So far identical to \SuccesiveElimination, the difference is how we
-%continue after elimination.
-In \SuccesiveEliminationReset we simply reset the algorithm with
-$A=A_s-A_{e,t}$, where $A_{e,t}$ is the set of eliminated actions
-after phase $t$. Namely, we restart $\hat{\mu}_{i,t}$ and ignore the
-old rewards before the elimination.
-
-\begin{lemma}
-Consider algorithm \SuccesiveEliminationReset.
-\begin{itemize}
-\item[(a)]
-With probability $1-\delta$, $\IReg(n)=O(\log(T/\delta)/\sqrt{n/K})$.
-
-\item[(b)]
-Assuming Bernoulli rewards, the algorithm is monotone.
-\ascomment{Steven: for each \MRV, or only in expectation?}
-\ascomment{Steven: can you derive optimal regret?}
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-\xhdr{Part (a).}
-Let the best action be $a^*=\arg\max_a \mu_a$. With probability
-$1-\delta$ at any time $n$ we have that for any action $i\in A_s$
-that $|\hat{\mu}_i -\mu_i|\leq \log(T/\delta)/\sqrt{n/K}$, and
-$a^*\in A_s$. This implies that any action $a$ such
-$\mu_{a^*}-\mu_{a}> 3\log(T/\delta)/\sqrt{n/K}$ is eliminated.
-Therefore, any action in $A_s$ has \BIR$(n)$ of at most
-$6\log(T/\delta)/\sqrt{n/K}$.
-
-\xhdr{Part (b).}
-Consider the first time $T$ an action is eliminated, and let
-$T=\tau$ be a realized value of $T$. Then, clearly for $n<\tau$ we
-have $\rew(n)=\rew(1)$ .
-
-Consider two actions $a_1,a_2\in A$, such that $\mu_{a_1} \geq
-\mu_{a_2}$. At time $T=\tau$, the probability that  $a_1$ is
-eliminated is smaller than the probability that $a_2$ is eliminated.
-This follows since $\hat{\mu}_{a_1}$ stochastically dominates
-$\hat{\mu}_{a_2}$, which implies that for any threshold $\theta$ we
-have $\Pr[\hat{\mu}_{a_1}\geq\theta]\geq
-\Pr[\hat{\mu}_{a_2}\geq\theta]$.
-
-After the elimination we consider the expected reward of the
-eliminated action $\sum_{i\in A} \mu_i q_i$, where $q_i$ is the
-probability that action $i$ was eliminated in time $T=\tau$. We have
-that $q_i \leq q_{i+1}$, from the probabilities of elimination.
-
-The sum $\sum_{i\in A} \mu_i q_i$ with $q_i \leq q_{i+1}$ and
-$\sum_i q_i=1$ is maximized by setting $q_i=1/|A|$. (We can see that
-if there are $q_i\neq 1/|A|$, then there are two $q_{i}< q_{i+1}$,
-and one can see that setting both to $(q_{i}+ q_{i+1})/2$ increases
-the value.) Therefore we have that the $\rew(\tau)\geq\rew(\tau-1)$.
-
-Now we can continue by induction. For the induction, we can show the
-property for {\em any} remaining set of at most $k-1$ actions. The
-main issue is that \SuccesiveEliminationReset restarts from scratch,
-so we can use induction.
-\end{proof}

### 2020-07-17 journal paper/journal_main.tex
Changed the options for the appendix.
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.
Changed "monotone" to \bmonotone.
added \bmonotonicity macro.
Split "monotonicity" into a separate appendix.
Removed the "small font" from the bib.
Added a sentence to the title note.

@@
 \usepackage{libertine}
-\usepackage[toc,page]{appendix}
+\usepackage[title,titletoc]{appendix}
 \usepackage{url}            % simple URL typesetting
@@
 
+% "monotone algorithms"
+\newcommand{\bmonotone}{Bayesian-monotone\xspace} 
+\newcommand{\bmonotonicity}{Bayesian-monotonicity\xspace}
+
 \newcommand{\termSub}[2]{\ensuremath{\mathtt{#1}_{#2}}\xspace}
@@
 with a unified and streamlined presentation and expanded background materials.
-All theoretical results are from \citet{CompetingBandits-itcs18}, and all experiments are from \citet{CompetingBandits-ec19}.}}
+All theoretical results are from \citet{CompetingBandits-itcs18}, and all experiments are from \citet{CompetingBandits-ec19}.
+Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.}}
 % \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
@@
 
-We consider a stylized duopoly model in which two firms face the same \gaedit{multi-armed bandit instance.} Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen.
+We consider a stylized duopoly model in which two firms face the same multi-armed bandit instance.
+Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen.
 We study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for consumers.
@@
 \bibliographystyle{plainnat}
-\begin{small}
+%\begin{small}
 \bibliography{bib-abbrv,bib-ML,refs,bib-bandits,bib-AGT,bib-slivkins, bib-random}
-\end{small}
+%\end{small}
 
 \clearpage
+
+\renewcommand{\appendixname}{Appendix}
+\renewcommand{\appendixtocname}{List of appendices}
+
 \begin{appendices}
@@
 \section{Background for non-specialists: multi-armed bandits}
-\label{sec:related-classes}
+\label{app:bg}
+%\label{sec:related-classes}
 \subfile{content/sec-bg}
 
-\section{Auxiliary theoretical results}
+\section{Monotone MAB algorithms}
 \label{app:examples}
@@
 
-\subsection{Non-degeneracy via a random perturbation}
+\section{Non-degeneracy via a random perturbation}
 \label{app:perturb}

### 2020-07-17 journal paper/content/sec-theory-proofs.tex
added \bmonotonicity macro.
Removed \newpage statements.

@@
 Fix agent $t$. Recall that $n_i(t)$ denotes the number of global rounds before $t$ in which principal $i$ is chosen. Let $\posteriorN{i}{t}$ denote the distribution of $n_i(t)$.
-\ascomment{Steven: in the ITCS paper, we defined $\posteriorN{i}{t}$ as the posterior distribution given all info available to agent $t$. But this is the same thing, isn't it?}
 
@@
 \end{align*}
-The above inequalities follow, resp., from algorithms' monotonicity and \eqref{eq:random-better-weaker}. Now,
+The above inequalities follow, resp., from algorithms' \bmonotonicity and \eqref{eq:random-better-weaker}. Now,
 \begin{align*}

### 2020-07-17 journal paper/itcs18paper/app-perturb.tex
Minor edits for space-saving.

@@
 We provide two examples when Assumption~\eqref{eq:assn-distinct} holds almost surely under a small random perturbation of the prior.
-We focus on problem instances with Bernoulli rewards, and assume that the prior $\priorMu$ is independent across arms.
+We posit Bernoulli rewards, and assume that the prior $\priorMu$ is independent across arms.
 
@@
   distribution $\mN$ over the bounded convex set
-  \[ Q = \left\{q \in \R^{d_a} \mid q \cdot \vec{1} = 0 \mbox{ and }
-      \|q\|_2 \leq \eps\right\}, \] where $\vec{1}$ denotes the all-1
+  $ Q = \left\{q \in \R^{d_a} \mid q \cdot \vec{1} = 0 \mbox{ and }
+      \|q\|_2 \leq \eps\right\}, $
+  where $\vec{1}$ denotes the all-1
   vector. If $x = a \vec{1}$ for some non-zero value of $a$, then
   \eqref{eq:app:perturb:noise-prop} holds because
-  $$x \cdot (p+q) = x \cdot p = a\neq 0.$$ Otherwise, denote
+  $x \cdot (p+q) = x \cdot p = a\neq 0.$
+Otherwise, denote
   $p=\vec{p}_a$ and observe that $x\cdot({p}+ {q}) = 0$ only if
@@
 
-To prove Theorem~\ref{thm:perturb}, it suffices to focus on two arms, and perturb one of them.
+To prove Theorem~\ref{thm:perturb}, it suffices to focus on two arms, and perturb one.
 Since realized rewards have finite support, there are only finitely many possible histories.
-Therefore, it suffices to focus on a fixed history $h$.
+So, it suffices to focus on a fixed history $h$.
 

### 2020-07-17 journal paper/content/sec-theory.tex
Fixed a reference.
Minor edits.

@@
 An algorithm is called \emph{\bmonotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing (and therefore $\BIR(\cdot)$ is non-decreasing). We argue this is a mild assumption, see
-Appendix~\ref{app:MAB-mono}.
+Appendix~\ref{app:examples}.
 
@@
 \xhdr{Time-discounted utility.}
-Corollary~\ref{cor:DG-dominance} holds under a more general model which allows time-discounting: namely, the utility of each principal $i$ in each global round $t$ is $U_{i,t}(r_t)$ if this principal is chosen, and $0$ otherwise, where $U_{i,t}(\cdot)$ is an arbitrary non-decreasing function with $U_{i,t}(0)>0$.
+Theorem~\ref{thm:DG-dominance} and Corollary~\ref{cor:DG-dominance} holds under a more general model which allows time-discounting: namely, the utility of each principal $i$ in each global round $t$ is $U_{i,t}(r_t)$ if this principal is chosen, and $0$ otherwise, where $U_{i,t}(\cdot)$ is an arbitrary non-decreasing function with $U_{i,t}(0)>0$.
 
@@
 \noindent Weak BIR-dominance can be redefined similarly.
-Theorem~\ref{thm:random-clean} and Theorem~\ref{thm:SoftMax-weak} extend to these generalized versions; the easy modifications are omitted.
+Theorem~\ref{thm:random-clean} and~\ref{thm:SoftMax-weak} easily extend to these versions.
 

### 2020-07-17 journal paper/content/sec-theory.tex
Minor edits.

@@
 \xhdr{Arbitrary reward distributions.}
-While Section~\ref{sec:model} posits Bernoulli rewards, we allow arbitrary reward distributions.
-Formally, for each arm $a\in A$ there is a parametric family $\psi_a(\cdot)$ of reward distributions, parameterized by the mean reward.
+Bernoulli rewards can be extended to arbitrary reward distributions.
+For each arm $a\in A$ there is a parametric family $\psi_a(\cdot)$ of reward distributions, parameterized by the mean reward.
 Whenever arm $a$ is chosen, the reward is drawn independently from distribution $\psi_a(\mu_a)$.

### 2020-07-18 journal paper/content/sec-intro.tex
Added a para on time-discounting.

@@
 Many platforms purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a well-known framework for exploration and making decisions under uncertainty.
+Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a standard and well-studied framework for exploration and making decisions under uncertainty
+\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
 
@@
 
-\xhdr{Assumptions.}
-Our models are stylized in several important  respects, some of which we discuss below.
+\xhdr{Modeling.}
+Our models are stylized in several important  respects.
 Firms compete only on the quality of service, rather than, say, pricing or the range of products.
@@
 Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.
+
+\asedit{Principals' utilities (and, accordingly, bandit rewards) are not discounted with time.
+While much of the early work focuses on  time-discounted bandits \citep{Gittins-book11}, non-discounted  formulations are prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
+Moreover, the distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.}
 

### 2020-07-18 journal paper/content/conclusion.tex
Full rewrite of conclusions.

--- 2019-08-18 journal paper/content/conclusion.tex
@@
 
-\section{Conclusion}\label{sec:conclusion}
+\section{Conclusions}\label{sec:conclusion}
 
-We considered a stylized duopoly setting where firms simultaneously learn from and compete for users.
-We analyzed two separate but connected variants of this model, one where consumers chose between firms based on their Bayesian expected reward and another where consumers chose between them based on their past performance.
-We showed in both cases that competition may not always induce firms to commit to better exploration algorithms, resulting in welfare losses for consumers.
-The primary mechanism in both was that consumers need to be incentivized to select a firm over its competitors, leading firms that engage in exploration to be starved of consumers before they make enough progress on their learning problem.
-We found that in order to incentivize ``better" exploration strategies the key intuition is that a firm needs to have some ``free" consumers that visit them without having to be incentivized to do so.
-We explored several economic mechanisms that can generate this.
-The first was by considering different variants of random utility models where stochastic choice gives firms enough free consumers to incentivize them to commit to ``better" algorithms and we explored how this varied based on the nature of the random utility model.
-The second was by considering giving one firm a first-mover advantage where all consumers in the market would go to this firm without having to be incentivized to do so.
-While this first-mover advantage lead to the incumbent firm being incentivized to commit to a ``better" algorithm it lead to a substantial market share going to this firm.
-We further analyzed to what extent this advantage came from having a more defined reputation or more data from the incumbency period.
-We found that even a small amount of ``data advantage" leads to substantial long-term market power due to the dynamics of competition.
+We study the tension between exploration and competition.
+We consider a stylized duopoly model in which two firms face an identical multi-armed bandit problem and compete for a stream of users.
+A firm makes progress on its learning problem if and only if it attracts users.
+We investigate two variants: an analytical variant where users do not observe any performance signals and choose between firms according to their Bayesian-expected rewards, and numerical simulations where a reputation score is observed for each firm, based on the average reward of its recent users.
+In both variants, firms are incentivized to adopt a ``greedy algorithm" which does no purposeful exploration and leads to welfare losses for users.
+We then consider two relaxations of competition: we relax users' rationality and give one of the firms a first-mover advantage.
+Both relaxations induce firms to adopt ``better" bandit algorithms, which benefits user welfare.
+
+Our results have two economic interpretations.
+Our results can be framed in terms of the classic inverted-U relationship between innovation and competition, where \innovation refers to the adoption of better bandit algorithms.
+Unlike other models in the literature, what prevents innovation is not its direct costs, but short-term reputation consequences of exploration.
+The second interpretation concerns the role of data in the digital economy.
+We find that even a small initial disparity in data or reputation gets amplified under competition to a very substantial difference in the eventual market share.
+Thus, we endogenously obtain ``network effects" without explicitly baking them into the model, and elucidate the role of data as a barrier to entry.
+
+With this paper as a common departure point, there are several exciting directions to explore:
+
+\begin{itemize}
+
+\item \emph{Prices.} 
+When the firms can set prices, they may be able to compensate early users for exploration, and potentially prevent the ``death spiral" effects.
+Whereas our paper zeroes in on competition between free, ad-supported platforms that primarily compete on quality.
+
+\item \emph{Horizontal differentiation.} 
+Horizontally differentiated user preferences may help explain how competition may encourage specialization, \ie how the firms may \emph{learn to specialize} under competition.
+
+\item \emph{Continuous learning.}
+While we focus on learning from scratch in a mainly stationary world, another well-motivated regime is continuous learning, when exploration counteracts change
+\citep{DynamicMAB-colt08}.
+The economic story would be about competition between relatively mature firms.
+
+\item \emph{Real-life data.}
+Numerical experiments based on real-life datasets would arguably be more realistic.
+Bandit algorithms would probably need to incorporate \emph{contexts} (auxiliary signals available before each round), as the real-life bandit problems tend to be contextual.
+
+\end{itemize}
+
+\noindent One difficulty with the last two directions is that the bandit ``side" of the model becomes considerably more complicated, both in terms of the rewards and in terms of the algorithms.
+Indeed, there are many reasonable ways to model a changing world (resp., contexts), and several  substantially different algorithmic approaches to deal with that.
+Even within the same approach, the distinctions between better and worse algorithms are not as clear, and the intuitions are not as well-established.
 

### 2020-07-18 journal paper/content/sec-intro.tex
Added a footnote re time-discounting.

@@
 
-\asedit{Principals' utilities (and, accordingly, bandit rewards) are not discounted with time.
+\asedit{Principals' utilities (and, accordingly, bandit rewards) are not discounted with time.%
+\footnote{However, our main theoretical result on the greedy algorithm extends to time-discounted rewards, see Section~\ref{sec:theory-extensions}.}
  While much of the early work focuses on  time-discounted bandits \citep{Gittins-book11}, non-discounted  formulations are prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.

### 2020-07-18 journal paper/content/conclusion.tex
typos / minor stylistic fixes

@@
 Our results have two economic interpretations.
-Our results can be framed in terms of the classic inverted-U relationship between innovation and competition, where \innovation refers to the adoption of better bandit algorithms.
-Unlike other models in the literature, what prevents innovation is not its direct costs, but short-term reputation consequences of exploration.
+\gaedit{The first is that they} can be framed in terms of the classic inverted-U relationship between innovation and competition, where \innovation refers to the adoption of better bandit algorithms.
+Unlike other models in the literature, what prevents innovation is not its direct costs, but \gaedit{the} short-term reputation consequences of exploration.
 The second interpretation concerns the role of data in the digital economy.

### 2020-07-18 journal paper/content/sec-intro.tex
typos / minor stylistic fixes

@@
 %whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
-The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved out of users (resp., take over most of the market).
+The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved of users (resp., take over most of the market).
 %\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}

### 2020-07-18 journal paper/itcs18paper/app-examples.tex
typos / minor stylistic fixes

@@
 
-This appendix provides some auxiliary results on \bmonotonicity of some standard algorithms: \DynGreedy, \DynamicEpsGreedy and Thompson Sampling.
+This appendix provides some auxiliary results on \bmonotonicity of some standard algorithms: \DynGreedy, \DynamicEpsGreedy and \Thompson.
 The former is needed in Section~\ref{sec:theory}, the last two merely add motivation for our theoretical story.
-The result on Thompson Sampling also implies that $\BIR(t)\leq \tildeO(t^{-1/2})$.
+The result on\Thompson also implies that $\BIR(t)\leq \tildeO(t^{-1/2})$.
 Recall that an algorithm is called \bmonotone if its Bayesian-expected reward is non-decreasing in time.
@@
 
-\xhdr{Monotonicity and \BIR of Thompson Sampling.} 
+\xhdr{Monotonicity and \BIR of \Thompson.} 
 Monotonicity follows from \citet{Selke-BE-2020}; the \BIR result is an immediate corollary given that
@@
 \begin{theorem}[\citet{Selke-BE-2020}]\label{thm:TS-mono}
-Thompson Sampling is \bmonotone.
+\Thompson is \bmonotone.
 \end{theorem}
@@
 \begin{corollary}
-Thompson Sampling satisfies $\BIR(t)\leq \tildeO(t^{-1/2})$.
+\Thompson satisfies $\BIR(t)\leq \tildeO(t^{-1/2})$.
 \end{corollary}
@@
 Let $a^*$ be the best arm: $a^*\in \argmax_a \mu_a$.
-In each round $t$, Thompson Sampling samples arm $a_t$ as follows:
+In each round $t$, \Thompson samples arm $a_t$ as follows:
 \begin{align}\label{eq:TS-defn}

### 2020-07-18 journal paper/content/related_work.tex
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.
fix broken ref

@@
 The literature on MAB is extremely vast and multi-threaded.
-The most relevant thread studies a basic model of MAB with IID rewards and no auxiliary structure, see Appendix~\ref{sec:related-classes}.
+The most relevant thread studies a basic model of MAB with IID rewards and no auxiliary structure, see Appendix~\ref{app:bg}.
 The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.} This literature is covered in many books and surveys.
@@
 \xhdr{Other related work in economics.} The competition vs.
-innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{Aghion-QJE05,Vives-08}.
+innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{aghion2005competition,Vives-08}.
 Production costs aside, this literature treats innovation as a priori beneficial for the firm.

### 2020-07-18 journal paper/content/sec-intro.tex
fix broken ref

@@
 Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
-This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{Aghion-QJE05,Vives-08}.
+This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{aghion2005competition,Vives-08}.
 

### 2020-07-18 journal paper/content/conclusion.tex
Minor edits in open q.

@@
 Numerical experiments based on real-life datasets would arguably be more realistic.
-Bandit algorithms would probably need to incorporate \emph{contexts} (auxiliary signals available before each round), as the real-life bandit problems tend to be contextual.
+\asedit{To deal with real-life datasets, bandit algorithms would probably need to accommodate change over time and} \emph{contexts} (auxiliary signals available before each round).
 
@@
 \noindent One difficulty with the last two directions is that the bandit ``side" of the model becomes considerably more complicated, both in terms of the rewards and in terms of the algorithms.
-Indeed, there are many reasonable ways to model a changing world (resp., contexts), and several  substantially different algorithmic approaches to deal with that.
-Even within the same approach, the distinctions between better and worse algorithms are not as clear, and the intuitions are not as well-established.
+Indeed, there are many reasonable ways to model a changing world (resp., dependence on  contexts), and several  substantially different algorithmic approaches to deal \asedit{these models.
+Consequently, the distinctions between better and worse algorithms are not as clear and well-established.}
 

### 2020-07-18 journal paper/itcs18paper/app-examples.tex
In "TS is monotone", added independence assumption.
In AppB: added prior into the setup.

@@
 
-Throughout, we consider a standard MAB setting.
+We consider a simple setting of Bayesian MAB with Bernoulli rewards.
 There are $T$ rounds and $K$ arms.
-In each round $t\in [T]$, the algorithm chooses an arm $a_t\in A$ and receives a reward $r_t\in[0,1]$ for this arm, drawn from a fixed but unknown distribution.
+In each round $t\in [T]$, the algorithm chooses an arm $a_t\in A$ and receives a reward $r_t\in\{0,1\}$ for this arm, drawn from a fixed but unknown distribution.
 The set of all arms is $A$; mean reward of arm $a$ is denoted $\mu_a$.
-
+\asedit{The mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$.}
 We let
@@
 
-Instantaneous regret is denoted
-    $\IReg(t) := \max_{a\in A} \mu_a - \mu_{a_t}$,
-so that
-    $\BIR(t) = \E[\IReg(t)]$.
+%Instantaneous regret is denoted
+%    $\IReg(t) := \max_{a\in A} \mu_a - \mu_{a_t}$,
+%so that
+%    $\BIR(t) = \E[\IReg(t)]$.
 
@@
 \begin{theorem}[\citet{Selke-BE-2020}]\label{thm:TS-mono}
-\Thompson is \bmonotone.
+\asedit{Assume the prior $\priorMu$ is independent across arms.} Then \Thompson is \bmonotone. Consequently, it satisfies $\BIR(t)\leq \tildeO(t^{-1/2})$.
 \end{theorem}
-
-\begin{corollary}
-\Thompson satisfies $\BIR(t)\leq \tildeO(t^{-1/2})$.
-\end{corollary}
 

### 2020-07-19 journal paper/content/conclusion.tex
fix typo in conclusion

@@
 \noindent One difficulty with the last two directions is that the bandit ``side" of the model becomes considerably more complicated, both in terms of the rewards and in terms of the algorithms.
-Indeed, there are many reasonable ways to model a changing world (resp., dependence on  contexts), and several  substantially different algorithmic approaches to deal \asedit{these models.
+Indeed, there are many reasonable ways to model a changing world (resp., dependence on  contexts), and several substantially different algorithmic approaches to deal \asedit{with these models.
 Consequently, the distinctions between better and worse algorithms are not as clear and well-established.}

### 2020-07-19 journal paper/content/sec-bg.tex
typos / minor stylistic fixes
Minor edits.

@@
 There are $T$ rounds and $K$ \emph{arms} to choose from.
-In each round $t\in [T]$, the algorithm chooses an arm and receives a reward $r_t\in[0,1]$ for this arm, drawn from a fixed but unknown distribution.
-Algorithm's goal is to maximize the total reward.
+In each round $t\in [T]$, the algorithm chooses an arm and receives a reward $r_t\in[0,1]$ for this arm, drawn from a fixed but unknown distribution.%
+\footnote{\asedit{All ``negative" results (\ie lower bounds on regret) assume reward distributions with constant variance.}}
+The algorithm's goal is to maximize the total reward.
 
@@
     $O(\tfrac{K}{\Delta}\log T)$
-for all problem instances with $\text{gap}\geq \Delta$, without knowing the $\Delta$ in advance
+for all problem instances with $\text{gap}\geq \Delta$ (``easy" instances), without knowing the $\Delta$ in advance
 \citep{Lai-Robbins-85,bandits-ucb1,bandits-exp3}.%
-\footnote{The \emph{gap} is the difference in mean reward between the best arm and the second-best arm.
-It is a standard notion of difficulty for MAB instances: instances with large gap are deemed  ``easy".}
+\footnote{The \emph{gap} is the difference in mean reward between the best arm and the second-best arm.}
 Exploration-separating algorithms can only achieve regret $\tildeO(T^{2/3})$ across all problem instances.
-They can achieve the ``gap-dependent" regret rate stated above, but \emph{only} if they know the $\Delta$ in advance, and at the cost of a terrible regret $\Omega(\Delta T)$ for some other problem instances \citep{MechMAB-ec09}.
+They can achieve the ``gap-dependent" regret rate stated above, but \emph{only} if they know the $\Delta$ in advance, and with terrible regret $\Omega(\Delta T)$ for some other problem instances \citep{MechMAB-ec09}.
 Finally, the greedy algorithm is terrible on a wide variety of problem instances, in the sense that with constant probability it fails to try the best arm even once, and therefore suffers regret $\Omega(T)$
@@
 The general definition of ``exploration-separating algorithms" and the associated lower bounds trace back to \citet{MechMAB-ec09}.%
-\footnote{\citet{MechMAB-ec09} considers a closely related, but technically different setting, which can be easily ``translated" into ours (either as a corollary or as another application of the same proof technique).
+\footnote{\citet{MechMAB-ec09} consider a closely related, but technically different setting, which can be easily ``translated" into ours (either as a corollary or as another application of the same proof technique).
 Detailing this is beyond the scope of this appendix.}
@@
 
+\asedit{Some MAB algorithms are Bayesian, \eg Thompson Sampling,  in the sense that they input a prior on mean rewards, and attain strong Bayesian guarantees (in expectation over the prior) when the prior is correct.
+Such algorithms can also be initialized with some simple `fake' priors; in fact, Thompson Sampling satisfies the optimal regret bounds listed above.}
+
 The intuition on (the separation between) the three algorithm classes applies more generally, far beyond the basic MAB model discussed above.
@@
 \begin{align*}%\label{eq:bg-BIR-defn}
-\BIR(t) := \E_{\text{prior}}\sbr{ \max_{a\in A} \mu_a - r_t}.
+\BIR(t) := \E_{\text{prior}}\sbr{ \max_{\text{arms $a$}} \mu_a - r_t}.
 \end{align*}
@@
  $\BReg(t)\geq \Omega(t^{2/3})$,
-the \BIR rate of $t^{-1/3}$  cannot be improved, in the following sense:
+the \BIR rate of $t^{-1/3}$  cannot be improved by \eqref{eq:bg-BReg}, in the following sense:
 if
@@
 
-This theoretical intuition is supported by our numerical simulations: see 
+\noindent This theoretical intuition is supported by our numerical simulations: see
 Figure~\ref{relative_rep_plots} and Appendix~\ref{app:isol}.

### 2020-07-19 journal paper/itcs18paper/app-examples.tex
Minor edits.
fix missing space
small comment on ts
small comment

@@
 The former is needed in Section~\ref{sec:theory}, the last two merely add motivation for our theoretical story.
-The result on\Thompson also implies that $\BIR(t)\leq \tildeO(t^{-1/2})$.
+The result on \Thompson also implies that $\BIR(t)\leq \tildeO(t^{-1/2})$.
 Recall that an algorithm is called \bmonotone if its Bayesian-expected reward is non-decreasing in time.
 
-We consider a simple setting of Bayesian MAB with Bernoulli rewards.
+We consider Bayesian MAB with Bernoulli rewards.
 There are $T$ rounds and $K$ arms.
@@
 \end{align}
+\swcomment{do we want to say $\Pr$ is over sampling first from prior and then generate history?}
 
@@
 \end{proof}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../journal_main"
+%%% End:

### 2020-08-22 journal paper/code_abstract.tex
add CODE abstract

--- 2020-07-18 journal paper/content/sec-intro.tex
@@
+\documentclass[10pt]{article}
+
+\usepackage{color,
+setspace,sectsty,comment,footmisc,caption,
+pdflscape,subcaption,array,hyperref,adjustbox,threeparttable
+}
+
+\usepackage{fullpage}
+%\usepackage{geometry}
+%\geometry{left=1.0in,right=1.0in,top=1.0in,bottom=1.0in}
+
+\usepackage{subfiles}
+\usepackage{libertine}
+\usepackage[title,titletoc]{appendix}
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{amsmath,amssymb,amsfonts,amsthm}
+\usepackage[round]{natbib} % cannot use with AAAI! A.S.
+
+\usepackage{float}
+\usepackage{sgame, tikz} % Game theory packages
+\usepackage{algorithm,algpseudocode}
+\usepackage{makecell}
+ \usepackage{multirow}
+ \usepackage{graphicx}
+
+\usepackage{booktabs} % For formal tables
+
+%\theoremstyle{definition}
+%\newtheorem{definition}{Definition}
+\newtheorem{finding}{Finding}
+%\newtheorem{conjecture}{Conjecture}
+\newtheorem{puzzle}{Puzzle}
+\captionsetup{font=footnotesize}
+
+\usepackage{slivkins-setup,slivkins-theorems}
+
+% model variants
+\newcommand{\TheoryModel}{Bayesian-choice model\xspace}
+\newcommand{\ExptsModel}{reputation-choice model\xspace}
+
+% Commands from EC 19 paper
+\newcommand{\term}[1]{\ensuremath{\mathtt{#1}}\xspace}
+
+% shorthand for algorithms
+\newcommand{\TS}{\term{TS}}    % Thompson Sampling
+\newcommand{\DEG}{\term{BEG}}  % "Dynamic eps-greedy"
+\newcommand{\DG}{\term{BG}}    % "Dynamic Greedy"
+
+% shorthand for models
+\newcommand{\HMR}{\term{HMR}} % HardMaxRandom
+\newcommand{\HM}{\term{HM}}    % HardMax
+
+\newcommand{\Beta}{\term{Beta}} % for Beta distribution
+\newcommand{\Eeog}{\term{EoG}} % shorthand for "effective end of game"
+
+\newcommand{\MRV}{mean reward vector\xspace} % mean reward vector
+\newcommand{\MRVs}{mean reward vectors\xspace} % mean reward vector
+
+% commands from "theory"
+
+% in connection with "a little greedy goes a long way"
+%     stuff for the "greedy modification".
+\newcommand{\gr}{\texttt{gr}}
+\newcommand{\BIRgr}{\BIR^{\gr}}
+\newcommand{\alggr}{\alg[\gr]}
+\newcommand{\rewgr}{\rew_{\gr}}
+
+\DeclareMathOperator*{\Expectation}{\mathbb{E}}
+\newcommand{\Ex}[2]{\Expectation_{#1}\left[#2\right]}
+\newcommand{\FiniteGame}{finite competition game\xspace}
+
+% notation
+%%% advanced notation
+\newcommand{\OPT}{\term{OPT}}
+\newcommand{\rew}{\term{rew}}  % Bayesian-expected reward after n local rounds
+\newcommand{\PMR}{\term{PMR}} % posterior mean reward
+\newcommand{\support}{\term{support}}
+
+\newcommand{\BIR}{\term{BIR}} % Bayesian Instantaneous Regret
+
+% response function
+\newcommand{\respF}{f_{\term{resp}}}
+\newcommand{\respEps}{\eps_\term{resp}}
+
+\newcommand{\BReg}{\term{BReg}}
+\newcommand{\HardMax}{\term{HardMax}}
+\newcommand{\HardMaxRandom}{\term{HardMax\&Random}}
+\newcommand{\SoftMaxRandom}{\term{SoftMax}}
+\newcommand{\Uniform}{\term{Uniform}}
+\newcommand{\Random}{\term{Random}}
+
+% algorithms
+\newcommand{\StaticGreedy}{\term{StaticGreedy}}
+\newcommand{\DynGreedy}{\term{BayesianGreedy}}
+\newcommand{\DynamicGreedy}{\DynGreedy}
+\newcommand{\DynamicEpsGreedy}{\term{BayesianEpsilonGreedy}}
+\newcommand{\Thompson}{\term{ThompsonSampling}}
+
+% "monotone algorithms"
+\newcommand{\bmonotone}{Bayesian-monotone\xspace}
+\newcommand{\bmonotonicity}{Bayesian-monotonicity\xspace}
+
+\newcommand{\termSub}[2]{\ensuremath{\mathtt{#1}_{#2}}\xspace}
+\newcommand{\alg}[1][]{\termSub{alg}{#1}}
+%\newcommand{\prin}[1][]{\termSub{prin}{#1}}  % principal
+\newcommand{\agent}[1][]{\termSub{agent}{#1}}
+
+% priors and posteriors
+\newcommand{\prior}{\ensuremath{\mP}\xspace}
+\newcommand{\priorMu}{\ensuremath{\prior_\mathtt{mean}}\xspace}
+\newcommand{\posteriorN}[2]{\mN_{#1,#2}}  % \posteriorN{principal}{round}
+
+% rationality / innovation / competition
+\newcommand{\termTXT}[1]{{\em {#1}}\xspace}
+
+\newcommand{\rationality}{\termTXT{rationality}}
+\newcommand{\Rationality}{\termTXT{Rationality}}
+\newcommand{\innovation}{\termTXT{innovation}}
+\newcommand{\Innovation}{\termTXT{Innovation}}
+\newcommand{\competition}{\termTXT{competition}}
+\newcommand{\Competition}{\termTXT{Competition}}
+\newcommand{\competitiveness}{\termTXT{competitiveness}}
+\newcommand{\exploration}{\termTXT{exploration}}
+\newcommand{\Exploration}{\termTXT{Exploration}}
+
+% a very useful package for edits and comments, from David Kempe (USC)
+\usepackage[suppress]{color-edits}
+%\usepackage[suppress]{color-edits}  % use this to suppress the package
+\addauthor{as}{red}    % as for Alex
+\addauthor{ga}{blue}  % ga for Guy
+\addauthor{sw}{orange} % sw for Steven
+\addauthor{ym}{madenta} % sw for Steven
+% e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
+
+\setlength{\tabcolsep}{8pt} % Default value: 6pt
+\renewcommand{\arraystretch}{1.5} % Default value: 1
+
+\begin{document}
+
+%\begin{titlepage}
+
+\title{Competing Bandits:\\
+The Perils of Exploration under Competition%
+\thanks{This is a merged and final version of two conference papers,
+\citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19},
+with a unified and streamlined presentation and expanded background materials.}}
+% \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
+
+\author{Guy Aridor%
+\footnote{Columbia University, Department of Economics. Email: g.aridor@columbia.edu}
+\and
+%\rule{0.2in}{0pt}
+Yishay Mansour%
+\footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com}
+\and
+%\rule{0.2in}{0pt}
+Aleksandrs Slivkins%
+\footnote{Microsoft Research New York City. Email: slivkins@microsoft.com}
+\and
+%\rule{0.2in}{0pt}
+Zhiwei Steven Wu%
+\footnote{University of Minnesota - Twin Cities, Department of Computer Science. Email: zsw@umn.edu\newline
+Part of the research was done when Z.S. Wu was an intern and a postdoc at Microsoft Research NYC.
+}}
+\date{August 2020}
+\maketitle
+
 Learning from interactions with users is ubiquitous in modern customer-facing platforms, from product recommendations to web search to content selection to fine-tuning user interfaces.
@@
 
-\subsection{Our model: competing bandits}
+\subsection*{Our model: competing bandits}
 \label{sec:intro-model}
@@
 
-While we build on a vast and active research area, we present sufficient background in Appendix~\ref{app:bg} so as to keep this paper accessible to non-specialists.
-
-\subsection{Our Results}
+\subsection*{Our Results}
 \label{sec:intro-results}
@@
 
-All results extend to a much more general version of the multi-armed bandit problem in which the principal may observe additional feedback before and/or after each decision, as long as the feedback distribution does not change over time.
-In most results, principal's utility may depend on both the market share and agents' rewards.
-
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
-      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
-      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
-      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
- \end{tikzpicture}
-
-\caption{Inverted-U relationship between competitiveness and algorithms.}
-\label{fig:inverted-U}
-\end{center}
-\end{figure}
-
-\xhdr{Economic interpretation: the inverted-U relationship.}
-Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the inverted-U relationship between competition and innovation, as in Figure~\ref{fig:inverted-U}.%
-\footnote{Here, ``innovation" refers to adoption of a better technology at a substantial R\&D expense to a given firm.
-It is not salient whether similar ideas and/or technologies already exist elsewhere.
-Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
-This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{Aghion-QJE05,Vives-08}.
-
-Our decision rules differ in terms of rationality: from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
-The same distinctions also control the severity of competition between the principals: from cut-throat competition with \HardMax to a more relaxed competition with \HardMaxRandom, to an even more relaxed competition with \SoftMaxRandom.
-Indeed, with \HardMax you lose all customers as soon as you fall behind in performance, with \HardMaxRandom you get some small market share no matter what, and with \SoftMaxRandom you are further guaranteed a market share close to $\tfrac12$ as long as your performance is not much worse than the competition.
-The uniform choice among principals corresponds to no rationality and no competition.
-While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.
-
-We identify the inverted-U relationship driven by the rationality/competitiveness distinctions outlined above: from \HardMax to \HardMaxRandom to \SoftMaxRandom to \Uniform.
-We also find another, technically different, inverted-U relationship which zeroes in on the \HardMaxRandom model.
-We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
-
-These inverted-U relationships are driven by different aspects in our model than the ones in the existing literature in economics.
-The latter focuses on the tradeoff between the R\&D costs and the benefits that the improved technology provides in the competition.
-In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
-
 \xhdr{Numerical simulations.}
@@
 
-\item We investigate the algorithms' performance ``in isolation", \ie without competition.
-We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.\footnote{Note the discrepancy with our theoretical results on \HardMax: there, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations the prior is not available).}
-     We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes \emph{not} a good predictor for the outcomes under competition.
-
 \item We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}).
@@
 
-\ascomment{One downside of the story above: we may need to explain why we didn't simulate \SoftMaxRandom.}
+\vspace{0.2in}
 
-\ascomment{The xhdr below is basically from Guy, with some rewordings by Alex.}
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below.
+%
 
-\xhdr{Economic interpretation: network effects of data.}
-Our model speaks to policy discussions on regulating data-intensive digital platforms \citep{furman2019unlocking, scott2019committee}, and particularly to the ongoing debate on the role of data in the digital economy.
-One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.
-%whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
-The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved of users (resp., take over most of the market).
-%\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}
-A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.
-Thus, we do not explicitly model the network effects, but they arise endogenously from our setup.
+\addcontentsline{toc}{section}{References}
 
-Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
-Approaches such as \citet{lambrecht2015can, bajari2018impact, varian2018artificial} argue that the diminishing returns to scale and scope of data in isolation mitigate such data feedback loops,
-%as non-existent
-but ignore the differences induced by learning in isolation versus under competition.
-Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.
+\bibliographystyle{plainnat}
+%\begin{small}
+\bibliography{bib-abbrv,bib-ML,refs,bib-bandits,bib-AGT,bib-slivkins, bib-random}
+%\end{small}
 
-\ascomment{"Econ interpretation" xhdrs fit incely in "our results", as opposed to a separate subsection.
-Because: we can explain the "inverted-U" stuff right after "theory", before we put the simulation results on the stack.
-And the "network effects" stuff goes nicely right after the numerical results, without a big subsection break.}
+\clearpage
+%\subfile{content/old-intro}
 
-\subsection{Further discussion}
-\label{sec:discussion}
-
-\xhdr{Significance.}
-Our theory takes a basic Bayesian approach, a standard perspective in economic theory, and discovers several strong asymptotic results.
-The numerical simulations provide a more nuanced and ``non-asymptotic" perspective.
-In essence, we look for substantial effects within relevant time scales.
-In fact, we start our investigation by determining what time scales are relevant in the context of our model.
-
-\ascomment{The next para makes a point that seems important (to me). Is this point clear?}
-
-Our study has a dual purpose: (i) shed light on real-world implications of some typical scenarios, and (ii) investigate the space of models for describing the real world.
-As an example to clarify the latter point, consider the \HardMax model with simultaneous entry.
-It is is not necessarily the most realistic model, but arguably the most natural one to study \emph{a priori}.
-However, our results elucidate the need for more refined models that allow for ``free exploration" (\eg via random agents or early entry).
-
-\xhdr{Modeling.}
-Our models are stylized in several important  respects.
-Firms compete only on the quality of service, rather than, say, pricing or the range of products.
-Agents are myopic: they do not worry about how their actions impact their future utility.\footnote{In particular, agents do not attempt to learn over time, to second-guess or game future agents, or to manipulate the principals' learning algorithms.
-This is arguably typical in practice, in part because one agent's influence tends to be relatively small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores", and further simplified to average performance over a time window.
-On the machine learning side, our setup is geared to distinguish between ``simple" vs.
-``better" vs.
-``smart" bandit algorithms; we are not interested in state-of-art algorithms for very realistic bandit settings.
-
-We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs.
-For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior.
-Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
-Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
-It is unclear how to incorporate performance signals in a theoretically tractable model.
-For the numerical results, the \ExptsModel accounts for competition in a more direct way, allows to separate reputation vs.
-data advantage, and makes our model amenable to numerical simulations.
-Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.
-
-\asedit{Principals' utilities (and, accordingly, bandit rewards) are not discounted with time.%
-\footnote{However, our main theoretical result on the greedy algorithm extends to time-discounted rewards, see Section~\ref{sec:theory-extensions}.}
- While much of the early work focuses on  time-discounted bandits \citep{Gittins-book11}, non-discounted  formulations are prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
-Moreover, the distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.}
-
-\xhdr{Challenges.}
-Much of the challenge in the theoretical results, both conceptual and technical, was in setting up the model and the theorems.
-Apart from zeroing in on the \TheoryModel, it was crucial to interpret the results and intuitions from the literature on multi-armed bandits so as to formulate meaningful and productive assumptions on bandit algorithms and Bayesian priors.
-
-The numerical investigation is quite challenging even with a stylized model such as ours.
-An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given instance of a multi-armed bandit problem (and each such experiment is run many times to reduce variance).
-Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit instances\}, and we are looking for findings that are consistent across this entire space.
-It is essential to keep each of the three dimensions small yet representative.
-In particular, we need to capture a huge variety of bandit instances with only a few representative examples.
-Further, we need a succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).
-
-While amenable to simulations, the \ExptsModel appears difficult to analyze. This is for several reasons:
-%
-(i) intricate feedback loop from performance to reputations to users to performance;
-%
-(ii) mean reputation, most connected to our intuition, is sometimes a bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited}); and
-%
-(iii)
-mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice.
-Given the theoretical results on the \TheoryModel, and the fact that we are in the realm of stylized economic models, we believe that resolving first-order theoretical questions about the \ExptsModel would not add much value to this paper.
-
+\end{document}
 %%% Local Variables:
 %%% mode: latex
-%%% TeX-master: "main"
+%%% TeX-master: t
 %%% End:

### 2020-08-24 journal paper/content/related_work.tex
add keller-rady reference and address a couple of comments on intro

@@
 \citet{Bolton-econometrica99,Keller-econometrica05,Johari-ec12} studied models with self-interested agents jointly performing exploration, with no principal to coordinate them.
+\gaedit{\cite{keller2003price} studied a model where firms compete via prices where there is a similar exploration vs.
+competition tension.
+Besides the difference in competition for quality vs.
+prices, a crucial difference between our setting and theirs is that in their setting exploration decisions of firms have no bearing on future choices made by consumers.}
 

### 2020-08-24 journal paper/content/sec-intro.tex
add keller-rady reference and address a couple of comments on intro

@@
 
-We distinguish three classes of bandit algorithms, based on
-the prevalent intuition in the area.
+\gaedit{While our model allows for the firms to pick from an arbitrary set of algorithms}, we distinguish between three classes of bandit algorithms, based on the prevalent intuition in the area.
 The distinction concerns the way in which they resolve the fundamental tradeoff between  exploration and \emph{exploitation} (making optimal myopic decisions using the available data).
-Going from more primitive to more sophisticated, these three classes are as follows:
+\gaedit{This classification is important for interpreting the equilibrium algorithm choice of the firms as it allows for a ``low" to ``medium" to ``high" technology interpretation.
+We further use the classification to pick a representative algorithm from each class for our numerical experiments.
+Going from low to medium to high, these three classes are as follows:} \gadelete{Going from more primitive to more sophisticated, these three classes are as follows:}
 
@@
 
-\subsection{Our Results}
+\subsection{Our results}
 \label{sec:intro-results}
@@
 Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
-Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
+It is important that  Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
 It is unclear how to incorporate performance signals in a theoretically tractable model.
-For the numerical results, the \ExptsModel accounts for competition in a more direct way, allows to separate reputation vs.
-data advantage, and makes our model amenable to numerical simulations.
-Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.
+For the numerical results, the \ExptsModel accounts for competition in a more direct way \gaedit{by positing a more realistic model of consumer behavior that does not require them to have direct information on the algorithms deployed by or the bandit problem faced by the firm.
+It further allows us to separate reputation vs.
+data advantage and makes our model amenable to numerical simulations.} Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.
 
@@
 Moreover, the distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.}
+
+\gaedit{Finally, the assumption that firms commit to the algorithms at the start of the game deserves further justification.
+We impose this commitment assumption since it is more realistic than considering dynamic strategies and algorithms are, by definition, a commitment to a set of rules.
+In many contexts where such exploration algorithms are typically deployed, consumers arrive at such a rapid volume that it it is infeasible to consider a firm dynamically adapting its algorithm to each consumer that arrives.
+Indeed, \cite{DS-arxiv} points out how difficult it is to deploy standard exploration algorithms in practice and such dynamic strategies are substantially more difficult to deploy in practice.
+While it is possible to consider share-adaptive algorithms that would still keep the commitment assumption but allow for indirect dynamic responses to market conditions, such algorithms do not exist currently in the literature but are a good avenue for future work.}
 

### 2020-08-24 journal paper/content/sec-intro.tex
update

@@
 We further use the classification to pick a representative algorithm from each class for our numerical experiments.
-Going from low to medium to high, these three classes are as follows:} \gadelete{Going from more primitive to more sophisticated, these three classes are as follows:}
+Going from low to medium to high technology, these three classes are as follows:} \gadelete{Going from more primitive to more sophisticated, these three classes are as follows:}
 
@@
 For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior.
-Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
-It is important that  Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
+Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.\gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}}  Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
 It is unclear how to incorporate performance signals in a theoretically tractable model.
-For the numerical results, the \ExptsModel accounts for competition in a more direct way \gaedit{by positing a more realistic model of consumer behavior that does not require them to have direct information on the algorithms deployed by or the bandit problem faced by the firm.
+For the numerical results, the \ExptsModel accounts for competition in a more direct way \gaedit{by positing a more realistic model of consumer behavior that does not require them to have direct information on the algorithms deployed or the bandit problem faced by the firm.
 It further allows us to separate reputation vs.

### 2020-08-24 journal paper/content/sec-intro.tex
move comment to model seciton

@@
 For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior.
-Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.\gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}}  Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
+Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
+Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
 It is unclear how to incorporate performance signals in a theoretically tractable model.

### 2020-08-24 journal paper/content/sec-model.tex
Added a label.
Changed the placement of a figure.
Moved "background" to Appendix A.
move comment to model seciton

@@
 
-\xhdr{Agents' response.} Each agent $t$ forms a reward estimate $\est_i(t)\in [0,1]$ for each principal $i$, and chooses the principal based on these reward estimates.
-Specifically, principal $1$ is chosen with probability
-\begin{align}
+\xhdr{Agents' response.} Each agent $t$ forms a reward estimate $\est_i(t)\in [0,1]$ for each principal $i$, and chooses the principal based on these reward estimates.gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}} Specifically, principal $1$ is chosen with probability
+\begin{align}\label{eq:model-f}
 p_t = \respF\rbr{ \est_1(t) - \est_2(t) },
@@
 \item In the \emph{\TheoryModel}, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$.
-Agents know the principals' algorithms.
-Their reward estimates are defined as posterior mean rewards:
+Agents know the principals' algorithms.\gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}} Their reward estimates are defined as posterior mean rewards:
         $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i }$

### 2020-08-24 journal paper/code_abstract.tex
get rid of technology section and bring back data network effect section

@@
 
-\subsection*{Our model: competing bandits}
+\subsection*{Our Model: competing bandits}
 \label{sec:intro-model}
@@
 \noindent We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
-In all variants, agents have little or no information about other agents' choices and rewards.
 
@@
 
-\xhdr{Technology: multi-armed bandit algorithms.}
-To flesh out \eqref{eq:main-Q}, let us elaborate what we mean by `better' bandit algorithms.
-In general, comparisons between bandit algorithms are rather subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
-
-We distinguish three classes of bandit algorithms, based on
-the prevalent intuition in the area.
-The distinction concerns the way in which they resolve the fundamental tradeoff between  exploration and \emph{exploitation} (making optimal myopic decisions using the available data).
-Going from more primitive to more sophisticated, these three classes are as follows:
-
-\begin{itemize}
-\item \emph{Greedy algorithms} that strive to maximize the reward for the next round given the available information. Thus, they always ``exploit" and never explicitly ``explore".
-
-\item \emph{Exploration-separating algorithms}
- that separate exploration and exploitation: essentially, each round is dedicated to one and completely ignores the other.
-
-\item \emph{Adaptive-exploration} algorithms that combine exploration and exploitation, and gradually sway the exploration choices towards more promising alternatives.
-\end{itemize}
-
-In isolation, \ie in the absence of competition, these classes are fairly well-understood.
-Greedy algorithms are terrible for a wide variety of problem instances, precisely because they never explore.
-Exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances.
-Adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances.
-Generally,  ``better" algorithms are better in the long run, but could be worse initially.
-
 \subsection*{Our Results}
@@
 We offer a mix of theoretical results and numerical simulations.
-We are mainly interested in qualitative differences between the three classes of algorithms in Section~\ref{sec:intro-model}.
-For numerical simulations, we pick one representative algorithm from each class.
-For theoretical results, we allow arbitrary algorithms and focus on asymptotic differences in the algorithms' performance.
 
@@
 
+\xhdr{Economic interpretation: network effects of data.}
+Our model speaks to policy discussions on regulating data-intensive digital platforms \citep{furman2019unlocking, scott2019committee}, and particularly to the ongoing debate on the role of data in the digital economy.
+One fundamental question in this debate is whether data can serve a similar role as traditional ``network effects", creating scenarios when only one firm can function in the market \citep{Rysman09, jullien2019economics}.
+%whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
+The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved of users (resp., take over most of the market).
+%\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}
+A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.
+Thus, we do not explicitly model the network effects, but they arise endogenously from our setup.
+
+Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
+Approaches such as \citet{lambrecht2015can, bajari2018impact, varian2018artificial} argue that the diminishing returns to scale and scope of data in isolation mitigate such data feedback loops,
+%as non-existent
+but ignore the differences induced by learning in isolation versus under competition.
+Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.
+
 \vspace{0.2in}

### 2020-08-25 journal paper/code_abstract.tex
Full pass.

@@
 The Perils of Exploration under Competition%
-\thanks{This is a merged and final version of two conference papers,
-\citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19},
-with a unified and streamlined presentation and expanded background materials.}}
+\thanks{This is an extended abstract of \citet{CompetingBandits-merged}, which in turn subsumes conference papers
+\citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}.}}
 % \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
@@
 Zhiwei Steven Wu%
-\footnote{University of Minnesota - Twin Cities, Department of Computer Science. Email: zsw@umn.edu\newline
-Part of the research was done when Z.S. Wu was an intern and a postdoc at Microsoft Research NYC.
+\footnote{Carnegie-Mellon University, Pittsburgh, PA.
+Email: zstevenwu@cmu.edu. 
+Partially done as an intern and a postdoc at MSR-NYC.
 }}
@@
 
-The main high-level question that we focus on in this paper is:
-\begin{align}\label{eq:main-Q}
+The main high-level question is:
+%\begin{align}\label{eq:main-Q}
 \textbf{How does competition incentivize the adoption of better exploration algorithms?}
-\end{align}
-
-\noindent This translates into a number of more concrete questions.
+%\end{align}
+This translates into a number of more concrete questions.
 While it is commonly assumed that better technology always helps, is this so under competition?
@@
 --- and how they relate to the anti-trust considerations?
-Finding formalizations that admit meaningful answers is a major part of the overall challenge.
-
+%Finding formalizations that admit meaningful answers is a major part of the overall challenge.
 To answer these questions, we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
@@
 Instead, we model the algorithms explicitly, and investigate how they play out in competition over an extended period of time.
+We offer a mix of theoretical results and numerical simulations, in a range of models.
 
@@
 
-\subsection*{Our Model: competing bandits}
-\label{sec:intro-model}
+%\subsection*{Our Model: competing bandits}
+%\label{sec:intro-model}
 
 %We investigate these questions with
-\xhdr{Competition game.} We consider a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
+\xhdr{Model: competition game.} We consider a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
 We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
@@
 Each piece can get rather complicated in isolation, let alone jointly, so we strive for simplicity.
-Thus, the key features of our model are as follows:
+%Thus, the key features of our model are as follows.
+%\begin{itemize}
 
-\begin{itemize}
-
-\item A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals.
+\textbf{(i)} 
+A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals.
 The principal chooses an action (\eg a list of web search results to show to the agent), the user experiences this action, and reports a reward.
@@
 
-\item Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
+\textbf{(ii)}
+Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
 \emph{arms}) and receives a reward drawn independently from a fixed distribution specific to this action.
-The reward distributions are initially unknown (but can be estimated over time from the data).
+The reward distributions are initially unknown.
+% (but can be estimated over time from the data).
 
-\item Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
+\textbf{(iii)}
+Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
 Each principal's objective is to maximize its market share (the fraction of users choosing this principal).
@@
 %, but may have access to each platforms' reputation score (more on this in    Section~\ref{sec:intro-discussion}).
-\end{itemize}
+%\end{itemize}
 
-\noindent We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
+We investigate several model variants, as we vary agents' decision rule and/or posit a first-mover advantage.
 
@@
 
-\subsection*{Our Results}
-\label{sec:intro-results}
+\xhdr{Technology: multi-armed bandit algorithms.}
+To compare between bandit algorithms, we build on prevalent intuition in the literature.
+We focus on standard notions of regret, and distinguish between three classes of bandit algorithms: ones that never explicitly explore (\emph{greedy algorithms}), ones that explore without looking at the data (\emph{exploration-separating algorithms}), and ones where exploration gradually zooms in on the best arm (\emph{adaptive-exploration algorithms}).
+In isolation, \ie in the absence of competition, these classes are fairly well-understood.
+Greedy algorithms are terrible for a wide variety of problem instances, precisely because they never explore.
+Exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances.
+Adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances.
+Generally,  ``better" algorithms are better in the long run, but could be worse initially.
 
-We offer a mix of theoretical results and numerical simulations.
+%\subsection*{Our Results}
+%\label{sec:intro-results}
 
@@
 %We endow agents with Bayesian rationality, a common modeling approach for a theoretical investigation.
-We consider a basic Bayesian model.
-We posit that agents have a common Bayesian prior on reward distributions, and know the principals' algorithms.
-For simplicity, agents do not receive any information about the previous agents' choices and rewards.
+We consider a basic Bayesian model, where agents have a common Bayesian prior on reward distributions, and know the principals' algorithms.
+For tractability, we posit that agents do not receive any information about the previous agents' choices and rewards.
 Each agent knows the round (s)he arrives in, computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
-We refer to this variant as the \emph{\TheoryModel}.
+%We refer to this variant as the \emph{\TheoryModel}.
 
-Our results depend crucially on the agents' decision rule.
-\begin{itemize}
+Our results depend crucially on the agents' decision rule:
+%\begin{itemize}
 
-\item The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
+\textbf{(i)}  
+The most obvious decision rule (\HardMax) maximizes the Bayesian-expected reward.
 We find that \HardMax is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
@@
 
-\item We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
+\textbf{(ii)} 
+We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
 We call this model \HardMaxRandom.
@@
 
-\item We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
-For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
+\textbf{(iii)} 
+We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
+Then, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
 This is the most technical result of the paper.
 The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm would attract slightly more).
-\end{itemize}
+
+\xhdr{Economic interpretation: the inverted-U relationship.}
+Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the \emph{inverted-U relationship} between competition and innovation.% 
+%\footnote{Here, ``innovation" refers to adoption of a better technology at a substantial R\&D expense to a given firm.
+It is not salient whether similar ideas and/or technologies already exist elsewhere.
+Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
+\footnote{This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{aghion2005competition,Vives-08}.}
+
+Our decision rules differ in terms of rationality: from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
+The same distinctions also control the severity of competition between the principals: from cut-throat competition with \HardMax to a more relaxed competition with \HardMaxRandom, to an even more relaxed competition with \SoftMaxRandom.
+Indeed, with \HardMax you lose all customers as soon as you fall behind in performance, with \HardMaxRandom you get some small market share no matter what, and with \SoftMaxRandom you are further guaranteed a market share close to $\tfrac12$ as long as your performance is not much worse than the competition.
+The uniform choice among principals corresponds to no rationality and no competition.
+While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.
+
+We identify the inverted-U relationship driven by the rationality/competitiveness distinctions outlined above: from \HardMax to \HardMaxRandom to \SoftMaxRandom to \Uniform.
+We also find another, technically different, inverted-U relationship which zeroes in on the \HardMaxRandom model.
+We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
+
+These inverted-U relationships are driven by different aspects in our model than the ones in the existing literature in economics.
+The latter focuses on the tradeoff between the R\&D costs and the benefits that the improved technology provides in the competition.
+In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
 
@@
 Thus, agents' decision rule depends only on the two reputation scores.
-We refer to this variant as the \emph{\ExptsModel}.%
+%We refer to this variant as the \emph{\ExptsModel}.%
 %\footnote{In comparison, the theoretical results focus on another extreme, with Bayesian rationality and no performance signals.}
-
 We refine and expand the theoretical results in several ways.
 
-\begin{itemize}
-\item We compare \HardMax and \HardMaxRandom decision rules. We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.
+%\begin{itemize}
+\textbf{(i)}  
+We compare \HardMax and \HardMaxRandom decision rules. We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.
  As predicted by the theory, better algorithms prevail if the expected number of ``random" users is sufficiently large. However, this effect is negligible for smaller parameter values.
@@
 
-\item We investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
+\textbf{(ii)}  
+ We investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
 (We focus on the \HardMax decision rule.) We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous arrival).
@@
 
-\item We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}).
+\textbf{(iii)}  
+We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}).
 We run additional experiments so as to isolate and compare these two effects.
@@
 
-\end{itemize}
+%\end{itemize}
 
@@
 
-\vspace{0.2in}
+\newpage
 

### 2020-08-25 journal paper/code_abstract.tex
slight edits to abstract

@@
 \textbf{(i)}  
-The most obvious decision rule (\HardMax) maximizes the Bayesian-expected reward.
+The most obvious decision rule (\HardMax) maximizes the current agent's Bayesian-expected reward.
 We find that \HardMax is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
-Further, we show that \HardMax is very sensitive to tie-breaking: if the tie-breaking is probabilistically biased in favor of one principal, then this principal has a simple ``winning strategy" no matter what the other principal does.
+Further, we show that \HardMax is sensitive to tie-breaking: if the tie-breaking is probabilistically biased in favor of one principal, then this principal has a simple ``winning strategy" no matter what.
 
@@
 
+\xhdr{Significance.}
+Our theory takes a Bayesian approach and discovers several strong asymptotic results.
+The numerical simulations provide a more nuanced and ``non-asymptotic" perspective.
+In essence, we look for substantial effects within relevant time scales.
+In fact, we start our investigation by determining what time scales are relevant in the context of our model.
+Our study has a dual purpose: (i) shed light on real-world implications of some typical scenarios, and (ii) investigate the space of models for describing the real world.
+As an example to clarify the latter point, consider the \HardMax model with simultaneous entry.
+It is is not necessarily the most realistic model.
+However, our results elucidate the need for more refined models that allow for ``free exploration" (\eg via random agents or early entry).
+
 \newpage

### 2020-08-26 journal paper/code_abstract.tex
Minor edit

@@
 
-Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
+Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely because competition leads to the endogenous generation of data observed by the firms.
 Approaches such as \citet{lambrecht2015can, bajari2018impact, varian2018artificial} argue that the diminishing returns to scale and scope of data in isolation mitigate such data feedback loops,

### 2020-09-02 journal paper/journal_main.tex
Removed \newpage statements.
Suppressed color-edits.
Restored color-edits.
Conclusion -> conclusions.
add keller-rady reference and address a couple of comments on intro
update
Minor formatting edit in the intro.

@@
 Zhiwei Steven Wu%
-\footnote{University of Minnesota - Twin Cities, Department of Computer Science. Email: zsw@umn.edu\newline
-Part of the research was done when Z.S. Wu was an intern and a postdoc at Microsoft Research NYC.
-}}
-\date{July 2020}
+\footnote{Carnegie-Mellon University, Pittsburgh, PA.
+Email: zstevenwu@cmu.edu.\newline
+Research done when Z.S. Wu was an intern and a postdoc at Microsoft Research NYC.}
+}
+\date{August 2020}
 \maketitle
@@
 
-\newpage
 \section{Related work}
@@
 
-\newpage
 \section{Theoretical results: the \TheoryModel}
@@
 \subfile{content/sec-theory}
-\newpage
 
@@
 \newpage
-\gaedit{
 \subfile{content/conclusion}
-}
 
@@
 
-\newpage
 \section{Full proofs for Section~\ref{sec:theory}}
@@
 
-\newpage
 \section{Full experimental results}

### 2020-09-02 journal paper/content/sec-intro.tex
Revising the intro ...

@@
 Many platforms purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a standard and well-studied framework for exploration and making decisions under uncertainty
-\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
+Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a standard and well-studied framework for exploration and making decisions under uncertainty.
+%\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
 
@@
 
-In this paper, we initiate a study of the interplay between \exploration and \competition.
+In this paper, we study the interplay between \exploration and \competition.
 Platforms that engage in exploration typically need to compete against one another; most importantly, they compete for users.
@@
 
-The main high-level question that we focus on in this paper is:
-\begin{align}\label{eq:main-Q}
+The main high-level question is: 
+%that we focus on in this paper is:
+%\begin{align}\label{eq:main-Q}
 \textbf{How does competition incentivize the adoption of better exploration algorithms?}
-\end{align}
-
-\noindent This translates into a number of more concrete questions.
+%\end{align}
+This translates into a number of more concrete questions.
 While it is commonly assumed that better technology always helps, is this so under competition?
@@
 Instead, we model the algorithms explicitly, and investigate how they play out in competition over an extended period of time.
+We offer a mix of theoretical results and numerical simulations.
 
@@
 
-\subsection{Our model: competing bandits}
-\label{sec:intro-model}
+%\subsection{Our model: competing bandits}
+%\label{sec:intro-model}
 
 %We investigate these questions with
-\xhdr{Competition game.} We consider a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
+\xhdr{Our model: Competition game.} We consider a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
 We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
@@
 
-\begin{itemize}
+%\begin{itemize}
 
-\item A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals.
+\textbf{(i)} A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals.
 The principal chooses an action (\eg a list of web search results to show to the agent), the user experiences this action, and reports a reward.
@@
 
-\item Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
+\textbf{(ii)} Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
 \emph{arms}) and receives a reward drawn independently from a fixed distribution specific to this action.
-The reward distributions are initially unknown (but can be estimated over time from the data).
+The reward distributions are initially unknown.
+% (but can be estimated over time from the data).
 
-\item Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
+\textbf{(iii)} Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
 Each principal's objective is to maximize its market share (the fraction of users choosing this principal).
@@
 %, but may have access to each platforms' reputation score (more on this in    Section~\ref{sec:intro-discussion}).
-\end{itemize}
+%\end{itemize}
 
-\noindent We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
+We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
 In all variants, agents have little or no information about other agents' choices and rewards.
@@
 The former variant used for the theoretical results, the latter for simulations.
+
+\ascomment{revising ...}
+
+\xhdr{Technology: multi-armed bandit algorithms.}
+To compare between bandit algorithms, we build on prevalent intuition in the literature.
+We focus on standard notions of regret, and distinguish between three classes of bandit algorithms: ones that never explicitly explore (\emph{greedy algorithms}), ones that explore without looking at the data (\emph{exploration-separating algorithms}), and ones where exploration gradually zooms in on the best arm (\emph{adaptive-exploration algorithms}).
+In the absence of competition, these classes are fairly well-understood: greedy algorithms are terrible for a wide variety of problem instances, exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances, and adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances.
+Generally,  ``better" algorithms are better in the long run, but could be worse initially.
 

### 2020-09-02 journal paper/content/sec-intro.tex
Continuing the revision ...

@@
 %Finding formalizations that admit meaningful answers is a major part of the overall challenge.
-To answer these questions, we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
+
+We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
 The choice of a particular technology (exploration algorithm) is no longer an abstract, static choice with a predetermined outcome for the platform.
 Instead, we model the algorithms explicitly, and investigate how they play out in competition over an extended period of time.
-We offer a mix of theoretical results and numerical simulations.
 
@@
 
-\textbf{(ii)} Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
+\textbf{(ii)} Each principal
+%learning problem is as follows: 
+faces a basic and well-studied version of the multi-armed bandit problem: 
+for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
 \emph{arms}) and receives a reward drawn independently from a fixed distribution specific to this action.
@@
 
+We draw on the rich literature on multi-armed bandits to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.%
+\footnote{Comparisons between bandit algorithms are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+More on this in Preliminaries.} One baseline is the \emph{greedy algorithm} that does not explicitly explore; in the absence of competition, it is known to be terrible for a wide variety of problem instances.
+
 We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
@@
 
-\ascomment{revising ...}
+%\ascomment{revising ...}
 
+\OMIT{ %%%%%%%%%% MIT CODE abstract:
 \xhdr{Technology: multi-armed bandit algorithms.}
@@
 Generally,  ``better" algorithms are better in the long run, but could be worse initially.
+} %%%%%%%%%%%
 
+\OMIT{ %%%%%%%%%% PREVIOUS VERSION
 \xhdr{Technology: multi-armed bandit algorithms.}
@@
 While we build on a vast and active research area, we present sufficient background in Appendix~\ref{app:bg} so as to keep this paper accessible to non-specialists.
+} %%%%%%%%%
 
-\subsection{Our results}
-\label{sec:intro-results}
+%\subsection{Our results}
+%\label{sec:intro-results}
 
-We offer a mix of theoretical results and numerical simulations.
+%We offer a mix of theoretical results and numerical simulations.
 We are mainly interested in qualitative differences between the three classes of algorithms in Section~\ref{sec:intro-model}.
@@
 In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
+
+\ascomment{stopped here.}
 

### 2020-09-03 journal paper/content/sec-intro.tex
A pass on the intro: compressed the model & algorithms, expanded on dynamic vs static strategies. Moved "discussion of the model" to the model section.

@@
 %We investigate these questions with
-\xhdr{Our model: Competition game.} We consider a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
-We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
-These two processes are interlinked, as exploration decisions are experienced by users and informed by their feedback.
+\xhdr{Our model: competition game.} We consider a stylized duopoly model in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
+Principals compete on quality rather than on prices.
+Agents arrive sequentially, and choose among the two principals according to some decision rule.
+The chosen principal chooses an action (\eg a list of web search results to show to the user), the user experiences this action, and reports a reward.
+Each principal only observes its own users.
+Principals commit to their strategies simultaneously and in advance, with objective to maximize the market share.
+We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
+In all variants, agents have little or no information about other agents' choices and rewards.
+
+%These two processes are interlinked, as exploration decisions are experienced by users and informed by their feedback.
 We need to specify several conceptual pieces: how the principals and agents interact, what is the machine learning problem faced by each principal, and what is the information structure.
@@
 %\begin{itemize}
-
-\textbf{(i)} A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals.
+%\textbf{(i)} A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals.
 The principal chooses an action (\eg a list of web search results to show to the agent), the user experiences this action, and reports a reward.
@@
 
-\textbf{(ii)} Each principal
-%learning problem is as follows: 
-faces a basic and well-studied version of the multi-armed bandit problem: 
-for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
+%\textbf{(ii)} Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
 \emph{arms}) and receives a reward drawn independently from a fixed distribution specific to this action.
@@
 
-\textbf{(iii)} Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
+%\textbf{(iii)} Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
 Each principal's objective is to maximize its market share (the fraction of users choosing this principal).
@@
 
-We draw on the rich literature on multi-armed bandits to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.%
+Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
+Each principal's strategy is a bandit algorithm; as such, it adjusts itself to the observed rewards, but not to the competition.
+This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning.
+Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
+We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.%
 \footnote{Comparisons between bandit algorithms are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
-More on this in Preliminaries.} One baseline is the \emph{greedy algorithm} that does not explicitly explore; in the absence of competition, it is known to be terrible for a wide variety of problem instances.
-
-We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
-In all variants, agents have little or no information about other agents' choices and rewards.
+Generally,  ``better" algorithms are better in the long run, but could be worse initially.
+More on this in Preliminaries.} One baseline is \emph{greedy algorithms} that do not explicitly explore; in the absence of competition, it is known to be terrible for a wide variety of problem instances.
 
@@
 We refer to this variant as the \emph{\TheoryModel}.
+We focus on qualitative, asymptotic differences in the algorithms' performance.
 
@@
 
-\ascomment{stopped here.}
-
 \xhdr{Numerical simulations.}
@@
 
-\ascomment{One downside of the story above: we may need to explain why we didn't simulate \SoftMaxRandom.}
-
-\ascomment{The xhdr below is basically from Guy, with some rewordings by Alex.}
-
 \xhdr{Economic interpretation: network effects of data.}
@@
 
-\ascomment{"Econ interpretation" xhdrs fit incely in "our results", as opposed to a separate subsection.
-Because: we can explain the "inverted-U" stuff right after "theory", before we put the simulation results on the stack.
-And the "network effects" stuff goes nicely right after the numerical results, without a big subsection break.}
-
-\subsection{Further discussion}
-\label{sec:discussion}
-
 \xhdr{Significance.}
@@
 
-\ascomment{The next para makes a point that seems important (to me). Is this point clear?}
-
 Our study has a dual purpose: (i) shed light on real-world implications of some typical scenarios, and (ii) investigate the space of models for describing the real world.
@@
 However, our results elucidate the need for more refined models that allow for ``free exploration" (\eg via random agents or early entry).
-
-\xhdr{Modeling.}
-Our models are stylized in several important  respects.
-Firms compete only on the quality of service, rather than, say, pricing or the range of products.
-Agents are myopic: they do not worry about how their actions impact their future utility.\footnote{In particular, agents do not attempt to learn over time, to second-guess or game future agents, or to manipulate the principals' learning algorithms.
-This is arguably typical in practice, in part because one agent's influence tends to be relatively small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores", and further simplified to average performance over a time window.
-On the machine learning side, our setup is geared to distinguish between ``simple" vs.
-``better" vs.
-``smart" bandit algorithms; we are not interested in state-of-art algorithms for very realistic bandit settings.
-
-We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs.
-For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior.
-Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
-Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
-It is unclear how to incorporate performance signals in a theoretically tractable model.
-For the numerical results, the \ExptsModel accounts for competition in a more direct way \gaedit{by positing a more realistic model of consumer behavior that does not require them to have direct information on the algorithms deployed or the bandit problem faced by the firm.
-It further allows us to separate reputation vs.
-data advantage and makes our model amenable to numerical simulations.} Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.
-
-\asedit{Principals' utilities (and, accordingly, bandit rewards) are not discounted with time.%
-\footnote{However, our main theoretical result on the greedy algorithm extends to time-discounted rewards, see Section~\ref{sec:theory-extensions}.}
- While much of the early work focuses on  time-discounted bandits \citep{Gittins-book11}, non-discounted  formulations are prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
-Moreover, the distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.}
-
-\gaedit{Finally, the assumption that firms commit to the algorithms at the start of the game deserves further justification.
-We impose this commitment assumption since it is more realistic than considering dynamic strategies and algorithms are, by definition, a commitment to a set of rules.
-In many contexts where such exploration algorithms are typically deployed, consumers arrive at such a rapid volume that it it is infeasible to consider a firm dynamically adapting its algorithm to each consumer that arrives.
-Indeed, \cite{DS-arxiv} points out how difficult it is to deploy standard exploration algorithms in practice and such dynamic strategies are substantially more difficult to deploy in practice.
-While it is possible to consider share-adaptive algorithms that would still keep the commitment assumption but allow for indirect dynamic responses to market conditions, such algorithms do not exist currently in the literature but are a good avenue for future work.}
 

### 2020-09-03 journal paper/content/sec-model.tex
A pass on the intro: compressed the model & algorithms, expanded on dynamic vs static strategies. Moved "discussion of the model" to the model section.

@@
 Principals are risk-neutral, in the sense that they optimize their expected utility.
+
+\subsection{Discussion}
+\label{sec:discussion}
+
+\ascomment{copied from the intro, to be expanded}
+
+Our models are stylized in several important  respects.
+Firms compete only on the quality of service, rather than, say, pricing or the range of products.
+Agents are myopic: they do not worry about how their actions impact their future utility.\footnote{In particular, agents do not attempt to learn over time, to second-guess or game future agents, or to manipulate the principals' learning algorithms.
+This is arguably typical in practice, in part because one agent's influence tends to be relatively small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores", and further simplified to average performance over a time window.
+On the machine learning side, our setup is geared to distinguish between ``simple" vs.
+``better" vs.
+``smart" bandit algorithms; we are not interested in state-of-art algorithms for very realistic bandit settings.
+
+We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs.
+For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior.
+Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
+Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
+It is unclear how to incorporate performance signals in a theoretically tractable model.
+For the numerical results, the \ExptsModel accounts for competition in a more direct way \gaedit{by positing a more realistic model of consumer behavior that does not require them to have direct information on the algorithms deployed or the bandit problem faced by the firm.
+It further allows us to separate reputation vs.
+data advantage and makes our model amenable to numerical simulations.} Indeed, simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable.
+
+\asedit{Principals' utilities (and, accordingly, bandit rewards) are not discounted with time.%
+\footnote{However, our main theoretical result on the greedy algorithm extends to time-discounted rewards, see Section~\ref{sec:theory-extensions}.}
+ While much of the early work focuses on  time-discounted bandits \citep{Gittins-book11}, non-discounted  formulations are prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
+Moreover, the distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.}
+
+\gaedit{Finally, the assumption that firms commit to the algorithms at the start of the game deserves further justification.
+We impose this commitment assumption since it is more realistic than considering dynamic strategies and algorithms are, by definition, a commitment to a set of rules.
+In many contexts where such exploration algorithms are typically deployed, consumers arrive at such a rapid volume that it it is infeasible to consider a firm dynamically adapting its algorithm to each consumer that arrives.
+Indeed, \cite{DS-arxiv} points out how difficult it is to deploy standard exploration algorithms in practice and such dynamic strategies are substantially more difficult to deploy in practice.
+While it is possible to consider share-adaptive algorithms that would still keep the commitment assumption but allow for indirect dynamic responses to market conditions, such algorithms do not exist currently in the literature but are a good avenue for future work.}

### 2020-09-03 journal paper/journal_main.tex
Changed the font size and margins, adjusted the title page.

@@
-\documentclass[11pt]{article}
+\documentclass[12pt]{article}
+\usepackage[margin=1.25in]{geometry}
 
@@
 }
-
-\usepackage{fullpage}
-%\usepackage{geometry}
-%\geometry{left=1.0in,right=1.0in,top=1.0in,bottom=1.0in}
 
@@
 
-\title{Competing Bandits:\\
+\title{\vspace{-12mm}Competing Bandits:\\
 The Perils of Exploration under Competition%
@@
 \and
-%\rule{0.2in}{0pt}
+\rule{0.2in}{0pt}
 Yishay Mansour%
@@
 \and
-%\rule{0.2in}{0pt}
+\rule{0.2in}{0pt}
 Aleksandrs Slivkins%
@@
 \and
-%\rule{0.2in}{0pt}
+\rule{0.2in}{0pt}
 Zhiwei Steven Wu%
@@
 }
-\date{August 2020}
+\date{First version: August 2020\\ This version: September 2020}
+
 \maketitle
+
+\vspace{-5mm}
 \begin{abstract}
-Most online platforms strive to learn from interactions with consumers, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-We initiate a study of the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for consumers.
-Here consumers play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
+Most online platforms strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
+We initiate a study of the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users.
+Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
 
@@
 Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen.
-We study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for consumers.
-We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low consumer welfare.
-However, we find that weakening competition by providing firms with some ``free" consumers incentivizes better exploration strategies and increases consumer welfare.
-We investigate two channels for weakening the competition: relaxing the rationality of consumers and giving one firm a first-mover advantage.
-We provide a mix of theoretical results and numerical simulations.
-
+Through a mix of theoretical results and numerical simulations, we study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for users.
+We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low welfare.
+However, weakening competition by providing firms with some ``free" users incentivizes better exploration strategies and increases welfare.
+We investigate two channels for weakening the competition: relaxing the rationality of users and giving one firm a first-mover advantage.
 Our findings are closely related to the ``competition vs.
-innovation" relationship, a well-studied theme in economics.
-They also elucidate the first-mover advantage in the digital economy by exploring the role that data can play as a barrier to entry in online markets.
+innovation" relationship, and elucidate the first-mover advantage in the digital economy.
 

### 2020-09-03 journal paper/content/sec-intro.tex
Lots of small edits because of the new font size.

@@
 
-Our results depend crucially on the agents' decision rule.
-\begin{itemize}
+Our results depend crucially on the agents' decision rule:
+%\begin{itemize}
 
-\item The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
+\textbf{(i)} The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
 We find that \HardMax is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
@@
 
-\item We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
+\textbf{(ii)} We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
 We call this model \HardMaxRandom.
@@
 
-\item We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
+\textbf{(iii)} We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
 For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
@@
 The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm would attract slightly more).
-\end{itemize}
+%\end{itemize}
 
@@
 
-\begin{itemize}
-\item We compare \HardMax and \HardMaxRandom decision rules. We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.
+%\begin{itemize}
+\textbf{(i)} We compare \HardMax and \HardMaxRandom decision rules. We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.
  As predicted by the theory, better algorithms prevail if the expected number of ``random" users is sufficiently large. However, this effect is negligible for smaller parameter values.
@@
 
-\item We investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
+\textbf{(ii)} We investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
 (We focus on the \HardMax decision rule.) We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous arrival).
@@
 
-\item We investigate the algorithms' performance ``in isolation", \ie without competition.
+\textbf{(iii)} We investigate the algorithms' performance ``in isolation", \ie without competition.
 We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.\footnote{Note the discrepancy with our theoretical results on \HardMax: there, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations the prior is not available).}
@@
 
-\item We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}).
+\textbf{(iv)} We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}).
 We run additional experiments so as to isolate and compare these two effects.

### 2020-09-03 journal paper/content/sec-model.tex
Lots of small edits because of the new font size.

@@
-Let us define the basic model; our theoretical results admit some extensions, detailed in Section~\ref{sec:theory-extensions}.
+Let us define the basic model (some extensions are detailed in Section~\ref{sec:theory-extensions}).
 
@@
 The principal chooses action $a_t\in A$, where $A$ is a fixed set of actions (same for both principals and all rounds).%
-\footnote{Throughout, we use `action' and `arm' interchangeably, as common in the literature on multi-armed bandits.}
+\footnote{We use `action' and `arm' interchangeably, as common in the literature on multi-armed bandits.}
  The agent experiences this action, receives a reward $r_t\in \{ 0,1\}$, and reports it back to the principal.
@@
 
-Each principal $i$ commits to a learning algorithm \alg[i] before round $1$, and uses this algorithm throughout.
+Each principal $i$ commits to a learning algorithm \alg[i] before round $1$, and uses it throughout.
 The algorithm follows the protocol of \emph{multi-armed bandits} (\emph{MAB}).
@@
 We will omit the global vs.
-local distinction when clear from the context.} each time it is called, it outputs an action from $A$, and inputs a reward for this action.
+local distinction when clear from the context.} each time it is called, it outputs an arm from $A$, and inputs a reward for this action.
 \alg[i] is called only in global rounds when principal $i$ is chosen.
@@
 where $\respF:[-1,1]\to [0,1]$ is the \emph{response function}, same for all agents and known to everyone.
+We assume that $\respF$ is monotonically non-decreasing, is larger than $\nicefrac12$ on the interval $(0,1]$, and smaller than $\nicefrac12$ on the interval $[-1,0)$.
 
-We assume that $\respF$ is monotonically non-decreasing, is larger than $\nicefrac12$ on the interval $(0,1]$, and smaller than $\nicefrac12$ on the interval $[-1,0)$.
 We consider three specific models (see \reffig{fig:response-functions}):
@@
 
-\item \SoftMaxRandom: $\respF(\cdot)$ is bounded away from $0$ and $1$, and ``smooth" around $0$ (see Section~\ref{sec:theory-SoftMax}).
+\item \SoftMaxRandom: $\respF$ lies in $[\eps_0,1-\eps_0]$ and is ``smooth" around $0$ (see Section~\ref{sec:theory-SoftMax}).
 \end{itemize}
@@
 \asedit{Principals' utilities (and, accordingly, bandit rewards) are not discounted with time.%
-\footnote{However, our main theoretical result on the greedy algorithm extends to time-discounted rewards, see Section~\ref{sec:theory-extensions}.}
+\footnote{However, our main theoretical result on the greedy algorithm extends to time-discounted rewards.}
  While much of the early work focuses on  time-discounted bandits \citep{Gittins-book11}, non-discounted  formulations are prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.

### 2020-09-03 journal paper/content/sec-theory-proofs.tex
Lots of small edits because of the new font size.

@@
 %%%%%%%%%%%%
-\subsection{For \HardMaxRandom, a little greedy goes a long way
+\subsection{A little greedy goes a long way
 (Proof of Theorem~\ref{thm:random-greedy})}
@@
 \end{align}
-Here inequality \eqref{eq:pf:thm:random-greedy-1} holds because $\rew_1(\cdot)$ is \bmonotone and $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$, and inequality \eqref{eq:pf:thm:random-greedy-2} holds because $\rew_1(n_0)<\rew_2(n_0)$ and $\mM_{2,t}(n_0)>0$.%
-\footnote{If $\rew_1(\cdot)$ is strictly increasing, then inequality \eqref{eq:pf:thm:random-greedy-1} is strict, too; this is because $\mM_{2,t}(t-1)>\posteriorN{1}{t}(t-1)$.  }
+Here \eqref{eq:pf:thm:random-greedy-1} holds because $\rew_1(\cdot)$ is \bmonotone and $\mM_{2,t}$ stochastically dominates $\posteriorN{1}{t}$, and inequality \eqref{eq:pf:thm:random-greedy-2} holds because $\rew_1(n_0)<\rew_2(n_0)$ and $\mM_{2,t}(n_0)>0$.%
+\footnote{If $\rew_1(\cdot)$ is strictly increasing, then  \eqref{eq:pf:thm:random-greedy-1} is strict, too; this is because $\mM_{2,t}(t-1)>\posteriorN{1}{t}(t-1)$.  }
 %\end{proof}
@@
 $$
-Note that the number of agents received by \alg[2] is at most
-$T_1 + (1/2 + \beta_1)t + (1/2 - \beta_1)(1 - \beta_0/5)t$.
+The number of agents received by \alg[2] is at most
+$T_1 + (\nicefrac{1}{2} + \beta_1)t + (\nicefrac{1}{2} - \beta_1)(1 - \beta_0/5)t$.
 

### 2020-09-03 journal paper/content/sec-theory.tex
Minor edit.
Lots of small edits because of the new font size.

@@
 
-\xhdr{Assumptions.} We make two mild assumptions on the prior. First, each arm $a$ can be the best:
+\xhdr{Assumptions.} We make two mild assumptions on the prior. First, each arm $a$ can be best:
 \begin{align}\label{eq:assn-prob}
 \forall a\in A:\;\;\;
-\Pr_{\mu\sim \priorMu}\sbr{\mu_a  > \mu_{a'}
-\;\forall a'\in A}
+\Pr\sbr{\mu_a  > \mu_{a'}
+\quad \forall a'\in A}
 > 0.
@@
 
-\noindent Second, actions' posterior mean rewards are pairwise distinct given any feasible history $h$:%
+\noindent Second, posterior mean rewards are pairwise distinct given any feasible history $h$:%
 \footnote{The \emph{history} of an MAB algorithm at a given step
@@
 \end{align}
-In particular, \emph{prior} mean rewards of actions are pairwise distinct:
+In particular, \emph{prior} mean rewards are pairwise distinct:
 $\E[\mu_a] \neq \E[\mu_a']$ for any $a,a'\in A$.
@@
 
-An algorithm is called \emph{\bmonotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing (and therefore $\BIR(\cdot)$ is non-decreasing). We argue this is a mild assumption, see
+An algorithm is called \emph{\bmonotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing, and therefore $\BIR(\cdot)$ is non-decreasing. This is a mild assumption, see
 Appendix~\ref{app:examples}.
@@
 \end{align}
-In fact, a weaker version of \eqref{eq:random-better-clean} suffices:
+In fact, a weaker version suffices:
 denoting $\eps_0 = \respF(-1)$, for some constant $n_0$ we have
@@
 
-We also need a mild technical assumption that $\BIR_2(\cdot)$ is not extremely small: for some constant $m_0$,
+We also need a mild technical assumption that $\BIR_2(\cdot)$ is not extremely small: 
+%for some constant $m_0$,
 \begin{align}\label{eq:random-assn}
- (\forall n\geq m_0) \qquad
+ (\exists m_0\;\forall n\geq m_0) \qquad
   \BIR_2(n) > 4\,e^{-\eps_0\ n/12}.
@@
 
-We find that the greedy modification prevails in competition if the mixing parameter is small enough.
+We find that the greedy modification prevails in competition if $p$ is small enough.
 
@@
 Whenever arm $a$ is chosen, the reward is drawn independently from distribution $\psi_a(\mu_a)$.
-The prior $\priorMu$ and the distributions $(\psi_a(\cdot)\colon a\in A)$ constitute the (full) Bayesian prior on rewards, denoted $\prior$.
+The prior $\priorMu$ and the distributions $(\psi_a(\cdot)\colon a\in A)$ constitute the (full) Bayesian prior on rewards.% denoted $\prior$.
 

### 2020-09-03 journal paper/content/sec-intro.tex
Commented out inverted-U figure.

@@
 
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
-      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
-      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
-      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
- \end{tikzpicture}
-
-\caption{Inverted-U relationship between competitiveness and algorithms.}
-\label{fig:inverted-U}
-\end{center}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\begin{tikzpicture}
+%      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
+%      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
+%      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
+%      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
+% \end{tikzpicture}
+%
+%\caption{Inverted-U relationship between competitiveness and algorithms.}
+%\label{fig:inverted-U}
+%\end{center}
+%\end{figure}
 
 \xhdr{Economic interpretation: the inverted-U relationship.}
-Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the inverted-U relationship between competition and innovation, as in Figure~\ref{fig:inverted-U}.%
+Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the inverted-U relationship between competition and innovation.%
+%, as in Figure~\ref{fig:inverted-U}.
 \footnote{Here, ``innovation" refers to adoption of a better technology at a substantial R\&D expense to a given firm.
@@
 \textbf{(iii)} We investigate the algorithms' performance ``in isolation", \ie without competition.
-We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.\footnote{Note the discrepancy with our theoretical results on \HardMax: there, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations the prior is not available).}
+We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).}
      We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes \emph{not} a good predictor for the outcomes under competition.

### 2020-09-03 journal paper/content/sec-intro.tex
More small edits.

@@
 %Finding formalizations that admit meaningful answers is a major part of the overall challenge.
+We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
 
-We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
-The choice of a particular technology (exploration algorithm) is no longer an abstract, static choice with a predetermined outcome for the platform.
+%The choice of a particular technology (exploration algorithm) not an abstract, static choice with a predetermined outcome for the platform.
 Instead, we model the algorithms explicitly, and investigate how they play out in competition over an extended period of time.
@@
 Generally,  ``better" algorithms are better in the long run, but could be worse initially.
-More on this in Preliminaries.} One baseline is \emph{greedy algorithms} that do not explicitly explore; in the absence of competition, it is known to be terrible for a wide variety of problem instances.
+More on this in Preliminaries.} One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.
 
@@
 %\begin{itemize}
-\textbf{(i)} We compare \HardMax and \HardMaxRandom decision rules. We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.
- As predicted by the theory, better algorithms prevail if the expected number of ``random" users is sufficiently large. However, this effect is negligible for smaller parameter values.
+\textbf{(i)} 
+%We compare \HardMax and \HardMaxRandom decision rules. 
+We find that the greedy algorithm often wins under the \HardMax decision rule, with a strong evidence of the ``death spiral" effect mentioned earlier.
+ As predicted by the theory, better algorithms prevail under \HardMaxRandom if the expected number of ``random" users is sufficiently large.
+However, this effect is negligible for smaller parameter values.
 
@@
 \xhdr{Challenges.}
-Much of the challenge in the theoretical results, both conceptual and technical, was in setting up the model and the theorems.
+Much of the difficulty in the theoretical results, both conceptual and technical, was in setting up the model and the theorems.
 Apart from zeroing in on the \TheoryModel, it was crucial to interpret the results and intuitions from the literature on multi-armed bandits so as to formulate meaningful and productive assumptions on bandit algorithms and Bayesian priors.
@@
 It is essential to keep each of the three dimensions small yet representative.
-In particular, we need to capture a huge variety of bandit instances with only a few representative examples.
+In particular, we need to capture a huge variety of bandit algorithms and bandit instances with only a few representative examples.
 Further, we need a succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).

### 2020-09-03 journal paper/content/sim_details.tex
Added a para on the three classes.

@@
 
-\xhdr{MAB algorithms.} We consider the three classes of bandit algorithms discussed in Section~\ref{sec:intro-model} (and in more detail in Appendix~\ref{app:bg}), and look for qualitative differences between them.
+\xhdr{MAB algorithms.} \asedit{We consider three essential classes of bandit algorithms: ones that never explicitly explore (\emph{greedy algorithms}), ones that explore without looking at the data (\emph{exploration-separating algorithms}), and ones where exploration gradually zooms in on the best arm (\emph{adaptive-exploration algorithms}).
+In the absence of competition, these classes are fairly well-understood: greedy algorithms are terrible for a wide variety of problem instances, exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances, and adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances (see Appendix~\ref{app:bg} for more details).}
+
+We look for qualitative differences between these three classes.
 We take a representative algorithm from each class.

### 2020-09-03 journal paper/content/sec-intro.tex
Added some coloring.

@@
 %Finding formalizations that admit meaningful answers is a major part of the overall challenge.
-We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
+\asedit{We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.}
 
@@
 %We investigate these questions with
+\asedit{
 \xhdr{Our model: competition game.} We consider a stylized duopoly model in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
@@
 In all variants, agents have little or no information about other agents' choices and rewards.
+
+Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
+Each principal's strategy is a bandit algorithm; as such, it adjusts itself to the observed rewards, but not to the competition.
+This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning.
+Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
+We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.%
+\footnote{Comparisons between bandit algorithms are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+Generally,  ``better" algorithms are better in the long run, but could be worse initially.
+More on this in Preliminaries.} One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.}
 
@@
 %\end{itemize}
-
-Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
-Each principal's strategy is a bandit algorithm; as such, it adjusts itself to the observed rewards, but not to the competition.
-This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning.
-Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
-We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.%
-\footnote{Comparisons between bandit algorithms are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
-Generally,  ``better" algorithms are better in the long run, but could be worse initially.
-More on this in Preliminaries.} One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.
 

### 2020-09-03 journal paper/content/sec-intro.tex
More small edits.

@@
 Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
-Each principal's strategy is a bandit algorithm; as such, it adjusts itself to the observed rewards, but not to the competition.
+Each principal's strategy is a bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.
 This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning.
@@
 We refer to this variant as the \emph{\TheoryModel}.
+
 We focus on qualitative, asymptotic differences in the algorithms' performance.
-
 Our results depend crucially on the agents' decision rule:
@@
 We find that \HardMax is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
-Further, we show that \HardMax is very sensitive to tie-breaking: if the tie-breaking is probabilistically biased in favor of one principal, then this principal has a simple ``winning strategy" no matter what the other principal does.
+Further,
+%\HardMax is very sensitive to tie-breaking:
+if the tie-breaking is probabilistically biased in favor of one principal, then this principal has a simple ``winning strategy" no matter what the other principal does.
 
@@
 We call this model \HardMaxRandom.
-We find that better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
+Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
 However, there is a substantial caveat: one can defeat any algorithm by interleaving it with the greedy algorithm.
@@
 
-\xhdr{Economic interpretation: the inverted-U relationship.}
+\xhdr{Interpretation: the inverted-U relationship.}
 Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the inverted-U relationship between competition and innovation.%
@@
 
-\textbf{(ii)} We investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
-(We focus on the \HardMax decision rule.) We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous arrival).
+\textbf{(ii)} Focusing on \HardMax, we investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
+We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous arrival).
 Consumer welfare is higher under early/late arrival than under simultaneous entry.
@@
 
-\xhdr{Economic interpretation: network effects of data.}
+\xhdr{Interpretation: network effects of data.}
 Our model speaks to policy discussions on regulating data-intensive digital platforms \citep{furman2019unlocking, scott2019committee}, and particularly to the ongoing debate on the role of data in the digital economy.

### 2020-09-03 journal paper/content/sec-intro.tex
Another pass at the intro. Shortened to 3.5 pages!

@@
 while exploration may be essential for improving the service tomorrow, it may degrade quality and make users leave \emph{today}, in which case there will be less users to learn from.
-This may further degrade the platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
-Taken to the extreme, such dynamics may create a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
+This may create a ``data feedback loop" when the platform's performance further degrades relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+Taken to the extreme, such dynamics may cause a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
 Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.
@@
 Does increased competition lead to higher consumer welfare?
-How significant are ``data feedback loops" --- when more data leads to more users, which leads to even more data, etc.
---- and how they relate to the anti-trust considerations?
+How significant are the data feedback loops
+%--- when more data leads to more users, which leads to even more data, etc. --- 
+and how they relate to the anti-trust considerations?
 %Finding formalizations that admit meaningful answers is a major part of the overall challenge.
@@
 Principals compete on quality rather than on prices.
-Agents arrive sequentially, and choose among the two principals according to some decision rule.
-The chosen principal chooses an action (\eg a list of web search results to show to the user), the user experiences this action, and reports a reward.
+A new agent arrives and chooses a principal according to some decision rule.
+The principal selects an action (\eg a list of web search results), the user experiences it and reports a reward.
 Each principal only observes its own users.
-Principals commit to their strategies simultaneously and in advance, with objective to maximize the market share.
-We investigate several model variants within this framework, where we vary agents' decision rule and/or give a first-mover advantage to one of the principals.
-In all variants, agents have little or no information about other agents' choices and rewards.
+Principals commit to their strategies in advance, with objective to maximize the market share.
+We investigate several model variants, %within this framework,
+where we vary agents' decision rule or allow a first-mover advantage. %to one of the principals.
+%In all variants, agents have little or no information about other agents' choices and rewards.
 
@@
 We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.%
-\footnote{Comparisons between bandit algorithms are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
-Generally,  ``better" algorithms are better in the long run, but could be worse initially.
-More on this in Preliminaries.} One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.}
+\footnote{Such comparisons are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others. More on this in Preliminaries.}
+%Generally,  ``better" algorithms are better in the long run, but could be worse initially. 
+ One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.}
 
@@
 %We endow agents with Bayesian rationality, a common modeling approach for a theoretical investigation.
-We consider a basic Bayesian model.
-We posit that agents have a common Bayesian prior on reward distributions, and know the principals' algorithms.
-For simplicity, agents do not receive any information about the previous agents' choices and rewards.
-Each agent knows the round (s)he arrives in, computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
-We refer to this variant as the \emph{\TheoryModel}.
-
-We focus on qualitative, asymptotic differences in the algorithms' performance.
+We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms, but do not receive any information about the previous agents.
+Each agent computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
+%We focus on qualitative, asymptotic differences in the algorithms' performance.
 Our results depend crucially on the agents' decision rule:
@@
 \textbf{(i)} The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
-We find that \HardMax is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
+We find that it is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
 Further,
 %\HardMax is very sensitive to tie-breaking:
-if the tie-breaking is probabilistically biased in favor of one principal, then this principal has a simple ``winning strategy" no matter what the other principal does.
+if the tie-breaking is probabilistically biased in favor of one principal, then this principal can always prevail in competition.
+% has a simple ``winning strategy" no matter what the other principal does.
 
@@
 Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
-However, there is a substantial caveat: one can defeat any algorithm by interleaving it with the greedy algorithm.
-This has two undesirable corollaries: a better algorithm may sometimes lose in competition, and a pure Nash equilibrium typically does not exist.
+There is a caveat, however: any algorithm can be defeated by interleaving it with the greedy algorithm.
+This has two undesirable consequences: a better algorithm may sometimes lose in competition, and a pure Nash equilibrium typically does not exist.
 
-\textbf{(iii)} We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
-For this model, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
-This is the most technical result of the paper.
-The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm would attract slightly more).
+\textbf{(iii)} We further relax the decision rule so that the selection probabilities vary smoothly in terms of principal's Bayesian-expected rewards; 
+%as a function of the difference between  principals' Bayesian-expected rewards; 
+we call it \SoftMaxRandom. For this model, better algorithm prevails under much weaker assumptions.
+% on what constitutes a better algorithm. 
+(This is the most technical result of the paper.) The competition is more relaxed: the better algorithm receives only a small advantage in the long run.
+
+%both principals attract approximately half of the agents as time goes by, but a better algorithm attracts slightly more.
 %\end{itemize}
 
-All results extend to a much more general version of the multi-armed bandit problem in which the principal may observe additional feedback before and/or after each decision, as long as the feedback distribution does not change over time.
+All results admit generalizations 
+%extend to a much more general version of the multi-armed bandit problem 
+in which the principal may observe additional feedback before and/or after each decision, as long as the feedback distribution does not change over time.
 In most results, principal's utility may depend on both the market share and agents' rewards.
@@
 \xhdr{Interpretation: the inverted-U relationship.}
-Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the inverted-U relationship between competition and innovation.%
-%, as in Figure~\ref{fig:inverted-U}.
-\footnote{Here, ``innovation" refers to adoption of a better technology at a substantial R\&D expense to a given firm.
-It is not salient whether similar ideas and/or technologies already exist elsewhere.
-Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
-This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{aghion2005competition,Vives-08}.
+\asedit{Our findings can be framed in terms of the inverted-U relationship between competition and innovation.
+This is a well-established concept, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{aghion2005competition,Vives-08}.
+We interpret innovation as the adoption of better algorithms,%
+\footnote{Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms are well-known and/or similar technologies already exist elsewhere  \citep[\eg see][]{DS-arxiv}.}
+%, even if the algorithms themselves are well-known in the research literature }
+and control the severity of the competition by varying the agents' decision rule from \HardMax (cut-throat competition) to \HardMaxRandom to \SoftMaxRandom and all the way to the uniform selection.%
+\footnote{Agents' decision rule also controls the agents' rationality.
+While agents' rationality and severity of competition are often modeled separately, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
+We also find another, technically different, inverted-U relationship which zeroes in on the \HardMaxRandom model. 
+%We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
+While the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs, in our work the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.}
 
-Our decision rules differ in terms of rationality: from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
+%Our decision rules differ in terms of rationality: from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
 The same distinctions also control the severity of competition between the principals: from cut-throat competition with \HardMax to a more relaxed competition with \HardMaxRandom, to an even more relaxed competition with \SoftMaxRandom.
@@
 
-We identify the inverted-U relationship driven by the rationality/competitiveness distinctions outlined above: from \HardMax to \HardMaxRandom to \SoftMaxRandom to \Uniform.
-We also find another, technically different, inverted-U relationship which zeroes in on the \HardMaxRandom model.
-We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
-
-These inverted-U relationships are driven by different aspects in our model than the ones in the existing literature in economics.
-The latter focuses on the tradeoff between the R\&D costs and the benefits that the improved technology provides in the competition.
+% These inverted-U relationships are driven by different aspects in our model than the ones in prior work.
+The latter focus on the tradeoff between the R\&D costs and the benefits that the improved technology provides in the competition.
 In our case, the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
@@
 \xhdr{Numerical simulations.}
-We consider a basic frequentist model.
-We posit that the agents observe signals about the principals' past performance, and base their decisions on these signals alone, without invoking any prior knowledge or beliefs.
+We consider a basic frequentist model (called the \emph{\ExptsModel}), where the agents observe signals about the principals' past performance, and base their decisions on these signals alone, without invoking any prior knowledge or beliefs.
 The performance signals are abstracted and aggregated as a scalar \emph{reputation score} for each principal, modeled as a sliding window average of its rewards.
 Thus, agents' decision rule depends only on the two reputation scores.
-We refer to this variant as the \emph{\ExptsModel}.%
+% We refer to this variant as the \emph{\ExptsModel}.%
 %\footnote{In comparison, the theoretical results focus on another extreme, with Bayesian rationality and no performance signals.}
-
-We refine and expand the theoretical results in several ways.
+We refine and expand the theoretical results in several ways:
 
@@
 
-\textbf{(iii)} We investigate the algorithms' performance ``in isolation", \ie without competition.
+\textbf{(iii)} We investigate the algorithms' performance without competition.
 We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).}
-     We find that mean reputation -- arguably, the most natural performance measure ``in isolation" -- is sometimes \emph{not} a good predictor for the outcomes under competition.
+     We find that mean reputation -- arguably, the most natural performance measure -- is sometimes \emph{not} a good predictor for the outcomes under competition.
 
-\textbf{(iv)} We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}).
-We run additional experiments so as to isolate and compare these two effects.
+\textbf{(iv)} We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}), and run additional experiments to separate and compare them.
 We find that either effect alone leads to a significant advantage under competition.
@@
 %whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
-The death spiral/amplification effects mentioned above have a similar flavor as network effects: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved of users (resp., take over most of the market).
+The death spiral/amplification effects mentioned above have a similar flavor: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved of users (resp., take over most of the market).
 %\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}
-A distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.
+However, a distinctive feature of our approach is that we explicitly model the learning problem of the firms and consider them deploying algorithms for solving this problem.
 Thus, we do not explicitly model the network effects, but they arise endogenously from our setup.
 
-Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of data observed by the firms.
+Our results highlight that understanding the performance of learning algorithms in isolation does not necessarily translate to understanding their impact in competition, precisely due to the fact that competition leads to the endogenous generation of observable data.
 Approaches such as \citet{lambrecht2015can, bajari2018impact, varian2018artificial} argue that the diminishing returns to scale and scope of data in isolation mitigate such data feedback loops,
@@
 but ignore the differences induced by learning in isolation versus under competition.
-Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.
+Furthermore, explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by \asedit{the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.}
 
 \xhdr{Significance.}
-Our theory takes a basic Bayesian approach, a standard perspective in economic theory, and discovers several strong asymptotic results.
+Our results have a dual purpose: shed light on real-world implications of some typical scenarios, and investigate the space of models for describing the real world.
+As an example for the latter, the \HardMax model with simultaneous entry is arguably the most natural model to study \emph{a priori}.
+However, our results elucidate the need for more refined models that allow for ``free exploration" (\eg via random agents or early entry).
+
+\asedit{Our theory takes a basic Bayesian approach, a standard perspective in economic theory, and discovers several strong asymptotic results.
+Much of the difficulty, both conceptual and technical, is in setting up the model and the theorems.
+Apart from zeroing in on the \TheoryModel, it was crucial to interpret the results and intuitions from the literature on multi-armed bandits so as to formulate meaningful and productive assumptions on bandit algorithms and Bayesian priors.
+
 The numerical simulations provide a more nuanced and ``non-asymptotic" perspective.
 In essence, we look for substantial effects within relevant time scales.
-In fact, we start our investigation by determining what time scales are relevant in the context of our model.
+(In fact, we start our investigation by determining what time scales are relevant in the context of our model.) The central challenge is to capture a huge variety of bandit algorithms and bandit problem instances with only a few representative examples, and arrive at findings that are consistent across the entire space.
 
-Our study has a dual purpose: (i) shed light on real-world implications of some typical scenarios, and (ii) investigate the space of models for describing the real world.
-As an example to clarify the latter point, consider the \HardMax model with simultaneous entry.
-It is is not necessarily the most realistic model, but arguably the most natural one to study \emph{a priori}.
-However, our results elucidate the need for more refined models that allow for ``free exploration" (\eg via random agents or early entry).
+One model we studied is suitable for analysis and another for simulations, but not vice versa. A natural implementation of the \TheoryModel requires running time quadratic in the number of rounds,%
+\footnote{\Eg this is because at each round $t$, one needs to recompute, and integrate over, a discrete distribution with $t$ possible values, namely the number of agents that have chosen principal $1$ so far.}
+which precludes experiments at a sufficient scale.
+The \ExptsModel features an intricate feedback loop between algorithms' performance, their reputations and agents' choices, which simplifies the simulations but does not appear analytically tractable.}
 
-\xhdr{Challenges.}
-Much of the difficulty in the theoretical results, both conceptual and technical, was in setting up the model and the theorems.
-Apart from zeroing in on the \TheoryModel, it was crucial to interpret the results and intuitions from the literature on multi-armed bandits so as to formulate meaningful and productive assumptions on bandit algorithms and Bayesian priors.
-
-The numerical investigation is quite challenging even with a stylized model such as ours.
+%The numerical investigation is quite challenging even with a stylized model such as ours.
 An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given instance of a multi-armed bandit problem (and each such experiment is run many times to reduce variance).
@@
 
-While amenable to simulations, the \ExptsModel appears difficult to analyze. This is for several reasons:
-%
-(i) intricate feedback loop from performance to reputations to users to performance;
-%
-(ii) mean reputation, most connected to our intuition, is sometimes a bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited}); and
-%
-(iii)
-mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice.
-Given the theoretical results on the \TheoryModel, and the fact that we are in the realm of stylized economic models, we believe that resolving first-order theoretical questions about the \ExptsModel would not add much value to this paper.
+%While amenable to simulations, the \ExptsModel appears difficult to analyze.
+This is for several reasons: (i) intricate feedback loop from performance to reputations to users to performance; (ii) mean reputation, most connected to our intuition, is sometimes a bad predictor in competition (see Sections~\ref{sec:isolation} and~\ref{sec:revisited}); and (iii) mathematical tools from regret-minimization would only produce ``asymptotic" results, which do not seem to suffice.
+Given the theoretical results on the \TheoryModel, and the fact that we are in the realm of stylized economic models, resolving similar first-order theoretical questions about the \ExptsModel does not appear essential.
 

### 2020-09-03 journal paper/content/sec-bg.tex
small comment on ts
Small edits.

@@
 We present self-contained background on multi-armed bandits (MAB), to make the paper accessible to researchers who are not experts on MAB.
-We focus on the three classes of MAB algorithms listed in Section~\ref{sec:intro-model}: adaptive-exploration algorithms, exploration-separating algorithms, and greedy algorithms.
+We focus on the three classes of MAB algorithms listed in Section~\ref{expts-prelims}: adaptive-exploration algorithms, exploration-separating algorithms, and greedy algorithms.
 We list the precise upper and lower bounds on the regret rates only to illustrate how the three algorithm classes are separated from one another; the exact results are not essential for the paper.
@@
 %\xhdr{}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../journal_main"
+%%% End:

### 2020-09-03 journal paper/content/sec-intro.tex
Small edits.

@@
 Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
-We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.%
-\footnote{Such comparisons are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others. More on this in Preliminaries.}
-%Generally,  ``better" algorithms are better in the long run, but could be worse initially. 
+We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.
+%\footnote{Such comparisons are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+More on this in Appendix~\ref{app:bg}.
+Generally,  ``better" algorithms are better in the long run, but could be worse initially.
+}
 One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.}
@@
 
-All results admit generalizations 
+Most results admit generalizations 
 %extend to a much more general version of the multi-armed bandit problem 
-in which the principal may observe additional feedback before and/or after each decision, as long as the feedback distribution does not change over time.
-In most results, principal's utility may depend on both the market share and agents' rewards.
+in which the principal may observe additional feedback before and/or after each decision, 
+%as long as the feedback distribution does not change over time. In most results,
+ and principal's utility may include agents' rewards.
 

### 2020-09-04 journal paper/ec19paper/content/inverted_u.tex
update figures

@@
 % Thu Aug 16 13:13:01 2018
-\footnotesize
 \begin{table*}[t]
 \centering
+\begin{adjustbox}{width=\textwidth,center}
 \begin{tabular}{|c|c|c|c||c|c|c|}
@@
 \end{tabular}
-\normalsize
-\caption{{\bf Simultaneous Entry}, for Heavy-Tail and Needle-in-Haystack instances.
+\end{adjustbox}
+\caption{\footnotesize {\bf Simultaneous Entry}, for Heavy-Tail and Needle-in-Haystack instances.
 Each cell describes a game between two algorithms, call them Alg1 vs.
@@
 \end{table*}
-
 \normalsize
@@
 \end{tabular}
-\caption{{\bf Simultaneous entry}, for the Uniform MAB instance. Semantics are the same as in Table \ref{sim_table}.}
+\caption{\footnotesize {\bf Simultaneous entry}, for the Uniform MAB instance. Semantics are the same as in Table \ref{sim_table}.}
 \label{tab:duopoly_unif}
@@
 \captionsetup{labelformat=empty}
-\caption{Market share of row player (entrant), 200 round head-start, Heavy-Tail Instance}
+\caption{\footnotesize Market share of row player (entrant), 200 round head-start, Heavy-Tail Instance}
 %\caption{{\bf Temporary monopoly}, with $X=200$ (and $T_0=20$), for the Heavy-Tail MAB instance.
@@
  \end{tikzpicture}
- \caption{A stylized ``inverted-U relationship" between strength of competition and ``level of innovation".}
+ \caption{\footnotesize A stylized ``inverted-U relationship" between strength of competition and ``level of innovation".}
 \label{fig:inverted-U-expts}
@@
 \includegraphics[scale=0.35]{ec19paper/figures/ht_eq_welfare}
-\caption{Smoothed welfare plots resulting from equilibrium strategies in the different market structures.
+\caption{\footnotesize Smoothed welfare plots resulting from equilibrium strategies in the different market structures.
 Note that welfare at $t = 0$ incorporates the regret incurred during the incumbent and warm start periods.

### 2020-09-04 journal paper/ec19paper/content/non_greedy_choice.tex
update figures

@@
 \centering
+\begin{adjustbox}{width=\textwidth,center}
 \begin{tabular}{|c|c|c|c||c|c|c|}
@@
 \end{tabular}
-\normalsize
-\caption{\HardMax (\HM) and \HardMaxRandom (\HMR) choice models on the Heavy-Tail MAB instance.
+\end{adjustbox}
+\caption{\footnotesize \HardMax (\HM) and \HardMaxRandom (\HMR) choice models on the Heavy-Tail MAB instance.
 Each cell describes the market shares in a game between two algorithms, call them Alg1 vs.
@@
 \end{tabular}
-\caption{\HardMaxRandom (\HMR) choice model for Uniform and Needle-In-Haystack MAB instances.}
+\end{adjustbox}
+\caption{\footnotesize \HardMaxRandom (\HMR) choice model for Uniform and Needle-In-Haystack MAB instances.}
 %Same semantics as in Table \ref{tab:non_greedy_table}.

### 2020-09-04 journal paper/ec19paper/content/perf_in_iso.tex
update figures

@@
 The shaded area display 95\% confidence intervals.}}
-\caption{Mean reputation trajectory (left) and mean instantaneous reward trajectory (right) for Needle-in-Haystack. The shaded area shows 95\% confidence intervals.}
+\caption{\footnotesize Mean reputation trajectory (left) and mean instantaneous reward trajectory (right) for Needle-in-Haystack. The shaded area shows 95\% confidence intervals.}
 \label{prelim_means}
@@
 The shaded area display 95\% confidence intervals.}}
-\caption{Relative reputation trajectory for $\Thompson$ vs $\DynamicGreedy$, on Uniform instance (left) and Needle-in-Haystack instance (right). Shaded area display 95\% confidence intervals.}
+\caption{\footnotesize Relative reputation trajectory for $\Thompson$ vs $\DynamicGreedy$, on Uniform instance (left) and Needle-in-Haystack instance (right).
+Shaded area display 95\% confidence intervals.}
 \label{relative_rep_plots}

### 2020-09-04 journal paper/ec19paper/content/revisited.tex
update figures

@@
 \includegraphics[scale=0.35]{ec19paper/figures/ts_dg_rep_diff_nih}
-\caption{Reputation scores for Needle-in-Haystack at $t=500$ (left), Reputation difference $\Thompson - \DynamicGreedy$ for Needle-in-Haystack (right).
+\caption{\footnotesize Reputation scores for Needle-in-Haystack at $t=500$ (left), Reputation difference $\Thompson - \DynamicGreedy$ for Needle-in-Haystack (right).
 Both are smoothed using a kernel density estimate.}

### 2020-09-04 journal paper/content/sec-intro.tex
one small thing in intro

@@
 \asedit{
-\xhdr{Our model: competition game.} We consider a stylized duopoly model in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
-Principals compete on quality rather than on prices.
+\xhdr{Our model: competition game.} We consider a stylized duopoly model in which two firms (\emph{principals}) compete for users (\emph{agents}).
+Principals compete on quality rather than on prices, \gaedit{and, crucially, need to engage in exploration in order to learn which actions lead to high quality products}.
 A new agent arrives and chooses a principal according to some decision rule.
@@
 Each principal only observes its own users.
-Principals commit to their strategies in advance, with objective to maximize the market share.
+Principals commit to their strategies in advance, with \gaedit{the} objective to maximize the market share.
 We investigate several model variants, %within this framework,

### 2020-09-04 journal paper/journal_main.tex
Removed a \newpage somewhere.
Redefined the \xhdr macro.
Moved back to 1-inch margins.
fix tables and main page
fix font for rest of paper
Formatting.
change bibtex formatting
add the customary acknowledgements...

@@
 \documentclass[12pt]{article}
-\usepackage[margin=1.25in]{geometry}
+\usepackage[margin=1in]{geometry}
 
@@
 setspace,sectsty,comment,footmisc,caption,
-pdflscape,subcaption,array,hyperref,adjustbox,threeparttable
+pdflscape,subcaption,array,hyperref,adjustbox
 }
@@
 \usepackage{slivkins-setup,slivkins-theorems}
+\renewcommand{\xhdr}[1]{\vspace{1mm} \noindent{\bf #1}}
 
@@
 All theoretical results are from \citet{CompetingBandits-itcs18}, and all experiments are from \citet{CompetingBandits-ec19}.
-Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.}}
+Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.
+We would like to thank Yeon-Koo Che, Glen Weyl, Sven Rady, and seminar participants at Columbia, ACM Economics and Computation, and Innovations in Theoretical Computer Science for helpful comments and conversations.
+All errors are our own.} }
 % \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
 
-\author{Guy Aridor%
+\author{Guy Aridor
 \footnote{Columbia University, Department of Economics. Email: g.aridor@columbia.edu}
-\and
-\rule{0.2in}{0pt}
-Yishay Mansour%
+\and \hspace{-0.75cm}
+\rule{0.0in}{0pt}
+Yishay Mansour
 \footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com}
-\and
-\rule{0.2in}{0pt}
+\and \hspace{-0.75cm}
+\rule{0.0in}{0pt}
 Aleksandrs Slivkins%
 \footnote{Microsoft Research New York City. Email: slivkins@microsoft.com}
-\and
-\rule{0.2in}{0pt}
+\and \hspace{-0.75cm}
+\rule{0.0in}{0pt}
 Zhiwei Steven Wu%
-\footnote{Carnegie-Mellon University, Pittsburgh, PA.
+\footnote{Carnegie Mellon University, Pittsburgh, PA.
 Email: zstevenwu@cmu.edu.\newline
@@
 }
-\date{First version: August 2020\\ This version: September 2020}
+\date{First version: July 2020\\ This version: September 2020}
 
@@
 \newpage
-
 %
@@
 
-\newpage
 \subfile{content/conclusion}
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{econometrica}
 %\begin{small}
@@
 
-\section{Background for non-specialists: multi-armed bandits}
+\section{Background on multi-armed bandits}
 \label{app:bg}

### 2020-09-04 journal paper/content/sec-intro.tex

@@
 %Finding formalizations that admit meaningful answers is a major part of the overall challenge.
-\asedit{We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.}
+We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
 
@@
 %We investigate these questions with
-\asedit{
 \xhdr{Our model: competition game.} We consider a stylized duopoly model in which two firms (\emph{principals}) compete for users (\emph{agents}).
-Principals compete on quality rather than on prices, \gaedit{and, crucially, need to engage in exploration in order to learn which actions lead to high quality products}.
+Principals compete on quality rather than on prices, and engage in exploration in order to learn which actions lead to high quality products.
 A new agent arrives and chooses a principal according to some decision rule.
@@
 Each principal only observes its own users.
-Principals commit to their strategies in advance, with \gaedit{the} objective to maximize the market share.
+Principals commit to their strategies in advance, with the objective to maximize the market share.
 We investigate several model variants, %within this framework,
@@
 }
-One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.}
+One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.
+
+\newpage
 
@@
 \xhdr{Interpretation: the inverted-U relationship.}
-\asedit{Our findings can be framed in terms of the inverted-U relationship between competition and innovation.
+Our findings can be framed in terms of the inverted-U relationship between competition and innovation.
 This is a well-established concept, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{aghion2005competition,Vives-08}.
@@
 \footnote{Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms are well-known and/or similar technologies already exist elsewhere  \citep[\eg see][]{DS-arxiv}.}
-%, even if the algorithms themselves are well-known in the research literature }
+%, even if the algorithms themselves are well-known in the research literature
 and control the severity of the competition by varying the agents' decision rule from \HardMax (cut-throat competition) to \HardMaxRandom to \SoftMaxRandom and all the way to the uniform selection.%
@@
 %We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
-While the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs, in our work the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.}
+While the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs, in our work the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
 
@@
 but ignore the differences induced by learning in isolation versus under competition.
-Furthermore, explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by \asedit{the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.}
+Furthermore, explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.
 
@@
 
-\asedit{Our theory takes a basic Bayesian approach, a standard perspective in economic theory, and discovers several strong asymptotic results.
+Our theory takes a basic Bayesian approach, a standard perspective in economic theory, and discovers several strong asymptotic results.
 Much of the difficulty, both conceptual and technical, is in setting up the model and the theorems.
@@
 which precludes experiments at a sufficient scale.
-The \ExptsModel features an intricate feedback loop between algorithms' performance, their reputations and agents' choices, which simplifies the simulations but does not appear analytically tractable.}
+The \ExptsModel features an intricate feedback loop between algorithms' performance, their reputations and agents' choices, which simplifies the simulations but does not appear analytically tractable.
 

### 2020-09-04 journal paper/content/sec-intro.tex
Minor edits.

@@
 Our findings can be framed in terms of the inverted-U relationship between competition and innovation.
-This is a well-established concept, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{aghion2005competition,Vives-08}.
+This is a well-established concept, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
+%\citep[\eg][]{aghion2005competition,Vives-08}.
 We interpret innovation as the adoption of better algorithms,%
@@
 and control the severity of the competition by varying the agents' decision rule from \HardMax (cut-throat competition) to \HardMaxRandom to \SoftMaxRandom and all the way to the uniform selection.%
-\footnote{Agents' decision rule also controls the agents' rationality.
+%\footnote{Agents' decision rule also controls the agents' rationality.
 While agents' rationality and severity of competition are often modeled separately, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
-We also find another, technically different, inverted-U relationship which zeroes in on the \HardMaxRandom model.
+We also find another, technically different inverted-U relationship which zeroes in on the \HardMaxRandom model.
 %We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
-While the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs, in our work the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
+\asedit{The barriers for innovation arise entirely from the reputational consequences of exploration in competition, whereas the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs.}
+
+%While the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs, in our work the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
 

### 2020-09-04 journal paper/content/related_work.tex
A pass: revised the para on Keller-Rady, and refactored a bit.

--- 2020-07-18 journal paper/content/related_work.tex
@@
-\xhdr{Bandits and exploration.} Multi-armed bandits (\emph{MAB}) is an elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information.
-MAB problems have been studied for many decades \asedit{by researchers from computer science, operations research, statistics and economics.
-The literature on MAB is extremely vast and multi-threaded.
-The most relevant thread studies a basic model of MAB with IID rewards and no auxiliary structure, see Appendix~\ref{app:bg}.
-The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.} This literature is covered in many books and surveys.
-Particularly, see
-\citet{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} for background on regret-minimizing formulations, \citet{Gittins-book11} for Bayesian and Markovian formulations, and \citet{CesaBL-book,slivkins-MABbook} for connections to economics and game theory.
-A discussion of industrial applications of MAB can be found in \citep{DS-arxiv}.
+\xhdr{Exploration.} Multi-armed bandits (\emph{MAB}) is an elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information.
+MAB problems have been studied for many decades by researchers from computer science, operations research, statistics and economics, generating a vast and multi-threaded literature.
+The most relevant thread concerns MAB with stochastic rewards and no auxiliary structure, which is the problem faced by each principal in our model; see Appendix~\ref{app:bg} for self-contained background.
+The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.
+This literature is covered in several dedicated monographs,
+\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.%
+\footnote{\citet{Gittins-book11} covers Markovian formulations, the other three are concerned with regret-minimization.}
+Various connections to economics and game theory are detailed in 
+books \citep{CesaBL-book,slivkins-MABbook} and surveys \citep{Bergemann-survey06,Horner-survey16}. A discussion of industrial applications can be found in \citep{DS-arxiv}.
 
-The three-way tradeoff between exploration, exploitation and incentives has been studied in several settings other than ours:
+The three-way tradeoff between exploration, exploitation and incentives has been studied in several settings very different from ours:
 incentivizing exploration in a recommendation system
@@
 dynamic auctions
-    \cite[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
+    \citep[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
 pay-per-click ad auctions with unknown click probabilities
-    \cite[\eg][]{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
+    \citep[\eg][]{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
 coordinating search and matching by self-interested agents
     \citep{Bobby-Glen-ec16},
-as well as human computation
-    \cite[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
-\citet{Bolton-econometrica99,Keller-econometrica05,Johari-ec12} studied models with self-interested agents jointly performing exploration, with no principal to coordinate them.
+human computation
+    \citep[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13},
+and social learning 
+    \citep[\eg][]{Bolton-econometrica99,Keller-econometrica05,Johari-ec12}. 
+A literature review of this work can be found in 
+\citep[Ch. 11.6]{slivkins-MABbook}.
+
+\asedit{\citet{keller2003price} studied the interplay of exploration and competition \emph{on prices}.
+This distinction leads to several important differences.
+First, the data externality -- that each customer brings a new data point to a given principal if and only if he chooses this principal -- is absolutely crucial to our setting, and absent in theirs.
+Second, the inputs to agents' decision rule are prices (which are directly controlled by the principals), rather than the quality of the chosen actions (which is not known a priori).
+Their focus on learning the decision rule, whereas ours is on learning quality.
+Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria, whereas we focus on adoption of better bandit algorithms.}
 
@@
 
-Our setting is closely related to the ``dueling algorithms" framework of \citet{DuelingAlgs-stoc11}, which studies competition between two principals, each running an algorithm for the same problem.
-However, \citet{DuelingAlgs-stoc11} consider algorithms for offline / full input scenarios, \asedit{and posit binary (win/lose) payoffs for the principals.
-Similarly, \citet{ben2017best, ben2019regression} study competition between offline learning algorithms.
-Whereas we focus on bandit algorithms and competition for users.}
-
-\xhdr{Other related work in economics.} The competition vs.
+\xhdr{Competition.} The competition vs.
 innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{aghion2005competition,Vives-08}.
@@
 
-\gaedit{There is a literature that studies the interaction between learning-by-doing and competition \citep{fudenberg1983learning, dasgupta1988learning, cabral1994learning}.
-In these models, firms learn while competing against each other where a firm attracting more consumers reduces its per-unit production costs.
-However, our model focuses on firms learning product quality (as opposed to reducing production costs) and the impact that this learning has on attracting consumers.}
+The literature on learning-by-doing vs.
+competition \citep[\eg][]{fudenberg1983learning, dasgupta1988learning, cabral1994learning} studies firms that learn while competing against each other, so that a firm attracting more consumers reduces its per-unit production costs.
+However, our model focuses on firms learning product quality (as opposed to reducing production costs) and the impact that this learning has on attracting consumers.
 
-A line of work on \emph{platform competition}, starting with \cite{Rysman09}, concerns competition between firms (\emph{platforms}) that improve as they attract more users (\emph{network effect}); see \citet{Weyl-White-14} for a recent survey.
-This literature is not concerned with \innovation, and typically models network effects exogenously, whereas in our model network effects are endogenous: they are created by MAB algorithms, an essential part of the model.
-\gaedit{A nascent literature studies whether and when network effects manifest themselves in data-intensive markets \citep{prufer2017competing, hagiu2020data}, but typically models learning as a reduced-form function of past consumer history and focus on the role of prices.}
+A line of work on \emph{platform competition} (starting with \cite{Rysman09}, see \citet{Weyl-White-14} for a survey) concerns competition between firms that improve as they attract more users.
+This literature is not concerned with \innovation, and typically models network effects exogenously, whereas they are endogenous in our model.
+%: they are created by MAB algorithms, an essential part of the model.
+A nascent literature studies 
+%whether and when network effects manifest themselves 
+\asedit{network effects}
+in data-intensive markets \citep{prufer2017competing, hagiu2020data}, but typically models learning as a reduced-form function of past consumer history and focuses on the role of prices.
 %as opposed to the reputational consequences of learning.
@@
 We observe a similar effect when we investigate the role that reputation can serve as a barrier to entry.
+\asmargincomment{I can't parse this sentence, could you pls simplify?}
 However, in our model this effect is further strengthened by the fact that the firms have to learn while competing adding that the incumbent may not only have a reputational advantage, but additionally has a further advantage due to the data it acquires in the incumbency period.
 Thus, our model also highlights the role that data can serve as a barrier to entry in online markets which has similarly been noted in \cite{de2020data}.
-\gaedit{For an extensive overview of the other channels through which first-mover advantages can lead to a competitive advantage, see \cite{kerin1992first}.}
+For an extensive overview of the other channels through which first-mover advantages can lead to a competitive advantage, see \cite{kerin1992first}.
 
+We use first-mover advantage and relaxed versions of rationality to model varying competition, instead of classic ``market competitiveness" measures such as the Lerner Index or the Herfindahl-Hirschman Index
+\citep{tirole1988theory}.
+The latter measures rely on ex-post observable attributes of a market such as prices or market shares.
+Neither is applicable to our setting, since there are no prices, and market shares are endogenous.
+
+The ``dueling algorithms" framework of \citet{DuelingAlgs-stoc11} studies competition between two principals, each running an algorithm for the same problem.
+However, they consider algorithms for offline / full input scenarios, and posit binary (win/lose) payoffs for the principals.
+Similarly, \citet{ben2017best, ben2019regression} study competition between offline learning algorithms.
+Whereas we focus on bandit algorithms and competition for users.
+
+\xhdr{Relaxed rationality.}
 Relaxed versions of rationality similar to ours are found in several notable lines of work.
@@
 Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom.
-
-We use first-mover advantage and relaxed versions of rationality to model varying competition, instead of classic ``market competitiveness" measures such as the Lerner Index or the Herfindahl-Hirschman Index
-\citep{tirole1988theory}.
-The latter measures rely on ex-post observable attributes of a market such as prices or market shares.
-Neither is applicable to our setting, since there are no prices, and market shares are endogenous.
-
-% moved this thought to the intro.
-\OMIT{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
+\asedit{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
 

### 2020-09-04 journal paper/content/related_work.tex
Minor edits.

@@
 MAB problems have been studied for many decades by researchers from computer science, operations research, statistics and economics, generating a vast and multi-threaded literature.
-The most relevant thread concerns MAB with stochastic rewards and no auxiliary structure, which is the problem faced by each principal in our model; see Appendix~\ref{app:bg} for self-contained background.
+The most relevant thread concerns MAB with stochastic rewards and no auxiliary structure, which is the problem faced by each principal in our model (more on this in Appendix~\ref{app:bg}).
 The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.

### 2020-09-04 journal paper/content/sec-intro.tex
Minor edits.

@@
 
-One model we studied is suitable for analysis and another for simulations, but not vice versa. A natural implementation of the \TheoryModel requires running time quadratic in the number of rounds,%
-\footnote{\Eg this is because at each round $t$, one needs to recompute, and integrate over, a discrete distribution with $t$ possible values, namely the number of agents that have chosen principal $1$ so far.}
+One model we study is suitable for analysis and another for simulations, but not vice versa. A natural implementation of the \TheoryModel requires running time quadratic in the number of rounds,%
+\footnote{\label{fn:Tsquared}\Eg this is because at each round $t$, one needs to recompute, and integrate over, a discrete distribution with $t$ possible values, namely the number of agents that have chosen principal $1$ so far.}
 which precludes experiments at a sufficient scale.

### 2020-09-04 journal paper/content/related_work.tex
Minor space-saving edits.

@@
 MAB problems have been studied for many decades by researchers from computer science, operations research, statistics and economics, generating a vast and multi-threaded literature.
-The most relevant thread concerns MAB with stochastic rewards and no auxiliary structure, which is the problem faced by each principal in our model (more on this in Appendix~\ref{app:bg}).
+The most relevant thread concerns MAB with stochastic rewards and no auxiliary structure, which is the problem faced by each principal in our model (see Appendix~\ref{app:bg}).
 The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.
-This literature is covered in several dedicated monographs,
+This literature is covered in dedicated monographs,
 \citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.%
@@
 Various connections to economics and game theory are detailed in 
-books \citep{CesaBL-book,slivkins-MABbook} and surveys \citep{Bergemann-survey06,Horner-survey16}. A discussion of industrial applications can be found in \citep{DS-arxiv}.
+books \citep{CesaBL-book,slivkins-MABbook} and surveys \citep{Bergemann-survey06,Horner-survey16}. Industrial applications are discussed in \citep{DS-arxiv}.
 

### 2020-09-04 journal paper/content/sec-intro.tex
An edit re dynamic strategies.

@@
 Each principal's strategy is a bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.
-This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning.
-Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
+\asedit{(This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.)}
+%Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical. 
 We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.
@@
 
-\newpage
+%\newpage
 
@@
 
-Most results admit generalizations
+%Most results admit generalizations
 %extend to a much more general version of the multi-armed bandit problem
-in which the principal may observe additional feedback before and/or after each decision,
+%in which the principal may observe additional feedback before and/or after each decision,
 %as long as the feedback distribution does not change over time. In most results,
- and principal's utility may include agents' rewards.
+% and principal's utility may include agents' rewards.
 

### 2020-09-04 journal paper/content/sec-theory.tex
A bunch of small edits.

@@
-In this section, we present our theoretical results for the \TheoryModel. While we provide intuition and (some) proof sketches, the detailed proofs are deferred to Appendix~\ref{sec:theory-proofs}.
+In this section, we present our theoretical results for the \TheoryModel. While we provide intuition and proof sketches, the detailed proofs are deferred to Appendix~\ref{sec:theory-proofs}.
 
@@
 
-Consider the \HardMaxRandom response model, where each principal is chosen with some positive baseline probability. Informally, we find that:
-\begin{align}\label{eq:theory-HMR-informal}
-\text{A much better algorithm wins big.}
-\end{align}
+Consider the \HardMaxRandom response model, where each principal is chosen with some positive baseline probability. Informally, we find that 
+%\begin{align}\label{eq:theory-HMR-informal}
+\emph{a much better algorithm wins big}.
+%\end{align}
 In more detail, a principal with asymptotically better \BIR wins by a large margin: after a ``learning phase" of constant duration, all agents choose this principal with maximal possible probability $\respF(1)$.
@@
 
-\asedit{We'd like to use Theorem~\ref{thm:random-clean} to conclude that a (much) better algorithm prevails in equilibrium.
-Indeed, this holds if the principals can only choose from a finite set of MAB algorithms.}
+We'd like to use Theorem~\ref{thm:random-clean} to conclude that a (much) better algorithm prevails in equilibrium.
+Indeed, this holds if the principals can only choose from a finite set of MAB algorithms.
 
 \begin{definition}\label{def:restricted-competition}
-A \emph{\FiniteGame} is a version of the competition game between the two principals in which they can only choose from a finite set of MAB algorithms, which are called \emph{feasible}.
+A \emph{\FiniteGame} is the competition game between the two principals in which they can only choose from a finite set of MAB algorithms (called \emph{feasible}).
 All feasible algorithms are assumed to be
@@
 Consider the \FiniteGame with a feasible algorithm \alg which BIR-dominates all other feasible algorithms.
-Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium: both principals choose \alg.
+Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
 \end{corollary}
@@
 Given any \bmonotone MAB algorithm \alg other than \DynGreedy, we design a modified algorithm which ``mixes in" some greedy choices (and consequently learns at a slower rate), yet prevails over \alg in the competition game.
-Thus, we have a counterpoint to \eqref{eq:theory-HMR-informal}: even under \HardMaxRandom, a better algorithm may lose in competition.
+Thus, we have a counterpoint to ``much better algorithms win": even under \HardMaxRandom, a better algorithm may lose in competition.
 The corresponding counterpoint to Corollary~\ref{cor:random} states that non-greedy algorithms are \emph{not} chosen in equilibrium.
@@
 Suppose \alg deviates from \DynGreedy starting from some (local) step $n_0$.
-Then
-\begin{enumerate}
-\item The modified algorithm coincides with \DynGreedy
-for the first $n_0-1$ steps;
-\item In each step $n\geq n_0$, \alg is invoked with probability
+%\begin{enumerate}
+%\item 
+The modified algorithm coincides with \DynGreedy
+for the first $n_0-1$ steps.
+%\item 
+In each step $n\geq n_0$, \alg is invoked with probability
   $1-p$, and with the remaining probability $p$ does the ``greedy
@@
   given the current information collected by \alg.
+%\item 
+The data from the ``greedy choice" steps are not recorded.%
+\footnote{In other words: the algorithm proceeds as if the ``greedy choice" steps have never happened.
+While it is usually more efficient to consider all available data, this modification simplifies analysis.}
 
-\item The modified algorithm does not record the ``greedy choice" steps in its data set.%
-    \footnote{In other words: in the subsequent rounds, as far as the modified algorithm is concerned, the ``greedy choice" steps have never happened.
+%    \footnote{In other words: in the subsequent rounds, as far as the modified algorithm is concerned, the ``greedy choice" steps have never happened.
 While it is usually more efficient to consider all available data, this simplification enables a cleaner comparison between the two algorithms.
@@
 
-\end{enumerate}
+%\end{enumerate}
 %The modified algorithm is called \emph{greedy deviation}, with probability parameter $p$.
@@
 
-\noindent Note that the modified algorithm is another pure strategy in the competition game between the two principals, not a mixed strategy that randomizes between \alg and the greedy algorithm.
+Note that the modified algorithm is another pure strategy in the competition game between the two principals, not a mixed strategy that randomizes between \alg and the greedy algorithm.
 
@@
 \end{theorem}
-
-\ascomment{This whole blurb before the corollary is new.}
 
@@
 Moreover, no pure Nash equilibrium exists when some algorithm BIR-dominates \DynGreedy and $T$ is sufficiently large.
-\asedit{These conclusions hold even if each principal's strategy set is a subset of \bmonotone MAB algorithms that is closed under the greedy modification.}%
+These conclusions hold even if each principal's strategy set is a subset of \bmonotone MAB algorithms that is closed under the greedy modification.%
 \footnote{That is, the strategy set of each principal consists of some (but not necessarily all) \bmonotone MAB algorithms, and contains the greedy modification for each algorithms in this set and each mixing parameter $p$.}

### 2020-09-04 journal paper/content/sec-model.tex
Rewrote model/discussion.

--- 2020-08-24 journal paper/content/sec-model.tex
@@
-Let us define the basic model; our theoretical results admit some extensions, detailed in Section~\ref{sec:theory-extensions}.
-
 \xhdr{Principals and agents.} There are two principals and $T$ agents,
@@
 
-\xhdr{Agents' response.} Each agent $t$ forms a reward estimate $\est_i(t)\in [0,1]$ for each principal $i$, and chooses the principal based on these reward estimates.gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}} Specifically, principal $1$ is chosen with probability
+%\gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}}
+
+\xhdr{Agents' response.} Each agent $t$ forms reward estimates $\est_i(t)\in [0,1]$ for each principal $i$, and uses them to choose the principal.
+Specifically, principal $1$ is chosen with probability
 \begin{align}\label{eq:model-f}
@@
 \end{align}
-where $\respF:[-1,1]\to [0,1]$ is the \emph{response function}, same for all agents and known to everyone.
-
+where $\respF:[-1,1]\to [0,1]$ is the \emph{response function}, same for all agents.
 We assume that $\respF$ is monotonically non-decreasing, is larger than $\nicefrac12$ on the interval $(0,1]$, and smaller than $\nicefrac12$ on the interval $[-1,0)$.
-We consider three specific models (see \reffig{fig:response-functions}):
+We consider three variants for $\respF$ (see \reffig{fig:response-functions}):
 
@@
 \end{center}
-\caption{The three models for $\respF$: \HardMax is thick blue, \HardMaxRandom is slim red, and \SoftMaxRandom is the dashed curve.}
+\caption{The three models for $\respF$: \HardMax is thick blue, \HardMaxRandom is red, and \SoftMaxRandom is dashed.}
 \label{fig:response-functions}
@@
 \item In the \emph{\TheoryModel}, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$. 
-Agents know the principals' algorithms.\gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}} Their reward estimates are defined as posterior mean rewards:
-        $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i }$
-    for each agent $t$ and principal $i$.
+    Agents' reward estimates are defined as posterior mean rewards:
+        $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i}$
+    for each agent $t$ and principal $i$.%
+\footnote{\asedit{Formally, agents do the Bayesian update knowing $t$, the principals' algorithms, the prior $\priorMu$, and $\respF$.}}
 
-\item In the \emph{\ExptsModel}, agents' reward estimate for a given principal is the average reward of the last $M$ agents that chose this principal (interpreted as a \emph{reputation score}).
-To make the reputation score meaningful initially, each principal enjoys a ``warm start": additional $T_0$ agents arrive before the game starts, and interact with the principal as described above.
+%Agents know the principals' algorithms.\gaedit{\footnote{While our model assumes that agents have correct beliefs about the performance of the algorithms, only the beliefs themselves matter and so our results hold as long as all agents have the same initial beliefs, even if they are not aligned with the true prior.}}
+
+\item In the \emph{\ExptsModel}, agents' reward estimate for a given principal is the average reward of the last $M$ agents that chose this principal, interpreted as a \emph{reputation score}.
+To make it meaningful initially, each principal enjoys a ``warm start": additional $T_0$ agents arrive before the game starts, and interact with the principal as described above.
 
@@
 Principals are risk-neutral, in the sense that they optimize their expected utility.
+
+\xhdr{Extensions.} \asedit{The \TheoryModel admits several extensions, detailed in Section~\ref{sec:theory-extensions}.
+First, all/most results extend to arbitrary reward distributions, allow reward-dependent utility, and carry over to a more general version of multi-armed bandits.
+Second, tather than knowing the actual tuple$(\alg[1],\alg[2],\priorMu,\respF)$, agents could have respective common beliefs $(\alg[1]',\alg[2]',\priorMu',\respF')$ than need not be correct; then, all results carry over with respect to these beliefs.
+Third, all results carry over \asmargincomment{check}  if each agent $t$ has a prior on her arrival time $t$, rather than know it exactly.
+Fourth, we can handle a limited amount of non-stationarity in $\respF$.
+\asmargincomment{check} Finally, the main result on \HardMax extends to time-discounted utilities.
+}
+
+\subsection{Discussion}
+\label{sec:discussion}
+
+\asedit{Reflecting reality, we posit that principals' exploration strategies are ``non-strategic" in nature.
+Even though the principals play a multi-step game, they do not react to each other's moves or to the agents' strategic choices.
+This is how industry approaches exploration algorithms, and for several good reasons.
+"Non-strategic" exploration is well-studied in machine learning, and yet it remains a very complex subject in research (as evidenced, \eg by the huge amount of activity therein).
+Even the seemingly simple algorithms are not straightforward to deploy in practice, and require a substantial investment in infrastructure \cite[\eg see the discussions in][]{DS-arxiv}.
+Responding to the competition represents another layer of complexity which has not been previously studied in this context, to the best of our knowledge, let alone made even remotely practical.
+Besides, the competitor's exploration strategy is typically not public, and understanding its exploration behavior via observations appears challenging even as a  research problem.}%
+\footnote{\asedit{Principals could potentially react to the market share or (the difference in) reputation scores.
+However, baking these signals into one's exploration strategy runs the risk of over-interpreting our competition model, as they may change for exogenous reasons.
+Alternatively, one could use such signals, as well as the intuitions coming from this paper, to guide the platform's decisions regarding exploration.}}
+
+%\gaedit{Finally, the assumption that firms commit to the algorithms at the start of the game deserves further justification.
+We impose this commitment assumption since it is more realistic than considering dynamic strategies and algorithms are, by definition, a commitment to a set of rules.
+In many contexts where such exploration algorithms are typically deployed, consumers arrive at such a rapid volume that it it is infeasible to consider a firm dynamically adapting its algorithm to each consumer that arrives.
+Indeed, \cite{DS-arxiv} points out how difficult it is to deploy standard exploration algorithms in practice and such dynamic strategies are substantially more difficult to deploy in practice.
+While it is possible to consider share-adaptive algorithms that would still keep the commitment assumption but allow for indirect dynamic responses to market conditions, such algorithms do not exist currently in the literature but are a good avenue for future work.}
+
+Our models are stylized in several important  respects.
+Firms compete only on the quality of service, rather than, say, pricing or the range of products.
+Agents are myopic: they do not worry about how their actions impact their future utility.\footnote{So, agents do not attempt to learn over time, second-guess or game future agents, or manipulate the principals' learning algorithms.
+This is arguably typical in practice, in part because one agent's influence tends to be small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores", and further simplified to average performance over a time window.
+
+We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs.
+For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior.
+Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
+Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
+It is unclear how to incorporate performance signals in a theoretically tractable model.
+For the numerical results, the \ExptsModel accounts for competition in a more direct way,
+% \gaedit{by positing a more realistic model of consumer behavior that does not
+not requiring users to have direct information on the algorithms deployed or the bandit problem faced by the firms.
+It further allows us to separate reputation vs.
+data advantage and makes our model amenable to numerical simulations.
+%(Whereas simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable, see Footnote~\ref{fn:Tsquared}.)
+
+\asedit{On the machine learning side, our model captures big, qualitative differences between bandit algorithms, building on the well-established intuition in the literature.
+Comparisons between bandit algorithms are generally somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+In particular, ``better" algorithms are better in the long run, but could be worse initially.
+We focus on a standard model of stochastic bandits.
+In Appendix~\ref{app:bg}, we present sufficient background on this model, accessible to non-specialists.
+However, we are less interested in state-of-art algorithms for realistic applications.}
+
+\asedit{Principals' utilities are not discounted with time.%
+\footnote{However, our main theoretical result on the greedy algorithm extends to time-discounted rewards.}
+ While much of the early work on bandits focuses on  time-discounted rewards \citep{Gittins-book11}, non-discounted  formulations are prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
+The distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.}

### 2020-09-04 journal paper/content/conclusion.tex
Lots of small edits because of the new font size.
Removed the coloring.

--- 2020-07-18 journal paper/content/conclusion.tex
@@
 Our results have two economic interpretations.
-\gaedit{The first is that they} can be framed in terms of the classic inverted-U relationship between innovation and competition, where \innovation refers to the adoption of better bandit algorithms.
-Unlike other models in the literature, what prevents innovation is not its direct costs, but \gaedit{the} short-term reputation consequences of exploration.
+The first is that they can be framed in terms of the classic inverted-U relationship between innovation and competition, where \innovation refers to the adoption of better bandit algorithms.
+Unlike other models in the literature, what prevents innovation is not its direct costs, but the short-term reputation consequences of exploration.
 The second interpretation concerns the role of data in the digital economy.
@@
 
-With this paper as a common departure point, there are several exciting directions to explore:
+With this paper as a departure point, there are several exciting directions to explore:
 
@@
 Numerical experiments based on real-life datasets would arguably be more realistic.
-\asedit{To deal with real-life datasets, bandit algorithms would probably need to accommodate change over time and} \emph{contexts} (auxiliary signals available before each round).
+To deal with real-life datasets, bandit algorithms would probably need to accommodate change over time and \emph{contexts} (auxiliary signals available before each round).
 
@@
 \noindent One difficulty with the last two directions is that the bandit ``side" of the model becomes considerably more complicated, both in terms of the rewards and in terms of the algorithms.
-Indeed, there are many reasonable ways to model a changing world (resp., dependence on  contexts), and several  substantially different algorithmic approaches to deal \asedit{these models.
-Consequently, the distinctions between better and worse algorithms are not as clear and well-established.}
+Indeed, there are many reasonable ways to model a changing world (resp., dependence on  contexts), and several substantially different algorithmic approaches to deal with these models.
+Consequently, the distinctions between better and worse algorithms are not as clear and well-established.
 

### 2020-09-04 journal paper/content/sec-model.tex
Minor edit.

@@
 First, all/most results extend to arbitrary reward distributions, allow reward-dependent utility, and carry over to a more general version of multi-armed bandits.
-Second, tather than knowing the actual tuple$(\alg[1],\alg[2],\priorMu,\respF)$, agents could have respective common beliefs $(\alg[1]',\alg[2]',\priorMu',\respF')$ than need not be correct; then, all results carry over with respect to these beliefs.
+Second, agents could have beliefs on $(\alg[1],\alg[2],\priorMu,\respF)$ that need not be correct; then, all results carry over with respect to these beliefs.
 Third, all results carry over \asmargincomment{check}  if each agent $t$ has a prior on her arrival time $t$, rather than know it exactly.

### 2020-09-04 journal paper/content/sec-theory.tex
In theory/extensions, added xhdrs re beliefs and non-stationary response function.

@@
 
+\asedit{\xhdr{Beliefs.} Rather than knowing the principals' algorithms $(\alg[1],\alg[2])$, the Bayesian prior $\priorMu$, and the response function $\respF$, agents could have beliefs on these objects that need not be correct.
+If agents have common ``point beliefs", respectively, algorithms $(\alg[1]',\alg[2]')$ prior $\priorMu'$ and response function $\respF'$,  then all our results carry over with respect to these beliefs.}
+
+\asedit{\xhdr{Limited non-stationarity in $\respF$.} 
+Different agents can have different response functions, in the following sense.
+For \HardMaxRandom, our results carry over if each agent $t$ has a \HardMaxRandom response function $\respF$ with parameter $\eps_t\geq \eps_0$.
+For \SoftMaxRandom, different agents can have different response functions that satisfy Definition~\ref{def:SoftMax} (with the same parameters).}
+
 \xhdr{MAB extensions.} Our results carry over, with little or no modification of the proofs, to much more general versions of MAB, as long as it satisfies the i.i.d.

### 2020-09-04 journal paper/content/sec-model.tex
Small edits.

@@
 Second, agents could have beliefs on $(\alg[1],\alg[2],\priorMu,\respF)$ that need not be correct; then, all results carry over with respect to these beliefs.
-Third, all results carry over \asmargincomment{check}  if each agent $t$ has a prior on her arrival time $t$, rather than know it exactly.
-Fourth, we can handle a limited amount of non-stationarity in $\respF$.
-\asmargincomment{check} Finally, the main result on \HardMax extends to time-discounted utilities.
+%Third, all results carry over \asmargincomment{check}  if each agent $t$ has a prior on her arrival time $t$, rather than know it exactly.
+Third, we can handle a limited amount of non-stationarity in $\respF$. \asmargincomment{check} Finally, the main result on \HardMax extends to time-discounted utilities.
 }
@@
 
-\asedit{Principals' utilities are not discounted with time.%
-\footnote{However, our main theoretical result on the greedy algorithm extends to time-discounted rewards.}
- While much of the early work on bandits focuses on  time-discounted rewards \citep{Gittins-book11}, non-discounted  formulations are prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
-The distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.}
+\asedit{Bandit rewards are not discounted with time, reflecting the fact that non-discounted, regret-minimizing formulations has been prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
+Also, the distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.}

### 2020-09-04 journal paper/content/conclusion.tex
Minor edits.

@@
 
-\item \emph{Horizontal differentiation.}
-Horizontally differentiated user preferences may help explain how competition may encourage specialization, \ie how the firms may \emph{learn to specialize} under competition.
+\item 
+\emph{Horizontally differentiated} user preferences may help explain how competition may encourage specialization, \ie how the firms may \emph{learn to specialize} under competition.
 
 \item \emph{Continuous learning.}
-While we focus on learning from scratch in a mainly stationary world, another well-motivated regime is continuous learning, when exploration counteracts change
+While we focus on a stationary world, another well-motivated regime is continuous learning, when exploration counteracts change
 \citep{DynamicMAB-colt08}.

### 2020-09-05 journal paper/ec19paper/content/inverted_u.tex
Minor edits.

@@
 We do this for each ``intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
-For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs \cite{DS-arxiv}).
-Note that \DynamicGreedy is trivially the dominant strategy under monopoly.
+For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs, see \cite{DS-arxiv}).
+\DynamicGreedy is trivially the dominant strategy under monopoly.
 
@@
 Our intuition for why incumbency creates more incentives for exploration is as follows.
-\gaedit{During the period in which the incumbent is the only firm in the market}, reputation consequences of exploration vanish.
+During the period in which the incumbent is the only firm in the market, reputation consequences of exploration vanish.
 Instead, the firm wants to improve its performance as much as possible by the time competition starts.

### 2020-09-05 journal paper/content/sim_details.tex
Copied two paragraphs from the intro.

--- 2020-07-17 journal paper/content/sim_details.tex
@@
 
+\xhdr{Challenges.} An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given instance of a multi-armed bandit problem (and each such experiment is run many times to reduce variance).
+Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit instances\}, and we are looking for findings that are consistent across this entire space.
+It is essential to keep each of the three dimensions small yet representative.
+In particular, we need to capture a huge variety of bandit algorithms and bandit instances with only a few representative examples.
+Further, we need a succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).
+\asmargincomment{copied from intro}
+
 \xhdr{Competition model.} All experiments use \HardMax response function (without mentioning it), except Section~\ref{sec:non_greedy} where we use \HardMaxRandom agents.
@@
 
-\xhdr{MAB algorithms.} We consider the three classes of bandit algorithms discussed in Section~\ref{sec:intro-model} (and in more detail in Appendix~\ref{app:bg}), and look for qualitative differences between them.
+\xhdr{MAB algorithms.} In abstract terms, we posit three types of technology, from ``low" to ``medium" to ``high".
+Concretely, we consider three essential classes of bandit algorithms: ones that never explicitly explore (\emph{greedy algorithms}), ones that explore without looking at the data (\emph{exploration-separating algorithms}), and ones where exploration gradually zooms in on the best arm (\emph{adaptive-exploration algorithms}).
+In the absence of competition, these classes are fairly well-understood: greedy algorithms are terrible for a wide variety of problem instances, exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances, and adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances (see Appendix~\ref{app:bg} for more details).
+\asmargincomment{copied from CODE submission}
+
+We look for qualitative differences between these three classes under competition.
 We take a representative algorithm from each class.

### 2020-09-05 journal paper/itcs18paper/app-examples.tex
Minor edit.
Removed some coloring.

@@
 The set of all arms is $A$; mean reward of arm $a$ is denoted $\mu_a$.
-\asedit{The mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$.}
+The mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$.
 We let
@@
 \begin{theorem}[\citet{Selke-BE-2020}]\label{thm:TS-mono}
-\asedit{Assume the prior $\priorMu$ is independent across arms.} Then \Thompson is \bmonotone. Consequently, it satisfies $\BIR(t)\leq \tildeO(t^{-1/2})$.
+Assume the prior $\priorMu$ is independent across arms. Then \Thompson is \bmonotone. Consequently, it satisfies $\BIR(t)\leq \tildeO(t^{-1/2})$.
 \end{theorem}
@@
 Let $a^*$ be the best arm: $a^*\in \argmax_a \mu_a$.
-In each round $t$, \Thompson samples arm $a_t$ as follows:
-\begin{align}\label{eq:TS-defn}
-    \Pr[a_t=a] := {\Pr}^t[a^*=a]
-    \quad\text{for each arm $a$}.
-\end{align}
-\swcomment{do we want to say $\Pr$ is over sampling first from prior and then generate history?}
+In each round $t$, \Thompson samples arm $a_t$ from the posterior distribution $p_t(a) = {\Pr}^t[a^*=a]$.
 

### 2020-09-05 journal paper/content/sec-bg.tex
App/bg: copied the 3 classes list from the intro.

@@
 We present self-contained background on multi-armed bandits (MAB), to make the paper accessible to researchers who are not experts on MAB.
-We focus on the three classes of MAB algorithms listed in Section~\ref{expts-prelims}: adaptive-exploration algorithms, exploration-separating algorithms, and greedy algorithms.
+We focus on three algorithm classes, as in Section~\ref{sec:sim}:
+
+\begin{itemize}
+\item \emph{Greedy algorithms} that strive to maximize the reward for the next round given the available information. Thus, they always ``exploit" and never explicitly ``explore".
+    \asmargincomment{copied the list from the intro.}
+
+\item \emph{Exploration-separating algorithms}
+ that separate exploration and exploitation: essentially, each round is dedicated to one and completely ignores the other.
+
+\item \emph{Adaptive-exploration} algorithms that combine exploration and exploitation, and gradually sway the exploration choices towards more promising alternatives.
+\end{itemize}
+
 We list the precise upper and lower bounds on the regret rates only to illustrate how the three algorithm classes are separated from one another; the exact results are not essential for the paper.
-
 We use standard conventions: $O(f(t))$ and $\Omega(f(t))$ means at most (resp., at least) $f(n)$, up to constant factors, starting from large enough $t$.
@@
 In each round $t\in [T]$, the algorithm chooses an arm and receives a reward $r_t\in[0,1]$ for this arm, drawn from a fixed but unknown distribution.%
-\footnote{\asedit{All ``negative" results (\ie lower bounds on regret) assume reward distributions with constant variance.}}
+\footnote{All ``negative" results (\ie lower bounds on regret) assume reward distributions with constant variance.}
 The algorithm's goal is to maximize the total reward.
@@
 
-\asedit{Some MAB algorithms are Bayesian, \eg Thompson Sampling,  in the sense that they input a prior on mean rewards, and attain strong Bayesian guarantees (in expectation over the prior) when the prior is correct.
-Such algorithms can also be initialized with some simple `fake' priors; in fact, Thompson Sampling satisfies the optimal regret bounds listed above.}
+Some MAB algorithms are Bayesian, \eg Thompson Sampling,  in the sense that they input a prior on mean rewards, and attain strong Bayesian guarantees (in expectation over the prior) when the prior is correct.
+Such algorithms can also be initialized with some simple `fake' priors; in fact, Thompson Sampling satisfies the optimal regret bounds listed above.
 
@@
 
-The three classes of algorithms are well-separated in terms of \BIR, much like they are in terms of regret.
+The three classes are well-separated in terms of \BIR, much like they are in terms of regret.
 %(The precise regret rates are listed below only to illustrate this point.)
@@
         $\BIR(t) = \tildeO\rbr{t^{-\gamma}}$
-    for all $t$, then $\gamma \leq \nicefrac{1}{2}$.
+    for all rounds $t$ and some constant $\gamma$, then $\gamma \leq \nicefrac{1}{2}$.
 

### 2020-09-05 journal paper/ec19paper/content/perf_in_iso.tex
Removed some coloring.

@@
 We plot the \emph{mean reputation trajectory}: how this score evolves over time.
-\asedit{We also plot the trajectory for instantaneous rewards (\emph{not} averaged over the previous time-periods), which provides a better view into algorithm's performance at a given time.}%
+We also plot the trajectory for instantaneous rewards (\emph{not} averaged over the previous time-periods), which provides a better view into algorithm's performance at a given time.%
 \footnote{For ``instantaneous reward" at a given time $t$, we report the average (over all \MRVs) of the mean rewards at this time, instead of the average of the \emph{realized} rewards, so as to decrease the noise.}
@@
 \begin{finding}
-\textit{The mean reputation trajectories \asedit{and the instantaneous reward trajectories} are arranged as predicted by prior work:
+\textit{The mean reputation trajectories and the instantaneous reward trajectories are arranged as predicted by prior work:
     $\Thompson > \DynamicEpsGreedy > \DynamicGreedy$ for a sufficiently long time horizon $T$.}

### 2020-09-05 journal paper/content/sec-intro.tex
one small tweak to intro and use newer econometrica bst

@@
 Each principal's strategy is a bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.
-\asedit{(This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.)}
+\asedit{\gadelete{(}This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.\gadelete{)}}
 %Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical. 

### 2020-09-05 journal paper/content/related_work.tex
small fixes in related work

@@
 Second, the inputs to agents' decision rule are prices (which are directly controlled by the principals), rather than the quality of the chosen actions (which is not known a priori).
-Their focus on learning the decision rule, whereas ours is on learning quality.
-Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria, whereas we focus on adoption of better bandit algorithms.}
+Their focus \gaedit{is} on learning the decision rule, whereas ours is on learning \gaedit{product} quality.
+Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria, whereas we focus on \gaedit{the} adoption of better bandit algorithms.}
 
@@
 We observe a similar effect when we investigate the role that reputation can serve as a barrier to entry.
-\asmargincomment{I can't parse this sentence, could you pls simplify?}
-However, in our model this effect is further strengthened by the fact that the firms have to learn while competing adding that the incumbent may not only have a reputational advantage, but additionally has a further advantage due to the data it acquires in the incumbency period.
-Thus, our model also highlights the role that data can serve as a barrier to entry in online markets which has similarly been noted in \cite{de2020data}.
+\asmargincomment{I can't parse this sentence, could you pls simplify?}\gaedit{In our model the first-mover advantage further provides the incumbent with a ``head start" on data collection relative to the late arrivers.}
+\gadelete{However, in our model this effect is further strengthened by the fact that the firms have to learn while competing adding that the incumbent may not only have a reputational advantage, but additionally has a further advantage due to the data it acquires in the incumbency period.}Thus, our model also highlights the role that data can serve as a barrier to entry in online markets which has similarly been noted in \cite{de2020data}.
 For an extensive overview of the other channels through which first-mover advantages can lead to a competitive advantage, see \cite{kerin1992first}.

### 2020-09-05 journal paper/content/sec-model.tex
minor points

@@
 %Third, all results carry over \asmargincomment{check}  if each agent $t$ has a prior on her arrival time $t$, rather than know it exactly.
-Third, we can handle a limited amount of non-stationarity in $\respF$. \asmargincomment{check} Finally, the main result on \HardMax extends to time-discounted utilities.
+Third, we can handle a limited amount of non-stationarity in $\respF$ \gaedit{for the \HardMaxRandom and \SoftMaxRandom decision rules.} \asmargincomment{check} Finally, the main result on \HardMax extends to time-discounted utilities.
 }
@@
 "Non-strategic" exploration is well-studied in machine learning, and yet it remains a very complex subject in research (as evidenced, \eg by the huge amount of activity therein).
-Even the seemingly simple algorithms are not straightforward to deploy in practice, and require a substantial investment in infrastructure \cite[\eg see the discussions in][]{DS-arxiv}.
+Even \gadelete{the} seemingly simple algorithms are not straightforward to deploy in practice, and require a substantial investment in infrastructure \cite[\eg see the discussions in][]{DS-arxiv}.
 Responding to the competition represents another layer of complexity which has not been previously studied in this context, to the best of our knowledge, let alone made even remotely practical.

### 2020-09-05 journal paper/content/related_work.tex
remove color edits and fix formatting of tables in the appendix

--- 2020-09-04 journal paper/content/related_work.tex
@@
 
-\asedit{\citet{keller2003price} studied the interplay of exploration and competition \emph{on prices}.
+\citet{keller2003price} studied the interplay of exploration and competition \emph{on prices}.
 This distinction leads to several important differences.
@@
 Second, the inputs to agents' decision rule are prices (which are directly controlled by the principals), rather than the quality of the chosen actions (which is not known a priori).
-Their focus on learning the decision rule, whereas ours is on learning quality.
-Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria, whereas we focus on adoption of better bandit algorithms.}
+Their focus is on learning the decision rule, whereas ours is on learning product quality.
+Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria, whereas we focus on the adoption of better bandit algorithms.
 
@@
 %whether and when network effects manifest themselves 
-\asedit{network effects}
+network effects
 in data-intensive markets \citep{prufer2017competing, hagiu2020data}, but typically models learning as a reduced-form function of past consumer history and focuses on the role of prices.
@@
 We observe a similar effect when we investigate the role that reputation can serve as a barrier to entry.
-\asmargincomment{I can't parse this sentence, could you pls simplify?}
-However, in our model this effect is further strengthened by the fact that the firms have to learn while competing adding that the incumbent may not only have a reputational advantage, but additionally has a further advantage due to the data it acquires in the incumbency period.
+In our model the first-mover advantage further provides the incumbent with a ``head start" on data collection relative to the late arrivers.
 Thus, our model also highlights the role that data can serve as a barrier to entry in online markets which has similarly been noted in \cite{de2020data}.

### 2020-09-05 journal paper/content/sec-intro.tex
remove color edits and fix formatting of tables in the appendix

--- 2020-09-04 journal paper/content/sec-intro.tex
@@
 Each principal's strategy is a bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.
-\asedit{(This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.)}
+This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.
 %Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical. 

### 2020-09-05 journal paper/content/related_work.tex
add bergemann valimaki paper comparisons

@@
 The most relevant thread concerns MAB with stochastic rewards and no auxiliary structure, which is the problem faced by each principal in our model (see Appendix~\ref{app:bg}).
-The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.
-This literature is covered in dedicated monographs,
+The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.\footnote{There is a superficial similarity --- in name only --- between this paper and the line of work on ``dueling bandits" \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
+The latter is not about competing bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".} This literature is covered in dedicated monographs,
 \citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.%
@@
 Their focus is on learning the decision rule, whereas ours is on learning product quality.
-Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria, whereas we focus on the adoption of better bandit algorithms.
-
-There is a superficial similarity --- in name only --- between this paper and the line of work on ``dueling bandits"
-    \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
-The latter is not about competing bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".
+Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria (MPE), whereas we focus on the adoption of better bandit algorithms.
+\gaedit{\citet{bergemann1997market,bergemann2000experimentation} studied a model where there is an incumbent with known product quality and an entrant with unknown, but exogenously determined, quality.
+Similar to \citet{keller2003price}, the firms compete on prices and they characterize the MPE of dynamic competition between the two firms.
+In their model, consumer purchases of the entrant's product result in an informational externality that allows both the firms and consumers to gather information on the entrant's product quality.
+However, we focus on a model where firms' product quality is endogenous and the focus of exploration in a context where the price mechanism is absent.}
 

### 2020-09-07 journal paper/content/sec-intro.tex
add line about product quality in intro
Intro: revised Guy's sentence in para 1, and added a sentence in "significance".

@@
 Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a standard and well-studied framework for exploration and making decisions under uncertainty.
+\asedit{This trend has been stimulated by two factors: almost-zero cost of deploying iterations of a product (provided an initial infrastructure investment), and the fact that many online platforms primarily compete on product quality, rather than price.
+(\eg because they are supported by ads or cheap subscriptions).}
 %\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
@@
 \xhdr{Significance.}
+\asedit{This paper paper connects two huge literatures: on regret-minimizing bandits in computer science and on competition in economics, while staying true to each.}
 Our results have a dual purpose: shed light on real-world implications of some typical scenarios, and investigate the space of models for describing the real world.

### 2020-09-07 journal paper/content/related_work.tex
Started another pass.

@@
 MAB problems have been studied for many decades by researchers from computer science, operations research, statistics and economics, generating a vast and multi-threaded literature.
-The most relevant thread concerns MAB with stochastic rewards and no auxiliary structure, which is the problem faced by each principal in our model (see Appendix~\ref{app:bg}).
-The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.\footnote{There is a superficial similarity --- in name only --- between this paper and the line of work on ``dueling bandits" \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
-The latter is not about competing bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".} This literature is covered in dedicated monographs,
-\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.%
-\footnote{\citet{Gittins-book11} covers Markovian formulations, the other three are concerned with regret-minimization.}
+The most relevant thread concerns the basic model of regret-minimizating bandits with stochastic rewards and no auxiliary structure (which is the problem faced by each principal in our model), see Appendix~\ref{app:bg} for background.
+This basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.\footnote{There is a superficial similarity --- in name only --- between this paper and the line of work on ``dueling bandits" \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
+The latter is not about competing bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".}
+Dedicated monographs \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} cover the work on regret-minimizing formulations (which mainly comes from computer science), and the classic book 
+\citep{Gittins-book11} focuses on the Markovian formulations (which predate regret-minimization and were mainly studied in operations research and statistics). 
 Various connections to economics and game theory are detailed in
@@
 
-The three-way tradeoff between exploration, exploitation and incentives has been studied in several settings very different from ours:
+The three-way tradeoff between exploration, exploitation and incentives has been studied in mechanism design, in several scenarios very different from ours:
 incentivizing exploration in a recommendation system
@@
     \citep{Bobby-Glen-ec16},
-human computation
-    \citep[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13},
-and social learning 
-    \citep[\eg][]{Bolton-econometrica99,Keller-econometrica05,Johari-ec12}. 
+and human computation
+    \citep[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}. 
 A literature review of this work can be found in
 \citep[Ch. 11.6]{slivkins-MABbook}.
+
+\xhdr{Exploration and competition.}
+\ascomment{TODO: revise/expand this xhdr.}
 
@@
 However, we focus on a model where firms' product quality is endogenous and the focus of exploration in a context where the price mechanism is absent.}
+
+\ascomment{to elaborate}
+ social learning
+    \citep[\eg][]{Bolton-econometrica99,Keller-econometrica05,Johari-ec12}.
 

### 2020-09-07 journal paper/content/sec-intro.tex
Suppressed a sentence.

--- 2020-09-05 journal paper/content/sec-intro.tex
@@
 Our results have a dual purpose: shed light on real-world implications of some typical scenarios, and investigate the space of models for describing the real world.
-As an example for the latter, the \HardMax model with simultaneous entry is arguably the most natural model to study \emph{a priori}.
-However, our results elucidate the need for more refined models that allow for ``free exploration" (\eg via random agents or early entry).
+As an example for the latter: while the \HardMax model with simultaneous entry is arguably the most natural model to study \emph{a priori}, our results elucidate the need for more refined models with ``free exploration" (\eg via random agents or early entry).
+\asedit{On a technical level, we connect a literature on regret-minimizing bandits in computer science and that on competition in economics.}
 

### 2020-09-07 journal paper/content/related_work.tex
In "related work", added a page on "competition & exploration".

@@
 The most relevant thread concerns the basic model of regret-minimizating bandits with stochastic rewards and no auxiliary structure (which is the problem faced by each principal in our model), see Appendix~\ref{app:bg} for background.
-This basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.\footnote{There is a superficial similarity --- in name only --- between this paper and the line of work on ``dueling bandits" \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
-The latter is not about competing bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".}
+This basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.%
+\footnote{There is a superficial similarity, in name only, between this paper and the work on ``dueling bandits" (starting from \citet{Yue-dueling12,Yue-dueling-icml09}).
+In the latter, an algorithm chooses two arms in each round, and the algorithm only observes which arm has ``won the duel".}
 Dedicated monographs \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} cover the work on regret-minimizing formulations (which mainly comes from computer science), and the classic book
@@
 \xhdr{Exploration and competition.}
-\ascomment{TODO: revise/expand this xhdr.}
+\asmargincomment{This whole xhdr is new.}
+\citet{bergemann1997market,bergemann2000experimentation,keller2003price} studied the interplay of exploration and competition for users when the competing firms experiment with \emph{prices}, rather than design alternatives.
+All three papers consider strategies that respond to competition, and analyze Markov Perfect Equilibria (MPE), whereas we focus on the adoption of better bandit algorithms.
+We discuss them in more detail below.
 
-\citet{keller2003price} studied the interplay of exploration and competition \emph{on prices}.
+In \citet{keller2003price}, the competition is entirely on prices, rather than product quality.
+This distinction leads to several important differences.
+First, the data externality -- that each customer brings a new data point to a firm if and only if he chooses this firm -- is absolutely crucial to our setting, and absent in theirs.
+Second, the inputs to agents' decision rule are prices, rather than the quality of the chosen alternatives; the former is directly controlled by the firms, whereas the latter is not known a priori.
+Third, the goal of exploration is to learn the agents' decision rule.
+
+The earlier work by \citet{bergemann1997market,bergemann2000experimentation} studies competition on both prices and quality, and allows for the data externality mentioned above.
+Specifically, the ``entrant" firm offers a new product of unknown quality, which is ``explored" if and only if the entrant attracts customers.
+However, because the entrant can only control the prices but not the product itself, the nature of exploration is fundamentally different.
+First, the goal of exploration is to reveal information about a fixed product.
+Second, the entrant can only affect the \emph{quantity} of the said information, but not \emph{which} information is being revealed.
+Third, revealing \emph{more} information is not necessarily better for the entrant (if the product's quality is actually low).
+Fourth, the social planner's exploration problem is very different from the firms'.%
+\footnote{The planner faces a two-armed bandit problem, rather than a pricing problem.
+This is because the planner can directly choose a firm for each agent (by choosing a very low price for this firm, and a very high price for the other firm, because payments cancel out in social welfare).}
+Consequently, when one compares competition to the planner, one can compare the outcomes but not the algorithms.
+
+%In \citet{keller2003price}, the competition is entirely on prices, rather than product quality.
 This distinction leads to several important differences.
@@
 Third, they consider strategies that respond to competition and analyze Markov Perfect Equilibria (MPE), whereas we focus on the adoption of better bandit algorithms.
-\gaedit{\citet{bergemann1997market,bergemann2000experimentation} studied a model where there is an incumbent with known product quality and an entrant with unknown, but exogenously determined, quality.
+
+%\gaedit{\citet{bergemann1997market,bergemann2000experimentation} studied a model where there is an incumbent with known product quality and an entrant with unknown, but exogenously determined, quality.
 Similar to \citet{keller2003price}, the firms compete on prices and they characterize the MPE of dynamic competition between the two firms.
@@
 
-\ascomment{to elaborate}
- social learning
-    \citep[\eg][]{Bolton-econometrica99,Keller-econometrica05,Johari-ec12}.
+\cite{Ufuk-jeea15} study exploration in a ``research competition".
+Two firms race towards a big discovery such as a new drug.
+Each firm solves a bandit problem with two arms, corresponding to safe and risky lines of research.
+The two algorithms do not interact, directly or indirectly, until one firm makes the discovery and wins the game.
+In comparison, we study a ``product competition" in which the two firms interact continuously (via the customers' choices), accrue rewards incrementally, and compete for individual customers.
+
+In the line of work on \emph{strategic experimentation} (starting from \citet{Bolton-econometrica99,Keller-econometrica05}, see \citet{Horner-survey16} for a survey), agents explore and learn over time in a shared environment.
+Thus, we have exploration algorithms which interact with each other strategically, \eg each agent would prefer to free-ride on the exploration done by others.
+However, the agents do not compete with each other in any meaningful sense.
+
+A long line of work from electrical engineering and computer science, starting from
+\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, targets an application to \emph{cognitive radios}, where multiple radios transmit simultaneously in a shared medium and compete for bandwidth.
+Each radio chooses channels over time using a multi-armed bandit algorithm.
+This work studies a repeated game between bandit algorithms, and focuses on designing algorithms which work well in this game, under various assumptions on communication, synchronization and collisions.
+
+%and continuing to, \eg \citet{MultiPlayerMAB-Mannor14,MultiPlayerMAB-Shamir-icml16,MultiPlayerMAB-Perchet-18,MultiPlayerMAB-Sellke-19}.
 

### 2020-09-07 journal paper/content/related_work.tex
Minor edit.

@@
 \footnote{The planner faces a two-armed bandit problem, rather than a pricing problem.
-This is because the planner can directly choose a firm for each agent (by choosing a very low price for this firm, and a very high price for the other firm, because payments cancel out in social welfare).}
+This is because the planner can directly choose a firm for each agent (by choosing a very low price for this firm, and a very high price for the other).}
+% firm, because payments cancel out in social welfare).}
 Consequently, when one compares competition to the planner, one can compare the outcomes but not the algorithms.

### 2020-09-08 journal paper/content/related_work.tex
Another pass on related work.

@@
 This basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.%
-\footnote{There is a superficial similarity, in name only, between this paper and the work on ``dueling bandits" (starting from \citet{Yue-dueling12,Yue-dueling-icml09}).
-In the latter, an algorithm chooses two arms in each round, and the algorithm only observes which arm has ``won the duel".}
-Dedicated monographs \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} cover the work on regret-minimizing formulations (which mainly comes from computer science), and the classic book
+\footnote{There is a superficial similarity, in name only, between this paper and the work on ``dueling bandits" \citep[starting from][]{Yue-dueling12,Yue-dueling-icml09}.
+In this work, there is only one bandit algorithm which chooses two arms in each round, and the only observes which arm has ``won the duel".}
+Dedicated monographs \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} cover the work on regret-minimizing formulations (which mainly comes from computer science). The classic book
 \citep{Gittins-book11} focuses on the Markovian formulations (which predate regret-minimization and were mainly studied in operations research and statistics).
@@
 
-\cite{Ufuk-jeea15} study exploration in a ``research competition".
-Two firms race towards a big discovery such as a new drug.
-Each firm solves a bandit problem with two arms, corresponding to safe and risky lines of research.
-The two algorithms do not interact, directly or indirectly, until one firm makes the discovery and wins the game.
-In comparison, we study a ``product competition" in which the two firms interact continuously (via the customers' choices), accrue rewards incrementally, and compete for individual customers.
+\vspace{2mm}
 
+\asmargincomment{It serves us to discuss this work here, IMO, and we have an excuse.}
 In the line of work on \emph{strategic experimentation} (starting from \citet{Bolton-econometrica99,Keller-econometrica05}, see \citet{Horner-survey16} for a survey), agents explore and learn over time in a shared environment.
@@
 
+Several papers study competition between two principals who run algorithms but do not interact, directly or indirectly, until the very end of the game.
+\cite{Ufuk-jeea15} consider a ``research competition" between two firms racing towards a big discovery.
+\asmargincomment{This way, this paper seems not that relevant, doesn't it?}
+Each firm deploys a bandit algorithm with two arms, corresponding to safe and risky lines of research.
+The firms do not interact untl one of them makes the discovery and wins the game.
+In the ``dueling algorithms" framework of \citet{DuelingAlgs-stoc11}, each principals runs an algorithm for the same problem.
+All inputs are observable at once, and principals' payoffs are binary (win/lose).
+\citet{ben2017best, ben2019regression} study competition between ``offline" machine learning algorithms.
+In comparison, we study a ``product competition" in which the two firms interact continuously (via the customers' choices), accrue rewards incrementally, and compete for individual customers.
+
+\asmargincomment{This happens to be about exploration \& competition ...}
 A long line of work from electrical engineering and computer science, starting from
@@
 
-\xhdr{Competition.} The competition vs.
+\xhdr{Competition in economics.} The competition vs.
 innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{aghion2005competition,Vives-08}.
@@
 
-\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers.
+\asedit{\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers; we find a similar effect with ``reputation advantage".
+The role of data as a barrier to entry in online markets has been noted in \cite{de2020data}; we find a similar effect with ``data advantage".
+\citet{kerin1992first} overview the other channels through which first-mover advantage can lead to a competitive advantage.}
+
+%\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers.
 We observe a similar effect when we investigate the role that reputation can serve as a barrier to entry.
@@
 Neither is applicable to our setting, since there are no prices, and market shares are endogenous.
-
-The ``dueling algorithms" framework of \citet{DuelingAlgs-stoc11} studies competition between two principals, each running an algorithm for the same problem.
-However, they consider algorithms for offline / full input scenarios, and posit binary (win/lose) payoffs for the principals.
-Similarly, \citet{ben2017best, ben2019regression} study competition between offline learning algorithms.
-Whereas we focus on bandit algorithms and competition for users.
 

### 2020-09-10 journal paper/journal_main.tex
Reverted the section title for app/bg.
Inserted some \newpage macro.
one small tweak to intro and use newer econometrica bst
Minor edits in the title note.

@@
 \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19},
-with a unified and streamlined presentation and expanded background materials.
+with a unified and streamlined presentation and expanded related work and background materials.
 All theoretical results are from \citet{CompetingBandits-itcs18}, and all experiments are from \citet{CompetingBandits-ec19}.
 Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.
-We would like to thank Yeon-Koo Che, Glen Weyl, Sven Rady, and seminar participants at Columbia, ACM Economics and Computation, and Innovations in Theoretical Computer Science for helpful comments and conversations.
+We would like to thank Ian Ball, Yeon-Koo Che, Glen Weyl and Sven Rady, as well as seminar participants at Columbia and conference participants at Innovations in Theoretical Computer Science 2018 and ACM Economics and Computation 2019, for helpful comments and conversations.
 All errors are our own.} }
@@
 
+\newpage
 \section{Numerical simulations: the \ExptsModel}
@@
 
-\bibliographystyle{econometrica}
+\bibliographystyle{ecta.bst}
 %\begin{small}
@@
 
-\section{Background on multi-armed bandits}
+\section{Background for non-specialists: \\multi-armed bandits}
 \label{app:bg}
@@
 \subfile{content/sec-bg}
+
+\newpage
 

### 2020-09-10 journal paper/content/sec-bg.tex
Minor edits.

@@
-We present self-contained background on multi-armed bandits (MAB), to make the paper accessible to researchers who are not experts on MAB.
+We present self-contained background on multi-armed bandits (MAB), to make the paper accessible to researchers who are not experts on MAB. More details can be found in any of the monographs
+\citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
+
 We focus on three algorithm classes, as in Section~\ref{sec:sim}:
@@
 
-We list the precise upper and lower bounds on the regret rates only to illustrate how the three algorithm classes are separated from one another; the exact results are not essential for the paper.
-We use standard conventions: $O(f(t))$ and $\Omega(f(t))$ means at most (resp., at least) $f(n)$, up to constant factors, starting from large enough $t$.
+While we list precise upper and lower bounds on the regret rates, the main goal is to illustrate how the three algorithm classes are separated from one another; the exact results are not essential for this paper.
+For ease of presentation, we use standard asymptotic notation from computer science: $O(f(t))$ and $\Omega(f(t))$ means at most (resp., at least) $f(n)$, up to constant factors, starting from large enough $t$.
 Likewise, $\tildeO(f(t))$ notation suppresses the $\polylog(t)$ factors.
@@
 Thompson Sampling \citep{Thompson-1933,TS-survey-FTML18},%
-\footnote{While Thompson Sampling is the first and arguably best-known bandit algorithm, the corresponding regret bounds have been proved relatively recently \citep{Shipra-colt12,Kaufmann-alt12,Shipra-aistats13}.}
+\footnote{While Thompson Sampling dates back to 1933 and is probably the best-known bandit algorithm, its regret has not been understood until recently \citep{Shipra-colt12,Kaufmann-alt12,Shipra-aistats13}.}
 UCB1 \citep{bandits-ucb1},
-and Successive Elimination \citep{EvenDar-icml06}. There is a substantial amount of follow-up work on refining these regret rates, but it is not as relevant to this paper.
+and Successive Elimination \citep{EvenDar-icml06}.%
+\footnote{A substantial follow-up work on more ``refined" regret rates is not as relevant to this paper.}
+These algorithms are very simple to describe.
+Focus on one round and consider the posterior distribution and/or the confidence interval on each arm's mean reward.
+Thompson Sampling draws a sample (``score") from each arm's posterior distribution, and picks an arm with the largest score.
+UCB1 picks an arm with the largest upper confidence bound.
+Successive Elimination eliminates an arm once it is worse than some other arm with high confidence, and chooses uniformly among the remaining arms.
 
@@
 Both algorithms, and the associated $\tildeO(T^{2/3})$ regret bounds, have been ``folklore knowledge" for decades.
-The general definition of ``exploration-separating algorithms" and the associated lower bounds trace back to \citet{MechMAB-ec09}.%
-\footnote{\citet{MechMAB-ec09} consider a closely related, but technically different setting, which can be easily ``translated" into ours (either as a corollary or as another application of the same proof technique).
-Detailing this is beyond the scope of this appendix.}
+The general definition and lower bounds trace back to \citet{MechMAB-ec09}.%
+\footnote{\citet{MechMAB-ec09} consider a closely related, but technically different setting, which can be easily ``translated" into ours (either as a corollary or as another application of the same proof technique).}
 
@@
 Switching from ``greedy" to ``exploration-separating" to ``adaptive-exploration" algorithms involves substantial adoption costs in infrastructure and personnel training \citep{DS-arxiv}.
-Inserting exploration into a complex decision-making pipeline necessitates a substantial awareness of the technology and a certain change in a mindset for the personnel, as well as an infrastructure to collect and analyze the data.
+Inserting exploration into a complex decision-making pipeline necessitates a substantial awareness of the technology and a certain change in mindset, as well as an infrastructure to collect and analyze the data.
 Adaptive exploration requires the said infrastructure to propagate the data analysis back to the ``front-end" where the decisions are made, and do it on a sufficiently fast and regular cadence.
-Framing the problem (in terms of the appropriate modeling assumptions and action features) and debugging the machine learning code tend to be quite subtle, too.
+Framing the problem (\eg choosing modeling assumptions and action features) and debugging the machine learning algorithms tend to be quite subtle, too.
 
-The lower bounds mentioned above are fairly typical, in the sense that they hold for a wide variety of problem instances, even though they are usually (and most cleanly) presented as worst-case.
+The lower bounds mentioned above are fairly typical: while they are usually (and most cleanly) presented as worst-case, the actually hold for a wide variety of problem instances.
 The $\Omega(\sqrt{T})$ lower bound from \citet{bandits-exp3} can be extended to hold for most problem instances, in the following sense: for each instance $\mI$ there exists a ``decoy instance" $\mI'$ such that any algorithm incurs regret $\Omega(\sqrt{T})$ on at least one of them.
@@
 in fact holds for all problem instances and all algorithms that are not \emph{terrible} on the large-gap instances \citep{Lai-Robbins-85}.
-The $\Omega(T^{2/3})$ lower bound for exploration-separating algorithms in fact applies to all problem instances, as long as the algorithm achieves $\tildeO(T^{2/3})$ regret rate in the worst case \citep{MechMAB-ec09}.
-Moreover, there is a tradeoff between the worst-case upper bound on the regret rate and a lower bound that applies for all problem instances \citep[Theorem 4.3 in][using the same proof technique]{MechMAB-ec09}.
+The $\Omega(T^{2/3})$ lower bound for exploration-separating algorithms in fact applies to all problem instances, as long as the algorithm achieves $\tildeO(T^{2/3})$ regret rate in the worst case \citep{MechMAB-ec09}.%
+\footnote{Moreover, there is a tradeoff between the worst-case upper bound on the regret rate and a lower bound that applies for all problem instances \citep[Theorem 4.3 in][]{MechMAB-ec09}.}
 
-Some MAB algorithms are Bayesian, \eg Thompson Sampling,  in the sense that they input a prior on mean rewards, and attain strong Bayesian guarantees (in expectation over the prior) when the prior is correct.
-Such algorithms can also be initialized with some simple `fake' priors; in fact, Thompson Sampling satisfies the optimal regret bounds listed above.
+Some MAB algorithms, \eg Thompson Sampling, are Bayesian: they input a prior on mean rewards, and attain strong Bayesian guarantees (in expectation over the prior) when the prior is correct.
+Such algorithms can also be initialized with some simple `fake' priors; in fact, this is how Thompson Sampling can be made to satisfy the optimal regret bounds listed above.
 
@@
 \begin{itemize}
-\item \DynGreedy has at least a constant \BIR for many reasonable priors (where the constant can depend on $K$ and the prior, but not on $t$).
-The reason and the proof are exactly the same as for regret.
+\item \DynGreedy has at least a constant \BIR for many reasonable priors (where the constant can depend on $K$ and the prior, but not on $t$). The reason / proof is the same as for regret.
 

### 2020-09-10 journal paper/content/related_work.tex
a pass through related work and remaining color edits

@@
 \xhdr{Exploration and competition.}
-\asmargincomment{This whole xhdr is new.}
 \citet{bergemann1997market,bergemann2000experimentation,keller2003price} studied the interplay of exploration and competition for users when the competing firms experiment with \emph{prices}, rather than design alternatives.
@@
 The earlier work by \citet{bergemann1997market,bergemann2000experimentation} studies competition on both prices and quality, and allows for the data externality mentioned above.
-Specifically, the ``entrant" firm offers a new product of unknown quality, which is ``explored" if and only if the entrant attracts customers.
+Specifically, the ``entrant" firm offers a new product of unknown, \gaedit{exogenously determined}, quality, which is ``explored" if and only if the entrant attracts customers.
 However, because the entrant can only control the prices but not the product itself, the nature of exploration is fundamentally different.
@@
 Second, the entrant can only affect the \emph{quantity} of the said information, but not \emph{which} information is being revealed.
-Third, revealing \emph{more} information is not necessarily better for the entrant (if the product's quality is actually low).
-Fourth, the social planner's exploration problem is very different from the firms'.%
-\footnote{The planner faces a two-armed bandit problem, rather than a pricing problem.
-This is because the planner can directly choose a firm for each agent (by choosing a very low price for this firm, and a very high price for the other).}
+\gadelete{Third, revealing more information is not necessarily better for the entrant (if the product's quality is actually low).} \gaedit{Third}, the social planner's exploration problem is very different from the firms' \gaedit{so that when one compares competition to the planner, one can only compare outcomes, not the algorithms.}\footnote{The planner faces a two-armed bandit problem, rather than a pricing problem \gaedit{because the planner can directly choose a firm for each agent.}}
 % firm, because payments cancel out in social welfare).}
-Consequently, when one compares competition to the planner, one can compare the outcomes but not the algorithms.
 
@@
 However, we focus on a model where firms' product quality is endogenous and the focus of exploration in a context where the price mechanism is absent.}
-
-\vspace{2mm}
-
-\asmargincomment{It serves us to discuss this work here, IMO, and we have an excuse.}
 In the line of work on \emph{strategic experimentation} (starting from \citet{Bolton-econometrica99,Keller-econometrica05}, see \citet{Horner-survey16} for a survey), agents explore and learn over time in a shared environment.
@@
 \cite{Ufuk-jeea15} consider a ``research competition" between two firms racing towards a big discovery.
-\asmargincomment{This way, this paper seems not that relevant, doesn't it?}
 Each firm deploys a bandit algorithm with two arms, corresponding to safe and risky lines of research.
-The firms do not interact untl one of them makes the discovery and wins the game.
+The firms do not interact until one of them makes the discovery and wins the game.
 In the ``dueling algorithms" framework of \citet{DuelingAlgs-stoc11}, each principals runs an algorithm for the same problem.
@@
 
-\asmargincomment{This happens to be about exploration \& competition ...}
 A long line of work from electrical engineering and computer science, starting from
-\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, targets an application to \emph{cognitive radios}, where multiple radios transmit simultaneously in a shared medium and compete for bandwidth.
+\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, \gaedit{focuses on competition for resources, not competition for consumers.
+Specifically, this literature targets an application} to \emph{cognitive radios}, where multiple radios transmit simultaneously in a shared medium and compete for bandwidth.
 Each radio chooses channels over time using a multi-armed bandit algorithm.
@@
 
-\asedit{\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers; we find a similar effect with ``reputation advantage".
+\cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers; we find a similar effect with ``reputation advantage".
 The role of data as a barrier to entry in online markets has been noted in \cite{de2020data}; we find a similar effect with ``data advantage".
-\citet{kerin1992first} overview the other channels through which first-mover advantage can lead to a competitive advantage.}
+\citet{kerin1992first} overview the other channels through which first-mover advantage can lead to a competitive advantage.
 
@@
 Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom.
-\asedit{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
+While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.
 

### 2020-09-10 journal paper/content/sec-intro.tex
a pass through related work and remaining color edits

@@
 Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a standard and well-studied framework for exploration and making decisions under uncertainty.
-\asedit{This trend has been stimulated by two factors: almost-zero cost of deploying iterations of a product (provided an initial infrastructure investment), and the fact that many online platforms primarily compete on product quality, rather than price.
-(\eg because they are supported by ads or cheap subscriptions).}
+This trend has been stimulated by two factors: almost-zero cost of deploying iterations of a product (provided an initial infrastructure investment), and the fact that many online platforms primarily compete on product quality, rather than price.
+(\eg because they are supported by ads or cheap subscriptions).
 %\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
@@
 %We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
-\asedit{The barriers for innovation arise entirely from the reputational consequences of exploration in competition, whereas the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs.}
+The barriers for innovation arise entirely from the reputational consequences of exploration in competition, whereas the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs.
 
@@
 As an example for the latter: while the \HardMax model with simultaneous entry is arguably the most natural model to study \emph{a priori}, our results elucidate the need for more refined models with ``free exploration" (\eg via random agents or early entry).
-\asedit{On a technical level, we connect a literature on regret-minimizing bandits in computer science and that on competition in economics.}
+On a technical level, we connect a literature on regret-minimizing bandits in computer science and that on competition in economics.
 

### 2020-09-10 journal paper/content/sec-model.tex
remove color edits and fix formatting of tables in the appendix
a pass through related work and remaining color edits

--- 2020-09-04 journal paper/content/sec-model.tex
@@
 %Third, all results carry over \asmargincomment{check}  if each agent $t$ has a prior on her arrival time $t$, rather than know it exactly.
-Third, we can handle a limited amount of non-stationarity in $\respF$. \asmargincomment{check} Finally, the main result on \HardMax extends to time-discounted utilities.
+Third, we can handle a limited amount of non-stationarity in $\respF$ for the \HardMaxRandom and \SoftMaxRandom decision rules.
+\asmargincomment{check} Finally, the main result on \HardMax extends to time-discounted utilities.
 }
@@
 
-\asedit{Reflecting reality, we posit that principals' exploration strategies are ``non-strategic" in nature.
+Reflecting reality, we posit that principals' exploration strategies are ``non-strategic" in nature.
 Even though the principals play a multi-step game, they do not react to each other's moves or to the agents' strategic choices.
@@
 Responding to the competition represents another layer of complexity which has not been previously studied in this context, to the best of our knowledge, let alone made even remotely practical.
-Besides, the competitor's exploration strategy is typically not public, and understanding its exploration behavior via observations appears challenging even as a  research problem.}%
-\footnote{\asedit{Principals could potentially react to the market share or (the difference in) reputation scores.
+Besides, the competitor's exploration strategy is typically not public, and understanding its exploration behavior via observations appears challenging even as a  research problem.%
+\footnote{Principals could potentially react to the market share or (the difference in) reputation scores.
 However, baking these signals into one's exploration strategy runs the risk of over-interpreting our competition model, as they may change for exogenous reasons.
-Alternatively, one could use such signals, as well as the intuitions coming from this paper, to guide the platform's decisions regarding exploration.}}
+Alternatively, one could use such signals, as well as the intuitions coming from this paper, to guide the platform's decisions regarding exploration.}
 
@@
 
-\asedit{On the machine learning side, our model captures big, qualitative differences between bandit algorithms, building on the well-established intuition in the literature.
+On the machine learning side, our model captures big, qualitative differences between bandit algorithms, building on the well-established intuition in the literature.
 Comparisons between bandit algorithms are generally somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
@@
 In Appendix~\ref{app:bg}, we present sufficient background on this model, accessible to non-specialists.
-However, we are less interested in state-of-art algorithms for realistic applications.}
+However, we are less interested in state-of-art algorithms for realistic applications.
 
-\asedit{Bandit rewards are not discounted with time, reflecting the fact that non-discounted, regret-minimizing formulations has been prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
-Also, the distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.}
+Bandit rewards are not discounted with time, reflecting the fact that non-discounted, regret-minimizing formulations has been prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
+Also, the distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.

### 2020-09-10 journal paper/content/sec-theory.tex
a pass through related work and remaining color edits

--- 2020-09-04 journal paper/content/sec-theory.tex
@@
 
+\xhdr{Beliefs.} Rather than knowing the principals' algorithms $(\alg[1],\alg[2])$, the Bayesian prior $\priorMu$, and the response function $\respF$, agents could have beliefs on these objects that need not be correct.
+If agents have common ``point beliefs", respectively, algorithms $(\alg[1]',\alg[2]')$ prior $\priorMu'$ and response function $\respF'$,  then all our results carry over with respect to these beliefs.
+
+\xhdr{Limited non-stationarity in $\respF$.} 
+Different agents can have different response functions, in the following sense.
+For \HardMaxRandom, our results carry over if each agent $t$ has a \HardMaxRandom response function $\respF$ with parameter $\eps_t\geq \eps_0$.
+For \SoftMaxRandom, different agents can have different response functions that satisfy Definition~\ref{def:SoftMax} (with the same parameters).
+
 \xhdr{MAB extensions.} Our results carry over, with little or no modification of the proofs, to much more general versions of MAB, as long as it satisfies the i.i.d.

### 2020-09-10 journal paper/content/related_work.tex
Related work: removed the coloring, and restored one sentence.

@@
 The earlier work by \citet{bergemann1997market,bergemann2000experimentation} studies competition on both prices and quality, and allows for the data externality mentioned above.
-Specifically, the ``entrant" firm offers a new product of unknown, \gaedit{exogenously determined}, quality, which is ``explored" if and only if the entrant attracts customers.
+Specifically, the ``entrant" firm offers a new product of unknown, exogenously determined, quality, which is ``explored" if and only if the entrant attracts customers.
 However, because the entrant can only control the prices but not the product itself, the nature of exploration is fundamentally different.
@@
 Second, the entrant can only affect the \emph{quantity} of the said information, but not \emph{which} information is being revealed.
-\gadelete{Third, revealing more information is not necessarily better for the entrant (if the product's quality is actually low).} \gaedit{Third}, the social planner's exploration problem is very different from the firms' \gaedit{so that when one compares competition to the planner, one can only compare outcomes, not the algorithms.}\footnote{The planner faces a two-armed bandit problem, rather than a pricing problem \gaedit{because the planner can directly choose a firm for each agent.}}
+Third, revealing more information is not necessarily better for the entrant (if the product's quality is actually low).
+Fourth, the social planner's exploration problem is very different from the firms', so that when one compares competition to the planner, one can only compare outcomes, but not the algorithms.\footnote{The planner faces a two-armed bandit problem, rather than a pricing problem, because the planner can directly choose a firm for each agent.}
 % firm, because payments cancel out in social welfare).}
@@
 A long line of work from electrical engineering and computer science, starting from
-\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, \gaedit{focuses on competition for resources, not competition for consumers.
-Specifically, this literature targets an application} to \emph{cognitive radios}, where multiple radios transmit simultaneously in a shared medium and compete for bandwidth.
+\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, focuses on competition for resources, not competition for consumers.
+Specifically, this literature targets an application to \emph{cognitive radios}, where multiple radios transmit simultaneously in a shared medium and compete for bandwidth.
 Each radio chooses channels over time using a multi-armed bandit algorithm.

### 2020-09-16 journal paper/ec19paper/content/inverted_u.tex
fix formatting issues in experiments section

@@
 Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship as in Figure~\ref{fig:inverted-U}.
+As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship.
 More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DynamicGreedy, \DynamicEpsGreedy and \Thompson.
@@
 \end{tabular}
-\captionsetup{labelformat=empty}
 \caption{\footnotesize Market share of row player (entrant), 200 round head-start, Heavy-Tail Instance}

### 2020-09-16 journal paper/ec19paper/content/non_greedy_choice.tex
Minor edit.
fix formatting issues in experiments section

@@
 
-\subsection{Non-deterministic choice Model (HardMaxRandom)}\label{sec:non_greedy}
+\subsection{Non-deterministic choice Model (\HardMaxRandom)}\label{sec:non_greedy}
 
@@
 If this quantity is large enough, we expect better algorithms to reach their long-term performance and prevail in competition.
-We confirm this intuition; we also find that this effect is negligible for smaller (but still relevant) values of $\eps$ or $T$.
+We confirm this intuition; we also find that this effect is negligible for smaller (but relevant) values of $\eps$ or $T$.
 

### 2020-09-16 journal paper/content/sec-model.tex

@@
     for each agent $t$ and principal $i$.%
-\footnote{\asedit{Formally, agents do the Bayesian update knowing $t$, the principals' algorithms, the prior $\priorMu$, and $\respF$.}}
+\footnote{Formally, agents do the Bayesian update knowing $t$, the principals' algorithms, the prior $\priorMu$, and $\respF$.}
 
@@
 
-\xhdr{Extensions.} \asedit{The \TheoryModel admits several extensions, detailed in Section~\ref{sec:theory-extensions}.
+\xhdr{Extensions.} The \TheoryModel admits several extensions, detailed in Section~\ref{sec:theory-extensions}.
 First, all/most results extend to arbitrary reward distributions, allow reward-dependent utility, and carry over to a more general version of multi-armed bandits.
@@
 Third, we can handle a limited amount of non-stationarity in $\respF$ for the \HardMaxRandom and \SoftMaxRandom decision rules.
-\asmargincomment{check} Finally, the main result on \HardMax extends to time-discounted utilities.
-}
+Finally, the main result on \HardMax extends to time-discounted utilities.
 

### 2020-09-16 journal paper/content/sec-theory.tex

@@
 \begin{corollary}\label{cor:DG-dominance}
-The competition game has a unique Nash equilibrium: both principals choose \DynGreedy.
-\asedit{This holds even if the principals can only choose from restricted strategy sets which include \DynGreedy.}
+The competition game has a unique Nash equilibrium: both principals choose \DynGreedy. This holds even if the principals can only choose from restricted strategy sets which include \DynGreedy.
 \end{corollary}
@@
 
-\ascomment{Steven: can you come up with a short proof sketch / intuition why this thm works?}
+%\ascomment{Steven: can you come up with a short proof sketch / intuition why this thm works?}
 

### 2020-09-22 journal paper/content/sec-intro.tex
Lots of edits to address Yishay's comments (all except limited rationality).

@@
 
-In this paper, we study the interplay between \exploration and \competition.
-Platforms that engage in exploration typically need to compete against one another; most importantly, they compete for users.
-This creates a tension:
+In this paper, we study the interplay between \exploration and \competition.%
+\footnote{\asedit{\Ie we add \competition to the standard exploration-exploitation tradeoff studied in multi-armed bandits.}}
+Platforms that engage in exploration typically need to compete against one another; most importantly, they compete for users. This creates a tension:
 %between \exploration and \competition.
@@
 %Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
-We draw on the machine learning literature to compare bandit algorithms to one another in the absence of competition, so that we can talk about \emph{better} bandit algorithms in a principled way.
+We draw on the machine learning literature to compare bandit algorithms to one another \asedit{in isolation},%
+\footnote{\asedit{As opposed to ``in competition", as solutions to a stand-alone exploration problem without competition.}}
+ so that we can talk about \emph{better} bandit algorithms in a principled way.
 %\footnote{Such comparisons are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
@@
 }
-One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in the absence of competition, they are known to be terrible for a wide variety of problem instances.
+One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); \asedit{in isolation}, they are known to be terrible for a wide variety of problem instances.
 
@@
 %We endow agents with Bayesian rationality, a common modeling approach for a theoretical investigation.
-We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms, but do not receive any information about the previous agents.
+We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms, but do not receive any information about the previous agents' \asedit{choices or rewards}.
 Each agent computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
@@
  As predicted by the theory, better algorithms prevail under \HardMaxRandom if the expected number of ``random" users is sufficiently large.
-However, this effect is negligible for smaller parameter values.
+\asdelete{However, this effect is negligible for smaller parameter values.}
 

### 2020-09-22 journal paper/content/sec-model.tex
Lots of edits to address Yishay's comments (all except limited rationality).

@@
 
-\item In the \emph{\ExptsModel}, agents' reward estimate for a given principal is the average reward of the last $M$ agents that chose this principal, interpreted as a \emph{reputation score}.
+\item In the \emph{\ExptsModel}, agents' reward estimate for a given principal is the average reward of the last $M$ agents that chose this principal, interpreted as a \asedit{current} \emph{reputation score}.
 To make it meaningful initially, each principal enjoys a ``warm start": additional $T_0$ agents arrive before the game starts, and interact with the principal as described above.
@@
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-Thus, choosing a pure strategy in this game corresponds to choosing an MAB algorithm.
-Principal's utility is defined as the market share, \ie the number of agents that chose this principal.
-Principals are risk-neutral, in the sense that they optimize their expected utility.
+\asedit{Pure strategies in this game correspond to MAB algorithms.}%
+\footnote{\asedit{Note that a mixed strategy is simply a randomized MAB algorithm, \ie it is also a pure strategy.}}
+ Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral: they optimize their expected utility.
 

### 2020-09-22 journal paper/content/sec-theory.tex
Lots of edits to address Yishay's comments (all except limited rationality).

@@
 The second algorithm, called \StaticGreedy, chooses an arm $a$ with the largest prior mean reward $\E[\mu_a]$, and uses this arm in all rounds.
-In both algorithms, ties are broken arbitrarily.
+\asdelete{In both algorithms, ties are broken arbitrarily.}
 
@@
 \begin{theorem}\label{thm:DG-dominance}
-Assume \HardMax response function with fair tie-breaking.
+Assume \HardMax response function with fair tie-breaking: \asedit{$\respF(0)=\nicefrac{1}{2}$}.
 Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$.
@@
 
-The remainder of the proof of Theorem~\ref{thm:DG-dominance} uses the conclusion of one lemma to derive the precondition for another, \ie goes from $\rew_1(n_0)>\rew_2(n_0)$ to $\PMR_1(n_0)>\PMR_2(n_0)$.
+The remainder of the proof of Theorem~\ref{thm:DG-dominance} uses the conclusion of Lemma~\ref{lm:DG-rew} to derive the precondition for Lemma~\ref{lm:DG-sudden}, \ie goes from $\rew_1(n_0)>\rew_2(n_0)$ to $\PMR_1(n_0)>\PMR_2(n_0)$.
 The subtlety one needs to deal with is that the principal's ``local" round corresponding to a given ``global" round is a random quantity due to the random tie-breaking.
@@
 \xhdr{Biased tie-breaking.}
-The \HardMax model is very sensitive to the tie-breaking rule.
+The \HardMax model is very sensitive to tie-breaking \asedit{between the principals}.
 For starters, if ties are  broken deterministically in favor of principal $1$, then principal 1 can get all agents no matter what the other principal does, simply by using \StaticGreedy.
@@
 While it is usually more efficient to consider all available data, this modification simplifies analysis.}
+This completes the specification of the modified algorithm; note that it is not merely
+%another pure strategy in the competition game between the two principals, not 
+a mixed strategy that randomizes between \alg and the greedy algorithm.
 
@@
 %Parameter $p>0$ is the same for all steps.
-
-Note that the modified algorithm is another pure strategy in the competition game between the two principals, not a mixed strategy that randomizes between \alg and the greedy algorithm.
 
@@
 
-\begin{proof}[Proof Sketch]
-We follow the steps in the proof of Theorem~\ref{thm:random-clean} to derive \begin{align*}
+%\begin{proof}[Proof Sketch]
+To prove this theorem, we follow the steps in the proof of Theorem~\ref{thm:random-clean} to derive \begin{align*}
 \PMR_1(t) - \PMR_2(t)
@@
 This is at least $\BIR_2(t)/4$ by \eqref{eq:random-assn}. Then \eqref{eq:thm:SoftMax-weak} follows by the smoothness condition \eqref{eq:SoftMax-smooth}.
-\end{proof}
+%\end{proof}
 

### 2020-09-22 journal paper/ec19paper/content/inverted_u.tex
Lots of edits to address Yishay's comments (all except limited rationality).

@@
 Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship.
+As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship \asedit{as in Section~\ref{sec:theory-welfare}}.
 More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DynamicGreedy, \DynamicEpsGreedy and \Thompson.

### 2020-09-22 journal paper/ec19paper/content/revisited.tex
Lots of edits to address Yishay's comments (all except limited rationality).

@@
 We find that this also holds for $K = 3$ arms and longer time horizons, see the supplement for more details.
-We conclude:
+We conclude \asedit{that Conjecture~\ref{conj:mean-trajectories} is false:}
 
@@
 \textit{
-Conjecture~\ref{conj:mean-trajectories} is false: mean reputation trajectories do not suffice to explain the outcomes under competition.}
+Mean reputation trajectories do not suffice to explain the outcomes under competition.}
 \end{finding}

### 2020-09-22 journal paper/content/sec-model.tex
Lots of edits re equilibria (mixed vs pure, finite vs infinite, etc.)

@@
 \asedit{Pure strategies in this game correspond to MAB algorithms.}%
-\footnote{\asedit{Note that a mixed strategy is simply a randomized MAB algorithm, \ie it is also a pure strategy.}}
+\footnote{\asedit{Note that a mixed strategy is also a (randomized) MAB algorithm, \ie also a pure strategy.}}
  Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral: they optimize their expected utility.

### 2020-09-22 journal paper/content/sec-theory.tex
Lots of edits re equilibria (mixed vs pure, finite vs infinite, etc.)

@@
 We show that principals are not incentivized to \emph{explore}--- \ie to deviate from \DynGreedy.
-The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates.
-
-To make this more precise, let us define what it means for MAB algorithms to deviate from each other.
+The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates, in the sense defined below.
 
@@
 
+\asedit{\DynGreedy is a weakly dominant strategy in the competition game, and a unique Nash equilibrium.
+This is because \DynGreedy receives, in expectation, at least half of the agents before global round $n_0$, and all agents after that; both are the best possible against \alg[2].
+Moreover, \DynGreedy guarantees at least $T/2$ agents in expectation, and any other strategy can receive strictly less than that (\eg if the opponent chooses \DynGreedy).}
+
 \begin{corollary}\label{cor:DG-dominance}
-The competition game has a unique Nash equilibrium: both principals choose \DynGreedy. This holds even if the principals can only choose from restricted strategy sets which include \DynGreedy.
+\asedit{\DynGreedy is a weakly dominant strategy in the competition game.} The game has a unique Nash equilibrium: both principals choose \DynGreedy.
 \end{corollary}
 
-The proof of Theorem~\ref{thm:DG-dominance} relies on two key lemmas: that deviating from \DynGreedy implies a strictly smaller Bayesian-expected reward, and that \HardMax implies a ``sudden-death" property: if one agent chooses principal $1$ with certainty, then so do all subsequent agents.
+The proof of Theorem~\ref{thm:DG-dominance} relies on two key lemmas: that deviating from \DynGreedy implies a strictly smaller Bayesian-expected reward, and that \HardMax implies a ``sudden-death" property: if one agent chooses principal $1$ with certainty, so do all subsequent agents.
 We re-use both lemmas in later results, so we state them in sufficient generality.
-In particular, Lemma~\ref{lm:DG-rew} works for any response function because it only considers the performance of each algorithm without competition.
+
+%In particular, Lemma~\ref{lm:DG-rew} works for any response function, as it only considers the stand-alone performance of each algorithm.
 
 \begin{lemma}\label{lm:DG-rew}
-Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$. Then $\rew_1(n_0)>\rew_2(n_0)$. This holds for any response function $\respF$.
+Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$.
+Then $\rew_1(n_0)>\rew_2(n_0)$.
+The lemma holds for any response function $\respF$ (as it only considers the stand-alone performance of each algorithm).
 \end{lemma}
@@
 \begin{theorem}\label{thm:random-clean}
-Assume \HardMaxRandom response function.
+Fix a \HardMaxRandom response function $\respF$.
 Suppose algorithms \alg[1], \alg[2] are \bmonotone and $T$-oblivious, and \eqref{eq:random-assn} holds.
-If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible).
+If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible selection probability for this response function).
 \end{theorem}
@@
 We'd like to use Theorem~\ref{thm:random-clean} to conclude that a (much) better algorithm prevails in equilibrium.
-Indeed, this holds if the principals can only choose from a finite set of MAB algorithms.
+\asedit{We consider a version of the competition game in which the principals are restricted to choosing from a given set of MAB algorithms; the algorithms in this set are called \emph{feasible}.}
 
-\begin{definition}\label{def:restricted-competition}
-A \emph{\FiniteGame} is the competition game between the two principals in which they can only choose from a finite set of MAB algorithms (called \emph{feasible}).
-All feasible algorithms are assumed to be
-\bmonotone and $T$-oblivious, and to satisfy \eqref{eq:random-assn}.
-\end{definition}
+%Indeed, this holds if the principals can only choose from a finite set of MAB algorithms.
+%\begin{definition}\label{def:restricted-competition}
+%A \emph{\FiniteGame} is the competition game between the two principals in which they can only choose from a finite set of MAB algorithms (called \emph{feasible}).
+All feasible algorithms are assumed to be \bmonotone and $T$-oblivious, and to satisfy \eqref{eq:random-assn}.
+%\end{definition}
 
 \begin{corollary}\label{cor:random}
-Assume \HardMaxRandom response function.
-Consider the \FiniteGame with a feasible algorithm \alg which BIR-dominates all other feasible algorithms.
-Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
+Fix a \HardMaxRandom response function $\respF$.
+\asedit{Consider the competition game in which all feasible MAB algorithms are $T$-oblivious, \bmonotone, and satisfy \eqref{eq:random-assn} for some fixed $m_0$.
+Suppose some feasible algorithm \alg  BIR-dominates all other feasible algorithms, starting from some local step $n_0$.} Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
 \end{corollary}
@@
 
-We find that the greedy modification prevails in competition if $p$ is small enough.
+We find that the greedy modification prevails in competition if $p$ is small enough. We focus on \emph{symmetric} response functions: ones that satisfy 
+$\respF(x)+\respF(-x)=1$ for any $x\in[0,1]$.
 
@@
 
-Let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2].
+First, let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2].
 We need to assume that \alg[1] is (much) better than the greedy step \alggr if the latter is computed on a fraction of the data.
@@
 \begin{corollary}\label{cor:random-greedy}
-Posit a symmetric \HardMaxRandom response function.
-Suppose both principals can only choose from \bmonotone MAB algorithms.
-Then for any time horizon $T$, the only possible \emph{pure} Nash equilibrium is one where both principals choose \DynGreedy.
-Moreover, no pure Nash equilibrium exists when some algorithm BIR-dominates \DynGreedy and $T$ is sufficiently large.
-These conclusions hold even if each principal's strategy set is a subset of \bmonotone MAB algorithms that is closed under the greedy modification.%
-\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) \bmonotone MAB algorithms, and contains the greedy modification for each algorithms in this set and each mixing parameter $p$.}
+Fix a symmetric \HardMaxRandom response function $\respF$. \asedit{Consider the competition game in which algorithms are feasible if and only if they are \bmonotone.}%
+\footnote{\asedit{More generally, this corollary holds if feasible algorithms are \bmonotone and closed under the greedy modification (\ie the greedy modification of any feasible algorithm is also feasible, for any mixing parameter $p$).}}
+\asedit{Then there are no Nash equilibria other than (\DynGreedy,\,\DynGreedy).}
+%The latter is not a Nash equilibrium when some algorithm BIR-dominates \DynGreedy and time horizon $T$ is sufficiently large.
 \end{corollary}
+
+\asedit{Recall that by Theorem~\ref{thm:random-clean} \DynGreedy cannot be played in equilibrium when it is BIR-dominated by some feasible, $T$-oblivious algorithm and the time horizon $T$ is large enough.}
+
+%These conclusions hold even if each principal's strategy set is a subset of \bmonotone MAB algorithms that is closed under the greedy modification.%
+%\footnote{That is, the strategy set of each principal consists of some (but not necessarily all) \bmonotone MAB algorithms, and contains the greedy modification for each algorithms in this set and each mixing parameter $p$.}
 
@@
 
-As in the previous subsection, we can state a game-theoretic corollary for the \FiniteGame (as in Definition~\ref{def:restricted-competition}).
-We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity.
+\asedit{It follows that a weakly-BIR-dominating algorithm prevails in equilibrium.} We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity.
 \BReg is a standard notion from the literature (along with \BIR):
@@
 \begin{corollary}\label{cor:SoftMax-strong}
-Assume \SoftMaxRandom response function. Consider the \FiniteGame in which all feasible algorithms satisfy $\BReg(n)\to \infty$.%
+Assume \SoftMaxRandom response function. Consider the competition game in which all feasible algorithms are \bmonotone, $T$-oblivious, and satisfy $\BReg(n)\to \infty$.%
 \footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition, see Appendix~\ref{app:bg} for background.}
-Suppose a feasible algorithm \alg BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.
+Suppose some feasible algorithm \alg weakly-BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.
 \end{corollary}

### 2020-09-22 journal paper/content/conclusion.tex
A bunch of edits, mainly to remove references to "rationality".

@@
 In both variants, firms are incentivized to adopt a ``greedy algorithm" which does no purposeful exploration and leads to welfare losses for users.
-We then consider two relaxations of competition: we relax users' rationality and give one of the firms a first-mover advantage.
+We then consider two relaxations of competition: we soften users' decision rule and give one of the firms a first-mover advantage.
 Both relaxations induce firms to adopt ``better" bandit algorithms, which benefits user welfare.
@@
 \item \emph{Continuous learning.}
-While we focus on a stationary world, another well-motivated regime is continuous learning, when exploration counteracts change
-\citep{DynamicMAB-colt08}.
-The economic story would be about competition between relatively mature firms.
+While we focus on a stationary world, another well-motivated regime is continuous learning, when exploration continuously counteracts change.
+The economic story would be about competition between relatively mature firms.%
+\footnote{One difficulty is that the ``bandit model" becomes considerably more complicated: there are many reasonable ways to deal with a continuously changing world, starting from \citet{DynamicMAB-colt08}, and the distinctions between better and worse algorithms are not as clear and well-established.}
 
-\item \emph{Real-life data.}
-Numerical experiments based on real-life datasets would arguably be more realistic.
+%\item \emph{Real-life data.}
+%Numerical experiments based on real-life datasets would arguably be more realistic.
 To deal with real-life datasets, bandit algorithms would probably need to accommodate change over time and \emph{contexts} (auxiliary signals available before each round).
@@
 
-\noindent One difficulty with the last two directions is that the bandit ``side" of the model becomes considerably more complicated, both in terms of the rewards and in terms of the algorithms.
+%\noindent One difficulty with the last two directions is that the bandit ``side" of the model becomes considerably more complicated, both in terms of the rewards and in terms of the algorithms.
 Indeed, there are many reasonable ways to model a changing world (resp., dependence on  contexts), and several substantially different algorithmic approaches to deal with these models.

### 2020-09-22 journal paper/content/related_work.tex
Lots of edits to address Yishay's comments (all except limited rationality).
A bunch of edits, mainly to remove references to "rationality".

@@
 
-\xhdr{Competition in economics.} The competition vs.
+\xhdr{Competition.} The competition vs.
 innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{aghion2005competition,Vives-08}.
@@
 
-We use first-mover advantage and relaxed versions of rationality to model varying competition, instead of classic ``market competitiveness" measures such as the Lerner Index or the Herfindahl-Hirschman Index
+We use first-mover advantage and \asedit{agents' decision rule} 
+%relaxed versions of rationality 
+to model varying competition, instead of classic ``market competitiveness" measures such as the Lerner Index or the Herfindahl-Hirschman Index
 \citep{tirole1988theory}.
@@
 
-\xhdr{Relaxed rationality.}
-Relaxed versions of rationality similar to ours are found in several notable lines of work.
-For example, ``random agents" (a.k.a.
+\asedit{\xhdr{Choice models.}
+Relaxed choice models} similar to ours are found in several notable lines of work.
+First, ``random agents" (a.k.a.
 noise traders) can side-step the ``no-trade theorem'' \citep{Milgrom-Stokey-82}, a famous impossibility result in financial economics.
-The \SoftMaxRandom model is closely related to the literature on \emph{product differentiation}, starting from \cite{Hotelling-29}, see \cite{Perloff-Salop-85} for a notable later paper.
-There is a large literature on non-existence of equilibria due to small deviations   (which is related to the corresponding result for \HardMaxRandom), starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets.
-Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom.
-While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.
+Second, the \SoftMaxRandom model is closely related to the literature on \emph{product differentiation}, starting from \cite{Hotelling-29}, see \cite{Perloff-Salop-85} for a notable later paper.
+Third, a large literature on non-existence of equilibria due to small deviations (starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets) is related to the corresponding non-existence result for \HardMaxRandom.
+Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom in this context.
+\asdelete{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
 

### 2020-09-22 journal paper/content/sec-intro.tex
A bunch of edits, mainly to remove references to "rationality".

@@
 %We focus on qualitative, asymptotic differences in the algorithms' performance.
-Our results depend crucially on the agents' decision rule:
+Our results depend crucially on agents' decision rule:
 %\begin{itemize}
@@
 
-\textbf{(iii)} We further relax the decision rule so that the selection probabilities vary smoothly in terms of principal's Bayesian-expected rewards;
+\textbf{(iii)} We further soften the decision rule so that the selection probabilities vary smoothly in terms of principal's Bayesian-expected rewards;
 %as a function of the difference between  principals' Bayesian-expected rewards;
-we call it \SoftMaxRandom. For this model, better algorithm prevails under much weaker assumptions.
+we call it \SoftMaxRandom.%
+\footnote{\asedit{In particular, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}}
+ Then a better algorithm prevails under much weaker assumptions.
 % on what constitutes a better algorithm.
-(This is the most technical result of the paper.) The competition is more relaxed: the better algorithm receives only a small advantage in the long run.
+This is the most technical result of the paper.
+\asdelete{The competition is more relaxed: the better algorithm receives only a small advantage in the long run.}
 
@@
 %, even if the algorithms themselves are well-known in the research literature
-and control the severity of the competition by varying the agents' decision rule from \HardMax (cut-throat competition) to \HardMaxRandom to \SoftMaxRandom and all the way to the uniform selection.%
+and control the severity of the competition by varying the agents' decision rule from \HardMax (cut-throat competition) to \HardMaxRandom to \SoftMaxRandom and all the way to the uniform selection. 
 %\footnote{Agents' decision rule also controls the agents' rationality.

### 2020-09-22 journal paper/content/sec-model.tex
A bunch of edits, mainly to remove references to "rationality".

@@
 We assume that $\respF$ is monotonically non-decreasing, is larger than $\nicefrac12$ on the interval $(0,1]$, and smaller than $\nicefrac12$ on the interval $[-1,0)$.
-We consider three variants for $\respF$ (see \reffig{fig:response-functions}):
+We consider three variants for $\respF$, depicted in  \reffig{fig:response-functions}:
 
@@
 
-\item \SoftMaxRandom: $\respF$ lies in $[\eps_0,1-\eps_0]$ and is ``smooth" around $0$ (see Section~\ref{sec:theory-SoftMax}).
+\item \SoftMaxRandom: $\respF$ lies in $[\eps_0,1-\eps_0]$, breaks ties fairly, and has a bounded derivative around $0$ (see Definition~\ref{def:SoftMax} for a formal definition).
 \end{itemize}
+
+\asedit{\HardMaxRandom and \SoftMaxRandom can be realized as mixtures over more ``basic" agent types. Indeed,
+\HardMaxRandom is a mixture of \HardMax and ``random" agents (which choose a principal uniformly at random).
+One can obtain a \SoftMaxRandom response function using agent types that choose a principal with a largest $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
+Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.}
+
 % \asedit{(We assume $\respF(-1)+\respF(1)=1$ in \HardMaxRandom and \SoftMaxRandom only to simplify notation.)}
@@
 \item In the \emph{\TheoryModel}, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$.
-    Agents' reward estimates are defined as posterior mean rewards:
+\asedit{Each agent knows its global round $t$, but not the performance signals such as the current market shares.} Agents' reward estimates are defined as posterior mean rewards:
         $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i}$

### 2020-09-22 journal paper/content/sec-theory.tex
A bunch of edits, mainly to remove references to "rationality".

@@
 
-\subsection{Full rationality: \HardMax}
+\subsection{\HardMax response function}
 \label{sec:theory-HM}
 
-We posit fully rational agents, in the sense that their response function is \HardMax.
-We show that principals are not incentivized to \emph{explore}--- \ie to deviate from \DynGreedy.
-The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates, in the sense defined below.
+%We posit fully rational agents, in the sense that their response function is \HardMax.
+
+We consider agents with \HardMax response function, and show that principals are not incentivized to \emph{explore}--- \ie to deviate from \DynGreedy.
+The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates.
+
+To make this formal, let's define what it means for MAB algorithms to deviate.
 
@@
 %%%%%%%%%%
-\subsection{Relaxed rationality: \HardMaxRandom}
+\subsection{\HardMax with random agents}
 \label{sec:theory-HMR}
 
-Consider the \HardMaxRandom response model, where each principal is chosen with some positive baseline probability. Informally, we find that
+Consider the \HardMaxRandom response model, \ie \HardMax mixed with ``random agents".
+Informally, we find that
 %\begin{align}\label{eq:theory-HMR-informal}
@@
 In particular, their \BIR does not depend on the time horizon $T$ of the game.
-Then this property can be formalized as:
+Then this property can be formalized as follows:
 \begin{align}\label{eq:random-better-clean}
@@
 
-{We frame our contributions in terms of the relationship between \competitiveness and \rationality on one side, and adoption of better algorithms on the other.
-Recall that both \competitiveness (of the game between the two principals) and \rationality (of the agents) are controlled by the response function $\respF$.}
+We frame our contributions in terms of the relationship between \competitiveness \asedit{(as expressed by the ``hardness" of the agents' response function $\respF$)},
+ %and \rationality on one side, 
+ and adoption of better algorithms. 
+\asdelete{Recall that both \competitiveness (of the game between the two principals) and \rationality (of the agents) are controlled by the response function $\respF$.}
 
@@
 Our main story concerns the \FiniteGame between the two principals where one allowed algorithm \alg is ``better" than the others.
-{We track whether and when \alg is chosen in an equilibrium.} We vary \competitiveness/\rationality by changing the response function from \HardMax (full rationality, very competitive environment) to \HardMaxRandom to  \SoftMaxRandom (less rationality and competition).
+{We track whether and when \alg is chosen in an equilibrium.} We vary \competitiveness by changing the response function from \HardMax (very competitive environment) to \HardMaxRandom to  \SoftMaxRandom (less competition).
 Our conclusions are as follows:
@@
 \item Under \SoftMaxRandom, more innovation: \alg is chosen as long as it weakly-BIR-dominates.%
-\footnote{This is a weaker condition, the better algorithm is chosen in a broader range of scenarios.}
 \end{OneLiners}
-These conclusions follow, respectively, from Corollaries~\ref{cor:DG-dominance}, \ref{cor:random} and \ref{cor:SoftMax-strong}.
-Further, {we consider the uniform choice between the principals.
-It corresponds to the least amount of rationality and competition, and (when principals' utility is the number of agents) uniform choice provides no incentives to innovate.}%
+These conclusions follow from Corollaries~\ref{cor:DG-dominance}, \ref{cor:random} and \ref{cor:SoftMax-strong}, respectively.
+Recall that weak-BIR-dominance is a weaker condition, so that a better algorithm is chosen in a broader range of scenarios.
+Further, {we consider the uniform choice between the principals, which corresponds to the least amount of competition and (when principals' utility is the number of agents) uniform choice provides no incentives to innovate.}%
 \footnote{On the other hand, if principals' utility is aligned with agents' welfare, then a monopolist principal is incentivized to choose the best possible MAB algorithm (namely, to minimize cumulative Bayesian regret $\BReg(T)$).
@@
       \draw[->] (-.5,0) -- (9.5,0) node[above]
-        {\qquad\qquad Competitiveness/Rationality};
+        {\qquad\qquad Competitiveness};
       \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithm in equilibrium};
@@
      \node[below] at (3.9, 0) {\footnotesize \SoftMaxRandom};
-     \node[below] at (6, 0) {\footnotesize \HardMaxRandom};
+     \node[below] at (6, -.5) {\footnotesize \HardMaxRandom};
      \node[below] at (8, 0) {\footnotesize \HardMax};
@@
 Let us zoom in on the symmetric  \HardMaxRandom model.
-{Competitiveness and rationality within this model are controlled by the baseline probability $\eps_0 = \respF(- 1)$, which goes smoothly between the two extremes of \HardMax ($\eps_0=0$) and the uniform choice ($\eps_0=\tfrac12$).
-Smaller $\eps_0$ corresponds to increased rationality and increased competitiveness.} For clarity, we assume that principal's utility is the number of agents.
+Competitiveness within this model are controlled by the baseline probability $\eps_0 = \respF(- 1)$, which goes smoothly between the two extremes of \HardMax ($\eps_0=0$, tough competition) and the uniform choice ($\eps_0=\tfrac12$, no competition).
+For clarity, we assume that principal's utility is the number of agents.
+
+\newcommand{\deltaU}{\Delta U}
 
@@
 Suppose initially both principals use some algorithm \alg, and principal 1 ponders switching to another algorithm \alg' which BIR-dominates \alg.
-{We are interested in the marginal utility of this switch.
-Then:}
+We are interested in the marginal utility $\deltaU$ of this switch.
+Then:
 
 \begin{OneLiners}
-\item $\eps_0 = 0$ (\HardMax):~~~~ the marginal utility can be negative if \alg is \DynGreedy.
+\item if $\eps_0 = 0$ then $\deltaU$ can be negative if \alg is \DynGreedy.
 
-\item $\eps_0$ near $0$:~~~~ only a small marginal utility can be guaranteed, as it may take a long time for $\alg'$ to ``catch up" with \alg, and hence less time to reap the benefits.
+\item if $\eps_0$ is near $0$ then only a small $\deltaU$ can be guaranteed, as it may take a long time for $\alg'$ to ``catch up" with \alg, and hence less time to reap the benefits.
 
-\item ``medium-range" $\eps_0$:~~~~ large marginal utility, as $\alg'$ learns fast and gets most agents.
+\item if $\eps_0$ is medium-range, then $\deltaU$ is large, as $\alg'$ learns fast and gets most agents.
 
-\item $\eps_0$ near $\tfrac12$:~~~~ small marginal utility, as principal 1 gets most agents for free no matter what.
+\item if $\eps_0$ is near $\tfrac12$, then $\deltaU$ is small, as principal 1 gets most agents for free no matter what.
 \end{OneLiners}
-The familiar inverted-U shape is depicted in Figure~\ref{fig:inverted-U3}.
+These findings can also be organized as an inverted-U relationship, see Figure~\ref{fig:inverted-U3}.
 

### 2020-09-23 journal paper/content/sec-intro.tex
Edits to address some of the Ian's comments.

@@
 We investigate several model variants, %within this framework,
-where we vary agents' decision rule or allow a first-mover advantage. %to one of the principals.
+where we vary agents' decision rule or allow a first-mover advantage.%
+\footnote{\asedit{In particular, we use different model variants (resp., Bayesian and frequentist) for theory and for simulations so as to ensure tractability.
+Our main findings are similar for both.}}
+
+%\asedit{For tractability, our theoretical results and our numerical simulations adopt closely related but technically different model variants (respectively, Bayesian and frequentist), with similar findings.}
+
+%to one of the principals.
 %In all variants, agents have little or no information about other agents' choices and rewards.
@@
 %We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
-The barriers for innovation arise entirely from the reputational consequences of exploration in competition, whereas the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs.
+The barriers for innovation arise entirely from the reputational consequences of exploration in competition, \asedit{without invoking monetary costs and benefits.
+Whereas the inverted-U relationships in prior work are due to the monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.}
+
+%the tradeoff between the benefits of the technology and its R\&D costs.
 

### 2020-09-23 journal paper/content/sec-model.tex
Edits to address some of the Ian's comments.

@@
 The mean rewards $\mu_a$ are fixed over time, and initially not known to anybody.
-Each principal is completely unaware of the rounds when the other principal is chosen.
+The principals can only react to the rewards in rounds when they are chosen, and are completely unaware of the rounds when the opponent is chosen.
 
@@
 local distinction when clear from the context.} each time it is called, it outputs an arm from $A$, and inputs a reward for this action.
-\alg[i] is called only in global rounds when principal $i$ is chosen.
+The algorothm is called only in global rounds when principal $i$ is chosen.
 

### 2020-09-23 journal paper/content/sec-intro.tex
A few more edits.

@@
 \footnote{\asedit{\Ie we add \competition to the standard exploration-exploitation tradeoff studied in multi-armed bandits.}}
-Platforms that engage in exploration typically need to compete against one another; most importantly, they compete for users. This creates a tension:
+Platforms that engage in exploration typically need to compete against one another. \asedit{Most importantly, platforms compete for users, who benefit them in two ways:
+generate revenue and provide data for learning.} This creates a tension:
 %between \exploration and \competition.
-while exploration may be essential for improving the service tomorrow, it may degrade quality and make users leave \emph{today}, in which case there will be less users to learn from.
+while exploration may be essential for improving the service tomorrow, it may degrade quality \emph{today}, in which case some of the users can leave and there will be less users to learn from.
 This may create a ``data feedback loop" when the platform's performance further degrades relative to competitors who keep learning and improving from \emph{their} users, and so forth.
 Taken to the extreme, such dynamics may cause a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
-Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.
+\asdelete{Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.}
 
-The main high-level question is:
+The main high-level question we ask is:
 %that we focus on in this paper is:

### 2020-09-23 journal paper/content/sec-model.tex
A few more edits.

@@
 Agents are myopic: they do not worry about how their actions impact their future utility.\footnote{So, agents do not attempt to learn over time, second-guess or game future agents, or manipulate the principals' learning algorithms.
-This is arguably typical in practice, in part because one agent's influence tends to be small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores", and further simplified to average performance over a time window.
+This is arguably typical in practice, in part because one agent's influence tends to be small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores" reflecting the current reputation, and further simplified to average performance over a sliding time window.
 

### 2020-09-23 journal paper/content/sec-model.tex
A few more edits re Ian's comments.

@@
 \footnote{We use `action' and `arm' interchangeably, as common in the literature on multi-armed bandits.}
- The agent experiences this action, receives a reward $r_t\in \{ 0,1\}$, and reports it back to the principal.
+The agent experiences this action and receives a reward $r_t\in \{ 0,1\}$ \asedit{for this action, which is then observed by the principal}.
 We posit \emph{stochastic rewards}: whenever a given action $a\in A$ is chosen, the reward is an independent draw from Bernoulli distribution with mean $\mu_a$.
 The mean rewards $\mu_a$ are fixed over time, and initially not known to anybody.
-The principals can only react to the rewards in rounds when they are chosen, and are completely unaware of the rounds when the opponent is chosen.
+The principals are completely unaware of the rounds when the opponent is chosen.
 
- Each principal $i$ commits to a learning algorithm \alg[i] before round $1$, and uses it throughout.
-The algorithm follows the protocol of \emph{multi-armed bandits} (\emph{MAB}).
-Namely, it proceeds in time-steps:%
+\asedit{Thus, each principal follows the protocol of \emph{multi-armed bandits} (\emph{MAB}): in each round when it is chosen, it chooses an arm from $A$ and observes a reward for this action (and nothing else).}
+
+Each principal $i$ commits to an MAB algorithm \alg[i] before round $1$, and uses it throughout. The algorithm proceeds in time-steps:%
 \footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before.
@@
 
-Reflecting reality, we posit that principals' exploration strategies are ``non-strategic" in nature.
+Reflecting reality, we posit that \asedit{the exploration strategies available to the principals} are ``non-strategic" in nature.
 Even though the principals play a multi-step game, they do not react to each other's moves or to the agents' strategic choices.

### 2020-09-23 journal paper/content/sec-theory.tex
A few more edits re Ian's comments.

@@
 \Pr\sbr{\mu_a  > \mu_{a'}
-\quad \forall a'\in A}
+\quad \forall a'\in A\setminus \{a\}}
 > 0.
@@
 
-An algorithm is called \emph{\bmonotone} if it can only get better over time: $\rew_i(\cdot)$ is non-decreasing, and therefore $\BIR(\cdot)$ is non-decreasing. This is a mild assumption, see
+An algorithm is called \emph{\bmonotone} if it can only get better over time, in the Bayesian sense: namely, if $\E[\rew_i(\cdot)]$ is non-decreasing, and therefore $\BIR(\cdot)$ is non-decreasing.
+This is a mild assumption, see
 Appendix~\ref{app:examples}.

### 2020-09-23 journal paper/journal_main.tex
Changed App title for App A.
a pass through related work and remaining color edits
Fixed the ToC entry for App A.
Removed some coloring, and a \newpage.
Added a \newpage.
A few more edits re Ian's comments.

@@
 
-We consider a stylized duopoly model in which two firms face the same multi-armed bandit instance.
+We consider a stylized duopoly model in which two firms face the same problem instance of multi-armed bandits.
 Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen.
@@
 
-\newpage
 \section{Numerical simulations: the \ExptsModel}
@@
 
-\newpage
+%\newpage
 \addcontentsline{toc}{section}{References}
@@
 
-\section{Background for non-specialists: \\multi-armed bandits}
+\section[Background for non-specialists: multi-armed bandits]{Background for non-specialists: \\multi-armed bandits}
 \label{app:bg}

### 2020-09-24 journal paper/content/related_work.tex
A few more edits.
edits to related work

@@
 
-\asedit{\xhdr{Choice models.}
-Relaxed choice models} similar to ours are found in several notable lines of work.
+\asedit{\xhdr{Choice models.}}
+\gaedit{Relaxed choice models similar to ours are found in several notable lines of work.
 First, ``random agents" (a.k.a.
-noise traders) can side-step the ``no-trade theorem'' \citep{Milgrom-Stokey-82}, a famous impossibility result in financial economics.
-Second, the \SoftMaxRandom model is closely related to the literature on \emph{product differentiation}, starting from \cite{Hotelling-29}, see \cite{Perloff-Salop-85} for a notable later paper.
-Third, a large literature on non-existence of equilibria due to small deviations (starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets) is related to the corresponding non-existence result for \HardMaxRandom.
-Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom in this context.
+noise traders) can side-step the ``no-trade theorem'' \citep{Milgrom-Stokey-82}, a famous impossibility result in financial economics, and play a similar role in our model as we move from \HardMax to \HardMaxRandom.
+Furthermore, a large literature on non-existence of equilibria due to small deviations (starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets) is related to the corresponding non-existence result for \HardMaxRandom.
+Second, the stochastic choice rules we consider have been widely utilized across economics.
+For instance, \SoftMaxRandom includes the standard logit choice rule, which has been widely used throughout economics with strong micro-foundations \citep{luce1959choice, mcfadden1974measurement, matvejka2015rational}, as a special case.
+The functional form for the \SoftMaxRandom model is similar to that in standard horizontal production differentiation models (e.g.
+\cite{Hotelling-29, Perloff-Salop-85}).
+\cite{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom.}
 \asdelete{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}

### 2020-09-24 journal paper/content/related_work.tex
Another revision for "choice models".

--- 2020-09-22 journal paper/content/related_work.tex
@@
 Third, revealing more information is not necessarily better for the entrant (if the product's quality is actually low).
-Fourth, the social planner's exploration problem is very different from the firms', so that when one compares competition to the planner, one can only compare outcomes, but not the algorithms.\footnote{The planner faces a two-armed bandit problem, rather than a pricing problem, because the planner can directly choose a firm for each agent.}
+Fourth, the social planner's exploration problem is very different from the firms', so that when one compares competition to the planner, one can only compare outcomes, but not the algorithms.\footnote{The planner faces a two-armed bandit problem, because the planner can directly choose a firm for each agent.}
 % firm, because payments cancel out in social welfare).}
@@
 
-We use first-mover advantage and \asedit{agents' decision rule} 
+%We use first-mover advantage and \asedit{agents' decision rule}
 %relaxed versions of rationality
-to model varying competition, instead of classic ``market competitiveness" measures such as the Lerner Index or the Herfindahl-Hirschman Index
-\citep{tirole1988theory}.
-The latter measures rely on ex-post observable attributes of a market such as prices or market shares.
-Neither is applicable to our setting, since there are no prices, and market shares are endogenous.
+%to model varying competition, instead of
+
+\asedit{Classic ``market competitiveness" measures,such as the Lerner Index or the Herfindahl-Hirschman Index
+\citep{tirole1988theory} are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).}
+
+%Neither is applicable to our setting (since there are no prices, and market shares are endogenous).
 
 \asedit{\xhdr{Choice models.}
-Relaxed choice models} similar to ours are found in several notable lines of work.
+Stochastic choice models similar to ours have been widely used throughout economics.
 First, ``random agents" (a.k.a.
 noise traders) can side-step the ``no-trade theorem'' \citep{Milgrom-Stokey-82}, a famous impossibility result in financial economics.
-Second, the \SoftMaxRandom model is closely related to the literature on \emph{product differentiation}, starting from \cite{Hotelling-29}, see \cite{Perloff-Salop-85} for a notable later paper.
-Third, a large literature on non-existence of equilibria due to small deviations (starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets) is related to the corresponding non-existence result for \HardMaxRandom.
-Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom in this context.
-\asdelete{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
+They play a similar role in our model, side-stepping the dominance of the greedy algorithm.
+%as we move from \HardMax to \HardMaxRandom.  
+Second, there a large literature on non-existence of equilibria due to small deviations, starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets.\footnote{\cite{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom in this context.}
+This is superficially similar to how small deviations towards the greedy algorithm rule out equilibria under \HardMaxRandom.
+%Second, the stochastic choice rules we consider have been widely utilized across economics. For instance, 
+Third, \SoftMaxRandom subsumes the logit choice rule, a standard behavioral model with strong empirical and microeconomic foundations 
+\citep[\eg][]{luce1959choice, mcfadden1974measurement, matvejka2015rational}.
+Fourth, choice models similar to \SoftMaxRandom are used to explain  horizontal product differentiation \citep[\eg][]{Hotelling-29, Perloff-Salop-85}.}
+%\asdelete{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
 

### 2020-09-24 journal paper/content/sec-model.tex
Added a sentence re agent models.

@@
 
-\asedit{\HardMaxRandom and \SoftMaxRandom can be realized as mixtures over more ``basic" agent types. Indeed,
-\HardMaxRandom is a mixture of \HardMax and ``random" agents (which choose a principal uniformly at random).
-One can obtain a \SoftMaxRandom response function using agent types that choose a principal with a largest $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
+\asedit{\HardMaxRandom and \SoftMaxRandom can be seen as agent types that, resp., give some chance to the other principal and avoid sharp transitions in their probabilistic response.
+Alternatively, they can be realized as mixtures over more ``basic" agent types.
+Indeed, \HardMaxRandom is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
+One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
 Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.}

### 2020-09-24 journal paper/ec19paper/content/inverted_u.tex
add EoG values for everything

@@
 \begin{adjustbox}{width=\textwidth,center}
-\begin{tabular}{|c|c|c|c||c|c|c|}
+\begin{tabular}{|c|c|c|c||c|c|c||c|c|c|}
   \hline
   & \multicolumn{3}{c||}{Heavy-Tail}
-  & \multicolumn{3}{c|}{Needle-in-Haystack}\\
+  & \multicolumn{3}{c|}{Needle-in-Haystack} 
+  & \multicolumn{3}{c|}{Uniform}\\
   \hline
+  & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500
    & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500
@@
 \TS vs \DG
-  & \makecell{\textbf{0.29} $\pm$0.03\\ \Eeog 55 (0)}
-  & \makecell{\textbf{0.72} $\pm$0.02\\ \Eeog 570 (0)}
-  & \makecell{\textbf{0.76} $\pm$0.02\\ \Eeog 620 (99)}
+  & \textbf{0.31} $\pm$0.03
+  & \textbf{0.72} $\pm$0.02
+  & \textbf{0.75} $\pm$0.02
   %%
-  & \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 200 (27)}
-  & \makecell{\textbf{0.6} $\pm$0.03\\ \Eeog 370 (0)}
-  & \makecell{\textbf{0.64} $\pm$0.03\\ \Eeog 580 (122)}
-  \\
+  & \textbf{0.68} $\pm$0.03
+  & \textbf{0.62} $\pm$0.03
+  & \textbf{0.65} $\pm$0.03 
+  %%
+  & \makecell{\textbf{0.44} $\pm$0.03} 
+ & \makecell{\textbf{0.52} $\pm$0.02} 
+ & \makecell{\textbf{0.58} $\pm$0.02} \\
 \hline
   $\TS$ vs $\DEG$  
-  & \makecell{\textbf{0.3} $\pm$0.03\\ \Eeog 37 (0)}
-  & \makecell{\textbf{0.88} $\pm$0.01\\ \Eeog 480 (0)}
-  & \makecell{\textbf{0.9} $\pm$0.01\\ \Eeog 570 (114)}
+  & \textbf{0.3} $\pm$0.03
+  & \textbf{0.89} $\pm$0.01 
+  & \textbf{0.9} $\pm$0.01
   %%
-  & \makecell{\textbf{0.57} $\pm$0.03\\ \Eeog 150 (14)}
-  & \makecell{\textbf{0.52} $\pm$0.03\\ \Eeog 460 (79)}
-  & \makecell{\textbf{0.56} $\pm$0.02\\ \Eeog 740 (628)}
-  \\
-\hline
+  & \textbf{0.6} $\pm$0.03
+  & \textbf{0.52} $\pm$0.03
+  & \textbf{0.55} $\pm$0.02 
+  %%
+ & \makecell{\textbf{0.41} $\pm$0.03} 
+ & \makecell{\textbf{0.47} $\pm$0.02} 
+ & \makecell{\textbf{0.55} $\pm$0.02} \\ \hline
   $\DG$ vs $\DEG$
-  & \makecell{\textbf{0.62} $\pm$0.03\\ \Eeog 410 (7)}
-  & \makecell{\textbf{0.6} $\pm$0.02\\ \Eeog 790 (762)}
-  & \makecell{\textbf{0.57} $\pm$0.03\\ \Eeog 730 (608)}
+  & \textbf{0.63} $\pm$0.03 
+  & \textbf{0.6} $\pm$0.02
+  & \textbf{0.56} $\pm$0.03
   %%
-  & \makecell{\textbf{0.46} $\pm$0.03\\ \Eeog 340 (129)}
-  & \makecell{\textbf{0.42} $\pm$0.02\\ \Eeog 650 (408)}
-  & \makecell{\textbf{0.42} $\pm$0.02\\ \Eeog 690 (467)}
-  \\
-   \hline
+  & \textbf{0.42} $\pm$0.03
+  & \textbf{0.41} $\pm$0.03
+  & \textbf{0.39} $\pm$0.02
+  %%
+   & \makecell{\textbf{0.5} $\pm$0.03} 
+ & \makecell{\textbf{0.46} $\pm$0.02} 
+ & \makecell{\textbf{0.45} $\pm$0.02} \\ \hline
 \end{tabular}
 \end{adjustbox}
-\caption{\footnotesize {\bf Simultaneous Entry}, for Heavy-Tail and Needle-in-Haystack instances.
+\caption{\footnotesize {\bf Simultaneous Entry, Market Share}.
 Each cell describes a game between two algorithms, call them Alg1 vs.
 Alg2, for a particular value of the warm start $T_0$.
-Line 1 in the cell is the market share of Alg 1: the average (in bold) and the 95\% confidence band.
+Each cell contains the market share of Alg 1: the average (in bold) and the 95\% confidence band.
 %For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
-Line 2 specifies the ``effective end of game" (\Eeog): the average and the median (in brackets). The time horizon is $T=2000$.}
-\label{sim_table}
+ The time horizon is $T=2000$.}
+\label{fig:market_share}
 \end{table*}
+
+\begin{table*}[t]
+\centering
+\begin{adjustbox}{width=\textwidth,center}
+\begin{tabular}{|c|c|c|c||c|c|c||c|c|c|}
+  \hline
+  & \multicolumn{3}{c||}{Heavy-Tail}
+  & \multicolumn{3}{c|}{Needle-in-Haystack} 
+  & \multicolumn{3}{c|}{Uniform}\\
+  \hline
+  & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500
+   & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500
+  & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 \\
+  \hline
+\TS vs \DG
+  & 68 (0)  & 560 (8.5)  & 610 (86.5)
+  %%
+ & 180 (30)  & 380 (0)  & 550 (6.5)
+  %%
+  &  260 (0)  
+  &  780 (676.5)  
+  &  880 (897.5) \\
+\hline
+  $\TS$ vs $\DEG$  
+ & 37 (0)  & 430 (0)  & 540 (105)
+  %%
+ & 150 (10)  & 460 (25)  & 780 (705)
+  %%
+ & 230 (0)  & 830 (772)  & 980 (1038) \\ \hline
+  $\DG$ vs $\DEG$
+ & 340 (110)  & 640 (393)  & 670 (425) 
+  %%
+  & 410 (8.5)  & 760 (666)  & 740 (646)
+  %%
+  & 530 (101)  & 990 (1058)  & 1000 (1059) \\ \hline
+\end{tabular}
+\end{adjustbox}
+\caption{\footnotesize {\bf Simultaneous Entry, \Eeog}.
+Each cell describes a game between two algorithms, call them Alg1 vs.
+Alg2, for a particular value of the warm start $T_0$.
+Each cell specifies the ``effective end of game" (\Eeog): the average and the median (in brackets).
+%For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG. 
+The time horizon is $T=2000$.}
+\label{fig:eog}
+\end{table*}
+
 \normalsize
@@
 We investigate the firms' market shares when they choose different algorithms (otherwise, by symmetry both firms get half of the agents).
-We report the market shares for Heavy-Tail and Needle-in-Haystack instances in Table~\ref{sim_table}  (see the first line in each cell), for a range of values of the warm start $T_0$.
-Table~\ref{tab:duopoly_unif} reports similarly on the Uniform instance.
+We report the market shares for \gaedit{each instance in} Table~\ref{fig:market_share}.
 We find that $\DG$ is a weakly dominant strategy for the Heavy-Tail and Uniform instances, as long as $T_0$ is sufficiently small.
@@
 By our tie-breaking rule, (\DynamicGreedy,\DynamicGreedy) is the only pure-strategy equilibrium.
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|c|c|c|c|}
- % \hline
- % & \multicolumn{3}{c|}{Uniform} \\
-\hline
-   & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 \\ \hline
-\TS vs \DG
-  & \makecell{\textbf{0.46} $\pm$0.03}
-    & \makecell{\textbf{0.52} $\pm$0.02}
-    & \makecell{\textbf{0.6} $\pm$0.02} \\ \hline
-\TS vs \DEG
-    & \makecell{\textbf{0.41} $\pm$0.03}
-    & \makecell{\textbf{0.51} $\pm$0.02}
-    & \makecell{\textbf{0.55} $\pm$0.02} \\ \hline
-\DG vs \DEG
-    & \makecell{\textbf{0.51} $\pm$0.03}
-    & \makecell{\textbf{0.48} $\pm$0.02}
-    & \makecell{\textbf{0.45} $\pm$0.02} \\\hline
-\end{tabular}
-\caption{\footnotesize {\bf Simultaneous entry}, for the Uniform MAB instance. Semantics are the same as in Table \ref{sim_table}.}
-\label{tab:duopoly_unif}
-\end{table}
 
@@
 We interpret low \Eeog as a strong evidence of the ``death spiral" effect.
-Focusing on the simultaneous entry scenario, we specify the \Eeog values in Table~\ref{sim_table} (the second line of each cell).
+Focusing on the simultaneous entry scenario, we specify the \Eeog values in Table~\ref{fig:eog} (the second line of each cell).
 We find that the \Eeog values are indeed small:

### 2020-09-24 analysis_code/eeog.R

@@ -14,8 +14,8 @@ concise_alg_rep <- function(alg) {
   }
 }
 
-WORKING_PATH <- "/Users/garidor/Desktop/CompetingBandits/results/tournament_raw_results/"
-f <- "tournament_experiment_longer_ws_raw.csv"
+WORKING_PATH <- "/Users/guyaridor/Dropbox/bandits_results/tournament_raw_results/"
+f <- "tournament_experiment_full_welfare_calc_raw.csv"
 dat <- read.csv(file=paste(WORKING_PATH, f, sep=""))
 dat <- filter(dat, Time.Horizon == 2000)
 priors <- as.list(unique(dat['Prior']))$Prior
@@ -25,9 +25,12 @@ warm_starts <- c(20, 250, 500)
 algs <- c("ThompsonSampling", "DynamicEpsilonGreedy", "DynamicGreedy")
 
 for (prior in priors) {
-  results <- matrix(nrow=3, ncol=length(warm_starts))
-  rownames(results) <- c("TS vs DG", "TS vs DEG", "DG vs DEG")
-  colnames(results) <- lapply(warm_starts, function (start) { return(paste("$T_0$ =", start))})
+  results_market_shares <- matrix(nrow=3, ncol=length(warm_starts))
+  results_eeog <- matrix(nrow=3, ncol=length(warm_starts))
+  rownames(results_market_shares) <- c("TS vs DG", "TS vs DEG", "DG vs DEG")
+  rownames(results_eeog) <- c("TS vs DG", "TS vs DEG", "DG vs DEG")
+  colnames(results_market_shares) <- lapply(warm_starts, function (start) { return(paste("$T_0$ =", start))})
+  colnames(results_eeog) <- lapply(warm_starts, function (start) { return(paste("$T_0$ =", start))})
   for (i in 1:length(warm_starts)) {
     warm_start = warm_starts[i]
     for (k in 1:length(algs)) {
@@ -41,20 +44,25 @@ for (prior in priors) {
         cin <- signif(test$conf.int[2] - mean(share), digits=1)
         eeog_mean <- signif(mean(filtered_dat$EEOG), digits=2)
         eeog_summary <- summary(filtered_dat$EEOG)
-        res_text <- paste("\\makecell{\\textbf{",signif(mean(share), digits=2), "} $\\pm$", cin, "\\\\ eeog \\\\ avg: ", eeog_mean, "\\\\ med: ", eeog_summary["Median"], "}" , sep="")
[…]

### 2020-09-24 journal paper/ec19paper/content/inverted_u.tex
change around spacing
Adjusted placement of figures and tables a bit.

@@
 
-% latex table generated in R 3.4.0 by xtable 1.8-2 package
-% Thu Aug 16 13:13:01 2018
+Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
+As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship \asedit{as in Section~\ref{sec:theory-welfare}}.
+More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DynamicGreedy, \DynamicEpsGreedy and \Thompson.
+We do this for each ``intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
+For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs, see \cite{DS-arxiv}).
+\DynamicGreedy is trivially the dominant strategy under monopoly.
+
+%(1) Permanent monopoly - indifferent between all algorithms but break indifference towards DG cause of deployment costs
+%(2) Temporary monopoly - TS is the dominant strategy for the incumbent
+%(3) Permanent duopoly -
+%   (a) Under NIH, TS is dominant
+%   (b) Under Heavy Tail, DG is weakly dominant
+%   (c) Under Uniform, DG "beats" TS, but DEG "beats" TS by more than DG and DEG vs DG leads to ~50% market share (it's slightly in favor of DG but 50/50 is within the confidence band).
+If we break indifference towards DG cause of deployment costs, then we have that the unique PSNE is (DG, DG).
+If we don't add the indifference breaking, then we have four PSNE, (DG, DEG), (DEG, DG), (DG, DG), (DEG, DEG)
+
+\xhdr{Simultaneous entry.}
+The basic scenario is when both firms are competing from round $1$. A crucial distinction is whether an MAB instance is exploration-disadvantaged:
+
+\begin{finding}\label{find:duopoly}
+\textit{Under simultaneous entry:
+\begin{itemize}
+\item[(a)] (\DynamicGreedy,\DynamicGreedy) is the unique pure-strategy Nash equilibrium for exploration-disadvantaged MAB instances with a sufficiently small ``warm start".
+\item[(b)] This is not necessarily the case for MAB instances that are not exploration-disadvantaged. In particular, \Thompson is a weakly dominant strategy for Needle-in-Haystack.
+\end{itemize}
+}
+\end{finding}
+
 \begin{table*}[t]
@@
 
-\begin{table*}[t]
-\centering
-\begin{adjustbox}{width=\textwidth,center}
-\begin{tabular}{|c|c|c|c||c|c|c||c|c|c|}
-  \hline
-  & \multicolumn{3}{c||}{Heavy-Tail}
-  & \multicolumn{3}{c|}{Needle-in-Haystack} 
-  & \multicolumn{3}{c|}{Uniform}\\
-  \hline
-  & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500
-   & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500
-  & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 \\
-  \hline
-\TS vs \DG
-  & 68 (0)  & 560 (8.5)  & 610 (86.5)
-  %%
- & 180 (30)  & 380 (0)  & 550 (6.5)
-  %%
-  &  260 (0)  
-  &  780 (676.5)  
-  &  880 (897.5) \\
-\hline
-  $\TS$ vs $\DEG$  
- & 37 (0)  & 430 (0)  & 540 (105)
-  %%
- & 150 (10)  & 460 (25)  & 780 (705)
-  %%
- & 230 (0)  & 830 (772)  & 980 (1038) \\ \hline
-  $\DG$ vs $\DEG$
- & 340 (110)  & 640 (393)  & 670 (425) 
-  %%
-  & 410 (8.5)  & 760 (666)  & 740 (646)
-  %%
-  & 530 (101)  & 990 (1058)  & 1000 (1059) \\ \hline
-\end{tabular}
-\end{adjustbox}
-\caption{\footnotesize {\bf Simultaneous Entry, \Eeog}.
-Each cell describes a game between two algorithms, call them Alg1 vs.
-Alg2, for a particular value of the warm start $T_0$.
-Each cell specifies the ``effective end of game" (\Eeog): the average and the median (in brackets).
-%For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG. 
-The time horizon is $T=2000$.}
-\label{fig:eog}
-\end{table*}
-
 \normalsize
-Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship \asedit{as in Section~\ref{sec:theory-welfare}}.
-More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DynamicGreedy, \DynamicEpsGreedy and \Thompson.
-We do this for each ``intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
-For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs, see \cite{DS-arxiv}).
-\DynamicGreedy is trivially the dominant strategy under monopoly.
-
-%(1) Permanent monopoly - indifferent between all algorithms but break indifference towards DG cause of deployment costs
-%(2) Temporary monopoly - TS is the dominant strategy for the incumbent
-%(3) Permanent duopoly -
-%   (a) Under NIH, TS is dominant
-%   (b) Under Heavy Tail, DG is weakly dominant
-%   (c) Under Uniform, DG "beats" TS, but DEG "beats" TS by more than DG and DEG vs DG leads to ~50% market share (it's slightly in favor of DG but 50/50 is within the confidence band).
-If we break indifference towards DG cause of deployment costs, then we have that the unique PSNE is (DG, DG).
-If we don't add the indifference breaking, then we have four PSNE, (DG, DEG), (DEG, DG), (DG, DG), (DEG, DEG)
-
-\xhdr{Simultaneous entry.}
-The basic scenario is when both firms are competing from round $1$. A crucial distinction is whether an MAB instance is exploration-disadvantaged:
-
-\begin{finding}\label{find:duopoly}
-\textit{Under simultaneous entry:
-\begin{itemize}
-\item[(a)] (\DynamicGreedy,\DynamicGreedy) is the unique pure-strategy Nash equilibrium for exploration-disadvantaged MAB instances with a sufficiently small ``warm start".
-\item[(b)] This is not necessarily the case for MAB instances that are not exploration-disadvantaged. In particular, \Thompson is a weakly dominant strategy for Needle-in-Haystack.
-\end{itemize}
-}
-\end{finding}
 
@@
 We interpret low \Eeog as a strong evidence of the ``death spiral" effect.
-Focusing on the simultaneous entry scenario, we specify the \Eeog values in Table~\ref{fig:eog} (the second line of each cell).
+Focusing on the simultaneous entry scenario, we specify the \Eeog values in Table~\ref{fig:eog}.
 We find that the \Eeog values are indeed small:
@@
 
+\begin{table*}[h]
+\centering
+\begin{adjustbox}{width=\textwidth,center}
+\begin{tabular}{|c|c|c|c||c|c|c||c|c|c|}
+  \hline
+  & \multicolumn{3}{c||}{Heavy-Tail}
+  & \multicolumn{3}{c|}{Needle-in-Haystack}
+  & \multicolumn{3}{c|}{Uniform}\\
+  \hline
+  & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500
+   & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500
+  & $T_0$ = 20 & $T_0$ = 250 & $T_0$ = 500 \\
+  \hline
+\TS vs \DG
+  & 68 (0)  & 560 (8.5)  & 610 (86.5)
+  %%
+ & 180 (30)  & 380 (0)  & 550 (6.5)
+  %%
+  &  260 (0)
+  &  780 (676.5)
+  &  880 (897.5) \\
+\hline
+  $\TS$ vs $\DEG$
+ & 37 (0)  & 430 (0)  & 540 (105)
+  %%
+ & 150 (10)  & 460 (25)  & 780 (705)
+  %%
+ & 230 (0)  & 830 (772)  & 980 (1038) \\ \hline
+  $\DG$ vs $\DEG$
+ & 340 (110)  & 640 (393)  & 670 (425)
+  %%
+  & 410 (8.5)  & 760 (666)  & 740 (646)
+  %%
+  & 530 (101)  & 990 (1058)  & 1000 (1059) \\ \hline
+\end{tabular}
+\end{adjustbox}
+\caption{\footnotesize {\bf Simultaneous Entry, \Eeog}.
+Each cell describes a game between two algorithms, call them Alg1 vs.
+Alg2, for a particular value of the warm start $T_0$.
+Each cell specifies the ``effective end of game" (\Eeog): the average and the median (in brackets).
+%For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
+The time horizon is $T=2000$.}
+\label{fig:eog}
+\end{table*}
+
 \xhdr{Welfare implications.}
@@
 
-\begin{figure}
+We assume that both firms play their respective equilibrium strategies.
+% for the corresponding competition level.
+As discussed previously, it is
+\DynamicGreedy in the monopoly scenario, and
+\DynamicGreedy for both firms in simultaneous entry scenario (Finding \ref{find:duopoly}).
+For the first-mover scenario, it is \Thompson for the incumbent (Finding \ref{find:temp-monopoly}) and \DynamicGreedy for the entrant (Finding \ref{find:temp-monopoly-entrant}).
+
+\begin{figure}[bth]
 \centering
@@
 \end{figure}
-
-We assume that both firms play their respective equilibrium strategies for the corresponding competition level. As discussed previously, these are:
-\begin{OneLiners}
-\item \DynamicGreedy in the monopoly,
-\item \DynamicGreedy for both firms in simultaneous entry (Finding \ref{find:duopoly}),
-\item \Thompson for the incumbent (Finding \ref{find:temp-monopoly}) and \DynamicGreedy for the entrant in first-mover (Finding \ref{find:temp-monopoly-entrant}).
-\end{OneLiners}
 

### 2020-09-24 journal paper/content/sim_details.tex
add bergemann valimaki paper comparisons
a pass through related work and remaining color edits
fix broken links

@@
 In particular, we need to capture a huge variety of bandit algorithms and bandit instances with only a few representative examples.
-Further, we need a succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{sim_table}).
-\asmargincomment{copied from intro}
+Further, we need a succinct and informative summarization of results within one atomic experiment and across multiple experiments (\eg see Table~\ref{fig:market_share}).
 
@@
 In the absence of competition, these classes are fairly well-understood: greedy algorithms are terrible for a wide variety of problem instances, exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances, and adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances (see Appendix~\ref{app:bg} for more details).
-\asmargincomment{copied from CODE submission}
 
@@
 
+\OMIT{
 \xhdr{Running time.}
@@
 Each experiment took about $10$ hours.
+}
 

### 2020-09-24 journal paper/ec19paper/content/revisited.tex
fix broken links

@@
 However, we find a more nuanced picture.
-For example, in Figure \ref{sim_table} we see that $\DynamicGreedy$ attains a larger market share than $\DynamicEpsGreedy$ even for large warm starts.
+For example, in Figure \ref{fig:market_share} we see that $\DynamicGreedy$ attains a larger market share than $\DynamicEpsGreedy$ even for large warm starts.
 We find that this also holds for $K = 3$ arms and longer time horizons, see the supplement for more details.

### 2020-09-24 journal paper/content/sec-intro.tex
A few more edits re Ian's comments.
Expanded a footnote.

@@
 Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
-Each principal's strategy is a bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.
+Each principal's strategy is a multi-armed bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.
 This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.
@@
 We draw on the machine learning literature to compare bandit algorithms to one another \asedit{in isolation},%
-\footnote{\asedit{As opposed to ``in competition", as solutions to a stand-alone exploration problem without competition.}}
+\footnote{\asedit{As opposed to ``in competition": as solutions to a stand-alone exploration problem which aims to maximize the total consumer welfare over time (see Appendix~\ref{app:bg} for self-contained backround).}}
  so that we can talk about \emph{better} bandit algorithms in a principled way.
@@
 There is a caveat, however: any algorithm can be defeated by interleaving it with the greedy algorithm.
-This has two undesirable consequences: a better algorithm may sometimes lose in competition, and a pure Nash equilibrium typically does not exist.
+This has two undesirable consequences: a better algorithm may sometimes lose in competition, and a Nash equilibrium typically does not exist.
 

### 2020-09-24 journal paper/content/related_work.tex
add mosteller-nogee ref
In related work / "incentivized exploration: replaced refs to a bunch of papers to a ref to the initial paper and my book chapter for a lit review.

@@
 incentivizing exploration in a recommendation system
-    \citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},
+(starting from \citet{Kremer-JPE14}, see \citet[Ch. 11.6]{slivkins-MABbook} for a literature review),
+    %\citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},
 dynamic auctions
@@
     \citep[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
-A literature review of this work can be found in
-\citep[Ch. 11.6]{slivkins-MABbook}.
+%A literature review of this work can be found in
+%\citep[Ch. 11.6]{slivkins-MABbook}.
 
@@
 Third, \SoftMaxRandom subsumes the logit choice rule, a standard behavioral model with strong empirical and microeconomic foundations
-\citep[\eg][]{luce1959choice, mcfadden1974measurement, matvejka2015rational}.
+\citep[\eg][]{mosteller1951experimental, luce1959choice, mcfadden1974measurement, matvejka2015rational}.
 Fourth, choice models similar to \SoftMaxRandom are used to explain  horizontal product differentiation \citep[\eg][]{Hotelling-29, Perloff-Salop-85}.}

### 2020-09-24 journal paper/content/sec-intro.tex
fix typo

@@
 We draw on the machine learning literature to compare bandit algorithms to one another \asedit{in isolation},%
-\footnote{\asedit{As opposed to ``in competition": as solutions to a stand-alone exploration problem which aims to maximize the total consumer welfare over time (see Appendix~\ref{app:bg} for self-contained backround).}}
+\footnote{\asedit{As opposed to ``in competition": as solutions to a stand-alone exploration problem which aims to maximize the total consumer welfare over time (see Appendix~\ref{app:bg} for self-contained background).}}
  so that we can talk about \emph{better} bandit algorithms in a principled way.

### 2020-09-24 journal paper/content/related_work.tex
One more edit in related work / exploration + incentives.
a tiny edit.
And one more small edit.

@@
 dynamic auctions
-    \citep[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
-pay-per-click ad auctions with unknown click probabilities
+    (\eg see \citet{DynAuctions-survey10} for a survey),
+    %\citep[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
+pay-per-click ad auctions 
     \citep[\eg][]{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
-coordinating search and matching by self-interested agents
+coordinating search and matching 
     \citep{Bobby-Glen-ec16},

### 2020-09-24 journal paper/content/conclusion.tex
Shortened the conclusions a bit.

@@
 A firm makes progress on its learning problem if and only if it attracts users.
-We investigate two variants: an analytical variant where users do not observe any performance signals and choose between firms according to their Bayesian-expected rewards, and numerical simulations where a reputation score is observed for each firm, based on the average reward of its recent users.
-In both variants, firms are incentivized to adopt a ``greedy algorithm" which does no purposeful exploration and leads to welfare losses for users.
+\asdelete{We investigate two variants: an analytical variant where users do not observe any performance signals and choose between firms according to their Bayesian-expected rewards, and numerical simulations where a reputation score is observed for each firm, based on the average reward of its recent users.
+In both variants,}
+We find that firms are incentivized to adopt a ``greedy algorithm" which does no purposeful exploration and leads to welfare losses for users.
 We then consider two relaxations of competition: we soften users' decision rule and give one of the firms a first-mover advantage.
@@
 
-With this paper as a departure point, there are several exciting directions to explore:
-
-\begin{itemize}
-
-\item \emph{Prices.}
-When the firms can set prices, they may be able to compensate early users for exploration, and potentially prevent the ``death spiral" effects.
-Whereas our paper zeroes in on competition between free, ad-supported platforms that primarily compete on quality.
-
-\item
-\emph{Horizontally differentiated} user preferences may help explain how competition may encourage specialization, \ie how the firms may \emph{learn to specialize} under competition.
-
-\item \emph{Continuous learning.}
-While we focus on a stationary world, another well-motivated regime is continuous learning, when exploration continuously counteracts change.
+With this paper as a departure point, there are several exciting directions to explore.
+%\begin{itemize}
+%\item \emph{Prices.}
+First, when the firms can set prices, they may be able to compensate early users for exploration, and potentially prevent the ``death spiral" effects.
+(Our paper zeroes in on competition between free, ad-supported platforms that primarily compete on quality.)
+%\item
+Second, horizontally differentiated user preferences may help explain how competition may encourage specialization, \ie how the firms may \emph{learn to specialize} under competition.
+%\item \emph{Continuous learning.}
+Third, while we focus on a stationary world, another well-motivated regime is ``continuous learning", when exploration continuously counteracts change.
 The economic story would be about competition between relatively mature firms.%

### 2020-09-24 journal paper/content/sim_details.tex
Reworded in expts / setup / terminology.

@@
 \xhdr{Terminology.}
-Following a standard game-theoretic terminology, algorithm Alg1 \emph{(weakly) dominates} algorithm Alg2 for a given firm if Alg1 provides a larger (or equal) market share than Alg2 at the end of the game.
-An algorithm is a (weakly) dominant strategy for the firm if it (weakly) dominates all other algorithms.
-This is for a particular MAB instance and a particular selection of the game parameters.
+A particular instance of the competition game is specified by the MAB instance and the game parameters, as described above.
+Recall that firms are interested in maximizing their expected market share at the end of the game.
+%Following a standard game-theoretic terminology, 
+Thus, for a given instance of the game and a given firm, algorithm Alg1 \emph{(weakly) dominates} algorithm Alg2 if Alg1 provides a larger (or equal) expected final market share than Alg2, no matter that the opponent does.
+An algorithm is a (weakly) dominant strategy for the firm if it (weakly) dominates the other two algorithms.
+%This is for a particular MAB instance and a particular selection of the game parameters.
 

### 2020-10-13 journal paper/code_abstract.tex
updated CODE abstract

--- 2020-08-25 journal paper/code_abstract.tex
@@
 \thanks{This is an extended abstract of \citet{CompetingBandits-merged}, which in turn subsumes conference papers
-\citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}.}}
+\citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}. Full version available at: \scriptsize \url{https://arxiv.org/abs/2007.10144}.}}
 % \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
@@
 }}
-\date{August 2020}
+\date{October  2020}
 \maketitle
@@
 Many platforms purposefully implement \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a standard and well-studied framework for exploration and making decisions under uncertainty
-\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
+Online platforms routinely deploy A/B tests, and are increasingly adopting  more sophisticated exploration methodologies based on \emph{multi-armed bandits}, a standard and well-studied framework for exploration and making decisions under uncertainty.
+This trend has been stimulated by two factors: almost-zero cost of deploying iterations of a product (provided an initial infrastructure investment), and the fact that many online platforms primarily compete on product quality, rather than price.
+(\eg because they are supported by ads or cheap subscriptions).
+%\citep{Gittins-book11,Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
 
@@
 
-In this paper, we initiate a study of the interplay between \exploration and \competition.
-Platforms that engage in exploration typically need to compete against one another; most importantly, they compete for users.
-This creates a tension:
+In this paper, we study the interplay between \exploration and \competition.%
+\footnote{\asedit{\Ie we add \competition to the standard exploration-exploitation tradeoff studied in multi-armed bandits.}}
+Platforms that engage in exploration typically need to compete against one another. \asedit{Most importantly, platforms compete for users, who benefit them in two ways:
+generate revenue and provide data for learning.} This creates a tension:
 %between \exploration and \competition.
-while exploration may be essential for improving the service tomorrow, it may degrade quality and make users leave \emph{today}, in which case there will be less users to learn from.
-This may further degrade the platform's performance relative to competitors who keep learning and improving from \emph{their} users, and so forth.
-Taken to the extreme, such dynamics may create a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
-Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.
+while exploration may be essential for improving the service tomorrow, it may degrade quality \emph{today}, in which case some of the users can leave and there will be less users to learn from.
+This may create a ``data feedback loop" when the platform's performance further degrades relative to competitors who keep learning and improving from \emph{their} users, and so forth.
+Taken to the extreme, such dynamics may cause a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
+\asdelete{Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.}
 
-The main high-level question is:
+The main high-level question we ask is:
+%that we focus on in this paper is:
 %\begin{align}\label{eq:main-Q}
@@
 Does increased competition lead to higher consumer welfare?
-How significant are ``data feedback loops" --- when more data leads to more users, which leads to even more data, etc.
---- and how they relate to the anti-trust considerations?
+How significant are the data feedback loops
+%--- when more data leads to more users, which leads to even more data, etc. ---
+and how they relate to the anti-trust considerations?
 %Finding formalizations that admit meaningful answers is a major part of the overall challenge.
-To answer these questions, we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
-The choice of a particular technology (exploration algorithm) is no longer an abstract, static choice with a predetermined outcome for the platform.
+We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
+
+%The choice of a particular technology (exploration algorithm) not an abstract, static choice with a predetermined outcome for the platform.
 Instead, we model the algorithms explicitly, and investigate how they play out in competition over an extended period of time.
-We offer a mix of theoretical results and numerical simulations, in a range of models.
 
@@
 
-%\subsection*{Our Model: competing bandits}
+%\subsection{Our model: competing bandits}
 %\label{sec:intro-model}
@@
 %We investigate these questions with
-\xhdr{Model: competition game.} We consider a stylized duopoly model where two firms commit to exploration strategies and compete for a stream of consumers.
-We define a game in which two firms (\emph{principals}) simultaneously engage in exploration and compete for users (\emph{agents}).
-These two processes are interlinked, as exploration decisions are experienced by users and informed by their feedback.
+\xhdr{Our model: competition game.} We consider a stylized duopoly model in which two firms (\emph{principals}) compete for users (\emph{agents}).
+Principals compete on quality rather than on prices, and engage in exploration in order to learn which actions lead to high quality products.
+A new agent arrives and chooses a principal according to some decision rule.
+The principal selects an action (\eg a list of web search results), the user experiences it and reports a reward.
+Each principal only observes its own users.
+Principals commit to their strategies in advance, with the objective to maximize the market share.
+We investigate several model variants, %within this framework,
+where we vary agents' decision rule or allow a first-mover advantage.%
+\footnote{\asedit{In particular, we use different model variants (resp., Bayesian and frequentist) for theory and for simulations so as to ensure tractability.
+Our main findings are similar for both.}}
+
+%\asedit{For tractability, our theoretical results and our numerical simulations adopt closely related but technically different model variants (respectively, Bayesian and frequentist), with similar findings.}
+
+%to one of the principals.
+%In all variants, agents have little or no information about other agents' choices and rewards.
+
+Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
+Each principal's strategy is a multi-armed bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.
+This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning.
+%Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
+We draw on the machine learning literature to compare bandit algorithms to one another \asedit{in isolation},%
+\footnote{\asedit{As opposed to ``in competition": as solutions to a stand-alone exploration problem which aims to maximize the total consumer welfare over time.}}
+ so that we can talk about \emph{better} bandit algorithms in a principled way.
+%\footnote{Such comparisons are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+More on this in Appendix~\ref{app:bg}.
+Generally,  ``better" algorithms are better in the long run, but could be worse initially.
+}
+One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); \asedit{in isolation}, they are known to be terrible for a wide variety of problem instances.
+
+%\newpage
+
+%These two processes are interlinked, as exploration decisions are experienced by users and informed by their feedback.
 We need to specify several conceptual pieces: how the principals and agents interact, what is the machine learning problem faced by each principal, and what is the information structure.
 Each piece can get rather complicated in isolation, let alone jointly, so we strive for simplicity.
-%Thus, the key features of our model are as follows.
+Thus, the key features of our model are as follows:
+
 %\begin{itemize}
-
-\textbf{(i)} 
-A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals.
+%\textbf{(i)} A new agent arrives in each round $t=1,2, \ldots$, and chooses among the two principals.
 The principal chooses an action (\eg a list of web search results to show to the agent), the user experiences this action, and reports a reward.
@@
 
-\textbf{(ii)}
-Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
+%\textbf{(ii)} Each principal faces a basic and well-studied version of the multi-armed bandit problem: for each arriving agent, it chooses from a fixed set of actions  (a.k.a.
 \emph{arms}) and receives a reward drawn independently from a fixed distribution specific to this action.
@@
 
-\textbf{(iii)}
-Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
+%\textbf{(iii)} Principals simultaneously announce their bandit algorithms before round $1$, and cannot change them afterwards.
 Each principal's objective is to maximize its market share (the fraction of users choosing this principal).
@@
 %\end{itemize}
-
-We investigate several model variants, as we vary agents' decision rule and/or posit a first-mover advantage.
 
@@
 
+%\ascomment{revising ...}
+
+\OMIT{ %%%%%%%%%% MIT CODE abstract:
 \xhdr{Technology: multi-armed bandit algorithms.}
@@
 We focus on standard notions of regret, and distinguish between three classes of bandit algorithms: ones that never explicitly explore (\emph{greedy algorithms}), ones that explore without looking at the data (\emph{exploration-separating algorithms}), and ones where exploration gradually zooms in on the best arm (\emph{adaptive-exploration algorithms}).
+In the absence of competition, these classes are fairly well-understood: greedy algorithms are terrible for a wide variety of problem instances, exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances, and adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances.
+Generally,  ``better" algorithms are better in the long run, but could be worse initially.
+} %%%%%%%%%%%
+
+\OMIT{ %%%%%%%%%% PREVIOUS VERSION
+\xhdr{Technology: multi-armed bandit algorithms.}
+To flesh out \eqref{eq:main-Q}, let us elaborate what we mean by `better' bandit algorithms.
+In general, comparisons between bandit algorithms are rather subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+
+\gaedit{While our model allows for the firms to pick from an arbitrary set of algorithms}, we distinguish between three classes of bandit algorithms, based on the prevalent intuition in the area.
+The distinction concerns the way in which they resolve the fundamental tradeoff between  exploration and \emph{exploitation} (making optimal myopic decisions using the available data).
+\gaedit{This classification is important for interpreting the equilibrium algorithm choice of the firms as it allows for a ``low" to ``medium" to ``high" technology interpretation.
+We further use the classification to pick a representative algorithm from each class for our numerical experiments.
+Going from low to medium to high technology, these three classes are as follows:} \gadelete{Going from more primitive to more sophisticated, these three classes are as follows:}
+
+\begin{itemize}
+\item \emph{Greedy algorithms} that strive to maximize the reward for the next round given the available information. Thus, they always ``exploit" and never explicitly ``explore".
+
+\item \emph{Exploration-separating algorithms}
+ that separate exploration and exploitation: essentially, each round is dedicated to one and completely ignores the other.
+
+\item \emph{Adaptive-exploration} algorithms that combine exploration and exploitation, and gradually sway the exploration choices towards more promising alternatives.
+\end{itemize}
+
 In isolation, \ie in the absence of competition, these classes are fairly well-understood.
@@
 
-%\subsection*{Our Results}
+While we build on a vast and active research area, we present sufficient background in Appendix~\ref{app:bg} so as to keep this paper accessible to non-specialists.
+} %%%%%%%%%
+
+%\subsection{Our results}
 %\label{sec:intro-results}
+
+%We offer a mix of theoretical results and numerical simulations.
+We are mainly interested in qualitative differences between the three classes of algorithms in Section~\ref{sec:intro-model}.
+For numerical simulations, we pick one representative algorithm from each class.
+For theoretical results, we allow arbitrary algorithms and focus on asymptotic differences in the algorithms' performance.
 
@@
 %We endow agents with Bayesian rationality, a common modeling approach for a theoretical investigation.
-We consider a basic Bayesian model, where agents have a common Bayesian prior on reward distributions, and know the principals' algorithms.
-For tractability, we posit that agents do not receive any information about the previous agents' choices and rewards.
-Each agent knows the round (s)he arrives in, computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
-%We refer to this variant as the \emph{\TheoryModel}.
-
-Our results depend crucially on the agents' decision rule:
+We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms, but do not receive any information about the previous agents' \asedit{choices or rewards}.
+Each agent computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
+%We focus on qualitative, asymptotic differences in the algorithms' performance.
+Our results depend crucially on agents' decision rule:
 %\begin{itemize}
 
-\textbf{(i)}  
-The most obvious decision rule (\HardMax) maximizes the Bayesian-expected reward.
-We find that \HardMax is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
-Further, we show that \HardMax is very sensitive to tie-breaking: if the tie-breaking is probabilistically biased in favor of one principal, then this principal has a simple ``winning strategy" no matter what the other principal does.
+\textbf{(i)} The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
+We find that it is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
+Further,
+%\HardMax is very sensitive to tie-breaking:
+if the tie-breaking is probabilistically biased in favor of one principal, then this principal can always prevail in competition.
+% has a simple ``winning strategy" no matter what the other principal does.
 
-\textbf{(ii)} 
-We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
+\textbf{(ii)} We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
 We call this model \HardMaxRandom.
-We find that better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
-However, there is a substantial caveat: one can defeat any algorithm by interleaving it with the greedy algorithm.
-This has two undesirable corollaries: a better algorithm may sometimes lose in competition, and a pure Nash equilibrium typically does not exist.
+Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
+There is a caveat, however: any algorithm can be defeated by interleaving it with the greedy algorithm.
+This has two undesirable consequences: a better algorithm may sometimes lose in competition, and a Nash equilibrium typically does not exist.
 
-\textbf{(iii)} 
-We further relax the decision rule so that the probability of choosing a given principal varies smoothly as a function of the difference between  principals' Bayesian-expected rewards; we call it \SoftMaxRandom.
-Then, the ``better algorithm wins" result holds under much weaker assumptions on what constitutes a better algorithm.
+\textbf{(iii)} We further soften the decision rule so that the selection probabilities vary smoothly in terms of principal's Bayesian-expected rewards;
+%as a function of the difference between  principals' Bayesian-expected rewards;
+we call it \SoftMaxRandom.%
+\footnote{\asedit{In particular, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}}
+ Then a better algorithm prevails under much weaker assumptions.
+% on what constitutes a better algorithm.
 This is the most technical result of the paper.
-The competition in this setting is necessarily much more relaxed: typically, both principals attract approximately half of the agents as time goes by (but a better algorithm would attract slightly more).
+\asdelete{The competition is more relaxed: the better algorithm receives only a small advantage in the long run.}
 
-\xhdr{Economic interpretation: the inverted-U relationship.}
-Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the \emph{inverted-U relationship} between competition and innovation.% 
-%\footnote{Here, ``innovation" refers to adoption of a better technology at a substantial R\&D expense to a given firm.
-It is not salient whether similar ideas and/or technologies already exist elsewhere.
-Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
-\footnote{This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{aghion2005competition,Vives-08}.}
+%both principals attract approximately half of the agents as time goes by, but a better algorithm attracts slightly more.
+%\end{itemize}
 
-Our decision rules differ in terms of rationality: from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
+%Most results admit generalizations
+%extend to a much more general version of the multi-armed bandit problem
+%in which the principal may observe additional feedback before and/or after each decision,
+%as long as the feedback distribution does not change over time. In most results,
+% and principal's utility may include agents' rewards.
+
+%\begin{figure}
+%\begin{center}
+%\begin{tikzpicture}
+%      \draw[->] (-.5,0) -- (3,0) node[right] {Competitiveness};
+%      \draw[->] (0,-.5) -- (0,3) node[above] {Better algorithms};
+%      \draw[scale=0.6,domain=0.5:4.5,smooth,variable=\x,blue, line width=0.3mm] plot ({\x},{4.5 - (\x - 2.5)^2});
+%      % \draw[scale=0.5,domain=-3:3,smooth,variable=\y,red]  plot ({\y*\y},{\y});
+% \end{tikzpicture}
+%
+%\caption{Inverted-U relationship between competitiveness and algorithms.}
+%\label{fig:inverted-U}
+%\end{center}
+%\end{figure}
+
+\xhdr{Interpretation: the inverted-U relationship.}
+Our findings can be framed in terms of the inverted-U relationship between competition and innovation.
+This is a well-established concept, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better.
+%\citep[\eg][]{aghion2005competition,Vives-08}.
+We interpret innovation as the adoption of better algorithms,%
+\footnote{Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms are well-known and/or similar technologies already exist elsewhere  \citep[\eg see][]{DS-arxiv}.}
+%, even if the algorithms themselves are well-known in the research literature
+and control the severity of the competition by varying the agents' decision rule from \HardMax (cut-throat competition) to \HardMaxRandom to \SoftMaxRandom and all the way to the uniform selection.
+%\footnote{Agents' decision rule also controls the agents' rationality.
+While agents' rationality and severity of competition are often modeled separately, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
+We also find another, technically different inverted-U relationship which zeroes in on the \HardMaxRandom model.
+%We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
+The barriers for innovation arise entirely from the reputational consequences of exploration in competition, \asedit{without invoking monetary costs and benefits.
+Whereas the inverted-U relationships in prior work are due to the monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.}
+
+%the tradeoff between the benefits of the technology and its R\&D costs.
+
+%While the inverted-U relationships in prior work focus on the tradeoff between the benefits of the technology and its R\&D costs, in our work the barriers for innovations arise entirely from the reputational consequences of exploration in competition, even in the absence of R\&D costs.
+
+%Our decision rules differ in terms of rationality: from fully rational decisions with \HardMax to relaxed rationality with \HardMaxRandom to an even more relaxed rationality with \SoftMaxRandom.
 The same distinctions also control the severity of competition between the principals: from cut-throat competition with \HardMax to a more relaxed competition with \HardMaxRandom, to an even more relaxed competition with \SoftMaxRandom.
@@
 \xhdr{Numerical simulations.}
-We consider a basic frequentist model.
-We posit that the agents observe signals about the principals' past performance, and base their decisions on these signals alone, without invoking any prior knowledge or beliefs.
+We consider a basic frequentist model (called the \emph{\ExptsModel}), where the agents observe signals about the principals' past performance, and base their decisions on these signals alone, without invoking any prior knowledge or beliefs.
 The performance signals are abstracted and aggregated as a scalar \emph{reputation score} for each principal, modeled as a sliding window average of its rewards.
 Thus, agents' decision rule depends only on the two reputation scores.
-%We refer to this variant as the \emph{\ExptsModel}.%
+% We refer to this variant as the \emph{\ExptsModel}.%
 %\footnote{In comparison, the theoretical results focus on another extreme, with Bayesian rationality and no performance signals.}
-We refine and expand the theoretical results in several ways.
+We refine and expand the theoretical results in several ways:
 
@@
 \textbf{(i)}
-We compare \HardMax and \HardMaxRandom decision rules. We find that the greedy algorithm often wins under \HardMax, with a strong evidence of the ``death spiral" effect mentioned earlier.
- As predicted by the theory, better algorithms prevail if the expected number of ``random" users is sufficiently large. However, this effect is negligible for smaller parameter values.
+%We compare \HardMax and \HardMaxRandom decision rules.
+We find that the greedy algorithm often wins under the \HardMax decision rule, with a strong evidence of the ``death spiral" effect mentioned earlier.
+ As predicted by the theory, better algorithms prevail under \HardMaxRandom if the expected number of ``random" users is sufficiently large.
+\asdelete{However, this effect is negligible for smaller parameter values.}
 
@@
 
-\textbf{(ii)}  
- We investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
-(We focus on the \HardMax decision rule.) We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous arrival).
+\textbf{(ii)} Focusing on \HardMax, we investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
+We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous arrival).
 Consumer welfare is higher under early/late arrival than under simultaneous entry.
@@
 
-\textbf{(iii)}  
-We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}).
-We run additional experiments so as to isolate and compare these two effects.
+\textbf{(iii)} We investigate the algorithms' performance without competition.
+We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).}
+     We find that mean reputation -- arguably, the most natural performance measure -- is sometimes \emph{not} a good predictor for the outcomes under competition.
+
+\textbf{(iv)} We decompose the first-mover advantage into two distinct effects: free data to learn from (\emph{data advantage}), and a more definite, and possibly better reputation compared to an entrant (\emph{reputation advantage}), and run additional experiments to separate and compare them.
 We find that either effect alone leads to a significant advantage under competition.
@@
 but ignore the differences induced by learning in isolation versus under competition.
-Furthermore, explicitly incorporating the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified not only by data quantity, but also the increased data quality gathered by better learning algorithms.
+Furthermore, explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.
 
-\newpage
+\xhdr{Significance.}
+Our results have a dual purpose: shed light on real-world implications of some typical scenarios, and investigate the space of models for describing the real world.
+As an example for the latter: while the \HardMax model with simultaneous entry is arguably the most natural model to study \emph{a priori}, our results elucidate the need for more refined models with ``free exploration" (\eg via random agents or early entry).
+On a technical level, we connect a literature on regret-minimizing bandits in computer science and that on competition in economics.
+
+Our theory takes a basic Bayesian approach, a standard perspective in economic theory, and discovers several strong asymptotic results.
+Much of the difficulty, both conceptual and technical, is in setting up the model and the theorems.
+Apart from zeroing in on the \TheoryModel, it was crucial to interpret the results and intuitions from the literature on multi-armed bandits so as to formulate meaningful and productive assumptions on bandit algorithms and Bayesian priors.
+
+The numerical simulations provide a more nuanced and ``non-asymptotic" perspective.
+In essence, we look for substantial effects within relevant time scales.
+(In fact, we start our investigation by determining what time scales are relevant in the context of our model.) The central challenge is to capture a huge variety of bandit algorithms and bandit problem instances with only a few representative examples, and arrive at findings that are consistent across the entire space.
+
+One model we study is suitable for analysis and another for simulations, but not vice versa. A natural implementation of the \TheoryModel requires running time quadratic in the number of rounds,%
+\footnote{\label{fn:Tsquared}\Eg this is because at each round $t$, one needs to recompute, and integrate over, a discrete distribution with $t$ possible values, namely the number of agents that have chosen principal $1$ so far.}
+which precludes experiments at a sufficient scale.
+The \ExptsModel features an intricate feedback loop between algorithms' performance, their reputations and agents' choices, which simplifies the simulations but does not appear analytically tractable.
+
+\xhdr{Discussion}: Reflecting reality, we posit that \asedit{the exploration strategies available to the principals} are ``non-strategic" in nature.
+Even though the principals play a multi-step game, they do not react to each other's moves or to the agents' strategic choices.
+This is how industry approaches exploration algorithms, and for several good reasons.
+"Non-strategic" exploration is well-studied in machine learning, and yet it remains a very complex subject in research (as evidenced, \eg by the huge amount of activity therein).
+Even the seemingly simple algorithms are not straightforward to deploy in practice, and require a substantial investment in infrastructure \cite[\eg see the discussions in][]{DS-arxiv}.
+Responding to the competition represents another layer of complexity which has not been previously studied in this context, to the best of our knowledge, let alone made even remotely practical.
+Besides, the competitor's exploration strategy is typically not public, and understanding its exploration behavior via observations appears challenging even as a  research problem.%
+\footnote{Principals could potentially react to the market share or (the difference in) reputation scores.
+However, baking these signals into one's exploration strategy runs the risk of over-interpreting our competition model, as they may change for exogenous reasons.
+Alternatively, one could use such signals, as well as the intuitions coming from this paper, to guide the platform's decisions regarding exploration.}
+
+%\gaedit{Finally, the assumption that firms commit to the algorithms at the start of the game deserves further justification.
+We impose this commitment assumption since it is more realistic than considering dynamic strategies and algorithms are, by definition, a commitment to a set of rules.
+In many contexts where such exploration algorithms are typically deployed, consumers arrive at such a rapid volume that it it is infeasible to consider a firm dynamically adapting its algorithm to each consumer that arrives.
+Indeed, \cite{DS-arxiv} points out how difficult it is to deploy standard exploration algorithms in practice and such dynamic strategies are substantially more difficult to deploy in practice.
+While it is possible to consider share-adaptive algorithms that would still keep the commitment assumption but allow for indirect dynamic responses to market conditions, such algorithms do not exist currently in the literature but are a good avenue for future work.}
+
+Our models are stylized in several important  respects.
+Firms compete only on the quality of service, rather than, say, pricing or the range of products.
+Agents are myopic: they do not worry about how their actions impact their future utility.\footnote{So, agents do not attempt to learn over time, second-guess or game future agents, or manipulate the principals' learning algorithms.
+This is arguably typical in practice, in part because one agent's influence tends to be small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores" reflecting the current reputation, and further simplified to average performance over a sliding time window.
+
+We consider two extremes: a simple Bayesian model with full Bayesian rationality and no performance signals, and a simple frequentist model with reputation scores and no prior knowledge or beliefs.
+For the theoretical results, the ``no-performance-signals" assumption makes agents' behavior independent of a particular realization of the prior.
+Therefore, we summarize each learning algorithm via its Bayesian-expected rewards, not worrying about its detailed performance on particular realizations of the prior.
+Such summarization is essential for formulating lucid and general analytic results, let alone proving them.
+It is unclear how to incorporate performance signals in a theoretically tractable model.
+For the numerical results, the \ExptsModel accounts for competition in a more direct way,
+% \gaedit{by positing a more realistic model of consumer behavior that does not
+not requiring users to have direct information on the algorithms deployed or the bandit problem faced by the firms.
+It further allows us to separate reputation vs.
+data advantage and makes our model amenable to numerical simulations.
+%(Whereas simulating the intricate interplay of learning dynamics and Bayesian priors appears computationally intractable, see Footnote~\ref{fn:Tsquared}.)
+
+On the machine learning side, our model captures big, qualitative differences between bandit algorithms, building on the well-established intuition in the literature.
+Comparisons between bandit algorithms are generally somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+In particular, ``better" algorithms are better in the long run, but could be worse initially.
+We focus on a standard model of stochastic bandits.
+However, we are less interested in state-of-art algorithms for realistic applications.
+
+Bandit rewards are not discounted with time, reflecting the fact that non-discounted, regret-minimizing formulations has been prevalent in the bandits literature over the past two decades \citep{slivkins-MABbook,LS19bandit-book}, and better correspond to practical deployments \citep[\eg][]{DS-arxiv}.
+Also, the distinctions between better and worse bandit algorithms are not as well-understood under time-discounting.
 
@@
 %
+
+\newpage
 

### 2020-11-21 journal paper/content/related_work.tex
updates

@@
 competition \citep[\eg][]{fudenberg1983learning, dasgupta1988learning, cabral1994learning} studies firms that learn while competing against each other, so that a firm attracting more consumers reduces its per-unit production costs.
-However, our model focuses on firms learning product quality (as opposed to reducing production costs) and the impact that this learning has on attracting consumers.
+\gaedit{However, our model differs from these models in two important respects.
+The first is that in our model firms focus on learning product quality (as opposed to reducing production costs) and the second is that firms' current period actions have reputational consequences that directly impact consumer choices in future periods.
+These features are absent from traditional learning-by-doing models, but important components of a model that tries to capture data-based innovation.}
 

### 2020-11-21 journal paper/content/sec-intro.tex
updates

@@
 Each principal only observes its own users.
-Principals commit to their strategies in advance, with the objective to maximize the market share.
+Principals commit to their strategies in advance, with the objective to maximize \gaedit{their} market share.
 We investigate several model variants, %within this framework,
@@
 }
-One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); \asedit{in isolation}, they are known to be terrible for a wide variety of problem instances.
+One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); \asedit{in isolation}, they are known to \gaedit{perform poorly}\gadelete{be terrible} for a wide variety of problem instances.
 
@@
 
-\textbf{(ii)} We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.
-We call this model \HardMaxRandom.
+\textbf{(ii)} We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random.\gaedit{\footnote{\gaedit{We provide an interpretation of the ``random" agents as reputation-agnostic consumers that are not aware of the reputation signals or the principals' strategies.}}} We call this model \HardMaxRandom.
 Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
@@
 %\citep[\eg][]{aghion2005competition,Vives-08}.
-We interpret innovation as the adoption of better algorithms,%
-\footnote{Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms are well-known and/or similar technologies already exist elsewhere  \citep[\eg see][]{DS-arxiv}.}
-%, even if the algorithms themselves are well-known in the research literature
-and control the severity of the competition by varying the agents' decision rule from \HardMax (cut-throat competition) to \HardMaxRandom to \SoftMaxRandom and all the way to the uniform selection.
+\gaedit{We recover a similar relationship even though our focus differs from the previous literature by analyzing data-based product innovation, which crucially depends on data that is generated by the consumers of the firm.
+We interpret the adoption of better exploration algorithms as innovation and control the severity of the competition by varying the agents' decision rule from \HardMax (cut-throat competition) to \HardMaxRandom to \SoftMaxRandom and all the way to the uniform selection.\footnote{Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms are well-known and/or similar technologies already exist elsewhere  \citep[\eg see][]{DS-arxiv}.}$^{,}$\gaedit{\footnote{\gaedit{This interpretation is what allows us to connect to the regret-minimizing bandit literature in order to discuss innovation in a principled manner by providing us with a disciplined comparison between the algorithms selected by the firm.}}}}
 %\footnote{Agents' decision rule also controls the agents' rationality.
@@
 %We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
-The barriers for innovation arise entirely from the reputational consequences of exploration in competition, \asedit{without invoking monetary costs and benefits.
-Whereas the inverted-U relationships in prior work are due to the monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.}
+\gaedit{Unlike traditional innovation models which study lab-based innovation, our focus on data-based innovation generates the traditional inverted-U relationship purely through the reputational consequences of exploration in competition, without explicitly invoking monetary costs and benefits.
+This is in contrast to the inverted-U relationships established in prior work which rely on monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.}
 
@@
 Furthermore, explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.
+\gaedit{Indeed, by exploring the first-mover advantage channel, we show how such incumbency may actually be good for innovation, but also lead to a barrier to entry precisely due to this feedback loop and how it's impacted by competition.}
 

### 2020-11-21 journal paper/content/sec-model.tex
updates

@@
 One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
-Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.}
+Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.} \gaedit{We provide an interpretation of the ``random agents" as reputation-agnostic consumers that are not aware of or do not seek out the reputation signals available to them in the respective choice model variants.}
 
@@
 \footnote{\asedit{Note that a mixed strategy is also a (randomized) MAB algorithm, \ie also a pure strategy.}}
- Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral: they optimize their expected utility.
+ Principal's utility is defined as the market share, \ie the number of agents that chose this principal.
+Principals are risk-neutral \gaedit{and aim to maximize their expected utility.}\gadelete{ : they optimize their expected utility.}
 

### 2020-11-21 journal paper/journal_main.tex
updates

@@
 with a unified and streamlined presentation and expanded related work and background materials.
-All theoretical results are from \citet{CompetingBandits-itcs18}, and all experiments are from \citet{CompetingBandits-ec19}.
+All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
 Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.
-We would like to thank Ian Ball, Yeon-Koo Che, Glen Weyl and Sven Rady, as well as seminar participants at Columbia and conference participants at Innovations in Theoretical Computer Science 2018 and ACM Economics and Computation 2019, for helpful comments and conversations.
+We would like to thank Ian Ball, Yeon-Koo Che, Glen Weyl and Sven Rady, as well as seminar participants at Columbia and conference participants at Innovations in Theoretical Computer Science 2018, ACM Economics and Computation 2019 and the MIT Conference on Digital Experimentation 2020, for helpful comments and conversations.
 All errors are our own.} }

### 2020-11-25 journal paper/content/related_work.tex
Removed some color-edit commands

@@
 
-\asedit{Classic ``market competitiveness" measures,such as the Lerner Index or the Herfindahl-Hirschman Index
-\citep{tirole1988theory} are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).}
+Classic ``market competitiveness" measures,such as the Lerner Index or the Herfindahl-Hirschman Index
+\citep{tirole1988theory} are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).
 
@@
 
-\asedit{\xhdr{Choice models.}
+\xhdr{Choice models.}
 Stochastic choice models similar to ours have been widely used throughout economics.
@@
 \citep[\eg][]{mosteller1951experimental, luce1959choice, mcfadden1974measurement, matvejka2015rational}.
-Fourth, choice models similar to \SoftMaxRandom are used to explain  horizontal product differentiation \citep[\eg][]{Hotelling-29, Perloff-Salop-85}.}
+Fourth, choice models similar to \SoftMaxRandom are used to explain  horizontal product differentiation \citep[\eg][]{Hotelling-29, Perloff-Salop-85}.
 %\asdelete{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}

### 2020-11-25 journal paper/content/sec-intro.tex
Some edits.

--- 2020-09-24 journal paper/content/sec-intro.tex
@@
 In this paper, we study the interplay between \exploration and \competition.%
-\footnote{\asedit{\Ie we add \competition to the standard exploration-exploitation tradeoff studied in multi-armed bandits.}}
-Platforms that engage in exploration typically need to compete against one another. \asedit{Most importantly, platforms compete for users, who benefit them in two ways:
-generate revenue and provide data for learning.} This creates a tension:
+\footnote{\Ie we add \competition to the standard exploration-exploitation tradeoff studied in multi-armed bandits.}
+Platforms that engage in exploration typically need to compete against one another. Most importantly, platforms compete for users, who benefit them in two ways:
+generate revenue and provide data for learning. This creates a tension:
 %between \exploration and \competition.
@@
 Taken to the extreme, such dynamics may cause a ``death spiral" effect when the vast majority of customers eventually switch to competitors.
-\asdelete{Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.}
+
+%\asdelete{Users therefore serve three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents who choose among the competing systems.}
 
@@
 Each principal only observes its own users.
-Principals commit to their strategies in advance, with the objective to maximize the market share.
+Principals commit to their strategies in advance, with the objective to maximize their market share.
 We investigate several model variants, %within this framework,
 where we vary agents' decision rule or allow a first-mover advantage.%
-\footnote{\asedit{In particular, we use different model variants (resp., Bayesian and frequentist) for theory and for simulations so as to ensure tractability.
-Our main findings are similar for both.}}
+\footnote{In particular, we use different model variants (resp., Bayesian and frequentist) for theory and for simulations so as to ensure tractability. Our main findings are similar for both.}
 
@@
 %Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
-We draw on the machine learning literature to compare bandit algorithms to one another \asedit{in isolation},%
-\footnote{\asedit{As opposed to ``in competition": as solutions to a stand-alone exploration problem which aims to maximize the total consumer welfare over time (see Appendix~\ref{app:bg} for self-contained backround).}}
+We draw on the machine learning literature to compare bandit algorithms to one another in isolation,%
+\footnote{As opposed to ``in competition": as solutions to a stand-alone exploration problem which aims to maximize the total consumer welfare over time (see Appendix~\ref{app:bg} for self-contained background).}
  so that we can talk about \emph{better} bandit algorithms in a principled way.
@@
 }
-One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); \asedit{in isolation}, they are known to be terrible for a wide variety of problem instances.
+One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in isolation, they are known to perform poorly for a wide variety of problem instances.
 
@@
 %We endow agents with Bayesian rationality, a common modeling approach for a theoretical investigation.
-We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms, but do not receive any information about the previous agents' \asedit{choices or rewards}.
+We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms, but do not receive any information about the previous agents' choices or rewards.
 Each agent computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
@@
 \textbf{(ii)} We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random. 
-We call this model \HardMaxRandom.
+\asedit{(One can interpret them as consumers that are not aware of the reputation signals or the principals' strategies.)} We call this model \HardMaxRandom.
 Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
@@
 
-\textbf{(iii)} We further soften the decision rule so that the selection probabilities vary smoothly in terms of principal's Bayesian-expected rewards;
+\textbf{(iii)} We further soften the decision rule so that the selection probabilities vary smoothly in terms of principals' Bayesian-expected rewards;
 %as a function of the difference between  principals' Bayesian-expected rewards;
-we call it \SoftMaxRandom.%
-\footnote{\asedit{In particular, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}}
- Then a better algorithm prevails under much weaker assumptions.
+We call it \SoftMaxRandom, \asedit{a realistic middle ground between \HardMax agents and random agents.}%
+\footnote{\asedit{Alternatively}, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}
+\asedit{In the most technical result of the paper, we find that} a better algorithm prevails under much weaker assumptions.
 % on what constitutes a better algorithm.
-This is the most technical result of the paper.
-\asdelete{The competition is more relaxed: the better algorithm receives only a small advantage in the long run.}
+%This is the most technical result of the paper.
+
+%\asdelete{The competition is more relaxed: the better algorithm receives only a small advantage in the long run.}
 
@@
 %\citep[\eg][]{aghion2005competition,Vives-08}.
-We interpret innovation as the adoption of better algorithms,%
+We interpret innovation as the adoption of better exploration algorithms,%
 \footnote{Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms are well-known and/or similar technologies already exist elsewhere \citep[\eg see][]{DS-arxiv}.}
-%, even if the algorithms themselves are well-known in the research literature
 and control the severity of the competition by varying the agents' decision rule from \HardMax (cut-throat competition) to \HardMaxRandom to \SoftMaxRandom and all the way to the uniform selection. 
@@
 %We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
-The barriers for innovation arise entirely from the reputational consequences of exploration in competition, \asedit{without invoking monetary costs and benefits.
-Whereas the inverted-U relationships in prior work are due to the monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.}
+
+\asedit{While traditional models of innovation study lab-based R\&D, we consider data-driven innovation: one that crucially depends on data generated by the firm's customers. 
+\asmargincomment{important sentence!}
+We focus on innovation in \emph{exploration technology} which systematically improves the firm's products, whereas prior work would define innovation as improvement in the products themselves.
+We recover the traditional inverted-U relationship purely through the reputational consequences of exploration, rather than monetary costs/benefits thereof.
+This is in contrast to the inverted-U relationships established in prior work, which rely on the monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.}
+
+% Alex -> Guy: this point has been made already!
+%We bring to bear the literature on multi-armed bandits so as to discuss and compare exploration technologies in a principled manner.
 
@@
  As predicted by the theory, better algorithms prevail under \HardMaxRandom if the expected number of ``random" users is sufficiently large. 
-\asdelete{However, this effect is negligible for smaller parameter values.}
+ %\asdelete{However, this effect is negligible for smaller parameter values.}
 

### 2020-11-25 journal paper/content/sec-model.tex
Removed my "old" color-edits.

--- 2020-09-24 journal paper/content/sec-model.tex
@@
 \footnote{We use `action' and `arm' interchangeably, as common in the literature on multi-armed bandits.}
-The agent experiences this action and receives a reward $r_t\in \{ 0,1\}$ \asedit{for this action, which is then observed by the principal}.
+The agent experiences this action and receives a reward $r_t\in \{ 0,1\}$ for this action, which is then observed by the principal.
 We posit \emph{stochastic rewards}: whenever a given action $a\in A$ is chosen, the reward is an independent draw from Bernoulli distribution with mean $\mu_a$.
@@
 
-\asedit{Thus, each principal follows the protocol of \emph{multi-armed bandits} (\emph{MAB}): in each round when it is chosen, it chooses an arm from $A$ and observes a reward for this action (and nothing else).}
+Thus, each principal follows the protocol of \emph{multi-armed bandits} (\emph{MAB}): in each round when it is chosen, it chooses an arm from $A$ and observes a reward for this action (and nothing else).
 
@@
 
-\asedit{\HardMaxRandom and \SoftMaxRandom can be seen as agent types that, resp., give some chance to the other principal and avoid sharp transitions in their probabilistic response.
+\HardMaxRandom and \SoftMaxRandom can be seen as agent types that, resp., give some chance to the other principal and avoid sharp transitions in their probabilistic response.
 Alternatively, they can be realized as mixtures over more ``basic" agent types.
@@
 One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
-Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.}
+Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.
+\gaedit{We provide an interpretation of the ``random agents" as reputation-agnostic consumers that are not aware of or do not seek out the reputation signals available to them in the respective choice model variants.}
 
@@
 \item In the \emph{\TheoryModel}, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$.
-\asedit{Each agent knows its global round $t$, but not the performance signals such as the current market shares.} Agents' reward estimates are defined as posterior mean rewards:
+Each agent knows its global round $t$, but not the performance signals such as the current market shares.
+Agents' reward estimates are defined as posterior mean rewards:
         $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i}$
@@
 
-\item In the \emph{\ExptsModel}, agents' reward estimate for a given principal is the average reward of the last $M$ agents that chose this principal, interpreted as a \asedit{current} \emph{reputation score}.
+\item In the \emph{\ExptsModel}, agents' reward estimate for a given principal is the average reward of the last $M$ agents that chose this principal, interpreted as a current \emph{reputation score}.
 To make it meaningful initially, each principal enjoys a ``warm start": additional $T_0$ agents arrive before the game starts, and interact with the principal as described above.
@@
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-\asedit{Pure strategies in this game correspond to MAB algorithms.}%
-\footnote{\asedit{Note that a mixed strategy is also a (randomized) MAB algorithm, \ie also a pure strategy.}}
- Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral: they optimize their expected utility.
+Pure strategies in this game correspond to MAB algorithms.%
+\footnote{Note that a mixed strategy is also a (randomized) MAB algorithm, \ie also a pure strategy.}
+ Principal's utility is defined as the market share, \ie the number of agents that chose this principal.
+Principals are risk-neutral \gaedit{and aim to maximize their expected utility.}\gadelete{ : they optimize their expected utility.}
 
@@
 
-Reflecting reality, we posit that \asedit{the exploration strategies available to the principals} are ``non-strategic" in nature.
+Reflecting reality, we posit that the exploration strategies available to the principals are ``non-strategic" in nature.
 Even though the principals play a multi-step game, they do not react to each other's moves or to the agents' strategic choices.

### 2020-11-25 journal paper/content/sec-intro.tex
Reworded Guy's sentence.

@@
 Furthermore, explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.
-\gaedit{Indeed, by exploring the first-mover advantage channel, we show how such incumbency may actually be good for innovation, but also lead to a barrier to entry precisely due to this feedback loop and how it's impacted by competition.}
+\asedit{In particular, we find that incumbency is actually good for innovation, but also creates a barrier to entry precisely due to this feedback loop.}
+% and how it is impacted by competition.
 

### 2020-11-25 journal paper/content/related_work.tex
Reworded/expanded Guy's new edit on LbD.

@@
 The literature on learning-by-doing vs.
-competition \citep[\eg][]{fudenberg1983learning, dasgupta1988learning, cabral1994learning} studies firms that learn while competing against each other, so that a firm attracting more consumers reduces its per-unit production costs.
-\gaedit{However, our model differs from these models in two important respects.
+competition \citep[\eg][]{fudenberg1983learning, dasgupta1988learning, cabral1994learning} studies firms that learn while competing against each other, so that a firm attracting more consumers reduces its production costs.
+\asedit{Our model differs in several important respects.
+First, firms learn to improve product quality rather than to reduce production costs.
+Second, the firms' current actions have consequences (via reputation and/or data collected by the algorithm) that directly impact consumer choices in the future.
+Third, we endogenize the technology behind learning-by-doing by explicitly considering bandit algorithms, and make comparisons between different such technologies.}
+
+%\gaedit{However, our model differs from these models in two important respects.
 The first is that in our model firms focus on learning product quality (as opposed to reducing production costs) and the second is that firms' current period actions have reputational consequences that directly impact consumer choices in future periods.

### 2020-11-25 journal paper/content/sec-intro.tex
Minor edits.

@@
 \textbf{(ii)} We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random. 
-\asedit{(One can interpret them as consumers that are not aware of the reputation signals or the principals' strategies.)} We call this model \HardMaxRandom.
+\asedit{(They can be interpreted as consumers that are oblivious to the principals' reputation.)} We call this model \HardMaxRandom.
 Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
@@
 %as a function of the difference between  principals' Bayesian-expected rewards;
-We call it \SoftMaxRandom, \asedit{a realistic middle ground between \HardMax agents and random agents.}%
+We call it \SoftMaxRandom, \asedit{a more realistic middle ground between \HardMax agents and random agents.}%
 \footnote{\asedit{Alternatively}, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}

### 2020-11-25 journal paper/content/sec-model.tex
Minor edits.

@@
 Indeed, \HardMaxRandom is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
-One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
+\asedit{The latter can be interpreted as consumers oblivious to principals' reputation, due to the lack of awareness or interest.} One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
 Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.
-\gaedit{We provide an interpretation of the ``random agents" as reputation-agnostic consumers that are not aware of or do not seek out the reputation signals available to them in the respective choice model variants.}
 
@@
 \footnote{Note that a mixed strategy is also a (randomized) MAB algorithm, \ie also a pure strategy.}
- Principal's utility is defined as the market share, \ie the number of agents that chose this principal.
-Principals are risk-neutral \gaedit{and aim to maximize their expected utility.}\gadelete{ : they optimize their expected utility.}
+ Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral and aim to maximize their expected utility.
 

### 2020-11-25 journal paper/content/related_work.tex
Minor edits.
add changes

@@
 competition \citep[\eg][]{fudenberg1983learning, dasgupta1988learning, cabral1994learning} studies firms that learn while competing against each other, so that a firm attracting more consumers reduces its production costs.
-\asedit{Our model differs in several important respects.
+Our model differs in several important respects.
 First, firms learn to improve product quality rather than to reduce production costs.
 Second, the firms' current actions have consequences (via reputation and/or data collected by the algorithm) that directly impact consumer choices in the future.
-Third, we endogenize the technology behind learning-by-doing by explicitly considering bandit algorithms, and make comparisons between different such technologies.}
+Third, we endogenize the technology behind learning-by-doing by explicitly considering bandit algorithms.
 

### 2020-11-25 journal paper/content/sec-intro.tex
add changes

@@
 \textbf{(ii)} We dilute the \HardMax agents with a small fraction of ``random agents" who choose a principal uniformly at random. 
-\asedit{(They can be interpreted as consumers that are oblivious to the principals' reputation.)} We call this model \HardMaxRandom.
+(They can be interpreted as consumers that are oblivious to the principals' reputation.) We call this model \HardMaxRandom.
 Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
@@
 %as a function of the difference between  principals' Bayesian-expected rewards;
-We call it \SoftMaxRandom, \asedit{a more realistic middle ground between \HardMax agents and random agents.}%
-\footnote{\asedit{Alternatively}, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}
-\asedit{In the most technical result of the paper, we find that} a better algorithm prevails under much weaker assumptions.
+We call it \SoftMaxRandom, a more realistic middle ground between \HardMax agents and random agents.%
+\footnote{Alternatively, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}
+In the most technical result of the paper, we find that a better algorithm prevails under much weaker assumptions.
 % on what constitutes a better algorithm.
@@
 
-\asedit{While traditional models of innovation study lab-based R\&D, we consider data-driven innovation: one that crucially depends on data generated by the firm's customers. 
+While traditional models of innovation study lab-based R\&D, we consider data-driven innovation: one that crucially depends on data generated by the firm's customers. 
 \asmargincomment{important sentence!}
@@
 We recover the traditional inverted-U relationship purely through the reputational consequences of exploration, rather than monetary costs/benefits thereof.
-This is in contrast to the inverted-U relationships established in prior work, which rely on the monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.}
+This is in contrast to the inverted-U relationships established in prior work, which rely on the monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.
 
@@
 Furthermore, explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.
-\asedit{In particular, we find that incumbency is actually good for innovation, but also creates a barrier to entry precisely due to this feedback loop.}
+In particular, we find that incumbency is actually good for innovation, but also creates a barrier to entry precisely due to this feedback loop.
 % and how it is impacted by competition.

### 2020-11-25 journal paper/content/sec-model.tex
add changes

--- 2020-11-25 journal paper/content/sec-model.tex
@@
 Indeed, \HardMaxRandom is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
+The latter can be interpreted as consumers oblivious to principals' reputation, due to the lack of awareness or interest.
 One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
 Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.
-\gaedit{We provide an interpretation of the ``random agents" as reputation-agnostic consumers that are not aware of or do not seek out the reputation signals available to them in the respective choice model variants.}
 

### 2020-11-25 journal paper/content/sec-theory.tex
add changes

@@
 \begin{theorem}\label{thm:DG-dominance}
-Assume \HardMax response function with fair tie-breaking: \asedit{$\respF(0)=\nicefrac{1}{2}$}.
+Assume \HardMax response function with fair tie-breaking: $\respF(0)=\nicefrac{1}{2}$.
 Assume that \alg[1] is \DynGreedy, and \alg[2] deviates from \DynGreedy starting from some (local) step $n_0<T$.
@@
 
-\asedit{\DynGreedy is a weakly dominant strategy in the competition game, and a unique Nash equilibrium.
+\DynGreedy is a weakly dominant strategy in the competition game, and a unique Nash equilibrium.
 This is because \DynGreedy receives, in expectation, at least half of the agents before global round $n_0$, and all agents after that; both are the best possible against \alg[2].
-Moreover, \DynGreedy guarantees at least $T/2$ agents in expectation, and any other strategy can receive strictly less than that (\eg if the opponent chooses \DynGreedy).}
+Moreover, \DynGreedy guarantees at least $T/2$ agents in expectation, and any other strategy can receive strictly less than that (\eg if the opponent chooses \DynGreedy).
 
 \begin{corollary}\label{cor:DG-dominance}
-\asedit{\DynGreedy is a weakly dominant strategy in the competition game.} The game has a unique Nash equilibrium: both principals choose \DynGreedy.
+\DynGreedy is a weakly dominant strategy in the competition game. The game has a unique Nash equilibrium: both principals choose \DynGreedy.
 \end{corollary}
@@
 \xhdr{Biased tie-breaking.}
-The \HardMax model is very sensitive to tie-breaking \asedit{between the principals}.
+The \HardMax model is very sensitive to tie-breaking between the principals.
 For starters, if ties are  broken deterministically in favor of principal $1$, then principal 1 can get all agents no matter what the other principal does, simply by using \StaticGreedy.
@@
 We'd like to use Theorem~\ref{thm:random-clean} to conclude that a (much) better algorithm prevails in equilibrium.
-\asedit{We consider a version of the competition game in which the principals are restricted to choosing from a given set of MAB algorithms; the algorithms in this set are called \emph{feasible}.}
+We consider a version of the competition game in which the principals are restricted to choosing from a given set of MAB algorithms; the algorithms in this set are called \emph{feasible}.
 
@@
 Fix a \HardMaxRandom response function $\respF$.
-\asedit{Consider the competition game in which all feasible MAB algorithms are $T$-oblivious, \bmonotone, and satisfy \eqref{eq:random-assn} for some fixed $m_0$.
-Suppose some feasible algorithm \alg  BIR-dominates all other feasible algorithms, starting from some local step $n_0$.} Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
+Consider the competition game in which all feasible MAB algorithms are $T$-oblivious, \bmonotone, and satisfy \eqref{eq:random-assn} for some fixed $m_0$.
+Suppose some feasible algorithm \alg  BIR-dominates all other feasible algorithms, starting from some local step $n_0$.
+Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
 \end{corollary}
@@
 \begin{corollary}\label{cor:random-greedy}
-Fix a symmetric \HardMaxRandom response function $\respF$. \asedit{Consider the competition game in which algorithms are feasible if and only if they are \bmonotone.}%
-\footnote{\asedit{More generally, this corollary holds if feasible algorithms are \bmonotone and closed under the greedy modification (\ie the greedy modification of any feasible algorithm is also feasible, for any mixing parameter $p$).}}
-\asedit{Then there are no Nash equilibria other than (\DynGreedy,\,\DynGreedy).}
+Fix a symmetric \HardMaxRandom response function $\respF$. Consider the competition game in which algorithms are feasible if and only if they are \bmonotone.%
+\footnote{More generally, this corollary holds if feasible algorithms are \bmonotone and closed under the greedy modification (\ie the greedy modification of any feasible algorithm is also feasible, for any mixing parameter $p$).}
+Then there are no Nash equilibria other than (\DynGreedy,\,\DynGreedy).
 %The latter is not a Nash equilibrium when some algorithm BIR-dominates \DynGreedy and time horizon $T$ is sufficiently large.
@@
 
-\asedit{Recall that by Theorem~\ref{thm:random-clean} \DynGreedy cannot be played in equilibrium when it is BIR-dominated by some feasible, $T$-oblivious algorithm and the time horizon $T$ is large enough.}
+Recall that by Theorem~\ref{thm:random-clean} \DynGreedy cannot be played in equilibrium when it is BIR-dominated by some feasible, $T$-oblivious algorithm and the time horizon $T$ is large enough.
 
@@
 
-\asedit{It follows that a weakly-BIR-dominating algorithm prevails in equilibrium.} We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity.
+It follows that a weakly-BIR-dominating algorithm prevails in equilibrium.
+We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity.
 \BReg is a standard notion from the literature (along with \BIR):
@@
 
-We frame our contributions in terms of the relationship between \competitiveness \asedit{(as expressed by the ``hardness" of the agents' response function $\respF$)},
+We frame our contributions in terms of the relationship between \competitiveness (as expressed by the ``hardness" of the agents' response function $\respF$),
  %and \rationality on one side, 

### 2020-11-25 journal paper/ec19paper/content/inverted_u.tex
Minor edit in a figure.
add changes

@@
 Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship \asedit{as in Section~\ref{sec:theory-welfare}}.
+As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship as in Section~\ref{sec:theory-welfare}.
 More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DynamicGreedy, \DynamicEpsGreedy and \Thompson.
@@
      \node[below] at (2.8, -0.2) {\footnotesize incumbent};
-     \node[below] at (4.8, -0.22) {\footnotesize sim. entry};
+     \node[below] at (4.5, -0.65) {\footnotesize simultaneous entry};
      \node[below] at (6.2, -0.2) {\footnotesize entrant};

### 2020-12-09 journal paper/content/sec-intro.tex
expand on the two footnotes

@@
 Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
-Each principal's strategy is a multi-armed bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.
-This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.
+Each principal's strategy is a multi-armed bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.\footnote{\gaedit{In particular, the algorithm does not respond to neither the competitors' rewards (which it does not observe) nor its current market share.}} This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.
 %Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
-We draw on the machine learning literature to compare bandit algorithms to one another in isolation,%
-\footnote{As opposed to ``in competition": as solutions to a stand-alone exploration problem which aims to maximize the total consumer welfare over time (see Appendix~\ref{app:bg} for self-contained background).}
- so that we can talk about \emph{better} bandit algorithms in a principled way.
+We draw on the machine learning literature to compare bandit algorithms to one another in isolation,
+ so that we can talk about \emph{better} bandit algorithms in a principled way.\footnote{\gaedit{This literature typically compares the performance of different algorithms in a stand-alone exploration problem according to their asymptotic regret, which can be interpreted as maximizing consumer welfare in our context (see Appendix~\ref{app:bg} for self-contained background).
+Thus, we can utilize this comparison measure to assess the quality of the algorithms adopted under competition.}}
 %\footnote{Such comparisons are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.

### 2020-12-20 journal paper/content/sec-bg.tex
Removed some coloring, and a \newpage.
Minor edits re student's comments.

@@
 \item \emph{Greedy algorithms} that strive to maximize the reward for the next round given the available information. Thus, they always ``exploit" and never explicitly ``explore".
-    \asmargincomment{copied the list from the intro.}
+    %\asmargincomment{copied the list from the intro.}
 
@@
 \end{itemize}
+
+\asedit{Below we discuss which algorithms are better than others, and what does it \emph{mean} for one bandit algorithm to be better than another.
+This is a rather subtle issue, because some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
+In particular, ``better" algorithms are better in the long run, but could be worse initially.}
 

### 2020-12-20 journal paper/content/sec-intro.tex
Minor edits re student's comments.

--- 2020-11-25 journal paper/content/sec-intro.tex
@@
 Platforms that engage in exploration typically need to compete against one another. Most importantly, platforms compete for users, who benefit them in two ways:
-generate revenue and provide data for learning. This creates a tension:
+generating revenue and providing data for learning. This creates a tension:
 %between \exploration and \competition.
-while exploration may be essential for improving the service tomorrow, it may degrade quality \emph{today}, in which case some of the users can leave and there will be less users to learn from.
+while exploration may be essential for improving the service tomorrow, it may degrade quality \emph{today}, in which case some of the users can leave and there will be fewer users to learn from.
 This may create a ``data feedback loop" when the platform's performance further degrades relative to competitors who keep learning and improving from \emph{their} users, and so forth.
@@
 %\begin{align}\label{eq:main-Q}
-\textbf{How does competition incentivize the adoption of better exploration algorithms?}
+\textbf{Whether and how does competition incentivize the adoption of better exploration algorithms?}
 %\end{align}
@@
 Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
-Each principal's strategy is a multi-armed bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the competition.
+Each principal's strategy is a multi-armed bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the \asedit{competitor's choices, rewards, or market share}.
+%\footnote{\gaedit{In particular, the algorithm does not respond to neither the competitors' rewards (which it does not observe) nor its current market share.}} 
 This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.
 %Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
-We draw on the machine learning literature to compare bandit algorithms to one another in isolation,%
-\footnote{As opposed to ``in competition": as solutions to a stand-alone exploration problem which aims to maximize the total consumer welfare over time (see Appendix~\ref{app:bg} for self-contained background).}
- so that we can talk about \emph{better} bandit algorithms in a principled way.
+
+\asedit{We draw on the machine learning literature to tell us how to compare bandit algorithms to one another, so that we can talk about \emph{better} bandit algorithms in a principled way.
+In particular, one considers algorithm's performance \emph{in isolation}: in a standalone exploration problem without competition.
+We focus on big, qualitative differences in algorithms' performance, see Appendix~\ref{app:bg} for background.}
+%We draw on the machine learning literature to compare bandit algorithms to one another in isolation, so that we can talk about \emph{better} bandit algorithms in a principled way.\footnote{\gaedit{This literature typically compares the performance of different algorithms in a stand-alone exploration problem according to their asymptotic regret, which can be interpreted as maximizing consumer welfare in our context (see Appendix~\ref{app:bg} for self-contained background).
+Thus, we can utilize this comparison measure to assess the quality of the algorithms adopted under competition.}}
 %\footnote{Such comparisons are somewhat subtle, as some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
@@
 \footnote{Alternatively, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}
-In the most technical result of the paper, we find that a better algorithm prevails under much weaker assumptions.
+In the most technical result of the paper, we find that a \asedit{sufficiently} better algorithm prevails under much weaker assumptions.
 % on what constitutes a better algorithm.

### 2020-12-20 journal paper/journal_main.tex
Changed the date, removed a \newpage.
Added a \newpage elsewhere!
Updated the date.
Revised the title note.

@@
 The Perils of Exploration under Competition%
-\thanks{This is a merged and final version of two conference papers,
+\thanks{This is a merged and final version of two conference publications,
 \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19},
@@
 All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
-Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.
-We would like to thank Ian Ball, Yeon-Koo Che, Glen Weyl and Sven Rady, as well as seminar participants at Columbia and conference participants at Innovations in Theoretical Computer Science 2018, ACM Economics and Computation 2019 and the MIT Conference on Digital Experimentation 2020, for helpful comments and conversations.
-All errors are our own.} }
+Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.\newline \indent
+We would like to thank Ian Ball, Yeon-Koo Che, Glen Weyl and Sven Rady for helpful comments and conversations. Likewise, we thank the audience of the conference and seminar talks on this work.
+%seminar participants at Columbia and conference participants at Innovations in Theoretical Computer Science 2018, ACM Economics and Computation 2019 and the MIT Conference on Digital Experimentation 2020, for helpful comments and conversations.
+All errors are our own.}}
 % \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
@@
 }
-\date{First version: July 2020\\ This version: September 2020}
+\date{First version: July 2020\\ This version: December 2020}
 
@@
 
-\newpage
+%\newpage
 \section{Theoretical results: the \TheoryModel}

### 2020-12-20 journal paper/content/related_work.tex
Minor edit to give credit to Che-Horner.

@@
 
-The three-way tradeoff between exploration, exploitation and incentives has been studied in mechanism design, in several scenarios very different from ours:
-incentivizing exploration in a recommendation system
-(starting from \citet{Kremer-JPE14}, see \citet[Ch. 11.6]{slivkins-MABbook} for a literature review),
+The three-way tradeoff between exploration, exploitation and incentives has been studied in (algorithmic) mechanism design, in several scenarios very different from ours. In particular: 
+incentivized exploration in a recommendation system
+(starting from \citet{Kremer-JPE14,Che-13}, see \citet[Ch. 11.6]{slivkins-MABbook} for a literature review),
     %\citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},

### 2020-12-20 journal paper/content/sec-intro.tex
Minor edits.

@@
 %between \exploration and \competition.
-while exploration may be essential for improving the service tomorrow, it may degrade quality \emph{today}, in which case some of the users can leave and there will be fewer users to learn from.
+while exploration may be essential for improving the service tomorrow, it may degrade the service quality \emph{today}, in which case some of the users can leave and there will be fewer users to learn from.
 This may create a ``data feedback loop" when the platform's performance further degrades relative to competitors who keep learning and improving from \emph{their} users, and so forth.
@@
 Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
-Each principal's strategy is a multi-armed bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the \asedit{competitor's choices, rewards, or market share}.
+Each principal's strategy is a multi-armed bandit algorithm.
+As such, it can dynamically adjust itself to the observed rewards, but not to \asedit{any aspect of competition (incl.
+the market share and the competitor's choices or rewards).}
 %\footnote{\gaedit{In particular, the algorithm does not respond to neither the competitors' rewards (which it does not observe) nor its current market share.}} 

### 2020-12-20 journal paper/journal_main.tex
Minor edits.

@@
 Most online platforms strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-We initiate a study of the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users.
+We \asedit{study} the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users.
 Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
 
-We consider a stylized duopoly model in which two firms face the same problem instance of multi-armed bandits.
+We consider a stylized duopoly model in which two firms face \asedit{the same multi-armed bandit problem}.
 Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen.

### 2020-12-21 journal paper/content/related_work.tex
A pass addressing YK comments.

--- 2020-11-25 journal paper/content/related_work.tex
@@
 
-The three-way tradeoff between exploration, exploitation and incentives has been studied in mechanism design, in several scenarios very different from ours:
-incentivizing exploration in a recommendation system
-(starting from \citet{Kremer-JPE14}, see \citet[Ch. 11.6]{slivkins-MABbook} for a literature review),
+The three-way tradeoff between exploration, exploitation and incentives has been studied in several scenarios very different from ours:
+incentivized exploration in recommendation systems 
+(starting from \citet{Kremer-JPE14,Che-13}, see \citet[Ch. 11.6]{slivkins-MABbook} for a literature review),
     %\citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},
@@
 \xhdr{Exploration and competition.}
-\citet{bergemann1997market,bergemann2000experimentation,keller2003price} studied the interplay of exploration and competition for users when the competing firms experiment with \emph{prices}, rather than design alternatives.
-All three papers consider strategies that respond to competition, and analyze Markov Perfect Equilibria (MPE), whereas we focus on the adoption of better bandit algorithms.
+\citet{bergemann1997market,bergemann2000experimentation,keller2003price} study the interplay of exploration and competition for users when the competing firms experiment with \emph{prices}
+\asedit{(whereas in our model they experiment with design alternatives).} 
+All three papers consider strategies that respond to competition and analyze Markov-perfect equilibria (whereas we focus on the adoption of better bandit algorithms).
 We discuss them in more detail below.
@@
 
-The literature on learning-by-doing vs.
-competition \citep[\eg][]{fudenberg1983learning, dasgupta1988learning, cabral1994learning} studies firms that learn while competing against each other, so that a firm attracting more consumers reduces its production costs.
+The literature on learning-by-doing vs. competition \citep[\eg][]{fudenberg1983learning, dasgupta1988learning, cabral1994learning} 
+\asmargincomment{Guy: pls cite Gilbert-Newberry, as per YK.}
+studies firms that learn while competing against each other, so that a firm attracting more consumers reduces its production costs.
 Our model differs in several important respects.
@@
 %as we move from \HardMax to \HardMaxRandom.
-Second, there a large literature on non-existence of equilibria due to small deviations, starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets.\footnote{\cite{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom in this context.}
+Second, there a large literature on non-existence of equilibria due to small deviations, starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets.
+\asdelete{\footnote{\cite{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom in this context.}}
 This is superficially similar to how small deviations towards the greedy algorithm rule out equilibria under \HardMaxRandom.

### 2020-12-21 journal paper/content/sec-intro.tex
A pass addressing YK comments.

--- 2020-12-20 journal paper/content/sec-intro.tex
@@
 %between \exploration and \competition.
-while exploration may be essential for improving the service tomorrow, it may degrade quality \emph{today}, in which case some of the users can leave and there will be fewer users to learn from.
+while exploration may be essential for improving the service tomorrow, it may degrade the \asedit{service} quality \emph{today}, in which case some of the users can leave and there will be fewer users to learn from.
 This may create a ``data feedback loop" when the platform's performance further degrades relative to competitors who keep learning and improving from \emph{their} users, and so forth.
@@
 A new agent arrives and chooses a principal according to some decision rule.
-The principal selects an action (\eg a list of web search results), the user experiences it and reports a reward.
+The principal selects an action \asedit{which affects the quality of service provided to this agent} (\eg a list of web search results).
+The agent experiences \asedit{this action} and reports a reward.
 Each principal only observes its own users.
-Principals commit to their strategies in advance, with the objective to maximize their market share.
-We investigate several model variants, %within this framework,
-where we vary agents' decision rule or allow a first-mover advantage.%
-\footnote{In particular, we use different model variants (resp., Bayesian and frequentist) for theory and for simulations so as to ensure tractability. Our main findings are similar for both.}
+Principals commit to their strategies in advance, so as to maximize their market share.
 
@@
 Each principal faces a basic and well-studied version of the multi-armed bandit problem, where each reward is drawn independently from a fixed, action-specific distribution.
-Each principal's strategy is a multi-armed bandit algorithm; as such, it can dynamically adjust itself to the observed rewards, but not to the \asedit{competitor's choices, rewards, or market share}.
+Each principal's strategy is a multi-armed bandit algorithm,
+\asedit{which dynamically adjusts to the observed rewards.}%
+\footnote{\asedit{We do not distinguish between pure and mixed strategies, because bandit algorithms can be randomized, and a distribution over bandit algorithms is also a bandit algorithm.}}
+\asedit{However, it is oblivious to all signals on competition (such as the market share or the competitor's choices or rewards), even when such signals are available.}
 %\footnote{\gaedit{In particular, the algorithm does not respond to neither the competitors' rewards (which it does not observe) nor its current market share.}} 
-This modeling choice reflects the reality of industrial applications, which build on a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}.
+This modeling choice reflects the reality of industrial applications, which follow a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}. 
+\asedit{Due to similar practical considerations, we expect the actual strategy choice to reflect the competition game only in a crude, qualitative way.
+Hence,  basic outcomes under competition are worth studying \emph{per se}, not only as a stepping stone to equilibrium characterization.}
 %Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
 
-\asedit{We draw on the machine learning literature to tell us how to compare bandit algorithms to one another, so that we can talk about \emph{better} bandit algorithms in a principled way.
+\asedit{To flesh out the meaning of \emph{better} exploration algorithms, as per the main question, we draw on the literature from machine learning which studies such algorithms.
 In particular, one considers algorithm's performance \emph{in isolation}: in a standalone exploration problem without competition.
-We focus on big, qualitative differences in algorithms' performance, see Appendix~\ref{app:bg} for background.}
+We focus on big, qualitative differences between algorithms that are well-established in the literature, see Appendix~\ref{app:bg} for self-contained background.
 %We draw on the machine learning literature to compare bandit algorithms to one another in isolation, so that we can talk about \emph{better} bandit algorithms in a principled way.\footnote{\gaedit{This literature typically compares the performance of different algorithms in a stand-alone exploration problem according to their asymptotic regret, which can be interpreted as maximizing consumer welfare in our context (see Appendix~\ref{app:bg} for self-contained background).
@@
 }
-One baseline is algorithms that do not explicitly explore (\emph{greedy algorithms}); in isolation, they are known to perform poorly for a wide variety of problem instances.
+One baseline is algorithms that do not purposefully explore, and instead strive to make myopically optimal decisions; we call them \emph{greedy algorithms}.
+In isolation, they are known to perform poorly for a wide variety of problem instances.}
+
+We investigate several model variants, where we vary agents' decision rule or allow a first-mover advantage.
+We use different model variants for theory and for simulations so as to ensure tractability; our main findings are similar for both.
 
@@
 %We endow agents with Bayesian rationality, a common modeling approach for a theoretical investigation.
-We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms, but do not receive any information about the previous agents' choices or rewards.
-Each agent computes the Bayesian-expected reward for each principal, and use these two numbers to decide which principal to choose.
+We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms \asedit{and their own arrival times}, but do not \asedit{observe} the previous agents' choices or rewards.
+Each agent computes the Bayesian-expected reward for each principal, and uses these two numbers to decide which principal to choose.
 %We focus on qualitative, asymptotic differences in the algorithms' performance.
@@
 \xhdr{Numerical simulations.}
-We consider a basic frequentist model (called the \emph{\ExptsModel}), where the agents observe signals about the principals' past performance, and base their decisions on these signals alone, without invoking any prior knowledge or beliefs.
+We consider a basic frequentist model (called the \emph{\ExptsModel}), where agents observe signals about the principals' past performance and 
+\asedit{make their decisions naively,}
+%base their decisions on these signals alone, 
+without invoking any prior knowledge or beliefs.
 The performance signals are abstracted and aggregated as a scalar \emph{reputation score} for each principal, modeled as a sliding window average of its rewards.
@@
 
-\textbf{(iii)} We investigate the algorithms' performance without competition.
-We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).}
-     We find that mean reputation -- arguably, the most natural performance measure -- is sometimes \emph{not} a good predictor for the outcomes under competition.
+\textbf{(iii)} 
+\asedit{However, the greedy algorithm is sometimes \emph{not} the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).}
+We revisit algorithms' performance without competition, and we find that the most natural performance measure does not explain this phenomenon.
+We suggest a new, more nuanced performance measure that does.}
+
+%\textbf{(iii)} We investigate the algorithms' performance without competition.
+We suggest a new performance measure to explain why the greedy algorithm is sometimes not the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).} We find that mean reputation -- arguably, the most natural performance measure -- is sometimes \emph{not} a good predictor for the outcomes under competition.
 
@@
 %whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
-The death spiral/amplification effects mentioned above have a similar flavor: a relatively small amount of exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved of users (resp., take over most of the market).
+The death spiral/amplification effects mentioned above have a similar flavor: a relatively small \asedit{performance loss due to} exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved of users (resp., take over most of the market).
 %\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}
@@
 but ignore the differences induced by learning in isolation versus under competition.
-Furthermore, explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.
-In particular, we find that incumbency is actually good for innovation, but also creates a barrier to entry precisely due to this feedback loop.
-% and how it is impacted by competition.
+Explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.
+In particular, we find that incumbency is good for innovation \asedit{and welfare, \emph{and}} % but also
+creates a barrier to entry, 
+\asedit{all} % precisely
+due to data feedback loops.
+
+\asedit{Interestingly, we find that data feedback loops can be bad for innovation (as in ``death spiral"), or good for innovation (as in ``amplification effect"), depending on the level of competition.}
 

### 2020-12-21 journal paper/content/sec-model.tex
A pass addressing YK comments.

@@
-\xhdr{Principals and agents.} There are two principals and $T$ agents,
-denoted, resp., principal $i\in \{1,2\}$ and agent $t\in [T]$.
-The game proceeds in (global) rounds.
+\xhdr{Principals and agents.} There are two principals and $T$ agents. We denote them, resp., principal $i\in \{1,2\}$ and agent $t\in [T]$,
+\asedit{where $[T] := \{1,2\LDOTS T\}$}.
+The game proceeds in $T$ rounds.
+
 In each round $t\in [T]$, the following  interaction takes place.
 Agent $t$ arrives and chooses a principal $i_t\in \{1,2\}$.
-The principal chooses action $a_t\in A$, where $A$ is a fixed set of actions (same for both principals and all rounds).%
+The principal chooses action $a_t\in A$, where $A$ is a fixed set of actions.%
 \footnote{We use `action' and `arm' interchangeably, as common in the literature on multi-armed bandits.}
-The agent experiences this action and receives a reward $r_t\in \{ 0,1\}$ for this action, which is then observed by the principal.
+The agent experiences this action and receives an associated reward $r_t\in \{ 0,1\}$, which is then observed by the principal.
 We posit \emph{stochastic rewards}: whenever a given action $a\in A$ is chosen, the reward is an independent draw from Bernoulli distribution with mean $\mu_a$.
-The mean rewards $\mu_a$ are fixed over time, and initially not known to anybody.
+\asedit{In particular, the mean rewards $\mu_a$, as well as the action set $A$, are the same for both principals and all rounds.} The mean rewards are initially not known to anybody.
 The principals are completely unaware of the rounds when the opponent is chosen.
 
-Thus, each principal follows the protocol of \emph{multi-armed bandits} (\emph{MAB}): in each round when it is chosen, it chooses an arm from $A$ and observes a reward for this action (and nothing else).
+% (which is the same for both principals and all rounds)
 
-Each principal $i$ commits to an MAB algorithm \alg[i] before round $1$, and uses it throughout. The algorithm proceeds in time-steps:%
-\footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before.
+Thus, each principal follows the protocol of \emph{multi-armed bandits} (henceforth, \emph{MAB}).
+Specifically, in each round when it is chosen, it chooses an arm from $A$ and observes a reward for this action.
+\asedit{It does not observe anything else, particularly rewards for other actions.}
+
+Each principal $i$ commits to an MAB algorithm \alg[i] before round $1$, and uses this algorithm throughout the game. The algorithm proceeds in time-steps:
+%\footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before.
 We will omit the global vs.
-local distinction when clear from the context.} each time it is called, it outputs an arm from $A$, and inputs a reward for this action.
-The algorothm is called only in global rounds when principal $i$ is chosen.
+local distinction when clear from the context.}
+each time it is called, it outputs an arm from $A$, and inputs a reward for this action. The algorithm is called only in game rounds when principal $i$ is chosen.
+\asedit{When the distinction between algorithm's time-steps and game rounds is unclear from the context, we will refer to them as, resp., \emph{local steps/rounds} and \emph{global rounds}.}
 
@@
 
-\xhdr{Agents' response.} Each agent $t$ forms reward estimates $\est_i(t)\in [0,1]$ for each principal $i$, and uses them to choose the principal.
-Specifically, principal $1$ is chosen with probability
+\xhdr{Agents' response.} Each agent $t$ forms reward estimates $\est_i(t)\in [0,1]$ for each principal $i$.
+\asedit{(What these estimates are, and how much the agents know in order to form them, depends on the Bayesian vs.
+frequentist model variant.)
+The reward estimates determine the choice of the principal.} Specifically, agent $t$ chooses principal $1$ with probability
 \begin{align}\label{eq:model-f}
@@
 \HardMaxRandom and \SoftMaxRandom can be seen as agent types that, resp., give some chance to the other principal and avoid sharp transitions in their probabilistic response. 
-Alternatively, they can be realized as mixtures over more ``basic" agent types.
-Indeed, \HardMaxRandom is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
+\asmargincomment{I don't get YK's beef with this.}
+Alternatively, they can be realized as a \asedit{distribution} over more ``basic" agent types.
+Indeed, \HardMaxRandom \asedit{distribution} is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
 The latter can be interpreted as consumers oblivious to principals' reputation, due to the lack of awareness or interest.
@@
 \xhdr{Bayesian vs. frequentist.}
-We consider two model variants, Bayesian and frequentist. The main difference between the two concerns the agents' reward estimates $\est_i(t)$.
+We consider two model variants, Bayesian and frequentist \asedit{(which we use, resp., for theoretical results and numerical simulations)}.
+The main difference between the two concerns the agents' reward estimates $\est_i(t)$.
 
@@
 
-\item In the \emph{\ExptsModel}, agents' reward estimate for a given principal is the average reward of the last $M$ agents that chose this principal, interpreted as a current \emph{reputation score}.
+\item In the \emph{\ExptsModel}, agents' reward estimate for a given principal is the average reward of the last $M$ agents that chose this principal.
+We call it \emph{reputation score}, and interpret it as the current reputation.
 To make it meaningful initially, each principal enjoys a ``warm start": additional $T_0$ agents arrive before the game starts, and interact with the principal as described above.
@@
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-Pure strategies in this game correspond to MAB algorithms.%
-\footnote{Note that a mixed strategy is also a (randomized) MAB algorithm, \ie also a pure strategy.}
+\asedit{(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.)}
  Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral and aim to maximize their expected utility.
@@
 
+\asedit{Now that we laid out our model, let us discuss several aspects thereof that are unclear from the bare definitions. In particular, we flesh out several points made in the Introduction.}
+
 Reflecting reality, we posit that the exploration strategies available to the principals are ``non-strategic" in nature.
@@
 This is how industry approaches exploration algorithms, and for several good reasons.
-"Non-strategic" exploration is well-studied in machine learning, and yet it remains a very complex subject in research (as evidenced, \eg by the huge amount of activity therein).
+"Non-strategic" exploration is well-studied in machine learning, and yet it remains a very complex and actively studied subject in research.
 Even the seemingly simple algorithms are not straightforward to deploy in practice, and require a substantial investment in infrastructure \cite[\eg see the discussions in][]{DS-arxiv}.

### 2020-12-21 journal paper/content/sec-theory.tex
A pass addressing YK comments.

@@
 
-\xhdr{Notation.} Let $\rew_i(n)$ denote the realized reward of principal $i$ at local step $n$.
-For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen.
+\xhdr{Notation.} Let $\rew_i(n)$ denote the \asedit{agent's realized reward observed by} principal $i$ at local step $n$.
+\asedit{Put differently, it is the reward collected by algorithm $\alg[i]$ in this local step.} For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen.
 Note that
@@
 \xhdr{MAB algorithms.}
-We consider two (Bayesian) greedy algorithms.
+\asedit{We consider two baseline algorithms.
+The main one,} called  \DynGreedy, chooses an arm $a$ with the largest posterior mean reward $\E[\mu_a \mid \cdot ]$ given all information currently available to the algorithm.
+\asedit{A more primitive baseline,} called \StaticGreedy, chooses an arm $a$ with the largest prior mean reward $\E[\mu_a]$, and uses this arm in all rounds.
+
+%We consider two (Bayesian) greedy algorithms.
 The first one, called  \DynGreedy, chooses an arm $a$ with the largest posterior mean reward $\E[\mu_a \mid \cdot ]$ given all information currently available to the algorithm.
@@
 \begin{definition}
-Two MAB algorithms \emph{deviate} at (local) step $n$ if there is a set $H$ of histories over the previous local steps such that both algorithms lead to $H$ with positive probability, and choose different distributions over arms given any history $h\in H$.
-The two algorithms deviate \emph{starting from} step $n_0$ if $n=n_0$ is the smallest step when they deviate.
+\asedit{One MAB algorithm \emph{deviates} from another} at (local) step $n$ if there is a set $H$ of histories over the previous local steps such that both algorithms lead to $H$ with positive probability, and choose different distributions over arms given any history $h\in H$.
+\asedit{We say $\alg$ deviates from $\alg'$} \emph{starting from} step $n_0$ if $n=n_0$ is the smallest step \asedit{when this happens.}
 \end{definition}
@@
 \end{lemma}
+
+\asedit{The sudden-death property in Lemma~\ref{lm:DG-sudden} holds because principal $1$ appears at least as good or better to the next agent (by the Bayesian-monotonicity property), whereas principal $2$ appears the same as before.
+Principal $2$ needs new data in order to improve, and it does not receive new data unless the response function is randomized.}
 
@@
 
+\asedit{Note that we make no positive prediction if one MAB algorithm is only slightly better than the other. Next we show that such predictions are essentially impossible.}
+
 %\subsection{A little greedy goes a long way}
@@
 
-Let us analyze the \BIR of the greedy modification (we will use it for what follows).
+\asedit{We argue below that \alg[1] can be a better algorithm than \alg[2]. This is consistent with Theorem~\ref{thm:random-clean}, since the BIR-dominance condition does not hold.}
+
+Let us analyze the \BIR of the greedy modification (we will use it twice afterwards).
 Use \alg[1] and \alg[2] as in the theorem.
@@
 
-First, let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2].
+Now, let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2].
 We need to assume that \alg[1] is (much) better than the greedy step \alggr if the latter is computed on a fraction of the data.
@@
 
-Second, let us argue that \alg[2] is \bmonotone.
+\asedit{We also use \refeq{eq:BIR-modification} to argue that} \alg[2] is \bmonotone.
 This follows from \eqref{eq:BIR-modification} since both \alg[1] and \alggr are \bmonotone.
-The latter follows from a ``monotonicity-in-information" property of the ``greedy step": essentially, it can only get better with more information
-(see Lemma~\ref{lm:MII}).
+The latter follows from a ``monotonicity-in-information" property of the ``greedy step": it can only get better with more information, see Lemma~\ref{lm:MII}.
 
@@
 \begin{corollary}\label{cor:random-greedy}
-Fix a symmetric \HardMaxRandom response function $\respF$. Consider the competition game in which algorithms are feasible if and only if they are \bmonotone.%
-\footnote{More generally, this corollary holds if feasible algorithms are \bmonotone and closed under the greedy modification (\ie the greedy modification of any feasible algorithm is also feasible, for any mixing parameter $p$).}
+Fix a symmetric \HardMaxRandom response function $\respF$. Consider the competition game in which algorithms are feasible if and only if they are \bmonotone.
 Then there are no Nash equilibria other than (\DynGreedy,\,\DynGreedy).
@@
 Recall that by Theorem~\ref{thm:random-clean} \DynGreedy cannot be played in equilibrium when it is BIR-dominated by some feasible, $T$-oblivious algorithm and the time horizon $T$ is large enough.
+\asedit{Also recall that pure and mixed Nash equilibria are the same in our model.}
+
+\asedit{Corollary~\ref{cor:random-greedy} holds more generally: it suffices to assume that feasible algorithms are \bmonotone and closed under the greedy modification.
+The latter means that the greedy modification of any feasible algorithm is also feasible, for any mixing parameter $p$.}
 
@@
 Competitiveness within this model are controlled by the baseline probability $\eps_0 = \respF(- 1)$, which goes smoothly between the two extremes of \HardMax ($\eps_0=0$, tough competition) and the uniform choice ($\eps_0=\tfrac12$, no competition).
-For clarity, we assume that principal's utility is the number of agents.
+\asedit{Recall} % For clarity, we assume 
+that principal's utility is the number of agents.
 

### 2020-12-21 journal paper/content/sim_details.tex
A pass addressing YK comments.

@@
 
-\xhdr{Challenges.} An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given instance of a multi-armed bandit problem (and each such experiment is run many times to reduce variance).
-Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit instances\}, and we are looking for findings that are consistent across this entire space.
+\xhdr{Challenges.} An ``atomic experiment" is a competition game between a given pair of bandit algorithms, in a given competition model, on a given 
+%instance of a 
+multi-armed bandit problem (and each such experiment is run many times to reduce variance).
+Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit \asedit{problems}\}, and we are looking for findings that are consistent across this entire space.
 It is essential to keep each of the three dimensions small yet representative.
@@
 \xhdr{MAB instances.}
-We consider instances with $K=10$ arms.
-Since we focus on 0-1 rewards, an instance of the MAB problem is specified by the \emph{\MRV} $(\mu(a):\; a\in A)$.
-Initially this vector is drawn from some distribution, termed \emph{MAB instance}.
+We consider \asedit{bandit problems} with $K=10$ arms and Bernoulli rewards.
+\asedit{The \emph{\MRV} $(\mu(a):\; a\in A)$ is initially} drawn from some distribution, termed \emph{MAB instance}.
 We consider three MAB instances:

### 2020-12-21 journal paper/content/related_work.tex
pass through alex's edits

@@
 \citet{bergemann1997market,bergemann2000experimentation,keller2003price} study the interplay of exploration and competition for users when the competing firms experiment with \emph{prices}
-\asedit{(whereas in our model they experiment with design alternatives).} 
+(whereas in our model they experiment with design alternatives).
 All three papers consider strategies that respond to competition and analyze Markov-perfect equilibria (whereas we focus on the adoption of better bandit algorithms).
@@
 Second, the inputs to agents' decision rule are prices, rather than the quality of the chosen alternatives; the former is directly controlled by the firms, whereas the latter is not known a priori.
-Third, the goal of exploration is to learn the agents' decision rule.
+Third, the goal of exploration is to learn the agents' decision rule, \gaedit{not the best product alternative}.
 
@@
 Specifically, the ``entrant" firm offers a new product of unknown, exogenously determined, quality, which is ``explored" if and only if the entrant attracts customers.
-However, because the entrant can only control the prices but not the product itself, the nature of exploration is fundamentally different.
+However, because the entrant can only control the prices but not the product itself, the nature of exploration is fundamentally different \gaedit{than in our setting}.
 First, the goal of exploration is to reveal information about a fixed product.
@@
 
-The literature on learning-by-doing vs. competition \citep[\eg][]{fudenberg1983learning, dasgupta1988learning, cabral1994learning} 
+The literature on learning-by-doing vs. competition \citep[\eg][]{gilbert1982preemptive, fudenberg1983learning, dasgupta1988learning, cabral1994learning} 
 \asmargincomment{Guy: pls cite Gilbert-Newberry, as per YK.}
@@
 %as we move from \HardMax to \HardMaxRandom.
-Second, there a large literature on non-existence of equilibria due to small deviations, starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets.
+\gadelete{Second, there a large literature on non-existence of equilibria due to small deviations, starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets.
 \asdelete{\footnote{\cite{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom in this context.}}
-This is superficially similar to how small deviations towards the greedy algorithm rule out equilibria under \HardMaxRandom.
+This is superficially similar to how small deviations towards the greedy algorithm rule out equilibria under \HardMaxRandom.}
 %Second, the stochastic choice rules we consider have been widely utilized across economics. For instance,
-Third, \SoftMaxRandom subsumes the logit choice rule, a standard behavioral model with strong empirical and microeconomic foundations
+Second, \SoftMaxRandom subsumes the logit choice rule, a standard behavioral model with strong empirical and microeconomic foundations
 \citep[\eg][]{mosteller1951experimental, luce1959choice, mcfadden1974measurement, matvejka2015rational}.
-Fourth, choice models similar to \SoftMaxRandom are used to explain  horizontal product differentiation \citep[\eg][]{Hotelling-29, Perloff-Salop-85}.
+Third, choice models similar to \SoftMaxRandom are used to explain  horizontal product differentiation \citep[\eg][]{Hotelling-29, Perloff-Salop-85}.
 %\asdelete{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}

### 2020-12-21 journal paper/content/sec-bg.tex
pass through alex's edits

@@
 
-\asedit{Below we discuss which algorithms are better than others, and what does it \emph{mean} for one bandit algorithm to be better than another.
+Below we discuss which algorithms are better than others, and what does it \emph{mean} for one bandit algorithm to be better than another.
 This is a rather subtle issue, because some algorithms may be better for some problem instances and/or time intervals, and worse for some others.
-In particular, ``better" algorithms are better in the long run, but could be worse initially.}
+In particular, ``better" algorithms are better in the long run, but could be worse initially.
 

### 2020-12-21 journal paper/content/sec-intro.tex
pass through alex's edits

@@
 %between \exploration and \competition.
-while exploration may be essential for improving the service tomorrow, it may degrade the \asedit{service} quality \emph{today}, in which case some of the users can leave and there will be fewer users to learn from.
+while exploration may be essential for improving the service tomorrow, it may degrade the service quality \emph{today}, in which case some of the users can leave and there will be fewer users to learn from.
 This may create a ``data feedback loop" when the platform's performance further degrades relative to competitors who keep learning and improving from \emph{their} users, and so forth.
@@
 Principals compete on quality rather than on prices, and engage in exploration in order to learn which actions lead to high quality products.
-A new agent arrives and chooses a principal according to some decision rule.
-The principal selects an action \asedit{which affects the quality of service provided to this agent} (\eg a list of web search results).
-The agent experiences \asedit{this action} and reports a reward.
-Each principal only observes its own users.
+A new agent arrives and chooses a principal according to some decision rule \gaedit{which depends on the principal's exploration strategy}.
+The principal selects an action which affects the quality of service provided to this agent (\eg a list of web search results).
+The agent experiences this action \gaedit{and the resulting utility from this action is observed by the principal.} \gadelete{and reports a reward.} Each principal only observes its own users.
 Principals commit to their strategies in advance, so as to maximize their market share.
@@
 Each principal's strategy is a multi-armed bandit algorithm,
-\asedit{which dynamically adjusts to the observed rewards.}%
-\footnote{\asedit{We do not distinguish between pure and mixed strategies, because bandit algorithms can be randomized, and a distribution over bandit algorithms is also a bandit algorithm.}}
-\asedit{However, it is oblivious to all signals on competition (such as the market share or the competitor's choices or rewards), even when such signals are available.}
+which dynamically adjusts to the observed rewards.%
+\footnote{We do not distinguish between pure and mixed strategies, because bandit algorithms can be randomized, and a distribution over bandit algorithms is also a bandit algorithm.}
+However, it is oblivious to all signals on competition (such as the market share or the competitor's choices or rewards), even when such signals are available.
 %\footnote{\gaedit{In particular, the algorithm does not respond to neither the competitors' rewards (which it does not observe) nor its current market share.}} 
 This modeling choice reflects the reality of industrial applications, which follow a huge body of knowledge in machine learning; more on this in Section~\ref{sec:discussion}. 
-\asedit{Due to similar practical considerations, we expect the actual strategy choice to reflect the competition game only in a crude, qualitative way.
-Hence,  basic outcomes under competition are worth studying \emph{per se}, not only as a stepping stone to equilibrium characterization.}
+Due to similar practical considerations, we expect the actual strategy choice to reflect the competition game only in a crude, qualitative way.
+Hence,  basic outcomes under competition are worth studying \emph{per se}, not only as a stepping stone to equilibrium characterization.
 %Responding to the competition is another layer of complexity which has not been previously studied in this context, let alone made practical.
 
-\asedit{To flesh out the meaning of \emph{better} exploration algorithms, as per the main question, we draw on the literature from machine learning which studies such algorithms.
+To flesh out the meaning of \emph{better} exploration algorithms, as per the main question, we draw on the literature from machine learning which studies such algorithms.
 In particular, one considers algorithm's performance \emph{in isolation}: in a standalone exploration problem without competition.
@@
 One baseline is algorithms that do not purposefully explore, and instead strive to make myopically optimal decisions; we call them \emph{greedy algorithms}.
-In isolation, they are known to perform poorly for a wide variety of problem instances.}
+In isolation, they are known to perform poorly for a wide variety of problem instances.
 
@@
 %We endow agents with Bayesian rationality, a common modeling approach for a theoretical investigation.
-We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms \asedit{and their own arrival times}, but do not \asedit{observe} the previous agents' choices or rewards.
+We consider a basic Bayesian model (called the \emph{\TheoryModel}), where agents have a common Bayesian prior on reward distributions, know the principals' algorithms and their own arrival times, but do not observe the previous agents' choices or rewards.
 Each agent computes the Bayesian-expected reward for each principal, and uses these two numbers to decide which principal to choose.
@@
 \footnote{Alternatively, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}
-In the most technical result of the paper, we find that a \asedit{sufficiently} better algorithm prevails under much weaker assumptions.
+In the most technical result of the paper, we find that a sufficiently better algorithm prevails under much weaker assumptions.
 % on what constitutes a better algorithm.
@@
 We consider a basic frequentist model (called the \emph{\ExptsModel}), where agents observe signals about the principals' past performance and 
-\asedit{make their decisions naively,}
+make their decisions naively,
 %base their decisions on these signals alone, 
@@
 \textbf{(iii)} 
-\asedit{However, the greedy algorithm is sometimes \emph{not} the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).}
-We revisit algorithms' performance without competition, and we find that the most natural performance measure does not explain this phenomenon.
-We suggest a new, more nuanced performance measure that does.}
+However, the greedy algorithm is sometimes \emph{not} the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).}
+We revisit algorithms' performance \gaedit{in a standalone bandit problem, \ie without competition, and} we find that the most natural performance measure does not explain this phenomenon.
+We suggest a new, more nuanced performance measure that does.
 
@@
 %whereby, when these effects are present, in many cases only one firm can function in the market, leading to competition \emph{for} the market being more important than competition \emph{in} the market \citep{Rysman09, jullien2019economics}.
-The death spiral/amplification effects mentioned above have a similar flavor: a relatively small \asedit{performance loss due to} exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved of users (resp., take over most of the market).
+The death spiral/amplification effects mentioned above have a similar flavor: a relatively small performance loss due to exploration (resp., data advantage)  gets amplified under competition and causes the firm to be starved of users (resp., take over most of the market).
 %\gaedit{We further find that a small data advantage for one firm gets amplified under competition and leads to that firm taking the entire market, showing that data can provide a similar incumbency advantage as those provided by traditional network effects and can serve as a barrier to entry in online markets.}
@@
 Explicitly modeling the interaction between learning technology and data creation allows us to speak on how data advantages are characterized and amplified by the increased \emph{quality} of data gathered by better learning algorithms, not just the quantity thereof.
-In particular, we find that incumbency is good for innovation \asedit{and welfare, \emph{and}} % but also
+In particular, we find that incumbency is good for innovation and welfare, \emph{and} % but also
 creates a barrier to entry, 
-\asedit{all} % precisely
+all % precisely
 due to data feedback loops.
 
-\asedit{Interestingly, we find that data feedback loops can be bad for innovation (as in ``death spiral"), or good for innovation (as in ``amplification effect"), depending on the level of competition.}
+Interestingly, we find that data feedback loops can be bad for innovation (as in ``death spiral"), or good for innovation (as in ``amplification effect"), depending on the level of competition.
 

### 2020-12-21 journal paper/content/sec-model.tex
pass through alex's edits

@@
 \xhdr{Principals and agents.} There are two principals and $T$ agents. We denote them, resp., principal $i\in \{1,2\}$ and agent $t\in [T]$,
-\asedit{where $[T] := \{1,2\LDOTS T\}$}.
+where $[T] := \{1,2\LDOTS T\}$.
 The game proceeds in $T$ rounds.
@@
 We posit \emph{stochastic rewards}: whenever a given action $a\in A$ is chosen, the reward is an independent draw from Bernoulli distribution with mean $\mu_a$.
-\asedit{In particular, the mean rewards $\mu_a$, as well as the action set $A$, are the same for both principals and all rounds.} The mean rewards are initially not known to anybody.
+In particular, the mean rewards $\mu_a$, as well as the action set $A$, are the same for both principals and all rounds.
+The mean rewards are initially not known to anybody.
 The principals are completely unaware of the rounds when the opponent is chosen.
@@
 Specifically, in each round when it is chosen, it chooses an arm from $A$ and observes a reward for this action.
-\asedit{It does not observe anything else, particularly rewards for other actions.}
+It does not observe anything else, particularly rewards for other actions.
 
@@
 each time it is called, it outputs an arm from $A$, and inputs a reward for this action. The algorithm is called only in game rounds when principal $i$ is chosen.
-\asedit{When the distinction between algorithm's time-steps and game rounds is unclear from the context, we will refer to them as, resp., \emph{local steps/rounds} and \emph{global rounds}.}
+When the distinction between algorithm's time-steps and game rounds is unclear from the context, we will refer to them as, resp., \emph{local steps/rounds} and \emph{global rounds}.
 
@@
 \xhdr{Agents' response.} Each agent $t$ forms reward estimates $\est_i(t)\in [0,1]$ for each principal $i$.
-\asedit{(What these estimates are, and how much the agents know in order to form them, depends on the Bayesian vs.
+(What these estimates are, and how much the agents know in order to form them, depends on the Bayesian vs.
 frequentist model variant.)
-The reward estimates determine the choice of the principal.} Specifically, agent $t$ chooses principal $1$ with probability
+The reward estimates determine the choice of the principal. Specifically, agent $t$ chooses principal $1$ with probability
 \begin{align}\label{eq:model-f}
@@
 
-\HardMaxRandom and \SoftMaxRandom can be seen as agent types that, resp., give some chance to the other principal and avoid sharp transitions in their probabilistic response. 
+\HardMaxRandom and \SoftMaxRandom can be seen as agent types that\gadelete{, resp., give some chance to the other principal and avoid sharp transitions in their probabilistic response.} \gaedit{make mistakes due to being oblivious to the principals' reputation, due to lack of awareness or interest.
+\HardMaxRandom models users as either being aware of the reputation or not, whereas \SoftMaxRandom models user mistakes as following a smooth density.}
 \asmargincomment{I don't get YK's beef with this.}
-Alternatively, they can be realized as a \asedit{distribution} over more ``basic" agent types.
-Indeed, \HardMaxRandom \asedit{distribution} is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
-The latter can be interpreted as consumers oblivious to principals' reputation, due to the lack of awareness or interest.
-One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
+Alternatively, they can be realized as a distribution over more ``basic" agent types.
+Indeed, \gaedit{the} \HardMaxRandom distribution is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
+\gadelete{The latter can be interpreted as consumers oblivious to principals' reputation, due to the lack of awareness or interest.} One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
 Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.
@@
 \xhdr{Bayesian vs. frequentist.}
-We consider two model variants, Bayesian and frequentist \asedit{(which we use, resp., for theoretical results and numerical simulations)}.
+We consider two model variants, Bayesian and frequentist (which we use, resp., for theoretical results and numerical simulations).
 The main difference between the two concerns the agents' reward estimates $\est_i(t)$.
@@
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-\asedit{(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.)}
- Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral and aim to maximize their expected utility.
+\gaedit{Principals are risk-neutral and their utility is defined as their market share, \ie the number of agents that chose this principal, and thus they aim to select the algorithm that maximizes their expected utility.
+(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.)}
+\gadelete{Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral and aim to maximize their expected utility.}
 
@@
 
-\asedit{Now that we laid out our model, let us discuss several aspects thereof that are unclear from the bare definitions. In particular, we flesh out several points made in the Introduction.}
+Now that we laid out our model, let us discuss several aspects thereof that are unclear from the bare definitions. In particular, we flesh out several points made in the Introduction.
 

### 2020-12-21 journal paper/content/sec-theory.tex
pass through alex's edits

@@
 
-\xhdr{Notation.} Let $\rew_i(n)$ denote the \asedit{agent's realized reward observed by} principal $i$ at local step $n$.
-\asedit{Put differently, it is the reward collected by algorithm $\alg[i]$ in this local step.} For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen.
+\xhdr{Notation.} Let $\rew_i(n)$ denote the agent's realized reward observed by principal $i$ at local step $n$.
+Put differently, it is the reward collected by algorithm $\alg[i]$ in this local step.
+For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen.
 Note that
@@
 \xhdr{MAB algorithms.}
-\asedit{We consider two baseline algorithms.
-The main one,} called  \DynGreedy, chooses an arm $a$ with the largest posterior mean reward $\E[\mu_a \mid \cdot ]$ given all information currently available to the algorithm.
-\asedit{A more primitive baseline,} called \StaticGreedy, chooses an arm $a$ with the largest prior mean reward $\E[\mu_a]$, and uses this arm in all rounds.
+We consider two baseline algorithms.
+The main one, called  \DynGreedy, chooses an arm $a$ with the largest posterior mean reward $\E[\mu_a \mid \cdot ]$ given all information currently available to the algorithm.
+A more primitive baseline, called \StaticGreedy, chooses an arm $a$ with the largest prior mean reward $\E[\mu_a]$, and uses this arm in all rounds.
 
@@
 \begin{definition}
-\asedit{One MAB algorithm \emph{deviates} from another} at (local) step $n$ if there is a set $H$ of histories over the previous local steps such that both algorithms lead to $H$ with positive probability, and choose different distributions over arms given any history $h\in H$.
-\asedit{We say $\alg$ deviates from $\alg'$} \emph{starting from} step $n_0$ if $n=n_0$ is the smallest step \asedit{when this happens.}
+One MAB algorithm \emph{deviates} from another at (local) step $n$ if there is a set $H$ of histories over the previous local steps such that both algorithms lead to $H$ with positive probability, and choose different distributions over arms given any history $h\in H$.
+We say $\alg$ deviates from $\alg'$ \emph{starting from} step $n_0$ if $n=n_0$ is the smallest step when this happens.
 \end{definition}
@@
 
-\asedit{The sudden-death property in Lemma~\ref{lm:DG-sudden} holds because principal $1$ appears at least as good or better to the next agent (by the Bayesian-monotonicity property), whereas principal $2$ appears the same as before.
-Principal $2$ needs new data in order to improve, and it does not receive new data unless the response function is randomized.}
+The sudden-death property in Lemma~\ref{lm:DG-sudden} holds because principal $1$ appears at least as good or better to the next agent (by the Bayesian-monotonicity property), whereas principal $2$ appears the same as before.
+Principal $2$ needs new data in order to improve, and it does not receive new data unless the response function is randomized.
 
@@
 
-\asedit{Note that we make no positive prediction if one MAB algorithm is only slightly better than the other. Next we show that such predictions are essentially impossible.}
+Note that we make no positive prediction if one MAB algorithm is only slightly better than the other. Next we show that such predictions are essentially impossible.
 
@@
 
-\asedit{We argue below that \alg[1] can be a better algorithm than \alg[2]. This is consistent with Theorem~\ref{thm:random-clean}, since the BIR-dominance condition does not hold.}
+We argue below that \alg[1] can be a better algorithm than \alg[2]. This is consistent with Theorem~\ref{thm:random-clean}, since the BIR-dominance condition does not hold.
 
@@
 
-\asedit{We also use \refeq{eq:BIR-modification} to argue that} \alg[2] is \bmonotone.
+We also use \refeq{eq:BIR-modification} to argue that \alg[2] is \bmonotone.
 This follows from \eqref{eq:BIR-modification} since both \alg[1] and \alggr are \bmonotone.
@@
 Recall that by Theorem~\ref{thm:random-clean} \DynGreedy cannot be played in equilibrium when it is BIR-dominated by some feasible, $T$-oblivious algorithm and the time horizon $T$ is large enough.
-\asedit{Also recall that pure and mixed Nash equilibria are the same in our model.}
+Also recall that pure and mixed Nash equilibria are the same in our model.
 
-\asedit{Corollary~\ref{cor:random-greedy} holds more generally: it suffices to assume that feasible algorithms are \bmonotone and closed under the greedy modification.
-The latter means that the greedy modification of any feasible algorithm is also feasible, for any mixing parameter $p$.}
+Corollary~\ref{cor:random-greedy} holds more generally: it suffices to assume that feasible algorithms are \bmonotone and closed under the greedy modification.
+The latter means that the greedy modification of any feasible algorithm is also feasible, for any mixing parameter $p$.
 
@@
  and adoption of better algorithms.
-\asdelete{Recall that both \competitiveness (of the game between the two principals) and \rationality (of the agents) are controlled by the response function $\respF$.}
 
@@
 Competitiveness within this model are controlled by the baseline probability $\eps_0 = \respF(- 1)$, which goes smoothly between the two extremes of \HardMax ($\eps_0=0$, tough competition) and the uniform choice ($\eps_0=\tfrac12$, no competition).
-\asedit{Recall} % For clarity, we assume 
+Recall % For clarity, we assume 
 that principal's utility is the number of agents.

### 2020-12-21 journal paper/content/sim_details.tex
pass through alex's edits

@@
 multi-armed bandit problem (and each such experiment is run many times to reduce variance).
-Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit \asedit{problems}\}, and we are looking for findings that are consistent across this entire space.
+Accordingly, we have a three-dimensional space of atomic experiments one needs to run and interpret: \{pairs of algorithms\} x \{competition models\} x \{bandit problems\}, and we are looking for findings that are consistent across this entire space.
 It is essential to keep each of the three dimensions small yet representative.
@@
 Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.
-We use \Thompson (\TS) from the ``adaptive-exploration" algorithms, \DynamicEpsGreedy (\DEG) from the ``exploration-separating" algorithms,%
-\footnote{In each round, \DynamicEpsGreedy algorithm explores uniformly with a predetermined probability $\eps$, and ``exploits" with the remaining probability, choosing an arm with maximal posterior mean reward given the current data.
+We use \Thompson (\TS) from the ``adaptive-exploration" algorithms\footnote{\gaedit{We provide details on \Thompson in Appendix~\ref{app:bg}.}} \DynamicEpsGreedy (\DEG) from the ``exploration-separating" algorithms,%
+\footnote{\gaedit{\DynamicEpsGreedy proceeds as follows.
+In each round, with probability $\eps$ it uniformly at random selects an arm to pull from the full set of arms and} ``exploits" with the remaining probability, choosing an arm with maximal posterior mean reward given the current data.
 We use $\eps=5\%$ throughout.
@@
 \xhdr{MAB instances.}
-We consider \asedit{bandit problems} with $K=10$ arms and Bernoulli rewards.
-\asedit{The \emph{\MRV} $(\mu(a):\; a\in A)$ is initially} drawn from some distribution, termed \emph{MAB instance}.
+We consider bandit problems with $K=10$ arms and Bernoulli rewards.
+The \emph{\MRV} $(\mu(a):\; a\in A)$ is initially drawn from some distribution, termed \emph{MAB instance}.
 We consider three MAB instances:

### 2020-12-21 journal paper/journal_main.tex
A pass addressing YK comments.
pass through alex's edits

@@
 Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.\newline \indent
-We would like to thank Ian Ball, Yeon-Koo Che, Glen Weyl and Sven Rady for helpful comments and conversations. Likewise, we thank the audience of the conference and seminar talks on this work.
+We would like to thank Ian Ball, Yeon-Koo Che, Sara Shahanaghi, Glen Weyl and Sven Rady for helpful comments and conversations.
+Likewise, we thank the audience of the conference and seminar talks on this work.
 %seminar participants at Columbia and conference participants at Innovations in Theoretical Computer Science 2018, ACM Economics and Computation 2019 and the MIT Conference on Digital Experimentation 2020, for helpful comments and conversations.
@@
 Most online platforms strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-We \asedit{study} the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users.
+We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users.
 Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
 
-We consider a stylized duopoly model in which two firms face \asedit{the same multi-armed bandit problem}.
-Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit instance only if it is chosen.
+We consider a stylized duopoly model in which two firms face the same multi-armed bandit problem.
+Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit problem only if it is chosen.
 Through a mix of theoretical results and numerical simulations, we study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for users.
@@
 
-%\newpage
 \section{Theoretical results: the \TheoryModel}

### 2020-12-21 journal paper/content/related_work.tex
Some edits after Guy's pass, removing some color-edits.

@@
 Second, the inputs to agents' decision rule are prices, rather than the quality of the chosen alternatives; the former is directly controlled by the firms, whereas the latter is not known a priori.
-Third, the goal of exploration is to learn the agents' decision rule, \gaedit{not the best product alternative}.
+Third, the goal of exploration is to learn the agents' decision rule, not the best product alternative.
 
@@
 Specifically, the ``entrant" firm offers a new product of unknown, exogenously determined, quality, which is ``explored" if and only if the entrant attracts customers.
-However, because the entrant can only control the prices but not the product itself, the nature of exploration is fundamentally different \gaedit{than in our setting}.
+However, because the entrant can only control the prices but not the product itself, the nature of exploration is fundamentally different than in our setting.
 First, the goal of exploration is to reveal information about a fixed product.

### 2020-12-21 journal paper/content/sec-intro.tex
Some edits after Guy's pass, removing some color-edits.

@@
 Principals compete on quality rather than on prices, and engage in exploration in order to learn which actions lead to high quality products.
-A new agent arrives and chooses a principal according to some decision rule \gaedit{which depends on the principal's exploration strategy}.
+A new agent arrives and chooses a principal according to some decision rule \asedit{(which may depend on the principal's exploration strategy, either directly or indirectly)}.
 The principal selects an action which affects the quality of service provided to this agent (\eg a list of web search results).
@@
 However, the greedy algorithm is sometimes \emph{not} the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).}
-We revisit algorithms' performance \gaedit{in a standalone bandit problem, \ie without competition, and} we find that the most natural performance measure does not explain this phenomenon.
+We revisit algorithms' performance in a standalone bandit problem, \ie without competition, and we find that the most natural performance measure does not explain this phenomenon.
 We suggest a new, more nuanced performance measure that does.

### 2020-12-21 journal paper/content/sec-model.tex
Some edits after Guy's pass, removing some color-edits.

@@
 \item \SoftMaxRandom: $\respF$  lies in $[\eps_0,1-\eps_0]$, breaks ties fairly, and has a bounded derivative around $0$ (see Definition~\ref{def:SoftMax} for a formal definition).
+\asedit{Intuitively, $\respF$ is a smoothed version of \HardMaxRandom, without a sharp threshold therein.}
 \end{itemize}
 
-\HardMaxRandom and \SoftMaxRandom can be seen as agent types that\gadelete{, resp., give some chance to the other principal and avoid sharp transitions in their probabilistic response.} \gaedit{make mistakes due to being oblivious to the principals' reputation, due to lack of awareness or interest.
+\asedit{\HardMaxRandom and \SoftMaxRandom agents can be interpreted in several ways.
+First, they make mistakes, due to lack of awareness or interest.
+Second, they give some chance to the non-preferred principal, due to irrational effects or curiosity.}
+%\gaedit{make mistakes due to being oblivious to the principals' reputation, due to lack of awareness or interest.
 \HardMaxRandom models users as either being aware of the reputation or not, whereas \SoftMaxRandom models user mistakes as following a smooth density.}
-\asmargincomment{I don't get YK's beef with this.}
-Alternatively, they can be realized as a distribution over more ``basic" agent types.
-Indeed, \gaedit{the} \HardMaxRandom distribution is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
-\gadelete{The latter can be interpreted as consumers oblivious to principals' reputation, due to the lack of awareness or interest.} One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
+Third, they can be realized as a distribution over more ``basic" agent types.
+Indeed, the \HardMaxRandom distribution is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
+\asedit{The latter can be interpreted as consumers that are completely oblivious to principals' reputation.}
+%, due to the lack of awareness or interest.} 
+One can obtain a \SoftMaxRandom response function using agent types that choose a principal $i$ with a largest reward estimate $\est_i$, unless $|\est_1-\est_2|$ is upper-bounded by some parameter $\theta$, in which case they choose uniformly.
 Then, we obtain \SoftMaxRandom as a mixture of random agents and these ``$\theta$-\HardMax'' agents, for a suitable distribution over $\theta$.
@@
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-\gaedit{Principals are risk-neutral and their utility is defined as their market share, \ie the number of agents that chose this principal, and thus they aim to select the algorithm that maximizes their expected utility.
-(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.)}
-\gadelete{Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral and aim to maximize their expected utility.}
+\asedit{(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.) Principals are risk-neutral and their utility is defined as their market share, \ie the number of agents that chose this principal.
+Thus they aim to select the algorithm that maximizes their expected market share.}
+
+%Principal's utility is defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral and aim to maximize their expected utility.
 

### 2020-12-21 journal paper/content/sim_details.tex
Some edits after Guy's pass, removing some color-edits.

@@
 Our pilot experiments indicate that our findings do not change substantially if other representative algorithms are chosen.
-We use \Thompson (\TS) from the ``adaptive-exploration" algorithms\footnote{\gaedit{We provide details on \Thompson in Appendix~\ref{app:bg}.}} \DynamicEpsGreedy (\DEG) from the ``exploration-separating" algorithms,%
-\footnote{\gaedit{\DynamicEpsGreedy proceeds as follows.
-In each round, with probability $\eps$ it uniformly at random selects an arm to pull from the full set of arms and} ``exploits" with the remaining probability, choosing an arm with maximal posterior mean reward given the current data.
+We use
+\DynGreedy (\DG) algorithm  as in Section~\ref{sec:theory-prelims},
+\DynamicEpsGreedy (\DEG) from the ``exploration-separating" algorithms,
+and \Thompson (\TS) from the ``adaptive-exploration" algorithms.%
+\footnote{\asedit{In each round $t$, \Thompson computes a Bayesian posterior on $\mu_a$ for each arm $a$ and draws an independent sample $\tilde{\mu}_{a,t}$ from this posterior; it chooses an arm which maximizes $\tilde{\mu}_{a,t}$.
+    \newline\indent 
+\DynamicEpsGreedy proceeds as follows.
+In each round, with probability $\eps$ it explores by choosing an arm from the full set of arms uniformly at random.
+With the remaining probability,} it ``exploits" by choosing an arm with maximal posterior mean reward given the current data.
 We use $\eps=5\%$ throughout.
-Our pilot experiments show that choosing a different $\eps$ does not qualitatively change the results.}  and \DynGreedy (\DG) algorithm  as in Section~\ref{sec:theory-prelims}.
+Our pilot experiments show that choosing a different $\eps$ does not qualitatively change the results.}
 For ease of comparison, all three algorithms are parameterized with the same ``fake" Bayesian prior: namely, the mean reward of each arm is drawn independently from a $\Beta(1,1)$ distribution.

### 2020-12-21 journal paper/content/sec-intro.tex
minor edits.

@@
 Principals compete on quality rather than on prices, and engage in exploration in order to learn which actions lead to high quality products.
-A new agent arrives and chooses a principal according to some decision rule \asedit{(which may depend on the principal's exploration strategy, either directly or indirectly)}.
+A new agent arrives and chooses a principal according to some decision rule (which may depend on the principal's exploration strategy, either directly or indirectly).
 The principal selects an action which affects the quality of service provided to this agent (\eg a list of web search results).
-The agent experiences this action \gaedit{and the resulting utility from this action is observed by the principal.} \gadelete{and reports a reward.} Each principal only observes its own users.
+The agent experiences this action and the resulting reward from this action is observed by the principal.
+Each principal only observes its own users.
 Principals commit to their strategies in advance, so as to maximize their market share.

### 2020-12-22 journal paper/content/sec-intro.tex
Minor edit.

@@
 A new agent arrives and chooses a principal according to some decision rule (which may depend on the principal's exploration strategy, either directly or indirectly).
-The principal selects an action which affects the quality of service provided to this agent (\eg a list of web search results).
+The principal selects an action which affects the quality of service provided to this agent, \eg a list of web search results.
 The agent experiences this action and the resulting reward from this action is observed by the principal.
@@
 (They can be interpreted as consumers that are oblivious to the principals' reputation.) We call this model \HardMaxRandom.
-Then better algorithms help in a big way: a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
+Then better algorithms help in a big way: \asedit{when two algorithms compete against one another,} a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
 There is a caveat, however: any algorithm can be defeated by interleaving it with the greedy algorithm.

### 2020-12-22 journal paper/journal_main.tex
\newpage and small for the bib.
Fixed ref to the ec19 paper.

@@
 \thanks{This is a merged and final version of two conference publications,
-\citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19},
+\citet*{CompetingBandits-itcs18} and \citet*{CompetingBandits-ec19},
 with a unified and streamlined presentation and expanded related work and background materials.
@@
 
-%\newpage
+\newpage
 \addcontentsline{toc}{section}{References}
@@
 \bibliographystyle{ecta.bst}
-%\begin{small}
+\begin{small}
 \bibliography{bib-abbrv,bib-ML,refs,bib-bandits,bib-AGT,bib-slivkins, bib-random}
-%\end{small}
+\end{small}
 

### 2020-12-23 journal paper/ecta.tex
add econometrica version

--- 2020-12-22 journal paper/journal_main.tex
@@
 \documentclass[12pt]{article}
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=1.25in]{geometry}
 
@@
 Zhiwei Steven Wu%
-\footnote{Carnegie Mellon University, Pittsburgh, PA.
-Email: zstevenwu@cmu.edu.\newline
+\footnote{Carnegie Mellon University, Pittsburgh, PA
+Email: zstevenwu@cmu.edu\newline
 Research done when Z.S. Wu was an intern and a postdoc at Microsoft Research NYC.}
@@
 \begin{abstract}
+\gadelete{
 Most online platforms strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
@@
 innovation" relationship, and elucidate the first-mover advantage in the digital economy.
+}
 
-\vspace{0.2in}
+\gaedit{
+Most online platforms strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
+We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users.
+
+We consider a stylized duopoly model in which two firms face the same multi-armed bandit problem.
+Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit problem only if it is chosen.
+We study whether competition incentivizes the adoption of better bandit algorithms.
+We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low welfare.
+However, weakening competition by providing firms with some ``free" users incentivizes better exploration strategies and increases welfare.
+We investigate two channels for weakening the competition: stochastic choice models for the users and giving one firm a first-mover advantage.
+} \\
+
+\vspace{0.1in}
+\noindent\textbf{Keywords:}  Multi-armed bandits, Competition vs. Innovation
 
@@
 \newpage
-\begin{small}
-\setcounter{tocdepth}{2}
-\tableofcontents
-\end{small}
+%\begin{small}
+%\setcounter{tocdepth}{2}
+%\tableofcontents
+%\end{small}
 \setcounter{page}{0}
-\thispagestyle{empty}
-\newpage
+%\thispagestyle{empty}
+%\newpage
+
+\onehalfspacing
 %

### 2020-12-23 journal paper/content/conclusion.tex
Removed the coloring.

@@
 A firm makes progress on its learning problem if and only if it attracts users.
-\asdelete{We investigate two variants: an analytical variant where users do not observe any performance signals and choose between firms according to their Bayesian-expected rewards, and numerical simulations where a reputation score is observed for each firm, based on the average reward of its recent users.
+%\asdelete{We investigate two variants: an analytical variant where users do not observe any performance signals and choose between firms according to their Bayesian-expected rewards, and numerical simulations where a reputation score is observed for each firm, based on the average reward of its recent users.
 In both variants,}

### 2020-12-23 journal paper/content/related_work.tex
Removed the coloring.

@@
 The literature on learning-by-doing vs. competition \citep[\eg][]{gilbert1982preemptive, fudenberg1983learning, dasgupta1988learning, cabral1994learning}
-\asmargincomment{Guy: pls cite Gilbert-Newberry, as per YK.}
 studies firms that learn while competing against each other, so that a firm attracting more consumers reduces its production costs.
@@
 %as we move from \HardMax to \HardMaxRandom.
-\gadelete{Second, there a large literature on non-existence of equilibria due to small deviations, starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets.
-\asdelete{\footnote{\cite{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom in this context.}}
-This is superficially similar to how small deviations towards the greedy algorithm rule out equilibria under \HardMaxRandom.}
-%Second, the stochastic choice rules we consider have been widely utilized across economics. For instance,
+%\gadelete{Second, there a large literature on non-existence of equilibria due to small deviations, starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets.
+\footnote{\cite{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom in this context.} This is superficially similar to how small deviations towards the greedy algorithm rule out equilibria under \HardMaxRandom.}
 Second, \SoftMaxRandom subsumes the logit choice rule, a standard behavioral model with strong empirical and microeconomic foundations

### 2020-12-23 journal paper/content/sec-intro.tex
Removed the coloring.

@@
 (They can be interpreted as consumers that are oblivious to the principals' reputation.) We call this model \HardMaxRandom.
-Then better algorithms help in a big way: \asedit{when two algorithms compete against one another,} a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
+Then better algorithms help in a big way: when two algorithms compete against one another, a sufficiently better algorithm is guaranteed to win all non-random agents after an initial learning phase.
 There is a caveat, however: any algorithm can be defeated by interleaving it with the greedy algorithm.
@@
 While traditional models of innovation study lab-based R\&D, we consider data-driven innovation: one that crucially depends on data generated by the firm's customers.
-\asmargincomment{important sentence!}
 We focus on innovation in \emph{exploration technology} which systematically improves the firm's products, whereas prior work would define innovation as improvement in the products themselves.

### 2020-12-23 journal paper/content/sec-model.tex
minor edits.
Removed the coloring.

@@
 \item \SoftMaxRandom: $\respF$  lies in $[\eps_0,1-\eps_0]$, breaks ties fairly, and has a bounded derivative around $0$ (see Definition~\ref{def:SoftMax} for a formal definition).
-\asedit{Intuitively, $\respF$ is a smoothed version of \HardMaxRandom, without a sharp threshold therein.}
+Intuitively, $\respF$ is a smoothed version of \HardMaxRandom, without a sharp threshold therein.
 \end{itemize}
 
-\asedit{\HardMaxRandom and \SoftMaxRandom agents can be interpreted in several ways.
+\HardMaxRandom and \SoftMaxRandom agents can be interpreted in several ways.
 First, they make mistakes, due to lack of awareness or interest.
-Second, they give some chance to the non-preferred principal, due to irrational effects or curiosity.}
+Second, they give some chance to the non-preferred principal, due to curiosity or a behavioral effect like probability matching.
 %\gaedit{make mistakes due to being oblivious to the principals' reputation, due to lack of awareness or interest.
@@
 Indeed, the \HardMaxRandom distribution is a mixture of \HardMax and ``random agents" (which choose a principal uniformly at random).
-\asedit{The latter can be interpreted as consumers that are completely oblivious to principals' reputation.}
+The latter can be interpreted as consumers that are completely oblivious to principals' reputation.
 %, due to the lack of awareness or interest.}
@@
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-\asedit{(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.) Principals are risk-neutral and their utility is defined as their market share, \ie the number of agents that chose this principal.
-Thus they aim to select the algorithm that maximizes their expected market share.}
+(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.) Principals are risk-neutral and their utility is defined as their market share, \ie the number of agents that chose this principal.
+Thus they aim to select the algorithm that maximizes their expected market share.
 

### 2020-12-23 journal paper/content/sim_details.tex
Removed the coloring.

@@
 and \Thompson (\TS) from the ``adaptive-exploration" algorithms.%
-\footnote{\asedit{In each round $t$, \Thompson computes a Bayesian posterior on $\mu_a$ for each arm $a$ and draws an independent sample $\tilde{\mu}_{a,t}$ from this posterior; it chooses an arm which maximizes $\tilde{\mu}_{a,t}$.
+\footnote{In each round $t$, \Thompson computes a Bayesian posterior on $\mu_a$ for each arm $a$ and draws an independent sample $\tilde{\mu}_{a,t}$ from this posterior; it chooses an arm which maximizes $\tilde{\mu}_{a,t}$.
     \newline\indent
@@
 In each round, with probability $\eps$ it explores by choosing an arm from the full set of arms uniformly at random.
-With the remaining probability,} it ``exploits" by choosing an arm with maximal posterior mean reward given the current data.
+With the remaining probability, it ``exploits" by choosing an arm with maximal posterior mean reward given the current data.
 We use $\eps=5\%$ throughout.

### 2020-12-23 journal paper/content/sec-bg.tex
Removed some \newpage, minor edit.

@@
-We present self-contained background on multi-armed bandits (MAB), to make the paper accessible to researchers who are not experts on MAB. More details can be found in any of the monographs
+We present self-contained background on multi-armed bandits (MAB), to make the paper accessible to researchers who are not experts on MAB. More details can be found in books 
 \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
@@
 Some MAB algorithms, \eg Thompson Sampling, are Bayesian: they input a prior on mean rewards, and attain strong Bayesian guarantees (in expectation over the prior) when the prior is correct.
-Such algorithms can also be initialized with some simple `fake' priors; in fact, this is how Thompson Sampling can be made to satisfy the optimal regret bounds listed above.
+Such algorithms can also be initialized with some simple `fake' priors; in fact, this is how Thompson Sampling can be made to satisfy the optimal regret bounds.
 

### 2020-12-23 journal paper/ecta.tex
Removed some \newpage, minor edit.
Edits to the title page.

--- 2020-12-21 journal paper/journal_main.tex
@@
 The Perils of Exploration under Competition%
-\thanks{This is a merged and final version of two conference publications,
-\citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19},
-with a unified and streamlined presentation and expanded related work and background materials.
+\thanks{Preliminary abstracts have appeared as 
+\citet*{CompetingBandits-itcs18} and \citet*{CompetingBandits-ec19}.
+This is a merged version, with a unified and streamlined presentation and expanded related work and background materials.
 All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
-Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.\newline \indent
+Appendices~\ref{app:bg},\ref{app:examples} are new.
+\newline \indent
 We would like to thank Ian Ball, Yeon-Koo Che, Sara Shahanaghi, Glen Weyl and Sven Rady for helpful comments and conversations.
@@
 Zhiwei Steven Wu%
-\footnote{Carnegie Mellon University, Pittsburgh, PA.
-Email: zstevenwu@cmu.edu.\newline
+\footnote{Carnegie Mellon University, Pittsburgh, PA
+Email: zstevenwu@cmu.edu\newline
 Research done when Z.S. Wu was an intern and a postdoc at Microsoft Research NYC.}
 }
-\date{First version: July 2020\\ This version: December 2020}
+\date{}
+%\date{First version: July 2020\\ This version: December 2020}
 
@@
 We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users.
-Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
+%Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
 
@@
 Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit problem only if it is chosen.
-Through a mix of theoretical results and numerical simulations, we study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for users.
+We study whether competition incentivizes the adoption of better bandit algorithms.
 We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low welfare.
 However, weakening competition by providing firms with some ``free" users incentivizes better exploration strategies and increases welfare.
-We investigate two channels for weakening the competition: relaxing the rationality of users and giving one firm a first-mover advantage.
-Our findings are closely related to the ``competition vs.
-innovation" relationship, and elucidate the first-mover advantage in the digital economy.
+We investigate two channels for weakening the competition: stochastic choice models for the users and giving one firm a first-mover advantage.
+%Our findings are closely related to the ``competition vs. innovation" relationship, and elucidate the first-mover advantage in the digital economy.
+\\
 
-\vspace{0.2in}
+\vspace{0.1in}
+\noindent\textbf{Keywords:}  Competition vs. innovation, exploration-exploitation tradeoff, multi-armed bandits, regret.
 
@@
 \subfile{content/sec-bg}
-
-\newpage
 

### 2021-01-16 journal paper/ecta.tex
make sure abstract is <150 words

@@
 This is a merged version, with a unified and streamlined presentation and expanded related work and background materials.
-All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
+\newline \indent %All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
 Appendices~\ref{app:bg},\ref{app:examples} are new.
-\newline \indent
 We would like to thank Ian Ball, Yeon-Koo Che, Sara Shahanaghi, Glen Weyl and Sven Rady for helpful comments and conversations.
@@
 \author{Guy Aridor
-\footnote{Columbia University, Department of Economics. Email: g.aridor@columbia.edu}
+\footnote{Columbia University, Department of Economics. Email: g.aridor@columbia.edu.}
 \and \hspace{-0.75cm}
@@
 Yishay Mansour
-\footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com}
+\footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com.}
 \and \hspace{-0.75cm}
@@
 Aleksandrs Slivkins%
-\footnote{Microsoft Research New York City. Email: slivkins@microsoft.com}
+\footnote{Microsoft Research New York City. Email: slivkins@microsoft.com.}
 \and \hspace{-0.75cm}
@@
 \footnote{Carnegie Mellon University, Pittsburgh, PA
-Email: zstevenwu@cmu.edu\newline
+Email: zstevenwu@cmu.edu.\newline
 Research done when Z.S. Wu was an intern and a postdoc at Microsoft Research NYC.}
@@
 \begin{abstract}
-Most online platforms strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users.
+Most online platforms strive to learn from interactions with users, and engage in \emph{exploration}: making potentially suboptimal choices in order to acquire new information.
+We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and competition for users.
 %Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
@@
 Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit problem only if it is chosen.
-We study whether competition incentivizes the adoption of better bandit algorithms.
-We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low welfare.
+We study whether competition incentivizes the adoption of better algorithms.
+We find that stark competition induces firms to commit to a ``greedy" algorithm that leads to low welfare.
 However, weakening competition by providing firms with some ``free" users incentivizes better exploration strategies and increases welfare.
-We investigate two channels for weakening the competition: stochastic choice models for the users and giving one firm a first-mover advantage.
+We investigate two channels for weakening the competition: stochastic user choice models and giving one firm a first-mover advantage.
 %Our findings are closely related to the ``competition vs. innovation" relationship, and elucidate the first-mover advantage in the digital economy.

### 2021-01-16 journal paper/content/sec-intro.tex
small tweaks

@@
 \gaedit{This classification is important for interpreting the equilibrium algorithm choice of the firms as it allows for a ``low" to ``medium" to ``high" technology interpretation.
-We further use the classification to pick a representative algorithm from each class for our numerical experiments.
+We further use the classification to pick a representative algorithm from each class for our numerical \gaedit{simulations}.
 Going from low to medium to high technology, these three classes are as follows:} \gadelete{Going from more primitive to more sophisticated, these three classes are as follows:}
@@
 \footnote{\label{fn:Tsquared}\Eg this is because at each round $t$, one needs to recompute, and integrate over, a discrete distribution with $t$ possible values, namely the number of agents that have chosen principal $1$ so far.}
-which precludes experiments at a sufficient scale.
+which precludes \gaedit{numerical simulations} at a sufficient scale.
 The \ExptsModel features an intricate feedback loop between algorithms' performance, their reputations and agents' choices, which simplifies the simulations but does not appear analytically tractable.

### 2021-01-18 journal paper/content/sec-intro.tex
Minor edits for space.

@@
 
-The main high-level question we ask is:
+%The main high-level question we ask is:
 %that we focus on in this paper is:
 %\begin{align}\label{eq:main-Q}
+\asedit{We ask:}
 \textbf{Whether and how does competition incentivize the adoption of better exploration algorithms?}
@@
 
-\textbf{(i)} The most obvious decision rule maximizes the Bayesian-expected reward; we refer to it as \HardMax.
+\textbf{(i)} The most obvious decision rule maximizes the Bayesian-expected reward; we call it \HardMax.
 We find that it is not conducive to adopting better algorithms: each principal's dominant strategy is to choose the greedy algorithm.
@@
 %\HardMax is very sensitive to tie-breaking:
-if the tie-breaking is probabilistically biased in favor of one principal, then this principal can always prevail in competition.
+if the tie-breaking is probabilistically biased in favor of one principal, the latter 
+%then this principal 
+can always prevail in competition.
 % has a simple ``winning strategy" no matter what the other principal does.
@@
 %as a function of the difference between  principals' Bayesian-expected rewards;
-We call it \SoftMaxRandom, a more realistic middle ground between \HardMax agents and random agents.%
+We call it \SoftMaxRandom, a more realistic middle ground between \HardMax and random agents.%
 \footnote{Alternatively, one can obtain a \SoftMaxRandom decision rule using a mixture of more ``basic" agent types that follow \HardMax unless the principal's Bayesian-expected rewards are too close to each other.}
@@
 We focus on innovation in \emph{exploration technology} which systematically improves the firm's products, whereas prior work would define innovation as improvement in the products themselves.
-We recover the traditional inverted-U relationship purely through the reputational consequences of exploration, rather than monetary costs/benefits thereof.
+We recover the traditional inverted-U relationship purely through the reputational consequences of exploration.
+\asdelete{rather than monetary costs/benefits thereof.}
 This is in contrast to the inverted-U relationships established in prior work, which rely on the monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.
@@
 without invoking any prior knowledge or beliefs.
-The performance signals are abstracted and aggregated as a scalar \emph{reputation score} for each principal, modeled as a sliding window average of its rewards.
+The performance signals are aggregated as a scalar \emph{reputation score} for each principal, modeled as a sliding window average of its rewards.
 Thus, agents' decision rule depends only on the two reputation scores.
@@
 We find that the greedy algorithm often wins under the \HardMax decision rule, with a strong evidence of the ``death spiral" effect mentioned earlier.
- As predicted by the theory, better algorithms prevail under \HardMaxRandom if the expected number of ``random" users is sufficiently large.
+As predicted by the theory, better algorithms prevail under \HardMaxRandom
+\asedit{with enough ``random" users.}
+%if the expected number of ``random" users is sufficiently large.
  %\asdelete{However, this effect is negligible for smaller parameter values.}
@@
 However, the greedy algorithm is sometimes \emph{not} the best strategy under high levels of competition.\footnote{In our theoretical results on \HardMax, the greedy algorithm is always the best strategy, mainly because it is aware of the Bayesian prior (whereas in the simulations  the prior is not available).}
-We revisit algorithms' performance in a standalone bandit problem, \ie without competition, and we find that the most natural performance measure does not explain this phenomenon.
-We suggest a new, more nuanced performance measure that does.
+We revisit algorithms' performance in a standalone bandit problem, \ie without competition.
+We find that the most natural performance measure does not explain this phenomenon, and suggest a new, more nuanced \asedit{one} that does.
+%performance measure that does.
 

### 2021-01-18 journal paper/content/related_work.tex
Minor space-saving.

@@
 The most relevant thread concerns the basic model of regret-minimizating bandits with stochastic rewards and no auxiliary structure (which is the problem faced by each principal in our model), see Appendix~\ref{app:bg} for background.
-This basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.%
-\footnote{There is a superficial similarity, in name only, between this paper and the work on ``dueling bandits" \citep[starting from][]{Yue-dueling12,Yue-dueling-icml09}.
+This basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.
+%\footnote{There is a superficial similarity, in name only, between this paper and the work on ``dueling bandits" \citep[starting from][]{Yue-dueling12,Yue-dueling-icml09}.
 In this work, there is only one bandit algorithm which chooses two arms in each round, and the only observes which arm has ``won the duel".}
 Dedicated monographs \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} cover the work on regret-minimizing formulations (which mainly comes from computer science). The classic book
-\citep{Gittins-book11} focuses on the Markovian formulations (which predate regret-minimization and were mainly studied in operations research and statistics).
-Various connections to economics and game theory are detailed in
+\citep{Gittins-book11} focuses on the Markovian formulations, which predate regret-minimization. Various connections to economics and game theory are detailed in
 books \citep{CesaBL-book,slivkins-MABbook} and surveys \citep{Bergemann-survey06,Horner-survey16}. Industrial applications are discussed in \citep{DS-arxiv}.
@@
 incentivized exploration in recommendation systems
-(starting from \citet{Kremer-JPE14,Che-13}, see \citet[Ch. 11.6]{slivkins-MABbook} for a literature review),
+(starting from \citet{Kremer-JPE14,Che-13}, see \citet[Ch. 11]{slivkins-MABbook}),
     %\citep[\eg][]{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,Jieming-unbiased18},
@@
 Third, revealing more information is not necessarily better for the entrant (if the product's quality is actually low).
-Fourth, the social planner's exploration problem is very different from the firms', so that when one compares competition to the planner, one can only compare outcomes, but not the algorithms.\footnote{The planner faces a two-armed bandit problem, because the planner can directly choose a firm for each agent.}
+Fourth, the social planner's exploration problem is very different from the firms', so that when one compares competition to the planner, one can only compare outcomes, but not the algorithms.\footnote{The planner faces a two-armed bandit problem, since it can directly choose a firm for each agent.}
 % firm, because payments cancel out in social welfare).}
@@
 In the line of work on \emph{strategic experimentation} (starting from \citet{Bolton-econometrica99,Keller-econometrica05}, see \citet{Horner-survey16} for a survey), agents explore and learn over time in a shared environment.
-Thus, we have exploration algorithms which interact with each other strategically, \eg each agent would prefer to free-ride on the exploration done by others.
+Thus, we have exploration algorithms which interact with each other strategically, \eg each agent prefers to free-ride on someone else's exploration.
 However, the agents do not compete with each other in any meaningful sense.
@@
 \cite{schmalensee1982product, bagwell1990informational} investigate how buyer uncertainty about product quality can serve as a barrier to entry for late arrivers; we find a similar effect with ``reputation advantage".
-The role of data as a barrier to entry in online markets has been noted in \cite{de2020data}; we find a similar effect with ``data advantage".
-\citet{kerin1992first} overview the other channels through which first-mover advantage can lead to a competitive advantage.
+\cite{de2020data} note the role of data as a barrier to entry in online markets; we find a similar effect with ``data advantage".
+\citet{kerin1992first} overview other channels through which first-mover advantage can \asedit{affect competition}.% lead to a competitive advantage.
 
@@
 
-Classic ``market competitiveness" measures,such as the Lerner Index or the Herfindahl-Hirschman Index
-\citep{tirole1988theory} are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).
+\asedit{While we use first-mover advantage and agents' decision rule,} classic ``market competitiveness" measures, such as the Lerner Index or the Herfindahl-Hirschman Index
+\citep{tirole1988theory}, are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).
 
@@
 \citep[\eg][]{mosteller1951experimental, luce1959choice, mcfadden1974measurement, matvejka2015rational}.
-Third, choice models similar to \SoftMaxRandom are used to explain  horizontal product differentiation \citep[\eg][]{Hotelling-29, Perloff-Salop-85}.
+Choice models similar to \SoftMaxRandom are used to explain  horizontal product differentiation \citep[\eg][]{Hotelling-29, Perloff-Salop-85}.
 %\asdelete{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}

### 2021-01-18 journal paper/content/sec-intro.tex
Minor space-saving.

@@
 
-Interestingly, we find that data feedback loops can be bad for innovation (as in ``death spiral"), or good for innovation (as in ``amplification effect"), depending on the level of competition.
+Interestingly, 
+\asdelete{we find that}
+data feedback loops can be bad for innovation (as in ``death spiral"), or good for innovation (as in ``amplification effect"), depending on the level of competition.
+
+%Interestingly, we find that data feedback loops can be bad for innovation (as in ``death spiral"), or good for innovation (as in ``amplification effect"), depending on the level of competition.
 
@@
 \footnote{\label{fn:Tsquared}\Eg this is because at each round $t$, one needs to recompute, and integrate over, a discrete distribution with $t$ possible values, namely the number of agents that have chosen principal $1$ so far.}
-which precludes \gaedit{numerical simulations} at a sufficient scale.
+which precludes numerical simulations at a sufficient scale.
 The \ExptsModel features an intricate feedback loop between algorithms' performance, their reputations and agents' choices, which simplifies the simulations but does not appear analytically tractable.

### 2021-01-19 journal paper/content/sec-model.tex
Minor edits for space-saving.

@@
 where $[T] := \{1,2\LDOTS T\}$.
-The game proceeds in $T$ rounds.
+%The game proceeds in $T$ rounds.
 
@@
 Thus, each principal follows the protocol of \emph{multi-armed bandits} (henceforth, \emph{MAB}).
-Specifically, in each round when it is chosen, it chooses an arm from $A$ and observes a reward for this action.
+In each round when it is chosen, it chooses an arm from $A$ and observes a reward for this action.
 It does not observe anything else, particularly rewards for other actions.
@@
 
-\xhdr{Agents' response.} Each agent $t$ forms reward estimates $\est_i(t)\in [0,1]$ for each principal $i$.
+\xhdr{Agent response.} Each agent $t$ forms a reward estimate $\est_i(t)\in [0,1]$ for each principal $i$.
 (What these estimates are, and how much the agents know in order to form them, depends on the Bayesian vs.
@@
 
-\begin{itemize}
+\begin{OneLiners}
 \item \HardMax: $\respF$ equals $0$ on the interval $[-1,0)$ and $1$
@@
 Intuitively, $\respF$ is a smoothed version of \HardMaxRandom, without a sharp threshold therein.
-\end{itemize}
+\end{OneLiners}
 
@@
 
-\begin{itemize}
+\begin{OneLiners}
 \item In the \emph{\TheoryModel}, the mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from a common Bayesian prior $\priorMu$.
@@
     for each agent $t$ and principal $i$.%
-\footnote{Formally, agents do the Bayesian update knowing $t$, the principals' algorithms, the prior $\priorMu$, and $\respF$.}
+\footnote{Agents do the Bayesian update knowing $t$, the principals' algorithms, the prior $\priorMu$, and $\respF$.}
 
@@
 
-\end{itemize}
+\end{OneLiners}
 

### 2021-01-19 journal paper/content/related_work.tex
space saver

@@
 pay-per-click ad auctions
-    \citep[\eg][]{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
+    \citep[\eg][]{MechMAB-ec09,DevanurK09},
 coordinating search and matching
@@
 and human computation
-    \citep[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.
+    \citep[\eg][]{RepeatedPA-ec14,Ghosh-itcs13}.
 %A literature review of this work can be found in
@@
 All inputs are observable at once, and principals' payoffs are binary (win/lose).
-\citet{ben2017best, ben2019regression} study competition between ``offline" machine learning algorithms.
+\citet{ben2019regression} study competition between ``offline" machine learning algorithms.
 In comparison, we study a ``product competition" in which the two firms interact continuously (via the customers' choices), accrue rewards incrementally, and compete for individual customers.
 
-A long line of work from electrical engineering and computer science, starting from
-\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, focuses on competition for resources, not competition for consumers.
+%A long line of work from electrical engineering and computer science, starting from
+%\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, focuses on competition for resources, not competition for consumers.
 Specifically, this literature targets an application to \emph{cognitive radios}, where multiple radios transmit simultaneously in a shared medium and compete for bandwidth.
@@
 
-The literature on learning-by-doing vs. competition \citep[\eg][]{gilbert1982preemptive, fudenberg1983learning, dasgupta1988learning, cabral1994learning}
+The literature on learning-by-doing vs. competition \citep[\eg][]{fudenberg1983learning, dasgupta1988learning, cabral1994learning}
 studies firms that learn while competing against each other, so that a firm attracting more consumers reduces its production costs.
@@
 
-\asedit{While we use first-mover advantage and agents' decision rule,} classic ``market competitiveness" measures, such as the Lerner Index or the Herfindahl-Hirschman Index
-\citep{tirole1988theory}, are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).
+%\asedit{While we use first-mover advantage and agents' decision rule,} classic ``market competitiveness" measures, such as the Lerner Index or the Herfindahl-Hirschman Index
+%\citep{tirole1988theory}, are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).
 
@@
 Second, \SoftMaxRandom subsumes the logit choice rule, a standard behavioral model with strong empirical and microeconomic foundations
-\citep[\eg][]{mosteller1951experimental, luce1959choice, mcfadden1974measurement, matvejka2015rational}.
+\citep[\eg][]{mosteller1951experimental, luce1959choice, matvejka2015rational}.
 Choice models similar to \SoftMaxRandom are used to explain  horizontal product differentiation \citep[\eg][]{Hotelling-29, Perloff-Salop-85}.

### 2021-01-19 journal paper/content/sec-bg.tex
space saver

@@
 Thompson Sampling \citep{Thompson-1933,TS-survey-FTML18},%
-\footnote{While Thompson Sampling dates back to 1933 and is probably the best-known bandit algorithm, its regret has not been understood until recently \citep{Shipra-colt12,Kaufmann-alt12,Shipra-aistats13}.}
+\footnote{While Thompson Sampling dates back to 1933 and is probably the best-known bandit algorithm, its regret has not been understood until recently \citep{Shipra-colt12}.}
 UCB1 \citep{bandits-ucb1},
@@
 This phenomenon has been observed in practice
-\citep{practicalCB-arxiv18}, and in theory \citep{kannan2018smoothed,bastani2017exploiting,externalities-colt18} under (very) substantial assumptions.
+\citep{practicalCB-arxiv18}, and in theory \citep{kannan2018smoothed,externalities-colt18} under (very) substantial assumptions.
 The prevalent intuition is that the diversity of contexts can --- under some conditions and to a limited extent --- substitute for explicit exploration.

### 2021-01-19 journal paper/ec19paper/content/inverted_u.tex
Some edits after Guy's pass, removing some color-edits.
space saver

@@
 We investigate the firms' market shares when they choose different algorithms (otherwise, by symmetry both firms get half of the agents).
-We report the market shares for \gaedit{each instance in} Table~\ref{fig:market_share}.
+We report the market shares for each instance in Table~\ref{fig:market_share}.
 We find that $\DG$ is a weakly dominant strategy for the Heavy-Tail and Uniform instances, as long as $T_0$ is sufficiently small.
@@
 Increasing the warm start length $T_0$ makes this period shorter: indeed, considering the relative reputation trajectory in Figure~\ref{relative_rep_plots} (left), increasing $T_0$ effectively shifts the starting time point to the right.
-This is why it helps \DynamicGreedy if $T_0$ is small.
+This is why it helps \DynamicGreedy if $T_0$ is small.\footnote{\gaedit{In Appendix Section~\ref{sec:perf_in_iso_revisited} we further explore the reason why the mean reputation trajectories do not suffice to explain the results observed under competition.
+The primary conclusion of the exercise is that the distribution of rewards of the algorithm is important and in some scenarios the mean is not always a good measure of the typical value of the distribution of rewards.
+As a result, it is not always predictive of results under competition whereas the relative reputation statistic is more closely related to how agents decide between the principals and thus more predictive of outcomes in the competition game.}}
 

### 2021-01-19 journal paper/ec19paper/content/revisited.tex
add changes
space saver

@@
 
-\subsection{Performance in Isolation, Revisited}\label{sec:revisited}
-
-We saw in Section~\ref{sec:competition} that mean reputation trajectories do not suffice to explain the outcomes under competition. Let us provide more evidence and intuition for this.
+We saw in Section~\ref{sec:competition} that mean reputation trajectories do not suffice to explain the outcomes under competition. In this section we provide more evidence and intuition for this.
 
@@
 We find that this also holds for $K = 3$ arms and longer time horizons, see the supplement for more details.
-We conclude \asedit{that Conjecture~\ref{conj:mean-trajectories} is false:}
+We conclude that Conjecture~\ref{conj:mean-trajectories} is false:
 

### 2021-01-19 journal paper/content/sec-theory.tex
Space-saving.

@@
 
-\xhdr{Notation.} Let $\rew_i(n)$ denote the agent's realized reward observed by principal $i$ at local step $n$.
-Put differently, it is the reward collected by algorithm $\alg[i]$ in this local step.
+%\xhdr{Notation.} 
+Let $\rew_i(n)$ denote the agent's realized reward observed by principal $i$ at local step $n$, \ie the reward collected by algorithm $\alg[i]$ in this local step.
 For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen.
@@
 
-\xhdr{Assumptions.} We make two mild assumptions on the prior. First, each arm $a$ can be best:
+\noindent\textbf{Assumptions.} We make two mild assumptions on the prior. First, each arm $a$ can be best:
 \begin{align}\label{eq:assn-prob}
@@
 \noindent Second, posterior mean rewards are pairwise distinct given any feasible history $h$:%
-\footnote{The \emph{history} of an MAB algorithm at a given step
-  comprises the chosen actions and the observed rewards in all
-  previous steps. The history is \emph{feasible} if for each arm-reward pair $(a,r)$ in the history, $r$ is in the support of the reward distribution for $a$.}
+\footnote{The \emph{history} of an MAB algorithm at a given step $t$
+  comprises actions $a_s$ and rewards $r_s$ in all
+  previous steps $s<t$. The history is \emph{feasible} if for each $s$, reward $r_s$ is in the support of the reward distribution for $a_s$.}
 \begin{align}\label{eq:assn-distinct}
@@
 \begin{align}\label{eq:theory-BIR-defn}
-\BIR_i(n) := \E\sbr{ \max_{a\in A} \mu_a - \rew_i(n)}.
+\BIR_i(n) := \textstyle \E[\; \max_{a\in A} \mu_a - \rew_i(n) \;].
 \end{align}
@@
 
-We consider agents with \HardMax response function, and show that principals are not incentivized to \emph{explore}--- \ie to deviate from \DynGreedy.
-The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates.
+We consider agents with \HardMax response function, and show that principals are not incentivized to \emph{explore}, \ie to deviate from \DynGreedy.
+The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates therefrom.
+\asedit{We make this formal below.}
 
-To make this formal, let's define what it means for MAB algorithms to deviate.
+%To make this formal, let's define what it means for MAB algorithms to deviate.
 
@@
 One MAB algorithm \emph{deviates} from another at (local) step $n$ if there is a set $H$ of histories over the previous local steps such that both algorithms lead to $H$ with positive probability, and choose different distributions over arms given any history $h\in H$.
-We say $\alg$ deviates from $\alg'$ \emph{starting from} step $n_0$ if $n=n_0$ is the smallest step when this happens.
+If $n=n_0$ is the smallest such step, we say $\alg$ deviates from $\alg'$ \emph{starting from} step $n_0$.
 \end{definition}
@@
 This is because \DynGreedy receives, in expectation, at least half of the agents before global round $n_0$, and all agents after that; both are the best possible against \alg[2].
-Moreover, \DynGreedy guarantees at least $T/2$ agents in expectation, and any other strategy can receive strictly less than that (\eg if the opponent chooses \DynGreedy).
+Moreover, \DynGreedy guarantees at least $T/2$ agents in expectation, and any other strategy can receive strictly less (\eg if the opponent chooses \DynGreedy).
 
@@
 The \HardMax model is very sensitive to tie-breaking between the principals. 
-For starters, if ties are  broken deterministically in favor of principal $1$, then principal 1 can get all agents no matter what the other principal does, simply by using \StaticGreedy.
+%For starters, 
+If ties are  broken deterministically in favor of principal $1$, this principal can get all agents no matter what the other principal does, simply by using \StaticGreedy.
 
@@
 Then this principal also has a ``winning strategy" no matter what the other principal does.
-Specifically, principal 1 can get all but the first few agents, under a mild technical assumption that \DynGreedy deviates from \StaticGreedy.
-Principal 1 can use \DynGreedy, or any other \bmonotone MAB algorithm that coincides with \DynGreedy in the first few steps.
+Specifically, principal 1 can get all but the first few agents, under a mild assumption that \DynGreedy deviates from \StaticGreedy.
 
@@
 
+Thus, Principal 1 can use \DynGreedy, or any other \bmonotone MAB algorithm that coincides with \DynGreedy in the first few steps.
 The proof re-uses Lemmas~\ref{lm:DG-rew} and~\ref{lm:DG-sudden}, which do not rely on fair tie-breaking.
@@
 In more detail, a principal with asymptotically better \BIR wins by a large margin: after a ``learning phase" of constant duration, all agents choose this principal with maximal possible probability $\respF(1)$.
-For example, a principal with $\BIR(n)\leq \tilde{O}(n^{-1/2})$ prevails over a principal with $\BIR(n)\geq \Omega(n^{-1/3})$.
+For example, a principal with $\BIR(n)\leq \tilde{O}(n^{-1/2})$ prevails over one with $\BIR(n)\geq \Omega(n^{-1/3})$.
 
@@
 (\forall \eps>0)\qquad
-\frac{\BIR_1(\eps n)}{\BIR_2(n)} \to 0.
+%\frac{\BIR_1(\eps n)}{\BIR_2(n)} \to 0.
+\BIR_1(\eps n)\,/\,\BIR_2(n) \to 0.
 \end{align}
@@
 (\forall n\geq n_0) \qquad
-\frac{\BIR_1(\eps_0\, n/2)}{\BIR_2(n)} <\nicefrac12.
+%\frac{\BIR_1(\eps_0\, n/2)}{\BIR_2(n)} <\nicefrac12.
+\BIR_1\rbr{\nicefrac{1}{2}\;\eps_0\, n}\,/\, \BIR_2(n)<\nicefrac12.
 \end{align}
@@
 
-Thus, the main result is stated as follows:
+%Thus, the main result is stated as follows:
 
@@
 
-We'd like to use Theorem~\ref{thm:random-clean} to conclude that a (much) better algorithm prevails in equilibrium.
-We consider a version of the competition game in which the principals are restricted to choosing from a given set of MAB algorithms; the algorithms in this set are called \emph{feasible}.
+%We'd like to use Theorem~\ref{thm:random-clean} 
+To conclude that a (much) better algorithm prevails in equilibrium, we consider a version of the competition game in which the principals are restricted to choosing from a given set of MAB algorithms; the algorithms in this set are called \emph{feasible}.
 
@@
 Thus, we have a counterpoint to ``much better algorithms win": even under \HardMaxRandom, a better algorithm may lose in competition.
-The corresponding counterpoint to Corollary~\ref{cor:random} states that non-greedy algorithms are \emph{not} chosen in equilibrium.
+A similar counterpoint to Corollary~\ref{cor:random} states that non-greedy algorithms are \emph{not} chosen in equilibrium.
 
@@
 While it is usually more efficient to consider all available data, this modification simplifies analysis.}
-This completes the specification of the modified algorithm; note that it is not merely
+This completes the specification; note that it is not merely
 %another pure strategy in the competition game between the two principals, not
@@
 
-We find that the greedy modification prevails in competition if $p$ is small enough. We focus on \emph{symmetric} response functions: ones that satisfy
-$\respF(x)+\respF(-x)=1$ for any $x\in[0,1]$.
+We find that the greedy modification prevails in competition if $p$ is small enough. We focus on \emph{symmetric} response functions: ones with 
+$f(x)+f(-x)=1$ for any $x\in[0,1]$.
 
@@
 Now, let us show how \alg[1] can be a better algorithm than its greedy modification \alg[2].
-We need to assume that \alg[1] is (much) better than the greedy step \alggr if the latter is computed on a fraction of the data.
+We assume that \alg[1] is (much) better than the greedy step \alggr if the latter is computed on a fraction of the data.
 Formally, we posit a convex function $f(\cdot)$ and $m_0>0$ such that
@@
 This follows from \eqref{eq:BIR-modification} since both \alg[1] and \alggr are \bmonotone.
-The latter follows from a ``monotonicity-in-information" property of the ``greedy step": it can only get better with more information, see Lemma~\ref{lm:MII}.
+The latter follows from the monotonicity of the ``greedy step": it can only get better with more information, see Lemma~\ref{lm:MII}.
 
-Thus, the greedy modification is beneficial in competition \emph{and} keeps us inside the set of \bmonotone MAB algorithms.
-Consequently, if the principals are restricted to choosing \bmonotone MAB algorithms, then non-greedy algorithms cannot be chosen in equilibrium.
+Thus, the greedy modification is beneficial in competition \emph{and} keeps us inside the set of \bmonotone MAB algorithms. Consequently, if the principals are restricted \asedit{to the latter set,}
+%to choosing \bmonotone MAB algorithms, 
+then non-greedy algorithms cannot be chosen in equilibrium.
 
@@
 
-Corollary~\ref{cor:random-greedy} holds more generally: it suffices to assume that feasible algorithms are \bmonotone and closed under the greedy modification.
+%Corollary~\ref{cor:random-greedy} holds more generally: it suffices to assume that feasible algorithms are \bmonotone and closed under the greedy modification.
 The latter means that the greedy modification of any feasible algorithm is also feasible, for any mixing parameter $p$.
@@
     $\respF(\cdot)\in [\eps,1-\eps]$ for some $\eps\in (0,\tfrac12)$,
+\item fair tie-breaking: $\respF(0)=\tfrac12$.
 \item  the response function
@@
  \end{align}
-\item fair tie-breaking: $\respF(0)=\tfrac12$.
 \end{OneLiners}
@@
 \begin{remark}
-This definition is fruitful when parameters $c_0$ and $c_0'$ are close to $\tfrac12$.
+This definition is fruitful when $c_0$ and $c_0'$ are close to $\tfrac12$.
 Throughout, we assume that \alg[1] is better than \alg[2], and obtain results parameterized by $c_0$.
-By symmetry, one could assume that \alg[2] is better than \alg[1], and obtain similar results parameterized by $c_0'$.
+By symmetry, one could assume that \alg[2] is better than \alg[1], and obtain similar results in terms of $c_0'$.
 \end{remark}
 
-For the sake of intuition, let us derive a version of Theorem~\ref{thm:random-clean}, with the same assumptions about the algorithms and essentially the same proof.
+For the sake of intuition, let us derive a version of Theorem~\ref{thm:random-clean}, with the same assumptions and a similar proof.
 The conclusion is much weaker, though: we can only guarantee that each agent $t\geq n_0$ chooses principal 1 with probability slightly larger than $\tfrac12$.
-This is essentially unavoidable in a typical case when both algorithms satisfy $\BIR(n)\to 0$, by Definition~\ref{def:SoftMax}.
+This is essentially unavoidable in a typical case when both algorithms satisfy $\BIR(n)\to 0$.
+%by Definition~\ref{def:SoftMax}.
 
@@
 %\begin{proof}[Proof Sketch]
-To prove this theorem, we follow the steps in the proof of Theorem~\ref{thm:random-clean} to derive \begin{align*}
-\PMR_1(t) - \PMR_2(t)
-    &\geq \BIR_2(t)/2 -q,
-    \quad \text{where $q = \exp(-\eps_0 t/12)$.}
-\end{align*}
-This is at least $\BIR_2(t)/4$ by \eqref{eq:random-assn}. Then \eqref{eq:thm:SoftMax-weak} follows by the smoothness condition \eqref{eq:SoftMax-smooth}.
+To prove this theorem, we follow the steps in the proof of Theorem~\ref{thm:random-clean} to derive 
+%\begin{align*}
+$\PMR_1(t) - \PMR_2(t)
+    \geq \BIR_2(t)/2 -\exp(-\eps_0\, t/12)$.
+%\end{align*}
+This is at least $\BIR_2(t)/4$ by \eqref{eq:random-assn}. Then \refeq{eq:thm:SoftMax-weak} follows by the smoothness condition \eqref{eq:SoftMax-smooth}.
 %\end{proof}
@@
 \begin{definition}
-Suppose MAB algorithms \alg[1],  \alg[2] are $T$-oblivious. Say that
-\alg[1] \emph{weakly BIR-dominates} \alg[2] if there exist absolute  constants $\beta_0, \alpha_0\in (0, 1/2)$ and $n_0\in\N$ such that
+Let \alg[1],  \alg[2] be $T$-oblivious MAB algorithms. Say that
+\alg[1] \emph{weakly BIR-dominates} \alg[2] if there are absolute  constants $\beta_0, \alpha_0\in (0, 1/2)$ and $n_0\in\N$ such that
  \begin{align}\label{eq:SoftMax-better}
@@
 (\exists m_0\; \forall n\geq m_0) \qquad
-\BIR_2(n) \geq \frac{4}{\alpha_0}
-\exp \left( \frac{-n\,\min\{\eps_0, \nicefrac{1}{8}\}}{12}\right).
+\BIR_2(n) \geq \tfrac{4}{\alpha_0}\;
+\exp \rbr{ - \tfrac{1}{12}\, n\,\min\{\eps_0, \nicefrac{1}{8}\} } .
 \end{align}
@@
 It follows that a weakly-BIR-dominating algorithm prevails in equilibrium.
+
+%It follows that a weakly-BIR-dominating algorithm prevails in equilibrium.
 We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity.
 \BReg is a standard notion from the literature (along with \BIR):
-\begin{align}\label{eq:SoftMax-BReg}
-\BReg(n) := n\cdot \E_{\mu\sim\priorMu}
-    \left[ \max_{a\in A} \mu_a\right] -
-    \E\sbr{ \sum_{m=1}^n \rew(m)}
-    = \sum_{m=1}^n \BIR(m).
-\end{align}
+%\begin{align}\label{eq:SoftMax-BReg}
+%\BReg(n) := \textstyle n\cdot \E_{\mu\sim\priorMu}
+%    \left[ \max_{a\in A} \mu_a\right] -
+%    \E\sbr{ \sum_{m=1}^n \rew(m)}
+%    = \sum_{m=1}^n \BIR(m).
+%\end{align}
 
 \begin{corollary}\label{cor:SoftMax-strong}
-Assume \SoftMaxRandom response function. Consider the competition game in which all feasible algorithms are \bmonotone, $T$-oblivious, and satisfy $\BReg(n)\to \infty$.%
-\footnote{$\BReg(n)\to \infty$ is a mild non-degeneracy condition, see Appendix~\ref{app:bg} for background.}
+%Assume \SoftMaxRandom response function. 
+Consider the competition game in which all feasible algorithms are \bmonotone, $T$-oblivious, and satisfy 
+    $\sum_{m=1}^n \BIR(m)\to_n \infty$.%
+\footnote{This is a very mild non-degeneracy condition, see Appendix~\ref{app:bg} for background.}
 Suppose some feasible algorithm \alg weakly-BIR-dominates all others. Then, for any sufficiently large time horizon $T$, there is a unique Nash equilibrium: both principals choose \alg.
@@
 
-We frame our contributions in terms of the relationship between \competitiveness (as expressed by the ``hardness" of the agents' response function $\respF$),
+We frame our contributions in terms of the relationship between \competitiveness (as expressed by the ``hardness" of the response function $\respF$),
  %and \rationality on one side,
@@
 Recall that weak-BIR-dominance is a weaker condition, so that a better algorithm is chosen in a broader range of scenarios.
-Further, {we consider the uniform choice between the principals, which corresponds to the least amount of competition and (when principals' utility is the number of agents) uniform choice provides no incentives to innovate.}%
-\footnote{On the other hand, if principals' utility is aligned with agents' welfare, then a monopolist principal is incentivized to choose the best possible MAB algorithm (namely, to minimize cumulative Bayesian regret $\BReg(T)$).
+We also consider the uniform choice between the principals, which entails the least amount of competition and (when principals optimize market share) provides no incentives to innovate.%
+\footnote{However, if principals' utility is aligned with agents' welfare, then a monopolist principal is incentivized to choose the best possible MAB algorithm (namely, to minimize cumulative Bayesian regret $\BReg(T)$).
 Accordingly, monopoly would result in better social welfare than competition, as the latter is likely to split the market and cause each principal to learn more slowly.
-This is a very generic and well-known effect regarding economies of scale.}
+This is a well-known effect of economies of scale.}
 Thus, we have an inverted-U relationship, see \reffig{fig:inverted-U2}.
@@
 Let us zoom in on the symmetric  \HardMaxRandom model.
-Competitiveness within this model are controlled by the baseline probability $\eps_0 = \respF(- 1)$, which goes smoothly between the two extremes of \HardMax ($\eps_0=0$, tough competition) and the uniform choice ($\eps_0=\tfrac12$, no competition).
-Recall % For clarity, we assume 
-that principal's utility is the number of agents.
+Competitiveness within this model are controlled by the baseline probability $\eps_0 = \respF(- 1)$, which varies  smoothly between the two extremes: \HardMax ($\eps_0=0$, tough competition) and the uniform choice ($\eps_0=\tfrac12$, no competition).
+Principals' utility is the number of agents.
 
@@
 Suppose initially both principals use some algorithm \alg, and principal 1 ponders switching to another algorithm \alg' which BIR-dominates \alg.
-We are interested in the marginal utility $\deltaU$ of this switch.
-Then:
+\asedit{What is the marginal utility $\deltaU$ of this switch?}
 
@@
 
-\item if $\eps_0$ is near $\tfrac12$, then $\deltaU$ is small, as principal 1 gets most agents for free no matter what.
+\item if $\eps_0$ is near $\tfrac12$, the algorithm matters less, so $\deltaU$ is small.
+%as principal 1 gets most agents for free no matter what.
 \end{OneLiners}
@@
 
-\xhdr{Beliefs.} Rather than knowing the principals' algorithms $(\alg[1],\alg[2])$, the Bayesian prior $\priorMu$, and the response function $\respF$, agents could have beliefs on these objects that need not be correct.
-If agents have common ``point beliefs", respectively, algorithms $(\alg[1]',\alg[2]')$ prior $\priorMu'$ and response function $\respF'$,  then all our results carry over with respect to these beliefs.
+\xhdr{Beliefs.} 
+Instead of knowing the principals' algorithms $(\alg[1],\alg[2])$, the Bayesian prior $\priorMu$, and the response function $\respF$, agents could have beliefs on these objects that need not be correct.
+If agents have common ``point beliefs" %respectively, algorithms $(\alg[1]',\alg[2]')$ prior $\priorMu'$ and response function $\respF'$,
+\asedit{on these objects}, then all our results carry over with respect to these beliefs.
 
 \xhdr{Limited non-stationarity in $\respF$.}
-Different agents can have different response functions, in the following sense.
+Different agents can have different response functions.
 For \HardMaxRandom, our results carry over if each agent $t$ has a \HardMaxRandom response function $\respF$ with parameter $\eps_t\geq \eps_0$.
-For \SoftMaxRandom, different agents can have different response functions that satisfy Definition~\ref{def:SoftMax} (with the same parameters).
+For \SoftMaxRandom, different agents can have different response functions that satisfy Definition~\ref{def:SoftMax} with the same parameters.
 
@@
 \xhdr{\BIR can depend on $T$.}
-Many MAB algorithms are parameterized by the time horizon $T$, and their regret bounds usually include $\polylog(T)$. In particular,  a typical regret bound for \BIR is
+Many MAB algorithms are parameterized by the time horizon $T$, and their regret bounds include $\polylog(T)$. In particular,  a typical regret bound for \BIR is
 \begin{align}
@@
 \end{align}
-Here we write $\BIR(n\mid T)$ to emphasize the dependence on $T$. Accordingly, BIR-dominance can be redefined as follows: there exists a number $T_0$ and a function $n_0(T)\in \polylog(T)$
+We write $\BIR(n\mid T)$ to emphasize the dependence on $T$. Accordingly, BIR-dominance can be redefined: there exists a number $T_0$ and a function $n_0(T)\in \polylog(T)$
 such that
@@
 (\forall T\geq T_0,\; n\geq n_0(T)) \quad
-\frac{\BIR_1(\eps_0 n /2\mid T)}{\BIR_2(n\mid T)} <\frac12.
+\BIR_1(\eps_0 n /2\mid T)\,/\, \BIR_2(n\mid T) <\nicefrac12.
 \end{align}
@@
 
-\noindent Weak BIR-dominance can be redefined similarly.
-Theorem~\ref{thm:random-clean} and~\ref{thm:SoftMax-weak} easily extend to these versions.
+\noindent Weak BIR-dominance extends similarly.
+Theorem~\ref{thm:random-clean} and~\ref{thm:SoftMax-weak} easily carry over.
 

### 2021-01-19 journal paper/content/sim_details.tex
Space-saving.

@@
 Concretely, we consider three essential classes of bandit algorithms: ones that never explicitly explore (\emph{greedy algorithms}), ones that explore without looking at the data (\emph{exploration-separating algorithms}), and ones where exploration gradually zooms in on the best arm (\emph{adaptive-exploration algorithms}).
-In the absence of competition, these classes are fairly well-understood: greedy algorithms are terrible for a wide variety of problem instances, exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances, and adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances (see Appendix~\ref{app:bg} for more details).
+In the absence of competition, these classes are fairly well-understood: greedy algorithms are terrible for a wide variety of problem instances, exploration-separated algorithms learn at a reasonable but mediocre rate across all problem instances, and adaptive-exploration algorithms are optimal in the worst case, and exponentially improve for ``easy" problem instances (see Appendix~\ref{app:bg}).
 

### 2021-01-19 journal paper/ec19paper/content/barriers.tex
update figures
Space-saving.

@@
 \end{tabular}
-\caption{Data advantage vs.
+\caption{\footnotesize Data advantage vs.
 reputation advantage experiment, on Heavy-Tail MAB instance.
@@
 \begin{finding}\label{barrier-find}
-\textit{\\
+\textit{
 (a) Data advantage and reputation advantage alone are large, across all algorithms and MAB instances.
-\\(b) The data advantage is larger than the reputation advantage when the incumbent chooses \Thompson.
-\\(c) The two advantages are similar in magnitude when the incumbent chooses \DynamicEpsGreedy or \DynamicGreedy.
+(b) The data advantage is larger than the reputation advantage when the incumbent chooses \Thompson.
+(c) The two advantages are similar in magnitude when the incumbent chooses \DynamicEpsGreedy or \DynamicGreedy.
 }

### 2021-01-19 journal paper/ec19paper/content/inverted_u.tex
Space-saving.

--- 2020-11-25 journal paper/ec19paper/content/inverted_u.tex
@@
 
-Our main experiments are with the duopoly game defined in Section~\ref{sec:model}.
-As the ``intensity of competition" varies from  monopoly to ``incumbent" to simultaneous entry duopoly to ``late entrant", we find a stylized inverted-U relationship as in Section~\ref{sec:theory-welfare}.
-More formally, we look for equilibria in the duopoly game, where each firm's choices are limited to \DynamicGreedy, \DynamicEpsGreedy and \Thompson.
+Our main experiments concern the duopoly game defined in Section~\ref{sec:model}.
+As the ``intensity of competition" varies from  monopoly to ``incumbent" to ``simultaneous entry" to ``late entrant", we find a stylized inverted-U relationship as in Section~\ref{sec:theory-welfare}.
+We look for equilibria in the duopoly game, where each firm's choices are limited to \DynamicGreedy, \DynamicEpsGreedy and \Thompson.
 We do this for each ``intensity level" and each MAB instance, and look for findings that are consistent across MAB instances.
-For cleaner results, we break ties towards less advanced algorithms (as they tend to have lower adoption costs, see \cite{DS-arxiv}).
-\DynamicGreedy is trivially the dominant strategy under monopoly.
+%For cleaner results, 
+We break ties towards less advanced algorithms, as they tend to have lower adoption costs \citep{DS-arxiv}. \DynamicGreedy is then the dominant strategy under monopoly.
 
@@
 In fact, incentives for \DynamicGreedy get stronger when the model transitions from simultaneous entry to ``late entrant".%
-\footnote{For the Heavy-Tail instance, \DynamicGreedy goes from a weakly dominant strategy to a strictly dominant one.
-For the Uniform instance, \DynamicGreedy goes from a Nash equilibrium strategy to a weakly dominant one.}
+\footnote{For the Heavy-Tail instance, \DynamicGreedy goes from a weakly dominant strategy to a strictly dominant.
+For the Uniform instance, \DynamicGreedy goes from a Nash equilibrium strategy to a weakly dominant.}
 Finally, the middle level of competition, ``incumbent" in the first-mover regime creates strong incentives for \Thompson.
@@
 \textit{
-Under simultaneous entry, \Eeog values tend to be much smaller than the time horizon $T$.
+Under simultaneous entry, \Eeog values tend to be much smaller than $T$.
 }
@@
 We conjecture this is because larger $T_0$ tends to be more beneficial for a better algorithm (as it tends to follow a better learning curve).
-Indeed, we know that the ``effective end of game" in this scenario typically occurs when a better algorithm loses, and helping it delays the loss.
+Indeed, we know that the \Eeog in this scenario typically occurs when a better algorithm loses, and helping it merely delays the loss.
 
@@
 \emph{market regret}:
-\[ \textstyle T\, \max_a \mu(a) - \sum_{t\in [T]} \mu(a_t), \]
+$ \textstyle T\, \max_a \mu(a) - \sum_{t\in [T]} \mu(a_t), $
 where $a_t$ is the arm chosen by agent $t$. This is a standard performance measure in the literature on multi-armed bandits. Note that smaller regret means higher welfare.
@@
 \DynamicGreedy in the monopoly scenario, and
-\DynamicGreedy for both firms in simultaneous entry scenario (Finding \ref{find:duopoly}).
+\DynamicGreedy for both firms for simultaneous entry (Finding \ref{find:duopoly}).
 For the first-mover scenario, it is \Thompson for the incumbent (Finding \ref{find:temp-monopoly}) and \DynamicGreedy for the entrant (Finding \ref{find:temp-monopoly-entrant}).

### 2021-01-19 journal paper/ec19paper/content/non_greedy_choice.tex
Lots of edits to address Yishay's comments (all except limited rationality).
Space-saving.

@@
 
-\subsection{Non-deterministic choice Model (\HardMaxRandom)}\label{sec:non_greedy}
+\subsection{Non-deterministic choice model (\HardMaxRandom)}\label{sec:non_greedy}
 
@@
 \begin{finding}\label{find:non_greedy_choice}
-\textit{\TS is weakly dominant under \HardMaxRandom, if and only if $\eps T$ is sufficiently large. Moreover, \HardMaxRandom leads to lower variance in market share, compared to \HardMax.}
+\textit{\Thompson is weakly dominant under \HardMaxRandom, if and only if $\eps T$ is sufficiently large. Moreover, \HardMaxRandom leads to lower variance in market share, compared to \HardMax.}
 \end{finding}

### 2021-01-19 journal paper/ec19paper/content/perf_in_iso.tex
pass through alex's edits
space saver
Space-saving.

@@
 Figure \ref{prelim_means} shows these trajectories for the Needle-in-Haystack instance; for other MAB instances the plots are similar.
-
 We summarize this finding as follows:
@@
 \caption{\footnotesize Relative reputation trajectory for $\Thompson$ vs $\DynamicGreedy$, on Uniform instance (left) and Needle-in-Haystack instance (right).
-Shaded area display 95\% confidence intervals.}
+Shaded area display 95\% confidence intervals.
+The relative reputation at time $t$ is the fraction of \MRVs for which, at time $t$, $\Thompson$ has a higher reputation score than $\DynamicGreedy$.}
 \label{relative_rep_plots}
@@
 \DynamicGreedy dips below $\tfrac12$ can be seen as an explanation for the competitive disadvantage of exploration.
-Such period also exists for the Heavy-Tail MAB instance.
+Such period also exists for the Heavy-Tail instance.
 However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots}.%
-\footnote{We see two explanations for this: $\Thompson$ identifies the best arm faster for the Needle-in-Haystack instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}
+\footnote{We see two explanations for this: $\Thompson$ identifies the best arm faster for the Needle-in-Haystack instance, and there are no ``very bad" arms to make exploration expensive in the near term.}
 
@@
 
-\noindent Uniform and Heavy-tail instance are exploration-disadvantaged, but Needle-in-Haystack is not.
+\noindent Note that Uniform and Heavy-tail instances are exploration-disadvantaged, but Needle-in-Haystack instance is not.
 

### 2021-01-19 journal paper/ec19paper/content/revisited.tex
Space-saving.

--- 2020-09-24 journal paper/ec19paper/content/revisited.tex
@@
 \includegraphics[scale=0.35]{ec19paper/figures/ts_dg_rep_diff_nih}
-\caption{\footnotesize Reputation scores for Needle-in-Haystack at $t=500$ (left), Reputation difference $\Thompson - \DynamicGreedy$ for Needle-in-Haystack (right).
-Both are smoothed using a kernel density estimate.}
+\caption{\footnotesize Needle-in-Haystack: reputation scores at $t=500$ (left), reputation difference $\Thompson - \DynamicGreedy$ (right). Both are smoothed using a kernel density estimate.}
 \label{fig:rep_dist_nih}
@@
 That is, consider the empirical distribution of this score over different \MRVs.%
-\footnote{Recall that each \MRV in our experimental setup comes with one specific realization table.} For concreteness, consider the Needle-in-Haystack instance at time $t=500$, plotted in Figure~\ref{fig:rep_dist_nih} (left).
+\footnote{Recall that each \MRV in our experimental setup comes with a separate realization table.} For concreteness, consider the Needle-in-Haystack instance at time $t=500$, plotted in Figure~\ref{fig:rep_dist_nih} (left).
 (The other MAB instances lead to a similar intuition.)
@@
 We observe that the distribution is skewed to the right, precisely due to the fact that $\DynamicGreedy$ either does slightly better than $\Thompson$ or does substantially worse.
-Therefore, the mean is not a good measure of the central tendency, or typical value, of this distribution.
+So, the mean is not a good measure of the central tendency of this distribution.
 

### 2021-01-19 journal paper/content/related_work.tex
trim some refs

@@
 
-%A long line of work from electrical engineering and computer science, starting from
-%\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10,MultiPlayerMAB-Anima11}, focuses on competition for resources, not competition for consumers.
+A long line of work from electrical engineering and computer science, starting from
+\citet{MultiPlayerMAB-Poor08,MultiPlayerMAB-Liu10}, focuses on competition for resources, not competition for consumers.
 Specifically, this literature targets an application to \emph{cognitive radios}, where multiple radios transmit simultaneously in a shared medium and compete for bandwidth.

### 2021-01-19 journal paper/content/related_work.tex
Edits for spacing.

@@
 Dedicated monographs \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} cover the work on regret-minimizing formulations (which mainly comes from computer science). The classic book
-\citep{Gittins-book11} focuses on the Markovian formulations, which predate regret-minimization. Various connections to economics and game theory are detailed in
+\citep{Gittins-book11} focuses on the Markovian formulations, which predate regret-minimization. Connections to economics are detailed in
 books \citep{CesaBL-book,slivkins-MABbook} and surveys \citep{Bergemann-survey06,Horner-survey16}. Industrial applications are discussed in \citep{DS-arxiv}.
@@
 dynamic auctions
-    (\eg see \citet{DynAuctions-survey10} for a survey),
+    \citep{DynAuctions-survey10},
     %\citep[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},

### 2021-01-19 journal paper/content/sec-intro.tex
Edits for spacing.

@@
 While agents' rationality and severity of competition are often modeled separately, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}
-We also find another, technically different inverted-U relationship which zeroes in on the \HardMaxRandom model.
+Another, technically different inverted-U relationship zeroes in on the \HardMaxRandom model.
 %We vary rationality/competitiveness inside this model, and track the marginal utility of switching to a better algorithm.
 
-While traditional models of innovation study lab-based R\&D, we consider data-driven innovation: one that crucially depends on data generated by the firm's customers.
+While traditional models of innovation study lab-based R\&D, we consider data-driven innovation, which crucially depends on data generated by the firm's customers.
 We focus on innovation in \emph{exploration technology} which systematically improves the firm's products, whereas prior work would define innovation as improvement in the products themselves.
-We recover the traditional inverted-U relationship purely through the reputational consequences of exploration.
+We recover the \asdelete{traditional} inverted-U relationship purely through the reputational consequences of exploration.
 \asdelete{rather than monetary costs/benefits thereof.}
-This is in contrast to the inverted-U relationships established in prior work, which rely on the monetary aspects: how much can a firm invest into  R\&D and how much can it profit from the innovation.
+By contrast, the inverted-U relationships from prior work rely on the monetary aspects: \asedit{investments into  R\&D and profits from innovation.}
+
+%how much can a firm invest into  R\&D and how much can it profit from the innovation.
 
@@
 
-\textbf{(ii)} Focusing on \HardMax, we investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous arrival to late-arriver.
-We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous arrival).
+\textbf{(ii)} Focusing on \HardMax, we investigate the first-mover advantage as a different channel to vary the intensity of competition: from the first-mover to simultaneous entry to late-arriver.
+We find that the first-mover is incentivized to choose a more advanced exploration algorithm, whereas the late-arriver is often incentivized to choose the ``greedy algorithm" (more so than under simultaneous entry).
 Consumer welfare is higher under early/late arrival than under simultaneous entry.

### 2021-01-19 journal paper/content/conclusion.tex
space saving edits

@@
 We study the tension between exploration and competition.
-We consider a stylized duopoly model in which two firms face an identical multi-armed bandit problem and compete for a stream of users.
-A firm makes progress on its learning problem if and only if it attracts users.
+We consider a stylized duopoly model in which two firms face \gadelete{an identical}\gaedit{the same} multi-armed bandit problem and compete for a stream of users.
+A firm makes progress on its learning problem\gadelete{if and} only if it attracts users.
 %\asdelete{We investigate two variants: an analytical variant where users do not observe any performance signals and choose between firms according to their Bayesian-expected rewards, and numerical simulations where a reputation score is observed for each firm, based on the average reward of its recent users.
@@
 We find that firms are incentivized to adopt a ``greedy algorithm" which does no purposeful exploration and leads to welfare losses for users.
-We then consider two relaxations of competition: we soften users' decision rule and give one of the firms a first-mover advantage.
+We then consider two relaxations of competition: we soften users' decision rule and give one\gadelete{of the firms} firm a first-mover advantage.
 Both relaxations induce firms to adopt ``better" bandit algorithms, which benefits user welfare.

### 2021-01-19 journal paper/content/sim_details.tex
space saving edits

@@
 For each \MRV we draw a table of realized rewards (\emph{realization table}), and use this same table for all experiments on this \MRV.
-This ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
+This ensures that differences in algorithm performance are \gadelete{not due to noise in the realizations but }\gaedit{solely due to} differences in the algorithms in the different experimental settings.
 

### 2021-01-19 journal paper/ec19paper/content/inverted_u.tex
space saving edits

@@
 In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U-expts}.%
-\footnote{We consider the monopoly scenario for comparison only, without presenting any findings. We just assume that a monopolist chooses
+\footnote{We consider the monopoly scenario for comparison only.\gadelete{, without presenting any findings.}We just assume that a monopolist chooses
 the greedy algorithm, because it is easier to deploy in practice.
@@
 Instead, the firm wants to improve its performance as much as possible by the time competition starts.
-Essentially, the firm only faces a classical explore-exploit trade-off, and is incentivized to choose algorithms that are best at optimizing this trade-off.
+Essentially, the firm only faces a classical explore-exploit trade-off, and \gadelete{is incentivized to }choose\gaedit{s} algorithms that are best at optimizing this trade-off.
 

### 2021-01-19 journal paper/ec19paper/content/revisited.tex
space saving edits

@@
 In the former case \DynamicGreedy does slightly better than $\Thompson$, and in the latter case it does substantially worse.
-However, the mean reputation trajectory may fail to capture this complexity since it simply takes average over different \MRVs.
-This may be inadequate for explaining the outcome of the duopoly game, given that the latter is determined by a simple comparison between the firm's reputation scores.
+However, the mean reputation trajectory \gadelete{may fail}\gaedit{fails} to capture this complexity since it \gadelete{simply }takes average over different \MRVs.
+This \gaedit{is}\gadelete{may be} inadequate for explaining the outcome of the duopoly game, given that the latter is determined by a \gadelete{simple} comparison between the firm's reputation scores.
 

### 2021-01-19 journal paper/content/related_work.tex
more space saving edits

@@
 Each radio chooses channels over time using a multi-armed bandit algorithm.
-This work studies a repeated game between bandit algorithms, and focuses on designing algorithms which work well in this game, under various assumptions on communication, synchronization and collisions.
+This work studies a repeated game between bandit algorithms, and focuses on designing algorithms which work well in this game\gadelete{, under various assumptions on communication, synchronization and collisions}.
 
@@
 
-%\asedit{While we use first-mover advantage and agents' decision rule,} classic ``market competitiveness" measures, such as the Lerner Index or the Herfindahl-Hirschman Index
-%\citep{tirole1988theory}, are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).
+\asedit{While we use first-mover advantage and agents' decision rule,} classic ``market competitiveness" measures, such as the Lerner Index or the Herfindahl-Hirschman Index
+\citep{tirole1988theory}, are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).
 

### 2021-01-19 journal paper/content/sec-model.tex
more space saving edits

@@
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.) Principals are risk-neutral and their utility is defined as their market share, \ie the number of agents that chose this principal.
-Thus they aim to select the algorithm that maximizes their expected market share.
+(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.) Principals are risk-neutral and their utility is defined as their market share, \ie the number of agents that \gaedit{choose them.}\gadelete{chose this principal.} Thus they aim to select the algorithm that maximizes their expected market share.
 

### 2021-01-19 journal paper/content/sec-theory.tex
more space saving edits

@@
 Suppose algorithms \alg[1], \alg[2] are \bmonotone and $T$-oblivious, and \eqref{eq:random-assn} holds.
-If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible selection probability for this response function).
+If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (\gadelete{which is }the largest possible selection probability for this response function).
 \end{theorem}
@@
 Fix a \HardMaxRandom response function $\respF$.
-Consider the competition game in which all feasible MAB algorithms are $T$-oblivious, \bmonotone, and satisfy \eqref{eq:random-assn} for some fixed $m_0$.
+Consider the competition game \gaedit{where}\gadelete{in which} all feasible MAB algorithms are $T$-oblivious, \bmonotone, and satisfy \eqref{eq:random-assn} for some fixed $m_0$.
 Suppose some feasible algorithm \alg  BIR-dominates all other feasible algorithms, starting from some local step $n_0$.
-Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
+Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium\gadelete{, in which}\gaedit{ where} both principals choose algorithm \alg.
 \end{corollary}

### 2021-01-19 journal paper/content/related_work.tex
Space-saving.

@@
 
-The three-way tradeoff between exploration, exploitation and incentives has been studied in several scenarios very different from ours:
+\asedit{The tension between exploration and incentives}
+%The three-way tradeoff between exploration, exploitation and incentives 
+has been studied in several scenarios very different from ours:
 incentivized exploration in recommendation systems

### 2021-01-19 journal paper/content/sec-model.tex
Space-saving.

--- 2021-01-19 journal paper/content/sec-model.tex
@@
 \end{center}
-\caption{The three models for $\respF$: \HardMax is thick blue, \HardMaxRandom is red, and \SoftMaxRandom is dashed.}
+\caption{The models for $\respF$: \HardMax is thick blue, \HardMaxRandom is red, and \SoftMaxRandom is dashed.}
 \label{fig:response-functions}
@@
 Each agent knows its global round $t$, but not the performance signals such as the current market shares.
-Agents' reward estimates are defined as posterior mean rewards:
+Her reward estimates are defined as posterior mean rewards:
         $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i}$
-    for each agent $t$ and principal $i$.%
-\footnote{Agents do the Bayesian update knowing $t$, the principals' algorithms, the prior $\priorMu$, and $\respF$.}
+    for each principal $i$, \asedit{where the agent knows $t$, the principals' algorithms, $\priorMu$, and $\respF$.}
+%\footnote{Agents do the Bayesian update knowing $t$, the principals' algorithms, the prior $\priorMu$, and $\respF$.}
 
@@
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.) Principals are risk-neutral and their utility is defined as their market share, \ie the number of agents that chose this principal.
-Thus they aim to select the algorithm that maximizes their expected market share.
+(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.) Principals are risk-neutral; their utility is defined as their market share, \ie the number of agents \asedit{they attract}.
+Thus, they aim to select the algorithm that maximizes their expected market share.
 

### 2021-01-19 journal paper/content/sec-model.tex
Space-saving

@@
 Firms compete only on the quality of service, rather than, say, pricing or the range of products.
-Agents are myopic: they do not worry about how their actions impact their future utility.\footnote{So, agents do not attempt to learn over time, second-guess or game future agents, or manipulate the principals' learning algorithms.
+Agents are myopic: they do not worry about how their actions impact their future utility.\footnote{So, agents do not attempt to learn over time, game future agents, or manipulate the principals' learning algorithms.
 This is arguably typical in practice, in part because one agent's influence tends to be small.}  Various performance signals available to the users, from personal experience to word-of-mouth to consumer reports, are abstracted as persistent ``reputation scores" reflecting the current reputation, and further simplified to average performance over a sliding time window.

### 2021-01-19 journal paper/content/sec-theory.tex
Space-saving
space-saving.

--- 2021-01-19 journal paper/content/sec-theory.tex
@@
 For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen.
-Note that
-    \[ \PMR_i(t):=  \E\sbr{ r_t\mid i_t = i }
-        = \E\sbr{ \rew_i(n_i(t)+1) }. \]
+We will use the fact that
+    $ \PMR_i(t):=  \E\sbr{ r_t\mid i_t = i }
+        = \E\sbr{ \rew_i(n_i(t)+1) }. $
 
@@
 (We treat the number of arms as a constant.) Intuitively, (much) better MAB algorithms tend to have a (much) smaller BIR, see Appendix~\ref{app:bg} for background.
-
 An algorithm is called \emph{\bmonotone} if it can only get better over time, in the Bayesian sense: namely, if $\E[\rew_i(\cdot)]$ is non-decreasing, and therefore $\BIR(\cdot)$ is non-decreasing.
@@
 Suppose algorithms \alg[1], \alg[2] are \bmonotone and $T$-oblivious, and \eqref{eq:random-assn} holds.
-If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible selection probability for this response function).
+If \alg[1] \emph{BIR-dominates} \alg[2] starting from step $n_0$, then each agent $t\geq \max(n_0,m_0)$ chooses principal $1$ with probability $\respF(1) = 1- \eps_0$ (which is the largest possible probability for this response function).
 \end{theorem}
@@
 Suppose some feasible algorithm \alg  BIR-dominates all other feasible algorithms, starting from some local step $n_0$.
-Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium, in which both principals choose algorithm \alg.
+Then, for any sufficiently large time horizon $T$, this game has a unique Nash equilibrium: both principals choose algorithm \alg.
 \end{corollary}

### 2021-01-19 journal paper/content/sec-theory.tex

@@
 Mean rewards may also have a known structure, such as Lipschitzness, convexity, or linearity; such structure can be incorporated via $\prior$.
-All these extensions have been studied extensively in the literature on MAB, and account for a substantial segment thereof.
-Background can be found in any of the recent books on MAB \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
+All these extensions have been studied extensively in the literature on MAB, and account for a substantial segment thereof;
+see \citep{slivkins-MABbook,LS19bandit-book} for background.
+
+%Background can be found in any of the recent books on MAB \citep{Bubeck-survey12,slivkins-MABbook,LS19bandit-book}.
 

### 2021-01-20 journal paper/ecma.tex
updates to TOC
Edits in the abstract.

--- 2021-01-16 journal paper/ecta.tex
@@
 }
-
+\usepackage{minitoc}
+\usepackage{times}
+\usepackage[toc,page,header]{appendix}
 \usepackage{subfiles}
-\usepackage{libertine}
-\usepackage[title,titletoc]{appendix}
 \usepackage{url}            % simple URL typesetting
@@
 \newcommand{\StaticGreedy}{\term{StaticGreedy}}
-\newcommand{\DynGreedy}{\term{BayesianGreedy}}
+\newcommand{\DynGreedy}{\term{BayesGreedy}}
 \newcommand{\DynamicGreedy}{\DynGreedy}
-\newcommand{\DynamicEpsGreedy}{\term{BayesianEpsilonGreedy}}
+\newcommand{\DynamicEpsGreedy}{\term{BayesEpsilonGreedy}}
 \newcommand{\Thompson}{\term{ThompsonSampling}}
@@
 \addauthor{sw}{orange} % sw for Steven
-\addauthor{ym}{madenta} % sw for Steven
+\addauthor{ym}{magenta} % ym for Yishay
 % e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
@@
 Appendices~\ref{app:bg},\ref{app:examples} are new.
-We would like to thank Ian Ball, Yeon-Koo Che, Sara Shahanaghi, Glen Weyl and Sven Rady for helpful comments and conversations.
+We would like to thank Ian Ball, Yeon-Koo Che, Sven Rady, Sara Shahanaghi and Glen Weyl for helpful comments and conversations.
 Likewise, we thank the audience of the conference and seminar talks on this work.
@@
 \footnote{Columbia University, Department of Economics. Email: g.aridor@columbia.edu.}
-\and \hspace{-0.75cm}
+\and \hspace{-0.5cm}
 \rule{0.0in}{0pt}
@@
 \footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com.}
-\and \hspace{-0.75cm}
+\and \hspace{-0.5cm}
 \rule{0.0in}{0pt}
-Aleksandrs Slivkins%
+Aleksandrs Slivkins
 \footnote{Microsoft Research New York City. Email: slivkins@microsoft.com.}
-\and \hspace{-0.75cm}
+\and \hspace{-0.5cm}
 \rule{0.0in}{0pt}
@@
 \begin{abstract}
-Most online platforms strive to learn from interactions with users, and engage in \emph{exploration}: making potentially suboptimal choices in order to acquire new information.
+Most online platforms learn from interactions with users, and engage in \emph{exploration}: making potentially suboptimal choices in order to acquire new information.
 We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and competition for users.
@@
 
-We consider a stylized duopoly model in which two firms face the same multi-armed bandit problem.
+We consider a stylized duopoly in which two firms face the same multi-armed bandit problem.
 Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit problem only if it is chosen.
 We study whether competition incentivizes the adoption of better algorithms.
-We find that stark competition induces firms to commit to a ``greedy" algorithm that leads to low welfare.
-However, weakening competition by providing firms with some ``free" users incentivizes better exploration strategies and increases welfare.
-We investigate two channels for weakening the competition: stochastic user choice models and giving one firm a first-mover advantage.
+We find that stark competition disincentivizes exploration, leading to low welfare.
+%induces firms to commit to a ``greedy" algorithm that leads to low welfare.
+However, weaker competition 
+%by providing firms with some ``free" users 
+incentivizes better exploration algorithms and increases welfare.  We investigate two channels for weakening the competition: stochastic user choice models and a first-mover advantage.
+Our findings speak to the competition-innovation relationship and the first-mover advantage in the digital economy.
 %Our findings are closely related to the ``competition vs. innovation" relationship, and elucidate the first-mover advantage in the digital economy.
@@
 \newpage
+\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}
 %\begin{small}
@@
 %\end{small}
-\setcounter{page}{0}
+\setcounter{page}{1}
 %\thispagestyle{empty}
@@
 
-\begin{appendices}
+      % body goes here
+\appendix
+\renewcommand{\contentsname}{Appendix, For Online Publication Only}
+
+\tableofcontents
+\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}
+      % appendices go here
+
+\newpage
 
@@
 
-\end{appendices}
 \end{document}

### 2021-03-02 journal paper/journal_main.tex
fix journal_main

--- 2020-12-21 journal paper/journal_main.tex
@@
 The Perils of Exploration under Competition%
-\thanks{This is a merged and final version of two conference publications,
-\citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19},
-with a unified and streamlined presentation and expanded related work and background materials.
-All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
-Appendices~\ref{app:bg},\ref{app:examples} are completely new compared to the conference versions.\newline \indent
-We would like to thank Ian Ball, Yeon-Koo Che, Sara Shahanaghi, Glen Weyl and Sven Rady for helpful comments and conversations.
+\thanks{Preliminary abstracts have appeared as
+\citet*{CompetingBandits-itcs18} and \citet*{CompetingBandits-ec19}.
+This is a merged version, with a unified and streamlined presentation and expanded related work and background materials.
+\newline \indent %All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
+Appendices~\ref{app:bg},\ref{app:examples} are new.
+We would like to thank Ian Ball, Yeon-Koo Che, Sven Rady, Sara Shahanaghi and Glen Weyl for helpful comments and conversations.
 Likewise, we thank the audience of the conference and seminar talks on this work.
@@
 Yishay Mansour
-\footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com}
+\footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com.\newline
+Supported in part by Israel Science Foundation grant number 993/17.}
 \and \hspace{-0.75cm}
 \rule{0.0in}{0pt}
-Aleksandrs Slivkins%
-\footnote{Microsoft Research New York City. Email: slivkins@microsoft.com}
+Aleksandrs Slivkins
+\footnote{Microsoft Research New York City. Email: slivkins@microsoft.com.}
 \and \hspace{-0.75cm}
@@
 Zhiwei Steven Wu%
-\footnote{Carnegie Mellon University, Pittsburgh, PA.
+\footnote{Carnegie Mellon University, Pittsburgh, PA
 Email: zstevenwu@cmu.edu.\newline
@@
 }
-\date{First version: July 2020\\ This version: December 2020}
+\date{}
+%\date{First version: July 2020\\ This version: December 2020}
 
@@
 \begin{abstract}
-Most online platforms strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information.
-We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users.
-Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
+Most online platforms learn from interactions with users, and engage in \emph{exploration}: making potentially suboptimal choices in order to acquire new information.
+We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and competition for users.
+%Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
 
-We consider a stylized duopoly model in which two firms face the same multi-armed bandit problem.
+We consider a stylized duopoly in which two firms face the same multi-armed bandit problem.
 Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit problem only if it is chosen.
-Through a mix of theoretical results and numerical simulations, we study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for users.
-We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low welfare.
-However, weakening competition by providing firms with some ``free" users incentivizes better exploration strategies and increases welfare.
-We investigate two channels for weakening the competition: relaxing the rationality of users and giving one firm a first-mover advantage.
-Our findings are closely related to the ``competition vs.
-innovation" relationship, and elucidate the first-mover advantage in the digital economy.
+We study whether competition incentivizes the adoption of better algorithms.
+We find that stark competition disincentivizes exploration, leading to low welfare.
+%induces firms to commit to a ``greedy" algorithm that leads to low welfare.
+However, weaker competition
+%by providing firms with some ``free" users
+incentivizes better exploration algorithms and increases welfare.  We investigate two channels for weakening the competition: stochastic user choice models and a first-mover advantage.
+Our findings speak to the competition-innovation relationship and the first-mover advantage in the digital economy.
+%Our findings are closely related to the ``competition vs. innovation" relationship, and elucidate the first-mover advantage in the digital economy.
+\\
 
-\vspace{0.2in}
+\vspace{0.1in}
+\noindent\textbf{Keywords:}  Competition vs. innovation, exploration-exploitation tradeoff, multi-armed bandits, regret. \\
+\noindent\textbf{JEL Codes}: D83, L15, O31
 
@@
 \newpage
-\begin{small}
-\setcounter{tocdepth}{2}
-\tableofcontents
-\end{small}
-\setcounter{page}{0}
-\thispagestyle{empty}
-\newpage
+\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}
+%\begin{small}
+%\setcounter{tocdepth}{2}
+%\tableofcontents
+%\end{small}
+\setcounter{page}{1}
+%\thispagestyle{empty}
+%\newpage
 %
@@
 
-\begin{appendices}
+      % body goes here
+\appendix
+\renewcommand{\contentsname}{Appendix}
+
+\tableofcontents
+\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}
+      % appendices go here
+
+\newpage
 

### 2021-03-02 journal paper/abstract-long.tex
Moved abstract(s) to separate file(s).

Most online platforms strive to learn from interactions with users, and many engage in \emph{exploration}: making potentially suboptimal choices for the sake of acquiring new information. We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and the competition for users. Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.

We consider a stylized duopoly model in which two firms face the same multi-armed bandit problem. Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit problem only if it is chosen. Through a mix of theoretical results and numerical simulations, we study whether and to what extent competition incentivizes the adoption of better bandit algorithms, and whether it leads to welfare increases for users. We find that stark competition induces firms to commit to a ``greedy" bandit algorithm that leads to low welfare. However, weakening competition by providing firms with some ``free" users incentivizes better exploration strategies and increases welfare.  We investigate two channels for weakening the competition: relaxing the rationality of users and giving one firm a first-mover advantage. Our findings are closely related to the ``competition vs. innovation" relationship, and elucidate the first-mover advantage in the digital economy.

### 2021-03-02 journal paper/abstract-short.tex
Moved abstract(s) to separate file(s).

--- 2021-01-20 journal paper/ecma.tex
@@
-\documentclass[12pt]{article}
-\usepackage[margin=1.25in]{geometry}
-
-\usepackage{color,
-setspace,sectsty,comment,footmisc,caption,
-pdflscape,subcaption,array,hyperref,adjustbox
-}
-\usepackage{minitoc}
-\usepackage{times}
-\usepackage[toc,page,header]{appendix}
-\usepackage{subfiles}
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{amsmath,amssymb,amsfonts,amsthm}
-\usepackage[round]{natbib} % cannot use with AAAI! A.S.
-
-\usepackage{float}
-\usepackage{sgame, tikz} % Game theory packages
-\usepackage{algorithm,algpseudocode}
-\usepackage{makecell}
- \usepackage{multirow}
- \usepackage{graphicx}
-
-\usepackage{booktabs} % For formal tables
-
-%\theoremstyle{definition}
-%\newtheorem{definition}{Definition}
-\newtheorem{finding}{Finding}
-%\newtheorem{conjecture}{Conjecture}
-\newtheorem{puzzle}{Puzzle}
-\captionsetup{font=footnotesize}
-
-\usepackage{slivkins-setup,slivkins-theorems}
-\renewcommand{\xhdr}[1]{\vspace{1mm} \noindent{\bf #1}}
-
-% model variants
-\newcommand{\TheoryModel}{Bayesian-choice model\xspace}
-\newcommand{\ExptsModel}{reputation-choice model\xspace}
-
-% Commands from EC 19 paper
-\newcommand{\term}[1]{\ensuremath{\mathtt{#1}}\xspace}
-
-% shorthand for algorithms
-\newcommand{\TS}{\term{TS}}    % Thompson Sampling
-\newcommand{\DEG}{\term{BEG}}  % "Dynamic eps-greedy"
-\newcommand{\DG}{\term{BG}}    % "Dynamic Greedy"
-
-% shorthand for models
-\newcommand{\HMR}{\term{HMR}} % HardMaxRandom
-\newcommand{\HM}{\term{HM}}    % HardMax
-
-\newcommand{\Beta}{\term{Beta}} % for Beta distribution
-\newcommand{\Eeog}{\term{EoG}} % shorthand for "effective end of game"
-
-\newcommand{\MRV}{mean reward vector\xspace} % mean reward vector
-\newcommand{\MRVs}{mean reward vectors\xspace} % mean reward vector
-
-% commands from "theory"
-
-% in connection with "a little greedy goes a long way"
-%     stuff for the "greedy modification".
-\newcommand{\gr}{\texttt{gr}}
-\newcommand{\BIRgr}{\BIR^{\gr}}
-\newcommand{\alggr}{\alg[\gr]}
-\newcommand{\rewgr}{\rew_{\gr}}
-
-\DeclareMathOperator*{\Expectation}{\mathbb{E}}
-\newcommand{\Ex}[2]{\Expectation_{#1}\left[#2\right]}
-\newcommand{\FiniteGame}{finite competition game\xspace}
-
-% notation
-%%% advanced notation
-\newcommand{\OPT}{\term{OPT}}
-\newcommand{\rew}{\term{rew}}  % Bayesian-expected reward after n local rounds
-\newcommand{\PMR}{\term{PMR}} % posterior mean reward
-\newcommand{\support}{\term{support}}
-
-\newcommand{\BIR}{\term{BIR}} % Bayesian Instantaneous Regret
-
-% response function
-\newcommand{\respF}{f_{\term{resp}}}
-\newcommand{\respEps}{\eps_\term{resp}}
-
-\newcommand{\BReg}{\term{BReg}}
-\newcommand{\HardMax}{\term{HardMax}}
-\newcommand{\HardMaxRandom}{\term{HardMax\&Random}}
-\newcommand{\SoftMaxRandom}{\term{SoftMax}}
-\newcommand{\Uniform}{\term{Uniform}}
-\newcommand{\Random}{\term{Random}}
-
-% algorithms
-\newcommand{\StaticGreedy}{\term{StaticGreedy}}
-\newcommand{\DynGreedy}{\term{BayesGreedy}}
-\newcommand{\DynamicGreedy}{\DynGreedy}
-\newcommand{\DynamicEpsGreedy}{\term{BayesEpsilonGreedy}}
-\newcommand{\Thompson}{\term{ThompsonSampling}}
-
-% "monotone algorithms"
-\newcommand{\bmonotone}{Bayesian-monotone\xspace}
-\newcommand{\bmonotonicity}{Bayesian-monotonicity\xspace}
-
-\newcommand{\termSub}[2]{\ensuremath{\mathtt{#1}_{#2}}\xspace}
-\newcommand{\alg}[1][]{\termSub{alg}{#1}}
-%\newcommand{\prin}[1][]{\termSub{prin}{#1}}  % principal
-\newcommand{\agent}[1][]{\termSub{agent}{#1}}
-
-% priors and posteriors
-\newcommand{\prior}{\ensuremath{\mP}\xspace}
-\newcommand{\priorMu}{\ensuremath{\prior_\mathtt{mean}}\xspace}
-\newcommand{\posteriorN}[2]{\mN_{#1,#2}}  % \posteriorN{principal}{round}
-
-% rationality / innovation / competition
-\newcommand{\termTXT}[1]{{\em {#1}}\xspace}
-
-\newcommand{\rationality}{\termTXT{rationality}}
-\newcommand{\Rationality}{\termTXT{Rationality}}
-\newcommand{\innovation}{\termTXT{innovation}}
-\newcommand{\Innovation}{\termTXT{Innovation}}
-\newcommand{\competition}{\termTXT{competition}}
-\newcommand{\Competition}{\termTXT{Competition}}
-\newcommand{\competitiveness}{\termTXT{competitiveness}}
-\newcommand{\exploration}{\termTXT{exploration}}
-\newcommand{\Exploration}{\termTXT{Exploration}}
-
-% a very useful package for edits and comments, from David Kempe (USC)
-%\usepackage{color-edits}
-\usepackage[suppress]{color-edits}  % use this to suppress the package
-\addauthor{as}{red}    % as for Alex
-\addauthor{ga}{blue}  % ga for Guy
-\addauthor{sw}{orange} % sw for Steven
-\addauthor{ym}{magenta} % ym for Yishay
-% e.g. for Alex, provides \asedit{}, \ascomment{} and \asdelete{}.
-
-\setlength{\tabcolsep}{8pt} % Default value: 6pt
-\renewcommand{\arraystretch}{1.5} % Default value: 1
-
-\begin{document}
-
-%\begin{titlepage}
-
-\title{\vspace{-12mm}Competing Bandits:\\
-The Perils of Exploration under Competition%
-\thanks{Preliminary abstracts have appeared as
-\citet*{CompetingBandits-itcs18} and \citet*{CompetingBandits-ec19}.
-This is a merged version, with a unified and streamlined presentation and expanded related work and background materials.
-\newline \indent %All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
-Appendices~\ref{app:bg},\ref{app:examples} are new.
-We would like to thank Ian Ball, Yeon-Koo Che, Sven Rady, Sara Shahanaghi and Glen Weyl for helpful comments and conversations.
-Likewise, we thank the audience of the conference and seminar talks on this work.
-%seminar participants at Columbia and conference participants at Innovations in Theoretical Computer Science 2018, ACM Economics and Computation 2019 and the MIT Conference on Digital Experimentation 2020, for helpful comments and conversations.
-All errors are our own.}}
-% \textit{Competing Bandits: Learning under Competition}, at Innovations in Theoretical Computer Science 2018, and \textit{The Perils of Exploration under Competition: A Computational Modeling Approach}, at ACM Economics and Computation 2019.}}
-
-\author{Guy Aridor
-\footnote{Columbia University, Department of Economics. Email: g.aridor@columbia.edu.}
-\and \hspace{-0.5cm}
-\rule{0.0in}{0pt}
-Yishay Mansour
-\footnote{Google and Tel Aviv University, Department of Computer Science. Email: mansour.yishay@gmail.com.}
-\and \hspace{-0.5cm}
-\rule{0.0in}{0pt}
-Aleksandrs Slivkins
-\footnote{Microsoft Research New York City. Email: slivkins@microsoft.com.}
-\and \hspace{-0.5cm}
-\rule{0.0in}{0pt}
-Zhiwei Steven Wu%
-\footnote{Carnegie Mellon University, Pittsburgh, PA
-Email: zstevenwu@cmu.edu.\newline
-Research done when Z.S. Wu was an intern and a postdoc at Microsoft Research NYC.}
-}
-\date{}
-%\date{First version: July 2020\\ This version: December 2020}
-
-\maketitle
-
-\vspace{-5mm}
-\begin{abstract}
 Most online platforms learn from interactions with users, and engage in \emph{exploration}: making potentially suboptimal choices in order to acquire new information.
@@
 %Our findings are closely related to the ``competition vs. innovation" relationship, and elucidate the first-mover advantage in the digital economy.
-\\
-
-\vspace{0.1in}
-\noindent\textbf{Keywords:}  Competition vs. innovation, exploration-exploitation tradeoff, multi-armed bandits, regret.
-
-\bigskip
-\end{abstract}
-\setcounter{page}{0}
-\thispagestyle{empty}
-%\end{titlepage}
-
-\newpage
-\addtocontents{toc}{\protect\setcounter{tocdepth}{0}}
-%\begin{small}
-%\setcounter{tocdepth}{2}
-%\tableofcontents
-%\end{small}
-\setcounter{page}{1}
-%\thispagestyle{empty}
-%\newpage
-
-\onehalfspacing
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below.
-%
-\section{Introduction}
-\label{sec:intro}
-
-\subfile{content/sec-intro}
-
-\section{Related work}
-\label{sec:related-work}
-\subfile{content/related_work}
-
-\section{Our model in detail}
-\label{sec:model}
-\subfile{content/sec-model}
-
-\section{Theoretical results: the \TheoryModel}
-\label{sec:theory}
-\subfile{content/sec-theory}
-
-%\section{Full rationality (HardMax)}
-%\label{sec:rational}
-%\subfile{itcs18paper/sec-rational}
-
-%\section{Relaxed rationality: HardMax \& Random}
-%\label{sec:random}
-%\subfile{itcs18paper/sec-random}
-
-%\section{SoftMax response function}
-%\label{sec:soft}
-%\subfile{itcs18paper/sec-soft}
-
-%\section{Economic implications}
-%\label{sec:welfare}
-%\subfile{itcs18paper/sec-welfare}
-
-\section{Numerical simulations: the \ExptsModel}
-\label{sec:sim}
-%\label{sec:sim_details}
-
-\subfile{content/sim_details}
-
-\subfile{ec19paper/content/perf_in_iso}
-
-\subfile{ec19paper/content/inverted_u}
-
-\subfile{ec19paper/content/barriers}
-
-\subfile{ec19paper/content/non_greedy_choice}
-
-\subfile{ec19paper/content/revisited}
-
-\subfile{content/conclusion}
-
-%\newpage
-\addcontentsline{toc}{section}{References}
-
-\bibliographystyle{ecta.bst}
-\begin{small}
-\bibliography{bib-abbrv,bib-ML,refs,bib-bandits,bib-AGT,bib-slivkins, bib-random}
-\end{small}
-
-\clearpage
-
-\renewcommand{\appendixname}{Appendix}
-\renewcommand{\appendixtocname}{List of appendices}
-
-      % body goes here
-\appendix
-\renewcommand{\contentsname}{Appendix, For Online Publication Only}
-
-\tableofcontents
-\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}
-      % appendices go here
-
-\newpage
-
-%In the appendix, we provide background on multi-armed bandits as well as several discussions and proofs omitted from the main text.
-Furthermore, we provide plots and tables for our experiments, which were omitted from the main text.
-In all cases, the plots and tables here are in line with those in the main text, and lead to similar qualitative conclusions.
-
-\section[Background for non-specialists: multi-armed bandits]{Background for non-specialists: \\multi-armed bandits}
-\label{app:bg}
-%\label{sec:related-classes}
-\subfile{content/sec-bg}
-
-\section{Monotone MAB algorithms}
-\label{app:examples}
-\subfile{itcs18paper/app-examples}
-
-\section{Non-degeneracy via a random perturbation}
-\label{app:perturb}
-\subfile{itcs18paper/app-perturb}
-
-\section{Full proofs for Section~\ref{sec:theory}}
-\label{sec:theory-proofs}
-\subfile{content/sec-theory-proofs}
-
-\section{Full experimental results}
-\label{app:expts}
-
-\subfile{ec19paper/content/appendix_for_one_version}
-
-%\subfile{content/old-intro}
-
-\end{document}
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2021-03-02 journal paper/journal_main.tex
Revised the main file a bit, in particular, to input a "long" abstract file.

@@
 \citet*{CompetingBandits-itcs18} and \citet*{CompetingBandits-ec19}.
-This is a merged version, with a unified and streamlined presentation and expanded related work and background materials.
-\newline \indent %All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
+This version features a unified and streamlined presentation and expanded related work and background materials.
+All theoretical results are from \citet{CompetingBandits-itcs18}, and all numerical simulations are from \citet{CompetingBandits-ec19}.
 Appendices~\ref{app:bg},\ref{app:examples} are new.
-We would like to thank Ian Ball, Yeon-Koo Che, Sven Rady, Sara Shahanaghi and Glen Weyl for helpful comments and conversations.
-Likewise, we thank the audience of the conference and seminar talks on this work.
+\newline \indent 
+We thank Ian Ball, Yeon-Koo Che, Sven Rady, Sara Shahanaghi and Glen Weyl for helpful comments and conversations. We also thank the audience of the conference and seminar talks on this work.
 %seminar participants at Columbia and conference participants at Innovations in Theoretical Computer Science 2018, ACM Economics and Computation 2019 and the MIT Conference on Digital Experimentation 2020, for helpful comments and conversations.
@@
 }
-\date{}
-%\date{First version: July 2020\\ This version: December 2020}
+
+\date{First version: July 2020\\ This version: March 2021}
 
@@
 \begin{abstract}
-Most online platforms learn from interactions with users, and engage in \emph{exploration}: making potentially suboptimal choices in order to acquire new information.
-We study the interplay between \emph{exploration} and \emph{competition}: how such platforms balance the exploration for learning and competition for users.
-%Here users play three distinct roles: they are customers that generate revenue, they are sources of data for learning, and they are self-interested agents which choose among the competing platforms.
+\input{abstract-long}
+\end{abstract}
 
-We consider a stylized duopoly in which two firms face the same multi-armed bandit problem.
-Users arrive one by one and choose between the two firms, so that each firm makes progress on its bandit problem only if it is chosen.
-We study whether competition incentivizes the adoption of better algorithms.
-We find that stark competition disincentivizes exploration, leading to low welfare.
-%induces firms to commit to a ``greedy" algorithm that leads to low welfare.
-However, weaker competition
-%by providing firms with some ``free" users
-incentivizes better exploration algorithms and increases welfare.  We investigate two channels for weakening the competition: stochastic user choice models and a first-mover advantage.
-Our findings speak to the competition-innovation relationship and the first-mover advantage in the digital economy.
-%Our findings are closely related to the ``competition vs. innovation" relationship, and elucidate the first-mover advantage in the digital economy.
-\\
-
-\vspace{0.1in}
-\noindent\textbf{Keywords:}  Competition vs. innovation, exploration-exploitation tradeoff, multi-armed bandits, regret. \\
+\noindent\textbf{Keywords:}  Competition vs. innovation, exploration vs. exploitation, multi-armed bandits, regret. \\
 \noindent\textbf{JEL Codes}: D83, L15, O31
 
-\bigskip
-\end{abstract}
-\setcounter{page}{0}
+%\setcounter{page}{0}
 \thispagestyle{empty}
@@
 %\end{small}
-\setcounter{page}{1}
+%\setcounter{page}{1}
 %\thispagestyle{empty}

### 2021-03-02 journal paper/ec19paper/content/inverted_u.tex
fix weird bug

--- 2021-01-19 journal paper/ec19paper/content/inverted_u.tex
@@
 \end{table}
-\OMIT{
-\begin{table}[h]
-\centering
-\caption{Heavy Tail}
-\begin{tabular}{|c|c|c|c|}
-\hline
-   & $\TS$  & $\DEG$  & $\DG$ \\ \hline
-$\TS$
-    & \makecell{\textbf{0.50, 0.50}}
-    & \makecell{\textbf{0.3, 0.7}}
-    & \makecell{\textbf{0.29, 0.71}} \\\hline
-$\DEG$
-    & \makecell{\textbf{0.7, 0.3}}
-    & \makecell{\textbf{0.50, 0.50}}
-    & \makecell{\textbf{0.38, 0.62}} \\\hline
-$\DG$
-    & \makecell{\textbf{0.71, 0.29}}
-    & \makecell{\textbf{0.62, 0.38}}
-    & \makecell{\textbf{0.50, 0.50}} \\\hline
-\end{tabular}
-\end{table}
-
-\begin{table}[h]
-\centering
-\caption{Needle In Haystack}
-\begin{tabular}{|c|c|c|c|}
-\hline
-   & $\TS$  & $\DEG$  & $\DG$ \\ \hline
-$\TS$
-    & \makecell{\textbf{0.50, 0.50}}
-    & \makecell{\textbf{0.57, 0.43}}
-    & \makecell{\textbf{0.64, 0.36}} \\\hline
-$\DEG$
-    & \makecell{\textbf{0.43, 0.57}}
-    & \makecell{\textbf{0.50, 0.50}}
-    & \makecell{\textbf{0.54, 0.46}} \\\hline
-$\DG$
-    & \makecell{\textbf{0.36, 0.64}}
-    & \makecell{\textbf{0.43, 0.57}}
-    & \makecell{\textbf{0.50, 0.50}} \\\hline
-\end{tabular}
-\end{table}
-}
 
@@
 In stylized form, this relationship is captured in Figure~\ref{fig:inverted-U-expts}.%
-\footnote{We consider the monopoly scenario for comparison only, without presenting any findings. We just assume that a monopolist chooses
-the greedy algorithm, because it is easier to deploy in practice.
+\footnote{We consider the monopoly scenario for comparison only.
+We just assume that a monopolist chooses the greedy algorithm, because it is easier to deploy in practice.
 Implicitly, users have no ``outside option": the service provided is an improvement over not having it (and therefore the monopolist is not incentivized to deploy better learning algorithms).
@@
 Instead, the firm wants to improve its performance as much as possible by the time competition starts.
-Essentially, the firm only faces a classical explore-exploit trade-off, and is incentivized to choose algorithms that are best at optimizing this trade-off.
+Essentially, the firm only faces a classical explore-exploit trade-off, and chooses algorithms that are best at optimizing this trade-off.
 
@@
 
-\begin{table*}[h]
+\begin{table*}[ht]
 \centering
@@
 
-\OMIT{As we increase the number of firms, a single
-individual firm gets fewer consumers for learning as consumers switch
-between firms more often. Additionally, bad luck with rewards or arm
-selection is punished even more harshly as we increase the number of
-firms. This further strengthens the intuition from our previous
-result, which is that learning may be hard under competition since
-exploration and mistakes are punished harshly but are necessary for
-learning.}
-
 \end{document}

### 2021-03-02 journal paper/content/sec-intro.tex

@@
 
-%The main high-level question we ask is:
+\asedit{The main high-level question we ask is:}
 %that we focus on in this paper is:
 %\begin{align}\label{eq:main-Q}
-\asedit{We ask:}
-\textbf{Whether and how does competition incentivize the adoption of better exploration algorithms?}
+%\asedit{We ask:}
+\textbf{Whether and how does competition \asedit{between platforms} incentivize the adoption of better exploration algorithms?}
 %\end{align}
@@
 We offer a mix of theoretical results and numerical simulations, in which we study complex interactions between platforms' learning dynamics and users' self-interested behavior.
+\asedit{Prior work on exploration vs.
+competition targets technically very different models of competition which are not amenable to our high-level question (as we discuss in Section~\ref{sec:related-work}).}
 

### 2021-03-02 journal paper/content/related_work.tex
Minor edits in related work.

@@
 
-\asedit{The tension between exploration and incentives}
+\asedit{A monopolistic bandit algorithm may interact with self-interested parties, leading to a tension between exploration and incentives. This tension has been studied in several scenarios:}
+%\asedit{The tension between exploration and incentives}
 %The three-way tradeoff between exploration, exploitation and incentives
-has been studied in several scenarios very different from ours:
+%has been studied in several scenarios very different from ours:
 incentivized exploration in recommendation systems
@@
     \citep[\eg][]{RepeatedPA-ec14,Ghosh-itcs13}.
+\asedit{Unlike this work, we focus on incentives created in a competition.}
 %A literature review of this work can be found in
@@
 \xhdr{Exploration and competition.}
-\citet{bergemann1997market,bergemann2000experimentation,keller2003price} study the interplay of exploration and competition for users when the competing firms experiment with \emph{prices}
-(whereas in our model they experiment with design alternatives).
+\asedit{Several papers consider exploration algorithms in scenarios when the explorer is not a monopolist.
+The technical models are very different, and not amenable to the high-level question articulated in the Introduction.}
+
+Most relevant are \citet{bergemann1997market,bergemann2000experimentation,keller2003price}.
+They study the interplay of exploration and competition for users when the competing firms experiment with \emph{prices} (whereas in our model the firms experiment with design alternatives).
 All three papers consider strategies that respond to competition and analyze Markov-perfect equilibria (whereas we focus on the adoption of better bandit algorithms).
@@
 Thus, we have exploration algorithms which interact with each other strategically, \eg each agent prefers to free-ride on someone else's exploration.
-However, the agents do not compete with each other in any meaningful sense.
+\asedit{However, this work is all about cooperation (or lack thereof), rather than competition.}
+%the agents do not compete with each other in any meaningful sense.
 

### 2021-03-02 journal paper/content/conclusion.tex
remove old color-edits

--- 2020-12-23 journal paper/content/conclusion.tex
@@
 We study the tension between exploration and competition.
-We consider a stylized duopoly model in which two firms face an identical multi-armed bandit problem and compete for a stream of users.
-A firm makes progress on its learning problem if and only if it attracts users.
+We consider a stylized duopoly model in which two firms face the same multi-armed bandit problem and compete for a stream of users.
+A firm makes progress on its learning problem only if it attracts users.
 %\asdelete{We investigate two variants: an analytical variant where users do not observe any performance signals and choose between firms according to their Bayesian-expected rewards, and numerical simulations where a reputation score is observed for each firm, based on the average reward of its recent users.
@@
 We find that firms are incentivized to adopt a ``greedy algorithm" which does no purposeful exploration and leads to welfare losses for users.
-We then consider two relaxations of competition: we soften users' decision rule and give one of the firms a first-mover advantage.
+We then consider two relaxations of competition: we soften users' decision rule and give one firm a first-mover advantage.
 Both relaxations induce firms to adopt ``better" bandit algorithms, which benefits user welfare.

### 2021-03-02 journal paper/content/related_work.tex
remove old color-edits

@@
 \cite{de2020data} note the role of data as a barrier to entry in online markets; we find a similar effect with ``data advantage".
-\citet{kerin1992first} overview other channels through which first-mover advantage can \asedit{affect competition}.% lead to a competitive advantage.
+\citet{kerin1992first} overview other channels through which first-mover advantage can affect competition.% lead to a competitive advantage.
 
@@
 
-\asedit{While we use first-mover advantage and agents' decision rule,} classic ``market competitiveness" measures, such as the Lerner Index or the Herfindahl-Hirschman Index
+While we use first-mover advantage and agents' decision rule, classic ``market competitiveness" measures, such as the Lerner Index or the Herfindahl-Hirschman Index
 \citep{tirole1988theory}, are not applicable to our setting, as they rely on ex-post observable market attributes such as prices or market shares (which are, resp., absent and endogenous for us).

### 2021-03-02 journal paper/content/sec-intro.tex
remove old color-edits

@@
 We focus on innovation in \emph{exploration technology} which systematically improves the firm's products, whereas prior work would define innovation as improvement in the products themselves.
-We recover the \asdelete{traditional} inverted-U relationship purely through the reputational consequences of exploration.
+We recover the traditional inverted-U relationship purely through the reputational consequences of exploration.
 \asdelete{rather than monetary costs/benefits thereof.}
@@
 We find that the greedy algorithm often wins under the \HardMax decision rule, with a strong evidence of the ``death spiral" effect mentioned earlier.
-As predicted by the theory, better algorithms prevail under \HardMaxRandom
-\asedit{with enough ``random" users.}
+As predicted by the theory, better algorithms prevail under \HardMaxRandom with enough ``random" users.
 %if the expected number of ``random" users is sufficiently large.
@@
 We revisit algorithms' performance in a standalone bandit problem, \ie without competition.
-We find that the most natural performance measure does not explain this phenomenon, and suggest a new, more nuanced \asedit{one} that does.
+We find that the most natural performance measure does not explain this phenomenon, and suggest a new, more nuanced one that does.
 %performance measure that does.
@@
 
-Interestingly,
-\asdelete{we find that}
-data feedback loops can be bad for innovation (as in ``death spiral"), or good for innovation (as in ``amplification effect"), depending on the level of competition.
+Interestingly, data feedback loops can be bad for innovation (as in ``death spiral"), or good for innovation (as in ``amplification effect"), depending on the level of competition.
 

### 2021-03-02 journal paper/content/sec-model.tex
remove old color-edits

@@
         $\est_i(t) = \PMR_i(t) := \E\sbr{ r_t\mid i_t = i}$
-    for each principal $i$, \asedit{where the agent knows $t$, the principals' algorithms, $\priorMu$, and $\respF$.}
+    for each principal $i$, where the agent knows $t$, the principals' algorithms, $\priorMu$, and $\respF$.
 %\footnote{Agents do the Bayesian update knowing $t$, the principals' algorithms, the prior $\priorMu$, and $\respF$.}
@@
 We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm.
-(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.) Principals are risk-neutral; their utility is defined as their market share, \ie the number of agents \asedit{they attract}.
+(Since MAB algorithms can be randomized, we make no distinction between pure and mixed strategies.) Principals are risk-neutral; their utility is defined as their market share, \ie the number of agents they attract.
 Thus, they aim to select the algorithm that maximizes their expected market share.

### 2021-03-02 journal paper/content/sec-theory.tex
remove old color-edits

@@
 The core technical result is that if one principal adopts \DynGreedy, then the other principal loses all agents as soon as he deviates therefrom.
-\asedit{We make this formal below.}
+We make this formal below.
 
@@
 
-Thus, the greedy modification is beneficial in competition \emph{and} keeps us inside the set of \bmonotone MAB algorithms. Consequently, if the principals are restricted \asedit{to the latter set,}
+Thus, the greedy modification is beneficial in competition \emph{and} keeps us inside the set of \bmonotone MAB algorithms. Consequently, if the principals are restricted to the latter set,
 %to choosing \bmonotone MAB algorithms,
@@
 Suppose initially both principals use some algorithm \alg, and principal 1 ponders switching to another algorithm \alg' which BIR-dominates \alg.
-\asedit{What is the marginal utility $\deltaU$ of this switch?}
+What is the marginal utility $\deltaU$ of this switch?
 
@@
 If agents have common ``point beliefs" %respectively, algorithms $(\alg[1]',\alg[2]')$ prior $\priorMu'$ and response function $\respF'$,
-\asedit{on these objects}, then all our results carry over with respect to these beliefs.
+on these objects, then all our results carry over with respect to these beliefs.
 

### 2021-03-02 journal paper/content/sim_details.tex
remove old color-edits

--- 2021-01-19 journal paper/content/sim_details.tex
@@
 For each \MRV we draw a table of realized rewards (\emph{realization table}), and use this same table for all experiments on this \MRV.
-This ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings.
+This ensures that differences in algorithm performance are solely due to differences in the algorithms in the different experimental settings.
 

### 2021-03-02 journal paper/ec19paper/content/revisited.tex
more space saving edits
remove old color-edits

--- 2021-01-19 journal paper/ec19paper/content/revisited.tex
@@
 \textit{
-Mean reputation trajectories do not suffice to explain the outcomes under competition.}
+Mean reputation trajectories do not explain the outcomes under competition.}
 \end{finding}
@@
 In the former case \DynamicGreedy does slightly better than $\Thompson$, and in the latter case it does substantially worse.
-However, the mean reputation trajectory may fail to capture this complexity since it simply takes average over different \MRVs.
-This may be inadequate for explaining the outcome of the duopoly game, given that the latter is determined by a simple comparison between the firm's reputation scores.
+However, the mean reputation trajectory fails to capture this complexity since it takes average over different \MRVs.
+This is inadequate for explaining the outcome of the duopoly game, given that the latter is determined by a comparison between the firm's reputation scores.
 

### 2021-03-02 journal paper/content/related_work.tex
Removed some old coloring in a few more places.

@@
 Each radio chooses channels over time using a multi-armed bandit algorithm.
-This work studies a repeated game between bandit algorithms, and focuses on designing algorithms which work well in this game\gadelete{, under various assumptions on communication, synchronization and collisions}.
+This work studies a repeated game between bandit algorithms, and focuses on designing algorithms which work well in this game.
+%\gadelete{, under various assumptions on communication, synchronization and collisions}.
 

### 2021-03-02 journal paper/content/sec-intro.tex
missed this one
Removed some old coloring in a few more places.

@@
 We focus on innovation in \emph{exploration technology} which systematically improves the firm's products, whereas prior work would define innovation as improvement in the products themselves.
-We recover the traditional inverted-U relationship purely through the reputational consequences of exploration.
-\asdelete{rather than monetary costs/benefits thereof.}
-By contrast, the inverted-U relationships from prior work rely on the monetary aspects: \asedit{investments into  R\&D and profits from innovation.}
+We recover the inverted-U relationship purely through the reputational consequences of exploration.
+%\asdelete{rather than monetary costs/benefits thereof.}
+By contrast, the inverted-U relationships from prior work rely on the monetary aspects: investments into  R\&D and profits from innovation.
 
